### 2011-11-15 src/synopsis.tex
setting up with github
v0.1 of synopsis.

\documentclass[MS]{iitmdiss}

\usepackage{times} 
\usepackage{t1enc}
\usepackage{../lib/mylatexlib}

\title{\mythesistitle} 
\author{\myname} 
\date{\MakeUppercase
  {\mysubmissionmonth}} 
\department{\MakeUppercase {\mydept}}

\begin{document}

\nocite{*}
\maketitle

\begin{singlespace}
\tableofcontents
\thispagestyle{empty}
% \listoftables
% \addcontentsline{toc}{chapter}{LIST OF TABLES}
% \listoffigures
% \addcontentsline{toc}{chapter}{LIST OF FIGURES}
\end{singlespace}

% The main text will follow from this point so set the page numbering
% to arabic from here on.
\pagenumbering{arabic}
\setcounter{page}{0}
%\newpage

%\chapter{\MakeUppercase{Introduction}}
\chapter{INTRODUCTION}
Consecutive ones property (COP) of binary matrices is a widely studied
combinatorial problem. The problem is to rearrange rows (columns) of a
binary matrix in such a way that every column (row) has its $1$s occur
consecutively. If this is possible the matrix is said to have the COP.
\begin{figure}[h]
  \centering

  \caption{(i) matrix with COP (ii) without COP}
  \label{fig:cop-matrix}
\end{figure}

This problem has several practical applications in diverse fields
including scheduling \cite{hl06}, information retrieval \cite{k77} and
computational biology \cite{abh98}.  Further, it is a tool in graph
theory \cite{mcg04} for interval graph recognition, characterization
of Hamiltonian graphs, and in integer linear programming
\cite{ht02,hl06}.

\tnote[a]{define ``COP permutation''}

\section{Background}
Different aspects of COP were studied as part of this thesis. They are
described in the following subsections.

\subsection{Consecutive Ones Testing}
The obvious first questions after being introduced to the consecutive
ones property of binary matrices are if COP can be detected
efficiently in a binary matrix and if so, can the COP permutation of
the matrix also be computed efficiently?  Recognition of COP in a
binary matrix is polynomial time solvable and the first such algorithm
was given by \cite{fg65}.  Since then several algorithms have been
invented; PQ trees \cite{bl76}, variations of PQ trees
\cite{mm09,wlh01,wlh02,mcc04}, ICPIA \cite{nsnrs09} are some of them.

\cite{bl76} showed for the first time that COP testing (COT) is indeed
polynomial time solvable by describing the construction of a data
structure called PQ trees. This paper draws on the close relationship
of COP matrices to interval graphs. A PQ tree of a matrix is one that
stores all row (column) permutations of the matrix that give the COP
orders (there could be multiple) of the matrix. This is constructed using an elaborate procedure in the linear
time algorithm described in \cite{bl76}.\tnote[a]{verify if it is
  linear}. Since this result, several improvements have been made to
the procedural complexity of the algorithm itself and also derivative
data structures that are easier to construct. PQR trees is one such
data structure \cite{mm96}, \cite{mpt97}\tnote{also, mpt97 $==$
  elaborate version of mm96}.  \cite{tm03} describes an improved
version of PQR trees. \tnote[a]{improvement in terms of
  what?}. \cite{h01a} describes the simpler algorithm for
COT. \tnote[a]{This result first apprearted in proc ISAAC 92} Hsu also
invented PC trees \cite{h01b} which is claimed to be much easier to
implement. \cite{nsnrs09} describes a characterization of consecutive
ones property solely based on the cardinality properties of the set representations
of the columns (rows); every column (row) is equivalent to a set that
has the row (column) indices of the rows (columns) that have one
entries in this column (row). This is interesting and relevant,
especially to this thesis because it simplifies COT to a great
degree. \tnote[a]{because it reduces the search space for the solution
  from set of all permutations of rows to set of all-- get this
  figured out!}

\cite{mcc04} describes a different approach to COT. While all previous
COT algorithms gave the COP order if the matrix has the property and
if not it exited stating negative, this algorithm gives an evidence by
way of a certificate of
matrix being devoid of COP thereby enabling a user to check the
algorithm's result. This is significant from an implementation
perspective because automated program verification is hard and manual
verification is more viable. Hence having a certificate
reinforces its credibility. Note that when the matrix has COP the COP
order is the certificate.  It may also be noted that the internal
machinery of this algorithm \annote[a] {is related to the weighted
  betweenness problem addressed } {in what way??} in \cite{co97}.
\tnote[a]{expand on the COP order graph creation and it having to be
  bipartite for M to have COP. and thus an odd cycle being an evidence
  of no COP.}

\tnote[a]{THESIS: where should this go?: \cite{jlm97} (application of
  PQ trees in graphics).}

% \noindent
To elaborate on what was mentioned earlier about \cite{nsnrs09}, the
problem of COT can be seen as a constraint satisfaction problem
involving a system of sets from a universe as follows. Every column of
the binary matrix can be converted into a set of integers which are
the indices of rows with $1$s in that column. It is apparent that if
the matrix has COP, then constructing such sets after applying the COP
permutation to the matrix will result in all sets with consecutive
integers. In other words, after application of COP reordering, the
sets are intervals. Indeed the problem now becomes finding interval
assignments to a given set system such that there exists a permutation
of the universe (set of row indices) which converts each set to its
assigned interval. The result in \cite{nsnrs09} characterizes interval
assignments to the sets which can be obtained from a single
permutation of the rows.  They show that for each set, the cardinality
of the interval assigned to it must be same as the cardinality of the
set, and the intersection cardinality of any two sets must be same as
the intersection cardinality of the corresponding intervals.  While
this is obviously a necessary condition, this result shows this is
also sufficient.  \cite{nsnrs09} calls such an interval assignment an
Intersection Cardinality Preserving Interval Assignment (ICPIA).  This
paper generalizes the idea from \cite{wlh02} of decomposing a given
binary matrix into prime matrices for COT and describes an algorithm
to test if an ICPIA exists for a given set system.

The literature also has results on parallel COT algorithms --
\cite{as95} \cite{bs03} \tnote[a]{is this parallel?}
\cite{cy91}\tnote[a]{is this parallel?}

\tnote[a]{THESIS: add helly's theorem citation
  19XXdgk-Hellystheorem-Danzer-Gruenbaum-Klee}

\subsection{Matrix modification to attain COP or ``almost'' COP}
\tnote[a]{THESIS: results and survey from \cite{d08phd} Chapters 2 and
  3}

So far we have been concerned about matrices that have the consecutive
ones property. However in real life applications, it is rare that data
sets represented by binary matrices have COP. Nevertheless, the COP is
not arbirary and there are several interesting problems when a matrix
does not have COP but is ``close'' to having COP or can be altered to
have COP.

-- sect 4.1 in \cite{d08phd} has many results surveyed. hardness results,
approx. results. results are usually for a class of matrices $(a,b)$
where number columns and rows are restriced to $a$ and $b$
\tnote[a]{check if it is the other way around}.\\
-- problem of flipping at most $k$ entries of $M$ to make it attain
COP. this is NP complete \cite{b75-phd}\\

The first step in this direction is to analyze matrices without
COP. \cite{at72} showed that a matrix that does not have COP have
certain forbidden substructures that prevent it from having
COP. Tucker classified these substructures into five classes of
submatrices. This result is presented in the context of AT-free
bipartite graphs \tnote[a]{THESIS: check this up. give details.} which
is equivalent to matrices with COP. A binary matrix is converted into
a bipartite graph as follows. The set of rows and the set of columns
form the two partitions of the graph. Each row node is adjacent to
those nodes that represent the columns that have ones in the
corresponding row. \cite{at72} proves that this bipartite graph has no
asteroidal triple if and only if the matrix has COP and goes on to
identify the forbidden substructures for these bipartite graphs. The
matrices corresponding to these substructures are the forbidden
submatrices. Once a matrix has been detected to not have COP (using
any of the COT algorithms mentioned earlier), it is naturally of
interest to find out the smallest\tnote[a]{define ``smallest''}
forbidden substructure. \cite{d08phd} discusses an exact algorithm
with time complexity \annote[a]{parameterized by the number of ones in
  the matrix}{check this.}.

{\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are similarly
inspired problems. In {\sc Min COS-R (Min COS-C)} the question is to
find the minimum number of rows (columns) that must be deleted to
result in a matrix with COP.  In the dual problem {\sc Max COS-R (Max
  COS-C)} the search is for the maximum number of rows (columns) that
induces a submatrix with COP. Given a no-COP matrix $M$,
\cite{b75-phd} shows that finding a submatrix $M'$ with all columns
but a maxmimum cardinality subset of rows such that $M'$ has COP is NP
complete. \tnote[a]{check if b75 deals with COP col or COP row. also
  is it any submatrix with k less than r rows or submatrix must have
  all columns?}. \cite{hgxx} \tnote[a]{get year XX} corrects an error of the abridged proof
of this reduction as given in \cite{gj79}.  \cite{d08phd} discusses
all these problems in detail giving an extensive survey of the
previously existing results which are almost exhaustively all
approximation results and hardness results. Taking this further,
\cite{d08phd} presents new results in the area of parameterized
algorithms for this problem. \tnote[a]{THESIS: elaborate - what are
  the results?}.

Another problem is to find the minumum number of entries in the matrix
that can be flipped to result in a matrix with COP.  \cite{v85}
discusses approximation of {\sc COP Augmentation} which is the problem
of changing of the minimum number of zero entries to ones so that the
resulting matrix has COP. As mentioned earlier, this problem had been known
to be NP complete due to \cite{b75-phd}. \cite{v85} also proves, using
a reduction to the longest path problem, \tnote[a]{or is it a survey
  of another result?  check.} that finding a Tucker's forbidden
submatrix of at least $k$
rows is NP complete. \tnote[a]{how is this different from booth's 75 result??} \tnote{[tag:appoximation,bitFlipping,matrixBreaking]}\\

\tnote[a]{THESIS: where should this go? \cite{tz03} (approx submatrix with COP sparse matrices)}

%\subsection{Further extensions of the COP problem}

\cite{jkckv04} \tnote{[tag:storage]} discusses the use of matrices
with ``near'' COP in the storage of very large databases.
The problem is that of $k$-run reordering of binary matrices which
they prove to be NP hard using a reduction from the hamiltonian path problem.
A connection of COP problem to the travelling salesman problem is also
introduced. \tnote[a]{THESIS? what does this mean? -- COP can be used as a tool to reorder $0.5T \le runs(M) \le T$}

The optimization version of the $k$-run problem, i.e. minimization of
number of blocks of ones is proven to be NP complete by
\cite{kou77}. \tnote[a]{THESIS: what is the reduction?}

\tnote[a]{THESIS: 
-- {\em other problems similar to COP}
  \begin{enumerate}
  \item \cite{ckl96} (ILP, circ ones, one drop)
  \item \cite{th99} (generalization of COP - minimax, biotonic column) Tucker
  \end{enumerate}
}

\subsection{Graph Isomorphism}
As it will be described in detail later in this document, isomorphism
of certain classes of graphs, namely chordal graphs, have a close
relationship with consecutive ones property and generalizations of it.
This is perhaps because of how closely COP of a matrix relates to
properties of graphs derived from matrices as seen in the following
results. %-- Other relations of COP with Graphs: \\
\tnote[a]{THESIS: peo exists iff chordal. lexicographic BFS
  [tag:chordalGraph]} 
A well known result in perfect graph theory is that the maximal
cliques of an interval graph $G$ can be linearly ordered such that for
all $v \in V(G)$, cliques containing $v$ are consecutive in the
ordering\cite{gh64}. This clearly means that a graph $G$ is an
interval graph if and only if \tnote[a]{verify from paper the
  statement of claim.} maximal clique vertex incidence matrix of $G$
has COP.  Also, maximal cliques of any chordal graph can be enumerated
in polytime $O(m+n)$ \tnote[a]{citation?!!}.  \cite{fg65} uses these
results to give the first polynomial time algorithm for
COT. \tnote[a]{THESIS: check. how do they use it?}  A bipartite graph
is convex (on $R$) if and only if its half adjacency matrix has COP on
rows.  The results in \cite{bl62} on COT are based on the result that
interval graphs are AT-free chordal graphs. \tnote[a]{the latter being
  Tucker's?}

\tnote[a]{THESIS: -- How COP (and TPL) relates to interval graph
  isomorphism (and path graph isomorphism) move/summarize what is
  already in Motivation section.}

\tnote[a]{THESIS: TBD survey -- see blue notes under Graph
  Isomorphism. namely citations in \cite{aas93}}

\tnote[a]{THESIS: -- -- a brief on heirarchy of chordal graphs, path
  graphs, interval graphs, peo, clique tree etc. and results.}

\tnote[a]{THESIS: -- interval graphs are incomparability graphs. see
  Golumbic. [tag:classification]}

\tnote[a]{THESIS: finding min length hole in bipartite graph is
  polynomial time [Sec 3.3.2 in \cite{d08phd}]}

\tnote[a]{THESIS: theorem 2.2 in \cite{d08phd} - $G$ is union of
  v.d. caterpillars iff $M$ has COP in rows. $M$ is edge vertex
  incidence matrix $(*,2)$ of $G$}

\subsection{Logspace complexity of graph canonization using COP}

Canonization is an important tool in graph isomorphism. Invention of a
deterministic method of canonization for any class of graphs naturally
results in an algorithm for isomorphism. All that is required is to
check if the canons of two graphs are the same. Thus complexity of
graph isomorphism can be studied by studying canonization
methods. While general graph ismorphism remains elusive in terms of
hardness, canonization has been studied for smaller classes of graphs
thus giving complexity/hardness results for them.

In 1992, \cite{sl92} made an important discovery that tree isomorphism
is in logspace. It was done by inventing a method of canonization of
trees using a logspace depth first traversal
algorithm. \tnote[a]{confirm the statement of claim}.

\cite{kklv10} proved that interval graph canonization is also in
logspace thus drawing logspace conslusions about COT. 
% This result is
% built on the fact that 
Interval graphs are FP+C (fixed point with
counting) definable \tnote[a]{elucidate} and \cite{l10} showed that
this implies that it captures PTIME. \tnote[a]{elucidate} This result
along with that of undirected graph reachability being in logspace
\cite{r08}, \cite{kklv10} proved their logspace result.\\

\tnote[a]{THESIS: -- -- the interesting survey in conclusion section
  of kklv10}

\tnote[a]{THESIS:
{\bf COP helping some problems with hardness}
having COP in the input structure makes some problems less hard than
in a general input.

-- min set cover problem is known to be NP hard. so is min hitting
set. these two problems are equivalent \cite{adp80}\\
-- -- complexity results: log factor polynomial time approximable \cite{f98}, W[2]-complete with solution size as parameter \cite{df99}\\
-- a couple of generalizations:\\
-- -- min degree hypergraph (MDH) problem\\
-- -- red blue covering (RBSC) problem\\
}

\tnote[a]{THESIS: \cite{d08phd} Chapter 5.}

\tnote[a]{THESIS: what are \cite{d08phd} results}

\tnote[a]{THESIS: -- Rectangle stabbing \tnote[a]{see \cite{d08phd} Chapter 6?}}

\tnote[a]{THESIS: -- {\em \cite{dgnw06} (min membership hitting set + COP)}}

\section{Motivation, Objective and Scope}

A natural generalization of the interval assignment problem is
feasible tree path labeling problem of a set system. The problem is
defined as follows -- given a set system $\cF$ from a universe $U$ and
a tree $T$, does there exist a bijection from $U$ to the vertices of
$T$ such that each set in the system maps to a path in $T$.  We refer
to this as the {\em tree path labeling problem} for an input set
system, target tree pair -- $(\cF,T)$. As a special case if the tree
$T$ is a path, the problem becomes the interval assignment problem.
We focus on the question of generalizing the notion of an ICPIA
\cite{nsnrs09} to characterize feasible path assignments.  We show
that for a given set system $\cF$, a tree $T$, and an assignment of
paths from $T$ to the sets, there is a feasible bijection between $U$
and $V(T)$ if and only if all intersection cardinalities among any
three sets (not necessarily distinct) is same as the intersection
cardinality of the paths assigned to them and the input passes a
filtering algorithm (described in this paper) successfully.  This
characterization gives a natural data structure that stores all the
relevant feasible bijections between $U$ and $V(T)$. This reduces the
search space for the solution considerably from the universe of all
possible bijections between $U$ and $V(T)$ to only those bijections
that maintain the characterization\tnote[a]{REWORD}.  Further, the
filtering algorithm is also an efficient algorithm to test if a tree
path labeling to the set system is feasible.  This generalizes the
result in \cite{nsnrs09}.

Our results also have close connection to recognition of path graphs
and connection to path graph isomorphism.  Graphs which can be
represented as the intersection graph of paths in a tree are called
{\em path graphs}\cite{mcg04}. Thus given a hypergraph $\cF$, it can
be viewed as paths in a tree, if and only if the intersection graph of
$\cF$ is a {\em path graph}. Path graphs are a subclass of chordal
graphs since chordal graphs are combinatorially characterized as the
intersection graphs of subtrees of a tree.  Path graphs are well
studied in the literature \cite{plr70,gav78,bp93,mcg04}.  Checking if
a graph is a path graph can be done in polynomial time by the results
of \cite{gav78,aas93}.  However, this is only a necessary condition
for our question.  \remove[a]{Indeed, our work provides a sufficient
  condition for a related problem: that is whether a path assignment
  to be feasible?}  On the other hand, path graph isomorphism is known
be isomorphism-complete; see for example \cite{kklv10}. Therefore,
\annote[a]{it is unlikely that we can solve the problem of finding
  feasible path labeling $\cl$ for a given $\cF$ and }{WHY?} tree $T$.
\annote[a]{It is definitely interesting to classify the kinds of trees
  and hypergraphs for which feasible path labelings can be
}{ELABORATE.} found efficiently.  \annote[a]{These results would form
  a natural generalization of COP testing and interval graph
  isomorphism, culminating in Graph Isomorphism }{REWORD} itself.

%\noindent 
\section{Summary of Results}
Given a path labeling $\cl$ to $\cF$ from a tree $T$, we give a
necessary and sufficient condition for it to be a feasible path
labeling.  This necessary and sufficient condition can be tested in
polynomial time.  The most interesting consequence is that in our
constructive procedure, it is sufficient to iteratively check if
three-way intersection cardinalities are preserved.  In other words,
in each iteration, it is sufficient to check if the intersection of
any three sets is of the same cardinality as the intersection of the
corresponding paths.  This generalizes the well studied question of
the case when the given tree $T$ is a path \cite{wlh02,nsnrs09}.

%\item 
In the Section 4, \annote[a]{we initiate an exploration of}{WTF?}
finding feasible path labeling of set systems in a special kind of
tree which we call the $k$-subdivided star.  \annote[a]{This question
  is an attempt to generalize}{REWORD.} the problem of testing if a
matrix has the consecutive ones property.  However, we restrict the
hypergraph $\cF$ to be such that all hyperedges have at most $k+2$
elements.  \annote[a]{In spite of this restricted case we consider, we
  believe that our results are of significant interest in
  understanding the nature of Graph Isomorphism}{REWORD.} which is
polynomial time solvable in interval graphs and is hard on path
graphs.
%\end{enumerate}

% In the later part of this paper, we decompose our search for a
% bijection between $U$ and $V(T)$ into subproblems.  Each subproblem
% is on a set subsystem in which for each set, there is another set in
% the set subsystem with which the intersection is {\em strict}, i.e.,
% there is a non-empty intersection, but neither is contained in the
% other.  This is in the spirit of results in \cite{wlh02,nsnrs09}
% where to test for the COP in a given matrix, the COP problem is
% solved on an equivalent set of prime matrices.

\section{Organization of document} 
\tnote[a]{THESIS: UNCOMMENT} 
% The necessary preliminaries with definitions etc. are presented in
% Section~\ref{sec:prelims}. 
Section~\ref{sec:feasible} documents the
characterization of a feasible path labeling and finally,
Section~\ref{sec:ksubdivstar} describes a polynomial time algorithm to
find the tree path labeling of a given set system from a given
$k$-subdivided tree.

%\chapter{\MakeUppercase{Results}}
\chapter{RESULTS}
% \section{Preliminaries} 
% \label{sec:prelims}

% \subsection{Hypergraph} 
% The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
%   system} of a universe $U$ with $|U| = n$.  The {\em support} of a
% set system $\F$ denoted by $supp(\cF)$ is the union of all the sets in
% $\F$; $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this
% paper, a set system is required to ``cover'' the universe; $ supp(\cF)
% = U$.
% % A set system $\cF$ can be alternatively represented by a {\em
% %   hypergraph} $\H_\cF$ whose vertex set is $supp(\cF)$ and
% % hyperedges are the sets in $\cF$. This is a known representation for
% % interval systems in literature \cite{bls99,kklv10}.  We extend this
% % definition here to path systems.
% A set system $\cF$ can also be visualized as a {\em
%   hypergraph}\, vertex set is $supp(\cF)$ and hyperedges are the sets
% in $\cF$.  The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph
% $\cF$ is a graph such that its vertex set has a bijection to $\cF$ and
% there exists an edge between two vertices \iff their corresponding
% hyperedges have a non-empty intersection \cite{mcg04}.  Two
% hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to each
% other, denoted by $\cF' \cong \cF''$, \iff there exists a bijection
% $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for all sets $A
% \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ \iff  $B$ is a
% hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
% \cite{kklv10}, written as $B=\phi(A)$.
% % This is called {\em hypergraph isomorphism}.

% %\noindent
% \subsection{Path Hypergraph from a Tree} 
% We consider trees $T$ such that $|V(T)|=|U|=n$.  A {\em path system}\,
% $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P \mid P
% \subseteq V, \text{ } T[P] \text{ is a path} \}$.  This generalizes
% the fact, from the literature \cite{bls99,kklv10}, that intervals can
% be viewed as sub-paths of a path.
% % Due to the equivalence of set system and hypergraph in the scope of
% % this paper, we drop the subscript $_H$ in the notation and refer to
% % both the structures by $\cF$.
% % \vspace{\prelimspace}%\tnote[a]{also extend $\phi$ to hyperedges -- see if
% % required } \vspace{\prelimspace} \vspace{\prelimspace}
% % \noindent
% If the intersection graphs of $\cF$ and $\cP$( a path system) are
% isomorphic, $\bI(\cF) \cong \bI(\cP)$, then the associated bijection
% $\cl: \cF \rightarrow \cP$ due to this isomorphism is called a {\em
%   path labeling} of the hypergraph $\cF$.  Note that there are two
% kinds of isomorphisms here.  We are concerned about the isomrophisms
% intersection graphs on $\cF$ and $\cP$, and also the isomorphism
% between the hypergraph $\cF$ and $\cP$.
% % To illustrate further, let $\cg: V(\cF) \rightarrow V(\cP)$ be the
% % above mentioned isomorphism where $V(\cF)$ and $V(\cP)$ are the
% % vertex sets that represent the hyperedges for each hypergraph
% % respectively, $V(\cF) = \{ v_S \mid S \in \cF\}$ and $V(\cP) = \{
% % v_P \mid P \in \cP\}$. Then the path labeling $\cl$ is defined as
% % follows: $\cl(S_1) = P_1$ \iff $\cg (v_{S_1}) = v_{P_1}$.  Just to
% % emphasize, for a path labeling $\cl$ of $\cF$ with $\cP$ as the path
% % system, $\cF^\cl$ is same as $\cP$.  The path system $\cP$ may be
% % alternatively denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In
% % most scenarios in this paper, what is given are the pair $(\cF,
% % \cl)$ and the target tree $T$; hence this notation will be used more
% % often.
% If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$ is called
% a {\em path hypergraph} and $\cP$ is called {\em path representation}
% of $\cF$. If this isomorphism is $\phi: supp(\cF) \rightarrow V(T)$,
% then it is clear that there is a path labeling $\cl_\phi: \cF
% \rightarrow \cP$ to the set system; $\cl_\phi(S) = \set{y \mid y =
%   \phi(x), x \in S}$ for all $S \in \cF$. In other words, if $\cF
% \cong \cP$, we get a path labeling.  Recall that $supp(\cP) = V(T)$.
% % \vspace{\prelimspace} A graph $G$ is a {\em path graph} if it is
% % isomorphic to the intersection graph $\bI(\cP)$ of a path system
% % $\cP$.  This isomorphism gives a bijection $\cl': V(G) \rightarrow
% % \cP$. Moreover, for the purposes of this paper, we require that in a
% % path labeling, $supp(\cP) = V(T)$.  If graph $G$ is also isomorphic
% % to $\bI(\cF)$ for some hypergraph $\cF$, then clearly there is a
% % bijection $\cl: \cF \rightarrow \cP$ such that $\cl(S) = \cl'(v_S)$
% % where $v_S$ is the vertex corresponding to set $S$ in $\bI(\cF)$ for
% % any $S \in \cF$. This bijection $\cl$ is called the {\em path
% %   labeling} of the hypergraph $\cF$ and the path system $\cP$ may be
% % alternatively denoted as $\cF^\cl$.  \vspace{\prelimspace}
% In this work, we are given as input $\cF$ and a tree $T$, and the
% question is whether there is a path labeling $\cl$ to a set of paths
% in $T$.  We refer to such a solution path system by $\cF^\cl$.  A path
% labeling $(\cF, \cl)$ is defined to be {\em feasible} if
% % $\cF \cong \cF^\cl$ and this
% there is a hypergraph isomorphism $\phi: supp(\cF) \rightarrow
% supp(\cF^\cl)=V(T)$ induces a path labeling $\cl_\phi: \cF \rightarrow
% \cF^\cl$ such that $\cl_\phi = \cl$.
% % \vspace{\prelimspace}

% \subsection{Overlap Graphs and Marginal Hyperedges} 
% An {\em
%   overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is a graph such
% that its vertex set has a bijection to $\cF$ and there exists an edge
% between two of its vertices \iff their corresponding hyperedges
% overlap. Two hyperedges $S$ and $S'$ are said to {\em overlap},
% denoted by $S \overlap S'$, if they have a non-empty intersection and
% neither is contained in the other; $S \overlap S' \text{ iff } S \cap
% S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq S$. Thus $\bO(\cF)$
% is a spanning subgraph of $\bI(\cF)$ and not necessarily
% connected. Each connected component of $\bO(\cF)$ is called an {\em
%   overlap component}.
% % If there are $d$ overlap components in $\bO(\cF)$, the set
% % subsystems are denoted by $\cO_1, \cO_2, \ldots \cO_d$. Clearly
% % $\cO_i \subseteq \F, i \in [d]$. For any $i, j \in [d]$, it can be
% % verified that one of the following is true.
% % \begin{enumerate}[a) ]
% % \item $supp(\cO_i)$ and $supp(\cO_j)$ are disjoint
% % \item $supp(\cO_i)$ is a subset of a set in $\cO_j$
% % \item $supp(\cO_j)$ is a subset of a set in $\cO_i$
% % \end{enumerate}
% % \vspace{\prelimspace} \xnoindent
% A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
% \overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
% \cite{kklv10}. Additionally, if $S$ is such that it is contained in no
% other marginal hyperedge in $\cF$, then it is called {\em
%   super-marginal}.
% % i.e., it is inclusion maximal then it is called {\em
% %   super-marginal}.

% \subsection{$k$-subdivided Star -- A Special Tree} 
% A {\em star} graph
% is a complete bipartite graph $K_{1,p}$ which is clearly a tree and
% $p$ is the number of leaves. The vertex with maximum degree is called
% the {\em center} of the star and the edges are called {\em rays} of
% the star. A {\em $k$-subdivided star} is a star with all its rays
% subdivided exactly $k$ times. The definition of a {\em ray of a
%   $k$-subdivided star} is extended to the path from the center to a
% leaf. It is clear that all rays are of length $k+2$. See
% fig.~\ref{fig:kstar} for examples.

% \begin{figure}[h] 
%   \centering
%   (i) \includegraphics{cop-tree_thesis.1} \ \ \ \ \ \  
%   (ii)  \includegraphics{cop-tree_thesis.2}
%   \caption{(i) $8$-subdivided star with 7 rays, (ii) 3-subdivided
%     star with 3 rays}
%   \label{fig:kstar}
% \end{figure}

\section{Characterization of Feasible Tree Path  Labelings} 
\label{sec:feasible} 
In this section we give an algorithmic characterization of a
feasibility of tree path labeling.
% Consider a path labeling $\cl: \cF \rightarrow \cP$ for set system $\cF$
% and path system $\cP$ on the given tree $T$. We
% call $\cl$ an {\em Intersection Cardinality Preserving Path Labeling
%   (ICPPL)} if it has the following properties.
Consider a path labeling $(\cF, \cl)$ on the
given tree $T$. We call $(\cF, \cl)$ an {\em Intersection Cardinality
  Preserving Path Labeling (ICPPL)} if it has the following
properties.

\begin{enumerate}[{\icpplpr\ }i. ]
\item \label{pr:i} $|S| = |\cl(S)|$, \hfill for all $S \in \cF$
  % \vspace{\topshrink}
\item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$, \hfill
  for all distinct $S_1, S_2 \in \cF$
  % \vspace{\topshrink}
\item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
  \cap \cl(S_3)|$, \hfill for all distinct $S_1, S_2, S_3 \in \cF$
\end{enumerate}

\noindent
The following lemma is useful in subsequent arguments. 
\begin{lemma}
  \label{lem:setminuscard}
  If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
  (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
  \cl(S_3))|$.
\end{lemma}
\tnote[a]{THESIS: UNCOMMENT}
% \begin{proof}%[Proof of Lemma~\ref{lem:setminuscard}]
%   Let $P_i = \cl(S_i)$, for all $i \in \set{1,2,3}$. We have $|S_1
%   \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| = |S_1
%   \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to properties
%   (\ref{pr:ii}) and (\ref{pr:iii}) of ICPPL, $|S_1 \cap S_2| - |S_1
%   \cap S_2 \cap S_3| = |P_1 \cap P_2| - |P_1 \cap P_2 \cap P_3| =
%   |(P_1 \cap P_2) \setminus P_3| = |P_1 \cap (P_2 \setminus
%   P_3)|$. Thus lemma is proved. \qed
% \end{proof}

In the remaining part of this section we show that $(\cF, \cl)$ is
feasible if and only if it is an ICPPL and
Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two
levels of filtering of $(\cF, \cl)$ to make it simpler while retaining
the set of isomorphisms, if any, between $\cF$ and $\cF^\cl$.
% One direction of this claim isclear: that if a path
% labeling is feasible, then all intersection cardinalities are
% preserved, i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
% \annote[a]{has no premature exit condition hence any input will go
%   through it}{Prove that the filtered sets has ICPPL \iff input PL
%   has ICPPL?}. Algorithm~\ref{leafasgn} has an exit condition at
% line~\ref{xempty}. It can be easily verified that $X$ cannot be
% empty if $\cl$ is a feasible path labeling. The reason is that a
% feasible path labeling has an associated bijection between
% $supp(\cF)$ and $V(T)$ \remove[a]{i.e. $supp(\cF^{\cl})$} such that
% the sets map to paths, ``preserving'' the path labeling.  The rest
% of the section is devoted to constructively proving that it is
% sufficient for a path labeling to be an ICPPL and pass the two
% filtering algorithms.  To describe in brief, the constructive
% approaches refine an ICPPL iteratively, such that at the end of each
% iteration we have a ``filtered'' path labeling, and finally we have
% a path labeling that defines a family of bijections from $supp(\cF)$
% to $V(T)$\remove[a]{ i.e. $supp(\cF^{\cl})$}.

First, we present Algorithm~\ref{perms} called {\tt filter\_1}
\tnote[a]{THESIS: UNCOMMENT}
% , and
% prove its correctness.  
This algorithm refines the path labeling by
processing pairs of paths in $\cF^\cl$ that share a leaf until no two
paths in the new path labeling share any leaf.

\begin{algorithm}[h]
  \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
  \label{perms}
  \begin{algorithmic}[\lndisplay]
    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$\\
    \STATE $j \assign 1$\\
    \WHILE {there is $S_1, S_2 \in \cF_{j-1}$ such that
      $\cl_{j-1}(S_1)$ and $\cl_{j-1}(S_2)$ have a common leaf in
      $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
    \{S_1, S_2\})
    \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1 \}$ \label{setbreak} 
    \COMMENT {Remove $S_1$, $S_2$ and add the ``filtered'' sets}
    \STATE {\bf for} every $S \in \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
    S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\

    \STATE $\cl_j(S_1 \cap S_2) \assign \cl_{j-1}(S_1) \cap
    \cl_{j-1}(S_2)$
    \COMMENT {Carry forward the path labeling for all existing sets other than
      $S_1$, $S_2$}
    \STATE $\cl_j(S_1 \setminus S_2) \assign \cl_{j-1}(S_1) \setminus
    \cl_{j-1}(S_2)$ 
    \COMMENT {Define path labeling for new sets}
    \STATE $\cl_j(S_2 \setminus S_1) \assign \cl_{j-1}(S_2) \setminus
    \cl_{j-1}(S_1)$

    \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of ICPPL}
    \label{ln:3waycheck}
    \STATE {\bf exit} \label{ln:exit1} \\
    \ENDIF

    \STATE $j \assign j+1$\\
    \ENDWHILE
    \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
    \RETURN $(\cF', \cl')$
  \end{algorithmic}
\end{algorithm}

\begin{lemma} 
 \label{lem:feasible} 
 In Algorithm~\ref{perms}, if input $(\cF, \cl)$ is a feasible path
 assignment then at the end of $j$th iteration of the {\bf while}
 loop, $j \ge 0$, $(\cF_j, \cl_j)$ is a feasible path assignment.
\end{lemma}
\tnote[a]{THESIS: UNCOMMENT}

% \begin{proof}%[Proof of Lemma~\ref{lem:feasible}]
%   We will prove this by mathematical induction on the number of
%   iterations. The base case $(\cF_0, \cl_0)$ is feasible since it is
%   the input itself which is given to be feasible. Assume the lemma is
%   true till $j-1$th iteration. i.e. every hypergraph isomorphism
%   $\phi: supp\left(\cF_{j-1}\right) \rightarrow V\left(T \right)$ that
%   defines $(\cF, \cl)$'s feasibility, is such that the induced path
%   labeling on $\cF_{j-1}$, $\cl_{\phi[{\cF_{j-1}}]}$ is equal to
%   $\cl_{j-1}$. We will prove that $\phi$ is also the bijection that
%   makes $(\cF_j, \cl_j)$ feasible. Note that $supp(\cF_{j-1}) =
%   supp(\cF_{j})$ since the new sets in $\cF_j$ are created from basic
%   set operations to the sets in $\cF_{j-1}$. For the same reason and
%   $\phi$ being a bijection, it is clear that when applying the $\phi$
%   induced path labeling on $\cF_j$, $ \cl_{\phi[{\cF_{j}}]}(S_1
%   \setminus S_2) = \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
%   \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Now observe that $ \cl_j(S_1
%   \setminus S_2) = \cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2) =
%   \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
%   \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Thus the induced path labeling
%   $\cl_{\phi[{\cF_{j}}]} = \cl_{j}$. Therefore lemma is proved.  \qed
% \end{proof}

\begin{lemma}
  \label{lem:invar1} In Algorithm~\ref{perms}, at the end of $j$th
  iteration, $j \ge 0$, of the {\bf while} loop, the following
  invariants are maintained.
  \begin{enumerate}[I {\ }] %\vspace{\topshrink}
  \item $\cl_j(R)$ is a path in $T$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
    \ \ \ \ \ \ \ \ \ \ for all $R \in \cF_j$%\vspace{\topshrink}
  \item $|R| = |\cl_j(R)|$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
    \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $R \in
    \cF_j$%\vspace{\topshrink}
  \item $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, \ \ \ \ \ \ \ \ \ \
    \ \ \ \ \ \ \ \ \ \ for all $R, R' \in \cF_j$%\vspace{\topshrink}
  \item $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
    \cl_j(R'')|$, \ \ \ for all $R, R', R'' \in \cF_j$
  \end{enumerate}
\end{lemma}
\tnote[a]{THESIS: UNCOMMENT}

% \begin{proof}
%  Proof is by induction on the number of iterations, $j$. In this
%   proof, the term ``new sets'' will refer to the sets added to $\cF_j$
%   in $j$th iteration in line~\ref{setbreak} of Algorithm~\ref{perms},
%   $S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1$ and its
%   images in $\cl_j$ where $\cl_{j-1}(S_1)$
%   and $\cl_{j-1}(S_2)$ intersect and share a leaf.\\
%   The invariants are true in the base case $(\cF_0, \cl_0)$, since it
%   is the input ICPPL.  Assume the lemma is true till the $j-1$th
%   iteration. Let us consider the possible cases for each of the above invariants for
%   the $j$th iteration.

%   \begin{enumerate}[\textreferencemark]
%   \item {\em Invariant} I/II
%     \begin{enumerate}[{I/II}a $|$] % \textbullet 
%     \item {\em $R$ is not a new set.} It is in $\cF_{j-1}$. Thus
%       trivially true by induction hypothesis.
%     \item {\em $R$ is a new set.} If $R$ is in $\cF_{j}$ and not in
%       $\cF_{j-1}$, then it must be one of the new sets added in
%       $\cF_j$. In this case, it is clear that for each new set, the
%       image under $\cl_j$ is a path since by definition the chosen
%       sets $S_1$, $S_2$ are from $\cF_{j-1}$ and due to the while loop
%       condition, $\cl_{j-1}(S_1)$, $\cl_{j-1}(S_2)$ have a
%       common leaf. Thus invariant I is proved.\\
%       Moreover, due to induction hypothesis of invariant III and the
%       definition of $l_j$ in terms of $l_{j-1}$, invariant II is
%       indeed true in the $j$th iteration for any of the new sets.  If
%       $R = S_1 \cap S_2$, $|R| = |S_1 \cap S_2| = |\cl_{j-1}(S_1) \cap
%       \cl_{j-1}(S_2)| = |\cl_j(S_1 \cap S_2)| = |\cl_j(R)|$.
%       If $R = S_1 \setminus S_2$, $|R| = |S_1 \setminus S_2| = |S_1| -
%       |S_1 \cap S_2| = |\cl_{j-1}(S_1)| - |\cl_{j-1}(S_1) \cap
%       \cl_{j-1}(S_2)| = |\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)| =
%       |\cl_j(S_1 \setminus S_2)|
%       = |\cl_j(R)|$. Similarly if $R = S_2 \setminus S_1$.\\
%     \end{enumerate}
%   \item {\em Invariant} III
%     \begin{enumerate}[{III}a $|$]
%     \item {\em $R$ and $R'$ are not new sets.} It is in
%       $\cF_{j-1}$. Thus trivially true by induction hypothesis.
%     \item {\em Only one, say $R$, is a new set.} Due to invariant IV
%       induction hypothesis, Lemma~\ref{lem:setminuscard} and
%       definition of $\cl_j$, it follows that invariant III is true no
%       matter which of the new sets $R$ is equal to. If $R = S_1 \cap
%       S_2$, $|R \cap R'| = |S_1 \cap S_2 \cap R'| = |\cl_{j-1}(S_1)
%       \cap \cl_{j-1}(S_2) \cap \cl_{j-1}(R')| = |\cl_j(S_1 \cap S_2)
%       \cap \cl_j(R')| = |\cl_j(R) \cap \cl_j(R')|$.  If $R = S_1
%       \setminus S_2$, $|R \cap R'| = |(S_1 \setminus S_2) \cap R'| =
%       |(\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)) \cap \cl_{j-1}(R')| =
%       |\cl_{j}(S_1 \cap S_2) \cap \cl_{j}(R')| = |\cl_{j}(R) \cap
%       \cl_{j}(R')|$. Similarly, if $R = S_2 \setminus
%       S_1$. Note $R'$ is not a new set.\\

%     \item {\em $R$ and $R'$ are new sets.} By definition, the new
%       sets and their path images in path label $\cl_j$ are disjoint so
%       $|R \cap R'| = |\cl_j(R) \cap \cl_j(R)| = 0$. Thus case proved.
%     \end{enumerate}
%   \item {\em Invariant} IV

%     Due to the condition in line~\ref{ln:3waycheck}, this invariant is
%     ensured at the end of every iteration.
% %     \begin{enumerate} [{Case 3.}1:]
% %     \item {\em $R$, $R'$ and $R''$ are not new sets.} Trivially
% %       true by induction hypothesis.
% %     \item {\em Only one, say $R$, is a new set.}
% %       If $R = S_1 \cap S_2$,  from Lemma~\ref{lem:fourpaths} and
% %       invariant III hypothesis,  this case is proven. Similarly if $R$
% %       is any of the other new  sets, the case is proven by also using
% %       Lemma ~\ref{lem:setminuscard}.
% %     \item {\em At least two of $R, R', R''$ are new sets.}
% %       The new sets are disjoint hence this case is vacuously true.
% %     \end{enumerate}
%   \end{enumerate} \qed
% %\vspace{-6mm}

% \end{proof}

\begin{lemma}
  \label{lem:noexit1}
  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{perms} is
  feasible, then the set of hypergraph isomorphism functions that
  defines $(\cF, \cl)$'s feasibility is the same as the set that
  defines $(\cF_j, \cl_j)$'s feasibility, if any.  Secondly, for any
  iteration $j > 0$ of the {\em \bf while} loop, the {\em \bf exit}
  statement in line~\ref{ln:exit1} will not execute.
\end{lemma}
\tnote[a]{THESIS: UNCOMMENT}
% \begin{proof}
%   Since $(\cF,\cl)$ is feasible, by Lemma~\ref{lem:feasible}
%   $(\cF_j,\cl_j)$ for every iteration $j > 0$ is feasible.  % Therefore,
% %   every hypergraph isomorphism $\phi: supp(\cF) \rightarrow V(T)$ that
% %   induces $\cl$ on $\cF$ also induces $\cl_{j-1}$ and $\cl_{j}$ on
% %   $\cF_{j-1}$ and $\cF_{j}$ respectively, i.e., $\cl_{\phi[\cF_{j-1}]}
% %   = \cl_{j-1}$ and $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
% %   that for all $x \in supp(\cF)$, for all $v \in V(T)$ the following
% %   hold true.
%   Also, every hypergraph isomorphism $\phi: supp(\cF) \rightarrow
%   V(T)$ that induces $\cl$ on $\cF$ also induces $\cl_{j}$ on
%   $\cF_{j}$, i.e., $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
%   that for all $x \in supp(\cF)$, for all $v \in V(T)$, if $(x,v) \in
%   \phi$ then $v \in \cl_{j}(S)$ for all $S \in \cF_{j}$ such that $x
%   \in S$.
% % the following
% %   hold true.
% %   \begin{enumerate}[i. ]
% %   \item If $(x,v) \in \phi$ then $v \in \cl_{j-1}(S)$ for all $S \in
% %     \cF_{j-1}$ such that $x \in S$.
% %   \item If $(x,v) \in \phi$ then $v \in \cl_{j}(S)$ for all $S \in
% %     \cF_{j}$ such that $x \in S$
% %   \end{enumerate}
%   In other words, filter 1 outputs a filtered path labeling that
%   ``preserves''
%   hypergraph isomorphisms of the original path labeling.\\
%   Secondly, line~\ref{ln:exit1} will execute \iff the exit condition in
%   line~\ref{ln:3waycheck}, i.e. failure of three way intersection
%   preservation, becomes true in any iteration of the {\em \bf while}
%   loop.  Due to Lemma~\ref{lem:invar1} Invariant IV, the exit
%   condition does not occur if the input is a feasible ICPPL.\qed

% %   such that $\phi(x) = v$ where $v$ is the leaf considered in the
% %   first iterations of while. Clearly, $\phi$ is a renaming of
% %   vertices in hypergraph $\cF$ to those in hypergraph $\cF^\cl$. Thus
% %   the following facts can be observed in every iteration of the loop.

% %   \begin{enumerate}[\hspace{2mm}i. ] \vspace{\topshrink}
% %   \item all intersection cardinalities are preserved in this path
% %     labeling \vspace{\topshrink}
% %   \item element $x$ is exclusive in a hyperedge in $\cF$ since $v$ is
% %     exclusive in a hyperedge in $\cF^\cl$.
% %   \end{enumerate}

% %   Thus the exit condition is never rendered true after $x$ and $v$ are
% %   removed from their respective hyperedges. \qed

% % \noindent
% % This proof uses mathematical induction on the number
% %   of iterations $j$, $j \ge 0$, of the loop that executed
% %   without exiting. The base case, $j = 0$ is obviously true since the
% %   input is an ICPPL and the exit condition cannot hold true due to
% %   ICPPL property (iii).  Assume the algorithm executes till the end
% %   of $j-1$th iteration without exiting at line
% %  ~\ref{ln:3waycheck}. Consider the $j$th iteration. From Lemma
% %  ~\ref{lem:feasible} we know that $(\cF_j, \cl_j)$ and $(\cF_{j-1},
% %   \cl_{j-1})$ are feasible\remove[AS]{and from the proof in lemma
% %     lem:invar1 we know that $(\cF_{j-1}, \cl_{j-1})$ satisfies all the
% %     invariants defined in the lemma}.  Thus there exists a bijection
% %   $\phi: supp(\cF) \rightarrow V(T)$ such that the induced path
% %   % labeling on $\cF_{j-1}$ $\cl_{\phi[\cF_{j-1}]} = \cl_{j-1}$.
% %   labeling on $\cF_{j}$, $\cl_{\phi[\cF_{j}]}$ and on $\cF_{j-1}$,
% %   $\cl_{\phi[\cF_{j-1}]}$ are equal to $\cl_{j}$ and $\cl_{j-1}$
% %   respectively.  We need to prove that for any $R, R', R'' \in
% %   \cF_{j}$, $|R \cap R' \cap R''| = |\cl_j(R) \cap \cl_j(R') \cap
% %   \cl_j(R'')|$.
% %   The following are the possible cases that could arise. From argument
% %   above, $|\cl_j(R) \cap \cl_j(R') \cap \cl_j(R'')| =
% %   |\cl_{\phi[\cF_{j}]}(R) \cap \cl_{\phi[\cF_{j}]} (R') \cap
% %   \cl_{\phi[\cF_{j}]} (R'')|$

% %   \begin{enumerate}[a $|$]
% %   \item {\em None of the sets are new. $R, R', R'' \in \cF_{j-1}$.}
% %     We know $(\cF_{j-1}, \cl_{j-1})$ is feasible. Thus $|R \cap R'
% %     \cap R''| = |\cl_{j-1}(R) \cap \cl_{j-1}(R') \cap \cl_{j-1}(R'')|
% %     = |\cl_{j}(R) \cap \cl_{j}(R') \cap \cl_{j}(R'')|$.
% %   \item {\em Only one, say $R$, is a new set.}  Let $R = S_1 \cap S_2$
% %     ($S_1, S_2$ are defined in the proof of lemma
% %    ~\ref{lem:invar1}). Now we have $|R \cap R' \cap R''| = |S_1 \cap
% %     S_2 \cap R' \cap R''| = |\cl_{j-1}(S_1) \cap \cl_{j-1}(S_2) \cap
% %     \cl_{j-1}(R') \cap \cl_{j-1}(R'')| = |\cl_{j}(R) \cap \cl_{j}(R')
% %     \cap \cl_{j}(R'')|$. Thus proven. If $R$ is any of the other new
% %     sets, the same claim can be verified using lemma
% %    ~\ref{lem:setminuscard}.
% %     % \item []{\bf Case 3:}
% %   \item {\em At least two of $R, R', R''$ are new sets.}  The new sets
% %     are disjoint hence this case is vacuously true.
% %   \end{enumerate}
% %   \qed
% %   \tnote[E2]{remove the induction proof. just text saying x and v are
% %     exclusive in these sets therefore the intersection cardinalities
% %     don't change thus all invariants are still true} 
% \end{proof}

As a result of Algorithm~\ref{perms} each leaf $v$ in $T$
is such that there is exactly one set in $\cF$ with $v$ as a vertex in
the path assigned to it.  In Algorithm~\ref{leafasgn} we identify
elements in $supp(\cF)$ whose images are leaves in a hypergraph
isomorphism if one exists.  Let $S \in \cF$ be such that $\cl(S)$ is a
path with leaf and $v \in V(T)$ is the unique leaf incident on it.  We
define a new path labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) =
\set{v}$ where $x$ an arbitrary element from $S \setminus \bigcup_{\hS
  \ne S} \hS$. In other words, $x$ is an element present in no other
set in $\cF$ except $S$. This is intuitive since $v$ is present in no
other path image under $\cl$ other than $\cl(S)$.  The element $x$ and
leaf $v$ are then removed from the set $S$ and path $\cl(S)$
respectively. After doing this for all leaves in $T$, all path images
in the new path labeling $\cl_{new}$ except leaf labels (a path that
has only a leaf is called the {\em leaf label} for the corresponding
single element hyperedge or set) are paths from a new pruned tree $T_0
= T \setminus \{v \mid v \text{ is a leaf in }
T\}$. Algorithm~\ref{leafasgn} is now presented with details.

\begin{algorithm}[h]
  \caption{Leaf labeling from an ICPPL {\tt filter\_2($\cF, \cl, T$)}}
  \label{leafasgn}
  \begin{algorithmic}[\lndisplay]
    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$
    \COMMENT {Path images are such that no two path images share a
      leaf.}
    \STATE $j \assign 1$\\
    \WHILE {there is a leaf $v$ in $T$ and a unique $S_1 \in
      \cF_{j-1}$ such that $v \in \cl_{j-1}(S_1)$ }\label{uniqueleaf}
    \STATE $\cF_j \assign \cF_{j-1} \setminus \{S_1\}$\\
    \STATE for all $S \in \cF_{j-1}$ such that $S \ne S_1$ set
    $\cl_j(S) \assign
    \cl_{j-1}(S)$\\
    \STATE $X \assign S_1 \setminus \bigcup_{S \in \cF_{j-1}, S \ne S_1}S$\\
    \IF{$X$ is empty} \label{xempty} \STATE {\bf exit} \label{ln:exit2} \ENDIF
    \STATE $x \assign $ arbitrary element from $X$\\
    \STATE $\cF_j \assign \cF_j \cup \{\{x\}, S_1 \setminus \{x\}\} $\\
    \STATE $\cl_j(\{x\}) \assign \{v\}$\\
    \STATE $\cl_j(S_1 \setminus \{x\}) \assign \cl_{j-1}(S_1) \setminus \{v\}$\\
    \STATE $j \assign j+1$\\
    \ENDWHILE
    \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
    \RETURN $(\cF', \cl')$
  \end{algorithmic}
\end{algorithm}

Suppose the input ICPPL $(\cF, \cl)$ is feasible, yet set $X$ in
Algorithm~\ref{leafasgn} is empty in some iteration of the {\bf while}
loop. This will abort our procedure of finding the hypergraph
isomorphism. The following lemma shows that this will not happen.

\begin{lemma}
  \label{lem:xnotempty}
  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{leafasgn} is
  feasible, then for all iterations $j > 0$ of the {\em \bf while}
  loop, the {\em \bf exit} statement in line~\ref{ln:exit2} does not
  execute.
\end{lemma}
\tnote[a]{THESIS: UNCOMMENT}
% \begin{proof}
%   Assume $X$ is empty for some iteration $j > 0$. We know that $v$ is
%   an element of $\cl_{j-1}(S_1)$. Since it is uniquely present in
%   $\cl_{j-1}(S_1)$, it is clear that $v \in \cl_{j-1}(S_1) \setminus
%   \bigcup_{(S \in \cF_{j-1}) \wedge (S \ne S_1)}\cl_{j-1}(S)$.  Note
%   that for any $x \in S_1$ it is contained in at least two sets due to
%   our assumption about cardinality of $X$. Let $S_2 \in \cF_{j-1}$ be
%   another set that contains $x$. From the above argument, we know $v
%   \notin \cl_{j-1}(S_2)$. Therefore there cannot exist a hypergraph
%   isomorphism bijection that maps elements in $S_2$ to those in
%   $\cl_{j-1}(S_2)$. This contradicts our assumption that the input is
%   feasible. Thus $X$ cannot be empty if input is ICPPL and feasible.
%   \qed
% \end{proof}

\begin{lemma}
  \label{lem:invar3}
  In Algorithm~\ref{leafasgn}, for all $j > 0$, at the end of the
  $j$th iteration of the {\bf while} loop the four invariants given in
  Lemma~\ref{lem:invar1} hold.
\end{lemma}
\tnote[a]{THESIS: UNCOMMENT}
% \begin{proof}
%   By Lemma~\ref{lem:xnotempty} we know that set $X$ will not be empty
%   in any iteration of the {\em \bf while} loop if input ICPPL $(\cF,
%   \cl)$ is feasible and $\cl_j$ is always computed for all $j >
%   0$. Also note that removing a leaf from any path keeps the new path
%   connected. Thus invariant I is obviously true. In every iteration $j
%   > 0$, we remove exactly one element $x$ from one set $S$ in $\cF$
%   and exactly one vertex $v$ which is a leaf from one path
%   $\cl_{j-1}(S)$ in $T$. This is because as seen in
%   Lemma~\ref{lem:xnotempty}, $x$ is exclusive to $S$ and $v$ is
%   exclusive to $\cl_{j-1}(S)$. Due to this fact, it is clear that the
%   intersection cardinality equations do not change, i.e., invariants
%   II, III, IV remain true. On the other hand, if the input ICPPL is
%   not feasible the invariants are vacuously true. \qed
% \end{proof}

%% OLD TEXT TB DELETED?
% \textcolor{cyan}{
% \begin{lemma}
%   \label{lem:notfeasibleexit}
%   \tnote[a]{IS THIS CORRECT?}  If input ICPPL $(\cF, \cl)$ is not
%   feasible, then in one of the recursive calls to Algorithm --3--, the
%   {\em \bf exit} statement in line x in Algorithm --1 or line y in
%   Algorithm --2 will get executed.
% \end{lemma}
% }
%% END OLD TEXT TB DELETED?

We have seen two filtering algorithms above, namely,
Algorithm~\ref{perms} {\tt filter\_1} and Algorithm~\ref{leafasgn}
{\tt filter\_2} which when executed serially respectively result in a
new ICPPL on the same universe $U$ and tree $T$. We also proved that
if the input is indeed feasible, these algorithms do indeed output the
filtered ICPPL. Now we present the algorithmic characterization of a
feasible tree path labeling by way of Algorithm~\ref{al:icppl-find-isomorph}.

Algorithm~\ref{al:icppl-find-isomorph} computes a
hypergraph isomorphism $\phi$ recursively using Algorithm~\ref{perms}
and Algorithm~\ref{leafasgn} and pruning the leaves of the input
tree. In brief, it is done as follows. Algorithm~\ref{leafasgn} gives
us the leaf labels in $\cF_2$, i.e., the elements in $supp(\cF)$ that
map to leaves in $T$, where $(\cF_2, \cl_2)$ is the output of
Algorithm~\ref{leafasgn}. All leaves in $T$ are then pruned away. The
leaf labels are removed from the path labeling $\cl_2$ and the
corresponding elements are removed from the corresponding sets in
$\cF_2$. This tree pruning algorithm is recursively called on the
altered hypergraph $\cF'$, path label $\cl'$ and tree $T'$. The
recursive call returns the bijection $\phi''$ for the rest of the
elements in $supp(\cF)$ which along with the leaf labels $\phi'$ gives
us the hypergraph isomorphism $\phi$.  The following lemma formalizes
the characeterization of feasible path labeling.

\begin{algorithm}[h]
  \caption{{\tt get-hypergraph-isomorphism($\cF, \cl, T$)}}
  \label{al:icppl-find-isomorph}
  \begin{algorithmic}[\lndisplay]

    \IF{$T$ is empty}
    \RETURN $\emptyset$\\
    \ENDIF
    \STATE $L \assign \{v \mid v \text{ is a leaf in }      T\}$\\
    \STATE $(\cF_1, \cl_1) \assign$ {\tt filter\_1($\cF, \cl,
      T$)}\\
    \STATE $(\cF_2, \cl_2) \assign$ {\tt filter\_2($\cF_1,
      \cl_1, T$)}\\

    \STATE $(\cF', \cl') \assign (\cF_2, \cl_2)$\\
    \STATE $\phi' \leftarrow \emptyset$

    \FOR {every $v \in L$}
    \STATE $\phi'(x) \assign v$ where $x \in \cl_2^{-1}(\{v\})$
    \COMMENT {Copy the leaf labels to a one to one function $\phi':
      supp(\cF) \rightarrow L$
      }\\
    \STATE Remove $\{x\}$ and $\{v\}$ from $\cF'$, $\cl'$  appropriately\\
    \ENDFOR

    \STATE $T' \assign T \setminus L$

    \STATE $\phi'' \assign$ {\tt
      get-hypergraph-isomorphism($\cF', \cl', T'$)}
    \STATE $\phi \assign \phi'' \cup \phi'$ \\
    \RETURN $\phi$
  \end{algorithmic}
\end{algorithm}

\begin{lemma}
  \label{lem:hyperiso}  %{lem:perm}
  If $(\cF, \cl)$ is an ICPPL from a tree $T$ and
  Algorithm~\ref{al:icppl-find-isomorph}, {\tt
    get-hypergraph-isomorphism ($\cF, \cl, T$)} returns a non-empty
  function, then there exists a hypergraph isomorphism $\phi :
  supp(\cF) \rightarrow V(T)$ such that the $\phi$-induced tree path
  labeling is equal to $\cl$ or $\cl_\phi = \cl$.
\end{lemma}
\tnote[a]{THESIS: UNCOMMENT}
% \begin{proof}
%   It is clear that in the end of every recursive call to
%   Algorithm~\ref{al:icppl-find-isomorph}, the function $\phi'$ is one
%   to one involving all the leaves in the tree passed as input to that
%   recursive call. Moreover, by Lemma~\ref{lem:noexit1} and
%   Lemma~\ref{lem:xnotempty} it is consistent with the tree path
%   labeling $\cl$ passed. The tree pruning is done by only removing
%   leaves in each call to the function and is done till the tree
%   becomes empty. Thus the returned function $\phi: supp(\cF)
%   \rightarrow V(T)$ is a union of mutually exclusive one to one
%   functions exhausting all vertices of the tree. In other words, it is
%   a bijection from $supp(\cF)$ to $V(T)$ inducing the given path
%   labeling $\cl$ and thus a hypergraph isomorphism. \qed
% \end{proof}

\begin{theorem}
  \label{th:charac}
  A path labeling $(\cF, \cl)$ on tree $T$ is feasible \iff it is an
  ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
  T)$ as input returns a non-empty function.
\end{theorem}
\tnote[a]{THESIS: UNCOMMENT}
% \begin{proof}
%   From Lemma~\ref{lem:hyperiso}, we know that if $(\cF, \cl)$ is an
%   ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
%   T)$ as input returns a non-empty function, $(\cF, \cl)$ is feasible.
%   Now consider the case where $(\cF, \cl)$ is feasible. i.e. there
%   exists a hypergraph isomorphism $\phi$ such that $\cl_\phi =
%   \cl$. Lemma~\ref{lem:noexit1} and Lemma~\ref{lem:xnotempty} show us
%   that filter 1 and filter 2 do not exit if input is feasible. Thus
%   Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
%   function.\qed
% \end{proof}

\subsection*{ICPPL when given tree is a path}
\label{subsec:icpplicpia}
Consider a special case of ICPPL with the following properties when
the tree $T$ is a path.  Hence, all path labels are can be viewed as
intervals assigned to the sets in $\cF$.  It is shown, in
\cite{nsnrs09}, that the filtering algorithms outlined above need only
preserve pairwise intersection cardinalities, and higher level
intersection cardinalities are preserved by the Helly Property of
intervals.  Consequently, the filter algorithms do not need to ever
evaluate the additional check to {\em \bf exit}.
%\begin{enumerate}
%\item Given tree $T$ is a path. Hence, all path labels are interval labels.
%\item Only pairwise intersection cardinality
%  preservation is sufficient. i.e. property (iii) in ICPPL is not enforced.
%\item The filter algorithms do not have {\em \bf exit} statements.
%\end{enumerate}
%This is called an Intersection Cardinality Preservation Interval
%Assignment (ICPIA) \cite{nsnrs09}. 
This structure and its algorithm is
used in the next section for finding tree path labeling from a
$k$-subdivided star due to this graph's close relationship with
intervals.

\section{Case of a Special Tree -- The $k$-subdivided star}
\label{sec:ksubdivstar}
In this section we consider the problem of assigning paths from a
$k$-subdivided star $T$ to a given set system $\cF$ such that each set
$X \in \cF$ is of cardinality at most $k+2$.  Secondly, we present our
results only for the case when overlap graph $\bO(\cF)$ is connected.
%The
%overlap graph is well-known from the work of
%\cite{kklv10,nsnrs09,wlh02}.  We use the notation in
%\cite{kklv10}. Recall from Section~\ref{sec:prelims} that hyperedges
%$S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$ and $S'$
%have a non-empty intersection but neither of them is contained in the
%other. The overlap graph $\bO(\cF)$ is a graph in which the vertices
%correspond to the sets in $\cF$, and the vertices corresponding to the
%hyperedges $S$ and $S'$ are adjacent if and only if they overlap.  Note
%that the intersection graph of $\cF$, $\bI(\cF)$ is different from
%$\bO(\cF)$ and $\bO(\cF) \subseteq \bI(\cF)$.  
A connected component of $\bO(\cF)$ is called an overlap component of
$\cF$.  An interesting property of the overlap components is that any
two distinct overlap components, say $\cO_1$ and $\cO_2$, are such
that any two sets $S_1 \in \cO_1$ and $S_2 \in \cO_2$ are disjoint,
or, w.l.o.g, all the sets in $\cO_1$ are contained within one set in
$\cO_2$.  This containment relation naturally determines a
decomposition of the overlap components into rooted containment trees.
We consider the case when there is only one rooted containment tree,
and we first present our algorithm when $\bO(\cF)$ is connected.  It
is easy to see that once the path labeling to the overlap component in
the root of the containment tree is achieved, the path labeling to the
other overlap components in the rooted containment tree is essentially
finding a path labeling when the target tree is a path: each target
path is a path that is allocated to sets in the root overlap
component.  Therefore, for the rest of this section, $\bO(\cF)$ is a
connected graph. Recall that we also consider the special case when
all hyperedges are of cardinality at most $k+2$.  By definition, a
$k$-subdivided star has a central vertex which we call the {\em root},
and each root to leaf path is called a {\em ray}.  First, we observe
that by removing the root $r$ from $T$, we get a collection of $p$
vertex disjoint paths of length $k+1$, $p$ being the number of leaves
in $T$.  We denote the rays by $R_1, \ldots, R_p$, and the number of
vertices in $R_i$, $i \in [p]$ is $k+2$.  Let
$\seq{v_{i1},\ldots,v_{i(k+2)}=r}$ denote the sequence of vertices in
$R_i$, where $v_{i1}$ is the leaf. Note that $r$ is a common vertex to
all $R_i$.

\subsection{Description of the Algorithm}
In this section the given hypergraph $\cF$, the $k$-subdivided star
and the root of the star are denoted by $\cO$, $T$ and vertex $r$,
respectively.  In particular, note that the vertices of $\cO$
correspond to the sets in $\cF$, and the edges correspond to the
overlap relation.

%\noindent
For each hyperedge $X \in \cO$, we will maintain a 2-tuple of non-negative
 numbers $\seq{p_1(X), p_2(X)}$.  The numbers satisfy the property that
 $p_1(X) + p_2(X) \leq |X|$, and at the end of path labeling, for each
 $X$, $p_1(X) + p_2(X) = |X|$.  This signifies the algorithm tracking
 the lengths of subpaths of the path assigned to $X$ from at most two
 rays. We also maintain another parameter called the {\em residue} of
 $X$ denoted by $s(X)=|X| - p_1(X)$. This signifies the residue path
 length that must be assigned to $X$ which must be from another
 ray. For instance, if $X$ is labeled a path from only one ray, then
 $p_1(X) = |X|$, $p_2(X) = 0$ and $s(X) = 0$.

 The algorithm proceeds in iterations, and in the $i$-th
 iteration, $i > 1$, a single hyperedge $X$ that overlaps with a
 hyperedge that has been assigned a path is considered.  At the
 beginning of each iteration hyperedges of $\cO$ are classifed into
 the following disjoint sets.
 \begin{enumerate}
 \item [$\cL_1^i$] {\em Labeled without $r$.} Those that have been
   labeled with a path which does not contain $r$ in one of the
   previous iterations.\\  $\cL_1^i = \set{ X \mid p_1(X) = |X| \text{ and
     } p_2(X) = 0 \text{ and } s(X) = 0, X \in \cO}$
 \item [$\cL_2^i$] {\em Labeled with $r$.} Those that have been labeled
   with two subpaths of $\cl(X)$ containing $r$ from two different rays
   in two previous iterations.\\ $\cL_2^i = \set{X \mid 0 < p_1\left(X\right),
     p_2\left(X\right) < |X|=p_1(X)+p_2(X) \text{ and } s(X) = 0, X \in \cO}$
   \item [$\cT_1^i$] {\em Type 1 / partially labeled.} Those that have
     been labeled with one path containing $r$ from a single ray in one
     of the previous iterations. Here, $p_1(X)$ denotes the length of
     the subpath of $\cl(X)$ that $X$ has been so far labeled
     with. \\
     $\cT_1^i = \set{ X \mid 0 < p_1(X) < |X| \text{ and } p_2(X) = 0
       \text{ and } s(X) = |X|-p_1(X), X \in \cO}$
   \item [$\cT_2^i$] {\em Type 2 / not labeled.} Those that have not been
     labeled with a path in any previous iteration.\\
     $\cT_2^i = \set{ X \mid p_1(X) = p_2(X) = 0 \text{ and } s(X) = |X|,
       X \in \cO}$
 \end{enumerate}
% \vspace{-2mm}
% \begin{align*}
%   \cO &= \cL_1^i \cup \cL_2^i \cup \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]\\
%   \cO_i &= \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]
% \end{align*}

%\noindent
The set $\cO_i$ refers to the set of hyperedges $\cT_1^i \cup \cT_2^i$
in the $i$th iteration.  Note that $\cO_1 = \cO$.  In the $i$th
iteration, hyperedges from $\cO_i$ are assigned paths from $T$ using
the following rules. Also the end of the iteration, $\cL_1^{i+1},
\cL_2^{i+1}, \cT_1^{i+1}, \cT_2^{i+1}$ are set to $\cL_1^{i},
\cL_2^{i}, \cT_1^{i}, \cT_2^{i}$ respectively, along with some
case-specific changes mentioned in the rules below.

%\noindent
\begin{enumerate}[I.]
\item {\bf Iteration 1:} Let $S=\{X_1,\ldots,X_s\}$ denote the
  super-marginal hyperedges from $\cO_1$.  If $|S|=s \neq p$, then
  exit reporting failure.  Else, assign to each $X_j \in S$, the path
  from $R_j$ such that the path contains the leaf in $R_j$.  This path
  is refered to as $\cl(X_j)$.  Set $p_1(X_j)=|X|, p_2(X_j)=s(X_j)=0$.
  Hyperedges in $S$ are not added to $\cO_2$ but are added to
  $\cL_1^2$ and all other hyperedges are added to
  $\cO_2$.
\item {\bf Iteration $i$:} Let $X$ be a hyperedge from $\cO_i$ such
  that there exists $Y \in \cL_1^i \cup \cL_2^i$ and $X \overlap
  Y$. Further let $Z \in \cL_1^i \cup L_2^i$ such that $Z \overlap Y$.
  If $X \in \cT_2^i$, and if there are multiple $Y$ candidates then
  any $Y$ is selected.  On the other hand, if $X \in \cT_1^i$, then
  $X$ has a partial path assignment, $\cl'(X)$ from a previous
  iteration, say from ray $R_j$. Then, $Y$ is
  chosen such that $X \cap Y$ has a non-empty intersection with a ray
  different from $R_j$.  The key things that are done in assigning a
  path to $X$ are as follows. The {\em end} of path $\cl(Y)$ where
  $\cl(X)$ would overlap is found, and then based on this the
  existence of a feasible assignment is decided.  It is important to
  note that since $X \overlap Y$, $\cl(X) \overlap \cl(Y)$ in any
  feasible assignment.  Therefore, the notion of the {\em end} at
  which $\cl(X)$ and $\cl(Y)$ overlap is unambiguous, since for any
  path, there are two end points.
  \begin{enumerate}
  \item \label{iendpoint} {\em End point of $\cl(Y)$ where $\cl(X)$ overlaps
      depends on $X \cap Z$:} If $X \cap Z
    \neq \emptyset$, then $\cl(X)$ has an overlap of $|X \cap Y|$ at that
    end of $\cl(Y)$ at which $\cl(Y)$ and $\cl(Z)$ overlap.  If $X
    \cap Z = \emptyset$, then $\cl(X)$ has an overlap of $|X \cap Y|$ at
    that end of $\cl(Y)$ where $\cl(Y)$ and $\cl(Z)$ do not intersect.
  \item {\em Any path of length $s(X)$ at the appropriate end contains
      $r$:} If $X \in \cT_1^i$ then after finding the appropriate end
    as in step~\ref{iendpoint} this the unique path of length $s(X)$
    should end at $r$.  If not, we exit reporting failure.  Else,
    $\cl(X)$ is computed as union of $\cl'(X)$ and this path. If any
    three-way intersection cardinality is violated with this new
    assignment, then exit, reporting failure.  Otherwise, $X$ is added
    to $\cL_2^{i+1}$.  On the other hand, if $X \in \cT_2^i$, then
    after step~\ref{iendpoint}, $\cl(X)$ or $\cl'(X)$ is unique up to
    the root and including it. Clearly, the vertices $\cl(X)$ or
    $\cl'(X)$ contains depends on $|X|$ and $|X \cap Y|$.  If any
    three way intersection cardinality is violated due to this
    assignment, exit, reporting failure.  Otherwise,
    $p_1(X)$ is updated as the length of the assigned path, and $s(X)
    = |X|-p_1(X)$.  If $s(X) > 0$, then $X$ is added to $\cT_1^{i+1}$.
    If $s(X)=0$, then $X$ is added to $\cL_1^{i+1}$.
  \item {\em The unique path of length $s(X)$ overlapping at the
      appropriate end of $Y$ does not contain $r$:} In this case,
    $\cl(X)$ is updated to include this path.  If any three way
    intersection cardinality is violated, exit, reporting failure.
    Otherwise, update $p_1(X)$ and $p_2(X)$ are appropriate, $X$ is
    added to $\cL_1^{i+1}$ or $\cL_2^{i+1}$, as appropriate.
  \end{enumerate}
\end{enumerate}

%\noindent 
\tnote[a]{THESIS: UNCOMMENT}
% \subsection*{Proof of Correctness and Analysis of Running Time} 
% It
% is clear that the algorithm runs in polynomial time, as at each step,
% at most three-way intersection cardinalities need to be checked.
% Further, finding super-marginal hyperedges can also be done in
% polynomial time, as it involves considering the overlap regions and
% checking if the inclusion partial order contains a single minimal
% element.  In particular, once the super-marginal edges are identified,
% each iteration involes finding the next hyperedge to consider, and
% testing for a path to that hyperedge.  To identify the next hyperedge
% to consider, we consider the breadth first layering of the hyperedges
% with the zeroeth layer consisting of the super-marginal hyperedges.
% Since $\cO$ is connected, it follows that all hyperedges of $\cO$ will
% be considered by the algorithm.  Once a hyperedge is considered, the
% path to be assigned to it can also be computed in constant time.  In
% particular, in the algorithm the path to be assigned to $X$ depends on
% $\cl(Y), \cl(Z)$, $s(X)$ and the presence or absence of $r$ in the
% candidate partial path $\cl'(X)$.  Therefore, once the super-marginal
% edges are identified, the running time of the algorithm is linear in
% the size of the input.  By the technique used for constructing prime
% matrices \cite{wlh02}, the super-marginal edges can be found in linear
% time in the input size.  Therefore, the algorithm can be implemented
% to run in linear time in the input size.

% %\noindent
% The proof of correctness uses the following main properties:
% \begin{enumerate}
% \item The $k$-subdivided star has a very symmetric structure.  This
%   symmetry is quantified based on the following observation -- either
%   there are no feasible path labelings of $\cO$ using paths from $T$,
%   or there are exactly $p!$ feasible path labelings.  In other words,
%   there is either no feasible assignment, or effectively a unique
%   assignment modulo symmetry.
% \item The $p$ super-marginal hyperedges, if they exist, will each be
%   assigned a path from distinct rays, and each such path contains the
%   leaf.
% \item For a candidate hyperedge $X$, the partial path assignment
%   $\cl'(X)$ is decided by its overlap with $\cl(Y)$ and cardinality of
%   intersection with $\cl(Z)$.
% \end{enumerate}
% These properties are formalized as follows:
% \begin{lemma}
%   \label{lem:sup-mar}
%   If $X \in \cF$ is super-marginal and $\cl$ is a feasible tree path
%   labeling to tree $T$, then $\cl(X)$ will contain a leaf in $T$.
% \end{lemma}
% \begin{proof}
%   Suppose $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
%   path labeling from $T$.  Assume $\cl(X)$ does not have a leaf.  Let
%   $R_i$ be one of the rays (or the only ray) $\cl(X)$ is part of.
%   Since $X$ is in a connected overlap component, there exists $Y_1 \in
%   \cF$ and $X \nsubseteq Y_1$ such that $Y_1 \overlap X$ and $Y_1$ has
%   at least one vertex closer to the leaf in $R_i$ than any vertex in
%   $X$. Similarly with the same argument there exists $Y_2 \in \cF$
%   with same subset and overlap relation with $X$ except it has has at
%   least one vertex farther away from the leaf in $R_i$ than any vertex
%   in $X$. Clearly $Y_1 \cap X$ and $Y_2 \cap X$ cannot be part of same
%   inclusion chain which contradicts that assumption $X$ is
%   super-marginal. Thus the claim is proved.\qed
% \end{proof}
% \begin{lemma}
%   If $\cO$ does not have any super-marginal edges, then in any
%   feasible path labeling $\cl$ of $\cO$ with paths from $T$ is such
%   that, for any hyperedge $X$ for which $\cl(X)$ contains a leaf, $|X|
%   \geq k+3$.
% \end{lemma}
% \begin{proof}
%   The proof of this lemma is by contradiction.  Let $X$ be a
%   hyperedges such that $|X| \leq k+2$ and that $\cl(X)$ has a leaf.
%   This implies that the overlap regions with $X$, which are captured by
%   the overlap regions with $\cl(X)$, will form a single inclusion
%   chain. This shows that $X$ is a marginal hyperedge which
%   contradicts the assumption that $\cO$ does not have super-marginal
%   hyperedges. \qed
% \end{proof}
% This lemma is used to prove the next lemma for the case when for all
% $X \in \cO$, $|X| \leq k+2$.  The proof is left out as it just uses
% the previous lemma and the fact that the hyperedges in $X$ have at
% most $k+2$ elements.
% \begin{lemma}
%   If there is a feasible path labeling for $\cO$ in $T$, then there
%   are exactly $p$ super-marginal hyperedges.
% \end{lemma}
% These lemmas now are used to prove the following theorem.
% \begin{theorem}
%   Given $\cO$ and a $k$-subdivided star $T$, the above algorithm
%   decides correctly if there is a feasible path labeling $\cl$.
% \end{theorem}
% \begin{proof} {\em Outline.}
% %   If the algorithm outputs a path labeling $\cl$, then it is clear
% %   that it is an ICPPL. The reason is that the algorithm checks that
% %   three-way intersection cardinalities are preserved in each iteration
% %   which ensures \icpplpr~\ref{pr:iii}. Moreover, it is clear that
% %   $\cl(X)$ for any $X \in \cO$ is computed by maintaining
% %   \icpplpr~\ref{pr:i} and \icpplpr~\ref{pr:ii}. For such a labeling
% %   $\cl$, the proof that it is feasible is by induction on $k$. What
% %   needs to be shown is that Algorithm~\ref{al:icppl-find-isomorph}
% %   successfully runs on input $(\cO, \cl)$. In base case $k=0$, $T$ is
% %   a star. The claim is clear by observing that after Filter 1 and one
% %   iteration of Filter 2, all the leaves have found their pre-images
% %   from support $\cO$.  Therefore, in the induction step, it is clear
% %   that after Filter 1 and one iteration of Filter 2, the leaves are
% %   assigned pre-images.  Removing the leaves from $T$ and the
% %   pre-images from support of $\cO$, results in an ICPPL to a
% %   $(k-1)$-subdivided star.  Now we apply the induction hypothesis, and
% %   we get a isomorphism between the hypergraph $\cO$ and $\cO^\cl$.
%   If the algorithm outputs a path labeling $\cl$, then it is clear
%   that it is an ICPPL. The reason is that the algorithm checks that
%   three-way intersection cardinalities are preserved in each iteration
%   which ensures ICPPL \icpplpr~\ref{pr:iii}. Moreover, it is clear that
%   $\cl(X)$ for any $X \in \cO$ is computed by maintaining
%   ICPPL \icpplpr~\ref{pr:i} and ICPPL \icpplpr~\ref{pr:ii}. For such a labeling
%   $\cl$, the proof that it is feasible is by induction on $k$. What
%   needs to be shown is that Algorithm~\ref{al:icppl-find-isomorph}
%   successfully runs on input $(\cO, \cl)$. In base case $k=0$, $T$ is
%   a star. Also every set is at most size 2 ($k+2$) size and thus
%   overlaps are at most 1. If two paths share a leaf in {\tt filter\_1}
%   one must be of length 2 and the other of length 1. Thus the exit
%   condition is not met. Now it is trivial that the exit condition in
%   {\tt filter\_2} is also not met. Thus claim proven for base case.
%   Now assume the claim to be true when target tree is a
%   $(k-1)$-subdivided star. Consider the case of a $k$-subdivided
%   star. It is clear that after {\tt filter\_1} and one iteration of
%   {\tt filter\_2}, the leaves are assigned pre-images.  Removing the
%   leaves from $T$ and the pre-images from support of $\cO$, results in
%   an ICPPL to a $(k-1)$-subdivided star.  Now we apply the induction
%   hypothesis, and we get a isomorphism between the hypergraph $\cO$
%   and $\cO^\cl$.

% %  \noindent
%   In the reverse direction if there is a feasible path labeling $\cl$,
%   then we know that $\cl$ is unique up to isomorphism.  Therefore,
%   again by induction on $k$ it follows that the algorithm finds $\cl$.
%   \qed
% \end{proof}

%\section{Conclusions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Bibliography.
\begin{singlespace}

{\footnotesize
  \bibliographystyle{alpha} 
  \bibliography{../lib/cop-variants__thesis}
  \addcontentsline{toc}{chapter}{REFERENCES}
}
\end{singlespace}

%%%%%%

\chapter*{Proposed Contents of the Thesis}
\addcontentsline{toc}{chapter}{Proposed Contents of the Thesis}
The outline of the thesis is as follows:

\begin{enumerate}[{Chapter} 1]
\item - Introduction
  \begin{enumerate} [{Section 1.}1]
  \item Consecutive Ones Testing
  \item Matrix modification to attain COP or ``almost'' COP
  \item Graph Isomorphism
  \item Logspace complexity of graph canonization using COP
  \item Motivation, Objective and Scope of Thesis
  \item Summary of Results
  \item Organization of document
  \end{enumerate}

\item - Consecutive ones property % ( survey of cop includes mainly *
  % dom's thesis (perhaps minus chap 4?) * NSNSR09 * McConnel's paper
  % *
  % KKLV10 to name a few. sections will be as follows)\\
  \begin{enumerate}[{Section 2.}1]
  \item Characterization of COP
  \item Recognition of COP
  \item Alteration to matrices to get COP
  \item Complexity of certain COP variations
  \end{enumerate}

\item - Other problems related to COP %(see literature)

\item Research %- Extension to trees
  \begin{enumerate}[{Section 4.}1]
  \item Tree path labeling of path
    hypergraphs%: ( stuff in latin paper,)
  \item Extension of po theory from \cite{nsnrs09}
    % (the matrix theory submitted in walcom, any other work - need to
    % see notes.)
  \end{enumerate}

\item - Conclusion

\item - Bibliography

\end{enumerate}

%\vskip 4cm
% \chapter{PUBLICATIONS}
% \section{Papers in Refereed Journals}
% \begin{enumerate}
% \item Title \\
% 	{\bf Author 1}, Author 2...\\
% 	{\it Journal title.}, {\bf Volume}, Page (Year)
% \end{enumerate}

% \section{Presentations in Conferences}
% \begin{enumerate}
% \item  Presented titled  {\em Title} at the {\bf Conference on...}  place, date..
% \end{enumerate}

\end{document}

### 2011-11-15 src/thesis.tex
setting up with github
v0.1 of synopsis.

\documentclass[synopsis,MS]             % MS thesis dissertation.
              {iitmdiss}

\usepackage{times}             % Font?
\usepackage{t1enc}             % TI encoding??  
                            %%% Graphics related.                 %

\usepackage{../lib/mylatexlib}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title page
\title{\mythesistitle}

\author{\myname}

\date{\MakeUppercase {\mysubmissionmonth}}
\department{\MakeUppercase {\mydept}}

%\nocite{*}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Certificate
\certificate

\vspace*{0.5in}

\noindent 
This is to certify that the thesis titled {\bf \mythesistitle},
submitted by {\bf \myname}, to the Indian Institute of Technology,
Madras, for the award of the degree of {\bf \mydegree}, is
a bona fide record of the research work done by her under our
supervision.  The contents of this thesis, in full or in parts, have
not been submitted to any other Institute or University for the award
of any degree or diploma.

\vspace*{1.5in}

\begin{singlespacing}
  \hspace*{-0.25in}
  \parbox{2.5in}{
    \noindent {\bf Prof.~\myadvisor} \\
    \noindent Research Guide \\
    \noindent Associate Professor \\
    \noindent Dept. of \mydept\\
    \noindent IIT Madras -- 600 036} \hfill
  \parbox{1.25in}{               % {1.75in}
    \vspace*{0.75in} 
    \noindent Chennai\\          %{\em Place:}
    \noindent \mysubmissiondate} %{\em Date:}
\end{singlespacing}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Acknowledgements
\acknowledgements

\tnote[a]{work in progress.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Abstract

\begin{abstract}
  \noindent {\bf Keywords}: \hspace*{0.5em} \parbox[t]{4.4in}{consecutive
    ones property, algorithmic graph theory, hypergraph isomorphism,
    interval labeling} \vspace*{24pt}

\def \tem {}
  \noindent 
  \annote[a]{In this paper, we explore a natural generalization of
  results on binary matrices with the |\tem consecutive ones property|. 
  We consider the following constraint satisfaction problem. Given (i)
  a set system $\F \subseteq$ $(2^{U} \setminus \emptyset)$ of a
  finite set $U$ of cardinality $n$, (ii) a tree $T$ of size $n$ and
  (iii) a bijection called |\tem tree path labeling|, $\cl$ mapping the
  sets in $\cF$ to paths in $T$, does there exist at least one
  bijection $\phi:U \rightarrow V(T)$ such that for each $S \in \cF$,
  $\{\phi(x) \mid x \in S\} = \cl(S)$?  A tree path labeling of a set
  system is called |\tem feasible| if there exists such a bijection
  $\phi$.  We present an algorithmic characterization of feasible tree
  path labeling. COP is a special instance of tree path labeling
  problem when $T$ is a path.  We conclude with a polynomial time
  algorithm to find a feasible tree path labeling of a given set
  system when $T$ is a |\tem $k$-subdivided star|, set system has a
  single containment tree of overlap components and set size is
  limited to at most $k+2$. }{this is the abstract of the
  paper. expand to fit thesis.}
\end{abstract}

%\pagebreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Table of contents etc.

\begin{singlespace}
\tableofcontents
\thispagestyle{empty}

\listoftables
\addcontentsline{toc}{chapter}{LIST OF TABLES}
\listoffigures
\addcontentsline{toc}{chapter}{LIST OF FIGURES}
\end{singlespace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Abbreviations
\abbreviations

\noindent 
\begin{tabbing}
xxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
\textbf{COP}   \> Consecutive Ones Property \\
\textbf{COT}   \> Consecutive Ones Testing \\
\textbf{ICPIA}   \> Intersection Cardinality Preservation Interval Assignment \\
\textbf{ICPPL}   \> Intersection Cardinality Preserved Path Labeling \\

\end{tabbing}

\pagebreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Notation

\chapter*{\centerline{NOTATION}}
\addcontentsline{toc}{chapter}{NOTATION}

\begin{singlespace}
\begin{tabbing}
xxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
\textbf{$2^{U}$}  \> Powerset of set $U$ \\
\end{tabbing}
\end{singlespace}

\pagebreak
\clearpage

% The main text will follow from this point so set the page numbering
% to arabic from here on.
\pagenumbering{arabic}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Introduction.

\chapter{Introduction}
\label{chap:intro}
\tnote[a]{ Some things to remember - have this somewhere.

  \begin{itemize}
  \item Examples: Illustration of the problem examples. both COP and
    extension
  \item Basic preliminaries: (definitions theorems etc) needed if any
    - graph theory
  \item Organization: Outline of document
  \end{itemize}

  Abstract must be a brief about what results we have and how it fits
  in the body of research.

  \begin{enumerate}
  \item \emph{Area:} - Broad to Specialized. i.e. Combinatorial
    algorithms -> Matrix reorganization -> general data reorganization
    (interval assignment) -> path assignment
  \item \emph{Class of problems:} say, data reorganization.
  \item \emph{Nature of results:} Is a generalization. We have a
    Polynomial algorithm for a subset of the generalization.
  \end{enumerate}

Introduction is a magnified version of Abstract with some brief
citations etc. mini survey. Here talk about canonization how it
connects these problems to graph isomorphism. Introduce our problem,
explain why it is important etc. How canonization is a
posisbility. Brief about results obtained.

Someone who doesn't want to go through the whole thesis must get the
gist of the whole document from the introduction chapter. that should
be the outlook.

once this is done, the rest of the chapters will fall into place.

{\em Survey chapter} is the full fledged expansion of the survey in
introduction with details, observations, theorems etc.

  {\em The main themes} of the thesis that must not be left out:
  \begin{enumerate}
  \item connection of COP to {\bf set systems} and thus to general
    data bases. \cite{nsnrs09}
  \item {\bf graph isomorphism} by canonization \cite{kklv10}
  \item {\bf certificate for a problem} \cite{mcc04}
  \end{enumerate}

Also see purple book notes.
}

\section{to area}
\section{to problem}
\section{to your thesis}
\section{Organization of document}

\chapter{Consecutive Ones property}
\tnote[a]{survey of cop includes mainly 
\begin{enumerate}
\item \cite{d08phd} (perhaps minus chap 4?)
\item \cite{nsnrs09}
\item \cite{mcc04}
\item \cite{kklv10}
\end{enumerate}
.. to name a few.

sections will be as follows:}
\section{Characterization of COP}

\tnote[a]{\cite{at72}, \cite{mcc04}, prime submatrix \cite{wlh02}?}
\section{Recognition of COP}
\subsection{\tnote[a]{Polynomial time algorithms}}
\tnote[a]{ (pq tree, pqr, icpia)}
\subsection{\tnote[a]{Tucker's submatrices, Dom's algo to find them}}
\subsection{\tnote[a]{McConnell's certificate}}

\section{Alteration to matrices to get COP}
\tnote[a]{algorithms to make a matrix into one with COP
  (min/max-cos-r/c) \cite{d08phd} }
\section{brief analysis of hard problems with input having COP (chap
  5, 6 in dom)}
\label{sec:briefanal}
\tnote[a]{Do I need this?}
\section{Complexity of certain COP variations}
\tnote[a]{maybe combine with Section~\ref{sec:briefanal}\\
complexity - poly, logspace}

\chapter{Other problems related to COP }
\tnote[a]{(see literature)}

\chapter{Research - Tree path labeling of path hypergraphs}
\tnote[a]{Extension to trees. ( stuff in latin paper,)}

\tnote[a]{UPDATE TO FINAL VERSION. BELOW IS OLDER VERSION WITH
  INCOMPLETE SEC 4.}
\section{Introduction}
Consecutive ones property (COP) of binary matrices is a widely studied
combinatorial problem. The problem is to rearrange rows (columns) of a
binary matrix in such a way that every column (row) has its $1$s occur
consecutively. If this is possible the matrix is said to have the COP.
This problem has several practical applications in diverse fields
including scheduling \cite{hl06}, information retrieval \cite{k77} and
computational biology \cite{abh98}.  Further, it is a tool in graph
theory \cite{mcg04} for interval graph recognition, characterization
of hamiltonian graphs, and in integer linear programming
\cite{ht02,hl06}.  Recognition of COP is polynomial time solvable by
several algorithms. PQ trees \cite{bl76}, variations of PQ trees
\cite{mm09,wlh01,wlh02,mcc04}, ICPIA \cite{nsnrs09} are the main ones.

%\noindent
The problem of COP testing can be easily seen as a simple constraint
satisfaction problem involving a system of sets from a universe. Every
column of the binary matrix can be converted into a set of integers
which are the indices of the rows with $1$s in that column. When
observed in this context, if the matrix has the COP, a reordering of
its rows will result in sets that have only consecutive integers. In
other words, the sets after reordering are intervals. Indeed the
problem now becomes finding interval assignments to the given set
system \cite{nsnrs09} with a single permutation of the universe (set
of row indices) which permutes each set to its interval. The result in
\cite{nsnrs09} characterizes interval assignments to the sets which
can be obtained from a single permutation of the rows.  They show that
for each set, the cardinality of the interval assigned to it must be
same as the cardinality of the set, and the intersection cardinality
of any two sets must be same as the interesction cardinality of the
corresponding intervals.  While this is naturally a necessary
condition, \cite{nsnrs09} shows this is indeed sufficient.  Such an
interval assignment is called an Intersection Cardinality Preserving
Interval Assignment (ICPIA).  Finally, the idea of decomposing a given
0-1 matrix into prime matrices to check for COP is adopted from
\cite{wlh02} to test if an ICPIA exists for a given set system.

\noindent 
{\bf Our Work.}  A natural generalization of the interval
assignment problem is feasible tree path labeling problem of a set
system. The problem is defined as follows -- given a set system $\cF$
from a universe $U$ and a tree $T$, does there exist a bijection from
$U$ to the vertices of $T$ such that each set in the system maps to a
path in $T$.  We refer to this as the {\em tree path labeling problem}
for an input set system, target tree pair -- $(\cF,T)$. As a special
case if the tree $T$ is a path, the problem becomes the interval
assignment problem.  We focus on the question of generalizing the
notion of an ICPIA \cite{nsnrs09} to characterize feasible path
assignments.  We show that for a given set system $\cF$, a tree $T$,
and an assignment of paths from $T$ to the sets, there is a feasible
bijection between $U$ and $V(T)$ if and only if all intersection
cardinalities among any three sets (not necessarily distinct) is same
as the intersection cardinality of the paths assigned to them and the
input runs a filtering algorithm (described in this paper)
successfully.  This characterization is proved constructively and it
gives a natural data structure that stores all the relevant feasible
bijections between $U$ and $V(T)$.  Further, the filtering algorithm
is also an efficient algorithm to test if a tree path labeling to the
set system is feasible.  This generalizes the result in
\cite{nsnrs09}.

%\noindent 
\tnote[a]{the following text could be in Further Study/Conclusion}It
is an interesting fact that for a matrix with the COP, the
intersection graph of the corresponding set system is an interval
graph.  A similar connection to a subclass of chordal graphs and a
superclass of interval graphs exists for the generalization of COP.
In this case, the intersection graph of the corresponding set system
must be a {\em path graph}. Chordal graphs are of great significance,
extensively studied, and have several applications.  One of the well
known and interesting properties of a chordal graphs is its connection
with intersection graphs \cite{mcg04}. For every chordal graph, there
exists a tree and a family of subtrees of this tree such that the
intersection graph of this family is isomorphic to the chordal graph
\cite{plr70,gav78,bp93}.  These trees when in a certain format, are
called {\em clique trees} \cite{apy92} of the chordal graph. This is a
compact representation of the chordal graph. There has also been work
done on the generalization of clique trees to clique hypergraphs
\cite{km02}.  If the chordal graph can be represented as the
intersection graph of paths in a tree, then the graph is called path
graph \cite{mcg04}.  Therefore, it is clear that if there is a
bijection from $U$ to $V(T)$ such that for every set, the elements in
it map to vertices of a unique path in $T$, then the intersection
graph of the set system is indeed a path graph.  However, this is only
a necessary condition and can be checked efficiently because path
graph recognition is polynomial time
solvable\cite{gav78,aas93}. Indeed, it is possible to construct a set
system and tree, such that the intersection graph is a path graph, but
there is no bijection between $U$ and $V(T)$ such that the sets map to
paths. Path graph isomorphism is known be isomorphism-complete, see
for example \cite{kklv10}. An interesting area of research would be to
see what this result tells us about the complexity of the tree path
labeling problem (not covered in this paper).

%\noindent
In the later part of this paper, we focus on a new special case of the
tree path labeling problem. Here the set system is such that every set
is at most $k+2$ in size and for every pair of sets in it there exists
a sequence of sets between them with consecutive sets in this sequence
having a strict intersection -- i.e., non-empty intersection with
neither being contained in the other. Moreover, the given tree is a
{\em $k$-subdivided star}. We demonstrate a polynomial time algorithm
to find a feasible path labeling in this case.

% In the later part of this paper, we decompose our search for a
% bijection between $U$ and $V(T)$ into subproblems.  Each subproblem is
% on a set subsystem in which for each set, there is another set in the
% set subsystem with which the intersection is {\em strict}, i.e., there
% is a non-empty intersection, but neither is contained in the other.
% This is in the spirit of results in \cite{wlh02,nsnrs09} where to test
% for the COP in a given matrix, the COP problem is solved on an
% equivalent set of prime matrices.

\noindent {\bf Roadmap.} The necessary preliminaries with definitions
etc. are presented in
Section~\ref{sec:prelims}. Section~\ref{sec:feasible} documents the
characterization of a feasible path labeling and finally,
Section~\ref{sec:ksubdivstar} describes a polynomial time algorithm to
find the tree path labeling of a given set system from a given
$k$-subdivided tree.

\section{Preliminaries} 
\label{sec:prelims}

This section states definitions and basic facts necessary in the scope
of this document.

\def\xnoindent {}

The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
  system} of a universe $U$ with $|U| = n$.  The {\em support} of a
set system $\F$ denoted by $supp(\cF)$ is the union of all the sets in
$\F$; $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this
paper, a set system is required to ``cover'' the universe; $
supp(\cF) = U$.
%\vspace{\prelimspace}

\xnoindent The graph $T$ represents a {\em target tree} with same
number of vertices as elements in $U$; $|V(T)| = n$.  A {\em path
  system}\, $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P
\mid P \subseteq V, \text{ } T[P] \text{ is a path} \}$.
%\vspace{\prelimspace}

% \xnoindent
% A set system $\cF$ can be alternatively represented by a {\em
%   hypergraph} $\H_\cF$ whose vertex set is $supp(\cF)$ and hyperedges
% are the sets in $\cF$. This is a known representation for interval
% systems in literature \cite{bls99,kklv10}.  We extend this definition here to
% path systems.  
\xnoindent
A set system $\cF$ can be alternatively represented by a {\em
  hypergraph}\, $\cF_H$ whose vertex set is $supp(\cF)$ and hyperedges
are the sets in $\cF$. This is a known representation for interval
systems in literature \cite{bls99,kklv10}.  We extend this definition here to
path systems. Due to the equivalence of set system and hypergraph in
the scope of this paper, we drop the
subscript $_H$ in the notation and refer to both the structures by $\cF$.
%\vspace{\prelimspace}

\xnoindent Two hypergraphs $\cF'$, $\cF''$ are said to be {\em
  isomorphic} to each other, denoted by $\cF' \cong \cF''$, iff there
exists a bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such
that for all sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in
$\cF'$ iff $B$ is a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x
\in A\}$ \cite{kklv10}. This is called {\em hypergraph
  isomorphism}. \tnote[a]{also extend $\phi$ to hyperedges -- see if
  required }
%\vspace{\prelimspace}

\xnoindent The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph
$\cF$ is a graph such that its vertex set has a bijection to $\cF$ and
there exists an edge between two vertices iff their corresponding
hyperedges have a non-empty intersection \cite{mcg04}.
%\vspace{\prelimspace}

%\noindent 
If the intersection graphs of two hypergraphs are isomorphic, $\bI(\cF)
\cong \bI(\cP)$ where $\cP$ is also a path system, then the bijection
$\cl: \cF \rightarrow \cP$ due to this isomorphism is called a {\em
  path labeling} of the hypergraph $\cF$. To illustrate further, let
$\cg: V(\cF) \rightarrow V(\cP)$ be the above mentioned isomorphism
where $V(\cF)$ and $V(\cP)$ are the vertex sets that represent the
hyperedges for each hypergraph respectively, $V(\cF) = \{ v_S \mid S
\in \cF\}$ and $V(\cP) = \{ v_P \mid P \in \cP\}$. Then the path
labeling $\cl$ is defined as follows: $\cl(S_1) = P_1$ iff $\cg
(v_{S_1}) = v_{P_1}$. The path system $\cP$ may be alternatively
denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In most scenarios in
this paper, what is given are the pair $(\cF, \cl)$ and the target
tree $T$; hence this notation will be used more often.

\xnoindent If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$
is called a {\em path hypergraph} and $\cP$ is called {\em path
  representation} of $\cF$. If this isomorphism is $\phi: supp(\cF)
\rightarrow V(T)$, then it is clear that there is an {\em induced path
labeling} $\cl_\phi: \cF \rightarrow \cP$ to the set system;
$\cl_\phi(S) = \set{y \mid y = \phi(x), x \in S}$ for all $S \in \cF$. Recall that
$supp(\cP) = V(T)$.
%\vspace{\prelimspace}

% A graph $G$ is a {\em path graph} if it is isomorphic to the intersection graph
% $\bI(\cP)$ of a path system $\cP$. 
% This
% isomorphism gives a bijection $\cl': V(G) \rightarrow \cP$. Moreover,
% for the purposes of this paper, we require that in a path labeling,
% $supp(\cP) = V(T)$.  If graph $G$ is also isomorphic to $\bI(\cF)$ for some hypergraph
% $\cF$, then clearly there is a bijection $\cl: \cF \rightarrow \cP$
% such that $\cl(S) = \cl'(v_S)$ where $v_S$ is the vertex corresponding
% to set $S$ in $\bI(\cF)$ for any $S \in \cF$. This bijection $\cl$ is
% called the {\em path labeling} of the hypergraph $\cF$ and the path
% system $\cP$ may be alternatively denoted as $\cF^\cl$.
%\vspace{\prelimspace}

\xnoindent
A path labeling $(\cF, \cl)$ is defined to be {\em
  feasible} if
$\cF \cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
\rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
\rightarrow \cF^\cl$ such that $\cl_\phi = \cl$. 
%\vspace{\prelimspace}

\xnoindent An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is
a graph such that its vertex set has a bijection to $\cF$ and there
exists an edge between two of its vertices iff their corresponding
hyperedges overlap. Two hyperedges $S$ and $S'$ are said to {\em
  overlap}, denoted by $S \overlap S'$, if they have a non-empty
intersection and neither is contained in the other; $S \overlap S'
\text{ iff } S \cap S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq
S$. Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$ and not
necessarily connected. Each connected component of $\bO(\cF)$ is
called an {\em overlap component}.
% If there are $d$ overlap components
% in $\bO(\cF)$, the set subsystems are denoted by $\cO_1, \cO_2, \ldots
% \cO_d$. Clearly $\cO_i \subseteq \F, i \in [d]$. For any $i, j \in [d]$,
% it can be verified that one of the following is true.
% \begin{enumerate}[a) ]
% \item $supp(\cO_i)$ and $supp(\cO_j)$ are disjoint
% \item $supp(\cO_i)$ is a subset of a set in $\cO_j$
% \item $supp(\cO_j)$ is a subset of a set in $\cO_i$
% \end{enumerate}
%\vspace{\prelimspace}

\xnoindent
A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
\overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
\cite{kklv10}. Additionally, if $S$ is such that it is contained in no
other hyperedge in $\cF$, i.e., it is inclusion maximal then it is called
{\em super-marginal}.

\xnoindent A {\em star} graph is a complete bipartite graph $K_{1,p}$
which is clearly a tree and $p$ is the number of leaves. The vertex
with maximum degree is called the {\em center} of the star and the
edges are called {\em rays} of the star. A {\em $k$-subdivided star}
is a star with all its rays subdivided exactly $k$ times. The
definition of a {\em ray of a $k$-subdivided star} is extended to the path
from the center to a leaf. It is clear that all rays are of length $k+2$.

\section{Characterization of Feasible Tree Path  Labeling} 
\label{sec:feasible}

In this section we give an algorithmic characterization of a
feasibility of tree path labeling.
% Consider a path labeling $\cl: \cF \rightarrow \cP$ for set system $\cF$
% and path system $\cP$ on the given tree $T$. We
% call $\cl$ an {\em Intersection Cardinality Preserving Path Labeling
%   (ICPPL)} if it has the following properties.
Consider a path labeling $(\cF, \cl)$ on the
given tree $T$. We call $(\cF, \cl)$ an {\em Intersection Cardinality
  Preserving Path Labeling (ICPPL)} if it has the following
properties.

%\vspace{\topshrink}

\begin{enumerate}[{(\icpplpr\ }i) \ \ \ ]
\item \label{pr:i} $|S| = |\cl(S)|$ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $S \in \cF$
%  \vspace{\topshrink}
\item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ \ \ \ \ \ \ \ \ \ \
  \ \ \ \ \ \ \ \ \ \ for all distinct
  $S_1, S_2 \in \cF$
%  \vspace{\topshrink}
\item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2) \cap
  \cl(S_3)|$ \ \ \ for all distinct $S_1, S_2, S_3 \in
  \cF$
\end{enumerate}

\noindent
The following lemma is useful in subsequent arguments. 
\begin{lemma}
  \label{lem:setminuscard}
  If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
  (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
  \cl(S_3))|$.
\end{lemma}
\begin{proof}%[Proof of Lemma~\ref{lem:setminuscard}]
  Let $P_i = \cl(S_i)$, for all $1 \le i \le  3$.
  $|S_1 \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| =
  |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to properties (ii)
  and (iii) of ICPPL, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
  \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
  = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. \qed
\end{proof}

\def \xnoindent  {}

\xnoindent In the remaining part of this section we show that $(\cF,
\cl)$ is feasible if and only if it is an ICPPL and
Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two levels of
filtering of $(\cF, \cl)$ to make it simpler while retaining the set
of isomorphisms, if any, between $\cF$ and $\cF^\cl$.
% One direction of this claim isclear: that if a path
% labeling is feasible, then all intersection cardinalities are
% preserved, i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
% \annote[a]{has no premature exit condition hence any input will go
%   through it}{Prove that the filtered sets has ICPPL iff input PL
%   has ICPPL?}. Algorithm~\ref{leafasgn} has an exit condition at
% line~\ref{xempty}. It can be easily verified that $X$ cannot be
% empty if $\cl$ is a feasible path labeling. The reason is that a
% feasible path labeling has an associated bijection between
% $supp(\cF)$ and $V(T)$ \remove[a]{i.e. $supp(\cF^{\cl})$} such that
% the sets map to paths, ``preserving'' the path labeling.  The rest
% of the section is devoted to constructively proving that it is
% sufficient for a path labeling to be an ICPPL and pass the two
% filtering algorithms.  To describe in brief, the constructive
% approaches refine an ICPPL iteratively, such that at the end of each
% iteration we have a ``filtered'' path labeling, and finally we have
% a path labeling that defines a family of bijections from $supp(\cF)$
% to $V(T)$\remove[a]{ i.e. $supp(\cF^{\cl})$}.
First, we present Algorithm~\ref{perms} or {\tt filter\_1}, and prove
its correctness.  This algorithm refines the path labeling by
processing pairs of paths in $\cF^\cl$ that share a leaf until no two
paths in the new path labeling share any leaf.

\begin{algorithm}[h]
  \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
  \label{perms}
  \begin{algorithmic}[\lndisplay]
    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$\\
    \STATE $j \assign 1$\\
    \WHILE {there is $S_1, S_2 \in \cF_{j-1}$ such that
      $\cl_{j-1}(S_1)$ and $\cl_{j-1}(S_2)$ have a common leaf in
      $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
    \{S_1, S_2\})
    \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1 \}$ \label{setbreak} 
    \COMMENT {Remove $S_1$, $S_2$ and add the ``filtered'' sets}
    \STATE {\bf for} every $S \in \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
    S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\

    \STATE $\cl_j(S_1 \cap S_2) \assign \cl_{j-1}(S_1) \cap
    \cl_{j-1}(S_2)$
    \COMMENT {Carry forward the path labeling for all existing sets other than
      $S_1$, $S_2$}
    \STATE $\cl_j(S_1 \setminus S_2) \assign \cl_{j-1}(S_1) \setminus
    \cl_{j-1}(S_2)$ 
    \COMMENT {Define path labeling for new sets}
    \STATE $\cl_j(S_2 \setminus S_1) \assign \cl_{j-1}(S_2) \setminus
    \cl_{j-1}(S_1)$

    \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of ICPPL}
    \label{ln:3waycheck}
    \STATE {\bf exit} \label{ln:exit1} \\
    \ENDIF

    \STATE $j \assign j+1$\\
    \ENDWHILE
    \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
    \RETURN $(\cF', \cl')$
  \end{algorithmic}
\end{algorithm}

\begin{lemma} 
 \label{lem:feasible} 
 In Algorithm~\ref{perms}, if input $(\cF, \cl)$ is a feasible path
 assignment then at the end of $j$th iteration of the {\bf while}
 loop, $j \ge 0$, $(\cF_j, \cl_j)$ is a feasible path assignment.
\end{lemma}
\begin{proof}%[Proof of Lemma~\ref{lem:feasible}]
  We will prove this by mathematical induction on the number of
  iterations. The base case $(\cF_0, \cl_0)$ is feasible since it is
  the input itself which is given to be feasible. Assume the lemma is
  true till $j-1$th iteration. i.e. every hypergraph isomorphism
  $\phi: supp\left(\cF_{j-1}\right) \rightarrow V\left(T \right)$ that
  defines $(\cF, \cl)$'s feasibility, is such that the induced path
  labeling on $\cF_{j-1}$, $\cl_{\phi[{\cF_{j-1}}]}$ is equal to
  $\cl_{j-1}$. We will prove that $\phi$ is also the bijection that
  makes $(\cF_j, \cl_j)$ feasible. Note that $supp(\cF_{j-1}) =
  supp(\cF_{j})$ since the new sets in $\cF_j$ are created from basic
  set operations to the sets in $\cF_{j-1}$. For the same reason and
  $\phi$ being a bijection, it is clear that when applying the $\phi$
  induced path labeling on $\cF_j$, $ \cl_{\phi[{\cF_{j}}]}(S_1
  \setminus S_2) = \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
  \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Now observe that $ \cl_j(S_1
  \setminus S_2) = \cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2) =
  \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
  \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Thus the induced path labeling
  $\cl_{\phi[{\cF_{j}}]} = \cl_{j}$. Therefore lemma is proven.  \qed
\end{proof}

\begin{lemma}
  \label{lem:invar1} In Algorithm~\ref{perms}, at the end of $j$th
  iteration, $j \ge 0$, of the {\bf while} loop, the following
  invariants are maintained.
  \begin{enumerate}[I {\ }] %\vspace{\topshrink}
  \item $\cl_j(R)$ is a path in $T$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
    \ \ \ \ \ \ \ \ \ \ for all $R \in \cF_j$%\vspace{\topshrink}
  \item $|R| = |\cl_j(R)|$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
    \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $R \in
    \cF_j$%\vspace{\topshrink}
  \item $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, \ \ \ \ \ \ \ \ \ \
    \ \ \ \ \ \ \ \ \ \ for all $R, R' \in \cF_j$%\vspace{\topshrink}
  \item $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
    \cl_j(R'')|$, \ \ \ for all $R, R', R'' \in \cF_j$
  \end{enumerate}
\end{lemma}

\begin{proof}
 Proof is by induction on the number of iterations, $j$. In this
  proof, the term ``new sets'' will refer to the sets added to $\cF_j$
  in $j$th iteration in line~\ref{setbreak} of Algorithm~\ref{perms},
  $S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1$ and its
  images in $\cl_j$ where $\cl_{j-1}(S_1)$
  and $\cl_{j-1}(S_2)$ intersect and share a leaf.\\
  The invariants are true in the base case $(\cF_0, \cl_0)$, since it
  is the input ICPPL.  Assume the lemma is true till the $j-1$th
  iteration. Let us consider the possible cases for each of the above invariants for
  the $j$th iteration.

  \xnoindent
 \begin{enumerate}[\xbullet]
  \item {\em Invariant} I/II
    \begin{enumerate}[{I/II}a $|$] % \textbullet 
    \item {\em $R$ is not a new set.} It is in $\cF_{j-1}$. Thus
      trivially true by induction hypothesis.
    \item {\em $R$ is a new set.} If $R$ is in $\cF_{j}$ and not in
      $\cF_{j-1}$, then it must be one of the new sets added in
      $\cF_j$. In this case, it is clear that for each new set, the
      image under $\cl_j$ is a path since by definition the chosen
      sets $S_1$, $S_2$ are from $\cF_{j-1}$ and due to the while loop
      condition, $\cl_{j-1}(S_1)$, $\cl_{j-1}(S_2)$ have a
      common leaf. Thus invariant I is proven.\\
      Moreover, due to induction hypothesis of invariant III and the
      definition of $l_j$ in terms of $l_{j-1}$, invariant II is
      indeed true in the $j$th iteration for any of the new sets.  If
      $R = S_1 \cap S_2$, $|R| = |S_1 \cap S_2| = |\cl_{j-1}(S_1) \cap
      \cl_{j-1}(S_2)| = |\cl_j(S_1 \cap S_2)| = |\cl_j(R)|$.
      If $R = S_1 \setminus S_2$, $|R| = |S_1 \setminus S_2| = |S_1| -
      |S_1 \cap S_2| = |\cl_{j-1}(S_1)| - |\cl_{j-1}(S_1) \cap
      \cl_{j-1}(S_2)| = |\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)| =
      |\cl_j(S_1 \setminus S_2)|
      = |\cl_j(R)|$. Similarly if $R = S_2 \setminus S_1$.\\
    \end{enumerate}
  \item {\em Invariant} III
    \begin{enumerate}[{III}a $|$]
    \item {\em $R$ and $R'$ are not new sets.} It is in
      $\cF_{j-1}$. Thus trivially true by induction hypothesis.
    \item {\em Only one, say $R$, is a new set.} Due to invariant IV
      induction hypothesis, Lemma~\ref{lem:setminuscard} and
      definition of $\cl_j$, it follows that invariant III is true no
      matter which of the new sets $R$ is equal to. If $R = S_1 \cap
      S_2$, $|R \cap R'| = |S_1 \cap S_2 \cap R'| = |\cl_{j-1}(S_1)
      \cap \cl_{j-1}(S_2) \cap \cl_{j-1}(R')| = |\cl_j(S_1 \cap S_2)
      \cap \cl_j(R')| = |\cl_j(R) \cap \cl_j(R')|$.  If $R = S_1
      \setminus S_2$, $|R \cap R'| = |(S_1 \setminus S_2) \cap R'| =
      |(\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)) \cap \cl_{j-1}(R')| =
      |\cl_{j}(S_1 \cap S_2) \cap \cl_{j}(R')| = |\cl_{j}(R) \cap
      \cl_{j}(R')|$. Similarly, if $R = S_2 \setminus
      S_1$. Note $R'$ is not a new set.\\

    \item {\em $R$ and $R'$ are new sets.} By definition, the new
      sets and their path images in path label $\cl_j$ are disjoint so
      $|R \cap R'| = |\cl_j(R) \cap \cl_j(R)| = 0$. Thus case proven.
    \end{enumerate}
  \item {\em Invariant} IV

    Due to the condition in line~\ref{ln:3waycheck}, this invariant is
    ensured at the end of every iteration.
%     \begin{enumerate} [{Case 3.}1:]
%     \item {\em $R$, $R'$ and $R''$ are not new sets.} Trivially
%       true by induction hypothesis.
%     \item {\em Only one, say $R$, is a new set.}
%       If $R = S_1 \cap S_2$,  from Lemma~\ref{lem:fourpaths} and
%       invariant III hypothesis,  this case is proven. Similarly if $R$
%       is any of the other new  sets, the case is proven by also using
%       Lemma ~\ref{lem:setminuscard}.
%     \item {\em At least two of $R, R', R''$ are new sets.}
%       The new sets are disjoint hence this case is vacuously true.
%     \end{enumerate}
  \end{enumerate} \qed
%\vspace{-6mm}

\end{proof}

\begin{lemma}
  \label{lem:noexit1}
  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{perms} is
  feasible, then the set of hypergraph isomorphism functions that
  defines $(\cF, \cl)$'s feasibility is the same as the set that
  defines $(\cF_j, \cl_j)$'s feasibility, if any.  Secondly, for any
  iteration $j > 0$ of the {\em \bf while} loop, the {\em \bf exit}
  statement in line~\ref{ln:exit1} will not execute.
\end{lemma}
\begin{proof}
  Since $(\cF,\cl)$ is feasible, by Lemma~\ref{lem:feasible}
  $(\cF_j,\cl_j)$ for every iteration $j > 0$ is feasible.  % Therefore,
%   every hypergraph isomorphism $\phi: supp(\cF) \rightarrow V(T)$ that
%   induces $\cl$ on $\cF$ also induces $\cl_{j-1}$ and $\cl_{j}$ on
%   $\cF_{j-1}$ and $\cF_{j}$ respectively, i.e., $\cl_{\phi[\cF_{j-1}]}
%   = \cl_{j-1}$ and $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
%   that for all $x \in supp(\cF)$, for all $v \in V(T)$ the following
%   hold true.
  Also, every hypergraph isomorphism $\phi: supp(\cF) \rightarrow
  V(T)$ that induces $\cl$ on $\cF$ also induces $\cl_{j}$ on
  $\cF_{j}$, i.e., $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
  that for all $x \in supp(\cF)$, for all $v \in V(T)$, if $(x,v) \in
  \phi$ then $v \in \cl_{j}(S)$ for all $S \in \cF_{j}$ such that $x
  \in S$.
% the following
%   hold true.
%   \begin{enumerate}[i. ]
%   \item If $(x,v) \in \phi$ then $v \in \cl_{j-1}(S)$ for all $S \in
%     \cF_{j-1}$ such that $x \in S$.
%   \item If $(x,v) \in \phi$ then $v \in \cl_{j}(S)$ for all $S \in
%     \cF_{j}$ such that $x \in S$
%   \end{enumerate}
  In other words, filter 1 outputs a filtered path labeling that
  ``preserves''
  hypergraph isomorphisms of the original path labeling.\\
  Secondly, line~\ref{ln:exit1} will execute iff the exit condition in
  line~\ref{ln:3waycheck}, i.e. failure of three way intersection
  preservation, becomes true in any iteration of the {\em \bf while}
  loop.  Due to Lemma~\ref{lem:invar1} Invariant IV, the exit
  condition does not occur if the input is a feasible ICPPL.\qed

%   such that $\phi(x) = v$ where $v$ is the leaf considered in the
%   first iterations of while. Clearly, $\phi$ is a renaming of
%   vertices in hypergraph $\cF$ to those in hypergraph $\cF^\cl$. Thus
%   the following facts can be observed in every iteration of the loop.

%   \begin{enumerate}[\hspace{2mm}i. ] \vspace{\topshrink}
%   \item all intersection cardinalities are preserved in this path
%     labeling \vspace{\topshrink}
%   \item element $x$ is exclusive in a hyperedge in $\cF$ since $v$ is
%     exclusive in a hyperedge in $\cF^\cl$.
%   \end{enumerate}

%   Thus the exit condition is never rendered true after $x$ and $v$ are
%   removed from their respective hyperedges. \qed

% \noindent
% This proof uses mathematical induction on the number
%   of iterations $j$, $j \ge 0$, of the loop that executed
%   without exiting. The base case, $j = 0$ is obviously true since the
%   input is an ICPPL and the exit condition cannot hold true due to
%   ICPPL property (iii).  Assume the algorithm executes till the end
%   of $j-1$th iteration without exiting at line
%  ~\ref{ln:3waycheck}. Consider the $j$th iteration. From Lemma
%  ~\ref{lem:feasible} we know that $(\cF_j, \cl_j)$ and $(\cF_{j-1},
%   \cl_{j-1})$ are feasible\remove[AS]{and from the proof in lemma
%     lem:invar1 we know that $(\cF_{j-1}, \cl_{j-1})$ satisfies all the
%     invariants defined in the lemma}.  Thus there exists a bijection
%   $\phi: supp(\cF) \rightarrow V(T)$ such that the induced path
%   % labeling on $\cF_{j-1}$ $\cl_{\phi[\cF_{j-1}]} = \cl_{j-1}$.
%   labeling on $\cF_{j}$, $\cl_{\phi[\cF_{j}]}$ and on $\cF_{j-1}$,
%   $\cl_{\phi[\cF_{j-1}]}$ are equal to $\cl_{j}$ and $\cl_{j-1}$
%   respectively.  We need to prove that for any $R, R', R'' \in
%   \cF_{j}$, $|R \cap R' \cap R''| = |\cl_j(R) \cap \cl_j(R') \cap
%   \cl_j(R'')|$.
%   The following are the possible cases that could arise. From argument
%   above, $|\cl_j(R) \cap \cl_j(R') \cap \cl_j(R'')| =
%   |\cl_{\phi[\cF_{j}]}(R) \cap \cl_{\phi[\cF_{j}]} (R') \cap
%   \cl_{\phi[\cF_{j}]} (R'')|$

%   \begin{enumerate}[a $|$]
%   \item {\em None of the sets are new. $R, R', R'' \in \cF_{j-1}$.}
%     We know $(\cF_{j-1}, \cl_{j-1})$ is feasible. Thus $|R \cap R'
%     \cap R''| = |\cl_{j-1}(R) \cap \cl_{j-1}(R') \cap \cl_{j-1}(R'')|
%     = |\cl_{j}(R) \cap \cl_{j}(R') \cap \cl_{j}(R'')|$.
%   \item {\em Only one, say $R$, is a new set.}  Let $R = S_1 \cap S_2$
%     ($S_1, S_2$ are defined in the proof of lemma
%    ~\ref{lem:invar1}). Now we have $|R \cap R' \cap R''| = |S_1 \cap
%     S_2 \cap R' \cap R''| = |\cl_{j-1}(S_1) \cap \cl_{j-1}(S_2) \cap
%     \cl_{j-1}(R') \cap \cl_{j-1}(R'')| = |\cl_{j}(R) \cap \cl_{j}(R')
%     \cap \cl_{j}(R'')|$. Thus proven. If $R$ is any of the other new
%     sets, the same claim can be verified using lemma
%    ~\ref{lem:setminuscard}.
%     % \item []{\bf Case 3:}
%   \item {\em At least two of $R, R', R''$ are new sets.}  The new sets
%     are disjoint hence this case is vacuously true.
%   \end{enumerate}
%   \qed
%   \tnote[E2]{remove the induction proof. just text saying x and v are
%     exclusive in these sets therefore the intersection cardinalities
%     don't change thus all invariants are still true} 
\end{proof}

\xnoindent As a result of Algorithm~\ref{perms} each leaf $v$ in $T$
is such that there is exactly one set in $\cF$ with $v$ as a vertex in
the path assigned to it.  In Algorithm~\ref{leafasgn} we identify
elements in $supp(\cF)$ whose images are leaves in a hypergraph
isomorphism if one exists.  Let $S \in \cF$ be such that $\cl(S)$ is a
path with leaf and $v \in V(T)$ is the unique leaf incident on it.  We
define a new path labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) =
\set{v}$ where $x$ an arbitrary element from $S \setminus \bigcup_{\hS
  \ne S} \hS$. In other words, $x$ is an element present in no other
set in $\cF$ except $S$. This is intuitive since $v$ is present in no
other path image under $\cl$ other than $\cl(S)$.  The element $x$ and
leaf $v$ are then removed from the set $S$ and path $\cl(S)$
respectively. After doing this for all leaves in $T$, all path images
in the new path labeling $\cl_{new}$ except leaf labels (a path that
has only a leaf is called the {\em leaf label} for the corresponding
single element hyperedge or set) are paths from a new pruned tree $T_0
= T \setminus \{v \mid v \text{ is a leaf in }
T\}$. Algorithm~\ref{leafasgn} is now presented with details.

\begin{algorithm}[h]
  \caption{Leaf labeling from an ICPPL {\tt filter\_2($\cF, \cl, T$)}}
  \label{leafasgn}
  \begin{algorithmic}[\lndisplay]
    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$
    \COMMENT {Path images are such that no two path images share a
      leaf.}
    \STATE $j \assign 1$\\
    \WHILE {there is a leaf $v$ in $T$ and a unique $S_1 \in
      \cF_{j-1}$ such that $v \in \cl_{j-1}(S_1)$ }\label{uniqueleaf}
    \STATE $\cF_j \assign \cF_{j-1} \setminus \{S_1\}$\\
    \STATE for all $S \in \cF_{j-1}$ such that $S \ne S_1$ set
    $\cl_j(S) \assign
    \cl_{j-1}(S)$\\
    \STATE $X \assign S_1 \setminus \bigcup_{S \in \cF_{j-1}, S \ne S_1}S$\\
    \IF{$X$ is empty} \label{xempty} \STATE {\bf exit} \label{ln:exit2} \ENDIF
    \STATE $x \assign $ arbitrary element from $X$\\
    \STATE $\cF_j \assign \cF_j \cup \{\{x\}, S_1 \setminus \{x\}\} $\\
    \STATE $\cl_j(\{x\}) \assign \{v\}$\\
    \STATE $\cl_j(S_1 \setminus \{x\}) \assign \cl_{j-1}(S_1) \setminus \{v\}$\\
    \STATE $j \assign j+1$\\
    \ENDWHILE
    \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
    \RETURN $(\cF', \cl')$
  \end{algorithmic}
\end{algorithm}

\xnoindent
Suppose the input ICPPL $(\cF, \cl)$ is feasible, yet set $X$ in
Algorithm~\ref{leafasgn} is empty in some iteration of the {\bf while}
loop. This will abort our procedure of finding the hypergraph
isomorphism. The following lemma shows that this will not happen.

\begin{lemma}
  \label{lem:xnotempty}
  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{leafasgn} is
  feasible, then for all iterations $j > 0$ of the {\em \bf while}
  loop, the {\em \bf exit} statement in line~\ref{ln:exit2} does not
  execute.
\end{lemma}
\begin{proof}
  Assume $X$ is empty for some iteration $j > 0$. We know that $v$ is
  an element of $\cl_{j-1}(S_1)$. Since it is uniquely present in
  $\cl_{j-1}(S_1)$, it is clear that $v \in \cl_{j-1}(S_1) \setminus
  \bigcup_{(S \in \cF_{j-1}) \wedge (S \ne S_1)}\cl_{j-1}(S)$.  Note
  that for any $x \in S_1$ it is contained in at least two sets due to
  our assumption about cardinality of $X$. Let $S_2 \in \cF_{j-1}$ be
  another set that contains $x$. From the above argument, we know $v
  \notin \cl_{j-1}(S_2)$. Therefore there cannot exist a hypergraph
  isomorphism bijection that maps elements in $S_2$ to those in
  $\cl_{j-1}(S_2)$. This contradicts our assumption that the input is
  feasible. Thus $X$ cannot be empty if input is ICPPL and feasible.
  \qed
\end{proof}

\begin{lemma}
  \label{lem:invar3}
  In Algorithm~\ref{leafasgn}, for all $j > 0$, at the end of the
  $j$th iteration of the {\bf while} loop the four invariants given in
  Lemma~\ref{lem:invar1} hold.
\end{lemma}
\begin{proof}
  By Lemma~\ref{lem:xnotempty} we know that set $X$ will not be empty
  in any iteration of the {\em \bf while} loop if input ICPPL $(\cF,
  \cl)$ is feasible and $\cl_j$ is always computed for all $j >
  0$. Also note that removing a leaf from any path keeps the new path
  connected. Thus invariant I is obviously true. In every iteration $j
  > 0$, we remove exactly one element $x$ from one set $S$ in $\cF$
  and exactly one vertex $v$ which is a leaf from one path
  $\cl_{j-1}(S)$ in $T$. This is because as seen in
  Lemma~\ref{lem:xnotempty}, $x$ is exclusive to $S$ and $v$ is
  exclusive to $\cl_{j-1}(S)$. Due to this fact, it is clear that the
  intersection cardinality equations do not change, i.e., invariants
  II, III, IV remain true. On the other hand, if the input ICPPL is
  not feasible the invariants are vacuously true. \qed
\end{proof}

% \textcolor{cyan}{
% \begin{lemma}
%   \label{lem:notfeasibleexit}
%   \tnote[a]{IS THIS CORRECT?}  If input ICPPL $(\cF, \cl)$ is not
%   feasible, then in one of the recursive calls to Algorithm --3--, the
%   {\em \bf exit} statement in line x in Algorithm --1 or line y in
%   Algorithm --2 will get executed.
% \end{lemma}
% }

\xnoindent We have seen two filtering algorithms above, namely,
Algorithm~\ref{perms} {\tt filter\_1} and Algorithm~\ref{leafasgn}
{\tt filter\_2} which when executed serially repectively result in a
new ICPPL on the same universe $U$ and tree $T$. We also proved that
if the input is indeed feasible, these algorithms do indeed output the
filtered ICPPL. Now we present the algorithmic characterization of a
feasible tree path labeling by way of Algorithm~\ref{al:icppl-find-isomorph}.

\xnoindent Algorithm~\ref{al:icppl-find-isomorph} computes a
hypergraph isomorphism $\phi$ recursively using Algorithm~\ref{perms}
and Algorithm~\ref{leafasgn} and pruning the leaves of the input
tree. In brief, it is done as follows. Algorithm~\ref{leafasgn} gives
us the leaf labels in $\cF_2$, i.e., the elements in $supp(\cF)$ that
map to leaves in $T$, where $(\cF_2, \cl_2)$ is the output of
Algorithm~\ref{leafasgn}. All leaves in $T$ are then pruned away. The
leaf labels are removed from the path labeling $\cl_2$ and the
corresponding elements are removed from the corresponding sets in
$\cF_2$. This tree pruning algorithm is recursively called on the
altered hypergraph $\cF'$, path label $\cl'$ and tree $T'$. The
recursive call returns the bijection $\phi''$ for the rest of the
elements in $supp(\cF)$ which along with the leaf labels $\phi'$ gives
us the hypergraph isomorphism $\phi$.  The following lemma formalizes
the characeterization of feasible path labeling.

\begin{algorithm}[h]
  \caption{{\tt get-hypergraph-isomorphism($\cF, \cl, T$)}}
  \label{al:icppl-find-isomorph}
  \begin{algorithmic}[\lndisplay]

    \IF{$T$ is empty}
    \RETURN $\emptyset$\\
    \ENDIF
    \STATE $L \assign \{v \mid v \text{ is a leaf in }      T\}$\\
    \STATE $(\cF_1, \cl_1) \assign$ {\tt filter\_1($\cF, \cl,
      T$)}\\
    \STATE $(\cF_2, \cl_2) \assign$ {\tt filter\_2($\cF_1,
      \cl_1, T$)}\\

    \STATE $(\cF', \cl') \assign (\cF_2, \cl_2)$\\
    \STATE $\phi' \leftarrow \emptyset$

    \FOR {every $v \in L$}
    \STATE $\phi'(x) \assign v$ where $x \in \cl_2^{-1}(\{v\})$
    \COMMENT {Copy the leaf labels to a one to one function $\phi':
      supp(\cF) \rightarrow L$
      }\\
    \STATE Remove $\{x\}$ and $\{v\}$ from $\cF'$, $\cl'$  appropriately\\
    \ENDFOR

    \STATE $T' \assign T \setminus L$

    \STATE $\phi'' \assign$ {\tt
      get-hypergraph-isomorphism($\cF', \cl', T'$)}
    \STATE $\phi \assign \phi'' \cup \phi'$ \\
    \RETURN $\phi$
  \end{algorithmic}
\end{algorithm}

\begin{lemma}
  \label{lem:hyperiso}  %{lem:perm}
  If $(\cF, \cl)$ is an ICPPL from a tree $T$ and
  Algorithm~\ref{al:icppl-find-isomorph}, {\tt
    get-hypergraph-isomorphism ($\cF, \cl, T$)} returns a non-empty
  function, then there exists a hypergraph isomorphism $\phi :
  supp(\cF) \rightarrow V(T)$ such that the $\phi$-induced tree path
  labeling is equal to $\cl$ or $\cl_\phi = \cl$.
\end{lemma}
\begin{proof}
  It is clear that in the end of every recursive call to
  Algorithm~\ref{al:icppl-find-isomorph}, the function $\phi'$ is one
  to one involving all the leaves in the tree passed as input to that
  recursive call. Moreover, by Lemma~\ref{lem:noexit1} and
  Lemma~\ref{lem:xnotempty} it is consistent with the tree path
  labeling $\cl$ passed. The tree pruning is done by only removing
  leaves in each call to the function and is done till the tree
  becomes empty. Thus the returned function $\phi: supp(\cF)
  \rightarrow V(T)$ is a union of mutually exclusive one to one
  functions exhausting all vertices of the tree. In other words, it is
  a bijection from $supp(\cF)$ to $V(T)$ inducing the given path
  labeling $\cl$ and thus a hypergraph isomorphism. \qed
\end{proof}

\begin{theorem}
  \label{th:charac}
  A path labeling $(\cF, \cl)$ on tree $T$ is feasible iff it is an
  ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
  T)$ as input returns a non-empty function.
\end{theorem}
\begin{proof}
  From Lemma~\ref{lem:hyperiso}, we know that if $(\cF, \cl)$ is an
  ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
  T)$ as input returns a non-empty function, $(\cF, \cl)$ is feasible.
  Now consider the case where $(\cF, \cl)$ is feasible. i.e. there
  exists a hypergraph isomorphism $\phi$ such that $\cl_\phi =
  \cl$. Lemma~\ref{lem:noexit1} and Lemma~\ref{lem:xnotempty} show us
  that filter 1 and filter 2 do not exit if input is feasible. Thus
  Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty function.\qed
\end{proof}

\subsection{ICPPL when given tree is a path}
\label{subsec:icpplicpia}
\xnoindent
Consider a special case of ICPPL with the following properties.
\begin{enumerate}
\item Given tree $T$ is a path. Hence, all path labels are interval labels.
\item Only pairwise intersection cardinality
  preservation is sufficient. i.e. property (iii) in ICPPL is not enforced.
\item The filter algorithms do not have {\em \bf exit} statements.
\end{enumerate}
This is called an Intersection Cardinality Preservation Interval
Assignment (ICPIA) \cite{nsnrs09}. This structure and its algorithm is
used in the next section for finding tree path labeling from a
$k$-subdivided star due to this graph's close relationship with
intervals.

\section{Testing for feasible path assignments to $k$-subdivided star}
\label{sec:ksubdivstar}
In this section we consider the problem of assigning paths from a
$k$-subdivided star $T$ to a given set system $\cF$.  We consider
$\cF$ for which the overlap graph $\bO(\cF)$ is connected.  The
overlap graph is well-known from the work of
\cite{kklv10,nsnrs09,wlh02}.  We use the notation in
\cite{kklv10}. Recall from Section~\ref{sec:prelims} that hyperedges
$S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$ and $S'$
have a non-empty intersection but neither of them is contained in the
other. The overlap graph $\bO(\cF)$ is a graph in which the vertices
correspond to the sets in $\cF$, and the vertices corresponding to the
hyperedges $S$ and $S'$ are adjacent if and only if they overlap.  Note
that the intersection graph of $\cF$, $\bI(\cF)$ is different from
$\bO(\cF)$ and $\bO(\cF) \subseteq \bI(\cF)$.  A connected component
of $\bO(\cF)$ is called an overlap component of $\cF$.  An interesting
property of the overlap components is that any two distinct overlap
components, say $\cO_1$ and $\cO_2$, are such that any two sets $S_1
\in \cO_1$ and $S_2 \in \cO_2$ are disjoint, or, w.l.o.g, all the sets
in $\cO_1$ are contained within one set in $\cO_2$.  This containment
relation naturally determines a decomposition of the overlap
components into rooted containment trees.  We consider the case when
there is only one rooted containment tree, and we first present our
algorithm when $\bO(\cF)$ is connected.  It is easy to see that once
the path labeling to the overlap component in the root of the
containment tree is achieved, the path labeling to the other overlap
components in the rooted containment tree is essentially finding a
path labeling when the target tree is a path: each target path is a
path that is allocated to sets in the root overlap component.
Therefore, for the rest of this section, $\bO(\cF)$ is a connected
graph. We also assume that all hyperedges are of cardinality at most
$k+2$.

\xnoindent
Recall from Section~\ref{sec:prelims} that a $k$-subdivided star is a
star with each edge subdivided $k$ times.  Therefore, a $k$-subdivided
star has a central vertex which we call the {\em root}, and each root
to leaf path is called a {\em ray}.  First, we observe that by
removing the root $r$ from $T$, we get a collection of $p$ vertex
disjoint paths of length $k+1$, $p$ being the number of leaves in $T$.
We denote the rays by $R_1, \ldots, R_p$, and the number of vertices in
$R_i$, $i \in [p]$ is $k+2$.  Let
$\seq{v_{i1},\ldots,v_{i(k+2)}=r}$ denote the sequence of vertices in
$R_i$, where $v_{i1}$ is the leaf. Note that $r$ is a common vertex to all $R_i$.

\xnoindent In this section the given hypergraph, the
$k$-subdivided star and the root of the star are denoted by $\cO$, $T$ and vertex
$r$, respectively.
% For each hyperedge $X \in \cO$, we will maintain a 2-tuple of non-negative
% numbers $\seq{p_1(X), p_2(X)}$.  The numbers satisfy the property that
% $p_1(X) + p_2(X) \leq |X|$, and at the end of path labeling, for each
% $X$, $p_1(X) + p_2(X) = |X|$.  This signifies the algorithm tracking
% the lengths of subpaths of the path assigned to $X$ from at most two
% rays. We also maintain another parameter called the {\em residue} of
% $X$ denoted by $s(X)=|X| - p_1(X)$. This signifies the residue path
% length that must be assigned to $X$ which must be from another
% ray. For instance, if $X$ is labeled a path from only one ray, then
% $p_1(X) = |X|$, $p_2(X) = 0$ and $s(X) = 0$.

% \xnoindent
% We iteratively consider each ray from which paths will be assigned to
% hyperedges. At the beginning of each iteration hyperedges of $\cO$ are
% classifed into the following disjoint sets.
% \begin{enumerate}
% \item [$\cL_1^i$] {\em Labeled without $r$.} Those that have been
%   labeled with a path which does not contain $r$ in one of the
%   previous iterations.\\  $\cL_1^i = \set{ X \mid p_1(X) = |X| \text{ and
%     } p_2(X) = 0 \text{ and } s(X) = 0, X \in \cO}$
% \item [$\cL_2^i$] {\em Labeled with $r$.} Those that have been labeled
%   with two subpaths of $\cl(X)$ containing $r$ from two different rays
%   in two previous iterations.\\ $\cL_2^i = \set{X \mid 0 < p_1\left(X\right),
%     p_2\left(X\right) < |X| \text{ and } s(X) = 0, X \in \cO}$
%   \item [$\cT_1^i$] {\em Type 1 / partially labeled.} Those that have
%     been labeled with one path containing $r$ from a single ray in one
%     of the previous iterations. Here, $p_1(X)$ denotes the length of
%     the subpath of $\cl(X)$ that $X$ has been so far labeled
%     with. Also, it is clear that such a path must start at $r$ and
%     must be in a ray different from the one corresponding to $p_1(X)$.\\
%     $\cT_1^i = \set{ X \mid 0 < p_1(X) < |X| \text{ and } p_2(X) = 0
%       \text{ and } s(X) > 0, X \in \cO}$
%   \item [$\cT_2^i$] {\em Type 2 / not labeled.} Those that have not been
%     labeled with a path in any previous iteration.\\
%     $\cT_2^i = \set{ X \mid p_1(X) = p_2(X) = 0 \text{ and } s(X) = |X|,
%       X \in \cO}$
% \end{enumerate}
% \vspace{-2mm}
% \begin{align*}
%   \cO &= \cL_1^i \cup \cL_2^i \cup \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]\\
%   \cO_i &= \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]
% \end{align*}

\xnoindent
The set $\cO_i$ refers to the set of hyperedges $\cT_1^i \cup \cT_2^i$ in
the $i$th iteration.  Note that $\cO_1 = \cO$.  In the $i$th
iteration, hyperedges from $\cO_i$ are assigned paths from $R_i$ using
the following rules. Also the end of the iteration, $\cL_1^{i+1}, \cL_2^{i+1}, \cT_1^{i+1},
\cT_2^{i+1}$ are set to $\cL_1^{i}, \cL_2^{i}, \cT_1^{i},
\cT_2^{i}$ respectively, along with some case-specific changes mentioned in the
rules below.

\noindent
Suppose the given hypergraph has a feasible path labeling to the given
tree $T$. Let this labeling be $\cl$. Now we present an algorithm that
lists the sequence of edges that maintain a set of properties
necessary for them to have a feasible tree path labeling.
\begin{enumerate}[I.]
\item {\bf Step 1}
  \begin{enumerate}
  \item {\bf {There are type 1
        edges, $\mathbf{|\cT_1^i| > 0}$}:} 
% , and for each such $X$ assign the
%     path in $R_i$ of length $s(X)$ starting at $v_{i,k+2}=r$.  $X$ is
%     then removed from $\cO_i$, and $p_2(X)=s(X)$, and $s(X)=0$.If any
%     intersection cardinality is violated because of this assignment,
%     EXIT reporting the non-existence of a feasible assignment.
  {If there is only one $\cT_1^i$ hyperedge, label it to the unique
    path in $R_i$ of length $s(X)$ starting at $v_{i(k+2)}=r$. If
    there are more than one $\cT_1^i$ hyperedges, for any pair of
    hyperedges $S_1, S_2 \in \cT_1^i$, one of the following is
    true. Let $\cl'(X) \subseteq R_j$ denote the subpath of length
    $p_1(X)$ that has been assigned to $X$ from a previously
    considered ray $R_j$.}
  \begin{enumerate}[{\em Case }1. ]
  \item $|\cl'(S_1) \cap \cl'(S_2)| = 1$ and $|S_1 \cap S_2| > 1$
    i.e. they are in different rays and their residues must contain
    their intersection. Hence assign both to $R_i$.
  \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| >
    |\cl'(S_1) \cap \cl'(S_2)|$ i.e. they are in the same rays and
    their residues must contain their intersection. Hence assign both
    to $R_i$.
  \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| =
    |\cl'(S_1) \cap \cl'(S_2)|$. i.e. the intersection cardinality has
    been met. Hence assign only $S_1$ to $R_i$ and $S_2$ must be
    assigned to some $i \ge p$.
  \end{enumerate}
  Remove the labeled edge or edges from $\cO_i$ and add it to
  $\cO_{i+1}$. Also do not add it to $\cT_1^{i+1}$ and add to
  $\cL_2^{i+1}$.
\item {\bf {There are no type-1 edges nor} super-marginal type-2
    hyperedges:} If all type-2 edges of length at most $k+1$, EXIT by
  saying there is no ICPPL.  Otherwise, Find an $X$ of type-2 such
  that $|X| \geq k+2$, assign it the unique path starting at
  $v_{i(k+1)}=r$ of length $k+2$, and set $p_1(X)=k+2,
  s(X)=|X|-(k+2)$, assign it the unique path starting at
  $v_{i(k+1)}=r$ of length $k+2$, mark this as a type-1 hyperedge, and
  add it to $\cO_{i+1}$ after removing it from $\cO_i$.
  % \remove[a]{Next, iteratively, consider each type-1 edge $X \in
  %   \cO_i$, and if it intersects a previously assigned $Y$ such that
  %   $\cl(Y) \subseteq R_i$, assign a path of length $s(X)$
  %   containing $v_{i(k+2)}=r$, set $p_2(X)=s(X), s(X)=0$, and remove
  %   $X$ from $\cO_i$.}\tnote[a]{covered in step 2}

\item {\bf There are super-marginal type-2 hyperedges {and no type-1
      edges:}} Let $X$ be a super-marginal hyperedge. If $|X| \leq
  k+2$, then $X$ is assigned the path of length $|X|$ starting at
  $v_{i1}$, and $p_1(X) = |X|, p_2(X)=0$. $X$ is removed from $\cO_i$
  and not added to $\cO_{i+1}$.  If $|X| > k+2$, then $p_1(X)=k+2$,
  assign it the unique path starting at $v_{i(k+1)}=r$ of length $k+2$
  and $s(X) = |X| - (k+2)$.  In this case, $X$ is classified as a
  Type-1 edge, removed from $\cO_i$, and added to $\cO_{i+1}$.
\end{enumerate}
\item {\bf Step 2:} Iteratively, a hyperedge $X$ is selected from
  $\cO_i$ that has an overlap with one of the hyperedges $Y$ such that
  $\cl(Y) \subseteq R_i$, and a unique path is assigned to $X$.  The
  path, say $U(X) \subseteq R_i$, that is assigned can be decided
  unambiguously since the $X \overlap Y$, and all intersection
  cardinalities can be preserved only at one of the ends of $\cl(Y)$.

  \noindent
  Let $\cl(X)$ denote the unique path assigned to $X$.  If $X$ is a
  type-2 hyperedge and if the unique path of length $|X|$ does not
  contain $r$, then $p_1(X) = |X|, p_2(X)=0$, and $X$ is removed from
  $\cO_i$.  In the case when $\cl(X)$ has to contain $r$, then
  $p_1(X)$ is the length of the path, $p_2(X)=0$, and $s(X) =
  |X|-p_1(X)$. Further $X$ is classified as a type-1 hyperedge, added
  to $\cO_{i+1}$ and removed from $\cO_i$.  In the case when $X$ is a
  type-1 hyperedge, then we check if $U(X)$, which is of length $s(X)$
  contains $r$. If it does, then we assign $\cl(X) \leftarrow \cl(X)
  \cup U(X)$, remove $X$ from $\cO_i$ and do not add it to
  $\cO_{i+1}$.  If not, then we {\bf Exit} reporting that an
  assignment cannot be found.  The iteration ends when no hyperedge in
  $\cO_i$ has an overlap with a hyperedge assigned to $R_i$.
\end{enumerate}

\noindent
The order in which they are assigned is the sequence of hyperedges
with the following properties. Note that if such a sequence cannot be
computed, $\cO$ cannot have a feasible path labeling from $T$.

\noindent
\tnote[a]{need to formalize the below properties} In the following
lemmas we identify a set of necessary conditions for $\cF$ to have an
ICPPL in the $k$-subdivided star $T$.  If during the execution of the
above algorithm, one of these necessary conditions is not satisfied,
the algorithm exits reporting the non-existence of an ICPPL.
\begin{lemma}
  Let all hyperedges in $\cO_i$ be type-1 edges.  Then there is a
  maximal subset $\ccT_i \subseteq \cO_i$ with the following
  properties:
  \begin{enumerate}
  \item $\ccT_i$ form an inclusion chain.
  \item For all $X \in \ccT_i$, $s(X) \leq k+2$, and There is a an $X
    \in \ccT_i$ such that $s(X)=k+2$.
  \end{enumerate}
\end{lemma}
\begin{lemma}
  If there are no \marginal type-2 edges in $\cO_i$, then there exists
  at least $p-i$ type-2 hyperedges $X \in \cO_i$ such that $|X| \geq
  k+2$.
  % is a exist type-2 hyperedges of length at least $k+2$, and there
  % must be at least $r-i$ such type-2 hyperedge, basically at least
  % one per ray.  Each hyperedge contained in such an hyperedges of
  % length at least $k+2$, is transitively related under the overlap
  % relation to one that gets assigned a path that contains
  % $r$. Basically, the proof will argue that otherwise, there will be
  % a \marginal type-2 hyperedge, or that it is infeasible.  Also, it
  % is here we assume that we assume that the size of the two
  % universes is the same.
\end{lemma}
\begin{lemma}
  At the end of Step-1 in the $i$-th iteration, if one hyperedge $X$
  of type-2 is such that $\cl(X) \subseteq R_i$, then all other
  hyperedges in $\cO_i$ are connected to $X$ in the overlap component.
\end{lemma}
\begin{lemma}
  At the end of Step-2, If control has exit at any time, there is no
  ICPPL. If control has not exit, then $R_i$ is saturated.  No
  hyperedge of $\cO_{i+1}$ will get a path from $R_i$ in the future
  iterations.  No type-2 hyperedge of $\cO_{i+1}$ will get a path from
  $R_i$.  Basically $R_i$ is done.
\end{lemma}
\begin{lemma}
  Finally, we need to prove that the assignment is an ICPPL. Secondly,
  if there is a permutation then maps sets to paths, then it is indeed
  an ICPPL, and our algorithm will basically find it.  It is a unique
  assignment upto permutation of the leaves.
\end{lemma}

\begin{lemma}
  \label{lem:sup-mar}
  If $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
  tree path labeling to tree $T$, then $\cl(X)$ will
  contain a leaf in $T$.\\
  On the otherhand, if $\cl(X)$ has a leaf in
  $T$, $X$ is marginal but may not be super-marginal.
\end{lemma}
\begin{proof}
  Suppose $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
  path labeling from $T$.  Assume $\cl(X)$ does not have a leaf.  Let
  $R_i$ be one of the rays (or the only ray) $\cl(X)$ is part of.
  Since $X$ is in a connected overlap component, there exists $Y_1 \in
  \cF$ and $X \nsubseteq Y_1$ such that $Y_1 \overlap X$ and $Y_1$ has
  at least one vertex closer to the leaf in $R_i$ than any vertex in
  $X$. Similarly with the same argument there exists $Y_2 \in \cF$
  with same subset and overlap relation with $X$ except it has has at
  least one vertex farther away from the leaf in $R_i$ than any vertex
  in $X$. Clearly $Y_1 \cap X$ and $Y_2 \cap X$ cannot be part of same
  inclusion chain which contradicts that assumption $X$ is
  super-marginal. Thus the claim is proven.\qed
\end{proof}

\noindent
Consider the overlap graph $\bO(\cF)$ of the given hypergraph
$\cF$. Let $S_{sm} \in \cF$ be such that it is a super-marginal
hyperedge.  Algorithm~\ref{al:ktree-label} uses $S_{sm}$ along with
the overlap graph $\bO(\cF)$ to calculate the feasible tree path
labeling to the $k$-subdivded tree $T$.

\begin{algorithm}[h]
  \caption{{\tt compute-ksubtree-path-labeling($X, \cF, T$)}}
  \label{al:ktree-label}
  \begin{algorithmic}[\lndisplay]
    \IF{$X = S_{sm}$}
    \STATE -- TBD --\\
    \ELSE 
    \STATE -- TBD --\\
    \ENDIF

 \end{algorithmic}
\end{algorithm}

% \subsection{temp section from nsnsr09}

% \begin{algorithm}
%   \caption{Basic step in an algorithm to find an ICPIA for a prime
%     matrix $M'$}
%   \label{ds-algo}
%   ICPIA(Set $S$, Integer $p > 0$)

% \noindent
% /* {\tt $S \cap S^i \not = \phi$ for some $i \in \{1,\ldots,p\}$, but
%   $S \not\subseteq S^i$, $S^i \not\subseteq S$.  \\ Assigns to $S$ an
%   interval $I$ such that $\{I^1,\ldots,I^p,I\}$ forms an ICPIA for
%   $\{S^1,\ldots,S^p,S\}$.} */
% \begin{algorithmic}
%   \STATE Let $|S \cap S^i| = z$. \\
%   \STATE Let $I_l$ be the interval such that $|I_l \cap I^i| = z$, $|I_l| = |S|$ and the $z$ common elements are the smallest elements of $I^i$. \\
%   \STATE Let $I_r$ be the interval such that $|I_r \cap I^i| = z$, $|I_r| = |S|$, and the $z$ common elements are the largest elements of $I^i$. \\
%   \IF {$p == 1$}
%   \STATE Assign $I_l$ to $S$ \\
%   \STATE /* {\tt In this case, $I_r$ could also be assigned to $S$.  This will yield the {\bf other} ICPIA} */\\
%   \ELSE \IF{$|I_l \cap I^q| = |S \cap S^q|$ for each $q \in
%     \{1,\ldots,p\}$}
%   \STATE Assign $I_l$ to $S$ and exit.\\
%   \ENDIF \IF{$|I_r \cap I^q| = |S \cap S^q|$ for each $q \in
%     \{1,\ldots,p\}$}
%   \STATE Assign $I_r$ to $S$ and exit.\\
%   \ENDIF \ENDIF \STATE Report no ICPIA and exit.
% \end{algorithmic}
% \end{algorithm}

% \begin{theorem}
%   Algorithm~\ref{ds-algo} outputs an ICPIA to a prime matrix $M'$ iff
%   there is an ICPIA for $M'$.
% \end{theorem}
% \begin{proof}
%   The only-if part of the theorem is straightforward.  We now show
%   that if there is an ICPIA for $M'$, then Algorithm~\ref{ds-algo}
%   will indeed discover it.  The key fact is that in $M'$ for each set
%   $S$, there is another set $T \in M'$ such that $S \cap T \not =
%   \phi$, and $S$ and $T$ are not contained in each other.  Due to this
%   fact, there are exactly two ICPIAs for $M'$.  The two distinct
%   ICPIAs differ based on the interval assigned to $S_1$, see Algorithm
%  ~\ref{ds-algo}.  If $I_l$ is assigned to $S_1$, then we get one, and
%   the other ICPIA is obtained by assigning $I_r$ to $S_1$.  For each
%   subsequent set, say $S^j$, the interval to be assigned is forced.
%   It is forced due to the fact that the interval assigned to $S^j$ is
%   based on the interval assigned to $S^i$, where $S^i \cap S^j \not =
%   \phi$, and $S^i \not\subseteq S^j$, and $S^j \not\subseteq S^i$.
%   Given the fact that the algorithm is an exact implementation of
%   these observations, it follows that Algorithm~\ref{ds-algo} finds an
%   ICPIA if there is one.
% \end{proof}

\chapter{ Research - [title TBD]} 
\tnote[a]{the matrix theory submitted in walcom, any other work - need to see notes.}

\chapter{Conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Appendices.

\appendix

\chapter{Appendix A}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Bibliography.

\begin{singlespace}
  \bibliography{../lib/cop-variants__thesis}
\end{singlespace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% List of papers

\listofpapers

\begin{enumerate}  
\item Authors....  \newblock
 Title...
  \newblock {\em Journal}, Volume,
  Page, (year).
\end{enumerate}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DOCUMENTATION ON THE THESIS CLASS FILE ETC.

% Before reading any further please note that you are strongly advised
% against changing any of the formatting options used in the class
% provided in this directory, unless you are absolutely sure that it
% does not violate the IITM formatting guidelines.  \emph{Please do not
%   change the margins or the spacing.}  If you do change the formatting
% you are on your own (don't blame me if you need to reprint your entire
% thesis).  In the case that you do change the formatting despite these
% warnings, the least I ask is that you do not redistribute your style
% files to your friends (or enemies).

% It is also a good idea to take a quick look at the formatting
% guidelines.  Your office or advisor should have a copy.  If they
% don't, pester them, they really should have the formatting guidelines
% readily available somewhere.

% To generate PDF's with the links from the \verb+hyperref+ package use
% the following command:
% \begin{verbatim}
% % dvipdfm -o thesis.pdf thesis.dvi
% \end{verbatim}

% \section{Package Options}

% Use this thesis as a basic template to format your thesis.  The
% \verb+iitmdiss+ class can be used by simply using something like this:
% \begin{verbatim}
% \documentclass[PhD]{iitmdiss}  
% \end{verbatim}

% To change the title page for different degrees just change the option
% from \verb+PhD+ to one of \verb+MS+, \verb+MTech+ or \verb+BTech+.
% The dual degree pages are not supported yet but should be quite easy
% to add.  The title page formatting really depends on how large or
% small your thesis title is.  Consequently it might require some hand
% tuning.  Edit your version of \verb+iitmdiss.cls+ suitably to do this.
% I recommend that this be done once your title is final.

% To write a synopsis simply use the \verb+synopsis.tex+ file as a
% simple template.  The synopsis option turns this on and can be used as
% shown below.
% \begin{verbatim}
% \documentclass[PhD,synopsis]{iitmdiss}                                
% \end{verbatim}

% Once again the title page may require some small amount of fine
% tuning.  This is again easily done by editing the class file.

% This sample file uses the \verb+hyperref+ package that makes all
% labels and references clickable in both the generated DVI and PDF
% files.  These are very useful when reading the document online and do
% not affect the output when the files are printed.

% \section{Example Figures and tables}

% Fig.~\ref{fig:iitm} shows a simple figure for illustration along with
% a long caption.  The formatting of the caption text is automatically
% single spaced and indented.  Table~\ref{tab:sample} shows a sample
% table with the caption placed correctly.  The caption for this should
% always be placed before the table as shown in the example.

% \begin{figure}[htpb]
%   \begin{center}
%     \resizebox{50mm}{!} {\includegraphics *{iitm.eps}}
%     \resizebox{50mm}{!} {\includegraphics *{iitm.eps}}
%     \caption {Two IITM logos in a row.  This is also an
%       illustration of a very long figure caption that wraps around two
%       two lines.  Notice that the caption is single-spaced.}
%   \label{fig:iitm}
%   \end{center}
% \end{figure}

% \begin{table}[htbp]
%   \caption{A sample table with a table caption placed
%     appropriately. This caption is also very long and is
%     single-spaced.  Also notice how the text is aligned.}
%   \begin{center}
%   \begin{tabular}[c]{|c|r|} \hline
%     $x$ & $x^2$ \\ \hline
%     1  &  1   \\
%     2  &  4  \\
%     3  &  9  \\
%     4  &  16  \\
%     5  &  25  \\
%     6  &  36  \\
%     7  &  49  \\
%     8  &  64  \\ \hline
%   \end{tabular}
%   \label{tab:sample}
%   \end{center}
% \end{table}

% \section{Bibliography with BIB\TeX}

% I strongly recommend that you use BIB\TeX\ to automatically generate
% your bibliography.  It makes managing your references much easier.  It
% is an excellent way to organize your references and reuse them.  You
% can use one set of entries for your references and cite them in your
% thesis, papers and reports.  If you haven't used it anytime before
% please invest some time learning how to use it.

% I've included a simple example BIB\TeX\ file along in this directory
% called \verb+refs.bib+.  The \verb+iitmdiss.cls+ class package which
% is used in this thesis and for the synopsis uses the \verb+natbib+
% package to format the references along with a customized bibliography
% style provided as the \verb+iitm.bst+ file in the directory containing
% \verb+thesis.tex+.  Documentation for the \verb+natbib+ package should
% be available in your distribution of \LaTeX.  Basically, to cite the
% author along with the author name and year use \verb+\cite{key}+ where
% \verb+key+ is the citation key for your bibliography entry.  You can
% also use \verb+\citet{key}+ to get the same effect.  To make the
% citation without the author name in the main text but inside the
% parenthesis use \verb+\citep{key}+.  The following paragraph shows how
% citations can be used in text effectively.

% More information on BIB\TeX\ is available in the book by
% \cite{lamport:86}.  There are many
% references~\citep{lamport:86,prabhu:xx} that explain how to use
% BIB\TeX.  Read the \verb+natbib+ package documentation for more
% details on how to cite things differently.

% Here are other references for example.  \citet{viz:mayavi} presents a
% Python based visualization system called MayaVi in a conference paper.
% \citet{pan:pr:flat-fst} illustrates a journal article with multiple
% authors.  Python~\citep{py:python} is a programming language and is
% cited here to show how to cite something that is best identified with
% a URL.

% \section{Other useful \LaTeX\ packages}

% The following packages might be useful when writing your thesis.

% \begin{itemize}  
% \item It is very useful to include line numbers in your document.
%   That way, it is very easy for people to suggest corrections to your
%   text.  I recommend the use of the \texttt{lineno} package for this
%   purpose.  This is not a standard package but can be obtained on the
%   internet.  The directory containing this file should contain a
%   lineno directory that includes the package along with documentation
%   for it.

% \item The \texttt{listings} package should be available with your
%   distribution of \LaTeX.  This package is very useful when one needs
%   to list source code or pseudo-code.

% \item For special figure captions the \texttt{ccaption} package may be
%   useful.  This is specially useful if one has a figure that spans
%   more than two pages and you need to use the same figure number.

% \item The notation page can be entered manually or automatically
%   generated using the \texttt{nomencl} package.

% \end{itemize}

% More details on how to use these specific packages are available along
% with the documentation of the respective packages.

% END OF DOCUMENTATION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2011-11-15 src/thesis_outline_OLD.tex
setting up with github
v0.1 of synopsis.

\documentstyle[epsfig,algorithm,algorithmic,graphicx,graphics]{MS}

% \usepackage{fullpage} \usepackage{graphicx} \usepackage{epsfig}
% \usepackage{graphics} \usepackage{setspace} \usepackage{fancyhdr}
% \usepackage{algorithmic} \usepackage{algorithm} \date{}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary} \newtheorem{fact}{Fact}
% \newenvironment{proof}[1][Proof]{\begin{trivlist}
%   \item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{proof}{Proof:}{$\Box$}
\newtheorem{definition}{Definition}[section]
%%%%

%%%%

\begin{document}

\pagenumbering{roman}

\input{frontpage.tex} \input{certificate.tex} \input{dedication.tex}
\input{ack.tex} \input{abstract.tex}

\tableofcontents
\listoffigures

\chapter{Intro}
description, org of paper
\section{to area}
\section{to problem}
\section{to your thesis}
\section{thesis outline}

\chapter{Survey of COP and related problems}
\pagenumbering{arabic}

\section{characterization of COP - (tucker, mcconnell, prime submatrix)}
\section{Recognition of COP}
\subsection{poly algorithms (pq tree, pqr, icpia)}
\subsection{tuckers submatrices, Dom's algo to find them}
\subsection{mc connells certificate}
\section{algorithms to make a matrix into one with COP (min/max-cos-r/c) (dom)}
\section{brief analysis of hard problems with input having COP (chap 5, 6 in dom)}
\section{complexity - poly, logspace}

\chapter{Research}
\section{old and new cop tree results}

\end{document}

### 2011-11-16 src/synopsis.tex
added illustration for v0.2. peanut story and all that. need to complete it - will do by next commit.`

@@
-\documentclass[MS]{iitmdiss}
+\documentclass[MS,]{iitmdiss} 
+\usepackage[usenames,dvipsnames]{color}
 
@@
 %\newpage
+
+\chapter{Introduction}
+\label{chap:intro}
+
+give some basic gyan about the problem itself? or screw it move to
+section directly
+
+\section{Illustration of the problem}
+\label{sec:problem}
+
+\def \LLL {{\em ``{\bf L}aughing with Kafka''}} %{Consider the {\bf
+                                %L}obster} 
+\def \GGG {{\em ``{\bf G}irl with Curious Hair''}} % {``{\bf G}od Bless You,
+                                % Mr. Franzen''} 
+\def \BBB {{\em ``{\bf B}rief Interviews with Hideous Men''}} %{{\bf B}room
+                                %of the System} 
+\def \TTT {{\em ``{\bf T}his is Water''}} %{``{\bf T}ennis, Trigonometry,
+                                %Tornadoes: A Midwestern  Boyhood''}
+                                %%{{\bf T}he Pale King}
+
+\def \Pa {{\em {\bf Pa}tricia}} 
+\def \Pi {{\em {\bf Pi}gpen}} %{{\em {\bf Ma}rcie}}
+\def \Sn {{\em {\bf Sn}oopy}}
+\def \Wo {{\em {\bf Wo}odstock}}
+\def \Vi {{\em {\bf Vi}olet}} 
+\def \Li {{\em {\bf Li}nus}} 
+\def \Ch {{\em {\bf Ch}arlie}}
+\def \Sa {{\em {\bf Sa}lly}}
+\def \Fr {{\em {\bf Fr}anklin}}  %{{\em {\bf Re}run}} 
+\def \Sc {{\em {\bf Sc}hr{\"o}eder}} 
+\def \Lu {{\em {\bf Lu}cy}}
+
+\def \residenceblock {{\em Infinite Loop}}
+
+A group of students, \Pa, \Pi, \Sn, \Wo, \Vi, \Li, \Ch, \Sa, \Fr,
+  \Sc\ and \Lu\ enroll at the
+% {\bf Wo}odstock, {\bf Vi}olet, {\bf Li}nus, {\bf Ch}arlie, {\bf
+%   Sa}lly, {\bf Re}run, {\bf Sc}hr{\"o}eder} and {\em {\bf Lu}cy}
+% enroll at the
+{\em Wallace Studies Institute} for a liberal arts programme.  As part
+of their semester thesis, they pick a body of work to study and form
+the namesake study groups, {\LLL}, {\GGG}, {\BBB} and {\TTT}. A student
+will be in at least one study group and may be in more than one. For
+instance, as will be seen later, {\Fr} studies both {\LLL} and {\TTT}
+while \Wo\ studies only \BBB.
+
+Let $U$ and $\cF$ represent the set of students and the set of study
+groups respectively and the integers $n$ and $m$ denote the total
+number students and study groups respectively. In relation to this
+example, these are defined as follows. Also given below is the study group
+allocation to students. \tnote{use tables}
+\tnote{use other font for elements of sets}
+\begin{enumerate}[]
+\item {$U = \{$Pa,$\; $Pi,$\; $Sn,$\; $Wo,$\; $Vi,$\; $Li,$\; $Ch,$\;
+    $Sa,$\; $Fr,$\; $Sc,$\; $Lu$\}$}
+\item $\cF = \{L, G, B, T\}$
+ % \item ${R = \{g, h, i, j, k\}}$
+  % \item ${B = \{a, b, e, g\}}$
+  % \item ${O = \{c, b, d\}}$
+  % \item ${G = \{e, f , g, i\}}$
+\item $L = \{$ Ch,$\; $ Sa,$\; $ Fr,$\; $ Sc,$\; $ Lu $\}$
+\item $G = \{$ Pa,$\; $ Pi,$\; $ Vi,$\; $ Ch $\}$
+\item $B = \{$ Su,$\; $ Pi,$\; $ Wo $\}$
+\item $T = \{$ Vi,$\; $ Li,$\; $ Ch,$\; $ Fr $\}$
+\item $n = |U| = 11$, $m = |\cF| = 4$
+\end{enumerate}
+
+% The administration faces a small challenge with their
+% accommodation. 
+The campus has a residential area {\residenceblock} that has $n$
+single occupancy apartments reserved for the study groups'
+accommodation.  All these apartments are located such that the streets
+connecting them do {\em not} form loops. Fig~\ref{fig:streetmap} shows
+the street map for {\residenceblock}. It may be noted that as a graph, it
+classifies as a tree.
+
+A natural question would be to find how the students should be
+allocated apartments such that each study group has the {\em least
+  distance to travel} for a discussion? More specifically, we are
+interested in enforcing additional conditions, namely, that all the
+students in a study group must be next to each other; in other words,
+for one student to reach another fellow study group member's apartment
+(for all study groups the student is part of), she must not have to
+pass the apartment of any student who is not in that study group. To
+further elucidate, the apartments of students of any study group must
+be arranged in an exclusive unfragmented path on the street
+map. Exclusivity here means that the path must not have apartments
+from other study groups (unless that apartment is also part of {\em
+  this} study group).
+
+An intuitive approach to this problem would be to first find the paths
+that each study group decides to inhabit and then refine the
+allocation to individual students. One possible allocation of
+exclusive routes to study groups is illustrated in
+Fig~\ref{fig:streetmap+path}.
+
+\begin{figure}[h]
+  % \centering
+  (a) \includegraphics[scale=0.3]{../img/a1.pdf}\label{fig:streetmap}
+  (b) \includegraphics[scale=0.3]{../img/a2.pdf}\label{fig:streetmap+path}
+  \caption{(a) {\residenceblock} street map (b) Street map with study
+    group routes allocated. Routes are color coded as follows: red for
+    \textcolor{red}{\em \LLL} group, blue for \textcolor{blue}{\em
+      \GGG} group, orange for \textcolor{YellowOrange}{\em \BBB} group,
+    green for \textcolor{green}{\em \TTT} group}
+\end{figure}
+
+\section{Motivation and background}
+\label{sec:motivation}
+
+\chapter{Summary of results}
+\label{chap:results}
 

### 2011-11-19 src/synopsis.tex
a whole bunch of changes. - put up the peanut dfw story. added images etc needed. i think some reorg of files were done too - organized the bib file chronologically. and fixed all the TBDs (except bl62 which is elusive and probably i'll delete it.

@@
 
-give some basic gyan about the problem itself? or screw it move to
-section directly
+\tnote{give some basic gyan about the problem itself? or screw it move to
+section directly}
 
@@
 
+\def \WSI {{\em Wallace Studies Institute}}
+
 A group of students, \Pa, \Pi, \Sn, \Wo, \Vi, \Li, \Ch, \Sa, \Fr,
@@
 % enroll at the
-{\em Wallace Studies Institute} for a liberal arts programme.  As part
+{\WSI} for a liberal arts programme.  As part
 of their semester thesis, they pick a body of work to study and form
@@
 that each study group decides to inhabit and then refine the
-allocation to individual students. One possible allocation of
-exclusive routes to study groups is illustrated in
-Fig~\ref{fig:streetmap+path}.
+allocation to individual students. A feasible allocation of exclusive
+routes to study groups is illustrated in Fig~\ref{fig:streetmappath}
+and the students' allocation of apartments that obey's this route
+allocation is shown in Fig~\ref{fig:streetmappathpeople}. How this is
+algorithmically computed is the focus of this thesis.
 
 \begin{figure}[h]
-  % \centering
-  (a) \includegraphics[scale=0.3]{../img/a1.pdf}\label{fig:streetmap}
-  (b) \includegraphics[scale=0.3]{../img/a2.pdf}\label{fig:streetmap+path}
-  \caption{(a) {\residenceblock} street map (b) Street map with study
-    group routes allocated. Routes are color coded as follows: red for
-    \textcolor{red}{\em \LLL} group, blue for \textcolor{blue}{\em
-      \GGG} group, orange for \textcolor{YellowOrange}{\em \BBB} group,
-    green for \textcolor{green}{\em \TTT} group}
+  \centering
+  \includegraphics[scale=0.3]{../img/1_infinite_loop.png}
+  \caption{{\residenceblock} street map}
+  \label{fig:streetmap}
 \end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.5]{../img/2_infinite_loop.png}
+  \caption{{\residenceblock} street map with study group routes
+    allocated. Routes are color coded as follows: red for
+    \textcolor{red}{\bf L} group, blue for \textcolor{blue}{\bf G}
+    group, orange for \textcolor{YellowOrange}{\bf B} group, green for
+    \textcolor{green}{\bf T} group}
+  \label{fig:streetmappath}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
+  \caption{Individual allocation of apartments to students in
+    {\residenceblock} that meets the requirements stated before.}
+  \label{fig:streetmappathpeople}
+\end{figure}
+
+If all the apartments are on the same street or if they are all lined
+up on a single path, the street map becomes a tree that is just a
+path. Then the problem becomes what is called an {\em interval
+  assignment problem}. The idea of interval assignment may not be
+obvious here; hence to see this, consider a different problem in
+{\WSI} where the classes for these study groups courses need to be
+scheduled during a day (or a week or any time period). Each study
+group has a bunch of courses associated with it some of which may be
+shared by two or more study groups. It is mandatory that a student who
+is a member of a study group takes all the courses associated with
+that group. There are slots during the day for classes to be held and
+the problem is to allocate class slots to courses such that all the
+classes of a study group are consecutive. It is debatable if this will
+not hamper the attention span and memory retention rate of the students but
+that is, regrettably, out of the scope of this thesis. The parallels between this
+class allocation problem and the accommodation problem can be seen as
+follows. The set $U$ here, are the courses offered (say Course 101 ``Post
+modernism in Wallace's work'', Course 102 ``A study on fragmented prose
+method'' and so on). The collection $\cF$ is the set of study groups but the
+study groups are filled by course IDs and not students like in the
+earlier example. For instance, Course 101 is mandatory for all study
+groups {\bf L, G, B, T} and Course 102 is mandatory for only the {\bf B}
+group) and so on. The sequence of class slots for the day (or week or
+any time period) is analogous to the street map in the accommodation
+problem. It is quite obvious now why this version of the problem
+(where the ``target graph'' is a path and not any tree) is called an
+interval assignment problem.
+
+\tnote{TBD: BEGIN REVIEW FROM BELOW}
+
+The interval assignment problem to a set system is equivalent to the
+consecutive ones problem (COP) in binary matrices\cite{wlh02,
+  nsnrs09}.  The COP problem is to rearrange rows (columns) of a
+binary matrix in such a way that every column (row) has its $1$s occur
+consecutively. If this is possible the matrix is said to have the COP.
+Consecutive ones property (COP) of binary matrices is a well
+researched combinatorial problem and has several positive results on
+tests for it and computing the COP permutation (i.e. the class slot
+allocation to course IDs in the above illustration) which will be
+surveyed later in this document. Hence we are interested in extensions
+of COP, more specifically, the extension of interval assignment
+problem to tree path assignment problem (which is illustrated by the
+study group accommodation problem).
+
+\tnote[a]{THESIS: UNCOMMENT} \tnote{include the rest of the sections as
+  the organization of the doc}
+% The necessary preliminaries with definitions etc. are presented in
+% Section~\ref{sec:prelims}. 
+% Section~\ref{sec:feasible} documents the
+% characterization of a feasible path labeling and finally,
+% Section~\ref{sec:ksubdivstar} describes a polynomial time algorithm to
+% find the tree path labeling of a given set system from a given
+% $k$-subdivided tree.
+
+Section~\ref{sec:motivation} gives a brief survey of COP and relatives.
+Chapter~\ref{chap:results} presents a summary of our results on the
+tree path labeling problem.
 
@@
 
-\chapter{Summary of results}
-\label{chap:results}
-
-%\chapter{\MakeUppercase{Introduction}}
-\chapter{INTRODUCTION}
-Consecutive ones property (COP) of binary matrices is a widely studied
-combinatorial problem. The problem is to rearrange rows (columns) of a
-binary matrix in such a way that every column (row) has its $1$s occur
-consecutively. If this is possible the matrix is said to have the COP.
+As descirbed earlier, a matrix with COP is one whose rows (columns)
+can be rearranged so that the 1s in every column (row) are in
+consecutive rows (columns).
+
 \begin{figure}[h]
   \centering
-
-  \caption{(i) matrix with COP (ii) without COP}
+  \includegraphics[scale=0.4]{../img/0_cop_nocop.png}
+
+  \caption{{\em Example for the COP:} The matrix on the left has the
+    COP because by permuting its columns (labeled with $c_1$-$c_4$)
+    one can obtain the matrix shown in the middle where the 1s in each
+    row appear consecutively. The matrix on the right, in contrast,
+    does not have the COP \cite{at72,d08phd}.}
   \label{fig:cop-matrix}
@@
 
-This problem has several practical applications in diverse fields
-including scheduling \cite{hl06}, information retrieval \cite{k77} and
-computational biology \cite{abh98}.  Further, it is a tool in graph
-theory \cite{mcg04} for interval graph recognition, characterization
-of Hamiltonian graphs, and in integer linear programming
-\cite{ht02,hl06}.
-
-\tnote[a]{define ``COP permutation''}
-
-\section{Background}
-Different aspects of COP were studied as part of this thesis. They are
-described in the following subsections.
-
-\subsection{Consecutive Ones Testing}
+COP in binary matrices has several practical applications in diverse
+fields including scheduling \cite{hl06}, information retrieval
+\cite{k77} and computational biology \cite{abh98}.  Further, it is a
+tool in graph theory \cite{mcg04} for interval graph recognition,
+characterization of Hamiltonian graphs, planarity testing \cite{bl76}
+and in integer linear programming \cite{ht02,hl06}.
+
 The obvious first questions after being introduced to the consecutive
@@
 was given by \cite{fg65}.  Since then several algorithms have been
-invented; PQ trees \cite{bl76}, variations of PQ trees
-\cite{mm09,wlh01,wlh02,mcc04}, ICPIA \cite{nsnrs09} are some of them.
-
-\cite{bl76} showed for the first time that COP testing (COT) is indeed
-polynomial time solvable by describing the construction of a data
-structure called PQ trees. This paper draws on the close relationship
-of COP matrices to interval graphs. A PQ tree of a matrix is one that
-stores all row (column) permutations of the matrix that give the COP
-orders (there could be multiple) of the matrix. This is constructed using an elaborate procedure in the linear
-time algorithm described in \cite{bl76}.\tnote[a]{verify if it is
-  linear}. Since this result, several improvements have been made to
-the procedural complexity of the algorithm itself and also derivative
-data structures that are easier to construct. PQR trees is one such
-data structure \cite{mm96}, \cite{mpt97}\tnote{also, mpt97 $==$
-  elaborate version of mm96}.  \cite{tm03} describes an improved
-version of PQR trees. \tnote[a]{improvement in terms of
-  what?}. \cite{h01a} describes the simpler algorithm for
-COT. \tnote[a]{This result first apprearted in proc ISAAC 92} Hsu also
-invented PC trees \cite{h01b} which is claimed to be much easier to
-implement. \cite{nsnrs09} describes a characterization of consecutive
-ones property solely based on the cardinality properties of the set representations
-of the columns (rows); every column (row) is equivalent to a set that
-has the row (column) indices of the rows (columns) that have one
-entries in this column (row). This is interesting and relevant,
-especially to this thesis because it simplifies COT to a great
-degree. \tnote[a]{because it reduces the search space for the solution
-  from set of all permutations of rows to set of all-- get this
-  figured out!}
+invented; PQ trees \cite{bl76} gave the first linear time
+algorithm. Variations of PQ trees \cite{mm96,wlh01,wlh02,mcc04} and
+ICPIA \cite{nsnrs09} further improve either the time complexity and/or
+the complexity of implementation.
+
+\cite{bl76} describes a linear time construction of a data structure
+called PQ trees. This paper draws on the close relationship of COP
+matrices to interval graphs. A PQ tree of a matrix is one that stores
+all row (column) permutations of the matrix that give the COP orders
+(there could be multiple) of the matrix. This is constructed using an
+elaborate linear time procedure. Since this
+result, several improvements have been made to the procedural
+complexity of the algorithm itself and also derivative data structures
+that are easier to construct. PQR trees is one such data structure
+\cite{mm96}, \cite{mpt98}\tnote{also, mpt98 $==$ elaborate version of
+  mm96}.  \cite{tm05} describes an improved version of PQR
+trees. \tnote[a]{improvement in terms of what?}\cite{wlh02} describes
+the simpler algorithm for COT. \tnote[a]{This result first apprearted
+  in proc ISAAC 92} Hsu also invented PC trees \cite{wlh01} which is
+claimed to be much easier to implement. \cite{nsnrs09} describes a
+characterization of consecutive ones property solely based on the
+cardinality properties of the set representations of the columns
+(rows); every column (row) is equivalent to a set that has the row
+(column) indices of the rows (columns) that have one entries in this
+column (row). This is interesting and relevant, especially to this
+thesis because it simplifies COT to a great degree. \tnote[a]{because
+  it reduces the search space for the solution from set of all
+  permutations of rows to set of all-- get this figured out!}
 
@@
 machinery of this algorithm \annote[a] {is related to the weighted
-  betweenness problem addressed } {in what way??} in \cite{co97}.
+  betweenness problem addressed } {in what way??} in \cite{co98}.
 \tnote[a]{expand on the COP order graph creation and it having to be
@@
 
-% \noindent
-To elaborate on what was mentioned earlier about \cite{nsnrs09}, the
-problem of COT can be seen as a constraint satisfaction problem
-involving a system of sets from a universe as follows. Every column of
-the binary matrix can be converted into a set of integers which are
-the indices of rows with $1$s in that column. It is apparent that if
-the matrix has COP, then constructing such sets after applying the COP
-permutation to the matrix will result in all sets with consecutive
-integers. In other words, after application of COP reordering, the
-sets are intervals. Indeed the problem now becomes finding interval
-assignments to a given set system such that there exists a permutation
-of the universe (set of row indices) which converts each set to its
-assigned interval. The result in \cite{nsnrs09} characterizes interval
-assignments to the sets which can be obtained from a single
-permutation of the rows.  They show that for each set, the cardinality
-of the interval assigned to it must be same as the cardinality of the
-set, and the intersection cardinality of any two sets must be same as
-the intersection cardinality of the corresponding intervals.  While
-this is obviously a necessary condition, this result shows this is
-also sufficient.  \cite{nsnrs09} calls such an interval assignment an
-Intersection Cardinality Preserving Interval Assignment (ICPIA).  This
-paper generalizes the idea from \cite{wlh02} of decomposing a given
-binary matrix into prime matrices for COT and describes an algorithm
-to test if an ICPIA exists for a given set system.
-
 The literature also has results on parallel COT algorithms --
 \cite{as95} \cite{bs03} \tnote[a]{is this parallel?}
-\cite{cy91}\tnote[a]{is this parallel?}
+\cite{ly91}\tnote[a]{is this parallel?}
 
@@
 
-\subsection{Matrix modification to attain COP or ``almost'' COP}
-\tnote[a]{THESIS: results and survey from \cite{d08phd} Chapters 2 and
-  3}
-
-So far we have been concerned about matrices that have the consecutive
+{\bf Matrices with no COP}: So far we have been concerned about matrices that have the consecutive
 ones property. However in real life applications, it is rare that data
@@
   is it any submatrix with k less than r rows or submatrix must have
-  all columns?}. \cite{hgxx} \tnote[a]{get year XX} corrects an error of the abridged proof
+  all columns?}. \cite{hg02} corrects an error of the abridged proof
 of this reduction as given in \cite{gj79}.  \cite{d08phd} discusses
@@
 
-\tnote[a]{THESIS: where should this go? \cite{tz03} (approx submatrix with COP sparse matrices)}
+\tnote[a]{THESIS: where should this go? \cite{tz04} (approx submatrix with COP sparse matrices)}
 
@@
 number of blocks of ones is proven to be NP complete by
-\cite{kou77}. \tnote[a]{THESIS: what is the reduction?}
+\cite{k77}. \tnote[a]{THESIS: what is the reduction?}
 
@@
   \item \cite{ckl96} (ILP, circ ones, one drop)
-  \item \cite{th99} (generalization of COP - minimax, biotonic column) Tucker
+  \item \cite{th98} (generalization of COP - minimax, biotonic column) Tucker
   \end{enumerate}
@@
 
-\subsection{Graph Isomorphism}
-As it will be described in detail later in this document, isomorphism
-of certain classes of graphs, namely chordal graphs, have a close
-relationship with consecutive ones property and generalizations of it.
-This is perhaps because of how closely COP of a matrix relates to
-properties of graphs derived from matrices as seen in the following
-results. %-- Other relations of COP with Graphs: \\
-\tnote[a]{THESIS: peo exists iff chordal. lexicographic BFS
-  [tag:chordalGraph]} 
-A well known result in perfect graph theory is that the maximal
-cliques of an interval graph $G$ can be linearly ordered such that for
-all $v \in V(G)$, cliques containing $v$ are consecutive in the
-ordering\cite{gh64}. This clearly means that a graph $G$ is an
-interval graph if and only if \tnote[a]{verify from paper the
-  statement of claim.} maximal clique vertex incidence matrix of $G$
-has COP.  Also, maximal cliques of any chordal graph can be enumerated
-in polytime $O(m+n)$ \tnote[a]{citation?!!}.  \cite{fg65} uses these
-results to give the first polynomial time algorithm for
-COT. \tnote[a]{THESIS: check. how do they use it?}  A bipartite graph
-is convex (on $R$) if and only if its half adjacency matrix has COP on
-rows.  The results in \cite{bl62} on COT are based on the result that
-interval graphs are AT-free chordal graphs. \tnote[a]{the latter being
-  Tucker's?}
-
-\tnote[a]{THESIS: -- How COP (and TPL) relates to interval graph
-  isomorphism (and path graph isomorphism) move/summarize what is
-  already in Motivation section.}
-
-\tnote[a]{THESIS: TBD survey -- see blue notes under Graph
-  Isomorphism. namely citations in \cite{aas93}}
-
-\tnote[a]{THESIS: -- -- a brief on heirarchy of chordal graphs, path
-  graphs, interval graphs, peo, clique tree etc. and results.}
-
-\tnote[a]{THESIS: -- interval graphs are incomparability graphs. see
-  Golumbic. [tag:classification]}
-
-\tnote[a]{THESIS: finding min length hole in bipartite graph is
-  polynomial time [Sec 3.3.2 in \cite{d08phd}]}
-
-\tnote[a]{THESIS: theorem 2.2 in \cite{d08phd} - $G$ is union of
-  v.d. caterpillars iff $M$ has COP in rows. $M$ is edge vertex
-  incidence matrix $(*,2)$ of $G$}
-
-\subsection{Logspace complexity of graph canonization using COP}
-
-Canonization is an important tool in graph isomorphism. Invention of a
-deterministic method of canonization for any class of graphs naturally
-results in an algorithm for isomorphism. All that is required is to
-check if the canons of two graphs are the same. Thus complexity of
-graph isomorphism can be studied by studying canonization
-methods. While general graph ismorphism remains elusive in terms of
-hardness, canonization has been studied for smaller classes of graphs
-thus giving complexity/hardness results for them.
-
-In 1992, \cite{sl92} made an important discovery that tree isomorphism
-is in logspace. It was done by inventing a method of canonization of
-trees using a logspace depth first traversal
-algorithm. \tnote[a]{confirm the statement of claim}.
-
-\cite{kklv10} proved that interval graph canonization is also in
-logspace thus drawing logspace conslusions about COT. 
-% This result is
-% built on the fact that 
-Interval graphs are FP+C (fixed point with
-counting) definable \tnote[a]{elucidate} and \cite{l10} showed that
-this implies that it captures PTIME. \tnote[a]{elucidate} This result
-along with that of undirected graph reachability being in logspace
-\cite{r08}, \cite{kklv10} proved their logspace result.\\
-
-\tnote[a]{THESIS: -- -- the interesting survey in conclusion section
-  of kklv10}
-
-\tnote[a]{THESIS:
-{\bf COP helping some problems with hardness}
-having COP in the input structure makes some problems less hard than
-in a general input.
-
--- min set cover problem is known to be NP hard. so is min hitting
-set. these two problems are equivalent \cite{adp80}\\
--- -- complexity results: log factor polynomial time approximable \cite{f98}, W[2]-complete with solution size as parameter \cite{df99}\\
--- a couple of generalizations:\\
--- -- min degree hypergraph (MDH) problem\\
--- -- red blue covering (RBSC) problem\\
-}
-
-\tnote[a]{THESIS: \cite{d08phd} Chapter 5.}
-
-\tnote[a]{THESIS: what are \cite{d08phd} results}
-
-\tnote[a]{THESIS: -- Rectangle stabbing \tnote[a]{see \cite{d08phd} Chapter 6?}}
-
-\tnote[a]{THESIS: -- {\em \cite{dgnw06} (min membership hitting set + COP)}}
-
-\section{Motivation, Objective and Scope}
-
-A natural generalization of the interval assignment problem is
+{\bf Motivation: } A natural generalization of the interval assignment problem is
 feasible tree path labeling problem of a set system. The problem is
@@
 
-%\noindent 
-\section{Summary of Results}
+How the problem of COT is related to the constraint statisfaction
+problem of interval assignment to sets in a system can be seen as
+follows. Every column of the binary matrix can be converted into a set
+of integers which are the indices of rows with $1$s in that column. It
+is apparent that if the matrix has COP, then constructing such sets
+after applying the COP permutation to the matrix will result in all
+sets with consecutive integers. In other words, after application of
+COP reordering, the sets are intervals. Indeed the problem now becomes
+finding interval assignments to a given set system such that there
+exists a permutation of the universe (set of row indices) which
+converts each set to its assigned interval. The result in
+\cite{nsnrs09} characterizes interval assignments to the sets which
+can be obtained from a single permutation of the rows.  They show that
+for each set, the cardinality of the interval assigned to it must be
+same as the cardinality of the set, and the intersection cardinality
+of any two sets must be same as the intersection cardinality of the
+corresponding intervals.  While this is obviously a necessary
+condition, this result shows this is also sufficient.  \cite{nsnrs09}
+calls such an interval assignment an Intersection Cardinality
+Preserving Interval Assignment (ICPIA).  This paper generalizes the
+idea from \cite{wlh02} of decomposing a given binary matrix into prime
+matrices for COT and describes an algorithm to test if an ICPIA exists
+for a given set system.
+
+\chapter{Summary of results}
+\label{chap:results}
+
 Given a path labeling $\cl$ to $\cF$ from a tree $T$, we give a
@@
 
-\section{Organization of document} 
-\tnote[a]{THESIS: UNCOMMENT} 
-% The necessary preliminaries with definitions etc. are presented in
-% Section~\ref{sec:prelims}. 
-Section~\ref{sec:feasible} documents the
-characterization of a feasible path labeling and finally,
-Section~\ref{sec:ksubdivstar} describes a polynomial time algorithm to
-find the tree path labeling of a given set system from a given
-$k$-subdivided tree.
-
-%\chapter{\MakeUppercase{Results}}
-\chapter{RESULTS}
+\section{RESULTS}
 % \section{Preliminaries} 
@@
 
-%\section{Conclusions}
+\chapter{Conclusion}
+\tnote{MAKE CONNECTION TO PATH GRAPH ISOMORPHISM HERE}
+As it will be described in detail later in this document, isomorphism
+of certain classes of graphs, namely chordal graphs, have a close
+relationship with consecutive ones property and generalizations of it.
+This is perhaps because of how closely COP of a matrix relates to
+properties of graphs derived from matrices as seen in the following
+results. %-- Other relations of COP with Graphs: \\
+\tnote[a]{THESIS: peo exists iff chordal. lexicographic BFS
+  [tag:chordalGraph]} 
+A well known result in perfect graph theory is that the maximal
+cliques of an interval graph $G$ can be linearly ordered such that for
+all $v \in V(G)$, cliques containing $v$ are consecutive in the
+ordering\cite{gh64}. This clearly means that a graph $G$ is an
+interval graph if and only if \tnote[a]{verify from paper the
+  statement of claim.} maximal clique vertex incidence matrix of $G$
+has COP.  Also, maximal cliques of any chordal graph can be enumerated
+in polytime $O(m+n)$ \tnote[a]{citation?!!}.  \cite{fg65} uses these
+results to give the first polynomial time algorithm for
+COT. \tnote[a]{THESIS: check. how do they use it?}  A bipartite graph
+is convex (on $R$) if and only if its half adjacency matrix has COP on
+rows.  The results in \cite{bl76}
+%\cite{bl62} 
+on COT are based on the result that
+interval graphs are AT-free chordal graphs. \tnote[a]{the latter being
+  Tucker's?}
+
+\tnote[a]{THESIS: -- How COP (and TPL) relates to interval graph
+  isomorphism (and path graph isomorphism) move/summarize what is
+  already in Motivation section.}
+
+\tnote[a]{THESIS: TBD survey -- see blue notes under Graph
+  Isomorphism. namely citations in \cite{aas93}}
+
+\tnote[a]{THESIS: -- -- a brief on heirarchy of chordal graphs, path
+  graphs, interval graphs, peo, clique tree etc. and results.}
+
+\tnote[a]{THESIS: -- interval graphs are incomparability graphs. see
+  Golumbic. [tag:classification]}
+
+\tnote[a]{THESIS: finding min length hole in bipartite graph is
+  polynomial time [Sec 3.3.2 in \cite{d08phd}]}
+
+\tnote[a]{THESIS: theorem 2.2 in \cite{d08phd} - $G$ is union of
+  v.d. caterpillars iff $M$ has COP in rows. $M$ is edge vertex
+  incidence matrix $(*,2)$ of $G$}
+
+{\bf Complexity challenges:} Canonization is an important tool in graph isomorphism. Invention of a
+deterministic method of canonization for any class of graphs naturally
+results in an algorithm for isomorphism. All that is required is to
+check if the canons of two graphs are the same. Thus complexity of
+graph isomorphism can be studied by studying canonization
+methods. While general graph ismorphism remains elusive in terms of
+hardness, canonization has been studied for smaller classes of graphs
+thus giving complexity/hardness results for them.
+
+In 1992, \cite{sl92} made an important discovery that tree isomorphism
+is in logspace. It was done by inventing a method of canonization of
+trees using a logspace depth first traversal
+algorithm. \tnote[a]{confirm the statement of claim}.
+
+\cite{kklv10} proved that interval graph canonization is also in
+logspace thus drawing logspace conslusions about COT. 
+% This result is
+% built on the fact that 
+Interval graphs are FP+C (fixed point with
+counting) definable \tnote[a]{elucidate} and \cite{l10} showed that
+this implies that it captures PTIME. \tnote[a]{elucidate} This result
+along with that of undirected graph reachability being in logspace
+\cite{rei08}, \cite{kklv10} proved their logspace result.\\
+
+\tnote[a]{THESIS: -- -- the interesting survey in conclusion section
+  of kklv10}
+
+\tnote[a]{THESIS:
+{\bf COP helping some problems with hardness}
+having COP in the input structure makes some problems less hard than
+in a general input.
+
+-- min set cover problem is known to be NP hard. so is min hitting
+set. these two problems are equivalent \cite{adp80}\\
+-- -- complexity results: log factor polynomial time approximable \cite{f98}, W[2]-complete with solution size as parameter \cite{df99}\\
+-- a couple of generalizations:\\
+-- -- min degree hypergraph (MDH) problem\\
+-- -- red blue covering (RBSC) problem\\
+}
+
+\tnote[a]{THESIS: \cite{d08phd} Chapter 5.}
+
+\tnote[a]{THESIS: what are \cite{d08phd} results}
+
+\tnote[a]{THESIS: -- Rectangle stabbing \tnote[a]{see \cite{d08phd} Chapter 6?}}
 

### 2011-11-20 src/synopsis.tex
1) added all the lit i have 2) changed the acronyms of study groups. fiiine, i obliged. but it is STILL IN THE CODE. 3) updated picture to match changes
shit just got real

@@
 
-\def \LLL {{\em ``{\bf L}aughing with Kafka''}} %{Consider the {\bf
-                                %L}obster} 
-\def \GGG {{\em ``{\bf G}irl with Curious Hair''}} % {``{\bf G}od Bless You,
-                                % Mr. Franzen''} 
-\def \BBB {{\em ``{\bf B}rief Interviews with Hideous Men''}} %{{\bf B}room
-                                %of the System} 
-\def \TTT {{\em ``{\bf T}his is Water''}} %{``{\bf T}ennis, Trigonometry,
-                                %Tornadoes: A Midwestern  Boyhood''}
-                                %%{{\bf T}he Pale King}
+% \def \LLL {{\em ``{\bf L}aughing with Kafka''}} 
+% \def \GGG {{\em ``{\bf G}irl with Curious Hair''}} 
+% \def \BBB {{\em ``{\bf B}rief Interviews with Hideous Men''}}
+% \def \TTT {{\em ``Everything and More: A Compact His{\bf T}ory of Infinity''}} 
+% \def \xLLL {{\bf L}}
+% \def \xGGG {{\bf G}} 
+% \def \xBBB {{\bf B}}
+% \def \xTTT {{\bf T}}
+
+%{\bf L}aughing {\bf G}irl {\bf B}rief {\bf T}his
+
+% \def \LLL {{\em ``{\bf L}aughing with Kafka''}} 
+% \def \GGG {{\em ``Brief {\bf I}nterviews with Hideous Men''}} 
+% \def \BBB {{\em ``Broom of the {\bf S}ystem''}}
+% \def \TTT {{\em ``Everything and More: {\bf A} Compact History of Infinity''}}
+
+% \def \xLLL {{\bf L}}
+% \def \xGGG {{\bf I}} 
+% \def \xBBB {{\bf S}}
+% \def \xTTT {{\bf A}}
+
+\def \LLL {{\em ``{\bf B}rief Interviews with Hideous Men''}} 
+\def \GGG {{\em ``The String {\bf T}heory''}} 
+\def \BBB {{\em ``[{\bf W}]Rhetoric and the Math Melodrama''}}
+\def \TTT {{\em ``{\bf F}ate, Time, and Language: An Essay on Free Will''}}
+
+\def \xLLL {{\bf B}}
+\def \xGGG {{\bf T}} 
+\def \xBBB {{\bf W}}
+\def \xTTT {{\bf F}}
+
+%{{\em ``Everything and More: A Compact History of In{\bf F}inity''}} 
+% {{\em ``Broom of {\bf T}he System''}} % {Consider the Lobster}
+% {{\em ``God Bless You, Mr. {\bf F}ranzen''}}
+% {{\em ``{\bf L}aughing with Kafka''}} 
+% {{\em ``{\bf G}irl with Curious Hair''}} 
+% {{\em ``Consider the {\bf L}obster|}}
+% {{\em ``{\bf G}od Bless You, Mr. Franzen''}}
+% {{\em ``{\bf T}ennis, Trigonometry,
+%                                Tornadoes: A Midwestern  Boyhood''}}
+% {{\em ``{\bf T}his is Water''}} %
+% {\em {{\bf T}he Pale King}}
 
@@
     $Sa,$\; $Fr,$\; $Sc,$\; $Lu$\}$}
-\item $\cF = \{L, G, B, T\}$
- % \item ${R = \{g, h, i, j, k\}}$
-  % \item ${B = \{a, b, e, g\}}$
-  % \item ${O = \{c, b, d\}}$
-  % \item ${G = \{e, f , g, i\}}$
-\item $L = \{$ Ch,$\; $ Sa,$\; $ Fr,$\; $ Sc,$\; $ Lu $\}$
-\item $G = \{$ Pa,$\; $ Pi,$\; $ Vi,$\; $ Ch $\}$
-\item $B = \{$ Su,$\; $ Pi,$\; $ Wo $\}$
-\item $T = \{$ Vi,$\; $ Li,$\; $ Ch,$\; $ Fr $\}$
+\item $\cF = \{\xLLL, \xGGG, \xBBB, \xTTT\}$
+\item $\xLLL = \{$ Ch,$\; $ Sa,$\; $ Fr,$\; $ Sc,$\; $ Lu $\}$
+\item $\xGGG = \{$ Pa,$\; $ Pi,$\; $ Vi,$\; $ Ch $\}$
+\item $\xBBB = \{$ Su,$\; $ Pi,$\; $ Wo $\}$
+\item $\xTTT = \{$ Vi,$\; $ Li,$\; $ Ch,$\; $ Fr $\}$
 \item $n = |U| = 11$, $m = |\cF| = 4$
@@
   \centering
-  \includegraphics[scale=0.5]{../img/2_infinite_loop.png}
+%  \includegraphics[scale=0.5]{../img/2_infinite_loop.png}
+  \includegraphics[scale=0.5]{../img/2_infinite_loop_BTWF.png}
   \caption{{\residenceblock} street map with study group routes
     allocated. Routes are color coded as follows: red for
-    \textcolor{red}{\bf L} group, blue for \textcolor{blue}{\bf G}
-    group, orange for \textcolor{YellowOrange}{\bf B} group, green for
-    \textcolor{green}{\bf T} group}
+    \textcolor{red}{\xLLL} group, blue for \textcolor{blue}{\xGGG}
+    group, orange for \textcolor{YellowOrange}{\xBBB} group, green for
+    \textcolor{green}{\xTTT} group}
   \label{fig:streetmappath}

### 2011-11-20 src/synopsis.tex
- made some formatting changes in intro - committing before making the major v0.2 review changes.

@@
   {\mysubmissionmonth}} 
-\department{\MakeUppercase {\mydept}}
+\department{\MakeUppercase{\mydept}}
 
@@
 \thispagestyle{empty}
-% \listoftables
-% \addcontentsline{toc}{chapter}{LIST OF TABLES}
-% \listoffigures
-% \addcontentsline{toc}{chapter}{LIST OF FIGURES}
+\listoftables
+\addcontentsline{toc}{chapter}{LIST OF TABLES}
+\listoffigures
+\addcontentsline{toc}{chapter}{LIST OF FIGURES}
 \end{singlespace}
@@
 
-\def \LLL {{\em ``{\bf B}rief Interviews with Hideous Men''}} 
-\def \GGG {{\em ``The String {\bf T}heory''}} 
-\def \BBB {{\em ``[{\bf W}]Rhetoric and the Math Melodrama''}}
-\def \TTT {{\em ``{\bf F}ate, Time, and Language: An Essay on Free Will''}}
-
-\def \xLLL {{\bf B}}
-\def \xGGG {{\bf T}} 
-\def \xBBB {{\bf W}}
-\def \xTTT {{\bf F}}
+% \def \LLL {{\em ``{\bf B}rief Interviews with Hideous Men''}} 
+% \def \GGG {{\em ``The String {\bf T}heory''}} 
+% \def \BBB {{\em ``[{\bf W}]Rhetoric and the Math Melodrama''}}
+% \def \TTT {{\em ``{\bf F}ate, Time, and Language: An Essay on Free Will''}}
+\def \LLL {{\em ``$\mathbb{B}$rief Interviews with Hideous Men''}}
+\def \GGG {{\em ``The String $\mathbb{T}$heory''}} 
+\def \BBB {{\em ``[$\mathbb{W}$]Rhetoric and the Math Melodrama''}}
+\def \TTT {{\em ``$\mathbb{F}$ate, Time, and Language: An Essay on Free Will''}}
+
+% \def \xLLL {{\bf B}}
+% \def \xGGG {{\bf T}} 
+% \def \xBBB {{\bf W}}
+% \def \xTTT {{\bf F}}
+\def \xLLL {\mathbb{B}}
+\def \xGGG {\mathbb{T}} 
+\def \xBBB {\mathbb{W}}
+\def \xTTT {\mathbb{F}}
 
@@
 
+\def \xPa {{\bf Pa}} 
+\def \xPi {{\bf Pi}} 
+\def \xSn {{\bf Sn}}
+\def \xWo {{\bf Wo}}
+\def \xVi {{\bf Vi}} 
+\def \xLi {{\bf Li}} 
+\def \xCh {{\bf Ch}}
+\def \xSa {{\bf Sa}}
+\def \xFr {{\bf Fr}}
+\def \xSc {{\bf Sc}} 
+\def \xLu {{\bf Lu}}
+
 \def \residenceblock {{\em Infinite Loop}}
@@
 \def \WSI {{\em Wallace Studies Institute}}
+
+\def \coneohone {{\em ``Influence of post modernism in Wallace's work''}}
+\def \coneohtwo {{\em ``A study on fragmented prose method''}}
 
@@
 \tnote{use other font for elements of sets}
-\begin{enumerate}[]
-\item {$U = \{$Pa,$\; $Pi,$\; $Sn,$\; $Wo,$\; $Vi,$\; $Li,$\; $Ch,$\;
-    $Sa,$\; $Fr,$\; $Sc,$\; $Lu$\}$}
-\item $\cF = \{\xLLL, \xGGG, \xBBB, \xTTT\}$
-\item $\xLLL = \{$ Ch,$\; $ Sa,$\; $ Fr,$\; $ Sc,$\; $ Lu $\}$
-\item $\xGGG = \{$ Pa,$\; $ Pi,$\; $ Vi,$\; $ Ch $\}$
-\item $\xBBB = \{$ Su,$\; $ Pi,$\; $ Wo $\}$
-\item $\xTTT = \{$ Vi,$\; $ Li,$\; $ Ch,$\; $ Fr $\}$
-\item $n = |U| = 11$, $m = |\cF| = 4$
-\end{enumerate}
+
+% \vspace{-3mm}
+% \begin{align*}
+% U &= \{\xPa,\; \xPi,\; \xSn,\; \xWo,\; \xVi,\; \xLi,\; \xCh,\;
+%     \xSa,\; \xFr,\; \xSc,\; \xLu\}\\
+% \cF &= \{\xLLL, \xGGG, \xBBB, \xTTT\}\\
+% \xLLL &= \{\xCh,\;  \xSa,\;  \xFr,\;  \xSc,\;  \xLu \}\\
+% \xGGG &= \{\xPa,\;  \xPi,\;  \xVi,\;  \xCh \}\\
+% \xBBB &= \{\xSn,\;  \xPi,\;  \xWo \}\\
+% \xTTT &= \{\xVi,\;  \xLi,\;  \xCh,\;  \xFr \}\\
+% n &= |U| = 11\\
+% m &= |\cF| = 4  
+% \end{align*}
+
+\begin{table}[htbp]
+  \centering
+  \begin{tabular}{ll}
+    $U $&$= \{\xPa,\; \xPi,\; \xSn,\; \xWo,\; \xVi,\; \xLi,\; \xCh,\;
+    \xSa,\; \xFr,\; \xSc,\; \xLu\}$\\
+    $\cF $&$= \{\xLLL, \xGGG, \xBBB, \xTTT\}$\\
+    $\xLLL $&$= \{\xCh,\;  \xSa,\;  \xFr,\;  \xSc,\;  \xLu \}$\\
+    $\xGGG $&$= \{\xPa,\;  \xPi,\;  \xVi,\;  \xCh \}$\\
+    $\xBBB $&$= \{\xSn,\;  \xPi,\;  \xWo \}$\\
+    $\xTTT $&$= \{\xVi,\;  \xLi,\;  \xCh,\;  \xFr \}$\\
+    $n $&$= |U| = 11$\\
+    $m $&$= |\cF| = 4$      
+  \end{tabular}
+  \caption{Students and study groups in \WSI}
+  \label{tab:wsigroups}
+\end{table}
 
@@
 routes to study groups is illustrated in Fig~\ref{fig:streetmappath}
-and the students' allocation of apartments that obey's this route
+and the students' allocation of apartments that obeys this route
 allocation is shown in Fig~\ref{fig:streetmappathpeople}. How this is
@@
 
-\begin{figure}[h]
+\begin{figure}[htbp]
   \centering
@@
     allocated. Routes are color coded as follows: red for
-    \textcolor{red}{\xLLL} group, blue for \textcolor{blue}{\xGGG}
-    group, orange for \textcolor{YellowOrange}{\xBBB} group, green for
-    \textcolor{green}{\xTTT} group}
+    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
+    group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
+    \textcolor{green}{$\xTTT$} group}
   \label{fig:streetmappath}
@@
 
-If all the apartments are on the same street or if they are all lined
-up on a single path, the street map becomes a tree that is just a
-path. Then the problem becomes what is called an {\em interval
-  assignment problem}. The idea of interval assignment may not be
-obvious here; hence to see this, consider a different problem in
-{\WSI} where the classes for these study groups courses need to be
-scheduled during a day (or a week or any time period). Each study
-group has a bunch of courses associated with it some of which may be
-shared by two or more study groups. It is mandatory that a student who
-is a member of a study group takes all the courses associated with
-that group. There are slots during the day for classes to be held and
-the problem is to allocate class slots to courses such that all the
-classes of a study group are consecutive. It is debatable if this will
-not hamper the attention span and memory retention rate of the students but
-that is, regrettably, out of the scope of this thesis. The parallels between this
-class allocation problem and the accommodation problem can be seen as
-follows. The set $U$ here, are the courses offered (say Course 101 ``Post
-modernism in Wallace's work'', Course 102 ``A study on fragmented prose
-method'' and so on). The collection $\cF$ is the set of study groups but the
-study groups are filled by course IDs and not students like in the
-earlier example. For instance, Course 101 is mandatory for all study
-groups {\bf L, G, B, T} and Course 102 is mandatory for only the {\bf B}
-group) and so on. The sequence of class slots for the day (or week or
-any time period) is analogous to the street map in the accommodation
-problem. It is quite obvious now why this version of the problem
-(where the ``target graph'' is a path and not any tree) is called an
-interval assignment problem.
+As a special case, suppose all the apartments are on the same
+street or if they are all lined up on a single path, the street map
+becomes a tree that is just a path. Then the problem becomes what is
+called an {\em interval assignment problem}. The idea of interval
+assignment may not be obvious here; hence to see this, consider a
+different problem in {\WSI} where the classes for these study groups
+courses need to be scheduled during a day (or a week or any time
+period). Each study group has a bunch of courses associated with it
+some of which may be shared by two or more study groups. It is
+mandatory that a student who is a member of a study group takes all
+the courses associated with that group. There are slots during the day
+for classes to be held and the problem is to allocate class slots to
+courses such that all the classes of a study group are consecutive. It
+is debatable if this will not hamper the attention span and memory
+retention rate of the students but that is, regrettably, out of the
+scope of this thesis. The parallels between this class allocation
+problem and the accommodation problem can be seen as follows. The set
+$U$ here, are the courses offered (say Course 101 {\coneohone}, Course
+102 {\coneohtwo} and so on). In this variation of the problem, the
+collection $\cF$ is the set of study groups but the study groups are
+filled by course IDs (in place of students in the earlier
+example). For instance, Course 101 is mandatory for all study groups
+$\xLLL$, $\xGGG$, $\xBBB$, $\xTTT$ and Course 102 is mandatory for
+only the $\xLLL$ group) and so on. The sequence of class slots for the
+day (or week or any time period) is analogous to the street map in the
+accommodation problem. It is quite obvious now why this version of the
+problem (where the ``target graph'' is a path and not any
+tree\tnote{\footnote{Allowing any tree in this example could be seen
+    as a scenario where there are parallel classes. A node falling in
+    the path between two other nodes would mean that the corresponding
+    is scheduled between the other two.}}) is called an interval
+assignment problem.
 
@@
 
-% \begin{figure}[h] 
+% \begin{figure}[htbp] 
 %   \centering

### 2011-11-20 src/synopsis.tex
- corrected the synopsis template usage. - deleted chapters. made sections - basic org done. - the main changes and final! changes are pending.

@@
-\documentclass[MS,]{iitmdiss} 
+\documentclass[MS,synopsis]{iitmdiss} 
 \usepackage[usenames,dvipsnames]{color}
@@
 
-\nocite{*}
+%\nocite{*}
 \maketitle
 
-\begin{singlespace}
-\tableofcontents
-\thispagestyle{empty}
-\listoftables
-\addcontentsline{toc}{chapter}{LIST OF TABLES}
-\listoffigures
-\addcontentsline{toc}{chapter}{LIST OF FIGURES}
-\end{singlespace}
+% \tnote{TBD: DELETE IF WE RUN OUT OF SPACE}
+% \begin{singlespace}
+% \tableofcontents
+% \thispagestyle{empty}
+% \listoffigures
+% \addcontentsline{toc}{chapter}{LIST OF FIGURES}
+% \listoftables
+% \addcontentsline{toc}{chapter}{LIST OF TABLES}
+% \end{singlespace}
 
@@
 
-\chapter{Introduction}
-\label{chap:intro}
-
-\tnote{give some basic gyan about the problem itself? or screw it move to
-section directly}
-
-\section{Illustration of the problem}
+\section{Introduction}
+\label{sec:intro}
+
+% \tnote{testing 1 2 3...}
+% \tnote[bogus1]{testing 1 2 3...}
+% \tnote[bogus2]{testing 1 2 3...}
+% \tnote[bogus3]{testing 1 2 3...}
+% \tnote[bogus4]{testing 1 2 3...}
+% \tnote[bogus5]{testing 1 2 3...}
+% \tnote[bogus6]{testing 1 2 3...}
+% \tnote[bogus7]{testing 1 2 3...}
+% \tnote[bogus8]{testing 1 2 3...}
+% \tnote[TH]{testing 1 2 3...}
+
+\subsection{Illustration of the problem}
 \label{sec:problem}
@@
 number students and study groups respectively. In relation to this
-example, these are defined as follows. Also given below is the study group
-allocation to students. \tnote{use tables}
-\tnote{use other font for elements of sets}
+example, these are defined in Table~\ref{tab:wsigroups}. Also given
+there is the study group allocation to students.
 
@@
   \centering
-  \begin{tabular}{ll}
-    $U $&$= \{\xPa,\; \xPi,\; \xSn,\; \xWo,\; \xVi,\; \xLi,\; \xCh,\;
+  \begin{tabular}{rcl}
+    $U $&$=$&$ \{\xPa,\; \xPi,\; \xSn,\; \xWo,\; \xVi,\; \xLi,\; \xCh,\;
     \xSa,\; \xFr,\; \xSc,\; \xLu\}$\\
-    $\cF $&$= \{\xLLL, \xGGG, \xBBB, \xTTT\}$\\
-    $\xLLL $&$= \{\xCh,\;  \xSa,\;  \xFr,\;  \xSc,\;  \xLu \}$\\
-    $\xGGG $&$= \{\xPa,\;  \xPi,\;  \xVi,\;  \xCh \}$\\
-    $\xBBB $&$= \{\xSn,\;  \xPi,\;  \xWo \}$\\
-    $\xTTT $&$= \{\xVi,\;  \xLi,\;  \xCh,\;  \xFr \}$\\
-    $n $&$= |U| = 11$\\
-    $m $&$= |\cF| = 4$      
+    $\cF $&$=$&$ \{\xLLL, \xGGG, \xBBB, \xTTT\}$\\
+    $\xLLL $&$=$&$ \{\xCh,\;  \xSa,\;  \xFr,\;  \xSc,\;  \xLu \}$\\
+    $\xGGG $&$=$&$ \{\xPa,\;  \xPi,\;  \xVi,\;  \xCh \}$\\
+    $\xBBB $&$=$&$ \{\xSn,\;  \xPi,\;  \xWo \}$\\
+    $\xTTT $&$=$&$ \{\xVi,\;  \xLi,\;  \xCh,\;  \xFr \}$\\
+    $n $&$=$&$ |U| = 11$\\
+    $m $&$=$&$ |\cF| = 4$      
   \end{tabular}
@@
   \centering
-%  \includegraphics[scale=0.5]{../img/2_infinite_loop.png}
-  \includegraphics[scale=0.5]{../img/2_infinite_loop_BTWF.png}
+  \includegraphics[scale=0.3]{../img/2_infinite_loop_BTWF.png}
   \caption{{\residenceblock} street map with study group routes
@@
 
-As a special case, suppose all the apartments are on the same
-street or if they are all lined up on a single path, the street map
-becomes a tree that is just a path. Then the problem becomes what is
-called an {\em interval assignment problem}. The idea of interval
-assignment may not be obvious here; hence to see this, consider a
-different problem in {\WSI} where the classes for these study groups
-courses need to be scheduled during a day (or a week or any time
-period). Each study group has a bunch of courses associated with it
-some of which may be shared by two or more study groups. It is
-mandatory that a student who is a member of a study group takes all
-the courses associated with that group. There are slots during the day
-for classes to be held and the problem is to allocate class slots to
-courses such that all the classes of a study group are consecutive. It
-is debatable if this will not hamper the attention span and memory
-retention rate of the students but that is, regrettably, out of the
-scope of this thesis. The parallels between this class allocation
-problem and the accommodation problem can be seen as follows. The set
-$U$ here, are the courses offered (say Course 101 {\coneohone}, Course
-102 {\coneohtwo} and so on). In this variation of the problem, the
+As a special case, suppose all the apartments are on the same street
+or if they are all lined up on a single path, the street map becomes a
+tree that is just a path. Then the problem becomes what is called an
+{\em interval assignment problem}. The idea of interval assignment may
+not be obvious here; hence to see this, consider a different problem
+in {\WSI} where the classes for these study groups courses need to be
+scheduled during a day (or a week or any time period). Each study
+group has a bunch of courses associated with it some of which may be
+shared by two or more study groups. It is mandatory that a student who
+is a member of a study group takes all the courses associated with
+that group. There are slots during the day for classes to be held and
+the problem is to allocate class slots to courses such that all the
+classes of a study group are consecutive. It is debatable if this will
+not hamper the attention span and memory retention rate of the
+students but that is, regrettably, out of the scope of this
+thesis. The parallels between this class allocation problem and the
+accommodation problem can be seen as follows. The set $U$ here, are
+the courses offered (say Course 101 {\coneohone}, Course 102
+{\coneohtwo} and so on). In this variation of the problem, the
 collection $\cF$ is the set of study groups but the study groups are
@@
 problem (where the ``target graph'' is a path and not any
-tree\tnote{\footnote{Allowing any tree in this example could be seen
-    as a scenario where there are parallel classes. A node falling in
-    the path between two other nodes would mean that the corresponding
-    is scheduled between the other two.}}) is called an interval
-assignment problem.
-
-\tnote{TBD: BEGIN REVIEW FROM BELOW}
+tree\tnote[TH]{Allowing any tree in this example could be seen as a
+  scenario where there are parallel classes. A node falling in the
+  path between two other nodes would mean that the corresponding is
+  scheduled between the other two.}) is called an interval assignment
+problem.
 
@@
 
-\tnote[a]{THESIS: UNCOMMENT} \tnote{include the rest of the sections as
-  the organization of the doc}
+\tnote{add a note on the organization of the doc}
 % The necessary preliminaries with definitions etc. are presented in
@@
 Section~\ref{sec:motivation} gives a brief survey of COP and relatives.
-Chapter~\ref{chap:results} presents a summary of our results on the
+Section~\ref{sec:results} presents a summary of our results on the
 tree path labeling problem.
@@
 
-As descirbed earlier, a matrix with COP is one whose rows (columns)
+As described earlier, a matrix with COP is one whose rows (columns)
 can be rearranged so that the 1s in every column (row) are in
@@
 (there could be multiple) of the matrix. This is constructed using an
-elaborate linear time procedure. Since this
-result, several improvements have been made to the procedural
-complexity of the algorithm itself and also derivative data structures
-that are easier to construct. PQR trees is one such data structure
-\cite{mm96}, \cite{mpt98}\tnote{also, mpt98 $==$ elaborate version of
-  mm96}.  \cite{tm05} describes an improved version of PQR
-trees. \tnote[a]{improvement in terms of what?}\cite{wlh02} describes
-the simpler algorithm for COT. \tnote[a]{This result first apprearted
-  in proc ISAAC 92} Hsu also invented PC trees \cite{wlh01} which is
-claimed to be much easier to implement. \cite{nsnrs09} describes a
-characterization of consecutive ones property solely based on the
-cardinality properties of the set representations of the columns
-(rows); every column (row) is equivalent to a set that has the row
-(column) indices of the rows (columns) that have one entries in this
-column (row). This is interesting and relevant, especially to this
-thesis because it simplifies COT to a great degree. \tnote[a]{because
-  it reduces the search space for the solution from set of all
-  permutations of rows to set of all-- get this figured out!}
+elaborate linear time procedure. Since this result, several
+improvements have been made to the procedural complexity of the
+algorithm itself and also derivative data structures that are easier
+to construct. PQR trees is one such data structure \cite{mm96},
+\cite{mpt98}\tnote[TH]{also, mpt98 $==$ elaborate version of mm96}.
+\cite{tm05} describes an improved version of PQR
+trees. \tnote{improv in terms of what?}\cite{wlh02} describes the
+simpler algorithm for COT. Hsu also invented PC trees
+\cite{wlh01}\tnote[TH]{This result first appeared inproc ISAAC92}
+which is claimed to be much easier to implement. \cite{nsnrs09}
+describes a characterization of consecutive ones property solely based
+on the cardinality properties of the set representations of the
+columns (rows); every column (row) is equivalent to a set that has the
+row (column) indices of the rows (columns) that have one entries in
+this column (row). This is interesting and relevant, especially to
+this thesis because it simplifies COT to a great
+degree. \tnote{get this
+  figured out! - because it reduces the search space for the solution
+  from set of all permutations of rows to set of all}
 
@@
 order is the certificate.  It may also be noted that the internal
-machinery of this algorithm \annote[a] {is related to the weighted
+machinery of this algorithm \annote {is related to the weighted
   betweenness problem addressed } {in what way??} in \cite{co98}.
-\tnote[a]{expand on the COP order graph creation and it having to be
+\tnote{expand on the COP order graph creation and it having to be
   bipartite for M to have COP. and thus an odd cycle being an evidence
   of no COP.}
-
-\tnote[a]{THESIS: where should this go?: \cite{jlm97} (application of
-  PQ trees in graphics).}
+\tnote[TH]{ where should this go?: cite|jlm97 (application of PQ trees
+  in graphics).\\ add helly's theorem citation
+  19XXdgk-Hellystheorem-Danzer-Gruenbaum-Klee}
 
 The literature also has results on parallel COT algorithms --
-\cite{as95} \cite{bs03} \tnote[a]{is this parallel?}
-\cite{ly91}\tnote[a]{is this parallel?}
-
-\tnote[a]{THESIS: add helly's theorem citation
-  19XXdgk-Hellystheorem-Danzer-Gruenbaum-Klee}
-
-{\bf Matrices with no COP}: So far we have been concerned about matrices that have the consecutive
-ones property. However in real life applications, it is rare that data
-sets represented by binary matrices have COP. Nevertheless, the COP is
-not arbirary and there are several interesting problems when a matrix
-does not have COP but is ``close'' to having COP or can be altered to
-have COP.
+\cite{as95} \cite{bs03} \cite{ly91}\tnote{add english to fit context}
+
+\subsection{Optimization problems in COP}
+So far we have been concerned about
+matrices that have the consecutive ones property. However in real life
+applications, it is rare that data sets represented by binary matrices
+have COP. Nevertheless, the COP is not arbirary and there are several
+interesting problems when a matrix does not have COP but is ``close''
+to having COP or can be altered to have COP.
 
@@
 where number columns and rows are restriced to $a$ and $b$
-\tnote[a]{check if it is the other way around}.\\
+\tnote{check if it is the other way around}.\\
 -- problem of flipping at most $k$ entries of $M$ to make it attain
-COP. this is NP complete \cite{b75-phd}\\
+COP. this is NP complete \cite{b75-phd}
 
@@
 submatrices. This result is presented in the context of AT-free
-bipartite graphs \tnote[a]{THESIS: check this up. give details.} which
-is equivalent to matrices with COP. A binary matrix is converted into
-a bipartite graph as follows. The set of rows and the set of columns
+bipartite graphs \tnote[TH]{ check this up. give details.} which is
+equivalent to matrices with COP. A binary matrix is converted into a
+bipartite graph as follows. The set of rows and the set of columns
 form the two partitions of the graph. Each row node is adjacent to
@@
 any of the COT algorithms mentioned earlier), it is naturally of
-interest to find out the smallest\tnote[a]{define ``smallest''}
-forbidden substructure. \cite{d08phd} discusses an exact algorithm
-with time complexity \annote[a]{parameterized by the number of ones in
-  the matrix}{check this.}.
+interest to find out the smallest (in terms of number of rows, columns
+and number of entries that are ones) forbidden
+substructure. \cite{d08phd} discusses an exact algorithm with time
+complexity \annote{parameterized by the number of ones in the
+  matrix}{check this.}.
 
@@
 but a maxmimum cardinality subset of rows such that $M'$ has COP is NP
-complete. \tnote[a]{check if b75 deals with COP col or COP row. also
+complete\tnote{check if b75 deals with COP col or COP row. also
   is it any submatrix with k less than r rows or submatrix must have
@@
 \cite{d08phd} presents new results in the area of parameterized
-algorithms for this problem. \tnote[a]{THESIS: elaborate - what are
+algorithms for this problem\tnote[TH]{elaborate - what are
   the results?}.
@@
 of changing of the minimum number of zero entries to ones so that the
-resulting matrix has COP. As mentioned earlier, this problem had been known
-to be NP complete due to \cite{b75-phd}. \cite{v85} also proves, using
-a reduction to the longest path problem, \tnote[a]{or is it a survey
-  of another result?  check.} that finding a Tucker's forbidden
-submatrix of at least $k$
-rows is NP complete. \tnote[a]{how is this different from booth's 75 result??} \tnote{[tag:appoximation,bitFlipping,matrixBreaking]}\\
-
-\tnote[a]{THESIS: where should this go? \cite{tz04} (approx submatrix with COP sparse matrices)}
-
-%\subsection{Further extensions of the COP problem}
-
-\cite{jkckv04} \tnote{[tag:storage]} discusses the use of matrices
-with ``near'' COP in the storage of very large databases.
-The problem is that of $k$-run reordering of binary matrices which
-they prove to be NP hard using a reduction from the hamiltonian path problem.
-A connection of COP problem to the travelling salesman problem is also
-introduced. \tnote[a]{THESIS? what does this mean? -- COP can be used as a tool to reorder $0.5T \le runs(M) \le T$}
+resulting matrix has COP. As mentioned earlier, this problem had been
+known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
+using a reduction to the longest path problem, \tnote{or is it a
+  survey of another result?  check.} that finding a Tucker's forbidden
+submatrix of at least $k$ rows is NP complete. \tnote{how is this
+  different from booth's 75 result??}  \tnote[TH]{where should this
+  go? cite|tz04 (approx submatrix with COP sparse matrices)}
+
+\cite{jkckv04} discusses the use of matrices with almost-COP in the
+storage of very large databases.  The problem is that of $k$-run
+reordering of binary matrices which they prove to be NP hard using a
+reduction from the hamiltonian path problem.  A connection of COP
+problem to the travelling salesman problem is also
+introduced. \tnote[TH]{what does this mean? -- COP can be used as a
+  tool to reorder $0.5T \le runs(M) \le T$}
 
@@
 number of blocks of ones is proven to be NP complete by
-\cite{k77}. \tnote[a]{THESIS: what is the reduction?}
-
-\tnote[a]{THESIS: 
--- {\em other problems similar to COP}
-  \begin{enumerate}
-  \item \cite{ckl96} (ILP, circ ones, one drop)
-  \item \cite{th98} (generalization of COP - minimax, biotonic column) Tucker
-  \end{enumerate}
-}
+\cite{k77}. \tnote[TH]{what is the reduction?}
+
+\tnote[TH]{other problems similar to COP -- cite:ckl96 (ILP, circ ones, one drop) -- cite:th98 (generalization of COP - minimax, biotonic column) Tucker}
 
@@
 possible bijections between $U$ and $V(T)$ to only those bijections
-that maintain the characterization\tnote[a]{REWORD}.  Further, the
+that maintain the characterization\tnote{REWORD}.  Further, the
 filtering algorithm is also an efficient algorithm to test if a tree
@@
 of \cite{gav78,aas93}.  However, this is only a necessary condition
-for our question.  \remove[a]{Indeed, our work provides a sufficient
+for our question.  \remove{Indeed, our work provides a sufficient
   condition for a related problem: that is whether a path assignment
@@
 be isomorphism-complete; see for example \cite{kklv10}. Therefore,
-\annote[a]{it is unlikely that we can solve the problem of finding
+\annote{it is unlikely that we can solve the problem of finding
   feasible path labeling $\cl$ for a given $\cF$ and }{WHY?} tree $T$.
-\annote[a]{It is definitely interesting to classify the kinds of trees
+\annote{It is definitely interesting to classify the kinds of trees
   and hypergraphs for which feasible path labelings can be
-}{ELABORATE.} found efficiently.  \annote[a]{These results would form
+}{ELABORATE.} found efficiently.  \annote{These results would form
   a natural generalization of COP testing and interval graph
@@
 
-\chapter{Summary of results}
-\label{chap:results}
+\section{Summary of results}
+\label{sec:results}
 
@@
 %\item 
-In the Section 4, \annote[a]{we initiate an exploration of}{WTF?}
+In the Section 4, \annote{we initiate an exploration of}{WTF?}
 finding feasible path labeling of set systems in a special kind of
-tree which we call the $k$-subdivided star.  \annote[a]{This question
+tree which we call the $k$-subdivided star.  \annote{This question
   is an attempt to generalize}{REWORD.} the problem of testing if a
@@
 hypergraph $\cF$ to be such that all hyperedges have at most $k+2$
-elements.  \annote[a]{In spite of this restricted case we consider, we
+elements.  \annote{In spite of this restricted case we consider, we
   believe that our results are of significant interest in
@@
 
-\section{RESULTS}
+%\subsection{RESULTS}
 % \section{Preliminaries} 
@@
 % % both the structures by $\cF$.
-% % \vspace{\prelimspace}%\tnote[a]{also extend $\phi$ to hyperedges -- see if
+% % \vspace{\prelimspace}%\tnote{also extend $\phi$ to hyperedges -- see if
 % % required } \vspace{\prelimspace} \vspace{\prelimspace}
@@
 
-\section{Characterization of Feasible Tree Path  Labelings} 
+\subsection{Characterization of Feasible Tree Path  Labelings} \tnote[TH]{UNCOMMENT all proofs}
 \label{sec:feasible}
+
 In this section we give an algorithmic characterization of a
@@
 
-\begin{enumerate}[{\icpplpr\ }i. ]
-\item \label{pr:i} $|S| = |\cl(S)|$, \hfill for all $S \in \cF$
-  % \vspace{\topshrink}
-\item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$, \hfill
-  for all distinct $S_1, S_2 \in \cF$
-  % \vspace{\topshrink}
-\item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
-  \cap \cl(S_3)|$, \hfill for all distinct $S_1, S_2, S_3 \in \cF$
-\end{enumerate}
+% \begin{enumerate}[{\icpplpr\ }i. ]
+% \item \label{pr:i} $|S| = |\cl(S)|$, \hfill for all $S \in \cF$
+%   % \vspace{\topshrink}
+% \item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$, \hfill
+%   for all distinct $S_1, S_2 \in \cF$
+%   % \vspace{\topshrink}
+% \item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
+%   \cap \cl(S_3)|$, \hfill for all distinct $S_1, S_2, S_3 \in \cF$
+% \end{enumerate}
+
+\begin{table}[h]
+  \centering
+  \begin{tabular}[l]{rrcll}
+    Pr i. & \label{xpr:i} $|S| $&$=$&$ |\cl(S)|$ & $\forall$ $S \in \cF$\\
+    Pr ii. & \label{xpr:ii}$|S_1 \cap S_2| $&$=$&$ |\cl(S_1) \cap
+    \cl(S_2)|$ &  $\forall$ distinct $S_1, S_2 \in \cF$\\
+    Pr iii. & \label{xpr:iii}$|S_1 \cap S_2 \cap S_3| $&$=$&$
+    |\cl(S_1) \cap \cl(S_2) \cap \cl(S_3)|$ & $\forall$ distinct $S_1,
+    S_2, S_3 \in \cF$ 
+  \end{tabular}
+  \caption{ICPPL properties}
+  \label{tab:icpplpr}
+\end{table}
 
@@
 \end{lemma}
-\tnote[a]{THESIS: UNCOMMENT}
+
 % \begin{proof}%[Proof of Lemma~\ref{lem:setminuscard}]
@@
 % preserved, i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
-% \annote[a]{has no premature exit condition hence any input will go
+% \annote{has no premature exit condition hence any input will go
 %   through it}{Prove that the filtered sets has ICPPL \iff input PL
@@
 % feasible path labeling has an associated bijection between
-% $supp(\cF)$ and $V(T)$ \remove[a]{i.e. $supp(\cF^{\cl})$} such that
+% $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that
 % the sets map to paths, ``preserving'' the path labeling.  The rest
@@
 % a path labeling that defines a family of bijections from $supp(\cF)$
-% to $V(T)$\remove[a]{ i.e. $supp(\cF^{\cl})$}.
+% to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.
 
 First, we present Algorithm~\ref{perms} called {\tt filter\_1}
-\tnote[a]{THESIS: UNCOMMENT}
 % , and
@@
 paths in the new path labeling share any leaf.
-
-\begin{algorithm}[h]
-  \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
-  \label{perms}
-  \begin{algorithmic}[\lndisplay]
-    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$\\
-    \STATE $j \assign 1$\\
-    \WHILE {there is $S_1, S_2 \in \cF_{j-1}$ such that
-      $\cl_{j-1}(S_1)$ and $\cl_{j-1}(S_2)$ have a common leaf in
-      $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
-    \{S_1, S_2\})
-    \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1 \}$ \label{setbreak} 
-    \COMMENT {Remove $S_1$, $S_2$ and add the ``filtered'' sets}
-    \STATE {\bf for} every $S \in \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
-    S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\
-
-    \STATE $\cl_j(S_1 \cap S_2) \assign \cl_{j-1}(S_1) \cap
-    \cl_{j-1}(S_2)$
-    \COMMENT {Carry forward the path labeling for all existing sets other than
-      $S_1$, $S_2$}
-    \STATE $\cl_j(S_1 \setminus S_2) \assign \cl_{j-1}(S_1) \setminus
-    \cl_{j-1}(S_2)$ 
-    \COMMENT {Define path labeling for new sets}
-    \STATE $\cl_j(S_2 \setminus S_1) \assign \cl_{j-1}(S_2) \setminus
-    \cl_{j-1}(S_1)$
-
-    \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of ICPPL}
-    \label{ln:3waycheck}
-    \STATE {\bf exit} \label{ln:exit1} \\
-    \ENDIF
-
-    \STATE $j \assign j+1$\\
-    \ENDWHILE
-    \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
-    \RETURN $(\cF', \cl')$
-  \end{algorithmic}
-\end{algorithm}
+\tnote{write a note summarizing filter 1}
+% \begin{algorithm}[h]
+%   \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
+%   \label{perms}
+%   \begin{algorithmic}[\lndisplay]
+%     \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$\\
+%     \STATE $j \assign 1$\\
+%     \WHILE {there is $S_1, S_2 \in \cF_{j-1}$ such that
+%       $\cl_{j-1}(S_1)$ and $\cl_{j-1}(S_2)$ have a common leaf in
+%       $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
+%     \{S_1, S_2\})
+%     \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1 \}$ \label{setbreak} 
+%     \COMMENT {Remove $S_1$, $S_2$ and add the ``filtered'' sets}
+%     \STATE {\bf for} every $S \in \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
+%     S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\
+
+%     \STATE $\cl_j(S_1 \cap S_2) \assign \cl_{j-1}(S_1) \cap
+%     \cl_{j-1}(S_2)$
+%     \COMMENT {Carry forward the path labeling for all existing sets other than
+%       $S_1$, $S_2$}
+%     \STATE $\cl_j(S_1 \setminus S_2) \assign \cl_{j-1}(S_1) \setminus
+%     \cl_{j-1}(S_2)$ 
+%     \COMMENT {Define path labeling for new sets}
+%     \STATE $\cl_j(S_2 \setminus S_1) \assign \cl_{j-1}(S_2) \setminus
+%     \cl_{j-1}(S_1)$
+
+%     \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of ICPPL}
+%     \label{ln:3waycheck}
+%     \STATE {\bf exit} \label{ln:exit1} \\
+%     \ENDIF
+
+%     \STATE $j \assign j+1$\\
+%     \ENDWHILE
+%     \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
+%     \RETURN $(\cF', \cl')$
+%   \end{algorithmic}
+% \end{algorithm}
 
@@
 \end{lemma}
-\tnote[a]{THESIS: UNCOMMENT}
 
@@
   invariants are maintained.
-  \begin{enumerate}[I {\ }] %\vspace{\topshrink}
-  \item $\cl_j(R)$ is a path in $T$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ for all $R \in \cF_j$%\vspace{\topshrink}
-  \item $|R| = |\cl_j(R)|$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $R \in
-    \cF_j$%\vspace{\topshrink}
-  \item $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ for all $R, R' \in \cF_j$%\vspace{\topshrink}
-  \item $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
-    \cl_j(R'')|$, \ \ \ for all $R, R', R'' \in \cF_j$
-  \end{enumerate}
+%   \begin{enumerate}[I {\ }] %\vspace{\topshrink}
+%   \item $\cl_j(R)$ is a path in $T$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
+%     \ \ \ \ \ \ \ \ \ \ for all $R \in \cF_j$%\vspace{\topshrink}
+%   \item $|R| = |\cl_j(R)|$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
+%     \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $R \in
+%     \cF_j$%\vspace{\topshrink}
+%   \item $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, \ \ \ \ \ \ \ \ \ \
+%     \ \ \ \ \ \ \ \ \ \ for all $R, R' \in \cF_j$%\vspace{\topshrink}
+%   \item $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
+%     \cl_j(R'')|$, \ \ \ for all $R, R', R'' \in \cF_j$
+%   \end{enumerate}
+
+  \begin{tabular}[h]{lll}
+    I & $\cl_j(R)$ is a path in $T$, & for all $R \in \cF_j$\\ 
+    II  & $|R| = |\cl_j(R)|$, & for all $R \in \cF_j$ \\ 
+    III  & $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, & for all $R, R'
+    \in \cF_j$\\ 
+    IV  & $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
+    \cl_j(R'')|$, & for all $R, R', R'' \in \cF_j$ 
+  \end{tabular}
 \end{lemma}
-\tnote[a]{THESIS: UNCOMMENT}
 
@@
 \end{lemma}
-\tnote[a]{THESIS: UNCOMMENT}
+\tnote[TH]{UNCOMMENT}
 % \begin{proof}
@@
 
-\begin{algorithm}[h]
-  \caption{Leaf labeling from an ICPPL {\tt filter\_2($\cF, \cl, T$)}}
-  \label{leafasgn}
-  \begin{algorithmic}[\lndisplay]
-    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$
-    \COMMENT {Path images are such that no two path images share a
-      leaf.}
-    \STATE $j \assign 1$\\
-    \WHILE {there is a leaf $v$ in $T$ and a unique $S_1 \in
-      \cF_{j-1}$ such that $v \in \cl_{j-1}(S_1)$ }\label{uniqueleaf}
-    \STATE $\cF_j \assign \cF_{j-1} \setminus \{S_1\}$\\
-    \STATE for all $S \in \cF_{j-1}$ such that $S \ne S_1$ set
-    $\cl_j(S) \assign
-    \cl_{j-1}(S)$\\
-    \STATE $X \assign S_1 \setminus \bigcup_{S \in \cF_{j-1}, S \ne S_1}S$\\
-    \IF{$X$ is empty} \label{xempty} \STATE {\bf exit} \label{ln:exit2} \ENDIF
-    \STATE $x \assign $ arbitrary element from $X$\\
-    \STATE $\cF_j \assign \cF_j \cup \{\{x\}, S_1 \setminus \{x\}\} $\\
-    \STATE $\cl_j(\{x\}) \assign \{v\}$\\
-    \STATE $\cl_j(S_1 \setminus \{x\}) \assign \cl_{j-1}(S_1) \setminus \{v\}$\\
-    \STATE $j \assign j+1$\\
-    \ENDWHILE
-    \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
-    \RETURN $(\cF', \cl')$
-  \end{algorithmic}
-\end{algorithm}
+\tnote{write a note summarizing filter 1}
+% \begin{algorithm}[h]
+%   \caption{Leaf labeling from an ICPPL {\tt filter\_2($\cF, \cl, T$)}}
+%   \label{leafasgn}
+%   \begin{algorithmic}[\lndisplay]
+%     \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$
+%     \COMMENT {Path images are such that no two path images share a
+%       leaf.}
+%     \STATE $j \assign 1$\\
+%     \WHILE {there is a leaf $v$ in $T$ and a unique $S_1 \in
+%       \cF_{j-1}$ such that $v \in \cl_{j-1}(S_1)$ }\label{uniqueleaf}
+%     \STATE $\cF_j \assign \cF_{j-1} \setminus \{S_1\}$\\
+%     \STATE for all $S \in \cF_{j-1}$ such that $S \ne S_1$ set
+%     $\cl_j(S) \assign
+%     \cl_{j-1}(S)$\\
+%     \STATE $X \assign S_1 \setminus \bigcup_{S \in \cF_{j-1}, S \ne S_1}S$\\
+%     \IF{$X$ is empty} \label{xempty} \STATE {\bf exit} \label{ln:exit2} \ENDIF
+%     \STATE $x \assign $ arbitrary element from $X$\\
+%     \STATE $\cF_j \assign \cF_j \cup \{\{x\}, S_1 \setminus \{x\}\} $\\
+%     \STATE $\cl_j(\{x\}) \assign \{v\}$\\
+%     \STATE $\cl_j(S_1 \setminus \{x\}) \assign \cl_{j-1}(S_1) \setminus \{v\}$\\
+%     \STATE $j \assign j+1$\\
+%     \ENDWHILE
+%     \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
+%     \RETURN $(\cF', \cl')$
+%   \end{algorithmic}
+% \end{algorithm}
 
@@
 \end{lemma}
-\tnote[a]{THESIS: UNCOMMENT}
+
 % \begin{proof}
@@
 %   \label{lem:notfeasibleexit}
-%   \tnote[a]{IS THIS CORRECT?}  If input ICPPL $(\cF, \cl)$ is not
+%   \tnote{IS THIS CORRECT?}  If input ICPPL $(\cF, \cl)$ is not
 %   feasible, then in one of the recursive calls to Algorithm --3--, the
@@
 
-\begin{algorithm}[h]
-  \caption{{\tt get-hypergraph-isomorphism($\cF, \cl, T$)}}
-  \label{al:icppl-find-isomorph}
-  \begin{algorithmic}[\lndisplay]
-
-    \IF{$T$ is empty}
-    \RETURN $\emptyset$\\
-    \ENDIF
-    \STATE $L \assign \{v \mid v \text{ is a leaf in }      T\}$\\
-    \STATE $(\cF_1, \cl_1) \assign$ {\tt filter\_1($\cF, \cl,
-      T$)}\\
-    \STATE $(\cF_2, \cl_2) \assign$ {\tt filter\_2($\cF_1,
-      \cl_1, T$)}\\
-
-    \STATE $(\cF', \cl') \assign (\cF_2, \cl_2)$\\
-    \STATE $\phi' \leftarrow \emptyset$
-
-    \FOR {every $v \in L$}
-    \STATE $\phi'(x) \assign v$ where $x \in \cl_2^{-1}(\{v\})$
-    \COMMENT {Copy the leaf labels to a one to one function $\phi':
-      supp(\cF) \rightarrow L$
-      }\\
-    \STATE Remove $\{x\}$ and $\{v\}$ from $\cF'$, $\cl'$  appropriately\\
-    \ENDFOR
-
-    \STATE $T' \assign T \setminus L$
-
-    \STATE $\phi'' \assign$ {\tt
-      get-hypergraph-isomorphism($\cF', \cl', T'$)}
-    \STATE $\phi \assign \phi'' \cup \phi'$ \\
-    \RETURN $\phi$
-  \end{algorithmic}
-\end{algorithm}
+\tnote{write a note summarizing filter 1}
+% \begin{algorithm}[h]
+%   \caption{{\tt get-hypergraph-isomorphism($\cF, \cl, T$)}}
+%   \label{al:icppl-find-isomorph}
+%   \begin{algorithmic}[\lndisplay]
+
+%     \IF{$T$ is empty}
+%     \RETURN $\emptyset$\\
+%     \ENDIF
+%     \STATE $L \assign \{v \mid v \text{ is a leaf in }      T\}$\\
+%     \STATE $(\cF_1, \cl_1) \assign$ {\tt filter\_1($\cF, \cl,
+%       T$)}\\
+%     \STATE $(\cF_2, \cl_2) \assign$ {\tt filter\_2($\cF_1,
+%       \cl_1, T$)}\\
+
+%     \STATE $(\cF', \cl') \assign (\cF_2, \cl_2)$\\
+%     \STATE $\phi' \leftarrow \emptyset$
+
+%     \FOR {every $v \in L$}
+%     \STATE $\phi'(x) \assign v$ where $x \in \cl_2^{-1}(\{v\})$
+%     \COMMENT {Copy the leaf labels to a one to one function $\phi':
+%       supp(\cF) \rightarrow L$
+%       }\\
+%     \STATE Remove $\{x\}$ and $\{v\}$ from $\cF'$, $\cl'$  appropriately\\
+%     \ENDFOR
+
+%     \STATE $T' \assign T \setminus L$
+
+%     \STATE $\phi'' \assign$ {\tt
+%       get-hypergraph-isomorphism($\cF', \cl', T'$)}
+%     \STATE $\phi \assign \phi'' \cup \phi'$ \\
+%     \RETURN $\phi$
+%   \end{algorithmic}
+% \end{algorithm}
 
@@
   Algorithm~\ref{al:icppl-find-isomorph}, {\tt
-    get-hypergraph-isomorphism ($\cF, \cl, T$)} returns a non-empty
+    get- hypergraph- isomorphism ($\cF, \cl, T$)} returns a non-empty
   function, then there exists a hypergraph isomorphism $\phi :
@@
 \end{lemma}
-\tnote[a]{THESIS: UNCOMMENT}
 % \begin{proof}
@@
 \end{theorem}
-\tnote[a]{THESIS: UNCOMMENT}
 % \begin{proof}
@@
 \subsection*{ICPPL when given tree is a path}
-\label{subsec:icpplicpia}
+\label{sec:icpplicpia}
 Consider a special case of ICPPL with the following properties when
@@
 
-\section{Case of a Special Tree -- The $k$-subdivided star}
+\subsection{Case of a Special Tree -- The $k$-subdivided star}
 \label{sec:ksubdivstar}
@@
 
-\subsection{Description of the Algorithm}
+\subsubsection{Description of the Algorithm}
 In this section the given hypergraph $\cF$, the $k$-subdivided star
@@
 %\noindent 
-\tnote[a]{THESIS: UNCOMMENT}
+\tnote[TH]{UNCOMMENT}
 % \subsection*{Proof of Correctness and Analysis of Running Time} 
@@
 
-\chapter{Conclusion}
+\section{Conclusion}
 \tnote{MAKE CONNECTION TO PATH GRAPH ISOMORPHISM HERE}
@@
 results. %-- Other relations of COP with Graphs: \\
-\tnote[a]{THESIS: peo exists iff chordal. lexicographic BFS
+\tnote[TH]{peo exists iff chordal. lexicographic BFS
   [tag:chordalGraph]} 
@@
 all $v \in V(G)$, cliques containing $v$ are consecutive in the
+
 ordering\cite{gh64}. This clearly means that a graph $G$ is an
-interval graph if and only if \tnote[a]{verify from paper the
+interval graph if and only if \tnote{verify from paper the
   statement of claim.} maximal clique vertex incidence matrix of $G$
 has COP.  Also, maximal cliques of any chordal graph can be enumerated
-in polytime $O(m+n)$ \tnote[a]{citation?!!}.  \cite{fg65} uses these
+in polytime $O(m+n)$ \tnote{citation?!!}.  \cite{fg65} uses these
 results to give the first polynomial time algorithm for
-COT. \tnote[a]{THESIS: check. how do they use it?}  A bipartite graph
+COT. \tnote[TH]{check. how do they use it?}  A bipartite graph
 is convex (on $R$) if and only if its half adjacency matrix has COP on
@@
 on COT are based on the result that
-interval graphs are AT-free chordal graphs. \tnote[a]{the latter being
+interval graphs are AT-free chordal graphs. \tnote{the latter being
   Tucker's?}
 
-\tnote[a]{THESIS: -- How COP (and TPL) relates to interval graph
+\tnote[TH]{How COP (and TPL) relates to interval graph
   isomorphism (and path graph isomorphism) move/summarize what is
@@
 
-\tnote[a]{THESIS: TBD survey -- see blue notes under Graph
-  Isomorphism. namely citations in \cite{aas93}}
-
-\tnote[a]{THESIS: -- -- a brief on heirarchy of chordal graphs, path
+\tnote[TH]{TBD survey -- see blue notes under Graph
+  Isomorphism. namely citations in cite:aas93}
+
+\tnote[TH]{a brief on heirarchy of chordal graphs, path
   graphs, interval graphs, peo, clique tree etc. and results.}
 
-\tnote[a]{THESIS: -- interval graphs are incomparability graphs. see
+\tnote[TH]{interval graphs are incomparability graphs. see
   Golumbic. [tag:classification]}
 
-\tnote[a]{THESIS: finding min length hole in bipartite graph is
-  polynomial time [Sec 3.3.2 in \cite{d08phd}]}
-
-\tnote[a]{THESIS: theorem 2.2 in \cite{d08phd} - $G$ is union of
+\tnote[TH]{finding min length hole in bipartite graph is
+  polynomial time [Sec 3.3.2 in cite:d08phd]}
+
+\tnote[TH]{theorem 2.2 in cite:d08phd - $G$ is union of
   v.d. caterpillars iff $M$ has COP in rows. $M$ is edge vertex
@@
 trees using a logspace depth first traversal
-algorithm. \tnote[a]{confirm the statement of claim}.
+algorithm. \tnote{confirm the statement of claim}.
 
@@
 Interval graphs are FP+C (fixed point with
-counting) definable \tnote[a]{elucidate} and \cite{l10} showed that
-this implies that it captures PTIME. \tnote[a]{elucidate} This result
+counting) definable \tnote{elucidate} and \cite{l10} showed that
+this implies that it captures PTIME. \tnote{elucidate} This result
 along with that of undirected graph reachability being in logspace
@@
 
-\tnote[a]{THESIS: -- -- the interesting survey in conclusion section
+\tnote[TH]{the interesting survey in conclusion section
   of kklv10}
 
-\tnote[a]{THESIS:
-{\bf COP helping some problems with hardness}
-having COP in the input structure makes some problems less hard than
-in a general input.
-
--- min set cover problem is known to be NP hard. so is min hitting
-set. these two problems are equivalent \cite{adp80}\\
--- -- complexity results: log factor polynomial time approximable \cite{f98}, W[2]-complete with solution size as parameter \cite{df99}\\
+\tnote[TH]{[[[COP helping some problems with hardness]]] having
+  COP in the input structure makes some problems less hard than in a
+  general input. -- min set cover problem is known to be NP hard. so
+  is min hitting
+  set. these two problems are equivalent cite-adp80\\
+  -- -- complexity results: log factor polynomial time approximable cite-f98, W[2]-complete with solution size as parameter cite-df99\\
   -- a couple of generalizations:\\
@@
   -- -- red blue covering (RBSC) problem\\
+----- cite-d08phd Chapter 5.\\
+----- what are cite-d08phd results\\
+----- Rectangle stabbing \\
+----- see cite-d08phd Chapter 6?
 }
 
-\tnote[a]{THESIS: \cite{d08phd} Chapter 5.}
-
-\tnote[a]{THESIS: what are \cite{d08phd} results}
-
-\tnote[a]{THESIS: -- Rectangle stabbing \tnote[a]{see \cite{d08phd} Chapter 6?}}
-
+\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{singlespace}
+%\begin{thebibliography}{10}
+\addcontentsline{toc}{section}{Bibliography}
 
@@
   \bibliography{../lib/cop-variants__thesis}
-  \addcontentsline{toc}{chapter}{REFERENCES}
 }
+%\end{thebibliography}
 \end{singlespace}
+\newpage
 
@@
 
-\chapter*{Proposed Contents of the Thesis}
-\addcontentsline{toc}{chapter}{Proposed Contents of the Thesis}
+%\chapter*{Proposed Contents of the Thesis}
+\section{Proposed Contents of the Thesis}
+%\addcontentsline{toc}{chapter}{Proposed Contents of the Thesis}
 The outline of the thesis is as follows:

### 2011-11-21 src/synopsis.tex
- formatted the COP example figure. have text instead of the png from doms.

@@
 researched combinatorial problem and has several positive results on
-tests for it and computing the COP permutation (i.e. the class slot
-allocation to course IDs in the above illustration) which will be
-surveyed later in this document. Hence we are interested in extensions
-of COP, more specifically, the extension of interval assignment
-problem to tree path assignment problem (which is illustrated by the
-study group accommodation problem).
+tests for it and computing the COP permutation (i.e. the course
+schedule in the above illustration) which will be surveyed later in
+this document. Hence we are interested in extensions of COP, more
+specifically, the extension of interval assignment problem to tree
+path assignment problem (which is illustrated by the study group
+accommodation problem).
 
@@
 
-Section~\ref{sec:motivation} gives a brief survey of COP and relatives.
-Section~\ref{sec:results} presents a summary of our results on the
+Section~\ref{sec:motivation} gives a brief survey of COP and
+optimization problems related to it.  Section~\ref{sec:results}
+presents a summary of our results on the extention of COP namely, the
 tree path labeling problem.
@@
 
-As described earlier, a matrix with COP is one whose rows (columns)
+%The course scheduling problem illustated earlier, as a special case of
+
+It is quite obvious now why this version of the
+problem (where the ``target graph'' is a path and not any
+tree\tnote[TH]{Allowing any tree in this example could be seen as a
+  scenario where there are parallel classes. A node falling in the
+  path between two other nodes would mean that the corresponding is
+  scheduled between the other two.}) is called an interval assignment
+problem.
+
+The interval assignment problem to a set system is equivalent to the
+consecutive ones problem (COP) in binary matrices\cite{wlh02,
+  nsnrs09}.  The COP problem is to rearrange rows (columns) of a
+binary matrix in such a way that every column (row) has its $1$s occur
+consecutively. If this is possible the matrix is said to have the COP.
+Consecutive ones property (COP) of binary matrices is a well
+researched combinatorial problem and has several positive results on
+tests for it and computing the COP permutation (i.e. the course
+schedule in the above illustration) which will be surveyed later in
+this document. Hence we are interested in extensions of COP, more
+specifically, the extension of interval assignment problem to tree
+path assignment problem (which is illustrated by the study group
+accommodation problem).
+
+As described earlier, 
+a matrix with COP is one whose rows (columns)
 can be rearranged so that the 1s in every column (row) are in
@@
 
+\def \un {\bf 1}
 \begin{figure}[htbp]
   \centering
-  \includegraphics[scale=0.4]{../img/0_cop_nocop.png}
-
-  \caption{{\em Example for the COP:} The matrix on the left has the
-    COP because by permuting its columns (labeled with $c_1$-$c_4$)
-    one can obtain the matrix shown in the middle where the 1s in each
-    row appear consecutively. The matrix on the right, in contrast,
-    does not have the COP \cite{at72,d08phd}.}
+  % \includegraphics[scale=0.4]{../img/0_cop_nocop.png}
+  \begin{tabular}[h]{l|lcccl}
+    $M_1$: & $M_1'$: &&&& $M_2$:\\
+    &&&&&\\
+    \begin{tabular}[h]{llll}
+      $c_1$ & $c_2$ &$c_3$ &$c_4$\\
+      &&&\\
+      \un & 0   & \un & 0\\
+      0   & \un & 0   & \un \\
+      \un & 0   & 0   & \un
+    \end{tabular}
+    &
+    \begin{tabular}[h]{llll}
+      $c_3$ &$c_1$ &$c_4$& $c_2$\\
+      &&&\\
+      \un & \un & 0 & 0\\
+      0 & 0 & \un & \un \\
+      0 & \un & \un & 0
+    \end{tabular}
+    &&&&
+    \begin{tabular}[h]{llll}
+      $d_1$ & $d_2$ &$d_3$ &$d_4$\\
+      % &&&\\
+      &&&\\
+      \un & \un & 0 & 0\\
+      0 & \un & \un & 0 \\
+      0 & \un & 0 & \un 
+    \end{tabular}
+  \end{tabular}
+  \caption{Matrices with and without COP. $M_1$ has COP because by
+    permuting its columns, $c_1$-$c_4$, one can obtain $M_1'$ where
+    the {\un}s in each row are consecutive. $M_2$, however, does not
+    have COP since no permutation of its columns, $d_1$-$d_4$, will
+    arrange {\un}s in each row consecutively \cite{d08phd}.}
   \label{fig:cop-matrix}
@@
 \subsection{Optimization problems in COP}
-So far we have been concerned about
-matrices that have the consecutive ones property. However in real life
-applications, it is rare that data sets represented by binary matrices
-have COP. Nevertheless, the COP is not arbirary and there are several
-interesting problems when a matrix does not have COP but is ``close''
-to having COP or can be altered to have COP.
-
--- sect 4.1 in \cite{d08phd} has many results surveyed. hardness results,
-approx. results. results are usually for a class of matrices $(a,b)$
-where number columns and rows are restriced to $a$ and $b$
+So far we have been concerned about matrices that have the consecutive
+ones property. However in real life applications, it is rare that data
+sets represented by binary matrices have COP. Nevertheless, the COP is
+not arbirary and there are several interesting problems when a matrix
+does not have COP but is ``close'' to having COP or can be altered to
+have COP.
+
+-- sect 4.1 in \cite{d08phd} has many results surveyed. hardness
+results, approx. results. results are usually for a class of matrices
+$(a,b)$ where number columns and rows are restriced to $a$ and $b$
 \tnote{check if it is the other way around}.\\

### 2011-11-22 src/synopsis.tex
a whole bunch of changes i cannot rememeber now. deal with it.

@@
 \department{\MakeUppercase{\mydept}}
+
+\def \un {\bf 1}
 
@@
 
-% \tnote{testing 1 2 3...}
-% \tnote[bogus1]{testing 1 2 3...}
-% \tnote[bogus2]{testing 1 2 3...}
-% \tnote[bogus3]{testing 1 2 3...}
-% \tnote[bogus4]{testing 1 2 3...}
-% \tnote[bogus5]{testing 1 2 3...}
-% \tnote[bogus6]{testing 1 2 3...}
-% \tnote[bogus7]{testing 1 2 3...}
-% \tnote[bogus8]{testing 1 2 3...}
-% \tnote[TH]{testing 1 2 3...}
+\tnote{testing 1 2 3...}
+\tnote[bogus1]{testing 1 2 3...}
+\tnote[bogus2]{testing 1 2 3...}
+\tnote[GTC]{testing 1 2 3...}
+\tnote[bogus4]{testing 1 2 3...}
+\tnote[TH]{testing 1 2 3...}
 
@@
 of their semester thesis, they pick a body of work to study and form
-the namesake study groups, {\LLL}, {\GGG}, {\BBB} and {\TTT}. A student
+the namesake study groups, {\LLL}, {\GGG}, {\BBB} and
+{\TTT}. \tnote[TH]{put bib entries for these works!} A student
 will be in at least one study group and may be in more than one. For
@@
 
-\begin{figure}[htbp]
+\begin{figure}[t]%[htbp]
  \centering
@@
 
-\begin{figure}[htbp]
+\begin{figure}[t]%[htbp]
   \centering
@@
 
-\begin{figure}[h]
+\begin{figure}[t] % [h]
   \centering
@@
   nsnrs09}.  The COP problem is to rearrange rows (columns) of a
-binary matrix in such a way that every column (row) has its $1$s occur
+binary matrix in such a way that every column (row) has its {\un}s occur
 consecutively. If this is possible the matrix is said to have the COP.
@@
 
-%The course scheduling problem illustated earlier, as a special case of
-
-It is quite obvious now why this version of the
-problem (where the ``target graph'' is a path and not any
-tree\tnote[TH]{Allowing any tree in this example could be seen as a
-  scenario where there are parallel classes. A node falling in the
-  path between two other nodes would mean that the corresponding is
-  scheduled between the other two.}) is called an interval assignment
-problem.
-
-The interval assignment problem to a set system is equivalent to the
-consecutive ones problem (COP) in binary matrices\cite{wlh02,
-  nsnrs09}.  The COP problem is to rearrange rows (columns) of a
-binary matrix in such a way that every column (row) has its $1$s occur
-consecutively. If this is possible the matrix is said to have the COP.
-Consecutive ones property (COP) of binary matrices is a well
-researched combinatorial problem and has several positive results on
-tests for it and computing the COP permutation (i.e. the course
-schedule in the above illustration) which will be surveyed later in
-this document. Hence we are interested in extensions of COP, more
-specifically, the extension of interval assignment problem to tree
-path assignment problem (which is illustrated by the study group
-accommodation problem).
-
-As described earlier, 
-a matrix with COP is one whose rows (columns)
-can be rearranged so that the 1s in every column (row) are in
-consecutive rows (columns).
-
-\def \un {\bf 1}
-\begin{figure}[htbp]
+As seen earlier, the interval assignment problem (illustrated as the
+course scheduling problem in Section~\ref{sec:problem}), is a special
+case of the problem we address in this thesis, namely the tree path
+labeling problem (illustrated as the study group accommodation
+problem). Interval assignment and COP are equivalent
+problems. In this section we will see some of the results that
+exists in the literature today towards solving the COP problem and
+optimization problems surrounding it.
+
+% The interval assignment problem to a set system is equivalent to the
+% consecutive ones problem (COP) in binary matrices\cite{wlh02,
+%   nsnrs09}.  The COP problem is to rearrange rows (columns) of a
+% binary matrix in such a way that every column (row) has its {\un}s occur
+% consecutively. If this is possible the matrix is said to have the COP.
+% Consecutive ones property (COP) of binary matrices is a well
+% researched combinatorial problem and has several positive results on
+% tests for it and computing the COP permutation (i.e. the course
+% schedule in the above illustration) which will be surveyed later in
+% this document. Hence we are interested in extensions of COP, more
+% specifically, the extension of interval assignment problem to tree
+% path assignment problem (which is illustrated by the study group
+% accommodation problem).
+
+\begin{figure}[t] %[htbp]
   \centering
@@
 
+Recall that a matrix with COP is one whose rows (columns)
+can be rearranged so that the {\un}s in every column (row) are in
+consecutive rows (columns). Figure~\ref{fig:cop-matrix} shows examples
+of this property.
 COP in binary matrices has several practical applications in diverse
@@
 binary matrix is polynomial time solvable and the first such algorithm
-was given by \cite{fg65}.  Since then several algorithms have been
-invented; PQ trees \cite{bl76} gave the first linear time
-algorithm. Variations of PQ trees \cite{mm96,wlh01,wlh02,mcc04} and
-ICPIA \cite{nsnrs09} further improve either the time complexity and/or
-the complexity of implementation.
-
-\cite{bl76} describes a linear time construction of a data structure
-called PQ trees. This paper draws on the close relationship of COP
-matrices to interval graphs. A PQ tree of a matrix is one that stores
-all row (column) permutations of the matrix that give the COP orders
-(there could be multiple) of the matrix. This is constructed using an
-elaborate linear time procedure. Since this result, several
-improvements have been made to the procedural complexity of the
-algorithm itself and also derivative data structures that are easier
-to construct. PQR trees is one such data structure \cite{mm96},
-\cite{mpt98}\tnote[TH]{also, mpt98 $==$ elaborate version of mm96}.
-\cite{tm05} describes an improved version of PQR
-trees. \tnote{improv in terms of what?}\cite{wlh02} describes the
-simpler algorithm for COT. Hsu also invented PC trees
+was given by \cite{fg65}.  A landmark result came a few years later
+when \cite{at72} discovered the families of forbidden submatrices that
+prevent a matrix from having COP and most, if not all, results that
+came later were based on this discovery which connected COP in binary
+matrices to convex bipartite graphs. In fact, the forbidden
+submatrices came as a corollary to the discovery that convex bipartite
+graphs are AT-free in \cite{at72}\tnote{check}. The first linear time
+algorithm for COP testing (COT) was invented by \cite{bl76} using a
+data structure called PQ trees.  Since then several COT algorithms
+have been invented -- some of which involved variations of PQ trees
+\cite{mm96,wlh01,mcc04}, some involved set theory and ICPIA
+\cite{wlh02,nsnrs09}, parallel COT algorithms\cite{as95,bs03,ly91} and
+certifying algorithms\cite{mcc04}. % These algorithms
+% improve over PQ trees on the time complexity and/or the complexity
+% of implementation.
+
+The construction of PQ trees in \cite{bl76} draws on the
+close relationship of COP matrices to interval graphs. A PQ tree of a
+matrix is one that stores all row (column) permutations of the matrix
+that give the COP orders (there could be multiple) of the matrix. This
+is constructed using an elaborate linear time procedure and is also
+a test for planarity\tnote{check check check. both interval
+  graph and planarity in this paper?}. % Since this
+% result, several improvements have been made to the procedural
+% complexity of the algorithm itself and also derivative data structures
+% that are easier to construct.
+PQR trees is a generalized data structure based on PQ trees
+\cite{mm96,mpt98}.  \cite{tm05} describes an improved algorithm to
+build PQR trees. \tnote[GTC]{improv in terms of what?}\cite{wlh02}
+describes the simpler algorithm for COT. Hsu also invented PC trees
 \cite{wlh01}\tnote[TH]{This result first appeared inproc ISAAC92}
@@
 this thesis because it simplifies COT to a great
-degree. \tnote{get this
-  figured out! - because it reduces the search space for the solution
-  from set of all permutations of rows to set of all}
+degree. \tnote[GTC]{it reduces the solution search space. fill in the blanks.}
 
 \cite{mcc04} describes a different approach to COT. While all previous
-COT algorithms gave the COP order if the matrix has the property and
-if not it exited stating negative, this algorithm gives an evidence by
-way of a certificate of
-matrix being devoid of COP thereby enabling a user to check the
-algorithm's result. This is significant from an implementation
-perspective because automated program verification is hard and manual
-verification is more viable. Hence having a certificate
-reinforces its credibility. Note that when the matrix has COP the COP
-order is the certificate.  It may also be noted that the internal
-machinery of this algorithm \annote {is related to the weighted
-  betweenness problem addressed } {in what way??} in \cite{co98}.
-\tnote{expand on the COP order graph creation and it having to be
-  bipartite for M to have COP. and thus an odd cycle being an evidence
-  of no COP.}
-\tnote[TH]{ where should this go?: cite|jlm97 (application of PQ trees
-  in graphics).\\ add helly's theorem citation
+COT algorithms gave the COP order if the matrix has the property but
+exited stating negative if otherwise, this algorithm gives an evidence
+by way of a certificate of matrix even when it has no COP. This
+enables a user to verify the algorithm's result even when the answer
+is negative. This is significant from an implementation perspective
+because automated program verification is hard and manual verification
+is more viable. Hence having a certificate reinforces an
+implementation's credibility. Note that when the matrix {\em has} COP,
+the COP order is the certificate.  The internal machinery of this
+algorithm is related to the weighted betweenness problem
+addressed\tnote[GTC]{in what way??} in \cite{co98}.  \tnote[GTC]{expand
+  on the COP order graph creation and it having to be bipartite for M
+  to have COP. and thus an odd cycle being an evidence of no COP.}
+\tnote[TH]{ where should this go?: (1) cite|jlm97 (application of PQ trees
+  in graphics). (2) helly's theorem citation
   19XXdgk-Hellystheorem-Danzer-Gruenbaum-Klee}
 
-The literature also has results on parallel COT algorithms --
-\cite{as95} \cite{bs03} \cite{ly91}\tnote{add english to fit context}
-
 \subsection{Optimization problems in COP}
+\label{sec:optcop}
 So far we have been concerned about matrices that have the consecutive
 ones property. However in real life applications, it is rare that data
-sets represented by binary matrices have COP. Nevertheless, the COP is
-not arbirary and there are several interesting problems when a matrix
-does not have COP but is ``close'' to having COP or can be altered to
-have COP.
-
+sets represented by binary matrices have COP, primarily due to the
+noisy nature of data available. At the same time, COP is not arbitrary
+and is a desirable property in practical data representation
+\cite{co98,jkckv04,k77}. In this context, there are several
+interesting problems when a matrix does not have COP but is ``close''
+to having COP or is allowed to be altered to have COP. These are the
+optimization problems related to a matrix which does not have
+COP. Some of the significant problems are surveyed in this section.
+
+MOVE TO THESIS?
 -- sect 4.1 in \cite{d08phd} has many results surveyed. hardness
@@
 COP. this is NP complete \cite{b75-phd}
-
-The first step in this direction is to analyze matrices without
-COP. \cite{at72} showed that a matrix that does not have COP have
-certain forbidden substructures that prevent it from having
-COP. Tucker classified these substructures into five classes of
-submatrices. This result is presented in the context of AT-free
-bipartite graphs \tnote[TH]{ check this up. give details.} which is
+END MOVE TO THESIS?
+
+\tnote[TH]{cite:lb62 - interval graphs are at-free chordal}
+
+\cite{at72} showed that a matrix that does not have COP have
+certain substructures that prevent it from having
+COP. Tucker classified these forbidden substructures into five classes of
+submatrices. This result is presented in the context of convex
+bipartite graphs which \cite{at72} proved to be
+AT-free\tnote[TH]{ check this up. give details. - doms'}. Convex bipartite
+graph have vertex incidence adjacency matrices is
 equivalent to matrices with COP. A binary matrix is converted into a
@@
 follows. Every column of the binary matrix can be converted into a set
-of integers which are the indices of rows with $1$s in that column. It
+of integers which are the indices of rows with {\un}s in that column. It
 is apparent that if the matrix has COP, then constructing such sets

### 2011-11-23 src/synopsis.tex
all sections except sec 4 onwards are done. rest will be done in less than an hour

@@
 
-\tnote{testing 1 2 3...}
-\tnote[bogus1]{testing 1 2 3...}
-\tnote[bogus2]{testing 1 2 3...}
-\tnote[GTC]{testing 1 2 3...}
-\tnote[bogus4]{testing 1 2 3...}
-\tnote[TH]{testing 1 2 3...}
+% \tnote{testing 1 2 3...}
+% \tnote[bogus1]{testing 1 2 3...}
+% \tnote[bogus2]{testing 1 2 3...}
+% \tnote[GTC]{testing 1 2 3...}
+% \tnote[bogus4]{testing 1 2 3...}
+% \tnote[TH]{testing 1 2 3...}
 
@@
 % enroll at the
-{\WSI} for a liberal arts programme.  As part
-of their semester thesis, they pick a body of work to study and form
-the namesake study groups, {\LLL}, {\GGG}, {\BBB} and
-{\TTT}. \tnote[TH]{put bib entries for these works!} A student
-will be in at least one study group and may be in more than one. For
-instance, as will be seen later, {\Fr} studies both {\LLL} and {\TTT}
-while \Wo\ studies only \BBB.
+  {\WSI} for a liberal arts programme.  As part of their semester
+  thesis, they pick a body of work to study and form the namesake
+  study groups, {\LLL}, {\GGG}, {\BBB} and {\TTT}\tnote[TH]{put bib
+    entries for these works!}. A student will be in at least one study
+  group and may be in more than one. For instance, as will be seen
+  later, {\Fr} studies both {\LLL} and {\TTT} while \Wo\ studies only
+  \BBB.
 
@@
 
+\def \figtabsize {\footnotesize}
+
 \begin{table}[htbp]
   \centering
+{
   \begin{tabular}{rcl}
@@
   \end{tabular}
-  \caption{Students and study groups in \WSI}
+}
+  \caption{\figtabsize Students and study groups in \WSI}
   \label{tab:wsigroups}
@@
 connecting them do {\em not} form loops. Fig~\ref{fig:streetmap} shows
-the street map for {\residenceblock}. It may be noted that as a graph, it
-classifies as a tree.
+the street map for {\residenceblock}. It may be noted that as a graph,
+it classifies as a tree.
 
@@
 and the students' allocation of apartments that obeys this route
-allocation is shown in Fig~\ref{fig:streetmappathpeople}. How this is
-algorithmically computed is the focus of this thesis.
-
-\begin{figure}[t]%[htbp]
+allocation is also shown\tnote[TH]{in
+  Fig~ref::fig:streetmappathpeople}. How this is algorithmically
+computed is the focus of this thesis.
+
+%%%%%%%%%%%% THESIS UNCOMMENT %%%%%%%%%%%%
+% \begin{figure}[t]%[htbp]
+%  \centering
+%   \includegraphics[scale=0.3]{../img/1_infinite_loop.png}
+%   \caption{{\residenceblock} street map}
+%   \label{fig:streetmap}
+% \end{figure}
+
+% \begin{figure}[t]%[htbp]
+%   \centering
+%   \includegraphics[scale=0.3]{../img/2_infinite_loop_BTWF.png}
+%   \caption{{\residenceblock} street map with study group routes
+%     allocated. Routes are color coded as follows: red for
+%     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
+%     group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
+%     \textcolor{green}{$\xTTT$} group}
+%   \label{fig:streetmappath}
+% \end{figure}
+% \begin{figure}[t] % [h]
+%   \centering
+%   \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
+%   \caption{Individual allocation of apartments to students in
+%     {\residenceblock} that meets the requirements stated before.}
+%   \label{fig:streetmappathpeople}
+% \end{figure}
+%%%%%%%%%%%% END THESIS UNCOMMENT %%%%%%%%%%%%
+\begin{figure}[htbp] %[t]%
  \centering
-  \includegraphics[scale=0.3]{../img/1_infinite_loop.png}
-  \caption{{\residenceblock} street map}
+  (a) \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
+  (b) \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
  \label{fig:streetmap}
-\end{figure}
-
-\begin{figure}[t]%[htbp]
-  \centering
-  \includegraphics[scale=0.3]{../img/2_infinite_loop_BTWF.png}
-  \caption{{\residenceblock} street map with study group routes
+  \caption{\figtabsize (a) {\residenceblock} street map (b) {\residenceblock} street map with study group routes
     allocated. Routes are color coded as follows: red for
@@
     \textcolor{green}{$\xTTT$} group}
-  \label{fig:streetmappath}
 \end{figure}
-
-\begin{figure}[t] % [h]
+\begin{figure}[htbp] %[t]%
   \centering
-  \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
-  \caption{Individual allocation of apartments to students in
+  \includegraphics[scale=0.3]{../img/3_infinite_loop.png}
+  \caption{\figtabsize Individual allocation of apartments to students in
     {\residenceblock} that meets the requirements stated before.}
@@
 
-\tnote{add a note on the organization of the doc}
-% The necessary preliminaries with definitions etc. are presented in
-% Section~\ref{sec:prelims}. 
-% Section~\ref{sec:feasible} documents the
-% characterization of a feasible path labeling and finally,
-% Section~\ref{sec:ksubdivstar} describes a polynomial time algorithm to
-% find the tree path labeling of a given set system from a given
-% $k$-subdivided tree.
-
-Section~\ref{sec:motivation} gives a brief survey of COP and
+Section~\ref{sec:background} gives a brief survey of COP and
 optimization problems related to it.  Section~\ref{sec:results}
 presents a summary of our results on the extention of COP namely, the
-tree path labeling problem.
-
-\section{Motivation and background}
-\label{sec:motivation}
-
+tree path labeling problem. \tnote[TH]{update for thesis}
+
+\section{Background}
+\label{sec:background}
+\subsection{Matrices with COP}
+\label{sec:copmatrices}
 As seen earlier, the interval assignment problem (illustrated as the
@@
   % \includegraphics[scale=0.4]{../img/0_cop_nocop.png}
+{\figtabsize
  \begin{tabular}[h]{l|lcccl}
@@
   \end{tabular}
-  \caption{Matrices with and without COP. $M_1$ has COP because by
+}
+
+  \caption{\figtabsize Matrices with and without COP. $M_1$ has COP because by
     permuting its columns, $c_1$-$c_4$, one can obtain $M_1'$ where
@@
     arrange {\un}s in each row consecutively \cite{d08phd}.}
+
   \label{fig:cop-matrix}
@@
 
-MOVE TO THESIS?
--- sect 4.1 in \cite{d08phd} has many results surveyed. hardness
-results, approx. results. results are usually for a class of matrices
-$(a,b)$ where number columns and rows are restriced to $a$ and $b$
-\tnote{check if it is the other way around}.\\
--- problem of flipping at most $k$ entries of $M$ to make it attain
-COP. this is NP complete \cite{b75-phd}
-END MOVE TO THESIS?
-
-\tnote[TH]{cite:lb62 - interval graphs are at-free chordal}
-
-\cite{at72} showed that a matrix that does not have COP have
-certain substructures that prevent it from having
-COP. Tucker classified these forbidden substructures into five classes of
-submatrices. This result is presented in the context of convex
-bipartite graphs which \cite{at72} proved to be
-AT-free\tnote[TH]{ check this up. give details. - doms'}. Convex bipartite
-graph have vertex incidence adjacency matrices is
-equivalent to matrices with COP. A binary matrix is converted into a
-bipartite graph as follows. The set of rows and the set of columns
-form the two partitions of the graph. Each row node is adjacent to
-those nodes that represent the columns that have ones in the
-corresponding row. \cite{at72} proves that this bipartite graph has no
-asteroidal triple if and only if the matrix has COP and goes on to
-identify the forbidden substructures for these bipartite graphs. The
-matrices corresponding to these substructures are the forbidden
-submatrices. Once a matrix has been detected to not have COP (using
-any of the COT algorithms mentioned earlier), it is naturally of
-interest to find out the smallest (in terms of number of rows, columns
-and number of entries that are ones) forbidden
-substructure. \cite{d08phd} discusses an exact algorithm with time
-complexity \annote{parameterized by the number of ones in the
-  matrix}{check this.}.
-
-{\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are similarly
-inspired problems. In {\sc Min COS-R (Min COS-C)} the question is to
-find the minimum number of rows (columns) that must be deleted to
-result in a matrix with COP.  In the dual problem {\sc Max COS-R (Max
-  COS-C)} the search is for the maximum number of rows (columns) that
-induces a submatrix with COP. Given a no-COP matrix $M$,
-\cite{b75-phd} shows that finding a submatrix $M'$ with all columns
-but a maxmimum cardinality subset of rows such that $M'$ has COP is NP
-complete\tnote{check if b75 deals with COP col or COP row. also
-  is it any submatrix with k less than r rows or submatrix must have
-  all columns?}. \cite{hg02} corrects an error of the abridged proof
-of this reduction as given in \cite{gj79}.  \cite{d08phd} discusses
-all these problems in detail giving an extensive survey of the
-previously existing results which are almost exhaustively all
-approximation results and hardness results. Taking this further,
-\cite{d08phd} presents new results in the area of parameterized
-algorithms for this problem\tnote[TH]{elaborate - what are
-  the results?}.
-
-Another problem is to find the minumum number of entries in the matrix
-that can be flipped to result in a matrix with COP.  \cite{v85}
+\tnote[TH]{ -- sect 4.1 in cite:d08phd has many results
+  surveyed. hardness results, approx. results. results are usually for
+  a class of matrices $(a,b)$ where number {\un}s in columns and rows
+  are restriced to $a$ and $b$ . -- problem of flipping at most $k$
+  entries of $M$ to make it attain COP. this is NP complete
+  cite:b75-phd}\tnote[TH]{(1) scite:lb62 showed that interval graphs
+  are AT-free.  describe AT (2) show the close relationship b/w COP
+  and graphs sec 2.2, pg 31} \cite{at72} showed that a matrix that
+does not have COP have certain substructures that prevent it from
+having COP. Tucker classified these forbidden substructures into five
+classes of submatrices. This result is presented in the context of
+convex bipartite graphs which \cite{at72} proved to be
+AT-free\tnote[TH]{ check this up. give details. - doms'}. By
+definition, convex bipartite graph have half adjacency matrices that
+have COP on either rows or columns (graph is biconvex if it has COP on
+both)\cite{d08phd}. A half adjacency matrix is a binary matrix
+representing a bipartite graph as follows. The set of rows and the set
+of columns form the two partitions of the graph. Each row node is
+adjacent to those nodes that represent the columns that have ones in
+the corresponding row. \cite{at72} proves that this bipartite graph
+has no asteroidal triple if and only if the matrix has COP and goes on
+to identify the forbidden substructures for these bipartite
+graphs. The matrices corresponding to these substructures are the
+forbidden submatrices.
+
+Once a matrix has been detected to not have COP (using any of the COT
+algorithms mentioned earlier), it is naturally of interest to find out
+the smallest (in terms of number of rows and/or columns and/or number
+of entries that are {\un}s) forbidden substructure. \cite{d08phd}
+discusses a couple of algorithms which are efficient if the number of
+ones in a row is small. This is of significance in the case of sparce
+matrices where this number is much lesser than the number of
+columns. $(*,\Delta)${\em -matrices} are matrices with no restriction
+on number of {\un}s in any column but has at most $\Delta$ {\un}s in
+any row. {\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are
+similar problems which deals with inducing COP on a matrix. In {\sc
+  Min COS-R (Min COS-C)} the question is to find the minimum number of
+rows (columns) that must be deleted to result in a matrix with COP.
+In the dual problem {\sc Max COS-R (Max COS-C)} the search is for the
+maximum number of rows (columns) that induces a submatrix with
+COP. Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
+a submatrix $M'$ with all columns\tnote{check if b75 deals with COP
+  col or COP row. also is it any submatrix with k less than r rows or
+  submatrix must have all columns?} but a maximum cardinality subset
+of rows such that $M'$ has COP is NP complete. \cite{hg02} corrects an
+error of the abridged proof of this reduction as given in \cite{gj79}.
+\cite{d08phd} discusses all these problems in detail giving an
+extensive survey of the previously existing results which are almost
+exhaustively all approximation results and hardness results. Taking
+this further, \cite{d08phd} presents new results in the area of
+parameterized algorithms for this problem\tnote[TH]{elaborate - what
+  are the results?}.
+
+Another problem is to find the minimum number of entries in the matrix
+that can be toggled to result in a matrix with COP.  \cite{v85}
 discusses approximation of {\sc COP Augmentation} which is the problem
-of changing of the minimum number of zero entries to ones so that the
-resulting matrix has COP. As mentioned earlier, this problem had been
+of changing of the minimum number of zero entries to {\un}s so that the
+resulting matrix has COP. As mentioned earlier, this problem is
 known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
@@
 
-\cite{jkckv04} discusses the use of matrices with almost-COP in the
-storage of very large databases.  The problem is that of $k$-run
-reordering of binary matrices which they prove to be NP hard using a
-reduction from the hamiltonian path problem.  A connection of COP
-problem to the travelling salesman problem is also
-introduced. \tnote[TH]{what does this mean? -- COP can be used as a
-  tool to reorder $0.5T \le runs(M) \le T$}
-
-The optimization version of the $k$-run problem, i.e. minimization of
-number of blocks of ones is proven to be NP complete by
-\cite{k77}. \tnote[TH]{what is the reduction?}
-
-\tnote[TH]{other problems similar to COP -- cite:ckl96 (ILP, circ ones, one drop) -- cite:th98 (generalization of COP - minimax, biotonic column) Tucker}
-
-{\bf Motivation: } A natural generalization of the interval assignment problem is
-feasible tree path labeling problem of a set system. The problem is
-defined as follows -- given a set system $\cF$ from a universe $U$ and
-a tree $T$, does there exist a bijection from $U$ to the vertices of
-$T$ such that each set in the system maps to a path in $T$.  We refer
-to this as the {\em tree path labeling problem} for an input set
-system, target tree pair -- $(\cF,T)$. As a special case if the tree
-$T$ is a path, the problem becomes the interval assignment problem.
-We focus on the question of generalizing the notion of an ICPIA
-\cite{nsnrs09} to characterize feasible path assignments.  We show
-that for a given set system $\cF$, a tree $T$, and an assignment of
-paths from $T$ to the sets, there is a feasible bijection between $U$
-and $V(T)$ if and only if all intersection cardinalities among any
-three sets (not necessarily distinct) is same as the intersection
-cardinality of the paths assigned to them and the input passes a
-filtering algorithm (described in this paper) successfully.  This
-characterization gives a natural data structure that stores all the
-relevant feasible bijections between $U$ and $V(T)$. This reduces the
-search space for the solution considerably from the universe of all
-possible bijections between $U$ and $V(T)$ to only those bijections
-that maintain the characterization\tnote{REWORD}.  Further, the
-filtering algorithm is also an efficient algorithm to test if a tree
-path labeling to the set system is feasible.  This generalizes the
-result in \cite{nsnrs09}.
-
-Our results also have close connection to recognition of path graphs
-and connection to path graph isomorphism.  Graphs which can be
-represented as the intersection graph of paths in a tree are called
-{\em path graphs}\cite{mcg04}. Thus given a hypergraph $\cF$, it can
-be viewed as paths in a tree, if and only if the intersection graph of
-$\cF$ is a {\em path graph}. Path graphs are a subclass of chordal
-graphs since chordal graphs are combinatorially characterized as the
-intersection graphs of subtrees of a tree.  Path graphs are well
-studied in the literature \cite{plr70,gav78,bp93,mcg04}.  Checking if
-a graph is a path graph can be done in polynomial time by the results
-of \cite{gav78,aas93}.  However, this is only a necessary condition
-for our question.  \remove{Indeed, our work provides a sufficient
-  condition for a related problem: that is whether a path assignment
-  to be feasible?}  On the other hand, path graph isomorphism is known
-be isomorphism-complete; see for example \cite{kklv10}. Therefore,
-\annote{it is unlikely that we can solve the problem of finding
-  feasible path labeling $\cl$ for a given $\cF$ and }{WHY?} tree $T$.
-\annote{It is definitely interesting to classify the kinds of trees
-  and hypergraphs for which feasible path labelings can be
-}{ELABORATE.} found efficiently.  \annote{These results would form
-  a natural generalization of COP testing and interval graph
-  isomorphism, culminating in Graph Isomorphism }{REWORD} itself.
-
-How the problem of COT is related to the constraint statisfaction
-problem of interval assignment to sets in a system can be seen as
-follows. Every column of the binary matrix can be converted into a set
-of integers which are the indices of rows with {\un}s in that column. It
-is apparent that if the matrix has COP, then constructing such sets
-after applying the COP permutation to the matrix will result in all
-sets with consecutive integers. In other words, after application of
-COP reordering, the sets are intervals. Indeed the problem now becomes
-finding interval assignments to a given set system such that there
-exists a permutation of the universe (set of row indices) which
-converts each set to its assigned interval. The result in
+\cite{jkckv04} discusses the use of matrices with almost-COP (instead
+of one block of consecutive {\un}s, they have $x$ blocks, or {\em
+  runs}, of consecutive {\un}s and $x$ is not too large) in the
+storage of very large databases.  The problem is that of reordering of
+a binary matrix such that the resulting matrix has at most $k$ runs of
+{\un}s. This is proven to be NP hard using a reduction from the
+hamiltonian path problem.\tnote[TH]{Theorem 2.1 in jkckv}
+\tnote[TH]{(1) A connection of COP problem to the travelling salesman
+  problem is also introduced. what does this mean? -- COP can be used
+  as a tool to reorder $0.5T \le runs(M) \le T$. (2) The optimization
+version of the $k$-run problem, i.e. minimization of number of blocks
+of ones is proven to be NP complete by cite:k77}\tnote{are these two
+  the same?} \tnote[TH]{what is the reduction?} \tnote[TH]{other
+  problems similar to COP -- cite:ckl96 (ILP, circ ones, one drop) --
+  cite:th98 (generalization of COP - minimax, biotonic column) Tucker}
+
+\section{Motivation}
+\label{sec:motive}
+As seen in Section~\ref{sec:copmatrices}, \cite{nsnrs09} describes a
+characterization of consecutive ones property based on the cardinality
+properties of the set representations of the columns. This result is
+very relevant to this thesis because aside from it simplifying COT to
+a great degree, our generalization problem is motivated by their
+results.
+
+The result in
 \cite{nsnrs09} characterizes interval assignments to the sets which
@@
 
+The tree path labeling problem is a natural generalization of the
+interval assignment problem or the COP problem. The problem is defined
+as follows -- given a set system $\cF$ from a universe $U$ and a tree
+$T$, does there exist a bijection from $U$ to the vertices of $T$ such
+that each set in the system maps to a path in $T$.  We refer to this
+as the {\em tree path labeling problem} for an input set system,
+target tree pair -- $(\cF,T)$. As a special case if the tree $T$ is a
+path, the problem becomes the interval assignment problem.  We focus
+on the question of generalizing the notion of an ICPIA \cite{nsnrs09}
+to characterize feasible path assignments.  We show that for a given
+set system $\cF$, a tree $T$, and an assignment of paths from $T$ to
+the sets, there is a feasible bijection between $U$ and $V(T)$ if and
+only if all intersection cardinalities among any three sets (not
+necessarily distinct) is same as the intersection cardinality of the
+paths assigned to them and the input passes a filtering algorithm
+(described in this paper) successfully.  This characterization gives a
+natural data structure that stores all the relevant feasible
+bijections between $U$ and $V(T)$. This reduces the search space for
+the solution considerably from the universe of all possible bijections
+between $U$ and $V(T)$ to only those bijections that maintain the
+characterization\tnote[TH]{REWORD}.  Further, the filtering algorithm
+is also an efficient algorithm to test if a tree path labeling to the
+set system is
+feasible.  %This generalizes the result in \cite{nsnrs09}.
+
+% How the problem of COT is related to the constraint statisfaction
+% problem of interval assignment to sets in a system can be seen as
+% follows. Every column of the binary matrix can be converted into a set
+% of integers which are the indices of rows with {\un}s in that column. It
+% is apparent that if the matrix has COP, then constructing such sets
+% after applying the COP permutation to the matrix will result in all
+% sets with consecutive integers. In other words, after application of
+% COP reordering, the sets are intervals. Indeed the problem now becomes
+% finding interval assignments to a given set system such that there
+% exists a permutation of the universe (set of row indices) which
+% converts each set to its assigned interval.
+
 \section{Summary of results}
 \label{sec:results}
+
+We address the computation of TPL\footnote{Tree Path Labeling} for a
+given set system and tree in this research. However optimization
+problems have not been explored.
 
@@
 corresponding paths.  This generalizes the well studied question of
-the case when the given tree $T$ is a path \cite{wlh02,nsnrs09}.
+the case when the given tree $T$ is a path \cite{wlh02,nsnrs09}, in
+other words the interval assignment problem or consecutive ones problem.
 
 %\item 
-In the Section 4, \annote{we initiate an exploration of}{WTF?}
-finding feasible path labeling of set systems in a special kind of
-tree which we call the $k$-subdivided star.  \annote{This question
-  is an attempt to generalize}{REWORD.} the problem of testing if a
-matrix has the consecutive ones property.  However, we restrict the
-hypergraph $\cF$ to be such that all hyperedges have at most $k+2$
-elements.  \annote{In spite of this restricted case we consider, we
-  believe that our results are of significant interest in
-  understanding the nature of Graph Isomorphism}{REWORD.} which is
-polynomial time solvable in interval graphs and is hard on path
-graphs.
+Aside from checking if a given TPL is feasible, we also attempt to
+compute the TPL for a given set system (hypergraph) and tree, if one
+exists. For the latter problem, we have an algorithm for general trees
+which runs in exponential time. We present a polynomial time
+algorithm for a special class of trees called $k$-subdivided trees and
+when the hypergraph $\cF$ is such that all hyperedges have at most
+$k+2$ elements.  In spite of this restricted case we consider, we
+believe that our results are of significant interest in understanding
+the nature of Graph Isomorphism which is polynomial time solvable in
+interval graphs and is hard on path graphs. The connection of TPL to graph
+isomorphism will be made later in the document
 %\end{enumerate}
@@
 \section{Conclusion}
+
+Our results also have close connection to recognition of path graphs
+and connection to path graph isomorphism.  Graphs which can be
+represented as the intersection graph of paths in a tree are called
+{\em path graphs}\cite{mcg04}. Thus given a hypergraph $\cF$, it can
+be viewed as paths in a tree, if and only if the intersection graph of
+$\cF$ is a {\em path graph}. Path graphs are a subclass of chordal
+graphs since chordal graphs are combinatorially characterized as the
+intersection graphs of subtrees of a tree.  Path graphs are well
+studied in the literature \cite{plr70,gav78,bp93,mcg04}.  Checking if
+a graph is a path graph can be done in polynomial time by the results
+of \cite{gav78,aas93}.  However, this is only a necessary condition
+for our question.  \remove{Indeed, our work provides a sufficient
+  condition for a related problem: that is whether a path assignment
+  to be feasible?}  On the other hand, path graph isomorphism is known
+be isomorphism-complete; see for example \cite{kklv10}. Therefore,
+\annote{it is unlikely that we can solve the problem of finding
+  feasible path labeling $\cl$ for a given $\cF$ and }{WHY?} tree $T$.
+\annote{It is definitely interesting to classify the kinds of trees
+  and hypergraphs for which feasible path labelings can be
+}{ELABORATE.} found efficiently.  \annote{These results would form
+  a natural generalization of COP testing and interval graph
+  isomorphism, culminating in Graph Isomorphism }{REWORD} itself.
+
 \tnote{MAKE CONNECTION TO PATH GRAPH ISOMORPHISM HERE}

### 2011-11-23 src/synopsis.tex
almost done. a wee bit more and i can submit it. almost submitted it today but i have a bus to catch. wooooohooooo bangalore, here i come.

@@
 \def \un {\bf 1}
+\def \secfirstparatrim {-5mm}
+\def \figtabsize {\footnotesize}
 
@@
 \label{sec:intro}
-
+\vspace{\secfirstparatrim}
+\tnote{set context: COP - generalization - TPL - no opt. just 2 or 3
+  sentences. use the space that the illustration does.}
 % \tnote{testing 1 2 3...}
@@
 \label{sec:problem}
-
+\vspace{\secfirstparatrim}
 % \def \LLL {{\em ``{\bf L}aughing with Kafka''}} 
@@
 % \end{align*}
-
-\def \figtabsize {\footnotesize}
 
@@
 %%%%%%%%%%%% END THESIS UNCOMMENT %%%%%%%%%%%%
-\begin{figure}[htbp] %[t]%
- \centering
-  (a) \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
-  (b) \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
- \label{fig:streetmap}
-  \caption{\figtabsize (a) {\residenceblock} street map (b) {\residenceblock} street map with study group routes
-    allocated. Routes are color coded as follows: red for
-    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
-    group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
-    \textcolor{green}{$\xTTT$} group}
-\end{figure}
-\begin{figure}[htbp] %[t]%
-  \centering
-  \includegraphics[scale=0.3]{../img/3_infinite_loop.png}
-  \caption{\figtabsize Individual allocation of apartments to students in
-    {\residenceblock} that meets the requirements stated before.}
-  \label{fig:streetmappathpeople}
-\end{figure}
+% \begin{figure}[htbp] %[t]%
+%  \centering
+%   (a) \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
+%   (b) \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
+%  \label{fig:streetmap}
+%   \caption{\figtabsize (a) {\residenceblock} street map (b) {\residenceblock} street map with study group routes
+%     allocated. Routes are color coded as follows: red for
+%     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
+%     group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
+%     \textcolor{green}{$\xTTT$} group}
+% \end{figure}
+% \begin{figure}[htbp] %[t]%
+%   \centering
+%   \includegraphics[scale=0.3]{../img/3_infinite_loop.png}
+%   \caption{\figtabsize Individual allocation of apartments to students in
+%     {\residenceblock} that meets the requirements stated before.}
+%   \label{fig:streetmappathpeople}
+% \end{figure}
 
@@
 The interval assignment problem to a set system is equivalent to the
-consecutive ones problem (COP) in binary matrices\cite{wlh02,
+consecutive ones property (COP) problem in binary matrices\cite{wlh02,
   nsnrs09}.  The COP problem is to rearrange rows (columns) of a
-binary matrix in such a way that every column (row) has its {\un}s occur
-consecutively. If this is possible the matrix is said to have the COP.
-Consecutive ones property (COP) of binary matrices is a well
-researched combinatorial problem and has several positive results on
-tests for it and computing the COP permutation (i.e. the course
-schedule in the above illustration) which will be surveyed later in
-this document. Hence we are interested in extensions of COP, more
-specifically, the extension of interval assignment problem to tree
-path assignment problem (which is illustrated by the study group
-accommodation problem).
+binary matrix in such a way that every column (row) has its {\un}s
+occur consecutively. If this is possible the matrix is said to have
+the COP.  COP is a well researched combinatorial problem and has
+several positive results on tests for it and computing the COP
+permutation (i.e. the course schedule in the above illustration) which
+will be surveyed later in this document. Hence we are interested in
+extensions of COP, more specifically, the extension of interval
+assignment problem to tree path assignment problem (which is
+illustrated by the study group accommodation problem).
 
@@
 
-\section{Background}
+\section{Consecutive Ones Property Testing - a Survey} %{Background}
 \label{sec:background}
+\vspace{\secfirstparatrim}
+In this section, a brief survey of the consecutive ones problem and its
+optimization problems is presented.
+
 \subsection{Matrices with COP}
 \label{sec:copmatrices}
+\vspace{\secfirstparatrim} 
 As seen earlier, the interval assignment problem (illustrated as the
@@
 labeling problem (illustrated as the study group accommodation
-problem). Interval assignment and COP are equivalent
+problem). The interval assignment problem and COP problem are equivalent
 problems. In this section we will see some of the results that
@@
 optimization problems surrounding it.
-
-% The interval assignment problem to a set system is equivalent to the
-% consecutive ones problem (COP) in binary matrices\cite{wlh02,
-%   nsnrs09}.  The COP problem is to rearrange rows (columns) of a
-% binary matrix in such a way that every column (row) has its {\un}s occur
-% consecutively. If this is possible the matrix is said to have the COP.
-% Consecutive ones property (COP) of binary matrices is a well
-% researched combinatorial problem and has several positive results on
-% tests for it and computing the COP permutation (i.e. the course
-% schedule in the above illustration) which will be surveyed later in
-% this document. Hence we are interested in extensions of COP, more
-% specifically, the extension of interval assignment problem to tree
-% path assignment problem (which is illustrated by the study group
-% accommodation problem).
 
@@
 The construction of PQ trees in \cite{bl76} draws on the
-close relationship of COP matrices to interval graphs. A PQ tree of a
+close relationship of matrices with COP to interval graphs. A PQ tree of a
 matrix is one that stores all row (column) permutations of the matrix
-that give the COP orders (there could be multiple) of the matrix. This
+that give the COP orders (there could be multiple orders of rows or columns) of the matrix. This
 is constructed using an elaborate linear time procedure and is also
@@
 \label{sec:optcop}
+\vspace{\secfirstparatrim}  
 So far we have been concerned about matrices that have the consecutive
@@
 
-\section{Motivation}
+\section{Generalization of COP - The Motivation}  %{Motivation}
 \label{sec:motive}
+\vspace{\secfirstparatrim}  
 As seen in Section~\ref{sec:copmatrices}, \cite{nsnrs09} describes a
@@
 
+\def \probdefwidth {0.85\linewidth}
+\def \FTPL {{\sc Feasbile Tree Path Labeling}}
+\def \CFTPL {\sc Compute Feasible Path Labeling}
+
 \section{Summary of results}
 \label{sec:results}
-
-We address the computation of TPL\footnote{Tree Path Labeling} for a
-given set system and tree in this research. However optimization
-problems have not been explored.
-
-Given a path labeling $\cl$ to $\cF$ from a tree $T$, we give a
-necessary and sufficient condition for it to be a feasible path
-labeling.  This necessary and sufficient condition can be tested in
-polynomial time.  The most interesting consequence is that in our
-constructive procedure, it is sufficient to iteratively check if
-three-way intersection cardinalities are preserved.  In other words,
-in each iteration, it is sufficient to check if the intersection of
-any three sets is of the same cardinality as the intersection of the
-corresponding paths.  This generalizes the well studied question of
-the case when the given tree $T$ is a path \cite{wlh02,nsnrs09}, in
-other words the interval assignment problem or consecutive ones problem.
-
-%\item 
-Aside from checking if a given TPL is feasible, we also attempt to
-compute the TPL for a given set system (hypergraph) and tree, if one
-exists. For the latter problem, we have an algorithm for general trees
-which runs in exponential time. We present a polynomial time
-algorithm for a special class of trees called $k$-subdivided trees and
-when the hypergraph $\cF$ is such that all hyperedges have at most
-$k+2$ elements.  In spite of this restricted case we consider, we
-believe that our results are of significant interest in understanding
-the nature of Graph Isomorphism which is polynomial time solvable in
-interval graphs and is hard on path graphs. The connection of TPL to graph
-isomorphism will be made later in the document
+\vspace{\secfirstparatrim}
+% We address the computation of TPL\footnote{Tree Path Labeling} for a
+% given hypergraph\footnote{The term {\em hypergraph} is used instead
+%   {\em set systems} are used interchangably in this
+%   section. Hypergraphs that have TPLs on some target tree are called
+%   {\em path hypergraphs}.} and target tree in this research. However
+% optimization problems in TPL, akin to problems in
+% Section~\ref{sec:optcop} for COP, is outside the scope of this thesis.
+As we saw earlier by the result in \cite{nsnrs09}, pairwise
+intersection cardinality preservation is necessary and sufficient for
+an interval assignment to be feasible for a given
+hypergraph\footnote{A {\em hypergraph} is an alternate representation
+  of a set system and will be used through out this section.} and thus
+is a characterization for COP. In our work we extend this
+characterization and find that trio-wise intersection cardinality
+preservation makes a tree path labeling (TPL) feasible, which is a
+generalization of the COP problem. This problem, {\FTPL}\footnote{A
+  {\em tree path labeling} $\cl$ is a bijection of paths from the
+  target tree $T$ to the hyperedges in given hypergraph $\cF$.}, is
+defined as follows.\\
+
+\begin{minipage}[h]{5in}
+  % \begin{singlespace}
+  {\large \FTPL}\\
+  % \hline
+  \begin{tabular}[t]{l|l}
+    {\tt Input} & 
+    \begin{minipage}[t]{\probdefwidth}
+      A hypergraph $\cF$ with vertex set $U$, a tree $T$ and a
+      bijection $\cl$~$:$~$\cF \rightarrow \cP$, where $\cP$ is a set
+      of paths from $T$.\\
+    \end{minipage}\\
+    {\tt Question} &
+    \begin{minipage}[t]{\probdefwidth}
+      Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
+      such that $\phi$ when applied on any hyperedge in $\cF$ will
+      give
+      the path mapped to it by the given tree path labeling $\cl$.\\
+      { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every hyperedge
+        $S \in \cF$.}
+    \end{minipage}\\
+  \end{tabular}
+  % \end{singlespace}
+\end{minipage}\\
+
+We give a necessary and sufficient condition for {\FTPL} to output in
+affirmative. This characterization can be checked in polynomial time.
+The most interesting consequence is that in our constructive
+procedure, it is sufficient to iteratively check if three-way
+intersection cardinalities are preserved.  In other words, in each
+iteration, it is sufficient to check if the intersection of any three
+hyperedges is of the same cardinality as the intersection of the
+corresponding paths.  Thus this generalizes the well studied question
+of the feasible interval assignment problem which is the special case
+when the target tree $T$ is simply a path \cite{wlh02,nsnrs09}.
+
+Aside from checking if a given TPL is feasible, we also solve the
+problem of computing a feasible TPL for a given hypergraph and target
+tree, if one exists. This problem, {\CFTPL}, is defined as follows. \\
+
+\begin{minipage}[h]{5in}
+  % \begin{singlespace}
+  {\large \CFTPL}\\
+  % \hline
+  \begin{tabular}[t]{l|l}
+    {\tt Input} & 
+    \begin{minipage}[t]{\probdefwidth}
+      A hypergraph $\cF$ with vertex set $U$ and a tree $T$.\\
+    \end{minipage}\\
+
+    {\tt Question} &
+    \begin{minipage}[t]{\probdefwidth}
+      Does there exist a set of paths $\cP$ from $T$ and a bijection
+      $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns {\bf
+        true} on $(\cF, T, \cl)$.
+    \end{minipage}\\
+  \end{tabular}
+  % \end{singlespace}
+\end{minipage}\\
+
+\def \CFTPLKTREE {\sc Compute $k$-subdivided Star Path Labeling}
+
+We present a polynomial time algorithm for {\CFTPL} when the target
+tree $T$ belongs to a special class of trees called {\em
+  $k$-subdivided stars} and when the hyperedges in the hypergraph
+$\cF$ have at most $k+2$ vertices. The examples of $k$-subdivided
+stars given in Figure~\ref{fig:ksubstar} makes the definition clear.
+
+%This problem, {\CFTPLKTREE}, is defined as follows. \\
+\begin{figure}[htbp]
+  \centering
+  (a) \includegraphics[scale=0.4]{../img/star.pdf}
+  (b) \includegraphics[scale=0.4]{../img/kstar.pdf}
+  \caption{\figtabsize Examples of $k$-subdivided stars. (a) $k = 0$ (b) $k = 2$ }
+  \label{fig:ksubstar}
+\end{figure}
+
+\begin{minipage}[h]{5in}
+  % \begin{singlespace}
+  {\large \CFTPLKTREE}\\
+  % \hline
+  \begin{tabular}[t]{l|l}
+    {\tt Input} & 
+    \begin{minipage}[t]{\probdefwidth}
+      A hypergraph $\cF$ with vertex set $U$ such that every hyperedge
+      $S \in \cF$ is of cardinality atmost $k+2$ and a $k$-subdivided
+      star $T$.\\
+    \end{minipage}\\
+
+    {\tt Question} &
+    \begin{minipage}[t]{\probdefwidth}
+      Does there exist a set of paths $\cP$ from $T$ and a bijection
+      $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns {\bf
+        true} on $(\cF, T, \cl)$.
+    \end{minipage}\\
+  \end{tabular}
+  % \end{singlespace}
+\end{minipage}\\
+
+In spite of this
+restricted case we consider, we believe that our results are of
+significant interest in understanding the nature of {\sc Graph
+  Isomorphism} which is polynomial time solvable in interval graphs
+while being hard on path graphs\cite{kklv10}.
+
+\tnote{give gyan on why ktree is so.}
+
+We also have an algorithm for solving {\CFTPL} with no restrictions
+which runs in exponential time and explores a partial order theory of
+matrices related to {FTPL}. This po theory gives interesting insight
+into the problem of finding path labeling from $T$ by decomposing it
+into smaller subproblems of finding path labeling of subsets
+of $\cF$ from subtrees of $T$.\tnote[TH]{The connection
+of TPL to graph isomorphism will be made later in the document}
 %\end{enumerate}
@@
 
-\subsection{Characterization of Feasible Tree Path  Labelings} \tnote[TH]{UNCOMMENT all proofs}
-\label{sec:feasible}
-
-In this section we give an algorithmic characterization of a
-feasibility of tree path labeling.
-% Consider a path labeling $\cl: \cF \rightarrow \cP$ for set system $\cF$
-% and path system $\cP$ on the given tree $T$. We
-% call $\cl$ an {\em Intersection Cardinality Preserving Path Labeling
-%   (ICPPL)} if it has the following properties.
-Consider a path labeling $(\cF, \cl)$ on the
-given tree $T$. We call $(\cF, \cl)$ an {\em Intersection Cardinality
-  Preserving Path Labeling (ICPPL)} if it has the following
-properties.
-
-% \begin{enumerate}[{\icpplpr\ }i. ]
-% \item \label{pr:i} $|S| = |\cl(S)|$, \hfill for all $S \in \cF$
-%   % \vspace{\topshrink}
-% \item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$, \hfill
-%   for all distinct $S_1, S_2 \in \cF$
-%   % \vspace{\topshrink}
-% \item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
-%   \cap \cl(S_3)|$, \hfill for all distinct $S_1, S_2, S_3 \in \cF$
+% \subsection{Characterization of Feasible Tree Path  Labelings} \tnote[TH]{UNCOMMENT all proofs}
+% \label{sec:feasible} 
+% \vspace{\secfirstparatrim}
+
+% In this section we give an algorithmic characterization of a
+% feasibility of tree path labeling.
+% % Consider a path labeling $\cl: \cF \rightarrow \cP$ for set system $\cF$
+% % and path system $\cP$ on the given tree $T$. We
+% % call $\cl$ an {\em Intersection Cardinality Preserving Path Labeling
+% %   (ICPPL)} if it has the following properties.
+% Consider a path labeling $(\cF, \cl)$ on the
+% given tree $T$. We call $(\cF, \cl)$ an {\em Intersection Cardinality
+%   Preserving Path Labeling (ICPPL)} if it has the following
+% properties.
+
+% % \begin{enumerate}[{\icpplpr\ }i. ]
+% % \item \label{pr:i} $|S| = |\cl(S)|$, \hfill for all $S \in \cF$
+% %   % \vspace{\topshrink}
+% % \item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$, \hfill
+% %   for all distinct $S_1, S_2 \in \cF$
+% %   % \vspace{\topshrink}
+% % \item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
+% %   \cap \cl(S_3)|$, \hfill for all distinct $S_1, S_2, S_3 \in \cF$
+% % \end{enumerate}
+
+% \begin{table}[h]
+%   \centering
+%   \begin{tabular}[l]{rrcll}
+%     Pr i. & \label{xpr:i} $|S| $&$=$&$ |\cl(S)|$ & $\forall$ $S \in \cF$\\
+%     Pr ii. & \label{xpr:ii}$|S_1 \cap S_2| $&$=$&$ |\cl(S_1) \cap
+%     \cl(S_2)|$ &  $\forall$ distinct $S_1, S_2 \in \cF$\\
+%     Pr iii. & \label{xpr:iii}$|S_1 \cap S_2 \cap S_3| $&$=$&$
+%     |\cl(S_1) \cap \cl(S_2) \cap \cl(S_3)|$ & $\forall$ distinct $S_1,
+%     S_2, S_3 \in \cF$ 
+%   \end{tabular}
+%   \caption{ICPPL properties}
+%   \label{tab:icpplpr}
+% \end{table}
+
+% \noindent
+% The following lemma is useful in subsequent arguments. 
+% \begin{lemma}
+%   \label{lem:setminuscard}
+%   If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
+%   (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
+%   \cl(S_3))|$.
+% \end{lemma}
+
+% % \begin{proof}%[Proof of Lemma~\ref{lem:setminuscard}]
+% %   Let $P_i = \cl(S_i)$, for all $i \in \set{1,2,3}$. We have $|S_1
+% %   \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| = |S_1
+% %   \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to properties
+% %   (\ref{pr:ii}) and (\ref{pr:iii}) of ICPPL, $|S_1 \cap S_2| - |S_1
+% %   \cap S_2 \cap S_3| = |P_1 \cap P_2| - |P_1 \cap P_2 \cap P_3| =
+% %   |(P_1 \cap P_2) \setminus P_3| = |P_1 \cap (P_2 \setminus
+% %   P_3)|$. Thus lemma is proved. \qed
+% % \end{proof}
+
+% In the remaining part of this section we show that $(\cF, \cl)$ is
+% feasible if and only if it is an ICPPL and
+% Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
+% function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two
+% levels of filtering of $(\cF, \cl)$ to make it simpler while retaining
+% the set of isomorphisms, if any, between $\cF$ and $\cF^\cl$.
+% % One direction of this claim isclear: that if a path
+% % labeling is feasible, then all intersection cardinalities are
+% % preserved, i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
+% % \annote{has no premature exit condition hence any input will go
+% %   through it}{Prove that the filtered sets has ICPPL \iff input PL
+% %   has ICPPL?}. Algorithm~\ref{leafasgn} has an exit condition at
+% % line~\ref{xempty}. It can be easily verified that $X$ cannot be
+% % empty if $\cl$ is a feasible path labeling. The reason is that a
+% % feasible path labeling has an associated bijection between
+% % $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that
+% % the sets map to paths, ``preserving'' the path labeling.  The rest
+% % of the section is devoted to constructively proving that it is
+% % sufficient for a path labeling to be an ICPPL and pass the two
+% % filtering algorithms.  To describe in brief, the constructive
+% % approaches refine an ICPPL iteratively, such that at the end of each
+% % iteration we have a ``filtered'' path labeling, and finally we have
+% % a path labeling that defines a family of bijections from $supp(\cF)$
+% % to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.
+
+% First, we present Algorithm~\ref{perms} called {\tt filter\_1}
+% % , and
+% % prove its correctness.  
+% This algorithm refines the path labeling by
+% processing pairs of paths in $\cF^\cl$ that share a leaf until no two
+% paths in the new path labeling share any leaf.
+% \tnote{write a note summarizing filter 1}
+% % \begin{algorithm}[h]
+% %   \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
+% %   \label{perms}
+% %   \begin{algorithmic}[\lndisplay]
+% %     \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$\\
+% %     \STATE $j \assign 1$\\
+% %     \WHILE {there is $S_1, S_2 \in \cF_{j-1}$ such that
+% %       $\cl_{j-1}(S_1)$ and $\cl_{j-1}(S_2)$ have a common leaf in
+% %       $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
+% %     \{S_1, S_2\})
+% %     \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1 \}$ \label{setbreak} 
+% %     \COMMENT {Remove $S_1$, $S_2$ and add the ``filtered'' sets}
+% %     \STATE {\bf for} every $S \in \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
+% %     S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\
+
+% %     \STATE $\cl_j(S_1 \cap S_2) \assign \cl_{j-1}(S_1) \cap
+% %     \cl_{j-1}(S_2)$
+% %     \COMMENT {Carry forward the path labeling for all existing sets other than
+% %       $S_1$, $S_2$}
+% %     \STATE $\cl_j(S_1 \setminus S_2) \assign \cl_{j-1}(S_1) \setminus
+% %     \cl_{j-1}(S_2)$ 
+% %     \COMMENT {Define path labeling for new sets}
+% %     \STATE $\cl_j(S_2 \setminus S_1) \assign \cl_{j-1}(S_2) \setminus
+% %     \cl_{j-1}(S_1)$
+
+% %     \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of ICPPL}
+% %     \label{ln:3waycheck}
+% %     \STATE {\bf exit} \label{ln:exit1} \\
+% %     \ENDIF
+
+% %     \STATE $j \assign j+1$\\
+% %     \ENDWHILE
+% %     \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
+% %     \RETURN $(\cF', \cl')$
+% %   \end{algorithmic}
+% % \end{algorithm}
+
+% \begin{lemma} 
+%  \label{lem:feasible} 
+%  In Algorithm~\ref{perms}, if input $(\cF, \cl)$ is a feasible path
+%  assignment then at the end of $j$th iteration of the {\bf while}
+%  loop, $j \ge 0$, $(\cF_j, \cl_j)$ is a feasible path assignment.
+% \end{lemma}
+
+% % \begin{proof}%[Proof of Lemma~\ref{lem:feasible}]
+% %   We will prove this by mathematical induction on the number of
+% %   iterations. The base case $(\cF_0, \cl_0)$ is feasible since it is
+% %   the input itself which is given to be feasible. Assume the lemma is
+% %   true till $j-1$th iteration. i.e. every hypergraph isomorphism
+% %   $\phi: supp\left(\cF_{j-1}\right) \rightarrow V\left(T \right)$ that
+% %   defines $(\cF, \cl)$'s feasibility, is such that the induced path
+% %   labeling on $\cF_{j-1}$, $\cl_{\phi[{\cF_{j-1}}]}$ is equal to
+% %   $\cl_{j-1}$. We will prove that $\phi$ is also the bijection that
+% %   makes $(\cF_j, \cl_j)$ feasible. Note that $supp(\cF_{j-1}) =
+% %   supp(\cF_{j})$ since the new sets in $\cF_j$ are created from basic
+% %   set operations to the sets in $\cF_{j-1}$. For the same reason and
+% %   $\phi$ being a bijection, it is clear that when applying the $\phi$
+% %   induced path labeling on $\cF_j$, $ \cl_{\phi[{\cF_{j}}]}(S_1
+% %   \setminus S_2) = \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
+% %   \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Now observe that $ \cl_j(S_1
+% %   \setminus S_2) = \cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2) =
+% %   \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
+% %   \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Thus the induced path labeling
+% %   $\cl_{\phi[{\cF_{j}}]} = \cl_{j}$. Therefore lemma is proved.  \qed
+% % \end{proof}
+
+% \begin{lemma}
+%   \label{lem:invar1} In Algorithm~\ref{perms}, at the end of $j$th
+%   iteration, $j \ge 0$, of the {\bf while} loop, the following
+%   invariants are maintained.
+% %   \begin{enumerate}[I {\ }] %\vspace{\topshrink}
+% %   \item $\cl_j(R)$ is a path in $T$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
+% %     \ \ \ \ \ \ \ \ \ \ for all $R \in \cF_j$%\vspace{\topshrink}
+% %   \item $|R| = |\cl_j(R)|$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
+% %     \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $R \in
+% %     \cF_j$%\vspace{\topshrink}
+% %   \item $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, \ \ \ \ \ \ \ \ \ \
+% %     \ \ \ \ \ \ \ \ \ \ for all $R, R' \in \cF_j$%\vspace{\topshrink}
+% %   \item $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
+% %     \cl_j(R'')|$, \ \ \ for all $R, R', R'' \in \cF_j$
+% %   \end{enumerate}
+
+%   \begin{tabular}[h]{lll}
+%     I & $\cl_j(R)$ is a path in $T$, & for all $R \in \cF_j$\\ 
+%     II  & $|R| = |\cl_j(R)|$, & for all $R \in \cF_j$ \\ 
+%     III  & $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, & for all $R, R'
+%     \in \cF_j$\\ 
+%     IV  & $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
+%     \cl_j(R'')|$, & for all $R, R', R'' \in \cF_j$ 
+%   \end{tabular}
+% \end{lemma}
+
+% % \begin{proof}
+% %  Proof is by induction on the number of iterations, $j$. In this
+% %   proof, the term ``new sets'' will refer to the sets added to $\cF_j$
+% %   in $j$th iteration in line~\ref{setbreak} of Algorithm~\ref{perms},
+% %   $S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1$ and its
+% %   images in $\cl_j$ where $\cl_{j-1}(S_1)$
+% %   and $\cl_{j-1}(S_2)$ intersect and share a leaf.\\
+% %   The invariants are true in the base case $(\cF_0, \cl_0)$, since it
+% %   is the input ICPPL.  Assume the lemma is true till the $j-1$th
+% %   iteration. Let us consider the possible cases for each of the above invariants for
+% %   the $j$th iteration.
+
+% %   \begin{enumerate}[\textreferencemark]
+% %   \item {\em Invariant} I/II
+% %     \begin{enumerate}[{I/II}a $|$] % \textbullet 
+% %     \item {\em $R$ is not a new set.} It is in $\cF_{j-1}$. Thus
+% %       trivially true by induction hypothesis.
+% %     \item {\em $R$ is a new set.} If $R$ is in $\cF_{j}$ and not in
+% %       $\cF_{j-1}$, then it must be one of the new sets added in
+% %       $\cF_j$. In this case, it is clear that for each new set, the
+% %       image under $\cl_j$ is a path since by definition the chosen
+% %       sets $S_1$, $S_2$ are from $\cF_{j-1}$ and due to the while loop
+% %       condition, $\cl_{j-1}(S_1)$, $\cl_{j-1}(S_2)$ have a
+% %       common leaf. Thus invariant I is proved.\\
+% %       Moreover, due to induction hypothesis of invariant III and the
+% %       definition of $l_j$ in terms of $l_{j-1}$, invariant II is
+% %       indeed true in the $j$th iteration for any of the new sets.  If
+% %       $R = S_1 \cap S_2$, $|R| = |S_1 \cap S_2| = |\cl_{j-1}(S_1) \cap
+% %       \cl_{j-1}(S_2)| = |\cl_j(S_1 \cap S_2)| = |\cl_j(R)|$.
+% %       If $R = S_1 \setminus S_2$, $|R| = |S_1 \setminus S_2| = |S_1| -
+% %       |S_1 \cap S_2| = |\cl_{j-1}(S_1)| - |\cl_{j-1}(S_1) \cap
+% %       \cl_{j-1}(S_2)| = |\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)| =
+% %       |\cl_j(S_1 \setminus S_2)|
+% %       = |\cl_j(R)|$. Similarly if $R = S_2 \setminus S_1$.\\
+% %     \end{enumerate}
+% %   \item {\em Invariant} III
+% %     \begin{enumerate}[{III}a $|$]
+% %     \item {\em $R$ and $R'$ are not new sets.} It is in
+% %       $\cF_{j-1}$. Thus trivially true by induction hypothesis.
+% %     \item {\em Only one, say $R$, is a new set.} Due to invariant IV
+% %       induction hypothesis, Lemma~\ref{lem:setminuscard} and
+% %       definition of $\cl_j$, it follows that invariant III is true no
+% %       matter which of the new sets $R$ is equal to. If $R = S_1 \cap
+% %       S_2$, $|R \cap R'| = |S_1 \cap S_2 \cap R'| = |\cl_{j-1}(S_1)
+% %       \cap \cl_{j-1}(S_2) \cap \cl_{j-1}(R')| = |\cl_j(S_1 \cap S_2)
+% %       \cap \cl_j(R')| = |\cl_j(R) \cap \cl_j(R')|$.  If $R = S_1
+% %       \setminus S_2$, $|R \cap R'| = |(S_1 \setminus S_2) \cap R'| =
+% %       |(\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)) \cap \cl_{j-1}(R')| =
+% %       |\cl_{j}(S_1 \cap S_2) \cap \cl_{j}(R')| = |\cl_{j}(R) \cap
+% %       \cl_{j}(R')|$. Similarly, if $R = S_2 \setminus
+% %       S_1$. Note $R'$ is not a new set.\\
+
+% %     \item {\em $R$ and $R'$ are new sets.} By definition, the new
+% %       sets and their path images in path label $\cl_j$ are disjoint so
+% %       $|R \cap R'| = |\cl_j(R) \cap \cl_j(R)| = 0$. Thus case proved.
+% %     \end{enumerate}
+% %   \item {\em Invariant} IV
+
+% %     Due to the condition in line~\ref{ln:3waycheck}, this invariant is
+% %     ensured at the end of every iteration.
+% % %     \begin{enumerate} [{Case 3.}1:]
+% % %     \item {\em $R$, $R'$ and $R''$ are not new sets.} Trivially
+% % %       true by induction hypothesis.
+% % %     \item {\em Only one, say $R$, is a new set.}
+% % %       If $R = S_1 \cap S_2$,  from Lemma~\ref{lem:fourpaths} and
+% % %       invariant III hypothesis,  this case is proven. Similarly if $R$
+% % %       is any of the other new  sets, the case is proven by also using
+% % %       Lemma ~\ref{lem:setminuscard}.
+% % %     \item {\em At least two of $R, R', R''$ are new sets.}
+% % %       The new sets are disjoint hence this case is vacuously true.
+% % %     \end{enumerate}
+% %   \end{enumerate} \qed
+% % %\vspace{-6mm}
+
+% % \end{proof}
+
+% \begin{lemma}
+%   \label{lem:noexit1}
+%   If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{perms} is
+%   feasible, then the set of hypergraph isomorphism functions that
+%   defines $(\cF, \cl)$'s feasibility is the same as the set that
+%   defines $(\cF_j, \cl_j)$'s feasibility, if any.  Secondly, for any
+%   iteration $j > 0$ of the {\em \bf while} loop, the {\em \bf exit}
+%   statement in line~\ref{ln:exit1} will not execute.
+% \end{lemma}
+% \tnote[TH]{UNCOMMENT}
+% % \begin{proof}
+% %   Since $(\cF,\cl)$ is feasible, by Lemma~\ref{lem:feasible}
+% %   $(\cF_j,\cl_j)$ for every iteration $j > 0$ is feasible.  % Therefore,
+% % %   every hypergraph isomorphism $\phi: supp(\cF) \rightarrow V(T)$ that
+% % %   induces $\cl$ on $\cF$ also induces $\cl_{j-1}$ and $\cl_{j}$ on
+% % %   $\cF_{j-1}$ and $\cF_{j}$ respectively, i.e., $\cl_{\phi[\cF_{j-1}]}
+% % %   = \cl_{j-1}$ and $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
+% % %   that for all $x \in supp(\cF)$, for all $v \in V(T)$ the following
+% % %   hold true.
+% %   Also, every hypergraph isomorphism $\phi: supp(\cF) \rightarrow
+% %   V(T)$ that induces $\cl$ on $\cF$ also induces $\cl_{j}$ on
+% %   $\cF_{j}$, i.e., $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
+% %   that for all $x \in supp(\cF)$, for all $v \in V(T)$, if $(x,v) \in
+% %   \phi$ then $v \in \cl_{j}(S)$ for all $S \in \cF_{j}$ such that $x
+% %   \in S$.
+% % % the following
+% % %   hold true.
+% % %   \begin{enumerate}[i. ]
+% % %   \item If $(x,v) \in \phi$ then $v \in \cl_{j-1}(S)$ for all $S \in
+% % %     \cF_{j-1}$ such that $x \in S$.
+% % %   \item If $(x,v) \in \phi$ then $v \in \cl_{j}(S)$ for all $S \in
+% % %     \cF_{j}$ such that $x \in S$
+% % %   \end{enumerate}
+% %   In other words, filter 1 outputs a filtered path labeling that
+% %   ``preserves''
+% %   hypergraph isomorphisms of the original path labeling.\\
+% %   Secondly, line~\ref{ln:exit1} will execute \iff the exit condition in
+% %   line~\ref{ln:3waycheck}, i.e. failure of three way intersection
+% %   preservation, becomes true in any iteration of the {\em \bf while}
+% %   loop.  Due to Lemma~\ref{lem:invar1} Invariant IV, the exit
+% %   condition does not occur if the input is a feasible ICPPL.\qed
+
+% % %   such that $\phi(x) = v$ where $v$ is the leaf considered in the
+% % %   first iterations of while. Clearly, $\phi$ is a renaming of
+% % %   vertices in hypergraph $\cF$ to those in hypergraph $\cF^\cl$. Thus
+% % %   the following facts can be observed in every iteration of the loop.
+
+% % %   \begin{enumerate}[\hspace{2mm}i. ] \vspace{\topshrink}
+% % %   \item all intersection cardinalities are preserved in this path
+% % %     labeling \vspace{\topshrink}
+% % %   \item element $x$ is exclusive in a hyperedge in $\cF$ since $v$ is
+% % %     exclusive in a hyperedge in $\cF^\cl$.
+% % %   \end{enumerate}
+
+% % %   Thus the exit condition is never rendered true after $x$ and $v$ are
+% % %   removed from their respective hyperedges. \qed
+
+% % % \noindent
+% % % This proof uses mathematical induction on the number
+% % %   of iterations $j$, $j \ge 0$, of the loop that executed
+% % %   without exiting. The base case, $j = 0$ is obviously true since the
+% % %   input is an ICPPL and the exit condition cannot hold true due to
+% % %   ICPPL property (iii).  Assume the algorithm executes till the end
+% % %   of $j-1$th iteration without exiting at line
+% % %  ~\ref{ln:3waycheck}. Consider the $j$th iteration. From Lemma
+% % %  ~\ref{lem:feasible} we know that $(\cF_j, \cl_j)$ and $(\cF_{j-1},
+% % %   \cl_{j-1})$ are feasible\remove[AS]{and from the proof in lemma
+% % %     lem:invar1 we know that $(\cF_{j-1}, \cl_{j-1})$ satisfies all the
+% % %     invariants defined in the lemma}.  Thus there exists a bijection
+% % %   $\phi: supp(\cF) \rightarrow V(T)$ such that the induced path
+% % %   % labeling on $\cF_{j-1}$ $\cl_{\phi[\cF_{j-1}]} = \cl_{j-1}$.
+% % %   labeling on $\cF_{j}$, $\cl_{\phi[\cF_{j}]}$ and on $\cF_{j-1}$,
+% % %   $\cl_{\phi[\cF_{j-1}]}$ are equal to $\cl_{j}$ and $\cl_{j-1}$
+% % %   respectively.  We need to prove that for any $R, R', R'' \in
+% % %   \cF_{j}$, $|R \cap R' \cap R''| = |\cl_j(R) \cap \cl_j(R') \cap
+% % %   \cl_j(R'')|$.
+% % %   The following are the possible cases that could arise. From argument
+% % %   above, $|\cl_j(R) \cap \cl_j(R') \cap \cl_j(R'')| =
+% % %   |\cl_{\phi[\cF_{j}]}(R) \cap \cl_{\phi[\cF_{j}]} (R') \cap
+% % %   \cl_{\phi[\cF_{j}]} (R'')|$
+
+% % %   \begin{enumerate}[a $|$]
+% % %   \item {\em None of the sets are new. $R, R', R'' \in \cF_{j-1}$.}
+% % %     We know $(\cF_{j-1}, \cl_{j-1})$ is feasible. Thus $|R \cap R'
+% % %     \cap R''| = |\cl_{j-1}(R) \cap \cl_{j-1}(R') \cap \cl_{j-1}(R'')|
+% % %     = |\cl_{j}(R) \cap \cl_{j}(R') \cap \cl_{j}(R'')|$.
+% % %   \item {\em Only one, say $R$, is a new set.}  Let $R = S_1 \cap S_2$
+% % %     ($S_1, S_2$ are defined in the proof of lemma
+% % %    ~\ref{lem:invar1}). Now we have $|R \cap R' \cap R''| = |S_1 \cap
+% % %     S_2 \cap R' \cap R''| = |\cl_{j-1}(S_1) \cap \cl_{j-1}(S_2) \cap
+% % %     \cl_{j-1}(R') \cap \cl_{j-1}(R'')| = |\cl_{j}(R) \cap \cl_{j}(R')
+% % %     \cap \cl_{j}(R'')|$. Thus proven. If $R$ is any of the other new
+% % %     sets, the same claim can be verified using lemma
+% % %    ~\ref{lem:setminuscard}.
+% % %     % \item []{\bf Case 3:}
+% % %   \item {\em At least two of $R, R', R''$ are new sets.}  The new sets
+% % %     are disjoint hence this case is vacuously true.
+% % %   \end{enumerate}
+% % %   \qed
+% % %   \tnote[E2]{remove the induction proof. just text saying x and v are
+% % %     exclusive in these sets therefore the intersection cardinalities
+% % %     don't change thus all invariants are still true} 
+% % \end{proof}
+
+% As a result of Algorithm~\ref{perms} each leaf $v$ in $T$
+% is such that there is exactly one set in $\cF$ with $v$ as a vertex in
+% the path assigned to it.  In Algorithm~\ref{leafasgn} we identify
+% elements in $supp(\cF)$ whose images are leaves in a hypergraph
+% isomorphism if one exists.  Let $S \in \cF$ be such that $\cl(S)$ is a
+% path with leaf and $v \in V(T)$ is the unique leaf incident on it.  We
+% define a new path labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) =
+% \set{v}$ where $x$ an arbitrary element from $S \setminus \bigcup_{\hS
+%   \ne S} \hS$. In other words, $x$ is an element present in no other
+% set in $\cF$ except $S$. This is intuitive since $v$ is present in no
+% other path image under $\cl$ other than $\cl(S)$.  The element $x$ and
+% leaf $v$ are then removed from the set $S$ and path $\cl(S)$
+% respectively. After doing this for all leaves in $T$, all path images
+% in the new path labeling $\cl_{new}$ except leaf labels (a path that
+% has only a leaf is called the {\em leaf label} for the corresponding
+% single element hyperedge or set) are paths from a new pruned tree $T_0
+% = T \setminus \{v \mid v \text{ is a leaf in }
+% T\}$. Algorithm~\ref{leafasgn} is now presented with details.
+
+% \tnote{write a note summarizing filter 1}
+% % \begin{algorithm}[h]
+% %   \caption{Leaf labeling from an ICPPL {\tt filter\_2($\cF, \cl, T$)}}
+% %   \label{leafasgn}
+% %   \begin{algorithmic}[\lndisplay]
+% %     \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$
+% %     \COMMENT {Path images are such that no two path images share a
+% %       leaf.}
+% %     \STATE $j \assign 1$\\
+% %     \WHILE {there is a leaf $v$ in $T$ and a unique $S_1 \in
+% %       \cF_{j-1}$ such that $v \in \cl_{j-1}(S_1)$ }\label{uniqueleaf}
+% %     \STATE $\cF_j \assign \cF_{j-1} \setminus \{S_1\}$\\
+% %     \STATE for all $S \in \cF_{j-1}$ such that $S \ne S_1$ set
+% %     $\cl_j(S) \assign
+% %     \cl_{j-1}(S)$\\
+% %     \STATE $X \assign S_1 \setminus \bigcup_{S \in \cF_{j-1}, S \ne S_1}S$\\
+% %     \IF{$X$ is empty} \label{xempty} \STATE {\bf exit} \label{ln:exit2} \ENDIF
+% %     \STATE $x \assign $ arbitrary element from $X$\\
+% %     \STATE $\cF_j \assign \cF_j \cup \{\{x\}, S_1 \setminus \{x\}\} $\\
+% %     \STATE $\cl_j(\{x\}) \assign \{v\}$\\
+% %     \STATE $\cl_j(S_1 \setminus \{x\}) \assign \cl_{j-1}(S_1) \setminus \{v\}$\\
+% %     \STATE $j \assign j+1$\\
+% %     \ENDWHILE
+% %     \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
+% %     \RETURN $(\cF', \cl')$
+% %   \end{algorithmic}
+% % \end{algorithm}
+
+% Suppose the input ICPPL $(\cF, \cl)$ is feasible, yet set $X$ in
+% Algorithm~\ref{leafasgn} is empty in some iteration of the {\bf while}
+% loop. This will abort our procedure of finding the hypergraph
+% isomorphism. The following lemma shows that this will not happen.
+
+% \begin{lemma}
+%   \label{lem:xnotempty}
+%   If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{leafasgn} is
+%   feasible, then for all iterations $j > 0$ of the {\em \bf while}
+%   loop, the {\em \bf exit} statement in line~\ref{ln:exit2} does not
+%   execute.
+% \end{lemma}
+
+% % \begin{proof}
+% %   Assume $X$ is empty for some iteration $j > 0$. We know that $v$ is
+% %   an element of $\cl_{j-1}(S_1)$. Since it is uniquely present in
+% %   $\cl_{j-1}(S_1)$, it is clear that $v \in \cl_{j-1}(S_1) \setminus
+% %   \bigcup_{(S \in \cF_{j-1}) \wedge (S \ne S_1)}\cl_{j-1}(S)$.  Note
+% %   that for any $x \in S_1$ it is contained in at least two sets due to
+% %   our assumption about cardinality of $X$. Let $S_2 \in \cF_{j-1}$ be
+% %   another set that contains $x$. From the above argument, we know $v
+% %   \notin \cl_{j-1}(S_2)$. Therefore there cannot exist a hypergraph
+% %   isomorphism bijection that maps elements in $S_2$ to those in
+% %   $\cl_{j-1}(S_2)$. This contradicts our assumption that the input is
+% %   feasible. Thus $X$ cannot be empty if input is ICPPL and feasible.
+% %   \qed
+% % \end{proof}
+
+% \begin{lemma}
+%   \label{lem:invar3}
+%   In Algorithm~\ref{leafasgn}, for all $j > 0$, at the end of the
+%   $j$th iteration of the {\bf while} loop the four invariants given in
+%   Lemma~\ref{lem:invar1} hold.
+% \end{lemma}
+
+% % \begin{proof}
+% %   By Lemma~\ref{lem:xnotempty} we know that set $X$ will not be empty
+% %   in any iteration of the {\em \bf while} loop if input ICPPL $(\cF,
+% %   \cl)$ is feasible and $\cl_j$ is always computed for all $j >
+% %   0$. Also note that removing a leaf from any path keeps the new path
+% %   connected. Thus invariant I is obviously true. In every iteration $j
+% %   > 0$, we remove exactly one element $x$ from one set $S$ in $\cF$
+% %   and exactly one vertex $v$ which is a leaf from one path
+% %   $\cl_{j-1}(S)$ in $T$. This is because as seen in
+% %   Lemma~\ref{lem:xnotempty}, $x$ is exclusive to $S$ and $v$ is
+% %   exclusive to $\cl_{j-1}(S)$. Due to this fact, it is clear that the
+% %   intersection cardinality equations do not change, i.e., invariants
+% %   II, III, IV remain true. On the other hand, if the input ICPPL is
+% %   not feasible the invariants are vacuously true. \qed
+% % \end{proof}
+
+% %% OLD TEXT TB DELETED?
+% % \textcolor{cyan}{
+% % \begin{lemma}
+% %   \label{lem:notfeasibleexit}
+% %   \tnote{IS THIS CORRECT?}  If input ICPPL $(\cF, \cl)$ is not
+% %   feasible, then in one of the recursive calls to Algorithm --3--, the
+% %   {\em \bf exit} statement in line x in Algorithm --1 or line y in
+% %   Algorithm --2 will get executed.
+% % \end{lemma}
+% % }
+% %% END OLD TEXT TB DELETED?
+
+% We have seen two filtering algorithms above, namely,
+% Algorithm~\ref{perms} {\tt filter\_1} and Algorithm~\ref{leafasgn}
+% {\tt filter\_2} which when executed serially respectively result in a
+% new ICPPL on the same universe $U$ and tree $T$. We also proved that
+% if the input is indeed feasible, these algorithms do indeed output the
+% filtered ICPPL. Now we present the algorithmic characterization of a
+% feasible tree path labeling by way of Algorithm~\ref{al:icppl-find-isomorph}.
+
+% Algorithm~\ref{al:icppl-find-isomorph} computes a
+% hypergraph isomorphism $\phi$ recursively using Algorithm~\ref{perms}
+% and Algorithm~\ref{leafasgn} and pruning the leaves of the input
+% tree. In brief, it is done as follows. Algorithm~\ref{leafasgn} gives
+% us the leaf labels in $\cF_2$, i.e., the elements in $supp(\cF)$ that
+% map to leaves in $T$, where $(\cF_2, \cl_2)$ is the output of
+% Algorithm~\ref{leafasgn}. All leaves in $T$ are then pruned away. The
+% leaf labels are removed from the path labeling $\cl_2$ and the
+% corresponding elements are removed from the corresponding sets in
+% $\cF_2$. This tree pruning algorithm is recursively called on the
+% altered hypergraph $\cF'$, path label $\cl'$ and tree $T'$. The
+% recursive call returns the bijection $\phi''$ for the rest of the
+% elements in $supp(\cF)$ which along with the leaf labels $\phi'$ gives
+% us the hypergraph isomorphism $\phi$.  The following lemma formalizes
+% the characeterization of feasible path labeling.
+
+% \tnote{write a note summarizing filter 1}
+% % \begin{algorithm}[h]
+% %   \caption{{\tt get-hypergraph-isomorphism($\cF, \cl, T$)}}
+% %   \label{al:icppl-find-isomorph}
+% %   \begin{algorithmic}[\lndisplay]
+
+% %     \IF{$T$ is empty}
+% %     \RETURN $\emptyset$\\
+% %     \ENDIF
+% %     \STATE $L \assign \{v \mid v \text{ is a leaf in }      T\}$\\
+% %     \STATE $(\cF_1, \cl_1) \assign$ {\tt filter\_1($\cF, \cl,
+% %       T$)}\\
+% %     \STATE $(\cF_2, \cl_2) \assign$ {\tt filter\_2($\cF_1,
+% %       \cl_1, T$)}\\
+
+% %     \STATE $(\cF', \cl') \assign (\cF_2, \cl_2)$\\
+% %     \STATE $\phi' \leftarrow \emptyset$
+
+% %     \FOR {every $v \in L$}
+% %     \STATE $\phi'(x) \assign v$ where $x \in \cl_2^{-1}(\{v\})$
+% %     \COMMENT {Copy the leaf labels to a one to one function $\phi':
+% %       supp(\cF) \rightarrow L$
+% %       }\\
+% %     \STATE Remove $\{x\}$ and $\{v\}$ from $\cF'$, $\cl'$  appropriately\\
+% %     \ENDFOR
+
+% %     \STATE $T' \assign T \setminus L$
+
+% %     \STATE $\phi'' \assign$ {\tt
+% %       get-hypergraph-isomorphism($\cF', \cl', T'$)}
+% %     \STATE $\phi \assign \phi'' \cup \phi'$ \\
+% %     \RETURN $\phi$
+% %   \end{algorithmic}
+% % \end{algorithm}
+
+% \begin{lemma}
+%   \label{lem:hyperiso}  %{lem:perm}
+%   If $(\cF, \cl)$ is an ICPPL from a tree $T$ and
+%   Algorithm~\ref{al:icppl-find-isomorph}, {\tt
+%     get- hypergraph- isomorphism ($\cF, \cl, T$)} returns a non-empty
+%   function, then there exists a hypergraph isomorphism $\phi :
+%   supp(\cF) \rightarrow V(T)$ such that the $\phi$-induced tree path
+%   labeling is equal to $\cl$ or $\cl_\phi = \cl$.
+% \end{lemma}
+% % \begin{proof}
+% %   It is clear that in the end of every recursive call to
+% %   Algorithm~\ref{al:icppl-find-isomorph}, the function $\phi'$ is one
+% %   to one involving all the leaves in the tree passed as input to that
+% %   recursive call. Moreover, by Lemma~\ref{lem:noexit1} and
+% %   Lemma~\ref{lem:xnotempty} it is consistent with the tree path
+% %   labeling $\cl$ passed. The tree pruning is done by only removing
+% %   leaves in each call to the function and is done till the tree
+% %   becomes empty. Thus the returned function $\phi: supp(\cF)
+% %   \rightarrow V(T)$ is a union of mutually exclusive one to one
+% %   functions exhausting all vertices of the tree. In other words, it is
+% %   a bijection from $supp(\cF)$ to $V(T)$ inducing the given path
+% %   labeling $\cl$ and thus a hypergraph isomorphism. \qed
+% % \end{proof}
+
+% \begin{theorem}
+%   \label{th:charac}
+%   A path labeling $(\cF, \cl)$ on tree $T$ is feasible \iff it is an
+%   ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
+%   T)$ as input returns a non-empty function.
+% \end{theorem}
+% % \begin{proof}
+% %   From Lemma~\ref{lem:hyperiso}, we know that if $(\cF, \cl)$ is an
+% %   ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
+% %   T)$ as input returns a non-empty function, $(\cF, \cl)$ is feasible.
+% %   Now consider the case where $(\cF, \cl)$ is feasible. i.e. there
+% %   exists a hypergraph isomorphism $\phi$ such that $\cl_\phi =
+% %   \cl$. Lemma~\ref{lem:noexit1} and Lemma~\ref{lem:xnotempty} show us
+% %   that filter 1 and filter 2 do not exit if input is feasible. Thus
+% %   Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
+% %   function.\qed
+% % \end{proof}
+
+% \subsection*{ICPPL when given tree is a path}
+% \label{sec:icpplicpia}
+% \vspace{\secfirstparatrim}  
+% Consider a special case of ICPPL with the following properties when
+% the tree $T$ is a path.  Hence, all path labels are can be viewed as
+% intervals assigned to the sets in $\cF$.  It is shown, in
+% \cite{nsnrs09}, that the filtering algorithms outlined above need only
+% preserve pairwise intersection cardinalities, and higher level
+% intersection cardinalities are preserved by the Helly Property of
+% intervals.  Consequently, the filter algorithms do not need to ever
+% evaluate the additional check to {\em \bf exit}.
+% %\begin{enumerate}
+% %\item Given tree $T$ is a path. Hence, all path labels are interval labels.
+% %\item Only pairwise intersection cardinality
+% %  preservation is sufficient. i.e. property (iii) in ICPPL is not enforced.
+% %\item The filter algorithms do not have {\em \bf exit} statements.
+% %\end{enumerate}
+% %This is called an Intersection Cardinality Preservation Interval
+% %Assignment (ICPIA) \cite{nsnrs09}. 
+% This structure and its algorithm is
+% used in the next section for finding tree path labeling from a
+% $k$-subdivided star due to this graph's close relationship with
+% intervals.
+
+% \subsection{Case of a Special Tree -- The $k$-subdivided star}
+% \label{sec:ksubdivstar}
+% \vspace{\secfirstparatrim}  
+% In this section we consider the problem of assigning paths from a
+% $k$-subdivided star $T$ to a given set system $\cF$ such that each set
+% $X \in \cF$ is of cardinality at most $k+2$.  Secondly, we present our
+% results only for the case when overlap graph $\bO(\cF)$ is connected.
+% %The
+% %overlap graph is well-known from the work of
+% %\cite{kklv10,nsnrs09,wlh02}.  We use the notation in
+% %\cite{kklv10}. Recall from Section~\ref{sec:prelims} that hyperedges
+% %$S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$ and $S'$
+% %have a non-empty intersection but neither of them is contained in the
+% %other. The overlap graph $\bO(\cF)$ is a graph in which the vertices
+% %correspond to the sets in $\cF$, and the vertices corresponding to the
+% %hyperedges $S$ and $S'$ are adjacent if and only if they overlap.  Note
+% %that the intersection graph of $\cF$, $\bI(\cF)$ is different from
+% %$\bO(\cF)$ and $\bO(\cF) \subseteq \bI(\cF)$.  
+% A connected component of $\bO(\cF)$ is called an overlap component of
+% $\cF$.  An interesting property of the overlap components is that any
+% two distinct overlap components, say $\cO_1$ and $\cO_2$, are such
+% that any two sets $S_1 \in \cO_1$ and $S_2 \in \cO_2$ are disjoint,
+% or, w.l.o.g, all the sets in $\cO_1$ are contained within one set in
+% $\cO_2$.  This containment relation naturally determines a
+% decomposition of the overlap components into rooted containment trees.
+% We consider the case when there is only one rooted containment tree,
+% and we first present our algorithm when $\bO(\cF)$ is connected.  It
+% is easy to see that once the path labeling to the overlap component in
+% the root of the containment tree is achieved, the path labeling to the
+% other overlap components in the rooted containment tree is essentially
+% finding a path labeling when the target tree is a path: each target
+% path is a path that is allocated to sets in the root overlap
+% component.  Therefore, for the rest of this section, $\bO(\cF)$ is a
+% connected graph. Recall that we also consider the special case when
+% all hyperedges are of cardinality at most $k+2$.  By definition, a
+% $k$-subdivided star has a central vertex which we call the {\em root},
+% and each root to leaf path is called a {\em ray}.  First, we observe
+% that by removing the root $r$ from $T$, we get a collection of $p$
+% vertex disjoint paths of length $k+1$, $p$ being the number of leaves
+% in $T$.  We denote the rays by $R_1, \ldots, R_p$, and the number of
+% vertices in $R_i$, $i \in [p]$ is $k+2$.  Let
+% $\seq{v_{i1},\ldots,v_{i(k+2)}=r}$ denote the sequence of vertices in
+% $R_i$, where $v_{i1}$ is the leaf. Note that $r$ is a common vertex to
+% all $R_i$.
+
+% \subsubsection{Description of the Algorithm}
+% \vspace{\secfirstparatrim}  
+% In this section the given hypergraph $\cF$, the $k$-subdivided star
+% and the root of the star are denoted by $\cO$, $T$ and vertex $r$,
+% respectively.  In particular, note that the vertices of $\cO$
+% correspond to the sets in $\cF$, and the edges correspond to the
+% overlap relation.
+
+% %\noindent
+% For each hyperedge $X \in \cO$, we will maintain a 2-tuple of non-negative
+%  numbers $\seq{p_1(X), p_2(X)}$.  The numbers satisfy the property that
+%  $p_1(X) + p_2(X) \leq |X|$, and at the end of path labeling, for each
+%  $X$, $p_1(X) + p_2(X) = |X|$.  This signifies the algorithm tracking
+%  the lengths of subpaths of the path assigned to $X$ from at most two
+%  rays. We also maintain another parameter called the {\em residue} of
+%  $X$ denoted by $s(X)=|X| - p_1(X)$. This signifies the residue path
+%  length that must be assigned to $X$ which must be from another
+%  ray. For instance, if $X$ is labeled a path from only one ray, then
+%  $p_1(X) = |X|$, $p_2(X) = 0$ and $s(X) = 0$.
+
+%  The algorithm proceeds in iterations, and in the $i$-th
+%  iteration, $i > 1$, a single hyperedge $X$ that overlaps with a
+%  hyperedge that has been assigned a path is considered.  At the
+%  beginning of each iteration hyperedges of $\cO$ are classifed into
+%  the following disjoint sets.
+%  \begin{enumerate}
+%  \item [$\cL_1^i$] {\em Labeled without $r$.} Those that have been
+%    labeled with a path which does not contain $r$ in one of the
+%    previous iterations.\\  $\cL_1^i = \set{ X \mid p_1(X) = |X| \text{ and
+%      } p_2(X) = 0 \text{ and } s(X) = 0, X \in \cO}$
+%  \item [$\cL_2^i$] {\em Labeled with $r$.} Those that have been labeled
+%    with two subpaths of $\cl(X)$ containing $r$ from two different rays
+%    in two previous iterations.\\ $\cL_2^i = \set{X \mid 0 < p_1\left(X\right),
+%      p_2\left(X\right) < |X|=p_1(X)+p_2(X) \text{ and } s(X) = 0, X \in \cO}$
+%    \item [$\cT_1^i$] {\em Type 1 / partially labeled.} Those that have
+%      been labeled with one path containing $r$ from a single ray in one
+%      of the previous iterations. Here, $p_1(X)$ denotes the length of
+%      the subpath of $\cl(X)$ that $X$ has been so far labeled
+%      with. \\
+%      $\cT_1^i = \set{ X \mid 0 < p_1(X) < |X| \text{ and } p_2(X) = 0
+%        \text{ and } s(X) = |X|-p_1(X), X \in \cO}$
+%    \item [$\cT_2^i$] {\em Type 2 / not labeled.} Those that have not been
+%      labeled with a path in any previous iteration.\\
+%      $\cT_2^i = \set{ X \mid p_1(X) = p_2(X) = 0 \text{ and } s(X) = |X|,
+%        X \in \cO}$
 %  \end{enumerate}
-
-\begin{table}[h]
-  \centering
-  \begin{tabular}[l]{rrcll}
-    Pr i. & \label{xpr:i} $|S| $&$=$&$ |\cl(S)|$ & $\forall$ $S \in \cF$\\
-    Pr ii. & \label{xpr:ii}$|S_1 \cap S_2| $&$=$&$ |\cl(S_1) \cap
-    \cl(S_2)|$ &  $\forall$ distinct $S_1, S_2 \in \cF$\\
-    Pr iii. & \label{xpr:iii}$|S_1 \cap S_2 \cap S_3| $&$=$&$
-    |\cl(S_1) \cap \cl(S_2) \cap \cl(S_3)|$ & $\forall$ distinct $S_1,
-    S_2, S_3 \in \cF$ 
-  \end{tabular}
-  \caption{ICPPL properties}
-  \label{tab:icpplpr}
-\end{table}
-
-\noindent
-The following lemma is useful in subsequent arguments. 
-\begin{lemma}
-  \label{lem:setminuscard}
-  If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
-  (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
-  \cl(S_3))|$.
-\end{lemma}
-
-% \begin{proof}%[Proof of Lemma~\ref{lem:setminuscard}]
-%   Let $P_i = \cl(S_i)$, for all $i \in \set{1,2,3}$. We have $|S_1
-%   \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| = |S_1
-%   \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to properties
-%   (\ref{pr:ii}) and (\ref{pr:iii}) of ICPPL, $|S_1 \cap S_2| - |S_1
-%   \cap S_2 \cap S_3| = |P_1 \cap P_2| - |P_1 \cap P_2 \cap P_3| =
-%   |(P_1 \cap P_2) \setminus P_3| = |P_1 \cap (P_2 \setminus
-%   P_3)|$. Thus lemma is proved. \qed
-% \end{proof}
-
-In the remaining part of this section we show that $(\cF, \cl)$ is
-feasible if and only if it is an ICPPL and
-Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
-function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two
-levels of filtering of $(\cF, \cl)$ to make it simpler while retaining
-the set of isomorphisms, if any, between $\cF$ and $\cF^\cl$.
-% One direction of this claim isclear: that if a path
-% labeling is feasible, then all intersection cardinalities are
-% preserved, i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
-% \annote{has no premature exit condition hence any input will go
-%   through it}{Prove that the filtered sets has ICPPL \iff input PL
-%   has ICPPL?}. Algorithm~\ref{leafasgn} has an exit condition at
-% line~\ref{xempty}. It can be easily verified that $X$ cannot be
-% empty if $\cl$ is a feasible path labeling. The reason is that a
-% feasible path labeling has an associated bijection between
-% $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that
-% the sets map to paths, ``preserving'' the path labeling.  The rest
-% of the section is devoted to constructively proving that it is
-% sufficient for a path labeling to be an ICPPL and pass the two
-% filtering algorithms.  To describe in brief, the constructive
-% approaches refine an ICPPL iteratively, such that at the end of each
-% iteration we have a ``filtered'' path labeling, and finally we have
-% a path labeling that defines a family of bijections from $supp(\cF)$
-% to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.
-
-First, we present Algorithm~\ref{perms} called {\tt filter\_1}
-% , and
-% prove its correctness.  
-This algorithm refines the path labeling by
-processing pairs of paths in $\cF^\cl$ that share a leaf until no two
-paths in the new path labeling share any leaf.
-\tnote{write a note summarizing filter 1}
-% \begin{algorithm}[h]
-%   \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
-%   \label{perms}
-%   \begin{algorithmic}[\lndisplay]
-%     \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$\\
-%     \STATE $j \assign 1$\\
-%     \WHILE {there is $S_1, S_2 \in \cF_{j-1}$ such that
-%       $\cl_{j-1}(S_1)$ and $\cl_{j-1}(S_2)$ have a common leaf in
-%       $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
-%     \{S_1, S_2\})
-%     \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1 \}$ \label{setbreak} 
-%     \COMMENT {Remove $S_1$, $S_2$ and add the ``filtered'' sets}
-%     \STATE {\bf for} every $S \in \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
-%     S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\
-
-%     \STATE $\cl_j(S_1 \cap S_2) \assign \cl_{j-1}(S_1) \cap
-%     \cl_{j-1}(S_2)$
-%     \COMMENT {Carry forward the path labeling for all existing sets other than
-%       $S_1$, $S_2$}
-%     \STATE $\cl_j(S_1 \setminus S_2) \assign \cl_{j-1}(S_1) \setminus
-%     \cl_{j-1}(S_2)$ 
-%     \COMMENT {Define path labeling for new sets}
-%     \STATE $\cl_j(S_2 \setminus S_1) \assign \cl_{j-1}(S_2) \setminus
-%     \cl_{j-1}(S_1)$
-
-%     \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of ICPPL}
-%     \label{ln:3waycheck}
-%     \STATE {\bf exit} \label{ln:exit1} \\
-%     \ENDIF
-
-%     \STATE $j \assign j+1$\\
-%     \ENDWHILE
-%     \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
-%     \RETURN $(\cF', \cl')$
-%   \end{algorithmic}
-% \end{algorithm}
-
-\begin{lemma} 
- \label{lem:feasible} 
- In Algorithm~\ref{perms}, if input $(\cF, \cl)$ is a feasible path
- assignment then at the end of $j$th iteration of the {\bf while}
- loop, $j \ge 0$, $(\cF_j, \cl_j)$ is a feasible path assignment.
-\end{lemma}
-
-% \begin{proof}%[Proof of Lemma~\ref{lem:feasible}]
-%   We will prove this by mathematical induction on the number of
-%   iterations. The base case $(\cF_0, \cl_0)$ is feasible since it is
-%   the input itself which is given to be feasible. Assume the lemma is
-%   true till $j-1$th iteration. i.e. every hypergraph isomorphism
-%   $\phi: supp\left(\cF_{j-1}\right) \rightarrow V\left(T \right)$ that
-%   defines $(\cF, \cl)$'s feasibility, is such that the induced path
-%   labeling on $\cF_{j-1}$, $\cl_{\phi[{\cF_{j-1}}]}$ is equal to
-%   $\cl_{j-1}$. We will prove that $\phi$ is also the bijection that
-%   makes $(\cF_j, \cl_j)$ feasible. Note that $supp(\cF_{j-1}) =
-%   supp(\cF_{j})$ since the new sets in $\cF_j$ are created from basic
-%   set operations to the sets in $\cF_{j-1}$. For the same reason and
-%   $\phi$ being a bijection, it is clear that when applying the $\phi$
-%   induced path labeling on $\cF_j$, $ \cl_{\phi[{\cF_{j}}]}(S_1
-%   \setminus S_2) = \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
-%   \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Now observe that $ \cl_j(S_1
-%   \setminus S_2) = \cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2) =
-%   \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
-%   \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Thus the induced path labeling
-%   $\cl_{\phi[{\cF_{j}}]} = \cl_{j}$. Therefore lemma is proved.  \qed
-% \end{proof}
-
-\begin{lemma}
-  \label{lem:invar1} In Algorithm~\ref{perms}, at the end of $j$th
-  iteration, $j \ge 0$, of the {\bf while} loop, the following
-  invariants are maintained.
-%   \begin{enumerate}[I {\ }] %\vspace{\topshrink}
-%   \item $\cl_j(R)$ is a path in $T$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-%     \ \ \ \ \ \ \ \ \ \ for all $R \in \cF_j$%\vspace{\topshrink}
-%   \item $|R| = |\cl_j(R)|$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-%     \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $R \in
-%     \cF_j$%\vspace{\topshrink}
-%   \item $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, \ \ \ \ \ \ \ \ \ \
-%     \ \ \ \ \ \ \ \ \ \ for all $R, R' \in \cF_j$%\vspace{\topshrink}
-%   \item $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
-%     \cl_j(R'')|$, \ \ \ for all $R, R', R'' \in \cF_j$
+% % \vspace{-2mm}
+% % \begin{align*}
+% %   \cO &= \cL_1^i \cup \cL_2^i \cup \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]\\
+% %   \cO_i &= \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]
+% % \end{align*}
+
+% %\noindent
+% The set $\cO_i$ refers to the set of hyperedges $\cT_1^i \cup \cT_2^i$
+% in the $i$th iteration.  Note that $\cO_1 = \cO$.  In the $i$th
+% iteration, hyperedges from $\cO_i$ are assigned paths from $T$ using
+% the following rules. Also the end of the iteration, $\cL_1^{i+1},
+% \cL_2^{i+1}, \cT_1^{i+1}, \cT_2^{i+1}$ are set to $\cL_1^{i},
+% \cL_2^{i}, \cT_1^{i}, \cT_2^{i}$ respectively, along with some
+% case-specific changes mentioned in the rules below.
+
+% %\noindent
+% \begin{enumerate}[I.]
+% \item {\bf Iteration 1:} Let $S=\{X_1,\ldots,X_s\}$ denote the
+%   super-marginal hyperedges from $\cO_1$.  If $|S|=s \neq p$, then
+%   exit reporting failure.  Else, assign to each $X_j \in S$, the path
+%   from $R_j$ such that the path contains the leaf in $R_j$.  This path
+%   is refered to as $\cl(X_j)$.  Set $p_1(X_j)=|X|, p_2(X_j)=s(X_j)=0$.
+%   Hyperedges in $S$ are not added to $\cO_2$ but are added to
+%   $\cL_1^2$ and all other hyperedges are added to
+%   $\cO_2$.
+% \item {\bf Iteration $i$:} Let $X$ be a hyperedge from $\cO_i$ such
+%   that there exists $Y \in \cL_1^i \cup \cL_2^i$ and $X \overlap
+%   Y$. Further let $Z \in \cL_1^i \cup L_2^i$ such that $Z \overlap Y$.
+%   If $X \in \cT_2^i$, and if there are multiple $Y$ candidates then
+%   any $Y$ is selected.  On the other hand, if $X \in \cT_1^i$, then
+%   $X$ has a partial path assignment, $\cl'(X)$ from a previous
+%   iteration, say from ray $R_j$. Then, $Y$ is
+%   chosen such that $X \cap Y$ has a non-empty intersection with a ray
+%   different from $R_j$.  The key things that are done in assigning a
+%   path to $X$ are as follows. The {\em end} of path $\cl(Y)$ where
+%   $\cl(X)$ would overlap is found, and then based on this the
+%   existence of a feasible assignment is decided.  It is important to
+%   note that since $X \overlap Y$, $\cl(X) \overlap \cl(Y)$ in any
+%   feasible assignment.  Therefore, the notion of the {\em end} at
+%   which $\cl(X)$ and $\cl(Y)$ overlap is unambiguous, since for any
+%   path, there are two end points.
+%   \begin{enumerate}
+%   \item \label{iendpoint} {\em End point of $\cl(Y)$ where $\cl(X)$ overlaps
+%       depends on $X \cap Z$:} If $X \cap Z
+%     \neq \emptyset$, then $\cl(X)$ has an overlap of $|X \cap Y|$ at that
+%     end of $\cl(Y)$ at which $\cl(Y)$ and $\cl(Z)$ overlap.  If $X
+%     \cap Z = \emptyset$, then $\cl(X)$ has an overlap of $|X \cap Y|$ at
+%     that end of $\cl(Y)$ where $\cl(Y)$ and $\cl(Z)$ do not intersect.
+%   \item {\em Any path of length $s(X)$ at the appropriate end contains
+%       $r$:} If $X \in \cT_1^i$ then after finding the appropriate end
+%     as in step~\ref{iendpoint} this the unique path of length $s(X)$
+%     should end at $r$.  If not, we exit reporting failure.  Else,
+%     $\cl(X)$ is computed as union of $\cl'(X)$ and this path. If any
+%     three-way intersection cardinality is violated with this new
+%     assignment, then exit, reporting failure.  Otherwise, $X$ is added
+%     to $\cL_2^{i+1}$.  On the other hand, if $X \in \cT_2^i$, then
+%     after step~\ref{iendpoint}, $\cl(X)$ or $\cl'(X)$ is unique up to
+%     the root and including it. Clearly, the vertices $\cl(X)$ or
+%     $\cl'(X)$ contains depends on $|X|$ and $|X \cap Y|$.  If any
+%     three way intersection cardinality is violated due to this
+%     assignment, exit, reporting failure.  Otherwise,
+%     $p_1(X)$ is updated as the length of the assigned path, and $s(X)
+%     = |X|-p_1(X)$.  If $s(X) > 0$, then $X$ is added to $\cT_1^{i+1}$.
+%     If $s(X)=0$, then $X$ is added to $\cL_1^{i+1}$.
+%   \item {\em The unique path of length $s(X)$ overlapping at the
+%       appropriate end of $Y$ does not contain $r$:} In this case,
+%     $\cl(X)$ is updated to include this path.  If any three way
+%     intersection cardinality is violated, exit, reporting failure.
+%     Otherwise, update $p_1(X)$ and $p_2(X)$ are appropriate, $X$ is
+%     added to $\cL_1^{i+1}$ or $\cL_2^{i+1}$, as appropriate.
 %   \end{enumerate}
-
-  \begin{tabular}[h]{lll}
-    I & $\cl_j(R)$ is a path in $T$, & for all $R \in \cF_j$\\ 
-    II  & $|R| = |\cl_j(R)|$, & for all $R \in \cF_j$ \\ 
-    III  & $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, & for all $R, R'
-    \in \cF_j$\\ 
-    IV  & $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
-    \cl_j(R'')|$, & for all $R, R', R'' \in \cF_j$ 
-  \end{tabular}
-\end{lemma}
-
-% \begin{proof}
-%  Proof is by induction on the number of iterations, $j$. In this
-%   proof, the term ``new sets'' will refer to the sets added to $\cF_j$
-%   in $j$th iteration in line~\ref{setbreak} of Algorithm~\ref{perms},
-%   $S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1$ and its
-%   images in $\cl_j$ where $\cl_{j-1}(S_1)$
-%   and $\cl_{j-1}(S_2)$ intersect and share a leaf.\\
-%   The invariants are true in the base case $(\cF_0, \cl_0)$, since it
-%   is the input ICPPL.  Assume the lemma is true till the $j-1$th
-%   iteration. Let us consider the possible cases for each of the above invariants for
-%   the $j$th iteration.
-
-%   \begin{enumerate}[\textreferencemark]
-%   \item {\em Invariant} I/II
-%     \begin{enumerate}[{I/II}a $|$] % \textbullet 
-%     \item {\em $R$ is not a new set.} It is in $\cF_{j-1}$. Thus
-%       trivially true by induction hypothesis.
-%     \item {\em $R$ is a new set.} If $R$ is in $\cF_{j}$ and not in
-%       $\cF_{j-1}$, then it must be one of the new sets added in
-%       $\cF_j$. In this case, it is clear that for each new set, the
-%       image under $\cl_j$ is a path since by definition the chosen
-%       sets $S_1$, $S_2$ are from $\cF_{j-1}$ and due to the while loop
-%       condition, $\cl_{j-1}(S_1)$, $\cl_{j-1}(S_2)$ have a
-%       common leaf. Thus invariant I is proved.\\
-%       Moreover, due to induction hypothesis of invariant III and the
-%       definition of $l_j$ in terms of $l_{j-1}$, invariant II is
-%       indeed true in the $j$th iteration for any of the new sets.  If
-%       $R = S_1 \cap S_2$, $|R| = |S_1 \cap S_2| = |\cl_{j-1}(S_1) \cap
-%       \cl_{j-1}(S_2)| = |\cl_j(S_1 \cap S_2)| = |\cl_j(R)|$.
-%       If $R = S_1 \setminus S_2$, $|R| = |S_1 \setminus S_2| = |S_1| -
-%       |S_1 \cap S_2| = |\cl_{j-1}(S_1)| - |\cl_{j-1}(S_1) \cap
-%       \cl_{j-1}(S_2)| = |\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)| =
-%       |\cl_j(S_1 \setminus S_2)|
-%       = |\cl_j(R)|$. Similarly if $R = S_2 \setminus S_1$.\\
 % \end{enumerate}
-%   \item {\em Invariant} III
-%     \begin{enumerate}[{III}a $|$]
-%     \item {\em $R$ and $R'$ are not new sets.} It is in
-%       $\cF_{j-1}$. Thus trivially true by induction hypothesis.
-%     \item {\em Only one, say $R$, is a new set.} Due to invariant IV
-%       induction hypothesis, Lemma~\ref{lem:setminuscard} and
-%       definition of $\cl_j$, it follows that invariant III is true no
-%       matter which of the new sets $R$ is equal to. If $R = S_1 \cap
-%       S_2$, $|R \cap R'| = |S_1 \cap S_2 \cap R'| = |\cl_{j-1}(S_1)
-%       \cap \cl_{j-1}(S_2) \cap \cl_{j-1}(R')| = |\cl_j(S_1 \cap S_2)
-%       \cap \cl_j(R')| = |\cl_j(R) \cap \cl_j(R')|$.  If $R = S_1
-%       \setminus S_2$, $|R \cap R'| = |(S_1 \setminus S_2) \cap R'| =
-%       |(\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)) \cap \cl_{j-1}(R')| =
-%       |\cl_{j}(S_1 \cap S_2) \cap \cl_{j}(R')| = |\cl_{j}(R) \cap
-%       \cl_{j}(R')|$. Similarly, if $R = S_2 \setminus
-%       S_1$. Note $R'$ is not a new set.\\
-
-%     \item {\em $R$ and $R'$ are new sets.} By definition, the new
-%       sets and their path images in path label $\cl_j$ are disjoint so
-%       $|R \cap R'| = |\cl_j(R) \cap \cl_j(R)| = 0$. Thus case proved.
-%     \end{enumerate}
-%   \item {\em Invariant} IV
-
-%     Due to the condition in line~\ref{ln:3waycheck}, this invariant is
-%     ensured at the end of every iteration.
-% %     \begin{enumerate} [{Case 3.}1:]
-% %     \item {\em $R$, $R'$ and $R''$ are not new sets.} Trivially
-% %       true by induction hypothesis.
-% %     \item {\em Only one, say $R$, is a new set.}
-% %       If $R = S_1 \cap S_2$,  from Lemma~\ref{lem:fourpaths} and
-% %       invariant III hypothesis,  this case is proven. Similarly if $R$
-% %       is any of the other new  sets, the case is proven by also using
-% %       Lemma ~\ref{lem:setminuscard}.
-% %     \item {\em At least two of $R, R', R''$ are new sets.}
-% %       The new sets are disjoint hence this case is vacuously true.
+
+% %\noindent 
+% \tnote[TH]{UNCOMMENT}
+% % \subsection*{Proof of Correctness and Analysis of Running Time} 
+% % It
+% % is clear that the algorithm runs in polynomial time, as at each step,
+% % at most three-way intersection cardinalities need to be checked.
+% % Further, finding super-marginal hyperedges can also be done in
+% % polynomial time, as it involves considering the overlap regions and
+% % checking if the inclusion partial order contains a single minimal
+% % element.  In particular, once the super-marginal edges are identified,
+% % each iteration involes finding the next hyperedge to consider, and
+% % testing for a path to that hyperedge.  To identify the next hyperedge
+% % to consider, we consider the breadth first layering of the hyperedges
+% % with the zeroeth layer consisting of the super-marginal hyperedges.
+% % Since $\cO$ is connected, it follows that all hyperedges of $\cO$ will
+% % be considered by the algorithm.  Once a hyperedge is considered, the
+% % path to be assigned to it can also be computed in constant time.  In
+% % particular, in the algorithm the path to be assigned to $X$ depends on
+% % $\cl(Y), \cl(Z)$, $s(X)$ and the presence or absence of $r$ in the
+% % candidate partial path $\cl'(X)$.  Therefore, once the super-marginal
+% % edges are identified, the running time of the algorithm is linear in
+% % the size of the input.  By the technique used for constructing prime
+% % matrices \cite{wlh02}, the super-marginal edges can be found in linear
+% % time in the input size.  Therefore, the algorithm can be implemented
+% % to run in linear time in the input size.
+
+% % %\noindent
+% % The proof of correctness uses the following main properties:
+% % \begin{enumerate}
+% % \item The $k$-subdivided star has a very symmetric structure.  This
+% %   symmetry is quantified based on the following observation -- either
+% %   there are no feasible path labelings of $\cO$ using paths from $T$,
+% %   or there are exactly $p!$ feasible path labelings.  In other words,
+% %   there is either no feasible assignment, or effectively a unique
+% %   assignment modulo symmetry.
+% % \item The $p$ super-marginal hyperedges, if they exist, will each be
+% %   assigned a path from distinct rays, and each such path contains the
+% %   leaf.
+% % \item For a candidate hyperedge $X$, the partial path assignment
+% %   $\cl'(X)$ is decided by its overlap with $\cl(Y)$ and cardinality of
+% %   intersection with $\cl(Z)$.
 % % \end{enumerate}
-%   \end{enumerate} \qed
-% %\vspace{-6mm}
-
-% \end{proof}
-
-\begin{lemma}
-  \label{lem:noexit1}
-  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{perms} is
-  feasible, then the set of hypergraph isomorphism functions that
-  defines $(\cF, \cl)$'s feasibility is the same as the set that
-  defines $(\cF_j, \cl_j)$'s feasibility, if any.  Secondly, for any
-  iteration $j > 0$ of the {\em \bf while} loop, the {\em \bf exit}
-  statement in line~\ref{ln:exit1} will not execute.
-\end{lemma}
-\tnote[TH]{UNCOMMENT}
-% \begin{proof}
-%   Since $(\cF,\cl)$ is feasible, by Lemma~\ref{lem:feasible}
-%   $(\cF_j,\cl_j)$ for every iteration $j > 0$ is feasible.  % Therefore,
-% %   every hypergraph isomorphism $\phi: supp(\cF) \rightarrow V(T)$ that
-% %   induces $\cl$ on $\cF$ also induces $\cl_{j-1}$ and $\cl_{j}$ on
-% %   $\cF_{j-1}$ and $\cF_{j}$ respectively, i.e., $\cl_{\phi[\cF_{j-1}]}
-% %   = \cl_{j-1}$ and $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
-% %   that for all $x \in supp(\cF)$, for all $v \in V(T)$ the following
-% %   hold true.
-%   Also, every hypergraph isomorphism $\phi: supp(\cF) \rightarrow
-%   V(T)$ that induces $\cl$ on $\cF$ also induces $\cl_{j}$ on
-%   $\cF_{j}$, i.e., $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
-%   that for all $x \in supp(\cF)$, for all $v \in V(T)$, if $(x,v) \in
-%   \phi$ then $v \in \cl_{j}(S)$ for all $S \in \cF_{j}$ such that $x
-%   \in S$.
-% % the following
-% %   hold true.
-% %   \begin{enumerate}[i. ]
-% %   \item If $(x,v) \in \phi$ then $v \in \cl_{j-1}(S)$ for all $S \in
-% %     \cF_{j-1}$ such that $x \in S$.
-% %   \item If $(x,v) \in \phi$ then $v \in \cl_{j}(S)$ for all $S \in
-% %     \cF_{j}$ such that $x \in S$
-% %   \end{enumerate}
-%   In other words, filter 1 outputs a filtered path labeling that
-%   ``preserves''
-%   hypergraph isomorphisms of the original path labeling.\\
-%   Secondly, line~\ref{ln:exit1} will execute \iff the exit condition in
-%   line~\ref{ln:3waycheck}, i.e. failure of three way intersection
-%   preservation, becomes true in any iteration of the {\em \bf while}
-%   loop.  Due to Lemma~\ref{lem:invar1} Invariant IV, the exit
-%   condition does not occur if the input is a feasible ICPPL.\qed
-
-% %   such that $\phi(x) = v$ where $v$ is the leaf considered in the
-% %   first iterations of while. Clearly, $\phi$ is a renaming of
-% %   vertices in hypergraph $\cF$ to those in hypergraph $\cF^\cl$. Thus
-% %   the following facts can be observed in every iteration of the loop.
-
-% %   \begin{enumerate}[\hspace{2mm}i. ] \vspace{\topshrink}
-% %   \item all intersection cardinalities are preserved in this path
-% %     labeling \vspace{\topshrink}
-% %   \item element $x$ is exclusive in a hyperedge in $\cF$ since $v$ is
-% %     exclusive in a hyperedge in $\cF^\cl$.
-% %   \end{enumerate}
-
-% %   Thus the exit condition is never rendered true after $x$ and $v$ are
-% %   removed from their respective hyperedges. \qed
-
-% % \noindent
-% % This proof uses mathematical induction on the number
-% %   of iterations $j$, $j \ge 0$, of the loop that executed
-% %   without exiting. The base case, $j = 0$ is obviously true since the
-% %   input is an ICPPL and the exit condition cannot hold true due to
-% %   ICPPL property (iii).  Assume the algorithm executes till the end
-% %   of $j-1$th iteration without exiting at line
-% %  ~\ref{ln:3waycheck}. Consider the $j$th iteration. From Lemma
-% %  ~\ref{lem:feasible} we know that $(\cF_j, \cl_j)$ and $(\cF_{j-1},
-% %   \cl_{j-1})$ are feasible\remove[AS]{and from the proof in lemma
-% %     lem:invar1 we know that $(\cF_{j-1}, \cl_{j-1})$ satisfies all the
-% %     invariants defined in the lemma}.  Thus there exists a bijection
-% %   $\phi: supp(\cF) \rightarrow V(T)$ such that the induced path
-% %   % labeling on $\cF_{j-1}$ $\cl_{\phi[\cF_{j-1}]} = \cl_{j-1}$.
-% %   labeling on $\cF_{j}$, $\cl_{\phi[\cF_{j}]}$ and on $\cF_{j-1}$,
-% %   $\cl_{\phi[\cF_{j-1}]}$ are equal to $\cl_{j}$ and $\cl_{j-1}$
-% %   respectively.  We need to prove that for any $R, R', R'' \in
-% %   \cF_{j}$, $|R \cap R' \cap R''| = |\cl_j(R) \cap \cl_j(R') \cap
-% %   \cl_j(R'')|$.
-% %   The following are the possible cases that could arise. From argument
-% %   above, $|\cl_j(R) \cap \cl_j(R') \cap \cl_j(R'')| =
-% %   |\cl_{\phi[\cF_{j}]}(R) \cap \cl_{\phi[\cF_{j}]} (R') \cap
-% %   \cl_{\phi[\cF_{j}]} (R'')|$
-
-% %   \begin{enumerate}[a $|$]
-% %   \item {\em None of the sets are new. $R, R', R'' \in \cF_{j-1}$.}
-% %     We know $(\cF_{j-1}, \cl_{j-1})$ is feasible. Thus $|R \cap R'
-% %     \cap R''| = |\cl_{j-1}(R) \cap \cl_{j-1}(R') \cap \cl_{j-1}(R'')|
-% %     = |\cl_{j}(R) \cap \cl_{j}(R') \cap \cl_{j}(R'')|$.
-% %   \item {\em Only one, say $R$, is a new set.}  Let $R = S_1 \cap S_2$
-% %     ($S_1, S_2$ are defined in the proof of lemma
-% %    ~\ref{lem:invar1}). Now we have $|R \cap R' \cap R''| = |S_1 \cap
-% %     S_2 \cap R' \cap R''| = |\cl_{j-1}(S_1) \cap \cl_{j-1}(S_2) \cap
-% %     \cl_{j-1}(R') \cap \cl_{j-1}(R'')| = |\cl_{j}(R) \cap \cl_{j}(R')
-% %     \cap \cl_{j}(R'')|$. Thus proven. If $R$ is any of the other new
-% %     sets, the same claim can be verified using lemma
-% %    ~\ref{lem:setminuscard}.
-% %     % \item []{\bf Case 3:}
-% %   \item {\em At least two of $R, R', R''$ are new sets.}  The new sets
-% %     are disjoint hence this case is vacuously true.
-% %   \end{enumerate}
-% %   \qed
-% %   \tnote[E2]{remove the induction proof. just text saying x and v are
-% %     exclusive in these sets therefore the intersection cardinalities
-% %     don't change thus all invariants are still true} 
-% \end{proof}
-
-As a result of Algorithm~\ref{perms} each leaf $v$ in $T$
-is such that there is exactly one set in $\cF$ with $v$ as a vertex in
-the path assigned to it.  In Algorithm~\ref{leafasgn} we identify
-elements in $supp(\cF)$ whose images are leaves in a hypergraph
-isomorphism if one exists.  Let $S \in \cF$ be such that $\cl(S)$ is a
-path with leaf and $v \in V(T)$ is the unique leaf incident on it.  We
-define a new path labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) =
-\set{v}$ where $x$ an arbitrary element from $S \setminus \bigcup_{\hS
-  \ne S} \hS$. In other words, $x$ is an element present in no other
-set in $\cF$ except $S$. This is intuitive since $v$ is present in no
-other path image under $\cl$ other than $\cl(S)$.  The element $x$ and
-leaf $v$ are then removed from the set $S$ and path $\cl(S)$
-respectively. After doing this for all leaves in $T$, all path images
-in the new path labeling $\cl_{new}$ except leaf labels (a path that
-has only a leaf is called the {\em leaf label} for the corresponding
-single element hyperedge or set) are paths from a new pruned tree $T_0
-= T \setminus \{v \mid v \text{ is a leaf in }
-T\}$. Algorithm~\ref{leafasgn} is now presented with details.
-
-\tnote{write a note summarizing filter 1}
-% \begin{algorithm}[h]
-%   \caption{Leaf labeling from an ICPPL {\tt filter\_2($\cF, \cl, T$)}}
-%   \label{leafasgn}
-%   \begin{algorithmic}[\lndisplay]
-%     \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$
-%     \COMMENT {Path images are such that no two path images share a
-%       leaf.}
-%     \STATE $j \assign 1$\\
-%     \WHILE {there is a leaf $v$ in $T$ and a unique $S_1 \in
-%       \cF_{j-1}$ such that $v \in \cl_{j-1}(S_1)$ }\label{uniqueleaf}
-%     \STATE $\cF_j \assign \cF_{j-1} \setminus \{S_1\}$\\
-%     \STATE for all $S \in \cF_{j-1}$ such that $S \ne S_1$ set
-%     $\cl_j(S) \assign
-%     \cl_{j-1}(S)$\\
-%     \STATE $X \assign S_1 \setminus \bigcup_{S \in \cF_{j-1}, S \ne S_1}S$\\
-%     \IF{$X$ is empty} \label{xempty} \STATE {\bf exit} \label{ln:exit2} \ENDIF
-%     \STATE $x \assign $ arbitrary element from $X$\\
-%     \STATE $\cF_j \assign \cF_j \cup \{\{x\}, S_1 \setminus \{x\}\} $\\
-%     \STATE $\cl_j(\{x\}) \assign \{v\}$\\
-%     \STATE $\cl_j(S_1 \setminus \{x\}) \assign \cl_{j-1}(S_1) \setminus \{v\}$\\
-%     \STATE $j \assign j+1$\\
-%     \ENDWHILE
-%     \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
-%     \RETURN $(\cF', \cl')$
-%   \end{algorithmic}
-% \end{algorithm}
-
-Suppose the input ICPPL $(\cF, \cl)$ is feasible, yet set $X$ in
-Algorithm~\ref{leafasgn} is empty in some iteration of the {\bf while}
-loop. This will abort our procedure of finding the hypergraph
-isomorphism. The following lemma shows that this will not happen.
-
-\begin{lemma}
-  \label{lem:xnotempty}
-  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{leafasgn} is
-  feasible, then for all iterations $j > 0$ of the {\em \bf while}
-  loop, the {\em \bf exit} statement in line~\ref{ln:exit2} does not
-  execute.
-\end{lemma}
-
-% \begin{proof}
-%   Assume $X$ is empty for some iteration $j > 0$. We know that $v$ is
-%   an element of $\cl_{j-1}(S_1)$. Since it is uniquely present in
-%   $\cl_{j-1}(S_1)$, it is clear that $v \in \cl_{j-1}(S_1) \setminus
-%   \bigcup_{(S \in \cF_{j-1}) \wedge (S \ne S_1)}\cl_{j-1}(S)$.  Note
-%   that for any $x \in S_1$ it is contained in at least two sets due to
-%   our assumption about cardinality of $X$. Let $S_2 \in \cF_{j-1}$ be
-%   another set that contains $x$. From the above argument, we know $v
-%   \notin \cl_{j-1}(S_2)$. Therefore there cannot exist a hypergraph
-%   isomorphism bijection that maps elements in $S_2$ to those in
-%   $\cl_{j-1}(S_2)$. This contradicts our assumption that the input is
-%   feasible. Thus $X$ cannot be empty if input is ICPPL and feasible.
-%   \qed
-% \end{proof}
-
-\begin{lemma}
-  \label{lem:invar3}
-  In Algorithm~\ref{leafasgn}, for all $j > 0$, at the end of the
-  $j$th iteration of the {\bf while} loop the four invariants given in
-  Lemma~\ref{lem:invar1} hold.
-\end{lemma}
-
-% \begin{proof}
-%   By Lemma~\ref{lem:xnotempty} we know that set $X$ will not be empty
-%   in any iteration of the {\em \bf while} loop if input ICPPL $(\cF,
-%   \cl)$ is feasible and $\cl_j$ is always computed for all $j >
-%   0$. Also note that removing a leaf from any path keeps the new path
-%   connected. Thus invariant I is obviously true. In every iteration $j
-%   > 0$, we remove exactly one element $x$ from one set $S$ in $\cF$
-%   and exactly one vertex $v$ which is a leaf from one path
-%   $\cl_{j-1}(S)$ in $T$. This is because as seen in
-%   Lemma~\ref{lem:xnotempty}, $x$ is exclusive to $S$ and $v$ is
-%   exclusive to $\cl_{j-1}(S)$. Due to this fact, it is clear that the
-%   intersection cardinality equations do not change, i.e., invariants
-%   II, III, IV remain true. On the other hand, if the input ICPPL is
-%   not feasible the invariants are vacuously true. \qed
-% \end{proof}
-
-%% OLD TEXT TB DELETED?
-% \textcolor{cyan}{
-% \begin{lemma}
-%   \label{lem:notfeasibleexit}
-%   \tnote{IS THIS CORRECT?}  If input ICPPL $(\cF, \cl)$ is not
-%   feasible, then in one of the recursive calls to Algorithm --3--, the
-%   {\em \bf exit} statement in line x in Algorithm --1 or line y in
-%   Algorithm --2 will get executed.
-% \end{lemma}
-% }
-%% END OLD TEXT TB DELETED?
-
-We have seen two filtering algorithms above, namely,
-Algorithm~\ref{perms} {\tt filter\_1} and Algorithm~\ref{leafasgn}
-{\tt filter\_2} which when executed serially respectively result in a
-new ICPPL on the same universe $U$ and tree $T$. We also proved that
-if the input is indeed feasible, these algorithms do indeed output the
-filtered ICPPL. Now we present the algorithmic characterization of a
-feasible tree path labeling by way of Algorithm~\ref{al:icppl-find-isomorph}.
-
-Algorithm~\ref{al:icppl-find-isomorph} computes a
-hypergraph isomorphism $\phi$ recursively using Algorithm~\ref{perms}
-and Algorithm~\ref{leafasgn} and pruning the leaves of the input
-tree. In brief, it is done as follows. Algorithm~\ref{leafasgn} gives
-us the leaf labels in $\cF_2$, i.e., the elements in $supp(\cF)$ that
-map to leaves in $T$, where $(\cF_2, \cl_2)$ is the output of
-Algorithm~\ref{leafasgn}. All leaves in $T$ are then pruned away. The
-leaf labels are removed from the path labeling $\cl_2$ and the
-corresponding elements are removed from the corresponding sets in
-$\cF_2$. This tree pruning algorithm is recursively called on the
-altered hypergraph $\cF'$, path label $\cl'$ and tree $T'$. The
-recursive call returns the bijection $\phi''$ for the rest of the
-elements in $supp(\cF)$ which along with the leaf labels $\phi'$ gives
-us the hypergraph isomorphism $\phi$.  The following lemma formalizes
-the characeterization of feasible path labeling.
-
-\tnote{write a note summarizing filter 1}
-% \begin{algorithm}[h]
-%   \caption{{\tt get-hypergraph-isomorphism($\cF, \cl, T$)}}
-%   \label{al:icppl-find-isomorph}
-%   \begin{algorithmic}[\lndisplay]
-
-%     \IF{$T$ is empty}
-%     \RETURN $\emptyset$\\
-%     \ENDIF
-%     \STATE $L \assign \{v \mid v \text{ is a leaf in }      T\}$\\
-%     \STATE $(\cF_1, \cl_1) \assign$ {\tt filter\_1($\cF, \cl,
-%       T$)}\\
-%     \STATE $(\cF_2, \cl_2) \assign$ {\tt filter\_2($\cF_1,
-%       \cl_1, T$)}\\
-
-%     \STATE $(\cF', \cl') \assign (\cF_2, \cl_2)$\\
-%     \STATE $\phi' \leftarrow \emptyset$
-
-%     \FOR {every $v \in L$}
-%     \STATE $\phi'(x) \assign v$ where $x \in \cl_2^{-1}(\{v\})$
-%     \COMMENT {Copy the leaf labels to a one to one function $\phi':
-%       supp(\cF) \rightarrow L$
-%       }\\
-%     \STATE Remove $\{x\}$ and $\{v\}$ from $\cF'$, $\cl'$  appropriately\\
-%     \ENDFOR
-
-%     \STATE $T' \assign T \setminus L$
-
-%     \STATE $\phi'' \assign$ {\tt
-%       get-hypergraph-isomorphism($\cF', \cl', T'$)}
-%     \STATE $\phi \assign \phi'' \cup \phi'$ \\
-%     \RETURN $\phi$
-%   \end{algorithmic}
-% \end{algorithm}
-
-\begin{lemma}
-  \label{lem:hyperiso}  %{lem:perm}
-  If $(\cF, \cl)$ is an ICPPL from a tree $T$ and
-  Algorithm~\ref{al:icppl-find-isomorph}, {\tt
-    get- hypergraph- isomorphism ($\cF, \cl, T$)} returns a non-empty
-  function, then there exists a hypergraph isomorphism $\phi :
-  supp(\cF) \rightarrow V(T)$ such that the $\phi$-induced tree path
-  labeling is equal to $\cl$ or $\cl_\phi = \cl$.
-\end{lemma}
-% \begin{proof}
-%   It is clear that in the end of every recursive call to
-%   Algorithm~\ref{al:icppl-find-isomorph}, the function $\phi'$ is one
-%   to one involving all the leaves in the tree passed as input to that
-%   recursive call. Moreover, by Lemma~\ref{lem:noexit1} and
-%   Lemma~\ref{lem:xnotempty} it is consistent with the tree path
-%   labeling $\cl$ passed. The tree pruning is done by only removing
-%   leaves in each call to the function and is done till the tree
-%   becomes empty. Thus the returned function $\phi: supp(\cF)
-%   \rightarrow V(T)$ is a union of mutually exclusive one to one
-%   functions exhausting all vertices of the tree. In other words, it is
-%   a bijection from $supp(\cF)$ to $V(T)$ inducing the given path
-%   labeling $\cl$ and thus a hypergraph isomorphism. \qed
-% \end{proof}
-
-\begin{theorem}
-  \label{th:charac}
-  A path labeling $(\cF, \cl)$ on tree $T$ is feasible \iff it is an
-  ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
-  T)$ as input returns a non-empty function.
-\end{theorem}
-% \begin{proof}
-%   From Lemma~\ref{lem:hyperiso}, we know that if $(\cF, \cl)$ is an
-%   ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
-%   T)$ as input returns a non-empty function, $(\cF, \cl)$ is feasible.
-%   Now consider the case where $(\cF, \cl)$ is feasible. i.e. there
-%   exists a hypergraph isomorphism $\phi$ such that $\cl_\phi =
-%   \cl$. Lemma~\ref{lem:noexit1} and Lemma~\ref{lem:xnotempty} show us
-%   that filter 1 and filter 2 do not exit if input is feasible. Thus
-%   Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
-%   function.\qed
-% \end{proof}
-
-\subsection*{ICPPL when given tree is a path}
-\label{sec:icpplicpia}
-Consider a special case of ICPPL with the following properties when
-the tree $T$ is a path.  Hence, all path labels are can be viewed as
-intervals assigned to the sets in $\cF$.  It is shown, in
-\cite{nsnrs09}, that the filtering algorithms outlined above need only
-preserve pairwise intersection cardinalities, and higher level
-intersection cardinalities are preserved by the Helly Property of
-intervals.  Consequently, the filter algorithms do not need to ever
-evaluate the additional check to {\em \bf exit}.
-%\begin{enumerate}
-%\item Given tree $T$ is a path. Hence, all path labels are interval labels.
-%\item Only pairwise intersection cardinality
-%  preservation is sufficient. i.e. property (iii) in ICPPL is not enforced.
-%\item The filter algorithms do not have {\em \bf exit} statements.
-%\end{enumerate}
-%This is called an Intersection Cardinality Preservation Interval
-%Assignment (ICPIA) \cite{nsnrs09}. 
-This structure and its algorithm is
-used in the next section for finding tree path labeling from a
-$k$-subdivided star due to this graph's close relationship with
-intervals.
-
-\subsection{Case of a Special Tree -- The $k$-subdivided star}
-\label{sec:ksubdivstar}
-In this section we consider the problem of assigning paths from a
-$k$-subdivided star $T$ to a given set system $\cF$ such that each set
-$X \in \cF$ is of cardinality at most $k+2$.  Secondly, we present our
-results only for the case when overlap graph $\bO(\cF)$ is connected.
-%The
-%overlap graph is well-known from the work of
-%\cite{kklv10,nsnrs09,wlh02}.  We use the notation in
-%\cite{kklv10}. Recall from Section~\ref{sec:prelims} that hyperedges
-%$S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$ and $S'$
-%have a non-empty intersection but neither of them is contained in the
-%other. The overlap graph $\bO(\cF)$ is a graph in which the vertices
-%correspond to the sets in $\cF$, and the vertices corresponding to the
-%hyperedges $S$ and $S'$ are adjacent if and only if they overlap.  Note
-%that the intersection graph of $\cF$, $\bI(\cF)$ is different from
-%$\bO(\cF)$ and $\bO(\cF) \subseteq \bI(\cF)$.  
-A connected component of $\bO(\cF)$ is called an overlap component of
-$\cF$.  An interesting property of the overlap components is that any
-two distinct overlap components, say $\cO_1$ and $\cO_2$, are such
-that any two sets $S_1 \in \cO_1$ and $S_2 \in \cO_2$ are disjoint,
-or, w.l.o.g, all the sets in $\cO_1$ are contained within one set in
-$\cO_2$.  This containment relation naturally determines a
-decomposition of the overlap components into rooted containment trees.
-We consider the case when there is only one rooted containment tree,
-and we first present our algorithm when $\bO(\cF)$ is connected.  It
-is easy to see that once the path labeling to the overlap component in
-the root of the containment tree is achieved, the path labeling to the
-other overlap components in the rooted containment tree is essentially
-finding a path labeling when the target tree is a path: each target
-path is a path that is allocated to sets in the root overlap
-component.  Therefore, for the rest of this section, $\bO(\cF)$ is a
-connected graph. Recall that we also consider the special case when
-all hyperedges are of cardinality at most $k+2$.  By definition, a
-$k$-subdivided star has a central vertex which we call the {\em root},
-and each root to leaf path is called a {\em ray}.  First, we observe
-that by removing the root $r$ from $T$, we get a collection of $p$
-vertex disjoint paths of length $k+1$, $p$ being the number of leaves
-in $T$.  We denote the rays by $R_1, \ldots, R_p$, and the number of
-vertices in $R_i$, $i \in [p]$ is $k+2$.  Let
-$\seq{v_{i1},\ldots,v_{i(k+2)}=r}$ denote the sequence of vertices in
-$R_i$, where $v_{i1}$ is the leaf. Note that $r$ is a common vertex to
-all $R_i$.
-
-\subsubsection{Description of the Algorithm}
-In this section the given hypergraph $\cF$, the $k$-subdivided star
-and the root of the star are denoted by $\cO$, $T$ and vertex $r$,
-respectively.  In particular, note that the vertices of $\cO$
-correspond to the sets in $\cF$, and the edges correspond to the
-overlap relation.
-
-%\noindent
-For each hyperedge $X \in \cO$, we will maintain a 2-tuple of non-negative
- numbers $\seq{p_1(X), p_2(X)}$.  The numbers satisfy the property that
- $p_1(X) + p_2(X) \leq |X|$, and at the end of path labeling, for each
- $X$, $p_1(X) + p_2(X) = |X|$.  This signifies the algorithm tracking
- the lengths of subpaths of the path assigned to $X$ from at most two
- rays. We also maintain another parameter called the {\em residue} of
- $X$ denoted by $s(X)=|X| - p_1(X)$. This signifies the residue path
- length that must be assigned to $X$ which must be from another
- ray. For instance, if $X$ is labeled a path from only one ray, then
- $p_1(X) = |X|$, $p_2(X) = 0$ and $s(X) = 0$.
-
- The algorithm proceeds in iterations, and in the $i$-th
- iteration, $i > 1$, a single hyperedge $X$ that overlaps with a
- hyperedge that has been assigned a path is considered.  At the
- beginning of each iteration hyperedges of $\cO$ are classifed into
- the following disjoint sets.
- \begin{enumerate}
- \item [$\cL_1^i$] {\em Labeled without $r$.} Those that have been
-   labeled with a path which does not contain $r$ in one of the
-   previous iterations.\\  $\cL_1^i = \set{ X \mid p_1(X) = |X| \text{ and
-     } p_2(X) = 0 \text{ and } s(X) = 0, X \in \cO}$
- \item [$\cL_2^i$] {\em Labeled with $r$.} Those that have been labeled
-   with two subpaths of $\cl(X)$ containing $r$ from two different rays
-   in two previous iterations.\\ $\cL_2^i = \set{X \mid 0 < p_1\left(X\right),
-     p_2\left(X\right) < |X|=p_1(X)+p_2(X) \text{ and } s(X) = 0, X \in \cO}$
-   \item [$\cT_1^i$] {\em Type 1 / partially labeled.} Those that have
-     been labeled with one path containing $r$ from a single ray in one
-     of the previous iterations. Here, $p_1(X)$ denotes the length of
-     the subpath of $\cl(X)$ that $X$ has been so far labeled
-     with. \\
-     $\cT_1^i = \set{ X \mid 0 < p_1(X) < |X| \text{ and } p_2(X) = 0
-       \text{ and } s(X) = |X|-p_1(X), X \in \cO}$
-   \item [$\cT_2^i$] {\em Type 2 / not labeled.} Those that have not been
-     labeled with a path in any previous iteration.\\
-     $\cT_2^i = \set{ X \mid p_1(X) = p_2(X) = 0 \text{ and } s(X) = |X|,
-       X \in \cO}$
- \end{enumerate}
-% \vspace{-2mm}
-% \begin{align*}
-%   \cO &= \cL_1^i \cup \cL_2^i \cup \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]\\
-%   \cO_i &= \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]
-% \end{align*}
-
-%\noindent
-The set $\cO_i$ refers to the set of hyperedges $\cT_1^i \cup \cT_2^i$
-in the $i$th iteration.  Note that $\cO_1 = \cO$.  In the $i$th
-iteration, hyperedges from $\cO_i$ are assigned paths from $T$ using
-the following rules. Also the end of the iteration, $\cL_1^{i+1},
-\cL_2^{i+1}, \cT_1^{i+1}, \cT_2^{i+1}$ are set to $\cL_1^{i},
-\cL_2^{i}, \cT_1^{i}, \cT_2^{i}$ respectively, along with some
-case-specific changes mentioned in the rules below.
-
-%\noindent
-\begin{enumerate}[I.]
-\item {\bf Iteration 1:} Let $S=\{X_1,\ldots,X_s\}$ denote the
-  super-marginal hyperedges from $\cO_1$.  If $|S|=s \neq p$, then
-  exit reporting failure.  Else, assign to each $X_j \in S$, the path
-  from $R_j$ such that the path contains the leaf in $R_j$.  This path
-  is refered to as $\cl(X_j)$.  Set $p_1(X_j)=|X|, p_2(X_j)=s(X_j)=0$.
-  Hyperedges in $S$ are not added to $\cO_2$ but are added to
-  $\cL_1^2$ and all other hyperedges are added to
-  $\cO_2$.
-\item {\bf Iteration $i$:} Let $X$ be a hyperedge from $\cO_i$ such
-  that there exists $Y \in \cL_1^i \cup \cL_2^i$ and $X \overlap
-  Y$. Further let $Z \in \cL_1^i \cup L_2^i$ such that $Z \overlap Y$.
-  If $X \in \cT_2^i$, and if there are multiple $Y$ candidates then
-  any $Y$ is selected.  On the other hand, if $X \in \cT_1^i$, then
-  $X$ has a partial path assignment, $\cl'(X)$ from a previous
-  iteration, say from ray $R_j$. Then, $Y$ is
-  chosen such that $X \cap Y$ has a non-empty intersection with a ray
-  different from $R_j$.  The key things that are done in assigning a
-  path to $X$ are as follows. The {\em end} of path $\cl(Y)$ where
-  $\cl(X)$ would overlap is found, and then based on this the
-  existence of a feasible assignment is decided.  It is important to
-  note that since $X \overlap Y$, $\cl(X) \overlap \cl(Y)$ in any
-  feasible assignment.  Therefore, the notion of the {\em end} at
-  which $\cl(X)$ and $\cl(Y)$ overlap is unambiguous, since for any
-  path, there are two end points.
-  \begin{enumerate}
-  \item \label{iendpoint} {\em End point of $\cl(Y)$ where $\cl(X)$ overlaps
-      depends on $X \cap Z$:} If $X \cap Z
-    \neq \emptyset$, then $\cl(X)$ has an overlap of $|X \cap Y|$ at that
-    end of $\cl(Y)$ at which $\cl(Y)$ and $\cl(Z)$ overlap.  If $X
-    \cap Z = \emptyset$, then $\cl(X)$ has an overlap of $|X \cap Y|$ at
-    that end of $\cl(Y)$ where $\cl(Y)$ and $\cl(Z)$ do not intersect.
-  \item {\em Any path of length $s(X)$ at the appropriate end contains
-      $r$:} If $X \in \cT_1^i$ then after finding the appropriate end
-    as in step~\ref{iendpoint} this the unique path of length $s(X)$
-    should end at $r$.  If not, we exit reporting failure.  Else,
-    $\cl(X)$ is computed as union of $\cl'(X)$ and this path. If any
-    three-way intersection cardinality is violated with this new
-    assignment, then exit, reporting failure.  Otherwise, $X$ is added
-    to $\cL_2^{i+1}$.  On the other hand, if $X \in \cT_2^i$, then
-    after step~\ref{iendpoint}, $\cl(X)$ or $\cl'(X)$ is unique up to
-    the root and including it. Clearly, the vertices $\cl(X)$ or
-    $\cl'(X)$ contains depends on $|X|$ and $|X \cap Y|$.  If any
-    three way intersection cardinality is violated due to this
-    assignment, exit, reporting failure.  Otherwise,
-    $p_1(X)$ is updated as the length of the assigned path, and $s(X)
-    = |X|-p_1(X)$.  If $s(X) > 0$, then $X$ is added to $\cT_1^{i+1}$.
-    If $s(X)=0$, then $X$ is added to $\cL_1^{i+1}$.
-  \item {\em The unique path of length $s(X)$ overlapping at the
-      appropriate end of $Y$ does not contain $r$:} In this case,
-    $\cl(X)$ is updated to include this path.  If any three way
-    intersection cardinality is violated, exit, reporting failure.
-    Otherwise, update $p_1(X)$ and $p_2(X)$ are appropriate, $X$ is
-    added to $\cL_1^{i+1}$ or $\cL_2^{i+1}$, as appropriate.
-  \end{enumerate}
-\end{enumerate}
-
-%\noindent 
-\tnote[TH]{UNCOMMENT}
-% \subsection*{Proof of Correctness and Analysis of Running Time} 
-% It
-% is clear that the algorithm runs in polynomial time, as at each step,
-% at most three-way intersection cardinalities need to be checked.
-% Further, finding super-marginal hyperedges can also be done in
-% polynomial time, as it involves considering the overlap regions and
-% checking if the inclusion partial order contains a single minimal
-% element.  In particular, once the super-marginal edges are identified,
-% each iteration involes finding the next hyperedge to consider, and
-% testing for a path to that hyperedge.  To identify the next hyperedge
-% to consider, we consider the breadth first layering of the hyperedges
-% with the zeroeth layer consisting of the super-marginal hyperedges.
-% Since $\cO$ is connected, it follows that all hyperedges of $\cO$ will
-% be considered by the algorithm.  Once a hyperedge is considered, the
-% path to be assigned to it can also be computed in constant time.  In
-% particular, in the algorithm the path to be assigned to $X$ depends on
-% $\cl(Y), \cl(Z)$, $s(X)$ and the presence or absence of $r$ in the
-% candidate partial path $\cl'(X)$.  Therefore, once the super-marginal
-% edges are identified, the running time of the algorithm is linear in
-% the size of the input.  By the technique used for constructing prime
-% matrices \cite{wlh02}, the super-marginal edges can be found in linear
-% time in the input size.  Therefore, the algorithm can be implemented
-% to run in linear time in the input size.
-
-% %\noindent
-% The proof of correctness uses the following main properties:
-% \begin{enumerate}
-% \item The $k$-subdivided star has a very symmetric structure.  This
-%   symmetry is quantified based on the following observation -- either
-%   there are no feasible path labelings of $\cO$ using paths from $T$,
-%   or there are exactly $p!$ feasible path labelings.  In other words,
-%   there is either no feasible assignment, or effectively a unique
-%   assignment modulo symmetry.
-% \item The $p$ super-marginal hyperedges, if they exist, will each be
-%   assigned a path from distinct rays, and each such path contains the
-%   leaf.
-% \item For a candidate hyperedge $X$, the partial path assignment
-%   $\cl'(X)$ is decided by its overlap with $\cl(Y)$ and cardinality of
-%   intersection with $\cl(Z)$.
-% \end{enumerate}
-% These properties are formalized as follows:
-% \begin{lemma}
-%   \label{lem:sup-mar}
-%   If $X \in \cF$ is super-marginal and $\cl$ is a feasible tree path
-%   labeling to tree $T$, then $\cl(X)$ will contain a leaf in $T$.
-% \end{lemma}
-% \begin{proof}
-%   Suppose $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
-%   path labeling from $T$.  Assume $\cl(X)$ does not have a leaf.  Let
-%   $R_i$ be one of the rays (or the only ray) $\cl(X)$ is part of.
-%   Since $X$ is in a connected overlap component, there exists $Y_1 \in
-%   \cF$ and $X \nsubseteq Y_1$ such that $Y_1 \overlap X$ and $Y_1$ has
-%   at least one vertex closer to the leaf in $R_i$ than any vertex in
-%   $X$. Similarly with the same argument there exists $Y_2 \in \cF$
-%   with same subset and overlap relation with $X$ except it has has at
-%   least one vertex farther away from the leaf in $R_i$ than any vertex
-%   in $X$. Clearly $Y_1 \cap X$ and $Y_2 \cap X$ cannot be part of same
-%   inclusion chain which contradicts that assumption $X$ is
-%   super-marginal. Thus the claim is proved.\qed
-% \end{proof}
-% \begin{lemma}
-%   If $\cO$ does not have any super-marginal edges, then in any
-%   feasible path labeling $\cl$ of $\cO$ with paths from $T$ is such
-%   that, for any hyperedge $X$ for which $\cl(X)$ contains a leaf, $|X|
-%   \geq k+3$.
-% \end{lemma}
-% \begin{proof}
-%   The proof of this lemma is by contradiction.  Let $X$ be a
-%   hyperedges such that $|X| \leq k+2$ and that $\cl(X)$ has a leaf.
-%   This implies that the overlap regions with $X$, which are captured by
-%   the overlap regions with $\cl(X)$, will form a single inclusion
-%   chain. This shows that $X$ is a marginal hyperedge which
-%   contradicts the assumption that $\cO$ does not have super-marginal
-%   hyperedges. \qed
-% \end{proof}
-% This lemma is used to prove the next lemma for the case when for all
-% $X \in \cO$, $|X| \leq k+2$.  The proof is left out as it just uses
-% the previous lemma and the fact that the hyperedges in $X$ have at
-% most $k+2$ elements.
-% \begin{lemma}
-%   If there is a feasible path labeling for $\cO$ in $T$, then there
-%   are exactly $p$ super-marginal hyperedges.
-% \end{lemma}
-% These lemmas now are used to prove the following theorem.
-% \begin{theorem}
-%   Given $\cO$ and a $k$-subdivided star $T$, the above algorithm
-%   decides correctly if there is a feasible path labeling $\cl$.
-% \end{theorem}
-% \begin{proof} {\em Outline.}
+% % These properties are formalized as follows:
+% % \begin{lemma}
+% %   \label{lem:sup-mar}
+% %   If $X \in \cF$ is super-marginal and $\cl$ is a feasible tree path
+% %   labeling to tree $T$, then $\cl(X)$ will contain a leaf in $T$.
+% % \end{lemma}
+% % \begin{proof}
+% %   Suppose $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
+% %   path labeling from $T$.  Assume $\cl(X)$ does not have a leaf.  Let
+% %   $R_i$ be one of the rays (or the only ray) $\cl(X)$ is part of.
+% %   Since $X$ is in a connected overlap component, there exists $Y_1 \in
+% %   \cF$ and $X \nsubseteq Y_1$ such that $Y_1 \overlap X$ and $Y_1$ has
+% %   at least one vertex closer to the leaf in $R_i$ than any vertex in
+% %   $X$. Similarly with the same argument there exists $Y_2 \in \cF$
+% %   with same subset and overlap relation with $X$ except it has has at
+% %   least one vertex farther away from the leaf in $R_i$ than any vertex
+% %   in $X$. Clearly $Y_1 \cap X$ and $Y_2 \cap X$ cannot be part of same
+% %   inclusion chain which contradicts that assumption $X$ is
+% %   super-marginal. Thus the claim is proved.\qed
+% % \end{proof}
+% % \begin{lemma}
+% %   If $\cO$ does not have any super-marginal edges, then in any
+% %   feasible path labeling $\cl$ of $\cO$ with paths from $T$ is such
+% %   that, for any hyperedge $X$ for which $\cl(X)$ contains a leaf, $|X|
+% %   \geq k+3$.
+% % \end{lemma}
+% % \begin{proof}
+% %   The proof of this lemma is by contradiction.  Let $X$ be a
+% %   hyperedges such that $|X| \leq k+2$ and that $\cl(X)$ has a leaf.
+% %   This implies that the overlap regions with $X$, which are captured by
+% %   the overlap regions with $\cl(X)$, will form a single inclusion
+% %   chain. This shows that $X$ is a marginal hyperedge which
+% %   contradicts the assumption that $\cO$ does not have super-marginal
+% %   hyperedges. \qed
+% % \end{proof}
+% % This lemma is used to prove the next lemma for the case when for all
+% % $X \in \cO$, $|X| \leq k+2$.  The proof is left out as it just uses
+% % the previous lemma and the fact that the hyperedges in $X$ have at
+% % most $k+2$ elements.
+% % \begin{lemma}
+% %   If there is a feasible path labeling for $\cO$ in $T$, then there
+% %   are exactly $p$ super-marginal hyperedges.
+% % \end{lemma}
+% % These lemmas now are used to prove the following theorem.
+% % \begin{theorem}
+% %   Given $\cO$ and a $k$-subdivided star $T$, the above algorithm
+% %   decides correctly if there is a feasible path labeling $\cl$.
+% % \end{theorem}
+% % \begin{proof} {\em Outline.}
+% % %   If the algorithm outputs a path labeling $\cl$, then it is clear
+% % %   that it is an ICPPL. The reason is that the algorithm checks that
+% % %   three-way intersection cardinalities are preserved in each iteration
+% % %   which ensures \icpplpr~\ref{pr:iii}. Moreover, it is clear that
+% % %   $\cl(X)$ for any $X \in \cO$ is computed by maintaining
+% % %   \icpplpr~\ref{pr:i} and \icpplpr~\ref{pr:ii}. For such a labeling
+% % %   $\cl$, the proof that it is feasible is by induction on $k$. What
+% % %   needs to be shown is that Algorithm~\ref{al:icppl-find-isomorph}
+% % %   successfully runs on input $(\cO, \cl)$. In base case $k=0$, $T$ is
+% % %   a star. The claim is clear by observing that after Filter 1 and one
+% % %   iteration of Filter 2, all the leaves have found their pre-images
+% % %   from support $\cO$.  Therefore, in the induction step, it is clear
+% % %   that after Filter 1 and one iteration of Filter 2, the leaves are
+% % %   assigned pre-images.  Removing the leaves from $T$ and the
+% % %   pre-images from support of $\cO$, results in an ICPPL to a
+% % %   $(k-1)$-subdivided star.  Now we apply the induction hypothesis, and
+% % %   we get a isomorphism between the hypergraph $\cO$ and $\cO^\cl$.
 % %   If the algorithm outputs a path labeling $\cl$, then it is clear
@@
 % %   three-way intersection cardinalities are preserved in each iteration
-% %   which ensures \icpplpr~\ref{pr:iii}. Moreover, it is clear that
+% %   which ensures ICPPL \icpplpr~\ref{pr:iii}. Moreover, it is clear that
 % %   $\cl(X)$ for any $X \in \cO$ is computed by maintaining
-% %   \icpplpr~\ref{pr:i} and \icpplpr~\ref{pr:ii}. For such a labeling
+% %   ICPPL \icpplpr~\ref{pr:i} and ICPPL \icpplpr~\ref{pr:ii}. For such a labeling
 % %   $\cl$, the proof that it is feasible is by induction on $k$. What
@@
 % %   successfully runs on input $(\cO, \cl)$. In base case $k=0$, $T$ is
-% %   a star. The claim is clear by observing that after Filter 1 and one
-% %   iteration of Filter 2, all the leaves have found their pre-images
-% %   from support $\cO$.  Therefore, in the induction step, it is clear
-% %   that after Filter 1 and one iteration of Filter 2, the leaves are
-% %   assigned pre-images.  Removing the leaves from $T$ and the
-% %   pre-images from support of $\cO$, results in an ICPPL to a
-% %   $(k-1)$-subdivided star.  Now we apply the induction hypothesis, and
-% %   we get a isomorphism between the hypergraph $\cO$ and $\cO^\cl$.
-%   If the algorithm outputs a path labeling $\cl$, then it is clear
-%   that it is an ICPPL. The reason is that the algorithm checks that
-%   three-way intersection cardinalities are preserved in each iteration
-%   which ensures ICPPL \icpplpr~\ref{pr:iii}. Moreover, it is clear that
-%   $\cl(X)$ for any $X \in \cO$ is computed by maintaining
-%   ICPPL \icpplpr~\ref{pr:i} and ICPPL \icpplpr~\ref{pr:ii}. For such a labeling
-%   $\cl$, the proof that it is feasible is by induction on $k$. What
-%   needs to be shown is that Algorithm~\ref{al:icppl-find-isomorph}
-%   successfully runs on input $(\cO, \cl)$. In base case $k=0$, $T$ is
-%   a star. Also every set is at most size 2 ($k+2$) size and thus
-%   overlaps are at most 1. If two paths share a leaf in {\tt filter\_1}
-%   one must be of length 2 and the other of length 1. Thus the exit
-%   condition is not met. Now it is trivial that the exit condition in
-%   {\tt filter\_2} is also not met. Thus claim proven for base case.
-%   Now assume the claim to be true when target tree is a
-%   $(k-1)$-subdivided star. Consider the case of a $k$-subdivided
-%   star. It is clear that after {\tt filter\_1} and one iteration of
-%   {\tt filter\_2}, the leaves are assigned pre-images.  Removing the
-%   leaves from $T$ and the pre-images from support of $\cO$, results in
-%   an ICPPL to a $(k-1)$-subdivided star.  Now we apply the induction
-%   hypothesis, and we get a isomorphism between the hypergraph $\cO$
-%   and $\cO^\cl$.
+% %   a star. Also every set is at most size 2 ($k+2$) size and thus
+% %   overlaps are at most 1. If two paths share a leaf in {\tt filter\_1}
+% %   one must be of length 2 and the other of length 1. Thus the exit
+% %   condition is not met. Now it is trivial that the exit condition in
+% %   {\tt filter\_2} is also not met. Thus claim proven for base case.
+% %   Now assume the claim to be true when target tree is a
+% %   $(k-1)$-subdivided star. Consider the case of a $k$-subdivided
+% %   star. It is clear that after {\tt filter\_1} and one iteration of
+% %   {\tt filter\_2}, the leaves are assigned pre-images.  Removing the
+% %   leaves from $T$ and the pre-images from support of $\cO$, results in
+% %   an ICPPL to a $(k-1)$-subdivided star.  Now we apply the induction
+% %   hypothesis, and we get a isomorphism between the hypergraph $\cO$
+% %   and $\cO^\cl$.
 
@@
 
-\tnote[TH]{How COP (and TPL) relates to interval graph
-  isomorphism (and path graph isomorphism) move/summarize what is
-  already in Motivation section.}
-
-\tnote[TH]{TBD survey -- see blue notes under Graph
-  Isomorphism. namely citations in cite:aas93}
-
-\tnote[TH]{a brief on heirarchy of chordal graphs, path
-  graphs, interval graphs, peo, clique tree etc. and results.}
-
-\tnote[TH]{interval graphs are incomparability graphs. see
-  Golumbic. [tag:classification]}
-
-\tnote[TH]{finding min length hole in bipartite graph is
-  polynomial time [Sec 3.3.2 in cite:d08phd]}
-
-\tnote[TH]{theorem 2.2 in cite:d08phd - $G$ is union of
+\tnote[TH]{(1) How COP (and TPL) relates to interval graph isomorphism
+  (and path graph isomorphism) move/summarize what is already in
+  Motivation section.(2) TBD survey -- see blue notes under Graph
+  Isomorphism. namely citations in cite:aas93 (3) a brief on heirarchy
+  of chordal graphs, path graphs, interval graphs, peo, clique tree
+  etc. and results. (4) interval graphs are incomparability
+  graphs. see Golumbic. [tag:classification] (5) finding min length
+  hole in bipartite graph is polynomial time [Sec 3.3.2 in
+  cite:d08phd] (6) theorem 2.2 in cite:d08phd - $G$ is union of
   v.d. caterpillars iff $M$ has COP in rows. $M$ is edge vertex
@@
 
-\tnote[TH]{the interesting survey in conclusion section
-  of kklv10}
-
-\tnote[TH]{[[[COP helping some problems with hardness]]] having
-  COP in the input structure makes some problems less hard than in a
-  general input. -- min set cover problem is known to be NP hard. so
-  is min hitting
+\tnote[TH]{ (1) the interesting survey in conclusion section of kklv10
+  (2) [[[COP helping some problems with hardness]]] having COP in the
+  input structure makes some problems less hard than in a general
+  input. -- min set cover problem is known to be NP hard. so is min
+  hitting
   set. these two problems are equivalent cite-adp80\\
@@
   ----- Rectangle stabbing \\
------ see cite-d08phd Chapter 6?
-}
+  ----- see cite-d08phd Chapter 6?  }
 

### 2011-11-26 src/synopsis.tex
another bunch of last minute changes. THE NEXT COMMIT WILL BE THE FINAL FUCKING ONE!

@@
 \usepackage{t1enc}
+\usepackage{boxedminipage}
 \usepackage{../lib/mylatexlib}
@@
 
-\begin{document}
-
-%\nocite{*}
-\maketitle
-
-% \tnote{TBD: DELETE IF WE RUN OUT OF SPACE}
-% \begin{singlespace}
-% \tableofcontents
-% \thispagestyle{empty}
-% \listoffigures
-% \addcontentsline{toc}{chapter}{LIST OF FIGURES}
-% \listoftables
-% \addcontentsline{toc}{chapter}{LIST OF TABLES}
-% \end{singlespace}
-
-% The main text will follow from this point so set the page numbering
-% to arabic from here on.
-\pagenumbering{arabic}
-\setcounter{page}{0}
-%\newpage
-
-\section{Introduction}
-\label{sec:intro}
-\vspace{\secfirstparatrim}
-\tnote{set context: COP - generalization - TPL - no opt. just 2 or 3
-  sentences. use the space that the illustration does.}
-% \tnote{testing 1 2 3...}
-% \tnote[bogus1]{testing 1 2 3...}
-% \tnote[bogus2]{testing 1 2 3...}
-% \tnote[GTC]{testing 1 2 3...}
-% \tnote[bogus4]{testing 1 2 3...}
-% \tnote[TH]{testing 1 2 3...}
-
-\subsection{Illustration of the problem}
-\label{sec:problem}
-\vspace{\secfirstparatrim}
 % \def \LLL {{\em ``{\bf L}aughing with Kafka''}} 
@@
 % \def \xTTT {{\bf T}}
-
-%{\bf L}aughing {\bf G}irl {\bf B}rief {\bf T}his
-
+%%%%%%
 % \def \LLL {{\em ``{\bf L}aughing with Kafka''}} 
@@
 % \def \TTT {{\em ``Everything and More: {\bf A} Compact History of Infinity''}}
-
 % \def \xLLL {{\bf L}}
@@
 % \def \xTTT {{\bf A}}
-
-% \def \LLL {{\em ``{\bf B}rief Interviews with Hideous Men''}} 
-% \def \GGG {{\em ``The String {\bf T}heory''}} 
-% \def \BBB {{\em ``[{\bf W}]Rhetoric and the Math Melodrama''}}
-% \def \TTT {{\em ``{\bf F}ate, Time, and Language: An Essay on Free Will''}}
+%%%%%%
 \def \LLL {{\em ``$\mathbb{B}$rief Interviews with Hideous Men''}}
@@
 \def \TTT {{\em ``$\mathbb{F}$ate, Time, and Language: An Essay on Free Will''}}
-
-% \def \xLLL {{\bf B}}
-% \def \xGGG {{\bf T}} 
-% \def \xBBB {{\bf W}}
-% \def \xTTT {{\bf F}}
 \def \xLLL {\mathbb{B}}
@@
 \def \xTTT {\mathbb{F}}
-
-%{{\em ``Everything and More: A Compact History of In{\bf F}inity''}} 
-% {{\em ``Broom of {\bf T}he System''}} % {Consider the Lobster}
+% {{\em ``{\bf B}rief Interviews with Hideous Men''}} 
+% {{\em ``The String {\bf T}heory''}} 
+% {{\em ``[{\bf W}]Rhetoric and the Math Melodrama''}}
+% {{\em ``{\bf F}ate, Time, and Language: An Essay on Free Will''}}
+% {{\em ``Everything and More: A Compact History of In{\bf F}inity''}} 
+% {{\em ``Broom of {\bf T}he System''}} 
+% {{\em ``Consider the Lobster''}}
 % {{\em ``God Bless You, Mr. {\bf F}ranzen''}}
@@
 % {{\em ``{\bf G}irl with Curious Hair''}} 
-% {{\em ``Consider the {\bf L}obster|}}
+% {{\em ``Consider the {\bf L}obster''}}
 % {{\em ``{\bf G}od Bless You, Mr. Franzen''}}
@@
 %                                Tornadoes: A Midwestern  Boyhood''}}
-% {{\em ``{\bf T}his is Water''}} %
-% {\em {{\bf T}he Pale King}}
+% {{\em ``{\bf T}his is Water''}}
+% {{\em ``{\bf T}he Pale King''}}
 
 \def \Pa {{\em {\bf Pa}tricia}} 
-\def \Pi {{\em {\bf Pi}gpen}} %{{\em {\bf Ma}rcie}}
+\def \Pi {{\em {\bf Pi}gpen}}
 \def \Sn {{\em {\bf Sn}oopy}}
@@
 \def \Sa {{\em {\bf Sa}lly}}
-\def \Fr {{\em {\bf Fr}anklin}}  %{{\em {\bf Re}run}} 
+\def \Fr {{\em {\bf Fr}anklin}} 
 \def \Sc {{\em {\bf Sc}hr{\"o}eder}} 
 \def \Lu {{\em {\bf Lu}cy}}
-
 \def \xPa {{\bf Pa}} 
@@
 \def \xLu {{\bf Lu}}
-
 \def \residenceblock {{\em Infinite Loop}}
-
 \def \WSI {{\em Wallace Studies Institute}}
-
 \def \coneohone {{\em ``Influence of post modernism in Wallace's work''}}
 \def \coneohtwo {{\em ``A study on fragmented prose method''}}
+
+\begin{document}
+
+%\nocite{*}
+\maketitle
+
+% The main text will follow from this point so set the page numbering
+% to arabic from here on.
+\pagenumbering{arabic}
+\setcounter{page}{0}
+%\newpage
+
+\section{Introduction}
+\label{sec:intro}
+\vspace{\secfirstparatrim}
+
+Consecutive ones property is a non-trivial property of binary matrices
+that has been studied widely in the literature for over past 50
+years. Detection of COP in a matrix is possible efficiently and there
+are several algorithms that achieve the same. This thesis documents
+the work done on an extension of COP extended from the equivalent
+interval assignment problem in \cite{nsnrs09}. These new results
+rigorously prove a natural extension (to trees) of their
+characterization as well as makes connections to graph isomorphism,
+namely path graph isomorphism.
+
+% \tnote{testing 1 2 3...}
+% \tnote[bogus1]{testing 1 2 3...}
+% \tnote[bogus2]{testing 1 2 3...}
+% \tnote[GTC]{testing 1 2 3...}
+% \tnote[bogus4]{testing 1 2 3...}
+% \tnote[TH]{testing 1 2 3...}
+
+\subsection{Illustration of the problem}
+\label{sec:problem}
+\vspace{\secfirstparatrim}
 
@@
   \centering
-{
+  { \footnotesize
   \begin{tabular}{rcl}
@@
 accommodation.  All these apartments are located such that the streets
-connecting them do {\em not} form loops. Fig~\ref{fig:streetmap} shows
-the street map for {\residenceblock}. It may be noted that as a graph,
-it classifies as a tree.
+connecting them do {\em not} form
+loops. Fig~\ref{fig:streetmappathpeople} \tnote[TH]{fig:streetmap}
+shows the street map for {\residenceblock}. It may be noted that as a
+graph, it classifies as a tree.
+
+\begin{table}[htbp]
+  \centering
+  {\footnotesize    
+    \begin{tabular}{l||l}
+      \begin{tabular}{rcl}
+        $T $&$=$&$ \text{\em Street map tree of \residenceblock}$\\
+        $V(T) $&$=$&$ \{ 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11 \}$\\
+        % $\phi^{-1}(V(T)) $&$=$&$ \{ \xSa,\; \xPi,\; \xFr,\; \xWo,\;
+        % \xCh,\; \xVi,\;
+        % \xPa,\; \xSn,\; \xLu,\; \xLi,\; \xSc \}$\\
+        $\cP $&$=$&$ \{R\xLLL, R\xGGG, R\xBBB, R\xTTT\}$\\
+        $R\xLLL $&$=$&$ \{9, 1, 5, 3, 11\}$\\
+        $R\xGGG $&$=$&$ \{7, 2, 6, 5\}$\\
+        $R\xBBB $&$=$&$ \{8, 2, 4 \}$\\
+        $R\xTTT $&$=$&$ \{10, 6, 5, 3\}$\\
+        $n $&$=$&$ |V| = 11$\\
+        $m $&$=$&$ |\cP| = 4$\\\\
+        $\cl $&$=$& {\em Study group to route mapping}\\        
+        $\cl(\mathbb{X}) $&$=$&  $R\mathbb{X}$ for all $\mathbb{X} \in
+        \cF$
+      \end{tabular} & 
+      \begin{tabular}{c}
+       {\em Apartment allocation }($\phi$)\\
+                    \begin{tabular}{c|c}                      
+                      1  & \xSa\\
+                      2  & \xPi\\
+                      3  & \xFr\\
+                      4  & \xWo\\
+                      5  & \xCh\\
+                      6  & \xVi\\
+                      7  & \xPa\\
+                      8  & \xSn\\
+                      9  & \xLu\\
+                      10 & \xLi\\
+                      11 & \xSc                     
+                    \end{tabular}
+      \end{tabular}\\
+    \end{tabular}
+  }
+  \caption{\figtabsize A solution to study group accommodation problem}
+  \label{tab:iltree}
+\end{table}
 
@@
 allocation to individual students. A feasible allocation of exclusive
-routes to study groups is illustrated in Fig~\ref{fig:streetmappath}
-and the students' allocation of apartments that obeys this route
-allocation is also shown\tnote[TH]{in
-  Fig~ref::fig:streetmappathpeople}. How this is algorithmically
-computed is the focus of this thesis.
+routes to study groups is illustrated in
+Fig~\ref{fig:streetmappathpeople} \tnote[TH]{fig:streetmappath} and
+the students' allocation of apartments that obeys this route
+allocation is also shown. Table~\ref{tab:iltree} shows the same
+solution set theoretically.  How this is algorithmically computed is
+the focus of this thesis.
 
@@
 % \end{figure}
-% \begin{figure}[htbp] %[t]%
-%   \centering
-%   \includegraphics[scale=0.3]{../img/3_infinite_loop.png}
-%   \caption{\figtabsize Individual allocation of apartments to students in
-%     {\residenceblock} that meets the requirements stated before.}
-%   \label{fig:streetmappathpeople}
-% \end{figure}
+\begin{figure}[htbp] %[t]%
+  \centering
+  \includegraphics[scale=0.3]{../img/3_infinite_loop.png}
+  \caption{\figtabsize Individual allocation of apartments to students
+    in {\residenceblock} that meets the requirements stated before.
+    The routes are color coded as follows: red for
+    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
+    group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green
+    for \textcolor{green}{$\xTTT$} group. {\tiny {\em Peanuts images {\copyright} Charles
+    Schulz}}}
+  \label{fig:streetmappathpeople}
+\end{figure}
 
@@
   target tree $T$ to the hyperedges in given hypergraph $\cF$.}, is
-defined as follows.\\
-
+defined as follows.
+
+%\begin{tabular}[t]{c}
+{\small
 \begin{minipage}[h]{5in}
  % \begin{singlespace}
+ \vspace{2mm}
  {\large \FTPL}\\
-  % \hline
  \begin{tabular}[t]{l|l}
+ \hline\\
     {\tt Input} & 
@@
 \end{minipage}\\
+}
+%\end{tabular}
 
@@
 problem of computing a feasible TPL for a given hypergraph and target
-tree, if one exists. This problem, {\CFTPL}, is defined as follows. \\
-
+tree, if one exists. This problem, {\CFTPL}, is defined as follows.
+
+{\small
 \begin{minipage}[h]{5in}
  % \begin{singlespace}
+ \vspace{2mm}
   {\large \CFTPL}\\
-  % \hline
+  % 
   \begin{tabular}[t]{l|l}
+    \hline\\
     {\tt Input} & 
@@
 \end{minipage}\\
-
+}
 \def \CFTPLKTREE {\sc Compute $k$-subdivided Star Path Labeling}
@@
 $\cF$ have at most $k+2$ vertices. The examples of $k$-subdivided
-stars given in Figure~\ref{fig:ksubstar} makes the definition clear.
+stars are given in Figure~\ref{fig:ksubstar}.
 
 %This problem, {\CFTPLKTREE}, is defined as follows. \\
-\begin{figure}[htbp]
+\begin{figure}[t]
   \centering
-  (a) \includegraphics[scale=0.4]{../img/star.pdf}
-  (b) \includegraphics[scale=0.4]{../img/kstar.pdf}
-  \caption{\figtabsize Examples of $k$-subdivided stars. (a) $k = 0$ (b) $k = 2$ }
+  \begin{tabular}[h]{ccccc}
+  \includegraphics[scale=0.3]{../img/star.pdf} &&&&
+  \includegraphics[scale=0.3]{../img/kstar.pdf}\\
+  (a) &&&& (b)
+  \end{tabular}
+  \caption{\figtabsize Examples of $k$-subdivided stars. (a) $k = 0$
+    (b) $k = 2$ }
   \label{fig:ksubstar}
@@
 
+%\vspace{3mm}
+{\small
 \begin{minipage}[h]{5in}
  % \begin{singlespace}
+ \vspace{2mm}
   {\large \CFTPLKTREE}\\
-  % \hline
   \begin{tabular}[t]{l|l}
+    \hline\\
     {\tt Input} & 
@@
     \end{minipage}\\
-
     {\tt Question} &
@@
 \end{minipage}\\
-
-In spite of this
-restricted case we consider, we believe that our results are of
-significant interest in understanding the nature of {\sc Graph
+}
+
+In spite of this being a restricted case, we believe that our results
+are of significant interest in understanding the nature of {\sc Graph
   Isomorphism} which is polynomial time solvable in interval graphs

### 2011-11-27 src/synopsis.tex
sec 4/sumary of results is almost done.

@@
 characterization and find that trio-wise intersection cardinality
-preservation makes a tree path labeling (TPL) feasible, which is a
-generalization of the COP problem. This problem, {\FTPL}\footnote{A
+preservation makes a tree path labeling \footnote{A
   {\em tree path labeling} $\cl$ is a bijection of paths from the
-  target tree $T$ to the hyperedges in given hypergraph $\cF$.}, is
+  target tree $T$ to the hyperedges in given hypergraph $\cF$.} (TPL) feasible, which is a
+generalization of the COP problem. This problem is
 defined as follows.
@@
     \begin{minipage}[t]{\probdefwidth}
-      A hypergraph $\cF$ with vertex set $U$, a tree $T$ and a
-      bijection $\cl$~$:$~$\cF \rightarrow \cP$, where $\cP$ is a set
-      of paths from $T$.\\
+      A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
+      paths $\cP$ from $T$ and a
+      bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
     \end{minipage}\\
@@
 
-We give a necessary and sufficient condition for {\FTPL} to output in
+We give a necessary and sufficient condition (called ICPPL)\tnote{say a few
+  words about icppl} for {\FTPL} to output in
 affirmative. This characterization can be checked in polynomial time.
@@
 when the target tree $T$ is simply a path \cite{wlh02,nsnrs09}.
+
+\tnote{A few lines on the algo itself.}
 
@@
 }
+
+\def \kstar {$k$-subdivided star}
 \def \CFTPLKTREE {\sc Compute $k$-subdivided Star Path Labeling}
@@
 We present a polynomial time algorithm for {\CFTPL} when the target
-tree $T$ belongs to a special class of trees called {\em
-  $k$-subdivided stars} and when the hyperedges in the hypergraph
-$\cF$ have at most $k+2$ vertices. The examples of $k$-subdivided
-stars are given in Figure~\ref{fig:ksubstar}.
+tree $T$ belongs to a special class of trees called {\em \kstar s} and
+when the hyperedges in the hypergraph $\cF$ have at most $k+2$
+vertices. A couple of examples of {\kstar s} are given in
+Figure~\ref{fig:ksubstar}.
 
@@
   \end{tabular}
-  \caption{\figtabsize Examples of $k$-subdivided stars. (a) $k = 0$
+  \caption{\figtabsize Examples of {\kstar s}. (a) $k = 0$
     (b) $k = 2$ }
@@
       A hypergraph $\cF$ with vertex set $U$ such that every hyperedge
-      $S \in \cF$ is of cardinality atmost $k+2$ and a $k$-subdivided
-      star $T$.\\
+      $S \in \cF$ is of cardinality atmost $k+2$ and a {\kstar} $T$.\\
     \end{minipage}\\
@@
   Isomorphism} which is polynomial time solvable in interval graphs
-while being hard on path graphs\cite{kklv10}.
-
-\tnote{give gyan on why ktree is so.}
+while being hard on path graphs\cite{kklv10}. {\kstar s} are a class
+of trees which are in many ways very close to intervals or paths. Each
+ray\footnote{The path from a leaf to the root, the vertex with highest
+degree, is called a {\em ray} of the \kstar.} are independent except
+for the root\footnote{The vertex with maximum degree in a {\kstar} is
+  called {\em root}} and hence can be considered as an independent interval
+till the root. Our algorithm builds on this fact and uses the interval
+assignment algorithm\cite{nsnrs09} up until ``reaching'' the root and
+then uses the trio-wise intersection cardinality (the extra condition
+in ICPPL after ICPIA) check to resolve the
+ambiguity about which ray the algorithm should ``grow'' the solution
+into in the next iteration.
 

### 2011-12-01 src/synopsis.tex
final version before spell check

@@
     group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green
-    for \textcolor{green}{$\xTTT$} group. {\tiny {\em Peanuts images {\copyright} Charles
-    Schulz}}}
+    for \textcolor{green}{$\xTTT$} group. {\tiny {\em Peanuts images
+        {\copyright} Charles Schulz}}}
   \label{fig:streetmappathpeople}
 \end{figure}
+\tnote[TH]{make dashed/textured lines for routes. make it color agnostic.}
 
@@
 Section~\ref{sec:background} gives a brief survey of COP and
-optimization problems related to it.  Section~\ref{sec:results}
+optimization problems related to it followed by motivation for the
+thesis in Section~\ref{sec:motive}.  Section~\ref{sec:results}
 presents a summary of our results on the extention of COP namely, the
@@
 optimization problems is presented.
+
+%%%%%%                              %%%%%%
+%THX is the second level of ``tnote''-ing%
+%%%%%%                              %%%%%%
+\tnote[TH]{ADD: As it will be described in detail later in this
+  document, isomorphism of certain % classes of graphs, namely chordal
+%   graphs, have a close relationship with consecutive ones property and
+%   generalizations of it.  This is perhaps because of how closely COP
+%   of a matrix relates to properties of graphs derived from matrices as
+%   seen in the following results.
+}
+\tnote[THX]{ADD: peo exists iff chordal. lexicographic BFS
+  [tag:chordalGraph]} %
+\tnote[THX]{ADD: A well known result in %perfect graph theory is that
+%  the maximal cliques of  an interval graph $G$ can be linearly ordered
+%   such that for all $v \in V(G)$, cliques containing $v$ are
+%   consecutive in the ordering cite~{gh64}. This clearly means that a
+%   graph $G$ is an interval graph if and only if
+}% 
+\tnote[THX]{verify
+  from paper the statement of claim.}% 
+\tnote[TH]{maximal clique vertex
+  incidence matrix of%  $G$ has COP.  Also, maximal cliques of any
+%   chordal graph can be enumerated in polytime $O(m+n)$
+}
+\tnote[THX]{citation?!!}%  
+\tnote[TH]{cite~{fg65} uses these results to
+  give the first polynomial time algorithm for COT.}%
+\tnote[THX]{check. how do they use it?}%  
+\tnote[TH]{ A bipartite graph
+  is convex%  (on $R$) if and only if its half adjacency matrix has COP
+%   on rows.  The results in cite~{bl76} on COT are based on the result
+%   that interval graphs are AT-free chordal graphs.
+}%
+\tnote[THX]{the latter being Tucker's?}%  
+\tnote[THX]{(2) TBD survey
+  -- % see blue notes (in notebook) under Graph Isomorphism. namely
+%   citations in cite:aas93 (3) a brief on heirarchy of chordal graphs,
+%   path graphs, interval graphs, peo, clique tree etc. and results. (4)
+%   interval graphs are incomparability graphs. see
+%   Golumbic. [tag:classification] (5) finding min length hole in
+%   bipartite graph is polynomial time [Sec 3.3.2 in cite:d08phd] (6)
+%   theorem 2.2 in cite:d08phd - $G$ is union of v.d. caterpillars iff
+%   $M$ has COP in rows. $M$ is edge vertex incidence matrix $(*,2)$ of
+%   $G$
+}%
 
@@
 submatrices came as a corollary to the discovery that convex bipartite
-graphs are AT-free in \cite{at72}\tnote{check}. The first linear time
+graphs are AT-free in \cite{at72}\tnote[GTC]{check}. The first linear time
 algorithm for COP testing (COT) was invented by \cite{bl76} using a
@@
 is constructed using an elaborate linear time procedure and is also
-a test for planarity\tnote{check check check. both interval
+a test for planarity\tnote[GTC]{check check check. both interval
   graph and planarity in this paper?}. % Since this
@@
 of columns form the two partitions of the graph. Each row node is
-adjacent to those nodes that represent the columns that have ones in
+adjacent to those nodes that represent the columns that have {\un}s in
 the corresponding row. \cite{at72} proves that this bipartite graph
@@
 algorithms mentioned earlier), it is naturally of interest to find out
-the smallest (in terms of number of rows and/or columns and/or number
-of entries that are {\un}s) forbidden substructure. \cite{d08phd}
+the smallest forbidden substructure (in terms of number of rows and/or
+columns and/or number of entries that are {\un}s). \cite{d08phd}
 discusses a couple of algorithms which are efficient if the number of
-ones in a row is small. This is of significance in the case of sparce
+{\un}s in a row is small. This is of significance in the case of sparce
 matrices where this number is much lesser than the number of
@@
 COP. Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
-a submatrix $M'$ with all columns\tnote{check if b75 deals with COP
-  col or COP row. also is it any submatrix with k less than r rows or
-  submatrix must have all columns?} but a maximum cardinality subset
-of rows such that $M'$ has COP is NP complete. \cite{hg02} corrects an
-error of the abridged proof of this reduction as given in \cite{gj79}.
-\cite{d08phd} discusses all these problems in detail giving an
-extensive survey of the previously existing results which are almost
-exhaustively all approximation results and hardness results. Taking
-this further, \cite{d08phd} presents new results in the area of
-parameterized algorithms for this problem\tnote[TH]{elaborate - what
-  are the results?}.
+a submatrix $M'$ with all columns\tnote[GTC]{check if b75 deals with
+  COP col or COP row. also is it any submatrix with k less than r rows
+  or submatrix must have all columns?} but a maximum cardinality
+subset of rows such that $M'$ has COP is NP complete. \cite{hg02}
+corrects an error of the abridged proof of this reduction as given in
+\cite{gj79}.  \cite{d08phd} discusses all these problems in detail
+giving an extensive survey of the previously existing results which
+are almost exhaustively all approximation results and hardness
+results. Taking this further, \cite{d08phd} presents new results in
+the area of parameterized algorithms for this
+problem\tnote[TH]{elaborate - what are the results?}.
 
@@
 known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
-using a reduction to the longest path problem, \tnote{or is it a
+using a reduction to the longest path problem, \tnote[GTC]{or is it a
   survey of another result?  check.} that finding a Tucker's forbidden
-submatrix of at least $k$ rows is NP complete. \tnote{how is this
+submatrix of at least $k$ rows is NP complete. \tnote[GTC]{how is this
   different from booth's 75 result??}  \tnote[TH]{where should this
@@
 version of the $k$-run problem, i.e. minimization of number of blocks
-of ones is proven to be NP complete by cite:k77}\tnote{are these two
+of ones is proven to be NP complete by cite:k77}\tnote[GTC]{are these two
   the same?} \tnote[TH]{what is the reduction?} \tnote[TH]{other
@@
 characterization and find that trio-wise intersection cardinality
-preservation makes a tree path labeling \footnote{A
-  {\em tree path labeling} $\cl$ is a bijection of paths from the
-  target tree $T$ to the hyperedges in given hypergraph $\cF$.} (TPL) feasible, which is a
-generalization of the COP problem. This problem is
-defined as follows.
+preservation makes a tree path labeling \footnote{A {\em tree path
+    labeling} $\cl$ is a bijection of paths from the target tree $T$
+  to the hyperedges in given hypergraph $\cF$.} (TPL) feasible, which
+is a generalization of the COP problem. This problem is defined as
+follows.
 
@@
 
-We give a necessary and sufficient condition (called ICPPL)\tnote{say a few
-  words about icppl} for {\FTPL} to output in
-affirmative. This characterization can be checked in polynomial time.
-The most interesting consequence is that in our constructive
-procedure, it is sufficient to iteratively check if three-way
-intersection cardinalities are preserved.  In other words, in each
-iteration, it is sufficient to check if the intersection of any three
-hyperedges is of the same cardinality as the intersection of the
-corresponding paths.  Thus this generalizes the well studied question
-of the feasible interval assignment problem which is the special case
-when the target tree $T$ is simply a path \cite{wlh02,nsnrs09}.
+We give a necessary and sufficient condition (called ICPPL)\tnote[GTC]{say
+  a few words about icppl} for {\FTPL} to output in affirmative. This
+characterization can be checked in polynomial time.  The most
+interesting consequence is that in our constructive procedure, it is
+sufficient to iteratively check if three-way intersection
+cardinalities are preserved.  In other words, in each iteration, it is
+sufficient to check if the intersection of any three hyperedges is of
+the same cardinality as the intersection of the corresponding paths.
+Thus this generalizes the well studied question of the feasible
+interval assignment problem which is the special case when the target
+tree $T$ is simply a path \cite{wlh02,nsnrs09}.
 
@@
 We also have an algorithm for solving {\CFTPL} with no restrictions
-which runs in exponential time and explores a partial order theory of
-matrices related to {FTPL}. This po theory gives interesting insight
-into the problem of finding path labeling from $T$ by decomposing it
-into smaller subproblems of finding path labeling of subsets
-of $\cF$ from subtrees of $T$.\tnote[TH]{The connection
-of TPL to graph isomorphism will be made later in the document}
-%\end{enumerate}
+which runs in exponential time. 
+% and explores a partial order theory of
+% matrices related to {FTPL}.
+% gives interesting insight
+% into the problem of 
+This algorithm finds a path labeling from $T$ by decomposing the
+problem into smaller subproblems of finding path labeling of subsets
+of $\cF$ from subtrees of $T$. Given the fact that binary matrices
+naturally represent a set
+system % , we look at the matrix representation of subsets of $\cF$.
+and that the overlap relation between the sets involved is an obvious
+equivalence relation, $\cF$ naturally partitions into equivalence
+classes known as overlap components. In the context of COP, {\em overlap
+components} were used in \cite{wlh02} and \cite{kklv10}. Moreoever,
+\cite{nsnrs09} discovered that these equivalence classes form a total
+order. We extend this to TPL and find that when $\cF$ is a path
+hypergraph\footnote{If there exists an FTPL for a hypergraph $\cF$, it
+  is called a path hypergraph.}, the classes can be partially ordered
+as an in-tree in polynomial time. Once $\cF$ is
+``broken'' into overlap components, one must identify the subtree of
+$T$ that it needs to map to and this is the hard part which is
+currently open to be solved in polynomial time.
+
+\tnote[TH]{The connection of TPL to graph
+  isomorphism will be made later in the document}
 
@@
 \section{Conclusion}
-
-Our results also have close connection to recognition of path graphs
-and connection to path graph isomorphism.  Graphs which can be
-represented as the intersection graph of paths in a tree are called
-{\em path graphs}\cite{mcg04}. Thus given a hypergraph $\cF$, it can
-be viewed as paths in a tree, if and only if the intersection graph of
-$\cF$ is a {\em path graph}. Path graphs are a subclass of chordal
-graphs since chordal graphs are combinatorially characterized as the
-intersection graphs of subtrees of a tree.  Path graphs are well
-studied in the literature \cite{plr70,gav78,bp93,mcg04}.  Checking if
-a graph is a path graph can be done in polynomial time by the results
-of \cite{gav78,aas93}.  However, this is only a necessary condition
-for our question.  \remove{Indeed, our work provides a sufficient
-  condition for a related problem: that is whether a path assignment
-  to be feasible?}  On the other hand, path graph isomorphism is known
-be isomorphism-complete; see for example \cite{kklv10}. Therefore,
-\annote{it is unlikely that we can solve the problem of finding
-  feasible path labeling $\cl$ for a given $\cF$ and }{WHY?} tree $T$.
-\annote{It is definitely interesting to classify the kinds of trees
-  and hypergraphs for which feasible path labelings can be
-}{ELABORATE.} found efficiently.  \annote{These results would form
-  a natural generalization of COP testing and interval graph
-  isomorphism, culminating in Graph Isomorphism }{REWORD} itself.
-
-\tnote{MAKE CONNECTION TO PATH GRAPH ISOMORPHISM HERE}
-As it will be described in detail later in this document, isomorphism
-of certain classes of graphs, namely chordal graphs, have a close
-relationship with consecutive ones property and generalizations of it.
-This is perhaps because of how closely COP of a matrix relates to
-properties of graphs derived from matrices as seen in the following
-results. %-- Other relations of COP with Graphs: \\
-\tnote[TH]{peo exists iff chordal. lexicographic BFS
-  [tag:chordalGraph]} 
-A well known result in perfect graph theory is that the maximal
-cliques of an interval graph $G$ can be linearly ordered such that for
-all $v \in V(G)$, cliques containing $v$ are consecutive in the
-
-ordering\cite{gh64}. This clearly means that a graph $G$ is an
-interval graph if and only if \tnote{verify from paper the
-  statement of claim.} maximal clique vertex incidence matrix of $G$
-has COP.  Also, maximal cliques of any chordal graph can be enumerated
-in polytime $O(m+n)$ \tnote{citation?!!}.  \cite{fg65} uses these
-results to give the first polynomial time algorithm for
-COT. \tnote[TH]{check. how do they use it?}  A bipartite graph
-is convex (on $R$) if and only if its half adjacency matrix has COP on
-rows.  The results in \cite{bl76}
-%\cite{bl62} 
-on COT are based on the result that
-interval graphs are AT-free chordal graphs. \tnote{the latter being
-  Tucker's?}
-
-\tnote[TH]{(1) How COP (and TPL) relates to interval graph isomorphism
-  (and path graph isomorphism) move/summarize what is already in
-  Motivation section.(2) TBD survey -- see blue notes under Graph
-  Isomorphism. namely citations in cite:aas93 (3) a brief on heirarchy
-  of chordal graphs, path graphs, interval graphs, peo, clique tree
-  etc. and results. (4) interval graphs are incomparability
-  graphs. see Golumbic. [tag:classification] (5) finding min length
-  hole in bipartite graph is polynomial time [Sec 3.3.2 in
-  cite:d08phd] (6) theorem 2.2 in cite:d08phd - $G$ is union of
-  v.d. caterpillars iff $M$ has COP in rows. $M$ is edge vertex
-  incidence matrix $(*,2)$ of $G$}
-
-{\bf Complexity challenges:} Canonization is an important tool in graph isomorphism. Invention of a
-deterministic method of canonization for any class of graphs naturally
-results in an algorithm for isomorphism. All that is required is to
-check if the canons of two graphs are the same. Thus complexity of
-graph isomorphism can be studied by studying canonization
-methods. While general graph ismorphism remains elusive in terms of
-hardness, canonization has been studied for smaller classes of graphs
-thus giving complexity/hardness results for them.
-
-In 1992, \cite{sl92} made an important discovery that tree isomorphism
-is in logspace. It was done by inventing a method of canonization of
-trees using a logspace depth first traversal
-algorithm. \tnote{confirm the statement of claim}.
-
-\cite{kklv10} proved that interval graph canonization is also in
-logspace thus drawing logspace conslusions about COT. 
-% This result is
-% built on the fact that 
-Interval graphs are FP+C (fixed point with
-counting) definable \tnote{elucidate} and \cite{l10} showed that
-this implies that it captures PTIME. \tnote{elucidate} This result
-along with that of undirected graph reachability being in logspace
-\cite{rei08}, \cite{kklv10} proved their logspace result.\\
-
+\label{sec:conclusion}
+\vspace{\secfirstparatrim}
+
+We give a characterization for feasible tree path labeling of path
+hypergraphs. The proof for this is constructive and computes a
+feasiblility bijection mapping vertices of the hypergraph to vertices
+of the given target tree. This thesis also discovered an exponential algorithm
+that computes a feasible TPL to a given hypergraph if it is a path hypergraph.
+
+Our results have close connections to recognition of path graphs and
+to path graph isomorphism.  Graphs which can be represented as the
+intersection graph of paths in a tree are called {\em path
+  graphs}\cite{mcg04}. Thus, a hypergraph $\cF$ can be interpreted as
+paths in a tree, if and only if the intersection graph of $\cF$ is a
+{\em path graph}. Path graphs are a subclass of chordal graphs since
+chordal graphs are characterized as the intersection graphs of
+subtrees of a tree\cite{mcg04}.  Path graphs are well studied in the
+literature \cite{plr70}--\cite{mcg04}\tnote[TH]{Check if cite:bp93 is
+  indeed for path graphs}.  Path graph recognition can be done in
+polynomial time\cite{gav78,aas93}.  Clearly, this is a necessary
+condition in terms of the intersection graph of the input hypergraph
+$\cF$ in \FTPL. However, one can easily obtain a counterexample to
+show the insufficiency of this condition\tnote[GTC]{Give a figure
+  showing insuff.}.  Path graph isomorphism is known be
+isomorphism-complete\cite{kklv10}. Therefore, \annote[GTC]{it is unlikely
+  that we can solve the problem of finding feasible path labeling
+  $\cl$ for a given $\cF$ and }{WHY?} tree $T$.  It is
+  definitely interesting to classify the kinds of trees and
+  hypergraphs for which feasible path labelings can be\tnote[TH]{ELABORATE.}
+found efficiently.  \annote[TH]{These results would form a natural
+  generalization of COP testing and interval graph isomorphism,
+  culminating in Graph Isomorphism }{REWORD} itself.
+ To this effect we consider TPL on {\kstar} and give a polynomial time
+ solution for the same.
+
+Whether TPL on general trees is solvable in $P$ remains open. So do
+optimization opportunities in TPL (possible extentions to optimization for COP in the
+Section~\ref{sec:optcop}.\tnote[TH]{link to section}).
+
+\tnote[TH]{ ||Complexity challenges:|| Canonization is an important tool in graph isomorphism. Invention of a
+% deterministic method of canonization for any class of graphs naturally
+% results in an algorithm for isomorphism. All that is required is to
+% check if the canons of two graphs are the same. Thus complexity of
+% graph isomorphism can be studied by studying canonization
+% methods. While general graph ismorphism remains elusive in terms of
+% hardness, canonization has been studied for smaller classes of graphs
+% thus giving complexity/hardness results for them.
+
+% In 1992, \cite{sl92} made an important discovery that tree isomorphism
+% is in logspace. It was done by inventing a method of canonization of
+% trees using a logspace depth first traversal
+% algorithm. \tnote{confirm the statement of claim}.
+
+% \cite{kklv10} proved that interval graph canonization is also in
+% logspace thus drawing logspace conslusions about COT. 
+% % This result is
+% % built on the fact that 
+% Interval graphs are FP+C (fixed point with
+% counting) definable \tnote{elucidate} and \cite{l10} showed that
+% this implies that it captures PTIME. \tnote{elucidate} This result
+% along with that of undirected graph reachability being in logspace
+% \cite{rei08}, \cite{kklv10} proved their logspace result.
+}%
 \tnote[TH]{ (1) the interesting survey in conclusion section of kklv10
@@
   \item Tree path labeling of path
-    hypergraphs%: ( stuff in latin paper,)
-  \item Extension of po theory from \cite{nsnrs09}
+    hypergraphs %: ( stuff in latin paper,)
+  \item Extension of the structural characterization of matrices in \cite{nsnrs09}
     % (the matrix theory submitted in walcom, any other work - need to

### 2011-12-01 src/synopsis.tex
spell check done with aspell. however, previous commit was printed and circulated. some typos in there - deal with it.

@@
 thesis in Section~\ref{sec:motive}.  Section~\ref{sec:results}
-presents a summary of our results on the extention of COP namely, the
+presents a summary of our results on the extension of COP namely, the
 tree path labeling problem. \tnote[TH]{update for thesis}
@@
 discusses a couple of algorithms which are efficient if the number of
-{\un}s in a row is small. This is of significance in the case of sparce
+{\un}s in a row is small. This is of significance in the case of sparse
 matrices where this number is much lesser than the number of
@@
 a binary matrix such that the resulting matrix has at most $k$ runs of
-{\un}s. This is proven to be NP hard using a reduction from the
-hamiltonian path problem.\tnote[TH]{Theorem 2.1 in jkckv}
+{\un}s. This is proved to be NP hard using a reduction from the
+Hamiltonian path problem.\tnote[TH]{Theorem 2.1 in jkckv}
 \tnote[TH]{(1) A connection of COP problem to the travelling salesman
@@
 \def \probdefwidth {0.85\linewidth}
-\def \FTPL {{\sc Feasbile Tree Path Labeling}}
+\def \FTPL {{\sc Feasible Tree Path Labeling}}
 \def \CFTPL {\sc Compute Feasible Path Labeling}
@@
       A hypergraph $\cF$ with vertex set $U$ such that every hyperedge
-      $S \in \cF$ is of cardinality atmost $k+2$ and a {\kstar} $T$.\\
+      $S \in \cF$ is of cardinality at most $k+2$ and a {\kstar} $T$.\\
     \end{minipage}\\
@@
 classes known as overlap components. In the context of COP, {\em overlap
-components} were used in \cite{wlh02} and \cite{kklv10}. Moreoever,
+components} were used in \cite{wlh02} and \cite{kklv10}. Moreover,
 \cite{nsnrs09} discovered that these equivalence classes form a total
@@
 hypergraphs. The proof for this is constructive and computes a
-feasiblility bijection mapping vertices of the hypergraph to vertices
+feasibility bijection mapping vertices of the hypergraph to vertices
 of the given target tree. This thesis also discovered an exponential algorithm
@@
 Whether TPL on general trees is solvable in $P$ remains open. So do
-optimization opportunities in TPL (possible extentions to optimization for COP in the
+optimization opportunities in TPL (possible extensions to optimization for COP in the
 Section~\ref{sec:optcop}.\tnote[TH]{link to section}).

### 2011-12-21 src/synopsis.tex
commented out a package. NBD.
cleaned up library - new sty for thesis files. now Beamer and thesis files can coexist on mylatexlib.sty.

@@
 \usepackage{t1enc}
-\usepackage{boxedminipage}
+%\usepackage{boxedminipage}
+\usepackage{../lib/mythesislib}
 \usepackage{../lib/mylatexlib}
@@
 \date{\MakeUppercase
-  {\mysubmissionmonth}} 
+  {\mysynopsissubmissionmonth}} 
 \department{\MakeUppercase{\mydept}}
 
-\def \un {\bf 1}
-\def \secfirstparatrim {-5mm}
-\def \figtabsize {\footnotesize}
-
-% \def \LLL {{\em ``{\bf L}aughing with Kafka''}} 
-% \def \GGG {{\em ``{\bf G}irl with Curious Hair''}} 
-% \def \BBB {{\em ``{\bf B}rief Interviews with Hideous Men''}}
-% \def \TTT {{\em ``Everything and More: A Compact His{\bf T}ory of Infinity''}} 
-% \def \xLLL {{\bf L}}
-% \def \xGGG {{\bf G}} 
-% \def \xBBB {{\bf B}}
-% \def \xTTT {{\bf T}}
-%%%%%%
-% \def \LLL {{\em ``{\bf L}aughing with Kafka''}} 
-% \def \GGG {{\em ``Brief {\bf I}nterviews with Hideous Men''}} 
-% \def \BBB {{\em ``Broom of the {\bf S}ystem''}}
-% \def \TTT {{\em ``Everything and More: {\bf A} Compact History of Infinity''}}
-% \def \xLLL {{\bf L}}
-% \def \xGGG {{\bf I}} 
-% \def \xBBB {{\bf S}}
-% \def \xTTT {{\bf A}}
-%%%%%%
-\def \LLL {{\em ``$\mathbb{B}$rief Interviews with Hideous Men''}}
-\def \GGG {{\em ``The String $\mathbb{T}$heory''}} 
-\def \BBB {{\em ``[$\mathbb{W}$]Rhetoric and the Math Melodrama''}}
-\def \TTT {{\em ``$\mathbb{F}$ate, Time, and Language: An Essay on Free Will''}}
-\def \xLLL {\mathbb{B}}
-\def \xGGG {\mathbb{T}} 
-\def \xBBB {\mathbb{W}}
-\def \xTTT {\mathbb{F}}
-% {{\em ``{\bf B}rief Interviews with Hideous Men''}} 
-% {{\em ``The String {\bf T}heory''}} 
-% {{\em ``[{\bf W}]Rhetoric and the Math Melodrama''}}
-% {{\em ``{\bf F}ate, Time, and Language: An Essay on Free Will''}}
-% {{\em ``Everything and More: A Compact History of In{\bf F}inity''}} 
-% {{\em ``Broom of {\bf T}he System''}} 
-% {{\em ``Consider the Lobster''}}
-% {{\em ``God Bless You, Mr. {\bf F}ranzen''}}
-% {{\em ``{\bf L}aughing with Kafka''}} 
-% {{\em ``{\bf G}irl with Curious Hair''}} 
-% {{\em ``Consider the {\bf L}obster''}}
-% {{\em ``{\bf G}od Bless You, Mr. Franzen''}}
-% {{\em ``{\bf T}ennis, Trigonometry,
-%                                Tornadoes: A Midwestern  Boyhood''}}
-% {{\em ``{\bf T}his is Water''}}
-% {{\em ``{\bf T}he Pale King''}}
-
-\def \Pa {{\em {\bf Pa}tricia}} 
-\def \Pi {{\em {\bf Pi}gpen}}
-\def \Sn {{\em {\bf Sn}oopy}}
-\def \Wo {{\em {\bf Wo}odstock}}
-\def \Vi {{\em {\bf Vi}olet}} 
-\def \Li {{\em {\bf Li}nus}} 
-\def \Ch {{\em {\bf Ch}arlie}}
-\def \Sa {{\em {\bf Sa}lly}}
-\def \Fr {{\em {\bf Fr}anklin}} 
-\def \Sc {{\em {\bf Sc}hr{\"o}eder}} 
-\def \Lu {{\em {\bf Lu}cy}}
-\def \xPa {{\bf Pa}} 
-\def \xPi {{\bf Pi}} 
-\def \xSn {{\bf Sn}}
-\def \xWo {{\bf Wo}}
-\def \xVi {{\bf Vi}} 
-\def \xLi {{\bf Li}} 
-\def \xCh {{\bf Ch}}
-\def \xSa {{\bf Sa}}
-\def \xFr {{\bf Fr}}
-\def \xSc {{\bf Sc}} 
-\def \xLu {{\bf Lu}}
-\def \residenceblock {{\em Infinite Loop}}
-\def \WSI {{\em Wallace Studies Institute}}
-\def \coneohone {{\em ``Influence of post modernism in Wallace's work''}}
-\def \coneohtwo {{\em ``A study on fragmented prose method''}}
-
 \begin{document}
 
-%\nocite{*}
 \maketitle
@@
 
-A group of students, \Pa, \Pi, \Sn, \Wo, \Vi, \Li, \Ch, \Sa, \Fr,
-  \Sc\ and \Lu\ enroll at the
-% {\bf Wo}odstock, {\bf Vi}olet, {\bf Li}nus, {\bf Ch}arlie, {\bf
-%   Sa}lly, {\bf Re}run, {\bf Sc}hr{\"o}eder} and {\em {\bf Lu}cy}
-% enroll at the
-  {\WSI} for a liberal arts programme.  As part of their semester
-  thesis, they pick a body of work to study and form the namesake
-  study groups, {\LLL}, {\GGG}, {\BBB} and {\TTT}\tnote[TH]{put bib
-    entries for these works!}. A student will be in at least one study
-  group and may be in more than one. For instance, as will be seen
-  later, {\Fr} studies both {\LLL} and {\TTT} while \Wo\ studies only
-  \BBB.
+A group of students, \Pa, \Pi, \Sn, \Wo, \Vi, \Li, \Ch, \Sa, \Fr, \Sc\
+and \Lu\ enroll at the {\WSI} for a liberal arts programme.  As part
+of their semester thesis, they pick a body of work to study and form
+the namesake study groups, {\LLL}, {\GGG}, {\BBB} and
+{\TTT}\tnote[TH]{put bib entries for these works!}. A student will be
+in at least one study group and may be in more than one. For instance,
+as will be seen later, {\Fr} studies both {\LLL} and {\TTT} while \Wo\
+studies only \BBB.
 
@@
 there is the study group allocation to students.
-
-% \vspace{-3mm}
-% \begin{align*}
-% U &= \{\xPa,\; \xPi,\; \xSn,\; \xWo,\; \xVi,\; \xLi,\; \xCh,\;
-%     \xSa,\; \xFr,\; \xSc,\; \xLu\}\\
-% \cF &= \{\xLLL, \xGGG, \xBBB, \xTTT\}\\
-% \xLLL &= \{\xCh,\;  \xSa,\;  \xFr,\;  \xSc,\;  \xLu \}\\
-% \xGGG &= \{\xPa,\;  \xPi,\;  \xVi,\;  \xCh \}\\
-% \xBBB &= \{\xSn,\;  \xPi,\;  \xWo \}\\
-% \xTTT &= \{\xVi,\;  \xLi,\;  \xCh,\;  \xFr \}\\
-% n &= |U| = 11\\
-% m &= |\cF| = 4  
-% \end{align*}
 
@@
 
-% The administration faces a small challenge with their
-% accommodation. 
 The campus has a residential area {\residenceblock} that has $n$
@@
         $V(T) $&$=$&$ \{ 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11 \}$\\
-        % $\phi^{-1}(V(T)) $&$=$&$ \{ \xSa,\; \xPi,\; \xFr,\; \xWo,\;
-        % \xCh,\; \xVi,\;
-        % \xPa,\; \xSn,\; \xLu,\; \xLi,\; \xSc \}$\\
         $\cP $&$=$&$ \{R\xLLL, R\xGGG, R\xBBB, R\xTTT\}$\\
@@
 
-\def \probdefwidth {0.85\linewidth}
-\def \FTPL {{\sc Feasible Tree Path Labeling}}
-\def \CFTPL {\sc Compute Feasible Path Labeling}
-
 \section{Summary of results}
@@
 }
-
-\def \kstar {$k$-subdivided star}
-\def \CFTPLKTREE {\sc Compute $k$-subdivided Star Path Labeling}
 

### 2011-12-21 src/thesis.tex
Updated thesis and cop presentaiton files to sync up with mylatexlib.sty. modified latter for the same as well.
Fixed the font issue in presentation which had also manifested in synopsis+thesis. Package txfonts was the culprit.
cleaned up library - new sty for thesis files. now Beamer and thesis files can coexist on mylatexlib.sty.

@@
-\documentclass[synopsis,MS]             % MS thesis dissertation.
+\documentclass[MS]             % MS thesis dissertation.
               {iitmdiss}
@@
 \usepackage{times}             % Font?
-\usepackage{t1enc}             % TI encoding??  
-                            %%% Graphics related.                 %
-
+\usepackage{t1enc}             % T1 encoding??  
+%\usepackage{boxedminipage}
+\usepackage{../lib/mythesislib}
 \usepackage{../lib/mylatexlib}
 
+\def \mythesissubmissiondate {30 December 2011}
+
 \begin{document}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \title{\mythesistitle} 
-
 \author{\myname} 
-
-\date{\MakeUppercase {\mysubmissionmonth}}
-\department{\MakeUppercase {\mydept}}
+\date{\MakeUppercase
+  {\mythesissubmissionmonth}} 
+\department{\MakeUppercase{\mydept}}
 
@@
   \hspace*{-0.25in}
-  \parbox{2.5in}{
-    \noindent {\bf Prof.~\myadvisor} \\
-    \noindent Research Guide \\
-    \noindent Associate Professor \\
+%   \begin{tabular}[h]{lr}
+%     {\bf Dr.~\myadvisor} & \\
+%     {\small Research Guide} & \\
+%     {\small Associate Professor} & \\
+%     {\small Dept. of \mydept} & \hfill {\small Chennai}\\
+%     {\small IIT Madras -- 600 036} & \hfill {\small\mythesissubmissiondate}
+%  \end{tabular}
+  \parbox{3in}{
+    \noindent {\bf Dr.~\myadvisor} \\
+    \small
+    \noindent Research Guide\\
+    \noindent Associate Professor\\
     \noindent Dept. of \mydept\\
-    \noindent IIT Madras -- 600 036} \hfill
+    \noindent IIT Madras -- 600 036
+  } \hfill
   \parbox{1.25in}{               % {1.75in}
+    \small
     \vspace*{0.75in}
-    \noindent Chennai\\          %{\em Place:}
-    \noindent \mysubmissiondate} %{\em Date:}
+    \begin{tabular}[h]{r}
+    \noindent {Chennai}\\         
+    \noindent {\mythesissubmissiondate} 
+    \end{tabular}
+  }
 \end{singlespacing}
@@
 
-\tnote[a]{work in progress.}
+\tnote{WIP}
 
@@
   \noindent 
-  \annote[a]{In this paper, we explore a natural generalization of
+  \annote{In this paper, we explore a natural generalization of
   results on binary matrices with the |\tem consecutive ones property|. 
@@
 \textbf{ICPPL}   \> Intersection Cardinality Preserved Path Labeling \\
-
+% ADD MORE HERE
 \end{tabbing}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Introduction.
-
 \chapter{Introduction}
 \label{chap:intro}
-\tnote[a]{ Some things to remember - have this somewhere.
+
+\section{Introduction}
+\label{sec:intro}
+\vspace{\secfirstparatrim}
+
+Consecutive ones property is a non-trivial property of binary matrices
+that has been studied widely in the literature for over past 50
+years. Detection of COP in a matrix is possible efficiently and there
+are several algorithms that achieve the same. This thesis documents
+the work done on an extension of COP extended from the equivalent
+interval assignment problem in \cite{nsnrs09}. These new results
+rigorously prove a natural extension (to trees) of their
+characterization as well as makes connections to graph isomorphism,
+namely path graph isomorphism.
+
+% \tnote{testing 1 2 3...}
+% \tnote[bogus1]{testing 1 2 3...}
+% \tnote[bogus2]{testing 1 2 3...}
+% \tnote[GTC]{testing 1 2 3...}
+% \tnote[bogus4]{testing 1 2 3...}
+% \tnote[TH]{testing 1 2 3...}
+
+\subsection{Illustration of the problem}
+\label{sec:problem}
+\vspace{\secfirstparatrim}
+
+A group of students, \Pa, \Pi, \Sn, \Wo, \Vi, \Li, \Ch, \Sa, \Fr,
+  \Sc\ and \Lu\ enroll at the
+% {\bf Wo}odstock, {\bf Vi}olet, {\bf Li}nus, {\bf Ch}arlie, {\bf
+%   Sa}lly, {\bf Re}run, {\bf Sc}hr{\"o}eder} and {\em {\bf Lu}cy}
+% enroll at the
+  {\WSI} for a liberal arts programme.  As part of their semester
+  thesis, they pick a body of work to study and form the namesake
+  study groups, {\LLL}, {\GGG}, {\BBB} and {\TTT}\tnote[TH]{put bib
+    entries for these works!}. A student will be in at least one study
+  group and may be in more than one. For instance, as will be seen
+  later, {\Fr} studies both {\LLL} and {\TTT} while \Wo\ studies only
+  \BBB.
+
+Let $U$ and $\cF$ represent the set of students and the set of study
+groups respectively and the integers $n$ and $m$ denote the total
+number students and study groups respectively. In relation to this
+example, these are defined in Table~\ref{tab:wsigroups}. Also given
+there is the study group allocation to students.
+
+% \vspace{-3mm}
+% \begin{align*}
+% U &= \{\xPa,\; \xPi,\; \xSn,\; \xWo,\; \xVi,\; \xLi,\; \xCh,\;
+%     \xSa,\; \xFr,\; \xSc,\; \xLu\}\\
+% \cF &= \{\xLLL, \xGGG, \xBBB, \xTTT\}\\
+% \xLLL &= \{\xCh,\;  \xSa,\;  \xFr,\;  \xSc,\;  \xLu \}\\
+% \xGGG &= \{\xPa,\;  \xPi,\;  \xVi,\;  \xCh \}\\
+% \xBBB &= \{\xSn,\;  \xPi,\;  \xWo \}\\
+% \xTTT &= \{\xVi,\;  \xLi,\;  \xCh,\;  \xFr \}\\
+% n &= |U| = 11\\
+% m &= |\cF| = 4  
+% \end{align*}
+
+\begin{table}[htbp]
+  \centering
+  { \footnotesize
+  \begin{tabular}{rcl}
+    $U $&$=$&$ \{\xPa,\; \xPi,\; \xSn,\; \xWo,\; \xVi,\; \xLi,\; \xCh,\;
+    \xSa,\; \xFr,\; \xSc,\; \xLu\}$\\
+    $\cF $&$=$&$ \{\xLLL, \xGGG, \xBBB, \xTTT\}$\\
+    $\xLLL $&$=$&$ \{\xCh,\;  \xSa,\;  \xFr,\;  \xSc,\;  \xLu \}$\\
+    $\xGGG $&$=$&$ \{\xPa,\;  \xPi,\;  \xVi,\;  \xCh \}$\\
+    $\xBBB $&$=$&$ \{\xSn,\;  \xPi,\;  \xWo \}$\\
+    $\xTTT $&$=$&$ \{\xVi,\;  \xLi,\;  \xCh,\;  \xFr \}$\\
+    $n $&$=$&$ |U| = 11$\\
+    $m $&$=$&$ |\cF| = 4$      
+  \end{tabular}
+}
+  \caption{\figtabsize Students and study groups in \WSI}
+  \label{tab:wsigroups}
+\end{table}
+
+% The administration faces a small challenge with their
+% accommodation. 
+The campus has a residential area {\residenceblock} that has $n$
+single occupancy apartments reserved for the study groups'
+accommodation.  All these apartments are located such that the streets
+connecting them do {\em not} form
+loops. Fig~\ref{fig:streetmappathpeople} \tnote[TH]{fig:streetmap}
+shows the street map for {\residenceblock}. It may be noted that as a
+graph, it classifies as a tree.
+
+\begin{table}[htbp]
+  \centering
+  {\footnotesize    
+    \begin{tabular}{l||l}
+      \begin{tabular}{rcl}
+        $T $&$=$&$ \text{\em Street map tree of \residenceblock}$\\
+        $V(T) $&$=$&$ \{ 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11 \}$\\
+        % $\phi^{-1}(V(T)) $&$=$&$ \{ \xSa,\; \xPi,\; \xFr,\; \xWo,\;
+        % \xCh,\; \xVi,\;
+        % \xPa,\; \xSn,\; \xLu,\; \xLi,\; \xSc \}$\\
+        $\cP $&$=$&$ \{R\xLLL, R\xGGG, R\xBBB, R\xTTT\}$\\
+        $R\xLLL $&$=$&$ \{9, 1, 5, 3, 11\}$\\
+        $R\xGGG $&$=$&$ \{7, 2, 6, 5\}$\\
+        $R\xBBB $&$=$&$ \{8, 2, 4 \}$\\
+        $R\xTTT $&$=$&$ \{10, 6, 5, 3\}$\\
+        $n $&$=$&$ |V| = 11$\\
+        $m $&$=$&$ |\cP| = 4$\\\\
+        $\cl $&$=$& {\em Study group to route mapping}\\        
+        $\cl(\mathbb{X}) $&$=$&  $R\mathbb{X}$ for all $\mathbb{X} \in
+        \cF$
+      \end{tabular} & 
+      \begin{tabular}{c}
+       {\em Apartment allocation }($\phi$)\\
+                    \begin{tabular}{c|c}                      
+                      1  & \xSa\\
+                      2  & \xPi\\
+                      3  & \xFr\\
+                      4  & \xWo\\
+                      5  & \xCh\\
+                      6  & \xVi\\
+                      7  & \xPa\\
+                      8  & \xSn\\
+                      9  & \xLu\\
+                      10 & \xLi\\
+                      11 & \xSc                     
+                    \end{tabular}
+      \end{tabular}\\
+    \end{tabular}
+  }
+  \caption{\figtabsize A solution to study group accommodation problem}
+  \label{tab:iltree}
+\end{table}
+
+A natural question would be to find how the students should be
+allocated apartments such that each study group has the {\em least
+  distance to travel} for a discussion? More specifically, we are
+interested in enforcing additional conditions, namely, that all the
+students in a study group must be next to each other; in other words,
+for one student to reach another fellow study group member's apartment
+(for all study groups the student is part of), she must not have to
+pass the apartment of any student who is not in that study group. To
+further elucidate, the apartments of students of any study group must
+be arranged in an exclusive unfragmented path on the street
+map. Exclusivity here means that the path must not have apartments
+from other study groups (unless that apartment is also part of {\em
+  this} study group).
+
+An intuitive approach to this problem would be to first find the paths
+that each study group decides to inhabit and then refine the
+allocation to individual students. A feasible allocation of exclusive
+routes to study groups is illustrated in
+Fig~\ref{fig:streetmappathpeople} \tnote[TH]{fig:streetmappath} and
+the students' allocation of apartments that obeys this route
+allocation is also shown. Table~\ref{tab:iltree} shows the same
+solution set theoretically.  How this is algorithmically computed is
+the focus of this thesis.
+
+%%%%%%%%%%%% THESIS UNCOMMENT %%%%%%%%%%%%
+% \begin{figure}[t]%[htbp]
+%  \centering
+%   \includegraphics[scale=0.3]{../img/1_infinite_loop.png}
+%   \caption{{\residenceblock} street map}
+%   \label{fig:streetmap}
+% \end{figure}
+
+% \begin{figure}[t]%[htbp]
+%   \centering
+%   \includegraphics[scale=0.3]{../img/2_infinite_loop_BTWF.png}
+%   \caption{{\residenceblock} street map with study group routes
+%     allocated. Routes are color coded as follows: red for
+%     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
+%     group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
+%     \textcolor{green}{$\xTTT$} group}
+%   \label{fig:streetmappath}
+% \end{figure}
+% \begin{figure}[t] % [h]
+%   \centering
+%   \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
+%   \caption{Individual allocation of apartments to students in
+%     {\residenceblock} that meets the requirements stated before.}
+%   \label{fig:streetmappathpeople}
+% \end{figure}
+%%%%%%%%%%%% END THESIS UNCOMMENT %%%%%%%%%%%%
+% \begin{figure}[htbp] %[t]%
+%  \centering
+%   (a) \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
+%   (b) \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
+%  \label{fig:streetmap}
+%   \caption{\figtabsize (a) {\residenceblock} street map (b) {\residenceblock} street map with study group routes
+%     allocated. Routes are color coded as follows: red for
+%     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
+%     group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
+%     \textcolor{green}{$\xTTT$} group}
+% \end{figure}
+\begin{figure}[htbp] %[t]%
+  \centering
+  \includegraphics[scale=0.3]{../img/3_infinite_loop.png}
+  \caption{\figtabsize Individual allocation of apartments to students
+    in {\residenceblock} that meets the requirements stated before.
+    The routes are color coded as follows: red for
+    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
+    group, orange for \textcolor{yellow}{$\xBBB$} group, green
+    for \textcolor{green}{$\xTTT$} group. {\tiny {\em Peanuts images
+        {\copyright} Charles Schulz}}}% YellowOrange
+  \label{fig:streetmappathpeople}
+\end{figure}
+\tnote[TH]{make dashed/textured lines for routes. make it color agnostic.}
+
+As a special case, suppose all the apartments are on the same street
+or if they are all lined up on a single path, the street map becomes a
+tree that is just a path. Then the problem becomes what is called an
+{\em interval assignment problem}. The idea of interval assignment may
+not be obvious here; hence to see this, consider a different problem
+in {\WSI} where the classes for these study groups courses need to be
+scheduled during a day (or a week or any time period). Each study
+group has a bunch of courses associated with it some of which may be
+shared by two or more study groups. It is mandatory that a student who
+is a member of a study group takes all the courses associated with
+that group. There are slots during the day for classes to be held and
+the problem is to allocate class slots to courses such that all the
+classes of a study group are consecutive. It is debatable if this will
+not hamper the attention span and memory retention rate of the
+students but that is, regrettably, out of the scope of this
+thesis. The parallels between this class allocation problem and the
+accommodation problem can be seen as follows. The set $U$ here, are
+the courses offered (say Course 101 {\coneohone}, Course 102
+{\coneohtwo} and so on). In this variation of the problem, the
+collection $\cF$ is the set of study groups but the study groups are
+filled by course IDs (in place of students in the earlier
+example). For instance, Course 101 is mandatory for all study groups
+$\xLLL$, $\xGGG$, $\xBBB$, $\xTTT$ and Course 102 is mandatory for
+only the $\xLLL$ group) and so on. The sequence of class slots for the
+day (or week or any time period) is analogous to the street map in the
+accommodation problem. It is quite obvious now why this version of the
+problem (where the ``target graph'' is a path and not any
+tree\tnote[TH]{Allowing any tree in this example could be seen as a
+  scenario where there are parallel classes. A node falling in the
+  path between two other nodes would mean that the corresponding is
+  scheduled between the other two.}) is called an interval assignment
+problem.
+
+The interval assignment problem to a set system is equivalent to the
+consecutive ones property (COP) problem in binary matrices\cite{wlh02,
+  nsnrs09}.  The COP problem is to rearrange rows (columns) of a
+binary matrix in such a way that every column (row) has its {\un}s
+occur consecutively. If this is possible the matrix is said to have
+the COP.  COP is a well researched combinatorial problem and has
+several positive results on tests for it and computing the COP
+permutation (i.e. the course schedule in the above illustration) which
+will be surveyed later in this document. Hence we are interested in
+extensions of COP, more specifically, the extension of interval
+assignment problem to tree path assignment problem (which is
+illustrated by the study group accommodation problem).
+
+Section~\ref{sec:background} gives a brief survey of COP and
+optimization problems related to it followed by motivation for the
+thesis in Section~\ref{sec:motive}.  Section~\ref{sec:results}
+presents a summary of our results on the extension of COP namely, the
+tree path labeling problem. \tnote[TH]{update for thesis}
+
+\section{Consecutive Ones Property Testing - a Survey} %{Background}
+\label{sec:background}
+\vspace{\secfirstparatrim}
+In this section, a brief survey of the consecutive ones problem and its
+optimization problems is presented.
+
+%%%%%%                              %%%%%%
+%THX is the second level of ``tnote''-ing%
+%%%%%%                              %%%%%%
+\tnote[TH]{ADD: As it will be described in detail later in this
+  document, isomorphism of certain % classes of graphs, namely chordal
+%   graphs, have a close relationship with consecutive ones property and
+%   generalizations of it.  This is perhaps because of how closely COP
+%   of a matrix relates to properties of graphs derived from matrices as
+%   seen in the following results.
+}
+\tnote[THX]{ADD: peo exists iff chordal. lexicographic BFS
+  [tag:chordalGraph]} %
+\tnote[THX]{ADD: A well known result in %perfect graph theory is that
+%  the maximal cliques of  an interval graph $G$ can be linearly ordered
+%   such that for all $v \in V(G)$, cliques containing $v$ are
+%   consecutive in the ordering cite~{gh64}. This clearly means that a
+%   graph $G$ is an interval graph if and only if
+}% 
+\tnote[THX]{verify
+  from paper the statement of claim.}% 
+\tnote[TH]{maximal clique vertex
+  incidence matrix of%  $G$ has COP.  Also, maximal cliques of any
+%   chordal graph can be enumerated in polytime $O(m+n)$
+}
+\tnote[THX]{citation?!!}%  
+\tnote[TH]{cite~{fg65} uses these results to
+  give the first polynomial time algorithm for COT.}%
+\tnote[THX]{check. how do they use it?}%  
+\tnote[TH]{ A bipartite graph
+  is convex%  (on $R$) if and only if its half adjacency matrix has COP
+%   on rows.  The results in cite~{bl76} on COT are based on the result
+%   that interval graphs are AT-free chordal graphs.
+}%
+\tnote[THX]{the latter being Tucker's?}%  
+\tnote[THX]{(2) TBD survey
+  -- % see blue notes (in notebook) under Graph Isomorphism. namely
+%   citations in cite:aas93 (3) a brief on heirarchy of chordal graphs,
+%   path graphs, interval graphs, peo, clique tree etc. and results. (4)
+%   interval graphs are incomparability graphs. see
+%   Golumbic. [tag:classification] (5) finding min length hole in
+%   bipartite graph is polynomial time [Sec 3.3.2 in cite:d08phd] (6)
+%   theorem 2.2 in cite:d08phd - $G$ is union of v.d. caterpillars iff
+%   $M$ has COP in rows. $M$ is edge vertex incidence matrix $(*,2)$ of
+%   $G$
+}%
+
+\subsection{Matrices with COP}
+\label{sec:copmatrices}
+\vspace{\secfirstparatrim} 
+As seen earlier, the interval assignment problem (illustrated as the
+course scheduling problem in Section~\ref{sec:problem}), is a special
+case of the problem we address in this thesis, namely the tree path
+labeling problem (illustrated as the study group accommodation
+problem). The interval assignment problem and COP problem are equivalent
+problems. In this section we will see some of the results that
+exists in the literature today towards solving the COP problem and
+optimization problems surrounding it.
+
+\begin{figure}[t] %[htbp]
+  \centering
+  % \includegraphics[scale=0.4]{../img/0_cop_nocop.png}
+{\figtabsize
+ \begin{tabular}[h]{l|lcccl}
+    $M_1$: & $M_1'$: &&&& $M_2$:\\
+    &&&&&\\
+    \begin{tabular}[h]{llll}
+      $c_1$ & $c_2$ &$c_3$ &$c_4$\\
+      &&&\\
+      \un & 0   & \un & 0\\
+      0   & \un & 0   & \un \\
+      \un & 0   & 0   & \un
+    \end{tabular}
+    &
+    \begin{tabular}[h]{llll}
+      $c_3$ &$c_1$ &$c_4$& $c_2$\\
+      &&&\\
+      \un & \un & 0 & 0\\
+      0 & 0 & \un & \un \\
+      0 & \un & \un & 0
+    \end{tabular}
+    &&&&
+    \begin{tabular}[h]{llll}
+      $d_1$ & $d_2$ &$d_3$ &$d_4$\\
+      % &&&\\
+      &&&\\
+      \un & \un & 0 & 0\\
+      0 & \un & \un & 0 \\
+      0 & \un & 0 & \un 
+    \end{tabular}
+  \end{tabular}
+}
+
+  \caption{\figtabsize Matrices with and without COP. $M_1$ has COP because by
+    permuting its columns, $c_1$-$c_4$, one can obtain $M_1'$ where
+    the {\un}s in each row are consecutive. $M_2$, however, does not
+    have COP since no permutation of its columns, $d_1$-$d_4$, will
+    arrange {\un}s in each row consecutively \cite{d08phd}.}
+
+  \label{fig:cop-matrix}
+\end{figure}
+
+Recall that a matrix with COP is one whose rows (columns)
+can be rearranged so that the {\un}s in every column (row) are in
+consecutive rows (columns). Figure~\ref{fig:cop-matrix} shows examples
+of this property.
+COP in binary matrices has several practical applications in diverse
+fields including scheduling \cite{hl06}, information retrieval
+\cite{k77} and computational biology \cite{abh98}.  Further, it is a
+tool in graph theory \cite{mcg04} for interval graph recognition,
+characterization of Hamiltonian graphs, planarity testing \cite{bl76}
+and in integer linear programming \cite{ht02,hl06}.
+
+The obvious first questions after being introduced to the consecutive
+ones property of binary matrices are if COP can be detected
+efficiently in a binary matrix and if so, can the COP permutation of
+the matrix also be computed efficiently?  Recognition of COP in a
+binary matrix is polynomial time solvable and the first such algorithm
+was given by \cite{fg65}.  A landmark result came a few years later
+when \cite{at72} discovered the families of forbidden submatrices that
+prevent a matrix from having COP and most, if not all, results that
+came later were based on this discovery which connected COP in binary
+matrices to convex bipartite graphs. In fact, the forbidden
+submatrices came as a corollary to the discovery that convex bipartite
+graphs are AT-free in \cite{at72}\tnote[GTC]{check}. The first linear time
+algorithm for COP testing (COT) was invented by \cite{bl76} using a
+data structure called PQ trees.  Since then several COT algorithms
+have been invented -- some of which involved variations of PQ trees
+\cite{mm96,wlh01,mcc04}, some involved set theory and ICPIA
+\cite{wlh02,nsnrs09}, parallel COT algorithms\cite{as95,bs03,ly91} and
+certifying algorithms\cite{mcc04}. % These algorithms
+% improve over PQ trees on the time complexity and/or the complexity
+% of implementation.
+
+The construction of PQ trees in \cite{bl76} draws on the
+close relationship of matrices with COP to interval graphs. A PQ tree of a
+matrix is one that stores all row (column) permutations of the matrix
+that give the COP orders (there could be multiple orders of rows or columns) of the matrix. This
+is constructed using an elaborate linear time procedure and is also
+a test for planarity\tnote[GTC]{check check check. both interval
+  graph and planarity in this paper?}. % Since this
+% result, several improvements have been made to the procedural
+% complexity of the algorithm itself and also derivative data structures
+% that are easier to construct.
+PQR trees is a generalized data structure based on PQ trees
+\cite{mm96,mpt98}.  \cite{tm05} describes an improved algorithm to
+build PQR trees. \tnote[GTC]{improv in terms of what?}\cite{wlh02}
+describes the simpler algorithm for COT. Hsu also invented PC trees
+\cite{wlh01}\tnote[TH]{This result first appeared inproc ISAAC92}
+which is claimed to be much easier to implement. \cite{nsnrs09}
+describes a characterization of consecutive ones property solely based
+on the cardinality properties of the set representations of the
+columns (rows); every column (row) is equivalent to a set that has the
+row (column) indices of the rows (columns) that have one entries in
+this column (row). This is interesting and relevant, especially to
+this thesis because it simplifies COT to a great
+degree. \tnote[GTC]{it reduces the solution search space. fill in the blanks.}
+
+\cite{mcc04} describes a different approach to COT. While all previous
+COT algorithms gave the COP order if the matrix has the property but
+exited stating negative if otherwise, this algorithm gives an evidence
+by way of a certificate of matrix even when it has no COP. This
+enables a user to verify the algorithm's result even when the answer
+is negative. This is significant from an implementation perspective
+because automated program verification is hard and manual verification
+is more viable. Hence having a certificate reinforces an
+implementation's credibility. Note that when the matrix {\em has} COP,
+the COP order is the certificate.  The internal machinery of this
+algorithm is related to the weighted betweenness problem
+addressed\tnote[GTC]{in what way??} in \cite{co98}.  \tnote[GTC]{expand
+  on the COP order graph creation and it having to be bipartite for M
+  to have COP. and thus an odd cycle being an evidence of no COP.}
+\tnote[TH]{ where should this go?: (1) cite|jlm97 (application of PQ trees
+  in graphics). (2) helly's theorem citation
+  19XXdgk-Hellystheorem-Danzer-Gruenbaum-Klee}
+
+\subsection{Optimization problems in COP}
+\label{sec:optcop}
+\vspace{\secfirstparatrim}  
+So far we have been concerned about matrices that have the consecutive
+ones property. However in real life applications, it is rare that data
+sets represented by binary matrices have COP, primarily due to the
+noisy nature of data available. At the same time, COP is not arbitrary
+and is a desirable property in practical data representation
+\cite{co98,jkckv04,k77}. In this context, there are several
+interesting problems when a matrix does not have COP but is ``close''
+to having COP or is allowed to be altered to have COP. These are the
+optimization problems related to a matrix which does not have
+COP. Some of the significant problems are surveyed in this section.
+
+\tnote[TH]{ -- sect 4.1 in cite:d08phd has many results
+  surveyed. hardness results, approx. results. results are usually for
+  a class of matrices $(a,b)$ where number {\un}s in columns and rows
+  are restriced to $a$ and $b$ . -- problem of flipping at most $k$
+  entries of $M$ to make it attain COP. this is NP complete
+  cite:b75-phd}\tnote[TH]{(1) scite:lb62 showed that interval graphs
+  are AT-free.  describe AT (2) show the close relationship b/w COP
+  and graphs sec 2.2, pg 31} \cite{at72} showed that a matrix that
+does not have COP have certain substructures that prevent it from
+having COP. Tucker classified these forbidden substructures into five
+classes of submatrices. This result is presented in the context of
+convex bipartite graphs which \cite{at72} proved to be
+AT-free\tnote[TH]{ check this up. give details. - doms'}. By
+definition, convex bipartite graph have half adjacency matrices that
+have COP on either rows or columns (graph is biconvex if it has COP on
+both)\cite{d08phd}. A half adjacency matrix is a binary matrix
+representing a bipartite graph as follows. The set of rows and the set
+of columns form the two partitions of the graph. Each row node is
+adjacent to those nodes that represent the columns that have {\un}s in
+the corresponding row. \cite{at72} proves that this bipartite graph
+has no asteroidal triple if and only if the matrix has COP and goes on
+to identify the forbidden substructures for these bipartite
+graphs. The matrices corresponding to these substructures are the
+forbidden submatrices.
+
+Once a matrix has been detected to not have COP (using any of the COT
+algorithms mentioned earlier), it is naturally of interest to find out
+the smallest forbidden substructure (in terms of number of rows and/or
+columns and/or number of entries that are {\un}s). \cite{d08phd}
+discusses a couple of algorithms which are efficient if the number of
+{\un}s in a row is small. This is of significance in the case of sparse
+matrices where this number is much lesser than the number of
+columns. $(*,\Delta)${\em -matrices} are matrices with no restriction
+on number of {\un}s in any column but has at most $\Delta$ {\un}s in
+any row. {\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are
+similar problems which deals with inducing COP on a matrix. In {\sc
+  Min COS-R (Min COS-C)} the question is to find the minimum number of
+rows (columns) that must be deleted to result in a matrix with COP.
+In the dual problem {\sc Max COS-R (Max COS-C)} the search is for the
+maximum number of rows (columns) that induces a submatrix with
+COP. Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
+a submatrix $M'$ with all columns\tnote[GTC]{check if b75 deals with
+  COP col or COP row. also is it any submatrix with k less than r rows
+  or submatrix must have all columns?} but a maximum cardinality
+subset of rows such that $M'$ has COP is NP complete. \cite{hg02}
+corrects an error of the abridged proof of this reduction as given in
+\cite{gj79}.  \cite{d08phd} discusses all these problems in detail
+giving an extensive survey of the previously existing results which
+are almost exhaustively all approximation results and hardness
+results. Taking this further, \cite{d08phd} presents new results in
+the area of parameterized algorithms for this
+problem\tnote[TH]{elaborate - what are the results?}.
+
+Another problem is to find the minimum number of entries in the matrix
+that can be toggled to result in a matrix with COP.  \cite{v85}
+discusses approximation of {\sc COP Augmentation} which is the problem
+of changing of the minimum number of zero entries to {\un}s so that the
+resulting matrix has COP. As mentioned earlier, this problem is
+known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
+using a reduction to the longest path problem, \tnote[GTC]{or is it a
+  survey of another result?  check.} that finding a Tucker's forbidden
+submatrix of at least $k$ rows is NP complete. \tnote[GTC]{how is this
+  different from booth's 75 result??}  \tnote[TH]{where should this
+  go? cite|tz04 (approx submatrix with COP sparse matrices)}
+
+\cite{jkckv04} discusses the use of matrices with almost-COP (instead
+of one block of consecutive {\un}s, they have $x$ blocks, or {\em
+  runs}, of consecutive {\un}s and $x$ is not too large) in the
+storage of very large databases.  The problem is that of reordering of
+a binary matrix such that the resulting matrix has at most $k$ runs of
+{\un}s. This is proved to be NP hard using a reduction from the
+Hamiltonian path problem.\tnote[TH]{Theorem 2.1 in jkckv}
+\tnote[TH]{(1) A connection of COP problem to the travelling salesman
+  problem is also introduced. what does this mean? -- COP can be used
+  as a tool to reorder $0.5T \le runs(M) \le T$. (2) The optimization
+version of the $k$-run problem, i.e. minimization of number of blocks
+of ones is proven to be NP complete by cite:k77}\tnote[GTC]{are these two
+  the same?} \tnote[TH]{what is the reduction?} \tnote[TH]{other
+  problems similar to COP -- cite:ckl96 (ILP, circ ones, one drop) --
+  cite:th98 (generalization of COP - minimax, biotonic column) Tucker}
+
+\section{Generalization of COP - The Motivation}  %{Motivation}
+\label{sec:motive}
+\vspace{\secfirstparatrim}  
+As seen in Section~\ref{sec:copmatrices}, \cite{nsnrs09} describes a
+characterization of consecutive ones property based on the cardinality
+properties of the set representations of the columns. This result is
+very relevant to this thesis because aside from it simplifying COT to
+a great degree, our generalization problem is motivated by their
+results.
+
+The result in
+\cite{nsnrs09} characterizes interval assignments to the sets which
+can be obtained from a single permutation of the rows.  They show that
+for each set, the cardinality of the interval assigned to it must be
+same as the cardinality of the set, and the intersection cardinality
+of any two sets must be same as the intersection cardinality of the
+corresponding intervals.  While this is obviously a necessary
+condition, this result shows this is also sufficient.  \cite{nsnrs09}
+calls such an interval assignment an Intersection Cardinality
+Preserving Interval Assignment (ICPIA).  This paper generalizes the
+idea from \cite{wlh02} of decomposing a given binary matrix into prime
+matrices for COT and describes an algorithm to test if an ICPIA exists
+for a given set system.
+
+The tree path labeling problem is a natural generalization of the
+interval assignment problem or the COP problem. The problem is defined
+as follows -- given a set system $\cF$ from a universe $U$ and a tree
+$T$, does there exist a bijection from $U$ to the vertices of $T$ such
+that each set in the system maps to a path in $T$.  We refer to this
+as the {\em tree path labeling problem} for an input set system,
+target tree pair -- $(\cF,T)$. As a special case if the tree $T$ is a
+path, the problem becomes the interval assignment problem.  We focus
+on the question of generalizing the notion of an ICPIA \cite{nsnrs09}
+to characterize feasible path assignments.  We show that for a given
+set system $\cF$, a tree $T$, and an assignment of paths from $T$ to
+the sets, there is a feasible bijection between $U$ and $V(T)$ if and
+only if all intersection cardinalities among any three sets (not
+necessarily distinct) is same as the intersection cardinality of the
+paths assigned to them and the input passes a filtering algorithm
+(described in this paper) successfully.  This characterization gives a
+natural data structure that stores all the relevant feasible
+bijections between $U$ and $V(T)$. This reduces the search space for
+the solution considerably from the universe of all possible bijections
+between $U$ and $V(T)$ to only those bijections that maintain the
+characterization\tnote[TH]{REWORD}.  Further, the filtering algorithm
+is also an efficient algorithm to test if a tree path labeling to the
+set system is
+feasible.  %This generalizes the result in \cite{nsnrs09}.
+
+% How the problem of COT is related to the constraint statisfaction
+% problem of interval assignment to sets in a system can be seen as
+% follows. Every column of the binary matrix can be converted into a set
+% of integers which are the indices of rows with {\un}s in that column. It
+% is apparent that if the matrix has COP, then constructing such sets
+% after applying the COP permutation to the matrix will result in all
+% sets with consecutive integers. In other words, after application of
+% COP reordering, the sets are intervals. Indeed the problem now becomes
+% finding interval assignments to a given set system such that there
+% exists a permutation of the universe (set of row indices) which
+% converts each set to its assigned interval.
+
+\def \probdefwidth {0.85\linewidth}
+\def \FTPL {{\sc Feasible Tree Path Labeling}}
+\def \CFTPL {\sc Compute Feasible Path Labeling} 
+\section{Summary of results}
+\label{sec:results}
+\vspace{\secfirstparatrim}
+% We address the computation of TPL\footnote{Tree Path Labeling} for a
+% given hypergraph\footnote{The term {\em hypergraph} is used instead
+%   {\em set systems} are used interchangably in this
+%   section. Hypergraphs that have TPLs on some target tree are called
+%   {\em path hypergraphs}.} and target tree in this research. However
+% optimization problems in TPL, akin to problems in
+% Section~\ref{sec:optcop} for COP, is outside the scope of this thesis.
+As we saw earlier by the result in \cite{nsnrs09}, pairwise
+intersection cardinality preservation is necessary and sufficient for
+an interval assignment to be feasible for a given
+hypergraph\footnote{A {\em hypergraph} is an alternate representation
+  of a set system and will be used through out this section.} and thus
+is a characterization for COP. In our work we extend this
+characterization and find that trio-wise intersection cardinality
+preservation makes a tree path labeling \footnote{A {\em tree path
+    labeling} $\cl$ is a bijection of paths from the target tree $T$
+  to the hyperedges in given hypergraph $\cF$.} (TPL) feasible, which
+is a generalization of the COP problem. This problem is defined as
+follows.
+
+%\begin{tabular}[t]{c}
+{\small
+\begin{minipage}[h]{5in}
+ % \begin{singlespace}
+ \vspace{2mm}
+ {\large \FTPL}\\
+ \begin{tabular}[t]{l|l}
+ \hline\\
+    {\tt Input} & 
+    \begin{minipage}[t]{\probdefwidth}
+      A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
+      paths $\cP$ from $T$ and a
+      bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
+    \end{minipage}\\
+    {\tt Question} &
+    \begin{minipage}[t]{\probdefwidth}
+      Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
+      such that $\phi$ when applied on any hyperedge in $\cF$ will
+      give
+      the path mapped to it by the given tree path labeling $\cl$.\\
+      { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every hyperedge
+        $S \in \cF$.}
+    \end{minipage}\\
+  \end{tabular}
+  % \end{singlespace}
+\end{minipage}\\
+}
+%\end{tabular}
+
+We give a necessary and sufficient condition (called ICPPL)\tnote[GTC]{say
+  a few words about icppl} for {\FTPL} to output in affirmative. This
+characterization can be checked in polynomial time.  The most
+interesting consequence is that in our constructive procedure, it is
+sufficient to iteratively check if three-way intersection
+cardinalities are preserved.  In other words, in each iteration, it is
+sufficient to check if the intersection of any three hyperedges is of
+the same cardinality as the intersection of the corresponding paths.
+Thus this generalizes the well studied question of the feasible
+interval assignment problem which is the special case when the target
+tree $T$ is simply a path \cite{wlh02,nsnrs09}.
+
+\tnote{A few lines on the algo itself.}
+
+Aside from checking if a given TPL is feasible, we also solve the
+problem of computing a feasible TPL for a given hypergraph and target
+tree, if one exists. This problem, {\CFTPL}, is defined as follows.
+
+{\small
+\begin{minipage}[h]{5in}
+ % \begin{singlespace}
+ \vspace{2mm}
+  {\large \CFTPL}\\
+  % 
+  \begin{tabular}[t]{l|l}
+    \hline\\
+    {\tt Input} & 
+    \begin{minipage}[t]{\probdefwidth}
+      A hypergraph $\cF$ with vertex set $U$ and a tree $T$.\\
+    \end{minipage}\\
+
+    {\tt Question} &
+    \begin{minipage}[t]{\probdefwidth}
+      Does there exist a set of paths $\cP$ from $T$ and a bijection
+      $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns {\bf
+        true} on $(\cF, T, \cl)$.
+    \end{minipage}\\
+  \end{tabular}
+  % \end{singlespace}
+\end{minipage}\\
+}
+
+\def \kstar {$k$-subdivided star}
+\def \CFTPLKTREE {\sc Compute $k$-subdivided Star Path Labeling}
+
+We present a polynomial time algorithm for {\CFTPL} when the target
+tree $T$ belongs to a special class of trees called {\em \kstar s} and
+when the hyperedges in the hypergraph $\cF$ have at most $k+2$
+vertices. A couple of examples of {\kstar s} are given in
+Figure~\ref{fig:ksubstar}.
+
+%This problem, {\CFTPLKTREE}, is defined as follows. \\
+\begin{figure}[t]
+  \centering
+  \begin{tabular}[h]{ccccc}
+  \includegraphics[scale=0.3]{../img/star.pdf} &&&&
+  \includegraphics[scale=0.3]{../img/kstar.pdf}\\
+  (a) &&&& (b)
+  \end{tabular}
+  \caption{\figtabsize Examples of {\kstar s}. (a) $k = 0$
+    (b) $k = 2$ }
+  \label{fig:ksubstar}
+\end{figure}
+
+%\vspace{3mm}
+{\small
+\begin{minipage}[h]{5in}
+ % \begin{singlespace}
+ \vspace{2mm}
+  {\large \CFTPLKTREE}\\
+  \begin{tabular}[t]{l|l}
+    \hline\\
+    {\tt Input} & 
+    \begin{minipage}[t]{\probdefwidth}
+      A hypergraph $\cF$ with vertex set $U$ such that every hyperedge
+      $S \in \cF$ is of cardinality at most $k+2$ and a {\kstar} $T$.\\
+    \end{minipage}\\
+    {\tt Question} &
+    \begin{minipage}[t]{\probdefwidth}
+      Does there exist a set of paths $\cP$ from $T$ and a bijection
+      $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns {\bf
+        true} on $(\cF, T, \cl)$.
+    \end{minipage}\\
+  \end{tabular}
+  % \end{singlespace}
+\end{minipage}\\
+}
+
+In spite of this being a restricted case, we believe that our results
+are of significant interest in understanding the nature of {\sc Graph
+  Isomorphism} which is polynomial time solvable in interval graphs
+while being hard on path graphs\cite{kklv10}. {\kstar s} are a class
+of trees which are in many ways very close to intervals or paths. Each
+ray\footnote{The path from a leaf to the root, the vertex with highest
+degree, is called a {\em ray} of the \kstar.} are independent except
+for the root\footnote{The vertex with maximum degree in a {\kstar} is
+  called {\em root}} and hence can be considered as an independent interval
+till the root. Our algorithm builds on this fact and uses the interval
+assignment algorithm\cite{nsnrs09} up until ``reaching'' the root and
+then uses the trio-wise intersection cardinality (the extra condition
+in ICPPL after ICPIA) check to resolve the
+ambiguity about which ray the algorithm should ``grow'' the solution
+into in the next iteration.
+
+We also have an algorithm for solving {\CFTPL} with no restrictions
+which runs in exponential time. 
+% and explores a partial order theory of
+% matrices related to {FTPL}.
+% gives interesting insight
+% into the problem of 
+This algorithm finds a path labeling from $T$ by decomposing the
+problem into smaller subproblems of finding path labeling of subsets
+of $\cF$ from subtrees of $T$. Given the fact that binary matrices
+naturally represent a set
+system % , we look at the matrix representation of subsets of $\cF$.
+and that the overlap relation between the sets involved is an obvious
+equivalence relation, $\cF$ naturally partitions into equivalence
+classes known as overlap components. In the context of COP, {\em overlap
+components} were used in \cite{wlh02} and \cite{kklv10}. Moreover,
+\cite{nsnrs09} discovered that these equivalence classes form a total
+order. We extend this to TPL and find that when $\cF$ is a path
+hypergraph\footnote{If there exists an FTPL for a hypergraph $\cF$, it
+  is called a path hypergraph.}, the classes can be partially ordered
+as an in-tree in polynomial time. Once $\cF$ is
+``broken'' into overlap components, one must identify the subtree of
+$T$ that it needs to map to and this is the hard part which is
+currently open to be solved in polynomial time.
+
+\tnote[TH]{The connection of TPL to graph
+  isomorphism will be made later in the document}
+
+% In the later part of this paper, we decompose our search for a
+% bijection between $U$ and $V(T)$ into subproblems.  Each subproblem
+% is on a set subsystem in which for each set, there is another set in
+% the set subsystem with which the intersection is {\em strict}, i.e.,
+% there is a non-empty intersection, but neither is contained in the
+% other.  This is in the spirit of results in \cite{wlh02,nsnrs09}
+% where to test for the COP in a given matrix, the COP problem is
+% solved on an equivalent set of prime matrices.
+
+%\subsection{RESULTS}
+% \section{Preliminaries} 
+% \label{sec:prelims}
+
+% \subsection{Hypergraph} 
+% The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
+%   system} of a universe $U$ with $|U| = n$.  The {\em support} of a
+% set system $\F$ denoted by $supp(\cF)$ is the union of all the sets in
+% $\F$; $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this
+% paper, a set system is required to ``cover'' the universe; $ supp(\cF)
+% = U$.
+% % A set system $\cF$ can be alternatively represented by a {\em
+% %   hypergraph} $\H_\cF$ whose vertex set is $supp(\cF)$ and
+% % hyperedges are the sets in $\cF$. This is a known representation for
+% % interval systems in literature \cite{bls99,kklv10}.  We extend this
+% % definition here to path systems.
+% A set system $\cF$ can also be visualized as a {\em
+%   hypergraph}\, vertex set is $supp(\cF)$ and hyperedges are the sets
+% in $\cF$.  The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph
+% $\cF$ is a graph such that its vertex set has a bijection to $\cF$ and
+% there exists an edge between two vertices \iff their corresponding
+% hyperedges have a non-empty intersection \cite{mcg04}.  Two
+% hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to each
+% other, denoted by $\cF' \cong \cF''$, \iff there exists a bijection
+% $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for all sets $A
+% \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ \iff  $B$ is a
+% hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
+% \cite{kklv10}, written as $B=\phi(A)$.
+% % This is called {\em hypergraph isomorphism}.
+
+% %\noindent
+% \subsection{Path Hypergraph from a Tree} 
+% We consider trees $T$ such that $|V(T)|=|U|=n$.  A {\em path system}\,
+% $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P \mid P
+% \subseteq V, \text{ } T[P] \text{ is a path} \}$.  This generalizes
+% the fact, from the literature \cite{bls99,kklv10}, that intervals can
+% be viewed as sub-paths of a path.
+% % Due to the equivalence of set system and hypergraph in the scope of
+% % this paper, we drop the subscript $_H$ in the notation and refer to
+% % both the structures by $\cF$.
+% % \vspace{\prelimspace}%\tnote{also extend $\phi$ to hyperedges -- see if
+% % required } \vspace{\prelimspace} \vspace{\prelimspace}
+% % \noindent
+% If the intersection graphs of $\cF$ and $\cP$( a path system) are
+% isomorphic, $\bI(\cF) \cong \bI(\cP)$, then the associated bijection
+% $\cl: \cF \rightarrow \cP$ due to this isomorphism is called a {\em
+%   path labeling} of the hypergraph $\cF$.  Note that there are two
+% kinds of isomorphisms here.  We are concerned about the isomrophisms
+% intersection graphs on $\cF$ and $\cP$, and also the isomorphism
+% between the hypergraph $\cF$ and $\cP$.
+% % To illustrate further, let $\cg: V(\cF) \rightarrow V(\cP)$ be the
+% % above mentioned isomorphism where $V(\cF)$ and $V(\cP)$ are the
+% % vertex sets that represent the hyperedges for each hypergraph
+% % respectively, $V(\cF) = \{ v_S \mid S \in \cF\}$ and $V(\cP) = \{
+% % v_P \mid P \in \cP\}$. Then the path labeling $\cl$ is defined as
+% % follows: $\cl(S_1) = P_1$ \iff $\cg (v_{S_1}) = v_{P_1}$.  Just to
+% % emphasize, for a path labeling $\cl$ of $\cF$ with $\cP$ as the path
+% % system, $\cF^\cl$ is same as $\cP$.  The path system $\cP$ may be
+% % alternatively denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In
+% % most scenarios in this paper, what is given are the pair $(\cF,
+% % \cl)$ and the target tree $T$; hence this notation will be used more
+% % often.
+% If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$ is called
+% a {\em path hypergraph} and $\cP$ is called {\em path representation}
+% of $\cF$. If this isomorphism is $\phi: supp(\cF) \rightarrow V(T)$,
+% then it is clear that there is a path labeling $\cl_\phi: \cF
+% \rightarrow \cP$ to the set system; $\cl_\phi(S) = \set{y \mid y =
+%   \phi(x), x \in S}$ for all $S \in \cF$. In other words, if $\cF
+% \cong \cP$, we get a path labeling.  Recall that $supp(\cP) = V(T)$.
+% % \vspace{\prelimspace} A graph $G$ is a {\em path graph} if it is
+% % isomorphic to the intersection graph $\bI(\cP)$ of a path system
+% % $\cP$.  This isomorphism gives a bijection $\cl': V(G) \rightarrow
+% % \cP$. Moreover, for the purposes of this paper, we require that in a
+% % path labeling, $supp(\cP) = V(T)$.  If graph $G$ is also isomorphic
+% % to $\bI(\cF)$ for some hypergraph $\cF$, then clearly there is a
+% % bijection $\cl: \cF \rightarrow \cP$ such that $\cl(S) = \cl'(v_S)$
+% % where $v_S$ is the vertex corresponding to set $S$ in $\bI(\cF)$ for
+% % any $S \in \cF$. This bijection $\cl$ is called the {\em path
+% %   labeling} of the hypergraph $\cF$ and the path system $\cP$ may be
+% % alternatively denoted as $\cF^\cl$.  \vspace{\prelimspace}
+% In this work, we are given as input $\cF$ and a tree $T$, and the
+% question is whether there is a path labeling $\cl$ to a set of paths
+% in $T$.  We refer to such a solution path system by $\cF^\cl$.  A path
+% labeling $(\cF, \cl)$ is defined to be {\em feasible} if
+% % $\cF \cong \cF^\cl$ and this
+% there is a hypergraph isomorphism $\phi: supp(\cF) \rightarrow
+% supp(\cF^\cl)=V(T)$ induces a path labeling $\cl_\phi: \cF \rightarrow
+% \cF^\cl$ such that $\cl_\phi = \cl$.
+% % \vspace{\prelimspace}
+
+% \subsection{Overlap Graphs and Marginal Hyperedges} 
+% An {\em
+%   overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is a graph such
+% that its vertex set has a bijection to $\cF$ and there exists an edge
+% between two of its vertices \iff their corresponding hyperedges
+% overlap. Two hyperedges $S$ and $S'$ are said to {\em overlap},
+% denoted by $S \overlap S'$, if they have a non-empty intersection and
+% neither is contained in the other; $S \overlap S' \text{ iff } S \cap
+% S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq S$. Thus $\bO(\cF)$
+% is a spanning subgraph of $\bI(\cF)$ and not necessarily
+% connected. Each connected component of $\bO(\cF)$ is called an {\em
+%   overlap component}.
+% % If there are $d$ overlap components in $\bO(\cF)$, the set
+% % subsystems are denoted by $\cO_1, \cO_2, \ldots \cO_d$. Clearly
+% % $\cO_i \subseteq \F, i \in [d]$. For any $i, j \in [d]$, it can be
+% % verified that one of the following is true.
+% % \begin{enumerate}[a) ]
+% % \item $supp(\cO_i)$ and $supp(\cO_j)$ are disjoint
+% % \item $supp(\cO_i)$ is a subset of a set in $\cO_j$
+% % \item $supp(\cO_j)$ is a subset of a set in $\cO_i$
+% % \end{enumerate}
+% % \vspace{\prelimspace} \xnoindent
+% A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
+% \overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
+% \cite{kklv10}. Additionally, if $S$ is such that it is contained in no
+% other marginal hyperedge in $\cF$, then it is called {\em
+%   super-marginal}.
+% % i.e., it is inclusion maximal then it is called {\em
+% %   super-marginal}.
+
+% \subsection{$k$-subdivided Star -- A Special Tree} 
+% A {\em star} graph
+% is a complete bipartite graph $K_{1,p}$ which is clearly a tree and
+% $p$ is the number of leaves. The vertex with maximum degree is called
+% the {\em center} of the star and the edges are called {\em rays} of
+% the star. A {\em $k$-subdivided star} is a star with all its rays
+% subdivided exactly $k$ times. The definition of a {\em ray of a
+%   $k$-subdivided star} is extended to the path from the center to a
+% leaf. It is clear that all rays are of length $k+2$. See
+% fig.~\ref{fig:kstar} for examples.
+
+% \begin{figure}[htbp] 
+%   \centering
+%   (i) \includegraphics{cop-tree_thesis.1} \ \ \ \ \ \  
+%   (ii)  \includegraphics{cop-tree_thesis.2}
+%   \caption{(i) $8$-subdivided star with 7 rays, (ii) 3-subdivided
+%     star with 3 rays}
+%   \label{fig:kstar}
+% \end{figure}
+
+% \subsection{Characterization of Feasible Tree Path  Labelings} \tnote[TH]{UNCOMMENT all proofs}
+% \label{sec:feasible} 
+% \vspace{\secfirstparatrim}
+
+% In this section we give an algorithmic characterization of a
+% feasibility of tree path labeling.
+% % Consider a path labeling $\cl: \cF \rightarrow \cP$ for set system $\cF$
+% % and path system $\cP$ on the given tree $T$. We
+% % call $\cl$ an {\em Intersection Cardinality Preserving Path Labeling
+% %   (ICPPL)} if it has the following properties.
+% Consider a path labeling $(\cF, \cl)$ on the
+% given tree $T$. We call $(\cF, \cl)$ an {\em Intersection Cardinality
+%   Preserving Path Labeling (ICPPL)} if it has the following
+% properties.
+
+% % \begin{enumerate}[{\icpplpr\ }i. ]
+% % \item \label{pr:i} $|S| = |\cl(S)|$, \hfill for all $S \in \cF$
+% %   % \vspace{\topshrink}
+% % \item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$, \hfill
+% %   for all distinct $S_1, S_2 \in \cF$
+% %   % \vspace{\topshrink}
+% % \item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
+% %   \cap \cl(S_3)|$, \hfill for all distinct $S_1, S_2, S_3 \in \cF$
+% % \end{enumerate}
+
+% \begin{table}[h]
+%   \centering
+%   \begin{tabular}[l]{rrcll}
+%     Pr i. & \label{xpr:i} $|S| $&$=$&$ |\cl(S)|$ & $\forall$ $S \in \cF$\\
+%     Pr ii. & \label{xpr:ii}$|S_1 \cap S_2| $&$=$&$ |\cl(S_1) \cap
+%     \cl(S_2)|$ &  $\forall$ distinct $S_1, S_2 \in \cF$\\
+%     Pr iii. & \label{xpr:iii}$|S_1 \cap S_2 \cap S_3| $&$=$&$
+%     |\cl(S_1) \cap \cl(S_2) \cap \cl(S_3)|$ & $\forall$ distinct $S_1,
+%     S_2, S_3 \in \cF$ 
+%   \end{tabular}
+%   \caption{ICPPL properties}
+%   \label{tab:icpplpr}
+% \end{table}
+
+% \noindent
+% The following lemma is useful in subsequent arguments. 
+% \begin{lemma}
+%   \label{lem:setminuscard}
+%   If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
+%   (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
+%   \cl(S_3))|$.
+% \end{lemma}
+
+% % \begin{proof}%[Proof of Lemma~\ref{lem:setminuscard}]
+% %   Let $P_i = \cl(S_i)$, for all $i \in \set{1,2,3}$. We have $|S_1
+% %   \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| = |S_1
+% %   \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to properties
+% %   (\ref{pr:ii}) and (\ref{pr:iii}) of ICPPL, $|S_1 \cap S_2| - |S_1
+% %   \cap S_2 \cap S_3| = |P_1 \cap P_2| - |P_1 \cap P_2 \cap P_3| =
+% %   |(P_1 \cap P_2) \setminus P_3| = |P_1 \cap (P_2 \setminus
+% %   P_3)|$. Thus lemma is proved. \qed
+% % \end{proof}
+
+% In the remaining part of this section we show that $(\cF, \cl)$ is
+% feasible if and only if it is an ICPPL and
+% Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
+% function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two
+% levels of filtering of $(\cF, \cl)$ to make it simpler while retaining
+% the set of isomorphisms, if any, between $\cF$ and $\cF^\cl$.
+% % One direction of this claim isclear: that if a path
+% % labeling is feasible, then all intersection cardinalities are
+% % preserved, i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
+% % \annote{has no premature exit condition hence any input will go
+% %   through it}{Prove that the filtered sets has ICPPL \iff input PL
+% %   has ICPPL?}. Algorithm~\ref{leafasgn} has an exit condition at
+% % line~\ref{xempty}. It can be easily verified that $X$ cannot be
+% % empty if $\cl$ is a feasible path labeling. The reason is that a
+% % feasible path labeling has an associated bijection between
+% % $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that
+% % the sets map to paths, ``preserving'' the path labeling.  The rest
+% % of the section is devoted to constructively proving that it is
+% % sufficient for a path labeling to be an ICPPL and pass the two
+% % filtering algorithms.  To describe in brief, the constructive
+% % approaches refine an ICPPL iteratively, such that at the end of each
+% % iteration we have a ``filtered'' path labeling, and finally we have
+% % a path labeling that defines a family of bijections from $supp(\cF)$
+% % to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.
+
+% First, we present Algorithm~\ref{perms} called {\tt filter\_1}
+% % , and
+% % prove its correctness.  
+% This algorithm refines the path labeling by
+% processing pairs of paths in $\cF^\cl$ that share a leaf until no two
+% paths in the new path labeling share any leaf.
+% \tnote{write a note summarizing filter 1}
+% % \begin{algorithm}[h]
+% %   \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
+% %   \label{perms}
+% %   \begin{algorithmic}[\lndisplay]
+% %     \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$\\
+% %     \STATE $j \assign 1$\\
+% %     \WHILE {there is $S_1, S_2 \in \cF_{j-1}$ such that
+% %       $\cl_{j-1}(S_1)$ and $\cl_{j-1}(S_2)$ have a common leaf in
+% %       $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
+% %     \{S_1, S_2\})
+% %     \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1 \}$ \label{setbreak} 
+% %     \COMMENT {Remove $S_1$, $S_2$ and add the ``filtered'' sets}
+% %     \STATE {\bf for} every $S \in \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
+% %     S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\
+
+% %     \STATE $\cl_j(S_1 \cap S_2) \assign \cl_{j-1}(S_1) \cap
+% %     \cl_{j-1}(S_2)$
+% %     \COMMENT {Carry forward the path labeling for all existing sets other than
+% %       $S_1$, $S_2$}
+% %     \STATE $\cl_j(S_1 \setminus S_2) \assign \cl_{j-1}(S_1) \setminus
+% %     \cl_{j-1}(S_2)$ 
+% %     \COMMENT {Define path labeling for new sets}
+% %     \STATE $\cl_j(S_2 \setminus S_1) \assign \cl_{j-1}(S_2) \setminus
+% %     \cl_{j-1}(S_1)$
+
+% %     \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of ICPPL}
+% %     \label{ln:3waycheck}
+% %     \STATE {\bf exit} \label{ln:exit1} \\
+% %     \ENDIF
+
+% %     \STATE $j \assign j+1$\\
+% %     \ENDWHILE
+% %     \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
+% %     \RETURN $(\cF', \cl')$
+% %   \end{algorithmic}
+% % \end{algorithm}
+
+% \begin{lemma} 
+%  \label{lem:feasible} 
+%  In Algorithm~\ref{perms}, if input $(\cF, \cl)$ is a feasible path
+%  assignment then at the end of $j$th iteration of the {\bf while}
+%  loop, $j \ge 0$, $(\cF_j, \cl_j)$ is a feasible path assignment.
+% \end{lemma}
+
+% % \begin{proof}%[Proof of Lemma~\ref{lem:feasible}]
+% %   We will prove this by mathematical induction on the number of
+% %   iterations. The base case $(\cF_0, \cl_0)$ is feasible since it is
+% %   the input itself which is given to be feasible. Assume the lemma is
+% %   true till $j-1$th iteration. i.e. every hypergraph isomorphism
+% %   $\phi: supp\left(\cF_{j-1}\right) \rightarrow V\left(T \right)$ that
+% %   defines $(\cF, \cl)$'s feasibility, is such that the induced path
+% %   labeling on $\cF_{j-1}$, $\cl_{\phi[{\cF_{j-1}}]}$ is equal to
+% %   $\cl_{j-1}$. We will prove that $\phi$ is also the bijection that
+% %   makes $(\cF_j, \cl_j)$ feasible. Note that $supp(\cF_{j-1}) =
+% %   supp(\cF_{j})$ since the new sets in $\cF_j$ are created from basic
+% %   set operations to the sets in $\cF_{j-1}$. For the same reason and
+% %   $\phi$ being a bijection, it is clear that when applying the $\phi$
+% %   induced path labeling on $\cF_j$, $ \cl_{\phi[{\cF_{j}}]}(S_1
+% %   \setminus S_2) = \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
+% %   \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Now observe that $ \cl_j(S_1
+% %   \setminus S_2) = \cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2) =
+% %   \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
+% %   \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Thus the induced path labeling
+% %   $\cl_{\phi[{\cF_{j}}]} = \cl_{j}$. Therefore lemma is proved.  \qed
+% % \end{proof}
+
+% \begin{lemma}
+%   \label{lem:invar1} In Algorithm~\ref{perms}, at the end of $j$th
+%   iteration, $j \ge 0$, of the {\bf while} loop, the following
+%   invariants are maintained.
+% %   \begin{enumerate}[I {\ }] %\vspace{\topshrink}
+% %   \item $\cl_j(R)$ is a path in $T$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
+% %     \ \ \ \ \ \ \ \ \ \ for all $R \in \cF_j$%\vspace{\topshrink}
+% %   \item $|R| = |\cl_j(R)|$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
+% %     \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $R \in
+% %     \cF_j$%\vspace{\topshrink}
+% %   \item $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, \ \ \ \ \ \ \ \ \ \
+% %     \ \ \ \ \ \ \ \ \ \ for all $R, R' \in \cF_j$%\vspace{\topshrink}
+% %   \item $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
+% %     \cl_j(R'')|$, \ \ \ for all $R, R', R'' \in \cF_j$
+% %   \end{enumerate}
+
+%   \begin{tabular}[h]{lll}
+%     I & $\cl_j(R)$ is a path in $T$, & for all $R \in \cF_j$\\ 
+%     II  & $|R| = |\cl_j(R)|$, & for all $R \in \cF_j$ \\ 
+%     III  & $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, & for all $R, R'
+%     \in \cF_j$\\ 
+%     IV  & $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
+%     \cl_j(R'')|$, & for all $R, R', R'' \in \cF_j$ 
+%   \end{tabular}
+% \end{lemma}
+
+% % \begin{proof}
+% %  Proof is by induction on the number of iterations, $j$. In this
+% %   proof, the term ``new sets'' will refer to the sets added to $\cF_j$
+% %   in $j$th iteration in line~\ref{setbreak} of Algorithm~\ref{perms},
+% %   $S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1$ and its
+% %   images in $\cl_j$ where $\cl_{j-1}(S_1)$
+% %   and $\cl_{j-1}(S_2)$ intersect and share a leaf.\\
+% %   The invariants are true in the base case $(\cF_0, \cl_0)$, since it
+% %   is the input ICPPL.  Assume the lemma is true till the $j-1$th
+% %   iteration. Let us consider the possible cases for each of the above invariants for
+% %   the $j$th iteration.
+
+% %   \begin{enumerate}[\textreferencemark]
+% %   \item {\em Invariant} I/II
+% %     \begin{enumerate}[{I/II}a $|$] % \textbullet 
+% %     \item {\em $R$ is not a new set.} It is in $\cF_{j-1}$. Thus
+% %       trivially true by induction hypothesis.
+% %     \item {\em $R$ is a new set.} If $R$ is in $\cF_{j}$ and not in
+% %       $\cF_{j-1}$, then it must be one of the new sets added in
+% %       $\cF_j$. In this case, it is clear that for each new set, the
+% %       image under $\cl_j$ is a path since by definition the chosen
+% %       sets $S_1$, $S_2$ are from $\cF_{j-1}$ and due to the while loop
+% %       condition, $\cl_{j-1}(S_1)$, $\cl_{j-1}(S_2)$ have a
+% %       common leaf. Thus invariant I is proved.\\
+% %       Moreover, due to induction hypothesis of invariant III and the
+% %       definition of $l_j$ in terms of $l_{j-1}$, invariant II is
+% %       indeed true in the $j$th iteration for any of the new sets.  If
+% %       $R = S_1 \cap S_2$, $|R| = |S_1 \cap S_2| = |\cl_{j-1}(S_1) \cap
+% %       \cl_{j-1}(S_2)| = |\cl_j(S_1 \cap S_2)| = |\cl_j(R)|$.
+% %       If $R = S_1 \setminus S_2$, $|R| = |S_1 \setminus S_2| = |S_1| -
+% %       |S_1 \cap S_2| = |\cl_{j-1}(S_1)| - |\cl_{j-1}(S_1) \cap
+% %       \cl_{j-1}(S_2)| = |\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)| =
+% %       |\cl_j(S_1 \setminus S_2)|
+% %       = |\cl_j(R)|$. Similarly if $R = S_2 \setminus S_1$.\\
+% %     \end{enumerate}
+% %   \item {\em Invariant} III
+% %     \begin{enumerate}[{III}a $|$]
+% %     \item {\em $R$ and $R'$ are not new sets.} It is in
+% %       $\cF_{j-1}$. Thus trivially true by induction hypothesis.
+% %     \item {\em Only one, say $R$, is a new set.} Due to invariant IV
+% %       induction hypothesis, Lemma~\ref{lem:setminuscard} and
+% %       definition of $\cl_j$, it follows that invariant III is true no
+% %       matter which of the new sets $R$ is equal to. If $R = S_1 \cap
+% %       S_2$, $|R \cap R'| = |S_1 \cap S_2 \cap R'| = |\cl_{j-1}(S_1)
+% %       \cap \cl_{j-1}(S_2) \cap \cl_{j-1}(R')| = |\cl_j(S_1 \cap S_2)
+% %       \cap \cl_j(R')| = |\cl_j(R) \cap \cl_j(R')|$.  If $R = S_1
+% %       \setminus S_2$, $|R \cap R'| = |(S_1 \setminus S_2) \cap R'| =
+% %       |(\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)) \cap \cl_{j-1}(R')| =
+% %       |\cl_{j}(S_1 \cap S_2) \cap \cl_{j}(R')| = |\cl_{j}(R) \cap
+% %       \cl_{j}(R')|$. Similarly, if $R = S_2 \setminus
+% %       S_1$. Note $R'$ is not a new set.\\
+
+% %     \item {\em $R$ and $R'$ are new sets.} By definition, the new
+% %       sets and their path images in path label $\cl_j$ are disjoint so
+% %       $|R \cap R'| = |\cl_j(R) \cap \cl_j(R)| = 0$. Thus case proved.
+% %     \end{enumerate}
+% %   \item {\em Invariant} IV
+
+% %     Due to the condition in line~\ref{ln:3waycheck}, this invariant is
+% %     ensured at the end of every iteration.
+% % %     \begin{enumerate} [{Case 3.}1:]
+% % %     \item {\em $R$, $R'$ and $R''$ are not new sets.} Trivially
+% % %       true by induction hypothesis.
+% % %     \item {\em Only one, say $R$, is a new set.}
+% % %       If $R = S_1 \cap S_2$,  from Lemma~\ref{lem:fourpaths} and
+% % %       invariant III hypothesis,  this case is proven. Similarly if $R$
+% % %       is any of the other new  sets, the case is proven by also using
+% % %       Lemma ~\ref{lem:setminuscard}.
+% % %     \item {\em At least two of $R, R', R''$ are new sets.}
+% % %       The new sets are disjoint hence this case is vacuously true.
+% % %     \end{enumerate}
+% %   \end{enumerate} \qed
+% % %\vspace{-6mm}
+
+% % \end{proof}
+
+% \begin{lemma}
+%   \label{lem:noexit1}
+%   If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{perms} is
+%   feasible, then the set of hypergraph isomorphism functions that
+%   defines $(\cF, \cl)$'s feasibility is the same as the set that
+%   defines $(\cF_j, \cl_j)$'s feasibility, if any.  Secondly, for any
+%   iteration $j > 0$ of the {\em \bf while} loop, the {\em \bf exit}
+%   statement in line~\ref{ln:exit1} will not execute.
+% \end{lemma}
+% \tnote[TH]{UNCOMMENT}
+% % \begin{proof}
+% %   Since $(\cF,\cl)$ is feasible, by Lemma~\ref{lem:feasible}
+% %   $(\cF_j,\cl_j)$ for every iteration $j > 0$ is feasible.  % Therefore,
+% % %   every hypergraph isomorphism $\phi: supp(\cF) \rightarrow V(T)$ that
+% % %   induces $\cl$ on $\cF$ also induces $\cl_{j-1}$ and $\cl_{j}$ on
+% % %   $\cF_{j-1}$ and $\cF_{j}$ respectively, i.e., $\cl_{\phi[\cF_{j-1}]}
+% % %   = \cl_{j-1}$ and $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
+% % %   that for all $x \in supp(\cF)$, for all $v \in V(T)$ the following
+% % %   hold true.
+% %   Also, every hypergraph isomorphism $\phi: supp(\cF) \rightarrow
+% %   V(T)$ that induces $\cl$ on $\cF$ also induces $\cl_{j}$ on
+% %   $\cF_{j}$, i.e., $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
+% %   that for all $x \in supp(\cF)$, for all $v \in V(T)$, if $(x,v) \in
+% %   \phi$ then $v \in \cl_{j}(S)$ for all $S \in \cF_{j}$ such that $x
+% %   \in S$.
+% % % the following
+% % %   hold true.
+% % %   \begin{enumerate}[i. ]
+% % %   \item If $(x,v) \in \phi$ then $v \in \cl_{j-1}(S)$ for all $S \in
+% % %     \cF_{j-1}$ such that $x \in S$.
+% % %   \item If $(x,v) \in \phi$ then $v \in \cl_{j}(S)$ for all $S \in
+% % %     \cF_{j}$ such that $x \in S$
+% % %   \end{enumerate}
+% %   In other words, filter 1 outputs a filtered path labeling that
+% %   ``preserves''
+% %   hypergraph isomorphisms of the original path labeling.\\
+% %   Secondly, line~\ref{ln:exit1} will execute \iff the exit condition in
+% %   line~\ref{ln:3waycheck}, i.e. failure of three way intersection
+% %   preservation, becomes true in any iteration of the {\em \bf while}
+% %   loop.  Due to Lemma~\ref{lem:invar1} Invariant IV, the exit
+% %   condition does not occur if the input is a feasible ICPPL.\qed
+
+% % %   such that $\phi(x) = v$ where $v$ is the leaf considered in the
+% % %   first iterations of while. Clearly, $\phi$ is a renaming of
+% % %   vertices in hypergraph $\cF$ to those in hypergraph $\cF^\cl$. Thus
+% % %   the following facts can be observed in every iteration of the loop.
+
+% % %   \begin{enumerate}[\hspace{2mm}i. ] \vspace{\topshrink}
+% % %   \item all intersection cardinalities are preserved in this path
+% % %     labeling \vspace{\topshrink}
+% % %   \item element $x$ is exclusive in a hyperedge in $\cF$ since $v$ is
+% % %     exclusive in a hyperedge in $\cF^\cl$.
+% % %   \end{enumerate}
+
+% % %   Thus the exit condition is never rendered true after $x$ and $v$ are
+% % %   removed from their respective hyperedges. \qed
+
+% % % \noindent
+% % % This proof uses mathematical induction on the number
+% % %   of iterations $j$, $j \ge 0$, of the loop that executed
+% % %   without exiting. The base case, $j = 0$ is obviously true since the
+% % %   input is an ICPPL and the exit condition cannot hold true due to
+% % %   ICPPL property (iii).  Assume the algorithm executes till the end
+% % %   of $j-1$th iteration without exiting at line
+% % %  ~\ref{ln:3waycheck}. Consider the $j$th iteration. From Lemma
+% % %  ~\ref{lem:feasible} we know that $(\cF_j, \cl_j)$ and $(\cF_{j-1},
+% % %   \cl_{j-1})$ are feasible\remove[AS]{and from the proof in lemma
+% % %     lem:invar1 we know that $(\cF_{j-1}, \cl_{j-1})$ satisfies all the
+% % %     invariants defined in the lemma}.  Thus there exists a bijection
+% % %   $\phi: supp(\cF) \rightarrow V(T)$ such that the induced path
+% % %   % labeling on $\cF_{j-1}$ $\cl_{\phi[\cF_{j-1}]} = \cl_{j-1}$.
+% % %   labeling on $\cF_{j}$, $\cl_{\phi[\cF_{j}]}$ and on $\cF_{j-1}$,
+% % %   $\cl_{\phi[\cF_{j-1}]}$ are equal to $\cl_{j}$ and $\cl_{j-1}$
+% % %   respectively.  We need to prove that for any $R, R', R'' \in
+% % %   \cF_{j}$, $|R \cap R' \cap R''| = |\cl_j(R) \cap \cl_j(R') \cap
+% % %   \cl_j(R'')|$.
+% % %   The following are the possible cases that could arise. From argument
+% % %   above, $|\cl_j(R) \cap \cl_j(R') \cap \cl_j(R'')| =
+% % %   |\cl_{\phi[\cF_{j}]}(R) \cap \cl_{\phi[\cF_{j}]} (R') \cap
+% % %   \cl_{\phi[\cF_{j}]} (R'')|$
+
+% % %   \begin{enumerate}[a $|$]
+% % %   \item {\em None of the sets are new. $R, R', R'' \in \cF_{j-1}$.}
+% % %     We know $(\cF_{j-1}, \cl_{j-1})$ is feasible. Thus $|R \cap R'
+% % %     \cap R''| = |\cl_{j-1}(R) \cap \cl_{j-1}(R') \cap \cl_{j-1}(R'')|
+% % %     = |\cl_{j}(R) \cap \cl_{j}(R') \cap \cl_{j}(R'')|$.
+% % %   \item {\em Only one, say $R$, is a new set.}  Let $R = S_1 \cap S_2$
+% % %     ($S_1, S_2$ are defined in the proof of lemma
+% % %    ~\ref{lem:invar1}). Now we have $|R \cap R' \cap R''| = |S_1 \cap
+% % %     S_2 \cap R' \cap R''| = |\cl_{j-1}(S_1) \cap \cl_{j-1}(S_2) \cap
+% % %     \cl_{j-1}(R') \cap \cl_{j-1}(R'')| = |\cl_{j}(R) \cap \cl_{j}(R')
+% % %     \cap \cl_{j}(R'')|$. Thus proven. If $R$ is any of the other new
+% % %     sets, the same claim can be verified using lemma
+% % %    ~\ref{lem:setminuscard}.
+% % %     % \item []{\bf Case 3:}
+% % %   \item {\em At least two of $R, R', R''$ are new sets.}  The new sets
+% % %     are disjoint hence this case is vacuously true.
+% % %   \end{enumerate}
+% % %   \qed
+% % %   \tnote[E2]{remove the induction proof. just text saying x and v are
+% % %     exclusive in these sets therefore the intersection cardinalities
+% % %     don't change thus all invariants are still true} 
+% % \end{proof}
+
+% As a result of Algorithm~\ref{perms} each leaf $v$ in $T$
+% is such that there is exactly one set in $\cF$ with $v$ as a vertex in
+% the path assigned to it.  In Algorithm~\ref{leafasgn} we identify
+% elements in $supp(\cF)$ whose images are leaves in a hypergraph
+% isomorphism if one exists.  Let $S \in \cF$ be such that $\cl(S)$ is a
+% path with leaf and $v \in V(T)$ is the unique leaf incident on it.  We
+% define a new path labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) =
+% \set{v}$ where $x$ an arbitrary element from $S \setminus \bigcup_{\hS
+%   \ne S} \hS$. In other words, $x$ is an element present in no other
+% set in $\cF$ except $S$. This is intuitive since $v$ is present in no
+% other path image under $\cl$ other than $\cl(S)$.  The element $x$ and
+% leaf $v$ are then removed from the set $S$ and path $\cl(S)$
+% respectively. After doing this for all leaves in $T$, all path images
+% in the new path labeling $\cl_{new}$ except leaf labels (a path that
+% has only a leaf is called the {\em leaf label} for the corresponding
+% single element hyperedge or set) are paths from a new pruned tree $T_0
+% = T \setminus \{v \mid v \text{ is a leaf in }
+% T\}$. Algorithm~\ref{leafasgn} is now presented with details.
+
+% \tnote{write a note summarizing filter 1}
+% % \begin{algorithm}[h]
+% %   \caption{Leaf labeling from an ICPPL {\tt filter\_2($\cF, \cl, T$)}}
+% %   \label{leafasgn}
+% %   \begin{algorithmic}[\lndisplay]
+% %     \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$
+% %     \COMMENT {Path images are such that no two path images share a
+% %       leaf.}
+% %     \STATE $j \assign 1$\\
+% %     \WHILE {there is a leaf $v$ in $T$ and a unique $S_1 \in
+% %       \cF_{j-1}$ such that $v \in \cl_{j-1}(S_1)$ }\label{uniqueleaf}
+% %     \STATE $\cF_j \assign \cF_{j-1} \setminus \{S_1\}$\\
+% %     \STATE for all $S \in \cF_{j-1}$ such that $S \ne S_1$ set
+% %     $\cl_j(S) \assign
+% %     \cl_{j-1}(S)$\\
+% %     \STATE $X \assign S_1 \setminus \bigcup_{S \in \cF_{j-1}, S \ne S_1}S$\\
+% %     \IF{$X$ is empty} \label{xempty} \STATE {\bf exit} \label{ln:exit2} \ENDIF
+% %     \STATE $x \assign $ arbitrary element from $X$\\
+% %     \STATE $\cF_j \assign \cF_j \cup \{\{x\}, S_1 \setminus \{x\}\} $\\
+% %     \STATE $\cl_j(\{x\}) \assign \{v\}$\\
+% %     \STATE $\cl_j(S_1 \setminus \{x\}) \assign \cl_{j-1}(S_1) \setminus \{v\}$\\
+% %     \STATE $j \assign j+1$\\
+% %     \ENDWHILE
+% %     \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
+% %     \RETURN $(\cF', \cl')$
+% %   \end{algorithmic}
+% % \end{algorithm}
+
+% Suppose the input ICPPL $(\cF, \cl)$ is feasible, yet set $X$ in
+% Algorithm~\ref{leafasgn} is empty in some iteration of the {\bf while}
+% loop. This will abort our procedure of finding the hypergraph
+% isomorphism. The following lemma shows that this will not happen.
+
+% \begin{lemma}
+%   \label{lem:xnotempty}
+%   If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{leafasgn} is
+%   feasible, then for all iterations $j > 0$ of the {\em \bf while}
+%   loop, the {\em \bf exit} statement in line~\ref{ln:exit2} does not
+%   execute.
+% \end{lemma}
+
+% % \begin{proof}
+% %   Assume $X$ is empty for some iteration $j > 0$. We know that $v$ is
+% %   an element of $\cl_{j-1}(S_1)$. Since it is uniquely present in
+% %   $\cl_{j-1}(S_1)$, it is clear that $v \in \cl_{j-1}(S_1) \setminus
+% %   \bigcup_{(S \in \cF_{j-1}) \wedge (S \ne S_1)}\cl_{j-1}(S)$.  Note
+% %   that for any $x \in S_1$ it is contained in at least two sets due to
+% %   our assumption about cardinality of $X$. Let $S_2 \in \cF_{j-1}$ be
+% %   another set that contains $x$. From the above argument, we know $v
+% %   \notin \cl_{j-1}(S_2)$. Therefore there cannot exist a hypergraph
+% %   isomorphism bijection that maps elements in $S_2$ to those in
+% %   $\cl_{j-1}(S_2)$. This contradicts our assumption that the input is
+% %   feasible. Thus $X$ cannot be empty if input is ICPPL and feasible.
+% %   \qed
+% % \end{proof}
+
+% \begin{lemma}
+%   \label{lem:invar3}
+%   In Algorithm~\ref{leafasgn}, for all $j > 0$, at the end of the
+%   $j$th iteration of the {\bf while} loop the four invariants given in
+%   Lemma~\ref{lem:invar1} hold.
+% \end{lemma}
+
+% % \begin{proof}
+% %   By Lemma~\ref{lem:xnotempty} we know that set $X$ will not be empty
+% %   in any iteration of the {\em \bf while} loop if input ICPPL $(\cF,
+% %   \cl)$ is feasible and $\cl_j$ is always computed for all $j >
+% %   0$. Also note that removing a leaf from any path keeps the new path
+% %   connected. Thus invariant I is obviously true. In every iteration $j
+% %   > 0$, we remove exactly one element $x$ from one set $S$ in $\cF$
+% %   and exactly one vertex $v$ which is a leaf from one path
+% %   $\cl_{j-1}(S)$ in $T$. This is because as seen in
+% %   Lemma~\ref{lem:xnotempty}, $x$ is exclusive to $S$ and $v$ is
+% %   exclusive to $\cl_{j-1}(S)$. Due to this fact, it is clear that the
+% %   intersection cardinality equations do not change, i.e., invariants
+% %   II, III, IV remain true. On the other hand, if the input ICPPL is
+% %   not feasible the invariants are vacuously true. \qed
+% % \end{proof}
+
+% %% OLD TEXT TB DELETED?
+% % \textcolor{cyan}{
+% % \begin{lemma}
+% %   \label{lem:notfeasibleexit}
+% %   \tnote{IS THIS CORRECT?}  If input ICPPL $(\cF, \cl)$ is not
+% %   feasible, then in one of the recursive calls to Algorithm --3--, the
+% %   {\em \bf exit} statement in line x in Algorithm --1 or line y in
+% %   Algorithm --2 will get executed.
+% % \end{lemma}
+% % }
+% %% END OLD TEXT TB DELETED?
+
+% We have seen two filtering algorithms above, namely,
+% Algorithm~\ref{perms} {\tt filter\_1} and Algorithm~\ref{leafasgn}
+% {\tt filter\_2} which when executed serially respectively result in a
+% new ICPPL on the same universe $U$ and tree $T$. We also proved that
+% if the input is indeed feasible, these algorithms do indeed output the
+% filtered ICPPL. Now we present the algorithmic characterization of a
+% feasible tree path labeling by way of Algorithm~\ref{al:icppl-find-isomorph}.
+
+% Algorithm~\ref{al:icppl-find-isomorph} computes a
+% hypergraph isomorphism $\phi$ recursively using Algorithm~\ref{perms}
+% and Algorithm~\ref{leafasgn} and pruning the leaves of the input
+% tree. In brief, it is done as follows. Algorithm~\ref{leafasgn} gives
+% us the leaf labels in $\cF_2$, i.e., the elements in $supp(\cF)$ that
+% map to leaves in $T$, where $(\cF_2, \cl_2)$ is the output of
+% Algorithm~\ref{leafasgn}. All leaves in $T$ are then pruned away. The
+% leaf labels are removed from the path labeling $\cl_2$ and the
+% corresponding elements are removed from the corresponding sets in
+% $\cF_2$. This tree pruning algorithm is recursively called on the
+% altered hypergraph $\cF'$, path label $\cl'$ and tree $T'$. The
+% recursive call returns the bijection $\phi''$ for the rest of the
+% elements in $supp(\cF)$ which along with the leaf labels $\phi'$ gives
+% us the hypergraph isomorphism $\phi$.  The following lemma formalizes
+% the characeterization of feasible path labeling.
+
+% \tnote{write a note summarizing filter 1}
+% % \begin{algorithm}[h]
+% %   \caption{{\tt get-hypergraph-isomorphism($\cF, \cl, T$)}}
+% %   \label{al:icppl-find-isomorph}
+% %   \begin{algorithmic}[\lndisplay]
+
+% %     \IF{$T$ is empty}
+% %     \RETURN $\emptyset$\\
+% %     \ENDIF
+% %     \STATE $L \assign \{v \mid v \text{ is a leaf in }      T\}$\\
+% %     \STATE $(\cF_1, \cl_1) \assign$ {\tt filter\_1($\cF, \cl,
+% %       T$)}\\
+% %     \STATE $(\cF_2, \cl_2) \assign$ {\tt filter\_2($\cF_1,
+% %       \cl_1, T$)}\\
+
+% %     \STATE $(\cF', \cl') \assign (\cF_2, \cl_2)$\\
+% %     \STATE $\phi' \leftarrow \emptyset$
+
+% %     \FOR {every $v \in L$}
+% %     \STATE $\phi'(x) \assign v$ where $x \in \cl_2^{-1}(\{v\})$
+% %     \COMMENT {Copy the leaf labels to a one to one function $\phi':
+% %       supp(\cF) \rightarrow L$
+% %       }\\
+% %     \STATE Remove $\{x\}$ and $\{v\}$ from $\cF'$, $\cl'$  appropriately\\
+% %     \ENDFOR
+
+% %     \STATE $T' \assign T \setminus L$
+
+% %     \STATE $\phi'' \assign$ {\tt
+% %       get-hypergraph-isomorphism($\cF', \cl', T'$)}
+% %     \STATE $\phi \assign \phi'' \cup \phi'$ \\
+% %     \RETURN $\phi$
+% %   \end{algorithmic}
+% % \end{algorithm}
+
+% \begin{lemma}
+%   \label{lem:hyperiso}  %{lem:perm}
+%   If $(\cF, \cl)$ is an ICPPL from a tree $T$ and
+%   Algorithm~\ref{al:icppl-find-isomorph}, {\tt
+%     get- hypergraph- isomorphism ($\cF, \cl, T$)} returns a non-empty
+%   function, then there exists a hypergraph isomorphism $\phi :
+%   supp(\cF) \rightarrow V(T)$ such that the $\phi$-induced tree path
+%   labeling is equal to $\cl$ or $\cl_\phi = \cl$.
+% \end{lemma}
+% % \begin{proof}
+% %   It is clear that in the end of every recursive call to
+% %   Algorithm~\ref{al:icppl-find-isomorph}, the function $\phi'$ is one
+% %   to one involving all the leaves in the tree passed as input to that
+% %   recursive call. Moreover, by Lemma~\ref{lem:noexit1} and
+% %   Lemma~\ref{lem:xnotempty} it is consistent with the tree path
+% %   labeling $\cl$ passed. The tree pruning is done by only removing
+% %   leaves in each call to the function and is done till the tree
+% %   becomes empty. Thus the returned function $\phi: supp(\cF)
+% %   \rightarrow V(T)$ is a union of mutually exclusive one to one
+% %   functions exhausting all vertices of the tree. In other words, it is
+% %   a bijection from $supp(\cF)$ to $V(T)$ inducing the given path
+% %   labeling $\cl$ and thus a hypergraph isomorphism. \qed
+% % \end{proof}
+
+% \begin{theorem}
+%   \label{th:charac}
+%   A path labeling $(\cF, \cl)$ on tree $T$ is feasible \iff it is an
+%   ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
+%   T)$ as input returns a non-empty function.
+% \end{theorem}
+% % \begin{proof}
+% %   From Lemma~\ref{lem:hyperiso}, we know that if $(\cF, \cl)$ is an
+% %   ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
+% %   T)$ as input returns a non-empty function, $(\cF, \cl)$ is feasible.
+% %   Now consider the case where $(\cF, \cl)$ is feasible. i.e. there
+% %   exists a hypergraph isomorphism $\phi$ such that $\cl_\phi =
+% %   \cl$. Lemma~\ref{lem:noexit1} and Lemma~\ref{lem:xnotempty} show us
+% %   that filter 1 and filter 2 do not exit if input is feasible. Thus
+% %   Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
+% %   function.\qed
+% % \end{proof}
+
+% \subsection*{ICPPL when given tree is a path}
+% \label{sec:icpplicpia}
+% \vspace{\secfirstparatrim}  
+% Consider a special case of ICPPL with the following properties when
+% the tree $T$ is a path.  Hence, all path labels are can be viewed as
+% intervals assigned to the sets in $\cF$.  It is shown, in
+% \cite{nsnrs09}, that the filtering algorithms outlined above need only
+% preserve pairwise intersection cardinalities, and higher level
+% intersection cardinalities are preserved by the Helly Property of
+% intervals.  Consequently, the filter algorithms do not need to ever
+% evaluate the additional check to {\em \bf exit}.
+% %\begin{enumerate}
+% %\item Given tree $T$ is a path. Hence, all path labels are interval labels.
+% %\item Only pairwise intersection cardinality
+% %  preservation is sufficient. i.e. property (iii) in ICPPL is not enforced.
+% %\item The filter algorithms do not have {\em \bf exit} statements.
+% %\end{enumerate}
+% %This is called an Intersection Cardinality Preservation Interval
+% %Assignment (ICPIA) \cite{nsnrs09}. 
+% This structure and its algorithm is
+% used in the next section for finding tree path labeling from a
+% $k$-subdivided star due to this graph's close relationship with
+% intervals.
+
+% \subsection{Case of a Special Tree -- The $k$-subdivided star}
+% \label{sec:ksubdivstar}
+% \vspace{\secfirstparatrim}  
+% In this section we consider the problem of assigning paths from a
+% $k$-subdivided star $T$ to a given set system $\cF$ such that each set
+% $X \in \cF$ is of cardinality at most $k+2$.  Secondly, we present our
+% results only for the case when overlap graph $\bO(\cF)$ is connected.
+% %The
+% %overlap graph is well-known from the work of
+% %\cite{kklv10,nsnrs09,wlh02}.  We use the notation in
+% %\cite{kklv10}. Recall from Section~\ref{sec:prelims} that hyperedges
+% %$S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$ and $S'$
+% %have a non-empty intersection but neither of them is contained in the
+% %other. The overlap graph $\bO(\cF)$ is a graph in which the vertices
+% %correspond to the sets in $\cF$, and the vertices corresponding to the
+% %hyperedges $S$ and $S'$ are adjacent if and only if they overlap.  Note
+% %that the intersection graph of $\cF$, $\bI(\cF)$ is different from
+% %$\bO(\cF)$ and $\bO(\cF) \subseteq \bI(\cF)$.  
+% A connected component of $\bO(\cF)$ is called an overlap component of
+% $\cF$.  An interesting property of the overlap components is that any
+% two distinct overlap components, say $\cO_1$ and $\cO_2$, are such
+% that any two sets $S_1 \in \cO_1$ and $S_2 \in \cO_2$ are disjoint,
+% or, w.l.o.g, all the sets in $\cO_1$ are contained within one set in
+% $\cO_2$.  This containment relation naturally determines a
+% decomposition of the overlap components into rooted containment trees.
+% We consider the case when there is only one rooted containment tree,
+% and we first present our algorithm when $\bO(\cF)$ is connected.  It
+% is easy to see that once the path labeling to the overlap component in
+% the root of the containment tree is achieved, the path labeling to the
+% other overlap components in the rooted containment tree is essentially
+% finding a path labeling when the target tree is a path: each target
+% path is a path that is allocated to sets in the root overlap
+% component.  Therefore, for the rest of this section, $\bO(\cF)$ is a
+% connected graph. Recall that we also consider the special case when
+% all hyperedges are of cardinality at most $k+2$.  By definition, a
+% $k$-subdivided star has a central vertex which we call the {\em root},
+% and each root to leaf path is called a {\em ray}.  First, we observe
+% that by removing the root $r$ from $T$, we get a collection of $p$
+% vertex disjoint paths of length $k+1$, $p$ being the number of leaves
+% in $T$.  We denote the rays by $R_1, \ldots, R_p$, and the number of
+% vertices in $R_i$, $i \in [p]$ is $k+2$.  Let
+% $\seq{v_{i1},\ldots,v_{i(k+2)}=r}$ denote the sequence of vertices in
+% $R_i$, where $v_{i1}$ is the leaf. Note that $r$ is a common vertex to
+% all $R_i$.
+
+% \subsubsection{Description of the Algorithm}
+% \vspace{\secfirstparatrim}  
+% In this section the given hypergraph $\cF$, the $k$-subdivided star
+% and the root of the star are denoted by $\cO$, $T$ and vertex $r$,
+% respectively.  In particular, note that the vertices of $\cO$
+% correspond to the sets in $\cF$, and the edges correspond to the
+% overlap relation.
+
+% %\noindent
+% For each hyperedge $X \in \cO$, we will maintain a 2-tuple of non-negative
+%  numbers $\seq{p_1(X), p_2(X)}$.  The numbers satisfy the property that
+%  $p_1(X) + p_2(X) \leq |X|$, and at the end of path labeling, for each
+%  $X$, $p_1(X) + p_2(X) = |X|$.  This signifies the algorithm tracking
+%  the lengths of subpaths of the path assigned to $X$ from at most two
+%  rays. We also maintain another parameter called the {\em residue} of
+%  $X$ denoted by $s(X)=|X| - p_1(X)$. This signifies the residue path
+%  length that must be assigned to $X$ which must be from another
+%  ray. For instance, if $X$ is labeled a path from only one ray, then
+%  $p_1(X) = |X|$, $p_2(X) = 0$ and $s(X) = 0$.
+
+%  The algorithm proceeds in iterations, and in the $i$-th
+%  iteration, $i > 1$, a single hyperedge $X$ that overlaps with a
+%  hyperedge that has been assigned a path is considered.  At the
+%  beginning of each iteration hyperedges of $\cO$ are classifed into
+%  the following disjoint sets.
+%  \begin{enumerate}
+%  \item [$\cL_1^i$] {\em Labeled without $r$.} Those that have been
+%    labeled with a path which does not contain $r$ in one of the
+%    previous iterations.\\  $\cL_1^i = \set{ X \mid p_1(X) = |X| \text{ and
+%      } p_2(X) = 0 \text{ and } s(X) = 0, X \in \cO}$
+%  \item [$\cL_2^i$] {\em Labeled with $r$.} Those that have been labeled
+%    with two subpaths of $\cl(X)$ containing $r$ from two different rays
+%    in two previous iterations.\\ $\cL_2^i = \set{X \mid 0 < p_1\left(X\right),
+%      p_2\left(X\right) < |X|=p_1(X)+p_2(X) \text{ and } s(X) = 0, X \in \cO}$
+%    \item [$\cT_1^i$] {\em Type 1 / partially labeled.} Those that have
+%      been labeled with one path containing $r$ from a single ray in one
+%      of the previous iterations. Here, $p_1(X)$ denotes the length of
+%      the subpath of $\cl(X)$ that $X$ has been so far labeled
+%      with. \\
+%      $\cT_1^i = \set{ X \mid 0 < p_1(X) < |X| \text{ and } p_2(X) = 0
+%        \text{ and } s(X) = |X|-p_1(X), X \in \cO}$
+%    \item [$\cT_2^i$] {\em Type 2 / not labeled.} Those that have not been
+%      labeled with a path in any previous iteration.\\
+%      $\cT_2^i = \set{ X \mid p_1(X) = p_2(X) = 0 \text{ and } s(X) = |X|,
+%        X \in \cO}$
+%  \end{enumerate}
+% % \vspace{-2mm}
+% % \begin{align*}
+% %   \cO &= \cL_1^i \cup \cL_2^i \cup \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]\\
+% %   \cO_i &= \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]
+% % \end{align*}
+
+% %\noindent
+% The set $\cO_i$ refers to the set of hyperedges $\cT_1^i \cup \cT_2^i$
+% in the $i$th iteration.  Note that $\cO_1 = \cO$.  In the $i$th
+% iteration, hyperedges from $\cO_i$ are assigned paths from $T$ using
+% the following rules. Also the end of the iteration, $\cL_1^{i+1},
+% \cL_2^{i+1}, \cT_1^{i+1}, \cT_2^{i+1}$ are set to $\cL_1^{i},
+% \cL_2^{i}, \cT_1^{i}, \cT_2^{i}$ respectively, along with some
+% case-specific changes mentioned in the rules below.
+
+% %\noindent
+% \begin{enumerate}[I.]
+% \item {\bf Iteration 1:} Let $S=\{X_1,\ldots,X_s\}$ denote the
+%   super-marginal hyperedges from $\cO_1$.  If $|S|=s \neq p$, then
+%   exit reporting failure.  Else, assign to each $X_j \in S$, the path
+%   from $R_j$ such that the path contains the leaf in $R_j$.  This path
+%   is refered to as $\cl(X_j)$.  Set $p_1(X_j)=|X|, p_2(X_j)=s(X_j)=0$.
+%   Hyperedges in $S$ are not added to $\cO_2$ but are added to
+%   $\cL_1^2$ and all other hyperedges are added to
+%   $\cO_2$.
+% \item {\bf Iteration $i$:} Let $X$ be a hyperedge from $\cO_i$ such
+%   that there exists $Y \in \cL_1^i \cup \cL_2^i$ and $X \overlap
+%   Y$. Further let $Z \in \cL_1^i \cup L_2^i$ such that $Z \overlap Y$.
+%   If $X \in \cT_2^i$, and if there are multiple $Y$ candidates then
+%   any $Y$ is selected.  On the other hand, if $X \in \cT_1^i$, then
+%   $X$ has a partial path assignment, $\cl'(X)$ from a previous
+%   iteration, say from ray $R_j$. Then, $Y$ is
+%   chosen such that $X \cap Y$ has a non-empty intersection with a ray
+%   different from $R_j$.  The key things that are done in assigning a
+%   path to $X$ are as follows. The {\em end} of path $\cl(Y)$ where
+%   $\cl(X)$ would overlap is found, and then based on this the
+%   existence of a feasible assignment is decided.  It is important to
+%   note that since $X \overlap Y$, $\cl(X) \overlap \cl(Y)$ in any
+%   feasible assignment.  Therefore, the notion of the {\em end} at
+%   which $\cl(X)$ and $\cl(Y)$ overlap is unambiguous, since for any
+%   path, there are two end points.
+%   \begin{enumerate}
+%   \item \label{iendpoint} {\em End point of $\cl(Y)$ where $\cl(X)$ overlaps
+%       depends on $X \cap Z$:} If $X \cap Z
+%     \neq \emptyset$, then $\cl(X)$ has an overlap of $|X \cap Y|$ at that
+%     end of $\cl(Y)$ at which $\cl(Y)$ and $\cl(Z)$ overlap.  If $X
+%     \cap Z = \emptyset$, then $\cl(X)$ has an overlap of $|X \cap Y|$ at
+%     that end of $\cl(Y)$ where $\cl(Y)$ and $\cl(Z)$ do not intersect.
+%   \item {\em Any path of length $s(X)$ at the appropriate end contains
+%       $r$:} If $X \in \cT_1^i$ then after finding the appropriate end
+%     as in step~\ref{iendpoint} this the unique path of length $s(X)$
+%     should end at $r$.  If not, we exit reporting failure.  Else,
+%     $\cl(X)$ is computed as union of $\cl'(X)$ and this path. If any
+%     three-way intersection cardinality is violated with this new
+%     assignment, then exit, reporting failure.  Otherwise, $X$ is added
+%     to $\cL_2^{i+1}$.  On the other hand, if $X \in \cT_2^i$, then
+%     after step~\ref{iendpoint}, $\cl(X)$ or $\cl'(X)$ is unique up to
+%     the root and including it. Clearly, the vertices $\cl(X)$ or
+%     $\cl'(X)$ contains depends on $|X|$ and $|X \cap Y|$.  If any
+%     three way intersection cardinality is violated due to this
+%     assignment, exit, reporting failure.  Otherwise,
+%     $p_1(X)$ is updated as the length of the assigned path, and $s(X)
+%     = |X|-p_1(X)$.  If $s(X) > 0$, then $X$ is added to $\cT_1^{i+1}$.
+%     If $s(X)=0$, then $X$ is added to $\cL_1^{i+1}$.
+%   \item {\em The unique path of length $s(X)$ overlapping at the
+%       appropriate end of $Y$ does not contain $r$:} In this case,
+%     $\cl(X)$ is updated to include this path.  If any three way
+%     intersection cardinality is violated, exit, reporting failure.
+%     Otherwise, update $p_1(X)$ and $p_2(X)$ are appropriate, $X$ is
+%     added to $\cL_1^{i+1}$ or $\cL_2^{i+1}$, as appropriate.
+%   \end{enumerate}
+% \end{enumerate}
+
+% %\noindent 
+% \tnote[TH]{UNCOMMENT}
+% % \subsection*{Proof of Correctness and Analysis of Running Time} 
+% % It
+% % is clear that the algorithm runs in polynomial time, as at each step,
+% % at most three-way intersection cardinalities need to be checked.
+% % Further, finding super-marginal hyperedges can also be done in
+% % polynomial time, as it involves considering the overlap regions and
+% % checking if the inclusion partial order contains a single minimal
+% % element.  In particular, once the super-marginal edges are identified,
+% % each iteration involes finding the next hyperedge to consider, and
+% % testing for a path to that hyperedge.  To identify the next hyperedge
+% % to consider, we consider the breadth first layering of the hyperedges
+% % with the zeroeth layer consisting of the super-marginal hyperedges.
+% % Since $\cO$ is connected, it follows that all hyperedges of $\cO$ will
+% % be considered by the algorithm.  Once a hyperedge is considered, the
+% % path to be assigned to it can also be computed in constant time.  In
+% % particular, in the algorithm the path to be assigned to $X$ depends on
+% % $\cl(Y), \cl(Z)$, $s(X)$ and the presence or absence of $r$ in the
+% % candidate partial path $\cl'(X)$.  Therefore, once the super-marginal
+% % edges are identified, the running time of the algorithm is linear in
+% % the size of the input.  By the technique used for constructing prime
+% % matrices \cite{wlh02}, the super-marginal edges can be found in linear
+% % time in the input size.  Therefore, the algorithm can be implemented
+% % to run in linear time in the input size.
+
+% % %\noindent
+% % The proof of correctness uses the following main properties:
+% % \begin{enumerate}
+% % \item The $k$-subdivided star has a very symmetric structure.  This
+% %   symmetry is quantified based on the following observation -- either
+% %   there are no feasible path labelings of $\cO$ using paths from $T$,
+% %   or there are exactly $p!$ feasible path labelings.  In other words,
+% %   there is either no feasible assignment, or effectively a unique
+% %   assignment modulo symmetry.
+% % \item The $p$ super-marginal hyperedges, if they exist, will each be
+% %   assigned a path from distinct rays, and each such path contains the
+% %   leaf.
+% % \item For a candidate hyperedge $X$, the partial path assignment
+% %   $\cl'(X)$ is decided by its overlap with $\cl(Y)$ and cardinality of
+% %   intersection with $\cl(Z)$.
+% % \end{enumerate}
+% % These properties are formalized as follows:
+% % \begin{lemma}
+% %   \label{lem:sup-mar}
+% %   If $X \in \cF$ is super-marginal and $\cl$ is a feasible tree path
+% %   labeling to tree $T$, then $\cl(X)$ will contain a leaf in $T$.
+% % \end{lemma}
+% % \begin{proof}
+% %   Suppose $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
+% %   path labeling from $T$.  Assume $\cl(X)$ does not have a leaf.  Let
+% %   $R_i$ be one of the rays (or the only ray) $\cl(X)$ is part of.
+% %   Since $X$ is in a connected overlap component, there exists $Y_1 \in
+% %   \cF$ and $X \nsubseteq Y_1$ such that $Y_1 \overlap X$ and $Y_1$ has
+% %   at least one vertex closer to the leaf in $R_i$ than any vertex in
+% %   $X$. Similarly with the same argument there exists $Y_2 \in \cF$
+% %   with same subset and overlap relation with $X$ except it has has at
+% %   least one vertex farther away from the leaf in $R_i$ than any vertex
+% %   in $X$. Clearly $Y_1 \cap X$ and $Y_2 \cap X$ cannot be part of same
+% %   inclusion chain which contradicts that assumption $X$ is
+% %   super-marginal. Thus the claim is proved.\qed
+% % \end{proof}
+% % \begin{lemma}
+% %   If $\cO$ does not have any super-marginal edges, then in any
+% %   feasible path labeling $\cl$ of $\cO$ with paths from $T$ is such
+% %   that, for any hyperedge $X$ for which $\cl(X)$ contains a leaf, $|X|
+% %   \geq k+3$.
+% % \end{lemma}
+% % \begin{proof}
+% %   The proof of this lemma is by contradiction.  Let $X$ be a
+% %   hyperedges such that $|X| \leq k+2$ and that $\cl(X)$ has a leaf.
+% %   This implies that the overlap regions with $X$, which are captured by
+% %   the overlap regions with $\cl(X)$, will form a single inclusion
+% %   chain. This shows that $X$ is a marginal hyperedge which
+% %   contradicts the assumption that $\cO$ does not have super-marginal
+% %   hyperedges. \qed
+% % \end{proof}
+% % This lemma is used to prove the next lemma for the case when for all
+% % $X \in \cO$, $|X| \leq k+2$.  The proof is left out as it just uses
+% % the previous lemma and the fact that the hyperedges in $X$ have at
+% % most $k+2$ elements.
+% % \begin{lemma}
+% %   If there is a feasible path labeling for $\cO$ in $T$, then there
+% %   are exactly $p$ super-marginal hyperedges.
+% % \end{lemma}
+% % These lemmas now are used to prove the following theorem.
+% % \begin{theorem}
+% %   Given $\cO$ and a $k$-subdivided star $T$, the above algorithm
+% %   decides correctly if there is a feasible path labeling $\cl$.
+% % \end{theorem}
+% % \begin{proof} {\em Outline.}
+% % %   If the algorithm outputs a path labeling $\cl$, then it is clear
+% % %   that it is an ICPPL. The reason is that the algorithm checks that
+% % %   three-way intersection cardinalities are preserved in each iteration
+% % %   which ensures \icpplpr~\ref{pr:iii}. Moreover, it is clear that
+% % %   $\cl(X)$ for any $X \in \cO$ is computed by maintaining
+% % %   \icpplpr~\ref{pr:i} and \icpplpr~\ref{pr:ii}. For such a labeling
+% % %   $\cl$, the proof that it is feasible is by induction on $k$. What
+% % %   needs to be shown is that Algorithm~\ref{al:icppl-find-isomorph}
+% % %   successfully runs on input $(\cO, \cl)$. In base case $k=0$, $T$ is
+% % %   a star. The claim is clear by observing that after Filter 1 and one
+% % %   iteration of Filter 2, all the leaves have found their pre-images
+% % %   from support $\cO$.  Therefore, in the induction step, it is clear
+% % %   that after Filter 1 and one iteration of Filter 2, the leaves are
+% % %   assigned pre-images.  Removing the leaves from $T$ and the
+% % %   pre-images from support of $\cO$, results in an ICPPL to a
+% % %   $(k-1)$-subdivided star.  Now we apply the induction hypothesis, and
+% % %   we get a isomorphism between the hypergraph $\cO$ and $\cO^\cl$.
+% %   If the algorithm outputs a path labeling $\cl$, then it is clear
+% %   that it is an ICPPL. The reason is that the algorithm checks that
+% %   three-way intersection cardinalities are preserved in each iteration
+% %   which ensures ICPPL \icpplpr~\ref{pr:iii}. Moreover, it is clear that
+% %   $\cl(X)$ for any $X \in \cO$ is computed by maintaining
+% %   ICPPL \icpplpr~\ref{pr:i} and ICPPL \icpplpr~\ref{pr:ii}. For such a labeling
+% %   $\cl$, the proof that it is feasible is by induction on $k$. What
+% %   needs to be shown is that Algorithm~\ref{al:icppl-find-isomorph}
+% %   successfully runs on input $(\cO, \cl)$. In base case $k=0$, $T$ is
+% %   a star. Also every set is at most size 2 ($k+2$) size and thus
+% %   overlaps are at most 1. If two paths share a leaf in {\tt filter\_1}
+% %   one must be of length 2 and the other of length 1. Thus the exit
+% %   condition is not met. Now it is trivial that the exit condition in
+% %   {\tt filter\_2} is also not met. Thus claim proven for base case.
+% %   Now assume the claim to be true when target tree is a
+% %   $(k-1)$-subdivided star. Consider the case of a $k$-subdivided
+% %   star. It is clear that after {\tt filter\_1} and one iteration of
+% %   {\tt filter\_2}, the leaves are assigned pre-images.  Removing the
+% %   leaves from $T$ and the pre-images from support of $\cO$, results in
+% %   an ICPPL to a $(k-1)$-subdivided star.  Now we apply the induction
+% %   hypothesis, and we get a isomorphism between the hypergraph $\cO$
+% %   and $\cO^\cl$.
+
+% %  \noindent
+%   In the reverse direction if there is a feasible path labeling $\cl$,
+%   then we know that $\cl$ is unique up to isomorphism.  Therefore,
+%   again by induction on $k$ it follows that the algorithm finds $\cl$.
+%   \qed
+% \end{proof}
+
+\section{Conclusion}
+\label{sec:conclusion}
+\vspace{\secfirstparatrim}
+
+We give a characterization for feasible tree path labeling of path
+hypergraphs. The proof for this is constructive and computes a
+feasibility bijection mapping vertices of the hypergraph to vertices
+of the given target tree. This thesis also discovered an exponential algorithm
+that computes a feasible TPL to a given hypergraph if it is a path hypergraph.
+
+Our results have close connections to recognition of path graphs and
+to path graph isomorphism.  Graphs which can be represented as the
+intersection graph of paths in a tree are called {\em path
+  graphs}\cite{mcg04}. Thus, a hypergraph $\cF$ can be interpreted as
+paths in a tree, if and only if the intersection graph of $\cF$ is a
+{\em path graph}. Path graphs are a subclass of chordal graphs since
+chordal graphs are characterized as the intersection graphs of
+subtrees of a tree\cite{mcg04}.  Path graphs are well studied in the
+literature \cite{plr70}--\cite{mcg04}\tnote[TH]{Check if cite:bp93 is
+  indeed for path graphs}.  Path graph recognition can be done in
+polynomial time\cite{gav78,aas93}.  Clearly, this is a necessary
+condition in terms of the intersection graph of the input hypergraph
+$\cF$ in \FTPL. However, one can easily obtain a counterexample to
+show the insufficiency of this condition\tnote[GTC]{Give a figure
+  showing insuff.}.  Path graph isomorphism is known be
+isomorphism-complete\cite{kklv10}. Therefore, \annote[GTC]{it is unlikely
+  that we can solve the problem of finding feasible path labeling
+  $\cl$ for a given $\cF$ and }{WHY?} tree $T$.  It is
+  definitely interesting to classify the kinds of trees and
+  hypergraphs for which feasible path labelings can be\tnote[TH]{ELABORATE.}
+found efficiently.  \annote[TH]{These results would form a natural
+  generalization of COP testing and interval graph isomorphism,
+  culminating in Graph Isomorphism }{REWORD} itself.
+ To this effect we consider TPL on {\kstar} and give a polynomial time
+ solution for the same.
+
+Whether TPL on general trees is solvable in $P$ remains open. So do
+optimization opportunities in TPL (possible extensions to optimization for COP in the
+Section~\ref{sec:optcop}.\tnote[TH]{link to section}).
+
+\tnote[TH]{ ||Complexity challenges:|| Canonization is an important tool in graph isomorphism. Invention of a
+% deterministic method of canonization for any class of graphs naturally
+% results in an algorithm for isomorphism. All that is required is to
+% check if the canons of two graphs are the same. Thus complexity of
+% graph isomorphism can be studied by studying canonization
+% methods. While general graph ismorphism remains elusive in terms of
+% hardness, canonization has been studied for smaller classes of graphs
+% thus giving complexity/hardness results for them.
+
+% In 1992, \cite{sl92} made an important discovery that tree isomorphism
+% is in logspace. It was done by inventing a method of canonization of
+% trees using a logspace depth first traversal
+% algorithm. \tnote{confirm the statement of claim}.
+
+% \cite{kklv10} proved that interval graph canonization is also in
+% logspace thus drawing logspace conslusions about COT. 
+% % This result is
+% % built on the fact that 
+% Interval graphs are FP+C (fixed point with
+% counting) definable \tnote{elucidate} and \cite{l10} showed that
+% this implies that it captures PTIME. \tnote{elucidate} This result
+% along with that of undirected graph reachability being in logspace
+% \cite{rei08}, \cite{kklv10} proved their logspace result.
+}%
+\tnote[TH]{ (1) the interesting survey in conclusion section of kklv10
+  (2) [[[COP helping some problems with hardness]]] having COP in the
+  input structure makes some problems less hard than in a general
+  input. -- min set cover problem is known to be NP hard. so is min
+  hitting
+  set. these two problems are equivalent cite-adp80\\
+  -- -- complexity results: log factor polynomial time approximable cite-f98, W[2]-complete with solution size as parameter cite-df99\\
+  -- a couple of generalizations:\\
+  -- -- min degree hypergraph (MDH) problem\\
+  -- -- red blue covering (RBSC) problem\\
+  ----- cite-d08phd Chapter 5.\\
+  ----- what are cite-d08phd results\\
+  ----- Rectangle stabbing \\
+  ----- see cite-d08phd Chapter 6?  }
+
+\tnote{ Some things to remember - have this somewhere.
 
@@
 \chapter{Consecutive Ones property}
-\tnote[a]{survey of cop includes mainly 
+\tnote{survey of cop includes mainly 
 \begin{enumerate}
@@
 
-\tnote[a]{\cite{at72}, \cite{mcc04}, prime submatrix \cite{wlh02}?}
+\tnote{\cite{at72}, \cite{mcc04}, prime submatrix \cite{wlh02}?}
 \section{Recognition of COP}
-\subsection{\tnote[a]{Polynomial time algorithms}}
-\tnote[a]{ (pq tree, pqr, icpia)}
-\subsection{\tnote[a]{Tucker's submatrices, Dom's algo to find them}}
-\subsection{\tnote[a]{McConnell's certificate}}
+\subsection{\tnote{Polynomial time algorithms}}
+\tnote{ (pq tree, pqr, icpia)}
+\subsection{\tnote{Tucker's submatrices, Dom's algo to find them}}
+\subsection{\tnote{McConnell's certificate}}
 
 \section{Alteration to matrices to get COP}
-\tnote[a]{algorithms to make a matrix into one with COP
+\tnote{algorithms to make a matrix into one with COP
   (min/max-cos-r/c) \cite{d08phd} }
@@
 \label{sec:briefanal}
-\tnote[a]{Do I need this?}
+\tnote{Do I need this?}
 \section{Complexity of certain COP variations}
-\tnote[a]{maybe combine with Section~\ref{sec:briefanal}\\
+\tnote{maybe combine with Section~\ref{sec:briefanal}\\
 complexity - poly, logspace}
@@
 \chapter{Other problems related to COP }
-\tnote[a]{(see literature)}
+\tnote{(see literature)}
 
 \chapter{Research - Tree path labeling of path hypergraphs}
-\tnote[a]{Extension to trees. ( stuff in latin paper,)}
-
-\tnote[a]{UPDATE TO FINAL VERSION. BELOW IS OLDER VERSION WITH
+\tnote{Extension to trees. ( stuff in latin paper,)}
+
+\tnote{UPDATE TO FINAL VERSION. BELOW IS OLDER VERSION WITH
   INCOMPLETE SEC 4.}
@@
 several algorithms. PQ trees \cite{bl76}, variations of PQ trees
-\cite{mm09,wlh01,wlh02,mcc04}, ICPIA \cite{nsnrs09} are the main ones.
+\cite{mm96,wlh01,wlh02,mcc04}, ICPIA \cite{nsnrs09} are the main ones.
 
@@
 %\noindent 
-\tnote[a]{the following text could be in Further Study/Conclusion}It
+\tnote{the following text could be in Further Study/Conclusion}It
 is an interesting fact that for a matrix with the COP, the
@@
 \in A\}$ \cite{kklv10}. This is called {\em hypergraph
-  isomorphism}. \tnote[a]{also extend $\phi$ to hyperedges -- see if
+  isomorphism}. \tnote{also extend $\phi$ to hyperedges -- see if
   required }
@@
 % preserved, i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
-% \annote[a]{has no premature exit condition hence any input will go
+% \annote{has no premature exit condition hence any input will go
 %   through it}{Prove that the filtered sets has ICPPL iff input PL
@@
 % a path labeling that defines a family of bijections from $supp(\cF)$
-% to $V(T)$\remove[a]{ i.e. $supp(\cF^{\cl})$}.
+% to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.
 First, we present Algorithm~\ref{perms} or {\tt filter\_1}, and prove
@@
   add it to $\cO_{i+1}$ after removing it from $\cO_i$.
-  % \remove[a]{Next, iteratively, consider each type-1 edge $X \in
+  % \remove{Next, iteratively, consider each type-1 edge $X \in
   %   \cO_i$, and if it intersects a previously assigned $Y$ such that
@@
   %   containing $v_{i(k+2)}=r$, set $p_2(X)=s(X), s(X)=0$, and remove
-  %   $X$ from $\cO_i$.}\tnote[a]{covered in step 2}
+  %   $X$ from $\cO_i$.}\tnote{covered in step 2}
 
@@
 \noindent
-\tnote[a]{need to formalize the below properties} In the following
+\tnote{need to formalize the below properties} In the following
 lemmas we identify a set of necessary conditions for $\cF$ to have an
@@
 \chapter{ Research - [title TBD]} 
-\tnote[a]{the matrix theory submitted in walcom, any other work - need to see notes.}
+\tnote{the matrix theory submitted in walcom, any other work - need to see notes.}
 
@@
 
+\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Bibliography.
-
 \begin{singlespace}
+%\begin{thebibliography}{10}
+\addcontentsline{toc}{section}{Bibliography}
+
+{\footnotesize
+  \bibliographystyle{alpha} 
   \bibliography{../lib/cop-variants__thesis}
+}
+%\end{thebibliography}
 \end{singlespace}
+\newpage
 

### 2011-12-21 src/thesis.tex
Basic organization of doc. First real version of TOC. To be sent to n for comments.

@@
 \documentclass[MS]             % MS thesis dissertation.
-              {iitmdiss}
+              {iitmdiss_as}    % Hacked by hyperbolicme.
+%              {iitmdiss}      % If the powers to be are finicky.
 
@@
 \def \mythesissubmissiondate {30 December 2011}
+\def \mythesissubmissionmonth {\MakeUppercase {December 2011}}
 
@@
 % Title page
-\title{\mythesistitle} 
+\title{\mythesistitleNOTSYNPOSIS} 
 \author{\myname} 
-\date{\MakeUppercase
+\date{
   {\mythesissubmissionmonth}} 
-\department{\MakeUppercase{\mydept}}
+\department{\MakeUppercase{\mythesisdept}}
+%\department{\mythesisdept}
 
@@
 \noindent 
-This is to certify that the thesis titled {\bf \mythesistitle},
-submitted by {\bf \myname}, to the Indian Institute of Technology,
-Madras, for the award of the degree of {\bf \mydegree}, is
+This is to certify that the thesis titled {\bf \mythesistitleNOTSYNPOSIS},
+submitted by {\bf \myname}, to the {\bf Indian Institute of Technology
+Madras}, for the award of the degree of {\bf \mydegree}, is 
 a bona fide record of the research work done by her under our
@@
 \begin{abstract}
-  \noindent {\bf Keywords}: \hspace*{0.5em} \parbox[t]{4.4in}{consecutive
+  \noindent {\em {\bf Keywords}: \hspace*{0.5em} \parbox[t]{4.4in}{consecutive
     ones property, algorithmic graph theory, hypergraph isomorphism,
-    interval labeling} \vspace*{24pt}
+    interval labeling}} \vspace*{24pt}
 
@@
   \noindent 
-  \annote{In this paper, we explore a natural generalization of
-  results on binary matrices with the |\tem consecutive ones property|. 
-  We consider the following constraint satisfaction problem. Given (i)
-  a set system $\F \subseteq$ $(2^{U} \setminus \emptyset)$ of a
-  finite set $U$ of cardinality $n$, (ii) a tree $T$ of size $n$ and
-  (iii) a bijection called |\tem tree path labeling|, $\cl$ mapping the
-  sets in $\cF$ to paths in $T$, does there exist at least one
-  bijection $\phi:U \rightarrow V(T)$ such that for each $S \in \cF$,
-  $\{\phi(x) \mid x \in S\} = \cl(S)$?  A tree path labeling of a set
-  system is called |\tem feasible| if there exists such a bijection
-  $\phi$.  We present an algorithmic characterization of feasible tree
-  path labeling. COP is a special instance of tree path labeling
-  problem when $T$ is a path.  We conclude with a polynomial time
-  algorithm to find a feasible tree path labeling of a given set
-  system when $T$ is a |\tem $k$-subdivided star|, set system has a
-  single containment tree of overlap components and set size is
-  limited to at most $k+2$. }{this is the abstract of the
-  paper. expand to fit thesis.}
+  Consecutive-ones property is a non-trivial property of binary
+  matrices that has been studied widely in the literature for over
+  past 50 years. Detection of COP in a matrix is possible efficiently
+  and there are several algorithms that achieve the same. This thesis
+  documents the work done on an extension of COP extended from the
+  equivalent interval assignment problem in \cite{nsnrs09}. These new
+  results rigorously prove a natural extension (to trees) of their
+  characterization as well as makes connections to graph isomorphism,
+  namely path graph isomorphism.
+
+  \tnote{EXPAND. 
+  Abstract must be a brief about what results we have and how it fits
+  in the body of research.\\
+  -- Area: Broad to Specialized. i.e. Combinatorial
+    algorithms -> Matrix reorganization -> general data reorganization
+    (interval assignment) -> path assignment\\
+  -- Class of problems: say, data reorganization.\\
+  -- Nature of results: Is a generalization. We have a
+    Polynomial algorithm for a subset of the generalization.\\
+% maybe add - we explore a natural
+%     generalization of results on binary matrices with the |\tem
+%     consecutive-ones property|.  We consider the following constraint
+%     satisfaction problem. Given (i) a set system $\F \subseteq$
+%     $(2^{U} \setminus \emptyset)$ of a finite set $U$ of cardinality
+%     $n$, (ii) a tree $T$ of size $n$ and (iii) a bijection called
+%     |\tem tree path labeling|, $\cl$ mapping the sets in $\cF$ to
+%     paths in $T$, does there exist at least one bijection $\phi:U
+%     \rightarrow V(T)$ such that for each $S \in \cF$, $\{\phi(x) \mid
+%     x \in S\} = \cl(S)$?  A tree path labeling of a set system is
+%     called |\tem feasible| if there exists such a bijection $\phi$.
+%     We present an algorithmic characterization of feasible tree path
+%     labeling. COP is a special instance of tree path labeling problem
+%     when $T$ is a path.  We conclude with a polynomial time algorithm
+%     to find a feasible tree path labeling of a given set system when
+%     $T$ is a |\tem $k$-subdivided star|, set system has a single
+%     containment tree of overlap components and set size is limited to
+%     at most $k+2$.
+  }
 \end{abstract}
@@
 xxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
-\textbf{COP}   \> Consecutive Ones Property \\
-\textbf{COT}   \> Consecutive Ones Testing \\
+\textbf{COP}   \> Consecutive-Ones Property \\
+\textbf{COT}   \> Consecutive-Ones Testing \\
 \textbf{ICPIA}   \> Intersection Cardinality Preservation Interval Assignment \\
@@
 
-\section{Introduction}
-\label{sec:intro}
-\vspace{\secfirstparatrim}
-
-Consecutive ones property is a non-trivial property of binary matrices
+% \section{Introduction}
+% \label{sec:intro}
+% \vspace{\secfirstparatrim}
+
+\tnote{ The main themes of the thesis that must not be left out:\\
+  connection of COP to ||set systems|| and thus to general
+  data bases. cite:nsnrs09\\
+  ||graph isomorphism|| by canonization cite:kklv10\\
+  ||certificate for a problem|| cite:mcc04\\
+  Also see purple book notes.  }
+
+Consecutive-ones property is a non-trivial property of binary matrices
 that has been studied widely in the literature for over past 50
@@
 
+\tnote{Have a para on Organization: Outline of document}
+Section~\ref{sec:background} gives a brief survey of COP and
+optimization problems related to it followed by motivation for the
+thesis in Section~\ref{sec:motive}.  Section~\ref{sec:results}
+presents a summary of our results on the extension of COP namely, the
+tree path labeling problem. \tnote[TH]{update for thesis}
+
 % \tnote{testing 1 2 3...}
@@
 
-\subsection{Illustration of the problem}
+\section{Illustration of the problem}
 \label{sec:problem}
-\vspace{\secfirstparatrim}
+%\vspace{\secfirstparatrim}
 
@@
 
-%%%%%%%%%%%% THESIS UNCOMMENT %%%%%%%%%%%%
 % \begin{figure}[t]%[htbp]
@@
 %     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
-%     group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
+%     group, orange for \textcolor{yellow}{$\xBBB$} group, green for
 %     \textcolor{green}{$\xTTT$} group}
@@
 % \end{figure}
-%%%%%%%%%%%% END THESIS UNCOMMENT %%%%%%%%%%%%
+
+\begin{figure}[htbp] %[t]%
+ \centering
+  (a) \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
+  (b) \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
+ \label{fig:streetmap}
+ \caption{\figtabsize (a) {\residenceblock} street map (b)
+   {\residenceblock} street map with study group routes
+   allocated. Routes are color coded as follows: red for
+   \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
+   group, orange for \textcolor{yellow}{$\xBBB$} group, green for
+   \textcolor{green}{$\xTTT$} group}
+% \end{figure}
 % \begin{figure}[htbp] %[t]%
 %   \centering
-%   (a) \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
-%   (b) \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
-%  \label{fig:streetmap}
-%   \caption{\figtabsize (a) {\residenceblock} street map (b) {\residenceblock} street map with study group routes
-%     allocated. Routes are color coded as follows: red for
-%     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
-%     group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
-%     \textcolor{green}{$\xTTT$} group}
-% \end{figure}
-\begin{figure}[htbp] %[t]%
-  \centering
-  \includegraphics[scale=0.3]{../img/3_infinite_loop.png}
+  \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
   \caption{\figtabsize Individual allocation of apartments to students
@@
 The interval assignment problem to a set system is equivalent to the
-consecutive ones property (COP) problem in binary matrices\cite{wlh02,
+consecutive-ones property (COP) problem in binary matrices\cite{wlh02,
   nsnrs09}.  The COP problem is to rearrange rows (columns) of a
@@
 
-Section~\ref{sec:background} gives a brief survey of COP and
-optimization problems related to it followed by motivation for the
-thesis in Section~\ref{sec:motive}.  Section~\ref{sec:results}
-presents a summary of our results on the extension of COP namely, the
-tree path labeling problem. \tnote[TH]{update for thesis}
-
-\section{Consecutive Ones Property Testing - a Survey} %{Background}
+\section{Basic preliminaries - general definitions and nomenclature}
+\tnote{(definitions theorems etc) needed if any
+    - graph theory}
+
+\section{Consecutive-Ones Property Testing - a Brief Survey} %{Background}
 \label{sec:background}
-\vspace{\secfirstparatrim}
-In this section, a brief survey of the consecutive ones problem and its
+%\vspace{\secfirstparatrim}
+In this section, a brief survey of the consecutive-ones problem and its
 optimization problems is presented.
@@
   document, isomorphism of certain % classes of graphs, namely chordal
-%   graphs, have a close relationship with consecutive ones property and
+%   graphs, have a close relationship with consecutive-ones property and
 %   generalizations of it.  This is perhaps because of how closely COP
@@
 \label{sec:copmatrices}
-\vspace{\secfirstparatrim} 
+%\vspace{\secfirstparatrim} 
 As seen earlier, the interval assignment problem (illustrated as the
@@
 which is claimed to be much easier to implement. \cite{nsnrs09}
-describes a characterization of consecutive ones property solely based
+describes a characterization of consecutive-ones property solely based
 on the cardinality properties of the set representations of the
@@
 \label{sec:optcop}
-\vspace{\secfirstparatrim}  
+%\vspace{\secfirstparatrim}  
 So far we have been concerned about matrices that have the consecutive
@@
 \label{sec:motive}
-\vspace{\secfirstparatrim}  
+%\vspace{\secfirstparatrim}  
 As seen in Section~\ref{sec:copmatrices}, \cite{nsnrs09} describes a
-characterization of consecutive ones property based on the cardinality
+characterization of consecutive-ones property based on the cardinality
 properties of the set representations of the columns. This result is
@@
 
-\def \probdefwidth {0.85\linewidth}
-\def \FTPL {{\sc Feasible Tree Path Labeling}}
-\def \CFTPL {\sc Compute Feasible Path Labeling} 
+% \def \probdefwidth {0.85\linewidth}
+% \def \FTPL {{\sc Feasible Tree Path Labeling}}
+% \def \CFTPL {\sc Compute Feasible Path Labeling} 
 \section{Summary of results}
 \label{sec:results}
-\vspace{\secfirstparatrim}
+%\vspace{\secfirstparatrim}
 % We address the computation of TPL\footnote{Tree Path Labeling} for a
@@
 
-\def \kstar {$k$-subdivided star}
-\def \CFTPLKTREE {\sc Compute $k$-subdivided Star Path Labeling}
+% \def \kstar {$k$-subdivided star}
+% \def \CFTPLKTREE {\sc Compute $k$-subdivided Star Path Labeling}
 
@@
 
-\section{Conclusion}
-\label{sec:conclusion}
-\vspace{\secfirstparatrim}
-
-We give a characterization for feasible tree path labeling of path
-hypergraphs. The proof for this is constructive and computes a
-feasibility bijection mapping vertices of the hypergraph to vertices
-of the given target tree. This thesis also discovered an exponential algorithm
-that computes a feasible TPL to a given hypergraph if it is a path hypergraph.
-
-Our results have close connections to recognition of path graphs and
-to path graph isomorphism.  Graphs which can be represented as the
-intersection graph of paths in a tree are called {\em path
-  graphs}\cite{mcg04}. Thus, a hypergraph $\cF$ can be interpreted as
-paths in a tree, if and only if the intersection graph of $\cF$ is a
-{\em path graph}. Path graphs are a subclass of chordal graphs since
-chordal graphs are characterized as the intersection graphs of
-subtrees of a tree\cite{mcg04}.  Path graphs are well studied in the
-literature \cite{plr70}--\cite{mcg04}\tnote[TH]{Check if cite:bp93 is
-  indeed for path graphs}.  Path graph recognition can be done in
-polynomial time\cite{gav78,aas93}.  Clearly, this is a necessary
-condition in terms of the intersection graph of the input hypergraph
-$\cF$ in \FTPL. However, one can easily obtain a counterexample to
-show the insufficiency of this condition\tnote[GTC]{Give a figure
-  showing insuff.}.  Path graph isomorphism is known be
-isomorphism-complete\cite{kklv10}. Therefore, \annote[GTC]{it is unlikely
-  that we can solve the problem of finding feasible path labeling
-  $\cl$ for a given $\cF$ and }{WHY?} tree $T$.  It is
-  definitely interesting to classify the kinds of trees and
-  hypergraphs for which feasible path labelings can be\tnote[TH]{ELABORATE.}
-found efficiently.  \annote[TH]{These results would form a natural
-  generalization of COP testing and interval graph isomorphism,
-  culminating in Graph Isomorphism }{REWORD} itself.
- To this effect we consider TPL on {\kstar} and give a polynomial time
- solution for the same.
-
-Whether TPL on general trees is solvable in $P$ remains open. So do
-optimization opportunities in TPL (possible extensions to optimization for COP in the
-Section~\ref{sec:optcop}.\tnote[TH]{link to section}).
-
-\tnote[TH]{ ||Complexity challenges:|| Canonization is an important tool in graph isomorphism. Invention of a
-% deterministic method of canonization for any class of graphs naturally
-% results in an algorithm for isomorphism. All that is required is to
-% check if the canons of two graphs are the same. Thus complexity of
-% graph isomorphism can be studied by studying canonization
-% methods. While general graph ismorphism remains elusive in terms of
-% hardness, canonization has been studied for smaller classes of graphs
-% thus giving complexity/hardness results for them.
-
-% In 1992, \cite{sl92} made an important discovery that tree isomorphism
-% is in logspace. It was done by inventing a method of canonization of
-% trees using a logspace depth first traversal
-% algorithm. \tnote{confirm the statement of claim}.
-
-% \cite{kklv10} proved that interval graph canonization is also in
-% logspace thus drawing logspace conslusions about COT. 
-% % This result is
-% % built on the fact that 
-% Interval graphs are FP+C (fixed point with
-% counting) definable \tnote{elucidate} and \cite{l10} showed that
-% this implies that it captures PTIME. \tnote{elucidate} This result
-% along with that of undirected graph reachability being in logspace
-% \cite{rei08}, \cite{kklv10} proved their logspace result.
-}%
-\tnote[TH]{ (1) the interesting survey in conclusion section of kklv10
-  (2) [[[COP helping some problems with hardness]]] having COP in the
-  input structure makes some problems less hard than in a general
-  input. -- min set cover problem is known to be NP hard. so is min
-  hitting
-  set. these two problems are equivalent cite-adp80\\
-  -- -- complexity results: log factor polynomial time approximable cite-f98, W[2]-complete with solution size as parameter cite-df99\\
-  -- a couple of generalizations:\\
-  -- -- min degree hypergraph (MDH) problem\\
-  -- -- red blue covering (RBSC) problem\\
-  ----- cite-d08phd Chapter 5.\\
-  ----- what are cite-d08phd results\\
-  ----- Rectangle stabbing \\
-  ----- see cite-d08phd Chapter 6?  }
-
-\tnote{ Some things to remember - have this somewhere.
-
-  \begin{itemize}
-  \item Examples: Illustration of the problem examples. both COP and
-    extension
-  \item Basic preliminaries: (definitions theorems etc) needed if any
-    - graph theory
-  \item Organization: Outline of document
-  \end{itemize}
-
-  Abstract must be a brief about what results we have and how it fits
-  in the body of research.
-
-  \begin{enumerate}
-  \item \emph{Area:} - Broad to Specialized. i.e. Combinatorial
-    algorithms -> Matrix reorganization -> general data reorganization
-    (interval assignment) -> path assignment
-  \item \emph{Class of problems:} say, data reorganization.
-  \item \emph{Nature of results:} Is a generalization. We have a
-    Polynomial algorithm for a subset of the generalization.
-  \end{enumerate}
-
-Introduction is a magnified version of Abstract with some brief
-citations etc. mini survey. Here talk about canonization how it
-connects these problems to graph isomorphism. Introduce our problem,
-explain why it is important etc. How canonization is a
-posisbility. Brief about results obtained.
-
-Someone who doesn't want to go through the whole thesis must get the
-gist of the whole document from the introduction chapter. that should
-be the outlook.
-
-once this is done, the rest of the chapters will fall into place.
-
-{\em Survey chapter} is the full fledged expansion of the survey in
-introduction with details, observations, theorems etc.
-
-  {\em The main themes} of the thesis that must not be left out:
-  \begin{enumerate}
-  \item connection of COP to {\bf set systems} and thus to general
-    data bases. \cite{nsnrs09}
-  \item {\bf graph isomorphism} by canonization \cite{kklv10}
-  \item {\bf certificate for a problem} \cite{mcc04}
-  \end{enumerate}
-
-Also see purple book notes.
-}
-
-\section{to area}
-\section{to problem}
-\section{to your thesis}
-\section{Organization of document}
-
-\chapter{Consecutive Ones property}
-\tnote{survey of cop includes mainly 
-\begin{enumerate}
-\item \cite{d08phd} (perhaps minus chap 4?)
-\item \cite{nsnrs09}
-\item \cite{mcc04}
-\item \cite{kklv10}
-\end{enumerate}
-.. to name a few.
-
-sections will be as follows:}
-\section{Characterization of COP}
-
-\tnote{\cite{at72}, \cite{mcc04}, prime submatrix \cite{wlh02}?}
-\section{Recognition of COP}
-\subsection{\tnote{Polynomial time algorithms}}
-\tnote{ (pq tree, pqr, icpia)}
-\subsection{\tnote{Tucker's submatrices, Dom's algo to find them}}
-\subsection{\tnote{McConnell's certificate}}
-
-\section{Alteration to matrices to get COP}
-\tnote{algorithms to make a matrix into one with COP
-  (min/max-cos-r/c) \cite{d08phd} }
-\section{brief analysis of hard problems with input having COP (chap
-  5, 6 in dom)}
-\label{sec:briefanal}
-\tnote{Do I need this?}
-\section{Complexity of certain COP variations}
-\tnote{maybe combine with Section~\ref{sec:briefanal}\\
-complexity - poly, logspace}
-
-\chapter{Other problems related to COP }
-\tnote{(see literature)}
-
-\chapter{Research - Tree path labeling of path hypergraphs}
-\tnote{Extension to trees. ( stuff in latin paper,)}
+\chapter{Consecutive-Ones property - Important Results from Literature}
+\tnote{TBD: have a few lines about organization of chapter}
+
+\tnote{ Survey chapter: is the full fledged expansion of the survey in
+  introduction with details, observations, theorems etc.}
+
+\section{Matrices with COP}
+\tnote{Expand on ref:sec:copmatrices}
+
+\section{Optimization problems in COP}
+\tnote{Expand on ref:sec:optcop}
+
+\chapter{Preliminaries to new results}
+%\section{Preliminaries} 
+\label{sec:prelims}
+
+\tnote{TBD: have a few lines about organization of chapter}
+
+This section states definitions and basic facts necessary in the scope
+of this document.
+
+\def\xnoindent {}
+
+The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
+  system} of a universe $U$ with $|U| = n$.  The {\em support} of a
+set system $\F$ denoted by $supp(\cF)$ is the union of all the sets in
+$\F$; $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this
+paper, a set system is required to ``cover'' the universe; $
+supp(\cF) = U$.
+%\vspace{\prelimspace}
+
+\xnoindent The graph $T$ represents a {\em target tree} with same
+number of vertices as elements in $U$; $|V(T)| = n$.  A {\em path
+  system}\, $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P
+\mid P \subseteq V, \text{ } T[P] \text{ is a path} \}$.
+%\vspace{\prelimspace}
+
+% \xnoindent
+% A set system $\cF$ can be alternatively represented by a {\em
+%   hypergraph} $\H_\cF$ whose vertex set is $supp(\cF)$ and hyperedges
+% are the sets in $\cF$. This is a known representation for interval
+% systems in literature \cite{bls99,kklv10}.  We extend this definition here to
+% path systems.  
+\xnoindent
+A set system $\cF$ can be alternatively represented by a {\em
+  hypergraph}\, $\cF_H$ whose vertex set is $supp(\cF)$ and hyperedges
+are the sets in $\cF$. This is a known representation for interval
+systems in literature \cite{bls99,kklv10}.  We extend this definition here to
+path systems. Due to the equivalence of set system and hypergraph in
+the scope of this paper, we drop the
+subscript $_H$ in the notation and refer to both the structures by $\cF$.
+%\vspace{\prelimspace}
+
+\xnoindent Two hypergraphs $\cF'$, $\cF''$ are said to be {\em
+  isomorphic} to each other, denoted by $\cF' \cong \cF''$, iff there
+exists a bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such
+that for all sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in
+$\cF'$ iff $B$ is a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x
+\in A\}$ \cite{kklv10}. This is called {\em hypergraph
+  isomorphism}. \tnote{also extend $\phi$ to hyperedges -- see if
+  required }
+%\vspace{\prelimspace}
+
+\xnoindent The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph
+$\cF$ is a graph such that its vertex set has a bijection to $\cF$ and
+there exists an edge between two vertices iff their corresponding
+hyperedges have a non-empty intersection \cite{mcg04}.
+%\vspace{\prelimspace}
+
+%\noindent 
+If the intersection graphs of two hypergraphs are isomorphic, $\bI(\cF)
+\cong \bI(\cP)$ where $\cP$ is also a path system, then the bijection
+$\cl: \cF \rightarrow \cP$ due to this isomorphism is called a {\em
+  path labeling} of the hypergraph $\cF$. To illustrate further, let
+$\cg: V(\cF) \rightarrow V(\cP)$ be the above mentioned isomorphism
+where $V(\cF)$ and $V(\cP)$ are the vertex sets that represent the
+hyperedges for each hypergraph respectively, $V(\cF) = \{ v_S \mid S
+\in \cF\}$ and $V(\cP) = \{ v_P \mid P \in \cP\}$. Then the path
+labeling $\cl$ is defined as follows: $\cl(S_1) = P_1$ iff $\cg
+(v_{S_1}) = v_{P_1}$. The path system $\cP$ may be alternatively
+denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In most scenarios in
+this paper, what is given are the pair $(\cF, \cl)$ and the target
+tree $T$; hence this notation will be used more often.
+
+\xnoindent If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$
+is called a {\em path hypergraph} and $\cP$ is called {\em path
+  representation} of $\cF$. If this isomorphism is $\phi: supp(\cF)
+\rightarrow V(T)$, then it is clear that there is an {\em induced path
+labeling} $\cl_\phi: \cF \rightarrow \cP$ to the set system;
+$\cl_\phi(S) = \set{y \mid y = \phi(x), x \in S}$ for all $S \in \cF$. Recall that
+$supp(\cP) = V(T)$.
+%\vspace{\prelimspace}
+
+% A graph $G$ is a {\em path graph} if it is isomorphic to the intersection graph
+% $\bI(\cP)$ of a path system $\cP$. 
+% This
+% isomorphism gives a bijection $\cl': V(G) \rightarrow \cP$. Moreover,
+% for the purposes of this paper, we require that in a path labeling,
+% $supp(\cP) = V(T)$.  If graph $G$ is also isomorphic to $\bI(\cF)$ for some hypergraph
+% $\cF$, then clearly there is a bijection $\cl: \cF \rightarrow \cP$
+% such that $\cl(S) = \cl'(v_S)$ where $v_S$ is the vertex corresponding
+% to set $S$ in $\bI(\cF)$ for any $S \in \cF$. This bijection $\cl$ is
+% called the {\em path labeling} of the hypergraph $\cF$ and the path
+% system $\cP$ may be alternatively denoted as $\cF^\cl$.
+%\vspace{\prelimspace}
+
+\xnoindent
+A path labeling $(\cF, \cl)$ is defined to be {\em
+  feasible} if
+$\cF \cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
+\rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
+\rightarrow \cF^\cl$ such that $\cl_\phi = \cl$. 
+%\vspace{\prelimspace}
+
+\xnoindent An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is
+a graph such that its vertex set has a bijection to $\cF$ and there
+exists an edge between two of its vertices iff their corresponding
+hyperedges overlap. Two hyperedges $S$ and $S'$ are said to {\em
+  overlap}, denoted by $S \overlap S'$, if they have a non-empty
+intersection and neither is contained in the other; $S \overlap S'
+\text{ iff } S \cap S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq
+S$. Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$ and not
+necessarily connected. Each connected component of $\bO(\cF)$ is
+called an {\em overlap component}.
+% If there are $d$ overlap components
+% in $\bO(\cF)$, the set subsystems are denoted by $\cO_1, \cO_2, \ldots
+% \cO_d$. Clearly $\cO_i \subseteq \F, i \in [d]$. For any $i, j \in [d]$,
+% it can be verified that one of the following is true.
+% \begin{enumerate}[a) ]
+% \item $supp(\cO_i)$ and $supp(\cO_j)$ are disjoint
+% \item $supp(\cO_i)$ is a subset of a set in $\cO_j$
+% \item $supp(\cO_j)$ is a subset of a set in $\cO_i$
+% \end{enumerate}
+%\vspace{\prelimspace}
+
+\xnoindent
+A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
+\overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
+\cite{kklv10}. Additionally, if $S$ is such that it is contained in no
+other hyperedge in $\cF$, i.e., it is inclusion maximal then it is called
+{\em super-marginal}.
+
+\xnoindent A {\em star} graph is a complete bipartite graph $K_{1,p}$
+which is clearly a tree and $p$ is the number of leaves. The vertex
+with maximum degree is called the {\em center} of the star and the
+edges are called {\em rays} of the star. A {\em $k$-subdivided star}
+is a star with all its rays subdivided exactly $k$ times. The
+definition of a {\em ray of a $k$-subdivided star} is extended to the path
+from the center to a leaf. It is clear that all rays are of length $k+2$.
+
+\chapter{Tree Path Labeling of Path Hypergraphs - the New Results}
 
@@
   INCOMPLETE SEC 4.}
-\section{Introduction}
-Consecutive ones property (COP) of binary matrices is a widely studied
-combinatorial problem. The problem is to rearrange rows (columns) of a
-binary matrix in such a way that every column (row) has its $1$s occur
-consecutively. If this is possible the matrix is said to have the COP.
-This problem has several practical applications in diverse fields
-including scheduling \cite{hl06}, information retrieval \cite{k77} and
-computational biology \cite{abh98}.  Further, it is a tool in graph
-theory \cite{mcg04} for interval graph recognition, characterization
-of hamiltonian graphs, and in integer linear programming
-\cite{ht02,hl06}.  Recognition of COP is polynomial time solvable by
-several algorithms. PQ trees \cite{bl76}, variations of PQ trees
-\cite{mm96,wlh01,wlh02,mcc04}, ICPIA \cite{nsnrs09} are the main ones.
-
-%\noindent
 The problem of COP testing can be easily seen as a simple constraint
@@
 
-\noindent 
-{\bf Our Work.}  A natural generalization of the interval
+\tnote{TBD: have a few lines about organization of chapter}
+The necessary preliminaries with definitions
+etc. are presented in
+Section~\ref{sec:prelims}. Section~\ref{sec:feasible} documents the
+characterization of a feasible path labeling and finally,
+Section~\ref{sec:ksubdivstar} describes a polynomial time algorithm to
+find the tree path labeling of a given set system from a given
+$k$-subdivided tree.
+
+\section{Introduction} 
+A natural generalization of the interval
 assignment problem is feasible tree path labeling problem of a set
@@
 % equivalent set of prime matrices.
-
-\noindent {\bf Roadmap.} The necessary preliminaries with definitions
-etc. are presented in
-Section~\ref{sec:prelims}. Section~\ref{sec:feasible} documents the
-characterization of a feasible path labeling and finally,
-Section~\ref{sec:ksubdivstar} describes a polynomial time algorithm to
-find the tree path labeling of a given set system from a given
-$k$-subdivided tree.
-
-\section{Preliminaries} 
-\label{sec:prelims}
-
-This section states definitions and basic facts necessary in the scope
-of this document.
-
-\def\xnoindent {}
-
-The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
-  system} of a universe $U$ with $|U| = n$.  The {\em support} of a
-set system $\F$ denoted by $supp(\cF)$ is the union of all the sets in
-$\F$; $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this
-paper, a set system is required to ``cover'' the universe; $
-supp(\cF) = U$.
-%\vspace{\prelimspace}
-
-\xnoindent The graph $T$ represents a {\em target tree} with same
-number of vertices as elements in $U$; $|V(T)| = n$.  A {\em path
-  system}\, $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P
-\mid P \subseteq V, \text{ } T[P] \text{ is a path} \}$.
-%\vspace{\prelimspace}
-
-% \xnoindent
-% A set system $\cF$ can be alternatively represented by a {\em
-%   hypergraph} $\H_\cF$ whose vertex set is $supp(\cF)$ and hyperedges
-% are the sets in $\cF$. This is a known representation for interval
-% systems in literature \cite{bls99,kklv10}.  We extend this definition here to
-% path systems.  
-\xnoindent
-A set system $\cF$ can be alternatively represented by a {\em
-  hypergraph}\, $\cF_H$ whose vertex set is $supp(\cF)$ and hyperedges
-are the sets in $\cF$. This is a known representation for interval
-systems in literature \cite{bls99,kklv10}.  We extend this definition here to
-path systems. Due to the equivalence of set system and hypergraph in
-the scope of this paper, we drop the
-subscript $_H$ in the notation and refer to both the structures by $\cF$.
-%\vspace{\prelimspace}
-
-\xnoindent Two hypergraphs $\cF'$, $\cF''$ are said to be {\em
-  isomorphic} to each other, denoted by $\cF' \cong \cF''$, iff there
-exists a bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such
-that for all sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in
-$\cF'$ iff $B$ is a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x
-\in A\}$ \cite{kklv10}. This is called {\em hypergraph
-  isomorphism}. \tnote{also extend $\phi$ to hyperedges -- see if
-  required }
-%\vspace{\prelimspace}
-
-\xnoindent The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph
-$\cF$ is a graph such that its vertex set has a bijection to $\cF$ and
-there exists an edge between two vertices iff their corresponding
-hyperedges have a non-empty intersection \cite{mcg04}.
-%\vspace{\prelimspace}
-
-%\noindent 
-If the intersection graphs of two hypergraphs are isomorphic, $\bI(\cF)
-\cong \bI(\cP)$ where $\cP$ is also a path system, then the bijection
-$\cl: \cF \rightarrow \cP$ due to this isomorphism is called a {\em
-  path labeling} of the hypergraph $\cF$. To illustrate further, let
-$\cg: V(\cF) \rightarrow V(\cP)$ be the above mentioned isomorphism
-where $V(\cF)$ and $V(\cP)$ are the vertex sets that represent the
-hyperedges for each hypergraph respectively, $V(\cF) = \{ v_S \mid S
-\in \cF\}$ and $V(\cP) = \{ v_P \mid P \in \cP\}$. Then the path
-labeling $\cl$ is defined as follows: $\cl(S_1) = P_1$ iff $\cg
-(v_{S_1}) = v_{P_1}$. The path system $\cP$ may be alternatively
-denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In most scenarios in
-this paper, what is given are the pair $(\cF, \cl)$ and the target
-tree $T$; hence this notation will be used more often.
-
-\xnoindent If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$
-is called a {\em path hypergraph} and $\cP$ is called {\em path
-  representation} of $\cF$. If this isomorphism is $\phi: supp(\cF)
-\rightarrow V(T)$, then it is clear that there is an {\em induced path
-labeling} $\cl_\phi: \cF \rightarrow \cP$ to the set system;
-$\cl_\phi(S) = \set{y \mid y = \phi(x), x \in S}$ for all $S \in \cF$. Recall that
-$supp(\cP) = V(T)$.
-%\vspace{\prelimspace}
-
-% A graph $G$ is a {\em path graph} if it is isomorphic to the intersection graph
-% $\bI(\cP)$ of a path system $\cP$. 
-% This
-% isomorphism gives a bijection $\cl': V(G) \rightarrow \cP$. Moreover,
-% for the purposes of this paper, we require that in a path labeling,
-% $supp(\cP) = V(T)$.  If graph $G$ is also isomorphic to $\bI(\cF)$ for some hypergraph
-% $\cF$, then clearly there is a bijection $\cl: \cF \rightarrow \cP$
-% such that $\cl(S) = \cl'(v_S)$ where $v_S$ is the vertex corresponding
-% to set $S$ in $\bI(\cF)$ for any $S \in \cF$. This bijection $\cl$ is
-% called the {\em path labeling} of the hypergraph $\cF$ and the path
-% system $\cP$ may be alternatively denoted as $\cF^\cl$.
-%\vspace{\prelimspace}
-
-\xnoindent
-A path labeling $(\cF, \cl)$ is defined to be {\em
-  feasible} if
-$\cF \cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
-\rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
-\rightarrow \cF^\cl$ such that $\cl_\phi = \cl$. 
-%\vspace{\prelimspace}
-
-\xnoindent An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is
-a graph such that its vertex set has a bijection to $\cF$ and there
-exists an edge between two of its vertices iff their corresponding
-hyperedges overlap. Two hyperedges $S$ and $S'$ are said to {\em
-  overlap}, denoted by $S \overlap S'$, if they have a non-empty
-intersection and neither is contained in the other; $S \overlap S'
-\text{ iff } S \cap S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq
-S$. Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$ and not
-necessarily connected. Each connected component of $\bO(\cF)$ is
-called an {\em overlap component}.
-% If there are $d$ overlap components
-% in $\bO(\cF)$, the set subsystems are denoted by $\cO_1, \cO_2, \ldots
-% \cO_d$. Clearly $\cO_i \subseteq \F, i \in [d]$. For any $i, j \in [d]$,
-% it can be verified that one of the following is true.
-% \begin{enumerate}[a) ]
-% \item $supp(\cO_i)$ and $supp(\cO_j)$ are disjoint
-% \item $supp(\cO_i)$ is a subset of a set in $\cO_j$
-% \item $supp(\cO_j)$ is a subset of a set in $\cO_i$
-% \end{enumerate}
-%\vspace{\prelimspace}
-
-\xnoindent
-A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
-\overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
-\cite{kklv10}. Additionally, if $S$ is such that it is contained in no
-other hyperedge in $\cF$, i.e., it is inclusion maximal then it is called
-{\em super-marginal}.
-
-\xnoindent A {\em star} graph is a complete bipartite graph $K_{1,p}$
-which is clearly a tree and $p$ is the number of leaves. The vertex
-with maximum degree is called the {\em center} of the star and the
-edges are called {\em rays} of the star. A {\em $k$-subdivided star}
-is a star with all its rays subdivided exactly $k$ times. The
-definition of a {\em ray of a $k$-subdivided star} is extended to the path
-from the center to a leaf. It is clear that all rays are of length $k+2$.
 
@@
 
-\section{Testing for feasible path assignments to $k$-subdivided star}
+\section{ TPL with special target trees}
+\tnote{give problem definition etc}
+
+\subsection{Testing for feasible path assignments to $k$-subdivided star}
 \label{sec:ksubdivstar}
@@
 
-\chapter{ Research - [title TBD]} 
-\tnote{the matrix theory submitted in walcom, any other work - need to see notes.}
+\section{ TPL with no restrictions}
+\tnote{ADD WALCOM PAPER TEXT}
 
 \chapter{Conclusion}
+% \section{Conclusion}
+\label{sec:conclusion}
+%\vspace{\secfirstparatrim}
+
+We give a characterization for feasible tree path labeling of path
+hypergraphs. The proof for this is constructive and computes a
+feasibility bijection mapping vertices of the hypergraph to vertices
+of the given target tree. This thesis also discovered an exponential algorithm
+that computes a feasible TPL to a given hypergraph if it is a path hypergraph.
+
+Our results have close connections to recognition of path graphs and
+to path graph isomorphism.  Graphs which can be represented as the
+intersection graph of paths in a tree are called {\em path
+  graphs}\cite{mcg04}. Thus, a hypergraph $\cF$ can be interpreted as
+paths in a tree, if and only if the intersection graph of $\cF$ is a
+{\em path graph}. Path graphs are a subclass of chordal graphs since
+chordal graphs are characterized as the intersection graphs of
+subtrees of a tree\cite{mcg04}.  Path graphs are well studied in the
+literature \cite{plr70}--\cite{mcg04}\tnote[TH]{Check if cite:bp93 is
+  indeed for path graphs}.  Path graph recognition can be done in
+polynomial time\cite{gav78,aas93}.  Clearly, this is a necessary
+condition in terms of the intersection graph of the input hypergraph
+$\cF$ in \FTPL. However, one can easily obtain a counterexample to
+show the insufficiency of this condition\tnote[GTC]{Give a figure
+  showing insuff.}.  Path graph isomorphism is known be
+isomorphism-complete\cite{kklv10}. Therefore, \annote[GTC]{it is unlikely
+  that we can solve the problem of finding feasible path labeling
+  $\cl$ for a given $\cF$ and }{WHY?} tree $T$.  It is
+  definitely interesting to classify the kinds of trees and
+  hypergraphs for which feasible path labelings can be\tnote[TH]{ELABORATE.}
+found efficiently.  \annote[TH]{These results would form a natural
+  generalization of COP testing and interval graph isomorphism,
+  culminating in Graph Isomorphism }{REWORD} itself.
+ To this effect we consider TPL on {\kstar} and give a polynomial time
+ solution for the same.
+
+Whether TPL on general trees is solvable in $P$ remains open. So do
+optimization opportunities in TPL (possible extensions to optimization for COP in the
+Section~\ref{sec:optcop}.\tnote[TH]{link to section}).
+
+\tnote[TH]{ ||Complexity challenges:|| Canonization is an important tool in graph isomorphism. Invention of a
+% deterministic method of canonization for any class of graphs naturally
+% results in an algorithm for isomorphism. All that is required is to
+% check if the canons of two graphs are the same. Thus complexity of
+% graph isomorphism can be studied by studying canonization
+% methods. While general graph ismorphism remains elusive in terms of
+% hardness, canonization has been studied for smaller classes of graphs
+% thus giving complexity/hardness results for them.
+
+% In 1992, \cite{sl92} made an important discovery that tree isomorphism
+% is in logspace. It was done by inventing a method of canonization of
+% trees using a logspace depth first traversal
+% algorithm. \tnote{confirm the statement of claim}.
+
+% \cite{kklv10} proved that interval graph canonization is also in
+% logspace thus drawing logspace conslusions about COT. 
+% % This result is
+% % built on the fact that 
+% Interval graphs are FP+C (fixed point with
+% counting) definable \tnote{elucidate} and \cite{l10} showed that
+% this implies that it captures PTIME. \tnote{elucidate} This result
+% along with that of undirected graph reachability being in logspace
+% \cite{rei08}, \cite{kklv10} proved their logspace result.
+}%
+\tnote[TH]{ (1) the interesting survey in conclusion section of kklv10
+  (2) [[[COP helping some problems with hardness]]] having COP in the
+  input structure makes some problems less hard than in a general
+  input. -- min set cover problem is known to be NP hard. so is min
+  hitting
+  set. these two problems are equivalent cite-adp80\\
+  -- -- complexity results: log factor polynomial time approximable cite-f98, W[2]-complete with solution size as parameter cite-df99\\
+  -- a couple of generalizations:\\
+  -- -- min degree hypergraph (MDH) problem\\
+  -- -- red blue covering (RBSC) problem\\
+  ----- cite-d08phd Chapter 5.\\
+  ----- what are cite-d08phd results\\
+  ----- Rectangle stabbing \\
+  ----- see cite-d08phd Chapter 6?  }
 

### 2011-12-22 src/thesis.tex
Reorg-ed lit. Added themes - set sys, canonization, certifying algo - to toc.

@@
 xxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
-\textbf{COP}   \> Consecutive-Ones Property \\
-\textbf{COT}   \> Consecutive-Ones Testing \\
+\textbf{COP}   \> Consecutive-ones Property \\
+\textbf{COT}   \> Consecutive-ones Testing \\
 \textbf{ICPIA}   \> Intersection Cardinality Preservation Interval Assignment \\
@@
 \tnote{ The main themes of the thesis that must not be left out:\\
-  connection of COP to ||set systems|| and thus to general
+  connection of COP to *SET SYSTEMS* and thus to general
   data bases. cite:nsnrs09\\
-  ||graph isomorphism|| by canonization cite:kklv10\\
-  ||certificate for a problem|| cite:mcc04\\
-  Also see purple book notes.  }
+  *GRAPH ISOMORPHSIM* by canonization cite:kklv10\\
+  *CERTIFICATE FOR A PROBLEM* cite:mcc04\\
+  Also see purple book notes.}
 
@@
 
-\section{Consecutive-Ones Property Testing - a Brief Survey} %{Background}
+\section{Consecutive-ones Property Testing - a Brief Survey} %{Background}
 \label{sec:background}
@@
   cite:th98 (generalization of COP - minimax, biotonic column) Tucker}
+
+\section{Application of COP in areas of Graph theory and algorithms}
+\subsection{COP in relational database model}
+\tnote{(set systems theme)}
+\subsection{COP in graph isomorphism}
+\tnote{(canonization theme)}
+\subsection{Invention of certifying algorithms using COP}
+\tnote{ (certification McC04 theme)}
 
@@
 % % \end{enumerate}
-% % \vspace{\prelimspace} \xnoindent
+% % \vspace{\prelimspace} 
 % A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
@@
 
-\chapter{Consecutive-Ones property - Important Results from Literature}
+\chapter{Consecutive-ones Property - a Survey of Important Results}
 \tnote{TBD: have a few lines about organization of chapter}
@@
 
+\tnote{ADD NEW SECTIONS - MATCHING WHAT'S IN INTRO CHAPTER}
+
 \chapter{Preliminaries to new results}
+\label{ch:prelims} 
 %\section{Preliminaries} 
-\label{sec:prelims}
-
-\tnote{TBD: have a few lines about organization of chapter}
-
-This section states definitions and basic facts necessary in the scope
-of this document.
-
-\def\xnoindent {}
+%\label{sec:prelims}
+
+\tnote{TBD: have a few lines about organization of chapter}This
+section states definitions and basic facts necessary in the scope of
+this document.
+
+\tnote{Move some of this to general prelim in chapter 1}
 
@@
 
-\xnoindent The graph $T$ represents a {\em target tree} with same
+The graph $T$ represents a {\em target tree} with same
 number of vertices as elements in $U$; $|V(T)| = n$.  A {\em path
@@
 
-% \xnoindent
+% 
 % A set system $\cF$ can be alternatively represented by a {\em
@@
 % path systems.
-\xnoindent
+
 A set system $\cF$ can be alternatively represented by a {\em
@@
 
-\xnoindent Two hypergraphs $\cF'$, $\cF''$ are said to be {\em
+Two hypergraphs $\cF'$, $\cF''$ are said to be {\em
   isomorphic} to each other, denoted by $\cF' \cong \cF''$, iff there
@@
 
-\xnoindent The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph
+The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph
 $\cF$ is a graph such that its vertex set has a bijection to $\cF$ and
@@
 
-\xnoindent If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$
+If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$
 is called a {\em path hypergraph} and $\cP$ is called {\em path
@@
 
-\xnoindent
 A path labeling $(\cF, \cl)$ is defined to be {\em
@@
 
-\xnoindent An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is
+An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is
 a graph such that its vertex set has a bijection to $\cF$ and there
@@
 
-\xnoindent
 A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
@@
 
-\xnoindent A {\em star} graph is a complete bipartite graph $K_{1,p}$
+A {\em star} graph is a complete bipartite graph $K_{1,p}$
 which is clearly a tree and $p$ is the number of leaves. The vertex
@@
 etc. are presented in
-Section~\ref{sec:prelims}. Section~\ref{sec:feasible} documents the
+Section~\ref{ch:prelims}. Section~\ref{sec:feasible} documents the
 characterization of a feasible path labeling and finally,
@@
 
-\def \xnoindent  {}
-
-\xnoindent In the remaining part of this section we show that $(\cF,
+In the remaining part of this section we show that $(\cF,
 \cl)$ is feasible if and only if it is an ICPPL and
@@
 
-  \xnoindent
  \begin{enumerate}[\xbullet]
@@
 
-\xnoindent As a result of Algorithm~\ref{perms} each leaf $v$ in $T$
+As a result of Algorithm~\ref{perms} each leaf $v$ in $T$
 is such that there is exactly one set in $\cF$ with $v$ as a vertex in
@@
 
-\xnoindent
 Suppose the input ICPPL $(\cF, \cl)$ is feasible, yet set $X$ in
@@
 
-\xnoindent We have seen two filtering algorithms above, namely,
+We have seen two filtering algorithms above, namely,
 Algorithm~\ref{perms} {\tt filter\_1} and Algorithm~\ref{leafasgn}
@@
 
-\xnoindent Algorithm~\ref{al:icppl-find-isomorph} computes a
+Algorithm~\ref{al:icppl-find-isomorph} computes a
 hypergraph isomorphism $\phi$ recursively using Algorithm~\ref{perms}
@@
 \label{subsec:icpplicpia}
-\xnoindent
+
 Consider a special case of ICPPL with the following properties.
@@
 \cite{kklv10,nsnrs09,wlh02}.  We use the notation in
-\cite{kklv10}. Recall from Section~\ref{sec:prelims} that hyperedges
+\cite{kklv10}. Recall from Section~\ref{ch:prelims} that hyperedges
 $S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$ and $S'$
@@
 
-\xnoindent
-Recall from Section~\ref{sec:prelims} that a $k$-subdivided star is a
+Recall from Section~\ref{ch:prelims} that a $k$-subdivided star is a
 star with each edge subdivided $k$ times.  Therefore, a $k$-subdivided
@@
 
-\xnoindent In this section the given hypergraph, the
+In this section the given hypergraph, the
 $k$-subdivided star and the root of the star are denoted by $\cO$, $T$ and vertex
@@
 
-% \xnoindent
+% 
 % We iteratively consider each ray from which paths will be assigned to
@@
 
-\xnoindent
 The set $\cO_i$ refers to the set of hyperedges $\cT_1^i \cup \cT_2^i$ in

### 2011-12-28 src/thesis.tex
Nightly update. Trivial changes.
Cool story bro but it's FUCKING TIME TO STEP ON THE GAS.

@@
 
-\def \mythesissubmissiondate {30 December 2011}
-\def \mythesissubmissionmonth {\MakeUppercase {December 2011}}
+\def \mythesissubmissiondate {10 January 2012}
+\def \mythesissubmissionmonth {\MakeUppercase {January 2012}}
 
@@
 
-\chapter{Preliminaries to new results}
-\label{ch:prelims} 
-%\section{Preliminaries} 
-%\label{sec:prelims}
-
-\tnote{TBD: have a few lines about organization of chapter}This
-section states definitions and basic facts necessary in the scope of
-this document.
-
-\tnote{Move some of this to general prelim in chapter 1}
-
-The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
-  system} of a universe $U$ with $|U| = n$.  The {\em support} of a
-set system $\F$ denoted by $supp(\cF)$ is the union of all the sets in
-$\F$; $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this
-paper, a set system is required to ``cover'' the universe; $
-supp(\cF) = U$.
-%\vspace{\prelimspace}
-
-The graph $T$ represents a {\em target tree} with same
-number of vertices as elements in $U$; $|V(T)| = n$.  A {\em path
-  system}\, $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P
-\mid P \subseteq V, \text{ } T[P] \text{ is a path} \}$.
-%\vspace{\prelimspace}
-
-% 
-% A set system $\cF$ can be alternatively represented by a {\em
-%   hypergraph} $\H_\cF$ whose vertex set is $supp(\cF)$ and hyperedges
-% are the sets in $\cF$. This is a known representation for interval
-% systems in literature \cite{bls99,kklv10}.  We extend this definition here to
-% path systems.
-
-A set system $\cF$ can be alternatively represented by a {\em
-  hypergraph}\, $\cF_H$ whose vertex set is $supp(\cF)$ and hyperedges
-are the sets in $\cF$. This is a known representation for interval
-systems in literature \cite{bls99,kklv10}.  We extend this definition here to
-path systems. Due to the equivalence of set system and hypergraph in
-the scope of this paper, we drop the
-subscript $_H$ in the notation and refer to both the structures by $\cF$.
-%\vspace{\prelimspace}
-
-Two hypergraphs $\cF'$, $\cF''$ are said to be {\em
-  isomorphic} to each other, denoted by $\cF' \cong \cF''$, iff there
-exists a bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such
-that for all sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in
-$\cF'$ iff $B$ is a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x
-\in A\}$ \cite{kklv10}. This is called {\em hypergraph
-  isomorphism}. \tnote{also extend $\phi$ to hyperedges -- see if
-  required }
-%\vspace{\prelimspace}
-
-The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph
-$\cF$ is a graph such that its vertex set has a bijection to $\cF$ and
-there exists an edge between two vertices iff their corresponding
-hyperedges have a non-empty intersection \cite{mcg04}.
-%\vspace{\prelimspace}
-
-%\noindent 
-If the intersection graphs of two hypergraphs are isomorphic, $\bI(\cF)
-\cong \bI(\cP)$ where $\cP$ is also a path system, then the bijection
-$\cl: \cF \rightarrow \cP$ due to this isomorphism is called a {\em
-  path labeling} of the hypergraph $\cF$. To illustrate further, let
-$\cg: V(\cF) \rightarrow V(\cP)$ be the above mentioned isomorphism
-where $V(\cF)$ and $V(\cP)$ are the vertex sets that represent the
-hyperedges for each hypergraph respectively, $V(\cF) = \{ v_S \mid S
-\in \cF\}$ and $V(\cP) = \{ v_P \mid P \in \cP\}$. Then the path
-labeling $\cl$ is defined as follows: $\cl(S_1) = P_1$ iff $\cg
-(v_{S_1}) = v_{P_1}$. The path system $\cP$ may be alternatively
-denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In most scenarios in
-this paper, what is given are the pair $(\cF, \cl)$ and the target
-tree $T$; hence this notation will be used more often.
-
-If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$
-is called a {\em path hypergraph} and $\cP$ is called {\em path
-  representation} of $\cF$. If this isomorphism is $\phi: supp(\cF)
-\rightarrow V(T)$, then it is clear that there is an {\em induced path
-labeling} $\cl_\phi: \cF \rightarrow \cP$ to the set system;
-$\cl_\phi(S) = \set{y \mid y = \phi(x), x \in S}$ for all $S \in \cF$. Recall that
-$supp(\cP) = V(T)$.
-%\vspace{\prelimspace}
-
-% A graph $G$ is a {\em path graph} if it is isomorphic to the intersection graph
-% $\bI(\cP)$ of a path system $\cP$. 
-% This
-% isomorphism gives a bijection $\cl': V(G) \rightarrow \cP$. Moreover,
-% for the purposes of this paper, we require that in a path labeling,
-% $supp(\cP) = V(T)$.  If graph $G$ is also isomorphic to $\bI(\cF)$ for some hypergraph
-% $\cF$, then clearly there is a bijection $\cl: \cF \rightarrow \cP$
-% such that $\cl(S) = \cl'(v_S)$ where $v_S$ is the vertex corresponding
-% to set $S$ in $\bI(\cF)$ for any $S \in \cF$. This bijection $\cl$ is
-% called the {\em path labeling} of the hypergraph $\cF$ and the path
-% system $\cP$ may be alternatively denoted as $\cF^\cl$.
-%\vspace{\prelimspace}
-
-A path labeling $(\cF, \cl)$ is defined to be {\em
-  feasible} if
-$\cF \cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
-\rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
-\rightarrow \cF^\cl$ such that $\cl_\phi = \cl$. 
-%\vspace{\prelimspace}
-
-An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is
-a graph such that its vertex set has a bijection to $\cF$ and there
-exists an edge between two of its vertices iff their corresponding
-hyperedges overlap. Two hyperedges $S$ and $S'$ are said to {\em
-  overlap}, denoted by $S \overlap S'$, if they have a non-empty
-intersection and neither is contained in the other; $S \overlap S'
-\text{ iff } S \cap S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq
-S$. Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$ and not
-necessarily connected. Each connected component of $\bO(\cF)$ is
-called an {\em overlap component}.
-% If there are $d$ overlap components
-% in $\bO(\cF)$, the set subsystems are denoted by $\cO_1, \cO_2, \ldots
-% \cO_d$. Clearly $\cO_i \subseteq \F, i \in [d]$. For any $i, j \in [d]$,
-% it can be verified that one of the following is true.
-% \begin{enumerate}[a) ]
-% \item $supp(\cO_i)$ and $supp(\cO_j)$ are disjoint
-% \item $supp(\cO_i)$ is a subset of a set in $\cO_j$
-% \item $supp(\cO_j)$ is a subset of a set in $\cO_i$
-% \end{enumerate}
-%\vspace{\prelimspace}
-
-A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
-\overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
-\cite{kklv10}. Additionally, if $S$ is such that it is contained in no
-other hyperedge in $\cF$, i.e., it is inclusion maximal then it is called
-{\em super-marginal}.
-
-A {\em star} graph is a complete bipartite graph $K_{1,p}$
-which is clearly a tree and $p$ is the number of leaves. The vertex
-with maximum degree is called the {\em center} of the star and the
-edges are called {\em rays} of the star. A {\em $k$-subdivided star}
-is a star with all its rays subdivided exactly $k$ times. The
-definition of a {\em ray of a $k$-subdivided star} is extended to the path
-from the center to a leaf. It is clear that all rays are of length $k+2$.
-
 \chapter{Tree Path Labeling of Path Hypergraphs - the New Results}
@@
   INCOMPLETE SEC 4.}
-The problem of COP testing can be easily seen as a simple constraint
-satisfaction problem involving a system of sets from a universe. Every
-column of the binary matrix can be converted into a set of integers
-which are the indices of the rows with $1$s in that column. When
-observed in this context, if the matrix has the COP, a reordering of
-its rows will result in sets that have only consecutive integers. In
-other words, the sets after reordering are intervals. Indeed the
-problem now becomes finding interval assignments to the given set
-system \cite{nsnrs09} with a single permutation of the universe (set
-of row indices) which permutes each set to its interval. The result in
-\cite{nsnrs09} characterizes interval assignments to the sets which
-can be obtained from a single permutation of the rows.  They show that
-for each set, the cardinality of the interval assigned to it must be
-same as the cardinality of the set, and the intersection cardinality
-of any two sets must be same as the interesction cardinality of the
-corresponding intervals.  While this is naturally a necessary
-condition, \cite{nsnrs09} shows this is indeed sufficient.  Such an
-interval assignment is called an Intersection Cardinality Preserving
-Interval Assignment (ICPIA).  Finally, the idea of decomposing a given
-0-1 matrix into prime matrices to check for COP is adopted from
-\cite{wlh02} to test if an ICPIA exists for a given set system.
-
-\tnote{TBD: have a few lines about organization of chapter}
-The necessary preliminaries with definitions
-etc. are presented in
-Section~\ref{ch:prelims}. Section~\ref{sec:feasible} documents the
-characterization of a feasible path labeling and finally,
-Section~\ref{sec:ksubdivstar} describes a polynomial time algorithm to
-find the tree path labeling of a given set system from a given
-$k$-subdivided tree.
-
-\section{Introduction} 
+As we saw earlier, tree path labeling problem of path hypergraphs is a
+generalization of consecutive-ones property testing.
+
+\remove{The problem of COP testing can be easily seen as a simple
+  constraint satisfaction problem involving a system of sets from a
+  universe (analogous to hypergraph). Every column of the binary
+  matrix can be converted into a set of integers which are the indices
+  of the rows with |un|s in that column. When observed in this
+  context, if the matrix has the COP, a reordering of its rows will
+  result in sets that have only consecutive integers. In other words,
+  the sets after reordering are intervals. Indeed the problem now
+  becomes finding interval assignments to the given set system
+  cite|nsnrs09| with a single permutation of the universe (set of row
+  indices) which permutes each set to its interval. The result in
+  cite|nsnrs09| characterizes interval assignments to the sets which
+  can be obtained from a single permutation of the rows.  They show
+  that for each set, the cardinality of the interval assigned to it
+  must be same as the cardinality of the set, and the intersection
+  cardinality of any two sets must be same as the interesction
+  cardinality of the corresponding intervals.  While this is naturally
+  a necessary condition, cite|nsnrs09| shows this is indeed
+  sufficient.  Such an interval assignment is called an Intersection
+  Cardinality Preserving Interval Assignment (ICPIA).  Finally, the
+  idea of decomposing a given 0-1 matrix into prime matrices to check
+  for COP is adopted from cite|wlh02| to test if an ICPIA exists for a
+  given set system.}\tnote{REPEATITION from intro chapter}
+
 A natural generalization of the interval
@@
 
+The necessary preliminaries with definitions
+etc. are presented in
+Section~\ref{ch:prelims}. Section~\ref{sec:feasible} documents the
+characterization of a feasible path labeling for any path
+labeling from any target trees. Section~\ref{sec:spltargettree}
+describes the special case where the target tree is of a particular
+family of trees and specifically,
+Section~\ref{sec:ksubdivstar} describes a polynomial time algorithm to
+find the tree path labeling of a given set system from a given
+$k$-subdivided tree. Section~\ref{sec:norestraint} describes the
+general situation where the target tree has no restrictions and the
+algorithm to find a TPL, if  any, in this case.
+
+\section{Introduction}
+
+%\begin{tabular}[t]{c}
+{\small
+\begin{minipage}[h]{5in}
+ % \begin{singlespace}
+ \vspace{2mm}
+ {\large \FTPL}\\
+ \begin{tabular}[t]{l|l}
+ \hline\\
+    {\tt Input} & 
+    \begin{minipage}[t]{\probdefwidth}
+      A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
+      paths $\cP$ from $T$ and a
+      bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
+    \end{minipage}\\
+    {\tt Question} &
+    \begin{minipage}[t]{\probdefwidth}
+      Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
+      such that $\phi$ when applied on any hyperedge in $\cF$ will
+      give
+      the path mapped to it by the given tree path labeling $\cl$.\\
+      { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every hyperedge
+        $S \in \cF$.}
+    \end{minipage}\\
+  \end{tabular}
+  % \end{singlespace}
+\end{minipage}\\
+}
+%\end{tabular}
+
 %\noindent 
-\tnote{the following text could be in Further Study/Conclusion}It
+\annote{It
 is an interesting fact that for a matrix with the COP, the
 intersection graph of the corresponding set system is an interval
-graph.  A similar connection to a subclass of chordal graphs and a
+graph.}{the following text could be in Further Study/Conclusion}  A similar connection to a subclass of chordal graphs and a
 superclass of interval graphs exists for the generalization of COP.
@@
 must be a {\em path graph}. Chordal graphs are of great significance,
-extensively studied, and have several applications.  One of the well
+are extensively studied, and have several applications.  One of the well
 known and interesting properties of a chordal graphs is its connection
@@
 % equivalent set of prime matrices.
+
+\section{Preliminaries to new results}
+\label{ch:prelims} 
+%\section{Preliminaries} 
+%\label{sec:prelims}
+
+\tnote{TBD: have a few lines about organization of chapter}This
+section states definitions and basic facts necessary in the scope of
+this document.
+
+\tnote{Move some of this to general prelim in chapter 1}
+
+The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
+  system} of a universe $U$ with $|U| = n$.  The {\em support} of a
+set system $\F$ denoted by $supp(\cF)$ is the union of all the sets in
+$\F$; $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this
+paper, a set system is required to ``cover'' the universe; $
+supp(\cF) = U$.
+%\vspace{\prelimspace}
+
+The graph $T$ represents a {\em target tree} with same
+number of vertices as elements in $U$; $|V(T)| = n$.  A {\em path
+  system}\, $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P
+\mid P \subseteq V, \text{ } T[P] \text{ is a path} \}$.
+%\vspace{\prelimspace}
+
+% 
+% A set system $\cF$ can be alternatively represented by a {\em
+%   hypergraph} $\H_\cF$ whose vertex set is $supp(\cF)$ and hyperedges
+% are the sets in $\cF$. This is a known representation for interval
+% systems in literature \cite{bls99,kklv10}.  We extend this definition here to
+% path systems.
+
+A set system $\cF$ can be alternatively represented by a {\em
+  hypergraph}\, $\cF_H$ whose vertex set is $supp(\cF)$ and hyperedges
+are the sets in $\cF$. This is a known representation for interval
+systems in literature \cite{bls99,kklv10}.  We extend this definition here to
+path systems. Due to the equivalence of set system and hypergraph in
+the scope of this paper, we drop the
+subscript $_H$ in the notation and refer to both the structures by $\cF$.
+%\vspace{\prelimspace}
+
+Two hypergraphs $\cF'$, $\cF''$ are said to be {\em
+  isomorphic} to each other, denoted by $\cF' \cong \cF''$, iff there
+exists a bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such
+that for all sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in
+$\cF'$ iff $B$ is a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x
+\in A\}$ \cite{kklv10}. This is called {\em hypergraph
+  isomorphism}. \tnote{also extend $\phi$ to hyperedges -- see if
+  required }
+%\vspace{\prelimspace}
+
+The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph
+$\cF$ is a graph such that its vertex set has a bijection to $\cF$ and
+there exists an edge between two vertices iff their corresponding
+hyperedges have a non-empty intersection \cite{mcg04}.
+%\vspace{\prelimspace}
+
+%\noindent 
+If the intersection graphs of two hypergraphs are isomorphic, $\bI(\cF)
+\cong \bI(\cP)$ where $\cP$ is also a path system, then the bijection
+$\cl: \cF \rightarrow \cP$ due to this isomorphism is called a {\em
+  path labeling} of the hypergraph $\cF$. To illustrate further, let
+$\cg: V(\cF) \rightarrow V(\cP)$ be the above mentioned isomorphism
+where $V(\cF)$ and $V(\cP)$ are the vertex sets that represent the
+hyperedges for each hypergraph respectively, $V(\cF) = \{ v_S \mid S
+\in \cF\}$ and $V(\cP) = \{ v_P \mid P \in \cP\}$. Then the path
+labeling $\cl$ is defined as follows: $\cl(S_1) = P_1$ iff $\cg
+(v_{S_1}) = v_{P_1}$. The path system $\cP$ may be alternatively
+denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In most scenarios in
+this paper, what is given are the pair $(\cF, \cl)$ and the target
+tree $T$; hence this notation will be used more often.
+
+If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$
+is called a {\em path hypergraph} and $\cP$ is called {\em path
+  representation} of $\cF$. If this isomorphism is $\phi: supp(\cF)
+\rightarrow V(T)$, then it is clear that there is an {\em induced path
+labeling} $\cl_\phi: \cF \rightarrow \cP$ to the set system;
+$\cl_\phi(S) = \set{y \mid y = \phi(x), x \in S}$ for all $S \in \cF$. Recall that
+$supp(\cP) = V(T)$.
+%\vspace{\prelimspace}
+
+% A graph $G$ is a {\em path graph} if it is isomorphic to the intersection graph
+% $\bI(\cP)$ of a path system $\cP$. 
+% This
+% isomorphism gives a bijection $\cl': V(G) \rightarrow \cP$. Moreover,
+% for the purposes of this paper, we require that in a path labeling,
+% $supp(\cP) = V(T)$.  If graph $G$ is also isomorphic to $\bI(\cF)$ for some hypergraph
+% $\cF$, then clearly there is a bijection $\cl: \cF \rightarrow \cP$
+% such that $\cl(S) = \cl'(v_S)$ where $v_S$ is the vertex corresponding
+% to set $S$ in $\bI(\cF)$ for any $S \in \cF$. This bijection $\cl$ is
+% called the {\em path labeling} of the hypergraph $\cF$ and the path
+% system $\cP$ may be alternatively denoted as $\cF^\cl$.
+%\vspace{\prelimspace}
+
+A path labeling $(\cF, \cl)$ is defined to be {\em
+  feasible} if
+$\cF \cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
+\rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
+\rightarrow \cF^\cl$ such that $\cl_\phi = \cl$. 
+%\vspace{\prelimspace}
+
+An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is
+a graph such that its vertex set has a bijection to $\cF$ and there
+exists an edge between two of its vertices iff their corresponding
+hyperedges overlap. Two hyperedges $S$ and $S'$ are said to {\em
+  overlap}, denoted by $S \overlap S'$, if they have a non-empty
+intersection and neither is contained in the other; $S \overlap S'
+\text{ iff } S \cap S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq
+S$. Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$ and not
+necessarily connected. Each connected component of $\bO(\cF)$ is
+called an {\em overlap component}.
+% If there are $d$ overlap components
+% in $\bO(\cF)$, the set subsystems are denoted by $\cO_1, \cO_2, \ldots
+% \cO_d$. Clearly $\cO_i \subseteq \F, i \in [d]$. For any $i, j \in [d]$,
+% it can be verified that one of the following is true.
+% \begin{enumerate}[a) ]
+% \item $supp(\cO_i)$ and $supp(\cO_j)$ are disjoint
+% \item $supp(\cO_i)$ is a subset of a set in $\cO_j$
+% \item $supp(\cO_j)$ is a subset of a set in $\cO_i$
+% \end{enumerate}
+%\vspace{\prelimspace}
+
+A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
+\overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
+\cite{kklv10}. Additionally, if $S$ is such that it is contained in no
+other hyperedge in $\cF$, i.e., it is inclusion maximal then it is called
+{\em super-marginal}.
+
+A {\em star} graph is a complete bipartite graph $K_{1,p}$
+which is clearly a tree and $p$ is the number of leaves. The vertex
+with maximum degree is called the {\em center} of the star and the
+edges are called {\em rays} of the star. A {\em $k$-subdivided star}
+is a star with all its rays subdivided exactly $k$ times. The
+definition of a {\em ray of a $k$-subdivided star} is extended to the path
+from the center to a leaf. It is clear that all rays are of length $k+2$.
 
@@
 \section{ TPL with special target trees}
+\label{sec:spltargettree}
 \tnote{give problem definition etc}
@@
 \section{ TPL with no restrictions}
+\label{sec:norestraint}
 \tnote{ADD WALCOM PAPER TEXT}

### 2011-12-30 src/thesis.tex
Tracking turned off to send to review
Fixed the \Pi bug. Pigpen is now \Pig.

Conceivably the first draft of Chapter 3 (ch:myresearch). Cleaned up a bit also.
- Chapter 1 is complete moreorless except for basic prelims and the 3 themes.
- chapter 2 needs to be done from scratch. is the expanded version of section 1.3 and 1.4.
- chapter 3. needs clean up of walcom. add pictures.

The file sent to nari for review. The source is the same as on this commit.

@@
 \usepackage{../lib/mylatexlib}
+% \usepackage{comment}           % DEBUG ONLY
 
@@
   \hspace*{-0.25in}
-%   \begin{tabular}[h]{lr}
-%     {\bf Dr.~\myadvisor} & \\
-%     {\small Research Guide} & \\
-%     {\small Associate Professor} & \\
-%     {\small Dept. of \mydept} & \hfill {\small Chennai}\\
-%     {\small IIT Madras -- 600 036} & \hfill {\small\mythesissubmissiondate}
-%  \end{tabular}
   \parbox{3in}{
@@
 \chapter{Introduction}
-\label{chap:intro}
-
-% \section{Introduction}
-% \label{sec:intro}
-% \vspace{\secfirstparatrim}
-
-\tnote{ The main themes of the thesis that must not be left out:\\
-  connection of COP to *SET SYSTEMS* and thus to general
-  data bases. cite:nsnrs09\\
-  *GRAPH ISOMORPHSIM* by canonization cite:kklv10\\
-  *CERTIFICATE FOR A PROBLEM* cite:mcc04\\
-  Also see purple book notes.}
+\label{ch:intro}
 
@@
 \label{sec:problem}
-%\vspace{\secfirstparatrim}
-
-A group of students, \Pa, \Pi, \Sn, \Wo, \Vi, \Li, \Ch, \Sa, \Fr,
-  \Sc\ and \Lu\ enroll at the
-% {\bf Wo}odstock, {\bf Vi}olet, {\bf Li}nus, {\bf Ch}arlie, {\bf
-%   Sa}lly, {\bf Re}run, {\bf Sc}hr{\"o}eder} and {\em {\bf Lu}cy}
-% enroll at the
-  {\WSI} for a liberal arts programme.  As part of their semester
-  thesis, they pick a body of work to study and form the namesake
-  study groups, {\LLL}, {\GGG}, {\BBB} and {\TTT}\tnote[TH]{put bib
-    entries for these works!}. A student will be in at least one study
-  group and may be in more than one. For instance, as will be seen
-  later, {\Fr} studies both {\LLL} and {\TTT} while \Wo\ studies only
-  \BBB.
+
+A group of students, \Pa, \Pig, \Sn, \Wo, \Vi, \Li, \Ch, \Sa, \Fr,
+\Sc\ and \Lu\ enroll at the {\WSI} for a liberal arts programme.  As
+part of their semester thesis, they pick a body of work to study and
+form the namesake study groups, {\LLL}, {\GGG}, {\BBB} and
+{\TTT}\tnote[TH]{put bib entries for these works!}. A student will be
+in at least one study group and may be in more than one. For instance,
+as will be seen later, {\Fr} studies both {\LLL} and {\TTT} while \Wo\
+studies only \BBB.
 
@@
 
-\begin{table}[htbp]
+\begin{table}[t]%[htbp]
   \centering
@@
 
-% \begin{figure}[t]%[htbp]
-%  \centering
-%   \includegraphics[scale=0.3]{../img/1_infinite_loop.png}
-%   \caption{{\residenceblock} street map}
-%   \label{fig:streetmap}
-% \end{figure}
-
-% \begin{figure}[t]%[htbp]
-%   \centering
-%   \includegraphics[scale=0.3]{../img/2_infinite_loop_BTWF.png}
-%   \caption{{\residenceblock} street map with study group routes
-%     allocated. Routes are color coded as follows: red for
-%     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
-%     group, orange for \textcolor{yellow}{$\xBBB$} group, green for
-%     \textcolor{green}{$\xTTT$} group}
-%   \label{fig:streetmappath}
-% \end{figure}
-% \begin{figure}[t] % [h]
-%   \centering
-%   \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
-%   \caption{Individual allocation of apartments to students in
-%     {\residenceblock} that meets the requirements stated before.}
-%   \label{fig:streetmappathpeople}
-% \end{figure}
-
 \begin{figure}[htbp] %[t]%
@@
   (b) \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
+
  \label{fig:streetmap}
@@
    \textcolor{green}{$\xTTT$} group}
-% \end{figure}
-% \begin{figure}[htbp] %[t]%
-%   \centering
+
  \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
@@
 \end{figure}
-\tnote[TH]{make dashed/textured lines for routes. make it color agnostic.}
+\tnote{UPDATE IMAGE (b) 2-infinite-loop-BTWF.png. REMOVE TITLES. ADD
+  {$\xLLL$} {$\xGGG$} {$\xBBB$} {$\xTTT$}}
+
+\tnote[TH]{make dashed/textured lines for routes. make it color
+  agnostic.}
 
@@
 
-\section{Consecutive-ones Property Testing - a Brief Survey} %{Background}
+\section{Consecutive-ones Property Testing - a Brief Survey}
 \label{sec:background}
-%\vspace{\secfirstparatrim}
+
 In this section, a brief survey of the consecutive-ones problem and its
@@
 \label{sec:copmatrices}
-%\vspace{\secfirstparatrim} 
 As seen earlier, the interval assignment problem (illustrated as the
@@
   \centering
-  % \includegraphics[scale=0.4]{../img/0_cop_nocop.png}
+
   {\figtabsize
@@
 
-  \caption{\figtabsize Matrices with and without COP. $M_1$ has COP because by
-    permuting its columns, $c_1$-$c_4$, one can obtain $M_1'$ where
-    the {\un}s in each row are consecutive. $M_2$, however, does not
-    have COP since no permutation of its columns, $d_1$-$d_4$, will
-    arrange {\un}s in each row consecutively \cite{d08phd}.}
+  \caption{\figtabsize Matrices with and without COP. $M_1$ has COP
+    because by permuting its columns, $c_1$-$c_4$, one can obtain
+    $M_1'$ where the {\un}s in each row are consecutive. $M_2$,
+    however, does not have COP since no permutation of its columns,
+    $d_1$-$d_4$, will arrange {\un}s in each row consecutively
+    \cite{d08phd}.}
 
@@
 
-Recall that a matrix with COP is one whose rows (columns)
-can be rearranged so that the {\un}s in every column (row) are in
-consecutive rows (columns). Figure~\ref{fig:cop-matrix} shows examples
-of this property.
-COP in binary matrices has several practical applications in diverse
-fields including scheduling \cite{hl06}, information retrieval
-\cite{k77} and computational biology \cite{abh98}.  Further, it is a
-tool in graph theory \cite{mcg04} for interval graph recognition,
-characterization of Hamiltonian graphs, planarity testing \cite{bl76}
-and in integer linear programming \cite{ht02,hl06}.
+Recall that a matrix with COP is one whose rows (columns) can be
+rearranged so that the {\un}s in every column (row) are in consecutive
+rows (columns). Figure~\ref{fig:cop-matrix} shows examples of this
+property.  COP in binary matrices has several practical applications
+in diverse fields including scheduling \cite{hl06}, information
+retrieval \cite{k77} and computational biology \cite{abh98}.  Further,
+it is a tool in graph theory \cite{mcg04} for interval graph
+recognition, characterization of Hamiltonian graphs, planarity testing
+\cite{bl76} and in integer linear programming \cite{ht02,hl06}.
 
@@
 submatrices came as a corollary to the discovery that convex bipartite
-graphs are AT-free in \cite{at72}\tnote[GTC]{check}. The first linear time
-algorithm for COP testing (COT) was invented by \cite{bl76} using a
-data structure called PQ trees.  Since then several COT algorithms
+graphs are AT-free in \cite{at72}\tnote[TH]{check}. The first linear
+time algorithm for COP testing (COT) was invented by \cite{bl76} using
+a data structure called PQ trees.  Since then several COT algorithms
 have been invented -- some of which involved variations of PQ trees
@@
 \cite{wlh02,nsnrs09}, parallel COT algorithms\cite{as95,bs03,ly91} and
-certifying algorithms\cite{mcc04}. % These algorithms
-% improve over PQ trees on the time complexity and/or the complexity
-% of implementation.
-
-The construction of PQ trees in \cite{bl76} draws on the
-close relationship of matrices with COP to interval graphs. A PQ tree of a
+certifying algorithms\cite{mcc04}.
+
+The construction of PQ trees in \cite{bl76} draws on the close
+relationship of matrices with COP to interval graphs. A PQ tree of a
 matrix is one that stores all row (column) permutations of the matrix
-that give the COP orders (there could be multiple orders of rows or columns) of the matrix. This
-is constructed using an elaborate linear time procedure and is also
-a test for planarity\tnote[GTC]{check check check. both interval
-  graph and planarity in this paper?}. % Since this
-% result, several improvements have been made to the procedural
-% complexity of the algorithm itself and also derivative data structures
-% that are easier to construct.
-PQR trees is a generalized data structure based on PQ trees
-\cite{mm96,mpt98}.  \cite{tm05} describes an improved algorithm to
-build PQR trees. \tnote[GTC]{improv in terms of what?}\cite{wlh02}
-describes the simpler algorithm for COT. Hsu also invented PC trees
+that give the COP orders (there could be multiple orders of rows or
+columns) of the matrix. This is constructed using an elaborate linear
+time procedure and is also a test for planarity\tnote[TH]{check check
+  check. both interval graph and planarity in this paper?}.  PQR trees
+is a generalized data structure based on PQ trees \cite{mm96,mpt98}.
+\cite{tm05} describes an improved algorithm to build PQR
+trees. \tnote[TH]{improv in terms of what?}\cite{wlh02} describes the
+simpler algorithm for COT. Hsu also invented PC trees
 \cite{wlh01}\tnote[TH]{This result first appeared inproc ISAAC92}
@@
 this column (row). This is interesting and relevant, especially to
-this thesis because it simplifies COT to a great
-degree. \tnote[GTC]{it reduces the solution search space. fill in the blanks.}
+this thesis because it simplifies COT to a great degree. \tnote[TH]{it
+  reduces the solution search space. fill in the blanks.}
 
@@
 algorithm is related to the weighted betweenness problem
-addressed\tnote[GTC]{in what way??} in \cite{co98}.  \tnote[GTC]{expand
+addressed\tnote[TH]{in what way??} in \cite{co98}.  \tnote[TH]{expand
   on the COP order graph creation and it having to be bipartite for M
@@
 \label{sec:optcop}
-%\vspace{\secfirstparatrim}  
+
 So far we have been concerned about matrices that have the consecutive
@@
 COP. Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
-a submatrix $M'$ with all columns\tnote[GTC]{check if b75 deals with
+a submatrix $M'$ with all columns\tnote[TH]{check if b75 deals with
   COP col or COP row. also is it any submatrix with k less than r rows
@@
 discusses approximation of {\sc COP Augmentation} which is the problem
-of changing of the minimum number of zero entries to {\un}s so that the
-resulting matrix has COP. As mentioned earlier, this problem is
+of changing of the minimum number of zero entries to {\un}s so that
+the resulting matrix has COP. As mentioned earlier, this problem is
 known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
-using a reduction to the longest path problem, \tnote[GTC]{or is it a
+using a reduction to the longest path problem, \tnote[TH]{or is it a
   survey of another result?  check.} that finding a Tucker's forbidden
-submatrix of at least $k$ rows is NP complete. \tnote[GTC]{how is this
+submatrix of at least $k$ rows is NP complete. \tnote[TH]{how is this
   different from booth's 75 result??}  \tnote[TH]{where should this
@@
   as a tool to reorder $0.5T \le runs(M) \le T$. (2) The optimization
-version of the $k$-run problem, i.e. minimization of number of blocks
-of ones is proven to be NP complete by cite:k77}\tnote[GTC]{are these two
-  the same?} \tnote[TH]{what is the reduction?} \tnote[TH]{other
-  problems similar to COP -- cite:ckl96 (ILP, circ ones, one drop) --
-  cite:th98 (generalization of COP - minimax, biotonic column) Tucker}
-
-\section{Application of COP in areas of Graph theory and algorithms}
-\subsection{COP in relational database model}
+  version of the $k$-run problem, i.e. minimization of number of
+  blocks of ones is proven to be NP complete by
+  cite:k77}\tnote[TH]{are these two the same?} \tnote[TH]{what is the
+  reduction?} \tnote[TH]{other problems similar to COP -- cite:ckl96
+  (ILP, circ ones, one drop) -- cite:th98 (generalization of COP -
+  minimax, biotonic column) Tucker}
+
+\section{********* Application of COP in Areas of Graph Theory and
+  Algorithms}
+\label{sec:appcopGTA}
+
+\subsection{********* COP in Relational Database Model}
+\label{sec:apprdbm}
 \tnote{(set systems theme)}
-\subsection{COP in graph isomorphism}
+
+\subsection{********* COP in Graph Isomorphism}
+\label{sec:appgraphiso}
 \tnote{(canonization theme)}
-\subsection{Invention of certifying algorithms using COP}
+
+\subsection{********* Certifying Algorithms}
+\label{sec:appcertalgo}
 \tnote{ (certification McC04 theme)}
 
-\section{Generalization of COP - The Motivation}  %{Motivation}
+\section{Generalization of COP - the Motivation}
 \label{sec:motive}
-%\vspace{\secfirstparatrim}  
-As seen in Section~\ref{sec:copmatrices}, \cite{nsnrs09} describes a
-characterization of consecutive-ones property based on the cardinality
-properties of the set representations of the columns. This result is
-very relevant to this thesis because aside from it simplifying COT to
-a great degree, our generalization problem is motivated by their
-results.
-
-The result in
+Section~\ref{sec:copmatrices} introduced a succinct characterization
+for consecutive-ones property which is solely based on the cardinality
+properties of the set representations of the matrix's columns
+\cite{nsnrs09}. This result is very relevant to this thesis because
+aside from it simplifying COT to a great degree, our generalization
+problem is motivated by their results.
+
 \cite{nsnrs09} characterizes interval assignments to the sets which
-can be obtained from a single permutation of the rows.  They show that
-for each set, the cardinality of the interval assigned to it must be
-same as the cardinality of the set, and the intersection cardinality
-of any two sets must be same as the intersection cardinality of the
-corresponding intervals.  While this is obviously a necessary
-condition, this result shows this is also sufficient.  \cite{nsnrs09}
-calls such an interval assignment an Intersection Cardinality
+can be obtained from a single permutation of the rows.  For an
+assignment to be feasible, the cardinality of the interval assigned to
+each set in the system must be same as the cardinality of the set, and
+the intersection cardinality of any two intervals must be same as the
+intersection cardinality of their corresponding sets.  While this is
+obviously a necessary condition, this result shows this is also
+sufficient.  \cite{nsnrs09} calls this an Intersection Cardinality
 Preserving Interval Assignment (ICPIA).  This paper generalizes the
@@
 
-The tree path labeling problem is a natural generalization of the
-interval assignment problem or the COP problem. The problem is defined
-as follows -- given a set system $\cF$ from a universe $U$ and a tree
-$T$, does there exist a bijection from $U$ to the vertices of $T$ such
-that each set in the system maps to a path in $T$.  We refer to this
-as the {\em tree path labeling problem} for an input set system,
-target tree pair -- $(\cF,T)$. As a special case if the tree $T$ is a
-path, the problem becomes the interval assignment problem.  We focus
-on the question of generalizing the notion of an ICPIA \cite{nsnrs09}
-to characterize feasible path assignments.  We show that for a given
-set system $\cF$, a tree $T$, and an assignment of paths from $T$ to
-the sets, there is a feasible bijection between $U$ and $V(T)$ if and
-only if all intersection cardinalities among any three sets (not
-necessarily distinct) is same as the intersection cardinality of the
-paths assigned to them and the input passes a filtering algorithm
-(described in this paper) successfully.  This characterization gives a
-natural data structure that stores all the relevant feasible
-bijections between $U$ and $V(T)$. This reduces the search space for
-the solution considerably from the universe of all possible bijections
-between $U$ and $V(T)$ to only those bijections that maintain the
-characterization\tnote[TH]{REWORD}.  Further, the filtering algorithm
-is also an efficient algorithm to test if a tree path labeling to the
-set system is
-feasible.  %This generalizes the result in \cite{nsnrs09}.
-
-% How the problem of COT is related to the constraint statisfaction
-% problem of interval assignment to sets in a system can be seen as
-% follows. Every column of the binary matrix can be converted into a set
-% of integers which are the indices of rows with {\un}s in that column. It
-% is apparent that if the matrix has COP, then constructing such sets
-% after applying the COP permutation to the matrix will result in all
-% sets with consecutive integers. In other words, after application of
-% COP reordering, the sets are intervals. Indeed the problem now becomes
-% finding interval assignments to a given set system such that there
-% exists a permutation of the universe (set of row indices) which
-% converts each set to its assigned interval.
-
-% \def \probdefwidth {0.85\linewidth}
-% \def \FTPL {{\sc Feasible Tree Path Labeling}}
-% \def \CFTPL {\sc Compute Feasible Path Labeling} 
-\section{Summary of results}
+The equivalence of the problem of testing for the consecutive-ones
+property to the constraint statisfaction problem of interval
+assignment \cite{nsnrs09} or interval labeling \cite{kklv10} is as
+follows. Every column (row) of the binary matrix can be converted into
+a set of non-negative integers which are the indices of rows (columns)
+with {\un}s in that column (row). It is apparent that if the matrix
+has COP in columns (rows), then constructing such sets after applying
+the COP permutation to the rows (columns) of the matrix will result in
+sets with consecutive integers. In other words, after application of
+COP reordering, the sets are intervals. Indeed the problem now becomes
+finding interval assignments to a given set system such that there
+exists a permutation of the universe of set of row indices (column
+indices) which converts each set to its assigned interval.
+
+The problem of interest in this thesis, namely, tree path labeling
+problem, is a natural generalization of the interval assignment
+problem or the COT problem. The problem is defined as follows -- given
+a set system $\cF$ from a universe $U$ and a target tree $T$, does
+there exist a bijection from $U$ to the vertices of $T$ such that each
+set in the system maps to a path in $T$.  We refer to this as the
+{\CFTPL} problem or simply {\em tree path labeling} problem for an
+input set system and target tree pair -- $(\cF,T)$. The special case
+of the target tree being a path, is the interval assignment problem.
+We focus on generalizing the notion of an ICPIA \cite{nsnrs09} to
+characterize feasible path assignments.  We show that for a given set
+system $\cF$, a tree $T$, and an assignment of paths from $T$ to the
+sets, there is a feasible\footnote{The notion of {\em feasibility} is
+  formally defined in Section~\ref{ch:prelims}.}  bijection between
+$U$ and $V(T)$ if and only if the intersection cardinalities among any
+three sets (not necessarily distinct) is equal to that of the
+corresponding paths assigned to them and the input passes a filtering
+algorithm (described in this paper) successfully.  This algorithmic
+characterization gives a natural data structure that stores all the
+\remove{relevant} feasible bijections between $U$ and $V(T)$. This
+reduces the search space for the solution considerably from the
+universe of all possible bijections between $U$ and $V(T)$ to only
+those bijections that maintain the characterization.  Further, the
+filtering algorithm is also an efficient algorithm to test if a tree
+path labeling\footnote{The terms {\em tree path labeling} and {\em
+    tree path assignment} are, in informal language,
+  synonyms. Formally, the former refers to the bijection $\cl: \cF
+  \rightarrow \cP$. The latter refers to the set of ordered pairs
+  $\{(S, P) \mid S \in \cF, P \in \cP\}$. $\cP$ is a set of paths on
+  $T$.} is feasible.
+
+\section{Summary of New Results in this Thesis}
 \label{sec:results}
-%\vspace{\secfirstparatrim}
-% We address the computation of TPL\footnote{Tree Path Labeling} for a
-% given hypergraph\footnote{The term {\em hypergraph} is used instead
-%   {\em set systems} are used interchangably in this
-%   section. Hypergraphs that have TPLs on some target tree are called
-%   {\em path hypergraphs}.} and target tree in this research. However
-% optimization problems in TPL, akin to problems in
-% Section~\ref{sec:optcop} for COP, is outside the scope of this thesis.
-As we saw earlier by the result in \cite{nsnrs09}, pairwise
-intersection cardinality preservation is necessary and sufficient for
-an interval assignment to be feasible for a given
-hypergraph\footnote{A {\em hypergraph} is an alternate representation
-  of a set system and will be used through out this section.} and thus
-is a characterization for COP. In our work we extend this
-characterization and find that trio-wise intersection cardinality
-preservation makes a tree path labeling \footnote{A {\em tree path
-    labeling} $\cl$ is a bijection of paths from the target tree $T$
-  to the hyperedges in given hypergraph $\cF$.} (TPL) feasible, which
-is a generalization of the COP problem. This problem is defined as
-follows.
-
+
+We see in Section~\ref{sec:motive} that pairwise intersection
+cardinality preservation is necessary and sufficient for an interval
+assignment to be feasible for a given hypergraph\footnote{A {\em
+    hypergraph} is an alternate representation of a set system and
+  will be used in this thesis. See Section~\ref{ch:prelims} for the
+  formal definition.} and thus is a characterization for COP
+\cite{nsnrs09}. \tnote{Refer to the survey section that will list the
+  theorems and/or lemma?} In our work we extend this characterization
+and find that trio-wise intersection cardinality preservation makes a
+tree path labeling \footnote{A {\em tree path labeling} $\cl$ is a
+  bijection of paths from the target tree $T$ to the hyperedges in
+  given hypergraph $\cF$.  See Section~\ref{ch:prelims} for the formal
+  definition.}  (TPL) feasible, which is a generalization of the COP
+problem. This problem is defined as follows.
+
+{\small
+  \begin{minipage}[h]{5in}
+    % \begin{singlespace}
+    \vspace{2mm}
+    {\large \FTPL}\\
+    \begin{tabular}[t]{l|l}
+      \hline\\
+      {\tt Input} & 
+      \begin{minipage}[t]{\probdefwidth}
+        A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
+        paths $\cP$ from $T$ and a
+        bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
+      \end{minipage}\\
+      {\tt Question} &
+      \begin{minipage}[t]{\probdefwidth}
+        Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
+        such that $\phi$ when applied on any hyperedge in $\cF$ will
+        give
+        the path mapped to it by the given tree path labeling $\cl$.\\
+        { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every
+          hyperedge $S \in \cF$.}
+      \end{minipage}\\
+    \end{tabular}
+    % \end{singlespace}
+  \end{minipage}\\
+}
+
+We give a necessary and sufficient condition by way of {\em
+  Intersection Cardinality Preservation Path Labeling} (ICPPL) and a
+filtering algorithm for {\FTPL} to output in affirmative. ICPPL
+captures the trio-wise cardinality property described
+earlier\footnote{See Section~\ref{sec:feasible} for the definition of
+  ICPPL.}. This characterization can be checked in polynomial time.  A
+relevant consequence of this constructive procedure is that it is
+sufficient to iteratively check if three-way intersection
+cardinalities are preserved.  In other words, in each iteration, it is
+sufficient to check if the intersection of any three hyperedges is of
+the same cardinality as the intersection of the corresponding paths.
+Thus this generalizes the well studied question of the feasible
+interval assignment problem which is the special case when the target
+tree $T$ is simply a path \cite{wlh02,nsnrs09}.
+
+Aside from checking if a given TPL is feasible, we also solve the
+problem of computing a feasible TPL for a given hypergraph and target
+tree, if one exists. This problem, {\CFTPL}, is defined as follows.
+
+{\small
+  \begin{minipage}[h]{5in}
+    % \begin{singlespace}
+    \vspace{2mm}
+    {\large \CFTPL}\\
+    \begin{tabular}[t]{l|l}
+      \hline\\
+      {\tt Input} & 
+      \begin{minipage}[t]{\probdefwidth}
+        A hypergraph $\cF$ with vertex set $U$ and a tree $T$.\\
+      \end{minipage}\\
+
+      {\tt Question} &
+      \begin{minipage}[t]{\probdefwidth}
+        Does there exist a set of paths $\cP$ from $T$ and a bijection
+        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
+        {\bf true} on $(\cF, T, \cl)$.
+      \end{minipage}\\
+    \end{tabular}
+    % \end{singlespace}
+  \end{minipage}\\
+}
+
+We present a polynomial time algorithm for {\CFTPL} when the target
+tree $T$ belongs to a special class of trees called {\em \kstar s} and
+when the hyperedges in the hypergraph $\cF$ have at most $k+2$
+vertices. A couple of examples of {\kstar s} are given in
+Figure~\ref{fig:ksubstar}.
+
+\begin{figure}[t]
+  \centering
+  \begin{tabular}[h]{ccccc}
+    \includegraphics[scale=0.3]{../img/star.pdf} &&&&
+    \includegraphics[scale=0.3]{../img/kstar.pdf}\\
+    (a) &&&& (b)
+  \end{tabular}
+  \caption{\figtabsize Examples of {\kstar s}. (a) $k = 0$ (b) $k = 2$
+  }
+  \label{fig:ksubstar}
+\end{figure}
+
+{\small
+  \begin{minipage}[h]{5in}
+    % \begin{singlespace}
+    \vspace{2mm}
+    {\large \CFTPLKTREE}\\
+    \begin{tabular}[t]{l|l}
+      \hline\\
+      {\tt Input} & 
+      \begin{minipage}[t]{\probdefwidth}
+        A hypergraph $\cF$ with vertex set $U$ such that every
+        hyperedge
+        $S \in \cF$ is of cardinality at most $k+2$ and a {\kstar} $T$.\\
+      \end{minipage}\\
+      {\tt Question} &
+      \begin{minipage}[t]{\probdefwidth}
+        Does there exist a set of paths $\cP$ from $T$ and a bijection
+        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
+        {\bf true} on $(\cF, T, \cl)$.
+      \end{minipage}\\
+    \end{tabular}
+    % \end{singlespace}
+  \end{minipage}\\
+}
+
+\tnote{The following sentence seems out of place in this para} In
+spite of this being a restricted case, we believe that our results are
+of significant interest in understanding the nature of {\sc Graph
+  Isomorphism} which is polynomial time solvable in interval graphs
+while being hard on path graphs\cite{kklv10}. {\kstar s} are a class
+of trees which are in many ways very close to intervals or paths. Each
+ray\footnote{The path from a leaf to the root, the vertex with highest
+  degree, is called a {\em ray} of the \kstar. See
+  Section~\ref{ch:prelims}.} are independent except for the
+root\footnote{The vertex with maximum degree in a {\kstar} is called
+  {\em root}. See Section~\ref{ch:prelims}.} and hence can be
+considered as an independent interval till the root. Our algorithm
+builds on this fact and uses the interval assignment
+algorithm\cite{nsnrs09} up until ``reaching'' the root and then uses
+the trio-wise intersection cardinality (the extra condition in ICPPL
+that generalizes ICPIA) check to resolve the ambiguity about which ray
+the algorithm should ``grow'' the solution into in the next iteration.
+
+We also have an algorithm for solving {\CFTPL} with no restrictions on
+the target tree or set size which runs in exponential time.  This
+algorithm finds a path labeling from $T$ by decomposing the problem
+into subproblems of finding path labeling of subsets of $\cF$ from
+subtrees of $T$. Given the fact that binary matrices naturally
+represent a set system (see Section~\ref{sec:motive}) and that the
+{\em overlap}\footnote{}\tnote{Give informal definition of ``overlap''
+  in footnote and link to sec:prelims. ADD the definition in prelim.}
+relation between the sets involved is an obvious equivalence relation,
+$\cF$ quite naturally partitions into equivalence classes known as
+{\em overlap components}\tnote{Give informal definition of ``overlap
+  components'' in footnote and link to sec:prelims. ADD the definition
+  in prelim.}. In the context of COP, overlap components were used in
+\cite{wlh02} and \cite{kklv10}. Moreover, \cite{nsnrs09} discovered
+that these equivalence classes form a total order \tnote{Give informal
+  definition of ``total order'' in footnote?}. We extend this to TPL
+and find that when $\cF$ is a path hypergraph\footnote{If there exists
+  an FTPL for a hypergraph $\cF$, it is called a path hypergraph.},
+the classes can be partially ordered as an in-tree in polynomial
+time. Once $\cF$ is ``broken'' into overlap components, one must
+identify the subtree of $T$ that it needs to map to and this is the
+hard part which is currently open to be solved in polynomial time.
+
+\tnote[TH]{The connection of TPL to graph isomorphism will be made
+  later in the document}
+
+\chapter{Consecutive-ones Property - a Survey of Important Results}
+\label{ch:copsurvey}
+\tnote{TBD: have a few lines about organization of chapter}
+
+\tnote{ Survey chapter: is the full fledged expansion of the survey in
+  introduction with details, observations, theorems etc.}
+
+\section{Matrices with COP}
+\label{sec:surveycoptest}
+\tnote{Expand on ref:sec:copmatrices}
+
+\section{Optimization problems in COP}
+\label{sec:surveycopopt}
+\tnote{Expand on ref:sec:optcop}
+
+\section{********* COP in Relational Database Model}
+\label{sec:surveyrdbm}
+\tnote{Expand on sec:apprdbm}
+\tnote{(set systems theme)}
+
+\subsection{********* COP in Graph Isomorphism}
+\label{sec:surveygraphiso}
+\tnote{Expand on sec:appgraphiso}
+\tnote{(canonization theme)}
+
+\subsection{********* Certifying Algorithms}
+\label{sec:surveycertalgo}
+\tnote{Expand on sec:appcertalgo}
+\tnote{ (certification McC04 theme)}
+
+\chapter{Tree Path Labeling of Path Hypergraphs - the New Results}
+\label{ch:myresearch}
+
+\tnote{UPDATE TO FINAL VERSION. BELOW IS OLDER VERSION WITH INCOMPLETE
+  SEC 4.}  This chapter documents all the new results obtained by us in the
+area of tree path labeling of path hypergraphs which is the parent
+problem addressed in this thesis. In Section~\ref{sec:motive} we see
+that consecutive-ones property and its equivalent problem of interval
+labeling of a hypergraph is a special case of the general problem of tree
+path labeling of path hypergraphs.
+
+\tnote{UPDATE the following para to current skeleton.}The necessary
+preliminaries with definitions etc. are presented in
+Section~\ref{ch:prelims}. Section~\ref{sec:feasible} documents the
+characterization of a feasible path labeling for any path labeling
+from any target trees. Section~\ref{sec:spltargettree} describes the
+special case where the target tree is of a particular family of trees
+and specifically, Section~\ref{sec:ksubdivstar} describes a polynomial
+time algorithm to find the tree path labeling of a given set system
+from a given $k$-subdivided tree. Section~\ref{sec:norestraint}
+describes the general situation where the target tree has no
+restrictions and the algorithm to find a TPL, if any, in this case.
+
+\section{Introduction}
+
+The problem of consecutive-ones property testing can be easily seen as
+a simple constraint satisfaction problem involving a hypergraph or a
+system of sets from a universe. Every column (row) of the binary
+matrix can be converted into a set of non-negative integers which are
+the indices of the rows (columns) with {\un}s in that column
+(row). When observed in this context, if the matrix has the COP on
+columns (rows), a reordering of its rows (columns) will result in sets
+that have only consecutive integers. In other words, the sets after
+applying the COP row (column) permutation are intervals.  In this
+form, one can see that this is indeed the problem of finding interval
+assignments to the given set system \cite{nsnrs09} with a single
+permutation of the universe (set of row or column indices for COP of
+columns or rows, respectively) which permutes each set to its
+interval. The result in \cite{nsnrs09} characterizes interval
+assignments to the sets which can be obtained from a single
+permutation of the rows (columns). They show that for each set, the
+cardinality of the interval assigned to it must be same as the
+cardinality of the set, and the intersection cardinality of any two
+sets must be same as the intersection cardinality of the corresponding
+intervals. This is a necessary and sufficient
+condition. \tnote{(ICPIA) put reference to survey chapter}.  Finally,
+the idea of decomposing a given binary matrix into prime matrices to
+check for COP is adopted from cite|wlh02| to test if an ICPIA exists
+for a given set system.\tnote{REPEATITION from intro chapter. but idea
+  recall is necessary. so keep it but reword it to avoid monotone.}
+
+A natural generalization of the interval assignment problem is
+feasible tree path labeling problem of a set system. The problem is
+defined as follows -- given a set system $\cF$ from a universe $U$ and
+a tree $T$, does there exist a bijection from $U$ to the vertices of
+$T$ such that each set in the system maps to a path in $T$.  We refer
+to this as the {\em tree path labeling problem} for an input set
+system, target tree pair -- $(\cF,T)$. \tnote{USE THE FORMAL PROBLEM DEF BLOCK}
 %\begin{tabular}[t]{c}
@@
 
-We give a necessary and sufficient condition (called ICPPL)\tnote[GTC]{say
-  a few words about icppl} for {\FTPL} to output in affirmative. This
-characterization can be checked in polynomial time.  The most
-interesting consequence is that in our constructive procedure, it is
-sufficient to iteratively check if three-way intersection
-cardinalities are preserved.  In other words, in each iteration, it is
-sufficient to check if the intersection of any three hyperedges is of
-the same cardinality as the intersection of the corresponding paths.
-Thus this generalizes the well studied question of the feasible
-interval assignment problem which is the special case when the target
-tree $T$ is simply a path \cite{wlh02,nsnrs09}.
-
-\tnote{A few lines on the algo itself.}
-
-Aside from checking if a given TPL is feasible, we also solve the
-problem of computing a feasible TPL for a given hypergraph and target
-tree, if one exists. This problem, {\CFTPL}, is defined as follows.
-
-{\small
-\begin{minipage}[h]{5in}
- % \begin{singlespace}
- \vspace{2mm}
-  {\large \CFTPL}\\
-  % 
-  \begin{tabular}[t]{l|l}
-    \hline\\
-    {\tt Input} & 
-    \begin{minipage}[t]{\probdefwidth}
-      A hypergraph $\cF$ with vertex set $U$ and a tree $T$.\\
-    \end{minipage}\\
-
-    {\tt Question} &
-    \begin{minipage}[t]{\probdefwidth}
-      Does there exist a set of paths $\cP$ from $T$ and a bijection
-      $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns {\bf
-        true} on $(\cF, T, \cl)$.
-    \end{minipage}\\
-  \end{tabular}
-  % \end{singlespace}
-\end{minipage}\\
-}
-
-% \def \kstar {$k$-subdivided star}
-% \def \CFTPLKTREE {\sc Compute $k$-subdivided Star Path Labeling}
-
-We present a polynomial time algorithm for {\CFTPL} when the target
-tree $T$ belongs to a special class of trees called {\em \kstar s} and
-when the hyperedges in the hypergraph $\cF$ have at most $k+2$
-vertices. A couple of examples of {\kstar s} are given in
-Figure~\ref{fig:ksubstar}.
-
-%This problem, {\CFTPLKTREE}, is defined as follows. \\
-\begin{figure}[t]
-  \centering
-  \begin{tabular}[h]{ccccc}
-  \includegraphics[scale=0.3]{../img/star.pdf} &&&&
-  \includegraphics[scale=0.3]{../img/kstar.pdf}\\
-  (a) &&&& (b)
-  \end{tabular}
-  \caption{\figtabsize Examples of {\kstar s}. (a) $k = 0$
-    (b) $k = 2$ }
-  \label{fig:ksubstar}
-\end{figure}
-
-%\vspace{3mm}
-{\small
-\begin{minipage}[h]{5in}
- % \begin{singlespace}
- \vspace{2mm}
-  {\large \CFTPLKTREE}\\
-  \begin{tabular}[t]{l|l}
-    \hline\\
-    {\tt Input} & 
-    \begin{minipage}[t]{\probdefwidth}
-      A hypergraph $\cF$ with vertex set $U$ such that every hyperedge
-      $S \in \cF$ is of cardinality at most $k+2$ and a {\kstar} $T$.\\
-    \end{minipage}\\
-    {\tt Question} &
-    \begin{minipage}[t]{\probdefwidth}
-      Does there exist a set of paths $\cP$ from $T$ and a bijection
-      $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns {\bf
-        true} on $(\cF, T, \cl)$.
-    \end{minipage}\\
-  \end{tabular}
-  % \end{singlespace}
-\end{minipage}\\
-}
-
-In spite of this being a restricted case, we believe that our results
-are of significant interest in understanding the nature of {\sc Graph
-  Isomorphism} which is polynomial time solvable in interval graphs
-while being hard on path graphs\cite{kklv10}. {\kstar s} are a class
-of trees which are in many ways very close to intervals or paths. Each
-ray\footnote{The path from a leaf to the root, the vertex with highest
-degree, is called a {\em ray} of the \kstar.} are independent except
-for the root\footnote{The vertex with maximum degree in a {\kstar} is
-  called {\em root}} and hence can be considered as an independent interval
-till the root. Our algorithm builds on this fact and uses the interval
-assignment algorithm\cite{nsnrs09} up until ``reaching'' the root and
-then uses the trio-wise intersection cardinality (the extra condition
-in ICPPL after ICPIA) check to resolve the
-ambiguity about which ray the algorithm should ``grow'' the solution
-into in the next iteration.
-
-We also have an algorithm for solving {\CFTPL} with no restrictions
-which runs in exponential time. 
-% and explores a partial order theory of
-% matrices related to {FTPL}.
-% gives interesting insight
-% into the problem of 
-This algorithm finds a path labeling from $T$ by decomposing the
-problem into smaller subproblems of finding path labeling of subsets
-of $\cF$ from subtrees of $T$. Given the fact that binary matrices
-naturally represent a set
-system % , we look at the matrix representation of subsets of $\cF$.
-and that the overlap relation between the sets involved is an obvious
-equivalence relation, $\cF$ naturally partitions into equivalence
-classes known as overlap components. In the context of COP, {\em overlap
-components} were used in \cite{wlh02} and \cite{kklv10}. Moreover,
-\cite{nsnrs09} discovered that these equivalence classes form a total
-order. We extend this to TPL and find that when $\cF$ is a path
-hypergraph\footnote{If there exists an FTPL for a hypergraph $\cF$, it
-  is called a path hypergraph.}, the classes can be partially ordered
-as an in-tree in polynomial time. Once $\cF$ is
-``broken'' into overlap components, one must identify the subtree of
-$T$ that it needs to map to and this is the hard part which is
-currently open to be solved in polynomial time.
-
-\tnote[TH]{The connection of TPL to graph
-  isomorphism will be made later in the document}
-
-% In the later part of this paper, we decompose our search for a
-% bijection between $U$ and $V(T)$ into subproblems.  Each subproblem
-% is on a set subsystem in which for each set, there is another set in
-% the set subsystem with which the intersection is {\em strict}, i.e.,
-% there is a non-empty intersection, but neither is contained in the
-% other.  This is in the spirit of results in \cite{wlh02,nsnrs09}
-% where to test for the COP in a given matrix, the COP problem is
-% solved on an equivalent set of prime matrices.
-
-%\subsection{RESULTS}
-% \section{Preliminaries} 
-% \label{sec:prelims}
-
-% \subsection{Hypergraph} 
-% The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
-%   system} of a universe $U$ with $|U| = n$.  The {\em support} of a
-% set system $\F$ denoted by $supp(\cF)$ is the union of all the sets in
-% $\F$; $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this
-% paper, a set system is required to ``cover'' the universe; $ supp(\cF)
-% = U$.
-% % A set system $\cF$ can be alternatively represented by a {\em
-% %   hypergraph} $\H_\cF$ whose vertex set is $supp(\cF)$ and
-% % hyperedges are the sets in $\cF$. This is a known representation for
-% % interval systems in literature \cite{bls99,kklv10}.  We extend this
-% % definition here to path systems.
-% A set system $\cF$ can also be visualized as a {\em
-%   hypergraph}\, vertex set is $supp(\cF)$ and hyperedges are the sets
-% in $\cF$.  The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph
-% $\cF$ is a graph such that its vertex set has a bijection to $\cF$ and
-% there exists an edge between two vertices \iff their corresponding
-% hyperedges have a non-empty intersection \cite{mcg04}.  Two
-% hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to each
-% other, denoted by $\cF' \cong \cF''$, \iff there exists a bijection
-% $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for all sets $A
-% \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ \iff  $B$ is a
-% hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
-% \cite{kklv10}, written as $B=\phi(A)$.
-% % This is called {\em hypergraph isomorphism}.
-
-% %\noindent
-% \subsection{Path Hypergraph from a Tree} 
-% We consider trees $T$ such that $|V(T)|=|U|=n$.  A {\em path system}\,
-% $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P \mid P
-% \subseteq V, \text{ } T[P] \text{ is a path} \}$.  This generalizes
-% the fact, from the literature \cite{bls99,kklv10}, that intervals can
-% be viewed as sub-paths of a path.
-% % Due to the equivalence of set system and hypergraph in the scope of
-% % this paper, we drop the subscript $_H$ in the notation and refer to
-% % both the structures by $\cF$.
-% % \vspace{\prelimspace}%\tnote{also extend $\phi$ to hyperedges -- see if
-% % required } \vspace{\prelimspace} \vspace{\prelimspace}
-% % \noindent
-% If the intersection graphs of $\cF$ and $\cP$( a path system) are
-% isomorphic, $\bI(\cF) \cong \bI(\cP)$, then the associated bijection
-% $\cl: \cF \rightarrow \cP$ due to this isomorphism is called a {\em
-%   path labeling} of the hypergraph $\cF$.  Note that there are two
-% kinds of isomorphisms here.  We are concerned about the isomrophisms
-% intersection graphs on $\cF$ and $\cP$, and also the isomorphism
-% between the hypergraph $\cF$ and $\cP$.
-% % To illustrate further, let $\cg: V(\cF) \rightarrow V(\cP)$ be the
-% % above mentioned isomorphism where $V(\cF)$ and $V(\cP)$ are the
-% % vertex sets that represent the hyperedges for each hypergraph
-% % respectively, $V(\cF) = \{ v_S \mid S \in \cF\}$ and $V(\cP) = \{
-% % v_P \mid P \in \cP\}$. Then the path labeling $\cl$ is defined as
-% % follows: $\cl(S_1) = P_1$ \iff $\cg (v_{S_1}) = v_{P_1}$.  Just to
-% % emphasize, for a path labeling $\cl$ of $\cF$ with $\cP$ as the path
-% % system, $\cF^\cl$ is same as $\cP$.  The path system $\cP$ may be
-% % alternatively denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In
-% % most scenarios in this paper, what is given are the pair $(\cF,
-% % \cl)$ and the target tree $T$; hence this notation will be used more
-% % often.
-% If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$ is called
-% a {\em path hypergraph} and $\cP$ is called {\em path representation}
-% of $\cF$. If this isomorphism is $\phi: supp(\cF) \rightarrow V(T)$,
-% then it is clear that there is a path labeling $\cl_\phi: \cF
-% \rightarrow \cP$ to the set system; $\cl_\phi(S) = \set{y \mid y =
-%   \phi(x), x \in S}$ for all $S \in \cF$. In other words, if $\cF
-% \cong \cP$, we get a path labeling.  Recall that $supp(\cP) = V(T)$.
-% % \vspace{\prelimspace} A graph $G$ is a {\em path graph} if it is
-% % isomorphic to the intersection graph $\bI(\cP)$ of a path system
-% % $\cP$.  This isomorphism gives a bijection $\cl': V(G) \rightarrow
-% % \cP$. Moreover, for the purposes of this paper, we require that in a
-% % path labeling, $supp(\cP) = V(T)$.  If graph $G$ is also isomorphic
-% % to $\bI(\cF)$ for some hypergraph $\cF$, then clearly there is a
-% % bijection $\cl: \cF \rightarrow \cP$ such that $\cl(S) = \cl'(v_S)$
-% % where $v_S$ is the vertex corresponding to set $S$ in $\bI(\cF)$ for
-% % any $S \in \cF$. This bijection $\cl$ is called the {\em path
-% %   labeling} of the hypergraph $\cF$ and the path system $\cP$ may be
-% % alternatively denoted as $\cF^\cl$.  \vspace{\prelimspace}
-% In this work, we are given as input $\cF$ and a tree $T$, and the
-% question is whether there is a path labeling $\cl$ to a set of paths
-% in $T$.  We refer to such a solution path system by $\cF^\cl$.  A path
-% labeling $(\cF, \cl)$ is defined to be {\em feasible} if
-% % $\cF \cong \cF^\cl$ and this
-% there is a hypergraph isomorphism $\phi: supp(\cF) \rightarrow
-% supp(\cF^\cl)=V(T)$ induces a path labeling $\cl_\phi: \cF \rightarrow
-% \cF^\cl$ such that $\cl_\phi = \cl$.
-% % \vspace{\prelimspace}
-
-% \subsection{Overlap Graphs and Marginal Hyperedges} 
-% An {\em
-%   overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is a graph such
-% that its vertex set has a bijection to $\cF$ and there exists an edge
-% between two of its vertices \iff their corresponding hyperedges
-% overlap. Two hyperedges $S$ and $S'$ are said to {\em overlap},
-% denoted by $S \overlap S'$, if they have a non-empty intersection and
-% neither is contained in the other; $S \overlap S' \text{ iff } S \cap
-% S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq S$. Thus $\bO(\cF)$
-% is a spanning subgraph of $\bI(\cF)$ and not necessarily
-% connected. Each connected component of $\bO(\cF)$ is called an {\em
-%   overlap component}.
-% % If there are $d$ overlap components in $\bO(\cF)$, the set
-% % subsystems are denoted by $\cO_1, \cO_2, \ldots \cO_d$. Clearly
-% % $\cO_i \subseteq \F, i \in [d]$. For any $i, j \in [d]$, it can be
-% % verified that one of the following is true.
-% % \begin{enumerate}[a) ]
-% % \item $supp(\cO_i)$ and $supp(\cO_j)$ are disjoint
-% % \item $supp(\cO_i)$ is a subset of a set in $\cO_j$
-% % \item $supp(\cO_j)$ is a subset of a set in $\cO_i$
-% % \end{enumerate}
-% % \vspace{\prelimspace} 
-% A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
-% \overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
-% \cite{kklv10}. Additionally, if $S$ is such that it is contained in no
-% other marginal hyperedge in $\cF$, then it is called {\em
-%   super-marginal}.
-% % i.e., it is inclusion maximal then it is called {\em
-% %   super-marginal}.
-
-% \subsection{$k$-subdivided Star -- A Special Tree} 
-% A {\em star} graph
-% is a complete bipartite graph $K_{1,p}$ which is clearly a tree and
-% $p$ is the number of leaves. The vertex with maximum degree is called
-% the {\em center} of the star and the edges are called {\em rays} of
-% the star. A {\em $k$-subdivided star} is a star with all its rays
-% subdivided exactly $k$ times. The definition of a {\em ray of a
-%   $k$-subdivided star} is extended to the path from the center to a
-% leaf. It is clear that all rays are of length $k+2$. See
-% fig.~\ref{fig:kstar} for examples.
-
-% \begin{figure}[htbp] 
-%   \centering
-%   (i) \includegraphics{cop-tree_thesis.1} \ \ \ \ \ \  
-%   (ii)  \includegraphics{cop-tree_thesis.2}
-%   \caption{(i) $8$-subdivided star with 7 rays, (ii) 3-subdivided
-%     star with 3 rays}
-%   \label{fig:kstar}
-% \end{figure}
-
-% \subsection{Characterization of Feasible Tree Path  Labelings} \tnote[TH]{UNCOMMENT all proofs}
-% \label{sec:feasible} 
-% \vspace{\secfirstparatrim}
-
-% In this section we give an algorithmic characterization of a
-% feasibility of tree path labeling.
-% % Consider a path labeling $\cl: \cF \rightarrow \cP$ for set system $\cF$
-% % and path system $\cP$ on the given tree $T$. We
-% % call $\cl$ an {\em Intersection Cardinality Preserving Path Labeling
-% %   (ICPPL)} if it has the following properties.
-% Consider a path labeling $(\cF, \cl)$ on the
-% given tree $T$. We call $(\cF, \cl)$ an {\em Intersection Cardinality
-%   Preserving Path Labeling (ICPPL)} if it has the following
-% properties.
-
-% % \begin{enumerate}[{\icpplpr\ }i. ]
-% % \item \label{pr:i} $|S| = |\cl(S)|$, \hfill for all $S \in \cF$
-% %   % \vspace{\topshrink}
-% % \item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$, \hfill
-% %   for all distinct $S_1, S_2 \in \cF$
-% %   % \vspace{\topshrink}
-% % \item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
-% %   \cap \cl(S_3)|$, \hfill for all distinct $S_1, S_2, S_3 \in \cF$
-% % \end{enumerate}
-
-% \begin{table}[h]
-%   \centering
-%   \begin{tabular}[l]{rrcll}
-%     Pr i. & \label{xpr:i} $|S| $&$=$&$ |\cl(S)|$ & $\forall$ $S \in \cF$\\
-%     Pr ii. & \label{xpr:ii}$|S_1 \cap S_2| $&$=$&$ |\cl(S_1) \cap
-%     \cl(S_2)|$ &  $\forall$ distinct $S_1, S_2 \in \cF$\\
-%     Pr iii. & \label{xpr:iii}$|S_1 \cap S_2 \cap S_3| $&$=$&$
-%     |\cl(S_1) \cap \cl(S_2) \cap \cl(S_3)|$ & $\forall$ distinct $S_1,
-%     S_2, S_3 \in \cF$ 
-%   \end{tabular}
-%   \caption{ICPPL properties}
-%   \label{tab:icpplpr}
-% \end{table}
-
-% \noindent
-% The following lemma is useful in subsequent arguments. 
-% \begin{lemma}
-%   \label{lem:setminuscard}
-%   If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
-%   (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
-%   \cl(S_3))|$.
-% \end{lemma}
-
-% % \begin{proof}%[Proof of Lemma~\ref{lem:setminuscard}]
-% %   Let $P_i = \cl(S_i)$, for all $i \in \set{1,2,3}$. We have $|S_1
-% %   \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| = |S_1
-% %   \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to properties
-% %   (\ref{pr:ii}) and (\ref{pr:iii}) of ICPPL, $|S_1 \cap S_2| - |S_1
-% %   \cap S_2 \cap S_3| = |P_1 \cap P_2| - |P_1 \cap P_2 \cap P_3| =
-% %   |(P_1 \cap P_2) \setminus P_3| = |P_1 \cap (P_2 \setminus
-% %   P_3)|$. Thus lemma is proved. \qed
-% % \end{proof}
-
-% In the remaining part of this section we show that $(\cF, \cl)$ is
-% feasible if and only if it is an ICPPL and
-% Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
-% function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two
-% levels of filtering of $(\cF, \cl)$ to make it simpler while retaining
-% the set of isomorphisms, if any, between $\cF$ and $\cF^\cl$.
-% % One direction of this claim isclear: that if a path
-% % labeling is feasible, then all intersection cardinalities are
-% % preserved, i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
-% % \annote{has no premature exit condition hence any input will go
-% %   through it}{Prove that the filtered sets has ICPPL \iff input PL
-% %   has ICPPL?}. Algorithm~\ref{leafasgn} has an exit condition at
-% % line~\ref{xempty}. It can be easily verified that $X$ cannot be
-% % empty if $\cl$ is a feasible path labeling. The reason is that a
-% % feasible path labeling has an associated bijection between
-% % $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that
-% % the sets map to paths, ``preserving'' the path labeling.  The rest
-% % of the section is devoted to constructively proving that it is
-% % sufficient for a path labeling to be an ICPPL and pass the two
-% % filtering algorithms.  To describe in brief, the constructive
-% % approaches refine an ICPPL iteratively, such that at the end of each
-% % iteration we have a ``filtered'' path labeling, and finally we have
-% % a path labeling that defines a family of bijections from $supp(\cF)$
-% % to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.
-
-% First, we present Algorithm~\ref{perms} called {\tt filter\_1}
-% % , and
-% % prove its correctness.  
-% This algorithm refines the path labeling by
-% processing pairs of paths in $\cF^\cl$ that share a leaf until no two
-% paths in the new path labeling share any leaf.
-% \tnote{write a note summarizing filter 1}
-% % \begin{algorithm}[h]
-% %   \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
-% %   \label{perms}
-% %   \begin{algorithmic}[\lndisplay]
-% %     \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$\\
-% %     \STATE $j \assign 1$\\
-% %     \WHILE {there is $S_1, S_2 \in \cF_{j-1}$ such that
-% %       $\cl_{j-1}(S_1)$ and $\cl_{j-1}(S_2)$ have a common leaf in
-% %       $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
-% %     \{S_1, S_2\})
-% %     \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1 \}$ \label{setbreak} 
-% %     \COMMENT {Remove $S_1$, $S_2$ and add the ``filtered'' sets}
-% %     \STATE {\bf for} every $S \in \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
-% %     S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\
-
-% %     \STATE $\cl_j(S_1 \cap S_2) \assign \cl_{j-1}(S_1) \cap
-% %     \cl_{j-1}(S_2)$
-% %     \COMMENT {Carry forward the path labeling for all existing sets other than
-% %       $S_1$, $S_2$}
-% %     \STATE $\cl_j(S_1 \setminus S_2) \assign \cl_{j-1}(S_1) \setminus
-% %     \cl_{j-1}(S_2)$ 
-% %     \COMMENT {Define path labeling for new sets}
-% %     \STATE $\cl_j(S_2 \setminus S_1) \assign \cl_{j-1}(S_2) \setminus
-% %     \cl_{j-1}(S_1)$
-
-% %     \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of ICPPL}
-% %     \label{ln:3waycheck}
-% %     \STATE {\bf exit} \label{ln:exit1} \\
-% %     \ENDIF
-
-% %     \STATE $j \assign j+1$\\
-% %     \ENDWHILE
-% %     \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
-% %     \RETURN $(\cF', \cl')$
-% %   \end{algorithmic}
-% % \end{algorithm}
-
-% \begin{lemma} 
-%  \label{lem:feasible} 
-%  In Algorithm~\ref{perms}, if input $(\cF, \cl)$ is a feasible path
-%  assignment then at the end of $j$th iteration of the {\bf while}
-%  loop, $j \ge 0$, $(\cF_j, \cl_j)$ is a feasible path assignment.
-% \end{lemma}
-
-% % \begin{proof}%[Proof of Lemma~\ref{lem:feasible}]
-% %   We will prove this by mathematical induction on the number of
-% %   iterations. The base case $(\cF_0, \cl_0)$ is feasible since it is
-% %   the input itself which is given to be feasible. Assume the lemma is
-% %   true till $j-1$th iteration. i.e. every hypergraph isomorphism
-% %   $\phi: supp\left(\cF_{j-1}\right) \rightarrow V\left(T \right)$ that
-% %   defines $(\cF, \cl)$'s feasibility, is such that the induced path
-% %   labeling on $\cF_{j-1}$, $\cl_{\phi[{\cF_{j-1}}]}$ is equal to
-% %   $\cl_{j-1}$. We will prove that $\phi$ is also the bijection that
-% %   makes $(\cF_j, \cl_j)$ feasible. Note that $supp(\cF_{j-1}) =
-% %   supp(\cF_{j})$ since the new sets in $\cF_j$ are created from basic
-% %   set operations to the sets in $\cF_{j-1}$. For the same reason and
-% %   $\phi$ being a bijection, it is clear that when applying the $\phi$
-% %   induced path labeling on $\cF_j$, $ \cl_{\phi[{\cF_{j}}]}(S_1
-% %   \setminus S_2) = \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
-% %   \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Now observe that $ \cl_j(S_1
-% %   \setminus S_2) = \cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2) =
-% %   \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
-% %   \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Thus the induced path labeling
-% %   $\cl_{\phi[{\cF_{j}}]} = \cl_{j}$. Therefore lemma is proved.  \qed
-% % \end{proof}
-
-% \begin{lemma}
-%   \label{lem:invar1} In Algorithm~\ref{perms}, at the end of $j$th
-%   iteration, $j \ge 0$, of the {\bf while} loop, the following
-%   invariants are maintained.
-% %   \begin{enumerate}[I {\ }] %\vspace{\topshrink}
-% %   \item $\cl_j(R)$ is a path in $T$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-% %     \ \ \ \ \ \ \ \ \ \ for all $R \in \cF_j$%\vspace{\topshrink}
-% %   \item $|R| = |\cl_j(R)|$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-% %     \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $R \in
-% %     \cF_j$%\vspace{\topshrink}
-% %   \item $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, \ \ \ \ \ \ \ \ \ \
-% %     \ \ \ \ \ \ \ \ \ \ for all $R, R' \in \cF_j$%\vspace{\topshrink}
-% %   \item $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
-% %     \cl_j(R'')|$, \ \ \ for all $R, R', R'' \in \cF_j$
-% %   \end{enumerate}
-
-%   \begin{tabular}[h]{lll}
-%     I & $\cl_j(R)$ is a path in $T$, & for all $R \in \cF_j$\\ 
-%     II  & $|R| = |\cl_j(R)|$, & for all $R \in \cF_j$ \\ 
-%     III  & $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, & for all $R, R'
-%     \in \cF_j$\\ 
-%     IV  & $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
-%     \cl_j(R'')|$, & for all $R, R', R'' \in \cF_j$ 
-%   \end{tabular}
-% \end{lemma}
-
-% % \begin{proof}
-% %  Proof is by induction on the number of iterations, $j$. In this
-% %   proof, the term ``new sets'' will refer to the sets added to $\cF_j$
-% %   in $j$th iteration in line~\ref{setbreak} of Algorithm~\ref{perms},
-% %   $S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1$ and its
-% %   images in $\cl_j$ where $\cl_{j-1}(S_1)$
-% %   and $\cl_{j-1}(S_2)$ intersect and share a leaf.\\
-% %   The invariants are true in the base case $(\cF_0, \cl_0)$, since it
-% %   is the input ICPPL.  Assume the lemma is true till the $j-1$th
-% %   iteration. Let us consider the possible cases for each of the above invariants for
-% %   the $j$th iteration.
-
-% %   \begin{enumerate}[\textreferencemark]
-% %   \item {\em Invariant} I/II
-% %     \begin{enumerate}[{I/II}a $|$] % \textbullet 
-% %     \item {\em $R$ is not a new set.} It is in $\cF_{j-1}$. Thus
-% %       trivially true by induction hypothesis.
-% %     \item {\em $R$ is a new set.} If $R$ is in $\cF_{j}$ and not in
-% %       $\cF_{j-1}$, then it must be one of the new sets added in
-% %       $\cF_j$. In this case, it is clear that for each new set, the
-% %       image under $\cl_j$ is a path since by definition the chosen
-% %       sets $S_1$, $S_2$ are from $\cF_{j-1}$ and due to the while loop
-% %       condition, $\cl_{j-1}(S_1)$, $\cl_{j-1}(S_2)$ have a
-% %       common leaf. Thus invariant I is proved.\\
-% %       Moreover, due to induction hypothesis of invariant III and the
-% %       definition of $l_j$ in terms of $l_{j-1}$, invariant II is
-% %       indeed true in the $j$th iteration for any of the new sets.  If
-% %       $R = S_1 \cap S_2$, $|R| = |S_1 \cap S_2| = |\cl_{j-1}(S_1) \cap
-% %       \cl_{j-1}(S_2)| = |\cl_j(S_1 \cap S_2)| = |\cl_j(R)|$.
-% %       If $R = S_1 \setminus S_2$, $|R| = |S_1 \setminus S_2| = |S_1| -
-% %       |S_1 \cap S_2| = |\cl_{j-1}(S_1)| - |\cl_{j-1}(S_1) \cap
-% %       \cl_{j-1}(S_2)| = |\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)| =
-% %       |\cl_j(S_1 \setminus S_2)|
-% %       = |\cl_j(R)|$. Similarly if $R = S_2 \setminus S_1$.\\
-% %     \end{enumerate}
-% %   \item {\em Invariant} III
-% %     \begin{enumerate}[{III}a $|$]
-% %     \item {\em $R$ and $R'$ are not new sets.} It is in
-% %       $\cF_{j-1}$. Thus trivially true by induction hypothesis.
-% %     \item {\em Only one, say $R$, is a new set.} Due to invariant IV
-% %       induction hypothesis, Lemma~\ref{lem:setminuscard} and
-% %       definition of $\cl_j$, it follows that invariant III is true no
-% %       matter which of the new sets $R$ is equal to. If $R = S_1 \cap
-% %       S_2$, $|R \cap R'| = |S_1 \cap S_2 \cap R'| = |\cl_{j-1}(S_1)
-% %       \cap \cl_{j-1}(S_2) \cap \cl_{j-1}(R')| = |\cl_j(S_1 \cap S_2)
-% %       \cap \cl_j(R')| = |\cl_j(R) \cap \cl_j(R')|$.  If $R = S_1
-% %       \setminus S_2$, $|R \cap R'| = |(S_1 \setminus S_2) \cap R'| =
-% %       |(\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)) \cap \cl_{j-1}(R')| =
-% %       |\cl_{j}(S_1 \cap S_2) \cap \cl_{j}(R')| = |\cl_{j}(R) \cap
-% %       \cl_{j}(R')|$. Similarly, if $R = S_2 \setminus
-% %       S_1$. Note $R'$ is not a new set.\\
-
-% %     \item {\em $R$ and $R'$ are new sets.} By definition, the new
-% %       sets and their path images in path label $\cl_j$ are disjoint so
-% %       $|R \cap R'| = |\cl_j(R) \cap \cl_j(R)| = 0$. Thus case proved.
-% %     \end{enumerate}
-% %   \item {\em Invariant} IV
-
-% %     Due to the condition in line~\ref{ln:3waycheck}, this invariant is
-% %     ensured at the end of every iteration.
-% % %     \begin{enumerate} [{Case 3.}1:]
-% % %     \item {\em $R$, $R'$ and $R''$ are not new sets.} Trivially
-% % %       true by induction hypothesis.
-% % %     \item {\em Only one, say $R$, is a new set.}
-% % %       If $R = S_1 \cap S_2$,  from Lemma~\ref{lem:fourpaths} and
-% % %       invariant III hypothesis,  this case is proven. Similarly if $R$
-% % %       is any of the other new  sets, the case is proven by also using
-% % %       Lemma ~\ref{lem:setminuscard}.
-% % %     \item {\em At least two of $R, R', R''$ are new sets.}
-% % %       The new sets are disjoint hence this case is vacuously true.
-% % %     \end{enumerate}
-% %   \end{enumerate} \qed
-% % %\vspace{-6mm}
-
-% % \end{proof}
-
-% \begin{lemma}
-%   \label{lem:noexit1}
-%   If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{perms} is
-%   feasible, then the set of hypergraph isomorphism functions that
-%   defines $(\cF, \cl)$'s feasibility is the same as the set that
-%   defines $(\cF_j, \cl_j)$'s feasibility, if any.  Secondly, for any
-%   iteration $j > 0$ of the {\em \bf while} loop, the {\em \bf exit}
-%   statement in line~\ref{ln:exit1} will not execute.
-% \end{lemma}
-% \tnote[TH]{UNCOMMENT}
-% % \begin{proof}
-% %   Since $(\cF,\cl)$ is feasible, by Lemma~\ref{lem:feasible}
-% %   $(\cF_j,\cl_j)$ for every iteration $j > 0$ is feasible.  % Therefore,
-% % %   every hypergraph isomorphism $\phi: supp(\cF) \rightarrow V(T)$ that
-% % %   induces $\cl$ on $\cF$ also induces $\cl_{j-1}$ and $\cl_{j}$ on
-% % %   $\cF_{j-1}$ and $\cF_{j}$ respectively, i.e., $\cl_{\phi[\cF_{j-1}]}
-% % %   = \cl_{j-1}$ and $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
-% % %   that for all $x \in supp(\cF)$, for all $v \in V(T)$ the following
-% % %   hold true.
-% %   Also, every hypergraph isomorphism $\phi: supp(\cF) \rightarrow
-% %   V(T)$ that induces $\cl$ on $\cF$ also induces $\cl_{j}$ on
-% %   $\cF_{j}$, i.e., $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
-% %   that for all $x \in supp(\cF)$, for all $v \in V(T)$, if $(x,v) \in
-% %   \phi$ then $v \in \cl_{j}(S)$ for all $S \in \cF_{j}$ such that $x
-% %   \in S$.
-% % % the following
-% % %   hold true.
-% % %   \begin{enumerate}[i. ]
-% % %   \item If $(x,v) \in \phi$ then $v \in \cl_{j-1}(S)$ for all $S \in
-% % %     \cF_{j-1}$ such that $x \in S$.
-% % %   \item If $(x,v) \in \phi$ then $v \in \cl_{j}(S)$ for all $S \in
-% % %     \cF_{j}$ such that $x \in S$
-% % %   \end{enumerate}
-% %   In other words, filter 1 outputs a filtered path labeling that
-% %   ``preserves''
-% %   hypergraph isomorphisms of the original path labeling.\\
-% %   Secondly, line~\ref{ln:exit1} will execute \iff the exit condition in
-% %   line~\ref{ln:3waycheck}, i.e. failure of three way intersection
-% %   preservation, becomes true in any iteration of the {\em \bf while}
-% %   loop.  Due to Lemma~\ref{lem:invar1} Invariant IV, the exit
-% %   condition does not occur if the input is a feasible ICPPL.\qed
-
-% % %   such that $\phi(x) = v$ where $v$ is the leaf considered in the
-% % %   first iterations of while. Clearly, $\phi$ is a renaming of
-% % %   vertices in hypergraph $\cF$ to those in hypergraph $\cF^\cl$. Thus
-% % %   the following facts can be observed in every iteration of the loop.
-
-% % %   \begin{enumerate}[\hspace{2mm}i. ] \vspace{\topshrink}
-% % %   \item all intersection cardinalities are preserved in this path
-% % %     labeling \vspace{\topshrink}
-% % %   \item element $x$ is exclusive in a hyperedge in $\cF$ since $v$ is
-% % %     exclusive in a hyperedge in $\cF^\cl$.
-% % %   \end{enumerate}
-
-% % %   Thus the exit condition is never rendered true after $x$ and $v$ are
-% % %   removed from their respective hyperedges. \qed
-
-% % % \noindent
-% % % This proof uses mathematical induction on the number
-% % %   of iterations $j$, $j \ge 0$, of the loop that executed
-% % %   without exiting. The base case, $j = 0$ is obviously true since the
-% % %   input is an ICPPL and the exit condition cannot hold true due to
-% % %   ICPPL property (iii).  Assume the algorithm executes till the end
-% % %   of $j-1$th iteration without exiting at line
-% % %  ~\ref{ln:3waycheck}. Consider the $j$th iteration. From Lemma
-% % %  ~\ref{lem:feasible} we know that $(\cF_j, \cl_j)$ and $(\cF_{j-1},
-% % %   \cl_{j-1})$ are feasible\remove[AS]{and from the proof in lemma
-% % %     lem:invar1 we know that $(\cF_{j-1}, \cl_{j-1})$ satisfies all the
-% % %     invariants defined in the lemma}.  Thus there exists a bijection
-% % %   $\phi: supp(\cF) \rightarrow V(T)$ such that the induced path
-% % %   % labeling on $\cF_{j-1}$ $\cl_{\phi[\cF_{j-1}]} = \cl_{j-1}$.
-% % %   labeling on $\cF_{j}$, $\cl_{\phi[\cF_{j}]}$ and on $\cF_{j-1}$,
-% % %   $\cl_{\phi[\cF_{j-1}]}$ are equal to $\cl_{j}$ and $\cl_{j-1}$
-% % %   respectively.  We need to prove that for any $R, R', R'' \in
-% % %   \cF_{j}$, $|R \cap R' \cap R''| = |\cl_j(R) \cap \cl_j(R') \cap
-% % %   \cl_j(R'')|$.
-% % %   The following are the possible cases that could arise. From argument
-% % %   above, $|\cl_j(R) \cap \cl_j(R') \cap \cl_j(R'')| =
-% % %   |\cl_{\phi[\cF_{j}]}(R) \cap \cl_{\phi[\cF_{j}]} (R') \cap
-% % %   \cl_{\phi[\cF_{j}]} (R'')|$
-
-% % %   \begin{enumerate}[a $|$]
-% % %   \item {\em None of the sets are new. $R, R', R'' \in \cF_{j-1}$.}
-% % %     We know $(\cF_{j-1}, \cl_{j-1})$ is feasible. Thus $|R \cap R'
-% % %     \cap R''| = |\cl_{j-1}(R) \cap \cl_{j-1}(R') \cap \cl_{j-1}(R'')|
-% % %     = |\cl_{j}(R) \cap \cl_{j}(R') \cap \cl_{j}(R'')|$.
-% % %   \item {\em Only one, say $R$, is a new set.}  Let $R = S_1 \cap S_2$
-% % %     ($S_1, S_2$ are defined in the proof of lemma
-% % %    ~\ref{lem:invar1}). Now we have $|R \cap R' \cap R''| = |S_1 \cap
-% % %     S_2 \cap R' \cap R''| = |\cl_{j-1}(S_1) \cap \cl_{j-1}(S_2) \cap
-% % %     \cl_{j-1}(R') \cap \cl_{j-1}(R'')| = |\cl_{j}(R) \cap \cl_{j}(R')
-% % %     \cap \cl_{j}(R'')|$. Thus proven. If $R$ is any of the other new
-% % %     sets, the same claim can be verified using lemma
-% % %    ~\ref{lem:setminuscard}.
-% % %     % \item []{\bf Case 3:}
-% % %   \item {\em At least two of $R, R', R''$ are new sets.}  The new sets
-% % %     are disjoint hence this case is vacuously true.
-% % %   \end{enumerate}
-% % %   \qed
-% % %   \tnote[E2]{remove the induction proof. just text saying x and v are
-% % %     exclusive in these sets therefore the intersection cardinalities
-% % %     don't change thus all invariants are still true} 
-% % \end{proof}
-
-% As a result of Algorithm~\ref{perms} each leaf $v$ in $T$
-% is such that there is exactly one set in $\cF$ with $v$ as a vertex in
-% the path assigned to it.  In Algorithm~\ref{leafasgn} we identify
-% elements in $supp(\cF)$ whose images are leaves in a hypergraph
-% isomorphism if one exists.  Let $S \in \cF$ be such that $\cl(S)$ is a
-% path with leaf and $v \in V(T)$ is the unique leaf incident on it.  We
-% define a new path labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) =
-% \set{v}$ where $x$ an arbitrary element from $S \setminus \bigcup_{\hS
-%   \ne S} \hS$. In other words, $x$ is an element present in no other
-% set in $\cF$ except $S$. This is intuitive since $v$ is present in no
-% other path image under $\cl$ other than $\cl(S)$.  The element $x$ and
-% leaf $v$ are then removed from the set $S$ and path $\cl(S)$
-% respectively. After doing this for all leaves in $T$, all path images
-% in the new path labeling $\cl_{new}$ except leaf labels (a path that
-% has only a leaf is called the {\em leaf label} for the corresponding
-% single element hyperedge or set) are paths from a new pruned tree $T_0
-% = T \setminus \{v \mid v \text{ is a leaf in }
-% T\}$. Algorithm~\ref{leafasgn} is now presented with details.
-
-% \tnote{write a note summarizing filter 1}
-% % \begin{algorithm}[h]
-% %   \caption{Leaf labeling from an ICPPL {\tt filter\_2($\cF, \cl, T$)}}
-% %   \label{leafasgn}
-% %   \begin{algorithmic}[\lndisplay]
-% %     \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$
-% %     \COMMENT {Path images are such that no two path images share a
-% %       leaf.}
-% %     \STATE $j \assign 1$\\
-% %     \WHILE {there is a leaf $v$ in $T$ and a unique $S_1 \in
-% %       \cF_{j-1}$ such that $v \in \cl_{j-1}(S_1)$ }\label{uniqueleaf}
-% %     \STATE $\cF_j \assign \cF_{j-1} \setminus \{S_1\}$\\
-% %     \STATE for all $S \in \cF_{j-1}$ such that $S \ne S_1$ set
-% %     $\cl_j(S) \assign
-% %     \cl_{j-1}(S)$\\
-% %     \STATE $X \assign S_1 \setminus \bigcup_{S \in \cF_{j-1}, S \ne S_1}S$\\
-% %     \IF{$X$ is empty} \label{xempty} \STATE {\bf exit} \label{ln:exit2} \ENDIF
-% %     \STATE $x \assign $ arbitrary element from $X$\\
-% %     \STATE $\cF_j \assign \cF_j \cup \{\{x\}, S_1 \setminus \{x\}\} $\\
-% %     \STATE $\cl_j(\{x\}) \assign \{v\}$\\
-% %     \STATE $\cl_j(S_1 \setminus \{x\}) \assign \cl_{j-1}(S_1) \setminus \{v\}$\\
-% %     \STATE $j \assign j+1$\\
-% %     \ENDWHILE
-% %     \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
-% %     \RETURN $(\cF', \cl')$
-% %   \end{algorithmic}
-% % \end{algorithm}
-
-% Suppose the input ICPPL $(\cF, \cl)$ is feasible, yet set $X$ in
-% Algorithm~\ref{leafasgn} is empty in some iteration of the {\bf while}
-% loop. This will abort our procedure of finding the hypergraph
-% isomorphism. The following lemma shows that this will not happen.
-
-% \begin{lemma}
-%   \label{lem:xnotempty}
-%   If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{leafasgn} is
-%   feasible, then for all iterations $j > 0$ of the {\em \bf while}
-%   loop, the {\em \bf exit} statement in line~\ref{ln:exit2} does not
-%   execute.
-% \end{lemma}
-
-% % \begin{proof}
-% %   Assume $X$ is empty for some iteration $j > 0$. We know that $v$ is
-% %   an element of $\cl_{j-1}(S_1)$. Since it is uniquely present in
-% %   $\cl_{j-1}(S_1)$, it is clear that $v \in \cl_{j-1}(S_1) \setminus
-% %   \bigcup_{(S \in \cF_{j-1}) \wedge (S \ne S_1)}\cl_{j-1}(S)$.  Note
-% %   that for any $x \in S_1$ it is contained in at least two sets due to
-% %   our assumption about cardinality of $X$. Let $S_2 \in \cF_{j-1}$ be
-% %   another set that contains $x$. From the above argument, we know $v
-% %   \notin \cl_{j-1}(S_2)$. Therefore there cannot exist a hypergraph
-% %   isomorphism bijection that maps elements in $S_2$ to those in
-% %   $\cl_{j-1}(S_2)$. This contradicts our assumption that the input is
-% %   feasible. Thus $X$ cannot be empty if input is ICPPL and feasible.
-% %   \qed
-% % \end{proof}
-
-% \begin{lemma}
-%   \label{lem:invar3}
-%   In Algorithm~\ref{leafasgn}, for all $j > 0$, at the end of the
-%   $j$th iteration of the {\bf while} loop the four invariants given in
-%   Lemma~\ref{lem:invar1} hold.
-% \end{lemma}
-
-% % \begin{proof}
-% %   By Lemma~\ref{lem:xnotempty} we know that set $X$ will not be empty
-% %   in any iteration of the {\em \bf while} loop if input ICPPL $(\cF,
-% %   \cl)$ is feasible and $\cl_j$ is always computed for all $j >
-% %   0$. Also note that removing a leaf from any path keeps the new path
-% %   connected. Thus invariant I is obviously true. In every iteration $j
-% %   > 0$, we remove exactly one element $x$ from one set $S$ in $\cF$
-% %   and exactly one vertex $v$ which is a leaf from one path
-% %   $\cl_{j-1}(S)$ in $T$. This is because as seen in
-% %   Lemma~\ref{lem:xnotempty}, $x$ is exclusive to $S$ and $v$ is
-% %   exclusive to $\cl_{j-1}(S)$. Due to this fact, it is clear that the
-% %   intersection cardinality equations do not change, i.e., invariants
-% %   II, III, IV remain true. On the other hand, if the input ICPPL is
-% %   not feasible the invariants are vacuously true. \qed
-% % \end{proof}
-
-% %% OLD TEXT TB DELETED?
-% % \textcolor{cyan}{
-% % \begin{lemma}
-% %   \label{lem:notfeasibleexit}
-% %   \tnote{IS THIS CORRECT?}  If input ICPPL $(\cF, \cl)$ is not
-% %   feasible, then in one of the recursive calls to Algorithm --3--, the
-% %   {\em \bf exit} statement in line x in Algorithm --1 or line y in
-% %   Algorithm --2 will get executed.
-% % \end{lemma}
-% % }
-% %% END OLD TEXT TB DELETED?
-
-% We have seen two filtering algorithms above, namely,
-% Algorithm~\ref{perms} {\tt filter\_1} and Algorithm~\ref{leafasgn}
-% {\tt filter\_2} which when executed serially respectively result in a
-% new ICPPL on the same universe $U$ and tree $T$. We also proved that
-% if the input is indeed feasible, these algorithms do indeed output the
-% filtered ICPPL. Now we present the algorithmic characterization of a
-% feasible tree path labeling by way of Algorithm~\ref{al:icppl-find-isomorph}.
-
-% Algorithm~\ref{al:icppl-find-isomorph} computes a
-% hypergraph isomorphism $\phi$ recursively using Algorithm~\ref{perms}
-% and Algorithm~\ref{leafasgn} and pruning the leaves of the input
-% tree. In brief, it is done as follows. Algorithm~\ref{leafasgn} gives
-% us the leaf labels in $\cF_2$, i.e., the elements in $supp(\cF)$ that
-% map to leaves in $T$, where $(\cF_2, \cl_2)$ is the output of
-% Algorithm~\ref{leafasgn}. All leaves in $T$ are then pruned away. The
-% leaf labels are removed from the path labeling $\cl_2$ and the
-% corresponding elements are removed from the corresponding sets in
-% $\cF_2$. This tree pruning algorithm is recursively called on the
-% altered hypergraph $\cF'$, path label $\cl'$ and tree $T'$. The
-% recursive call returns the bijection $\phi''$ for the rest of the
-% elements in $supp(\cF)$ which along with the leaf labels $\phi'$ gives
-% us the hypergraph isomorphism $\phi$.  The following lemma formalizes
-% the characeterization of feasible path labeling.
-
-% \tnote{write a note summarizing filter 1}
-% % \begin{algorithm}[h]
-% %   \caption{{\tt get-hypergraph-isomorphism($\cF, \cl, T$)}}
-% %   \label{al:icppl-find-isomorph}
-% %   \begin{algorithmic}[\lndisplay]
-
-% %     \IF{$T$ is empty}
-% %     \RETURN $\emptyset$\\
-% %     \ENDIF
-% %     \STATE $L \assign \{v \mid v \text{ is a leaf in }      T\}$\\
-% %     \STATE $(\cF_1, \cl_1) \assign$ {\tt filter\_1($\cF, \cl,
-% %       T$)}\\
-% %     \STATE $(\cF_2, \cl_2) \assign$ {\tt filter\_2($\cF_1,
-% %       \cl_1, T$)}\\
-
-% %     \STATE $(\cF', \cl') \assign (\cF_2, \cl_2)$\\
-% %     \STATE $\phi' \leftarrow \emptyset$
-
-% %     \FOR {every $v \in L$}
-% %     \STATE $\phi'(x) \assign v$ where $x \in \cl_2^{-1}(\{v\})$
-% %     \COMMENT {Copy the leaf labels to a one to one function $\phi':
-% %       supp(\cF) \rightarrow L$
-% %       }\\
-% %     \STATE Remove $\{x\}$ and $\{v\}$ from $\cF'$, $\cl'$  appropriately\\
-% %     \ENDFOR
-
-% %     \STATE $T' \assign T \setminus L$
-
-% %     \STATE $\phi'' \assign$ {\tt
-% %       get-hypergraph-isomorphism($\cF', \cl', T'$)}
-% %     \STATE $\phi \assign \phi'' \cup \phi'$ \\
-% %     \RETURN $\phi$
-% %   \end{algorithmic}
-% % \end{algorithm}
-
-% \begin{lemma}
-%   \label{lem:hyperiso}  %{lem:perm}
-%   If $(\cF, \cl)$ is an ICPPL from a tree $T$ and
-%   Algorithm~\ref{al:icppl-find-isomorph}, {\tt
-%     get- hypergraph- isomorphism ($\cF, \cl, T$)} returns a non-empty
-%   function, then there exists a hypergraph isomorphism $\phi :
-%   supp(\cF) \rightarrow V(T)$ such that the $\phi$-induced tree path
-%   labeling is equal to $\cl$ or $\cl_\phi = \cl$.
-% \end{lemma}
-% % \begin{proof}
-% %   It is clear that in the end of every recursive call to
-% %   Algorithm~\ref{al:icppl-find-isomorph}, the function $\phi'$ is one
-% %   to one involving all the leaves in the tree passed as input to that
-% %   recursive call. Moreover, by Lemma~\ref{lem:noexit1} and
-% %   Lemma~\ref{lem:xnotempty} it is consistent with the tree path
-% %   labeling $\cl$ passed. The tree pruning is done by only removing
-% %   leaves in each call to the function and is done till the tree
-% %   becomes empty. Thus the returned function $\phi: supp(\cF)
-% %   \rightarrow V(T)$ is a union of mutually exclusive one to one
-% %   functions exhausting all vertices of the tree. In other words, it is
-% %   a bijection from $supp(\cF)$ to $V(T)$ inducing the given path
-% %   labeling $\cl$ and thus a hypergraph isomorphism. \qed
-% % \end{proof}
-
-% \begin{theorem}
-%   \label{th:charac}
-%   A path labeling $(\cF, \cl)$ on tree $T$ is feasible \iff it is an
-%   ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
-%   T)$ as input returns a non-empty function.
-% \end{theorem}
-% % \begin{proof}
-% %   From Lemma~\ref{lem:hyperiso}, we know that if $(\cF, \cl)$ is an
-% %   ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
-% %   T)$ as input returns a non-empty function, $(\cF, \cl)$ is feasible.
-% %   Now consider the case where $(\cF, \cl)$ is feasible. i.e. there
-% %   exists a hypergraph isomorphism $\phi$ such that $\cl_\phi =
-% %   \cl$. Lemma~\ref{lem:noexit1} and Lemma~\ref{lem:xnotempty} show us
-% %   that filter 1 and filter 2 do not exit if input is feasible. Thus
-% %   Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
-% %   function.\qed
-% % \end{proof}
-
-% \subsection*{ICPPL when given tree is a path}
-% \label{sec:icpplicpia}
-% \vspace{\secfirstparatrim}  
-% Consider a special case of ICPPL with the following properties when
-% the tree $T$ is a path.  Hence, all path labels are can be viewed as
-% intervals assigned to the sets in $\cF$.  It is shown, in
-% \cite{nsnrs09}, that the filtering algorithms outlined above need only
-% preserve pairwise intersection cardinalities, and higher level
-% intersection cardinalities are preserved by the Helly Property of
-% intervals.  Consequently, the filter algorithms do not need to ever
-% evaluate the additional check to {\em \bf exit}.
-% %\begin{enumerate}
-% %\item Given tree $T$ is a path. Hence, all path labels are interval labels.
-% %\item Only pairwise intersection cardinality
-% %  preservation is sufficient. i.e. property (iii) in ICPPL is not enforced.
-% %\item The filter algorithms do not have {\em \bf exit} statements.
-% %\end{enumerate}
-% %This is called an Intersection Cardinality Preservation Interval
-% %Assignment (ICPIA) \cite{nsnrs09}. 
-% This structure and its algorithm is
-% used in the next section for finding tree path labeling from a
-% $k$-subdivided star due to this graph's close relationship with
-% intervals.
-
-% \subsection{Case of a Special Tree -- The $k$-subdivided star}
-% \label{sec:ksubdivstar}
-% \vspace{\secfirstparatrim}  
-% In this section we consider the problem of assigning paths from a
-% $k$-subdivided star $T$ to a given set system $\cF$ such that each set
-% $X \in \cF$ is of cardinality at most $k+2$.  Secondly, we present our
-% results only for the case when overlap graph $\bO(\cF)$ is connected.
-% %The
-% %overlap graph is well-known from the work of
-% %\cite{kklv10,nsnrs09,wlh02}.  We use the notation in
-% %\cite{kklv10}. Recall from Section~\ref{sec:prelims} that hyperedges
-% %$S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$ and $S'$
-% %have a non-empty intersection but neither of them is contained in the
-% %other. The overlap graph $\bO(\cF)$ is a graph in which the vertices
-% %correspond to the sets in $\cF$, and the vertices corresponding to the
-% %hyperedges $S$ and $S'$ are adjacent if and only if they overlap.  Note
-% %that the intersection graph of $\cF$, $\bI(\cF)$ is different from
-% %$\bO(\cF)$ and $\bO(\cF) \subseteq \bI(\cF)$.  
-% A connected component of $\bO(\cF)$ is called an overlap component of
-% $\cF$.  An interesting property of the overlap components is that any
-% two distinct overlap components, say $\cO_1$ and $\cO_2$, are such
-% that any two sets $S_1 \in \cO_1$ and $S_2 \in \cO_2$ are disjoint,
-% or, w.l.o.g, all the sets in $\cO_1$ are contained within one set in
-% $\cO_2$.  This containment relation naturally determines a
-% decomposition of the overlap components into rooted containment trees.
-% We consider the case when there is only one rooted containment tree,
-% and we first present our algorithm when $\bO(\cF)$ is connected.  It
-% is easy to see that once the path labeling to the overlap component in
-% the root of the containment tree is achieved, the path labeling to the
-% other overlap components in the rooted containment tree is essentially
-% finding a path labeling when the target tree is a path: each target
-% path is a path that is allocated to sets in the root overlap
-% component.  Therefore, for the rest of this section, $\bO(\cF)$ is a
-% connected graph. Recall that we also consider the special case when
-% all hyperedges are of cardinality at most $k+2$.  By definition, a
-% $k$-subdivided star has a central vertex which we call the {\em root},
-% and each root to leaf path is called a {\em ray}.  First, we observe
-% that by removing the root $r$ from $T$, we get a collection of $p$
-% vertex disjoint paths of length $k+1$, $p$ being the number of leaves
-% in $T$.  We denote the rays by $R_1, \ldots, R_p$, and the number of
-% vertices in $R_i$, $i \in [p]$ is $k+2$.  Let
-% $\seq{v_{i1},\ldots,v_{i(k+2)}=r}$ denote the sequence of vertices in
-% $R_i$, where $v_{i1}$ is the leaf. Note that $r$ is a common vertex to
-% all $R_i$.
-
-% \subsubsection{Description of the Algorithm}
-% \vspace{\secfirstparatrim}  
-% In this section the given hypergraph $\cF$, the $k$-subdivided star
-% and the root of the star are denoted by $\cO$, $T$ and vertex $r$,
-% respectively.  In particular, note that the vertices of $\cO$
-% correspond to the sets in $\cF$, and the edges correspond to the
-% overlap relation.
-
-% %\noindent
-% For each hyperedge $X \in \cO$, we will maintain a 2-tuple of non-negative
-%  numbers $\seq{p_1(X), p_2(X)}$.  The numbers satisfy the property that
-%  $p_1(X) + p_2(X) \leq |X|$, and at the end of path labeling, for each
-%  $X$, $p_1(X) + p_2(X) = |X|$.  This signifies the algorithm tracking
-%  the lengths of subpaths of the path assigned to $X$ from at most two
-%  rays. We also maintain another parameter called the {\em residue} of
-%  $X$ denoted by $s(X)=|X| - p_1(X)$. This signifies the residue path
-%  length that must be assigned to $X$ which must be from another
-%  ray. For instance, if $X$ is labeled a path from only one ray, then
-%  $p_1(X) = |X|$, $p_2(X) = 0$ and $s(X) = 0$.
-
-%  The algorithm proceeds in iterations, and in the $i$-th
-%  iteration, $i > 1$, a single hyperedge $X$ that overlaps with a
-%  hyperedge that has been assigned a path is considered.  At the
-%  beginning of each iteration hyperedges of $\cO$ are classifed into
-%  the following disjoint sets.
-%  \begin{enumerate}
-%  \item [$\cL_1^i$] {\em Labeled without $r$.} Those that have been
-%    labeled with a path which does not contain $r$ in one of the
-%    previous iterations.\\  $\cL_1^i = \set{ X \mid p_1(X) = |X| \text{ and
-%      } p_2(X) = 0 \text{ and } s(X) = 0, X \in \cO}$
-%  \item [$\cL_2^i$] {\em Labeled with $r$.} Those that have been labeled
-%    with two subpaths of $\cl(X)$ containing $r$ from two different rays
-%    in two previous iterations.\\ $\cL_2^i = \set{X \mid 0 < p_1\left(X\right),
-%      p_2\left(X\right) < |X|=p_1(X)+p_2(X) \text{ and } s(X) = 0, X \in \cO}$
-%    \item [$\cT_1^i$] {\em Type 1 / partially labeled.} Those that have
-%      been labeled with one path containing $r$ from a single ray in one
-%      of the previous iterations. Here, $p_1(X)$ denotes the length of
-%      the subpath of $\cl(X)$ that $X$ has been so far labeled
-%      with. \\
-%      $\cT_1^i = \set{ X \mid 0 < p_1(X) < |X| \text{ and } p_2(X) = 0
-%        \text{ and } s(X) = |X|-p_1(X), X \in \cO}$
-%    \item [$\cT_2^i$] {\em Type 2 / not labeled.} Those that have not been
-%      labeled with a path in any previous iteration.\\
-%      $\cT_2^i = \set{ X \mid p_1(X) = p_2(X) = 0 \text{ and } s(X) = |X|,
-%        X \in \cO}$
-%  \end{enumerate}
-% % \vspace{-2mm}
-% % \begin{align*}
-% %   \cO &= \cL_1^i \cup \cL_2^i \cup \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]\\
-% %   \cO_i &= \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]
-% % \end{align*}
-
-% %\noindent
-% The set $\cO_i$ refers to the set of hyperedges $\cT_1^i \cup \cT_2^i$
-% in the $i$th iteration.  Note that $\cO_1 = \cO$.  In the $i$th
-% iteration, hyperedges from $\cO_i$ are assigned paths from $T$ using
-% the following rules. Also the end of the iteration, $\cL_1^{i+1},
-% \cL_2^{i+1}, \cT_1^{i+1}, \cT_2^{i+1}$ are set to $\cL_1^{i},
-% \cL_2^{i}, \cT_1^{i}, \cT_2^{i}$ respectively, along with some
-% case-specific changes mentioned in the rules below.
-
-% %\noindent
-% \begin{enumerate}[I.]
-% \item {\bf Iteration 1:} Let $S=\{X_1,\ldots,X_s\}$ denote the
-%   super-marginal hyperedges from $\cO_1$.  If $|S|=s \neq p$, then
-%   exit reporting failure.  Else, assign to each $X_j \in S$, the path
-%   from $R_j$ such that the path contains the leaf in $R_j$.  This path
-%   is refered to as $\cl(X_j)$.  Set $p_1(X_j)=|X|, p_2(X_j)=s(X_j)=0$.
-%   Hyperedges in $S$ are not added to $\cO_2$ but are added to
-%   $\cL_1^2$ and all other hyperedges are added to
-%   $\cO_2$.
-% \item {\bf Iteration $i$:} Let $X$ be a hyperedge from $\cO_i$ such
-%   that there exists $Y \in \cL_1^i \cup \cL_2^i$ and $X \overlap
-%   Y$. Further let $Z \in \cL_1^i \cup L_2^i$ such that $Z \overlap Y$.
-%   If $X \in \cT_2^i$, and if there are multiple $Y$ candidates then
-%   any $Y$ is selected.  On the other hand, if $X \in \cT_1^i$, then
-%   $X$ has a partial path assignment, $\cl'(X)$ from a previous
-%   iteration, say from ray $R_j$. Then, $Y$ is
-%   chosen such that $X \cap Y$ has a non-empty intersection with a ray
-%   different from $R_j$.  The key things that are done in assigning a
-%   path to $X$ are as follows. The {\em end} of path $\cl(Y)$ where
-%   $\cl(X)$ would overlap is found, and then based on this the
-%   existence of a feasible assignment is decided.  It is important to
-%   note that since $X \overlap Y$, $\cl(X) \overlap \cl(Y)$ in any
-%   feasible assignment.  Therefore, the notion of the {\em end} at
-%   which $\cl(X)$ and $\cl(Y)$ overlap is unambiguous, since for any
-%   path, there are two end points.
-%   \begin{enumerate}
-%   \item \label{iendpoint} {\em End point of $\cl(Y)$ where $\cl(X)$ overlaps
-%       depends on $X \cap Z$:} If $X \cap Z
-%     \neq \emptyset$, then $\cl(X)$ has an overlap of $|X \cap Y|$ at that
-%     end of $\cl(Y)$ at which $\cl(Y)$ and $\cl(Z)$ overlap.  If $X
-%     \cap Z = \emptyset$, then $\cl(X)$ has an overlap of $|X \cap Y|$ at
-%     that end of $\cl(Y)$ where $\cl(Y)$ and $\cl(Z)$ do not intersect.
-%   \item {\em Any path of length $s(X)$ at the appropriate end contains
-%       $r$:} If $X \in \cT_1^i$ then after finding the appropriate end
-%     as in step~\ref{iendpoint} this the unique path of length $s(X)$
-%     should end at $r$.  If not, we exit reporting failure.  Else,
-%     $\cl(X)$ is computed as union of $\cl'(X)$ and this path. If any
-%     three-way intersection cardinality is violated with this new
-%     assignment, then exit, reporting failure.  Otherwise, $X$ is added
-%     to $\cL_2^{i+1}$.  On the other hand, if $X \in \cT_2^i$, then
-%     after step~\ref{iendpoint}, $\cl(X)$ or $\cl'(X)$ is unique up to
-%     the root and including it. Clearly, the vertices $\cl(X)$ or
-%     $\cl'(X)$ contains depends on $|X|$ and $|X \cap Y|$.  If any
-%     three way intersection cardinality is violated due to this
-%     assignment, exit, reporting failure.  Otherwise,
-%     $p_1(X)$ is updated as the length of the assigned path, and $s(X)
-%     = |X|-p_1(X)$.  If $s(X) > 0$, then $X$ is added to $\cT_1^{i+1}$.
-%     If $s(X)=0$, then $X$ is added to $\cL_1^{i+1}$.
-%   \item {\em The unique path of length $s(X)$ overlapping at the
-%       appropriate end of $Y$ does not contain $r$:} In this case,
-%     $\cl(X)$ is updated to include this path.  If any three way
-%     intersection cardinality is violated, exit, reporting failure.
-%     Otherwise, update $p_1(X)$ and $p_2(X)$ are appropriate, $X$ is
-%     added to $\cL_1^{i+1}$ or $\cL_2^{i+1}$, as appropriate.
-%   \end{enumerate}
-% \end{enumerate}
-
-% %\noindent 
-% \tnote[TH]{UNCOMMENT}
-% % \subsection*{Proof of Correctness and Analysis of Running Time} 
-% % It
-% % is clear that the algorithm runs in polynomial time, as at each step,
-% % at most three-way intersection cardinalities need to be checked.
-% % Further, finding super-marginal hyperedges can also be done in
-% % polynomial time, as it involves considering the overlap regions and
-% % checking if the inclusion partial order contains a single minimal
-% % element.  In particular, once the super-marginal edges are identified,
-% % each iteration involes finding the next hyperedge to consider, and
-% % testing for a path to that hyperedge.  To identify the next hyperedge
-% % to consider, we consider the breadth first layering of the hyperedges
-% % with the zeroeth layer consisting of the super-marginal hyperedges.
-% % Since $\cO$ is connected, it follows that all hyperedges of $\cO$ will
-% % be considered by the algorithm.  Once a hyperedge is considered, the
-% % path to be assigned to it can also be computed in constant time.  In
-% % particular, in the algorithm the path to be assigned to $X$ depends on
-% % $\cl(Y), \cl(Z)$, $s(X)$ and the presence or absence of $r$ in the
-% % candidate partial path $\cl'(X)$.  Therefore, once the super-marginal
-% % edges are identified, the running time of the algorithm is linear in
-% % the size of the input.  By the technique used for constructing prime
-% % matrices \cite{wlh02}, the super-marginal edges can be found in linear
-% % time in the input size.  Therefore, the algorithm can be implemented
-% % to run in linear time in the input size.
-
-% % %\noindent
-% % The proof of correctness uses the following main properties:
-% % \begin{enumerate}
-% % \item The $k$-subdivided star has a very symmetric structure.  This
-% %   symmetry is quantified based on the following observation -- either
-% %   there are no feasible path labelings of $\cO$ using paths from $T$,
-% %   or there are exactly $p!$ feasible path labelings.  In other words,
-% %   there is either no feasible assignment, or effectively a unique
-% %   assignment modulo symmetry.
-% % \item The $p$ super-marginal hyperedges, if they exist, will each be
-% %   assigned a path from distinct rays, and each such path contains the
-% %   leaf.
-% % \item For a candidate hyperedge $X$, the partial path assignment
-% %   $\cl'(X)$ is decided by its overlap with $\cl(Y)$ and cardinality of
-% %   intersection with $\cl(Z)$.
-% % \end{enumerate}
-% % These properties are formalized as follows:
-% % \begin{lemma}
-% %   \label{lem:sup-mar}
-% %   If $X \in \cF$ is super-marginal and $\cl$ is a feasible tree path
-% %   labeling to tree $T$, then $\cl(X)$ will contain a leaf in $T$.
-% % \end{lemma}
-% % \begin{proof}
-% %   Suppose $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
-% %   path labeling from $T$.  Assume $\cl(X)$ does not have a leaf.  Let
-% %   $R_i$ be one of the rays (or the only ray) $\cl(X)$ is part of.
-% %   Since $X$ is in a connected overlap component, there exists $Y_1 \in
-% %   \cF$ and $X \nsubseteq Y_1$ such that $Y_1 \overlap X$ and $Y_1$ has
-% %   at least one vertex closer to the leaf in $R_i$ than any vertex in
-% %   $X$. Similarly with the same argument there exists $Y_2 \in \cF$
-% %   with same subset and overlap relation with $X$ except it has has at
-% %   least one vertex farther away from the leaf in $R_i$ than any vertex
-% %   in $X$. Clearly $Y_1 \cap X$ and $Y_2 \cap X$ cannot be part of same
-% %   inclusion chain which contradicts that assumption $X$ is
-% %   super-marginal. Thus the claim is proved.\qed
-% % \end{proof}
-% % \begin{lemma}
-% %   If $\cO$ does not have any super-marginal edges, then in any
-% %   feasible path labeling $\cl$ of $\cO$ with paths from $T$ is such
-% %   that, for any hyperedge $X$ for which $\cl(X)$ contains a leaf, $|X|
-% %   \geq k+3$.
-% % \end{lemma}
-% % \begin{proof}
-% %   The proof of this lemma is by contradiction.  Let $X$ be a
-% %   hyperedges such that $|X| \leq k+2$ and that $\cl(X)$ has a leaf.
-% %   This implies that the overlap regions with $X$, which are captured by
-% %   the overlap regions with $\cl(X)$, will form a single inclusion
-% %   chain. This shows that $X$ is a marginal hyperedge which
-% %   contradicts the assumption that $\cO$ does not have super-marginal
-% %   hyperedges. \qed
-% % \end{proof}
-% % This lemma is used to prove the next lemma for the case when for all
-% % $X \in \cO$, $|X| \leq k+2$.  The proof is left out as it just uses
-% % the previous lemma and the fact that the hyperedges in $X$ have at
-% % most $k+2$ elements.
-% % \begin{lemma}
-% %   If there is a feasible path labeling for $\cO$ in $T$, then there
-% %   are exactly $p$ super-marginal hyperedges.
-% % \end{lemma}
-% % These lemmas now are used to prove the following theorem.
-% % \begin{theorem}
-% %   Given $\cO$ and a $k$-subdivided star $T$, the above algorithm
-% %   decides correctly if there is a feasible path labeling $\cl$.
-% % \end{theorem}
-% % \begin{proof} {\em Outline.}
-% % %   If the algorithm outputs a path labeling $\cl$, then it is clear
-% % %   that it is an ICPPL. The reason is that the algorithm checks that
-% % %   three-way intersection cardinalities are preserved in each iteration
-% % %   which ensures \icpplpr~\ref{pr:iii}. Moreover, it is clear that
-% % %   $\cl(X)$ for any $X \in \cO$ is computed by maintaining
-% % %   \icpplpr~\ref{pr:i} and \icpplpr~\ref{pr:ii}. For such a labeling
-% % %   $\cl$, the proof that it is feasible is by induction on $k$. What
-% % %   needs to be shown is that Algorithm~\ref{al:icppl-find-isomorph}
-% % %   successfully runs on input $(\cO, \cl)$. In base case $k=0$, $T$ is
-% % %   a star. The claim is clear by observing that after Filter 1 and one
-% % %   iteration of Filter 2, all the leaves have found their pre-images
-% % %   from support $\cO$.  Therefore, in the induction step, it is clear
-% % %   that after Filter 1 and one iteration of Filter 2, the leaves are
-% % %   assigned pre-images.  Removing the leaves from $T$ and the
-% % %   pre-images from support of $\cO$, results in an ICPPL to a
-% % %   $(k-1)$-subdivided star.  Now we apply the induction hypothesis, and
-% % %   we get a isomorphism between the hypergraph $\cO$ and $\cO^\cl$.
-% %   If the algorithm outputs a path labeling $\cl$, then it is clear
-% %   that it is an ICPPL. The reason is that the algorithm checks that
-% %   three-way intersection cardinalities are preserved in each iteration
-% %   which ensures ICPPL \icpplpr~\ref{pr:iii}. Moreover, it is clear that
-% %   $\cl(X)$ for any $X \in \cO$ is computed by maintaining
-% %   ICPPL \icpplpr~\ref{pr:i} and ICPPL \icpplpr~\ref{pr:ii}. For such a labeling
-% %   $\cl$, the proof that it is feasible is by induction on $k$. What
-% %   needs to be shown is that Algorithm~\ref{al:icppl-find-isomorph}
-% %   successfully runs on input $(\cO, \cl)$. In base case $k=0$, $T$ is
-% %   a star. Also every set is at most size 2 ($k+2$) size and thus
-% %   overlaps are at most 1. If two paths share a leaf in {\tt filter\_1}
-% %   one must be of length 2 and the other of length 1. Thus the exit
-% %   condition is not met. Now it is trivial that the exit condition in
-% %   {\tt filter\_2} is also not met. Thus claim proven for base case.
-% %   Now assume the claim to be true when target tree is a
-% %   $(k-1)$-subdivided star. Consider the case of a $k$-subdivided
-% %   star. It is clear that after {\tt filter\_1} and one iteration of
-% %   {\tt filter\_2}, the leaves are assigned pre-images.  Removing the
-% %   leaves from $T$ and the pre-images from support of $\cO$, results in
-% %   an ICPPL to a $(k-1)$-subdivided star.  Now we apply the induction
-% %   hypothesis, and we get a isomorphism between the hypergraph $\cO$
-% %   and $\cO^\cl$.
-
-% %  \noindent
-%   In the reverse direction if there is a feasible path labeling $\cl$,
-%   then we know that $\cl$ is unique up to isomorphism.  Therefore,
-%   again by induction on $k$ it follows that the algorithm finds $\cl$.
-%   \qed
-% \end{proof}
-
-\chapter{Consecutive-ones Property - a Survey of Important Results}
-\tnote{TBD: have a few lines about organization of chapter}
-
-\tnote{ Survey chapter: is the full fledged expansion of the survey in
-  introduction with details, observations, theorems etc.}
-
-\section{Matrices with COP}
-\tnote{Expand on ref:sec:copmatrices}
-
-\section{Optimization problems in COP}
-\tnote{Expand on ref:sec:optcop}
-
-\tnote{ADD NEW SECTIONS - MATCHING WHAT'S IN INTRO CHAPTER}
-
-\chapter{Tree Path Labeling of Path Hypergraphs - the New Results}
-
-\tnote{UPDATE TO FINAL VERSION. BELOW IS OLDER VERSION WITH
-  INCOMPLETE SEC 4.}
-As we saw earlier, tree path labeling problem of path hypergraphs is a
-generalization of consecutive-ones property testing.
-
-\remove{The problem of COP testing can be easily seen as a simple
-  constraint satisfaction problem involving a system of sets from a
-  universe (analogous to hypergraph). Every column of the binary
-  matrix can be converted into a set of integers which are the indices
-  of the rows with |un|s in that column. When observed in this
-  context, if the matrix has the COP, a reordering of its rows will
-  result in sets that have only consecutive integers. In other words,
-  the sets after reordering are intervals. Indeed the problem now
-  becomes finding interval assignments to the given set system
-  cite|nsnrs09| with a single permutation of the universe (set of row
-  indices) which permutes each set to its interval. The result in
-  cite|nsnrs09| characterizes interval assignments to the sets which
-  can be obtained from a single permutation of the rows.  They show
-  that for each set, the cardinality of the interval assigned to it
-  must be same as the cardinality of the set, and the intersection
-  cardinality of any two sets must be same as the interesction
-  cardinality of the corresponding intervals.  While this is naturally
-  a necessary condition, cite|nsnrs09| shows this is indeed
-  sufficient.  Such an interval assignment is called an Intersection
-  Cardinality Preserving Interval Assignment (ICPIA).  Finally, the
-  idea of decomposing a given 0-1 matrix into prime matrices to check
-  for COP is adopted from cite|wlh02| to test if an ICPIA exists for a
-  given set system.}\tnote{REPEATITION from intro chapter}
-
-A natural generalization of the interval
-assignment problem is feasible tree path labeling problem of a set
-system. The problem is defined as follows -- given a set system $\cF$
-from a universe $U$ and a tree $T$, does there exist a bijection from
-$U$ to the vertices of $T$ such that each set in the system maps to a
-path in $T$.  We refer to this as the {\em tree path labeling problem}
-for an input set system, target tree pair -- $(\cF,T)$. As a special
-case if the tree $T$ is a path, the problem becomes the interval
-assignment problem.  We focus on the question of generalizing the
-notion of an ICPIA \cite{nsnrs09} to characterize feasible path
-assignments.  We show that for a given set system $\cF$, a tree $T$,
-and an assignment of paths from $T$ to the sets, there is a feasible
-bijection between $U$ and $V(T)$ if and only if all intersection
-cardinalities among any three sets (not necessarily distinct) is same
-as the intersection cardinality of the paths assigned to them and the
-input runs a filtering algorithm (described in this paper)
-successfully.  This characterization is proved constructively and it
-gives a natural data structure that stores all the relevant feasible
-bijections between $U$ and $V(T)$.  Further, the filtering algorithm
-is also an efficient algorithm to test if a tree path labeling to the
-set system is feasible.  This generalizes the result in
-\cite{nsnrs09}.
-
-The necessary preliminaries with definitions
-etc. are presented in
-Section~\ref{ch:prelims}. Section~\ref{sec:feasible} documents the
-characterization of a feasible path labeling for any path
-labeling from any target trees. Section~\ref{sec:spltargettree}
-describes the special case where the target tree is of a particular
-family of trees and specifically,
-Section~\ref{sec:ksubdivstar} describes a polynomial time algorithm to
-find the tree path labeling of a given set system from a given
-$k$-subdivided tree. Section~\ref{sec:norestraint} describes the
-general situation where the target tree has no restrictions and the
-algorithm to find a TPL, if  any, in this case.
-
-\section{Introduction}
-
-%\begin{tabular}[t]{c}
-{\small
-\begin{minipage}[h]{5in}
- % \begin{singlespace}
- \vspace{2mm}
- {\large \FTPL}\\
- \begin{tabular}[t]{l|l}
- \hline\\
-    {\tt Input} & 
-    \begin{minipage}[t]{\probdefwidth}
-      A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
-      paths $\cP$ from $T$ and a
-      bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
-    \end{minipage}\\
-    {\tt Question} &
-    \begin{minipage}[t]{\probdefwidth}
-      Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
-      such that $\phi$ when applied on any hyperedge in $\cF$ will
-      give
-      the path mapped to it by the given tree path labeling $\cl$.\\
-      { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every hyperedge
-        $S \in \cF$.}
-    \end{minipage}\\
-  \end{tabular}
-  % \end{singlespace}
-\end{minipage}\\
-}
-%\end{tabular}
-
-%\noindent 
-\annote{It
-is an interesting fact that for a matrix with the COP, the
-intersection graph of the corresponding set system is an interval
-graph.}{the following text could be in Further Study/Conclusion}  A similar connection to a subclass of chordal graphs and a
-superclass of interval graphs exists for the generalization of COP.
-In this case, the intersection graph of the corresponding set system
-must be a {\em path graph}. Chordal graphs are of great significance,
-are extensively studied, and have several applications.  One of the well
-known and interesting properties of a chordal graphs is its connection
-with intersection graphs \cite{mcg04}. For every chordal graph, there
-exists a tree and a family of subtrees of this tree such that the
-intersection graph of this family is isomorphic to the chordal graph
-\cite{plr70,gav78,bp93}.  These trees when in a certain format, are
-called {\em clique trees} \cite{apy92} of the chordal graph. This is a
-compact representation of the chordal graph. There has also been work
-done on the generalization of clique trees to clique hypergraphs
-\cite{km02}.  If the chordal graph can be represented as the
-intersection graph of paths in a tree, then the graph is called path
-graph \cite{mcg04}.  Therefore, it is clear that if there is a
-bijection from $U$ to $V(T)$ such that for every set, the elements in
-it map to vertices of a unique path in $T$, then the intersection
-graph of the set system is indeed a path graph.  However, this is only
-a necessary condition and can be checked efficiently because path
-graph recognition is polynomial time
-solvable\cite{gav78,aas93}. Indeed, it is possible to construct a set
-system and tree, such that the intersection graph is a path graph, but
-there is no bijection between $U$ and $V(T)$ such that the sets map to
-paths. Path graph isomorphism is known be isomorphism-complete, see
-for example \cite{kklv10}. An interesting area of research would be to
-see what this result tells us about the complexity of the tree path
-labeling problem (not covered in this paper).
-
-%\noindent
+As a special case if the tree
+$T$ is a path, the problem becomes the interval assignment problem. \tnote{USE THE FORMAL PROBLEM DEF BLOCK}
+We focus on the question of generalizing the notion of an ICPIA
+\cite{nsnrs09} to characterize feasible path assignments.  We show
+that for a given set system $\cF$, a tree $T$, and an assignment of
+paths from $T$ to the sets, there is a feasible bijection between $U$
+and $V(T)$ if and only if all intersection cardinalities among any
+three sets (not necessarily distinct) is same as the intersection
+cardinality of the paths assigned to them and the input runs a
+filtering algorithm (described in this paper) successfully.  This
+characterization is proved constructively and it gives a natural data
+structure that stores all the relevant feasible bijections between $U$
+and $V(T)$.  Further, the filtering algorithm is also an efficient
+algorithm to test if a tree path labeling to the set system is
+feasible.  This generalizes the result in \cite{nsnrs09}.
+
 In the later part of this paper, we focus on a new special case of the
@@
 
-% In the later part of this paper, we decompose our search for a
-% bijection between $U$ and $V(T)$ into subproblems.  Each subproblem is
-% on a set subsystem in which for each set, there is another set in the
-% set subsystem with which the intersection is {\em strict}, i.e., there
-% is a non-empty intersection, but neither is contained in the other.
-% This is in the spirit of results in \cite{wlh02,nsnrs09} where to test
-% for the COP in a given matrix, the COP problem is solved on an
-% equivalent set of prime matrices.
-
 \section{Preliminaries to new results}
 \label{ch:prelims}
-%\section{Preliminaries} 
-%\label{sec:prelims}
 
@@
 supp(\cF) = U$.
-%\vspace{\prelimspace}
 
@@
 \mid P \subseteq V, \text{ } T[P] \text{ is a path} \}$.
-%\vspace{\prelimspace}
 
@@
 % A set system $\cF$ can be alternatively represented by a {\em
-%   hypergraph} $\H_\cF$ whose vertex set is $supp(\cF)$ and hyperedges
-% are the sets in $\cF$. This is a known representation for interval
-% systems in literature \cite{bls99,kklv10}.  We extend this definition here to
-% path systems.
+%   hypergraph} $\H_\cF$ whose vertex set is $supp(\cF)$ and
+% hyperedges are the sets in $\cF$. This is a known representation for
+% interval systems in literature \cite{bls99,kklv10}.  We extend this
+% definition here to path systems.
 
@@
 are the sets in $\cF$. This is a known representation for interval
-systems in literature \cite{bls99,kklv10}.  We extend this definition here to
-path systems. Due to the equivalence of set system and hypergraph in
-the scope of this paper, we drop the
-subscript $_H$ in the notation and refer to both the structures by $\cF$.
-%\vspace{\prelimspace}
-
-Two hypergraphs $\cF'$, $\cF''$ are said to be {\em
-  isomorphic} to each other, denoted by $\cF' \cong \cF''$, iff there
-exists a bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such
-that for all sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in
-$\cF'$ iff $B$ is a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x
-\in A\}$ \cite{kklv10}. This is called {\em hypergraph
+systems in literature \cite{bls99,kklv10}.  We extend this definition
+here to path systems. Due to the equivalence of set system and
+hypergraph in the scope of this paper, we drop the subscript $_H$ in
+the notation and refer to both the structures by $\cF$.
+
+Two hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to
+each other, denoted by $\cF' \cong \cF''$, iff there exists a
+bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for all
+sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ iff $B$ is
+a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
+\cite{kklv10}. This is called {\em hypergraph
   isomorphism}. \tnote{also extend $\phi$ to hyperedges -- see if
   required }
-%\vspace{\prelimspace}
-
-The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph
-$\cF$ is a graph such that its vertex set has a bijection to $\cF$ and
-there exists an edge between two vertices iff their corresponding
-hyperedges have a non-empty intersection \cite{mcg04}.
-%\vspace{\prelimspace}
-
-%\noindent 
-If the intersection graphs of two hypergraphs are isomorphic, $\bI(\cF)
-\cong \bI(\cP)$ where $\cP$ is also a path system, then the bijection
-$\cl: \cF \rightarrow \cP$ due to this isomorphism is called a {\em
-  path labeling} of the hypergraph $\cF$. To illustrate further, let
-$\cg: V(\cF) \rightarrow V(\cP)$ be the above mentioned isomorphism
-where $V(\cF)$ and $V(\cP)$ are the vertex sets that represent the
-hyperedges for each hypergraph respectively, $V(\cF) = \{ v_S \mid S
-\in \cF\}$ and $V(\cP) = \{ v_P \mid P \in \cP\}$. Then the path
-labeling $\cl$ is defined as follows: $\cl(S_1) = P_1$ iff $\cg
-(v_{S_1}) = v_{P_1}$. The path system $\cP$ may be alternatively
+
+The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph $\cF$ is a
+graph such that its vertex set has a bijection to $\cF$ and there
+exists an edge between two vertices iff their corresponding hyperedges
+have a non-empty intersection \cite{mcg04}.
+
+If the intersection graphs of two hypergraphs are isomorphic,
+$\bI(\cF) \cong \bI(\cP)$ where $\cP$ is also a path system, then the
+bijection $\cl: \cF \rightarrow \cP$ due to this isomorphism is called
+a {\em path labeling} of the hypergraph $\cF$. To illustrate further,
+let $\cg: V(\cF) \rightarrow V(\cP)$ be the above mentioned
+isomorphism where $V(\cF)$ and $V(\cP)$ are the vertex sets that
+represent the hyperedges for each hypergraph respectively, $V(\cF) =
+\{ v_S \mid S \in \cF\}$ and $V(\cP) = \{ v_P \mid P \in \cP\}$. Then
+the path labeling $\cl$ is defined as follows: $\cl(S_1) = P_1$ iff
+$\cg (v_{S_1}) = v_{P_1}$. The path system $\cP$ may be alternatively
 denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In most scenarios in
@@
 $supp(\cP) = V(T)$.
-%\vspace{\prelimspace}
-
-% A graph $G$ is a {\em path graph} if it is isomorphic to the intersection graph
-% $\bI(\cP)$ of a path system $\cP$. 
-% This
-% isomorphism gives a bijection $\cl': V(G) \rightarrow \cP$. Moreover,
-% for the purposes of this paper, we require that in a path labeling,
-% $supp(\cP) = V(T)$.  If graph $G$ is also isomorphic to $\bI(\cF)$ for some hypergraph
-% $\cF$, then clearly there is a bijection $\cl: \cF \rightarrow \cP$
-% such that $\cl(S) = \cl'(v_S)$ where $v_S$ is the vertex corresponding
-% to set $S$ in $\bI(\cF)$ for any $S \in \cF$. This bijection $\cl$ is
-% called the {\em path labeling} of the hypergraph $\cF$ and the path
-% system $\cP$ may be alternatively denoted as $\cF^\cl$.
-%\vspace{\prelimspace}
 
@@
 \rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.
-%\vspace{\prelimspace}
 
@@
 called an {\em overlap component}.
-% If there are $d$ overlap components
-% in $\bO(\cF)$, the set subsystems are denoted by $\cO_1, \cO_2, \ldots
-% \cO_d$. Clearly $\cO_i \subseteq \F, i \in [d]$. For any $i, j \in [d]$,
-% it can be verified that one of the following is true.
-% \begin{enumerate}[a) ]
-% \item $supp(\cO_i)$ and $supp(\cO_j)$ are disjoint
-% \item $supp(\cO_i)$ is a subset of a set in $\cO_j$
-% \item $supp(\cO_j)$ is a subset of a set in $\cO_i$
-% \end{enumerate}
-%\vspace{\prelimspace}
 
@@
 In this section we give an algorithmic characterization of a
-feasibility of tree path labeling.
-% Consider a path labeling $\cl: \cF \rightarrow \cP$ for set system $\cF$
-% and path system $\cP$ on the given tree $T$. We
-% call $\cl$ an {\em Intersection Cardinality Preserving Path Labeling
-%   (ICPPL)} if it has the following properties.
-Consider a path labeling $(\cF, \cl)$ on the
-given tree $T$. We call $(\cF, \cl)$ an {\em Intersection Cardinality
-  Preserving Path Labeling (ICPPL)} if it has the following
-properties.
-
-%\vspace{\topshrink}
+feasibility of tree path labeling.  Consider a path labeling $(\cF,
+\cl)$ on the given tree $T$. We call $(\cF, \cl)$ an {\em Intersection
+  Cardinality Preserving Path Labeling (ICPPL)} if it has the
+following properties.
 
 \begin{enumerate}[{(\icpplpr\ }i) \ \ \ ]
-\item \label{pr:i} $|S| = |\cl(S)|$ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $S \in \cF$
+\item \label{pr:i} $|S| = |\cl(S)|$ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
+  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $S \in \cF$
   % \vspace{\topshrink}
-\item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ \ \ \ \ \ \ \ \ \ \
-  \ \ \ \ \ \ \ \ \ \ for all distinct
-  $S_1, S_2 \in \cF$
+\item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ \ \ \ \
+  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all distinct $S_1, S_2 \in \cF$
   % \vspace{\topshrink}
-\item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2) \cap
-  \cl(S_3)|$ \ \ \ for all distinct $S_1, S_2, S_3 \in
-  \cF$
+\item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
+  \cap \cl(S_3)|$ \ \ \ for all distinct $S_1, S_2, S_3 \in \cF$
 \end{enumerate}
 
-\noindent
 The following lemma is useful in subsequent arguments. 
@@
 \begin{algorithm}[h]
-  \caption{{\tt get-hypergraph-isomorphism($\cF, \cl, T$)}}
+  \caption{{\tt get- hypergraph- isomorphism ($\cF, \cl, T$)}}
   \label{al:icppl-find-isomorph}
@@
   that filter 1 and filter 2 do not exit if input is feasible. Thus
-  Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty function.\qed
+  Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
+  function.\qed
 \end{proof}
 
-\subsection{ICPPL when given tree is a path}
-\label{subsec:icpplicpia}
+\section{ Computing feasible TPL with special target trees \tnote{give
+    problem definition etc}}
+\label{sec:spltargettree}
+
+Section~\ref{sec:feasible} described properties that a TPL must have
+for it to be feasible. The next problem of interest is to test if a
+given hypergraph is a path hypergraph with respect to a given target
+tree\footnote{A larger problem would be to find if a given hypergraph
+  is a path graph on any tree. This problem is not addressed in this
+  thesis.}. In other words, the problem is to find out if a feasible
+tree path labeling exists from a given target tree for a given
+hypergraph. In this section we will see two special cases of this
+problem where the target tree is from a particular family of
+trees. The first one, where the tree is a path, is a shown to be
+equivalent to the well studied problem of consecutive-ones in
+Section~\ref{sec:icpplicpia}. The second one, where the tree is a
+$k$-subdivided tree\footnote{See Section~\ref{ch:prelims} for the
+  formal definition.}, has been solved using a polynomial time
+algorithm. The latter problem enforces some conditions on the
+hypergraph too which will be seen in Section~\ref{sec:ksubdivstar}.
+
+\subsection{Target tree is a Path}
+\label{sec:icpplicpia}
 
@@
 \begin{enumerate}
-\item Given tree $T$ is a path. Hence, all path labels are interval labels.
-\item Only pairwise intersection cardinality
-  preservation is sufficient. i.e. property (iii) in ICPPL is not enforced.
+\item Given tree $T$ is a path. Hence, all path labels are interval
+  labels.
+\item Only pairwise intersection cardinality preservation is
+  sufficient. i.e. property (iii) in ICPPL is not enforced.
 \item The filter algorithms do not have {\em \bf exit} statements.
@@
 
-\section{ TPL with special target trees}
-\label{sec:spltargettree}
-\tnote{give problem definition etc}
-
-\subsection{Testing for feasible path assignments to $k$-subdivided star}
+\subsection{Target tree is a $k$-subdivided Star}
 \label{sec:ksubdivstar}
+
+\begin{figure}[t] %[htbp] 
+  \centering
+  \begin{tabular}{lr}
+    (a) \includegraphics{cop-tree_thesis.1}&
+    (b) \includegraphics{cop-tree_thesis.2}
+  \end{tabular}
+  \caption{\figtabsize (a) $8$-subdivided star with 7 rays (b)
+    3-subdivided star with 3 rays}
+  \label{fig:kstar}
+\end{figure}
+
 In this section we consider the problem of assigning paths from a
@@
 \cite{kklv10}. Recall from Section~\ref{ch:prelims} that hyperedges
-$S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$ and $S'$
-have a non-empty intersection but neither of them is contained in the
-other. The overlap graph $\bO(\cF)$ is a graph in which the vertices
-correspond to the sets in $\cF$, and the vertices corresponding to the
-hyperedges $S$ and $S'$ are adjacent if and only if they overlap.  Note
-that the intersection graph of $\cF$, $\bI(\cF)$ is different from
-$\bO(\cF)$ and $\bO(\cF) \subseteq \bI(\cF)$.  A connected component
-of $\bO(\cF)$ is called an overlap component of $\cF$.  An interesting
-property of the overlap components is that any two distinct overlap
-components, say $\cO_1$ and $\cO_2$, are such that any two sets $S_1
-\in \cO_1$ and $S_2 \in \cO_2$ are disjoint, or, w.l.o.g, all the sets
-in $\cO_1$ are contained within one set in $\cO_2$.  This containment
-relation naturally determines a decomposition of the overlap
-components into rooted containment trees.  We consider the case when
-there is only one rooted containment tree, and we first present our
-algorithm when $\bO(\cF)$ is connected.  It is easy to see that once
-the path labeling to the overlap component in the root of the
-containment tree is achieved, the path labeling to the other overlap
-components in the rooted containment tree is essentially finding a
-path labeling when the target tree is a path: each target path is a
-path that is allocated to sets in the root overlap component.
-Therefore, for the rest of this section, $\bO(\cF)$ is a connected
-graph. We also assume that all hyperedges are of cardinality at most
-$k+2$.
+$S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$
+and $S'$ have a non-empty intersection but neither of them is
+contained in the other. The overlap graph $\bO(\cF)$ is a graph in
+which the vertices correspond to the sets in $\cF$, and the vertices
+corresponding to the hyperedges $S$ and $S'$ are adjacent if and only
+if they overlap.  Note that the intersection graph of $\cF$,
+$\bI(\cF)$ is different from $\bO(\cF)$ and $\bO(\cF) \subseteq
+\bI(\cF)$.  A connected component of $\bO(\cF)$ is called an overlap
+component of $\cF$.  An interesting property of the overlap components
+is that any two distinct overlap components, say $\cO_1$ and $\cO_2$,
+are such that any two sets $S_1 \in \cO_1$ and $S_2 \in \cO_2$ are
+disjoint, or, w.l.o.g, all the sets in $\cO_1$ are contained within
+one set in $\cO_2$.  This containment relation naturally determines a
+decomposition of the overlap components into rooted containment trees.
+We consider the case when there is only one rooted containment tree,
+and we first present our algorithm when $\bO(\cF)$ is connected.  It
+is easy to see that once the path labeling to the overlap component in
+the root of the containment tree is achieved, the path labeling to the
+other overlap components in the rooted containment tree is essentially
+finding a path labeling when the target tree is a path: each target
+path is a path that is allocated to sets in the root overlap
+component.  Therefore, for the rest of this section, $\bO(\cF)$ is a
+connected graph. We also assume that all hyperedges are of cardinality
+at most $k+2$.
 
@@
 star has a central vertex which we call the {\em root}, and each root
-to leaf path is called a {\em ray}.  First, we observe that by
-removing the root $r$ from $T$, we get a collection of $p$ vertex
-disjoint paths of length $k+1$, $p$ being the number of leaves in $T$.
-We denote the rays by $R_1, \ldots, R_p$, and the number of vertices in
-$R_i$, $i \in [p]$ is $k+2$.  Let
-$\seq{v_{i1},\ldots,v_{i(k+2)}=r}$ denote the sequence of vertices in
-$R_i$, where $v_{i1}$ is the leaf. Note that $r$ is a common vertex to all $R_i$.
-
-In this section the given hypergraph, the
-$k$-subdivided star and the root of the star are denoted by $\cO$, $T$ and vertex
-$r$, respectively.
+to leaf path is called a {\em ray}. First, we observe that by removing
+the root $r$ from $T$, we get a collection of $p$ vertex disjoint
+paths of length $k+1$, $p$ being the number of leaves in $T$.  We
+denote the rays by $R_1, \ldots, R_p$, and the number of vertices in
+$R_i$, $i \in [p]$ is $k+2$.  Let $\seq{v_{i1},\ldots,v_{i(k+2)}=r}$
+denote the sequence of vertices in $R_i$, where $v_{i1}$ is the
+leaf. Note that $r$ is a common vertex to all $R_i$.
+
+In this section the given hypergraph, the $k$-subdivided star and the
+root of the star are denoted by $\cO$, $T$ and vertex $r$,
+respectively.
 % For each hyperedge $X \in \cO$, we will maintain a 2-tuple of non-negative
@@
 
-The set $\cO_i$ refers to the set of hyperedges $\cT_1^i \cup \cT_2^i$ in
-the $i$th iteration.  Note that $\cO_1 = \cO$.  In the $i$th
+The set $\cO_i$ refers to the set of hyperedges $\cT_1^i \cup \cT_2^i$
+in the $i$th iteration.  Note that $\cO_1 = \cO$.  In the $i$th
 iteration, hyperedges from $\cO_i$ are assigned paths from $R_i$ using
-the following rules. Also the end of the iteration, $\cL_1^{i+1}, \cL_2^{i+1}, \cT_1^{i+1},
-\cT_2^{i+1}$ are set to $\cL_1^{i}, \cL_2^{i}, \cT_1^{i},
-\cT_2^{i}$ respectively, along with some case-specific changes mentioned in the
-rules below.
+the following rules. Also the end of the iteration, $\cL_1^{i+1},
+\cL_2^{i+1}, \cT_1^{i+1}, \cT_2^{i+1}$ are set to $\cL_1^{i},
+\cL_2^{i}, \cT_1^{i}, \cT_2^{i}$ respectively, along with some
+case-specific changes mentioned in the rules below.
 
@@
 
-% \subsection{temp section from nsnsr09}
-
-% \begin{algorithm}
-%   \caption{Basic step in an algorithm to find an ICPIA for a prime
-%     matrix $M'$}
-%   \label{ds-algo}
-%   ICPIA(Set $S$, Integer $p > 0$)
-
-% \noindent
-% /* {\tt $S \cap S^i \not = \phi$ for some $i \in \{1,\ldots,p\}$, but
-%   $S \not\subseteq S^i$, $S^i \not\subseteq S$.  \\ Assigns to $S$ an
-%   interval $I$ such that $\{I^1,\ldots,I^p,I\}$ forms an ICPIA for
-%   $\{S^1,\ldots,S^p,S\}$.} */
-% \begin{algorithmic}
-%   \STATE Let $|S \cap S^i| = z$. \\
-%   \STATE Let $I_l$ be the interval such that $|I_l \cap I^i| = z$, $|I_l| = |S|$ and the $z$ common elements are the smallest elements of $I^i$. \\
-%   \STATE Let $I_r$ be the interval such that $|I_r \cap I^i| = z$, $|I_r| = |S|$, and the $z$ common elements are the largest elements of $I^i$. \\
-%   \IF {$p == 1$}
-%   \STATE Assign $I_l$ to $S$ \\
-%   \STATE /* {\tt In this case, $I_r$ could also be assigned to $S$.  This will yield the {\bf other} ICPIA} */\\
-%   \ELSE \IF{$|I_l \cap I^q| = |S \cap S^q|$ for each $q \in
-%     \{1,\ldots,p\}$}
-%   \STATE Assign $I_l$ to $S$ and exit.\\
-%   \ENDIF \IF{$|I_r \cap I^q| = |S \cap S^q|$ for each $q \in
-%     \{1,\ldots,p\}$}
-%   \STATE Assign $I_r$ to $S$ and exit.\\
-%   \ENDIF \ENDIF \STATE Report no ICPIA and exit.
-% \end{algorithmic}
-% \end{algorithm}
-
-% \begin{theorem}
-%   Algorithm~\ref{ds-algo} outputs an ICPIA to a prime matrix $M'$ iff
-%   there is an ICPIA for $M'$.
-% \end{theorem}
-% \begin{proof}
-%   The only-if part of the theorem is straightforward.  We now show
-%   that if there is an ICPIA for $M'$, then Algorithm~\ref{ds-algo}
-%   will indeed discover it.  The key fact is that in $M'$ for each set
-%   $S$, there is another set $T \in M'$ such that $S \cap T \not =
-%   \phi$, and $S$ and $T$ are not contained in each other.  Due to this
-%   fact, there are exactly two ICPIAs for $M'$.  The two distinct
-%   ICPIAs differ based on the interval assigned to $S_1$, see Algorithm
-%  ~\ref{ds-algo}.  If $I_l$ is assigned to $S_1$, then we get one, and
-%   the other ICPIA is obtained by assigning $I_r$ to $S_1$.  For each
-%   subsequent set, say $S^j$, the interval to be assigned is forced.
-%   It is forced due to the fact that the interval assigned to $S^j$ is
-%   based on the interval assigned to $S^i$, where $S^i \cap S^j \not =
-%   \phi$, and $S^i \not\subseteq S^j$, and $S^j \not\subseteq S^i$.
-%   Given the fact that the algorithm is an exact implementation of
-%   these observations, it follows that Algorithm~\ref{ds-algo} finds an
-%   ICPIA if there is one.
-% \end{proof}
-
 \section{ TPL with no restrictions}
 \label{sec:norestraint}
-\tnote{ADD WALCOM PAPER TEXT}
+
+{\tt abstract begin}
+  it is
+  known that if the given tree is a path a feasible assignment can be
+  found in polynomial time, and we observe that it can actually be
+  done in logspace. \tnote{[TRUE?]}
+{\tt abstract end}
+
+In Section \ref{prelims} we present the
+necessary preliminaries, in Section \ref{feasible} we present our
+characterization of feasible tree path assignments, and in Section
+\ref{decompos} we present the characterizing subproblems for finding a
+bijection between $U$ and $V(T)$ such that sets map to tree
+paths. Finally, in Section \ref{complexity} we conclude by showing
+that Tree Path Assignment is GI-Complete, and also observe that
+Consecutive Ones Testing is in Logspace.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC \subsection{Introduction}
+
+\annote{ We refer to this as the Tree Path Assignment
+problem for an input $(\cF,T)$ pair.}{Add this ``pair'' as a
+  definition in ch:prelims}
+
+\remove{It is an interesting fact that for a matrix with the COP, the
+intersection graph of the corresponding set system is an interval
+graph.  A similar connection to a subclass of chordal graphs, and this
+subclass contains interval graphs, exists for the generalization of
+COP.  In this case, the intersection graph of the corresponding set
+system must be a path graph. Chordal graphs are of great
+significance, extensively studied, and have several applications.  One
+of the well known and interesting properties of a chordal graphs is
+its connection with intersection graphs cite{mcg04}. For every chordal
+graph, there exists a tree and a family of subtrees of this tree such
+that the intersection graph of this family is isomorphic to the
+chordal graph cite{plr70,gav78,bp93}.  Certain format of these trees
+are called clique trees cite{apy92} of the graph which is a compact
+representation of the chordal graph. There has also been work done on
+the generalization of clique trees to clique hypergraphs cite{km02}.
+If the chordal graph can be represented as the intersection graph of
+paths in a tree, then the graph is called path graph cite{mcg04} or
+path chordal graph.  Therefore, it is clear that if there is a
+bijection from $U$ to $V(T)$ such that the sets map to paths, then the
+intersection graph of the set system is indeed a path chordal graph.
+This is, however, only a necessary condition and can be checked
+efficiently, as path graph recognition is polynomial
+solvable cite{gav78,aas93}.  Indeed, it is possible to construct a set
+system and tree, such that the intersection graph is a path chordal
+graph, but there is no bijection between $U$ and $V(T)$ such that the
+sets map to paths.  This connection indeed suggests that our problem
+is indeed as hard as path graph isomorphism.  Further path graph
+isomorphism is known be isomorphism-complete, see for example
+cite{kklv10}. }
+In the second part of this paper, we decompose our
+search for a bijection between $U$ and $V(T)$ into subproblems.  Each
+subproblem is on a set system in which for each set, there is another
+set in the set system with which the intersection is {\em strict}-
+there is a non-empty intersection, but neither is contained in the
+other.  This is in the spirit of results in \cite{wlh02,nsnrs09} where
+to test for COP in a given matrix, the COP problem is solved on an
+equivalent set of prime matrices.  \annote{Our decomposition localizes the
+challenge of path graph isomorphism to two problems.}{HOW?}
+
+Finally, we show that Tree Path Assignment is isomorphism-complete.
+We also point out Consecutive Ones Testing is in Logspace from two
+different results in the literature \cite{kklv10, mcc04}. To the best
+of our knowledge this observation has not been made earlier.
+
+% \noindent
+% It has been long known that interval graph recognition is in
+% logspace\cite{rei84}. Recently interval graph isomorphism was also
+% shown to be logspace decidable using a logspace canonization
+% algorithm by \cite{kklv10}.  This result is built on top of logspace
+% results of undirected graph connectivity \cite{rei08}, logspace
+% tractability using a certain logical formalism called FP+C and
+% modular decomposition of interval graphs\cite{lau10} etc.  Interval
+% graphs are closely connected to binary matrices with COP. The
+% maximal clique vertex incidence matrix (matrix with rows
+% representing maximal cliques and columns representing vertices of a
+% graph) has COP on columns iff the graph is an interval
+% graph\cite{fg65}. This follows from the interval graph
+% characterization by \cite{gh64}. Due to this close relation it is
+% natural to see if consecutive ones property can be
+% tested in logspace. \\
+% \noindent
+% We also explore some extensions of the interval assignment problem
+% in \cite{nsnrs09}, namely caterpillar path assignment problem.
+
+% We present a logspace algorithm here that uses the ICPIA
+% characterization of binary matrices with COP (set system associated
+% with such a matrix)\cite{nsnrs09}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC\subsection{Preliminaries} 
+\label{prelims} 
+In this paper, the collection
+$\F = \{S_i \mid S_i \subseteq U, S_i \ne \O, i \in I\}$ is a {\em set
+  system} of universe, $U = \{1, \ldots, n\}$. Moreoever, a set system
+is assumed to ``cover'' the universe,
+i.e. $ \bigcup_{i \in I}S_i = U$. \\
+
+\noindent
+The graph $T=(V,E)$ represents a given tree with $|V| = n$. Further,
+for simplicity, $V$ is defined as $\{1,\ldots,n\}$. All paths
+referred to in this paper are paths of $T$ unless explicitly specified. \\
+
+\noindent
+A {\em path assignment} $\A$ to $\F$ is defined as a set assignment
+where second universe is the vertex set $V$ of a given tree $T$ and
+every second subset in the ordered pairs is a path in this
+tree. Formally, the definition is as follows.
+\begin{align*}
+  \A = \{ (S_i,P_i) \mid S_i \in \F, P_i \subseteq V \text{
+    s.t. }T[P_i] \text{ is a path, } i \in I \}
+\end{align*}
+In other words, $P_i$ is the path on the tree $T$ assigned to $S_i$ in
+$\A$. As mentioned before for set systems, the paths cover the whole
+tree, i.e. $\bigcup_{i \in I}P_i = V$ \\
+
+\noindent
+Generalizing the definition of {\em feasibility} in \cite{nsnrs09} to
+a set assignment, a path assignment $\A$ is defined to be {\em
+  feasible} if there exists a bijection defined as follows.
+\begin{align}
+  \sigma: U \rightarrow V(T), \text{ such that }\sigma(S_i) = P_i
+  \text{ for all } i \in I, \sigma \text{ is a bijection}
+  \label{eq:stf}
+\end{align}
+% The assignment $\A$ is then called {\em Set Translation Feasible}.
+
+\noindent
+Let $X$ be a partially ordered set with $\preccurlyeq$ being the
+partial order on $X$.  $mub(X)$ represents an element in $X$ which is
+a maximal upper bound on $X$.  $X_m \in X$ is a maximal upper bound of
+$X$ if $\nexists X_q \in X$ such that $X_m
+\preccurlyeq X_q$. \\
+
+\noindent
+The set $I$ represents the index set $[m]$. If index $i$ is used
+without further qualification, it is meant to be $i \in I$. Any
+function, if not defined on a domain of sets, when applied on a set is
+understood as the function applied to each of its elements. i.e. for
+any function $f$ defined with domain $U$, the abuse of notation is as
+follows; $f(S)$ is used instead of $\hat f(S)$ where
+$\hat f(S) = \{y \mid y = f(x), x \in S\}$. \\
+
+\noindent
+When refering to a tree as $T$ it could be a reference to the tree
+itself, or the vertices of the tree. This will be clear from the
+context.\\
+
+\noindent
+Finally, an in-tree is a directed rooted tree in which all edges are
+directed toward to the root.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC\subsection{Characerization of Feasible Tree Path
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC  Assignment\tnote{Merge with the original charac section}} 
+\label{feasible} 
+Consider a path assignment $\cA =
+\{(S_i, P_i) \mid S_i \in \cF, P_i \text{ is a path from $T$}, i \in
+[m]\}$ to a set system $\cF = \{S_i \mid S_i \subseteq U, i \in
+[m]\}$, were $T$ is a given tree, $U$ is the set system's universe and
+$m$ is the number of sets in $\cF$. We call $\cA$ an {\em Intersection
+  Cardinality Preserving Path Assignment (ICPPA)} if it has the
+following properties.
+
+\begin{enumerate}
+\item [i.]  $|S_i| = |P_i|$ for all $i \in [m]$
+\item [ii.] $|S_i \cap S_j| = |P_i \cap P_j|$ for all $i,j \in [m]$
+  % \item [iii.] $|\bigcup_{i \in J} S_i| = |\bigcup_{i \in J} P_i|$
+  %   for all $J \subseteq I$\footnote{TBD: this is needed now to fix
+  %     some proofs}
+\item [iii.] $|S_i \cap S_j \cap S_k| = |P_i \cap P_j \cap P_k|$ for
+  all $i,j,k \in [m]$
+\end{enumerate}
+
+\begin{lemma}
+  \label{lem:setminuscard}
+  If $\cA$ is an ICPPA, and $(S_1, P_1),(S_2, P_2),(S_3, P_3) \in
+  \cA$, then $|S_1 \cap (S_2 \setminus S_3)| = |P_1 \cap (P_2
+  \setminus P_3)|$.
+\end{lemma}
+\begin{proof}
+  $|S_1 \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| =
+  |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to conditions (ii)
+  and (iii) of ICPPA, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
+  \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
+  = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. \qed
+\end{proof}
+
+\begin{lemma}
+  \label{lem:fourpaths} Consider four paths in a tree $Q_1, Q_2, Q_3,
+  Q_4$ such that they have nonempty pairwise intersection and $Q_1,
+  Q_2$ share a leaf. Then there exists distinct $i, j, k \in
+  \{1,2,3,4\}$ such that, $Q_1 \cap Q_2 \cap Q_3 \cap Q_4 = Q_i \cap
+  Q_j \cap Q_k$.
+\end{lemma}
+\begin{proof}
+  {\em Case 1:} w.l.o.g, consider $Q_3 \cap Q_4$ and let us call it
+  $Q$. This is clearly a path (intersection of two paths is a path).
+  % Since $Q_1, Q_2$ share a leaf, the following are paths $Q_1
+  % \setminus Q_2$, $Q_2 \setminus Q_1$, $Q_1 \cap Q_2$ and they are
+  % mutually disjoint.
+  Suppose $Q$ does not intersect with $Q_1 \setminus Q_2$, i.e. $Q
+  \cap (Q_1 \setminus Q_2) = \O$. Then $Q \cap Q_1 \cap Q_2 = Q \cap
+  Q_2$. Similarly, if $Q \cap (Q_2 \setminus Q_1) = \O$, $Q \cap Q_1
+  \cap Q_2 = Q \cap Q_1$. Thus it is clear that if the intersection of
+  any two paths does not intersect with any of the set differences of
+  the remaining two paths, the claim in the lemma is true.
+  % Note that $Q_1 \setminus Q_2$ and $Q_2
+  % \setminus Q_1$ are paths because $Q_1, Q_2$ share a leaf.\\
+  {\em Case 2:} Let us consider the compliment of the previous
+  case. i.e. the intersection of any two paths intersects with both
+  the set differences of the other two. First let us consider $Q \cap
+  (Q_1 \setminus Q_2) \ne \O$ and $Q \cap (Q_1 \setminus Q_2) \ne \O$,
+  where $Q = Q_3 \cap Q_4$. Since $Q_1$ and $Q_2$ share a leaf, there
+  is exactly one vertex at which they branch off from the path $Q_1
+  \cap Q_2$ into two paths $Q_1 \setminus Q_2$ and $Q_2 \setminus
+  Q_1$. Let this vertex be $v$. It is clear that if path $Q_3 \cap
+  Q_4$, must intersect with paths $Q_1 \setminus Q_2$ and $Q_2
+  \setminus Q_1$, it must contain $v$ since these are paths from a
+  tree. Moreover, $Q_3 \cap Q_4$ intersects with $Q_1 \cap Q_2$ at
+  exactly $v$ and only at $v$ which means that $Q_1 \cap Q_2$ does not
+  intersect with $Q_3 \setminus Q_4$ or $Q_4 \setminus Q_3$ which
+  contradicts initial condition of this case. Thus this
+  case cannot occur and case 1 is the only possible scenario. \\
+  Thus lemma is proven \qed
+\end{proof}
+
+\noindent
+In the remaining part of this section we show that a path assignment
+is feasible if and only if it is an ICPPA.  One direction of this
+claim is clear: that is a path assignment is feasible, then all
+intersection cardinalities are preserved, that is the path assignment
+is an ICPPA.  The reason is that a feasible path assignment has an
+associated bijection between $U$ and $V(T)$ such that the sets map to
+paths.  The rest of the section is devoted to constructively proving
+that it is sufficient, for the existence of a bijection, for a path
+assignment to be an ICPPA.  At a top-level, the constructive
+approaches refine the path assignment iteratively, such that at the
+end of each iteration we have a path assignment, and finally we have a
+family of bijections.  First we present and then prove the correctness
+of Algorithm \ref{perms}.  This algorithm refines the path assignment
+by considering pairs of paths that share a leaf.
+
+\begin{algorithm}[h]
+  \caption{Permutations from an ICPPA $\{(S_i,P_i) | i \in I\}$}
+  \label{perms}
+  \begin{algorithmic}
+    \STATE Let $\Pi_0=\{(S_i,P_i)| i \in I\}$\\
+    \STATE $j = 1$;
+    \label{shareleaf} \WHILE {There is $(P_1,Q_1), (P_2,Q_2) \in
+      \Pi_{j-1}$ with $Q_1$ and $Q_2$ having a common leaf} \STATE
+    $\Pi_j= \Pi_{j-1} \setminus \{(P_1,Q_1),(P_2,Q_2)\}$;
+    \label{setbreak}\STATE $\Pi_j = \Pi_j \cup \{(P_1 \cap P_2,Q_1
+    \cap Q_2), (P_1 \setminus P_2,Q_1 \setminus Q_2), (P_2 \setminus
+    P_1, Q_2 \setminus Q_1)\}$; \STATE $j = j+1$; \ENDWHILE \STATE
+    $\Pi = \Pi_j$; \STATE Return $\Pi$;
+  \end{algorithmic}
+\end{algorithm}
+
+\begin{lemma}
+  \label{lem:invar1}
+  In Algorithm \ref{perms}, at the end of $j$th iteration, $j \ge 0$,
+  of the while loop of Algorithm \ref{perms}, the following invariants
+  are maintained.
+  \begin{itemize}
+  \item {\em Invariant I:} $Q$ is a path in $T$ for each $(P,Q) \in
+    \Pi_j$
+  \item {\em Invariant II:} $|P|=|Q|$ for each $(P,Q) \in \Pi_j$
+  \item {\em Invariant III:} For any two $(P,Q), (P',Q') \in \Pi_j$,
+    $|P' \cap P''|=|Q' \cap Q''|$.
+  \item {\em Invariant IV:} For any three, $(P',Q'), (P'',Q''), (P, Q)
+    \in \Pi_j$, $|P' \cap P'' \cap P|=|Q' \cap Q'' \cap Q|$.
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  Proof is by induction on the number of iterations, $j$. In the rest
+  of the proof, the term ``new sets'' will refer to the new sets added
+  in $j$th iteration as defined in line \ref{setbreak} of Algorithm
+  \ref{perms}, i.e. the following three assignment pairs for some
+  $(P_1,Q_1), (P_2,Q_2) \in \Pi_{j-1}$ where $Q_1$ and $Q_2$ intersect
+  and share a leaf: $(P_1 \cap P_2, Q_1 \cap Q_2)$, or $(P_1 \setminus
+  P_2, Q_1 \setminus Q_2)$, or $(P_2 \setminus P_1, Q_2
+  \setminus Q_1)$.\\
+  \noindent
+  The base case, $\Pi_0 = \{(S_i,P_i) \mid i \in [m]\}$, is trivially
+  true since it is the input which is an ICPPA.  Assume the lemma is
+  true till the $j-1$ iteration. Consider $j$th
+  iteration:\\
+  \noindent
+  If $(P,Q)$, $(P',Q')$ and $(P'',Q'')$ are in $\Pi_{j}$ and
+  $\Pi_{j-1}$, all the invariants are
+  clearly true because they are from $j-1$ iteration.\\
+  If $(P,Q)$ is in $\Pi_{j}$ and not in $\Pi_{j-1}$, then it must be
+  one of the new sets added in $\Pi_j$. Since $(P_1,Q_1)$ and
+  $(P_2,Q_2)$ are from $\Pi_{j-1}$ and $Q_1,Q_2$ intersect and have a
+  common leaf, it can be verified that the
+  new sets are also paths. \\
+  By hypothesis for invariant III, invariant II also holds for $(P,Q)$
+  no matter which new set in $\Pi_j$ it
+  is.\\
+  To prove invariant III, if $(P,Q)$ and $(P',Q')$ are not in
+  $\Pi_{j-1}$, then they are both new sets and invariant III holds
+  trivially (new sets are disjoint). Next consider $(P,Q), (P',Q') \in
+  \Pi_j$ with only one of them, say $(P',Q')$, in $\Pi_{j-1}$. Then
+  $(P,Q)$ is one of the new sets added in line \ref{setbreak}. It is
+  easy to see that if $(P,Q)$ is $(P_1 \cap P_2, Q_1 \cap Q_2)$, then
+  due to invariant IV in hypothesis, invariant III becomes true in
+  this iteration. Similarly, using lemma \ref{lem:setminuscard}
+  invariant III is proven if $(P, Q)$ is $(P_1 \setminus P_2, Q_1
+  \setminus Q_2)$, or $(P_2 \setminus P_1, Q_2
+  \setminus Q_1)$.\\
+  To prove invariant IV, consider three assignments
+  $(P,Q),(P',Q'),(P'',Q'')$. If at least two of these pairs are in not
+  $\Pi_{j-1}$, then they are any two of the new sets. Note that these
+  new sets are disjoint and hence if $(P',Q'), (P'',Q'')$ are any of
+  these sets, $|P \cap P' \cap P''|=|Q \cap Q' \cap Q''|=0$ and
+  invariant IV is true. Now we consider the case if at most one of
+  $(P,Q),(P',Q'),(P'',Q'')$ is not in $\Pi_{j-1}$. If none of them are
+  not in $\Pi_{j-1}$ (i.e. all of them are in $\Pi_{j-1}$), invariant
+  IV is clearly true. Consider the case where exactly one of them is
+  not in $\Pi_{j-1}$. w.l.o.g let that be $(P,Q)$ and it could be any
+  one of the new sets. If $(P,Q)$ is $(P_1 \cap P_2, Q_1 \cap Q_2)$,
+  from lemma \ref{lem:fourpaths} and invariant III hypothesis,
+  invariant IV is proven. Similarly if $(P,Q)$ is any of the other new
+  sets, invariant IV is proven by also using lemma
+  \ref{lem:setminuscard}. \qed
+
+\end{proof}
+
+\noindent
+It can be observed that the output of algorithm \ref{perms} is such
+that every leaf is incident on at most a single path in the new set of
+assignments. This is due to the loop condition at line
+\ref{shareleaf}. Let $v_1$ be the leaf incident on path $P_i$. Assign
+to it any one element from $S_i \setminus \bigcup_{i \ne j}
+S_j$. Remove $(S_i, P_i)$ from assignments and add $(\{x_1\},
+\{v_1\}), (S_i \setminus \{x_1\}, P_i \setminus \{v_1\})$. Now all
+assignments except single leaf assignments are paths from the subtree
+$T_0 = T \setminus \{v \mid v \text{ is a leaf in } T\}$.
+
+% exclusively on the spine of the caterpillar. Consider the spine to
+% be an interval and run ICPIA algorithm on it to get the required
+% permutation. Algorithm \ref{leafasgn} makes this clear.
+
+\begin{algorithm}[h]
+  \caption{Leaf assignments from an ICPPA $\{(S_i,P_i) | i \in I\}$}
+  \label{leafasgn}
+  \begin{algorithmic}
+    \STATE Let $\Pi_0=\{(S_i,P_i)| i \in [m]\}$. Paths are such that
+    no
+    two paths $P_i, P_j, i \ne j$ share a leaf.\\
+    \STATE $j = 1$\\
+    \WHILE {there is a leaf $v$ and a unique $(S_{i_1}, P_{i_1})$ such
+      that $v \in P_{i_1}$}
+    \STATE $\Pi_j=   \Pi_{j-1} \setminus \{(S_{i_1}, P_{i_1})\}$\\
+    \STATE $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in I}S_i$
+    \IF{$X$ is empty} \STATE exit \ENDIF; \STATE Let $x = $ arbitrary
+    element from $X$ \vspace{2mm}
+
+    \STATE $\Pi_j = \Pi_j \cup \{(S_{i_1} \setminus \{x\}),(P_{i_1}
+    \setminus
+    \{v\}), (\{x\},\{v\})\}$\\
+    \STATE $j = j+1$\\
+    \ENDWHILE
+    \STATE $\Pi = \Pi_j$\\
+    \STATE Return $\Pi$\\
+  \end{algorithmic}
+\end{algorithm}
+
+\begin{lemma}
+  \label{lem:invar3}
+  In Algorithm \ref{leafasgn},for all $j \geq 0$, at the end of the
+  $j$th iteration the four invariants given in lemma \ref{lem:invar1}
+  are valid.
+  % Moreover, $X$ as defined in the algorithm is non-empty if this is
+  % an ICPPA.
+\end{lemma}
+\begin{proof}
+  First we see that $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in
+    I}S_i$ is non empty in every iteration for an ICPPA. Suppose $X$
+  is empty. We know that $v \in P_{i_1} \setminus \bigcup_{i \ne i_1,
+    i \in I}P_i$ since $v$ is in the unique path $P_{i_1}$. Since this
+  is an ICPPA $|S_{i_1}| = |P_{i_1}|$. For any $x \in S_{i_1}$ it is
+  contained in at least two sets due to our assumption. Let $S_{i_2}$
+  be a second set that contains $x$. We know $v \notin
+  P_{i_2}$. Therefore there cannot exist a permutation that maps
+  elements of $S_{i_2}$ to $P_{i_2}$. This contradicts our assumption
+  that this is an ICPPA. Thus $X$ cannot be empty.
+
+  We use mathematical induction on the number of iterations for this
+  proof. The term ``new sets'' will refer to the sets added in $\Pi_j$
+  in the $j$th iteration, i.e. $(P' \setminus \{x\},Q' \setminus
+  \{v\})$ and $(\{x\},\{v\})$ for some $(P',Q')$ in $\Pi_{j-1}$ such
+  that $v$ is a leaf and $Q'$ is the unique path
+  incident on it.\\
+  For $\Pi_0$ all invariants hold because it is output from algorithm
+  \ref{perms} which is an ICPPA. Hence base case is proved.  Assume
+  the lemma holds for $\Pi_{j-1}$. Consider $\Pi_j$ and any $(P,Q) \in
+  \Pi_j$. If $(P,Q) $ is in $ \Pi_j$ and $\Pi_{j-1}$ invariants I and
+  II are true because of induction assumption. If it is only in
+  $\Pi_j$, then it is $\{(P' \setminus \{x\}),(Q' \setminus \{v\})$ or
+  $(\{x\},\{v\})$ for some $(P',Q')$ in $\Pi_{j-1}$. By definition,
+  $x$ is an element in $P'$ (as defined in the algorithm) and $v$ is a
+  leaf in $Q'$. If $(P,Q)$ is $\{(P' \setminus \{x\}),(Q' \setminus
+  \{v\})$, $Q$ is a path since only a leaf is removed from path
+  $Q'$. We know $|P'| = |Q'|$, therefore $|P' \setminus \{x\}| = |Q'
+  \setminus \{v\}|$. Hence in this case invariants I and II are
+  obvious. It is easy to see these invariants hold if $(P,Q)$ is
+  $(\{x\},\{v\})$.
+
+  For invariant III consider $(P_1,Q_1),(P_2,Q_2)$ in $\Pi_j$. If both
+  of them are also in $\Pi_{j-1}$, claim is proved. If one of them is
+  not in $\Pi_{j-1}$ then it has to be $\{(P' \setminus \{x\}),(Q'
+  \setminus \{v\})$ or $(\{x\},\{v\})$ for some $(P',Q')$ in
+  $\Pi_{j-1}$. Since by definition, $Q'$ is the only path with $v$ and
+  $P'$ the only set with $x$ in the previous iteration, $|P_1 \cap (P'
+  \setminus \{x\})| = |P_1 \cap P'|$ and $|Q_1 \cap (Q' \setminus
+  \{v\})| = |Q_1 \cap Q'|$ and $|P_1 \cap \{x\}| = 0, Q_1 \cap \{v\} =
+  0$. Thus invariant III is also proven.
+
+  \noindent
+  To prove invariant IV, consider $(P_1,Q_1),(P_2,Q_2), (P_3,Q_3)$ in
+  $\Pi_j$. If exactly one of them, say $P_3 \notin \Pi_{j-1}$, it is
+  one of the new sets. By the same argument used to prove invariant
+  III, $|P_1 \cap P_2 \cap (P' \setminus \{x\})| = |P_1 \cap P_2 \cap
+  P'|$ and $|Q_1 \cap Q_2 \cap (Q' \setminus \{x\})| = |Q_1 \cap Q_2
+  \cap Q'|$. Since $P_1, P_2, P'$ are all in $\Pi_{j-1}$, by induction
+  hypothesis $|P_1 \cap P_2 \cap P'| = |Q_1 \cap Q_2 \cap Q'|$. Also
+  $|P_1 \cap P_2 \cap \{x\}| = 0, Q_1 \cap Q_2 \cap \{v\} = 0$.  If
+  two or more of them are not in $\Pi_{j-1}$, then it can be verified
+  that $|P_1 \cap P_2 \cap P_3| = |Q_1 \cap Q_2 \cap Q_3|$ since the
+  new sets in $\Pi_j$ are either disjoint or as follows: assuming
+  $P_1, P_2 \notin \Pi_{j-1}$ and new sets are derived from $(P', Q'),
+  (P'', Q'') \in \Pi_{j-1}$ with $x_1, x_2$ exclusively in $P_1, P_2$,
+  $(\{x_1\},\{v_1\}), (\{x_2\},\{v_2\}) \in \Pi_j $ thus $v_1, v_2$
+  are exclusively in $Q_1, Q_2$ resp. it follows that $|P_1 \cap P_2
+  \cap P_3| = |(P' \setminus \{x_1\}) \cap (P'' \setminus \{x_2\})
+  \cap P_3| = |P' \cap P'' \cap P_3| = |Q' \cap Q'' \cap Q_3| = |(Q'
+  \setminus \{v_1\} \cap Q'' \setminus \{v_2\} \cap Q_3| = |Q_1 \cap
+  Q_2 \cap Q_3|$. Thus invariant IV is also proven.  \qed
+\end{proof}
+
+\noindent
+Using algorithms \ref{perms} and \ref{leafasgn} we prove the following
+theorem.
+
+\begin{theorem}
+  \label{th:perm}
+  If $\cA$ is an ICPPA, then there exists a bijection $\sigma : U
+  \rightarrow V(T)$ such that $\sigma(S_i) = P_i$ for all $i \in I$
+\end{theorem}
+\begin{proof}
+  This is a contructive proof. First, the given ICPPA $\cA$ and tree
+  $T$ are given as input to Algorithm \ref{perms}. This yields a
+  ``filtered'' ICPPA as the output which is input to Algorithm
+  \ref{leafasgn}.  It can be observed that the output of Algorithm
+  \ref{leafasgn} is a set of interval assignments to sets and
+  one-to-one assignment of elements of $U$ to each leaf of $T$. To be
+  precise, it would be of the form $\cB_0 = \cA_0 \cup \cL_0$. The
+  leaf assignments are defined in $\cL_0 = \{ (x_i,v_i) \mid x_i \in
+  U, v_i \in T, x_i \ne x_j, v_i \ne v_j, i \ne j, i,j \in [k] \}$
+  where $k$ is the number of leaves in $T$. The path assignments are
+  defined in $\cA_0 \subseteq \{(S_i',P_i') \mid S_i' \subseteq U_0,
+  P_i' \text{ is a path from } T_0\}$ where $T_0$ is the tree obtained
+  by removing all the leaves in $T$ and $U_0 = U \setminus \{ x \mid x
+  \text{ is assigned to a leaf in }\cL_0 \}$. Now we have a subproblem
+  of finding the permutation for the path assignment $\cA_0$ which has
+  paths from tree $T_0$ and sets from universe $U_0$. Now we repeat
+  the procedure and the path assignment $\cA_0$ and tree $T_0$ is
+  given as input to Algorithm \ref{perms}. The output of this
+  algorithm is given to Algorithm \ref{leafasgn} to get a new union of
+  path and leaf assignments $\cB_1 = \cA_1 \cup \cL_1$ defined similar
+  to $\cB_0, \cL_0, \cA_0$. In general, the two algorithms are run on
+  path assignment $\cA_{i-1}$ with paths from tree $T_{i-1}$ to get a
+  new subproblem with path assignment $\cA_i$ and tree $T_{i}$. $T_i$
+  is the subtree of $T_{i-1}$ obtained by removing all its
+  leaves. More importantly, it gives leaf assignments $\cL_{i}$ to the
+  leaves in tree $T_{i-1}$. This is continued until we get a
+  subproblem with path assignment $\cA_{d-1}$ and tree $T_{d-1}$ for
+  some $d \le n$ which is just a path. From the last lemma we know
+  that $\cA_{d-1}$ is an ICPPA. Another observation is that an ICPPA
+  with all its tree paths being intervals (subpaths from a path) is
+  nothing but an ICPIA\cite{nsnrs09}.  Let $\cA_{d-1}$ be equal to
+  $\{(S_i'',P_i'') \mid S_i'' \subseteq U_{d-1}, P_i'' \text{ is a
+    path from } T_{d-1} \}$. It is true that the paths $P_i''$s may
+  not be precisely an interval in the sense of consecutive integers
+  because they are some nodes from a tree. However, it is easy to see
+  that the nodes of $T_{d-1}$ can be ordered from left to right and
+  ranked to get intervals $I_i$ for every path $P_i''$ as
+  follows. $I_i = \{[l,r] \mid l = \text{ the lowest rank of the nodes
+    in }P_i'', r = l+|P_i''|-1 \}$. Let asssignment $\cA_d$ be with
+  the renamed paths. $\cA_d = \{ (S_i'', I_i) \mid (S_i'', P_i'') \in
+  \cA_{d-1} \}$. What has been effectively done is renaming the nodes
+  in $T_{d-1}$ to get a tree $T_d$.  The ICPIA $\cA_d$ is now in the
+  format that the ICPIA algorithm requires which gives us the
+  permutation $\sigma' : U_{d-1} \rightarrow T_{d-1}$
+
+\noindent
+$\sigma'$ along with all the leaf assignments $\cL_i$ gives us the
+permutation for the original path assignment $\cA$.  More precisely,
+the permutation for tree path assignment $\cA$ is defined as
+follows. $\sigma: U \rightarrow T$ such that the following is
+maintained.
+\begin{align*}
+  \sigma(x) &= \sigma'(x),   \text{ if } x \in U_{d-1} \\
+  &= \cL_i(x), \text{ where $x$ is assigned to a leaf in a subproblem
+    $\cA_{i-1}, T_{i-1}$}
+\end{align*}
+
+\noindent
+To summarize, run algorithm \ref{perms} and \ref{leafasgn} on
+$T$. After the leaves have been assigned to specific elements from
+$U$, remove all leaves from $T$ to get new tree $T_0$. The leaf
+assignments are in $\cL_0$. Since only leaves were removed $T_0$ is
+indeed a tree. Repeat the algorithms on $T_0$ to get leaf assignments
+$\cL_{1}$. Remove the leaves in $T_0$ to get $T_1$ and so on until the
+pruned tree $T_d$ is a single path. Now run ICPIA algorithm on $T_d$
+to get permutation $\sigma'$. The relation $\cL_0 \cup \cL_1 \cup
+.. \cup \cL_{d} \cup \sigma'$ gives the bijection required in the
+original problem.\qed
+\end{proof}
+
+\subsection{Finding an assignment of tree paths to a set
+  system} \label{decompos} In the previous section we have shown that
+the problem of finding a Tree Path Asssignment to an input $(\cF,T)$
+is equivalent to finding an ICPPA to $\cF$ in tree $T$.  In this
+section we characterize those set systems that have an ICPPA in a
+given tree.  As a consequence of this characterization we identify two
+sub-problems that must be solved to obtain an ICPPA.  We do not solve
+the problem and in the next section show that finding an ICPPA in a
+given tree is GI-Complete.
+
+\noindent
+A set system can be concisely represented by a binary matrix where the
+row indices denote the universe of the set system and the column
+indices denote each of the sets. Let the binary matrix be $M$ with
+order $n \times m$, the set system be $\cF = \{S_i \mid i \in [m]\}$,
+universe of set system $U = \{x_1, \dots ,x_n\}$. If $M$ represents
+$\cF$, $|U| = n, |\cF| = m$. Thus $(i,j)$th element of $M$, $M_{ij} =
+1$ iff $x_i \in S_j$. If $\cF$ has a feasible tree path assignment
+(ICPPA) $\cA = \{(S_i,P_i) \mid i \in [m]\}$, then we say its
+corresponding matrix $M$ has an ICPPA. Conversly we say that a matrix
+$M$ has an ICPPA if there exists an ICPPA $\cA$ as defined
+above.\\
+\noindent
+We now define the strict intersection graph or overlap graph of
+$\cF$. This graph occurs at many places in the literature, see for
+example \cite{kklv10, wlh02, nsnrs09}.  The vertices of the graph
+correspond to the sets in $\cF$.  An edge is present between vertices
+of two sets iff the corresponding sets have a nonempty intersection
+and none is contained in the other. Formally, intersection graph is
+$G_f = (V_f, E_f)$ such that $V_f = \{v_i \mid S_i \in \cF\}$ and $E_f
+= \{(v_i, v_j) \mid S_i \cap S_j \ne \O \text{ and }S_i \nsubseteq
+S_j, S_j \nsubseteq S_i \}$.  We use this graph to decompose $M$ as
+described in \cite{wlh02,nsnrs09}.  A prime sub-matrix of $M$ is
+defined as the matrix formed by a set of columns of $M$ which
+correspond to a connected component of the graph $G_f$.  Let us denote
+the prime sub-matrices by $M_1,\ldots,M_p$ each corresponding to one
+of the $p$ components of $G_f$. Clearly, two distinct matrices have a
+distinct set of columns.  Let $col(M_i)$ be the set of columns in the
+sub-matrix $M_i$.  The support of a prime sub-matrix $M_i$ is defined
+as $supp(M_i) = \displaystyle \bigcup_{j \in col(M_i)}S_j$.  Note that
+for each $i$, $supp(M_i) \subseteq U$.  For a set of prime
+sub-matrices $X$ we define
+$supp(X) = \displaystyle \bigcup_{M \in X} supp(M)$. \\
+
+\noindent
+Consider the relation $\preccurlyeq$ on the prime sub-matrices $M_1,
+\ldots, M_p$ defined as follows:
+\begin{equation}
+  \nonumber \{(M_i,M_j) | \mbox{ A set } S \in
+  M_i \mbox{ is contained in a set } S' \in M_j\} \cup \{(M_i,M_i) | 1
+  \leq i \leq p\} 
+\end{equation}
+
+\noindent
+This relation is the same as that defined in \cite{nsnrs09}. The prime
+submatrices and the above relation can be defined for any set
+system. We will use this structure of prime submatrices to present our
+results on an an ICPPA for a set system $\cF$. Recall the following
+lemmas, and theorem that $\preccurlyeq$ is a partial order, from
+\cite{nsnrs09}.
+
+\begin{lemma} \label{lem:containment} Let $(M_i,M_j) \in
+  \preccurlyeq$.  Then there is a set $S' \in M_j$ such that for each
+  $S \in M_i$, $S \subseteq S'$.
+\end{lemma}
+\begin{lemma}
+  For each pair of prime sub-matrices, either $(M_i,M_j) \not\in
+  \preccurlyeq$ or $(M_j,M_i) \not\in \preccurlyeq$.
+  % If $(M_i,M_j) \in \preccurlyeq$ and $(M_j,M_i) \in \preccurlyeq$,
+  % then $i = j$ and $|M_i| = 1$.
+\end{lemma}
+\begin{lemma}
+  If $(M_i,M_j) \in \preccurlyeq $ and $(M_j,M_k) \in \preccurlyeq$,
+  then $(M_i,M_k) \in \preccurlyeq$.
+\end{lemma}
+\begin{lemma}
+  If $(M_i,M_j) \in \preccurlyeq$ and $(M_i,M_k) \in \preccurlyeq$,
+  then either $(M_j,M_k) \in \preccurlyeq$ or $(M_k,M_j) \in
+  \preccurlyeq$.
+\end{lemma}
+\begin{theorem} \label{thm:partitionold} $\preccurlyeq$ is a partial
+  order on the set of prime sub-matrices of $M$.  Further, it uniquely
+  partitions the prime sub-matrices of $M$ such that on each set in
+  the partition $\preccurlyeq$ induces a total order.
+\end{theorem}
+For the purposes of this paper, we refine the total order mentioned in
+Theorem \ref{thm:partitionold}. We do this by identifying an in-tree
+rooted at each maximal upper bound under $\preccurlyeq$.  Each of
+these in-trees will be on disjoint vertex sets, which in this case
+would be disjoint sets of prime-submatrices.  The in-trees are
+specified by selecting the appropriate edges from the Hasse diagram
+associated with $\preccurlyeq$.  Let $\cI$ be the following set:
+\begin{align*}
+  \cI = \{ (M_i,M_j) \in \preccurlyeq \mid \nexists M_k s.t. M_i
+  \preccurlyeq M_k, M_k \preccurlyeq M_j \} \cup \{ (M_i,M_i), i \in
+  [p] \}
+\end{align*}
+
+\begin{theorem} \label{thm:partition} Consider the directed graph $X$
+  whose vertices correspond to the prime sub-matrices, and the edges
+  are given by $\cI$.  Then, $X$ is a vertex disjoint collection of
+  in-trees and the root of each in-tree is a maximal upper bound in
+  $\preccurlyeq$.
+\end{theorem}
+\begin{proof}
+  To observe that $X$ is a collection of in-trees, we observe that for
+  vertices corresponding to maximal upper bounds, no out-going edge is
+  present in $I$.  Secondly, for each other element, exactly one
+  out-going edge is chosen, and for the minimal lower bound, there is
+  no in-coming edge.  Consequently, $X$ is acyclic, and since each
+  vertex has at most one edge leaving it, it follows that $X$ is a
+  collection of in-trees, and for each in-tree, the root is a maximal
+  upper bound in $\preccurlyeq$.  Hence the theorem.
+\end{proof}
+Let the partition of $X$ given by Theorem \ref{thm:partition} be
+$\{X_1,\ldots,X_r\}$.  Further, each in-tree itself can be layered
+based on the distance from the root.  The root is considered to be at
+level zero. For $j \geq 0$, Let $X_{i,j}$ denote the set of prime
+matrices in level $j$ of in-tree $X_i$.
+
+\begin{lemma}
+  \label{lem:subicppa}
+  Let $M$ be a matrix and let $X$ be the directed graph whose vertices
+  are in correspondence with the prime submatrices of $M$.  Further
+  let $\{X_1,\ldots,X_r\}$ be the partition of $X$ into in-trees as
+  defined above.  Then, matrix $M$ has an ICPPA in tree $T$ iff $T$
+  can be partitioned into vertex disjoint subtrees $\{T_1, T_2, \dots
+  T_r\}$ such that, for each $1 \leq i \leq r$, the set of prime
+  sub-matrices corresponding to vertices in $X_i$ has an ICPPA in
+  $T_i$.
+\end{lemma}
+\begin{proof}
+  Let us consider the reverse direction first.  Let us assume that $T$
+  can be partitioned into $T_1, \ldots, T_r$ such that for each $1
+  \leq i \leq r $, the set of prime sub-matrices corresponding to
+  vertices in $X_i$ has an ICPPA in $T_i$.  It is clear from the
+  properties of the partial order $\preccurlyeq$ that these ICPPAs
+  naturally yield an ICPPA of $M$ in $T$.  The main property used in
+  this inference is that for each $1 \leq i \neq j \leq r$, $supp(X_i)
+  \cap supp(X_j) = \phi$.
+
+\noindent
+To prove the forward direction, we show that if $M$ has an ICPPA, say
+$\cA$, in $T$, then there exists a partition of $T$ into vertex
+disjoint subtree $T_1, \ldots, T_r$ such that for each $1 \leq i \leq
+r$, the set of prime sub-matrices corresponding to vertices in $X_i$
+has an ICPPA in $T_i$.  For each $1 \leq i \leq r$, we define based on
+$\cA$ a subtree $T_i$ corresponding to $X_i$.  We then argue that the
+trees thus defined are vertex disjoint, and complete the proof.
+Consider $X_i$ and consider the prime sub-matrix in $X_{i,0}$.
+Consider the paths assigned under $\cA$ to the sets in the prime
+sub-matrix in $X_{i,0}$.  Since the component in $G_f$ corresponding
+to this matrix is a connected component, it follows that union of
+paths assigned to this prime-submatrix is a subtree of $T$.  We call
+this sub-tree $T_i$.  All other prime-submatrices in $X_i$ are
+assigned paths in $T_i$ since $\cA$ is an ICPPA, and the support of
+other prime sub-matrices in $X_i$ are contained in the support of the
+matrix in $X_{i,0}$.  Secondly, for each $1 \leq i \neq j \leq r$,
+$supp(X_i) \cap supp(X_j) = \phi$, and since $\cA$ is an ICPPA, it
+follows that $T_i$ and $T_j$ are vertex disjoint.  Finally, since $|U|
+= |V(T)|$, it follows that $T_1, \ldots, T_r$ is a partition of $T$
+into vertex disjoint sub-trees such that for each $1 \leq i \leq r$,
+the set of matrices corresponding to nodes in $X_i$ has an ICPPA in
+$T_i$.  Hence the lemma.
+\end{proof}
+The essence of the following lemma is that an ICPPA only needs to be
+assigned to the prime sub-matrix corresponding to the root of each
+in-tree, and all the other prime sub-matrices only need to have an
+Intersection Cardinality Preserving Interval Assignments (ICPIA).
+Recall, an ICPIA is an assignment of intervals to sets such that the
+cardinality of an assigned interval is same as the cardinality of the
+interval, and the cardinality of intersection of any two sets is same
+as the cardinality of the intersection of the corresponding intervals.
+It is shown in \cite{nsnrs09} that the existence of an ICPIA is a
+necessary and sufficient condition for a matrix to have COP.  We
+present the pseudo-code to test if $M$ has an ICPPA in $T$.
+\begin{lemma} \label{lem:rooticppa} Let $M$ be a matrix and let $X$ be
+  the directed graph whose vertices are in correspondence with the
+  prime submatrices of $M$.  Further let $\{X_1,\ldots,X_r\}$ be the
+  partition of $X$ into in-trees as defined earlier in this section.
+  Let $T$ be the given tree and let $\{T_1, \ldots, T_r\}$ be a given
+  partition of $T$ into vertex disjoint sub-trees.  Then, for each $1
+  \leq i \leq r$, the set of matrices corresponding to vertices of
+  $X_i$ has an ICPPA in $T_i$ if and only if the matrix in $X_{i,0}$
+  has an ICPPA in $T_i$ and all other matrices in $X_i$ have an {\bf
+    {\em ICPIA}}.
+\end{lemma}
+\begin{proof}
+  The proof is based on the following fact- $\preccurlyeq$ is a
+  partial order and $X$ is a digraph which is the disjoint union of
+  in-trees.  Each edge in the in-tree is a containment relationship
+  among the supports of the corresonding sub-matrices. Therefore, any
+  ICPPA to a prime sub-matrix that is not the root is contained in a
+  path assigned to the sets in the parent matrix.  Consequently, any
+  ICPPA to the prime sub-matrix that is not at the root is an ICPIA,
+  and any ICPIA can be used to construct an ICPPA to the matrices
+  corresponding to nodes in $X_i$ provided the matrix in the root has
+  an ICPPA in $T_i$.  Hence the lemma.
+\end{proof}
+Lemma \ref{lem:subicppa} and Lemma \ref{lem:rooticppa} point out two
+algorithmic challenges in finding an ICPPA for a given set system
+$\cF$ in a tree $T$.  Given $\cF$, finding $X$ and its partition
+$\{X_1,\ldots,X_r\}$ into in-trees can be done in polynomial time.  On
+the other hand, as per lemma \ref{lem:subicppa} we need to parition
+$T$ into vertex disjoint sub-trees $\{T_1, \ldots, T_r\}$ such that
+for each $i$, the set of matrices corresponding to nodes in $X_i$ have
+an ICPPA in $T_i$.  This seems to be a challenging step, and it must
+be remarked that this step is easy when $T$ itself is a path, as each
+individual $T_i$ would be sub-paths.  The second algorithmic challenge
+is identified by lemma \ref{lem:rooticppa} which is to assign an ICPPA
+from a given tree to the matrix associated with the root node of
+$X_i$.
+\begin{algorithm}[h]
+  \caption{Algorithm to find an ICPPA for a matrix $M$ on tree $T$:
+    $main\_ICPPA(M, T$)}
+  \label{al:icppa-main}
+  \begin{algorithmic}
+    \STATE Identify the prime sub-matrices. This is done by constructing the strict overlap graph and identify connected components.  Each connected component yields a prime sub-matrix.   \\
+    \STATE Construct the partial order $\preccurlyeq$ on the set of prime sub-matrices.  \\
+    \STATE Construct the partition $X_1,\ldots,X_r$ of the prime
+    sub-matrices induced by $\preccurlyeq$ \\
+    \STATE For each $1 \leq i \leq r$, Check if all matrices except
+    those in $X_{i,0}$ has an ICPIA.  If a matrix does not have ICPIA
+    exit with a negative answer.  To check for the existence of ICPIA,
+    use the result in \cite{nsnrs09}.  \STATE Find a partition of
+    $T_1, \ldots, T_r$ such that matrices in $X_{i,0}$ has an ICPPA in
+    $T_i$.  If not such partition exists, exit with negative answer.
+  \end{algorithmic}
+\end{algorithm}
+
+\subsection{Complexity of Tree Path Assignment-A
+  Discussion} 
+\label{complexity} 
+Recall that the input to the Tree
+Path Assignment question is an order pair $(\cF,T)$ where $\cF$ is a
+family of subsets of an universe $U$, and $T$ is a tree such that
+$|V(T)| = |U|$.  The question is to come up with a bijection from $U$
+to $V(T)$ such that the image of each set in $\cF$ is a path in $T$.
+We show that this problem is at least as hard as the problem of
+testing if two given path chordal graphs are isomorphic.
+\begin{theorem} \label{thm:isocomp} Tree Path Assignment is
+  isomorphism-complete.
+\end{theorem}
+\begin{proof}
+  It is well known (see for example \cite{kklv10}) that testing
+  isomorphism of path chordal graphs is isomorphism complete.  We show
+  a reduction of path chordal graph isomorphism to tree path
+  assignment.  Given $G_1$ and $G_2$ two path chordal graphs, let
+  $T_2$ be the clique tree of $G_2$ obtained from say \cite{gav78}.
+  The nodes of $T_2$ correspond to the maximal cliques of $G_2$ and
+  each vertex of $G_2$ corresponds to a path in $G_2$.  This is a
+  well-known characterization of path chordal graphs and $T_2$ can be
+  computed in polynomial time.  In $G_1$, let $S_v$ denote the maximal
+  cliques of $G_1$ that contain $v$.  This can be computed in
+  polynomial time as $G_1$ is path chordal, and all chordal graph only
+  have a linear number of maximal cliques.  The universe $U$
+  corresponds bijectively to the set of maximal cliques in $G_1$, and
+  $\cF = \{S_v | v \in V(G_1)\}$.  Now, we claim that $(\cF, T_2)$ has
+  a tree path assignment if and only if $G_1$ and $G_2$ are
+  isomorphic.  This is clear since for each vertex $v \in G_1$, there
+  is an associated $S_v$ which is the set of maximal cliques
+  containing $v$.  In $G_2$, each vertex corresponds to a path in
+  $T_2$, and the nodes on this path corresponds to the maximal cliques
+  in $G_2$.  Consequenlty, a tree path assignment will naturally yield
+  an isomorphism between $G_1$ and $G_2$, and vice versa.  Therefore,
+  Tree Path Assignment is isomorphism-complete.
+\end{proof}
+
+\subsubsection{Consecutive Ones Testing is in Logspace}
+While Tree Path Assignment is isomorphism-complete, it is polynomial
+time solvable when the given tree is a path.  Indeed, in this case we
+encounter a restatement of matrices with the COP.  The known
+approaches to testing for COP fall into two categories: those that
+provide a witness when the input matrix does not have COP, and those
+that do not provide a witness.  The first linear time algorithm for
+testing COP for a binary matrix was using a data structure called PQ
+trees, which represent all COP orderings of $M$, invented by
+\cite{bl76}. There is a PQ tree for a matrix iff the matrix has COP.
+Indeed, this is an algorithmic characterization of the consecutive
+ones property and the absence of the PQ-tree does not yield any
+witness to the reason for failure.  A closely related data structure
+is the generalized PQ tree in \cite{mcc04}.  In generalized PQ tree
+the P and Q nodes are called prime and linear nodes. Aside from that,
+it has a third type of node called degenerate nodes which is present
+only if the set system does not have COP \cite{mcc04}.  Using the idea
+of generalized PQ tree, \cite{mcc04} proves that checking for
+bipartiteness in the certain incomparability graph is sufficient to
+check for COP.  \cite{mcc04} invented a certificate to confirm when a
+binary matrix does not have COP.  \cite{mcc04} describes a graph
+called incompatibility graph of a set system $\cF$ which has vertices
+$(a,b), a \ne b$ for every $a, b \in U$, $U$ being the universe of the
+set system. There are edges $((a,b),(b,c))$ and $((b,a),(c,b))$ if
+there is a set $S \in \cF$ such that $a, c \in S$ and $b \notin S$. In
+other words the vertices of an edge in this graph represents two
+orderings that cannot occur in a consecutive ones ordering of $\cF$.
+\begin{theorem}[Theorem 6.1, \cite{mcc04}]
+  Let $\cF$ be an arbitrary set family on domain $V$. Then $\cF$ has
+  the consecutive ones property if and only if its incompatibility
+  graph is bipartite, and if it does not have the consecutive ones
+  property, the incompatibility graph has an odd cycle of length at
+  most $n+3$.
+\end{theorem}
+This theorem gives a certificate as to why a given matrix does not
+have COP.  Similarly, the approach of testing for an ICPIA in
+\cite{nsnrs09} also gives a different certificate- a prime sub-matrix
+that does not have an ICPIA.  Further, the above theorem can be used
+to check if a given matrix has COP in logspace by checking if its
+incompatibility graph is bipartite. \cite{rei84} showed that checking
+for bipartiteness can be done in logspace. Thus we conclude that
+consecutive ones testing can be done in logspace.
+
+\noindent
+More recently, \cite{kklv10} showed that interval graph isomorphism
+can be done in logspace. Their paper proves that a canon for interval
+graphs can be calculated in logspace using an interval hypergraph
+representation of the interval graph with each hyperedge being a set
+to which an interval shall be assigned by the canonization
+algorithm. An overlap graph (subgraph of intersection graph, edges
+define only strict intersections and no containment) of the hyperedges
+of the hypergraph is created and canons are computed for each overlap
+component. The overlap components define a tree like relation due to
+the fact that two overlap components are such that either all the
+hyperedges of one is disjoint from all in the other, or all of them
+are contained in one hyperedge in the other. This is similar to the
+containment tree defined in \cite{nsnrs09} and in this paper. Finally
+the canon for the whole graph is created using logspace tree
+canonization algorithm from \cite{sl92}. The interval labelling done
+in this process of canonization is exactly the same as the problem of
+assigning feasible intervals to a set system, and thus the problem of
+finding a COP ordering in a binary matrix \cite{nsnrs09}.
+
+\begin{theorem}[Theorem 4.7, \cite{kklv10}]
+  \label{th:canonlabel}
+  Given an interval hypergraph $\cH$, a canonical interval labeling
+  $l_H$ for $H$ can be computed in FL.
+\end{theorem}
+
+We present the following reduction to see that COP testing is indeed
+in logspace. Given a binary matrix $M$ of order $n \times m$, let $S_i
+= \{j \mid M[j,i]=1 \}$. Let $\cF = \{S_i \mid i \in [m] \}$ be this
+set system. Construct a hypergraph $\cH$ with its vertex set being
+$\{1, 2, \dots n\}$. The edge set of $\cH$ is isomorphic to
+$\cF$. Thus every edge in $\cH$ represents a set in the given set
+system $\cF$. Let this mapping be $\pi: E(\cH) \rightarrow \cF$. It is
+easy to see that if $M$ has COP, then $\cH$ is an interval
+hypergraph. From theorem \ref{th:canonlabel}, it is clear that the
+interval labeling $l_{\cH}: V(\cH) \rightarrow [n]$ can be calculated
+in logspace. Construct sets $I_i = \{ \l_{\cH}(x) \mid x \in E, E \in
+E(\cH), \pi(E) = S_i\}$, for all $i \in [m]$. Since $\cH$ is an
+interval hypergraph, $I_i$ is an interval for all $i \in [m]$, and is
+the interval assigned to $S_i$ if $M$ has COP.
+
+Now we have the following corollary.
+\begin{corollary}
+  \label{cor:coplog}
+  If a binary matrix $M$ has COP then the interval assignments to each
+  of its columns can be calculated in FL.
+\end{corollary}
+
+\noindent
+Finally, we conclude by asking about the complexity of Tree Path
+Assignment restricted to other subclasses of trees.  In particular, is
+Tree Path Assignment in caterpillars easier than Tree Path assignment
+in general trees.
 
 \chapter{Conclusion}
-% \section{Conclusion}
-\label{sec:conclusion}
-%\vspace{\secfirstparatrim}
+\label{ch:conclusion}
+
+\begin{center}
+{\tt MOVED FROM INTRO SUBSECTION IN ch:myresearch CHAPTER begin}
+\end{center}
+
+% \annote{
+  It is an interesting fact that for a matrix with the COP, the
+  intersection graph of the corresponding set system is an interval
+  graph. % }{the following text could be in Further Study/Conclusion}
+A
+similar connection to a subclass of chordal graphs and a superclass of
+interval graphs exists for the generalization of COP.  In this case,
+the intersection graph of the corresponding set system must be a {\em
+  path graph}. Chordal graphs are of great significance, are
+extensively studied, and have several applications.  One of the well
+known and interesting properties of a chordal graphs is its connection
+with intersection graphs \cite{mcg04}. For every chordal graph, there
+exists a tree and a family of subtrees of this tree such that the
+intersection graph of this family is isomorphic to the chordal graph
+\cite{plr70,gav78,bp93}.  These trees when in a certain format, are
+called {\em clique trees} \cite{apy92} of the chordal graph. This is a
+compact representation of the chordal graph. There has also been work
+done on the generalization of clique trees to clique hypergraphs
+\cite{km02}.  If the chordal graph can be represented as the
+intersection graph of paths in a tree, then the graph is called path
+graph \cite{mcg04}.  Therefore, it is clear that if there is a
+bijection from $U$ to $V(T)$ such that for every set, the elements in
+it map to vertices of a unique path in $T$, then the intersection
+graph of the set system is indeed a path graph.  However, this is only
+a necessary condition and can be checked efficiently because path
+graph recognition is polynomial time
+solvable\cite{gav78,aas93}. Indeed, it is possible to construct a set
+system and tree, such that the intersection graph is a path graph, but
+there is no bijection between $U$ and $V(T)$ such that the sets map to
+paths. Path graph isomorphism is known be isomorphism-complete, see
+for example \cite{kklv10}. An interesting area of research would be to
+see what this result tells us about the complexity of the tree path
+labeling problem (not covered in this paper).
+
+\begin{center}
+{\tt MOVED FROM INTRO SUBSECTION IN ch:myresearch CHAPTER\\ end}
+\end{center}
 
@@
 $\cF$ in \FTPL. However, one can easily obtain a counterexample to
-show the insufficiency of this condition\tnote[GTC]{Give a figure
+show the insufficiency of this condition\tnote[TH]{Give a figure
   showing insuff.}.  Path graph isomorphism is known be
-isomorphism-complete\cite{kklv10}. Therefore, \annote[GTC]{it is unlikely
-  that we can solve the problem of finding feasible path labeling
-  $\cl$ for a given $\cF$ and }{WHY?} tree $T$.  It is
-  definitely interesting to classify the kinds of trees and
-  hypergraphs for which feasible path labelings can be\tnote[TH]{ELABORATE.}
-found efficiently.  \annote[TH]{These results would form a natural
+isomorphism-complete\cite{kklv10}. Therefore, \annote[TH]{it is
+  unlikely that we can solve the problem of finding feasible path
+  labeling $\cl$ for a given $\cF$ and }{WHY?} tree $T$.  It is
+definitely interesting to classify the kinds of trees and hypergraphs
+for which feasible path labelings can be\tnote[TH]{ELABORATE.}  found
+efficiently.  \annote[TH]{These results would form a natural
   generalization of COP testing and interval graph isomorphism,
-  culminating in Graph Isomorphism }{REWORD} itself.
- To this effect we consider TPL on {\kstar} and give a polynomial time
- solution for the same.
-
-Whether TPL on general trees is solvable in $P$ remains open. So do
-optimization opportunities in TPL (possible extensions to optimization for COP in the
+  culminating in Graph Isomorphism }{REWORD} itself.  To this effect
+we consider TPL on {\kstar} and give a polynomial time solution for
+the same.
+
+While we address the computation of TPL
+for a given hypergraph% \footnote{The term {\em hypergraph} is used
+%   instead {\em set systems} are used interchangably in this
+%   section. Hypergraphs that have TPLs on some target tree are called
+%   {\em path hypergraphs}.}
+from a given target tree in this research, however optimization
+problems in TPL, akin to problems in Section~\ref{sec:optcop} for COP,
+is outside the scope of this thesis. Whether TPL on general trees is
+solvable in $P$ remains open. So do optimization opportunities in TPL
+(possible extensions to optimization for COP in the
 Section~\ref{sec:optcop}.\tnote[TH]{link to section}).
 
-\tnote[TH]{ ||Complexity challenges:|| Canonization is an important tool in graph isomorphism. Invention of a
+%\newpage %DUMMY
+\tnote[TH]{ ||Complexity challenges:|| Canonization is an important
+  tool in graph isomorphism. Invention of a
 % deterministic method of canonization for any class of graphs naturally
@@
 % \cite{rei08}, \cite{kklv10} proved their logspace result.
-}%
+}   %
 \tnote[TH]{ (1) the interesting survey in conclusion section of kklv10
@@
 
-\appendix
-
-\chapter{Appendix A}
+% \appendix %%% NON NEIN NAHI ILLA NONE
+
+% \chapter{A}
+% \tnote{DO WE NEED THIS?}
 
@@
 %\begin{thebibliography}{10}
-\addcontentsline{toc}{section}{Bibliography}
+\addcontentsline{toc}{chapter}{REFERENCES}
 
@@
 \end{singlespace}
-\newpage
+%\newpage
 
@@
 
-\listofpapers
-
-\begin{enumerate}  
-\item Authors....  \newblock
- Title...
-  \newblock {\em Journal}, Volume,
-  Page, (year).
-\end{enumerate}
+% \listofpapers %%% NON NEIN NAHI ILLA NONE
+
+% \begin{enumerate}  
+% \item Authors....  \newblock
+%  Title...
+%   \newblock {\em Journal}, Volume,
+%   Page, (year).
+% \end{enumerate}
 

### 2011-12-31 src/thesis.tex
Nightly commit (yesterdays). Changes are predominantly for chapter 3 - (ch:myresearch)

@@
 
-\subsection{********* COP in Graph Isomorphism}
+\section{********* COP in Graph Isomorphism}
 \label{sec:surveygraphiso}
@@
 
-\subsection{********* Certifying Algorithms}
+\section{********* Certifying Algorithms}
 \label{sec:surveycertalgo}
@@
 \tnote{UPDATE TO FINAL VERSION. BELOW IS OLDER VERSION WITH INCOMPLETE
-  SEC 4.}  This chapter documents all the new results obtained by us in the
-area of tree path labeling of path hypergraphs which is the parent
-problem addressed in this thesis. In Section~\ref{sec:motive} we see
-that consecutive-ones property and its equivalent problem of interval
-labeling of a hypergraph is a special case of the general problem of tree
-path labeling of path hypergraphs.
-
-\tnote{UPDATE the following para to current skeleton.}The necessary
-preliminaries with definitions etc. are presented in
-Section~\ref{ch:prelims}. Section~\ref{sec:feasible} documents the
-characterization of a feasible path labeling for any path labeling
-from any target trees. Section~\ref{sec:spltargettree} describes the
-special case where the target tree is of a particular family of trees
-and specifically, Section~\ref{sec:ksubdivstar} describes a polynomial
-time algorithm to find the tree path labeling of a given set system
-from a given $k$-subdivided tree. Section~\ref{sec:norestraint}
-describes the general situation where the target tree has no
-restrictions and the algorithm to find a TPL, if any, in this case.
+  SEC 4.}  This chapter documents all the new results obtained by us
+in the area of tree path labeling of path hypergraphs which is the
+parent problem addressed in this thesis.
+
+Section~\ref{sec:myresearchintro} recalls the idea of tree path
+labeling. The necessary preliminaries with definitions etc. are
+presented in Section~\ref{ch:prelims}. Section~\ref{sec:feasible}
+documents the characterization of a feasible path labeling for any
+path labeling from any target trees. Section~\ref{sec:spltargettree}
+describes two special cases where the target tree is of a particular
+family of trees. The first one is shown to be equivalent to COP
+testing in Section~\ref{sec:icpplicpia}. Section~\ref{sec:ksubdivstar}
+discusses the second special case and presents a polynomial time
+algorithm to find the tree path labeling of a given set system from a
+given \kstar. Section~\ref{sec:norestraint} discusses the plain
+vanilla version where the target tree has no restrictions and the
+algorithm to find a feasible TPL, if any, in this case.
 
 \section{Introduction} 
-
-The problem of consecutive-ones property testing can be easily seen as
-a simple constraint satisfaction problem involving a hypergraph or a
-system of sets from a universe. Every column (row) of the binary
-matrix can be converted into a set of non-negative integers which are
-the indices of the rows (columns) with {\un}s in that column
+\label{sec:myresearchintro}
+In Section~\ref{sec:motive} we see that consecutive-ones property and
+its equivalent problem of interval labeling of a hypergraph is a
+special case of the general problem of tree path labeling of path
+hypergraphs.  The problem of consecutive-ones property testing can be
+easily seen as a simple constraint satisfaction problem involving a
+hypergraph or a system of sets from a universe. Every column (row) of
+the binary matrix can be converted into a set of non-negative integers
+which are the indices of the rows (columns) with {\un}s in that column
 (row). When observed in this context, if the matrix has the COP on
@@
 assignments to the sets which can be obtained from a single
-permutation of the rows (columns). They show that for each set, the
-cardinality of the interval assigned to it must be same as the
-cardinality of the set, and the intersection cardinality of any two
-sets must be same as the intersection cardinality of the corresponding
-intervals. This is a necessary and sufficient
-condition. \tnote{(ICPIA) put reference to survey chapter}.  Finally,
-the idea of decomposing a given binary matrix into prime matrices to
-check for COP is adopted from cite|wlh02| to test if an ICPIA exists
-for a given set system.\tnote{REPEATITION from intro chapter. but idea
-  recall is necessary. so keep it but reword it to avoid monotone.}
-
-A natural generalization of the interval assignment problem is
-feasible tree path labeling problem of a set system. The problem is
-defined as follows -- given a set system $\cF$ from a universe $U$ and
-a tree $T$, does there exist a bijection from $U$ to the vertices of
-$T$ such that each set in the system maps to a path in $T$.  We refer
-to this as the {\em tree path labeling problem} for an input set
-system, target tree pair -- $(\cF,T)$. \tnote{USE THE FORMAL PROBLEM DEF BLOCK}
+permutation of the rows (columns). The cardinality of the interval
+assigned to it must be same as the cardinality of the set, and the
+intersection cardinality of any two sets must be same as the
+intersection cardinality of the corresponding intervals. This is a
+necessary and sufficient condition. \tnote{(ICPIA) put reference to
+  survey chapter}
+
+Naturally, intervals are paths from a tree with maximum degree
+two. Thus the interval assignment problem can be generalized into path
+assignment problem from any tree. We refer to this as the {\em tree
+  path labeling problem}. To elaborate, the problem is defined as
+follows -- given a hypergraph $\cF$ from universe (vertex set) $U$ and
+a target tree $T$, does there exist a bijection from $U$ to the
+vertices of $T$ such that when applied to elements of each hyperedge,
+it gives a path on $T$.  More formally, the problem definition is as
+defined by \CFTPL.
+
+{\small
+  \begin{minipage}[h]{5in}
+    \vspace{2mm}
+    {\large \CFTPL}\\
+    \begin{tabular}[t]{l|l}
+      \hline\\
+      {\tt Input} & 
+      \begin{minipage}[t]{\probdefwidth}
+        A hypergraph $\cF$ with vertex set $U$ and a tree $T$.\\
+      \end{minipage}\\
+
+      {\tt Question} &
+      \begin{minipage}[t]{\probdefwidth}
+        Does there exist a set of paths $\cP$ from $T$ and a bijection
+        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
+        {\bf true} on $(\cF, T, \cl)$.
+      \end{minipage}\\
+    \end{tabular}
+  \end{minipage}\\
+}
+
+To characterize a feasible TPL, we consider the case were a tree path
+labeling is also given as input and we are required to test if the
+given labeling is feasible. This is defined by the \FTPL problem.
+
 %\begin{tabular}[t]{c}
@@
 
-As a special case if the tree
-$T$ is a path, the problem becomes the interval assignment problem. \tnote{USE THE FORMAL PROBLEM DEF BLOCK}
-We focus on the question of generalizing the notion of an ICPIA
-\cite{nsnrs09} to characterize feasible path assignments.  We show
+Section~\ref{sec:feasible} discusses \FTPL and presents an algorithmic
+characterization for a feasible TPL.
+
+With respect to computing a feasible TPL, as suggested by \CFTPL
+problem, we were unable to discover an efficient algorithm for
+it. Hence we consider two special cases of \CFTPL -- namely, \CFTPLINT
+and \CFTPLKTREE on special target trees, intervals and \kstar, respectively.
+Section~\ref{sec:spltargettree}
+discusses these problems.
+
+{\small
+  \begin{minipage}[h]{5in}
+    % \begin{singlespace}
+    \vspace{2mm}
+    {\large \CFTPLINT}\\
+    \begin{tabular}[t]{l|l}
+      \hline\\
+      {\tt Input} & 
+      \begin{minipage}[t]{\probdefwidth}
+        A hypergraph $\cF$ with vertex set $U$ and a tree $T$ with
+        maximum degree 2.\\
+      \end{minipage}\\
+      {\tt Question} &
+      \begin{minipage}[t]{\probdefwidth}
+        Does there exist a set of paths $\cP$ from $T$ and a bijection
+        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
+        {\bf true} on $(\cF, T, \cl)$.
+      \end{minipage}\\
+    \end{tabular}
+    % \end{singlespace}
+  \end{minipage}\\
+}
+
+{\small
+  \begin{minipage}[h]{5in}
+    % \begin{singlespace}
+    \vspace{2mm}
+    {\large \CFTPLKTREE}\\
+    \begin{tabular}[t]{l|l}
+      \hline\\
+      {\tt Input} & 
+      \begin{minipage}[t]{\probdefwidth}
+        A hypergraph $\cF$ with vertex set $U$ such that every
+        hyperedge
+        $S \in \cF$ is of cardinality at most $k+2$ and a {\kstar} $T$.\\
+      \end{minipage}\\
+      {\tt Question} &
+      \begin{minipage}[t]{\probdefwidth}
+        Does there exist a set of paths $\cP$ from $T$ and a bijection
+        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
+        {\bf true} on $(\cF, T, \cl)$.
+      \end{minipage}\\
+    \end{tabular}
+    % \end{singlespace}
+  \end{minipage}\\
+}
+
+\CFTPLINT is nothing but the
+consecutive-ones property testing problem.
+
+An algorithm for \CFTPL on general trees which is less efficient than
+polynomial time is presented in
+Section~\ref{sec:norestraint}. \tnote{put some more description
+  here. refer walcom.}
+
+\remove{We focus on the question of generalizing the notion of an ICPIA
+cite-nsnrs09 to characterize feasible path assignments.  We show
 that for a given set system $\cF$, a tree $T$, and an assignment of
@@
 algorithm to test if a tree path labeling to the set system is
-feasible.  This generalizes the result in \cite{nsnrs09}.
-
-In the later part of this paper, we focus on a new special case of the
-tree path labeling problem. Here the set system is such that every set
-is at most $k+2$ in size and for every pair of sets in it there exists
-a sequence of sets between them with consecutive sets in this sequence
-having a strict intersection -- i.e., non-empty intersection with
-neither being contained in the other. Moreover, the given tree is a
-{\em $k$-subdivided star}. We demonstrate a polynomial time algorithm
-to find a feasible path labeling in this case.
+feasible.  This generalizes the result in cite-nsnrs09.}
 
@@
 
-\tnote{TBD: have a few lines about organization of chapter}This
-section states definitions and basic facts necessary in the scope of
-this document.
-
-\tnote{Move some of this to general prelim in chapter 1}
-
-The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
-  system} of a universe $U$ with $|U| = n$.  The {\em support} of a
-set system $\F$ denoted by $supp(\cF)$ is the union of all the sets in
-$\F$; $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this
-paper, a set system is required to ``cover'' the universe; $
-supp(\cF) = U$.
-
-The graph $T$ represents a {\em target tree} with same
-number of vertices as elements in $U$; $|V(T)| = n$.  A {\em path
-  system}\, $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P
-\mid P \subseteq V, \text{ } T[P] \text{ is a path} \}$.
-
-% 
-% A set system $\cF$ can be alternatively represented by a {\em
-%   hypergraph} $\H_\cF$ whose vertex set is $supp(\cF)$ and
-% hyperedges are the sets in $\cF$. This is a known representation for
-% interval systems in literature \cite{bls99,kklv10}.  We extend this
-% definition here to path systems.
-
-A set system $\cF$ can be alternatively represented by a {\em
+This section states definitions and basic facts necessary in the scope
+of this document.
+
+{\bf Hypergraph Preliminaries:} The set $\F \subseteq (2^{U} \setminus
+\emptyset)$ is a {\em set system} of a universe $U$ with $|U| = n$.
+The {\em support} of a set system $\F$ denoted by $supp(\cF)$ is the
+union of all the sets in $\F$; $supp(\F) = \bigcup_{S \in \F}S$. For
+the purposes of this paper, a set system is required to ``cover'' the
+universe; $ supp(\cF) = U$.  A set system $\cF$ can also be visualized
+as a {\em hypergraph}\, vertex set is $supp(\cF)$ and hyperedges are
+the sets in $\cF$.  The {\em intersection graph}\, $\bI(\cF)$ of a
+hypergraph $\cF$ is a graph such that its vertex set has a bijection
+to $\cF$ and there exists an edge between two vertices iff their
+corresponding hyperedges have a non-empty intersection \cite{mcg04}.
+
+Two hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to
+each other, denoted by $\cF' \cong \cF''$, iff there exists a
+bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for all
+sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ iff $B$ is
+a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
+\cite{kklv10}, written as $B=\phi(A)$.
+This is called {\em hypergraph isomorphism}.
+
+\add{A set system $\cF$ can be alternatively} represented by a {\em
   hypergraph}\, $\cF_H$ whose vertex set is $supp(\cF)$ and hyperedges
@@
 
-Two hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to
-each other, denoted by $\cF' \cong \cF''$, iff there exists a
-bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for all
-sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ iff $B$ is
-a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
-\cite{kklv10}. This is called {\em hypergraph
-  isomorphism}. \tnote{also extend $\phi$ to hyperedges -- see if
-  required }
-
-The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph $\cF$ is a
-graph such that its vertex set has a bijection to $\cF$ and there
-exists an edge between two vertices iff their corresponding hyperedges
-have a non-empty intersection \cite{mcg04}.
-
-If the intersection graphs of two hypergraphs are isomorphic,
-$\bI(\cF) \cong \bI(\cP)$ where $\cP$ is also a path system, then the
-bijection $\cl: \cF \rightarrow \cP$ due to this isomorphism is called
-a {\em path labeling} of the hypergraph $\cF$. To illustrate further,
-let $\cg: V(\cF) \rightarrow V(\cP)$ be the above mentioned
-isomorphism where $V(\cF)$ and $V(\cP)$ are the vertex sets that
-represent the hyperedges for each hypergraph respectively, $V(\cF) =
-\{ v_S \mid S \in \cF\}$ and $V(\cP) = \{ v_P \mid P \in \cP\}$. Then
-the path labeling $\cl$ is defined as follows: $\cl(S_1) = P_1$ iff
-$\cg (v_{S_1}) = v_{P_1}$. The path system $\cP$ may be alternatively
-denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In most scenarios in
-this paper, what is given are the pair $(\cF, \cl)$ and the target
-tree $T$; hence this notation will be used more often.
+{\bf Path Hypergraph from a Tree:} The graph $T$ represents a {\em
+  target tree} with same number of vertices as elements in $U$;
+$|V(T)|=|U|=n$.  A {\em path system}\, $\cP$ is a set system of paths
+from $T$; $\cP \subseteq \{P \mid P \subseteq V, \text{ } T[P] \text{
+  is a path} \}$.  This generalizes the fact, from the literature
+\cite{bls99,kklv10}, that intervals can be viewed as sub-paths of a
+path.
+
+\add{Due to the equivalence of set system and hypergraph in the scope of
+this paper, we drop the subscript $_H$ in the notation and refer to
+both the structures by $\cF$.}
+
+If the intersection graphs of $\cF$ and $\cP$ (a path system) are
+isomorphic, $\bI(\cF) \cong \bI(\cP)$, then the associated bijection
+$\cl: \cF \rightarrow \cP$ due to this isomorphism is called a {\em
+  path labeling} of the hypergraph $\cF$.  Note that there are two
+kinds of isomorphisms here.  One is the isomorphism of
+intersection graphs on $\cF$ and $\cP$, i.e. $\bI(\cF)$ and $\bI(\cP)$
+respectively. Second is the isomorphism
+between the hypergraphs $\cF$ and $\cP$.
+
+\add{To illustrate further, let $\cg: V(\cF) \rightarrow V(\cP)$ be the
+above mentioned isomorphism where $V(\cF)$ and $V(\cP)$ are the
+vertex sets that represent the hyperedges for each hypergraph
+respectively, $V(\cF) = \{ v_S \mid S \in \cF\}$ and $V(\cP) = \{
+v_P \mid P \in \cP\}$. Then the path labeling $\cl$ is defined as
+follows: $\cl(S_1) = P_1$ iff $\cg (v_{S_1}) = v_{P_1}$.  Just to
+emphasize, for a path labeling $\cl$ of $\cF$ with $\cP$ as the path
+system, $\cF^\cl$ is same as $\cP$.  The path system $\cP$ may be
+alternatively denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In
+most scenarios in this paper, what is given are the pair $(\cF,
+\cl)$ and the target tree $T$; hence this notation will be used more
+often.}
 
@@
 
+\add{A graph} $G$ is a {\em path graph} if it is
+isomorphic to the intersection graph $\bI(\cP)$ of a path system
+$\cP$.  This isomorphism gives a bijection $\cl': V(G) \rightarrow
+\cP$. Moreover, for the purposes of this paper, we require that in a
+path labeling, $supp(\cP) = V(T)$.  If graph $G$ is also isomorphic
+to $\bI(\cF)$ for some hypergraph $\cF$, then clearly there is a
+bijection $\cl: \cF \rightarrow \cP$ such that $\cl(S) = \cl'(v_S)$
+where $v_S$ is the vertex corresponding to set $S$ in $\bI(\cF)$ for
+any $S \in \cF$. This bijection $\cl$ is called the {\em path
+  labeling} of the hypergraph $\cF$ and the path system $\cP$ may be
+alternatively denoted as $\cF^\cl$.
+
 A path labeling $(\cF, \cl)$ is defined to be {\em
@@
 \rightarrow \cF^\cl$ such that $\cl_\phi = \cl$. 
-
-An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is
-a graph such that its vertex set has a bijection to $\cF$ and there
-exists an edge between two of its vertices iff their corresponding
-hyperedges overlap. Two hyperedges $S$ and $S'$ are said to {\em
-  overlap}, denoted by $S \overlap S'$, if they have a non-empty
-intersection and neither is contained in the other; $S \overlap S'
-\text{ iff } S \cap S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq
-S$. Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$ and not
-necessarily connected. Each connected component of $\bO(\cF)$ is
-called an {\em overlap component}.
+\add{In this work}, we are given as input $\cF$ and a tree $T$, and the
+question is whether there is a path labeling $\cl$ to a set of paths
+in $T$.  We refer to such a solution path system by $\cF^\cl$.  A path
+labeling $(\cF, \cl)$ is defined to be {\em feasible} if
+% $\cF \cong \cF^\cl$ and this
+there is a hypergraph isomorphism $\phi: supp(\cF) \rightarrow
+supp(\cF^\cl)=V(T)$ induces a path labeling $\cl_\phi: \cF \rightarrow
+\cF^\cl$ such that $\cl_\phi = \cl$.
+
+{\bf Overlap Graphs and Marginal Hyperedges:} An {\em
+  overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is a graph such
+that its vertex set has a bijection to $\cF$ and there exists an edge
+between two of its vertices iff their corresponding hyperedges
+overlap. Two hyperedges $S$ and $S'$ are said to {\em overlap},
+denoted by $S \overlap S'$, if they have a non-empty intersection and
+neither is contained in the other; $S \overlap S' \text{ iff } S \cap
+S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq S$. Thus $\bO(\cF)$
+is a spanning subgraph of $\bI(\cF)$ and not necessarily
+connected. Each connected component of $\bO(\cF)$ is called an {\em
+  overlap component}.
 
@@
 \cite{kklv10}. Additionally, if $S$ is such that it is contained in no
-other hyperedge in $\cF$, i.e., it is inclusion maximal then it is called
-{\em super-marginal}.
-
-A {\em star} graph is a complete bipartite graph $K_{1,p}$
-which is clearly a tree and $p$ is the number of leaves. The vertex
-with maximum degree is called the {\em center} of the star and the
-edges are called {\em rays} of the star. A {\em $k$-subdivided star}
-is a star with all its rays subdivided exactly $k$ times. The
-definition of a {\em ray of a $k$-subdivided star} is extended to the path
-from the center to a leaf. It is clear that all rays are of length $k+2$.
+other marginal hyperedge in $\cF$, then it is called {\em
+  super-marginal}.
+
+{\bf $k$-subdivided star -- a special tree} A {\em star} graph
+is a complete bipartite graph $K_{1,p}$ which is clearly a tree and
+$p$ is the number of leaves. The vertex with maximum degree is called
+the {\em center} of the star and the edges are called {\em rays} of
+the star. A {\em $k$-subdivided star} is a star with all its rays
+subdivided exactly $k$ times. The definition of a {\em ray of a
+  $k$-subdivided star} is extended to the path from the center to a
+leaf. It is clear that all rays are of length $k+2$.
+
+{\tt --- WALCOM}
+
+A {\em path assignment} $\A$ to $\F$ is defined as a set assignment
+where second universe is the vertex set $V$ of a given tree $T$ and
+every second subset in the ordered pairs is a path in this
+tree. Formally, the definition is as follows.
+\begin{align*}
+  \A = \{ (S_i,P_i) \mid S_i \in \F, P_i \subseteq V \text{
+    s.t. }T[P_i] \text{ is a path, } i \in I \}
+\end{align*}
+In other words, $P_i$ is the path on the tree $T$ assigned to $S_i$ in
+$\A$. As mentioned before for set systems, the paths cover the whole
+tree, i.e. $\bigcup_{i \in I}P_i = V$ \\
+
+\noindent
+Generalizing the definition of {\em feasibility} in \cite{nsnrs09} to
+a set assignment, a path assignment $\A$ is defined to be {\em
+  feasible} if there exists a bijection defined as follows.
+\begin{align}
+  \sigma: U \rightarrow V(T), \text{ such that }\sigma(S_i) = P_i
+  \text{ for all } i \in I, \sigma \text{ is a bijection}
+  \label{eq:stf}
+\end{align}
+
+Let $X$ be a partially ordered set with $\preccurlyeq$ being the
+partial order on $X$.  $mub(X)$ represents an element in $X$ which is
+a maximal upper bound on $X$.  $X_m \in X$ is a maximal upper bound of
+$X$ if $\nexists X_q \in X$ such that $X_m
+\preccurlyeq X_q$. \\
+
+\noindent
+The set $I$ represents the index set $[m]$. If index $i$ is used
+without further qualification, it is meant to be $i \in I$. Any
+function, if not defined on a domain of sets, when applied on a set is
+understood as the function applied to each of its elements. i.e. for
+any function $f$ defined with domain $U$, the abuse of notation is as
+follows; $f(S)$ is used instead of $\hat f(S)$ where
+$\hat f(S) = \{y \mid y = f(x), x \in S\}$. \\
+
+\noindent
+When refering to a tree as $T$ it could be a reference to the tree
+itself, or the vertices of the tree. This will be clear from the
+context.\\
+
+\noindent
+Finally, an in-tree is a directed rooted tree in which all edges are
+directed toward to the root.
 
@@
 
-\begin{lemma}
-  \label{lem:invar1} In Algorithm~\ref{perms}, at the end of $j$th
-  iteration, $j \ge 0$, of the {\bf while} loop, the following
-  invariants are maintained.
-  \begin{enumerate}[I {\ }] %\vspace{\topshrink}
-  \item $\cl_j(R)$ is a path in $T$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ for all $R \in \cF_j$%\vspace{\topshrink}
-  \item $|R| = |\cl_j(R)|$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $R \in
-    \cF_j$%\vspace{\topshrink}
-  \item $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ for all $R, R' \in \cF_j$%\vspace{\topshrink}
-  \item $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
-    \cl_j(R'')|$, \ \ \ for all $R, R', R'' \in \cF_j$
-  \end{enumerate}
-\end{lemma}
-
-\begin{proof}
- Proof is by induction on the number of iterations, $j$. In this
-  proof, the term ``new sets'' will refer to the sets added to $\cF_j$
-  in $j$th iteration in line~\ref{setbreak} of Algorithm~\ref{perms},
-  $S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1$ and its
-  images in $\cl_j$ where $\cl_{j-1}(S_1)$
-  and $\cl_{j-1}(S_2)$ intersect and share a leaf.\\
-  The invariants are true in the base case $(\cF_0, \cl_0)$, since it
-  is the input ICPPL.  Assume the lemma is true till the $j-1$th
-  iteration. Let us consider the possible cases for each of the above invariants for
-  the $j$th iteration.
-
- \begin{enumerate}[\xbullet]
-  \item {\em Invariant} I/II
-    \begin{enumerate}[{I/II}a $|$] % \textbullet 
-    \item {\em $R$ is not a new set.} It is in $\cF_{j-1}$. Thus
-      trivially true by induction hypothesis.
-    \item {\em $R$ is a new set.} If $R$ is in $\cF_{j}$ and not in
-      $\cF_{j-1}$, then it must be one of the new sets added in
-      $\cF_j$. In this case, it is clear that for each new set, the
-      image under $\cl_j$ is a path since by definition the chosen
-      sets $S_1$, $S_2$ are from $\cF_{j-1}$ and due to the while loop
-      condition, $\cl_{j-1}(S_1)$, $\cl_{j-1}(S_2)$ have a
-      common leaf. Thus invariant I is proven.\\
-      Moreover, due to induction hypothesis of invariant III and the
-      definition of $l_j$ in terms of $l_{j-1}$, invariant II is
-      indeed true in the $j$th iteration for any of the new sets.  If
-      $R = S_1 \cap S_2$, $|R| = |S_1 \cap S_2| = |\cl_{j-1}(S_1) \cap
-      \cl_{j-1}(S_2)| = |\cl_j(S_1 \cap S_2)| = |\cl_j(R)|$.
-      If $R = S_1 \setminus S_2$, $|R| = |S_1 \setminus S_2| = |S_1| -
-      |S_1 \cap S_2| = |\cl_{j-1}(S_1)| - |\cl_{j-1}(S_1) \cap
-      \cl_{j-1}(S_2)| = |\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)| =
-      |\cl_j(S_1 \setminus S_2)|
-      = |\cl_j(R)|$. Similarly if $R = S_2 \setminus S_1$.\\
-    \end{enumerate}
-  \item {\em Invariant} III
-    \begin{enumerate}[{III}a $|$]
-    \item {\em $R$ and $R'$ are not new sets.} It is in
-      $\cF_{j-1}$. Thus trivially true by induction hypothesis.
-    \item {\em Only one, say $R$, is a new set.} Due to invariant IV
-      induction hypothesis, Lemma~\ref{lem:setminuscard} and
-      definition of $\cl_j$, it follows that invariant III is true no
-      matter which of the new sets $R$ is equal to. If $R = S_1 \cap
-      S_2$, $|R \cap R'| = |S_1 \cap S_2 \cap R'| = |\cl_{j-1}(S_1)
-      \cap \cl_{j-1}(S_2) \cap \cl_{j-1}(R')| = |\cl_j(S_1 \cap S_2)
-      \cap \cl_j(R')| = |\cl_j(R) \cap \cl_j(R')|$.  If $R = S_1
-      \setminus S_2$, $|R \cap R'| = |(S_1 \setminus S_2) \cap R'| =
-      |(\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)) \cap \cl_{j-1}(R')| =
-      |\cl_{j}(S_1 \cap S_2) \cap \cl_{j}(R')| = |\cl_{j}(R) \cap
-      \cl_{j}(R')|$. Similarly, if $R = S_2 \setminus
-      S_1$. Note $R'$ is not a new set.\\
-
-    \item {\em $R$ and $R'$ are new sets.} By definition, the new
-      sets and their path images in path label $\cl_j$ are disjoint so
-      $|R \cap R'| = |\cl_j(R) \cap \cl_j(R)| = 0$. Thus case proven.
-    \end{enumerate}
-  \item {\em Invariant} IV
-
-    Due to the condition in line~\ref{ln:3waycheck}, this invariant is
-    ensured at the end of every iteration.
-%     \begin{enumerate} [{Case 3.}1:]
-%     \item {\em $R$, $R'$ and $R''$ are not new sets.} Trivially
-%       true by induction hypothesis.
-%     \item {\em Only one, say $R$, is a new set.}
-%       If $R = S_1 \cap S_2$,  from Lemma~\ref{lem:fourpaths} and
-%       invariant III hypothesis,  this case is proven. Similarly if $R$
-%       is any of the other new  sets, the case is proven by also using
-%       Lemma ~\ref{lem:setminuscard}.
-%     \item {\em At least two of $R, R', R''$ are new sets.}
-%       The new sets are disjoint hence this case is vacuously true.
-%     \end{enumerate}
-  \end{enumerate} \qed
-%\vspace{-6mm}
-
-\end{proof}
-
-\begin{lemma}
-  \label{lem:noexit1}
-  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{perms} is
-  feasible, then the set of hypergraph isomorphism functions that
-  defines $(\cF, \cl)$'s feasibility is the same as the set that
-  defines $(\cF_j, \cl_j)$'s feasibility, if any.  Secondly, for any
-  iteration $j > 0$ of the {\em \bf while} loop, the {\em \bf exit}
-  statement in line~\ref{ln:exit1} will not execute.
-\end{lemma}
-\begin{proof}
-  Since $(\cF,\cl)$ is feasible, by Lemma~\ref{lem:feasible}
-  $(\cF_j,\cl_j)$ for every iteration $j > 0$ is feasible.  % Therefore,
-%   every hypergraph isomorphism $\phi: supp(\cF) \rightarrow V(T)$ that
-%   induces $\cl$ on $\cF$ also induces $\cl_{j-1}$ and $\cl_{j}$ on
-%   $\cF_{j-1}$ and $\cF_{j}$ respectively, i.e., $\cl_{\phi[\cF_{j-1}]}
-%   = \cl_{j-1}$ and $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
-%   that for all $x \in supp(\cF)$, for all $v \in V(T)$ the following
-%   hold true.
-  Also, every hypergraph isomorphism $\phi: supp(\cF) \rightarrow
-  V(T)$ that induces $\cl$ on $\cF$ also induces $\cl_{j}$ on
-  $\cF_{j}$, i.e., $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
-  that for all $x \in supp(\cF)$, for all $v \in V(T)$, if $(x,v) \in
-  \phi$ then $v \in \cl_{j}(S)$ for all $S \in \cF_{j}$ such that $x
-  \in S$.
-% the following
-%   hold true.
-%   \begin{enumerate}[i. ]
-%   \item If $(x,v) \in \phi$ then $v \in \cl_{j-1}(S)$ for all $S \in
-%     \cF_{j-1}$ such that $x \in S$.
-%   \item If $(x,v) \in \phi$ then $v \in \cl_{j}(S)$ for all $S \in
-%     \cF_{j}$ such that $x \in S$
-%   \end{enumerate}
-  In other words, filter 1 outputs a filtered path labeling that
-  ``preserves''
-  hypergraph isomorphisms of the original path labeling.\\
-  Secondly, line~\ref{ln:exit1} will execute iff the exit condition in
-  line~\ref{ln:3waycheck}, i.e. failure of three way intersection
-  preservation, becomes true in any iteration of the {\em \bf while}
-  loop.  Due to Lemma~\ref{lem:invar1} Invariant IV, the exit
-  condition does not occur if the input is a feasible ICPPL.\qed
-
-%   such that $\phi(x) = v$ where $v$ is the leaf considered in the
-%   first iterations of while. Clearly, $\phi$ is a renaming of
-%   vertices in hypergraph $\cF$ to those in hypergraph $\cF^\cl$. Thus
-%   the following facts can be observed in every iteration of the loop.
-
-%   \begin{enumerate}[\hspace{2mm}i. ] \vspace{\topshrink}
-%   \item all intersection cardinalities are preserved in this path
-%     labeling \vspace{\topshrink}
-%   \item element $x$ is exclusive in a hyperedge in $\cF$ since $v$ is
-%     exclusive in a hyperedge in $\cF^\cl$.
-%   \end{enumerate}
-
-%   Thus the exit condition is never rendered true after $x$ and $v$ are
-%   removed from their respective hyperedges. \qed
-
-% \noindent
-% This proof uses mathematical induction on the number
-%   of iterations $j$, $j \ge 0$, of the loop that executed
-%   without exiting. The base case, $j = 0$ is obviously true since the
-%   input is an ICPPL and the exit condition cannot hold true due to
-%   ICPPL property (iii).  Assume the algorithm executes till the end
-%   of $j-1$th iteration without exiting at line
-%  ~\ref{ln:3waycheck}. Consider the $j$th iteration. From Lemma
-%  ~\ref{lem:feasible} we know that $(\cF_j, \cl_j)$ and $(\cF_{j-1},
-%   \cl_{j-1})$ are feasible\remove[AS]{and from the proof in lemma
-%     lem:invar1 we know that $(\cF_{j-1}, \cl_{j-1})$ satisfies all the
-%     invariants defined in the lemma}.  Thus there exists a bijection
-%   $\phi: supp(\cF) \rightarrow V(T)$ such that the induced path
-%   % labeling on $\cF_{j-1}$ $\cl_{\phi[\cF_{j-1}]} = \cl_{j-1}$.
-%   labeling on $\cF_{j}$, $\cl_{\phi[\cF_{j}]}$ and on $\cF_{j-1}$,
-%   $\cl_{\phi[\cF_{j-1}]}$ are equal to $\cl_{j}$ and $\cl_{j-1}$
-%   respectively.  We need to prove that for any $R, R', R'' \in
-%   \cF_{j}$, $|R \cap R' \cap R''| = |\cl_j(R) \cap \cl_j(R') \cap
-%   \cl_j(R'')|$.
-%   The following are the possible cases that could arise. From argument
-%   above, $|\cl_j(R) \cap \cl_j(R') \cap \cl_j(R'')| =
-%   |\cl_{\phi[\cF_{j}]}(R) \cap \cl_{\phi[\cF_{j}]} (R') \cap
-%   \cl_{\phi[\cF_{j}]} (R'')|$
-
-%   \begin{enumerate}[a $|$]
-%   \item {\em None of the sets are new. $R, R', R'' \in \cF_{j-1}$.}
-%     We know $(\cF_{j-1}, \cl_{j-1})$ is feasible. Thus $|R \cap R'
-%     \cap R''| = |\cl_{j-1}(R) \cap \cl_{j-1}(R') \cap \cl_{j-1}(R'')|
-%     = |\cl_{j}(R) \cap \cl_{j}(R') \cap \cl_{j}(R'')|$.
-%   \item {\em Only one, say $R$, is a new set.}  Let $R = S_1 \cap S_2$
-%     ($S_1, S_2$ are defined in the proof of lemma
-%    ~\ref{lem:invar1}). Now we have $|R \cap R' \cap R''| = |S_1 \cap
-%     S_2 \cap R' \cap R''| = |\cl_{j-1}(S_1) \cap \cl_{j-1}(S_2) \cap
-%     \cl_{j-1}(R') \cap \cl_{j-1}(R'')| = |\cl_{j}(R) \cap \cl_{j}(R')
-%     \cap \cl_{j}(R'')|$. Thus proven. If $R$ is any of the other new
-%     sets, the same claim can be verified using lemma
-%    ~\ref{lem:setminuscard}.
-%     % \item []{\bf Case 3:}
-%   \item {\em At least two of $R, R', R''$ are new sets.}  The new sets
-%     are disjoint hence this case is vacuously true.
-%   \end{enumerate}
-%   \qed
-%   \tnote[E2]{remove the induction proof. just text saying x and v are
-%     exclusive in these sets therefore the intersection cardinalities
-%     don't change thus all invariants are still true} 
-\end{proof}
-
-As a result of Algorithm~\ref{perms} each leaf $v$ in $T$
-is such that there is exactly one set in $\cF$ with $v$ as a vertex in
-the path assigned to it.  In Algorithm~\ref{leafasgn} we identify
-elements in $supp(\cF)$ whose images are leaves in a hypergraph
-isomorphism if one exists.  Let $S \in \cF$ be such that $\cl(S)$ is a
-path with leaf and $v \in V(T)$ is the unique leaf incident on it.  We
-define a new path labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) =
-\set{v}$ where $x$ an arbitrary element from $S \setminus \bigcup_{\hS
-  \ne S} \hS$. In other words, $x$ is an element present in no other
-set in $\cF$ except $S$. This is intuitive since $v$ is present in no
-other path image under $\cl$ other than $\cl(S)$.  The element $x$ and
-leaf $v$ are then removed from the set $S$ and path $\cl(S)$
-respectively. After doing this for all leaves in $T$, all path images
-in the new path labeling $\cl_{new}$ except leaf labels (a path that
-has only a leaf is called the {\em leaf label} for the corresponding
-single element hyperedge or set) are paths from a new pruned tree $T_0
-= T \setminus \{v \mid v \text{ is a leaf in }
-T\}$. Algorithm~\ref{leafasgn} is now presented with details.
-
-\begin{algorithm}[h]
-  \caption{Leaf labeling from an ICPPL {\tt filter\_2($\cF, \cl, T$)}}
-  \label{leafasgn}
-  \begin{algorithmic}[\lndisplay]
-    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$
-    \COMMENT {Path images are such that no two path images share a
-      leaf.}
-    \STATE $j \assign 1$\\
-    \WHILE {there is a leaf $v$ in $T$ and a unique $S_1 \in
-      \cF_{j-1}$ such that $v \in \cl_{j-1}(S_1)$ }\label{uniqueleaf}
-    \STATE $\cF_j \assign \cF_{j-1} \setminus \{S_1\}$\\
-    \STATE for all $S \in \cF_{j-1}$ such that $S \ne S_1$ set
-    $\cl_j(S) \assign
-    \cl_{j-1}(S)$\\
-    \STATE $X \assign S_1 \setminus \bigcup_{S \in \cF_{j-1}, S \ne S_1}S$\\
-    \IF{$X$ is empty} \label{xempty} \STATE {\bf exit} \label{ln:exit2} \ENDIF
-    \STATE $x \assign $ arbitrary element from $X$\\
-    \STATE $\cF_j \assign \cF_j \cup \{\{x\}, S_1 \setminus \{x\}\} $\\
-    \STATE $\cl_j(\{x\}) \assign \{v\}$\\
-    \STATE $\cl_j(S_1 \setminus \{x\}) \assign \cl_{j-1}(S_1) \setminus \{v\}$\\
-    \STATE $j \assign j+1$\\
-    \ENDWHILE
-    \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
-    \RETURN $(\cF', \cl')$
-  \end{algorithmic}
-\end{algorithm}
-
-Suppose the input ICPPL $(\cF, \cl)$ is feasible, yet set $X$ in
-Algorithm~\ref{leafasgn} is empty in some iteration of the {\bf while}
-loop. This will abort our procedure of finding the hypergraph
-isomorphism. The following lemma shows that this will not happen.
-
-\begin{lemma}
-  \label{lem:xnotempty}
-  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{leafasgn} is
-  feasible, then for all iterations $j > 0$ of the {\em \bf while}
-  loop, the {\em \bf exit} statement in line~\ref{ln:exit2} does not
-  execute.
-\end{lemma}
-\begin{proof}
-  Assume $X$ is empty for some iteration $j > 0$. We know that $v$ is
-  an element of $\cl_{j-1}(S_1)$. Since it is uniquely present in
-  $\cl_{j-1}(S_1)$, it is clear that $v \in \cl_{j-1}(S_1) \setminus
-  \bigcup_{(S \in \cF_{j-1}) \wedge (S \ne S_1)}\cl_{j-1}(S)$.  Note
-  that for any $x \in S_1$ it is contained in at least two sets due to
-  our assumption about cardinality of $X$. Let $S_2 \in \cF_{j-1}$ be
-  another set that contains $x$. From the above argument, we know $v
-  \notin \cl_{j-1}(S_2)$. Therefore there cannot exist a hypergraph
-  isomorphism bijection that maps elements in $S_2$ to those in
-  $\cl_{j-1}(S_2)$. This contradicts our assumption that the input is
-  feasible. Thus $X$ cannot be empty if input is ICPPL and feasible.
-  \qed
-\end{proof}
-
-\begin{lemma}
-  \label{lem:invar3}
-  In Algorithm~\ref{leafasgn}, for all $j > 0$, at the end of the
-  $j$th iteration of the {\bf while} loop the four invariants given in
-  Lemma~\ref{lem:invar1} hold.
-\end{lemma}
-\begin{proof}
-  By Lemma~\ref{lem:xnotempty} we know that set $X$ will not be empty
-  in any iteration of the {\em \bf while} loop if input ICPPL $(\cF,
-  \cl)$ is feasible and $\cl_j$ is always computed for all $j >
-  0$. Also note that removing a leaf from any path keeps the new path
-  connected. Thus invariant I is obviously true. In every iteration $j
-  > 0$, we remove exactly one element $x$ from one set $S$ in $\cF$
-  and exactly one vertex $v$ which is a leaf from one path
-  $\cl_{j-1}(S)$ in $T$. This is because as seen in
-  Lemma~\ref{lem:xnotempty}, $x$ is exclusive to $S$ and $v$ is
-  exclusive to $\cl_{j-1}(S)$. Due to this fact, it is clear that the
-  intersection cardinality equations do not change, i.e., invariants
-  II, III, IV remain true. On the other hand, if the input ICPPL is
-  not feasible the invariants are vacuously true. \qed
-\end{proof}
-
-% \textcolor{cyan}{
-% \begin{lemma}
-%   \label{lem:notfeasibleexit}
-%   \tnote{IS THIS CORRECT?}  If input ICPPL $(\cF, \cl)$ is not
-%   feasible, then in one of the recursive calls to Algorithm --3--, the
-%   {\em \bf exit} statement in line x in Algorithm --1 or line y in
-%   Algorithm --2 will get executed.
-% \end{lemma}
-% }
-
-We have seen two filtering algorithms above, namely,
-Algorithm~\ref{perms} {\tt filter\_1} and Algorithm~\ref{leafasgn}
-{\tt filter\_2} which when executed serially repectively result in a
-new ICPPL on the same universe $U$ and tree $T$. We also proved that
-if the input is indeed feasible, these algorithms do indeed output the
-filtered ICPPL. Now we present the algorithmic characterization of a
-feasible tree path labeling by way of Algorithm~\ref{al:icppl-find-isomorph}.
-
-Algorithm~\ref{al:icppl-find-isomorph} computes a
-hypergraph isomorphism $\phi$ recursively using Algorithm~\ref{perms}
-and Algorithm~\ref{leafasgn} and pruning the leaves of the input
-tree. In brief, it is done as follows. Algorithm~\ref{leafasgn} gives
-us the leaf labels in $\cF_2$, i.e., the elements in $supp(\cF)$ that
-map to leaves in $T$, where $(\cF_2, \cl_2)$ is the output of
-Algorithm~\ref{leafasgn}. All leaves in $T$ are then pruned away. The
-leaf labels are removed from the path labeling $\cl_2$ and the
-corresponding elements are removed from the corresponding sets in
-$\cF_2$. This tree pruning algorithm is recursively called on the
-altered hypergraph $\cF'$, path label $\cl'$ and tree $T'$. The
-recursive call returns the bijection $\phi''$ for the rest of the
-elements in $supp(\cF)$ which along with the leaf labels $\phi'$ gives
-us the hypergraph isomorphism $\phi$.  The following lemma formalizes
-the characeterization of feasible path labeling.
-
-\begin{algorithm}[h]
-  \caption{{\tt get- hypergraph- isomorphism ($\cF, \cl, T$)}}
-  \label{al:icppl-find-isomorph}
-  \begin{algorithmic}[\lndisplay]
-
-    \IF{$T$ is empty}
-    \RETURN $\emptyset$\\
-    \ENDIF
-    \STATE $L \assign \{v \mid v \text{ is a leaf in }      T\}$\\
-    \STATE $(\cF_1, \cl_1) \assign$ {\tt filter\_1($\cF, \cl,
-      T$)}\\
-    \STATE $(\cF_2, \cl_2) \assign$ {\tt filter\_2($\cF_1,
-      \cl_1, T$)}\\
-
-    \STATE $(\cF', \cl') \assign (\cF_2, \cl_2)$\\
-    \STATE $\phi' \leftarrow \emptyset$
-
-    \FOR {every $v \in L$}
-    \STATE $\phi'(x) \assign v$ where $x \in \cl_2^{-1}(\{v\})$
-    \COMMENT {Copy the leaf labels to a one to one function $\phi':
-      supp(\cF) \rightarrow L$
-      }\\
-    \STATE Remove $\{x\}$ and $\{v\}$ from $\cF'$, $\cl'$  appropriately\\
-    \ENDFOR
-
-    \STATE $T' \assign T \setminus L$
-
-    \STATE $\phi'' \assign$ {\tt
-      get-hypergraph-isomorphism($\cF', \cl', T'$)}
-    \STATE $\phi \assign \phi'' \cup \phi'$ \\
-    \RETURN $\phi$
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma}
-  \label{lem:hyperiso}  %{lem:perm}
-  If $(\cF, \cl)$ is an ICPPL from a tree $T$ and
-  Algorithm~\ref{al:icppl-find-isomorph}, {\tt
-    get- hypergraph- isomorphism ($\cF, \cl, T$)} returns a non-empty
-  function, then there exists a hypergraph isomorphism $\phi :
-  supp(\cF) \rightarrow V(T)$ such that the $\phi$-induced tree path
-  labeling is equal to $\cl$ or $\cl_\phi = \cl$.
-\end{lemma}
-\begin{proof}
-  It is clear that in the end of every recursive call to
-  Algorithm~\ref{al:icppl-find-isomorph}, the function $\phi'$ is one
-  to one involving all the leaves in the tree passed as input to that
-  recursive call. Moreover, by Lemma~\ref{lem:noexit1} and
-  Lemma~\ref{lem:xnotempty} it is consistent with the tree path
-  labeling $\cl$ passed. The tree pruning is done by only removing
-  leaves in each call to the function and is done till the tree
-  becomes empty. Thus the returned function $\phi: supp(\cF)
-  \rightarrow V(T)$ is a union of mutually exclusive one to one
-  functions exhausting all vertices of the tree. In other words, it is
-  a bijection from $supp(\cF)$ to $V(T)$ inducing the given path
-  labeling $\cl$ and thus a hypergraph isomorphism. \qed
-\end{proof}
-
-\begin{theorem}
-  \label{th:charac}
-  A path labeling $(\cF, \cl)$ on tree $T$ is feasible iff it is an
-  ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
-  T)$ as input returns a non-empty function.
-\end{theorem}
-\begin{proof}
-  From Lemma~\ref{lem:hyperiso}, we know that if $(\cF, \cl)$ is an
-  ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
-  T)$ as input returns a non-empty function, $(\cF, \cl)$ is feasible.
-  Now consider the case where $(\cF, \cl)$ is feasible. i.e. there
-  exists a hypergraph isomorphism $\phi$ such that $\cl_\phi =
-  \cl$. Lemma~\ref{lem:noexit1} and Lemma~\ref{lem:xnotempty} show us
-  that filter 1 and filter 2 do not exit if input is feasible. Thus
-  Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
-  function.\qed
-\end{proof}
-
-\section{ Computing feasible TPL with special target trees \tnote{give
-    problem definition etc}}
-\label{sec:spltargettree}
-
-Section~\ref{sec:feasible} described properties that a TPL must have
-for it to be feasible. The next problem of interest is to test if a
-given hypergraph is a path hypergraph with respect to a given target
-tree\footnote{A larger problem would be to find if a given hypergraph
-  is a path graph on any tree. This problem is not addressed in this
-  thesis.}. In other words, the problem is to find out if a feasible
-tree path labeling exists from a given target tree for a given
-hypergraph. In this section we will see two special cases of this
-problem where the target tree is from a particular family of
-trees. The first one, where the tree is a path, is a shown to be
-equivalent to the well studied problem of consecutive-ones in
-Section~\ref{sec:icpplicpia}. The second one, where the tree is a
-$k$-subdivided tree\footnote{See Section~\ref{ch:prelims} for the
-  formal definition.}, has been solved using a polynomial time
-algorithm. The latter problem enforces some conditions on the
-hypergraph too which will be seen in Section~\ref{sec:ksubdivstar}.
-
-\subsection{Target tree is a Path}
-\label{sec:icpplicpia}
-
-Consider a special case of ICPPL with the following properties.
-\begin{enumerate}
-\item Given tree $T$ is a path. Hence, all path labels are interval
-  labels.
-\item Only pairwise intersection cardinality preservation is
-  sufficient. i.e. property (iii) in ICPPL is not enforced.
-\item The filter algorithms do not have {\em \bf exit} statements.
-\end{enumerate}
-This is called an Intersection Cardinality Preservation Interval
-Assignment (ICPIA) \cite{nsnrs09}. This structure and its algorithm is
-used in the next section for finding tree path labeling from a
-$k$-subdivided star due to this graph's close relationship with
-intervals.
-
-\subsection{Target tree is a $k$-subdivided Star}
-\label{sec:ksubdivstar}
-
-\begin{figure}[t] %[htbp] 
-  \centering
-  \begin{tabular}{lr}
-    (a) \includegraphics{cop-tree_thesis.1}&
-    (b) \includegraphics{cop-tree_thesis.2}
-  \end{tabular}
-  \caption{\figtabsize (a) $8$-subdivided star with 7 rays (b)
-    3-subdivided star with 3 rays}
-  \label{fig:kstar}
-\end{figure}
-
-In this section we consider the problem of assigning paths from a
-$k$-subdivided star $T$ to a given set system $\cF$.  We consider
-$\cF$ for which the overlap graph $\bO(\cF)$ is connected.  The
-overlap graph is well-known from the work of
-\cite{kklv10,nsnrs09,wlh02}.  We use the notation in
-\cite{kklv10}. Recall from Section~\ref{ch:prelims} that hyperedges
-$S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$
-and $S'$ have a non-empty intersection but neither of them is
-contained in the other. The overlap graph $\bO(\cF)$ is a graph in
-which the vertices correspond to the sets in $\cF$, and the vertices
-corresponding to the hyperedges $S$ and $S'$ are adjacent if and only
-if they overlap.  Note that the intersection graph of $\cF$,
-$\bI(\cF)$ is different from $\bO(\cF)$ and $\bO(\cF) \subseteq
-\bI(\cF)$.  A connected component of $\bO(\cF)$ is called an overlap
-component of $\cF$.  An interesting property of the overlap components
-is that any two distinct overlap components, say $\cO_1$ and $\cO_2$,
-are such that any two sets $S_1 \in \cO_1$ and $S_2 \in \cO_2$ are
-disjoint, or, w.l.o.g, all the sets in $\cO_1$ are contained within
-one set in $\cO_2$.  This containment relation naturally determines a
-decomposition of the overlap components into rooted containment trees.
-We consider the case when there is only one rooted containment tree,
-and we first present our algorithm when $\bO(\cF)$ is connected.  It
-is easy to see that once the path labeling to the overlap component in
-the root of the containment tree is achieved, the path labeling to the
-other overlap components in the rooted containment tree is essentially
-finding a path labeling when the target tree is a path: each target
-path is a path that is allocated to sets in the root overlap
-component.  Therefore, for the rest of this section, $\bO(\cF)$ is a
-connected graph. We also assume that all hyperedges are of cardinality
-at most $k+2$.
-
-Recall from Section~\ref{ch:prelims} that a $k$-subdivided star is a
-star with each edge subdivided $k$ times. Therefore, a $k$-subdivided
-star has a central vertex which we call the {\em root}, and each root
-to leaf path is called a {\em ray}. First, we observe that by removing
-the root $r$ from $T$, we get a collection of $p$ vertex disjoint
-paths of length $k+1$, $p$ being the number of leaves in $T$.  We
-denote the rays by $R_1, \ldots, R_p$, and the number of vertices in
-$R_i$, $i \in [p]$ is $k+2$.  Let $\seq{v_{i1},\ldots,v_{i(k+2)}=r}$
-denote the sequence of vertices in $R_i$, where $v_{i1}$ is the
-leaf. Note that $r$ is a common vertex to all $R_i$.
-
-In this section the given hypergraph, the $k$-subdivided star and the
-root of the star are denoted by $\cO$, $T$ and vertex $r$,
-respectively.
-% For each hyperedge $X \in \cO$, we will maintain a 2-tuple of non-negative
-% numbers $\seq{p_1(X), p_2(X)}$.  The numbers satisfy the property that
-% $p_1(X) + p_2(X) \leq |X|$, and at the end of path labeling, for each
-% $X$, $p_1(X) + p_2(X) = |X|$.  This signifies the algorithm tracking
-% the lengths of subpaths of the path assigned to $X$ from at most two
-% rays. We also maintain another parameter called the {\em residue} of
-% $X$ denoted by $s(X)=|X| - p_1(X)$. This signifies the residue path
-% length that must be assigned to $X$ which must be from another
-% ray. For instance, if $X$ is labeled a path from only one ray, then
-% $p_1(X) = |X|$, $p_2(X) = 0$ and $s(X) = 0$.
-
-% 
-% We iteratively consider each ray from which paths will be assigned to
-% hyperedges. At the beginning of each iteration hyperedges of $\cO$ are
-% classifed into the following disjoint sets.
-% \begin{enumerate}
-% \item [$\cL_1^i$] {\em Labeled without $r$.} Those that have been
-%   labeled with a path which does not contain $r$ in one of the
-%   previous iterations.\\  $\cL_1^i = \set{ X \mid p_1(X) = |X| \text{ and
-%     } p_2(X) = 0 \text{ and } s(X) = 0, X \in \cO}$
-% \item [$\cL_2^i$] {\em Labeled with $r$.} Those that have been labeled
-%   with two subpaths of $\cl(X)$ containing $r$ from two different rays
-%   in two previous iterations.\\ $\cL_2^i = \set{X \mid 0 < p_1\left(X\right),
-%     p_2\left(X\right) < |X| \text{ and } s(X) = 0, X \in \cO}$
-%   \item [$\cT_1^i$] {\em Type 1 / partially labeled.} Those that have
-%     been labeled with one path containing $r$ from a single ray in one
-%     of the previous iterations. Here, $p_1(X)$ denotes the length of
-%     the subpath of $\cl(X)$ that $X$ has been so far labeled
-%     with. Also, it is clear that such a path must start at $r$ and
-%     must be in a ray different from the one corresponding to $p_1(X)$.\\
-%     $\cT_1^i = \set{ X \mid 0 < p_1(X) < |X| \text{ and } p_2(X) = 0
-%       \text{ and } s(X) > 0, X \in \cO}$
-%   \item [$\cT_2^i$] {\em Type 2 / not labeled.} Those that have not been
-%     labeled with a path in any previous iteration.\\
-%     $\cT_2^i = \set{ X \mid p_1(X) = p_2(X) = 0 \text{ and } s(X) = |X|,
-%       X \in \cO}$
-% \end{enumerate}
-% \vspace{-2mm}
-% \begin{align*}
-%   \cO &= \cL_1^i \cup \cL_2^i \cup \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]\\
-%   \cO_i &= \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]
-% \end{align*}
-
-The set $\cO_i$ refers to the set of hyperedges $\cT_1^i \cup \cT_2^i$
-in the $i$th iteration.  Note that $\cO_1 = \cO$.  In the $i$th
-iteration, hyperedges from $\cO_i$ are assigned paths from $R_i$ using
-the following rules. Also the end of the iteration, $\cL_1^{i+1},
-\cL_2^{i+1}, \cT_1^{i+1}, \cT_2^{i+1}$ are set to $\cL_1^{i},
-\cL_2^{i}, \cT_1^{i}, \cT_2^{i}$ respectively, along with some
-case-specific changes mentioned in the rules below.
-
-\noindent
-Suppose the given hypergraph has a feasible path labeling to the given
-tree $T$. Let this labeling be $\cl$. Now we present an algorithm that
-lists the sequence of edges that maintain a set of properties
-necessary for them to have a feasible tree path labeling.
-\begin{enumerate}[I.]
-\item {\bf Step 1}
-  \begin{enumerate}
-  \item {\bf {There are type 1
-        edges, $\mathbf{|\cT_1^i| > 0}$}:} 
-% , and for each such $X$ assign the
-%     path in $R_i$ of length $s(X)$ starting at $v_{i,k+2}=r$.  $X$ is
-%     then removed from $\cO_i$, and $p_2(X)=s(X)$, and $s(X)=0$.If any
-%     intersection cardinality is violated because of this assignment,
-%     EXIT reporting the non-existence of a feasible assignment.
-  {If there is only one $\cT_1^i$ hyperedge, label it to the unique
-    path in $R_i$ of length $s(X)$ starting at $v_{i(k+2)}=r$. If
-    there are more than one $\cT_1^i$ hyperedges, for any pair of
-    hyperedges $S_1, S_2 \in \cT_1^i$, one of the following is
-    true. Let $\cl'(X) \subseteq R_j$ denote the subpath of length
-    $p_1(X)$ that has been assigned to $X$ from a previously
-    considered ray $R_j$.}
-  \begin{enumerate}[{\em Case }1. ]
-  \item $|\cl'(S_1) \cap \cl'(S_2)| = 1$ and $|S_1 \cap S_2| > 1$
-    i.e. they are in different rays and their residues must contain
-    their intersection. Hence assign both to $R_i$.
-  \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| >
-    |\cl'(S_1) \cap \cl'(S_2)|$ i.e. they are in the same rays and
-    their residues must contain their intersection. Hence assign both
-    to $R_i$.
-  \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| =
-    |\cl'(S_1) \cap \cl'(S_2)|$. i.e. the intersection cardinality has
-    been met. Hence assign only $S_1$ to $R_i$ and $S_2$ must be
-    assigned to some $i \ge p$.
-  \end{enumerate}
-  Remove the labeled edge or edges from $\cO_i$ and add it to
-  $\cO_{i+1}$. Also do not add it to $\cT_1^{i+1}$ and add to
-  $\cL_2^{i+1}$.
-\item {\bf {There are no type-1 edges nor} super-marginal type-2
-    hyperedges:} If all type-2 edges of length at most $k+1$, EXIT by
-  saying there is no ICPPL.  Otherwise, Find an $X$ of type-2 such
-  that $|X| \geq k+2$, assign it the unique path starting at
-  $v_{i(k+1)}=r$ of length $k+2$, and set $p_1(X)=k+2,
-  s(X)=|X|-(k+2)$, assign it the unique path starting at
-  $v_{i(k+1)}=r$ of length $k+2$, mark this as a type-1 hyperedge, and
-  add it to $\cO_{i+1}$ after removing it from $\cO_i$.
-  % \remove{Next, iteratively, consider each type-1 edge $X \in
-  %   \cO_i$, and if it intersects a previously assigned $Y$ such that
-  %   $\cl(Y) \subseteq R_i$, assign a path of length $s(X)$
-  %   containing $v_{i(k+2)}=r$, set $p_2(X)=s(X), s(X)=0$, and remove
-  %   $X$ from $\cO_i$.}\tnote{covered in step 2}
-
-\item {\bf There are super-marginal type-2 hyperedges {and no type-1
-      edges:}} Let $X$ be a super-marginal hyperedge. If $|X| \leq
-  k+2$, then $X$ is assigned the path of length $|X|$ starting at
-  $v_{i1}$, and $p_1(X) = |X|, p_2(X)=0$. $X$ is removed from $\cO_i$
-  and not added to $\cO_{i+1}$.  If $|X| > k+2$, then $p_1(X)=k+2$,
-  assign it the unique path starting at $v_{i(k+1)}=r$ of length $k+2$
-  and $s(X) = |X| - (k+2)$.  In this case, $X$ is classified as a
-  Type-1 edge, removed from $\cO_i$, and added to $\cO_{i+1}$.
-\end{enumerate}
-\item {\bf Step 2:} Iteratively, a hyperedge $X$ is selected from
-  $\cO_i$ that has an overlap with one of the hyperedges $Y$ such that
-  $\cl(Y) \subseteq R_i$, and a unique path is assigned to $X$.  The
-  path, say $U(X) \subseteq R_i$, that is assigned can be decided
-  unambiguously since the $X \overlap Y$, and all intersection
-  cardinalities can be preserved only at one of the ends of $\cl(Y)$.
-
-  \noindent
-  Let $\cl(X)$ denote the unique path assigned to $X$.  If $X$ is a
-  type-2 hyperedge and if the unique path of length $|X|$ does not
-  contain $r$, then $p_1(X) = |X|, p_2(X)=0$, and $X$ is removed from
-  $\cO_i$.  In the case when $\cl(X)$ has to contain $r$, then
-  $p_1(X)$ is the length of the path, $p_2(X)=0$, and $s(X) =
-  |X|-p_1(X)$. Further $X$ is classified as a type-1 hyperedge, added
-  to $\cO_{i+1}$ and removed from $\cO_i$.  In the case when $X$ is a
-  type-1 hyperedge, then we check if $U(X)$, which is of length $s(X)$
-  contains $r$. If it does, then we assign $\cl(X) \leftarrow \cl(X)
-  \cup U(X)$, remove $X$ from $\cO_i$ and do not add it to
-  $\cO_{i+1}$.  If not, then we {\bf Exit} reporting that an
-  assignment cannot be found.  The iteration ends when no hyperedge in
-  $\cO_i$ has an overlap with a hyperedge assigned to $R_i$.
-\end{enumerate}
-
-\noindent
-The order in which they are assigned is the sequence of hyperedges
-with the following properties. Note that if such a sequence cannot be
-computed, $\cO$ cannot have a feasible path labeling from $T$.
-
-\noindent
-\tnote{need to formalize the below properties} In the following
-lemmas we identify a set of necessary conditions for $\cF$ to have an
-ICPPL in the $k$-subdivided star $T$.  If during the execution of the
-above algorithm, one of these necessary conditions is not satisfied,
-the algorithm exits reporting the non-existence of an ICPPL.
-\begin{lemma}
-  Let all hyperedges in $\cO_i$ be type-1 edges.  Then there is a
-  maximal subset $\ccT_i \subseteq \cO_i$ with the following
-  properties:
-  \begin{enumerate}
-  \item $\ccT_i$ form an inclusion chain.
-  \item For all $X \in \ccT_i$, $s(X) \leq k+2$, and There is a an $X
-    \in \ccT_i$ such that $s(X)=k+2$.
-  \end{enumerate}
-\end{lemma}
-\begin{lemma}
-  If there are no \marginal type-2 edges in $\cO_i$, then there exists
-  at least $p-i$ type-2 hyperedges $X \in \cO_i$ such that $|X| \geq
-  k+2$.
-  % is a exist type-2 hyperedges of length at least $k+2$, and there
-  % must be at least $r-i$ such type-2 hyperedge, basically at least
-  % one per ray.  Each hyperedge contained in such an hyperedges of
-  % length at least $k+2$, is transitively related under the overlap
-  % relation to one that gets assigned a path that contains
-  % $r$. Basically, the proof will argue that otherwise, there will be
-  % a \marginal type-2 hyperedge, or that it is infeasible.  Also, it
-  % is here we assume that we assume that the size of the two
-  % universes is the same.
-\end{lemma}
-\begin{lemma}
-  At the end of Step-1 in the $i$-th iteration, if one hyperedge $X$
-  of type-2 is such that $\cl(X) \subseteq R_i$, then all other
-  hyperedges in $\cO_i$ are connected to $X$ in the overlap component.
-\end{lemma}
-\begin{lemma}
-  At the end of Step-2, If control has exit at any time, there is no
-  ICPPL. If control has not exit, then $R_i$ is saturated.  No
-  hyperedge of $\cO_{i+1}$ will get a path from $R_i$ in the future
-  iterations.  No type-2 hyperedge of $\cO_{i+1}$ will get a path from
-  $R_i$.  Basically $R_i$ is done.
-\end{lemma}
-\begin{lemma}
-  Finally, we need to prove that the assignment is an ICPPL. Secondly,
-  if there is a permutation then maps sets to paths, then it is indeed
-  an ICPPL, and our algorithm will basically find it.  It is a unique
-  assignment upto permutation of the leaves.
-\end{lemma}
-
-\begin{lemma}
-  \label{lem:sup-mar}
-  If $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
-  tree path labeling to tree $T$, then $\cl(X)$ will
-  contain a leaf in $T$.\\
-  On the otherhand, if $\cl(X)$ has a leaf in
-  $T$, $X$ is marginal but may not be super-marginal.
-\end{lemma}
-\begin{proof}
-  Suppose $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
-  path labeling from $T$.  Assume $\cl(X)$ does not have a leaf.  Let
-  $R_i$ be one of the rays (or the only ray) $\cl(X)$ is part of.
-  Since $X$ is in a connected overlap component, there exists $Y_1 \in
-  \cF$ and $X \nsubseteq Y_1$ such that $Y_1 \overlap X$ and $Y_1$ has
-  at least one vertex closer to the leaf in $R_i$ than any vertex in
-  $X$. Similarly with the same argument there exists $Y_2 \in \cF$
-  with same subset and overlap relation with $X$ except it has has at
-  least one vertex farther away from the leaf in $R_i$ than any vertex
-  in $X$. Clearly $Y_1 \cap X$ and $Y_2 \cap X$ cannot be part of same
-  inclusion chain which contradicts that assumption $X$ is
-  super-marginal. Thus the claim is proven.\qed
-\end{proof}
-
-\noindent
-Consider the overlap graph $\bO(\cF)$ of the given hypergraph
-$\cF$. Let $S_{sm} \in \cF$ be such that it is a super-marginal
-hyperedge.  Algorithm~\ref{al:ktree-label} uses $S_{sm}$ along with
-the overlap graph $\bO(\cF)$ to calculate the feasible tree path
-labeling to the $k$-subdivded tree $T$.
-
-\begin{algorithm}[h]
-  \caption{{\tt compute-ksubtree-path-labeling($X, \cF, T$)}}
-  \label{al:ktree-label}
-  \begin{algorithmic}[\lndisplay]
-    \IF{$X = S_{sm}$}
-    \STATE -- TBD --\\
-    \ELSE 
-    \STATE -- TBD --\\
-    \ENDIF
-
- \end{algorithmic}
-\end{algorithm}
-
-\section{ TPL with no restrictions}
-\label{sec:norestraint}
-
-{\tt abstract begin}
-  it is
-  known that if the given tree is a path a feasible assignment can be
-  found in polynomial time, and we observe that it can actually be
-  done in logspace. \tnote{[TRUE?]}
-{\tt abstract end}
-
-In Section \ref{prelims} we present the
-necessary preliminaries, in Section \ref{feasible} we present our
-characterization of feasible tree path assignments, and in Section
-\ref{decompos} we present the characterizing subproblems for finding a
-bijection between $U$ and $V(T)$ such that sets map to tree
-paths. Finally, in Section \ref{complexity} we conclude by showing
-that Tree Path Assignment is GI-Complete, and also observe that
-Consecutive Ones Testing is in Logspace.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC \subsection{Introduction}
-
-\annote{ We refer to this as the Tree Path Assignment
-problem for an input $(\cF,T)$ pair.}{Add this ``pair'' as a
-  definition in ch:prelims}
-
-\remove{It is an interesting fact that for a matrix with the COP, the
-intersection graph of the corresponding set system is an interval
-graph.  A similar connection to a subclass of chordal graphs, and this
-subclass contains interval graphs, exists for the generalization of
-COP.  In this case, the intersection graph of the corresponding set
-system must be a path graph. Chordal graphs are of great
-significance, extensively studied, and have several applications.  One
-of the well known and interesting properties of a chordal graphs is
-its connection with intersection graphs cite{mcg04}. For every chordal
-graph, there exists a tree and a family of subtrees of this tree such
-that the intersection graph of this family is isomorphic to the
-chordal graph cite{plr70,gav78,bp93}.  Certain format of these trees
-are called clique trees cite{apy92} of the graph which is a compact
-representation of the chordal graph. There has also been work done on
-the generalization of clique trees to clique hypergraphs cite{km02}.
-If the chordal graph can be represented as the intersection graph of
-paths in a tree, then the graph is called path graph cite{mcg04} or
-path chordal graph.  Therefore, it is clear that if there is a
-bijection from $U$ to $V(T)$ such that the sets map to paths, then the
-intersection graph of the set system is indeed a path chordal graph.
-This is, however, only a necessary condition and can be checked
-efficiently, as path graph recognition is polynomial
-solvable cite{gav78,aas93}.  Indeed, it is possible to construct a set
-system and tree, such that the intersection graph is a path chordal
-graph, but there is no bijection between $U$ and $V(T)$ such that the
-sets map to paths.  This connection indeed suggests that our problem
-is indeed as hard as path graph isomorphism.  Further path graph
-isomorphism is known be isomorphism-complete, see for example
-cite{kklv10}. }
-In the second part of this paper, we decompose our
-search for a bijection between $U$ and $V(T)$ into subproblems.  Each
-subproblem is on a set system in which for each set, there is another
-set in the set system with which the intersection is {\em strict}-
-there is a non-empty intersection, but neither is contained in the
-other.  This is in the spirit of results in \cite{wlh02,nsnrs09} where
-to test for COP in a given matrix, the COP problem is solved on an
-equivalent set of prime matrices.  \annote{Our decomposition localizes the
-challenge of path graph isomorphism to two problems.}{HOW?}
-
-Finally, we show that Tree Path Assignment is isomorphism-complete.
-We also point out Consecutive Ones Testing is in Logspace from two
-different results in the literature \cite{kklv10, mcc04}. To the best
-of our knowledge this observation has not been made earlier.
-
-% \noindent
-% It has been long known that interval graph recognition is in
-% logspace\cite{rei84}. Recently interval graph isomorphism was also
-% shown to be logspace decidable using a logspace canonization
-% algorithm by \cite{kklv10}.  This result is built on top of logspace
-% results of undirected graph connectivity \cite{rei08}, logspace
-% tractability using a certain logical formalism called FP+C and
-% modular decomposition of interval graphs\cite{lau10} etc.  Interval
-% graphs are closely connected to binary matrices with COP. The
-% maximal clique vertex incidence matrix (matrix with rows
-% representing maximal cliques and columns representing vertices of a
-% graph) has COP on columns iff the graph is an interval
-% graph\cite{fg65}. This follows from the interval graph
-% characterization by \cite{gh64}. Due to this close relation it is
-% natural to see if consecutive ones property can be
-% tested in logspace. \\
-% \noindent
-% We also explore some extensions of the interval assignment problem
-% in \cite{nsnrs09}, namely caterpillar path assignment problem.
-
-% We present a logspace algorithm here that uses the ICPIA
-% characterization of binary matrices with COP (set system associated
-% with such a matrix)\cite{nsnrs09}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC\subsection{Preliminaries} 
-\label{prelims} 
-In this paper, the collection
-$\F = \{S_i \mid S_i \subseteq U, S_i \ne \O, i \in I\}$ is a {\em set
-  system} of universe, $U = \{1, \ldots, n\}$. Moreoever, a set system
-is assumed to ``cover'' the universe,
-i.e. $ \bigcup_{i \in I}S_i = U$. \\
-
-\noindent
-The graph $T=(V,E)$ represents a given tree with $|V| = n$. Further,
-for simplicity, $V$ is defined as $\{1,\ldots,n\}$. All paths
-referred to in this paper are paths of $T$ unless explicitly specified. \\
-
-\noindent
-A {\em path assignment} $\A$ to $\F$ is defined as a set assignment
-where second universe is the vertex set $V$ of a given tree $T$ and
-every second subset in the ordered pairs is a path in this
-tree. Formally, the definition is as follows.
-\begin{align*}
-  \A = \{ (S_i,P_i) \mid S_i \in \F, P_i \subseteq V \text{
-    s.t. }T[P_i] \text{ is a path, } i \in I \}
-\end{align*}
-In other words, $P_i$ is the path on the tree $T$ assigned to $S_i$ in
-$\A$. As mentioned before for set systems, the paths cover the whole
-tree, i.e. $\bigcup_{i \in I}P_i = V$ \\
-
-\noindent
-Generalizing the definition of {\em feasibility} in \cite{nsnrs09} to
-a set assignment, a path assignment $\A$ is defined to be {\em
-  feasible} if there exists a bijection defined as follows.
-\begin{align}
-  \sigma: U \rightarrow V(T), \text{ such that }\sigma(S_i) = P_i
-  \text{ for all } i \in I, \sigma \text{ is a bijection}
-  \label{eq:stf}
-\end{align}
-% The assignment $\A$ is then called {\em Set Translation Feasible}.
-
-\noindent
-Let $X$ be a partially ordered set with $\preccurlyeq$ being the
-partial order on $X$.  $mub(X)$ represents an element in $X$ which is
-a maximal upper bound on $X$.  $X_m \in X$ is a maximal upper bound of
-$X$ if $\nexists X_q \in X$ such that $X_m
-\preccurlyeq X_q$. \\
-
-\noindent
-The set $I$ represents the index set $[m]$. If index $i$ is used
-without further qualification, it is meant to be $i \in I$. Any
-function, if not defined on a domain of sets, when applied on a set is
-understood as the function applied to each of its elements. i.e. for
-any function $f$ defined with domain $U$, the abuse of notation is as
-follows; $f(S)$ is used instead of $\hat f(S)$ where
-$\hat f(S) = \{y \mid y = f(x), x \in S\}$. \\
-
-\noindent
-When refering to a tree as $T$ it could be a reference to the tree
-itself, or the vertices of the tree. This will be clear from the
-context.\\
-
-\noindent
-Finally, an in-tree is a directed rooted tree in which all edges are
-directed toward to the root.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC\subsection{Characerization of Feasible Tree Path
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC  Assignment\tnote{Merge with the original charac section}} 
-\label{feasible} 
-Consider a path assignment $\cA =
-\{(S_i, P_i) \mid S_i \in \cF, P_i \text{ is a path from $T$}, i \in
-[m]\}$ to a set system $\cF = \{S_i \mid S_i \subseteq U, i \in
-[m]\}$, were $T$ is a given tree, $U$ is the set system's universe and
-$m$ is the number of sets in $\cF$. We call $\cA$ an {\em Intersection
-  Cardinality Preserving Path Assignment (ICPPA)} if it has the
-following properties.
-
-\begin{enumerate}
-\item [i.]  $|S_i| = |P_i|$ for all $i \in [m]$
-\item [ii.] $|S_i \cap S_j| = |P_i \cap P_j|$ for all $i,j \in [m]$
-  % \item [iii.] $|\bigcup_{i \in J} S_i| = |\bigcup_{i \in J} P_i|$
-  %   for all $J \subseteq I$\footnote{TBD: this is needed now to fix
-  %     some proofs}
-\item [iii.] $|S_i \cap S_j \cap S_k| = |P_i \cap P_j \cap P_k|$ for
-  all $i,j,k \in [m]$
-\end{enumerate}
+{\tt WALCOM begin}
 
@@
 \end{proof}
-
-\noindent
-In the remaining part of this section we show that a path assignment
-is feasible if and only if it is an ICPPA.  One direction of this
-claim is clear: that is a path assignment is feasible, then all
-intersection cardinalities are preserved, that is the path assignment
-is an ICPPA.  The reason is that a feasible path assignment has an
-associated bijection between $U$ and $V(T)$ such that the sets map to
-paths.  The rest of the section is devoted to constructively proving
-that it is sufficient, for the existence of a bijection, for a path
-assignment to be an ICPPA.  At a top-level, the constructive
-approaches refine the path assignment iteratively, such that at the
-end of each iteration we have a path assignment, and finally we have a
-family of bijections.  First we present and then prove the correctness
-of Algorithm \ref{perms}.  This algorithm refines the path assignment
-by considering pairs of paths that share a leaf.
-
-\begin{algorithm}[h]
-  \caption{Permutations from an ICPPA $\{(S_i,P_i) | i \in I\}$}
-  \label{perms}
-  \begin{algorithmic}
-    \STATE Let $\Pi_0=\{(S_i,P_i)| i \in I\}$\\
-    \STATE $j = 1$;
-    \label{shareleaf} \WHILE {There is $(P_1,Q_1), (P_2,Q_2) \in
-      \Pi_{j-1}$ with $Q_1$ and $Q_2$ having a common leaf} \STATE
-    $\Pi_j= \Pi_{j-1} \setminus \{(P_1,Q_1),(P_2,Q_2)\}$;
-    \label{setbreak}\STATE $\Pi_j = \Pi_j \cup \{(P_1 \cap P_2,Q_1
-    \cap Q_2), (P_1 \setminus P_2,Q_1 \setminus Q_2), (P_2 \setminus
-    P_1, Q_2 \setminus Q_1)\}$; \STATE $j = j+1$; \ENDWHILE \STATE
-    $\Pi = \Pi_j$; \STATE Return $\Pi$;
-  \end{algorithmic}
-\end{algorithm}
 
@@
 \end{proof}
-
-\noindent
-It can be observed that the output of algorithm \ref{perms} is such
-that every leaf is incident on at most a single path in the new set of
-assignments. This is due to the loop condition at line
-\ref{shareleaf}. Let $v_1$ be the leaf incident on path $P_i$. Assign
-to it any one element from $S_i \setminus \bigcup_{i \ne j}
-S_j$. Remove $(S_i, P_i)$ from assignments and add $(\{x_1\},
-\{v_1\}), (S_i \setminus \{x_1\}, P_i \setminus \{v_1\})$. Now all
-assignments except single leaf assignments are paths from the subtree
-$T_0 = T \setminus \{v \mid v \text{ is a leaf in } T\}$.
-
-% exclusively on the spine of the caterpillar. Consider the spine to
-% be an interval and run ICPIA algorithm on it to get the required
-% permutation. Algorithm \ref{leafasgn} makes this clear.
+{\tt WALCOM end}
+
+\begin{lemma}
+  \label{lem:invar1} In Algorithm~\ref{perms}, at the end of $j$th
+  iteration, $j \ge 0$, of the {\bf while} loop, the following
+  invariants are maintained.
+  \begin{enumerate}[I {\ }] %\vspace{\topshrink}
+  \item $\cl_j(R)$ is a path in $T$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
+    \ \ \ \ \ \ \ \ \ \ for all $R \in \cF_j$%\vspace{\topshrink}
+  \item $|R| = |\cl_j(R)|$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
+    \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $R \in
+    \cF_j$%\vspace{\topshrink}
+  \item $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, \ \ \ \ \ \ \ \ \ \
+    \ \ \ \ \ \ \ \ \ \ for all $R, R' \in \cF_j$%\vspace{\topshrink}
+  \item $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
+    \cl_j(R'')|$, \ \ \ for all $R, R', R'' \in \cF_j$
+  \end{enumerate}
+\end{lemma}
+
+\begin{proof}
+ Proof is by induction on the number of iterations, $j$. In this
+  proof, the term ``new sets'' will refer to the sets added to $\cF_j$
+  in $j$th iteration in line~\ref{setbreak} of Algorithm~\ref{perms},
+  $S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1$ and its
+  images in $\cl_j$ where $\cl_{j-1}(S_1)$
+  and $\cl_{j-1}(S_2)$ intersect and share a leaf.\\
+  The invariants are true in the base case $(\cF_0, \cl_0)$, since it
+  is the input ICPPL.  Assume the lemma is true till the $j-1$th
+  iteration. Let us consider the possible cases for each of the above invariants for
+  the $j$th iteration.
+
+ \begin{enumerate}[\xbullet]
+  \item {\em Invariant} I/II
+    \begin{enumerate}[{I/II}a $|$] % \textbullet 
+    \item {\em $R$ is not a new set.} It is in $\cF_{j-1}$. Thus
+      trivially true by induction hypothesis.
+    \item {\em $R$ is a new set.} If $R$ is in $\cF_{j}$ and not in
+      $\cF_{j-1}$, then it must be one of the new sets added in
+      $\cF_j$. In this case, it is clear that for each new set, the
+      image under $\cl_j$ is a path since by definition the chosen
+      sets $S_1$, $S_2$ are from $\cF_{j-1}$ and due to the while loop
+      condition, $\cl_{j-1}(S_1)$, $\cl_{j-1}(S_2)$ have a
+      common leaf. Thus invariant I is proven.\\
+      Moreover, due to induction hypothesis of invariant III and the
+      definition of $l_j$ in terms of $l_{j-1}$, invariant II is
+      indeed true in the $j$th iteration for any of the new sets.  If
+      $R = S_1 \cap S_2$, $|R| = |S_1 \cap S_2| = |\cl_{j-1}(S_1) \cap
+      \cl_{j-1}(S_2)| = |\cl_j(S_1 \cap S_2)| = |\cl_j(R)|$.
+      If $R = S_1 \setminus S_2$, $|R| = |S_1 \setminus S_2| = |S_1| -
+      |S_1 \cap S_2| = |\cl_{j-1}(S_1)| - |\cl_{j-1}(S_1) \cap
+      \cl_{j-1}(S_2)| = |\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)| =
+      |\cl_j(S_1 \setminus S_2)|
+      = |\cl_j(R)|$. Similarly if $R = S_2 \setminus S_1$.\\
+    \end{enumerate}
+  \item {\em Invariant} III
+    \begin{enumerate}[{III}a $|$]
+    \item {\em $R$ and $R'$ are not new sets.} It is in
+      $\cF_{j-1}$. Thus trivially true by induction hypothesis.
+    \item {\em Only one, say $R$, is a new set.} Due to invariant IV
+      induction hypothesis, Lemma~\ref{lem:setminuscard} and
+      definition of $\cl_j$, it follows that invariant III is true no
+      matter which of the new sets $R$ is equal to. If $R = S_1 \cap
+      S_2$, $|R \cap R'| = |S_1 \cap S_2 \cap R'| = |\cl_{j-1}(S_1)
+      \cap \cl_{j-1}(S_2) \cap \cl_{j-1}(R')| = |\cl_j(S_1 \cap S_2)
+      \cap \cl_j(R')| = |\cl_j(R) \cap \cl_j(R')|$.  If $R = S_1
+      \setminus S_2$, $|R \cap R'| = |(S_1 \setminus S_2) \cap R'| =
+      |(\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)) \cap \cl_{j-1}(R')| =
+      |\cl_{j}(S_1 \cap S_2) \cap \cl_{j}(R')| = |\cl_{j}(R) \cap
+      \cl_{j}(R')|$. Similarly, if $R = S_2 \setminus
+      S_1$. Note $R'$ is not a new set.\\
+
+    \item {\em $R$ and $R'$ are new sets.} By definition, the new
+      sets and their path images in path label $\cl_j$ are disjoint so
+      $|R \cap R'| = |\cl_j(R) \cap \cl_j(R)| = 0$. Thus case proven.
+    \end{enumerate}
+  \item {\em Invariant} IV
+
+    Due to the condition in line~\ref{ln:3waycheck}, this invariant is
+    ensured at the end of every iteration.
+%     \begin{enumerate} [{Case 3.}1:]
+%     \item {\em $R$, $R'$ and $R''$ are not new sets.} Trivially
+%       true by induction hypothesis.
+%     \item {\em Only one, say $R$, is a new set.}
+%       If $R = S_1 \cap S_2$,  from Lemma~\ref{lem:fourpaths} and
+%       invariant III hypothesis,  this case is proven. Similarly if $R$
+%       is any of the other new  sets, the case is proven by also using
+%       Lemma ~\ref{lem:setminuscard}.
+%     \item {\em At least two of $R, R', R''$ are new sets.}
+%       The new sets are disjoint hence this case is vacuously true.
+%     \end{enumerate}
+  \end{enumerate} \qed
+%\vspace{-6mm}
+
+\end{proof}
+
+\begin{lemma}
+  \label{lem:noexit1}
+  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{perms} is
+  feasible, then the set of hypergraph isomorphism functions that
+  defines $(\cF, \cl)$'s feasibility is the same as the set that
+  defines $(\cF_j, \cl_j)$'s feasibility, if any.  Secondly, for any
+  iteration $j > 0$ of the {\em \bf while} loop, the {\em \bf exit}
+  statement in line~\ref{ln:exit1} will not execute.
+\end{lemma}
+\begin{proof}
+  Since $(\cF,\cl)$ is feasible, by Lemma~\ref{lem:feasible}
+  $(\cF_j,\cl_j)$ for every iteration $j > 0$ is feasible.  % Therefore,
+%   every hypergraph isomorphism $\phi: supp(\cF) \rightarrow V(T)$ that
+%   induces $\cl$ on $\cF$ also induces $\cl_{j-1}$ and $\cl_{j}$ on
+%   $\cF_{j-1}$ and $\cF_{j}$ respectively, i.e., $\cl_{\phi[\cF_{j-1}]}
+%   = \cl_{j-1}$ and $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
+%   that for all $x \in supp(\cF)$, for all $v \in V(T)$ the following
+%   hold true.
+  Also, every hypergraph isomorphism $\phi: supp(\cF) \rightarrow
+  V(T)$ that induces $\cl$ on $\cF$ also induces $\cl_{j}$ on
+  $\cF_{j}$, i.e., $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
+  that for all $x \in supp(\cF)$, for all $v \in V(T)$, if $(x,v) \in
+  \phi$ then $v \in \cl_{j}(S)$ for all $S \in \cF_{j}$ such that $x
+  \in S$.
+% the following
+%   hold true.
+%   \begin{enumerate}[i. ]
+%   \item If $(x,v) \in \phi$ then $v \in \cl_{j-1}(S)$ for all $S \in
+%     \cF_{j-1}$ such that $x \in S$.
+%   \item If $(x,v) \in \phi$ then $v \in \cl_{j}(S)$ for all $S \in
+%     \cF_{j}$ such that $x \in S$
+%   \end{enumerate}
+  In other words, filter 1 outputs a filtered path labeling that
+  ``preserves''
+  hypergraph isomorphisms of the original path labeling.\\
+  Secondly, line~\ref{ln:exit1} will execute iff the exit condition in
+  line~\ref{ln:3waycheck}, i.e. failure of three way intersection
+  preservation, becomes true in any iteration of the {\em \bf while}
+  loop.  Due to Lemma~\ref{lem:invar1} Invariant IV, the exit
+  condition does not occur if the input is a feasible ICPPL.\qed
+
+%   such that $\phi(x) = v$ where $v$ is the leaf considered in the
+%   first iterations of while. Clearly, $\phi$ is a renaming of
+%   vertices in hypergraph $\cF$ to those in hypergraph $\cF^\cl$. Thus
+%   the following facts can be observed in every iteration of the loop.
+
+%   \begin{enumerate}[\hspace{2mm}i. ] \vspace{\topshrink}
+%   \item all intersection cardinalities are preserved in this path
+%     labeling \vspace{\topshrink}
+%   \item element $x$ is exclusive in a hyperedge in $\cF$ since $v$ is
+%     exclusive in a hyperedge in $\cF^\cl$.
+%   \end{enumerate}
+
+%   Thus the exit condition is never rendered true after $x$ and $v$ are
+%   removed from their respective hyperedges. \qed
+
+% \noindent
+% This proof uses mathematical induction on the number
+%   of iterations $j$, $j \ge 0$, of the loop that executed
+%   without exiting. The base case, $j = 0$ is obviously true since the
+%   input is an ICPPL and the exit condition cannot hold true due to
+%   ICPPL property (iii).  Assume the algorithm executes till the end
+%   of $j-1$th iteration without exiting at line
+%  ~\ref{ln:3waycheck}. Consider the $j$th iteration. From Lemma
+%  ~\ref{lem:feasible} we know that $(\cF_j, \cl_j)$ and $(\cF_{j-1},
+%   \cl_{j-1})$ are feasible\remove[AS]{and from the proof in lemma
+%     lem:invar1 we know that $(\cF_{j-1}, \cl_{j-1})$ satisfies all the
+%     invariants defined in the lemma}.  Thus there exists a bijection
+%   $\phi: supp(\cF) \rightarrow V(T)$ such that the induced path
+%   % labeling on $\cF_{j-1}$ $\cl_{\phi[\cF_{j-1}]} = \cl_{j-1}$.
+%   labeling on $\cF_{j}$, $\cl_{\phi[\cF_{j}]}$ and on $\cF_{j-1}$,
+%   $\cl_{\phi[\cF_{j-1}]}$ are equal to $\cl_{j}$ and $\cl_{j-1}$
+%   respectively.  We need to prove that for any $R, R', R'' \in
+%   \cF_{j}$, $|R \cap R' \cap R''| = |\cl_j(R) \cap \cl_j(R') \cap
+%   \cl_j(R'')|$.
+%   The following are the possible cases that could arise. From argument
+%   above, $|\cl_j(R) \cap \cl_j(R') \cap \cl_j(R'')| =
+%   |\cl_{\phi[\cF_{j}]}(R) \cap \cl_{\phi[\cF_{j}]} (R') \cap
+%   \cl_{\phi[\cF_{j}]} (R'')|$
+
+%   \begin{enumerate}[a $|$]
+%   \item {\em None of the sets are new. $R, R', R'' \in \cF_{j-1}$.}
+%     We know $(\cF_{j-1}, \cl_{j-1})$ is feasible. Thus $|R \cap R'
+%     \cap R''| = |\cl_{j-1}(R) \cap \cl_{j-1}(R') \cap \cl_{j-1}(R'')|
+%     = |\cl_{j}(R) \cap \cl_{j}(R') \cap \cl_{j}(R'')|$.
+%   \item {\em Only one, say $R$, is a new set.}  Let $R = S_1 \cap S_2$
+%     ($S_1, S_2$ are defined in the proof of lemma
+%    ~\ref{lem:invar1}). Now we have $|R \cap R' \cap R''| = |S_1 \cap
+%     S_2 \cap R' \cap R''| = |\cl_{j-1}(S_1) \cap \cl_{j-1}(S_2) \cap
+%     \cl_{j-1}(R') \cap \cl_{j-1}(R'')| = |\cl_{j}(R) \cap \cl_{j}(R')
+%     \cap \cl_{j}(R'')|$. Thus proven. If $R$ is any of the other new
+%     sets, the same claim can be verified using lemma
+%    ~\ref{lem:setminuscard}.
+%     % \item []{\bf Case 3:}
+%   \item {\em At least two of $R, R', R''$ are new sets.}  The new sets
+%     are disjoint hence this case is vacuously true.
+%   \end{enumerate}
+%   \qed
+%   \tnote[E2]{remove the induction proof. just text saying x and v are
+%     exclusive in these sets therefore the intersection cardinalities
+%     don't change thus all invariants are still true} 
+\end{proof}
+
+As a result of Algorithm~\ref{perms} each leaf $v$ in $T$
+is such that there is exactly one set in $\cF$ with $v$ as a vertex in
+the path assigned to it.  In Algorithm~\ref{leafasgn} we identify
+elements in $supp(\cF)$ whose images are leaves in a hypergraph
+isomorphism if one exists.  Let $S \in \cF$ be such that $\cl(S)$ is a
+path with leaf and $v \in V(T)$ is the unique leaf incident on it.  We
+define a new path labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) =
+\set{v}$ where $x$ an arbitrary element from $S \setminus \bigcup_{\hS
+  \ne S} \hS$. In other words, $x$ is an element present in no other
+set in $\cF$ except $S$. This is intuitive since $v$ is present in no
+other path image under $\cl$ other than $\cl(S)$.  The element $x$ and
+leaf $v$ are then removed from the set $S$ and path $\cl(S)$
+respectively. After doing this for all leaves in $T$, all path images
+in the new path labeling $\cl_{new}$ except leaf labels (a path that
+has only a leaf is called the {\em leaf label} for the corresponding
+single element hyperedge or set) are paths from a new pruned tree $T_0
+= T \setminus \{v \mid v \text{ is a leaf in }
+T\}$. Algorithm~\ref{leafasgn} is now presented with details.
 
 \begin{algorithm}[h]
-  \caption{Leaf assignments from an ICPPA $\{(S_i,P_i) | i \in I\}$}
+  \caption{Leaf labeling from an ICPPL {\tt filter\_2($\cF, \cl, T$)}}
   \label{leafasgn}
-  \begin{algorithmic}
-    \STATE Let $\Pi_0=\{(S_i,P_i)| i \in [m]\}$. Paths are such that
-    no
-    two paths $P_i, P_j, i \ne j$ share a leaf.\\
-    \STATE $j = 1$\\
-    \WHILE {there is a leaf $v$ and a unique $(S_{i_1}, P_{i_1})$ such
-      that $v \in P_{i_1}$}
-    \STATE $\Pi_j=   \Pi_{j-1} \setminus \{(S_{i_1}, P_{i_1})\}$\\
-    \STATE $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in I}S_i$
-    \IF{$X$ is empty} \STATE exit \ENDIF; \STATE Let $x = $ arbitrary
-    element from $X$ \vspace{2mm}
-
-    \STATE $\Pi_j = \Pi_j \cup \{(S_{i_1} \setminus \{x\}),(P_{i_1}
-    \setminus
-    \{v\}), (\{x\},\{v\})\}$\\
-    \STATE $j = j+1$\\
+  \begin{algorithmic}[\lndisplay]
+    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$
+    \COMMENT {Path images are such that no two path images share a
+      leaf.}
+    \STATE $j \assign 1$\\
+    \WHILE {there is a leaf $v$ in $T$ and a unique $S_1 \in
+      \cF_{j-1}$ such that $v \in \cl_{j-1}(S_1)$ }\label{uniqueleaf}
+    \STATE $\cF_j \assign \cF_{j-1} \setminus \{S_1\}$\\
+    \STATE for all $S \in \cF_{j-1}$ such that $S \ne S_1$ set
+    $\cl_j(S) \assign
+    \cl_{j-1}(S)$\\
+    \STATE $X \assign S_1 \setminus \bigcup_{S \in \cF_{j-1}, S \ne S_1}S$\\
+    \IF{$X$ is empty} \label{xempty} \STATE {\bf exit} \label{ln:exit2} \ENDIF
+    \STATE $x \assign $ arbitrary element from $X$\\
+    \STATE $\cF_j \assign \cF_j \cup \{\{x\}, S_1 \setminus \{x\}\} $\\
+    \STATE $\cl_j(\{x\}) \assign \{v\}$\\
+    \STATE $\cl_j(S_1 \setminus \{x\}) \assign \cl_{j-1}(S_1) \setminus \{v\}$\\
+    \STATE $j \assign j+1$\\
     \ENDWHILE
-    \STATE $\Pi = \Pi_j$\\
-    \STATE Return $\Pi$\\
+    \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
+    \RETURN $(\cF', \cl')$
   \end{algorithmic}
@@
 
+Suppose the input ICPPL $(\cF, \cl)$ is feasible, yet set $X$ in
+Algorithm~\ref{leafasgn} is empty in some iteration of the {\bf while}
+loop. This will abort our procedure of finding the hypergraph
+isomorphism. The following lemma shows that this will not happen.
+
+{\tt WALCOM begin}
 \begin{lemma}
@@
 
-\noindent
 Using algorithms \ref{perms} and \ref{leafasgn} we prove the following
@@
 \end{proof}
+
+{\tt WALCOM end}
+
+\begin{lemma}
+  \label{lem:xnotempty}
+  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{leafasgn} is
+  feasible, then for all iterations $j > 0$ of the {\em \bf while}
+  loop, the {\em \bf exit} statement in line~\ref{ln:exit2} does not
+  execute.
+\end{lemma}
+\begin{proof}
+  Assume $X$ is empty for some iteration $j > 0$. We know that $v$ is
+  an element of $\cl_{j-1}(S_1)$. Since it is uniquely present in
+  $\cl_{j-1}(S_1)$, it is clear that $v \in \cl_{j-1}(S_1) \setminus
+  \bigcup_{(S \in \cF_{j-1}) \wedge (S \ne S_1)}\cl_{j-1}(S)$.  Note
+  that for any $x \in S_1$ it is contained in at least two sets due to
+  our assumption about cardinality of $X$. Let $S_2 \in \cF_{j-1}$ be
+  another set that contains $x$. From the above argument, we know $v
+  \notin \cl_{j-1}(S_2)$. Therefore there cannot exist a hypergraph
+  isomorphism bijection that maps elements in $S_2$ to those in
+  $\cl_{j-1}(S_2)$. This contradicts our assumption that the input is
+  feasible. Thus $X$ cannot be empty if input is ICPPL and feasible.
+  \qed
+\end{proof}
+
+\begin{lemma}
+  \label{lem:invar3}
+  In Algorithm~\ref{leafasgn}, for all $j > 0$, at the end of the
+  $j$th iteration of the {\bf while} loop the four invariants given in
+  Lemma~\ref{lem:invar1} hold.
+\end{lemma}
+\begin{proof}
+  By Lemma~\ref{lem:xnotempty} we know that set $X$ will not be empty
+  in any iteration of the {\em \bf while} loop if input ICPPL $(\cF,
+  \cl)$ is feasible and $\cl_j$ is always computed for all $j >
+  0$. Also note that removing a leaf from any path keeps the new path
+  connected. Thus invariant I is obviously true. In every iteration $j
+  > 0$, we remove exactly one element $x$ from one set $S$ in $\cF$
+  and exactly one vertex $v$ which is a leaf from one path
+  $\cl_{j-1}(S)$ in $T$. This is because as seen in
+  Lemma~\ref{lem:xnotempty}, $x$ is exclusive to $S$ and $v$ is
+  exclusive to $\cl_{j-1}(S)$. Due to this fact, it is clear that the
+  intersection cardinality equations do not change, i.e., invariants
+  II, III, IV remain true. On the other hand, if the input ICPPL is
+  not feasible the invariants are vacuously true. \qed
+\end{proof}
+
+% \textcolor{cyan}{
+% \begin{lemma}
+%   \label{lem:notfeasibleexit}
+%   \tnote{IS THIS CORRECT?}  If input ICPPL $(\cF, \cl)$ is not
+%   feasible, then in one of the recursive calls to Algorithm --3--, the
+%   {\em \bf exit} statement in line x in Algorithm --1 or line y in
+%   Algorithm --2 will get executed.
+% \end{lemma}
+% }
+
+We have seen two filtering algorithms above, namely,
+Algorithm~\ref{perms} {\tt filter\_1} and Algorithm~\ref{leafasgn}
+{\tt filter\_2} which when executed serially repectively result in a
+new ICPPL on the same universe $U$ and tree $T$. We also proved that
+if the input is indeed feasible, these algorithms do indeed output the
+filtered ICPPL. Now we present the algorithmic characterization of a
+feasible tree path labeling by way of Algorithm~\ref{al:icppl-find-isomorph}.
+
+Algorithm~\ref{al:icppl-find-isomorph} computes a
+hypergraph isomorphism $\phi$ recursively using Algorithm~\ref{perms}
+and Algorithm~\ref{leafasgn} and pruning the leaves of the input
+tree. In brief, it is done as follows. Algorithm~\ref{leafasgn} gives
+us the leaf labels in $\cF_2$, i.e., the elements in $supp(\cF)$ that
+map to leaves in $T$, where $(\cF_2, \cl_2)$ is the output of
+Algorithm~\ref{leafasgn}. All leaves in $T$ are then pruned away. The
+leaf labels are removed from the path labeling $\cl_2$ and the
+corresponding elements are removed from the corresponding sets in
+$\cF_2$. This tree pruning algorithm is recursively called on the
+altered hypergraph $\cF'$, path label $\cl'$ and tree $T'$. The
+recursive call returns the bijection $\phi''$ for the rest of the
+elements in $supp(\cF)$ which along with the leaf labels $\phi'$ gives
+us the hypergraph isomorphism $\phi$.  The following lemma formalizes
+the characeterization of feasible path labeling.
+
+\begin{algorithm}[h]
+  \caption{{\tt get- hypergraph- isomorphism ($\cF, \cl, T$)}}
+  \label{al:icppl-find-isomorph}
+  \begin{algorithmic}[\lndisplay]
+
+    \IF{$T$ is empty}
+    \RETURN $\emptyset$\\
+    \ENDIF
+    \STATE $L \assign \{v \mid v \text{ is a leaf in }      T\}$\\
+    \STATE $(\cF_1, \cl_1) \assign$ {\tt filter\_1($\cF, \cl,
+      T$)}\\
+    \STATE $(\cF_2, \cl_2) \assign$ {\tt filter\_2($\cF_1,
+      \cl_1, T$)}\\
+
+    \STATE $(\cF', \cl') \assign (\cF_2, \cl_2)$\\
+    \STATE $\phi' \leftarrow \emptyset$
+
+    \FOR {every $v \in L$}
+    \STATE $\phi'(x) \assign v$ where $x \in \cl_2^{-1}(\{v\})$
+    \COMMENT {Copy the leaf labels to a one to one function $\phi':
+      supp(\cF) \rightarrow L$
+      }\\
+    \STATE Remove $\{x\}$ and $\{v\}$ from $\cF'$, $\cl'$  appropriately\\
+    \ENDFOR
+
+    \STATE $T' \assign T \setminus L$
+
+    \STATE $\phi'' \assign$ {\tt
+      get-hypergraph-isomorphism($\cF', \cl', T'$)}
+    \STATE $\phi \assign \phi'' \cup \phi'$ \\
+    \RETURN $\phi$
+  \end{algorithmic}
+\end{algorithm}
+
+\begin{lemma}
+  \label{lem:hyperiso}  %{lem:perm}
+  If $(\cF, \cl)$ is an ICPPL from a tree $T$ and
+  Algorithm~\ref{al:icppl-find-isomorph}, {\tt
+    get- hypergraph- isomorphism ($\cF, \cl, T$)} returns a non-empty
+  function, then there exists a hypergraph isomorphism $\phi :
+  supp(\cF) \rightarrow V(T)$ such that the $\phi$-induced tree path
+  labeling is equal to $\cl$ or $\cl_\phi = \cl$.
+\end{lemma}
+\begin{proof}
+  It is clear that in the end of every recursive call to
+  Algorithm~\ref{al:icppl-find-isomorph}, the function $\phi'$ is one
+  to one involving all the leaves in the tree passed as input to that
+  recursive call. Moreover, by Lemma~\ref{lem:noexit1} and
+  Lemma~\ref{lem:xnotempty} it is consistent with the tree path
+  labeling $\cl$ passed. The tree pruning is done by only removing
+  leaves in each call to the function and is done till the tree
+  becomes empty. Thus the returned function $\phi: supp(\cF)
+  \rightarrow V(T)$ is a union of mutually exclusive one to one
+  functions exhausting all vertices of the tree. In other words, it is
+  a bijection from $supp(\cF)$ to $V(T)$ inducing the given path
+  labeling $\cl$ and thus a hypergraph isomorphism. \qed
+\end{proof}
+
+\begin{theorem}
+  \label{th:charac}
+  A path labeling $(\cF, \cl)$ on tree $T$ is feasible iff it is an
+  ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
+  T)$ as input returns a non-empty function.
+\end{theorem}
+\begin{proof}
+  From Lemma~\ref{lem:hyperiso}, we know that if $(\cF, \cl)$ is an
+  ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
+  T)$ as input returns a non-empty function, $(\cF, \cl)$ is feasible.
+  Now consider the case where $(\cF, \cl)$ is feasible. i.e. there
+  exists a hypergraph isomorphism $\phi$ such that $\cl_\phi =
+  \cl$. Lemma~\ref{lem:noexit1} and Lemma~\ref{lem:xnotempty} show us
+  that filter 1 and filter 2 do not exit if input is feasible. Thus
+  Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
+  function.\qed
+\end{proof}
+
+\section{ Computing feasible TPL with special target trees \tnote{give
+    problem definition etc}}
+\label{sec:spltargettree}
+
+Section~\ref{sec:feasible} described properties that a TPL must have
+for it to be feasible. The next problem of interest is to test if a
+given hypergraph is a path hypergraph with respect to a given target
+tree\footnote{A larger problem would be to find if a given hypergraph
+  is a path graph on any tree. This problem is not addressed in this
+  thesis.}. In other words, the problem is to find out if a feasible
+tree path labeling exists from a given target tree for a given
+hypergraph. In this section we will see two special cases of this
+problem where the target tree is from a particular family of
+trees. The first one, where the tree is a path, is a shown to be
+equivalent to the well studied problem of consecutive-ones in
+Section~\ref{sec:icpplicpia}. The second one, where the tree is a
+$k$-subdivided tree\footnote{See Section~\ref{ch:prelims} for the
+  formal definition.}, has been solved using a polynomial time
+algorithm. The latter problem enforces some conditions on the
+hypergraph too which will be seen in Section~\ref{sec:ksubdivstar}.
+
+\subsection{Target tree is a Path}
+\label{sec:icpplicpia}
+
+Consider a special case of ICPPL with the following properties.
+\begin{enumerate}
+\item Given tree $T$ is a path. Hence, all path labels are interval
+  labels.
+\item Only pairwise intersection cardinality preservation is
+  sufficient. i.e. property (iii) in ICPPL is not enforced.
+\item The filter algorithms do not have {\em \bf exit} statements.
+\end{enumerate}
+This is called an Intersection Cardinality Preservation Interval
+Assignment (ICPIA) \cite{nsnrs09}. This structure and its algorithm is
+used in the next section for finding tree path labeling from a
+$k$-subdivided star due to this graph's close relationship with
+intervals.
+
+\subsection{Target tree is a $k$-subdivided Star}
+\label{sec:ksubdivstar}
+
+\begin{figure}[t] %[htbp] 
+  \centering
+  \begin{tabular}{lr}
+    (a) \includegraphics{cop-tree_thesis.1}&
+    (b) \includegraphics{cop-tree_thesis.2}
+  \end{tabular}
+  \caption{\figtabsize (a) $8$-subdivided star with 7 rays (b)
+    3-subdivided star with 3 rays}
+  \label{fig:kstar}
+\end{figure}
+
+In this section we consider the problem of assigning paths from a
+$k$-subdivided star $T$ to a given set system $\cF$.  We consider
+$\cF$ for which the overlap graph $\bO(\cF)$ is connected.  The
+overlap graph is well-known from the work of
+\cite{kklv10,nsnrs09,wlh02}.  We use the notation in
+\cite{kklv10}. Recall from Section~\ref{ch:prelims} that hyperedges
+$S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$
+and $S'$ have a non-empty intersection but neither of them is
+contained in the other. The overlap graph $\bO(\cF)$ is a graph in
+which the vertices correspond to the sets in $\cF$, and the vertices
+corresponding to the hyperedges $S$ and $S'$ are adjacent if and only
+if they overlap.  Note that the intersection graph of $\cF$,
+$\bI(\cF)$ is different from $\bO(\cF)$ and $\bO(\cF) \subseteq
+\bI(\cF)$.  A connected component of $\bO(\cF)$ is called an overlap
+component of $\cF$.  An interesting property of the overlap components
+is that any two distinct overlap components, say $\cO_1$ and $\cO_2$,
+are such that any two sets $S_1 \in \cO_1$ and $S_2 \in \cO_2$ are
+disjoint, or, w.l.o.g, all the sets in $\cO_1$ are contained within
+one set in $\cO_2$.  This containment relation naturally determines a
+decomposition of the overlap components into rooted containment trees.
+We consider the case when there is only one rooted containment tree,
+and we first present our algorithm when $\bO(\cF)$ is connected.  It
+is easy to see that once the path labeling to the overlap component in
+the root of the containment tree is achieved, the path labeling to the
+other overlap components in the rooted containment tree is essentially
+finding a path labeling when the target tree is a path: each target
+path is a path that is allocated to sets in the root overlap
+component.  Therefore, for the rest of this section, $\bO(\cF)$ is a
+connected graph. We also assume that all hyperedges are of cardinality
+at most $k+2$.
+
+Recall from Section~\ref{ch:prelims} that a $k$-subdivided star is a
+star with each edge subdivided $k$ times. Therefore, a $k$-subdivided
+star has a central vertex which we call the {\em root}, and each root
+to leaf path is called a {\em ray}. First, we observe that by removing
+the root $r$ from $T$, we get a collection of $p$ vertex disjoint
+paths of length $k+1$, $p$ being the number of leaves in $T$.  We
+denote the rays by $R_1, \ldots, R_p$, and the number of vertices in
+$R_i$, $i \in [p]$ is $k+2$.  Let $\seq{v_{i1},\ldots,v_{i(k+2)}=r}$
+denote the sequence of vertices in $R_i$, where $v_{i1}$ is the
+leaf. Note that $r$ is a common vertex to all $R_i$.
+
+In this section the given hypergraph, the $k$-subdivided star and the
+root of the star are denoted by $\cO$, $T$ and vertex $r$,
+respectively.
+% For each hyperedge $X \in \cO$, we will maintain a 2-tuple of non-negative
+% numbers $\seq{p_1(X), p_2(X)}$.  The numbers satisfy the property that
+% $p_1(X) + p_2(X) \leq |X|$, and at the end of path labeling, for each
+% $X$, $p_1(X) + p_2(X) = |X|$.  This signifies the algorithm tracking
+% the lengths of subpaths of the path assigned to $X$ from at most two
+% rays. We also maintain another parameter called the {\em residue} of
+% $X$ denoted by $s(X)=|X| - p_1(X)$. This signifies the residue path
+% length that must be assigned to $X$ which must be from another
+% ray. For instance, if $X$ is labeled a path from only one ray, then
+% $p_1(X) = |X|$, $p_2(X) = 0$ and $s(X) = 0$.
+
+% 
+% We iteratively consider each ray from which paths will be assigned to
+% hyperedges. At the beginning of each iteration hyperedges of $\cO$ are
+% classifed into the following disjoint sets.
+% \begin{enumerate}
+% \item [$\cL_1^i$] {\em Labeled without $r$.} Those that have been
+%   labeled with a path which does not contain $r$ in one of the
+%   previous iterations.\\  $\cL_1^i = \set{ X \mid p_1(X) = |X| \text{ and
+%     } p_2(X) = 0 \text{ and } s(X) = 0, X \in \cO}$
+% \item [$\cL_2^i$] {\em Labeled with $r$.} Those that have been labeled
+%   with two subpaths of $\cl(X)$ containing $r$ from two different rays
+%   in two previous iterations.\\ $\cL_2^i = \set{X \mid 0 < p_1\left(X\right),
+%     p_2\left(X\right) < |X| \text{ and } s(X) = 0, X \in \cO}$
+%   \item [$\cT_1^i$] {\em Type 1 / partially labeled.} Those that have
+%     been labeled with one path containing $r$ from a single ray in one
+%     of the previous iterations. Here, $p_1(X)$ denotes the length of
+%     the subpath of $\cl(X)$ that $X$ has been so far labeled
+%     with. Also, it is clear that such a path must start at $r$ and
+%     must be in a ray different from the one corresponding to $p_1(X)$.\\
+%     $\cT_1^i = \set{ X \mid 0 < p_1(X) < |X| \text{ and } p_2(X) = 0
+%       \text{ and } s(X) > 0, X \in \cO}$
+%   \item [$\cT_2^i$] {\em Type 2 / not labeled.} Those that have not been
+%     labeled with a path in any previous iteration.\\
+%     $\cT_2^i = \set{ X \mid p_1(X) = p_2(X) = 0 \text{ and } s(X) = |X|,
+%       X \in \cO}$
+% \end{enumerate}
+% \vspace{-2mm}
+% \begin{align*}
+%   \cO &= \cL_1^i \cup \cL_2^i \cup \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]\\
+%   \cO_i &= \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]
+% \end{align*}
+
+The set $\cO_i$ refers to the set of hyperedges $\cT_1^i \cup \cT_2^i$
+in the $i$th iteration.  Note that $\cO_1 = \cO$.  In the $i$th
+iteration, hyperedges from $\cO_i$ are assigned paths from $R_i$ using
+the following rules. Also the end of the iteration, $\cL_1^{i+1},
+\cL_2^{i+1}, \cT_1^{i+1}, \cT_2^{i+1}$ are set to $\cL_1^{i},
+\cL_2^{i}, \cT_1^{i}, \cT_2^{i}$ respectively, along with some
+case-specific changes mentioned in the rules below.
+
+\noindent
+Suppose the given hypergraph has a feasible path labeling to the given
+tree $T$. Let this labeling be $\cl$. Now we present an algorithm that
+lists the sequence of edges that maintain a set of properties
+necessary for them to have a feasible tree path labeling.
+\begin{enumerate}[I.]
+\item {\bf Step 1}
+  \begin{enumerate}
+  \item {\bf {There are type 1
+        edges, $\mathbf{|\cT_1^i| > 0}$}:} 
+% , and for each such $X$ assign the
+%     path in $R_i$ of length $s(X)$ starting at $v_{i,k+2}=r$.  $X$ is
+%     then removed from $\cO_i$, and $p_2(X)=s(X)$, and $s(X)=0$.If any
+%     intersection cardinality is violated because of this assignment,
+%     EXIT reporting the non-existence of a feasible assignment.
+  {If there is only one $\cT_1^i$ hyperedge, label it to the unique
+    path in $R_i$ of length $s(X)$ starting at $v_{i(k+2)}=r$. If
+    there are more than one $\cT_1^i$ hyperedges, for any pair of
+    hyperedges $S_1, S_2 \in \cT_1^i$, one of the following is
+    true. Let $\cl'(X) \subseteq R_j$ denote the subpath of length
+    $p_1(X)$ that has been assigned to $X$ from a previously
+    considered ray $R_j$.}
+  \begin{enumerate}[{\em Case }1. ]
+  \item $|\cl'(S_1) \cap \cl'(S_2)| = 1$ and $|S_1 \cap S_2| > 1$
+    i.e. they are in different rays and their residues must contain
+    their intersection. Hence assign both to $R_i$.
+  \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| >
+    |\cl'(S_1) \cap \cl'(S_2)|$ i.e. they are in the same rays and
+    their residues must contain their intersection. Hence assign both
+    to $R_i$.
+  \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| =
+    |\cl'(S_1) \cap \cl'(S_2)|$. i.e. the intersection cardinality has
+    been met. Hence assign only $S_1$ to $R_i$ and $S_2$ must be
+    assigned to some $i \ge p$.
+  \end{enumerate}
+  Remove the labeled edge or edges from $\cO_i$ and add it to
+  $\cO_{i+1}$. Also do not add it to $\cT_1^{i+1}$ and add to
+  $\cL_2^{i+1}$.
+\item {\bf {There are no type-1 edges nor} super-marginal type-2
+    hyperedges:} If all type-2 edges of length at most $k+1$, EXIT by
+  saying there is no ICPPL.  Otherwise, Find an $X$ of type-2 such
+  that $|X| \geq k+2$, assign it the unique path starting at
+  $v_{i(k+1)}=r$ of length $k+2$, and set $p_1(X)=k+2,
+  s(X)=|X|-(k+2)$, assign it the unique path starting at
+  $v_{i(k+1)}=r$ of length $k+2$, mark this as a type-1 hyperedge, and
+  add it to $\cO_{i+1}$ after removing it from $\cO_i$.
+  % \remove{Next, iteratively, consider each type-1 edge $X \in
+  %   \cO_i$, and if it intersects a previously assigned $Y$ such that
+  %   $\cl(Y) \subseteq R_i$, assign a path of length $s(X)$
+  %   containing $v_{i(k+2)}=r$, set $p_2(X)=s(X), s(X)=0$, and remove
+  %   $X$ from $\cO_i$.}\tnote{covered in step 2}
+
+\item {\bf There are super-marginal type-2 hyperedges {and no type-1
+      edges:}} Let $X$ be a super-marginal hyperedge. If $|X| \leq
+  k+2$, then $X$ is assigned the path of length $|X|$ starting at
+  $v_{i1}$, and $p_1(X) = |X|, p_2(X)=0$. $X$ is removed from $\cO_i$
+  and not added to $\cO_{i+1}$.  If $|X| > k+2$, then $p_1(X)=k+2$,
+  assign it the unique path starting at $v_{i(k+1)}=r$ of length $k+2$
+  and $s(X) = |X| - (k+2)$.  In this case, $X$ is classified as a
+  Type-1 edge, removed from $\cO_i$, and added to $\cO_{i+1}$.
+\end{enumerate}
+\item {\bf Step 2:} Iteratively, a hyperedge $X$ is selected from
+  $\cO_i$ that has an overlap with one of the hyperedges $Y$ such that
+  $\cl(Y) \subseteq R_i$, and a unique path is assigned to $X$.  The
+  path, say $U(X) \subseteq R_i$, that is assigned can be decided
+  unambiguously since the $X \overlap Y$, and all intersection
+  cardinalities can be preserved only at one of the ends of $\cl(Y)$.
+
+  \noindent
+  Let $\cl(X)$ denote the unique path assigned to $X$.  If $X$ is a
+  type-2 hyperedge and if the unique path of length $|X|$ does not
+  contain $r$, then $p_1(X) = |X|, p_2(X)=0$, and $X$ is removed from
+  $\cO_i$.  In the case when $\cl(X)$ has to contain $r$, then
+  $p_1(X)$ is the length of the path, $p_2(X)=0$, and $s(X) =
+  |X|-p_1(X)$. Further $X$ is classified as a type-1 hyperedge, added
+  to $\cO_{i+1}$ and removed from $\cO_i$.  In the case when $X$ is a
+  type-1 hyperedge, then we check if $U(X)$, which is of length $s(X)$
+  contains $r$. If it does, then we assign $\cl(X) \leftarrow \cl(X)
+  \cup U(X)$, remove $X$ from $\cO_i$ and do not add it to
+  $\cO_{i+1}$.  If not, then we {\bf Exit} reporting that an
+  assignment cannot be found.  The iteration ends when no hyperedge in
+  $\cO_i$ has an overlap with a hyperedge assigned to $R_i$.
+\end{enumerate}
+
+\noindent
+The order in which they are assigned is the sequence of hyperedges
+with the following properties. Note that if such a sequence cannot be
+computed, $\cO$ cannot have a feasible path labeling from $T$.
+
+\noindent
+\tnote{need to formalize the below properties} In the following
+lemmas we identify a set of necessary conditions for $\cF$ to have an
+ICPPL in the $k$-subdivided star $T$.  If during the execution of the
+above algorithm, one of these necessary conditions is not satisfied,
+the algorithm exits reporting the non-existence of an ICPPL.
+\begin{lemma}
+  Let all hyperedges in $\cO_i$ be type-1 edges.  Then there is a
+  maximal subset $\ccT_i \subseteq \cO_i$ with the following
+  properties:
+  \begin{enumerate}
+  \item $\ccT_i$ form an inclusion chain.
+  \item For all $X \in \ccT_i$, $s(X) \leq k+2$, and There is a an $X
+    \in \ccT_i$ such that $s(X)=k+2$.
+  \end{enumerate}
+\end{lemma}
+\begin{lemma}
+  If there are no \marginal type-2 edges in $\cO_i$, then there exists
+  at least $p-i$ type-2 hyperedges $X \in \cO_i$ such that $|X| \geq
+  k+2$.
+  % is a exist type-2 hyperedges of length at least $k+2$, and there
+  % must be at least $r-i$ such type-2 hyperedge, basically at least
+  % one per ray.  Each hyperedge contained in such an hyperedges of
+  % length at least $k+2$, is transitively related under the overlap
+  % relation to one that gets assigned a path that contains
+  % $r$. Basically, the proof will argue that otherwise, there will be
+  % a \marginal type-2 hyperedge, or that it is infeasible.  Also, it
+  % is here we assume that we assume that the size of the two
+  % universes is the same.
+\end{lemma}
+\begin{lemma}
+  At the end of Step-1 in the $i$-th iteration, if one hyperedge $X$
+  of type-2 is such that $\cl(X) \subseteq R_i$, then all other
+  hyperedges in $\cO_i$ are connected to $X$ in the overlap component.
+\end{lemma}
+\begin{lemma}
+  At the end of Step-2, If control has exit at any time, there is no
+  ICPPL. If control has not exit, then $R_i$ is saturated.  No
+  hyperedge of $\cO_{i+1}$ will get a path from $R_i$ in the future
+  iterations.  No type-2 hyperedge of $\cO_{i+1}$ will get a path from
+  $R_i$.  Basically $R_i$ is done.
+\end{lemma}
+\begin{lemma}
+  Finally, we need to prove that the assignment is an ICPPL. Secondly,
+  if there is a permutation then maps sets to paths, then it is indeed
+  an ICPPL, and our algorithm will basically find it.  It is a unique
+  assignment upto permutation of the leaves.
+\end{lemma}
+
+\begin{lemma}
+  \label{lem:sup-mar}
+  If $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
+  tree path labeling to tree $T$, then $\cl(X)$ will
+  contain a leaf in $T$.\\
+  On the otherhand, if $\cl(X)$ has a leaf in
+  $T$, $X$ is marginal but may not be super-marginal.
+\end{lemma}
+\begin{proof}
+  Suppose $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
+  path labeling from $T$.  Assume $\cl(X)$ does not have a leaf.  Let
+  $R_i$ be one of the rays (or the only ray) $\cl(X)$ is part of.
+  Since $X$ is in a connected overlap component, there exists $Y_1 \in
+  \cF$ and $X \nsubseteq Y_1$ such that $Y_1 \overlap X$ and $Y_1$ has
+  at least one vertex closer to the leaf in $R_i$ than any vertex in
+  $X$. Similarly with the same argument there exists $Y_2 \in \cF$
+  with same subset and overlap relation with $X$ except it has has at
+  least one vertex farther away from the leaf in $R_i$ than any vertex
+  in $X$. Clearly $Y_1 \cap X$ and $Y_2 \cap X$ cannot be part of same
+  inclusion chain which contradicts that assumption $X$ is
+  super-marginal. Thus the claim is proven.\qed
+\end{proof}
+
+\noindent
+Consider the overlap graph $\bO(\cF)$ of the given hypergraph
+$\cF$. Let $S_{sm} \in \cF$ be such that it is a super-marginal
+hyperedge.  Algorithm~\ref{al:ktree-label} uses $S_{sm}$ along with
+the overlap graph $\bO(\cF)$ to calculate the feasible tree path
+labeling to the $k$-subdivded tree $T$.
+
+\begin{algorithm}[h]
+  \caption{{\tt compute-ksubtree-path-labeling($X, \cF, T$)}}
+  \label{al:ktree-label}
+  \begin{algorithmic}[\lndisplay]
+    \IF{$X = S_{sm}$}
+    \STATE -- TBD --\\
+    \ELSE 
+    \STATE -- TBD --\\
+    \ENDIF
+
+ \end{algorithmic}
+\end{algorithm}
+
+\section{ TPL with no restrictions}
+\label{sec:norestraint}
+
+{\tt abstract begin}
+  it is
+  known that if the given tree is a path a feasible assignment can be
+  found in polynomial time, and we observe that it can actually be
+  done in logspace. \tnote{[TRUE?]}
+{\tt abstract end}
+
+In Section \ref{prelims} we present the
+necessary preliminaries, in Section \ref{feasible} we present our
+characterization of feasible tree path assignments, and in Section
+\ref{decompos} we present the characterizing subproblems for finding a
+bijection between $U$ and $V(T)$ such that sets map to tree
+paths. Finally, in Section \ref{complexity} we conclude by showing
+that Tree Path Assignment is GI-Complete, and also observe that
+Consecutive Ones Testing is in Logspace.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC \subsection{Introduction}
+
+\annote{ We refer to this as the Tree Path Assignment
+problem for an input $(\cF,T)$ pair.}{Add this ``pair'' as a
+  definition in ch:prelims}
+
+\remove{It is an interesting fact that for a matrix with the COP, the
+intersection graph of the corresponding set system is an interval
+graph.  A similar connection to a subclass of chordal graphs, and this
+subclass contains interval graphs, exists for the generalization of
+COP.  In this case, the intersection graph of the corresponding set
+system must be a path graph. Chordal graphs are of great
+significance, extensively studied, and have several applications.  One
+of the well known and interesting properties of a chordal graphs is
+its connection with intersection graphs cite{mcg04}. For every chordal
+graph, there exists a tree and a family of subtrees of this tree such
+that the intersection graph of this family is isomorphic to the
+chordal graph cite{plr70,gav78,bp93}.  Certain format of these trees
+are called clique trees cite{apy92} of the graph which is a compact
+representation of the chordal graph. There has also been work done on
+the generalization of clique trees to clique hypergraphs cite{km02}.
+If the chordal graph can be represented as the intersection graph of
+paths in a tree, then the graph is called path graph cite{mcg04} or
+path chordal graph.  Therefore, it is clear that if there is a
+bijection from $U$ to $V(T)$ such that the sets map to paths, then the
+intersection graph of the set system is indeed a path chordal graph.
+This is, however, only a necessary condition and can be checked
+efficiently, as path graph recognition is polynomial
+solvable cite{gav78,aas93}.  Indeed, it is possible to construct a set
+system and tree, such that the intersection graph is a path chordal
+graph, but there is no bijection between $U$ and $V(T)$ such that the
+sets map to paths.  This connection indeed suggests that our problem
+is indeed as hard as path graph isomorphism.  Further path graph
+isomorphism is known be isomorphism-complete, see for example
+cite{kklv10}. }
+In the second part of this paper, we decompose our
+search for a bijection between $U$ and $V(T)$ into subproblems.  Each
+subproblem is on a set system in which for each set, there is another
+set in the set system with which the intersection is {\em strict}-
+there is a non-empty intersection, but neither is contained in the
+other.  This is in the spirit of results in \cite{wlh02,nsnrs09} where
+to test for COP in a given matrix, the COP problem is solved on an
+equivalent set of prime matrices.  \annote{Our decomposition localizes the
+challenge of path graph isomorphism to two problems.}{HOW?}
+
+Finally, we show that Tree Path Assignment is isomorphism-complete.
+We also point out Consecutive Ones Testing is in Logspace from two
+different results in the literature \cite{kklv10, mcc04}. To the best
+of our knowledge this observation has not been made earlier.
 
@@
 
-\subsection{Complexity of Tree Path Assignment-A
+\section{Complexity of Tree Path Assignment-A
   Discussion} 
@@
 to $V(T)$ such that the image of each set in $\cF$ is a path in $T$.
-We show that this problem is at least as hard as the problem of
-testing if two given path chordal graphs are isomorphic.
-\begin{theorem} \label{thm:isocomp} Tree Path Assignment is
-  isomorphism-complete.
-\end{theorem}
-\begin{proof}
-  It is well known (see for example \cite{kklv10}) that testing
-  isomorphism of path chordal graphs is isomorphism complete.  We show
-  a reduction of path chordal graph isomorphism to tree path
-  assignment.  Given $G_1$ and $G_2$ two path chordal graphs, let
-  $T_2$ be the clique tree of $G_2$ obtained from say \cite{gav78}.
-  The nodes of $T_2$ correspond to the maximal cliques of $G_2$ and
-  each vertex of $G_2$ corresponds to a path in $G_2$.  This is a
-  well-known characterization of path chordal graphs and $T_2$ can be
-  computed in polynomial time.  In $G_1$, let $S_v$ denote the maximal
-  cliques of $G_1$ that contain $v$.  This can be computed in
-  polynomial time as $G_1$ is path chordal, and all chordal graph only
-  have a linear number of maximal cliques.  The universe $U$
-  corresponds bijectively to the set of maximal cliques in $G_1$, and
-  $\cF = \{S_v | v \in V(G_1)\}$.  Now, we claim that $(\cF, T_2)$ has
-  a tree path assignment if and only if $G_1$ and $G_2$ are
-  isomorphic.  This is clear since for each vertex $v \in G_1$, there
-  is an associated $S_v$ which is the set of maximal cliques
-  containing $v$.  In $G_2$, each vertex corresponds to a path in
-  $T_2$, and the nodes on this path corresponds to the maximal cliques
-  in $G_2$.  Consequenlty, a tree path assignment will naturally yield
-  an isomorphism between $G_1$ and $G_2$, and vice versa.  Therefore,
-  Tree Path Assignment is isomorphism-complete.
-\end{proof}
-
-\subsubsection{Consecutive Ones Testing is in Logspace}
+
+\subsection{Consecutive Ones Testing is in Logspace}
 While Tree Path Assignment is isomorphism-complete, it is polynomial
@@
 
+% \noindent
+% It has been long known that interval graph recognition is in
+% logspace\cite{rei84}. Recently interval graph isomorphism was also
+% shown to be logspace decidable using a logspace canonization
+% algorithm by \cite{kklv10}.  This result is built on top of logspace
+% results of undirected graph connectivity \cite{rei08}, logspace
+% tractability using a certain logical formalism called FP+C and
+% modular decomposition of interval graphs\cite{lau10} etc.  Interval
+% graphs are closely connected to binary matrices with COP. The
+% maximal clique vertex incidence matrix (matrix with rows
+% representing maximal cliques and columns representing vertices of a
+% graph) has COP on columns iff the graph is an interval
+% graph\cite{fg65}. This follows from the interval graph
+% characterization by \cite{gh64}. Due to this close relation it is
+% natural to see if consecutive ones property can be
+% tested in logspace. \\
+% \noindent
+% We also explore some extensions of the interval assignment problem
+% in \cite{nsnrs09}, namely caterpillar path assignment problem.
+
+% We present a logspace algorithm here that uses the ICPIA
+% characterization of binary matrices with COP (set system associated
+% with such a matrix)\cite{nsnrs09}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC  Assignment\tnote{Merge with the original charac section}} 
+% \label{feasible} 
+% Consider a path assignment $\cA =
+% \{(S_i, P_i) \mid S_i \in \cF, P_i \text{ is a path from $T$}, i \in
+% [m]\}$ to a set system $\cF = \{S_i \mid S_i \subseteq U, i \in
+% [m]\}$, were $T$ is a given tree, $U$ is the set system's universe and
+% $m$ is the number of sets in $\cF$. We call $\cA$ an {\em Intersection
+%   Cardinality Preserving Path Assignment (ICPPA)} if it has the
+% following properties.
+
+% \begin{enumerate}
+% \item [i.]  $|S_i| = |P_i|$ for all $i \in [m]$
+% \item [ii.] $|S_i \cap S_j| = |P_i \cap P_j|$ for all $i,j \in [m]$
+%   % \item [iii.] $|\bigcup_{i \in J} S_i| = |\bigcup_{i \in J} P_i|$
+%   %   for all $J \subseteq I$\footnote{TBD: this is needed now to fix
+%   %     some proofs}
+% \item [iii.] $|S_i \cap S_j \cap S_k| = |P_i \cap P_j \cap P_k|$ for
+%   all $i,j,k \in [m]$
+% \end{enumerate}
+
+% \begin{lemma}
+%   \label{lem:setminuscard}
+%   If $\cA$ is an ICPPA, and $(S_1, P_1),(S_2, P_2),(S_3, P_3) \in
+%   \cA$, then $|S_1 \cap (S_2 \setminus S_3)| = |P_1 \cap (P_2
+%   \setminus P_3)|$.
+% \end{lemma}
+% \begin{proof}
+%   $|S_1 \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| =
+%   |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to conditions (ii)
+%   and (iii) of ICPPA, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
+%   \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
+%   = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. \qed
+% \end{proof}
+
+% \begin{lemma}
+%   \label{lem:fourpaths} Consider four paths in a tree $Q_1, Q_2, Q_3,
+%   Q_4$ such that they have nonempty pairwise intersection and $Q_1,
+%   Q_2$ share a leaf. Then there exists distinct $i, j, k \in
+%   \{1,2,3,4\}$ such that, $Q_1 \cap Q_2 \cap Q_3 \cap Q_4 = Q_i \cap
+%   Q_j \cap Q_k$.
+% \end{lemma}
+% \begin{proof}
+%   {\em Case 1:} w.l.o.g, consider $Q_3 \cap Q_4$ and let us call it
+%   $Q$. This is clearly a path (intersection of two paths is a path).
+%   % Since $Q_1, Q_2$ share a leaf, the following are paths $Q_1
+%   % \setminus Q_2$, $Q_2 \setminus Q_1$, $Q_1 \cap Q_2$ and they are
+%   % mutually disjoint.
+%   Suppose $Q$ does not intersect with $Q_1 \setminus Q_2$, i.e. $Q
+%   \cap (Q_1 \setminus Q_2) = \O$. Then $Q \cap Q_1 \cap Q_2 = Q \cap
+%   Q_2$. Similarly, if $Q \cap (Q_2 \setminus Q_1) = \O$, $Q \cap Q_1
+%   \cap Q_2 = Q \cap Q_1$. Thus it is clear that if the intersection of
+%   any two paths does not intersect with any of the set differences of
+%   the remaining two paths, the claim in the lemma is true.
+%   % Note that $Q_1 \setminus Q_2$ and $Q_2
+%   % \setminus Q_1$ are paths because $Q_1, Q_2$ share a leaf.\\
+%   {\em Case 2:} Let us consider the compliment of the previous
+%   case. i.e. the intersection of any two paths intersects with both
+%   the set differences of the other two. First let us consider $Q \cap
+%   (Q_1 \setminus Q_2) \ne \O$ and $Q \cap (Q_1 \setminus Q_2) \ne \O$,
+%   where $Q = Q_3 \cap Q_4$. Since $Q_1$ and $Q_2$ share a leaf, there
+%   is exactly one vertex at which they branch off from the path $Q_1
+%   \cap Q_2$ into two paths $Q_1 \setminus Q_2$ and $Q_2 \setminus
+%   Q_1$. Let this vertex be $v$. It is clear that if path $Q_3 \cap
+%   Q_4$, must intersect with paths $Q_1 \setminus Q_2$ and $Q_2
+%   \setminus Q_1$, it must contain $v$ since these are paths from a
+%   tree. Moreover, $Q_3 \cap Q_4$ intersects with $Q_1 \cap Q_2$ at
+%   exactly $v$ and only at $v$ which means that $Q_1 \cap Q_2$ does not
+%   intersect with $Q_3 \setminus Q_4$ or $Q_4 \setminus Q_3$ which
+%   contradicts initial condition of this case. Thus this
+%   case cannot occur and case 1 is the only possible scenario. \\
+%   Thus lemma is proven \qed
+% \end{proof}
+
+% \noindent
+% In the remaining part of this section we show that a path assignment
+% is feasible if and only if it is an ICPPA.  One direction of this
+% claim is clear: that is a path assignment is feasible, then all
+% intersection cardinalities are preserved, that is the path assignment
+% is an ICPPA.  The reason is that a feasible path assignment has an
+% associated bijection between $U$ and $V(T)$ such that the sets map to
+% paths.  The rest of the section is devoted to constructively proving
+% that it is sufficient, for the existence of a bijection, for a path
+% assignment to be an ICPPA.  At a top-level, the constructive
+% approaches refine the path assignment iteratively, such that at the
+% end of each iteration we have a path assignment, and finally we have a
+% family of bijections.  First we present and then prove the correctness
+% of Algorithm \ref{perms}.  This algorithm refines the path assignment
+% by considering pairs of paths that share a leaf.
+
+% \begin{algorithm}[h]
+%   \caption{Permutations from an ICPPA $\{(S_i,P_i) | i \in I\}$}
+%   \label{perms}
+%   \begin{algorithmic}
+%     \STATE Let $\Pi_0=\{(S_i,P_i)| i \in I\}$\\
+%     \STATE $j = 1$;
+%     \label{shareleaf} \WHILE {There is $(P_1,Q_1), (P_2,Q_2) \in
+%       \Pi_{j-1}$ with $Q_1$ and $Q_2$ having a common leaf} \STATE
+%     $\Pi_j= \Pi_{j-1} \setminus \{(P_1,Q_1),(P_2,Q_2)\}$;
+%     \label{setbreak}\STATE $\Pi_j = \Pi_j \cup \{(P_1 \cap P_2,Q_1
+%     \cap Q_2), (P_1 \setminus P_2,Q_1 \setminus Q_2), (P_2 \setminus
+%     P_1, Q_2 \setminus Q_1)\}$; \STATE $j = j+1$; \ENDWHILE \STATE
+%     $\Pi = \Pi_j$; \STATE Return $\Pi$;
+%   \end{algorithmic}
+% \end{algorithm}
+
+% \begin{lemma}
+%   \label{lem:invar1}
+%   In Algorithm \ref{perms}, at the end of $j$th iteration, $j \ge 0$,
+%   of the while loop of Algorithm \ref{perms}, the following invariants
+%   are maintained.
+%   \begin{itemize}
+%   \item {\em Invariant I:} $Q$ is a path in $T$ for each $(P,Q) \in
+%     \Pi_j$
+%   \item {\em Invariant II:} $|P|=|Q|$ for each $(P,Q) \in \Pi_j$
+%   \item {\em Invariant III:} For any two $(P,Q), (P',Q') \in \Pi_j$,
+%     $|P' \cap P''|=|Q' \cap Q''|$.
+%   \item {\em Invariant IV:} For any three, $(P',Q'), (P'',Q''), (P, Q)
+%     \in \Pi_j$, $|P' \cap P'' \cap P|=|Q' \cap Q'' \cap Q|$.
+%   \end{itemize}
+% \end{lemma}
+% \begin{proof}
+%   Proof is by induction on the number of iterations, $j$. In the rest
+%   of the proof, the term ``new sets'' will refer to the new sets added
+%   in $j$th iteration as defined in line \ref{setbreak} of Algorithm
+%   \ref{perms}, i.e. the following three assignment pairs for some
+%   $(P_1,Q_1), (P_2,Q_2) \in \Pi_{j-1}$ where $Q_1$ and $Q_2$ intersect
+%   and share a leaf: $(P_1 \cap P_2, Q_1 \cap Q_2)$, or $(P_1 \setminus
+%   P_2, Q_1 \setminus Q_2)$, or $(P_2 \setminus P_1, Q_2
+%   \setminus Q_1)$.\\
+%   \noindent
+%   The base case, $\Pi_0 = \{(S_i,P_i) \mid i \in [m]\}$, is trivially
+%   true since it is the input which is an ICPPA.  Assume the lemma is
+%   true till the $j-1$ iteration. Consider $j$th
+%   iteration:\\
+%   \noindent
+%   If $(P,Q)$, $(P',Q')$ and $(P'',Q'')$ are in $\Pi_{j}$ and
+%   $\Pi_{j-1}$, all the invariants are
+%   clearly true because they are from $j-1$ iteration.\\
+%   If $(P,Q)$ is in $\Pi_{j}$ and not in $\Pi_{j-1}$, then it must be
+%   one of the new sets added in $\Pi_j$. Since $(P_1,Q_1)$ and
+%   $(P_2,Q_2)$ are from $\Pi_{j-1}$ and $Q_1,Q_2$ intersect and have a
+%   common leaf, it can be verified that the
+%   new sets are also paths. \\
+%   By hypothesis for invariant III, invariant II also holds for $(P,Q)$
+%   no matter which new set in $\Pi_j$ it
+%   is.\\
+%   To prove invariant III, if $(P,Q)$ and $(P',Q')$ are not in
+%   $\Pi_{j-1}$, then they are both new sets and invariant III holds
+%   trivially (new sets are disjoint). Next consider $(P,Q), (P',Q') \in
+%   \Pi_j$ with only one of them, say $(P',Q')$, in $\Pi_{j-1}$. Then
+%   $(P,Q)$ is one of the new sets added in line \ref{setbreak}. It is
+%   easy to see that if $(P,Q)$ is $(P_1 \cap P_2, Q_1 \cap Q_2)$, then
+%   due to invariant IV in hypothesis, invariant III becomes true in
+%   this iteration. Similarly, using lemma \ref{lem:setminuscard}
+%   invariant III is proven if $(P, Q)$ is $(P_1 \setminus P_2, Q_1
+%   \setminus Q_2)$, or $(P_2 \setminus P_1, Q_2
+%   \setminus Q_1)$.\\
+%   To prove invariant IV, consider three assignments
+%   $(P,Q),(P',Q'),(P'',Q'')$. If at least two of these pairs are in not
+%   $\Pi_{j-1}$, then they are any two of the new sets. Note that these
+%   new sets are disjoint and hence if $(P',Q'), (P'',Q'')$ are any of
+%   these sets, $|P \cap P' \cap P''|=|Q \cap Q' \cap Q''|=0$ and
+%   invariant IV is true. Now we consider the case if at most one of
+%   $(P,Q),(P',Q'),(P'',Q'')$ is not in $\Pi_{j-1}$. If none of them are
+%   not in $\Pi_{j-1}$ (i.e. all of them are in $\Pi_{j-1}$), invariant
+%   IV is clearly true. Consider the case where exactly one of them is
+%   not in $\Pi_{j-1}$. w.l.o.g let that be $(P,Q)$ and it could be any
+%   one of the new sets. If $(P,Q)$ is $(P_1 \cap P_2, Q_1 \cap Q_2)$,
+%   from lemma \ref{lem:fourpaths} and invariant III hypothesis,
+%   invariant IV is proven. Similarly if $(P,Q)$ is any of the other new
+%   sets, invariant IV is proven by also using lemma
+%   \ref{lem:setminuscard}. \qed
+
+% \end{proof}
+
+% \noindent
+% It can be observed that the output of algorithm \ref{perms} is such
+% that every leaf is incident on at most a single path in the new set of
+% assignments. This is due to the loop condition at line
+% \ref{shareleaf}. Let $v_1$ be the leaf incident on path $P_i$. Assign
+% to it any one element from $S_i \setminus \bigcup_{i \ne j}
+% S_j$. Remove $(S_i, P_i)$ from assignments and add $(\{x_1\},
+% \{v_1\}), (S_i \setminus \{x_1\}, P_i \setminus \{v_1\})$. Now all
+% assignments except single leaf assignments are paths from the subtree
+% $T_0 = T \setminus \{v \mid v \text{ is a leaf in } T\}$.
+
+% \begin{algorithm}[h]
+%   \caption{Leaf assignments from an ICPPA $\{(S_i,P_i) | i \in I\}$}
+%   \label{leafasgn}
+%   \begin{algorithmic}
+%     \STATE Let $\Pi_0=\{(S_i,P_i)| i \in [m]\}$. Paths are such that
+%     no
+%     two paths $P_i, P_j, i \ne j$ share a leaf.\\
+%     \STATE $j = 1$\\
+%     \WHILE {there is a leaf $v$ and a unique $(S_{i_1}, P_{i_1})$ such
+%       that $v \in P_{i_1}$}
+%     \STATE $\Pi_j=   \Pi_{j-1} \setminus \{(S_{i_1}, P_{i_1})\}$\\
+%     \STATE $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in I}S_i$
+%     \IF{$X$ is empty} \STATE exit \ENDIF; \STATE Let $x = $ arbitrary
+%     element from $X$ \vspace{2mm}
+
+%     \STATE $\Pi_j = \Pi_j \cup \{(S_{i_1} \setminus \{x\}),(P_{i_1}
+%     \setminus
+%     \{v\}), (\{x\},\{v\})\}$\\
+%     \STATE $j = j+1$\\
+%     \ENDWHILE
+%     \STATE $\Pi = \Pi_j$\\
+%     \STATE Return $\Pi$\\
+%   \end{algorithmic}
+% \end{algorithm}
+
+% \begin{lemma}
+%   \label{lem:invar3}
+%   In Algorithm \ref{leafasgn},for all $j \geq 0$, at the end of the
+%   $j$th iteration the four invariants given in lemma \ref{lem:invar1}
+%   are valid.
+%   % Moreover, $X$ as defined in the algorithm is non-empty if this is
+%   % an ICPPA.
+% \end{lemma}
+% \begin{proof}
+%   First we see that $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in
+%     I}S_i$ is non empty in every iteration for an ICPPA. Suppose $X$
+%   is empty. We know that $v \in P_{i_1} \setminus \bigcup_{i \ne i_1,
+%     i \in I}P_i$ since $v$ is in the unique path $P_{i_1}$. Since this
+%   is an ICPPA $|S_{i_1}| = |P_{i_1}|$. For any $x \in S_{i_1}$ it is
+%   contained in at least two sets due to our assumption. Let $S_{i_2}$
+%   be a second set that contains $x$. We know $v \notin
+%   P_{i_2}$. Therefore there cannot exist a permutation that maps
+%   elements of $S_{i_2}$ to $P_{i_2}$. This contradicts our assumption
+%   that this is an ICPPA. Thus $X$ cannot be empty.
+
+%   We use mathematical induction on the number of iterations for this
+%   proof. The term ``new sets'' will refer to the sets added in $\Pi_j$
+%   in the $j$th iteration, i.e. $(P' \setminus \{x\},Q' \setminus
+%   \{v\})$ and $(\{x\},\{v\})$ for some $(P',Q')$ in $\Pi_{j-1}$ such
+%   that $v$ is a leaf and $Q'$ is the unique path
+%   incident on it.\\
+%   For $\Pi_0$ all invariants hold because it is output from algorithm
+%   \ref{perms} which is an ICPPA. Hence base case is proved.  Assume
+%   the lemma holds for $\Pi_{j-1}$. Consider $\Pi_j$ and any $(P,Q) \in
+%   \Pi_j$. If $(P,Q) $ is in $ \Pi_j$ and $\Pi_{j-1}$ invariants I and
+%   II are true because of induction assumption. If it is only in
+%   $\Pi_j$, then it is $\{(P' \setminus \{x\}),(Q' \setminus \{v\})$ or
+%   $(\{x\},\{v\})$ for some $(P',Q')$ in $\Pi_{j-1}$. By definition,
+%   $x$ is an element in $P'$ (as defined in the algorithm) and $v$ is a
+%   leaf in $Q'$. If $(P,Q)$ is $\{(P' \setminus \{x\}),(Q' \setminus
+%   \{v\})$, $Q$ is a path since only a leaf is removed from path
+%   $Q'$. We know $|P'| = |Q'|$, therefore $|P' \setminus \{x\}| = |Q'
+%   \setminus \{v\}|$. Hence in this case invariants I and II are
+%   obvious. It is easy to see these invariants hold if $(P,Q)$ is
+%   $(\{x\},\{v\})$.
+
+%   For invariant III consider $(P_1,Q_1),(P_2,Q_2)$ in $\Pi_j$. If both
+%   of them are also in $\Pi_{j-1}$, claim is proved. If one of them is
+%   not in $\Pi_{j-1}$ then it has to be $\{(P' \setminus \{x\}),(Q'
+%   \setminus \{v\})$ or $(\{x\},\{v\})$ for some $(P',Q')$ in
+%   $\Pi_{j-1}$. Since by definition, $Q'$ is the only path with $v$ and
+%   $P'$ the only set with $x$ in the previous iteration, $|P_1 \cap (P'
+%   \setminus \{x\})| = |P_1 \cap P'|$ and $|Q_1 \cap (Q' \setminus
+%   \{v\})| = |Q_1 \cap Q'|$ and $|P_1 \cap \{x\}| = 0, Q_1 \cap \{v\} =
+%   0$. Thus invariant III is also proven.
+
+%   \noindent
+%   To prove invariant IV, consider $(P_1,Q_1),(P_2,Q_2), (P_3,Q_3)$ in
+%   $\Pi_j$. If exactly one of them, say $P_3 \notin \Pi_{j-1}$, it is
+%   one of the new sets. By the same argument used to prove invariant
+%   III, $|P_1 \cap P_2 \cap (P' \setminus \{x\})| = |P_1 \cap P_2 \cap
+%   P'|$ and $|Q_1 \cap Q_2 \cap (Q' \setminus \{x\})| = |Q_1 \cap Q_2
+%   \cap Q'|$. Since $P_1, P_2, P'$ are all in $\Pi_{j-1}$, by induction
+%   hypothesis $|P_1 \cap P_2 \cap P'| = |Q_1 \cap Q_2 \cap Q'|$. Also
+%   $|P_1 \cap P_2 \cap \{x\}| = 0, Q_1 \cap Q_2 \cap \{v\} = 0$.  If
+%   two or more of them are not in $\Pi_{j-1}$, then it can be verified
+%   that $|P_1 \cap P_2 \cap P_3| = |Q_1 \cap Q_2 \cap Q_3|$ since the
+%   new sets in $\Pi_j$ are either disjoint or as follows: assuming
+%   $P_1, P_2 \notin \Pi_{j-1}$ and new sets are derived from $(P', Q'),
+%   (P'', Q'') \in \Pi_{j-1}$ with $x_1, x_2$ exclusively in $P_1, P_2$,
+%   $(\{x_1\},\{v_1\}), (\{x_2\},\{v_2\}) \in \Pi_j $ thus $v_1, v_2$
+%   are exclusively in $Q_1, Q_2$ resp. it follows that $|P_1 \cap P_2
+%   \cap P_3| = |(P' \setminus \{x_1\}) \cap (P'' \setminus \{x_2\})
+%   \cap P_3| = |P' \cap P'' \cap P_3| = |Q' \cap Q'' \cap Q_3| = |(Q'
+%   \setminus \{v_1\} \cap Q'' \setminus \{v_2\} \cap Q_3| = |Q_1 \cap
+%   Q_2 \cap Q_3|$. Thus invariant IV is also proven.  \qed
+% \end{proof}
+
+% Using algorithms \ref{perms} and \ref{leafasgn} we prove the following
+% theorem.
+
+% \begin{theorem}
+%   \label{th:perm}
+%   If $\cA$ is an ICPPA, then there exists a bijection $\sigma : U
+%   \rightarrow V(T)$ such that $\sigma(S_i) = P_i$ for all $i \in I$
+% \end{theorem}
+% \begin{proof}
+%   This is a contructive proof. First, the given ICPPA $\cA$ and tree
+%   $T$ are given as input to Algorithm \ref{perms}. This yields a
+%   ``filtered'' ICPPA as the output which is input to Algorithm
+%   \ref{leafasgn}.  It can be observed that the output of Algorithm
+%   \ref{leafasgn} is a set of interval assignments to sets and
+%   one-to-one assignment of elements of $U$ to each leaf of $T$. To be
+%   precise, it would be of the form $\cB_0 = \cA_0 \cup \cL_0$. The
+%   leaf assignments are defined in $\cL_0 = \{ (x_i,v_i) \mid x_i \in
+%   U, v_i \in T, x_i \ne x_j, v_i \ne v_j, i \ne j, i,j \in [k] \}$
+%   where $k$ is the number of leaves in $T$. The path assignments are
+%   defined in $\cA_0 \subseteq \{(S_i',P_i') \mid S_i' \subseteq U_0,
+%   P_i' \text{ is a path from } T_0\}$ where $T_0$ is the tree obtained
+%   by removing all the leaves in $T$ and $U_0 = U \setminus \{ x \mid x
+%   \text{ is assigned to a leaf in }\cL_0 \}$. Now we have a subproblem
+%   of finding the permutation for the path assignment $\cA_0$ which has
+%   paths from tree $T_0$ and sets from universe $U_0$. Now we repeat
+%   the procedure and the path assignment $\cA_0$ and tree $T_0$ is
+%   given as input to Algorithm \ref{perms}. The output of this
+%   algorithm is given to Algorithm \ref{leafasgn} to get a new union of
+%   path and leaf assignments $\cB_1 = \cA_1 \cup \cL_1$ defined similar
+%   to $\cB_0, \cL_0, \cA_0$. In general, the two algorithms are run on
+%   path assignment $\cA_{i-1}$ with paths from tree $T_{i-1}$ to get a
+%   new subproblem with path assignment $\cA_i$ and tree $T_{i}$. $T_i$
+%   is the subtree of $T_{i-1}$ obtained by removing all its
+%   leaves. More importantly, it gives leaf assignments $\cL_{i}$ to the
+%   leaves in tree $T_{i-1}$. This is continued until we get a
+%   subproblem with path assignment $\cA_{d-1}$ and tree $T_{d-1}$ for
+%   some $d \le n$ which is just a path. From the last lemma we know
+%   that $\cA_{d-1}$ is an ICPPA. Another observation is that an ICPPA
+%   with all its tree paths being intervals (subpaths from a path) is
+%   nothing but an ICPIA\cite{nsnrs09}.  Let $\cA_{d-1}$ be equal to
+%   $\{(S_i'',P_i'') \mid S_i'' \subseteq U_{d-1}, P_i'' \text{ is a
+%     path from } T_{d-1} \}$. It is true that the paths $P_i''$s may
+%   not be precisely an interval in the sense of consecutive integers
+%   because they are some nodes from a tree. However, it is easy to see
+%   that the nodes of $T_{d-1}$ can be ordered from left to right and
+%   ranked to get intervals $I_i$ for every path $P_i''$ as
+%   follows. $I_i = \{[l,r] \mid l = \text{ the lowest rank of the nodes
+%     in }P_i'', r = l+|P_i''|-1 \}$. Let asssignment $\cA_d$ be with
+%   the renamed paths. $\cA_d = \{ (S_i'', I_i) \mid (S_i'', P_i'') \in
+%   \cA_{d-1} \}$. What has been effectively done is renaming the nodes
+%   in $T_{d-1}$ to get a tree $T_d$.  The ICPIA $\cA_d$ is now in the
+%   format that the ICPIA algorithm requires which gives us the
+%   permutation $\sigma' : U_{d-1} \rightarrow T_{d-1}$
+
+% \noindent
+% $\sigma'$ along with all the leaf assignments $\cL_i$ gives us the
+% permutation for the original path assignment $\cA$.  More precisely,
+% the permutation for tree path assignment $\cA$ is defined as
+% follows. $\sigma: U \rightarrow T$ such that the following is
+% maintained.
+% \begin{align*}
+%   \sigma(x) &= \sigma'(x),   \text{ if } x \in U_{d-1} \\
+%   &= \cL_i(x), \text{ where $x$ is assigned to a leaf in a subproblem
+%     $\cA_{i-1}, T_{i-1}$}
+% \end{align*}
+
+% \noindent
+% To summarize, run algorithm \ref{perms} and \ref{leafasgn} on
+% $T$. After the leaves have been assigned to specific elements from
+% $U$, remove all leaves from $T$ to get new tree $T_0$. The leaf
+% assignments are in $\cL_0$. Since only leaves were removed $T_0$ is
+% indeed a tree. Repeat the algorithms on $T_0$ to get leaf assignments
+% $\cL_{1}$. Remove the leaves in $T_0$ to get $T_1$ and so on until the
+% pruned tree $T_d$ is a single path. Now run ICPIA algorithm on $T_d$
+% to get permutation $\sigma'$. The relation $\cL_0 \cup \cL_1 \cup
+% .. \cup \cL_{d} \cup \sigma'$ gives the bijection required in the
+% original problem.\qed
+% \end{proof}
+
 \chapter{Conclusion}

### 2012-01-03 src/001certificate.tex
Cleaned up thesis.tex. Split thesis.tex into chapterwise subfiles for easy edit tex management. Content changes are mainly in chapter 1 and 3 although nothing significant. To being working on chapter 2 after this commit.

--- 2011-12-30 src/thesis.tex
@@
-\documentclass[MS]             % MS thesis dissertation.
-              {iitmdiss_as}    % Hacked by hyperbolicme.
-%              {iitmdiss}      % If the powers to be are finicky.
-
-\usepackage{times}             % Font?
-\usepackage{t1enc}             % T1 encoding??  
-%\usepackage{boxedminipage}
-\usepackage{../lib/mythesislib}
-\usepackage{../lib/mylatexlib}
-% \usepackage{comment}           % DEBUG ONLY
-
-\def \mythesissubmissiondate {10 January 2012}
-\def \mythesissubmissionmonth {\MakeUppercase {January 2012}}
-
-\begin{document}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Title page
-\title{\mythesistitleNOTSYNPOSIS} 
-\author{\myname} 
-\date{
-  {\mythesissubmissionmonth}} 
-\department{\MakeUppercase{\mythesisdept}}
-%\department{\mythesisdept}
-
-%\nocite{*}
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Certificate
 \certificate
@@
 \noindent 
-This is to certify that the thesis titled {\bf \mythesistitleNOTSYNPOSIS},
-submitted by {\bf \myname}, to the {\bf Indian Institute of Technology
-Madras}, for the award of the degree of {\bf \mydegree}, is 
-a bona fide record of the research work done by her under our
-supervision.  The contents of this thesis, in full or in parts, have
-not been submitted to any other Institute or University for the award
-of any degree or diploma.
+This is to certify that the thesis titled {\bf
+  \mythesistitleNOTSYNPOSIS}, submitted by {\bf \myname}, to the {\bf
+  Indian Institute of Technology Madras}, for the award of the degree
+of {\bf \mydegree}, is a bona fide record of the research work done by
+her under our supervision.  The contents of this thesis, in full or in
+parts, have not been submitted to any other Institute or University
+for the award of any degree or diploma.
 
@@
     \noindent Dept. of \mydept\\
-    \noindent IIT Madras -- 600 036
-  } \hfill
+    \noindent IIT Madras -- 600 036 } \hfill
   \parbox{1.25in}{               % {1.75in}
@@
 \end{singlespacing}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Acknowledgements
-\acknowledgements
-
-\tnote{WIP}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Abstract
-
-\begin{abstract}
-  \noindent {\em {\bf Keywords}: \hspace*{0.5em} \parbox[t]{4.4in}{consecutive
-    ones property, algorithmic graph theory, hypergraph isomorphism,
-    interval labeling}} \vspace*{24pt}
-
-  \def \tem {}
-  \noindent 
-  Consecutive-ones property is a non-trivial property of binary
-  matrices that has been studied widely in the literature for over
-  past 50 years. Detection of COP in a matrix is possible efficiently
-  and there are several algorithms that achieve the same. This thesis
-  documents the work done on an extension of COP extended from the
-  equivalent interval assignment problem in \cite{nsnrs09}. These new
-  results rigorously prove a natural extension (to trees) of their
-  characterization as well as makes connections to graph isomorphism,
-  namely path graph isomorphism.
-
-  \tnote{EXPAND. 
-  Abstract must be a brief about what results we have and how it fits
-  in the body of research.\\
-  -- Area: Broad to Specialized. i.e. Combinatorial
-    algorithms -> Matrix reorganization -> general data reorganization
-    (interval assignment) -> path assignment\\
-  -- Class of problems: say, data reorganization.\\
-  -- Nature of results: Is a generalization. We have a
-    Polynomial algorithm for a subset of the generalization.\\
-% maybe add - we explore a natural
-%     generalization of results on binary matrices with the |\tem
-%     consecutive-ones property|.  We consider the following constraint
-%     satisfaction problem. Given (i) a set system $\F \subseteq$
-%     $(2^{U} \setminus \emptyset)$ of a finite set $U$ of cardinality
-%     $n$, (ii) a tree $T$ of size $n$ and (iii) a bijection called
-%     |\tem tree path labeling|, $\cl$ mapping the sets in $\cF$ to
-%     paths in $T$, does there exist at least one bijection $\phi:U
-%     \rightarrow V(T)$ such that for each $S \in \cF$, $\{\phi(x) \mid
-%     x \in S\} = \cl(S)$?  A tree path labeling of a set system is
-%     called |\tem feasible| if there exists such a bijection $\phi$.
-%     We present an algorithmic characterization of feasible tree path
-%     labeling. COP is a special instance of tree path labeling problem
-%     when $T$ is a path.  We conclude with a polynomial time algorithm
-%     to find a feasible tree path labeling of a given set system when
-%     $T$ is a |\tem $k$-subdivided star|, set system has a single
-%     containment tree of overlap components and set size is limited to
-%     at most $k+2$.
-  }
-\end{abstract}
-
-%\pagebreak
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Table of contents etc.
-
-\begin{singlespace}
-\tableofcontents
-\thispagestyle{empty}
-
-\listoftables
-\addcontentsline{toc}{chapter}{LIST OF TABLES}
-\listoffigures
-\addcontentsline{toc}{chapter}{LIST OF FIGURES}
-\end{singlespace}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Abbreviations
-\abbreviations
-
-\noindent 
-\begin{tabbing}
-xxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
-\textbf{COP}   \> Consecutive-ones Property \\
-\textbf{COT}   \> Consecutive-ones Testing \\
-\textbf{ICPIA}   \> Intersection Cardinality Preservation Interval Assignment \\
-\textbf{ICPPL}   \> Intersection Cardinality Preserved Path Labeling \\
-% ADD MORE HERE
-\end{tabbing}
-
-\pagebreak
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Notation
-
-\chapter*{\centerline{NOTATION}}
-\addcontentsline{toc}{chapter}{NOTATION}
-
-\begin{singlespace}
-\begin{tabbing}
-xxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
-\textbf{$2^{U}$}  \> Powerset of set $U$ \\
-\end{tabbing}
-\end{singlespace}
-
-\pagebreak
-\clearpage
-
-% The main text will follow from this point so set the page numbering
-% to arabic from here on.
-\pagenumbering{arabic}
-
-\chapter{Introduction}
-\label{ch:intro}
-
-Consecutive-ones property is a non-trivial property of binary matrices
-that has been studied widely in the literature for over past 50
-years. Detection of COP in a matrix is possible efficiently and there
-are several algorithms that achieve the same. This thesis documents
-the work done on an extension of COP extended from the equivalent
-interval assignment problem in \cite{nsnrs09}. These new results
-rigorously prove a natural extension (to trees) of their
-characterization as well as makes connections to graph isomorphism,
-namely path graph isomorphism.
-
-\tnote{Have a para on Organization: Outline of document}
-Section~\ref{sec:background} gives a brief survey of COP and
-optimization problems related to it followed by motivation for the
-thesis in Section~\ref{sec:motive}.  Section~\ref{sec:results}
-presents a summary of our results on the extension of COP namely, the
-tree path labeling problem. \tnote[TH]{update for thesis}
-
-% \tnote{testing 1 2 3...}
-% \tnote[bogus1]{testing 1 2 3...}
-% \tnote[bogus2]{testing 1 2 3...}
-% \tnote[GTC]{testing 1 2 3...}
-% \tnote[bogus4]{testing 1 2 3...}
-% \tnote[TH]{testing 1 2 3...}
-
-\section{Illustration of the problem}
-\label{sec:problem}
-
-A group of students, \Pa, \Pig, \Sn, \Wo, \Vi, \Li, \Ch, \Sa, \Fr,
-\Sc\ and \Lu\ enroll at the {\WSI} for a liberal arts programme.  As
-part of their semester thesis, they pick a body of work to study and
-form the namesake study groups, {\LLL}, {\GGG}, {\BBB} and
-{\TTT}\tnote[TH]{put bib entries for these works!}. A student will be
-in at least one study group and may be in more than one. For instance,
-as will be seen later, {\Fr} studies both {\LLL} and {\TTT} while \Wo\
-studies only \BBB.
-
-Let $U$ and $\cF$ represent the set of students and the set of study
-groups respectively and the integers $n$ and $m$ denote the total
-number students and study groups respectively. In relation to this
-example, these are defined in Table~\ref{tab:wsigroups}. Also given
-there is the study group allocation to students.
-
-\begin{table}[t]%[htbp]
-  \centering
-  { \footnotesize
-  \begin{tabular}{rcl}
-    $U $&$=$&$ \{\xPa,\; \xPi,\; \xSn,\; \xWo,\; \xVi,\; \xLi,\; \xCh,\;
-    \xSa,\; \xFr,\; \xSc,\; \xLu\}$\\
-    $\cF $&$=$&$ \{\xLLL, \xGGG, \xBBB, \xTTT\}$\\
-    $\xLLL $&$=$&$ \{\xCh,\;  \xSa,\;  \xFr,\;  \xSc,\;  \xLu \}$\\
-    $\xGGG $&$=$&$ \{\xPa,\;  \xPi,\;  \xVi,\;  \xCh \}$\\
-    $\xBBB $&$=$&$ \{\xSn,\;  \xPi,\;  \xWo \}$\\
-    $\xTTT $&$=$&$ \{\xVi,\;  \xLi,\;  \xCh,\;  \xFr \}$\\
-    $n $&$=$&$ |U| = 11$\\
-    $m $&$=$&$ |\cF| = 4$      
-  \end{tabular}
-}
-  \caption{\figtabsize Students and study groups in \WSI}
-  \label{tab:wsigroups}
-\end{table}
-
-The campus has a residential area {\residenceblock} that has $n$
-single occupancy apartments reserved for the study groups'
-accommodation.  All these apartments are located such that the streets
-connecting them do {\em not} form
-loops. Fig~\ref{fig:streetmappathpeople} \tnote[TH]{fig:streetmap}
-shows the street map for {\residenceblock}. It may be noted that as a
-graph, it classifies as a tree.
-
-\begin{table}[htbp]
-  \centering
-  {\footnotesize    
-    \begin{tabular}{l||l}
-      \begin{tabular}{rcl}
-        $T $&$=$&$ \text{\em Street map tree of \residenceblock}$\\
-        $V(T) $&$=$&$ \{ 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11 \}$\\
-        $\cP $&$=$&$ \{R\xLLL, R\xGGG, R\xBBB, R\xTTT\}$\\
-        $R\xLLL $&$=$&$ \{9, 1, 5, 3, 11\}$\\
-        $R\xGGG $&$=$&$ \{7, 2, 6, 5\}$\\
-        $R\xBBB $&$=$&$ \{8, 2, 4 \}$\\
-        $R\xTTT $&$=$&$ \{10, 6, 5, 3\}$\\
-        $n $&$=$&$ |V| = 11$\\
-        $m $&$=$&$ |\cP| = 4$\\\\
-        $\cl $&$=$& {\em Study group to route mapping}\\        
-        $\cl(\mathbb{X}) $&$=$&  $R\mathbb{X}$ for all $\mathbb{X} \in
-        \cF$
-      \end{tabular} & 
-      \begin{tabular}{c}
-       {\em Apartment allocation }($\phi$)\\
-                    \begin{tabular}{c|c}                      
-                      1  & \xSa\\
-                      2  & \xPi\\
-                      3  & \xFr\\
-                      4  & \xWo\\
-                      5  & \xCh\\
-                      6  & \xVi\\
-                      7  & \xPa\\
-                      8  & \xSn\\
-                      9  & \xLu\\
-                      10 & \xLi\\
-                      11 & \xSc                     
-                    \end{tabular}
-      \end{tabular}\\
-    \end{tabular}
-  }
-  \caption{\figtabsize A solution to study group accommodation problem}
-  \label{tab:iltree}
-\end{table}
-
-A natural question would be to find how the students should be
-allocated apartments such that each study group has the {\em least
-  distance to travel} for a discussion? More specifically, we are
-interested in enforcing additional conditions, namely, that all the
-students in a study group must be next to each other; in other words,
-for one student to reach another fellow study group member's apartment
-(for all study groups the student is part of), she must not have to
-pass the apartment of any student who is not in that study group. To
-further elucidate, the apartments of students of any study group must
-be arranged in an exclusive unfragmented path on the street
-map. Exclusivity here means that the path must not have apartments
-from other study groups (unless that apartment is also part of {\em
-  this} study group).
-
-An intuitive approach to this problem would be to first find the paths
-that each study group decides to inhabit and then refine the
-allocation to individual students. A feasible allocation of exclusive
-routes to study groups is illustrated in
-Fig~\ref{fig:streetmappathpeople} \tnote[TH]{fig:streetmappath} and
-the students' allocation of apartments that obeys this route
-allocation is also shown. Table~\ref{tab:iltree} shows the same
-solution set theoretically.  How this is algorithmically computed is
-the focus of this thesis.
-
-\begin{figure}[htbp] %[t]%
- \centering
-  (a) \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
-  (b) \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
-
- \label{fig:streetmap}
- \caption{\figtabsize (a) {\residenceblock} street map (b)
-   {\residenceblock} street map with study group routes
-   allocated. Routes are color coded as follows: red for
-   \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
-   group, orange for \textcolor{yellow}{$\xBBB$} group, green for
-   \textcolor{green}{$\xTTT$} group}
-
- \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
-  \caption{\figtabsize Individual allocation of apartments to students
-    in {\residenceblock} that meets the requirements stated before.
-    The routes are color coded as follows: red for
-    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
-    group, orange for \textcolor{yellow}{$\xBBB$} group, green
-    for \textcolor{green}{$\xTTT$} group. {\tiny {\em Peanuts images
-        {\copyright} Charles Schulz}}}% YellowOrange
-  \label{fig:streetmappathpeople}
-\end{figure}
-\tnote{UPDATE IMAGE (b) 2-infinite-loop-BTWF.png. REMOVE TITLES. ADD
-  {$\xLLL$} {$\xGGG$} {$\xBBB$} {$\xTTT$}}
-
-\tnote[TH]{make dashed/textured lines for routes. make it color
-  agnostic.}
-
-As a special case, suppose all the apartments are on the same street
-or if they are all lined up on a single path, the street map becomes a
-tree that is just a path. Then the problem becomes what is called an
-{\em interval assignment problem}. The idea of interval assignment may
-not be obvious here; hence to see this, consider a different problem
-in {\WSI} where the classes for these study groups courses need to be
-scheduled during a day (or a week or any time period). Each study
-group has a bunch of courses associated with it some of which may be
-shared by two or more study groups. It is mandatory that a student who
-is a member of a study group takes all the courses associated with
-that group. There are slots during the day for classes to be held and
-the problem is to allocate class slots to courses such that all the
-classes of a study group are consecutive. It is debatable if this will
-not hamper the attention span and memory retention rate of the
-students but that is, regrettably, out of the scope of this
-thesis. The parallels between this class allocation problem and the
-accommodation problem can be seen as follows. The set $U$ here, are
-the courses offered (say Course 101 {\coneohone}, Course 102
-{\coneohtwo} and so on). In this variation of the problem, the
-collection $\cF$ is the set of study groups but the study groups are
-filled by course IDs (in place of students in the earlier
-example). For instance, Course 101 is mandatory for all study groups
-$\xLLL$, $\xGGG$, $\xBBB$, $\xTTT$ and Course 102 is mandatory for
-only the $\xLLL$ group) and so on. The sequence of class slots for the
-day (or week or any time period) is analogous to the street map in the
-accommodation problem. It is quite obvious now why this version of the
-problem (where the ``target graph'' is a path and not any
-tree\tnote[TH]{Allowing any tree in this example could be seen as a
-  scenario where there are parallel classes. A node falling in the
-  path between two other nodes would mean that the corresponding is
-  scheduled between the other two.}) is called an interval assignment
-problem.
-
-The interval assignment problem to a set system is equivalent to the
-consecutive-ones property (COP) problem in binary matrices\cite{wlh02,
-  nsnrs09}.  The COP problem is to rearrange rows (columns) of a
-binary matrix in such a way that every column (row) has its {\un}s
-occur consecutively. If this is possible the matrix is said to have
-the COP.  COP is a well researched combinatorial problem and has
-several positive results on tests for it and computing the COP
-permutation (i.e. the course schedule in the above illustration) which
-will be surveyed later in this document. Hence we are interested in
-extensions of COP, more specifically, the extension of interval
-assignment problem to tree path assignment problem (which is
-illustrated by the study group accommodation problem).
-
-\section{Basic preliminaries - general definitions and nomenclature}
-\tnote{(definitions theorems etc) needed if any
-    - graph theory}
-
-\section{Consecutive-ones Property Testing - a Brief Survey}
-\label{sec:background}
-
-In this section, a brief survey of the consecutive-ones problem and its
-optimization problems is presented.
-
-%%%%%%                              %%%%%%
-%THX is the second level of ``tnote''-ing%
-%%%%%%                              %%%%%%
-\tnote[TH]{ADD: As it will be described in detail later in this
-  document, isomorphism of certain % classes of graphs, namely chordal
-%   graphs, have a close relationship with consecutive-ones property and
-%   generalizations of it.  This is perhaps because of how closely COP
-%   of a matrix relates to properties of graphs derived from matrices as
-%   seen in the following results.
-}
-\tnote[THX]{ADD: peo exists iff chordal. lexicographic BFS
-  [tag:chordalGraph]} %
-\tnote[THX]{ADD: A well known result in %perfect graph theory is that
-%  the maximal cliques of  an interval graph $G$ can be linearly ordered
-%   such that for all $v \in V(G)$, cliques containing $v$ are
-%   consecutive in the ordering cite~{gh64}. This clearly means that a
-%   graph $G$ is an interval graph if and only if
-}% 
-\tnote[THX]{verify
-  from paper the statement of claim.}% 
-\tnote[TH]{maximal clique vertex
-  incidence matrix of%  $G$ has COP.  Also, maximal cliques of any
-%   chordal graph can be enumerated in polytime $O(m+n)$
-}
-\tnote[THX]{citation?!!}%  
-\tnote[TH]{cite~{fg65} uses these results to
-  give the first polynomial time algorithm for COT.}%
-\tnote[THX]{check. how do they use it?}%  
-\tnote[TH]{ A bipartite graph
-  is convex%  (on $R$) if and only if its half adjacency matrix has COP
-%   on rows.  The results in cite~{bl76} on COT are based on the result
-%   that interval graphs are AT-free chordal graphs.
-}%
-\tnote[THX]{the latter being Tucker's?}%  
-\tnote[THX]{(2) TBD survey
-  -- % see blue notes (in notebook) under Graph Isomorphism. namely
-%   citations in cite:aas93 (3) a brief on heirarchy of chordal graphs,
-%   path graphs, interval graphs, peo, clique tree etc. and results. (4)
-%   interval graphs are incomparability graphs. see
-%   Golumbic. [tag:classification] (5) finding min length hole in
-%   bipartite graph is polynomial time [Sec 3.3.2 in cite:d08phd] (6)
-%   theorem 2.2 in cite:d08phd - $G$ is union of v.d. caterpillars iff
-%   $M$ has COP in rows. $M$ is edge vertex incidence matrix $(*,2)$ of
-%   $G$
-}%
-
-\subsection{Matrices with COP}
-\label{sec:copmatrices}
-As seen earlier, the interval assignment problem (illustrated as the
-course scheduling problem in Section~\ref{sec:problem}), is a special
-case of the problem we address in this thesis, namely the tree path
-labeling problem (illustrated as the study group accommodation
-problem). The interval assignment problem and COP problem are equivalent
-problems. In this section we will see some of the results that
-exists in the literature today towards solving the COP problem and
-optimization problems surrounding it.
-
-\begin{figure}[t] %[htbp]
-  \centering
-
-  {\figtabsize
-    \begin{tabular}[h]{l|lcccl}
-      $M_1$: & $M_1'$: &&&& $M_2$:\\
-      &&&&&\\
-      \begin{tabular}[h]{llll}
-        $c_1$ & $c_2$ &$c_3$ &$c_4$\\
-        &&&\\
-        \un & 0   & \un & 0\\
-        0   & \un & 0   & \un \\
-        \un & 0   & 0   & \un
-      \end{tabular}
-      &
-      \begin{tabular}[h]{llll}
-        $c_3$ &$c_1$ &$c_4$& $c_2$\\
-        &&&\\
-        \un & \un & 0 & 0\\
-        0 & 0 & \un & \un \\
-        0 & \un & \un & 0
-      \end{tabular}
-      &&&&
-      \begin{tabular}[h]{llll}
-        $d_1$ & $d_2$ &$d_3$ &$d_4$\\
-        % &&&\\
-        &&&\\
-        \un & \un & 0 & 0\\
-        0 & \un & \un & 0 \\
-        0 & \un & 0 & \un 
-      \end{tabular}
-    \end{tabular}
-  }
-
-  \caption{\figtabsize Matrices with and without COP. $M_1$ has COP
-    because by permuting its columns, $c_1$-$c_4$, one can obtain
-    $M_1'$ where the {\un}s in each row are consecutive. $M_2$,
-    however, does not have COP since no permutation of its columns,
-    $d_1$-$d_4$, will arrange {\un}s in each row consecutively
-    \cite{d08phd}.}
-
-  \label{fig:cop-matrix}
-\end{figure}
-
-Recall that a matrix with COP is one whose rows (columns) can be
-rearranged so that the {\un}s in every column (row) are in consecutive
-rows (columns). Figure~\ref{fig:cop-matrix} shows examples of this
-property.  COP in binary matrices has several practical applications
-in diverse fields including scheduling \cite{hl06}, information
-retrieval \cite{k77} and computational biology \cite{abh98}.  Further,
-it is a tool in graph theory \cite{mcg04} for interval graph
-recognition, characterization of Hamiltonian graphs, planarity testing
-\cite{bl76} and in integer linear programming \cite{ht02,hl06}.
-
-The obvious first questions after being introduced to the consecutive
-ones property of binary matrices are if COP can be detected
-efficiently in a binary matrix and if so, can the COP permutation of
-the matrix also be computed efficiently?  Recognition of COP in a
-binary matrix is polynomial time solvable and the first such algorithm
-was given by \cite{fg65}.  A landmark result came a few years later
-when \cite{at72} discovered the families of forbidden submatrices that
-prevent a matrix from having COP and most, if not all, results that
-came later were based on this discovery which connected COP in binary
-matrices to convex bipartite graphs. In fact, the forbidden
-submatrices came as a corollary to the discovery that convex bipartite
-graphs are AT-free in \cite{at72}\tnote[TH]{check}. The first linear
-time algorithm for COP testing (COT) was invented by \cite{bl76} using
-a data structure called PQ trees.  Since then several COT algorithms
-have been invented -- some of which involved variations of PQ trees
-\cite{mm96,wlh01,mcc04}, some involved set theory and ICPIA
-\cite{wlh02,nsnrs09}, parallel COT algorithms\cite{as95,bs03,ly91} and
-certifying algorithms\cite{mcc04}.
-
-The construction of PQ trees in \cite{bl76} draws on the close
-relationship of matrices with COP to interval graphs. A PQ tree of a
-matrix is one that stores all row (column) permutations of the matrix
-that give the COP orders (there could be multiple orders of rows or
-columns) of the matrix. This is constructed using an elaborate linear
-time procedure and is also a test for planarity\tnote[TH]{check check
-  check. both interval graph and planarity in this paper?}.  PQR trees
-is a generalized data structure based on PQ trees \cite{mm96,mpt98}.
-\cite{tm05} describes an improved algorithm to build PQR
-trees. \tnote[TH]{improv in terms of what?}\cite{wlh02} describes the
-simpler algorithm for COT. Hsu also invented PC trees
-\cite{wlh01}\tnote[TH]{This result first appeared inproc ISAAC92}
-which is claimed to be much easier to implement. \cite{nsnrs09}
-describes a characterization of consecutive-ones property solely based
-on the cardinality properties of the set representations of the
-columns (rows); every column (row) is equivalent to a set that has the
-row (column) indices of the rows (columns) that have one entries in
-this column (row). This is interesting and relevant, especially to
-this thesis because it simplifies COT to a great degree. \tnote[TH]{it
-  reduces the solution search space. fill in the blanks.}
-
-\cite{mcc04} describes a different approach to COT. While all previous
-COT algorithms gave the COP order if the matrix has the property but
-exited stating negative if otherwise, this algorithm gives an evidence
-by way of a certificate of matrix even when it has no COP. This
-enables a user to verify the algorithm's result even when the answer
-is negative. This is significant from an implementation perspective
-because automated program verification is hard and manual verification
-is more viable. Hence having a certificate reinforces an
-implementation's credibility. Note that when the matrix {\em has} COP,
-the COP order is the certificate.  The internal machinery of this
-algorithm is related to the weighted betweenness problem
-addressed\tnote[TH]{in what way??} in \cite{co98}.  \tnote[TH]{expand
-  on the COP order graph creation and it having to be bipartite for M
-  to have COP. and thus an odd cycle being an evidence of no COP.}
-\tnote[TH]{ where should this go?: (1) cite|jlm97 (application of PQ trees
-  in graphics). (2) helly's theorem citation
-  19XXdgk-Hellystheorem-Danzer-Gruenbaum-Klee}
-
-\subsection{Optimization problems in COP}
-\label{sec:optcop}
-
-So far we have been concerned about matrices that have the consecutive
-ones property. However in real life applications, it is rare that data
-sets represented by binary matrices have COP, primarily due to the
-noisy nature of data available. At the same time, COP is not arbitrary
-and is a desirable property in practical data representation
-\cite{co98,jkckv04,k77}. In this context, there are several
-interesting problems when a matrix does not have COP but is ``close''
-to having COP or is allowed to be altered to have COP. These are the
-optimization problems related to a matrix which does not have
-COP. Some of the significant problems are surveyed in this section.
-
-\tnote[TH]{ -- sect 4.1 in cite:d08phd has many results
-  surveyed. hardness results, approx. results. results are usually for
-  a class of matrices $(a,b)$ where number {\un}s in columns and rows
-  are restriced to $a$ and $b$ . -- problem of flipping at most $k$
-  entries of $M$ to make it attain COP. this is NP complete
-  cite:b75-phd}\tnote[TH]{(1) scite:lb62 showed that interval graphs
-  are AT-free.  describe AT (2) show the close relationship b/w COP
-  and graphs sec 2.2, pg 31} \cite{at72} showed that a matrix that
-does not have COP have certain substructures that prevent it from
-having COP. Tucker classified these forbidden substructures into five
-classes of submatrices. This result is presented in the context of
-convex bipartite graphs which \cite{at72} proved to be
-AT-free\tnote[TH]{ check this up. give details. - doms'}. By
-definition, convex bipartite graph have half adjacency matrices that
-have COP on either rows or columns (graph is biconvex if it has COP on
-both)\cite{d08phd}. A half adjacency matrix is a binary matrix
-representing a bipartite graph as follows. The set of rows and the set
-of columns form the two partitions of the graph. Each row node is
-adjacent to those nodes that represent the columns that have {\un}s in
-the corresponding row. \cite{at72} proves that this bipartite graph
-has no asteroidal triple if and only if the matrix has COP and goes on
-to identify the forbidden substructures for these bipartite
-graphs. The matrices corresponding to these substructures are the
-forbidden submatrices.
-
-Once a matrix has been detected to not have COP (using any of the COT
-algorithms mentioned earlier), it is naturally of interest to find out
-the smallest forbidden substructure (in terms of number of rows and/or
-columns and/or number of entries that are {\un}s). \cite{d08phd}
-discusses a couple of algorithms which are efficient if the number of
-{\un}s in a row is small. This is of significance in the case of sparse
-matrices where this number is much lesser than the number of
-columns. $(*,\Delta)${\em -matrices} are matrices with no restriction
-on number of {\un}s in any column but has at most $\Delta$ {\un}s in
-any row. {\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are
-similar problems which deals with inducing COP on a matrix. In {\sc
-  Min COS-R (Min COS-C)} the question is to find the minimum number of
-rows (columns) that must be deleted to result in a matrix with COP.
-In the dual problem {\sc Max COS-R (Max COS-C)} the search is for the
-maximum number of rows (columns) that induces a submatrix with
-COP. Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
-a submatrix $M'$ with all columns\tnote[TH]{check if b75 deals with
-  COP col or COP row. also is it any submatrix with k less than r rows
-  or submatrix must have all columns?} but a maximum cardinality
-subset of rows such that $M'$ has COP is NP complete. \cite{hg02}
-corrects an error of the abridged proof of this reduction as given in
-\cite{gj79}.  \cite{d08phd} discusses all these problems in detail
-giving an extensive survey of the previously existing results which
-are almost exhaustively all approximation results and hardness
-results. Taking this further, \cite{d08phd} presents new results in
-the area of parameterized algorithms for this
-problem\tnote[TH]{elaborate - what are the results?}.
-
-Another problem is to find the minimum number of entries in the matrix
-that can be toggled to result in a matrix with COP.  \cite{v85}
-discusses approximation of {\sc COP Augmentation} which is the problem
-of changing of the minimum number of zero entries to {\un}s so that
-the resulting matrix has COP. As mentioned earlier, this problem is
-known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
-using a reduction to the longest path problem, \tnote[TH]{or is it a
-  survey of another result?  check.} that finding a Tucker's forbidden
-submatrix of at least $k$ rows is NP complete. \tnote[TH]{how is this
-  different from booth's 75 result??}  \tnote[TH]{where should this
-  go? cite|tz04 (approx submatrix with COP sparse matrices)}
-
-\cite{jkckv04} discusses the use of matrices with almost-COP (instead
-of one block of consecutive {\un}s, they have $x$ blocks, or {\em
-  runs}, of consecutive {\un}s and $x$ is not too large) in the
-storage of very large databases.  The problem is that of reordering of
-a binary matrix such that the resulting matrix has at most $k$ runs of
-{\un}s. This is proved to be NP hard using a reduction from the
-Hamiltonian path problem.\tnote[TH]{Theorem 2.1 in jkckv}
-\tnote[TH]{(1) A connection of COP problem to the travelling salesman
-  problem is also introduced. what does this mean? -- COP can be used
-  as a tool to reorder $0.5T \le runs(M) \le T$. (2) The optimization
-  version of the $k$-run problem, i.e. minimization of number of
-  blocks of ones is proven to be NP complete by
-  cite:k77}\tnote[TH]{are these two the same?} \tnote[TH]{what is the
-  reduction?} \tnote[TH]{other problems similar to COP -- cite:ckl96
-  (ILP, circ ones, one drop) -- cite:th98 (generalization of COP -
-  minimax, biotonic column) Tucker}
-
-\section{********* Application of COP in Areas of Graph Theory and
-  Algorithms}
-\label{sec:appcopGTA}
-
-\subsection{********* COP in Relational Database Model}
-\label{sec:apprdbm}
-\tnote{(set systems theme)}
-
-\subsection{********* COP in Graph Isomorphism}
-\label{sec:appgraphiso}
-\tnote{(canonization theme)}
-
-\subsection{********* Certifying Algorithms}
-\label{sec:appcertalgo}
-\tnote{ (certification McC04 theme)}
-
-\section{Generalization of COP - the Motivation}
-\label{sec:motive}
-Section~\ref{sec:copmatrices} introduced a succinct characterization
-for consecutive-ones property which is solely based on the cardinality
-properties of the set representations of the matrix's columns
-\cite{nsnrs09}. This result is very relevant to this thesis because
-aside from it simplifying COT to a great degree, our generalization
-problem is motivated by their results.
-
-\cite{nsnrs09} characterizes interval assignments to the sets which
-can be obtained from a single permutation of the rows.  For an
-assignment to be feasible, the cardinality of the interval assigned to
-each set in the system must be same as the cardinality of the set, and
-the intersection cardinality of any two intervals must be same as the
-intersection cardinality of their corresponding sets.  While this is
-obviously a necessary condition, this result shows this is also
-sufficient.  \cite{nsnrs09} calls this an Intersection Cardinality
-Preserving Interval Assignment (ICPIA).  This paper generalizes the
-idea from \cite{wlh02} of decomposing a given binary matrix into prime
-matrices for COT and describes an algorithm to test if an ICPIA exists
-for a given set system.
-
-The equivalence of the problem of testing for the consecutive-ones
-property to the constraint statisfaction problem of interval
-assignment \cite{nsnrs09} or interval labeling \cite{kklv10} is as
-follows. Every column (row) of the binary matrix can be converted into
-a set of non-negative integers which are the indices of rows (columns)
-with {\un}s in that column (row). It is apparent that if the matrix
-has COP in columns (rows), then constructing such sets after applying
-the COP permutation to the rows (columns) of the matrix will result in
-sets with consecutive integers. In other words, after application of
-COP reordering, the sets are intervals. Indeed the problem now becomes
-finding interval assignments to a given set system such that there
-exists a permutation of the universe of set of row indices (column
-indices) which converts each set to its assigned interval.
-
-The problem of interest in this thesis, namely, tree path labeling
-problem, is a natural generalization of the interval assignment
-problem or the COT problem. The problem is defined as follows -- given
-a set system $\cF$ from a universe $U$ and a target tree $T$, does
-there exist a bijection from $U$ to the vertices of $T$ such that each
-set in the system maps to a path in $T$.  We refer to this as the
-{\CFTPL} problem or simply {\em tree path labeling} problem for an
-input set system and target tree pair -- $(\cF,T)$. The special case
-of the target tree being a path, is the interval assignment problem.
-We focus on generalizing the notion of an ICPIA \cite{nsnrs09} to
-characterize feasible path assignments.  We show that for a given set
-system $\cF$, a tree $T$, and an assignment of paths from $T$ to the
-sets, there is a feasible\footnote{The notion of {\em feasibility} is
-  formally defined in Section~\ref{ch:prelims}.}  bijection between
-$U$ and $V(T)$ if and only if the intersection cardinalities among any
-three sets (not necessarily distinct) is equal to that of the
-corresponding paths assigned to them and the input passes a filtering
-algorithm (described in this paper) successfully.  This algorithmic
-characterization gives a natural data structure that stores all the
-\remove{relevant} feasible bijections between $U$ and $V(T)$. This
-reduces the search space for the solution considerably from the
-universe of all possible bijections between $U$ and $V(T)$ to only
-those bijections that maintain the characterization.  Further, the
-filtering algorithm is also an efficient algorithm to test if a tree
-path labeling\footnote{The terms {\em tree path labeling} and {\em
-    tree path assignment} are, in informal language,
-  synonyms. Formally, the former refers to the bijection $\cl: \cF
-  \rightarrow \cP$. The latter refers to the set of ordered pairs
-  $\{(S, P) \mid S \in \cF, P \in \cP\}$. $\cP$ is a set of paths on
-  $T$.} is feasible.
-
-\section{Summary of New Results in this Thesis}
-\label{sec:results}
-
-We see in Section~\ref{sec:motive} that pairwise intersection
-cardinality preservation is necessary and sufficient for an interval
-assignment to be feasible for a given hypergraph\footnote{A {\em
-    hypergraph} is an alternate representation of a set system and
-  will be used in this thesis. See Section~\ref{ch:prelims} for the
-  formal definition.} and thus is a characterization for COP
-\cite{nsnrs09}. \tnote{Refer to the survey section that will list the
-  theorems and/or lemma?} In our work we extend this characterization
-and find that trio-wise intersection cardinality preservation makes a
-tree path labeling \footnote{A {\em tree path labeling} $\cl$ is a
-  bijection of paths from the target tree $T$ to the hyperedges in
-  given hypergraph $\cF$.  See Section~\ref{ch:prelims} for the formal
-  definition.}  (TPL) feasible, which is a generalization of the COP
-problem. This problem is defined as follows.
-
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \FTPL}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
-        paths $\cP$ from $T$ and a
-        bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
-      \end{minipage}\\
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
-        such that $\phi$ when applied on any hyperedge in $\cF$ will
-        give
-        the path mapped to it by the given tree path labeling $\cl$.\\
-        { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every
-          hyperedge $S \in \cF$.}
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
-
-We give a necessary and sufficient condition by way of {\em
-  Intersection Cardinality Preservation Path Labeling} (ICPPL) and a
-filtering algorithm for {\FTPL} to output in affirmative. ICPPL
-captures the trio-wise cardinality property described
-earlier\footnote{See Section~\ref{sec:feasible} for the definition of
-  ICPPL.}. This characterization can be checked in polynomial time.  A
-relevant consequence of this constructive procedure is that it is
-sufficient to iteratively check if three-way intersection
-cardinalities are preserved.  In other words, in each iteration, it is
-sufficient to check if the intersection of any three hyperedges is of
-the same cardinality as the intersection of the corresponding paths.
-Thus this generalizes the well studied question of the feasible
-interval assignment problem which is the special case when the target
-tree $T$ is simply a path \cite{wlh02,nsnrs09}.
-
-Aside from checking if a given TPL is feasible, we also solve the
-problem of computing a feasible TPL for a given hypergraph and target
-tree, if one exists. This problem, {\CFTPL}, is defined as follows.
-
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \CFTPL}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$ and a tree $T$.\\
-      \end{minipage}\\
-
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a set of paths $\cP$ from $T$ and a bijection
-        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
-        {\bf true} on $(\cF, T, \cl)$.
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
-
-We present a polynomial time algorithm for {\CFTPL} when the target
-tree $T$ belongs to a special class of trees called {\em \kstar s} and
-when the hyperedges in the hypergraph $\cF$ have at most $k+2$
-vertices. A couple of examples of {\kstar s} are given in
-Figure~\ref{fig:ksubstar}.
-
-\begin{figure}[t]
-  \centering
-  \begin{tabular}[h]{ccccc}
-    \includegraphics[scale=0.3]{../img/star.pdf} &&&&
-    \includegraphics[scale=0.3]{../img/kstar.pdf}\\
-    (a) &&&& (b)
-  \end{tabular}
-  \caption{\figtabsize Examples of {\kstar s}. (a) $k = 0$ (b) $k = 2$
-  }
-  \label{fig:ksubstar}
-\end{figure}
-
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \CFTPLKTREE}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$ such that every
-        hyperedge
-        $S \in \cF$ is of cardinality at most $k+2$ and a {\kstar} $T$.\\
-      \end{minipage}\\
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a set of paths $\cP$ from $T$ and a bijection
-        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
-        {\bf true} on $(\cF, T, \cl)$.
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
-
-\tnote{The following sentence seems out of place in this para} In
-spite of this being a restricted case, we believe that our results are
-of significant interest in understanding the nature of {\sc Graph
-  Isomorphism} which is polynomial time solvable in interval graphs
-while being hard on path graphs\cite{kklv10}. {\kstar s} are a class
-of trees which are in many ways very close to intervals or paths. Each
-ray\footnote{The path from a leaf to the root, the vertex with highest
-  degree, is called a {\em ray} of the \kstar. See
-  Section~\ref{ch:prelims}.} are independent except for the
-root\footnote{The vertex with maximum degree in a {\kstar} is called
-  {\em root}. See Section~\ref{ch:prelims}.} and hence can be
-considered as an independent interval till the root. Our algorithm
-builds on this fact and uses the interval assignment
-algorithm\cite{nsnrs09} up until ``reaching'' the root and then uses
-the trio-wise intersection cardinality (the extra condition in ICPPL
-that generalizes ICPIA) check to resolve the ambiguity about which ray
-the algorithm should ``grow'' the solution into in the next iteration.
-
-We also have an algorithm for solving {\CFTPL} with no restrictions on
-the target tree or set size which runs in exponential time.  This
-algorithm finds a path labeling from $T$ by decomposing the problem
-into subproblems of finding path labeling of subsets of $\cF$ from
-subtrees of $T$. Given the fact that binary matrices naturally
-represent a set system (see Section~\ref{sec:motive}) and that the
-{\em overlap}\footnote{}\tnote{Give informal definition of ``overlap''
-  in footnote and link to sec:prelims. ADD the definition in prelim.}
-relation between the sets involved is an obvious equivalence relation,
-$\cF$ quite naturally partitions into equivalence classes known as
-{\em overlap components}\tnote{Give informal definition of ``overlap
-  components'' in footnote and link to sec:prelims. ADD the definition
-  in prelim.}. In the context of COP, overlap components were used in
-\cite{wlh02} and \cite{kklv10}. Moreover, \cite{nsnrs09} discovered
-that these equivalence classes form a total order \tnote{Give informal
-  definition of ``total order'' in footnote?}. We extend this to TPL
-and find that when $\cF$ is a path hypergraph\footnote{If there exists
-  an FTPL for a hypergraph $\cF$, it is called a path hypergraph.},
-the classes can be partially ordered as an in-tree in polynomial
-time. Once $\cF$ is ``broken'' into overlap components, one must
-identify the subtree of $T$ that it needs to map to and this is the
-hard part which is currently open to be solved in polynomial time.
-
-\tnote[TH]{The connection of TPL to graph isomorphism will be made
-  later in the document}
-
-\chapter{Consecutive-ones Property - a Survey of Important Results}
-\label{ch:copsurvey}
-\tnote{TBD: have a few lines about organization of chapter}
-
-\tnote{ Survey chapter: is the full fledged expansion of the survey in
-  introduction with details, observations, theorems etc.}
-
-\section{Matrices with COP}
-\label{sec:surveycoptest}
-\tnote{Expand on ref:sec:copmatrices}
-
-\section{Optimization problems in COP}
-\label{sec:surveycopopt}
-\tnote{Expand on ref:sec:optcop}
-
-\section{********* COP in Relational Database Model}
-\label{sec:surveyrdbm}
-\tnote{Expand on sec:apprdbm}
-\tnote{(set systems theme)}
-
-\subsection{********* COP in Graph Isomorphism}
-\label{sec:surveygraphiso}
-\tnote{Expand on sec:appgraphiso}
-\tnote{(canonization theme)}
-
-\subsection{********* Certifying Algorithms}
-\label{sec:surveycertalgo}
-\tnote{Expand on sec:appcertalgo}
-\tnote{ (certification McC04 theme)}
-
-\chapter{Tree Path Labeling of Path Hypergraphs - the New Results}
-\label{ch:myresearch}
-
-\tnote{UPDATE TO FINAL VERSION. BELOW IS OLDER VERSION WITH INCOMPLETE
-  SEC 4.}  This chapter documents all the new results obtained by us in the
-area of tree path labeling of path hypergraphs which is the parent
-problem addressed in this thesis. In Section~\ref{sec:motive} we see
-that consecutive-ones property and its equivalent problem of interval
-labeling of a hypergraph is a special case of the general problem of tree
-path labeling of path hypergraphs.
-
-\tnote{UPDATE the following para to current skeleton.}The necessary
-preliminaries with definitions etc. are presented in
-Section~\ref{ch:prelims}. Section~\ref{sec:feasible} documents the
-characterization of a feasible path labeling for any path labeling
-from any target trees. Section~\ref{sec:spltargettree} describes the
-special case where the target tree is of a particular family of trees
-and specifically, Section~\ref{sec:ksubdivstar} describes a polynomial
-time algorithm to find the tree path labeling of a given set system
-from a given $k$-subdivided tree. Section~\ref{sec:norestraint}
-describes the general situation where the target tree has no
-restrictions and the algorithm to find a TPL, if any, in this case.
-
-\section{Introduction}
-
-The problem of consecutive-ones property testing can be easily seen as
-a simple constraint satisfaction problem involving a hypergraph or a
-system of sets from a universe. Every column (row) of the binary
-matrix can be converted into a set of non-negative integers which are
-the indices of the rows (columns) with {\un}s in that column
-(row). When observed in this context, if the matrix has the COP on
-columns (rows), a reordering of its rows (columns) will result in sets
-that have only consecutive integers. In other words, the sets after
-applying the COP row (column) permutation are intervals.  In this
-form, one can see that this is indeed the problem of finding interval
-assignments to the given set system \cite{nsnrs09} with a single
-permutation of the universe (set of row or column indices for COP of
-columns or rows, respectively) which permutes each set to its
-interval. The result in \cite{nsnrs09} characterizes interval
-assignments to the sets which can be obtained from a single
-permutation of the rows (columns). They show that for each set, the
-cardinality of the interval assigned to it must be same as the
-cardinality of the set, and the intersection cardinality of any two
-sets must be same as the intersection cardinality of the corresponding
-intervals. This is a necessary and sufficient
-condition. \tnote{(ICPIA) put reference to survey chapter}.  Finally,
-the idea of decomposing a given binary matrix into prime matrices to
-check for COP is adopted from cite|wlh02| to test if an ICPIA exists
-for a given set system.\tnote{REPEATITION from intro chapter. but idea
-  recall is necessary. so keep it but reword it to avoid monotone.}
-
-A natural generalization of the interval assignment problem is
-feasible tree path labeling problem of a set system. The problem is
-defined as follows -- given a set system $\cF$ from a universe $U$ and
-a tree $T$, does there exist a bijection from $U$ to the vertices of
-$T$ such that each set in the system maps to a path in $T$.  We refer
-to this as the {\em tree path labeling problem} for an input set
-system, target tree pair -- $(\cF,T)$. \tnote{USE THE FORMAL PROBLEM DEF BLOCK}
-%\begin{tabular}[t]{c}
-{\small
-\begin{minipage}[h]{5in}
- % \begin{singlespace}
- \vspace{2mm}
- {\large \FTPL}\\
- \begin{tabular}[t]{l|l}
- \hline\\
-    {\tt Input} & 
-    \begin{minipage}[t]{\probdefwidth}
-      A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
-      paths $\cP$ from $T$ and a
-      bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
-    \end{minipage}\\
-    {\tt Question} &
-    \begin{minipage}[t]{\probdefwidth}
-      Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
-      such that $\phi$ when applied on any hyperedge in $\cF$ will
-      give
-      the path mapped to it by the given tree path labeling $\cl$.\\
-      { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every hyperedge
-        $S \in \cF$.}
-    \end{minipage}\\
-  \end{tabular}
-  % \end{singlespace}
-\end{minipage}\\
-}
-%\end{tabular}
-
-As a special case if the tree
-$T$ is a path, the problem becomes the interval assignment problem. \tnote{USE THE FORMAL PROBLEM DEF BLOCK}
-We focus on the question of generalizing the notion of an ICPIA
-\cite{nsnrs09} to characterize feasible path assignments.  We show
-that for a given set system $\cF$, a tree $T$, and an assignment of
-paths from $T$ to the sets, there is a feasible bijection between $U$
-and $V(T)$ if and only if all intersection cardinalities among any
-three sets (not necessarily distinct) is same as the intersection
-cardinality of the paths assigned to them and the input runs a
-filtering algorithm (described in this paper) successfully.  This
-characterization is proved constructively and it gives a natural data
-structure that stores all the relevant feasible bijections between $U$
-and $V(T)$.  Further, the filtering algorithm is also an efficient
-algorithm to test if a tree path labeling to the set system is
-feasible.  This generalizes the result in \cite{nsnrs09}.
-
-In the later part of this paper, we focus on a new special case of the
-tree path labeling problem. Here the set system is such that every set
-is at most $k+2$ in size and for every pair of sets in it there exists
-a sequence of sets between them with consecutive sets in this sequence
-having a strict intersection -- i.e., non-empty intersection with
-neither being contained in the other. Moreover, the given tree is a
-{\em $k$-subdivided star}. We demonstrate a polynomial time algorithm
-to find a feasible path labeling in this case.
-
-\section{Preliminaries to new results}
-\label{ch:prelims}
-
-\tnote{TBD: have a few lines about organization of chapter}This
-section states definitions and basic facts necessary in the scope of
-this document.
-
-\tnote{Move some of this to general prelim in chapter 1}
-
-The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
-  system} of a universe $U$ with $|U| = n$.  The {\em support} of a
-set system $\F$ denoted by $supp(\cF)$ is the union of all the sets in
-$\F$; $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this
-paper, a set system is required to ``cover'' the universe; $
-supp(\cF) = U$.
-
-The graph $T$ represents a {\em target tree} with same
-number of vertices as elements in $U$; $|V(T)| = n$.  A {\em path
-  system}\, $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P
-\mid P \subseteq V, \text{ } T[P] \text{ is a path} \}$.
-
-% 
-% A set system $\cF$ can be alternatively represented by a {\em
-%   hypergraph} $\H_\cF$ whose vertex set is $supp(\cF)$ and
-% hyperedges are the sets in $\cF$. This is a known representation for
-% interval systems in literature \cite{bls99,kklv10}.  We extend this
-% definition here to path systems.
-
-A set system $\cF$ can be alternatively represented by a {\em
-  hypergraph}\, $\cF_H$ whose vertex set is $supp(\cF)$ and hyperedges
-are the sets in $\cF$. This is a known representation for interval
-systems in literature \cite{bls99,kklv10}.  We extend this definition
-here to path systems. Due to the equivalence of set system and
-hypergraph in the scope of this paper, we drop the subscript $_H$ in
-the notation and refer to both the structures by $\cF$.
-
-Two hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to
-each other, denoted by $\cF' \cong \cF''$, iff there exists a
-bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for all
-sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ iff $B$ is
-a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
-\cite{kklv10}. This is called {\em hypergraph
-  isomorphism}. \tnote{also extend $\phi$ to hyperedges -- see if
-  required }
-
-The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph $\cF$ is a
-graph such that its vertex set has a bijection to $\cF$ and there
-exists an edge between two vertices iff their corresponding hyperedges
-have a non-empty intersection \cite{mcg04}.
-
-If the intersection graphs of two hypergraphs are isomorphic,
-$\bI(\cF) \cong \bI(\cP)$ where $\cP$ is also a path system, then the
-bijection $\cl: \cF \rightarrow \cP$ due to this isomorphism is called
-a {\em path labeling} of the hypergraph $\cF$. To illustrate further,
-let $\cg: V(\cF) \rightarrow V(\cP)$ be the above mentioned
-isomorphism where $V(\cF)$ and $V(\cP)$ are the vertex sets that
-represent the hyperedges for each hypergraph respectively, $V(\cF) =
-\{ v_S \mid S \in \cF\}$ and $V(\cP) = \{ v_P \mid P \in \cP\}$. Then
-the path labeling $\cl$ is defined as follows: $\cl(S_1) = P_1$ iff
-$\cg (v_{S_1}) = v_{P_1}$. The path system $\cP$ may be alternatively
-denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In most scenarios in
-this paper, what is given are the pair $(\cF, \cl)$ and the target
-tree $T$; hence this notation will be used more often.
-
-If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$
-is called a {\em path hypergraph} and $\cP$ is called {\em path
-  representation} of $\cF$. If this isomorphism is $\phi: supp(\cF)
-\rightarrow V(T)$, then it is clear that there is an {\em induced path
-labeling} $\cl_\phi: \cF \rightarrow \cP$ to the set system;
-$\cl_\phi(S) = \set{y \mid y = \phi(x), x \in S}$ for all $S \in \cF$. Recall that
-$supp(\cP) = V(T)$.
-
-A path labeling $(\cF, \cl)$ is defined to be {\em
-  feasible} if
-$\cF \cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
-\rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
-\rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.
-
-An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is
-a graph such that its vertex set has a bijection to $\cF$ and there
-exists an edge between two of its vertices iff their corresponding
-hyperedges overlap. Two hyperedges $S$ and $S'$ are said to {\em
-  overlap}, denoted by $S \overlap S'$, if they have a non-empty
-intersection and neither is contained in the other; $S \overlap S'
-\text{ iff } S \cap S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq
-S$. Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$ and not
-necessarily connected. Each connected component of $\bO(\cF)$ is
-called an {\em overlap component}.
-
-A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
-\overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
-\cite{kklv10}. Additionally, if $S$ is such that it is contained in no
-other hyperedge in $\cF$, i.e., it is inclusion maximal then it is called
-{\em super-marginal}.
-
-A {\em star} graph is a complete bipartite graph $K_{1,p}$
-which is clearly a tree and $p$ is the number of leaves. The vertex
-with maximum degree is called the {\em center} of the star and the
-edges are called {\em rays} of the star. A {\em $k$-subdivided star}
-is a star with all its rays subdivided exactly $k$ times. The
-definition of a {\em ray of a $k$-subdivided star} is extended to the path
-from the center to a leaf. It is clear that all rays are of length $k+2$.
-
-\section{Characterization of Feasible Tree Path  Labeling} 
-\label{sec:feasible}
-
-In this section we give an algorithmic characterization of a
-feasibility of tree path labeling.  Consider a path labeling $(\cF,
-\cl)$ on the given tree $T$. We call $(\cF, \cl)$ an {\em Intersection
-  Cardinality Preserving Path Labeling (ICPPL)} if it has the
-following properties.
-
-\begin{enumerate}[{(\icpplpr\ }i) \ \ \ ]
-\item \label{pr:i} $|S| = |\cl(S)|$ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $S \in \cF$
-  % \vspace{\topshrink}
-\item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ \ \ \ \
-  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all distinct $S_1, S_2 \in \cF$
-  % \vspace{\topshrink}
-\item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
-  \cap \cl(S_3)|$ \ \ \ for all distinct $S_1, S_2, S_3 \in \cF$
-\end{enumerate}
-
-The following lemma is useful in subsequent arguments. 
-\begin{lemma}
-  \label{lem:setminuscard}
-  If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
-  (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
-  \cl(S_3))|$.
-\end{lemma}
-\begin{proof}%[Proof of Lemma~\ref{lem:setminuscard}]
-  Let $P_i = \cl(S_i)$, for all $1 \le i \le  3$.
-  $|S_1 \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| =
-  |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to properties (ii)
-  and (iii) of ICPPL, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
-  \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
-  = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. \qed
-\end{proof}
-
-In the remaining part of this section we show that $(\cF,
-\cl)$ is feasible if and only if it is an ICPPL and
-Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
-function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two levels of
-filtering of $(\cF, \cl)$ to make it simpler while retaining the set
-of isomorphisms, if any, between $\cF$ and $\cF^\cl$.
-% One direction of this claim isclear: that if a path
-% labeling is feasible, then all intersection cardinalities are
-% preserved, i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
-% \annote{has no premature exit condition hence any input will go
-%   through it}{Prove that the filtered sets has ICPPL iff input PL
-%   has ICPPL?}. Algorithm~\ref{leafasgn} has an exit condition at
-% line~\ref{xempty}. It can be easily verified that $X$ cannot be
-% empty if $\cl$ is a feasible path labeling. The reason is that a
-% feasible path labeling has an associated bijection between
-% $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that
-% the sets map to paths, ``preserving'' the path labeling.  The rest
-% of the section is devoted to constructively proving that it is
-% sufficient for a path labeling to be an ICPPL and pass the two
-% filtering algorithms.  To describe in brief, the constructive
-% approaches refine an ICPPL iteratively, such that at the end of each
-% iteration we have a ``filtered'' path labeling, and finally we have
-% a path labeling that defines a family of bijections from $supp(\cF)$
-% to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.
-First, we present Algorithm~\ref{perms} or {\tt filter\_1}, and prove
-its correctness.  This algorithm refines the path labeling by
-processing pairs of paths in $\cF^\cl$ that share a leaf until no two
-paths in the new path labeling share any leaf.
-
-\begin{algorithm}[h]
-  \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
-  \label{perms}
-  \begin{algorithmic}[\lndisplay]
-    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$\\
-    \STATE $j \assign 1$\\
-    \WHILE {there is $S_1, S_2 \in \cF_{j-1}$ such that
-      $\cl_{j-1}(S_1)$ and $\cl_{j-1}(S_2)$ have a common leaf in
-      $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
-    \{S_1, S_2\})
-    \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1 \}$ \label{setbreak} 
-    \COMMENT {Remove $S_1$, $S_2$ and add the ``filtered'' sets}
-    \STATE {\bf for} every $S \in \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
-    S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\
-
-    \STATE $\cl_j(S_1 \cap S_2) \assign \cl_{j-1}(S_1) \cap
-    \cl_{j-1}(S_2)$
-    \COMMENT {Carry forward the path labeling for all existing sets other than
-      $S_1$, $S_2$}
-    \STATE $\cl_j(S_1 \setminus S_2) \assign \cl_{j-1}(S_1) \setminus
-    \cl_{j-1}(S_2)$ 
-    \COMMENT {Define path labeling for new sets}
-    \STATE $\cl_j(S_2 \setminus S_1) \assign \cl_{j-1}(S_2) \setminus
-    \cl_{j-1}(S_1)$
-
-    \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of ICPPL}
-    \label{ln:3waycheck}
-    \STATE {\bf exit} \label{ln:exit1} \\
-    \ENDIF
-
-    \STATE $j \assign j+1$\\
-    \ENDWHILE
-    \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
-    \RETURN $(\cF', \cl')$
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma} 
- \label{lem:feasible} 
- In Algorithm~\ref{perms}, if input $(\cF, \cl)$ is a feasible path
- assignment then at the end of $j$th iteration of the {\bf while}
- loop, $j \ge 0$, $(\cF_j, \cl_j)$ is a feasible path assignment.
-\end{lemma}
-\begin{proof}%[Proof of Lemma~\ref{lem:feasible}]
-  We will prove this by mathematical induction on the number of
-  iterations. The base case $(\cF_0, \cl_0)$ is feasible since it is
-  the input itself which is given to be feasible. Assume the lemma is
-  true till $j-1$th iteration. i.e. every hypergraph isomorphism
-  $\phi: supp\left(\cF_{j-1}\right) \rightarrow V\left(T \right)$ that
-  defines $(\cF, \cl)$'s feasibility, is such that the induced path
-  labeling on $\cF_{j-1}$, $\cl_{\phi[{\cF_{j-1}}]}$ is equal to
-  $\cl_{j-1}$. We will prove that $\phi$ is also the bijection that
-  makes $(\cF_j, \cl_j)$ feasible. Note that $supp(\cF_{j-1}) =
-  supp(\cF_{j})$ since the new sets in $\cF_j$ are created from basic
-  set operations to the sets in $\cF_{j-1}$. For the same reason and
-  $\phi$ being a bijection, it is clear that when applying the $\phi$
-  induced path labeling on $\cF_j$, $ \cl_{\phi[{\cF_{j}}]}(S_1
-  \setminus S_2) = \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
-  \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Now observe that $ \cl_j(S_1
-  \setminus S_2) = \cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2) =
-  \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
-  \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Thus the induced path labeling
-  $\cl_{\phi[{\cF_{j}}]} = \cl_{j}$. Therefore lemma is proven.  \qed
-\end{proof}
-
-\begin{lemma}
-  \label{lem:invar1} In Algorithm~\ref{perms}, at the end of $j$th
-  iteration, $j \ge 0$, of the {\bf while} loop, the following
-  invariants are maintained.
-  \begin{enumerate}[I {\ }] %\vspace{\topshrink}
-  \item $\cl_j(R)$ is a path in $T$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ for all $R \in \cF_j$%\vspace{\topshrink}
-  \item $|R| = |\cl_j(R)|$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $R \in
-    \cF_j$%\vspace{\topshrink}
-  \item $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ for all $R, R' \in \cF_j$%\vspace{\topshrink}
-  \item $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
-    \cl_j(R'')|$, \ \ \ for all $R, R', R'' \in \cF_j$
-  \end{enumerate}
-\end{lemma}
-
-\begin{proof}
- Proof is by induction on the number of iterations, $j$. In this
-  proof, the term ``new sets'' will refer to the sets added to $\cF_j$
-  in $j$th iteration in line~\ref{setbreak} of Algorithm~\ref{perms},
-  $S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1$ and its
-  images in $\cl_j$ where $\cl_{j-1}(S_1)$
-  and $\cl_{j-1}(S_2)$ intersect and share a leaf.\\
-  The invariants are true in the base case $(\cF_0, \cl_0)$, since it
-  is the input ICPPL.  Assume the lemma is true till the $j-1$th
-  iteration. Let us consider the possible cases for each of the above invariants for
-  the $j$th iteration.
-
- \begin{enumerate}[\xbullet]
-  \item {\em Invariant} I/II
-    \begin{enumerate}[{I/II}a $|$] % \textbullet 
-    \item {\em $R$ is not a new set.} It is in $\cF_{j-1}$. Thus
-      trivially true by induction hypothesis.
-    \item {\em $R$ is a new set.} If $R$ is in $\cF_{j}$ and not in
-      $\cF_{j-1}$, then it must be one of the new sets added in
-      $\cF_j$. In this case, it is clear that for each new set, the
-      image under $\cl_j$ is a path since by definition the chosen
-      sets $S_1$, $S_2$ are from $\cF_{j-1}$ and due to the while loop
-      condition, $\cl_{j-1}(S_1)$, $\cl_{j-1}(S_2)$ have a
-      common leaf. Thus invariant I is proven.\\
-      Moreover, due to induction hypothesis of invariant III and the
-      definition of $l_j$ in terms of $l_{j-1}$, invariant II is
-      indeed true in the $j$th iteration for any of the new sets.  If
-      $R = S_1 \cap S_2$, $|R| = |S_1 \cap S_2| = |\cl_{j-1}(S_1) \cap
-      \cl_{j-1}(S_2)| = |\cl_j(S_1 \cap S_2)| = |\cl_j(R)|$.
-      If $R = S_1 \setminus S_2$, $|R| = |S_1 \setminus S_2| = |S_1| -
-      |S_1 \cap S_2| = |\cl_{j-1}(S_1)| - |\cl_{j-1}(S_1) \cap
-      \cl_{j-1}(S_2)| = |\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)| =
-      |\cl_j(S_1 \setminus S_2)|
-      = |\cl_j(R)|$. Similarly if $R = S_2 \setminus S_1$.\\
-    \end{enumerate}
-  \item {\em Invariant} III
-    \begin{enumerate}[{III}a $|$]
-    \item {\em $R$ and $R'$ are not new sets.} It is in
-      $\cF_{j-1}$. Thus trivially true by induction hypothesis.
-    \item {\em Only one, say $R$, is a new set.} Due to invariant IV
-      induction hypothesis, Lemma~\ref{lem:setminuscard} and
-      definition of $\cl_j$, it follows that invariant III is true no
-      matter which of the new sets $R$ is equal to. If $R = S_1 \cap
-      S_2$, $|R \cap R'| = |S_1 \cap S_2 \cap R'| = |\cl_{j-1}(S_1)
-      \cap \cl_{j-1}(S_2) \cap \cl_{j-1}(R')| = |\cl_j(S_1 \cap S_2)
-      \cap \cl_j(R')| = |\cl_j(R) \cap \cl_j(R')|$.  If $R = S_1
-      \setminus S_2$, $|R \cap R'| = |(S_1 \setminus S_2) \cap R'| =
-      |(\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)) \cap \cl_{j-1}(R')| =
-      |\cl_{j}(S_1 \cap S_2) \cap \cl_{j}(R')| = |\cl_{j}(R) \cap
-      \cl_{j}(R')|$. Similarly, if $R = S_2 \setminus
-      S_1$. Note $R'$ is not a new set.\\
-
-    \item {\em $R$ and $R'$ are new sets.} By definition, the new
-      sets and their path images in path label $\cl_j$ are disjoint so
-      $|R \cap R'| = |\cl_j(R) \cap \cl_j(R)| = 0$. Thus case proven.
-    \end{enumerate}
-  \item {\em Invariant} IV
-
-    Due to the condition in line~\ref{ln:3waycheck}, this invariant is
-    ensured at the end of every iteration.
-%     \begin{enumerate} [{Case 3.}1:]
-%     \item {\em $R$, $R'$ and $R''$ are not new sets.} Trivially
-%       true by induction hypothesis.
-%     \item {\em Only one, say $R$, is a new set.}
-%       If $R = S_1 \cap S_2$,  from Lemma~\ref{lem:fourpaths} and
-%       invariant III hypothesis,  this case is proven. Similarly if $R$
-%       is any of the other new  sets, the case is proven by also using
-%       Lemma ~\ref{lem:setminuscard}.
-%     \item {\em At least two of $R, R', R''$ are new sets.}
-%       The new sets are disjoint hence this case is vacuously true.
-%     \end{enumerate}
-  \end{enumerate} \qed
-%\vspace{-6mm}
-
-\end{proof}
-
-\begin{lemma}
-  \label{lem:noexit1}
-  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{perms} is
-  feasible, then the set of hypergraph isomorphism functions that
-  defines $(\cF, \cl)$'s feasibility is the same as the set that
-  defines $(\cF_j, \cl_j)$'s feasibility, if any.  Secondly, for any
-  iteration $j > 0$ of the {\em \bf while} loop, the {\em \bf exit}
-  statement in line~\ref{ln:exit1} will not execute.
-\end{lemma}
-\begin{proof}
-  Since $(\cF,\cl)$ is feasible, by Lemma~\ref{lem:feasible}
-  $(\cF_j,\cl_j)$ for every iteration $j > 0$ is feasible.  % Therefore,
-%   every hypergraph isomorphism $\phi: supp(\cF) \rightarrow V(T)$ that
-%   induces $\cl$ on $\cF$ also induces $\cl_{j-1}$ and $\cl_{j}$ on
-%   $\cF_{j-1}$ and $\cF_{j}$ respectively, i.e., $\cl_{\phi[\cF_{j-1}]}
-%   = \cl_{j-1}$ and $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
-%   that for all $x \in supp(\cF)$, for all $v \in V(T)$ the following
-%   hold true.
-  Also, every hypergraph isomorphism $\phi: supp(\cF) \rightarrow
-  V(T)$ that induces $\cl$ on $\cF$ also induces $\cl_{j}$ on
-  $\cF_{j}$, i.e., $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
-  that for all $x \in supp(\cF)$, for all $v \in V(T)$, if $(x,v) \in
-  \phi$ then $v \in \cl_{j}(S)$ for all $S \in \cF_{j}$ such that $x
-  \in S$.
-% the following
-%   hold true.
-%   \begin{enumerate}[i. ]
-%   \item If $(x,v) \in \phi$ then $v \in \cl_{j-1}(S)$ for all $S \in
-%     \cF_{j-1}$ such that $x \in S$.
-%   \item If $(x,v) \in \phi$ then $v \in \cl_{j}(S)$ for all $S \in
-%     \cF_{j}$ such that $x \in S$
-%   \end{enumerate}
-  In other words, filter 1 outputs a filtered path labeling that
-  ``preserves''
-  hypergraph isomorphisms of the original path labeling.\\
-  Secondly, line~\ref{ln:exit1} will execute iff the exit condition in
-  line~\ref{ln:3waycheck}, i.e. failure of three way intersection
-  preservation, becomes true in any iteration of the {\em \bf while}
-  loop.  Due to Lemma~\ref{lem:invar1} Invariant IV, the exit
-  condition does not occur if the input is a feasible ICPPL.\qed
-
-%   such that $\phi(x) = v$ where $v$ is the leaf considered in the
-%   first iterations of while. Clearly, $\phi$ is a renaming of
-%   vertices in hypergraph $\cF$ to those in hypergraph $\cF^\cl$. Thus
-%   the following facts can be observed in every iteration of the loop.
-
-%   \begin{enumerate}[\hspace{2mm}i. ] \vspace{\topshrink}
-%   \item all intersection cardinalities are preserved in this path
-%     labeling \vspace{\topshrink}
-%   \item element $x$ is exclusive in a hyperedge in $\cF$ since $v$ is
-%     exclusive in a hyperedge in $\cF^\cl$.
-%   \end{enumerate}
-
-%   Thus the exit condition is never rendered true after $x$ and $v$ are
-%   removed from their respective hyperedges. \qed
-
-% \noindent
-% This proof uses mathematical induction on the number
-%   of iterations $j$, $j \ge 0$, of the loop that executed
-%   without exiting. The base case, $j = 0$ is obviously true since the
-%   input is an ICPPL and the exit condition cannot hold true due to
-%   ICPPL property (iii).  Assume the algorithm executes till the end
-%   of $j-1$th iteration without exiting at line
-%  ~\ref{ln:3waycheck}. Consider the $j$th iteration. From Lemma
-%  ~\ref{lem:feasible} we know that $(\cF_j, \cl_j)$ and $(\cF_{j-1},
-%   \cl_{j-1})$ are feasible\remove[AS]{and from the proof in lemma
-%     lem:invar1 we know that $(\cF_{j-1}, \cl_{j-1})$ satisfies all the
-%     invariants defined in the lemma}.  Thus there exists a bijection
-%   $\phi: supp(\cF) \rightarrow V(T)$ such that the induced path
-%   % labeling on $\cF_{j-1}$ $\cl_{\phi[\cF_{j-1}]} = \cl_{j-1}$.
-%   labeling on $\cF_{j}$, $\cl_{\phi[\cF_{j}]}$ and on $\cF_{j-1}$,
-%   $\cl_{\phi[\cF_{j-1}]}$ are equal to $\cl_{j}$ and $\cl_{j-1}$
-%   respectively.  We need to prove that for any $R, R', R'' \in
-%   \cF_{j}$, $|R \cap R' \cap R''| = |\cl_j(R) \cap \cl_j(R') \cap
-%   \cl_j(R'')|$.
-%   The following are the possible cases that could arise. From argument
-%   above, $|\cl_j(R) \cap \cl_j(R') \cap \cl_j(R'')| =
-%   |\cl_{\phi[\cF_{j}]}(R) \cap \cl_{\phi[\cF_{j}]} (R') \cap
-%   \cl_{\phi[\cF_{j}]} (R'')|$
-
-%   \begin{enumerate}[a $|$]
-%   \item {\em None of the sets are new. $R, R', R'' \in \cF_{j-1}$.}
-%     We know $(\cF_{j-1}, \cl_{j-1})$ is feasible. Thus $|R \cap R'
-%     \cap R''| = |\cl_{j-1}(R) \cap \cl_{j-1}(R') \cap \cl_{j-1}(R'')|
-%     = |\cl_{j}(R) \cap \cl_{j}(R') \cap \cl_{j}(R'')|$.
-%   \item {\em Only one, say $R$, is a new set.}  Let $R = S_1 \cap S_2$
-%     ($S_1, S_2$ are defined in the proof of lemma
-%    ~\ref{lem:invar1}). Now we have $|R \cap R' \cap R''| = |S_1 \cap
-%     S_2 \cap R' \cap R''| = |\cl_{j-1}(S_1) \cap \cl_{j-1}(S_2) \cap
-%     \cl_{j-1}(R') \cap \cl_{j-1}(R'')| = |\cl_{j}(R) \cap \cl_{j}(R')
-%     \cap \cl_{j}(R'')|$. Thus proven. If $R$ is any of the other new
-%     sets, the same claim can be verified using lemma
-%    ~\ref{lem:setminuscard}.
-%     % \item []{\bf Case 3:}
-%   \item {\em At least two of $R, R', R''$ are new sets.}  The new sets
-%     are disjoint hence this case is vacuously true.
-%   \end{enumerate}
-%   \qed
-%   \tnote[E2]{remove the induction proof. just text saying x and v are
-%     exclusive in these sets therefore the intersection cardinalities
-%     don't change thus all invariants are still true} 
-\end{proof}
-
-As a result of Algorithm~\ref{perms} each leaf $v$ in $T$
-is such that there is exactly one set in $\cF$ with $v$ as a vertex in
-the path assigned to it.  In Algorithm~\ref{leafasgn} we identify
-elements in $supp(\cF)$ whose images are leaves in a hypergraph
-isomorphism if one exists.  Let $S \in \cF$ be such that $\cl(S)$ is a
-path with leaf and $v \in V(T)$ is the unique leaf incident on it.  We
-define a new path labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) =
-\set{v}$ where $x$ an arbitrary element from $S \setminus \bigcup_{\hS
-  \ne S} \hS$. In other words, $x$ is an element present in no other
-set in $\cF$ except $S$. This is intuitive since $v$ is present in no
-other path image under $\cl$ other than $\cl(S)$.  The element $x$ and
-leaf $v$ are then removed from the set $S$ and path $\cl(S)$
-respectively. After doing this for all leaves in $T$, all path images
-in the new path labeling $\cl_{new}$ except leaf labels (a path that
-has only a leaf is called the {\em leaf label} for the corresponding
-single element hyperedge or set) are paths from a new pruned tree $T_0
-= T \setminus \{v \mid v \text{ is a leaf in }
-T\}$. Algorithm~\ref{leafasgn} is now presented with details.
-
-\begin{algorithm}[h]
-  \caption{Leaf labeling from an ICPPL {\tt filter\_2($\cF, \cl, T$)}}
-  \label{leafasgn}
-  \begin{algorithmic}[\lndisplay]
-    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$
-    \COMMENT {Path images are such that no two path images share a
-      leaf.}
-    \STATE $j \assign 1$\\
-    \WHILE {there is a leaf $v$ in $T$ and a unique $S_1 \in
-      \cF_{j-1}$ such that $v \in \cl_{j-1}(S_1)$ }\label{uniqueleaf}
-    \STATE $\cF_j \assign \cF_{j-1} \setminus \{S_1\}$\\
-    \STATE for all $S \in \cF_{j-1}$ such that $S \ne S_1$ set
-    $\cl_j(S) \assign
-    \cl_{j-1}(S)$\\
-    \STATE $X \assign S_1 \setminus \bigcup_{S \in \cF_{j-1}, S \ne S_1}S$\\
-    \IF{$X$ is empty} \label{xempty} \STATE {\bf exit} \label{ln:exit2} \ENDIF
-    \STATE $x \assign $ arbitrary element from $X$\\
-    \STATE $\cF_j \assign \cF_j \cup \{\{x\}, S_1 \setminus \{x\}\} $\\
-    \STATE $\cl_j(\{x\}) \assign \{v\}$\\
-    \STATE $\cl_j(S_1 \setminus \{x\}) \assign \cl_{j-1}(S_1) \setminus \{v\}$\\
-    \STATE $j \assign j+1$\\
-    \ENDWHILE
-    \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
-    \RETURN $(\cF', \cl')$
-  \end{algorithmic}
-\end{algorithm}
-
-Suppose the input ICPPL $(\cF, \cl)$ is feasible, yet set $X$ in
-Algorithm~\ref{leafasgn} is empty in some iteration of the {\bf while}
-loop. This will abort our procedure of finding the hypergraph
-isomorphism. The following lemma shows that this will not happen.
-
-\begin{lemma}
-  \label{lem:xnotempty}
-  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{leafasgn} is
-  feasible, then for all iterations $j > 0$ of the {\em \bf while}
-  loop, the {\em \bf exit} statement in line~\ref{ln:exit2} does not
-  execute.
-\end{lemma}
-\begin{proof}
-  Assume $X$ is empty for some iteration $j > 0$. We know that $v$ is
-  an element of $\cl_{j-1}(S_1)$. Since it is uniquely present in
-  $\cl_{j-1}(S_1)$, it is clear that $v \in \cl_{j-1}(S_1) \setminus
-  \bigcup_{(S \in \cF_{j-1}) \wedge (S \ne S_1)}\cl_{j-1}(S)$.  Note
-  that for any $x \in S_1$ it is contained in at least two sets due to
-  our assumption about cardinality of $X$. Let $S_2 \in \cF_{j-1}$ be
-  another set that contains $x$. From the above argument, we know $v
-  \notin \cl_{j-1}(S_2)$. Therefore there cannot exist a hypergraph
-  isomorphism bijection that maps elements in $S_2$ to those in
-  $\cl_{j-1}(S_2)$. This contradicts our assumption that the input is
-  feasible. Thus $X$ cannot be empty if input is ICPPL and feasible.
-  \qed
-\end{proof}
-
-\begin{lemma}
-  \label{lem:invar3}
-  In Algorithm~\ref{leafasgn}, for all $j > 0$, at the end of the
-  $j$th iteration of the {\bf while} loop the four invariants given in
-  Lemma~\ref{lem:invar1} hold.
-\end{lemma}
-\begin{proof}
-  By Lemma~\ref{lem:xnotempty} we know that set $X$ will not be empty
-  in any iteration of the {\em \bf while} loop if input ICPPL $(\cF,
-  \cl)$ is feasible and $\cl_j$ is always computed for all $j >
-  0$. Also note that removing a leaf from any path keeps the new path
-  connected. Thus invariant I is obviously true. In every iteration $j
-  > 0$, we remove exactly one element $x$ from one set $S$ in $\cF$
-  and exactly one vertex $v$ which is a leaf from one path
-  $\cl_{j-1}(S)$ in $T$. This is because as seen in
-  Lemma~\ref{lem:xnotempty}, $x$ is exclusive to $S$ and $v$ is
-  exclusive to $\cl_{j-1}(S)$. Due to this fact, it is clear that the
-  intersection cardinality equations do not change, i.e., invariants
-  II, III, IV remain true. On the other hand, if the input ICPPL is
-  not feasible the invariants are vacuously true. \qed
-\end{proof}
-
-% \textcolor{cyan}{
-% \begin{lemma}
-%   \label{lem:notfeasibleexit}
-%   \tnote{IS THIS CORRECT?}  If input ICPPL $(\cF, \cl)$ is not
-%   feasible, then in one of the recursive calls to Algorithm --3--, the
-%   {\em \bf exit} statement in line x in Algorithm --1 or line y in
-%   Algorithm --2 will get executed.
-% \end{lemma}
-% }
-
-We have seen two filtering algorithms above, namely,
-Algorithm~\ref{perms} {\tt filter\_1} and Algorithm~\ref{leafasgn}
-{\tt filter\_2} which when executed serially repectively result in a
-new ICPPL on the same universe $U$ and tree $T$. We also proved that
-if the input is indeed feasible, these algorithms do indeed output the
-filtered ICPPL. Now we present the algorithmic characterization of a
-feasible tree path labeling by way of Algorithm~\ref{al:icppl-find-isomorph}.
-
-Algorithm~\ref{al:icppl-find-isomorph} computes a
-hypergraph isomorphism $\phi$ recursively using Algorithm~\ref{perms}
-and Algorithm~\ref{leafasgn} and pruning the leaves of the input
-tree. In brief, it is done as follows. Algorithm~\ref{leafasgn} gives
-us the leaf labels in $\cF_2$, i.e., the elements in $supp(\cF)$ that
-map to leaves in $T$, where $(\cF_2, \cl_2)$ is the output of
-Algorithm~\ref{leafasgn}. All leaves in $T$ are then pruned away. The
-leaf labels are removed from the path labeling $\cl_2$ and the
-corresponding elements are removed from the corresponding sets in
-$\cF_2$. This tree pruning algorithm is recursively called on the
-altered hypergraph $\cF'$, path label $\cl'$ and tree $T'$. The
-recursive call returns the bijection $\phi''$ for the rest of the
-elements in $supp(\cF)$ which along with the leaf labels $\phi'$ gives
-us the hypergraph isomorphism $\phi$.  The following lemma formalizes
-the characeterization of feasible path labeling.
-
-\begin{algorithm}[h]
-  \caption{{\tt get- hypergraph- isomorphism ($\cF, \cl, T$)}}
-  \label{al:icppl-find-isomorph}
-  \begin{algorithmic}[\lndisplay]
-
-    \IF{$T$ is empty}
-    \RETURN $\emptyset$\\
-    \ENDIF
-    \STATE $L \assign \{v \mid v \text{ is a leaf in }      T\}$\\
-    \STATE $(\cF_1, \cl_1) \assign$ {\tt filter\_1($\cF, \cl,
-      T$)}\\
-    \STATE $(\cF_2, \cl_2) \assign$ {\tt filter\_2($\cF_1,
-      \cl_1, T$)}\\
-
-    \STATE $(\cF', \cl') \assign (\cF_2, \cl_2)$\\
-    \STATE $\phi' \leftarrow \emptyset$
-
-    \FOR {every $v \in L$}
-    \STATE $\phi'(x) \assign v$ where $x \in \cl_2^{-1}(\{v\})$
-    \COMMENT {Copy the leaf labels to a one to one function $\phi':
-      supp(\cF) \rightarrow L$
-      }\\
-    \STATE Remove $\{x\}$ and $\{v\}$ from $\cF'$, $\cl'$  appropriately\\
-    \ENDFOR
-
-    \STATE $T' \assign T \setminus L$
-
-    \STATE $\phi'' \assign$ {\tt
-      get-hypergraph-isomorphism($\cF', \cl', T'$)}
-    \STATE $\phi \assign \phi'' \cup \phi'$ \\
-    \RETURN $\phi$
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma}
-  \label{lem:hyperiso}  %{lem:perm}
-  If $(\cF, \cl)$ is an ICPPL from a tree $T$ and
-  Algorithm~\ref{al:icppl-find-isomorph}, {\tt
-    get- hypergraph- isomorphism ($\cF, \cl, T$)} returns a non-empty
-  function, then there exists a hypergraph isomorphism $\phi :
-  supp(\cF) \rightarrow V(T)$ such that the $\phi$-induced tree path
-  labeling is equal to $\cl$ or $\cl_\phi = \cl$.
-\end{lemma}
-\begin{proof}
-  It is clear that in the end of every recursive call to
-  Algorithm~\ref{al:icppl-find-isomorph}, the function $\phi'$ is one
-  to one involving all the leaves in the tree passed as input to that
-  recursive call. Moreover, by Lemma~\ref{lem:noexit1} and
-  Lemma~\ref{lem:xnotempty} it is consistent with the tree path
-  labeling $\cl$ passed. The tree pruning is done by only removing
-  leaves in each call to the function and is done till the tree
-  becomes empty. Thus the returned function $\phi: supp(\cF)
-  \rightarrow V(T)$ is a union of mutually exclusive one to one
-  functions exhausting all vertices of the tree. In other words, it is
-  a bijection from $supp(\cF)$ to $V(T)$ inducing the given path
-  labeling $\cl$ and thus a hypergraph isomorphism. \qed
-\end{proof}
-
-\begin{theorem}
-  \label{th:charac}
-  A path labeling $(\cF, \cl)$ on tree $T$ is feasible iff it is an
-  ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
-  T)$ as input returns a non-empty function.
-\end{theorem}
-\begin{proof}
-  From Lemma~\ref{lem:hyperiso}, we know that if $(\cF, \cl)$ is an
-  ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
-  T)$ as input returns a non-empty function, $(\cF, \cl)$ is feasible.
-  Now consider the case where $(\cF, \cl)$ is feasible. i.e. there
-  exists a hypergraph isomorphism $\phi$ such that $\cl_\phi =
-  \cl$. Lemma~\ref{lem:noexit1} and Lemma~\ref{lem:xnotempty} show us
-  that filter 1 and filter 2 do not exit if input is feasible. Thus
-  Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
-  function.\qed
-\end{proof}
-
-\section{ Computing feasible TPL with special target trees \tnote{give
-    problem definition etc}}
-\label{sec:spltargettree}
-
-Section~\ref{sec:feasible} described properties that a TPL must have
-for it to be feasible. The next problem of interest is to test if a
-given hypergraph is a path hypergraph with respect to a given target
-tree\footnote{A larger problem would be to find if a given hypergraph
-  is a path graph on any tree. This problem is not addressed in this
-  thesis.}. In other words, the problem is to find out if a feasible
-tree path labeling exists from a given target tree for a given
-hypergraph. In this section we will see two special cases of this
-problem where the target tree is from a particular family of
-trees. The first one, where the tree is a path, is a shown to be
-equivalent to the well studied problem of consecutive-ones in
-Section~\ref{sec:icpplicpia}. The second one, where the tree is a
-$k$-subdivided tree\footnote{See Section~\ref{ch:prelims} for the
-  formal definition.}, has been solved using a polynomial time
-algorithm. The latter problem enforces some conditions on the
-hypergraph too which will be seen in Section~\ref{sec:ksubdivstar}.
-
-\subsection{Target tree is a Path}
-\label{sec:icpplicpia}
-
-Consider a special case of ICPPL with the following properties.
-\begin{enumerate}
-\item Given tree $T$ is a path. Hence, all path labels are interval
-  labels.
-\item Only pairwise intersection cardinality preservation is
-  sufficient. i.e. property (iii) in ICPPL is not enforced.
-\item The filter algorithms do not have {\em \bf exit} statements.
-\end{enumerate}
-This is called an Intersection Cardinality Preservation Interval
-Assignment (ICPIA) \cite{nsnrs09}. This structure and its algorithm is
-used in the next section for finding tree path labeling from a
-$k$-subdivided star due to this graph's close relationship with
-intervals.
-
-\subsection{Target tree is a $k$-subdivided Star}
-\label{sec:ksubdivstar}
-
-\begin{figure}[t] %[htbp] 
-  \centering
-  \begin{tabular}{lr}
-    (a) \includegraphics{cop-tree_thesis.1}&
-    (b) \includegraphics{cop-tree_thesis.2}
-  \end{tabular}
-  \caption{\figtabsize (a) $8$-subdivided star with 7 rays (b)
-    3-subdivided star with 3 rays}
-  \label{fig:kstar}
-\end{figure}
-
-In this section we consider the problem of assigning paths from a
-$k$-subdivided star $T$ to a given set system $\cF$.  We consider
-$\cF$ for which the overlap graph $\bO(\cF)$ is connected.  The
-overlap graph is well-known from the work of
-\cite{kklv10,nsnrs09,wlh02}.  We use the notation in
-\cite{kklv10}. Recall from Section~\ref{ch:prelims} that hyperedges
-$S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$
-and $S'$ have a non-empty intersection but neither of them is
-contained in the other. The overlap graph $\bO(\cF)$ is a graph in
-which the vertices correspond to the sets in $\cF$, and the vertices
-corresponding to the hyperedges $S$ and $S'$ are adjacent if and only
-if they overlap.  Note that the intersection graph of $\cF$,
-$\bI(\cF)$ is different from $\bO(\cF)$ and $\bO(\cF) \subseteq
-\bI(\cF)$.  A connected component of $\bO(\cF)$ is called an overlap
-component of $\cF$.  An interesting property of the overlap components
-is that any two distinct overlap components, say $\cO_1$ and $\cO_2$,
-are such that any two sets $S_1 \in \cO_1$ and $S_2 \in \cO_2$ are
-disjoint, or, w.l.o.g, all the sets in $\cO_1$ are contained within
-one set in $\cO_2$.  This containment relation naturally determines a
-decomposition of the overlap components into rooted containment trees.
-We consider the case when there is only one rooted containment tree,
-and we first present our algorithm when $\bO(\cF)$ is connected.  It
-is easy to see that once the path labeling to the overlap component in
-the root of the containment tree is achieved, the path labeling to the
-other overlap components in the rooted containment tree is essentially
-finding a path labeling when the target tree is a path: each target
-path is a path that is allocated to sets in the root overlap
-component.  Therefore, for the rest of this section, $\bO(\cF)$ is a
-connected graph. We also assume that all hyperedges are of cardinality
-at most $k+2$.
-
-Recall from Section~\ref{ch:prelims} that a $k$-subdivided star is a
-star with each edge subdivided $k$ times. Therefore, a $k$-subdivided
-star has a central vertex which we call the {\em root}, and each root
-to leaf path is called a {\em ray}. First, we observe that by removing
-the root $r$ from $T$, we get a collection of $p$ vertex disjoint
-paths of length $k+1$, $p$ being the number of leaves in $T$.  We
-denote the rays by $R_1, \ldots, R_p$, and the number of vertices in
-$R_i$, $i \in [p]$ is $k+2$.  Let $\seq{v_{i1},\ldots,v_{i(k+2)}=r}$
-denote the sequence of vertices in $R_i$, where $v_{i1}$ is the
-leaf. Note that $r$ is a common vertex to all $R_i$.
-
-In this section the given hypergraph, the $k$-subdivided star and the
-root of the star are denoted by $\cO$, $T$ and vertex $r$,
-respectively.
-% For each hyperedge $X \in \cO$, we will maintain a 2-tuple of non-negative
-% numbers $\seq{p_1(X), p_2(X)}$.  The numbers satisfy the property that
-% $p_1(X) + p_2(X) \leq |X|$, and at the end of path labeling, for each
-% $X$, $p_1(X) + p_2(X) = |X|$.  This signifies the algorithm tracking
-% the lengths of subpaths of the path assigned to $X$ from at most two
-% rays. We also maintain another parameter called the {\em residue} of
-% $X$ denoted by $s(X)=|X| - p_1(X)$. This signifies the residue path
-% length that must be assigned to $X$ which must be from another
-% ray. For instance, if $X$ is labeled a path from only one ray, then
-% $p_1(X) = |X|$, $p_2(X) = 0$ and $s(X) = 0$.
-
-% 
-% We iteratively consider each ray from which paths will be assigned to
-% hyperedges. At the beginning of each iteration hyperedges of $\cO$ are
-% classifed into the following disjoint sets.
-% \begin{enumerate}
-% \item [$\cL_1^i$] {\em Labeled without $r$.} Those that have been
-%   labeled with a path which does not contain $r$ in one of the
-%   previous iterations.\\  $\cL_1^i = \set{ X \mid p_1(X) = |X| \text{ and
-%     } p_2(X) = 0 \text{ and } s(X) = 0, X \in \cO}$
-% \item [$\cL_2^i$] {\em Labeled with $r$.} Those that have been labeled
-%   with two subpaths of $\cl(X)$ containing $r$ from two different rays
-%   in two previous iterations.\\ $\cL_2^i = \set{X \mid 0 < p_1\left(X\right),
-%     p_2\left(X\right) < |X| \text{ and } s(X) = 0, X \in \cO}$
-%   \item [$\cT_1^i$] {\em Type 1 / partially labeled.} Those that have
-%     been labeled with one path containing $r$ from a single ray in one
-%     of the previous iterations. Here, $p_1(X)$ denotes the length of
-%     the subpath of $\cl(X)$ that $X$ has been so far labeled
-%     with. Also, it is clear that such a path must start at $r$ and
-%     must be in a ray different from the one corresponding to $p_1(X)$.\\
-%     $\cT_1^i = \set{ X \mid 0 < p_1(X) < |X| \text{ and } p_2(X) = 0
-%       \text{ and } s(X) > 0, X \in \cO}$
-%   \item [$\cT_2^i$] {\em Type 2 / not labeled.} Those that have not been
-%     labeled with a path in any previous iteration.\\
-%     $\cT_2^i = \set{ X \mid p_1(X) = p_2(X) = 0 \text{ and } s(X) = |X|,
-%       X \in \cO}$
-% \end{enumerate}
-% \vspace{-2mm}
-% \begin{align*}
-%   \cO &= \cL_1^i \cup \cL_2^i \cup \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]\\
-%   \cO_i &= \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]
-% \end{align*}
-
-The set $\cO_i$ refers to the set of hyperedges $\cT_1^i \cup \cT_2^i$
-in the $i$th iteration.  Note that $\cO_1 = \cO$.  In the $i$th
-iteration, hyperedges from $\cO_i$ are assigned paths from $R_i$ using
-the following rules. Also the end of the iteration, $\cL_1^{i+1},
-\cL_2^{i+1}, \cT_1^{i+1}, \cT_2^{i+1}$ are set to $\cL_1^{i},
-\cL_2^{i}, \cT_1^{i}, \cT_2^{i}$ respectively, along with some
-case-specific changes mentioned in the rules below.
-
-\noindent
-Suppose the given hypergraph has a feasible path labeling to the given
-tree $T$. Let this labeling be $\cl$. Now we present an algorithm that
-lists the sequence of edges that maintain a set of properties
-necessary for them to have a feasible tree path labeling.
-\begin{enumerate}[I.]
-\item {\bf Step 1}
-  \begin{enumerate}
-  \item {\bf {There are type 1
-        edges, $\mathbf{|\cT_1^i| > 0}$}:} 
-% , and for each such $X$ assign the
-%     path in $R_i$ of length $s(X)$ starting at $v_{i,k+2}=r$.  $X$ is
-%     then removed from $\cO_i$, and $p_2(X)=s(X)$, and $s(X)=0$.If any
-%     intersection cardinality is violated because of this assignment,
-%     EXIT reporting the non-existence of a feasible assignment.
-  {If there is only one $\cT_1^i$ hyperedge, label it to the unique
-    path in $R_i$ of length $s(X)$ starting at $v_{i(k+2)}=r$. If
-    there are more than one $\cT_1^i$ hyperedges, for any pair of
-    hyperedges $S_1, S_2 \in \cT_1^i$, one of the following is
-    true. Let $\cl'(X) \subseteq R_j$ denote the subpath of length
-    $p_1(X)$ that has been assigned to $X$ from a previously
-    considered ray $R_j$.}
-  \begin{enumerate}[{\em Case }1. ]
-  \item $|\cl'(S_1) \cap \cl'(S_2)| = 1$ and $|S_1 \cap S_2| > 1$
-    i.e. they are in different rays and their residues must contain
-    their intersection. Hence assign both to $R_i$.
-  \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| >
-    |\cl'(S_1) \cap \cl'(S_2)|$ i.e. they are in the same rays and
-    their residues must contain their intersection. Hence assign both
-    to $R_i$.
-  \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| =
-    |\cl'(S_1) \cap \cl'(S_2)|$. i.e. the intersection cardinality has
-    been met. Hence assign only $S_1$ to $R_i$ and $S_2$ must be
-    assigned to some $i \ge p$.
-  \end{enumerate}
-  Remove the labeled edge or edges from $\cO_i$ and add it to
-  $\cO_{i+1}$. Also do not add it to $\cT_1^{i+1}$ and add to
-  $\cL_2^{i+1}$.
-\item {\bf {There are no type-1 edges nor} super-marginal type-2
-    hyperedges:} If all type-2 edges of length at most $k+1$, EXIT by
-  saying there is no ICPPL.  Otherwise, Find an $X$ of type-2 such
-  that $|X| \geq k+2$, assign it the unique path starting at
-  $v_{i(k+1)}=r$ of length $k+2$, and set $p_1(X)=k+2,
-  s(X)=|X|-(k+2)$, assign it the unique path starting at
-  $v_{i(k+1)}=r$ of length $k+2$, mark this as a type-1 hyperedge, and
-  add it to $\cO_{i+1}$ after removing it from $\cO_i$.
-  % \remove{Next, iteratively, consider each type-1 edge $X \in
-  %   \cO_i$, and if it intersects a previously assigned $Y$ such that
-  %   $\cl(Y) \subseteq R_i$, assign a path of length $s(X)$
-  %   containing $v_{i(k+2)}=r$, set $p_2(X)=s(X), s(X)=0$, and remove
-  %   $X$ from $\cO_i$.}\tnote{covered in step 2}
-
-\item {\bf There are super-marginal type-2 hyperedges {and no type-1
-      edges:}} Let $X$ be a super-marginal hyperedge. If $|X| \leq
-  k+2$, then $X$ is assigned the path of length $|X|$ starting at
-  $v_{i1}$, and $p_1(X) = |X|, p_2(X)=0$. $X$ is removed from $\cO_i$
-  and not added to $\cO_{i+1}$.  If $|X| > k+2$, then $p_1(X)=k+2$,
-  assign it the unique path starting at $v_{i(k+1)}=r$ of length $k+2$
-  and $s(X) = |X| - (k+2)$.  In this case, $X$ is classified as a
-  Type-1 edge, removed from $\cO_i$, and added to $\cO_{i+1}$.
-\end{enumerate}
-\item {\bf Step 2:} Iteratively, a hyperedge $X$ is selected from
-  $\cO_i$ that has an overlap with one of the hyperedges $Y$ such that
-  $\cl(Y) \subseteq R_i$, and a unique path is assigned to $X$.  The
-  path, say $U(X) \subseteq R_i$, that is assigned can be decided
-  unambiguously since the $X \overlap Y$, and all intersection
-  cardinalities can be preserved only at one of the ends of $\cl(Y)$.
-
-  \noindent
-  Let $\cl(X)$ denote the unique path assigned to $X$.  If $X$ is a
-  type-2 hyperedge and if the unique path of length $|X|$ does not
-  contain $r$, then $p_1(X) = |X|, p_2(X)=0$, and $X$ is removed from
-  $\cO_i$.  In the case when $\cl(X)$ has to contain $r$, then
-  $p_1(X)$ is the length of the path, $p_2(X)=0$, and $s(X) =
-  |X|-p_1(X)$. Further $X$ is classified as a type-1 hyperedge, added
-  to $\cO_{i+1}$ and removed from $\cO_i$.  In the case when $X$ is a
-  type-1 hyperedge, then we check if $U(X)$, which is of length $s(X)$
-  contains $r$. If it does, then we assign $\cl(X) \leftarrow \cl(X)
-  \cup U(X)$, remove $X$ from $\cO_i$ and do not add it to
-  $\cO_{i+1}$.  If not, then we {\bf Exit} reporting that an
-  assignment cannot be found.  The iteration ends when no hyperedge in
-  $\cO_i$ has an overlap with a hyperedge assigned to $R_i$.
-\end{enumerate}
-
-\noindent
-The order in which they are assigned is the sequence of hyperedges
-with the following properties. Note that if such a sequence cannot be
-computed, $\cO$ cannot have a feasible path labeling from $T$.
-
-\noindent
-\tnote{need to formalize the below properties} In the following
-lemmas we identify a set of necessary conditions for $\cF$ to have an
-ICPPL in the $k$-subdivided star $T$.  If during the execution of the
-above algorithm, one of these necessary conditions is not satisfied,
-the algorithm exits reporting the non-existence of an ICPPL.
-\begin{lemma}
-  Let all hyperedges in $\cO_i$ be type-1 edges.  Then there is a
-  maximal subset $\ccT_i \subseteq \cO_i$ with the following
-  properties:
-  \begin{enumerate}
-  \item $\ccT_i$ form an inclusion chain.
-  \item For all $X \in \ccT_i$, $s(X) \leq k+2$, and There is a an $X
-    \in \ccT_i$ such that $s(X)=k+2$.
-  \end{enumerate}
-\end{lemma}
-\begin{lemma}
-  If there are no \marginal type-2 edges in $\cO_i$, then there exists
-  at least $p-i$ type-2 hyperedges $X \in \cO_i$ such that $|X| \geq
-  k+2$.
-  % is a exist type-2 hyperedges of length at least $k+2$, and there
-  % must be at least $r-i$ such type-2 hyperedge, basically at least
-  % one per ray.  Each hyperedge contained in such an hyperedges of
-  % length at least $k+2$, is transitively related under the overlap
-  % relation to one that gets assigned a path that contains
-  % $r$. Basically, the proof will argue that otherwise, there will be
-  % a \marginal type-2 hyperedge, or that it is infeasible.  Also, it
-  % is here we assume that we assume that the size of the two
-  % universes is the same.
-\end{lemma}
-\begin{lemma}
-  At the end of Step-1 in the $i$-th iteration, if one hyperedge $X$
-  of type-2 is such that $\cl(X) \subseteq R_i$, then all other
-  hyperedges in $\cO_i$ are connected to $X$ in the overlap component.
-\end{lemma}
-\begin{lemma}
-  At the end of Step-2, If control has exit at any time, there is no
-  ICPPL. If control has not exit, then $R_i$ is saturated.  No
-  hyperedge of $\cO_{i+1}$ will get a path from $R_i$ in the future
-  iterations.  No type-2 hyperedge of $\cO_{i+1}$ will get a path from
-  $R_i$.  Basically $R_i$ is done.
-\end{lemma}
-\begin{lemma}
-  Finally, we need to prove that the assignment is an ICPPL. Secondly,
-  if there is a permutation then maps sets to paths, then it is indeed
-  an ICPPL, and our algorithm will basically find it.  It is a unique
-  assignment upto permutation of the leaves.
-\end{lemma}
-
-\begin{lemma}
-  \label{lem:sup-mar}
-  If $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
-  tree path labeling to tree $T$, then $\cl(X)$ will
-  contain a leaf in $T$.\\
-  On the otherhand, if $\cl(X)$ has a leaf in
-  $T$, $X$ is marginal but may not be super-marginal.
-\end{lemma}
-\begin{proof}
-  Suppose $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
-  path labeling from $T$.  Assume $\cl(X)$ does not have a leaf.  Let
-  $R_i$ be one of the rays (or the only ray) $\cl(X)$ is part of.
-  Since $X$ is in a connected overlap component, there exists $Y_1 \in
-  \cF$ and $X \nsubseteq Y_1$ such that $Y_1 \overlap X$ and $Y_1$ has
-  at least one vertex closer to the leaf in $R_i$ than any vertex in
-  $X$. Similarly with the same argument there exists $Y_2 \in \cF$
-  with same subset and overlap relation with $X$ except it has has at
-  least one vertex farther away from the leaf in $R_i$ than any vertex
-  in $X$. Clearly $Y_1 \cap X$ and $Y_2 \cap X$ cannot be part of same
-  inclusion chain which contradicts that assumption $X$ is
-  super-marginal. Thus the claim is proven.\qed
-\end{proof}
-
-\noindent
-Consider the overlap graph $\bO(\cF)$ of the given hypergraph
-$\cF$. Let $S_{sm} \in \cF$ be such that it is a super-marginal
-hyperedge.  Algorithm~\ref{al:ktree-label} uses $S_{sm}$ along with
-the overlap graph $\bO(\cF)$ to calculate the feasible tree path
-labeling to the $k$-subdivded tree $T$.
-
-\begin{algorithm}[h]
-  \caption{{\tt compute-ksubtree-path-labeling($X, \cF, T$)}}
-  \label{al:ktree-label}
-  \begin{algorithmic}[\lndisplay]
-    \IF{$X = S_{sm}$}
-    \STATE -- TBD --\\
-    \ELSE 
-    \STATE -- TBD --\\
-    \ENDIF
-
- \end{algorithmic}
-\end{algorithm}
-
-\section{ TPL with no restrictions}
-\label{sec:norestraint}
-
-{\tt abstract begin}
-  it is
-  known that if the given tree is a path a feasible assignment can be
-  found in polynomial time, and we observe that it can actually be
-  done in logspace. \tnote{[TRUE?]}
-{\tt abstract end}
-
-In Section \ref{prelims} we present the
-necessary preliminaries, in Section \ref{feasible} we present our
-characterization of feasible tree path assignments, and in Section
-\ref{decompos} we present the characterizing subproblems for finding a
-bijection between $U$ and $V(T)$ such that sets map to tree
-paths. Finally, in Section \ref{complexity} we conclude by showing
-that Tree Path Assignment is GI-Complete, and also observe that
-Consecutive Ones Testing is in Logspace.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC \subsection{Introduction}
-
-\annote{ We refer to this as the Tree Path Assignment
-problem for an input $(\cF,T)$ pair.}{Add this ``pair'' as a
-  definition in ch:prelims}
-
-\remove{It is an interesting fact that for a matrix with the COP, the
-intersection graph of the corresponding set system is an interval
-graph.  A similar connection to a subclass of chordal graphs, and this
-subclass contains interval graphs, exists for the generalization of
-COP.  In this case, the intersection graph of the corresponding set
-system must be a path graph. Chordal graphs are of great
-significance, extensively studied, and have several applications.  One
-of the well known and interesting properties of a chordal graphs is
-its connection with intersection graphs cite{mcg04}. For every chordal
-graph, there exists a tree and a family of subtrees of this tree such
-that the intersection graph of this family is isomorphic to the
-chordal graph cite{plr70,gav78,bp93}.  Certain format of these trees
-are called clique trees cite{apy92} of the graph which is a compact
-representation of the chordal graph. There has also been work done on
-the generalization of clique trees to clique hypergraphs cite{km02}.
-If the chordal graph can be represented as the intersection graph of
-paths in a tree, then the graph is called path graph cite{mcg04} or
-path chordal graph.  Therefore, it is clear that if there is a
-bijection from $U$ to $V(T)$ such that the sets map to paths, then the
-intersection graph of the set system is indeed a path chordal graph.
-This is, however, only a necessary condition and can be checked
-efficiently, as path graph recognition is polynomial
-solvable cite{gav78,aas93}.  Indeed, it is possible to construct a set
-system and tree, such that the intersection graph is a path chordal
-graph, but there is no bijection between $U$ and $V(T)$ such that the
-sets map to paths.  This connection indeed suggests that our problem
-is indeed as hard as path graph isomorphism.  Further path graph
-isomorphism is known be isomorphism-complete, see for example
-cite{kklv10}. }
-In the second part of this paper, we decompose our
-search for a bijection between $U$ and $V(T)$ into subproblems.  Each
-subproblem is on a set system in which for each set, there is another
-set in the set system with which the intersection is {\em strict}-
-there is a non-empty intersection, but neither is contained in the
-other.  This is in the spirit of results in \cite{wlh02,nsnrs09} where
-to test for COP in a given matrix, the COP problem is solved on an
-equivalent set of prime matrices.  \annote{Our decomposition localizes the
-challenge of path graph isomorphism to two problems.}{HOW?}
-
-Finally, we show that Tree Path Assignment is isomorphism-complete.
-We also point out Consecutive Ones Testing is in Logspace from two
-different results in the literature \cite{kklv10, mcc04}. To the best
-of our knowledge this observation has not been made earlier.
-
-% \noindent
-% It has been long known that interval graph recognition is in
-% logspace\cite{rei84}. Recently interval graph isomorphism was also
-% shown to be logspace decidable using a logspace canonization
-% algorithm by \cite{kklv10}.  This result is built on top of logspace
-% results of undirected graph connectivity \cite{rei08}, logspace
-% tractability using a certain logical formalism called FP+C and
-% modular decomposition of interval graphs\cite{lau10} etc.  Interval
-% graphs are closely connected to binary matrices with COP. The
-% maximal clique vertex incidence matrix (matrix with rows
-% representing maximal cliques and columns representing vertices of a
-% graph) has COP on columns iff the graph is an interval
-% graph\cite{fg65}. This follows from the interval graph
-% characterization by \cite{gh64}. Due to this close relation it is
-% natural to see if consecutive ones property can be
-% tested in logspace. \\
-% \noindent
-% We also explore some extensions of the interval assignment problem
-% in \cite{nsnrs09}, namely caterpillar path assignment problem.
-
-% We present a logspace algorithm here that uses the ICPIA
-% characterization of binary matrices with COP (set system associated
-% with such a matrix)\cite{nsnrs09}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC\subsection{Preliminaries} 
-\label{prelims} 
-In this paper, the collection
-$\F = \{S_i \mid S_i \subseteq U, S_i \ne \O, i \in I\}$ is a {\em set
-  system} of universe, $U = \{1, \ldots, n\}$. Moreoever, a set system
-is assumed to ``cover'' the universe,
-i.e. $ \bigcup_{i \in I}S_i = U$. \\
-
-\noindent
-The graph $T=(V,E)$ represents a given tree with $|V| = n$. Further,
-for simplicity, $V$ is defined as $\{1,\ldots,n\}$. All paths
-referred to in this paper are paths of $T$ unless explicitly specified. \\
-
-\noindent
-A {\em path assignment} $\A$ to $\F$ is defined as a set assignment
-where second universe is the vertex set $V$ of a given tree $T$ and
-every second subset in the ordered pairs is a path in this
-tree. Formally, the definition is as follows.
-\begin{align*}
-  \A = \{ (S_i,P_i) \mid S_i \in \F, P_i \subseteq V \text{
-    s.t. }T[P_i] \text{ is a path, } i \in I \}
-\end{align*}
-In other words, $P_i$ is the path on the tree $T$ assigned to $S_i$ in
-$\A$. As mentioned before for set systems, the paths cover the whole
-tree, i.e. $\bigcup_{i \in I}P_i = V$ \\
-
-\noindent
-Generalizing the definition of {\em feasibility} in \cite{nsnrs09} to
-a set assignment, a path assignment $\A$ is defined to be {\em
-  feasible} if there exists a bijection defined as follows.
-\begin{align}
-  \sigma: U \rightarrow V(T), \text{ such that }\sigma(S_i) = P_i
-  \text{ for all } i \in I, \sigma \text{ is a bijection}
-  \label{eq:stf}
-\end{align}
-% The assignment $\A$ is then called {\em Set Translation Feasible}.
-
-\noindent
-Let $X$ be a partially ordered set with $\preccurlyeq$ being the
-partial order on $X$.  $mub(X)$ represents an element in $X$ which is
-a maximal upper bound on $X$.  $X_m \in X$ is a maximal upper bound of
-$X$ if $\nexists X_q \in X$ such that $X_m
-\preccurlyeq X_q$. \\
-
-\noindent
-The set $I$ represents the index set $[m]$. If index $i$ is used
-without further qualification, it is meant to be $i \in I$. Any
-function, if not defined on a domain of sets, when applied on a set is
-understood as the function applied to each of its elements. i.e. for
-any function $f$ defined with domain $U$, the abuse of notation is as
-follows; $f(S)$ is used instead of $\hat f(S)$ where
-$\hat f(S) = \{y \mid y = f(x), x \in S\}$. \\
-
-\noindent
-When refering to a tree as $T$ it could be a reference to the tree
-itself, or the vertices of the tree. This will be clear from the
-context.\\
-
-\noindent
-Finally, an in-tree is a directed rooted tree in which all edges are
-directed toward to the root.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC\subsection{Characerization of Feasible Tree Path
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC  Assignment\tnote{Merge with the original charac section}} 
-\label{feasible} 
-Consider a path assignment $\cA =
-\{(S_i, P_i) \mid S_i \in \cF, P_i \text{ is a path from $T$}, i \in
-[m]\}$ to a set system $\cF = \{S_i \mid S_i \subseteq U, i \in
-[m]\}$, were $T$ is a given tree, $U$ is the set system's universe and
-$m$ is the number of sets in $\cF$. We call $\cA$ an {\em Intersection
-  Cardinality Preserving Path Assignment (ICPPA)} if it has the
-following properties.
-
-\begin{enumerate}
-\item [i.]  $|S_i| = |P_i|$ for all $i \in [m]$
-\item [ii.] $|S_i \cap S_j| = |P_i \cap P_j|$ for all $i,j \in [m]$
-  % \item [iii.] $|\bigcup_{i \in J} S_i| = |\bigcup_{i \in J} P_i|$
-  %   for all $J \subseteq I$\footnote{TBD: this is needed now to fix
-  %     some proofs}
-\item [iii.] $|S_i \cap S_j \cap S_k| = |P_i \cap P_j \cap P_k|$ for
-  all $i,j,k \in [m]$
-\end{enumerate}
-
-\begin{lemma}
-  \label{lem:setminuscard}
-  If $\cA$ is an ICPPA, and $(S_1, P_1),(S_2, P_2),(S_3, P_3) \in
-  \cA$, then $|S_1 \cap (S_2 \setminus S_3)| = |P_1 \cap (P_2
-  \setminus P_3)|$.
-\end{lemma}
-\begin{proof}
-  $|S_1 \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| =
-  |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to conditions (ii)
-  and (iii) of ICPPA, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
-  \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
-  = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. \qed
-\end{proof}
-
-\begin{lemma}
-  \label{lem:fourpaths} Consider four paths in a tree $Q_1, Q_2, Q_3,
-  Q_4$ such that they have nonempty pairwise intersection and $Q_1,
-  Q_2$ share a leaf. Then there exists distinct $i, j, k \in
-  \{1,2,3,4\}$ such that, $Q_1 \cap Q_2 \cap Q_3 \cap Q_4 = Q_i \cap
-  Q_j \cap Q_k$.
-\end{lemma}
-\begin{proof}
-  {\em Case 1:} w.l.o.g, consider $Q_3 \cap Q_4$ and let us call it
-  $Q$. This is clearly a path (intersection of two paths is a path).
-  % Since $Q_1, Q_2$ share a leaf, the following are paths $Q_1
-  % \setminus Q_2$, $Q_2 \setminus Q_1$, $Q_1 \cap Q_2$ and they are
-  % mutually disjoint.
-  Suppose $Q$ does not intersect with $Q_1 \setminus Q_2$, i.e. $Q
-  \cap (Q_1 \setminus Q_2) = \O$. Then $Q \cap Q_1 \cap Q_2 = Q \cap
-  Q_2$. Similarly, if $Q \cap (Q_2 \setminus Q_1) = \O$, $Q \cap Q_1
-  \cap Q_2 = Q \cap Q_1$. Thus it is clear that if the intersection of
-  any two paths does not intersect with any of the set differences of
-  the remaining two paths, the claim in the lemma is true.
-  % Note that $Q_1 \setminus Q_2$ and $Q_2
-  % \setminus Q_1$ are paths because $Q_1, Q_2$ share a leaf.\\
-  {\em Case 2:} Let us consider the compliment of the previous
-  case. i.e. the intersection of any two paths intersects with both
-  the set differences of the other two. First let us consider $Q \cap
-  (Q_1 \setminus Q_2) \ne \O$ and $Q \cap (Q_1 \setminus Q_2) \ne \O$,
-  where $Q = Q_3 \cap Q_4$. Since $Q_1$ and $Q_2$ share a leaf, there
-  is exactly one vertex at which they branch off from the path $Q_1
-  \cap Q_2$ into two paths $Q_1 \setminus Q_2$ and $Q_2 \setminus
-  Q_1$. Let this vertex be $v$. It is clear that if path $Q_3 \cap
-  Q_4$, must intersect with paths $Q_1 \setminus Q_2$ and $Q_2
-  \setminus Q_1$, it must contain $v$ since these are paths from a
-  tree. Moreover, $Q_3 \cap Q_4$ intersects with $Q_1 \cap Q_2$ at
-  exactly $v$ and only at $v$ which means that $Q_1 \cap Q_2$ does not
-  intersect with $Q_3 \setminus Q_4$ or $Q_4 \setminus Q_3$ which
-  contradicts initial condition of this case. Thus this
-  case cannot occur and case 1 is the only possible scenario. \\
-  Thus lemma is proven \qed
-\end{proof}
-
-\noindent
-In the remaining part of this section we show that a path assignment
-is feasible if and only if it is an ICPPA.  One direction of this
-claim is clear: that is a path assignment is feasible, then all
-intersection cardinalities are preserved, that is the path assignment
-is an ICPPA.  The reason is that a feasible path assignment has an
-associated bijection between $U$ and $V(T)$ such that the sets map to
-paths.  The rest of the section is devoted to constructively proving
-that it is sufficient, for the existence of a bijection, for a path
-assignment to be an ICPPA.  At a top-level, the constructive
-approaches refine the path assignment iteratively, such that at the
-end of each iteration we have a path assignment, and finally we have a
-family of bijections.  First we present and then prove the correctness
-of Algorithm \ref{perms}.  This algorithm refines the path assignment
-by considering pairs of paths that share a leaf.
-
-\begin{algorithm}[h]
-  \caption{Permutations from an ICPPA $\{(S_i,P_i) | i \in I\}$}
-  \label{perms}
-  \begin{algorithmic}
-    \STATE Let $\Pi_0=\{(S_i,P_i)| i \in I\}$\\
-    \STATE $j = 1$;
-    \label{shareleaf} \WHILE {There is $(P_1,Q_1), (P_2,Q_2) \in
-      \Pi_{j-1}$ with $Q_1$ and $Q_2$ having a common leaf} \STATE
-    $\Pi_j= \Pi_{j-1} \setminus \{(P_1,Q_1),(P_2,Q_2)\}$;
-    \label{setbreak}\STATE $\Pi_j = \Pi_j \cup \{(P_1 \cap P_2,Q_1
-    \cap Q_2), (P_1 \setminus P_2,Q_1 \setminus Q_2), (P_2 \setminus
-    P_1, Q_2 \setminus Q_1)\}$; \STATE $j = j+1$; \ENDWHILE \STATE
-    $\Pi = \Pi_j$; \STATE Return $\Pi$;
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma}
-  \label{lem:invar1}
-  In Algorithm \ref{perms}, at the end of $j$th iteration, $j \ge 0$,
-  of the while loop of Algorithm \ref{perms}, the following invariants
-  are maintained.
-  \begin{itemize}
-  \item {\em Invariant I:} $Q$ is a path in $T$ for each $(P,Q) \in
-    \Pi_j$
-  \item {\em Invariant II:} $|P|=|Q|$ for each $(P,Q) \in \Pi_j$
-  \item {\em Invariant III:} For any two $(P,Q), (P',Q') \in \Pi_j$,
-    $|P' \cap P''|=|Q' \cap Q''|$.
-  \item {\em Invariant IV:} For any three, $(P',Q'), (P'',Q''), (P, Q)
-    \in \Pi_j$, $|P' \cap P'' \cap P|=|Q' \cap Q'' \cap Q|$.
-  \end{itemize}
-\end{lemma}
-\begin{proof}
-  Proof is by induction on the number of iterations, $j$. In the rest
-  of the proof, the term ``new sets'' will refer to the new sets added
-  in $j$th iteration as defined in line \ref{setbreak} of Algorithm
-  \ref{perms}, i.e. the following three assignment pairs for some
-  $(P_1,Q_1), (P_2,Q_2) \in \Pi_{j-1}$ where $Q_1$ and $Q_2$ intersect
-  and share a leaf: $(P_1 \cap P_2, Q_1 \cap Q_2)$, or $(P_1 \setminus
-  P_2, Q_1 \setminus Q_2)$, or $(P_2 \setminus P_1, Q_2
-  \setminus Q_1)$.\\
-  \noindent
-  The base case, $\Pi_0 = \{(S_i,P_i) \mid i \in [m]\}$, is trivially
-  true since it is the input which is an ICPPA.  Assume the lemma is
-  true till the $j-1$ iteration. Consider $j$th
-  iteration:\\
-  \noindent
-  If $(P,Q)$, $(P',Q')$ and $(P'',Q'')$ are in $\Pi_{j}$ and
-  $\Pi_{j-1}$, all the invariants are
-  clearly true because they are from $j-1$ iteration.\\
-  If $(P,Q)$ is in $\Pi_{j}$ and not in $\Pi_{j-1}$, then it must be
-  one of the new sets added in $\Pi_j$. Since $(P_1,Q_1)$ and
-  $(P_2,Q_2)$ are from $\Pi_{j-1}$ and $Q_1,Q_2$ intersect and have a
-  common leaf, it can be verified that the
-  new sets are also paths. \\
-  By hypothesis for invariant III, invariant II also holds for $(P,Q)$
-  no matter which new set in $\Pi_j$ it
-  is.\\
-  To prove invariant III, if $(P,Q)$ and $(P',Q')$ are not in
-  $\Pi_{j-1}$, then they are both new sets and invariant III holds
-  trivially (new sets are disjoint). Next consider $(P,Q), (P',Q') \in
-  \Pi_j$ with only one of them, say $(P',Q')$, in $\Pi_{j-1}$. Then
-  $(P,Q)$ is one of the new sets added in line \ref{setbreak}. It is
-  easy to see that if $(P,Q)$ is $(P_1 \cap P_2, Q_1 \cap Q_2)$, then
-  due to invariant IV in hypothesis, invariant III becomes true in
-  this iteration. Similarly, using lemma \ref{lem:setminuscard}
-  invariant III is proven if $(P, Q)$ is $(P_1 \setminus P_2, Q_1
-  \setminus Q_2)$, or $(P_2 \setminus P_1, Q_2
-  \setminus Q_1)$.\\
-  To prove invariant IV, consider three assignments
-  $(P,Q),(P',Q'),(P'',Q'')$. If at least two of these pairs are in not
-  $\Pi_{j-1}$, then they are any two of the new sets. Note that these
-  new sets are disjoint and hence if $(P',Q'), (P'',Q'')$ are any of
-  these sets, $|P \cap P' \cap P''|=|Q \cap Q' \cap Q''|=0$ and
-  invariant IV is true. Now we consider the case if at most one of
-  $(P,Q),(P',Q'),(P'',Q'')$ is not in $\Pi_{j-1}$. If none of them are
-  not in $\Pi_{j-1}$ (i.e. all of them are in $\Pi_{j-1}$), invariant
-  IV is clearly true. Consider the case where exactly one of them is
-  not in $\Pi_{j-1}$. w.l.o.g let that be $(P,Q)$ and it could be any
-  one of the new sets. If $(P,Q)$ is $(P_1 \cap P_2, Q_1 \cap Q_2)$,
-  from lemma \ref{lem:fourpaths} and invariant III hypothesis,
-  invariant IV is proven. Similarly if $(P,Q)$ is any of the other new
-  sets, invariant IV is proven by also using lemma
-  \ref{lem:setminuscard}. \qed
-
-\end{proof}
-
-\noindent
-It can be observed that the output of algorithm \ref{perms} is such
-that every leaf is incident on at most a single path in the new set of
-assignments. This is due to the loop condition at line
-\ref{shareleaf}. Let $v_1$ be the leaf incident on path $P_i$. Assign
-to it any one element from $S_i \setminus \bigcup_{i \ne j}
-S_j$. Remove $(S_i, P_i)$ from assignments and add $(\{x_1\},
-\{v_1\}), (S_i \setminus \{x_1\}, P_i \setminus \{v_1\})$. Now all
-assignments except single leaf assignments are paths from the subtree
-$T_0 = T \setminus \{v \mid v \text{ is a leaf in } T\}$.
-
-% exclusively on the spine of the caterpillar. Consider the spine to
-% be an interval and run ICPIA algorithm on it to get the required
-% permutation. Algorithm \ref{leafasgn} makes this clear.
-
-\begin{algorithm}[h]
-  \caption{Leaf assignments from an ICPPA $\{(S_i,P_i) | i \in I\}$}
-  \label{leafasgn}
-  \begin{algorithmic}
-    \STATE Let $\Pi_0=\{(S_i,P_i)| i \in [m]\}$. Paths are such that
-    no
-    two paths $P_i, P_j, i \ne j$ share a leaf.\\
-    \STATE $j = 1$\\
-    \WHILE {there is a leaf $v$ and a unique $(S_{i_1}, P_{i_1})$ such
-      that $v \in P_{i_1}$}
-    \STATE $\Pi_j=   \Pi_{j-1} \setminus \{(S_{i_1}, P_{i_1})\}$\\
-    \STATE $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in I}S_i$
-    \IF{$X$ is empty} \STATE exit \ENDIF; \STATE Let $x = $ arbitrary
-    element from $X$ \vspace{2mm}
-
-    \STATE $\Pi_j = \Pi_j \cup \{(S_{i_1} \setminus \{x\}),(P_{i_1}
-    \setminus
-    \{v\}), (\{x\},\{v\})\}$\\
-    \STATE $j = j+1$\\
-    \ENDWHILE
-    \STATE $\Pi = \Pi_j$\\
-    \STATE Return $\Pi$\\
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma}
-  \label{lem:invar3}
-  In Algorithm \ref{leafasgn},for all $j \geq 0$, at the end of the
-  $j$th iteration the four invariants given in lemma \ref{lem:invar1}
-  are valid.
-  % Moreover, $X$ as defined in the algorithm is non-empty if this is
-  % an ICPPA.
-\end{lemma}
-\begin{proof}
-  First we see that $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in
-    I}S_i$ is non empty in every iteration for an ICPPA. Suppose $X$
-  is empty. We know that $v \in P_{i_1} \setminus \bigcup_{i \ne i_1,
-    i \in I}P_i$ since $v$ is in the unique path $P_{i_1}$. Since this
-  is an ICPPA $|S_{i_1}| = |P_{i_1}|$. For any $x \in S_{i_1}$ it is
-  contained in at least two sets due to our assumption. Let $S_{i_2}$
-  be a second set that contains $x$. We know $v \notin
-  P_{i_2}$. Therefore there cannot exist a permutation that maps
-  elements of $S_{i_2}$ to $P_{i_2}$. This contradicts our assumption
-  that this is an ICPPA. Thus $X$ cannot be empty.
-
-  We use mathematical induction on the number of iterations for this
-  proof. The term ``new sets'' will refer to the sets added in $\Pi_j$
-  in the $j$th iteration, i.e. $(P' \setminus \{x\},Q' \setminus
-  \{v\})$ and $(\{x\},\{v\})$ for some $(P',Q')$ in $\Pi_{j-1}$ such
-  that $v$ is a leaf and $Q'$ is the unique path
-  incident on it.\\
-  For $\Pi_0$ all invariants hold because it is output from algorithm
-  \ref{perms} which is an ICPPA. Hence base case is proved.  Assume
-  the lemma holds for $\Pi_{j-1}$. Consider $\Pi_j$ and any $(P,Q) \in
-  \Pi_j$. If $(P,Q) $ is in $ \Pi_j$ and $\Pi_{j-1}$ invariants I and
-  II are true because of induction assumption. If it is only in
-  $\Pi_j$, then it is $\{(P' \setminus \{x\}),(Q' \setminus \{v\})$ or
-  $(\{x\},\{v\})$ for some $(P',Q')$ in $\Pi_{j-1}$. By definition,
-  $x$ is an element in $P'$ (as defined in the algorithm) and $v$ is a
-  leaf in $Q'$. If $(P,Q)$ is $\{(P' \setminus \{x\}),(Q' \setminus
-  \{v\})$, $Q$ is a path since only a leaf is removed from path
-  $Q'$. We know $|P'| = |Q'|$, therefore $|P' \setminus \{x\}| = |Q'
-  \setminus \{v\}|$. Hence in this case invariants I and II are
-  obvious. It is easy to see these invariants hold if $(P,Q)$ is
-  $(\{x\},\{v\})$.
-
-  For invariant III consider $(P_1,Q_1),(P_2,Q_2)$ in $\Pi_j$. If both
-  of them are also in $\Pi_{j-1}$, claim is proved. If one of them is
-  not in $\Pi_{j-1}$ then it has to be $\{(P' \setminus \{x\}),(Q'
-  \setminus \{v\})$ or $(\{x\},\{v\})$ for some $(P',Q')$ in
-  $\Pi_{j-1}$. Since by definition, $Q'$ is the only path with $v$ and
-  $P'$ the only set with $x$ in the previous iteration, $|P_1 \cap (P'
-  \setminus \{x\})| = |P_1 \cap P'|$ and $|Q_1 \cap (Q' \setminus
-  \{v\})| = |Q_1 \cap Q'|$ and $|P_1 \cap \{x\}| = 0, Q_1 \cap \{v\} =
-  0$. Thus invariant III is also proven.
-
-  \noindent
-  To prove invariant IV, consider $(P_1,Q_1),(P_2,Q_2), (P_3,Q_3)$ in
-  $\Pi_j$. If exactly one of them, say $P_3 \notin \Pi_{j-1}$, it is
-  one of the new sets. By the same argument used to prove invariant
-  III, $|P_1 \cap P_2 \cap (P' \setminus \{x\})| = |P_1 \cap P_2 \cap
-  P'|$ and $|Q_1 \cap Q_2 \cap (Q' \setminus \{x\})| = |Q_1 \cap Q_2
-  \cap Q'|$. Since $P_1, P_2, P'$ are all in $\Pi_{j-1}$, by induction
-  hypothesis $|P_1 \cap P_2 \cap P'| = |Q_1 \cap Q_2 \cap Q'|$. Also
-  $|P_1 \cap P_2 \cap \{x\}| = 0, Q_1 \cap Q_2 \cap \{v\} = 0$.  If
-  two or more of them are not in $\Pi_{j-1}$, then it can be verified
-  that $|P_1 \cap P_2 \cap P_3| = |Q_1 \cap Q_2 \cap Q_3|$ since the
-  new sets in $\Pi_j$ are either disjoint or as follows: assuming
-  $P_1, P_2 \notin \Pi_{j-1}$ and new sets are derived from $(P', Q'),
-  (P'', Q'') \in \Pi_{j-1}$ with $x_1, x_2$ exclusively in $P_1, P_2$,
-  $(\{x_1\},\{v_1\}), (\{x_2\},\{v_2\}) \in \Pi_j $ thus $v_1, v_2$
-  are exclusively in $Q_1, Q_2$ resp. it follows that $|P_1 \cap P_2
-  \cap P_3| = |(P' \setminus \{x_1\}) \cap (P'' \setminus \{x_2\})
-  \cap P_3| = |P' \cap P'' \cap P_3| = |Q' \cap Q'' \cap Q_3| = |(Q'
-  \setminus \{v_1\} \cap Q'' \setminus \{v_2\} \cap Q_3| = |Q_1 \cap
-  Q_2 \cap Q_3|$. Thus invariant IV is also proven.  \qed
-\end{proof}
-
-\noindent
-Using algorithms \ref{perms} and \ref{leafasgn} we prove the following
-theorem.
-
-\begin{theorem}
-  \label{th:perm}
-  If $\cA$ is an ICPPA, then there exists a bijection $\sigma : U
-  \rightarrow V(T)$ such that $\sigma(S_i) = P_i$ for all $i \in I$
-\end{theorem}
-\begin{proof}
-  This is a contructive proof. First, the given ICPPA $\cA$ and tree
-  $T$ are given as input to Algorithm \ref{perms}. This yields a
-  ``filtered'' ICPPA as the output which is input to Algorithm
-  \ref{leafasgn}.  It can be observed that the output of Algorithm
-  \ref{leafasgn} is a set of interval assignments to sets and
-  one-to-one assignment of elements of $U$ to each leaf of $T$. To be
-  precise, it would be of the form $\cB_0 = \cA_0 \cup \cL_0$. The
-  leaf assignments are defined in $\cL_0 = \{ (x_i,v_i) \mid x_i \in
-  U, v_i \in T, x_i \ne x_j, v_i \ne v_j, i \ne j, i,j \in [k] \}$
-  where $k$ is the number of leaves in $T$. The path assignments are
-  defined in $\cA_0 \subseteq \{(S_i',P_i') \mid S_i' \subseteq U_0,
-  P_i' \text{ is a path from } T_0\}$ where $T_0$ is the tree obtained
-  by removing all the leaves in $T$ and $U_0 = U \setminus \{ x \mid x
-  \text{ is assigned to a leaf in }\cL_0 \}$. Now we have a subproblem
-  of finding the permutation for the path assignment $\cA_0$ which has
-  paths from tree $T_0$ and sets from universe $U_0$. Now we repeat
-  the procedure and the path assignment $\cA_0$ and tree $T_0$ is
-  given as input to Algorithm \ref{perms}. The output of this
-  algorithm is given to Algorithm \ref{leafasgn} to get a new union of
-  path and leaf assignments $\cB_1 = \cA_1 \cup \cL_1$ defined similar
-  to $\cB_0, \cL_0, \cA_0$. In general, the two algorithms are run on
-  path assignment $\cA_{i-1}$ with paths from tree $T_{i-1}$ to get a
-  new subproblem with path assignment $\cA_i$ and tree $T_{i}$. $T_i$
-  is the subtree of $T_{i-1}$ obtained by removing all its
-  leaves. More importantly, it gives leaf assignments $\cL_{i}$ to the
-  leaves in tree $T_{i-1}$. This is continued until we get a
-  subproblem with path assignment $\cA_{d-1}$ and tree $T_{d-1}$ for
-  some $d \le n$ which is just a path. From the last lemma we know
-  that $\cA_{d-1}$ is an ICPPA. Another observation is that an ICPPA
-  with all its tree paths being intervals (subpaths from a path) is
-  nothing but an ICPIA\cite{nsnrs09}.  Let $\cA_{d-1}$ be equal to
-  $\{(S_i'',P_i'') \mid S_i'' \subseteq U_{d-1}, P_i'' \text{ is a
-    path from } T_{d-1} \}$. It is true that the paths $P_i''$s may
-  not be precisely an interval in the sense of consecutive integers
-  because they are some nodes from a tree. However, it is easy to see
-  that the nodes of $T_{d-1}$ can be ordered from left to right and
-  ranked to get intervals $I_i$ for every path $P_i''$ as
-  follows. $I_i = \{[l,r] \mid l = \text{ the lowest rank of the nodes
-    in }P_i'', r = l+|P_i''|-1 \}$. Let asssignment $\cA_d$ be with
-  the renamed paths. $\cA_d = \{ (S_i'', I_i) \mid (S_i'', P_i'') \in
-  \cA_{d-1} \}$. What has been effectively done is renaming the nodes
-  in $T_{d-1}$ to get a tree $T_d$.  The ICPIA $\cA_d$ is now in the
-  format that the ICPIA algorithm requires which gives us the
-  permutation $\sigma' : U_{d-1} \rightarrow T_{d-1}$
-
-\noindent
-$\sigma'$ along with all the leaf assignments $\cL_i$ gives us the
-permutation for the original path assignment $\cA$.  More precisely,
-the permutation for tree path assignment $\cA$ is defined as
-follows. $\sigma: U \rightarrow T$ such that the following is
-maintained.
-\begin{align*}
-  \sigma(x) &= \sigma'(x),   \text{ if } x \in U_{d-1} \\
-  &= \cL_i(x), \text{ where $x$ is assigned to a leaf in a subproblem
-    $\cA_{i-1}, T_{i-1}$}
-\end{align*}
-
-\noindent
-To summarize, run algorithm \ref{perms} and \ref{leafasgn} on
-$T$. After the leaves have been assigned to specific elements from
-$U$, remove all leaves from $T$ to get new tree $T_0$. The leaf
-assignments are in $\cL_0$. Since only leaves were removed $T_0$ is
-indeed a tree. Repeat the algorithms on $T_0$ to get leaf assignments
-$\cL_{1}$. Remove the leaves in $T_0$ to get $T_1$ and so on until the
-pruned tree $T_d$ is a single path. Now run ICPIA algorithm on $T_d$
-to get permutation $\sigma'$. The relation $\cL_0 \cup \cL_1 \cup
-.. \cup \cL_{d} \cup \sigma'$ gives the bijection required in the
-original problem.\qed
-\end{proof}
-
-\subsection{Finding an assignment of tree paths to a set
-  system} \label{decompos} In the previous section we have shown that
-the problem of finding a Tree Path Asssignment to an input $(\cF,T)$
-is equivalent to finding an ICPPA to $\cF$ in tree $T$.  In this
-section we characterize those set systems that have an ICPPA in a
-given tree.  As a consequence of this characterization we identify two
-sub-problems that must be solved to obtain an ICPPA.  We do not solve
-the problem and in the next section show that finding an ICPPA in a
-given tree is GI-Complete.
-
-\noindent
-A set system can be concisely represented by a binary matrix where the
-row indices denote the universe of the set system and the column
-indices denote each of the sets. Let the binary matrix be $M$ with
-order $n \times m$, the set system be $\cF = \{S_i \mid i \in [m]\}$,
-universe of set system $U = \{x_1, \dots ,x_n\}$. If $M$ represents
-$\cF$, $|U| = n, |\cF| = m$. Thus $(i,j)$th element of $M$, $M_{ij} =
-1$ iff $x_i \in S_j$. If $\cF$ has a feasible tree path assignment
-(ICPPA) $\cA = \{(S_i,P_i) \mid i \in [m]\}$, then we say its
-corresponding matrix $M$ has an ICPPA. Conversly we say that a matrix
-$M$ has an ICPPA if there exists an ICPPA $\cA$ as defined
-above.\\
-\noindent
-We now define the strict intersection graph or overlap graph of
-$\cF$. This graph occurs at many places in the literature, see for
-example \cite{kklv10, wlh02, nsnrs09}.  The vertices of the graph
-correspond to the sets in $\cF$.  An edge is present between vertices
-of two sets iff the corresponding sets have a nonempty intersection
-and none is contained in the other. Formally, intersection graph is
-$G_f = (V_f, E_f)$ such that $V_f = \{v_i \mid S_i \in \cF\}$ and $E_f
-= \{(v_i, v_j) \mid S_i \cap S_j \ne \O \text{ and }S_i \nsubseteq
-S_j, S_j \nsubseteq S_i \}$.  We use this graph to decompose $M$ as
-described in \cite{wlh02,nsnrs09}.  A prime sub-matrix of $M$ is
-defined as the matrix formed by a set of columns of $M$ which
-correspond to a connected component of the graph $G_f$.  Let us denote
-the prime sub-matrices by $M_1,\ldots,M_p$ each corresponding to one
-of the $p$ components of $G_f$. Clearly, two distinct matrices have a
-distinct set of columns.  Let $col(M_i)$ be the set of columns in the
-sub-matrix $M_i$.  The support of a prime sub-matrix $M_i$ is defined
-as $supp(M_i) = \displaystyle \bigcup_{j \in col(M_i)}S_j$.  Note that
-for each $i$, $supp(M_i) \subseteq U$.  For a set of prime
-sub-matrices $X$ we define
-$supp(X) = \displaystyle \bigcup_{M \in X} supp(M)$. \\
-
-\noindent
-Consider the relation $\preccurlyeq$ on the prime sub-matrices $M_1,
-\ldots, M_p$ defined as follows:
-\begin{equation}
-  \nonumber \{(M_i,M_j) | \mbox{ A set } S \in
-  M_i \mbox{ is contained in a set } S' \in M_j\} \cup \{(M_i,M_i) | 1
-  \leq i \leq p\} 
-\end{equation}
-
-\noindent
-This relation is the same as that defined in \cite{nsnrs09}. The prime
-submatrices and the above relation can be defined for any set
-system. We will use this structure of prime submatrices to present our
-results on an an ICPPA for a set system $\cF$. Recall the following
-lemmas, and theorem that $\preccurlyeq$ is a partial order, from
-\cite{nsnrs09}.
-
-\begin{lemma} \label{lem:containment} Let $(M_i,M_j) \in
-  \preccurlyeq$.  Then there is a set $S' \in M_j$ such that for each
-  $S \in M_i$, $S \subseteq S'$.
-\end{lemma}
-\begin{lemma}
-  For each pair of prime sub-matrices, either $(M_i,M_j) \not\in
-  \preccurlyeq$ or $(M_j,M_i) \not\in \preccurlyeq$.
-  % If $(M_i,M_j) \in \preccurlyeq$ and $(M_j,M_i) \in \preccurlyeq$,
-  % then $i = j$ and $|M_i| = 1$.
-\end{lemma}
-\begin{lemma}
-  If $(M_i,M_j) \in \preccurlyeq $ and $(M_j,M_k) \in \preccurlyeq$,
-  then $(M_i,M_k) \in \preccurlyeq$.
-\end{lemma}
-\begin{lemma}
-  If $(M_i,M_j) \in \preccurlyeq$ and $(M_i,M_k) \in \preccurlyeq$,
-  then either $(M_j,M_k) \in \preccurlyeq$ or $(M_k,M_j) \in
-  \preccurlyeq$.
-\end{lemma}
-\begin{theorem} \label{thm:partitionold} $\preccurlyeq$ is a partial
-  order on the set of prime sub-matrices of $M$.  Further, it uniquely
-  partitions the prime sub-matrices of $M$ such that on each set in
-  the partition $\preccurlyeq$ induces a total order.
-\end{theorem}
-For the purposes of this paper, we refine the total order mentioned in
-Theorem \ref{thm:partitionold}. We do this by identifying an in-tree
-rooted at each maximal upper bound under $\preccurlyeq$.  Each of
-these in-trees will be on disjoint vertex sets, which in this case
-would be disjoint sets of prime-submatrices.  The in-trees are
-specified by selecting the appropriate edges from the Hasse diagram
-associated with $\preccurlyeq$.  Let $\cI$ be the following set:
-\begin{align*}
-  \cI = \{ (M_i,M_j) \in \preccurlyeq \mid \nexists M_k s.t. M_i
-  \preccurlyeq M_k, M_k \preccurlyeq M_j \} \cup \{ (M_i,M_i), i \in
-  [p] \}
-\end{align*}
-
-\begin{theorem} \label{thm:partition} Consider the directed graph $X$
-  whose vertices correspond to the prime sub-matrices, and the edges
-  are given by $\cI$.  Then, $X$ is a vertex disjoint collection of
-  in-trees and the root of each in-tree is a maximal upper bound in
-  $\preccurlyeq$.
-\end{theorem}
-\begin{proof}
-  To observe that $X$ is a collection of in-trees, we observe that for
-  vertices corresponding to maximal upper bounds, no out-going edge is
-  present in $I$.  Secondly, for each other element, exactly one
-  out-going edge is chosen, and for the minimal lower bound, there is
-  no in-coming edge.  Consequently, $X$ is acyclic, and since each
-  vertex has at most one edge leaving it, it follows that $X$ is a
-  collection of in-trees, and for each in-tree, the root is a maximal
-  upper bound in $\preccurlyeq$.  Hence the theorem.
-\end{proof}
-Let the partition of $X$ given by Theorem \ref{thm:partition} be
-$\{X_1,\ldots,X_r\}$.  Further, each in-tree itself can be layered
-based on the distance from the root.  The root is considered to be at
-level zero. For $j \geq 0$, Let $X_{i,j}$ denote the set of prime
-matrices in level $j$ of in-tree $X_i$.
-
-\begin{lemma}
-  \label{lem:subicppa}
-  Let $M$ be a matrix and let $X$ be the directed graph whose vertices
-  are in correspondence with the prime submatrices of $M$.  Further
-  let $\{X_1,\ldots,X_r\}$ be the partition of $X$ into in-trees as
-  defined above.  Then, matrix $M$ has an ICPPA in tree $T$ iff $T$
-  can be partitioned into vertex disjoint subtrees $\{T_1, T_2, \dots
-  T_r\}$ such that, for each $1 \leq i \leq r$, the set of prime
-  sub-matrices corresponding to vertices in $X_i$ has an ICPPA in
-  $T_i$.
-\end{lemma}
-\begin{proof}
-  Let us consider the reverse direction first.  Let us assume that $T$
-  can be partitioned into $T_1, \ldots, T_r$ such that for each $1
-  \leq i \leq r $, the set of prime sub-matrices corresponding to
-  vertices in $X_i$ has an ICPPA in $T_i$.  It is clear from the
-  properties of the partial order $\preccurlyeq$ that these ICPPAs
-  naturally yield an ICPPA of $M$ in $T$.  The main property used in
-  this inference is that for each $1 \leq i \neq j \leq r$, $supp(X_i)
-  \cap supp(X_j) = \phi$.
-
-\noindent
-To prove the forward direction, we show that if $M$ has an ICPPA, say
-$\cA$, in $T$, then there exists a partition of $T$ into vertex
-disjoint subtree $T_1, \ldots, T_r$ such that for each $1 \leq i \leq
-r$, the set of prime sub-matrices corresponding to vertices in $X_i$
-has an ICPPA in $T_i$.  For each $1 \leq i \leq r$, we define based on
-$\cA$ a subtree $T_i$ corresponding to $X_i$.  We then argue that the
-trees thus defined are vertex disjoint, and complete the proof.
-Consider $X_i$ and consider the prime sub-matrix in $X_{i,0}$.
-Consider the paths assigned under $\cA$ to the sets in the prime
-sub-matrix in $X_{i,0}$.  Since the component in $G_f$ corresponding
-to this matrix is a connected component, it follows that union of
-paths assigned to this prime-submatrix is a subtree of $T$.  We call
-this sub-tree $T_i$.  All other prime-submatrices in $X_i$ are
-assigned paths in $T_i$ since $\cA$ is an ICPPA, and the support of
-other prime sub-matrices in $X_i$ are contained in the support of the
-matrix in $X_{i,0}$.  Secondly, for each $1 \leq i \neq j \leq r$,
-$supp(X_i) \cap supp(X_j) = \phi$, and since $\cA$ is an ICPPA, it
-follows that $T_i$ and $T_j$ are vertex disjoint.  Finally, since $|U|
-= |V(T)|$, it follows that $T_1, \ldots, T_r$ is a partition of $T$
-into vertex disjoint sub-trees such that for each $1 \leq i \leq r$,
-the set of matrices corresponding to nodes in $X_i$ has an ICPPA in
-$T_i$.  Hence the lemma.
-\end{proof}
-The essence of the following lemma is that an ICPPA only needs to be
-assigned to the prime sub-matrix corresponding to the root of each
-in-tree, and all the other prime sub-matrices only need to have an
-Intersection Cardinality Preserving Interval Assignments (ICPIA).
-Recall, an ICPIA is an assignment of intervals to sets such that the
-cardinality of an assigned interval is same as the cardinality of the
-interval, and the cardinality of intersection of any two sets is same
-as the cardinality of the intersection of the corresponding intervals.
-It is shown in \cite{nsnrs09} that the existence of an ICPIA is a
-necessary and sufficient condition for a matrix to have COP.  We
-present the pseudo-code to test if $M$ has an ICPPA in $T$.
-\begin{lemma} \label{lem:rooticppa} Let $M$ be a matrix and let $X$ be
-  the directed graph whose vertices are in correspondence with the
-  prime submatrices of $M$.  Further let $\{X_1,\ldots,X_r\}$ be the
-  partition of $X$ into in-trees as defined earlier in this section.
-  Let $T$ be the given tree and let $\{T_1, \ldots, T_r\}$ be a given
-  partition of $T$ into vertex disjoint sub-trees.  Then, for each $1
-  \leq i \leq r$, the set of matrices corresponding to vertices of
-  $X_i$ has an ICPPA in $T_i$ if and only if the matrix in $X_{i,0}$
-  has an ICPPA in $T_i$ and all other matrices in $X_i$ have an {\bf
-    {\em ICPIA}}.
-\end{lemma}
-\begin{proof}
-  The proof is based on the following fact- $\preccurlyeq$ is a
-  partial order and $X$ is a digraph which is the disjoint union of
-  in-trees.  Each edge in the in-tree is a containment relationship
-  among the supports of the corresonding sub-matrices. Therefore, any
-  ICPPA to a prime sub-matrix that is not the root is contained in a
-  path assigned to the sets in the parent matrix.  Consequently, any
-  ICPPA to the prime sub-matrix that is not at the root is an ICPIA,
-  and any ICPIA can be used to construct an ICPPA to the matrices
-  corresponding to nodes in $X_i$ provided the matrix in the root has
-  an ICPPA in $T_i$.  Hence the lemma.
-\end{proof}
-Lemma \ref{lem:subicppa} and Lemma \ref{lem:rooticppa} point out two
-algorithmic challenges in finding an ICPPA for a given set system
-$\cF$ in a tree $T$.  Given $\cF$, finding $X$ and its partition
-$\{X_1,\ldots,X_r\}$ into in-trees can be done in polynomial time.  On
-the other hand, as per lemma \ref{lem:subicppa} we need to parition
-$T$ into vertex disjoint sub-trees $\{T_1, \ldots, T_r\}$ such that
-for each $i$, the set of matrices corresponding to nodes in $X_i$ have
-an ICPPA in $T_i$.  This seems to be a challenging step, and it must
-be remarked that this step is easy when $T$ itself is a path, as each
-individual $T_i$ would be sub-paths.  The second algorithmic challenge
-is identified by lemma \ref{lem:rooticppa} which is to assign an ICPPA
-from a given tree to the matrix associated with the root node of
-$X_i$.
-\begin{algorithm}[h]
-  \caption{Algorithm to find an ICPPA for a matrix $M$ on tree $T$:
-    $main\_ICPPA(M, T$)}
-  \label{al:icppa-main}
-  \begin{algorithmic}
-    \STATE Identify the prime sub-matrices. This is done by constructing the strict overlap graph and identify connected components.  Each connected component yields a prime sub-matrix.   \\
-    \STATE Construct the partial order $\preccurlyeq$ on the set of prime sub-matrices.  \\
-    \STATE Construct the partition $X_1,\ldots,X_r$ of the prime
-    sub-matrices induced by $\preccurlyeq$ \\
-    \STATE For each $1 \leq i \leq r$, Check if all matrices except
-    those in $X_{i,0}$ has an ICPIA.  If a matrix does not have ICPIA
-    exit with a negative answer.  To check for the existence of ICPIA,
-    use the result in \cite{nsnrs09}.  \STATE Find a partition of
-    $T_1, \ldots, T_r$ such that matrices in $X_{i,0}$ has an ICPPA in
-    $T_i$.  If not such partition exists, exit with negative answer.
-  \end{algorithmic}
-\end{algorithm}
-
-\subsection{Complexity of Tree Path Assignment-A
-  Discussion} 
-\label{complexity} 
-Recall that the input to the Tree
-Path Assignment question is an order pair $(\cF,T)$ where $\cF$ is a
-family of subsets of an universe $U$, and $T$ is a tree such that
-$|V(T)| = |U|$.  The question is to come up with a bijection from $U$
-to $V(T)$ such that the image of each set in $\cF$ is a path in $T$.
-We show that this problem is at least as hard as the problem of
-testing if two given path chordal graphs are isomorphic.
-\begin{theorem} \label{thm:isocomp} Tree Path Assignment is
-  isomorphism-complete.
-\end{theorem}
-\begin{proof}
-  It is well known (see for example \cite{kklv10}) that testing
-  isomorphism of path chordal graphs is isomorphism complete.  We show
-  a reduction of path chordal graph isomorphism to tree path
-  assignment.  Given $G_1$ and $G_2$ two path chordal graphs, let
-  $T_2$ be the clique tree of $G_2$ obtained from say \cite{gav78}.
-  The nodes of $T_2$ correspond to the maximal cliques of $G_2$ and
-  each vertex of $G_2$ corresponds to a path in $G_2$.  This is a
-  well-known characterization of path chordal graphs and $T_2$ can be
-  computed in polynomial time.  In $G_1$, let $S_v$ denote the maximal
-  cliques of $G_1$ that contain $v$.  This can be computed in
-  polynomial time as $G_1$ is path chordal, and all chordal graph only
-  have a linear number of maximal cliques.  The universe $U$
-  corresponds bijectively to the set of maximal cliques in $G_1$, and
-  $\cF = \{S_v | v \in V(G_1)\}$.  Now, we claim that $(\cF, T_2)$ has
-  a tree path assignment if and only if $G_1$ and $G_2$ are
-  isomorphic.  This is clear since for each vertex $v \in G_1$, there
-  is an associated $S_v$ which is the set of maximal cliques
-  containing $v$.  In $G_2$, each vertex corresponds to a path in
-  $T_2$, and the nodes on this path corresponds to the maximal cliques
-  in $G_2$.  Consequenlty, a tree path assignment will naturally yield
-  an isomorphism between $G_1$ and $G_2$, and vice versa.  Therefore,
-  Tree Path Assignment is isomorphism-complete.
-\end{proof}
-
-\subsubsection{Consecutive Ones Testing is in Logspace}
-While Tree Path Assignment is isomorphism-complete, it is polynomial
-time solvable when the given tree is a path.  Indeed, in this case we
-encounter a restatement of matrices with the COP.  The known
-approaches to testing for COP fall into two categories: those that
-provide a witness when the input matrix does not have COP, and those
-that do not provide a witness.  The first linear time algorithm for
-testing COP for a binary matrix was using a data structure called PQ
-trees, which represent all COP orderings of $M$, invented by
-\cite{bl76}. There is a PQ tree for a matrix iff the matrix has COP.
-Indeed, this is an algorithmic characterization of the consecutive
-ones property and the absence of the PQ-tree does not yield any
-witness to the reason for failure.  A closely related data structure
-is the generalized PQ tree in \cite{mcc04}.  In generalized PQ tree
-the P and Q nodes are called prime and linear nodes. Aside from that,
-it has a third type of node called degenerate nodes which is present
-only if the set system does not have COP \cite{mcc04}.  Using the idea
-of generalized PQ tree, \cite{mcc04} proves that checking for
-bipartiteness in the certain incomparability graph is sufficient to
-check for COP.  \cite{mcc04} invented a certificate to confirm when a
-binary matrix does not have COP.  \cite{mcc04} describes a graph
-called incompatibility graph of a set system $\cF$ which has vertices
-$(a,b), a \ne b$ for every $a, b \in U$, $U$ being the universe of the
-set system. There are edges $((a,b),(b,c))$ and $((b,a),(c,b))$ if
-there is a set $S \in \cF$ such that $a, c \in S$ and $b \notin S$. In
-other words the vertices of an edge in this graph represents two
-orderings that cannot occur in a consecutive ones ordering of $\cF$.
-\begin{theorem}[Theorem 6.1, \cite{mcc04}]
-  Let $\cF$ be an arbitrary set family on domain $V$. Then $\cF$ has
-  the consecutive ones property if and only if its incompatibility
-  graph is bipartite, and if it does not have the consecutive ones
-  property, the incompatibility graph has an odd cycle of length at
-  most $n+3$.
-\end{theorem}
-This theorem gives a certificate as to why a given matrix does not
-have COP.  Similarly, the approach of testing for an ICPIA in
-\cite{nsnrs09} also gives a different certificate- a prime sub-matrix
-that does not have an ICPIA.  Further, the above theorem can be used
-to check if a given matrix has COP in logspace by checking if its
-incompatibility graph is bipartite. \cite{rei84} showed that checking
-for bipartiteness can be done in logspace. Thus we conclude that
-consecutive ones testing can be done in logspace.
-
-\noindent
-More recently, \cite{kklv10} showed that interval graph isomorphism
-can be done in logspace. Their paper proves that a canon for interval
-graphs can be calculated in logspace using an interval hypergraph
-representation of the interval graph with each hyperedge being a set
-to which an interval shall be assigned by the canonization
-algorithm. An overlap graph (subgraph of intersection graph, edges
-define only strict intersections and no containment) of the hyperedges
-of the hypergraph is created and canons are computed for each overlap
-component. The overlap components define a tree like relation due to
-the fact that two overlap components are such that either all the
-hyperedges of one is disjoint from all in the other, or all of them
-are contained in one hyperedge in the other. This is similar to the
-containment tree defined in \cite{nsnrs09} and in this paper. Finally
-the canon for the whole graph is created using logspace tree
-canonization algorithm from \cite{sl92}. The interval labelling done
-in this process of canonization is exactly the same as the problem of
-assigning feasible intervals to a set system, and thus the problem of
-finding a COP ordering in a binary matrix \cite{nsnrs09}.
-
-\begin{theorem}[Theorem 4.7, \cite{kklv10}]
-  \label{th:canonlabel}
-  Given an interval hypergraph $\cH$, a canonical interval labeling
-  $l_H$ for $H$ can be computed in FL.
-\end{theorem}
-
-We present the following reduction to see that COP testing is indeed
-in logspace. Given a binary matrix $M$ of order $n \times m$, let $S_i
-= \{j \mid M[j,i]=1 \}$. Let $\cF = \{S_i \mid i \in [m] \}$ be this
-set system. Construct a hypergraph $\cH$ with its vertex set being
-$\{1, 2, \dots n\}$. The edge set of $\cH$ is isomorphic to
-$\cF$. Thus every edge in $\cH$ represents a set in the given set
-system $\cF$. Let this mapping be $\pi: E(\cH) \rightarrow \cF$. It is
-easy to see that if $M$ has COP, then $\cH$ is an interval
-hypergraph. From theorem \ref{th:canonlabel}, it is clear that the
-interval labeling $l_{\cH}: V(\cH) \rightarrow [n]$ can be calculated
-in logspace. Construct sets $I_i = \{ \l_{\cH}(x) \mid x \in E, E \in
-E(\cH), \pi(E) = S_i\}$, for all $i \in [m]$. Since $\cH$ is an
-interval hypergraph, $I_i$ is an interval for all $i \in [m]$, and is
-the interval assigned to $S_i$ if $M$ has COP.
-
-Now we have the following corollary.
-\begin{corollary}
-  \label{cor:coplog}
-  If a binary matrix $M$ has COP then the interval assignments to each
-  of its columns can be calculated in FL.
-\end{corollary}
-
-\noindent
-Finally, we conclude by asking about the complexity of Tree Path
-Assignment restricted to other subclasses of trees.  In particular, is
-Tree Path Assignment in caterpillars easier than Tree Path assignment
-in general trees.
-
-\chapter{Conclusion}
-\label{ch:conclusion}
-
-\begin{center}
-{\tt MOVED FROM INTRO SUBSECTION IN ch:myresearch CHAPTER begin}
-\end{center}
-
-% \annote{
-  It is an interesting fact that for a matrix with the COP, the
-  intersection graph of the corresponding set system is an interval
-  graph. % }{the following text could be in Further Study/Conclusion}
-A
-similar connection to a subclass of chordal graphs and a superclass of
-interval graphs exists for the generalization of COP.  In this case,
-the intersection graph of the corresponding set system must be a {\em
-  path graph}. Chordal graphs are of great significance, are
-extensively studied, and have several applications.  One of the well
-known and interesting properties of a chordal graphs is its connection
-with intersection graphs \cite{mcg04}. For every chordal graph, there
-exists a tree and a family of subtrees of this tree such that the
-intersection graph of this family is isomorphic to the chordal graph
-\cite{plr70,gav78,bp93}.  These trees when in a certain format, are
-called {\em clique trees} \cite{apy92} of the chordal graph. This is a
-compact representation of the chordal graph. There has also been work
-done on the generalization of clique trees to clique hypergraphs
-\cite{km02}.  If the chordal graph can be represented as the
-intersection graph of paths in a tree, then the graph is called path
-graph \cite{mcg04}.  Therefore, it is clear that if there is a
-bijection from $U$ to $V(T)$ such that for every set, the elements in
-it map to vertices of a unique path in $T$, then the intersection
-graph of the set system is indeed a path graph.  However, this is only
-a necessary condition and can be checked efficiently because path
-graph recognition is polynomial time
-solvable\cite{gav78,aas93}. Indeed, it is possible to construct a set
-system and tree, such that the intersection graph is a path graph, but
-there is no bijection between $U$ and $V(T)$ such that the sets map to
-paths. Path graph isomorphism is known be isomorphism-complete, see
-for example \cite{kklv10}. An interesting area of research would be to
-see what this result tells us about the complexity of the tree path
-labeling problem (not covered in this paper).
-
-\begin{center}
-{\tt MOVED FROM INTRO SUBSECTION IN ch:myresearch CHAPTER\\ end}
-\end{center}
-
-We give a characterization for feasible tree path labeling of path
-hypergraphs. The proof for this is constructive and computes a
-feasibility bijection mapping vertices of the hypergraph to vertices
-of the given target tree. This thesis also discovered an exponential algorithm
-that computes a feasible TPL to a given hypergraph if it is a path hypergraph.
-
-Our results have close connections to recognition of path graphs and
-to path graph isomorphism.  Graphs which can be represented as the
-intersection graph of paths in a tree are called {\em path
-  graphs}\cite{mcg04}. Thus, a hypergraph $\cF$ can be interpreted as
-paths in a tree, if and only if the intersection graph of $\cF$ is a
-{\em path graph}. Path graphs are a subclass of chordal graphs since
-chordal graphs are characterized as the intersection graphs of
-subtrees of a tree\cite{mcg04}.  Path graphs are well studied in the
-literature \cite{plr70}--\cite{mcg04}\tnote[TH]{Check if cite:bp93 is
-  indeed for path graphs}.  Path graph recognition can be done in
-polynomial time\cite{gav78,aas93}.  Clearly, this is a necessary
-condition in terms of the intersection graph of the input hypergraph
-$\cF$ in \FTPL. However, one can easily obtain a counterexample to
-show the insufficiency of this condition\tnote[TH]{Give a figure
-  showing insuff.}.  Path graph isomorphism is known be
-isomorphism-complete\cite{kklv10}. Therefore, \annote[TH]{it is
-  unlikely that we can solve the problem of finding feasible path
-  labeling $\cl$ for a given $\cF$ and }{WHY?} tree $T$.  It is
-definitely interesting to classify the kinds of trees and hypergraphs
-for which feasible path labelings can be\tnote[TH]{ELABORATE.}  found
-efficiently.  \annote[TH]{These results would form a natural
-  generalization of COP testing and interval graph isomorphism,
-  culminating in Graph Isomorphism }{REWORD} itself.  To this effect
-we consider TPL on {\kstar} and give a polynomial time solution for
-the same.
-
-While we address the computation of TPL
-for a given hypergraph% \footnote{The term {\em hypergraph} is used
-%   instead {\em set systems} are used interchangably in this
-%   section. Hypergraphs that have TPLs on some target tree are called
-%   {\em path hypergraphs}.}
-from a given target tree in this research, however optimization
-problems in TPL, akin to problems in Section~\ref{sec:optcop} for COP,
-is outside the scope of this thesis. Whether TPL on general trees is
-solvable in $P$ remains open. So do optimization opportunities in TPL
-(possible extensions to optimization for COP in the
-Section~\ref{sec:optcop}.\tnote[TH]{link to section}).
-
-%\newpage %DUMMY
-\tnote[TH]{ ||Complexity challenges:|| Canonization is an important
-  tool in graph isomorphism. Invention of a
-% deterministic method of canonization for any class of graphs naturally
-% results in an algorithm for isomorphism. All that is required is to
-% check if the canons of two graphs are the same. Thus complexity of
-% graph isomorphism can be studied by studying canonization
-% methods. While general graph ismorphism remains elusive in terms of
-% hardness, canonization has been studied for smaller classes of graphs
-% thus giving complexity/hardness results for them.
-
-% In 1992, \cite{sl92} made an important discovery that tree isomorphism
-% is in logspace. It was done by inventing a method of canonization of
-% trees using a logspace depth first traversal
-% algorithm. \tnote{confirm the statement of claim}.
-
-% \cite{kklv10} proved that interval graph canonization is also in
-% logspace thus drawing logspace conslusions about COT. 
-% % This result is
-% % built on the fact that 
-% Interval graphs are FP+C (fixed point with
-% counting) definable \tnote{elucidate} and \cite{l10} showed that
-% this implies that it captures PTIME. \tnote{elucidate} This result
-% along with that of undirected graph reachability being in logspace
-% \cite{rei08}, \cite{kklv10} proved their logspace result.
-}   %
-\tnote[TH]{ (1) the interesting survey in conclusion section of kklv10
-  (2) [[[COP helping some problems with hardness]]] having COP in the
-  input structure makes some problems less hard than in a general
-  input. -- min set cover problem is known to be NP hard. so is min
-  hitting
-  set. these two problems are equivalent cite-adp80\\
-  -- -- complexity results: log factor polynomial time approximable cite-f98, W[2]-complete with solution size as parameter cite-df99\\
-  -- a couple of generalizations:\\
-  -- -- min degree hypergraph (MDH) problem\\
-  -- -- red blue covering (RBSC) problem\\
-  ----- cite-d08phd Chapter 5.\\
-  ----- what are cite-d08phd results\\
-  ----- Rectangle stabbing \\
-  ----- see cite-d08phd Chapter 6?  }
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Appendices.
-
-% \appendix %%% NON NEIN NAHI ILLA NONE
-
-% \chapter{A}
-% \tnote{DO WE NEED THIS?}
-
-\newpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Bibliography.
-\begin{singlespace}
-%\begin{thebibliography}{10}
-\addcontentsline{toc}{chapter}{REFERENCES}
-
-{\footnotesize
-  \bibliographystyle{alpha} 
-  \bibliography{../lib/cop-variants__thesis}
-}
-%\end{thebibliography}
-\end{singlespace}
-%\newpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% List of papers
-
-% \listofpapers %%% NON NEIN NAHI ILLA NONE
-
-% \begin{enumerate}  
-% \item Authors....  \newblock
-%  Title...
-%   \newblock {\em Journal}, Volume,
-%   Page, (year).
-% \end{enumerate}
-
-\end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% DOCUMENTATION ON THE THESIS CLASS FILE ETC.
-
-% Before reading any further please note that you are strongly advised
-% against changing any of the formatting options used in the class
-% provided in this directory, unless you are absolutely sure that it
-% does not violate the IITM formatting guidelines.  \emph{Please do not
-%   change the margins or the spacing.}  If you do change the formatting
-% you are on your own (don't blame me if you need to reprint your entire
-% thesis).  In the case that you do change the formatting despite these
-% warnings, the least I ask is that you do not redistribute your style
-% files to your friends (or enemies).
-
-% It is also a good idea to take a quick look at the formatting
-% guidelines.  Your office or advisor should have a copy.  If they
-% don't, pester them, they really should have the formatting guidelines
-% readily available somewhere.
-
-% To generate PDF's with the links from the \verb+hyperref+ package use
-% the following command:
-% \begin{verbatim}
-% % dvipdfm -o thesis.pdf thesis.dvi
-% \end{verbatim}
-
-% \section{Package Options}
-
-% Use this thesis as a basic template to format your thesis.  The
-% \verb+iitmdiss+ class can be used by simply using something like this:
-% \begin{verbatim}
-% \documentclass[PhD]{iitmdiss}  
-% \end{verbatim}
-
-% To change the title page for different degrees just change the option
-% from \verb+PhD+ to one of \verb+MS+, \verb+MTech+ or \verb+BTech+.
-% The dual degree pages are not supported yet but should be quite easy
-% to add.  The title page formatting really depends on how large or
-% small your thesis title is.  Consequently it might require some hand
-% tuning.  Edit your version of \verb+iitmdiss.cls+ suitably to do this.
-% I recommend that this be done once your title is final.
-
-% To write a synopsis simply use the \verb+synopsis.tex+ file as a
-% simple template.  The synopsis option turns this on and can be used as
-% shown below.
-% \begin{verbatim}
-% \documentclass[PhD,synopsis]{iitmdiss}                                
-% \end{verbatim}
-
-% Once again the title page may require some small amount of fine
-% tuning.  This is again easily done by editing the class file.
-
-% This sample file uses the \verb+hyperref+ package that makes all
-% labels and references clickable in both the generated DVI and PDF
-% files.  These are very useful when reading the document online and do
-% not affect the output when the files are printed.
-
-% \section{Example Figures and tables}
-
-% Fig.~\ref{fig:iitm} shows a simple figure for illustration along with
-% a long caption.  The formatting of the caption text is automatically
-% single spaced and indented.  Table~\ref{tab:sample} shows a sample
-% table with the caption placed correctly.  The caption for this should
-% always be placed before the table as shown in the example.
-
-% \begin{figure}[htpb]
-%   \begin{center}
-%     \resizebox{50mm}{!} {\includegraphics *{iitm.eps}}
-%     \resizebox{50mm}{!} {\includegraphics *{iitm.eps}}
-%     \caption {Two IITM logos in a row.  This is also an
-%       illustration of a very long figure caption that wraps around two
-%       two lines.  Notice that the caption is single-spaced.}
-%   \label{fig:iitm}
-%   \end{center}
-% \end{figure}
-
-% \begin{table}[htbp]
-%   \caption{A sample table with a table caption placed
-%     appropriately. This caption is also very long and is
-%     single-spaced.  Also notice how the text is aligned.}
-%   \begin{center}
-%   \begin{tabular}[c]{|c|r|} \hline
-%     $x$ & $x^2$ \\ \hline
-%     1  &  1   \\
-%     2  &  4  \\
-%     3  &  9  \\
-%     4  &  16  \\
-%     5  &  25  \\
-%     6  &  36  \\
-%     7  &  49  \\
-%     8  &  64  \\ \hline
-%   \end{tabular}
-%   \label{tab:sample}
-%   \end{center}
-% \end{table}
-
-% \section{Bibliography with BIB\TeX}
-
-% I strongly recommend that you use BIB\TeX\ to automatically generate
-% your bibliography.  It makes managing your references much easier.  It
-% is an excellent way to organize your references and reuse them.  You
-% can use one set of entries for your references and cite them in your
-% thesis, papers and reports.  If you haven't used it anytime before
-% please invest some time learning how to use it.
-
-% I've included a simple example BIB\TeX\ file along in this directory
-% called \verb+refs.bib+.  The \verb+iitmdiss.cls+ class package which
-% is used in this thesis and for the synopsis uses the \verb+natbib+
-% package to format the references along with a customized bibliography
-% style provided as the \verb+iitm.bst+ file in the directory containing
-% \verb+thesis.tex+.  Documentation for the \verb+natbib+ package should
-% be available in your distribution of \LaTeX.  Basically, to cite the
-% author along with the author name and year use \verb+\cite{key}+ where
-% \verb+key+ is the citation key for your bibliography entry.  You can
-% also use \verb+\citet{key}+ to get the same effect.  To make the
-% citation without the author name in the main text but inside the
-% parenthesis use \verb+\citep{key}+.  The following paragraph shows how
-% citations can be used in text effectively.
-
-% More information on BIB\TeX\ is available in the book by
-% \cite{lamport:86}.  There are many
-% references~\citep{lamport:86,prabhu:xx} that explain how to use
-% BIB\TeX.  Read the \verb+natbib+ package documentation for more
-% details on how to cite things differently.
-
-% Here are other references for example.  \citet{viz:mayavi} presents a
-% Python based visualization system called MayaVi in a conference paper.
-% \citet{pan:pr:flat-fst} illustrates a journal article with multiple
-% authors.  Python~\citep{py:python} is a programming language and is
-% cited here to show how to cite something that is best identified with
-% a URL.
-
-% \section{Other useful \LaTeX\ packages}
-
-% The following packages might be useful when writing your thesis.
-
-% \begin{itemize}  
-% \item It is very useful to include line numbers in your document.
-%   That way, it is very easy for people to suggest corrections to your
-%   text.  I recommend the use of the \texttt{lineno} package for this
-%   purpose.  This is not a standard package but can be obtained on the
-%   internet.  The directory containing this file should contain a
-%   lineno directory that includes the package along with documentation
-%   for it.
-
-% \item The \texttt{listings} package should be available with your
-%   distribution of \LaTeX.  This package is very useful when one needs
-%   to list source code or pseudo-code.
-
-% \item For special figure captions the \texttt{ccaption} package may be
-%   useful.  This is specially useful if one has a figure that spans
-%   more than two pages and you need to use the same figure number.
-
-% \item The notation page can be entered manually or automatically
-%   generated using the \texttt{nomencl} package.
-
-% \end{itemize}
-
-% More details on how to use these specific packages are available along
-% with the documentation of the respective packages.
-
-% END OF DOCUMENTATION
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-01-03 src/0021ded.tex
Cleaned up thesis.tex. Split thesis.tex into chapterwise subfiles for easy edit tex management. Content changes are mainly in chapter 1 and 3 although nothing significant. To being working on chapter 2 after this commit.

%\vskip 15ex%
\vspace*{1in}

{
  \fontencoding{T1}
%  \fontfamily{cmdh} 
  \fontfamily{bch} %{put}  %{bch} % {pbk}
%  \fontfamily{pag} %
  \selectfont%
  {\large \noindent The process of preparing programs for the digital
    computer is especially attractive, not only because it can be
    economically and scientifically rewarding, but also because it can
    be an aesthetic experience much like composing poetry or
    music.%
    {\small \par \hfill \emph{The Art of Computer Programming},
      D.\;E.\;K.}%
  }%
  \vskip 5ex%
  {\Large \noindent The rhetoric and mythos of science create the
    comforting image of linear progression toward truth.%
    {\normalsize \par \hfill \emph{New Oxford American
        Dictionary, 2nd Ed.}}%\textbf{mythos};
  }%  
  \vskip 5ex%
  {\huge%
    \noindent Art is the solution to chaos. \par %
  }%
}

### 2012-01-03 src/003abstract.tex
Cleaned up thesis.tex. Split thesis.tex into chapterwise subfiles for easy edit tex management. Content changes are mainly in chapter 1 and 3 although nothing significant. To being working on chapter 2 after this commit.

--- 2011-12-30 src/thesis.tex
@@
-\documentclass[MS]             % MS thesis dissertation.
-              {iitmdiss_as}    % Hacked by hyperbolicme.
-%              {iitmdiss}      % If the powers to be are finicky.
-
-\usepackage{times}             % Font?
-\usepackage{t1enc}             % T1 encoding??  
-%\usepackage{boxedminipage}
-\usepackage{../lib/mythesislib}
-\usepackage{../lib/mylatexlib}
-% \usepackage{comment}           % DEBUG ONLY
-
-\def \mythesissubmissiondate {10 January 2012}
-\def \mythesissubmissionmonth {\MakeUppercase {January 2012}}
-
-\begin{document}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Title page
-\title{\mythesistitleNOTSYNPOSIS} 
-\author{\myname} 
-\date{
-  {\mythesissubmissionmonth}} 
-\department{\MakeUppercase{\mythesisdept}}
-%\department{\mythesisdept}
-
-%\nocite{*}
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Certificate
-\certificate
-
-\vspace*{0.5in}
-
-\noindent 
-This is to certify that the thesis titled {\bf \mythesistitleNOTSYNPOSIS},
-submitted by {\bf \myname}, to the {\bf Indian Institute of Technology
-Madras}, for the award of the degree of {\bf \mydegree}, is 
-a bona fide record of the research work done by her under our
-supervision.  The contents of this thesis, in full or in parts, have
-not been submitted to any other Institute or University for the award
-of any degree or diploma.
-
-\vspace*{1.5in}
-
-\begin{singlespacing}
-  \hspace*{-0.25in}
-  \parbox{3in}{
-    \noindent {\bf Dr.~\myadvisor} \\
-    \small
-    \noindent Research Guide\\
-    \noindent Associate Professor\\
-    \noindent Dept. of \mydept\\
-    \noindent IIT Madras -- 600 036
-  } \hfill
-  \parbox{1.25in}{               % {1.75in}
-    \small
-    \vspace*{0.75in}
-    \begin{tabular}[h]{r}
-    \noindent {Chennai}\\         
-    \noindent {\mythesissubmissiondate} 
-    \end{tabular}
-  }
-\end{singlespacing}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Acknowledgements
-\acknowledgements
-
-\tnote{WIP}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Abstract
-
 \begin{abstract}
-  \noindent {\em {\bf Keywords}: \hspace*{0.5em} \parbox[t]{4.4in}{consecutive
-    ones property, algorithmic graph theory, hypergraph isomorphism,
-    interval labeling}} \vspace*{24pt}
+  \noindent {\em {\bf Keywords}:
+    \hspace*{0.5em} \parbox[t]{4.4in}{consecutive ones property,
+      algorithmic graph theory, hypergraph isomorphism, interval
+      labeling}} \vspace*{24pt}
 
@@
 
-  \tnote{EXPAND. 
-  Abstract must be a brief about what results we have and how it fits
+  \tnote{EXPAND.  Abstract must be a brief about what results we have
+    and how it fits
     in the body of research.\\
-  -- Area: Broad to Specialized. i.e. Combinatorial
-    algorithms -> Matrix reorganization -> general data reorganization
+    -- Area: Broad to Specialized. i.e. Combinatorial algorithms ->
+    Matrix reorganization -> general data reorganization
     (interval assignment) -> path assignment\\
@@
     Polynomial algorithm for a subset of the generalization.\\
-% maybe add - we explore a natural
-%     generalization of results on binary matrices with the |\tem
-%     consecutive-ones property|.  We consider the following constraint
-%     satisfaction problem. Given (i) a set system $\F \subseteq$
-%     $(2^{U} \setminus \emptyset)$ of a finite set $U$ of cardinality
-%     $n$, (ii) a tree $T$ of size $n$ and (iii) a bijection called
-%     |\tem tree path labeling|, $\cl$ mapping the sets in $\cF$ to
-%     paths in $T$, does there exist at least one bijection $\phi:U
-%     \rightarrow V(T)$ such that for each $S \in \cF$, $\{\phi(x) \mid
-%     x \in S\} = \cl(S)$?  A tree path labeling of a set system is
-%     called |\tem feasible| if there exists such a bijection $\phi$.
-%     We present an algorithmic characterization of feasible tree path
-%     labeling. COP is a special instance of tree path labeling problem
-%     when $T$ is a path.  We conclude with a polynomial time algorithm
-%     to find a feasible tree path labeling of a given set system when
-%     $T$ is a |\tem $k$-subdivided star|, set system has a single
-%     containment tree of overlap components and set size is limited to
-%     at most $k+2$.
+    % maybe add - we explore a natural generalization of results on
+    % binary matrices with the |\tem consecutive-ones property|.  We
+    % consider the following constraint satisfaction problem. Given
+    % (i) a set system $\F \subseteq$ $(2^{U} \setminus \emptyset)$ of
+    % a finite set $U$ of cardinality $n$, (ii) a tree $T$ of size $n$
+    % and (iii) a bijection called |\tem tree path labeling|, $\cl$
+    % mapping the sets in $\cF$ to paths in $T$, does there exist at
+    % least one bijection $\phi:U \rightarrow V(T)$ such that for each
+    % $S \in \cF$, $\{\phi(x) \mid x \in S\} = \cl(S)$?  A tree path
+    % labeling of a set system is called |\tem feasible| if there
+    % exists such a bijection $\phi$.  We present an algorithmic
+    % characterization of feasible tree path labeling. COP is a
+    % special instance of tree path labeling problem when $T$ is a
+    % path.  We conclude with a polynomial time algorithm to find a
+    % feasible tree path labeling of a given set system when $T$ is a
+    % |\tem $k$-subdivided star|, set system has a single containment
+    % tree of overlap components and set size is limited to at most
+    % $k+2$.
   }
 \end{abstract}
-
-%\pagebreak
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Table of contents etc.
-
-\begin{singlespace}
-\tableofcontents
-\thispagestyle{empty}
-
-\listoftables
-\addcontentsline{toc}{chapter}{LIST OF TABLES}
-\listoffigures
-\addcontentsline{toc}{chapter}{LIST OF FIGURES}
-\end{singlespace}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Abbreviations
-\abbreviations
-
-\noindent 
-\begin{tabbing}
-xxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
-\textbf{COP}   \> Consecutive-ones Property \\
-\textbf{COT}   \> Consecutive-ones Testing \\
-\textbf{ICPIA}   \> Intersection Cardinality Preservation Interval Assignment \\
-\textbf{ICPPL}   \> Intersection Cardinality Preserved Path Labeling \\
-% ADD MORE HERE
-\end{tabbing}
-
-\pagebreak
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Notation
-
-\chapter*{\centerline{NOTATION}}
-\addcontentsline{toc}{chapter}{NOTATION}
-
-\begin{singlespace}
-\begin{tabbing}
-xxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
-\textbf{$2^{U}$}  \> Powerset of set $U$ \\
-\end{tabbing}
-\end{singlespace}
-
-\pagebreak
-\clearpage
-
-% The main text will follow from this point so set the page numbering
-% to arabic from here on.
-\pagenumbering{arabic}
-
-\chapter{Introduction}
-\label{ch:intro}
-
-Consecutive-ones property is a non-trivial property of binary matrices
-that has been studied widely in the literature for over past 50
-years. Detection of COP in a matrix is possible efficiently and there
-are several algorithms that achieve the same. This thesis documents
-the work done on an extension of COP extended from the equivalent
-interval assignment problem in \cite{nsnrs09}. These new results
-rigorously prove a natural extension (to trees) of their
-characterization as well as makes connections to graph isomorphism,
-namely path graph isomorphism.
-
-\tnote{Have a para on Organization: Outline of document}
-Section~\ref{sec:background} gives a brief survey of COP and
-optimization problems related to it followed by motivation for the
-thesis in Section~\ref{sec:motive}.  Section~\ref{sec:results}
-presents a summary of our results on the extension of COP namely, the
-tree path labeling problem. \tnote[TH]{update for thesis}
-
-% \tnote{testing 1 2 3...}
-% \tnote[bogus1]{testing 1 2 3...}
-% \tnote[bogus2]{testing 1 2 3...}
-% \tnote[GTC]{testing 1 2 3...}
-% \tnote[bogus4]{testing 1 2 3...}
-% \tnote[TH]{testing 1 2 3...}
-
-\section{Illustration of the problem}
-\label{sec:problem}
-
-A group of students, \Pa, \Pig, \Sn, \Wo, \Vi, \Li, \Ch, \Sa, \Fr,
-\Sc\ and \Lu\ enroll at the {\WSI} for a liberal arts programme.  As
-part of their semester thesis, they pick a body of work to study and
-form the namesake study groups, {\LLL}, {\GGG}, {\BBB} and
-{\TTT}\tnote[TH]{put bib entries for these works!}. A student will be
-in at least one study group and may be in more than one. For instance,
-as will be seen later, {\Fr} studies both {\LLL} and {\TTT} while \Wo\
-studies only \BBB.
-
-Let $U$ and $\cF$ represent the set of students and the set of study
-groups respectively and the integers $n$ and $m$ denote the total
-number students and study groups respectively. In relation to this
-example, these are defined in Table~\ref{tab:wsigroups}. Also given
-there is the study group allocation to students.
-
-\begin{table}[t]%[htbp]
-  \centering
-  { \footnotesize
-  \begin{tabular}{rcl}
-    $U $&$=$&$ \{\xPa,\; \xPi,\; \xSn,\; \xWo,\; \xVi,\; \xLi,\; \xCh,\;
-    \xSa,\; \xFr,\; \xSc,\; \xLu\}$\\
-    $\cF $&$=$&$ \{\xLLL, \xGGG, \xBBB, \xTTT\}$\\
-    $\xLLL $&$=$&$ \{\xCh,\;  \xSa,\;  \xFr,\;  \xSc,\;  \xLu \}$\\
-    $\xGGG $&$=$&$ \{\xPa,\;  \xPi,\;  \xVi,\;  \xCh \}$\\
-    $\xBBB $&$=$&$ \{\xSn,\;  \xPi,\;  \xWo \}$\\
-    $\xTTT $&$=$&$ \{\xVi,\;  \xLi,\;  \xCh,\;  \xFr \}$\\
-    $n $&$=$&$ |U| = 11$\\
-    $m $&$=$&$ |\cF| = 4$      
-  \end{tabular}
-}
-  \caption{\figtabsize Students and study groups in \WSI}
-  \label{tab:wsigroups}
-\end{table}
-
-The campus has a residential area {\residenceblock} that has $n$
-single occupancy apartments reserved for the study groups'
-accommodation.  All these apartments are located such that the streets
-connecting them do {\em not} form
-loops. Fig~\ref{fig:streetmappathpeople} \tnote[TH]{fig:streetmap}
-shows the street map for {\residenceblock}. It may be noted that as a
-graph, it classifies as a tree.
-
-\begin{table}[htbp]
-  \centering
-  {\footnotesize    
-    \begin{tabular}{l||l}
-      \begin{tabular}{rcl}
-        $T $&$=$&$ \text{\em Street map tree of \residenceblock}$\\
-        $V(T) $&$=$&$ \{ 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11 \}$\\
-        $\cP $&$=$&$ \{R\xLLL, R\xGGG, R\xBBB, R\xTTT\}$\\
-        $R\xLLL $&$=$&$ \{9, 1, 5, 3, 11\}$\\
-        $R\xGGG $&$=$&$ \{7, 2, 6, 5\}$\\
-        $R\xBBB $&$=$&$ \{8, 2, 4 \}$\\
-        $R\xTTT $&$=$&$ \{10, 6, 5, 3\}$\\
-        $n $&$=$&$ |V| = 11$\\
-        $m $&$=$&$ |\cP| = 4$\\\\
-        $\cl $&$=$& {\em Study group to route mapping}\\        
-        $\cl(\mathbb{X}) $&$=$&  $R\mathbb{X}$ for all $\mathbb{X} \in
-        \cF$
-      \end{tabular} & 
-      \begin{tabular}{c}
-       {\em Apartment allocation }($\phi$)\\
-                    \begin{tabular}{c|c}                      
-                      1  & \xSa\\
-                      2  & \xPi\\
-                      3  & \xFr\\
-                      4  & \xWo\\
-                      5  & \xCh\\
-                      6  & \xVi\\
-                      7  & \xPa\\
-                      8  & \xSn\\
-                      9  & \xLu\\
-                      10 & \xLi\\
-                      11 & \xSc                     
-                    \end{tabular}
-      \end{tabular}\\
-    \end{tabular}
-  }
-  \caption{\figtabsize A solution to study group accommodation problem}
-  \label{tab:iltree}
-\end{table}
-
-A natural question would be to find how the students should be
-allocated apartments such that each study group has the {\em least
-  distance to travel} for a discussion? More specifically, we are
-interested in enforcing additional conditions, namely, that all the
-students in a study group must be next to each other; in other words,
-for one student to reach another fellow study group member's apartment
-(for all study groups the student is part of), she must not have to
-pass the apartment of any student who is not in that study group. To
-further elucidate, the apartments of students of any study group must
-be arranged in an exclusive unfragmented path on the street
-map. Exclusivity here means that the path must not have apartments
-from other study groups (unless that apartment is also part of {\em
-  this} study group).
-
-An intuitive approach to this problem would be to first find the paths
-that each study group decides to inhabit and then refine the
-allocation to individual students. A feasible allocation of exclusive
-routes to study groups is illustrated in
-Fig~\ref{fig:streetmappathpeople} \tnote[TH]{fig:streetmappath} and
-the students' allocation of apartments that obeys this route
-allocation is also shown. Table~\ref{tab:iltree} shows the same
-solution set theoretically.  How this is algorithmically computed is
-the focus of this thesis.
-
-\begin{figure}[htbp] %[t]%
- \centering
-  (a) \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
-  (b) \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
-
- \label{fig:streetmap}
- \caption{\figtabsize (a) {\residenceblock} street map (b)
-   {\residenceblock} street map with study group routes
-   allocated. Routes are color coded as follows: red for
-   \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
-   group, orange for \textcolor{yellow}{$\xBBB$} group, green for
-   \textcolor{green}{$\xTTT$} group}
-
- \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
-  \caption{\figtabsize Individual allocation of apartments to students
-    in {\residenceblock} that meets the requirements stated before.
-    The routes are color coded as follows: red for
-    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
-    group, orange for \textcolor{yellow}{$\xBBB$} group, green
-    for \textcolor{green}{$\xTTT$} group. {\tiny {\em Peanuts images
-        {\copyright} Charles Schulz}}}% YellowOrange
-  \label{fig:streetmappathpeople}
-\end{figure}
-\tnote{UPDATE IMAGE (b) 2-infinite-loop-BTWF.png. REMOVE TITLES. ADD
-  {$\xLLL$} {$\xGGG$} {$\xBBB$} {$\xTTT$}}
-
-\tnote[TH]{make dashed/textured lines for routes. make it color
-  agnostic.}
-
-As a special case, suppose all the apartments are on the same street
-or if they are all lined up on a single path, the street map becomes a
-tree that is just a path. Then the problem becomes what is called an
-{\em interval assignment problem}. The idea of interval assignment may
-not be obvious here; hence to see this, consider a different problem
-in {\WSI} where the classes for these study groups courses need to be
-scheduled during a day (or a week or any time period). Each study
-group has a bunch of courses associated with it some of which may be
-shared by two or more study groups. It is mandatory that a student who
-is a member of a study group takes all the courses associated with
-that group. There are slots during the day for classes to be held and
-the problem is to allocate class slots to courses such that all the
-classes of a study group are consecutive. It is debatable if this will
-not hamper the attention span and memory retention rate of the
-students but that is, regrettably, out of the scope of this
-thesis. The parallels between this class allocation problem and the
-accommodation problem can be seen as follows. The set $U$ here, are
-the courses offered (say Course 101 {\coneohone}, Course 102
-{\coneohtwo} and so on). In this variation of the problem, the
-collection $\cF$ is the set of study groups but the study groups are
-filled by course IDs (in place of students in the earlier
-example). For instance, Course 101 is mandatory for all study groups
-$\xLLL$, $\xGGG$, $\xBBB$, $\xTTT$ and Course 102 is mandatory for
-only the $\xLLL$ group) and so on. The sequence of class slots for the
-day (or week or any time period) is analogous to the street map in the
-accommodation problem. It is quite obvious now why this version of the
-problem (where the ``target graph'' is a path and not any
-tree\tnote[TH]{Allowing any tree in this example could be seen as a
-  scenario where there are parallel classes. A node falling in the
-  path between two other nodes would mean that the corresponding is
-  scheduled between the other two.}) is called an interval assignment
-problem.
-
-The interval assignment problem to a set system is equivalent to the
-consecutive-ones property (COP) problem in binary matrices\cite{wlh02,
-  nsnrs09}.  The COP problem is to rearrange rows (columns) of a
-binary matrix in such a way that every column (row) has its {\un}s
-occur consecutively. If this is possible the matrix is said to have
-the COP.  COP is a well researched combinatorial problem and has
-several positive results on tests for it and computing the COP
-permutation (i.e. the course schedule in the above illustration) which
-will be surveyed later in this document. Hence we are interested in
-extensions of COP, more specifically, the extension of interval
-assignment problem to tree path assignment problem (which is
-illustrated by the study group accommodation problem).
-
-\section{Basic preliminaries - general definitions and nomenclature}
-\tnote{(definitions theorems etc) needed if any
-    - graph theory}
-
-\section{Consecutive-ones Property Testing - a Brief Survey}
-\label{sec:background}
-
-In this section, a brief survey of the consecutive-ones problem and its
-optimization problems is presented.
-
-%%%%%%                              %%%%%%
-%THX is the second level of ``tnote''-ing%
-%%%%%%                              %%%%%%
-\tnote[TH]{ADD: As it will be described in detail later in this
-  document, isomorphism of certain % classes of graphs, namely chordal
-%   graphs, have a close relationship with consecutive-ones property and
-%   generalizations of it.  This is perhaps because of how closely COP
-%   of a matrix relates to properties of graphs derived from matrices as
-%   seen in the following results.
-}
-\tnote[THX]{ADD: peo exists iff chordal. lexicographic BFS
-  [tag:chordalGraph]} %
-\tnote[THX]{ADD: A well known result in %perfect graph theory is that
-%  the maximal cliques of  an interval graph $G$ can be linearly ordered
-%   such that for all $v \in V(G)$, cliques containing $v$ are
-%   consecutive in the ordering cite~{gh64}. This clearly means that a
-%   graph $G$ is an interval graph if and only if
-}% 
-\tnote[THX]{verify
-  from paper the statement of claim.}% 
-\tnote[TH]{maximal clique vertex
-  incidence matrix of%  $G$ has COP.  Also, maximal cliques of any
-%   chordal graph can be enumerated in polytime $O(m+n)$
-}
-\tnote[THX]{citation?!!}%  
-\tnote[TH]{cite~{fg65} uses these results to
-  give the first polynomial time algorithm for COT.}%
-\tnote[THX]{check. how do they use it?}%  
-\tnote[TH]{ A bipartite graph
-  is convex%  (on $R$) if and only if its half adjacency matrix has COP
-%   on rows.  The results in cite~{bl76} on COT are based on the result
-%   that interval graphs are AT-free chordal graphs.
-}%
-\tnote[THX]{the latter being Tucker's?}%  
-\tnote[THX]{(2) TBD survey
-  -- % see blue notes (in notebook) under Graph Isomorphism. namely
-%   citations in cite:aas93 (3) a brief on heirarchy of chordal graphs,
-%   path graphs, interval graphs, peo, clique tree etc. and results. (4)
-%   interval graphs are incomparability graphs. see
-%   Golumbic. [tag:classification] (5) finding min length hole in
-%   bipartite graph is polynomial time [Sec 3.3.2 in cite:d08phd] (6)
-%   theorem 2.2 in cite:d08phd - $G$ is union of v.d. caterpillars iff
-%   $M$ has COP in rows. $M$ is edge vertex incidence matrix $(*,2)$ of
-%   $G$
-}%
-
-\subsection{Matrices with COP}
-\label{sec:copmatrices}
-As seen earlier, the interval assignment problem (illustrated as the
-course scheduling problem in Section~\ref{sec:problem}), is a special
-case of the problem we address in this thesis, namely the tree path
-labeling problem (illustrated as the study group accommodation
-problem). The interval assignment problem and COP problem are equivalent
-problems. In this section we will see some of the results that
-exists in the literature today towards solving the COP problem and
-optimization problems surrounding it.
-
-\begin{figure}[t] %[htbp]
-  \centering
-
-  {\figtabsize
-    \begin{tabular}[h]{l|lcccl}
-      $M_1$: & $M_1'$: &&&& $M_2$:\\
-      &&&&&\\
-      \begin{tabular}[h]{llll}
-        $c_1$ & $c_2$ &$c_3$ &$c_4$\\
-        &&&\\
-        \un & 0   & \un & 0\\
-        0   & \un & 0   & \un \\
-        \un & 0   & 0   & \un
-      \end{tabular}
-      &
-      \begin{tabular}[h]{llll}
-        $c_3$ &$c_1$ &$c_4$& $c_2$\\
-        &&&\\
-        \un & \un & 0 & 0\\
-        0 & 0 & \un & \un \\
-        0 & \un & \un & 0
-      \end{tabular}
-      &&&&
-      \begin{tabular}[h]{llll}
-        $d_1$ & $d_2$ &$d_3$ &$d_4$\\
-        % &&&\\
-        &&&\\
-        \un & \un & 0 & 0\\
-        0 & \un & \un & 0 \\
-        0 & \un & 0 & \un 
-      \end{tabular}
-    \end{tabular}
-  }
-
-  \caption{\figtabsize Matrices with and without COP. $M_1$ has COP
-    because by permuting its columns, $c_1$-$c_4$, one can obtain
-    $M_1'$ where the {\un}s in each row are consecutive. $M_2$,
-    however, does not have COP since no permutation of its columns,
-    $d_1$-$d_4$, will arrange {\un}s in each row consecutively
-    \cite{d08phd}.}
-
-  \label{fig:cop-matrix}
-\end{figure}
-
-Recall that a matrix with COP is one whose rows (columns) can be
-rearranged so that the {\un}s in every column (row) are in consecutive
-rows (columns). Figure~\ref{fig:cop-matrix} shows examples of this
-property.  COP in binary matrices has several practical applications
-in diverse fields including scheduling \cite{hl06}, information
-retrieval \cite{k77} and computational biology \cite{abh98}.  Further,
-it is a tool in graph theory \cite{mcg04} for interval graph
-recognition, characterization of Hamiltonian graphs, planarity testing
-\cite{bl76} and in integer linear programming \cite{ht02,hl06}.
-
-The obvious first questions after being introduced to the consecutive
-ones property of binary matrices are if COP can be detected
-efficiently in a binary matrix and if so, can the COP permutation of
-the matrix also be computed efficiently?  Recognition of COP in a
-binary matrix is polynomial time solvable and the first such algorithm
-was given by \cite{fg65}.  A landmark result came a few years later
-when \cite{at72} discovered the families of forbidden submatrices that
-prevent a matrix from having COP and most, if not all, results that
-came later were based on this discovery which connected COP in binary
-matrices to convex bipartite graphs. In fact, the forbidden
-submatrices came as a corollary to the discovery that convex bipartite
-graphs are AT-free in \cite{at72}\tnote[TH]{check}. The first linear
-time algorithm for COP testing (COT) was invented by \cite{bl76} using
-a data structure called PQ trees.  Since then several COT algorithms
-have been invented -- some of which involved variations of PQ trees
-\cite{mm96,wlh01,mcc04}, some involved set theory and ICPIA
-\cite{wlh02,nsnrs09}, parallel COT algorithms\cite{as95,bs03,ly91} and
-certifying algorithms\cite{mcc04}.
-
-The construction of PQ trees in \cite{bl76} draws on the close
-relationship of matrices with COP to interval graphs. A PQ tree of a
-matrix is one that stores all row (column) permutations of the matrix
-that give the COP orders (there could be multiple orders of rows or
-columns) of the matrix. This is constructed using an elaborate linear
-time procedure and is also a test for planarity\tnote[TH]{check check
-  check. both interval graph and planarity in this paper?}.  PQR trees
-is a generalized data structure based on PQ trees \cite{mm96,mpt98}.
-\cite{tm05} describes an improved algorithm to build PQR
-trees. \tnote[TH]{improv in terms of what?}\cite{wlh02} describes the
-simpler algorithm for COT. Hsu also invented PC trees
-\cite{wlh01}\tnote[TH]{This result first appeared inproc ISAAC92}
-which is claimed to be much easier to implement. \cite{nsnrs09}
-describes a characterization of consecutive-ones property solely based
-on the cardinality properties of the set representations of the
-columns (rows); every column (row) is equivalent to a set that has the
-row (column) indices of the rows (columns) that have one entries in
-this column (row). This is interesting and relevant, especially to
-this thesis because it simplifies COT to a great degree. \tnote[TH]{it
-  reduces the solution search space. fill in the blanks.}
-
-\cite{mcc04} describes a different approach to COT. While all previous
-COT algorithms gave the COP order if the matrix has the property but
-exited stating negative if otherwise, this algorithm gives an evidence
-by way of a certificate of matrix even when it has no COP. This
-enables a user to verify the algorithm's result even when the answer
-is negative. This is significant from an implementation perspective
-because automated program verification is hard and manual verification
-is more viable. Hence having a certificate reinforces an
-implementation's credibility. Note that when the matrix {\em has} COP,
-the COP order is the certificate.  The internal machinery of this
-algorithm is related to the weighted betweenness problem
-addressed\tnote[TH]{in what way??} in \cite{co98}.  \tnote[TH]{expand
-  on the COP order graph creation and it having to be bipartite for M
-  to have COP. and thus an odd cycle being an evidence of no COP.}
-\tnote[TH]{ where should this go?: (1) cite|jlm97 (application of PQ trees
-  in graphics). (2) helly's theorem citation
-  19XXdgk-Hellystheorem-Danzer-Gruenbaum-Klee}
-
-\subsection{Optimization problems in COP}
-\label{sec:optcop}
-
-So far we have been concerned about matrices that have the consecutive
-ones property. However in real life applications, it is rare that data
-sets represented by binary matrices have COP, primarily due to the
-noisy nature of data available. At the same time, COP is not arbitrary
-and is a desirable property in practical data representation
-\cite{co98,jkckv04,k77}. In this context, there are several
-interesting problems when a matrix does not have COP but is ``close''
-to having COP or is allowed to be altered to have COP. These are the
-optimization problems related to a matrix which does not have
-COP. Some of the significant problems are surveyed in this section.
-
-\tnote[TH]{ -- sect 4.1 in cite:d08phd has many results
-  surveyed. hardness results, approx. results. results are usually for
-  a class of matrices $(a,b)$ where number {\un}s in columns and rows
-  are restriced to $a$ and $b$ . -- problem of flipping at most $k$
-  entries of $M$ to make it attain COP. this is NP complete
-  cite:b75-phd}\tnote[TH]{(1) scite:lb62 showed that interval graphs
-  are AT-free.  describe AT (2) show the close relationship b/w COP
-  and graphs sec 2.2, pg 31} \cite{at72} showed that a matrix that
-does not have COP have certain substructures that prevent it from
-having COP. Tucker classified these forbidden substructures into five
-classes of submatrices. This result is presented in the context of
-convex bipartite graphs which \cite{at72} proved to be
-AT-free\tnote[TH]{ check this up. give details. - doms'}. By
-definition, convex bipartite graph have half adjacency matrices that
-have COP on either rows or columns (graph is biconvex if it has COP on
-both)\cite{d08phd}. A half adjacency matrix is a binary matrix
-representing a bipartite graph as follows. The set of rows and the set
-of columns form the two partitions of the graph. Each row node is
-adjacent to those nodes that represent the columns that have {\un}s in
-the corresponding row. \cite{at72} proves that this bipartite graph
-has no asteroidal triple if and only if the matrix has COP and goes on
-to identify the forbidden substructures for these bipartite
-graphs. The matrices corresponding to these substructures are the
-forbidden submatrices.
-
-Once a matrix has been detected to not have COP (using any of the COT
-algorithms mentioned earlier), it is naturally of interest to find out
-the smallest forbidden substructure (in terms of number of rows and/or
-columns and/or number of entries that are {\un}s). \cite{d08phd}
-discusses a couple of algorithms which are efficient if the number of
-{\un}s in a row is small. This is of significance in the case of sparse
-matrices where this number is much lesser than the number of
-columns. $(*,\Delta)${\em -matrices} are matrices with no restriction
-on number of {\un}s in any column but has at most $\Delta$ {\un}s in
-any row. {\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are
-similar problems which deals with inducing COP on a matrix. In {\sc
-  Min COS-R (Min COS-C)} the question is to find the minimum number of
-rows (columns) that must be deleted to result in a matrix with COP.
-In the dual problem {\sc Max COS-R (Max COS-C)} the search is for the
-maximum number of rows (columns) that induces a submatrix with
-COP. Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
-a submatrix $M'$ with all columns\tnote[TH]{check if b75 deals with
-  COP col or COP row. also is it any submatrix with k less than r rows
-  or submatrix must have all columns?} but a maximum cardinality
-subset of rows such that $M'$ has COP is NP complete. \cite{hg02}
-corrects an error of the abridged proof of this reduction as given in
-\cite{gj79}.  \cite{d08phd} discusses all these problems in detail
-giving an extensive survey of the previously existing results which
-are almost exhaustively all approximation results and hardness
-results. Taking this further, \cite{d08phd} presents new results in
-the area of parameterized algorithms for this
-problem\tnote[TH]{elaborate - what are the results?}.
-
-Another problem is to find the minimum number of entries in the matrix
-that can be toggled to result in a matrix with COP.  \cite{v85}
-discusses approximation of {\sc COP Augmentation} which is the problem
-of changing of the minimum number of zero entries to {\un}s so that
-the resulting matrix has COP. As mentioned earlier, this problem is
-known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
-using a reduction to the longest path problem, \tnote[TH]{or is it a
-  survey of another result?  check.} that finding a Tucker's forbidden
-submatrix of at least $k$ rows is NP complete. \tnote[TH]{how is this
-  different from booth's 75 result??}  \tnote[TH]{where should this
-  go? cite|tz04 (approx submatrix with COP sparse matrices)}
-
-\cite{jkckv04} discusses the use of matrices with almost-COP (instead
-of one block of consecutive {\un}s, they have $x$ blocks, or {\em
-  runs}, of consecutive {\un}s and $x$ is not too large) in the
-storage of very large databases.  The problem is that of reordering of
-a binary matrix such that the resulting matrix has at most $k$ runs of
-{\un}s. This is proved to be NP hard using a reduction from the
-Hamiltonian path problem.\tnote[TH]{Theorem 2.1 in jkckv}
-\tnote[TH]{(1) A connection of COP problem to the travelling salesman
-  problem is also introduced. what does this mean? -- COP can be used
-  as a tool to reorder $0.5T \le runs(M) \le T$. (2) The optimization
-  version of the $k$-run problem, i.e. minimization of number of
-  blocks of ones is proven to be NP complete by
-  cite:k77}\tnote[TH]{are these two the same?} \tnote[TH]{what is the
-  reduction?} \tnote[TH]{other problems similar to COP -- cite:ckl96
-  (ILP, circ ones, one drop) -- cite:th98 (generalization of COP -
-  minimax, biotonic column) Tucker}
-
-\section{********* Application of COP in Areas of Graph Theory and
-  Algorithms}
-\label{sec:appcopGTA}
-
-\subsection{********* COP in Relational Database Model}
-\label{sec:apprdbm}
-\tnote{(set systems theme)}
-
-\subsection{********* COP in Graph Isomorphism}
-\label{sec:appgraphiso}
-\tnote{(canonization theme)}
-
-\subsection{********* Certifying Algorithms}
-\label{sec:appcertalgo}
-\tnote{ (certification McC04 theme)}
-
-\section{Generalization of COP - the Motivation}
-\label{sec:motive}
-Section~\ref{sec:copmatrices} introduced a succinct characterization
-for consecutive-ones property which is solely based on the cardinality
-properties of the set representations of the matrix's columns
-\cite{nsnrs09}. This result is very relevant to this thesis because
-aside from it simplifying COT to a great degree, our generalization
-problem is motivated by their results.
-
-\cite{nsnrs09} characterizes interval assignments to the sets which
-can be obtained from a single permutation of the rows.  For an
-assignment to be feasible, the cardinality of the interval assigned to
-each set in the system must be same as the cardinality of the set, and
-the intersection cardinality of any two intervals must be same as the
-intersection cardinality of their corresponding sets.  While this is
-obviously a necessary condition, this result shows this is also
-sufficient.  \cite{nsnrs09} calls this an Intersection Cardinality
-Preserving Interval Assignment (ICPIA).  This paper generalizes the
-idea from \cite{wlh02} of decomposing a given binary matrix into prime
-matrices for COT and describes an algorithm to test if an ICPIA exists
-for a given set system.
-
-The equivalence of the problem of testing for the consecutive-ones
-property to the constraint statisfaction problem of interval
-assignment \cite{nsnrs09} or interval labeling \cite{kklv10} is as
-follows. Every column (row) of the binary matrix can be converted into
-a set of non-negative integers which are the indices of rows (columns)
-with {\un}s in that column (row). It is apparent that if the matrix
-has COP in columns (rows), then constructing such sets after applying
-the COP permutation to the rows (columns) of the matrix will result in
-sets with consecutive integers. In other words, after application of
-COP reordering, the sets are intervals. Indeed the problem now becomes
-finding interval assignments to a given set system such that there
-exists a permutation of the universe of set of row indices (column
-indices) which converts each set to its assigned interval.
-
-The problem of interest in this thesis, namely, tree path labeling
-problem, is a natural generalization of the interval assignment
-problem or the COT problem. The problem is defined as follows -- given
-a set system $\cF$ from a universe $U$ and a target tree $T$, does
-there exist a bijection from $U$ to the vertices of $T$ such that each
-set in the system maps to a path in $T$.  We refer to this as the
-{\CFTPL} problem or simply {\em tree path labeling} problem for an
-input set system and target tree pair -- $(\cF,T)$. The special case
-of the target tree being a path, is the interval assignment problem.
-We focus on generalizing the notion of an ICPIA \cite{nsnrs09} to
-characterize feasible path assignments.  We show that for a given set
-system $\cF$, a tree $T$, and an assignment of paths from $T$ to the
-sets, there is a feasible\footnote{The notion of {\em feasibility} is
-  formally defined in Section~\ref{ch:prelims}.}  bijection between
-$U$ and $V(T)$ if and only if the intersection cardinalities among any
-three sets (not necessarily distinct) is equal to that of the
-corresponding paths assigned to them and the input passes a filtering
-algorithm (described in this paper) successfully.  This algorithmic
-characterization gives a natural data structure that stores all the
-\remove{relevant} feasible bijections between $U$ and $V(T)$. This
-reduces the search space for the solution considerably from the
-universe of all possible bijections between $U$ and $V(T)$ to only
-those bijections that maintain the characterization.  Further, the
-filtering algorithm is also an efficient algorithm to test if a tree
-path labeling\footnote{The terms {\em tree path labeling} and {\em
-    tree path assignment} are, in informal language,
-  synonyms. Formally, the former refers to the bijection $\cl: \cF
-  \rightarrow \cP$. The latter refers to the set of ordered pairs
-  $\{(S, P) \mid S \in \cF, P \in \cP\}$. $\cP$ is a set of paths on
-  $T$.} is feasible.
-
-\section{Summary of New Results in this Thesis}
-\label{sec:results}
-
-We see in Section~\ref{sec:motive} that pairwise intersection
-cardinality preservation is necessary and sufficient for an interval
-assignment to be feasible for a given hypergraph\footnote{A {\em
-    hypergraph} is an alternate representation of a set system and
-  will be used in this thesis. See Section~\ref{ch:prelims} for the
-  formal definition.} and thus is a characterization for COP
-\cite{nsnrs09}. \tnote{Refer to the survey section that will list the
-  theorems and/or lemma?} In our work we extend this characterization
-and find that trio-wise intersection cardinality preservation makes a
-tree path labeling \footnote{A {\em tree path labeling} $\cl$ is a
-  bijection of paths from the target tree $T$ to the hyperedges in
-  given hypergraph $\cF$.  See Section~\ref{ch:prelims} for the formal
-  definition.}  (TPL) feasible, which is a generalization of the COP
-problem. This problem is defined as follows.
-
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \FTPL}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
-        paths $\cP$ from $T$ and a
-        bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
-      \end{minipage}\\
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
-        such that $\phi$ when applied on any hyperedge in $\cF$ will
-        give
-        the path mapped to it by the given tree path labeling $\cl$.\\
-        { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every
-          hyperedge $S \in \cF$.}
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
-
-We give a necessary and sufficient condition by way of {\em
-  Intersection Cardinality Preservation Path Labeling} (ICPPL) and a
-filtering algorithm for {\FTPL} to output in affirmative. ICPPL
-captures the trio-wise cardinality property described
-earlier\footnote{See Section~\ref{sec:feasible} for the definition of
-  ICPPL.}. This characterization can be checked in polynomial time.  A
-relevant consequence of this constructive procedure is that it is
-sufficient to iteratively check if three-way intersection
-cardinalities are preserved.  In other words, in each iteration, it is
-sufficient to check if the intersection of any three hyperedges is of
-the same cardinality as the intersection of the corresponding paths.
-Thus this generalizes the well studied question of the feasible
-interval assignment problem which is the special case when the target
-tree $T$ is simply a path \cite{wlh02,nsnrs09}.
-
-Aside from checking if a given TPL is feasible, we also solve the
-problem of computing a feasible TPL for a given hypergraph and target
-tree, if one exists. This problem, {\CFTPL}, is defined as follows.
-
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \CFTPL}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$ and a tree $T$.\\
-      \end{minipage}\\
-
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a set of paths $\cP$ from $T$ and a bijection
-        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
-        {\bf true} on $(\cF, T, \cl)$.
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
-
-We present a polynomial time algorithm for {\CFTPL} when the target
-tree $T$ belongs to a special class of trees called {\em \kstar s} and
-when the hyperedges in the hypergraph $\cF$ have at most $k+2$
-vertices. A couple of examples of {\kstar s} are given in
-Figure~\ref{fig:ksubstar}.
-
-\begin{figure}[t]
-  \centering
-  \begin{tabular}[h]{ccccc}
-    \includegraphics[scale=0.3]{../img/star.pdf} &&&&
-    \includegraphics[scale=0.3]{../img/kstar.pdf}\\
-    (a) &&&& (b)
-  \end{tabular}
-  \caption{\figtabsize Examples of {\kstar s}. (a) $k = 0$ (b) $k = 2$
-  }
-  \label{fig:ksubstar}
-\end{figure}
-
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \CFTPLKTREE}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$ such that every
-        hyperedge
-        $S \in \cF$ is of cardinality at most $k+2$ and a {\kstar} $T$.\\
-      \end{minipage}\\
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a set of paths $\cP$ from $T$ and a bijection
-        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
-        {\bf true} on $(\cF, T, \cl)$.
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
-
-\tnote{The following sentence seems out of place in this para} In
-spite of this being a restricted case, we believe that our results are
-of significant interest in understanding the nature of {\sc Graph
-  Isomorphism} which is polynomial time solvable in interval graphs
-while being hard on path graphs\cite{kklv10}. {\kstar s} are a class
-of trees which are in many ways very close to intervals or paths. Each
-ray\footnote{The path from a leaf to the root, the vertex with highest
-  degree, is called a {\em ray} of the \kstar. See
-  Section~\ref{ch:prelims}.} are independent except for the
-root\footnote{The vertex with maximum degree in a {\kstar} is called
-  {\em root}. See Section~\ref{ch:prelims}.} and hence can be
-considered as an independent interval till the root. Our algorithm
-builds on this fact and uses the interval assignment
-algorithm\cite{nsnrs09} up until ``reaching'' the root and then uses
-the trio-wise intersection cardinality (the extra condition in ICPPL
-that generalizes ICPIA) check to resolve the ambiguity about which ray
-the algorithm should ``grow'' the solution into in the next iteration.
-
-We also have an algorithm for solving {\CFTPL} with no restrictions on
-the target tree or set size which runs in exponential time.  This
-algorithm finds a path labeling from $T$ by decomposing the problem
-into subproblems of finding path labeling of subsets of $\cF$ from
-subtrees of $T$. Given the fact that binary matrices naturally
-represent a set system (see Section~\ref{sec:motive}) and that the
-{\em overlap}\footnote{}\tnote{Give informal definition of ``overlap''
-  in footnote and link to sec:prelims. ADD the definition in prelim.}
-relation between the sets involved is an obvious equivalence relation,
-$\cF$ quite naturally partitions into equivalence classes known as
-{\em overlap components}\tnote{Give informal definition of ``overlap
-  components'' in footnote and link to sec:prelims. ADD the definition
-  in prelim.}. In the context of COP, overlap components were used in
-\cite{wlh02} and \cite{kklv10}. Moreover, \cite{nsnrs09} discovered
-that these equivalence classes form a total order \tnote{Give informal
-  definition of ``total order'' in footnote?}. We extend this to TPL
-and find that when $\cF$ is a path hypergraph\footnote{If there exists
-  an FTPL for a hypergraph $\cF$, it is called a path hypergraph.},
-the classes can be partially ordered as an in-tree in polynomial
-time. Once $\cF$ is ``broken'' into overlap components, one must
-identify the subtree of $T$ that it needs to map to and this is the
-hard part which is currently open to be solved in polynomial time.
-
-\tnote[TH]{The connection of TPL to graph isomorphism will be made
-  later in the document}
-
-\chapter{Consecutive-ones Property - a Survey of Important Results}
-\label{ch:copsurvey}
-\tnote{TBD: have a few lines about organization of chapter}
-
-\tnote{ Survey chapter: is the full fledged expansion of the survey in
-  introduction with details, observations, theorems etc.}
-
-\section{Matrices with COP}
-\label{sec:surveycoptest}
-\tnote{Expand on ref:sec:copmatrices}
-
-\section{Optimization problems in COP}
-\label{sec:surveycopopt}
-\tnote{Expand on ref:sec:optcop}
-
-\section{********* COP in Relational Database Model}
-\label{sec:surveyrdbm}
-\tnote{Expand on sec:apprdbm}
-\tnote{(set systems theme)}
-
-\subsection{********* COP in Graph Isomorphism}
-\label{sec:surveygraphiso}
-\tnote{Expand on sec:appgraphiso}
-\tnote{(canonization theme)}
-
-\subsection{********* Certifying Algorithms}
-\label{sec:surveycertalgo}
-\tnote{Expand on sec:appcertalgo}
-\tnote{ (certification McC04 theme)}
-
-\chapter{Tree Path Labeling of Path Hypergraphs - the New Results}
-\label{ch:myresearch}
-
-\tnote{UPDATE TO FINAL VERSION. BELOW IS OLDER VERSION WITH INCOMPLETE
-  SEC 4.}  This chapter documents all the new results obtained by us in the
-area of tree path labeling of path hypergraphs which is the parent
-problem addressed in this thesis. In Section~\ref{sec:motive} we see
-that consecutive-ones property and its equivalent problem of interval
-labeling of a hypergraph is a special case of the general problem of tree
-path labeling of path hypergraphs.
-
-\tnote{UPDATE the following para to current skeleton.}The necessary
-preliminaries with definitions etc. are presented in
-Section~\ref{ch:prelims}. Section~\ref{sec:feasible} documents the
-characterization of a feasible path labeling for any path labeling
-from any target trees. Section~\ref{sec:spltargettree} describes the
-special case where the target tree is of a particular family of trees
-and specifically, Section~\ref{sec:ksubdivstar} describes a polynomial
-time algorithm to find the tree path labeling of a given set system
-from a given $k$-subdivided tree. Section~\ref{sec:norestraint}
-describes the general situation where the target tree has no
-restrictions and the algorithm to find a TPL, if any, in this case.
-
-\section{Introduction}
-
-The problem of consecutive-ones property testing can be easily seen as
-a simple constraint satisfaction problem involving a hypergraph or a
-system of sets from a universe. Every column (row) of the binary
-matrix can be converted into a set of non-negative integers which are
-the indices of the rows (columns) with {\un}s in that column
-(row). When observed in this context, if the matrix has the COP on
-columns (rows), a reordering of its rows (columns) will result in sets
-that have only consecutive integers. In other words, the sets after
-applying the COP row (column) permutation are intervals.  In this
-form, one can see that this is indeed the problem of finding interval
-assignments to the given set system \cite{nsnrs09} with a single
-permutation of the universe (set of row or column indices for COP of
-columns or rows, respectively) which permutes each set to its
-interval. The result in \cite{nsnrs09} characterizes interval
-assignments to the sets which can be obtained from a single
-permutation of the rows (columns). They show that for each set, the
-cardinality of the interval assigned to it must be same as the
-cardinality of the set, and the intersection cardinality of any two
-sets must be same as the intersection cardinality of the corresponding
-intervals. This is a necessary and sufficient
-condition. \tnote{(ICPIA) put reference to survey chapter}.  Finally,
-the idea of decomposing a given binary matrix into prime matrices to
-check for COP is adopted from cite|wlh02| to test if an ICPIA exists
-for a given set system.\tnote{REPEATITION from intro chapter. but idea
-  recall is necessary. so keep it but reword it to avoid monotone.}
-
-A natural generalization of the interval assignment problem is
-feasible tree path labeling problem of a set system. The problem is
-defined as follows -- given a set system $\cF$ from a universe $U$ and
-a tree $T$, does there exist a bijection from $U$ to the vertices of
-$T$ such that each set in the system maps to a path in $T$.  We refer
-to this as the {\em tree path labeling problem} for an input set
-system, target tree pair -- $(\cF,T)$. \tnote{USE THE FORMAL PROBLEM DEF BLOCK}
-%\begin{tabular}[t]{c}
-{\small
-\begin{minipage}[h]{5in}
- % \begin{singlespace}
- \vspace{2mm}
- {\large \FTPL}\\
- \begin{tabular}[t]{l|l}
- \hline\\
-    {\tt Input} & 
-    \begin{minipage}[t]{\probdefwidth}
-      A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
-      paths $\cP$ from $T$ and a
-      bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
-    \end{minipage}\\
-    {\tt Question} &
-    \begin{minipage}[t]{\probdefwidth}
-      Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
-      such that $\phi$ when applied on any hyperedge in $\cF$ will
-      give
-      the path mapped to it by the given tree path labeling $\cl$.\\
-      { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every hyperedge
-        $S \in \cF$.}
-    \end{minipage}\\
-  \end{tabular}
-  % \end{singlespace}
-\end{minipage}\\
-}
-%\end{tabular}
-
-As a special case if the tree
-$T$ is a path, the problem becomes the interval assignment problem. \tnote{USE THE FORMAL PROBLEM DEF BLOCK}
-We focus on the question of generalizing the notion of an ICPIA
-\cite{nsnrs09} to characterize feasible path assignments.  We show
-that for a given set system $\cF$, a tree $T$, and an assignment of
-paths from $T$ to the sets, there is a feasible bijection between $U$
-and $V(T)$ if and only if all intersection cardinalities among any
-three sets (not necessarily distinct) is same as the intersection
-cardinality of the paths assigned to them and the input runs a
-filtering algorithm (described in this paper) successfully.  This
-characterization is proved constructively and it gives a natural data
-structure that stores all the relevant feasible bijections between $U$
-and $V(T)$.  Further, the filtering algorithm is also an efficient
-algorithm to test if a tree path labeling to the set system is
-feasible.  This generalizes the result in \cite{nsnrs09}.
-
-In the later part of this paper, we focus on a new special case of the
-tree path labeling problem. Here the set system is such that every set
-is at most $k+2$ in size and for every pair of sets in it there exists
-a sequence of sets between them with consecutive sets in this sequence
-having a strict intersection -- i.e., non-empty intersection with
-neither being contained in the other. Moreover, the given tree is a
-{\em $k$-subdivided star}. We demonstrate a polynomial time algorithm
-to find a feasible path labeling in this case.
-
-\section{Preliminaries to new results}
-\label{ch:prelims}
-
-\tnote{TBD: have a few lines about organization of chapter}This
-section states definitions and basic facts necessary in the scope of
-this document.
-
-\tnote{Move some of this to general prelim in chapter 1}
-
-The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
-  system} of a universe $U$ with $|U| = n$.  The {\em support} of a
-set system $\F$ denoted by $supp(\cF)$ is the union of all the sets in
-$\F$; $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this
-paper, a set system is required to ``cover'' the universe; $
-supp(\cF) = U$.
-
-The graph $T$ represents a {\em target tree} with same
-number of vertices as elements in $U$; $|V(T)| = n$.  A {\em path
-  system}\, $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P
-\mid P \subseteq V, \text{ } T[P] \text{ is a path} \}$.
-
-% 
-% A set system $\cF$ can be alternatively represented by a {\em
-%   hypergraph} $\H_\cF$ whose vertex set is $supp(\cF)$ and
-% hyperedges are the sets in $\cF$. This is a known representation for
-% interval systems in literature \cite{bls99,kklv10}.  We extend this
-% definition here to path systems.
-
-A set system $\cF$ can be alternatively represented by a {\em
-  hypergraph}\, $\cF_H$ whose vertex set is $supp(\cF)$ and hyperedges
-are the sets in $\cF$. This is a known representation for interval
-systems in literature \cite{bls99,kklv10}.  We extend this definition
-here to path systems. Due to the equivalence of set system and
-hypergraph in the scope of this paper, we drop the subscript $_H$ in
-the notation and refer to both the structures by $\cF$.
-
-Two hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to
-each other, denoted by $\cF' \cong \cF''$, iff there exists a
-bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for all
-sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ iff $B$ is
-a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
-\cite{kklv10}. This is called {\em hypergraph
-  isomorphism}. \tnote{also extend $\phi$ to hyperedges -- see if
-  required }
-
-The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph $\cF$ is a
-graph such that its vertex set has a bijection to $\cF$ and there
-exists an edge between two vertices iff their corresponding hyperedges
-have a non-empty intersection \cite{mcg04}.
-
-If the intersection graphs of two hypergraphs are isomorphic,
-$\bI(\cF) \cong \bI(\cP)$ where $\cP$ is also a path system, then the
-bijection $\cl: \cF \rightarrow \cP$ due to this isomorphism is called
-a {\em path labeling} of the hypergraph $\cF$. To illustrate further,
-let $\cg: V(\cF) \rightarrow V(\cP)$ be the above mentioned
-isomorphism where $V(\cF)$ and $V(\cP)$ are the vertex sets that
-represent the hyperedges for each hypergraph respectively, $V(\cF) =
-\{ v_S \mid S \in \cF\}$ and $V(\cP) = \{ v_P \mid P \in \cP\}$. Then
-the path labeling $\cl$ is defined as follows: $\cl(S_1) = P_1$ iff
-$\cg (v_{S_1}) = v_{P_1}$. The path system $\cP$ may be alternatively
-denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In most scenarios in
-this paper, what is given are the pair $(\cF, \cl)$ and the target
-tree $T$; hence this notation will be used more often.
-
-If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$
-is called a {\em path hypergraph} and $\cP$ is called {\em path
-  representation} of $\cF$. If this isomorphism is $\phi: supp(\cF)
-\rightarrow V(T)$, then it is clear that there is an {\em induced path
-labeling} $\cl_\phi: \cF \rightarrow \cP$ to the set system;
-$\cl_\phi(S) = \set{y \mid y = \phi(x), x \in S}$ for all $S \in \cF$. Recall that
-$supp(\cP) = V(T)$.
-
-A path labeling $(\cF, \cl)$ is defined to be {\em
-  feasible} if
-$\cF \cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
-\rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
-\rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.
-
-An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is
-a graph such that its vertex set has a bijection to $\cF$ and there
-exists an edge between two of its vertices iff their corresponding
-hyperedges overlap. Two hyperedges $S$ and $S'$ are said to {\em
-  overlap}, denoted by $S \overlap S'$, if they have a non-empty
-intersection and neither is contained in the other; $S \overlap S'
-\text{ iff } S \cap S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq
-S$. Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$ and not
-necessarily connected. Each connected component of $\bO(\cF)$ is
-called an {\em overlap component}.
-
-A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
-\overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
-\cite{kklv10}. Additionally, if $S$ is such that it is contained in no
-other hyperedge in $\cF$, i.e., it is inclusion maximal then it is called
-{\em super-marginal}.
-
-A {\em star} graph is a complete bipartite graph $K_{1,p}$
-which is clearly a tree and $p$ is the number of leaves. The vertex
-with maximum degree is called the {\em center} of the star and the
-edges are called {\em rays} of the star. A {\em $k$-subdivided star}
-is a star with all its rays subdivided exactly $k$ times. The
-definition of a {\em ray of a $k$-subdivided star} is extended to the path
-from the center to a leaf. It is clear that all rays are of length $k+2$.
-
-\section{Characterization of Feasible Tree Path  Labeling} 
-\label{sec:feasible}
-
-In this section we give an algorithmic characterization of a
-feasibility of tree path labeling.  Consider a path labeling $(\cF,
-\cl)$ on the given tree $T$. We call $(\cF, \cl)$ an {\em Intersection
-  Cardinality Preserving Path Labeling (ICPPL)} if it has the
-following properties.
-
-\begin{enumerate}[{(\icpplpr\ }i) \ \ \ ]
-\item \label{pr:i} $|S| = |\cl(S)|$ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $S \in \cF$
-  % \vspace{\topshrink}
-\item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ \ \ \ \
-  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all distinct $S_1, S_2 \in \cF$
-  % \vspace{\topshrink}
-\item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
-  \cap \cl(S_3)|$ \ \ \ for all distinct $S_1, S_2, S_3 \in \cF$
-\end{enumerate}
-
-The following lemma is useful in subsequent arguments. 
-\begin{lemma}
-  \label{lem:setminuscard}
-  If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
-  (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
-  \cl(S_3))|$.
-\end{lemma}
-\begin{proof}%[Proof of Lemma~\ref{lem:setminuscard}]
-  Let $P_i = \cl(S_i)$, for all $1 \le i \le  3$.
-  $|S_1 \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| =
-  |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to properties (ii)
-  and (iii) of ICPPL, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
-  \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
-  = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. \qed
-\end{proof}
-
-In the remaining part of this section we show that $(\cF,
-\cl)$ is feasible if and only if it is an ICPPL and
-Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
-function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two levels of
-filtering of $(\cF, \cl)$ to make it simpler while retaining the set
-of isomorphisms, if any, between $\cF$ and $\cF^\cl$.
-% One direction of this claim isclear: that if a path
-% labeling is feasible, then all intersection cardinalities are
-% preserved, i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
-% \annote{has no premature exit condition hence any input will go
-%   through it}{Prove that the filtered sets has ICPPL iff input PL
-%   has ICPPL?}. Algorithm~\ref{leafasgn} has an exit condition at
-% line~\ref{xempty}. It can be easily verified that $X$ cannot be
-% empty if $\cl$ is a feasible path labeling. The reason is that a
-% feasible path labeling has an associated bijection between
-% $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that
-% the sets map to paths, ``preserving'' the path labeling.  The rest
-% of the section is devoted to constructively proving that it is
-% sufficient for a path labeling to be an ICPPL and pass the two
-% filtering algorithms.  To describe in brief, the constructive
-% approaches refine an ICPPL iteratively, such that at the end of each
-% iteration we have a ``filtered'' path labeling, and finally we have
-% a path labeling that defines a family of bijections from $supp(\cF)$
-% to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.
-First, we present Algorithm~\ref{perms} or {\tt filter\_1}, and prove
-its correctness.  This algorithm refines the path labeling by
-processing pairs of paths in $\cF^\cl$ that share a leaf until no two
-paths in the new path labeling share any leaf.
-
-\begin{algorithm}[h]
-  \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
-  \label{perms}
-  \begin{algorithmic}[\lndisplay]
-    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$\\
-    \STATE $j \assign 1$\\
-    \WHILE {there is $S_1, S_2 \in \cF_{j-1}$ such that
-      $\cl_{j-1}(S_1)$ and $\cl_{j-1}(S_2)$ have a common leaf in
-      $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
-    \{S_1, S_2\})
-    \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1 \}$ \label{setbreak} 
-    \COMMENT {Remove $S_1$, $S_2$ and add the ``filtered'' sets}
-    \STATE {\bf for} every $S \in \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
-    S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\
-
-    \STATE $\cl_j(S_1 \cap S_2) \assign \cl_{j-1}(S_1) \cap
-    \cl_{j-1}(S_2)$
-    \COMMENT {Carry forward the path labeling for all existing sets other than
-      $S_1$, $S_2$}
-    \STATE $\cl_j(S_1 \setminus S_2) \assign \cl_{j-1}(S_1) \setminus
-    \cl_{j-1}(S_2)$ 
-    \COMMENT {Define path labeling for new sets}
-    \STATE $\cl_j(S_2 \setminus S_1) \assign \cl_{j-1}(S_2) \setminus
-    \cl_{j-1}(S_1)$
-
-    \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of ICPPL}
-    \label{ln:3waycheck}
-    \STATE {\bf exit} \label{ln:exit1} \\
-    \ENDIF
-
-    \STATE $j \assign j+1$\\
-    \ENDWHILE
-    \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
-    \RETURN $(\cF', \cl')$
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma} 
- \label{lem:feasible} 
- In Algorithm~\ref{perms}, if input $(\cF, \cl)$ is a feasible path
- assignment then at the end of $j$th iteration of the {\bf while}
- loop, $j \ge 0$, $(\cF_j, \cl_j)$ is a feasible path assignment.
-\end{lemma}
-\begin{proof}%[Proof of Lemma~\ref{lem:feasible}]
-  We will prove this by mathematical induction on the number of
-  iterations. The base case $(\cF_0, \cl_0)$ is feasible since it is
-  the input itself which is given to be feasible. Assume the lemma is
-  true till $j-1$th iteration. i.e. every hypergraph isomorphism
-  $\phi: supp\left(\cF_{j-1}\right) \rightarrow V\left(T \right)$ that
-  defines $(\cF, \cl)$'s feasibility, is such that the induced path
-  labeling on $\cF_{j-1}$, $\cl_{\phi[{\cF_{j-1}}]}$ is equal to
-  $\cl_{j-1}$. We will prove that $\phi$ is also the bijection that
-  makes $(\cF_j, \cl_j)$ feasible. Note that $supp(\cF_{j-1}) =
-  supp(\cF_{j})$ since the new sets in $\cF_j$ are created from basic
-  set operations to the sets in $\cF_{j-1}$. For the same reason and
-  $\phi$ being a bijection, it is clear that when applying the $\phi$
-  induced path labeling on $\cF_j$, $ \cl_{\phi[{\cF_{j}}]}(S_1
-  \setminus S_2) = \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
-  \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Now observe that $ \cl_j(S_1
-  \setminus S_2) = \cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2) =
-  \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
-  \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Thus the induced path labeling
-  $\cl_{\phi[{\cF_{j}}]} = \cl_{j}$. Therefore lemma is proven.  \qed
-\end{proof}
-
-\begin{lemma}
-  \label{lem:invar1} In Algorithm~\ref{perms}, at the end of $j$th
-  iteration, $j \ge 0$, of the {\bf while} loop, the following
-  invariants are maintained.
-  \begin{enumerate}[I {\ }] %\vspace{\topshrink}
-  \item $\cl_j(R)$ is a path in $T$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ for all $R \in \cF_j$%\vspace{\topshrink}
-  \item $|R| = |\cl_j(R)|$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $R \in
-    \cF_j$%\vspace{\topshrink}
-  \item $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ for all $R, R' \in \cF_j$%\vspace{\topshrink}
-  \item $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
-    \cl_j(R'')|$, \ \ \ for all $R, R', R'' \in \cF_j$
-  \end{enumerate}
-\end{lemma}
-
-\begin{proof}
- Proof is by induction on the number of iterations, $j$. In this
-  proof, the term ``new sets'' will refer to the sets added to $\cF_j$
-  in $j$th iteration in line~\ref{setbreak} of Algorithm~\ref{perms},
-  $S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1$ and its
-  images in $\cl_j$ where $\cl_{j-1}(S_1)$
-  and $\cl_{j-1}(S_2)$ intersect and share a leaf.\\
-  The invariants are true in the base case $(\cF_0, \cl_0)$, since it
-  is the input ICPPL.  Assume the lemma is true till the $j-1$th
-  iteration. Let us consider the possible cases for each of the above invariants for
-  the $j$th iteration.
-
- \begin{enumerate}[\xbullet]
-  \item {\em Invariant} I/II
-    \begin{enumerate}[{I/II}a $|$] % \textbullet 
-    \item {\em $R$ is not a new set.} It is in $\cF_{j-1}$. Thus
-      trivially true by induction hypothesis.
-    \item {\em $R$ is a new set.} If $R$ is in $\cF_{j}$ and not in
-      $\cF_{j-1}$, then it must be one of the new sets added in
-      $\cF_j$. In this case, it is clear that for each new set, the
-      image under $\cl_j$ is a path since by definition the chosen
-      sets $S_1$, $S_2$ are from $\cF_{j-1}$ and due to the while loop
-      condition, $\cl_{j-1}(S_1)$, $\cl_{j-1}(S_2)$ have a
-      common leaf. Thus invariant I is proven.\\
-      Moreover, due to induction hypothesis of invariant III and the
-      definition of $l_j$ in terms of $l_{j-1}$, invariant II is
-      indeed true in the $j$th iteration for any of the new sets.  If
-      $R = S_1 \cap S_2$, $|R| = |S_1 \cap S_2| = |\cl_{j-1}(S_1) \cap
-      \cl_{j-1}(S_2)| = |\cl_j(S_1 \cap S_2)| = |\cl_j(R)|$.
-      If $R = S_1 \setminus S_2$, $|R| = |S_1 \setminus S_2| = |S_1| -
-      |S_1 \cap S_2| = |\cl_{j-1}(S_1)| - |\cl_{j-1}(S_1) \cap
-      \cl_{j-1}(S_2)| = |\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)| =
-      |\cl_j(S_1 \setminus S_2)|
-      = |\cl_j(R)|$. Similarly if $R = S_2 \setminus S_1$.\\
-    \end{enumerate}
-  \item {\em Invariant} III
-    \begin{enumerate}[{III}a $|$]
-    \item {\em $R$ and $R'$ are not new sets.} It is in
-      $\cF_{j-1}$. Thus trivially true by induction hypothesis.
-    \item {\em Only one, say $R$, is a new set.} Due to invariant IV
-      induction hypothesis, Lemma~\ref{lem:setminuscard} and
-      definition of $\cl_j$, it follows that invariant III is true no
-      matter which of the new sets $R$ is equal to. If $R = S_1 \cap
-      S_2$, $|R \cap R'| = |S_1 \cap S_2 \cap R'| = |\cl_{j-1}(S_1)
-      \cap \cl_{j-1}(S_2) \cap \cl_{j-1}(R')| = |\cl_j(S_1 \cap S_2)
-      \cap \cl_j(R')| = |\cl_j(R) \cap \cl_j(R')|$.  If $R = S_1
-      \setminus S_2$, $|R \cap R'| = |(S_1 \setminus S_2) \cap R'| =
-      |(\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)) \cap \cl_{j-1}(R')| =
-      |\cl_{j}(S_1 \cap S_2) \cap \cl_{j}(R')| = |\cl_{j}(R) \cap
-      \cl_{j}(R')|$. Similarly, if $R = S_2 \setminus
-      S_1$. Note $R'$ is not a new set.\\
-
-    \item {\em $R$ and $R'$ are new sets.} By definition, the new
-      sets and their path images in path label $\cl_j$ are disjoint so
-      $|R \cap R'| = |\cl_j(R) \cap \cl_j(R)| = 0$. Thus case proven.
-    \end{enumerate}
-  \item {\em Invariant} IV
-
-    Due to the condition in line~\ref{ln:3waycheck}, this invariant is
-    ensured at the end of every iteration.
-%     \begin{enumerate} [{Case 3.}1:]
-%     \item {\em $R$, $R'$ and $R''$ are not new sets.} Trivially
-%       true by induction hypothesis.
-%     \item {\em Only one, say $R$, is a new set.}
-%       If $R = S_1 \cap S_2$,  from Lemma~\ref{lem:fourpaths} and
-%       invariant III hypothesis,  this case is proven. Similarly if $R$
-%       is any of the other new  sets, the case is proven by also using
-%       Lemma ~\ref{lem:setminuscard}.
-%     \item {\em At least two of $R, R', R''$ are new sets.}
-%       The new sets are disjoint hence this case is vacuously true.
-%     \end{enumerate}
-  \end{enumerate} \qed
-%\vspace{-6mm}
-
-\end{proof}
-
-\begin{lemma}
-  \label{lem:noexit1}
-  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{perms} is
-  feasible, then the set of hypergraph isomorphism functions that
-  defines $(\cF, \cl)$'s feasibility is the same as the set that
-  defines $(\cF_j, \cl_j)$'s feasibility, if any.  Secondly, for any
-  iteration $j > 0$ of the {\em \bf while} loop, the {\em \bf exit}
-  statement in line~\ref{ln:exit1} will not execute.
-\end{lemma}
-\begin{proof}
-  Since $(\cF,\cl)$ is feasible, by Lemma~\ref{lem:feasible}
-  $(\cF_j,\cl_j)$ for every iteration $j > 0$ is feasible.  % Therefore,
-%   every hypergraph isomorphism $\phi: supp(\cF) \rightarrow V(T)$ that
-%   induces $\cl$ on $\cF$ also induces $\cl_{j-1}$ and $\cl_{j}$ on
-%   $\cF_{j-1}$ and $\cF_{j}$ respectively, i.e., $\cl_{\phi[\cF_{j-1}]}
-%   = \cl_{j-1}$ and $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
-%   that for all $x \in supp(\cF)$, for all $v \in V(T)$ the following
-%   hold true.
-  Also, every hypergraph isomorphism $\phi: supp(\cF) \rightarrow
-  V(T)$ that induces $\cl$ on $\cF$ also induces $\cl_{j}$ on
-  $\cF_{j}$, i.e., $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
-  that for all $x \in supp(\cF)$, for all $v \in V(T)$, if $(x,v) \in
-  \phi$ then $v \in \cl_{j}(S)$ for all $S \in \cF_{j}$ such that $x
-  \in S$.
-% the following
-%   hold true.
-%   \begin{enumerate}[i. ]
-%   \item If $(x,v) \in \phi$ then $v \in \cl_{j-1}(S)$ for all $S \in
-%     \cF_{j-1}$ such that $x \in S$.
-%   \item If $(x,v) \in \phi$ then $v \in \cl_{j}(S)$ for all $S \in
-%     \cF_{j}$ such that $x \in S$
-%   \end{enumerate}
-  In other words, filter 1 outputs a filtered path labeling that
-  ``preserves''
-  hypergraph isomorphisms of the original path labeling.\\
-  Secondly, line~\ref{ln:exit1} will execute iff the exit condition in
-  line~\ref{ln:3waycheck}, i.e. failure of three way intersection
-  preservation, becomes true in any iteration of the {\em \bf while}
-  loop.  Due to Lemma~\ref{lem:invar1} Invariant IV, the exit
-  condition does not occur if the input is a feasible ICPPL.\qed
-
-%   such that $\phi(x) = v$ where $v$ is the leaf considered in the
-%   first iterations of while. Clearly, $\phi$ is a renaming of
-%   vertices in hypergraph $\cF$ to those in hypergraph $\cF^\cl$. Thus
-%   the following facts can be observed in every iteration of the loop.
-
-%   \begin{enumerate}[\hspace{2mm}i. ] \vspace{\topshrink}
-%   \item all intersection cardinalities are preserved in this path
-%     labeling \vspace{\topshrink}
-%   \item element $x$ is exclusive in a hyperedge in $\cF$ since $v$ is
-%     exclusive in a hyperedge in $\cF^\cl$.
-%   \end{enumerate}
-
-%   Thus the exit condition is never rendered true after $x$ and $v$ are
-%   removed from their respective hyperedges. \qed
-
-% \noindent
-% This proof uses mathematical induction on the number
-%   of iterations $j$, $j \ge 0$, of the loop that executed
-%   without exiting. The base case, $j = 0$ is obviously true since the
-%   input is an ICPPL and the exit condition cannot hold true due to
-%   ICPPL property (iii).  Assume the algorithm executes till the end
-%   of $j-1$th iteration without exiting at line
-%  ~\ref{ln:3waycheck}. Consider the $j$th iteration. From Lemma
-%  ~\ref{lem:feasible} we know that $(\cF_j, \cl_j)$ and $(\cF_{j-1},
-%   \cl_{j-1})$ are feasible\remove[AS]{and from the proof in lemma
-%     lem:invar1 we know that $(\cF_{j-1}, \cl_{j-1})$ satisfies all the
-%     invariants defined in the lemma}.  Thus there exists a bijection
-%   $\phi: supp(\cF) \rightarrow V(T)$ such that the induced path
-%   % labeling on $\cF_{j-1}$ $\cl_{\phi[\cF_{j-1}]} = \cl_{j-1}$.
-%   labeling on $\cF_{j}$, $\cl_{\phi[\cF_{j}]}$ and on $\cF_{j-1}$,
-%   $\cl_{\phi[\cF_{j-1}]}$ are equal to $\cl_{j}$ and $\cl_{j-1}$
-%   respectively.  We need to prove that for any $R, R', R'' \in
-%   \cF_{j}$, $|R \cap R' \cap R''| = |\cl_j(R) \cap \cl_j(R') \cap
-%   \cl_j(R'')|$.
-%   The following are the possible cases that could arise. From argument
-%   above, $|\cl_j(R) \cap \cl_j(R') \cap \cl_j(R'')| =
-%   |\cl_{\phi[\cF_{j}]}(R) \cap \cl_{\phi[\cF_{j}]} (R') \cap
-%   \cl_{\phi[\cF_{j}]} (R'')|$
-
-%   \begin{enumerate}[a $|$]
-%   \item {\em None of the sets are new. $R, R', R'' \in \cF_{j-1}$.}
-%     We know $(\cF_{j-1}, \cl_{j-1})$ is feasible. Thus $|R \cap R'
-%     \cap R''| = |\cl_{j-1}(R) \cap \cl_{j-1}(R') \cap \cl_{j-1}(R'')|
-%     = |\cl_{j}(R) \cap \cl_{j}(R') \cap \cl_{j}(R'')|$.
-%   \item {\em Only one, say $R$, is a new set.}  Let $R = S_1 \cap S_2$
-%     ($S_1, S_2$ are defined in the proof of lemma
-%    ~\ref{lem:invar1}). Now we have $|R \cap R' \cap R''| = |S_1 \cap
-%     S_2 \cap R' \cap R''| = |\cl_{j-1}(S_1) \cap \cl_{j-1}(S_2) \cap
-%     \cl_{j-1}(R') \cap \cl_{j-1}(R'')| = |\cl_{j}(R) \cap \cl_{j}(R')
-%     \cap \cl_{j}(R'')|$. Thus proven. If $R$ is any of the other new
-%     sets, the same claim can be verified using lemma
-%    ~\ref{lem:setminuscard}.
-%     % \item []{\bf Case 3:}
-%   \item {\em At least two of $R, R', R''$ are new sets.}  The new sets
-%     are disjoint hence this case is vacuously true.
-%   \end{enumerate}
-%   \qed
-%   \tnote[E2]{remove the induction proof. just text saying x and v are
-%     exclusive in these sets therefore the intersection cardinalities
-%     don't change thus all invariants are still true} 
-\end{proof}
-
-As a result of Algorithm~\ref{perms} each leaf $v$ in $T$
-is such that there is exactly one set in $\cF$ with $v$ as a vertex in
-the path assigned to it.  In Algorithm~\ref{leafasgn} we identify
-elements in $supp(\cF)$ whose images are leaves in a hypergraph
-isomorphism if one exists.  Let $S \in \cF$ be such that $\cl(S)$ is a
-path with leaf and $v \in V(T)$ is the unique leaf incident on it.  We
-define a new path labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) =
-\set{v}$ where $x$ an arbitrary element from $S \setminus \bigcup_{\hS
-  \ne S} \hS$. In other words, $x$ is an element present in no other
-set in $\cF$ except $S$. This is intuitive since $v$ is present in no
-other path image under $\cl$ other than $\cl(S)$.  The element $x$ and
-leaf $v$ are then removed from the set $S$ and path $\cl(S)$
-respectively. After doing this for all leaves in $T$, all path images
-in the new path labeling $\cl_{new}$ except leaf labels (a path that
-has only a leaf is called the {\em leaf label} for the corresponding
-single element hyperedge or set) are paths from a new pruned tree $T_0
-= T \setminus \{v \mid v \text{ is a leaf in }
-T\}$. Algorithm~\ref{leafasgn} is now presented with details.
-
-\begin{algorithm}[h]
-  \caption{Leaf labeling from an ICPPL {\tt filter\_2($\cF, \cl, T$)}}
-  \label{leafasgn}
-  \begin{algorithmic}[\lndisplay]
-    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$
-    \COMMENT {Path images are such that no two path images share a
-      leaf.}
-    \STATE $j \assign 1$\\
-    \WHILE {there is a leaf $v$ in $T$ and a unique $S_1 \in
-      \cF_{j-1}$ such that $v \in \cl_{j-1}(S_1)$ }\label{uniqueleaf}
-    \STATE $\cF_j \assign \cF_{j-1} \setminus \{S_1\}$\\
-    \STATE for all $S \in \cF_{j-1}$ such that $S \ne S_1$ set
-    $\cl_j(S) \assign
-    \cl_{j-1}(S)$\\
-    \STATE $X \assign S_1 \setminus \bigcup_{S \in \cF_{j-1}, S \ne S_1}S$\\
-    \IF{$X$ is empty} \label{xempty} \STATE {\bf exit} \label{ln:exit2} \ENDIF
-    \STATE $x \assign $ arbitrary element from $X$\\
-    \STATE $\cF_j \assign \cF_j \cup \{\{x\}, S_1 \setminus \{x\}\} $\\
-    \STATE $\cl_j(\{x\}) \assign \{v\}$\\
-    \STATE $\cl_j(S_1 \setminus \{x\}) \assign \cl_{j-1}(S_1) \setminus \{v\}$\\
-    \STATE $j \assign j+1$\\
-    \ENDWHILE
-    \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
-    \RETURN $(\cF', \cl')$
-  \end{algorithmic}
-\end{algorithm}
-
-Suppose the input ICPPL $(\cF, \cl)$ is feasible, yet set $X$ in
-Algorithm~\ref{leafasgn} is empty in some iteration of the {\bf while}
-loop. This will abort our procedure of finding the hypergraph
-isomorphism. The following lemma shows that this will not happen.
-
-\begin{lemma}
-  \label{lem:xnotempty}
-  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{leafasgn} is
-  feasible, then for all iterations $j > 0$ of the {\em \bf while}
-  loop, the {\em \bf exit} statement in line~\ref{ln:exit2} does not
-  execute.
-\end{lemma}
-\begin{proof}
-  Assume $X$ is empty for some iteration $j > 0$. We know that $v$ is
-  an element of $\cl_{j-1}(S_1)$. Since it is uniquely present in
-  $\cl_{j-1}(S_1)$, it is clear that $v \in \cl_{j-1}(S_1) \setminus
-  \bigcup_{(S \in \cF_{j-1}) \wedge (S \ne S_1)}\cl_{j-1}(S)$.  Note
-  that for any $x \in S_1$ it is contained in at least two sets due to
-  our assumption about cardinality of $X$. Let $S_2 \in \cF_{j-1}$ be
-  another set that contains $x$. From the above argument, we know $v
-  \notin \cl_{j-1}(S_2)$. Therefore there cannot exist a hypergraph
-  isomorphism bijection that maps elements in $S_2$ to those in
-  $\cl_{j-1}(S_2)$. This contradicts our assumption that the input is
-  feasible. Thus $X$ cannot be empty if input is ICPPL and feasible.
-  \qed
-\end{proof}
-
-\begin{lemma}
-  \label{lem:invar3}
-  In Algorithm~\ref{leafasgn}, for all $j > 0$, at the end of the
-  $j$th iteration of the {\bf while} loop the four invariants given in
-  Lemma~\ref{lem:invar1} hold.
-\end{lemma}
-\begin{proof}
-  By Lemma~\ref{lem:xnotempty} we know that set $X$ will not be empty
-  in any iteration of the {\em \bf while} loop if input ICPPL $(\cF,
-  \cl)$ is feasible and $\cl_j$ is always computed for all $j >
-  0$. Also note that removing a leaf from any path keeps the new path
-  connected. Thus invariant I is obviously true. In every iteration $j
-  > 0$, we remove exactly one element $x$ from one set $S$ in $\cF$
-  and exactly one vertex $v$ which is a leaf from one path
-  $\cl_{j-1}(S)$ in $T$. This is because as seen in
-  Lemma~\ref{lem:xnotempty}, $x$ is exclusive to $S$ and $v$ is
-  exclusive to $\cl_{j-1}(S)$. Due to this fact, it is clear that the
-  intersection cardinality equations do not change, i.e., invariants
-  II, III, IV remain true. On the other hand, if the input ICPPL is
-  not feasible the invariants are vacuously true. \qed
-\end{proof}
-
-% \textcolor{cyan}{
-% \begin{lemma}
-%   \label{lem:notfeasibleexit}
-%   \tnote{IS THIS CORRECT?}  If input ICPPL $(\cF, \cl)$ is not
-%   feasible, then in one of the recursive calls to Algorithm --3--, the
-%   {\em \bf exit} statement in line x in Algorithm --1 or line y in
-%   Algorithm --2 will get executed.
-% \end{lemma}
-% }
-
-We have seen two filtering algorithms above, namely,
-Algorithm~\ref{perms} {\tt filter\_1} and Algorithm~\ref{leafasgn}
-{\tt filter\_2} which when executed serially repectively result in a
-new ICPPL on the same universe $U$ and tree $T$. We also proved that
-if the input is indeed feasible, these algorithms do indeed output the
-filtered ICPPL. Now we present the algorithmic characterization of a
-feasible tree path labeling by way of Algorithm~\ref{al:icppl-find-isomorph}.
-
-Algorithm~\ref{al:icppl-find-isomorph} computes a
-hypergraph isomorphism $\phi$ recursively using Algorithm~\ref{perms}
-and Algorithm~\ref{leafasgn} and pruning the leaves of the input
-tree. In brief, it is done as follows. Algorithm~\ref{leafasgn} gives
-us the leaf labels in $\cF_2$, i.e., the elements in $supp(\cF)$ that
-map to leaves in $T$, where $(\cF_2, \cl_2)$ is the output of
-Algorithm~\ref{leafasgn}. All leaves in $T$ are then pruned away. The
-leaf labels are removed from the path labeling $\cl_2$ and the
-corresponding elements are removed from the corresponding sets in
-$\cF_2$. This tree pruning algorithm is recursively called on the
-altered hypergraph $\cF'$, path label $\cl'$ and tree $T'$. The
-recursive call returns the bijection $\phi''$ for the rest of the
-elements in $supp(\cF)$ which along with the leaf labels $\phi'$ gives
-us the hypergraph isomorphism $\phi$.  The following lemma formalizes
-the characeterization of feasible path labeling.
-
-\begin{algorithm}[h]
-  \caption{{\tt get- hypergraph- isomorphism ($\cF, \cl, T$)}}
-  \label{al:icppl-find-isomorph}
-  \begin{algorithmic}[\lndisplay]
-
-    \IF{$T$ is empty}
-    \RETURN $\emptyset$\\
-    \ENDIF
-    \STATE $L \assign \{v \mid v \text{ is a leaf in }      T\}$\\
-    \STATE $(\cF_1, \cl_1) \assign$ {\tt filter\_1($\cF, \cl,
-      T$)}\\
-    \STATE $(\cF_2, \cl_2) \assign$ {\tt filter\_2($\cF_1,
-      \cl_1, T$)}\\
-
-    \STATE $(\cF', \cl') \assign (\cF_2, \cl_2)$\\
-    \STATE $\phi' \leftarrow \emptyset$
-
-    \FOR {every $v \in L$}
-    \STATE $\phi'(x) \assign v$ where $x \in \cl_2^{-1}(\{v\})$
-    \COMMENT {Copy the leaf labels to a one to one function $\phi':
-      supp(\cF) \rightarrow L$
-      }\\
-    \STATE Remove $\{x\}$ and $\{v\}$ from $\cF'$, $\cl'$  appropriately\\
-    \ENDFOR
-
-    \STATE $T' \assign T \setminus L$
-
-    \STATE $\phi'' \assign$ {\tt
-      get-hypergraph-isomorphism($\cF', \cl', T'$)}
-    \STATE $\phi \assign \phi'' \cup \phi'$ \\
-    \RETURN $\phi$
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma}
-  \label{lem:hyperiso}  %{lem:perm}
-  If $(\cF, \cl)$ is an ICPPL from a tree $T$ and
-  Algorithm~\ref{al:icppl-find-isomorph}, {\tt
-    get- hypergraph- isomorphism ($\cF, \cl, T$)} returns a non-empty
-  function, then there exists a hypergraph isomorphism $\phi :
-  supp(\cF) \rightarrow V(T)$ such that the $\phi$-induced tree path
-  labeling is equal to $\cl$ or $\cl_\phi = \cl$.
-\end{lemma}
-\begin{proof}
-  It is clear that in the end of every recursive call to
-  Algorithm~\ref{al:icppl-find-isomorph}, the function $\phi'$ is one
-  to one involving all the leaves in the tree passed as input to that
-  recursive call. Moreover, by Lemma~\ref{lem:noexit1} and
-  Lemma~\ref{lem:xnotempty} it is consistent with the tree path
-  labeling $\cl$ passed. The tree pruning is done by only removing
-  leaves in each call to the function and is done till the tree
-  becomes empty. Thus the returned function $\phi: supp(\cF)
-  \rightarrow V(T)$ is a union of mutually exclusive one to one
-  functions exhausting all vertices of the tree. In other words, it is
-  a bijection from $supp(\cF)$ to $V(T)$ inducing the given path
-  labeling $\cl$ and thus a hypergraph isomorphism. \qed
-\end{proof}
-
-\begin{theorem}
-  \label{th:charac}
-  A path labeling $(\cF, \cl)$ on tree $T$ is feasible iff it is an
-  ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
-  T)$ as input returns a non-empty function.
-\end{theorem}
-\begin{proof}
-  From Lemma~\ref{lem:hyperiso}, we know that if $(\cF, \cl)$ is an
-  ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
-  T)$ as input returns a non-empty function, $(\cF, \cl)$ is feasible.
-  Now consider the case where $(\cF, \cl)$ is feasible. i.e. there
-  exists a hypergraph isomorphism $\phi$ such that $\cl_\phi =
-  \cl$. Lemma~\ref{lem:noexit1} and Lemma~\ref{lem:xnotempty} show us
-  that filter 1 and filter 2 do not exit if input is feasible. Thus
-  Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
-  function.\qed
-\end{proof}
-
-\section{ Computing feasible TPL with special target trees \tnote{give
-    problem definition etc}}
-\label{sec:spltargettree}
-
-Section~\ref{sec:feasible} described properties that a TPL must have
-for it to be feasible. The next problem of interest is to test if a
-given hypergraph is a path hypergraph with respect to a given target
-tree\footnote{A larger problem would be to find if a given hypergraph
-  is a path graph on any tree. This problem is not addressed in this
-  thesis.}. In other words, the problem is to find out if a feasible
-tree path labeling exists from a given target tree for a given
-hypergraph. In this section we will see two special cases of this
-problem where the target tree is from a particular family of
-trees. The first one, where the tree is a path, is a shown to be
-equivalent to the well studied problem of consecutive-ones in
-Section~\ref{sec:icpplicpia}. The second one, where the tree is a
-$k$-subdivided tree\footnote{See Section~\ref{ch:prelims} for the
-  formal definition.}, has been solved using a polynomial time
-algorithm. The latter problem enforces some conditions on the
-hypergraph too which will be seen in Section~\ref{sec:ksubdivstar}.
-
-\subsection{Target tree is a Path}
-\label{sec:icpplicpia}
-
-Consider a special case of ICPPL with the following properties.
-\begin{enumerate}
-\item Given tree $T$ is a path. Hence, all path labels are interval
-  labels.
-\item Only pairwise intersection cardinality preservation is
-  sufficient. i.e. property (iii) in ICPPL is not enforced.
-\item The filter algorithms do not have {\em \bf exit} statements.
-\end{enumerate}
-This is called an Intersection Cardinality Preservation Interval
-Assignment (ICPIA) \cite{nsnrs09}. This structure and its algorithm is
-used in the next section for finding tree path labeling from a
-$k$-subdivided star due to this graph's close relationship with
-intervals.
-
-\subsection{Target tree is a $k$-subdivided Star}
-\label{sec:ksubdivstar}
-
-\begin{figure}[t] %[htbp] 
-  \centering
-  \begin{tabular}{lr}
-    (a) \includegraphics{cop-tree_thesis.1}&
-    (b) \includegraphics{cop-tree_thesis.2}
-  \end{tabular}
-  \caption{\figtabsize (a) $8$-subdivided star with 7 rays (b)
-    3-subdivided star with 3 rays}
-  \label{fig:kstar}
-\end{figure}
-
-In this section we consider the problem of assigning paths from a
-$k$-subdivided star $T$ to a given set system $\cF$.  We consider
-$\cF$ for which the overlap graph $\bO(\cF)$ is connected.  The
-overlap graph is well-known from the work of
-\cite{kklv10,nsnrs09,wlh02}.  We use the notation in
-\cite{kklv10}. Recall from Section~\ref{ch:prelims} that hyperedges
-$S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$
-and $S'$ have a non-empty intersection but neither of them is
-contained in the other. The overlap graph $\bO(\cF)$ is a graph in
-which the vertices correspond to the sets in $\cF$, and the vertices
-corresponding to the hyperedges $S$ and $S'$ are adjacent if and only
-if they overlap.  Note that the intersection graph of $\cF$,
-$\bI(\cF)$ is different from $\bO(\cF)$ and $\bO(\cF) \subseteq
-\bI(\cF)$.  A connected component of $\bO(\cF)$ is called an overlap
-component of $\cF$.  An interesting property of the overlap components
-is that any two distinct overlap components, say $\cO_1$ and $\cO_2$,
-are such that any two sets $S_1 \in \cO_1$ and $S_2 \in \cO_2$ are
-disjoint, or, w.l.o.g, all the sets in $\cO_1$ are contained within
-one set in $\cO_2$.  This containment relation naturally determines a
-decomposition of the overlap components into rooted containment trees.
-We consider the case when there is only one rooted containment tree,
-and we first present our algorithm when $\bO(\cF)$ is connected.  It
-is easy to see that once the path labeling to the overlap component in
-the root of the containment tree is achieved, the path labeling to the
-other overlap components in the rooted containment tree is essentially
-finding a path labeling when the target tree is a path: each target
-path is a path that is allocated to sets in the root overlap
-component.  Therefore, for the rest of this section, $\bO(\cF)$ is a
-connected graph. We also assume that all hyperedges are of cardinality
-at most $k+2$.
-
-Recall from Section~\ref{ch:prelims} that a $k$-subdivided star is a
-star with each edge subdivided $k$ times. Therefore, a $k$-subdivided
-star has a central vertex which we call the {\em root}, and each root
-to leaf path is called a {\em ray}. First, we observe that by removing
-the root $r$ from $T$, we get a collection of $p$ vertex disjoint
-paths of length $k+1$, $p$ being the number of leaves in $T$.  We
-denote the rays by $R_1, \ldots, R_p$, and the number of vertices in
-$R_i$, $i \in [p]$ is $k+2$.  Let $\seq{v_{i1},\ldots,v_{i(k+2)}=r}$
-denote the sequence of vertices in $R_i$, where $v_{i1}$ is the
-leaf. Note that $r$ is a common vertex to all $R_i$.
-
-In this section the given hypergraph, the $k$-subdivided star and the
-root of the star are denoted by $\cO$, $T$ and vertex $r$,
-respectively.
-% For each hyperedge $X \in \cO$, we will maintain a 2-tuple of non-negative
-% numbers $\seq{p_1(X), p_2(X)}$.  The numbers satisfy the property that
-% $p_1(X) + p_2(X) \leq |X|$, and at the end of path labeling, for each
-% $X$, $p_1(X) + p_2(X) = |X|$.  This signifies the algorithm tracking
-% the lengths of subpaths of the path assigned to $X$ from at most two
-% rays. We also maintain another parameter called the {\em residue} of
-% $X$ denoted by $s(X)=|X| - p_1(X)$. This signifies the residue path
-% length that must be assigned to $X$ which must be from another
-% ray. For instance, if $X$ is labeled a path from only one ray, then
-% $p_1(X) = |X|$, $p_2(X) = 0$ and $s(X) = 0$.
-
-% 
-% We iteratively consider each ray from which paths will be assigned to
-% hyperedges. At the beginning of each iteration hyperedges of $\cO$ are
-% classifed into the following disjoint sets.
-% \begin{enumerate}
-% \item [$\cL_1^i$] {\em Labeled without $r$.} Those that have been
-%   labeled with a path which does not contain $r$ in one of the
-%   previous iterations.\\  $\cL_1^i = \set{ X \mid p_1(X) = |X| \text{ and
-%     } p_2(X) = 0 \text{ and } s(X) = 0, X \in \cO}$
-% \item [$\cL_2^i$] {\em Labeled with $r$.} Those that have been labeled
-%   with two subpaths of $\cl(X)$ containing $r$ from two different rays
-%   in two previous iterations.\\ $\cL_2^i = \set{X \mid 0 < p_1\left(X\right),
-%     p_2\left(X\right) < |X| \text{ and } s(X) = 0, X \in \cO}$
-%   \item [$\cT_1^i$] {\em Type 1 / partially labeled.} Those that have
-%     been labeled with one path containing $r$ from a single ray in one
-%     of the previous iterations. Here, $p_1(X)$ denotes the length of
-%     the subpath of $\cl(X)$ that $X$ has been so far labeled
-%     with. Also, it is clear that such a path must start at $r$ and
-%     must be in a ray different from the one corresponding to $p_1(X)$.\\
-%     $\cT_1^i = \set{ X \mid 0 < p_1(X) < |X| \text{ and } p_2(X) = 0
-%       \text{ and } s(X) > 0, X \in \cO}$
-%   \item [$\cT_2^i$] {\em Type 2 / not labeled.} Those that have not been
-%     labeled with a path in any previous iteration.\\
-%     $\cT_2^i = \set{ X \mid p_1(X) = p_2(X) = 0 \text{ and } s(X) = |X|,
-%       X \in \cO}$
-% \end{enumerate}
-% \vspace{-2mm}
-% \begin{align*}
-%   \cO &= \cL_1^i \cup \cL_2^i \cup \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]\\
-%   \cO_i &= \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]
-% \end{align*}
-
-The set $\cO_i$ refers to the set of hyperedges $\cT_1^i \cup \cT_2^i$
-in the $i$th iteration.  Note that $\cO_1 = \cO$.  In the $i$th
-iteration, hyperedges from $\cO_i$ are assigned paths from $R_i$ using
-the following rules. Also the end of the iteration, $\cL_1^{i+1},
-\cL_2^{i+1}, \cT_1^{i+1}, \cT_2^{i+1}$ are set to $\cL_1^{i},
-\cL_2^{i}, \cT_1^{i}, \cT_2^{i}$ respectively, along with some
-case-specific changes mentioned in the rules below.
-
-\noindent
-Suppose the given hypergraph has a feasible path labeling to the given
-tree $T$. Let this labeling be $\cl$. Now we present an algorithm that
-lists the sequence of edges that maintain a set of properties
-necessary for them to have a feasible tree path labeling.
-\begin{enumerate}[I.]
-\item {\bf Step 1}
-  \begin{enumerate}
-  \item {\bf {There are type 1
-        edges, $\mathbf{|\cT_1^i| > 0}$}:} 
-% , and for each such $X$ assign the
-%     path in $R_i$ of length $s(X)$ starting at $v_{i,k+2}=r$.  $X$ is
-%     then removed from $\cO_i$, and $p_2(X)=s(X)$, and $s(X)=0$.If any
-%     intersection cardinality is violated because of this assignment,
-%     EXIT reporting the non-existence of a feasible assignment.
-  {If there is only one $\cT_1^i$ hyperedge, label it to the unique
-    path in $R_i$ of length $s(X)$ starting at $v_{i(k+2)}=r$. If
-    there are more than one $\cT_1^i$ hyperedges, for any pair of
-    hyperedges $S_1, S_2 \in \cT_1^i$, one of the following is
-    true. Let $\cl'(X) \subseteq R_j$ denote the subpath of length
-    $p_1(X)$ that has been assigned to $X$ from a previously
-    considered ray $R_j$.}
-  \begin{enumerate}[{\em Case }1. ]
-  \item $|\cl'(S_1) \cap \cl'(S_2)| = 1$ and $|S_1 \cap S_2| > 1$
-    i.e. they are in different rays and their residues must contain
-    their intersection. Hence assign both to $R_i$.
-  \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| >
-    |\cl'(S_1) \cap \cl'(S_2)|$ i.e. they are in the same rays and
-    their residues must contain their intersection. Hence assign both
-    to $R_i$.
-  \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| =
-    |\cl'(S_1) \cap \cl'(S_2)|$. i.e. the intersection cardinality has
-    been met. Hence assign only $S_1$ to $R_i$ and $S_2$ must be
-    assigned to some $i \ge p$.
-  \end{enumerate}
-  Remove the labeled edge or edges from $\cO_i$ and add it to
-  $\cO_{i+1}$. Also do not add it to $\cT_1^{i+1}$ and add to
-  $\cL_2^{i+1}$.
-\item {\bf {There are no type-1 edges nor} super-marginal type-2
-    hyperedges:} If all type-2 edges of length at most $k+1$, EXIT by
-  saying there is no ICPPL.  Otherwise, Find an $X$ of type-2 such
-  that $|X| \geq k+2$, assign it the unique path starting at
-  $v_{i(k+1)}=r$ of length $k+2$, and set $p_1(X)=k+2,
-  s(X)=|X|-(k+2)$, assign it the unique path starting at
-  $v_{i(k+1)}=r$ of length $k+2$, mark this as a type-1 hyperedge, and
-  add it to $\cO_{i+1}$ after removing it from $\cO_i$.
-  % \remove{Next, iteratively, consider each type-1 edge $X \in
-  %   \cO_i$, and if it intersects a previously assigned $Y$ such that
-  %   $\cl(Y) \subseteq R_i$, assign a path of length $s(X)$
-  %   containing $v_{i(k+2)}=r$, set $p_2(X)=s(X), s(X)=0$, and remove
-  %   $X$ from $\cO_i$.}\tnote{covered in step 2}
-
-\item {\bf There are super-marginal type-2 hyperedges {and no type-1
-      edges:}} Let $X$ be a super-marginal hyperedge. If $|X| \leq
-  k+2$, then $X$ is assigned the path of length $|X|$ starting at
-  $v_{i1}$, and $p_1(X) = |X|, p_2(X)=0$. $X$ is removed from $\cO_i$
-  and not added to $\cO_{i+1}$.  If $|X| > k+2$, then $p_1(X)=k+2$,
-  assign it the unique path starting at $v_{i(k+1)}=r$ of length $k+2$
-  and $s(X) = |X| - (k+2)$.  In this case, $X$ is classified as a
-  Type-1 edge, removed from $\cO_i$, and added to $\cO_{i+1}$.
-\end{enumerate}
-\item {\bf Step 2:} Iteratively, a hyperedge $X$ is selected from
-  $\cO_i$ that has an overlap with one of the hyperedges $Y$ such that
-  $\cl(Y) \subseteq R_i$, and a unique path is assigned to $X$.  The
-  path, say $U(X) \subseteq R_i$, that is assigned can be decided
-  unambiguously since the $X \overlap Y$, and all intersection
-  cardinalities can be preserved only at one of the ends of $\cl(Y)$.
-
-  \noindent
-  Let $\cl(X)$ denote the unique path assigned to $X$.  If $X$ is a
-  type-2 hyperedge and if the unique path of length $|X|$ does not
-  contain $r$, then $p_1(X) = |X|, p_2(X)=0$, and $X$ is removed from
-  $\cO_i$.  In the case when $\cl(X)$ has to contain $r$, then
-  $p_1(X)$ is the length of the path, $p_2(X)=0$, and $s(X) =
-  |X|-p_1(X)$. Further $X$ is classified as a type-1 hyperedge, added
-  to $\cO_{i+1}$ and removed from $\cO_i$.  In the case when $X$ is a
-  type-1 hyperedge, then we check if $U(X)$, which is of length $s(X)$
-  contains $r$. If it does, then we assign $\cl(X) \leftarrow \cl(X)
-  \cup U(X)$, remove $X$ from $\cO_i$ and do not add it to
-  $\cO_{i+1}$.  If not, then we {\bf Exit} reporting that an
-  assignment cannot be found.  The iteration ends when no hyperedge in
-  $\cO_i$ has an overlap with a hyperedge assigned to $R_i$.
-\end{enumerate}
-
-\noindent
-The order in which they are assigned is the sequence of hyperedges
-with the following properties. Note that if such a sequence cannot be
-computed, $\cO$ cannot have a feasible path labeling from $T$.
-
-\noindent
-\tnote{need to formalize the below properties} In the following
-lemmas we identify a set of necessary conditions for $\cF$ to have an
-ICPPL in the $k$-subdivided star $T$.  If during the execution of the
-above algorithm, one of these necessary conditions is not satisfied,
-the algorithm exits reporting the non-existence of an ICPPL.
-\begin{lemma}
-  Let all hyperedges in $\cO_i$ be type-1 edges.  Then there is a
-  maximal subset $\ccT_i \subseteq \cO_i$ with the following
-  properties:
-  \begin{enumerate}
-  \item $\ccT_i$ form an inclusion chain.
-  \item For all $X \in \ccT_i$, $s(X) \leq k+2$, and There is a an $X
-    \in \ccT_i$ such that $s(X)=k+2$.
-  \end{enumerate}
-\end{lemma}
-\begin{lemma}
-  If there are no \marginal type-2 edges in $\cO_i$, then there exists
-  at least $p-i$ type-2 hyperedges $X \in \cO_i$ such that $|X| \geq
-  k+2$.
-  % is a exist type-2 hyperedges of length at least $k+2$, and there
-  % must be at least $r-i$ such type-2 hyperedge, basically at least
-  % one per ray.  Each hyperedge contained in such an hyperedges of
-  % length at least $k+2$, is transitively related under the overlap
-  % relation to one that gets assigned a path that contains
-  % $r$. Basically, the proof will argue that otherwise, there will be
-  % a \marginal type-2 hyperedge, or that it is infeasible.  Also, it
-  % is here we assume that we assume that the size of the two
-  % universes is the same.
-\end{lemma}
-\begin{lemma}
-  At the end of Step-1 in the $i$-th iteration, if one hyperedge $X$
-  of type-2 is such that $\cl(X) \subseteq R_i$, then all other
-  hyperedges in $\cO_i$ are connected to $X$ in the overlap component.
-\end{lemma}
-\begin{lemma}
-  At the end of Step-2, If control has exit at any time, there is no
-  ICPPL. If control has not exit, then $R_i$ is saturated.  No
-  hyperedge of $\cO_{i+1}$ will get a path from $R_i$ in the future
-  iterations.  No type-2 hyperedge of $\cO_{i+1}$ will get a path from
-  $R_i$.  Basically $R_i$ is done.
-\end{lemma}
-\begin{lemma}
-  Finally, we need to prove that the assignment is an ICPPL. Secondly,
-  if there is a permutation then maps sets to paths, then it is indeed
-  an ICPPL, and our algorithm will basically find it.  It is a unique
-  assignment upto permutation of the leaves.
-\end{lemma}
-
-\begin{lemma}
-  \label{lem:sup-mar}
-  If $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
-  tree path labeling to tree $T$, then $\cl(X)$ will
-  contain a leaf in $T$.\\
-  On the otherhand, if $\cl(X)$ has a leaf in
-  $T$, $X$ is marginal but may not be super-marginal.
-\end{lemma}
-\begin{proof}
-  Suppose $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
-  path labeling from $T$.  Assume $\cl(X)$ does not have a leaf.  Let
-  $R_i$ be one of the rays (or the only ray) $\cl(X)$ is part of.
-  Since $X$ is in a connected overlap component, there exists $Y_1 \in
-  \cF$ and $X \nsubseteq Y_1$ such that $Y_1 \overlap X$ and $Y_1$ has
-  at least one vertex closer to the leaf in $R_i$ than any vertex in
-  $X$. Similarly with the same argument there exists $Y_2 \in \cF$
-  with same subset and overlap relation with $X$ except it has has at
-  least one vertex farther away from the leaf in $R_i$ than any vertex
-  in $X$. Clearly $Y_1 \cap X$ and $Y_2 \cap X$ cannot be part of same
-  inclusion chain which contradicts that assumption $X$ is
-  super-marginal. Thus the claim is proven.\qed
-\end{proof}
-
-\noindent
-Consider the overlap graph $\bO(\cF)$ of the given hypergraph
-$\cF$. Let $S_{sm} \in \cF$ be such that it is a super-marginal
-hyperedge.  Algorithm~\ref{al:ktree-label} uses $S_{sm}$ along with
-the overlap graph $\bO(\cF)$ to calculate the feasible tree path
-labeling to the $k$-subdivded tree $T$.
-
-\begin{algorithm}[h]
-  \caption{{\tt compute-ksubtree-path-labeling($X, \cF, T$)}}
-  \label{al:ktree-label}
-  \begin{algorithmic}[\lndisplay]
-    \IF{$X = S_{sm}$}
-    \STATE -- TBD --\\
-    \ELSE 
-    \STATE -- TBD --\\
-    \ENDIF
-
- \end{algorithmic}
-\end{algorithm}
-
-\section{ TPL with no restrictions}
-\label{sec:norestraint}
-
-{\tt abstract begin}
-  it is
-  known that if the given tree is a path a feasible assignment can be
-  found in polynomial time, and we observe that it can actually be
-  done in logspace. \tnote{[TRUE?]}
-{\tt abstract end}
-
-In Section \ref{prelims} we present the
-necessary preliminaries, in Section \ref{feasible} we present our
-characterization of feasible tree path assignments, and in Section
-\ref{decompos} we present the characterizing subproblems for finding a
-bijection between $U$ and $V(T)$ such that sets map to tree
-paths. Finally, in Section \ref{complexity} we conclude by showing
-that Tree Path Assignment is GI-Complete, and also observe that
-Consecutive Ones Testing is in Logspace.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC \subsection{Introduction}
-
-\annote{ We refer to this as the Tree Path Assignment
-problem for an input $(\cF,T)$ pair.}{Add this ``pair'' as a
-  definition in ch:prelims}
-
-\remove{It is an interesting fact that for a matrix with the COP, the
-intersection graph of the corresponding set system is an interval
-graph.  A similar connection to a subclass of chordal graphs, and this
-subclass contains interval graphs, exists for the generalization of
-COP.  In this case, the intersection graph of the corresponding set
-system must be a path graph. Chordal graphs are of great
-significance, extensively studied, and have several applications.  One
-of the well known and interesting properties of a chordal graphs is
-its connection with intersection graphs cite{mcg04}. For every chordal
-graph, there exists a tree and a family of subtrees of this tree such
-that the intersection graph of this family is isomorphic to the
-chordal graph cite{plr70,gav78,bp93}.  Certain format of these trees
-are called clique trees cite{apy92} of the graph which is a compact
-representation of the chordal graph. There has also been work done on
-the generalization of clique trees to clique hypergraphs cite{km02}.
-If the chordal graph can be represented as the intersection graph of
-paths in a tree, then the graph is called path graph cite{mcg04} or
-path chordal graph.  Therefore, it is clear that if there is a
-bijection from $U$ to $V(T)$ such that the sets map to paths, then the
-intersection graph of the set system is indeed a path chordal graph.
-This is, however, only a necessary condition and can be checked
-efficiently, as path graph recognition is polynomial
-solvable cite{gav78,aas93}.  Indeed, it is possible to construct a set
-system and tree, such that the intersection graph is a path chordal
-graph, but there is no bijection between $U$ and $V(T)$ such that the
-sets map to paths.  This connection indeed suggests that our problem
-is indeed as hard as path graph isomorphism.  Further path graph
-isomorphism is known be isomorphism-complete, see for example
-cite{kklv10}. }
-In the second part of this paper, we decompose our
-search for a bijection between $U$ and $V(T)$ into subproblems.  Each
-subproblem is on a set system in which for each set, there is another
-set in the set system with which the intersection is {\em strict}-
-there is a non-empty intersection, but neither is contained in the
-other.  This is in the spirit of results in \cite{wlh02,nsnrs09} where
-to test for COP in a given matrix, the COP problem is solved on an
-equivalent set of prime matrices.  \annote{Our decomposition localizes the
-challenge of path graph isomorphism to two problems.}{HOW?}
-
-Finally, we show that Tree Path Assignment is isomorphism-complete.
-We also point out Consecutive Ones Testing is in Logspace from two
-different results in the literature \cite{kklv10, mcc04}. To the best
-of our knowledge this observation has not been made earlier.
-
-% \noindent
-% It has been long known that interval graph recognition is in
-% logspace\cite{rei84}. Recently interval graph isomorphism was also
-% shown to be logspace decidable using a logspace canonization
-% algorithm by \cite{kklv10}.  This result is built on top of logspace
-% results of undirected graph connectivity \cite{rei08}, logspace
-% tractability using a certain logical formalism called FP+C and
-% modular decomposition of interval graphs\cite{lau10} etc.  Interval
-% graphs are closely connected to binary matrices with COP. The
-% maximal clique vertex incidence matrix (matrix with rows
-% representing maximal cliques and columns representing vertices of a
-% graph) has COP on columns iff the graph is an interval
-% graph\cite{fg65}. This follows from the interval graph
-% characterization by \cite{gh64}. Due to this close relation it is
-% natural to see if consecutive ones property can be
-% tested in logspace. \\
-% \noindent
-% We also explore some extensions of the interval assignment problem
-% in \cite{nsnrs09}, namely caterpillar path assignment problem.
-
-% We present a logspace algorithm here that uses the ICPIA
-% characterization of binary matrices with COP (set system associated
-% with such a matrix)\cite{nsnrs09}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC\subsection{Preliminaries} 
-\label{prelims} 
-In this paper, the collection
-$\F = \{S_i \mid S_i \subseteq U, S_i \ne \O, i \in I\}$ is a {\em set
-  system} of universe, $U = \{1, \ldots, n\}$. Moreoever, a set system
-is assumed to ``cover'' the universe,
-i.e. $ \bigcup_{i \in I}S_i = U$. \\
-
-\noindent
-The graph $T=(V,E)$ represents a given tree with $|V| = n$. Further,
-for simplicity, $V$ is defined as $\{1,\ldots,n\}$. All paths
-referred to in this paper are paths of $T$ unless explicitly specified. \\
-
-\noindent
-A {\em path assignment} $\A$ to $\F$ is defined as a set assignment
-where second universe is the vertex set $V$ of a given tree $T$ and
-every second subset in the ordered pairs is a path in this
-tree. Formally, the definition is as follows.
-\begin{align*}
-  \A = \{ (S_i,P_i) \mid S_i \in \F, P_i \subseteq V \text{
-    s.t. }T[P_i] \text{ is a path, } i \in I \}
-\end{align*}
-In other words, $P_i$ is the path on the tree $T$ assigned to $S_i$ in
-$\A$. As mentioned before for set systems, the paths cover the whole
-tree, i.e. $\bigcup_{i \in I}P_i = V$ \\
-
-\noindent
-Generalizing the definition of {\em feasibility} in \cite{nsnrs09} to
-a set assignment, a path assignment $\A$ is defined to be {\em
-  feasible} if there exists a bijection defined as follows.
-\begin{align}
-  \sigma: U \rightarrow V(T), \text{ such that }\sigma(S_i) = P_i
-  \text{ for all } i \in I, \sigma \text{ is a bijection}
-  \label{eq:stf}
-\end{align}
-% The assignment $\A$ is then called {\em Set Translation Feasible}.
-
-\noindent
-Let $X$ be a partially ordered set with $\preccurlyeq$ being the
-partial order on $X$.  $mub(X)$ represents an element in $X$ which is
-a maximal upper bound on $X$.  $X_m \in X$ is a maximal upper bound of
-$X$ if $\nexists X_q \in X$ such that $X_m
-\preccurlyeq X_q$. \\
-
-\noindent
-The set $I$ represents the index set $[m]$. If index $i$ is used
-without further qualification, it is meant to be $i \in I$. Any
-function, if not defined on a domain of sets, when applied on a set is
-understood as the function applied to each of its elements. i.e. for
-any function $f$ defined with domain $U$, the abuse of notation is as
-follows; $f(S)$ is used instead of $\hat f(S)$ where
-$\hat f(S) = \{y \mid y = f(x), x \in S\}$. \\
-
-\noindent
-When refering to a tree as $T$ it could be a reference to the tree
-itself, or the vertices of the tree. This will be clear from the
-context.\\
-
-\noindent
-Finally, an in-tree is a directed rooted tree in which all edges are
-directed toward to the root.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC\subsection{Characerization of Feasible Tree Path
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC  Assignment\tnote{Merge with the original charac section}} 
-\label{feasible} 
-Consider a path assignment $\cA =
-\{(S_i, P_i) \mid S_i \in \cF, P_i \text{ is a path from $T$}, i \in
-[m]\}$ to a set system $\cF = \{S_i \mid S_i \subseteq U, i \in
-[m]\}$, were $T$ is a given tree, $U$ is the set system's universe and
-$m$ is the number of sets in $\cF$. We call $\cA$ an {\em Intersection
-  Cardinality Preserving Path Assignment (ICPPA)} if it has the
-following properties.
-
-\begin{enumerate}
-\item [i.]  $|S_i| = |P_i|$ for all $i \in [m]$
-\item [ii.] $|S_i \cap S_j| = |P_i \cap P_j|$ for all $i,j \in [m]$
-  % \item [iii.] $|\bigcup_{i \in J} S_i| = |\bigcup_{i \in J} P_i|$
-  %   for all $J \subseteq I$\footnote{TBD: this is needed now to fix
-  %     some proofs}
-\item [iii.] $|S_i \cap S_j \cap S_k| = |P_i \cap P_j \cap P_k|$ for
-  all $i,j,k \in [m]$
-\end{enumerate}
-
-\begin{lemma}
-  \label{lem:setminuscard}
-  If $\cA$ is an ICPPA, and $(S_1, P_1),(S_2, P_2),(S_3, P_3) \in
-  \cA$, then $|S_1 \cap (S_2 \setminus S_3)| = |P_1 \cap (P_2
-  \setminus P_3)|$.
-\end{lemma}
-\begin{proof}
-  $|S_1 \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| =
-  |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to conditions (ii)
-  and (iii) of ICPPA, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
-  \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
-  = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. \qed
-\end{proof}
-
-\begin{lemma}
-  \label{lem:fourpaths} Consider four paths in a tree $Q_1, Q_2, Q_3,
-  Q_4$ such that they have nonempty pairwise intersection and $Q_1,
-  Q_2$ share a leaf. Then there exists distinct $i, j, k \in
-  \{1,2,3,4\}$ such that, $Q_1 \cap Q_2 \cap Q_3 \cap Q_4 = Q_i \cap
-  Q_j \cap Q_k$.
-\end{lemma}
-\begin{proof}
-  {\em Case 1:} w.l.o.g, consider $Q_3 \cap Q_4$ and let us call it
-  $Q$. This is clearly a path (intersection of two paths is a path).
-  % Since $Q_1, Q_2$ share a leaf, the following are paths $Q_1
-  % \setminus Q_2$, $Q_2 \setminus Q_1$, $Q_1 \cap Q_2$ and they are
-  % mutually disjoint.
-  Suppose $Q$ does not intersect with $Q_1 \setminus Q_2$, i.e. $Q
-  \cap (Q_1 \setminus Q_2) = \O$. Then $Q \cap Q_1 \cap Q_2 = Q \cap
-  Q_2$. Similarly, if $Q \cap (Q_2 \setminus Q_1) = \O$, $Q \cap Q_1
-  \cap Q_2 = Q \cap Q_1$. Thus it is clear that if the intersection of
-  any two paths does not intersect with any of the set differences of
-  the remaining two paths, the claim in the lemma is true.
-  % Note that $Q_1 \setminus Q_2$ and $Q_2
-  % \setminus Q_1$ are paths because $Q_1, Q_2$ share a leaf.\\
-  {\em Case 2:} Let us consider the compliment of the previous
-  case. i.e. the intersection of any two paths intersects with both
-  the set differences of the other two. First let us consider $Q \cap
-  (Q_1 \setminus Q_2) \ne \O$ and $Q \cap (Q_1 \setminus Q_2) \ne \O$,
-  where $Q = Q_3 \cap Q_4$. Since $Q_1$ and $Q_2$ share a leaf, there
-  is exactly one vertex at which they branch off from the path $Q_1
-  \cap Q_2$ into two paths $Q_1 \setminus Q_2$ and $Q_2 \setminus
-  Q_1$. Let this vertex be $v$. It is clear that if path $Q_3 \cap
-  Q_4$, must intersect with paths $Q_1 \setminus Q_2$ and $Q_2
-  \setminus Q_1$, it must contain $v$ since these are paths from a
-  tree. Moreover, $Q_3 \cap Q_4$ intersects with $Q_1 \cap Q_2$ at
-  exactly $v$ and only at $v$ which means that $Q_1 \cap Q_2$ does not
-  intersect with $Q_3 \setminus Q_4$ or $Q_4 \setminus Q_3$ which
-  contradicts initial condition of this case. Thus this
-  case cannot occur and case 1 is the only possible scenario. \\
-  Thus lemma is proven \qed
-\end{proof}
-
-\noindent
-In the remaining part of this section we show that a path assignment
-is feasible if and only if it is an ICPPA.  One direction of this
-claim is clear: that is a path assignment is feasible, then all
-intersection cardinalities are preserved, that is the path assignment
-is an ICPPA.  The reason is that a feasible path assignment has an
-associated bijection between $U$ and $V(T)$ such that the sets map to
-paths.  The rest of the section is devoted to constructively proving
-that it is sufficient, for the existence of a bijection, for a path
-assignment to be an ICPPA.  At a top-level, the constructive
-approaches refine the path assignment iteratively, such that at the
-end of each iteration we have a path assignment, and finally we have a
-family of bijections.  First we present and then prove the correctness
-of Algorithm \ref{perms}.  This algorithm refines the path assignment
-by considering pairs of paths that share a leaf.
-
-\begin{algorithm}[h]
-  \caption{Permutations from an ICPPA $\{(S_i,P_i) | i \in I\}$}
-  \label{perms}
-  \begin{algorithmic}
-    \STATE Let $\Pi_0=\{(S_i,P_i)| i \in I\}$\\
-    \STATE $j = 1$;
-    \label{shareleaf} \WHILE {There is $(P_1,Q_1), (P_2,Q_2) \in
-      \Pi_{j-1}$ with $Q_1$ and $Q_2$ having a common leaf} \STATE
-    $\Pi_j= \Pi_{j-1} \setminus \{(P_1,Q_1),(P_2,Q_2)\}$;
-    \label{setbreak}\STATE $\Pi_j = \Pi_j \cup \{(P_1 \cap P_2,Q_1
-    \cap Q_2), (P_1 \setminus P_2,Q_1 \setminus Q_2), (P_2 \setminus
-    P_1, Q_2 \setminus Q_1)\}$; \STATE $j = j+1$; \ENDWHILE \STATE
-    $\Pi = \Pi_j$; \STATE Return $\Pi$;
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma}
-  \label{lem:invar1}
-  In Algorithm \ref{perms}, at the end of $j$th iteration, $j \ge 0$,
-  of the while loop of Algorithm \ref{perms}, the following invariants
-  are maintained.
-  \begin{itemize}
-  \item {\em Invariant I:} $Q$ is a path in $T$ for each $(P,Q) \in
-    \Pi_j$
-  \item {\em Invariant II:} $|P|=|Q|$ for each $(P,Q) \in \Pi_j$
-  \item {\em Invariant III:} For any two $(P,Q), (P',Q') \in \Pi_j$,
-    $|P' \cap P''|=|Q' \cap Q''|$.
-  \item {\em Invariant IV:} For any three, $(P',Q'), (P'',Q''), (P, Q)
-    \in \Pi_j$, $|P' \cap P'' \cap P|=|Q' \cap Q'' \cap Q|$.
-  \end{itemize}
-\end{lemma}
-\begin{proof}
-  Proof is by induction on the number of iterations, $j$. In the rest
-  of the proof, the term ``new sets'' will refer to the new sets added
-  in $j$th iteration as defined in line \ref{setbreak} of Algorithm
-  \ref{perms}, i.e. the following three assignment pairs for some
-  $(P_1,Q_1), (P_2,Q_2) \in \Pi_{j-1}$ where $Q_1$ and $Q_2$ intersect
-  and share a leaf: $(P_1 \cap P_2, Q_1 \cap Q_2)$, or $(P_1 \setminus
-  P_2, Q_1 \setminus Q_2)$, or $(P_2 \setminus P_1, Q_2
-  \setminus Q_1)$.\\
-  \noindent
-  The base case, $\Pi_0 = \{(S_i,P_i) \mid i \in [m]\}$, is trivially
-  true since it is the input which is an ICPPA.  Assume the lemma is
-  true till the $j-1$ iteration. Consider $j$th
-  iteration:\\
-  \noindent
-  If $(P,Q)$, $(P',Q')$ and $(P'',Q'')$ are in $\Pi_{j}$ and
-  $\Pi_{j-1}$, all the invariants are
-  clearly true because they are from $j-1$ iteration.\\
-  If $(P,Q)$ is in $\Pi_{j}$ and not in $\Pi_{j-1}$, then it must be
-  one of the new sets added in $\Pi_j$. Since $(P_1,Q_1)$ and
-  $(P_2,Q_2)$ are from $\Pi_{j-1}$ and $Q_1,Q_2$ intersect and have a
-  common leaf, it can be verified that the
-  new sets are also paths. \\
-  By hypothesis for invariant III, invariant II also holds for $(P,Q)$
-  no matter which new set in $\Pi_j$ it
-  is.\\
-  To prove invariant III, if $(P,Q)$ and $(P',Q')$ are not in
-  $\Pi_{j-1}$, then they are both new sets and invariant III holds
-  trivially (new sets are disjoint). Next consider $(P,Q), (P',Q') \in
-  \Pi_j$ with only one of them, say $(P',Q')$, in $\Pi_{j-1}$. Then
-  $(P,Q)$ is one of the new sets added in line \ref{setbreak}. It is
-  easy to see that if $(P,Q)$ is $(P_1 \cap P_2, Q_1 \cap Q_2)$, then
-  due to invariant IV in hypothesis, invariant III becomes true in
-  this iteration. Similarly, using lemma \ref{lem:setminuscard}
-  invariant III is proven if $(P, Q)$ is $(P_1 \setminus P_2, Q_1
-  \setminus Q_2)$, or $(P_2 \setminus P_1, Q_2
-  \setminus Q_1)$.\\
-  To prove invariant IV, consider three assignments
-  $(P,Q),(P',Q'),(P'',Q'')$. If at least two of these pairs are in not
-  $\Pi_{j-1}$, then they are any two of the new sets. Note that these
-  new sets are disjoint and hence if $(P',Q'), (P'',Q'')$ are any of
-  these sets, $|P \cap P' \cap P''|=|Q \cap Q' \cap Q''|=0$ and
-  invariant IV is true. Now we consider the case if at most one of
-  $(P,Q),(P',Q'),(P'',Q'')$ is not in $\Pi_{j-1}$. If none of them are
-  not in $\Pi_{j-1}$ (i.e. all of them are in $\Pi_{j-1}$), invariant
-  IV is clearly true. Consider the case where exactly one of them is
-  not in $\Pi_{j-1}$. w.l.o.g let that be $(P,Q)$ and it could be any
-  one of the new sets. If $(P,Q)$ is $(P_1 \cap P_2, Q_1 \cap Q_2)$,
-  from lemma \ref{lem:fourpaths} and invariant III hypothesis,
-  invariant IV is proven. Similarly if $(P,Q)$ is any of the other new
-  sets, invariant IV is proven by also using lemma
-  \ref{lem:setminuscard}. \qed
-
-\end{proof}
-
-\noindent
-It can be observed that the output of algorithm \ref{perms} is such
-that every leaf is incident on at most a single path in the new set of
-assignments. This is due to the loop condition at line
-\ref{shareleaf}. Let $v_1$ be the leaf incident on path $P_i$. Assign
-to it any one element from $S_i \setminus \bigcup_{i \ne j}
-S_j$. Remove $(S_i, P_i)$ from assignments and add $(\{x_1\},
-\{v_1\}), (S_i \setminus \{x_1\}, P_i \setminus \{v_1\})$. Now all
-assignments except single leaf assignments are paths from the subtree
-$T_0 = T \setminus \{v \mid v \text{ is a leaf in } T\}$.
-
-% exclusively on the spine of the caterpillar. Consider the spine to
-% be an interval and run ICPIA algorithm on it to get the required
-% permutation. Algorithm \ref{leafasgn} makes this clear.
-
-\begin{algorithm}[h]
-  \caption{Leaf assignments from an ICPPA $\{(S_i,P_i) | i \in I\}$}
-  \label{leafasgn}
-  \begin{algorithmic}
-    \STATE Let $\Pi_0=\{(S_i,P_i)| i \in [m]\}$. Paths are such that
-    no
-    two paths $P_i, P_j, i \ne j$ share a leaf.\\
-    \STATE $j = 1$\\
-    \WHILE {there is a leaf $v$ and a unique $(S_{i_1}, P_{i_1})$ such
-      that $v \in P_{i_1}$}
-    \STATE $\Pi_j=   \Pi_{j-1} \setminus \{(S_{i_1}, P_{i_1})\}$\\
-    \STATE $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in I}S_i$
-    \IF{$X$ is empty} \STATE exit \ENDIF; \STATE Let $x = $ arbitrary
-    element from $X$ \vspace{2mm}
-
-    \STATE $\Pi_j = \Pi_j \cup \{(S_{i_1} \setminus \{x\}),(P_{i_1}
-    \setminus
-    \{v\}), (\{x\},\{v\})\}$\\
-    \STATE $j = j+1$\\
-    \ENDWHILE
-    \STATE $\Pi = \Pi_j$\\
-    \STATE Return $\Pi$\\
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma}
-  \label{lem:invar3}
-  In Algorithm \ref{leafasgn},for all $j \geq 0$, at the end of the
-  $j$th iteration the four invariants given in lemma \ref{lem:invar1}
-  are valid.
-  % Moreover, $X$ as defined in the algorithm is non-empty if this is
-  % an ICPPA.
-\end{lemma}
-\begin{proof}
-  First we see that $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in
-    I}S_i$ is non empty in every iteration for an ICPPA. Suppose $X$
-  is empty. We know that $v \in P_{i_1} \setminus \bigcup_{i \ne i_1,
-    i \in I}P_i$ since $v$ is in the unique path $P_{i_1}$. Since this
-  is an ICPPA $|S_{i_1}| = |P_{i_1}|$. For any $x \in S_{i_1}$ it is
-  contained in at least two sets due to our assumption. Let $S_{i_2}$
-  be a second set that contains $x$. We know $v \notin
-  P_{i_2}$. Therefore there cannot exist a permutation that maps
-  elements of $S_{i_2}$ to $P_{i_2}$. This contradicts our assumption
-  that this is an ICPPA. Thus $X$ cannot be empty.
-
-  We use mathematical induction on the number of iterations for this
-  proof. The term ``new sets'' will refer to the sets added in $\Pi_j$
-  in the $j$th iteration, i.e. $(P' \setminus \{x\},Q' \setminus
-  \{v\})$ and $(\{x\},\{v\})$ for some $(P',Q')$ in $\Pi_{j-1}$ such
-  that $v$ is a leaf and $Q'$ is the unique path
-  incident on it.\\
-  For $\Pi_0$ all invariants hold because it is output from algorithm
-  \ref{perms} which is an ICPPA. Hence base case is proved.  Assume
-  the lemma holds for $\Pi_{j-1}$. Consider $\Pi_j$ and any $(P,Q) \in
-  \Pi_j$. If $(P,Q) $ is in $ \Pi_j$ and $\Pi_{j-1}$ invariants I and
-  II are true because of induction assumption. If it is only in
-  $\Pi_j$, then it is $\{(P' \setminus \{x\}),(Q' \setminus \{v\})$ or
-  $(\{x\},\{v\})$ for some $(P',Q')$ in $\Pi_{j-1}$. By definition,
-  $x$ is an element in $P'$ (as defined in the algorithm) and $v$ is a
-  leaf in $Q'$. If $(P,Q)$ is $\{(P' \setminus \{x\}),(Q' \setminus
-  \{v\})$, $Q$ is a path since only a leaf is removed from path
-  $Q'$. We know $|P'| = |Q'|$, therefore $|P' \setminus \{x\}| = |Q'
-  \setminus \{v\}|$. Hence in this case invariants I and II are
-  obvious. It is easy to see these invariants hold if $(P,Q)$ is
-  $(\{x\},\{v\})$.
-
-  For invariant III consider $(P_1,Q_1),(P_2,Q_2)$ in $\Pi_j$. If both
-  of them are also in $\Pi_{j-1}$, claim is proved. If one of them is
-  not in $\Pi_{j-1}$ then it has to be $\{(P' \setminus \{x\}),(Q'
-  \setminus \{v\})$ or $(\{x\},\{v\})$ for some $(P',Q')$ in
-  $\Pi_{j-1}$. Since by definition, $Q'$ is the only path with $v$ and
-  $P'$ the only set with $x$ in the previous iteration, $|P_1 \cap (P'
-  \setminus \{x\})| = |P_1 \cap P'|$ and $|Q_1 \cap (Q' \setminus
-  \{v\})| = |Q_1 \cap Q'|$ and $|P_1 \cap \{x\}| = 0, Q_1 \cap \{v\} =
-  0$. Thus invariant III is also proven.
-
-  \noindent
-  To prove invariant IV, consider $(P_1,Q_1),(P_2,Q_2), (P_3,Q_3)$ in
-  $\Pi_j$. If exactly one of them, say $P_3 \notin \Pi_{j-1}$, it is
-  one of the new sets. By the same argument used to prove invariant
-  III, $|P_1 \cap P_2 \cap (P' \setminus \{x\})| = |P_1 \cap P_2 \cap
-  P'|$ and $|Q_1 \cap Q_2 \cap (Q' \setminus \{x\})| = |Q_1 \cap Q_2
-  \cap Q'|$. Since $P_1, P_2, P'$ are all in $\Pi_{j-1}$, by induction
-  hypothesis $|P_1 \cap P_2 \cap P'| = |Q_1 \cap Q_2 \cap Q'|$. Also
-  $|P_1 \cap P_2 \cap \{x\}| = 0, Q_1 \cap Q_2 \cap \{v\} = 0$.  If
-  two or more of them are not in $\Pi_{j-1}$, then it can be verified
-  that $|P_1 \cap P_2 \cap P_3| = |Q_1 \cap Q_2 \cap Q_3|$ since the
-  new sets in $\Pi_j$ are either disjoint or as follows: assuming
-  $P_1, P_2 \notin \Pi_{j-1}$ and new sets are derived from $(P', Q'),
-  (P'', Q'') \in \Pi_{j-1}$ with $x_1, x_2$ exclusively in $P_1, P_2$,
-  $(\{x_1\},\{v_1\}), (\{x_2\},\{v_2\}) \in \Pi_j $ thus $v_1, v_2$
-  are exclusively in $Q_1, Q_2$ resp. it follows that $|P_1 \cap P_2
-  \cap P_3| = |(P' \setminus \{x_1\}) \cap (P'' \setminus \{x_2\})
-  \cap P_3| = |P' \cap P'' \cap P_3| = |Q' \cap Q'' \cap Q_3| = |(Q'
-  \setminus \{v_1\} \cap Q'' \setminus \{v_2\} \cap Q_3| = |Q_1 \cap
-  Q_2 \cap Q_3|$. Thus invariant IV is also proven.  \qed
-\end{proof}
-
-\noindent
-Using algorithms \ref{perms} and \ref{leafasgn} we prove the following
-theorem.
-
-\begin{theorem}
-  \label{th:perm}
-  If $\cA$ is an ICPPA, then there exists a bijection $\sigma : U
-  \rightarrow V(T)$ such that $\sigma(S_i) = P_i$ for all $i \in I$
-\end{theorem}
-\begin{proof}
-  This is a contructive proof. First, the given ICPPA $\cA$ and tree
-  $T$ are given as input to Algorithm \ref{perms}. This yields a
-  ``filtered'' ICPPA as the output which is input to Algorithm
-  \ref{leafasgn}.  It can be observed that the output of Algorithm
-  \ref{leafasgn} is a set of interval assignments to sets and
-  one-to-one assignment of elements of $U$ to each leaf of $T$. To be
-  precise, it would be of the form $\cB_0 = \cA_0 \cup \cL_0$. The
-  leaf assignments are defined in $\cL_0 = \{ (x_i,v_i) \mid x_i \in
-  U, v_i \in T, x_i \ne x_j, v_i \ne v_j, i \ne j, i,j \in [k] \}$
-  where $k$ is the number of leaves in $T$. The path assignments are
-  defined in $\cA_0 \subseteq \{(S_i',P_i') \mid S_i' \subseteq U_0,
-  P_i' \text{ is a path from } T_0\}$ where $T_0$ is the tree obtained
-  by removing all the leaves in $T$ and $U_0 = U \setminus \{ x \mid x
-  \text{ is assigned to a leaf in }\cL_0 \}$. Now we have a subproblem
-  of finding the permutation for the path assignment $\cA_0$ which has
-  paths from tree $T_0$ and sets from universe $U_0$. Now we repeat
-  the procedure and the path assignment $\cA_0$ and tree $T_0$ is
-  given as input to Algorithm \ref{perms}. The output of this
-  algorithm is given to Algorithm \ref{leafasgn} to get a new union of
-  path and leaf assignments $\cB_1 = \cA_1 \cup \cL_1$ defined similar
-  to $\cB_0, \cL_0, \cA_0$. In general, the two algorithms are run on
-  path assignment $\cA_{i-1}$ with paths from tree $T_{i-1}$ to get a
-  new subproblem with path assignment $\cA_i$ and tree $T_{i}$. $T_i$
-  is the subtree of $T_{i-1}$ obtained by removing all its
-  leaves. More importantly, it gives leaf assignments $\cL_{i}$ to the
-  leaves in tree $T_{i-1}$. This is continued until we get a
-  subproblem with path assignment $\cA_{d-1}$ and tree $T_{d-1}$ for
-  some $d \le n$ which is just a path. From the last lemma we know
-  that $\cA_{d-1}$ is an ICPPA. Another observation is that an ICPPA
-  with all its tree paths being intervals (subpaths from a path) is
-  nothing but an ICPIA\cite{nsnrs09}.  Let $\cA_{d-1}$ be equal to
-  $\{(S_i'',P_i'') \mid S_i'' \subseteq U_{d-1}, P_i'' \text{ is a
-    path from } T_{d-1} \}$. It is true that the paths $P_i''$s may
-  not be precisely an interval in the sense of consecutive integers
-  because they are some nodes from a tree. However, it is easy to see
-  that the nodes of $T_{d-1}$ can be ordered from left to right and
-  ranked to get intervals $I_i$ for every path $P_i''$ as
-  follows. $I_i = \{[l,r] \mid l = \text{ the lowest rank of the nodes
-    in }P_i'', r = l+|P_i''|-1 \}$. Let asssignment $\cA_d$ be with
-  the renamed paths. $\cA_d = \{ (S_i'', I_i) \mid (S_i'', P_i'') \in
-  \cA_{d-1} \}$. What has been effectively done is renaming the nodes
-  in $T_{d-1}$ to get a tree $T_d$.  The ICPIA $\cA_d$ is now in the
-  format that the ICPIA algorithm requires which gives us the
-  permutation $\sigma' : U_{d-1} \rightarrow T_{d-1}$
-
-\noindent
-$\sigma'$ along with all the leaf assignments $\cL_i$ gives us the
-permutation for the original path assignment $\cA$.  More precisely,
-the permutation for tree path assignment $\cA$ is defined as
-follows. $\sigma: U \rightarrow T$ such that the following is
-maintained.
-\begin{align*}
-  \sigma(x) &= \sigma'(x),   \text{ if } x \in U_{d-1} \\
-  &= \cL_i(x), \text{ where $x$ is assigned to a leaf in a subproblem
-    $\cA_{i-1}, T_{i-1}$}
-\end{align*}
-
-\noindent
-To summarize, run algorithm \ref{perms} and \ref{leafasgn} on
-$T$. After the leaves have been assigned to specific elements from
-$U$, remove all leaves from $T$ to get new tree $T_0$. The leaf
-assignments are in $\cL_0$. Since only leaves were removed $T_0$ is
-indeed a tree. Repeat the algorithms on $T_0$ to get leaf assignments
-$\cL_{1}$. Remove the leaves in $T_0$ to get $T_1$ and so on until the
-pruned tree $T_d$ is a single path. Now run ICPIA algorithm on $T_d$
-to get permutation $\sigma'$. The relation $\cL_0 \cup \cL_1 \cup
-.. \cup \cL_{d} \cup \sigma'$ gives the bijection required in the
-original problem.\qed
-\end{proof}
-
-\subsection{Finding an assignment of tree paths to a set
-  system} \label{decompos} In the previous section we have shown that
-the problem of finding a Tree Path Asssignment to an input $(\cF,T)$
-is equivalent to finding an ICPPA to $\cF$ in tree $T$.  In this
-section we characterize those set systems that have an ICPPA in a
-given tree.  As a consequence of this characterization we identify two
-sub-problems that must be solved to obtain an ICPPA.  We do not solve
-the problem and in the next section show that finding an ICPPA in a
-given tree is GI-Complete.
-
-\noindent
-A set system can be concisely represented by a binary matrix where the
-row indices denote the universe of the set system and the column
-indices denote each of the sets. Let the binary matrix be $M$ with
-order $n \times m$, the set system be $\cF = \{S_i \mid i \in [m]\}$,
-universe of set system $U = \{x_1, \dots ,x_n\}$. If $M$ represents
-$\cF$, $|U| = n, |\cF| = m$. Thus $(i,j)$th element of $M$, $M_{ij} =
-1$ iff $x_i \in S_j$. If $\cF$ has a feasible tree path assignment
-(ICPPA) $\cA = \{(S_i,P_i) \mid i \in [m]\}$, then we say its
-corresponding matrix $M$ has an ICPPA. Conversly we say that a matrix
-$M$ has an ICPPA if there exists an ICPPA $\cA$ as defined
-above.\\
-\noindent
-We now define the strict intersection graph or overlap graph of
-$\cF$. This graph occurs at many places in the literature, see for
-example \cite{kklv10, wlh02, nsnrs09}.  The vertices of the graph
-correspond to the sets in $\cF$.  An edge is present between vertices
-of two sets iff the corresponding sets have a nonempty intersection
-and none is contained in the other. Formally, intersection graph is
-$G_f = (V_f, E_f)$ such that $V_f = \{v_i \mid S_i \in \cF\}$ and $E_f
-= \{(v_i, v_j) \mid S_i \cap S_j \ne \O \text{ and }S_i \nsubseteq
-S_j, S_j \nsubseteq S_i \}$.  We use this graph to decompose $M$ as
-described in \cite{wlh02,nsnrs09}.  A prime sub-matrix of $M$ is
-defined as the matrix formed by a set of columns of $M$ which
-correspond to a connected component of the graph $G_f$.  Let us denote
-the prime sub-matrices by $M_1,\ldots,M_p$ each corresponding to one
-of the $p$ components of $G_f$. Clearly, two distinct matrices have a
-distinct set of columns.  Let $col(M_i)$ be the set of columns in the
-sub-matrix $M_i$.  The support of a prime sub-matrix $M_i$ is defined
-as $supp(M_i) = \displaystyle \bigcup_{j \in col(M_i)}S_j$.  Note that
-for each $i$, $supp(M_i) \subseteq U$.  For a set of prime
-sub-matrices $X$ we define
-$supp(X) = \displaystyle \bigcup_{M \in X} supp(M)$. \\
-
-\noindent
-Consider the relation $\preccurlyeq$ on the prime sub-matrices $M_1,
-\ldots, M_p$ defined as follows:
-\begin{equation}
-  \nonumber \{(M_i,M_j) | \mbox{ A set } S \in
-  M_i \mbox{ is contained in a set } S' \in M_j\} \cup \{(M_i,M_i) | 1
-  \leq i \leq p\} 
-\end{equation}
-
-\noindent
-This relation is the same as that defined in \cite{nsnrs09}. The prime
-submatrices and the above relation can be defined for any set
-system. We will use this structure of prime submatrices to present our
-results on an an ICPPA for a set system $\cF$. Recall the following
-lemmas, and theorem that $\preccurlyeq$ is a partial order, from
-\cite{nsnrs09}.
-
-\begin{lemma} \label{lem:containment} Let $(M_i,M_j) \in
-  \preccurlyeq$.  Then there is a set $S' \in M_j$ such that for each
-  $S \in M_i$, $S \subseteq S'$.
-\end{lemma}
-\begin{lemma}
-  For each pair of prime sub-matrices, either $(M_i,M_j) \not\in
-  \preccurlyeq$ or $(M_j,M_i) \not\in \preccurlyeq$.
-  % If $(M_i,M_j) \in \preccurlyeq$ and $(M_j,M_i) \in \preccurlyeq$,
-  % then $i = j$ and $|M_i| = 1$.
-\end{lemma}
-\begin{lemma}
-  If $(M_i,M_j) \in \preccurlyeq $ and $(M_j,M_k) \in \preccurlyeq$,
-  then $(M_i,M_k) \in \preccurlyeq$.
-\end{lemma}
-\begin{lemma}
-  If $(M_i,M_j) \in \preccurlyeq$ and $(M_i,M_k) \in \preccurlyeq$,
-  then either $(M_j,M_k) \in \preccurlyeq$ or $(M_k,M_j) \in
-  \preccurlyeq$.
-\end{lemma}
-\begin{theorem} \label{thm:partitionold} $\preccurlyeq$ is a partial
-  order on the set of prime sub-matrices of $M$.  Further, it uniquely
-  partitions the prime sub-matrices of $M$ such that on each set in
-  the partition $\preccurlyeq$ induces a total order.
-\end{theorem}
-For the purposes of this paper, we refine the total order mentioned in
-Theorem \ref{thm:partitionold}. We do this by identifying an in-tree
-rooted at each maximal upper bound under $\preccurlyeq$.  Each of
-these in-trees will be on disjoint vertex sets, which in this case
-would be disjoint sets of prime-submatrices.  The in-trees are
-specified by selecting the appropriate edges from the Hasse diagram
-associated with $\preccurlyeq$.  Let $\cI$ be the following set:
-\begin{align*}
-  \cI = \{ (M_i,M_j) \in \preccurlyeq \mid \nexists M_k s.t. M_i
-  \preccurlyeq M_k, M_k \preccurlyeq M_j \} \cup \{ (M_i,M_i), i \in
-  [p] \}
-\end{align*}
-
-\begin{theorem} \label{thm:partition} Consider the directed graph $X$
-  whose vertices correspond to the prime sub-matrices, and the edges
-  are given by $\cI$.  Then, $X$ is a vertex disjoint collection of
-  in-trees and the root of each in-tree is a maximal upper bound in
-  $\preccurlyeq$.
-\end{theorem}
-\begin{proof}
-  To observe that $X$ is a collection of in-trees, we observe that for
-  vertices corresponding to maximal upper bounds, no out-going edge is
-  present in $I$.  Secondly, for each other element, exactly one
-  out-going edge is chosen, and for the minimal lower bound, there is
-  no in-coming edge.  Consequently, $X$ is acyclic, and since each
-  vertex has at most one edge leaving it, it follows that $X$ is a
-  collection of in-trees, and for each in-tree, the root is a maximal
-  upper bound in $\preccurlyeq$.  Hence the theorem.
-\end{proof}
-Let the partition of $X$ given by Theorem \ref{thm:partition} be
-$\{X_1,\ldots,X_r\}$.  Further, each in-tree itself can be layered
-based on the distance from the root.  The root is considered to be at
-level zero. For $j \geq 0$, Let $X_{i,j}$ denote the set of prime
-matrices in level $j$ of in-tree $X_i$.
-
-\begin{lemma}
-  \label{lem:subicppa}
-  Let $M$ be a matrix and let $X$ be the directed graph whose vertices
-  are in correspondence with the prime submatrices of $M$.  Further
-  let $\{X_1,\ldots,X_r\}$ be the partition of $X$ into in-trees as
-  defined above.  Then, matrix $M$ has an ICPPA in tree $T$ iff $T$
-  can be partitioned into vertex disjoint subtrees $\{T_1, T_2, \dots
-  T_r\}$ such that, for each $1 \leq i \leq r$, the set of prime
-  sub-matrices corresponding to vertices in $X_i$ has an ICPPA in
-  $T_i$.
-\end{lemma}
-\begin{proof}
-  Let us consider the reverse direction first.  Let us assume that $T$
-  can be partitioned into $T_1, \ldots, T_r$ such that for each $1
-  \leq i \leq r $, the set of prime sub-matrices corresponding to
-  vertices in $X_i$ has an ICPPA in $T_i$.  It is clear from the
-  properties of the partial order $\preccurlyeq$ that these ICPPAs
-  naturally yield an ICPPA of $M$ in $T$.  The main property used in
-  this inference is that for each $1 \leq i \neq j \leq r$, $supp(X_i)
-  \cap supp(X_j) = \phi$.
-
-\noindent
-To prove the forward direction, we show that if $M$ has an ICPPA, say
-$\cA$, in $T$, then there exists a partition of $T$ into vertex
-disjoint subtree $T_1, \ldots, T_r$ such that for each $1 \leq i \leq
-r$, the set of prime sub-matrices corresponding to vertices in $X_i$
-has an ICPPA in $T_i$.  For each $1 \leq i \leq r$, we define based on
-$\cA$ a subtree $T_i$ corresponding to $X_i$.  We then argue that the
-trees thus defined are vertex disjoint, and complete the proof.
-Consider $X_i$ and consider the prime sub-matrix in $X_{i,0}$.
-Consider the paths assigned under $\cA$ to the sets in the prime
-sub-matrix in $X_{i,0}$.  Since the component in $G_f$ corresponding
-to this matrix is a connected component, it follows that union of
-paths assigned to this prime-submatrix is a subtree of $T$.  We call
-this sub-tree $T_i$.  All other prime-submatrices in $X_i$ are
-assigned paths in $T_i$ since $\cA$ is an ICPPA, and the support of
-other prime sub-matrices in $X_i$ are contained in the support of the
-matrix in $X_{i,0}$.  Secondly, for each $1 \leq i \neq j \leq r$,
-$supp(X_i) \cap supp(X_j) = \phi$, and since $\cA$ is an ICPPA, it
-follows that $T_i$ and $T_j$ are vertex disjoint.  Finally, since $|U|
-= |V(T)|$, it follows that $T_1, \ldots, T_r$ is a partition of $T$
-into vertex disjoint sub-trees such that for each $1 \leq i \leq r$,
-the set of matrices corresponding to nodes in $X_i$ has an ICPPA in
-$T_i$.  Hence the lemma.
-\end{proof}
-The essence of the following lemma is that an ICPPA only needs to be
-assigned to the prime sub-matrix corresponding to the root of each
-in-tree, and all the other prime sub-matrices only need to have an
-Intersection Cardinality Preserving Interval Assignments (ICPIA).
-Recall, an ICPIA is an assignment of intervals to sets such that the
-cardinality of an assigned interval is same as the cardinality of the
-interval, and the cardinality of intersection of any two sets is same
-as the cardinality of the intersection of the corresponding intervals.
-It is shown in \cite{nsnrs09} that the existence of an ICPIA is a
-necessary and sufficient condition for a matrix to have COP.  We
-present the pseudo-code to test if $M$ has an ICPPA in $T$.
-\begin{lemma} \label{lem:rooticppa} Let $M$ be a matrix and let $X$ be
-  the directed graph whose vertices are in correspondence with the
-  prime submatrices of $M$.  Further let $\{X_1,\ldots,X_r\}$ be the
-  partition of $X$ into in-trees as defined earlier in this section.
-  Let $T$ be the given tree and let $\{T_1, \ldots, T_r\}$ be a given
-  partition of $T$ into vertex disjoint sub-trees.  Then, for each $1
-  \leq i \leq r$, the set of matrices corresponding to vertices of
-  $X_i$ has an ICPPA in $T_i$ if and only if the matrix in $X_{i,0}$
-  has an ICPPA in $T_i$ and all other matrices in $X_i$ have an {\bf
-    {\em ICPIA}}.
-\end{lemma}
-\begin{proof}
-  The proof is based on the following fact- $\preccurlyeq$ is a
-  partial order and $X$ is a digraph which is the disjoint union of
-  in-trees.  Each edge in the in-tree is a containment relationship
-  among the supports of the corresonding sub-matrices. Therefore, any
-  ICPPA to a prime sub-matrix that is not the root is contained in a
-  path assigned to the sets in the parent matrix.  Consequently, any
-  ICPPA to the prime sub-matrix that is not at the root is an ICPIA,
-  and any ICPIA can be used to construct an ICPPA to the matrices
-  corresponding to nodes in $X_i$ provided the matrix in the root has
-  an ICPPA in $T_i$.  Hence the lemma.
-\end{proof}
-Lemma \ref{lem:subicppa} and Lemma \ref{lem:rooticppa} point out two
-algorithmic challenges in finding an ICPPA for a given set system
-$\cF$ in a tree $T$.  Given $\cF$, finding $X$ and its partition
-$\{X_1,\ldots,X_r\}$ into in-trees can be done in polynomial time.  On
-the other hand, as per lemma \ref{lem:subicppa} we need to parition
-$T$ into vertex disjoint sub-trees $\{T_1, \ldots, T_r\}$ such that
-for each $i$, the set of matrices corresponding to nodes in $X_i$ have
-an ICPPA in $T_i$.  This seems to be a challenging step, and it must
-be remarked that this step is easy when $T$ itself is a path, as each
-individual $T_i$ would be sub-paths.  The second algorithmic challenge
-is identified by lemma \ref{lem:rooticppa} which is to assign an ICPPA
-from a given tree to the matrix associated with the root node of
-$X_i$.
-\begin{algorithm}[h]
-  \caption{Algorithm to find an ICPPA for a matrix $M$ on tree $T$:
-    $main\_ICPPA(M, T$)}
-  \label{al:icppa-main}
-  \begin{algorithmic}
-    \STATE Identify the prime sub-matrices. This is done by constructing the strict overlap graph and identify connected components.  Each connected component yields a prime sub-matrix.   \\
-    \STATE Construct the partial order $\preccurlyeq$ on the set of prime sub-matrices.  \\
-    \STATE Construct the partition $X_1,\ldots,X_r$ of the prime
-    sub-matrices induced by $\preccurlyeq$ \\
-    \STATE For each $1 \leq i \leq r$, Check if all matrices except
-    those in $X_{i,0}$ has an ICPIA.  If a matrix does not have ICPIA
-    exit with a negative answer.  To check for the existence of ICPIA,
-    use the result in \cite{nsnrs09}.  \STATE Find a partition of
-    $T_1, \ldots, T_r$ such that matrices in $X_{i,0}$ has an ICPPA in
-    $T_i$.  If not such partition exists, exit with negative answer.
-  \end{algorithmic}
-\end{algorithm}
-
-\subsection{Complexity of Tree Path Assignment-A
-  Discussion} 
-\label{complexity} 
-Recall that the input to the Tree
-Path Assignment question is an order pair $(\cF,T)$ where $\cF$ is a
-family of subsets of an universe $U$, and $T$ is a tree such that
-$|V(T)| = |U|$.  The question is to come up with a bijection from $U$
-to $V(T)$ such that the image of each set in $\cF$ is a path in $T$.
-We show that this problem is at least as hard as the problem of
-testing if two given path chordal graphs are isomorphic.
-\begin{theorem} \label{thm:isocomp} Tree Path Assignment is
-  isomorphism-complete.
-\end{theorem}
-\begin{proof}
-  It is well known (see for example \cite{kklv10}) that testing
-  isomorphism of path chordal graphs is isomorphism complete.  We show
-  a reduction of path chordal graph isomorphism to tree path
-  assignment.  Given $G_1$ and $G_2$ two path chordal graphs, let
-  $T_2$ be the clique tree of $G_2$ obtained from say \cite{gav78}.
-  The nodes of $T_2$ correspond to the maximal cliques of $G_2$ and
-  each vertex of $G_2$ corresponds to a path in $G_2$.  This is a
-  well-known characterization of path chordal graphs and $T_2$ can be
-  computed in polynomial time.  In $G_1$, let $S_v$ denote the maximal
-  cliques of $G_1$ that contain $v$.  This can be computed in
-  polynomial time as $G_1$ is path chordal, and all chordal graph only
-  have a linear number of maximal cliques.  The universe $U$
-  corresponds bijectively to the set of maximal cliques in $G_1$, and
-  $\cF = \{S_v | v \in V(G_1)\}$.  Now, we claim that $(\cF, T_2)$ has
-  a tree path assignment if and only if $G_1$ and $G_2$ are
-  isomorphic.  This is clear since for each vertex $v \in G_1$, there
-  is an associated $S_v$ which is the set of maximal cliques
-  containing $v$.  In $G_2$, each vertex corresponds to a path in
-  $T_2$, and the nodes on this path corresponds to the maximal cliques
-  in $G_2$.  Consequenlty, a tree path assignment will naturally yield
-  an isomorphism between $G_1$ and $G_2$, and vice versa.  Therefore,
-  Tree Path Assignment is isomorphism-complete.
-\end{proof}
-
-\subsubsection{Consecutive Ones Testing is in Logspace}
-While Tree Path Assignment is isomorphism-complete, it is polynomial
-time solvable when the given tree is a path.  Indeed, in this case we
-encounter a restatement of matrices with the COP.  The known
-approaches to testing for COP fall into two categories: those that
-provide a witness when the input matrix does not have COP, and those
-that do not provide a witness.  The first linear time algorithm for
-testing COP for a binary matrix was using a data structure called PQ
-trees, which represent all COP orderings of $M$, invented by
-\cite{bl76}. There is a PQ tree for a matrix iff the matrix has COP.
-Indeed, this is an algorithmic characterization of the consecutive
-ones property and the absence of the PQ-tree does not yield any
-witness to the reason for failure.  A closely related data structure
-is the generalized PQ tree in \cite{mcc04}.  In generalized PQ tree
-the P and Q nodes are called prime and linear nodes. Aside from that,
-it has a third type of node called degenerate nodes which is present
-only if the set system does not have COP \cite{mcc04}.  Using the idea
-of generalized PQ tree, \cite{mcc04} proves that checking for
-bipartiteness in the certain incomparability graph is sufficient to
-check for COP.  \cite{mcc04} invented a certificate to confirm when a
-binary matrix does not have COP.  \cite{mcc04} describes a graph
-called incompatibility graph of a set system $\cF$ which has vertices
-$(a,b), a \ne b$ for every $a, b \in U$, $U$ being the universe of the
-set system. There are edges $((a,b),(b,c))$ and $((b,a),(c,b))$ if
-there is a set $S \in \cF$ such that $a, c \in S$ and $b \notin S$. In
-other words the vertices of an edge in this graph represents two
-orderings that cannot occur in a consecutive ones ordering of $\cF$.
-\begin{theorem}[Theorem 6.1, \cite{mcc04}]
-  Let $\cF$ be an arbitrary set family on domain $V$. Then $\cF$ has
-  the consecutive ones property if and only if its incompatibility
-  graph is bipartite, and if it does not have the consecutive ones
-  property, the incompatibility graph has an odd cycle of length at
-  most $n+3$.
-\end{theorem}
-This theorem gives a certificate as to why a given matrix does not
-have COP.  Similarly, the approach of testing for an ICPIA in
-\cite{nsnrs09} also gives a different certificate- a prime sub-matrix
-that does not have an ICPIA.  Further, the above theorem can be used
-to check if a given matrix has COP in logspace by checking if its
-incompatibility graph is bipartite. \cite{rei84} showed that checking
-for bipartiteness can be done in logspace. Thus we conclude that
-consecutive ones testing can be done in logspace.
-
-\noindent
-More recently, \cite{kklv10} showed that interval graph isomorphism
-can be done in logspace. Their paper proves that a canon for interval
-graphs can be calculated in logspace using an interval hypergraph
-representation of the interval graph with each hyperedge being a set
-to which an interval shall be assigned by the canonization
-algorithm. An overlap graph (subgraph of intersection graph, edges
-define only strict intersections and no containment) of the hyperedges
-of the hypergraph is created and canons are computed for each overlap
-component. The overlap components define a tree like relation due to
-the fact that two overlap components are such that either all the
-hyperedges of one is disjoint from all in the other, or all of them
-are contained in one hyperedge in the other. This is similar to the
-containment tree defined in \cite{nsnrs09} and in this paper. Finally
-the canon for the whole graph is created using logspace tree
-canonization algorithm from \cite{sl92}. The interval labelling done
-in this process of canonization is exactly the same as the problem of
-assigning feasible intervals to a set system, and thus the problem of
-finding a COP ordering in a binary matrix \cite{nsnrs09}.
-
-\begin{theorem}[Theorem 4.7, \cite{kklv10}]
-  \label{th:canonlabel}
-  Given an interval hypergraph $\cH$, a canonical interval labeling
-  $l_H$ for $H$ can be computed in FL.
-\end{theorem}
-
-We present the following reduction to see that COP testing is indeed
-in logspace. Given a binary matrix $M$ of order $n \times m$, let $S_i
-= \{j \mid M[j,i]=1 \}$. Let $\cF = \{S_i \mid i \in [m] \}$ be this
-set system. Construct a hypergraph $\cH$ with its vertex set being
-$\{1, 2, \dots n\}$. The edge set of $\cH$ is isomorphic to
-$\cF$. Thus every edge in $\cH$ represents a set in the given set
-system $\cF$. Let this mapping be $\pi: E(\cH) \rightarrow \cF$. It is
-easy to see that if $M$ has COP, then $\cH$ is an interval
-hypergraph. From theorem \ref{th:canonlabel}, it is clear that the
-interval labeling $l_{\cH}: V(\cH) \rightarrow [n]$ can be calculated
-in logspace. Construct sets $I_i = \{ \l_{\cH}(x) \mid x \in E, E \in
-E(\cH), \pi(E) = S_i\}$, for all $i \in [m]$. Since $\cH$ is an
-interval hypergraph, $I_i$ is an interval for all $i \in [m]$, and is
-the interval assigned to $S_i$ if $M$ has COP.
-
-Now we have the following corollary.
-\begin{corollary}
-  \label{cor:coplog}
-  If a binary matrix $M$ has COP then the interval assignments to each
-  of its columns can be calculated in FL.
-\end{corollary}
-
-\noindent
-Finally, we conclude by asking about the complexity of Tree Path
-Assignment restricted to other subclasses of trees.  In particular, is
-Tree Path Assignment in caterpillars easier than Tree Path assignment
-in general trees.
-
-\chapter{Conclusion}
-\label{ch:conclusion}
-
-\begin{center}
-{\tt MOVED FROM INTRO SUBSECTION IN ch:myresearch CHAPTER begin}
-\end{center}
-
-% \annote{
-  It is an interesting fact that for a matrix with the COP, the
-  intersection graph of the corresponding set system is an interval
-  graph. % }{the following text could be in Further Study/Conclusion}
-A
-similar connection to a subclass of chordal graphs and a superclass of
-interval graphs exists for the generalization of COP.  In this case,
-the intersection graph of the corresponding set system must be a {\em
-  path graph}. Chordal graphs are of great significance, are
-extensively studied, and have several applications.  One of the well
-known and interesting properties of a chordal graphs is its connection
-with intersection graphs \cite{mcg04}. For every chordal graph, there
-exists a tree and a family of subtrees of this tree such that the
-intersection graph of this family is isomorphic to the chordal graph
-\cite{plr70,gav78,bp93}.  These trees when in a certain format, are
-called {\em clique trees} \cite{apy92} of the chordal graph. This is a
-compact representation of the chordal graph. There has also been work
-done on the generalization of clique trees to clique hypergraphs
-\cite{km02}.  If the chordal graph can be represented as the
-intersection graph of paths in a tree, then the graph is called path
-graph \cite{mcg04}.  Therefore, it is clear that if there is a
-bijection from $U$ to $V(T)$ such that for every set, the elements in
-it map to vertices of a unique path in $T$, then the intersection
-graph of the set system is indeed a path graph.  However, this is only
-a necessary condition and can be checked efficiently because path
-graph recognition is polynomial time
-solvable\cite{gav78,aas93}. Indeed, it is possible to construct a set
-system and tree, such that the intersection graph is a path graph, but
-there is no bijection between $U$ and $V(T)$ such that the sets map to
-paths. Path graph isomorphism is known be isomorphism-complete, see
-for example \cite{kklv10}. An interesting area of research would be to
-see what this result tells us about the complexity of the tree path
-labeling problem (not covered in this paper).
-
-\begin{center}
-{\tt MOVED FROM INTRO SUBSECTION IN ch:myresearch CHAPTER\\ end}
-\end{center}
-
-We give a characterization for feasible tree path labeling of path
-hypergraphs. The proof for this is constructive and computes a
-feasibility bijection mapping vertices of the hypergraph to vertices
-of the given target tree. This thesis also discovered an exponential algorithm
-that computes a feasible TPL to a given hypergraph if it is a path hypergraph.
-
-Our results have close connections to recognition of path graphs and
-to path graph isomorphism.  Graphs which can be represented as the
-intersection graph of paths in a tree are called {\em path
-  graphs}\cite{mcg04}. Thus, a hypergraph $\cF$ can be interpreted as
-paths in a tree, if and only if the intersection graph of $\cF$ is a
-{\em path graph}. Path graphs are a subclass of chordal graphs since
-chordal graphs are characterized as the intersection graphs of
-subtrees of a tree\cite{mcg04}.  Path graphs are well studied in the
-literature \cite{plr70}--\cite{mcg04}\tnote[TH]{Check if cite:bp93 is
-  indeed for path graphs}.  Path graph recognition can be done in
-polynomial time\cite{gav78,aas93}.  Clearly, this is a necessary
-condition in terms of the intersection graph of the input hypergraph
-$\cF$ in \FTPL. However, one can easily obtain a counterexample to
-show the insufficiency of this condition\tnote[TH]{Give a figure
-  showing insuff.}.  Path graph isomorphism is known be
-isomorphism-complete\cite{kklv10}. Therefore, \annote[TH]{it is
-  unlikely that we can solve the problem of finding feasible path
-  labeling $\cl$ for a given $\cF$ and }{WHY?} tree $T$.  It is
-definitely interesting to classify the kinds of trees and hypergraphs
-for which feasible path labelings can be\tnote[TH]{ELABORATE.}  found
-efficiently.  \annote[TH]{These results would form a natural
-  generalization of COP testing and interval graph isomorphism,
-  culminating in Graph Isomorphism }{REWORD} itself.  To this effect
-we consider TPL on {\kstar} and give a polynomial time solution for
-the same.
-
-While we address the computation of TPL
-for a given hypergraph% \footnote{The term {\em hypergraph} is used
-%   instead {\em set systems} are used interchangably in this
-%   section. Hypergraphs that have TPLs on some target tree are called
-%   {\em path hypergraphs}.}
-from a given target tree in this research, however optimization
-problems in TPL, akin to problems in Section~\ref{sec:optcop} for COP,
-is outside the scope of this thesis. Whether TPL on general trees is
-solvable in $P$ remains open. So do optimization opportunities in TPL
-(possible extensions to optimization for COP in the
-Section~\ref{sec:optcop}.\tnote[TH]{link to section}).
-
-%\newpage %DUMMY
-\tnote[TH]{ ||Complexity challenges:|| Canonization is an important
-  tool in graph isomorphism. Invention of a
-% deterministic method of canonization for any class of graphs naturally
-% results in an algorithm for isomorphism. All that is required is to
-% check if the canons of two graphs are the same. Thus complexity of
-% graph isomorphism can be studied by studying canonization
-% methods. While general graph ismorphism remains elusive in terms of
-% hardness, canonization has been studied for smaller classes of graphs
-% thus giving complexity/hardness results for them.
-
-% In 1992, \cite{sl92} made an important discovery that tree isomorphism
-% is in logspace. It was done by inventing a method of canonization of
-% trees using a logspace depth first traversal
-% algorithm. \tnote{confirm the statement of claim}.
-
-% \cite{kklv10} proved that interval graph canonization is also in
-% logspace thus drawing logspace conslusions about COT. 
-% % This result is
-% % built on the fact that 
-% Interval graphs are FP+C (fixed point with
-% counting) definable \tnote{elucidate} and \cite{l10} showed that
-% this implies that it captures PTIME. \tnote{elucidate} This result
-% along with that of undirected graph reachability being in logspace
-% \cite{rei08}, \cite{kklv10} proved their logspace result.
-}   %
-\tnote[TH]{ (1) the interesting survey in conclusion section of kklv10
-  (2) [[[COP helping some problems with hardness]]] having COP in the
-  input structure makes some problems less hard than in a general
-  input. -- min set cover problem is known to be NP hard. so is min
-  hitting
-  set. these two problems are equivalent cite-adp80\\
-  -- -- complexity results: log factor polynomial time approximable cite-f98, W[2]-complete with solution size as parameter cite-df99\\
-  -- a couple of generalizations:\\
-  -- -- min degree hypergraph (MDH) problem\\
-  -- -- red blue covering (RBSC) problem\\
-  ----- cite-d08phd Chapter 5.\\
-  ----- what are cite-d08phd results\\
-  ----- Rectangle stabbing \\
-  ----- see cite-d08phd Chapter 6?  }
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Appendices.
-
-% \appendix %%% NON NEIN NAHI ILLA NONE
-
-% \chapter{A}
-% \tnote{DO WE NEED THIS?}
-
-\newpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Bibliography.
-\begin{singlespace}
-%\begin{thebibliography}{10}
-\addcontentsline{toc}{chapter}{REFERENCES}
-
-{\footnotesize
-  \bibliographystyle{alpha} 
-  \bibliography{../lib/cop-variants__thesis}
-}
-%\end{thebibliography}
-\end{singlespace}
-%\newpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% List of papers
-
-% \listofpapers %%% NON NEIN NAHI ILLA NONE
-
-% \begin{enumerate}  
-% \item Authors....  \newblock
-%  Title...
-%   \newblock {\em Journal}, Volume,
-%   Page, (year).
-% \end{enumerate}
-
-\end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% DOCUMENTATION ON THE THESIS CLASS FILE ETC.
-
-% Before reading any further please note that you are strongly advised
-% against changing any of the formatting options used in the class
-% provided in this directory, unless you are absolutely sure that it
-% does not violate the IITM formatting guidelines.  \emph{Please do not
-%   change the margins or the spacing.}  If you do change the formatting
-% you are on your own (don't blame me if you need to reprint your entire
-% thesis).  In the case that you do change the formatting despite these
-% warnings, the least I ask is that you do not redistribute your style
-% files to your friends (or enemies).
-
-% It is also a good idea to take a quick look at the formatting
-% guidelines.  Your office or advisor should have a copy.  If they
-% don't, pester them, they really should have the formatting guidelines
-% readily available somewhere.
-
-% To generate PDF's with the links from the \verb+hyperref+ package use
-% the following command:
-% \begin{verbatim}
-% % dvipdfm -o thesis.pdf thesis.dvi
-% \end{verbatim}
-
-% \section{Package Options}
-
-% Use this thesis as a basic template to format your thesis.  The
-% \verb+iitmdiss+ class can be used by simply using something like this:
-% \begin{verbatim}
-% \documentclass[PhD]{iitmdiss}  
-% \end{verbatim}
-
-% To change the title page for different degrees just change the option
-% from \verb+PhD+ to one of \verb+MS+, \verb+MTech+ or \verb+BTech+.
-% The dual degree pages are not supported yet but should be quite easy
-% to add.  The title page formatting really depends on how large or
-% small your thesis title is.  Consequently it might require some hand
-% tuning.  Edit your version of \verb+iitmdiss.cls+ suitably to do this.
-% I recommend that this be done once your title is final.
-
-% To write a synopsis simply use the \verb+synopsis.tex+ file as a
-% simple template.  The synopsis option turns this on and can be used as
-% shown below.
-% \begin{verbatim}
-% \documentclass[PhD,synopsis]{iitmdiss}                                
-% \end{verbatim}
-
-% Once again the title page may require some small amount of fine
-% tuning.  This is again easily done by editing the class file.
-
-% This sample file uses the \verb+hyperref+ package that makes all
-% labels and references clickable in both the generated DVI and PDF
-% files.  These are very useful when reading the document online and do
-% not affect the output when the files are printed.
-
-% \section{Example Figures and tables}
-
-% Fig.~\ref{fig:iitm} shows a simple figure for illustration along with
-% a long caption.  The formatting of the caption text is automatically
-% single spaced and indented.  Table~\ref{tab:sample} shows a sample
-% table with the caption placed correctly.  The caption for this should
-% always be placed before the table as shown in the example.
-
-% \begin{figure}[htpb]
-%   \begin{center}
-%     \resizebox{50mm}{!} {\includegraphics *{iitm.eps}}
-%     \resizebox{50mm}{!} {\includegraphics *{iitm.eps}}
-%     \caption {Two IITM logos in a row.  This is also an
-%       illustration of a very long figure caption that wraps around two
-%       two lines.  Notice that the caption is single-spaced.}
-%   \label{fig:iitm}
-%   \end{center}
-% \end{figure}
-
-% \begin{table}[htbp]
-%   \caption{A sample table with a table caption placed
-%     appropriately. This caption is also very long and is
-%     single-spaced.  Also notice how the text is aligned.}
-%   \begin{center}
-%   \begin{tabular}[c]{|c|r|} \hline
-%     $x$ & $x^2$ \\ \hline
-%     1  &  1   \\
-%     2  &  4  \\
-%     3  &  9  \\
-%     4  &  16  \\
-%     5  &  25  \\
-%     6  &  36  \\
-%     7  &  49  \\
-%     8  &  64  \\ \hline
-%   \end{tabular}
-%   \label{tab:sample}
-%   \end{center}
-% \end{table}
-
-% \section{Bibliography with BIB\TeX}
-
-% I strongly recommend that you use BIB\TeX\ to automatically generate
-% your bibliography.  It makes managing your references much easier.  It
-% is an excellent way to organize your references and reuse them.  You
-% can use one set of entries for your references and cite them in your
-% thesis, papers and reports.  If you haven't used it anytime before
-% please invest some time learning how to use it.
-
-% I've included a simple example BIB\TeX\ file along in this directory
-% called \verb+refs.bib+.  The \verb+iitmdiss.cls+ class package which
-% is used in this thesis and for the synopsis uses the \verb+natbib+
-% package to format the references along with a customized bibliography
-% style provided as the \verb+iitm.bst+ file in the directory containing
-% \verb+thesis.tex+.  Documentation for the \verb+natbib+ package should
-% be available in your distribution of \LaTeX.  Basically, to cite the
-% author along with the author name and year use \verb+\cite{key}+ where
-% \verb+key+ is the citation key for your bibliography entry.  You can
-% also use \verb+\citet{key}+ to get the same effect.  To make the
-% citation without the author name in the main text but inside the
-% parenthesis use \verb+\citep{key}+.  The following paragraph shows how
-% citations can be used in text effectively.
-
-% More information on BIB\TeX\ is available in the book by
-% \cite{lamport:86}.  There are many
-% references~\citep{lamport:86,prabhu:xx} that explain how to use
-% BIB\TeX.  Read the \verb+natbib+ package documentation for more
-% details on how to cite things differently.
-
-% Here are other references for example.  \citet{viz:mayavi} presents a
-% Python based visualization system called MayaVi in a conference paper.
-% \citet{pan:pr:flat-fst} illustrates a journal article with multiple
-% authors.  Python~\citep{py:python} is a programming language and is
-% cited here to show how to cite something that is best identified with
-% a URL.
-
-% \section{Other useful \LaTeX\ packages}
-
-% The following packages might be useful when writing your thesis.
-
-% \begin{itemize}  
-% \item It is very useful to include line numbers in your document.
-%   That way, it is very easy for people to suggest corrections to your
-%   text.  I recommend the use of the \texttt{lineno} package for this
-%   purpose.  This is not a standard package but can be obtained on the
-%   internet.  The directory containing this file should contain a
-%   lineno directory that includes the package along with documentation
-%   for it.
-
-% \item The \texttt{listings} package should be available with your
-%   distribution of \LaTeX.  This package is very useful when one needs
-%   to list source code or pseudo-code.
-
-% \item For special figure captions the \texttt{ccaption} package may be
-%   useful.  This is specially useful if one has a figure that spans
-%   more than two pages and you need to use the same figure number.
-
-% \item The notation page can be entered manually or automatically
-%   generated using the \texttt{nomencl} package.
-
-% \end{itemize}
-
-% More details on how to use these specific packages are available along
-% with the documentation of the respective packages.
-
-% END OF DOCUMENTATION
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-01-03 src/010chap1.tex
Cleaned up thesis.tex. Split thesis.tex into chapterwise subfiles for easy edit tex management. Content changes are mainly in chapter 1 and 3 although nothing significant. To being working on chapter 2 after this commit.

--- 2011-12-30 src/thesis.tex
@@
-\documentclass[MS]             % MS thesis dissertation.
-              {iitmdiss_as}    % Hacked by hyperbolicme.
-%              {iitmdiss}      % If the powers to be are finicky.
-
-\usepackage{times}             % Font?
-\usepackage{t1enc}             % T1 encoding??  
-%\usepackage{boxedminipage}
-\usepackage{../lib/mythesislib}
-\usepackage{../lib/mylatexlib}
-% \usepackage{comment}           % DEBUG ONLY
-
-\def \mythesissubmissiondate {10 January 2012}
-\def \mythesissubmissionmonth {\MakeUppercase {January 2012}}
-
-\begin{document}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Title page
-\title{\mythesistitleNOTSYNPOSIS} 
-\author{\myname} 
-\date{
-  {\mythesissubmissionmonth}} 
-\department{\MakeUppercase{\mythesisdept}}
-%\department{\mythesisdept}
-
-%\nocite{*}
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Certificate
-\certificate
-
-\vspace*{0.5in}
-
-\noindent 
-This is to certify that the thesis titled {\bf \mythesistitleNOTSYNPOSIS},
-submitted by {\bf \myname}, to the {\bf Indian Institute of Technology
-Madras}, for the award of the degree of {\bf \mydegree}, is 
-a bona fide record of the research work done by her under our
-supervision.  The contents of this thesis, in full or in parts, have
-not been submitted to any other Institute or University for the award
-of any degree or diploma.
-
-\vspace*{1.5in}
-
-\begin{singlespacing}
-  \hspace*{-0.25in}
-  \parbox{3in}{
-    \noindent {\bf Dr.~\myadvisor} \\
-    \small
-    \noindent Research Guide\\
-    \noindent Associate Professor\\
-    \noindent Dept. of \mydept\\
-    \noindent IIT Madras -- 600 036
-  } \hfill
-  \parbox{1.25in}{               % {1.75in}
-    \small
-    \vspace*{0.75in}
-    \begin{tabular}[h]{r}
-    \noindent {Chennai}\\         
-    \noindent {\mythesissubmissiondate} 
-    \end{tabular}
-  }
-\end{singlespacing}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Acknowledgements
-\acknowledgements
-
-\tnote{WIP}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Abstract
-
-\begin{abstract}
-  \noindent {\em {\bf Keywords}: \hspace*{0.5em} \parbox[t]{4.4in}{consecutive
-    ones property, algorithmic graph theory, hypergraph isomorphism,
-    interval labeling}} \vspace*{24pt}
-
-  \def \tem {}
-  \noindent 
-  Consecutive-ones property is a non-trivial property of binary
-  matrices that has been studied widely in the literature for over
-  past 50 years. Detection of COP in a matrix is possible efficiently
-  and there are several algorithms that achieve the same. This thesis
-  documents the work done on an extension of COP extended from the
-  equivalent interval assignment problem in \cite{nsnrs09}. These new
-  results rigorously prove a natural extension (to trees) of their
-  characterization as well as makes connections to graph isomorphism,
-  namely path graph isomorphism.
-
-  \tnote{EXPAND. 
-  Abstract must be a brief about what results we have and how it fits
-  in the body of research.\\
-  -- Area: Broad to Specialized. i.e. Combinatorial
-    algorithms -> Matrix reorganization -> general data reorganization
-    (interval assignment) -> path assignment\\
-  -- Class of problems: say, data reorganization.\\
-  -- Nature of results: Is a generalization. We have a
-    Polynomial algorithm for a subset of the generalization.\\
-% maybe add - we explore a natural
-%     generalization of results on binary matrices with the |\tem
-%     consecutive-ones property|.  We consider the following constraint
-%     satisfaction problem. Given (i) a set system $\F \subseteq$
-%     $(2^{U} \setminus \emptyset)$ of a finite set $U$ of cardinality
-%     $n$, (ii) a tree $T$ of size $n$ and (iii) a bijection called
-%     |\tem tree path labeling|, $\cl$ mapping the sets in $\cF$ to
-%     paths in $T$, does there exist at least one bijection $\phi:U
-%     \rightarrow V(T)$ such that for each $S \in \cF$, $\{\phi(x) \mid
-%     x \in S\} = \cl(S)$?  A tree path labeling of a set system is
-%     called |\tem feasible| if there exists such a bijection $\phi$.
-%     We present an algorithmic characterization of feasible tree path
-%     labeling. COP is a special instance of tree path labeling problem
-%     when $T$ is a path.  We conclude with a polynomial time algorithm
-%     to find a feasible tree path labeling of a given set system when
-%     $T$ is a |\tem $k$-subdivided star|, set system has a single
-%     containment tree of overlap components and set size is limited to
-%     at most $k+2$.
-  }
-\end{abstract}
-
-%\pagebreak
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Table of contents etc.
-
-\begin{singlespace}
-\tableofcontents
-\thispagestyle{empty}
-
-\listoftables
-\addcontentsline{toc}{chapter}{LIST OF TABLES}
-\listoffigures
-\addcontentsline{toc}{chapter}{LIST OF FIGURES}
-\end{singlespace}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Abbreviations
-\abbreviations
-
-\noindent 
-\begin{tabbing}
-xxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
-\textbf{COP}   \> Consecutive-ones Property \\
-\textbf{COT}   \> Consecutive-ones Testing \\
-\textbf{ICPIA}   \> Intersection Cardinality Preservation Interval Assignment \\
-\textbf{ICPPL}   \> Intersection Cardinality Preserved Path Labeling \\
-% ADD MORE HERE
-\end{tabbing}
-
-\pagebreak
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Notation
-
-\chapter*{\centerline{NOTATION}}
-\addcontentsline{toc}{chapter}{NOTATION}
-
-\begin{singlespace}
-\begin{tabbing}
-xxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
-\textbf{$2^{U}$}  \> Powerset of set $U$ \\
-\end{tabbing}
-\end{singlespace}
-
-\pagebreak
-\clearpage
-
-% The main text will follow from this point so set the page numbering
-% to arabic from here on.
-\pagenumbering{arabic}
-
 \chapter{Introduction}
@@
 
-\tnote{Have a para on Organization: Outline of document}
-Section~\ref{sec:background} gives a brief survey of COP and
-optimization problems related to it followed by motivation for the
+\section{Organization of the document}
+\label{sec:orgofdoc}
+This chapter (Chapter~\ref{ch:intro}) introduces the area of research
+and the problems addressed in this thesis.  Section~\ref{sec:problem}
+introduces the main problem of this thesis by way of an
+illustration. Section~\ref{sec:basicprelim} lays out a few general
+definitions that are helpful in understanding the rest of the chapter.
+Section~\ref{sec:background} gives a brief survey of COP
+and optimization problems related to it followed by motivation for the
 thesis in Section~\ref{sec:motive}.  Section~\ref{sec:results}
 presents a summary of our results on the extension of COP namely, the
-tree path labeling problem. \tnote[TH]{update for thesis}
-
-% \tnote{testing 1 2 3...}
-% \tnote[bogus1]{testing 1 2 3...}
-% \tnote[bogus2]{testing 1 2 3...}
-% \tnote[GTC]{testing 1 2 3...}
-% \tnote[bogus4]{testing 1 2 3...}
-% \tnote[TH]{testing 1 2 3...}
+tree path labeling problem.
+
+Chapter~\ref{ch:copsurvey} gives a more detailed survey briefed in
+Section~\ref{sec:background}.  Chapter~\ref{ch:myresearch} details all
+the results obtained to the problems of this thesis and finally the
+conclusion of the thesis is discussed in Chapter~\ref{ch:conclusion}.
+
+% \tnote{testing 1 2 3...}%
+% \tnote[bogus1]{testing 1 2 3...}%
+% \tnote[important]{testing 1 2 3...}% 
+% \tnote[dire]{testing 1 2 3...}% 
+% \tnote[minor]{testing 1 2 3...}% 
+% \tnote[pressing]{testing 1 2 3...}% 
+% \tnote[UNDEF]{testing 1 2 3...}%
 
@@
 form the namesake study groups, {\LLL}, {\GGG}, {\BBB} and
-{\TTT}\tnote[TH]{put bib entries for these works!}. A student will be
+{\TTT}\tnote[minor]{put bib entries for these works!}. A student will be
 in at least one study group and may be in more than one. For instance,
@@
 accommodation.  All these apartments are located such that the streets
-connecting them do {\em not} form
-loops. Fig~\ref{fig:streetmappathpeople} \tnote[TH]{fig:streetmap}
-shows the street map for {\residenceblock}. It may be noted that as a
-graph, it classifies as a tree.
-
-\begin{table}[htbp]
+connecting them do {\em not} form loops. Figure~\ref{fig:streetmap} shows
+the street map for {\residenceblock}. It may be noted that as a graph,
+it classifies as a tree.
+
+\begin{table}[t] %[htbp]
   \centering
@@
   }
-  \caption{\figtabsize A solution to study group accommodation problem}
+  \caption{\figtabsize A solution to \illustrationproblem} %study group accommodation problem
   \label{tab:iltree}
@@
 A natural question would be to find how the students should be
-allocated apartments such that each study group has the {\em least
-  distance to travel} for a discussion? More specifically, we are
-interested in enforcing additional conditions, namely, that all the
-students in a study group must be next to each other; in other words,
-for one student to reach another fellow study group member's apartment
-(for all study groups the student is part of), she must not have to
-pass the apartment of any student who is not in that study group. To
+allocated apartments such that each study group has the least distance
+to travel for a discussion? More specifically, we are interested in
+the problem with additional conditions, namely, that all the students
+in a study group must be next to each other; in other words, for one
+student to reach another fellow study group member's apartment (for
+all study groups the student is part of), she must not have to pass
+the apartment of any student who is not in that study group. To
 further elucidate, the apartments of students of any study group must
@@
 allocation to individual students. A feasible allocation of exclusive
-routes to study groups is illustrated in
-Fig~\ref{fig:streetmappathpeople} \tnote[TH]{fig:streetmappath} and
-the students' allocation of apartments that obeys this route
-allocation is also shown. Table~\ref{tab:iltree} shows the same
-solution set theoretically.  How this is algorithmically computed is
-the focus of this thesis.
+routes to study groups is illustrated in Figure~\ref{fig:streetmap}.  The
+students' allocation of apartments that obeys this route allocation is
+shown in Figure~\ref{fig:streetmappathpeople}. Table~\ref{tab:iltree}
+shows the same solution set theoretically.  How this is
+algorithmically computed is the focus of this thesis.
+
+%D<debug> figures inside tables
+% \begin{tabular}{cc}
+%  \begin{figure} %[t]%
+% %    \centering
+%     \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
+% %    \label{fig:streetmap}
+%     \caption{--placeholder--}
+%   \end{figure}
+%     &
+%   \begin{figure} %[t]%
+% %    \centering
+%     \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
+% %    \label{fig:streetmappath}
+%     \caption{--placeholder--}
+%   \end{figure}
+% \end{tabular}
 
@@
   \centering
-  (a) \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
-  (b) \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
-
+  \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
+  \caption{\figtabsize {\residenceblock} street map.}
   \label{fig:streetmap}
- \caption{\figtabsize (a) {\residenceblock} street map (b)
-   {\residenceblock} street map with study group routes
-   allocated. Routes are color coded as follows: red for
+
+  \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
+  \caption[\figtabsize \residenceblock street map with study group
+    routes allocated.]{\figtabsize {\residenceblock} street map with study group
+    routes allocated. Routes are color coded as follows: red for
     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
@@
     \textcolor{green}{$\xTTT$} group}
+  \label{fig:streetmappath}
+%   \caption{\figtabsize (a) {\residenceblock} street map (b)
+%     {\residenceblock} street map with study group routes
+%     allocated. Routes are color coded as follows: red for
+%     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
+%     group, orange for \textcolor{yellow}{$\xBBB$} group, green for
+%     \textcolor{green}{$\xTTT$} group}
+%  \label{fig:streetmap}
 
   \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
-  \caption{\figtabsize Individual allocation of apartments to students
-    in {\residenceblock} that meets the requirements stated before.
-    The routes are color coded as follows: red for
+  \caption[\figtabsize Solution to the student accommodation
+  problem.]{\figtabsize Individual allocation of apartments to
+    students in {\residenceblock} that meets the requirements stated
+    before.  The routes are color coded as follows: red for
     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
-    group, orange for \textcolor{yellow}{$\xBBB$} group, green
-    for \textcolor{green}{$\xTTT$} group. {\tiny {\em Peanuts images
+    group, orange for \textcolor{yellow}{$\xBBB$} group, green for
+    \textcolor{green}{$\xTTT$} group.\\ {\tiny {\em Peanuts images
         {\copyright} Charles Schulz}}}% YellowOrange
   \label{fig:streetmappathpeople}
-\end{figure}
-\tnote{UPDATE IMAGE (b) 2-infinite-loop-BTWF.png. REMOVE TITLES. ADD
-  {$\xLLL$} {$\xGGG$} {$\xBBB$} {$\xTTT$}}
-
-\tnote[TH]{make dashed/textured lines for routes. make it color
-  agnostic.}
+\end{figure} \tnote{[i] UPDATE IMAGE (b) 2-infinite-loop-BTWF.png. REMOVE
+  TITLES. ADD {$\xLLL$} {$\xGGG$} {$\xBBB$} {$\xTTT$} [ii]make
+  dashed/textured lines for routes. make it color agnostic.}
 
@@
 problem (where the ``target graph'' is a path and not any
-tree\tnote[TH]{Allowing any tree in this example could be seen as a
+tree\tnote[minor]{Allowing any tree in this example could be seen as a
   scenario where there are parallel classes. A node falling in the
@@
 \section{Basic preliminaries - general definitions and nomenclature}
-\tnote{(definitions theorems etc) needed if any
-    - graph theory}
+\label{sec:basicprelim}
+\tnote{(definitions theorems etc) needed if any - graph theory}
 
@@
 
-In this section, a brief survey of the consecutive-ones problem and its
-optimization problems is presented.
-
-%%%%%%                              %%%%%%
-%THX is the second level of ``tnote''-ing%
-%%%%%%                              %%%%%%
-\tnote[TH]{ADD: As it will be described in detail later in this
-  document, isomorphism of certain % classes of graphs, namely chordal
-%   graphs, have a close relationship with consecutive-ones property and
-%   generalizations of it.  This is perhaps because of how closely COP
-%   of a matrix relates to properties of graphs derived from matrices as
-%   seen in the following results.
-}
-\tnote[THX]{ADD: peo exists iff chordal. lexicographic BFS
-  [tag:chordalGraph]} %
-\tnote[THX]{ADD: A well known result in %perfect graph theory is that
-%  the maximal cliques of  an interval graph $G$ can be linearly ordered
-%   such that for all $v \in V(G)$, cliques containing $v$ are
-%   consecutive in the ordering cite~{gh64}. This clearly means that a
-%   graph $G$ is an interval graph if and only if
-}% 
-\tnote[THX]{verify
-  from paper the statement of claim.}% 
-\tnote[TH]{maximal clique vertex
-  incidence matrix of%  $G$ has COP.  Also, maximal cliques of any
-%   chordal graph can be enumerated in polytime $O(m+n)$
-}
-\tnote[THX]{citation?!!}%  
-\tnote[TH]{cite~{fg65} uses these results to
-  give the first polynomial time algorithm for COT.}%
-\tnote[THX]{check. how do they use it?}%  
-\tnote[TH]{ A bipartite graph
-  is convex%  (on $R$) if and only if its half adjacency matrix has COP
-%   on rows.  The results in cite~{bl76} on COT are based on the result
-%   that interval graphs are AT-free chordal graphs.
-}%
-\tnote[THX]{the latter being Tucker's?}%  
-\tnote[THX]{(2) TBD survey
-  -- % see blue notes (in notebook) under Graph Isomorphism. namely
-%   citations in cite:aas93 (3) a brief on heirarchy of chordal graphs,
-%   path graphs, interval graphs, peo, clique tree etc. and results. (4)
-%   interval graphs are incomparability graphs. see
-%   Golumbic. [tag:classification] (5) finding min length hole in
-%   bipartite graph is polynomial time [Sec 3.3.2 in cite:d08phd] (6)
-%   theorem 2.2 in cite:d08phd - $G$ is union of v.d. caterpillars iff
-%   $M$ has COP in rows. $M$ is edge vertex incidence matrix $(*,2)$ of
-%   $G$
-}%
+In this section, a brief survey of the consecutive-ones problem and
+its optimization problems is presented.
 
@@
 case of the problem we address in this thesis, namely the tree path
-labeling problem (illustrated as the study group accommodation
-problem). The interval assignment problem and COP problem are equivalent
-problems. In this section we will see some of the results that
-exists in the literature today towards solving the COP problem and
+labeling problem (illustrated as the \illustrationproblem). The
+interval assignment problem and COP problem are equivalent
+problems. In this section we will see some of the results that exists
+in the literature today towards solving the COP problem and
 optimization problems surrounding it.
 
-\begin{figure}[t] %[htbp]
-  \centering
-
-  {\figtabsize
-    \begin{tabular}[h]{l|lcccl}
-      $M_1$: & $M_1'$: &&&& $M_2$:\\
-      &&&&&\\
-      \begin{tabular}[h]{llll}
-        $c_1$ & $c_2$ &$c_3$ &$c_4$\\
-        &&&\\
-        \un & 0   & \un & 0\\
-        0   & \un & 0   & \un \\
-        \un & 0   & 0   & \un
-      \end{tabular}
-      &
-      \begin{tabular}[h]{llll}
-        $c_3$ &$c_1$ &$c_4$& $c_2$\\
-        &&&\\
-        \un & \un & 0 & 0\\
-        0 & 0 & \un & \un \\
-        0 & \un & \un & 0
-      \end{tabular}
-      &&&&
-      \begin{tabular}[h]{llll}
-        $d_1$ & $d_2$ &$d_3$ &$d_4$\\
+% \begin{figure}[t] %[htbp]
+%   \centering
+
+%   {\figtabsize
+%     \begin{tabular}[h]{l|lcccl}
+%       $M_1$: & $M_1'$: &&&& $M_2$:\\
+%       &&&&&\\
+%       \begin{tabular}[h]{llll}
+%         $c_1$ & $c_2$ &$c_3$ &$c_4$\\
 %         &&&\\
-        &&&\\
-        \un & \un & 0 & 0\\
-        0 & \un & \un & 0 \\
-        0 & \un & 0 & \un 
-      \end{tabular}
-    \end{tabular}
-  }
-
-  \caption{\figtabsize Matrices with and without COP. $M_1$ has COP
-    because by permuting its columns, $c_1$-$c_4$, one can obtain
-    $M_1'$ where the {\un}s in each row are consecutive. $M_2$,
-    however, does not have COP since no permutation of its columns,
-    $d_1$-$d_4$, will arrange {\un}s in each row consecutively
-    \cite{d08phd}.}
-
-  \label{fig:cop-matrix}
-\end{figure}
+%         \un & 0   & \un & 0\\
+%         0   & \un & 0   & \un \\
+%         \un & 0   & 0   & \un
+%       \end{tabular}
+%       &
+%       \begin{tabular}[h]{llll}
+%         $c_3$ &$c_1$ &$c_4$& $c_2$\\
+%         &&&\\
+%         \un & \un & 0 & 0\\
+%         0 & 0 & \un & \un \\
+%         0 & \un & \un & 0
+%       \end{tabular}
+%       &&&&
+%       \begin{tabular}[h]{llll}
+%         $d_1$ & $d_2$ &$d_3$ &$d_4$\\
+%         % &&&\\
+%         &&&\\
+%         \un & \un & 0 & 0\\
+%         0 & \un & \un & 0 \\
+%         0 & \un & 0 & \un 
+%       \end{tabular}
+%     \end{tabular}
+%   }
+
+%   \caption[\figtabsize Matrices with and without COP.]{\figtabsize
+%     Matrices with and without COP. $M_1$ has COP because by permuting
+%     its columns, $c_1$-$c_4$, one can obtain $M_1'$ where the {\un}s
+%     in each row are consecutive. $M_2$, however, does not have COP
+%     since no permutation of its columns, $d_1$-$d_4$, will arrange
+%     {\un}s in each row consecutively \cite{d08phd}.}
+
+%   \label{fig:cop-matrix}
+% \end{figure}
 
@@
 rearranged so that the {\un}s in every column (row) are in consecutive
-rows (columns). Figure~\ref{fig:cop-matrix} shows examples of this
-property.  COP in binary matrices has several practical applications
+rows (columns). % Figure~\ref{fig:cop-matrix} shows examples of this
+% property. 
+COP in binary matrices has several practical applications
 in diverse fields including scheduling \cite{hl06}, information
@@
 submatrices came as a corollary to the discovery that convex bipartite
-graphs are AT-free in \cite{at72}\tnote[TH]{check}. The first linear
+graphs are AT-free in \cite{at72}\tnote{check}. The first linear
 time algorithm for COP testing (COT) was invented by \cite{bl76} using
@@
 columns) of the matrix. This is constructed using an elaborate linear
-time procedure and is also a test for planarity\tnote[TH]{check check
+time procedure and is also a test for planarity\tnote{check check
   check. both interval graph and planarity in this paper?}.  PQR trees
@@
 \cite{tm05} describes an improved algorithm to build PQR
-trees. \tnote[TH]{improv in terms of what?}\cite{wlh02} describes the
+trees. \tnote{improv in terms of what?}\cite{wlh02} describes the
 simpler algorithm for COT. Hsu also invented PC trees
-\cite{wlh01}\tnote[TH]{This result first appeared inproc ISAAC92}
+\cite{wlh01}\tnote{This result first appeared inproc ISAAC92}
 which is claimed to be much easier to implement. \cite{nsnrs09}
@@
 this column (row). This is interesting and relevant, especially to
-this thesis because it simplifies COT to a great degree. \tnote[TH]{it
+this thesis because it simplifies COT to a great degree. \tnote{it
   reduces the solution search space. fill in the blanks.}
@@
 algorithm is related to the weighted betweenness problem
-addressed\tnote[TH]{in what way??} in \cite{co98}.  \tnote[TH]{expand
+addressed\tnote{in what way??} in \cite{co98}.  \tnote{expand
   on the COP order graph creation and it having to be bipartite for M
   to have COP. and thus an odd cycle being an evidence of no COP.}
-\tnote[TH]{ where should this go?: (1) cite|jlm97 (application of PQ trees
-  in graphics). (2) helly's theorem citation
+\tnote{ where should this go?: (1) cite|jlm97 (application of PQ
+  trees in graphics). (2) helly's theorem citation
   19XXdgk-Hellystheorem-Danzer-Gruenbaum-Klee}
@@
 
-\tnote[TH]{ -- sect 4.1 in cite:d08phd has many results
+\tnote{ -- sect 4.1 in cite:d08phd has many results
   surveyed. hardness results, approx. results. results are usually for
@@
   entries of $M$ to make it attain COP. this is NP complete
-  cite:b75-phd}\tnote[TH]{(1) scite:lb62 showed that interval graphs
+  cite:b75-phd}\tnote{(1) scite:lb62 showed that interval graphs
   are AT-free.  describe AT (2) show the close relationship b/w COP
@@
 convex bipartite graphs which \cite{at72} proved to be
-AT-free\tnote[TH]{ check this up. give details. - doms'}. By
+AT-free\tnote{ check this up. give details. - doms'}. By
 definition, convex bipartite graph have half adjacency matrices that
@@
 discusses a couple of algorithms which are efficient if the number of
-{\un}s in a row is small. This is of significance in the case of sparse
-matrices where this number is much lesser than the number of
+{\un}s in a row is small. This is of significance in the case of
+sparse matrices where this number is much lesser than the number of
 columns. $(*,\Delta)${\em -matrices} are matrices with no restriction
@@
 COP. Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
-a submatrix $M'$ with all columns\tnote[TH]{check if b75 deals with
+a submatrix $M'$ with all columns\tnote{check if b75 deals with
   COP col or COP row. also is it any submatrix with k less than r rows
@@
 the area of parameterized algorithms for this
-problem\tnote[TH]{elaborate - what are the results?}.
+problem\tnote{elaborate - what are the results?}.
 
@@
 known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
-using a reduction to the longest path problem, \tnote[TH]{or is it a
+using a reduction to the longest path problem, \tnote{or is it a
   survey of another result?  check.} that finding a Tucker's forbidden
-submatrix of at least $k$ rows is NP complete. \tnote[TH]{how is this
-  different from booth's 75 result??}  \tnote[TH]{where should this
+submatrix of at least $k$ rows is NP complete. \tnote{how is this
+  different from booth's 75 result??}  \tnote{where should this
   go? cite|tz04 (approx submatrix with COP sparse matrices)}
@@
 {\un}s. This is proved to be NP hard using a reduction from the
-Hamiltonian path problem.\tnote[TH]{Theorem 2.1 in jkckv}
-\tnote[TH]{(1) A connection of COP problem to the travelling salesman
+Hamiltonian path problem.\tnote{Theorem 2.1 in jkckv}
+\tnote{(1) A connection of COP problem to the travelling salesman
   problem is also introduced. what does this mean? -- COP can be used
@@
   blocks of ones is proven to be NP complete by
-  cite:k77}\tnote[TH]{are these two the same?} \tnote[TH]{what is the
-  reduction?} \tnote[TH]{other problems similar to COP -- cite:ckl96
+  cite:k77}\tnote{are these two the same?} \tnote{what is the
+  reduction?} \tnote{other problems similar to COP -- cite:ckl96
   (ILP, circ ones, one drop) -- cite:th98 (generalization of COP -
@@
 system $\cF$, a tree $T$, and an assignment of paths from $T$ to the
-sets, there is a feasible\footnote{The notion of {\em feasibility} is
+sets, there is a feasible\endnote{The notion of {\em feasibility} is
   formally defined in Section~\ref{ch:prelims}.}  bijection between
@@
 filtering algorithm is also an efficient algorithm to test if a tree
-path labeling\footnote{The terms {\em tree path labeling} and {\em
+path labeling\endnote{The terms {\em tree path labeling} and {\em
     tree path assignment} are, in informal language,
@@
 cardinality preservation is necessary and sufficient for an interval
-assignment to be feasible for a given hypergraph\footnote{A {\em
+assignment to be feasible for a given hypergraph\endnote{A {\em
     hypergraph} is an alternate representation of a set system and
-  will be used in this thesis. See Section~\ref{ch:prelims} for the
+  will be used in this thesis.}~\endnote{See Section~\ref{ch:prelims} for the
   formal definition.} and thus is a characterization for COP
@@
 and find that trio-wise intersection cardinality preservation makes a
-tree path labeling \footnote{A {\em tree path labeling} $\cl$ is a
+tree path labeling\endnote{A {\em tree path labeling} $\cl$ is a
   bijection of paths from the target tree $T$ to the hyperedges in
-  given hypergraph $\cF$.  See Section~\ref{ch:prelims} for the formal
-  definition.}  (TPL) feasible, which is a generalization of the COP
+  given hypergraph $\cF$.}~\endnotemark[4]  (TPL) feasible, which is a generalization of the COP
 problem. This problem is defined as follows.
@@
 captures the trio-wise cardinality property described
-earlier\footnote{See Section~\ref{sec:feasible} for the definition of
+earlier\endnote{See Section~\ref{sec:feasible} for the definition of
   ICPPL.}. This characterization can be checked in polynomial time.  A
@@
 We present a polynomial time algorithm for {\CFTPL} when the target
-tree $T$ belongs to a special class of trees called {\em \kstar s} and
+tree $T$ belongs to a special class of trees called {\em \kstars} and
 when the hyperedges in the hypergraph $\cF$ have at most $k+2$
-vertices. A couple of examples of {\kstar s} are given in
+vertices. A couple of examples of {\kstars} are given in
 Figure~\ref{fig:ksubstar}.
@@
   \end{tabular}
-  \caption{\figtabsize Examples of {\kstar s}. (a) $k = 0$ (b) $k = 2$
+  \caption{\figtabsize Examples of {\kstars}. (a) $k = 0$ (b) $k = 2$
   }
@@
   Isomorphism} which is polynomial time solvable in interval graphs
-while being hard on path graphs\cite{kklv10}. {\kstar s} are a class
+while being hard on path graphs\cite{kklv10}. {\kstars} are a class
 of trees which are in many ways very close to intervals or paths. Each
-ray\footnote{The path from a leaf to the root, the vertex with highest
-  degree, is called a {\em ray} of the \kstar. See
-  Section~\ref{ch:prelims}.} are independent except for the
-root\footnote{The vertex with maximum degree in a {\kstar} is called
-  {\em root}. See Section~\ref{ch:prelims}.} and hence can be
+ray\endnote{The path from a leaf to the root, the vertex with highest
+  degree, is called a {\em ray} of the \kstar.}~\endnotemark[4] are independent except for the
+root\endnote{The vertex with maximum degree in a {\kstar} is called
+  {\em root}.}~\endnotemark[4] and hence can be
 considered as an independent interval till the root. Our algorithm
@@
 represent a set system (see Section~\ref{sec:motive}) and that the
-{\em overlap}\footnote{}\tnote{Give informal definition of ``overlap''
-  in footnote and link to sec:prelims. ADD the definition in prelim.}
+{\em overlap}\tnote{Give informal definition of ``overlap''
+  in footnote/endnote and link to sec:prelims. ADD the definition in prelim.}
 relation between the sets involved is an obvious equivalence relation,
@@
   definition of ``total order'' in footnote?}. We extend this to TPL
-and find that when $\cF$ is a path hypergraph\footnote{If there exists
+and find that when $\cF$ is a path hypergraph\endnote{If there exists
   an FTPL for a hypergraph $\cF$, it is called a path hypergraph.},
@@
 
-\tnote[TH]{The connection of TPL to graph isomorphism will be made
+\tnote{The connection of TPL to graph isomorphism will be made
   later in the document}
 
-\chapter{Consecutive-ones Property - a Survey of Important Results}
-\label{ch:copsurvey}
-\tnote{TBD: have a few lines about organization of chapter}
-
-\tnote{ Survey chapter: is the full fledged expansion of the survey in
-  introduction with details, observations, theorems etc.}
-
-\section{Matrices with COP}
-\label{sec:surveycoptest}
-\tnote{Expand on ref:sec:copmatrices}
-
-\section{Optimization problems in COP}
-\label{sec:surveycopopt}
-\tnote{Expand on ref:sec:optcop}
-
-\section{********* COP in Relational Database Model}
-\label{sec:surveyrdbm}
-\tnote{Expand on sec:apprdbm}
-\tnote{(set systems theme)}
-
-\subsection{********* COP in Graph Isomorphism}
-\label{sec:surveygraphiso}
-\tnote{Expand on sec:appgraphiso}
-\tnote{(canonization theme)}
-
-\subsection{********* Certifying Algorithms}
-\label{sec:surveycertalgo}
-\tnote{Expand on sec:appcertalgo}
-\tnote{ (certification McC04 theme)}
-
-\chapter{Tree Path Labeling of Path Hypergraphs - the New Results}
-\label{ch:myresearch}
-
-\tnote{UPDATE TO FINAL VERSION. BELOW IS OLDER VERSION WITH INCOMPLETE
-  SEC 4.}  This chapter documents all the new results obtained by us in the
-area of tree path labeling of path hypergraphs which is the parent
-problem addressed in this thesis. In Section~\ref{sec:motive} we see
-that consecutive-ones property and its equivalent problem of interval
-labeling of a hypergraph is a special case of the general problem of tree
-path labeling of path hypergraphs.
-
-\tnote{UPDATE the following para to current skeleton.}The necessary
-preliminaries with definitions etc. are presented in
-Section~\ref{ch:prelims}. Section~\ref{sec:feasible} documents the
-characterization of a feasible path labeling for any path labeling
-from any target trees. Section~\ref{sec:spltargettree} describes the
-special case where the target tree is of a particular family of trees
-and specifically, Section~\ref{sec:ksubdivstar} describes a polynomial
-time algorithm to find the tree path labeling of a given set system
-from a given $k$-subdivided tree. Section~\ref{sec:norestraint}
-describes the general situation where the target tree has no
-restrictions and the algorithm to find a TPL, if any, in this case.
-
-\section{Introduction}
-
-The problem of consecutive-ones property testing can be easily seen as
-a simple constraint satisfaction problem involving a hypergraph or a
-system of sets from a universe. Every column (row) of the binary
-matrix can be converted into a set of non-negative integers which are
-the indices of the rows (columns) with {\un}s in that column
-(row). When observed in this context, if the matrix has the COP on
-columns (rows), a reordering of its rows (columns) will result in sets
-that have only consecutive integers. In other words, the sets after
-applying the COP row (column) permutation are intervals.  In this
-form, one can see that this is indeed the problem of finding interval
-assignments to the given set system \cite{nsnrs09} with a single
-permutation of the universe (set of row or column indices for COP of
-columns or rows, respectively) which permutes each set to its
-interval. The result in \cite{nsnrs09} characterizes interval
-assignments to the sets which can be obtained from a single
-permutation of the rows (columns). They show that for each set, the
-cardinality of the interval assigned to it must be same as the
-cardinality of the set, and the intersection cardinality of any two
-sets must be same as the intersection cardinality of the corresponding
-intervals. This is a necessary and sufficient
-condition. \tnote{(ICPIA) put reference to survey chapter}.  Finally,
-the idea of decomposing a given binary matrix into prime matrices to
-check for COP is adopted from cite|wlh02| to test if an ICPIA exists
-for a given set system.\tnote{REPEATITION from intro chapter. but idea
-  recall is necessary. so keep it but reword it to avoid monotone.}
-
-A natural generalization of the interval assignment problem is
-feasible tree path labeling problem of a set system. The problem is
-defined as follows -- given a set system $\cF$ from a universe $U$ and
-a tree $T$, does there exist a bijection from $U$ to the vertices of
-$T$ such that each set in the system maps to a path in $T$.  We refer
-to this as the {\em tree path labeling problem} for an input set
-system, target tree pair -- $(\cF,T)$. \tnote{USE THE FORMAL PROBLEM DEF BLOCK}
-%\begin{tabular}[t]{c}
-{\small
-\begin{minipage}[h]{5in}
- % \begin{singlespace}
- \vspace{2mm}
- {\large \FTPL}\\
- \begin{tabular}[t]{l|l}
- \hline\\
-    {\tt Input} & 
-    \begin{minipage}[t]{\probdefwidth}
-      A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
-      paths $\cP$ from $T$ and a
-      bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
-    \end{minipage}\\
-    {\tt Question} &
-    \begin{minipage}[t]{\probdefwidth}
-      Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
-      such that $\phi$ when applied on any hyperedge in $\cF$ will
-      give
-      the path mapped to it by the given tree path labeling $\cl$.\\
-      { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every hyperedge
-        $S \in \cF$.}
-    \end{minipage}\\
-  \end{tabular}
-  % \end{singlespace}
-\end{minipage}\\
-}
-%\end{tabular}
-
-As a special case if the tree
-$T$ is a path, the problem becomes the interval assignment problem. \tnote{USE THE FORMAL PROBLEM DEF BLOCK}
-We focus on the question of generalizing the notion of an ICPIA
-\cite{nsnrs09} to characterize feasible path assignments.  We show
-that for a given set system $\cF$, a tree $T$, and an assignment of
-paths from $T$ to the sets, there is a feasible bijection between $U$
-and $V(T)$ if and only if all intersection cardinalities among any
-three sets (not necessarily distinct) is same as the intersection
-cardinality of the paths assigned to them and the input runs a
-filtering algorithm (described in this paper) successfully.  This
-characterization is proved constructively and it gives a natural data
-structure that stores all the relevant feasible bijections between $U$
-and $V(T)$.  Further, the filtering algorithm is also an efficient
-algorithm to test if a tree path labeling to the set system is
-feasible.  This generalizes the result in \cite{nsnrs09}.
-
-In the later part of this paper, we focus on a new special case of the
-tree path labeling problem. Here the set system is such that every set
-is at most $k+2$ in size and for every pair of sets in it there exists
-a sequence of sets between them with consecutive sets in this sequence
-having a strict intersection -- i.e., non-empty intersection with
-neither being contained in the other. Moreover, the given tree is a
-{\em $k$-subdivided star}. We demonstrate a polynomial time algorithm
-to find a feasible path labeling in this case.
-
-\section{Preliminaries to new results}
-\label{ch:prelims}
-
-\tnote{TBD: have a few lines about organization of chapter}This
-section states definitions and basic facts necessary in the scope of
-this document.
-
-\tnote{Move some of this to general prelim in chapter 1}
-
-The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
-  system} of a universe $U$ with $|U| = n$.  The {\em support} of a
-set system $\F$ denoted by $supp(\cF)$ is the union of all the sets in
-$\F$; $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this
-paper, a set system is required to ``cover'' the universe; $
-supp(\cF) = U$.
-
-The graph $T$ represents a {\em target tree} with same
-number of vertices as elements in $U$; $|V(T)| = n$.  A {\em path
-  system}\, $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P
-\mid P \subseteq V, \text{ } T[P] \text{ is a path} \}$.
-
-% 
-% A set system $\cF$ can be alternatively represented by a {\em
-%   hypergraph} $\H_\cF$ whose vertex set is $supp(\cF)$ and
-% hyperedges are the sets in $\cF$. This is a known representation for
-% interval systems in literature \cite{bls99,kklv10}.  We extend this
-% definition here to path systems.
-
-A set system $\cF$ can be alternatively represented by a {\em
-  hypergraph}\, $\cF_H$ whose vertex set is $supp(\cF)$ and hyperedges
-are the sets in $\cF$. This is a known representation for interval
-systems in literature \cite{bls99,kklv10}.  We extend this definition
-here to path systems. Due to the equivalence of set system and
-hypergraph in the scope of this paper, we drop the subscript $_H$ in
-the notation and refer to both the structures by $\cF$.
-
-Two hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to
-each other, denoted by $\cF' \cong \cF''$, iff there exists a
-bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for all
-sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ iff $B$ is
-a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
-\cite{kklv10}. This is called {\em hypergraph
-  isomorphism}. \tnote{also extend $\phi$ to hyperedges -- see if
-  required }
-
-The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph $\cF$ is a
-graph such that its vertex set has a bijection to $\cF$ and there
-exists an edge between two vertices iff their corresponding hyperedges
-have a non-empty intersection \cite{mcg04}.
-
-If the intersection graphs of two hypergraphs are isomorphic,
-$\bI(\cF) \cong \bI(\cP)$ where $\cP$ is also a path system, then the
-bijection $\cl: \cF \rightarrow \cP$ due to this isomorphism is called
-a {\em path labeling} of the hypergraph $\cF$. To illustrate further,
-let $\cg: V(\cF) \rightarrow V(\cP)$ be the above mentioned
-isomorphism where $V(\cF)$ and $V(\cP)$ are the vertex sets that
-represent the hyperedges for each hypergraph respectively, $V(\cF) =
-\{ v_S \mid S \in \cF\}$ and $V(\cP) = \{ v_P \mid P \in \cP\}$. Then
-the path labeling $\cl$ is defined as follows: $\cl(S_1) = P_1$ iff
-$\cg (v_{S_1}) = v_{P_1}$. The path system $\cP$ may be alternatively
-denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In most scenarios in
-this paper, what is given are the pair $(\cF, \cl)$ and the target
-tree $T$; hence this notation will be used more often.
-
-If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$
-is called a {\em path hypergraph} and $\cP$ is called {\em path
-  representation} of $\cF$. If this isomorphism is $\phi: supp(\cF)
-\rightarrow V(T)$, then it is clear that there is an {\em induced path
-labeling} $\cl_\phi: \cF \rightarrow \cP$ to the set system;
-$\cl_\phi(S) = \set{y \mid y = \phi(x), x \in S}$ for all $S \in \cF$. Recall that
-$supp(\cP) = V(T)$.
-
-A path labeling $(\cF, \cl)$ is defined to be {\em
-  feasible} if
-$\cF \cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
-\rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
-\rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.
-
-An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is
-a graph such that its vertex set has a bijection to $\cF$ and there
-exists an edge between two of its vertices iff their corresponding
-hyperedges overlap. Two hyperedges $S$ and $S'$ are said to {\em
-  overlap}, denoted by $S \overlap S'$, if they have a non-empty
-intersection and neither is contained in the other; $S \overlap S'
-\text{ iff } S \cap S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq
-S$. Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$ and not
-necessarily connected. Each connected component of $\bO(\cF)$ is
-called an {\em overlap component}.
-
-A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
-\overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
-\cite{kklv10}. Additionally, if $S$ is such that it is contained in no
-other hyperedge in $\cF$, i.e., it is inclusion maximal then it is called
-{\em super-marginal}.
-
-A {\em star} graph is a complete bipartite graph $K_{1,p}$
-which is clearly a tree and $p$ is the number of leaves. The vertex
-with maximum degree is called the {\em center} of the star and the
-edges are called {\em rays} of the star. A {\em $k$-subdivided star}
-is a star with all its rays subdivided exactly $k$ times. The
-definition of a {\em ray of a $k$-subdivided star} is extended to the path
-from the center to a leaf. It is clear that all rays are of length $k+2$.
-
-\section{Characterization of Feasible Tree Path  Labeling} 
-\label{sec:feasible}
-
-In this section we give an algorithmic characterization of a
-feasibility of tree path labeling.  Consider a path labeling $(\cF,
-\cl)$ on the given tree $T$. We call $(\cF, \cl)$ an {\em Intersection
-  Cardinality Preserving Path Labeling (ICPPL)} if it has the
-following properties.
-
-\begin{enumerate}[{(\icpplpr\ }i) \ \ \ ]
-\item \label{pr:i} $|S| = |\cl(S)|$ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $S \in \cF$
-  % \vspace{\topshrink}
-\item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ \ \ \ \
-  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all distinct $S_1, S_2 \in \cF$
-  % \vspace{\topshrink}
-\item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
-  \cap \cl(S_3)|$ \ \ \ for all distinct $S_1, S_2, S_3 \in \cF$
-\end{enumerate}
-
-The following lemma is useful in subsequent arguments. 
-\begin{lemma}
-  \label{lem:setminuscard}
-  If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
-  (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
-  \cl(S_3))|$.
-\end{lemma}
-\begin{proof}%[Proof of Lemma~\ref{lem:setminuscard}]
-  Let $P_i = \cl(S_i)$, for all $1 \le i \le  3$.
-  $|S_1 \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| =
-  |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to properties (ii)
-  and (iii) of ICPPL, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
-  \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
-  = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. \qed
-\end{proof}
-
-In the remaining part of this section we show that $(\cF,
-\cl)$ is feasible if and only if it is an ICPPL and
-Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
-function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two levels of
-filtering of $(\cF, \cl)$ to make it simpler while retaining the set
-of isomorphisms, if any, between $\cF$ and $\cF^\cl$.
-% One direction of this claim isclear: that if a path
-% labeling is feasible, then all intersection cardinalities are
-% preserved, i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
-% \annote{has no premature exit condition hence any input will go
-%   through it}{Prove that the filtered sets has ICPPL iff input PL
-%   has ICPPL?}. Algorithm~\ref{leafasgn} has an exit condition at
-% line~\ref{xempty}. It can be easily verified that $X$ cannot be
-% empty if $\cl$ is a feasible path labeling. The reason is that a
-% feasible path labeling has an associated bijection between
-% $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that
-% the sets map to paths, ``preserving'' the path labeling.  The rest
-% of the section is devoted to constructively proving that it is
-% sufficient for a path labeling to be an ICPPL and pass the two
-% filtering algorithms.  To describe in brief, the constructive
-% approaches refine an ICPPL iteratively, such that at the end of each
-% iteration we have a ``filtered'' path labeling, and finally we have
-% a path labeling that defines a family of bijections from $supp(\cF)$
-% to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.
-First, we present Algorithm~\ref{perms} or {\tt filter\_1}, and prove
-its correctness.  This algorithm refines the path labeling by
-processing pairs of paths in $\cF^\cl$ that share a leaf until no two
-paths in the new path labeling share any leaf.
-
-\begin{algorithm}[h]
-  \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
-  \label{perms}
-  \begin{algorithmic}[\lndisplay]
-    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$\\
-    \STATE $j \assign 1$\\
-    \WHILE {there is $S_1, S_2 \in \cF_{j-1}$ such that
-      $\cl_{j-1}(S_1)$ and $\cl_{j-1}(S_2)$ have a common leaf in
-      $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
-    \{S_1, S_2\})
-    \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1 \}$ \label{setbreak} 
-    \COMMENT {Remove $S_1$, $S_2$ and add the ``filtered'' sets}
-    \STATE {\bf for} every $S \in \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
-    S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\
-
-    \STATE $\cl_j(S_1 \cap S_2) \assign \cl_{j-1}(S_1) \cap
-    \cl_{j-1}(S_2)$
-    \COMMENT {Carry forward the path labeling for all existing sets other than
-      $S_1$, $S_2$}
-    \STATE $\cl_j(S_1 \setminus S_2) \assign \cl_{j-1}(S_1) \setminus
-    \cl_{j-1}(S_2)$ 
-    \COMMENT {Define path labeling for new sets}
-    \STATE $\cl_j(S_2 \setminus S_1) \assign \cl_{j-1}(S_2) \setminus
-    \cl_{j-1}(S_1)$
-
-    \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of ICPPL}
-    \label{ln:3waycheck}
-    \STATE {\bf exit} \label{ln:exit1} \\
-    \ENDIF
-
-    \STATE $j \assign j+1$\\
-    \ENDWHILE
-    \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
-    \RETURN $(\cF', \cl')$
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma} 
- \label{lem:feasible} 
- In Algorithm~\ref{perms}, if input $(\cF, \cl)$ is a feasible path
- assignment then at the end of $j$th iteration of the {\bf while}
- loop, $j \ge 0$, $(\cF_j, \cl_j)$ is a feasible path assignment.
-\end{lemma}
-\begin{proof}%[Proof of Lemma~\ref{lem:feasible}]
-  We will prove this by mathematical induction on the number of
-  iterations. The base case $(\cF_0, \cl_0)$ is feasible since it is
-  the input itself which is given to be feasible. Assume the lemma is
-  true till $j-1$th iteration. i.e. every hypergraph isomorphism
-  $\phi: supp\left(\cF_{j-1}\right) \rightarrow V\left(T \right)$ that
-  defines $(\cF, \cl)$'s feasibility, is such that the induced path
-  labeling on $\cF_{j-1}$, $\cl_{\phi[{\cF_{j-1}}]}$ is equal to
-  $\cl_{j-1}$. We will prove that $\phi$ is also the bijection that
-  makes $(\cF_j, \cl_j)$ feasible. Note that $supp(\cF_{j-1}) =
-  supp(\cF_{j})$ since the new sets in $\cF_j$ are created from basic
-  set operations to the sets in $\cF_{j-1}$. For the same reason and
-  $\phi$ being a bijection, it is clear that when applying the $\phi$
-  induced path labeling on $\cF_j$, $ \cl_{\phi[{\cF_{j}}]}(S_1
-  \setminus S_2) = \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
-  \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Now observe that $ \cl_j(S_1
-  \setminus S_2) = \cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2) =
-  \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
-  \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Thus the induced path labeling
-  $\cl_{\phi[{\cF_{j}}]} = \cl_{j}$. Therefore lemma is proven.  \qed
-\end{proof}
-
-\begin{lemma}
-  \label{lem:invar1} In Algorithm~\ref{perms}, at the end of $j$th
-  iteration, $j \ge 0$, of the {\bf while} loop, the following
-  invariants are maintained.
-  \begin{enumerate}[I {\ }] %\vspace{\topshrink}
-  \item $\cl_j(R)$ is a path in $T$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ for all $R \in \cF_j$%\vspace{\topshrink}
-  \item $|R| = |\cl_j(R)|$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $R \in
-    \cF_j$%\vspace{\topshrink}
-  \item $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ for all $R, R' \in \cF_j$%\vspace{\topshrink}
-  \item $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
-    \cl_j(R'')|$, \ \ \ for all $R, R', R'' \in \cF_j$
-  \end{enumerate}
-\end{lemma}
-
-\begin{proof}
- Proof is by induction on the number of iterations, $j$. In this
-  proof, the term ``new sets'' will refer to the sets added to $\cF_j$
-  in $j$th iteration in line~\ref{setbreak} of Algorithm~\ref{perms},
-  $S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1$ and its
-  images in $\cl_j$ where $\cl_{j-1}(S_1)$
-  and $\cl_{j-1}(S_2)$ intersect and share a leaf.\\
-  The invariants are true in the base case $(\cF_0, \cl_0)$, since it
-  is the input ICPPL.  Assume the lemma is true till the $j-1$th
-  iteration. Let us consider the possible cases for each of the above invariants for
-  the $j$th iteration.
-
- \begin{enumerate}[\xbullet]
-  \item {\em Invariant} I/II
-    \begin{enumerate}[{I/II}a $|$] % \textbullet 
-    \item {\em $R$ is not a new set.} It is in $\cF_{j-1}$. Thus
-      trivially true by induction hypothesis.
-    \item {\em $R$ is a new set.} If $R$ is in $\cF_{j}$ and not in
-      $\cF_{j-1}$, then it must be one of the new sets added in
-      $\cF_j$. In this case, it is clear that for each new set, the
-      image under $\cl_j$ is a path since by definition the chosen
-      sets $S_1$, $S_2$ are from $\cF_{j-1}$ and due to the while loop
-      condition, $\cl_{j-1}(S_1)$, $\cl_{j-1}(S_2)$ have a
-      common leaf. Thus invariant I is proven.\\
-      Moreover, due to induction hypothesis of invariant III and the
-      definition of $l_j$ in terms of $l_{j-1}$, invariant II is
-      indeed true in the $j$th iteration for any of the new sets.  If
-      $R = S_1 \cap S_2$, $|R| = |S_1 \cap S_2| = |\cl_{j-1}(S_1) \cap
-      \cl_{j-1}(S_2)| = |\cl_j(S_1 \cap S_2)| = |\cl_j(R)|$.
-      If $R = S_1 \setminus S_2$, $|R| = |S_1 \setminus S_2| = |S_1| -
-      |S_1 \cap S_2| = |\cl_{j-1}(S_1)| - |\cl_{j-1}(S_1) \cap
-      \cl_{j-1}(S_2)| = |\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)| =
-      |\cl_j(S_1 \setminus S_2)|
-      = |\cl_j(R)|$. Similarly if $R = S_2 \setminus S_1$.\\
-    \end{enumerate}
-  \item {\em Invariant} III
-    \begin{enumerate}[{III}a $|$]
-    \item {\em $R$ and $R'$ are not new sets.} It is in
-      $\cF_{j-1}$. Thus trivially true by induction hypothesis.
-    \item {\em Only one, say $R$, is a new set.} Due to invariant IV
-      induction hypothesis, Lemma~\ref{lem:setminuscard} and
-      definition of $\cl_j$, it follows that invariant III is true no
-      matter which of the new sets $R$ is equal to. If $R = S_1 \cap
-      S_2$, $|R \cap R'| = |S_1 \cap S_2 \cap R'| = |\cl_{j-1}(S_1)
-      \cap \cl_{j-1}(S_2) \cap \cl_{j-1}(R')| = |\cl_j(S_1 \cap S_2)
-      \cap \cl_j(R')| = |\cl_j(R) \cap \cl_j(R')|$.  If $R = S_1
-      \setminus S_2$, $|R \cap R'| = |(S_1 \setminus S_2) \cap R'| =
-      |(\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)) \cap \cl_{j-1}(R')| =
-      |\cl_{j}(S_1 \cap S_2) \cap \cl_{j}(R')| = |\cl_{j}(R) \cap
-      \cl_{j}(R')|$. Similarly, if $R = S_2 \setminus
-      S_1$. Note $R'$ is not a new set.\\
-
-    \item {\em $R$ and $R'$ are new sets.} By definition, the new
-      sets and their path images in path label $\cl_j$ are disjoint so
-      $|R \cap R'| = |\cl_j(R) \cap \cl_j(R)| = 0$. Thus case proven.
-    \end{enumerate}
-  \item {\em Invariant} IV
-
-    Due to the condition in line~\ref{ln:3waycheck}, this invariant is
-    ensured at the end of every iteration.
-%     \begin{enumerate} [{Case 3.}1:]
-%     \item {\em $R$, $R'$ and $R''$ are not new sets.} Trivially
-%       true by induction hypothesis.
-%     \item {\em Only one, say $R$, is a new set.}
-%       If $R = S_1 \cap S_2$,  from Lemma~\ref{lem:fourpaths} and
-%       invariant III hypothesis,  this case is proven. Similarly if $R$
-%       is any of the other new  sets, the case is proven by also using
-%       Lemma ~\ref{lem:setminuscard}.
-%     \item {\em At least two of $R, R', R''$ are new sets.}
-%       The new sets are disjoint hence this case is vacuously true.
-%     \end{enumerate}
-  \end{enumerate} \qed
-%\vspace{-6mm}
-
-\end{proof}
-
-\begin{lemma}
-  \label{lem:noexit1}
-  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{perms} is
-  feasible, then the set of hypergraph isomorphism functions that
-  defines $(\cF, \cl)$'s feasibility is the same as the set that
-  defines $(\cF_j, \cl_j)$'s feasibility, if any.  Secondly, for any
-  iteration $j > 0$ of the {\em \bf while} loop, the {\em \bf exit}
-  statement in line~\ref{ln:exit1} will not execute.
-\end{lemma}
-\begin{proof}
-  Since $(\cF,\cl)$ is feasible, by Lemma~\ref{lem:feasible}
-  $(\cF_j,\cl_j)$ for every iteration $j > 0$ is feasible.  % Therefore,
-%   every hypergraph isomorphism $\phi: supp(\cF) \rightarrow V(T)$ that
-%   induces $\cl$ on $\cF$ also induces $\cl_{j-1}$ and $\cl_{j}$ on
-%   $\cF_{j-1}$ and $\cF_{j}$ respectively, i.e., $\cl_{\phi[\cF_{j-1}]}
-%   = \cl_{j-1}$ and $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
-%   that for all $x \in supp(\cF)$, for all $v \in V(T)$ the following
-%   hold true.
-  Also, every hypergraph isomorphism $\phi: supp(\cF) \rightarrow
-  V(T)$ that induces $\cl$ on $\cF$ also induces $\cl_{j}$ on
-  $\cF_{j}$, i.e., $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
-  that for all $x \in supp(\cF)$, for all $v \in V(T)$, if $(x,v) \in
-  \phi$ then $v \in \cl_{j}(S)$ for all $S \in \cF_{j}$ such that $x
-  \in S$.
-% the following
-%   hold true.
-%   \begin{enumerate}[i. ]
-%   \item If $(x,v) \in \phi$ then $v \in \cl_{j-1}(S)$ for all $S \in
-%     \cF_{j-1}$ such that $x \in S$.
-%   \item If $(x,v) \in \phi$ then $v \in \cl_{j}(S)$ for all $S \in
-%     \cF_{j}$ such that $x \in S$
-%   \end{enumerate}
-  In other words, filter 1 outputs a filtered path labeling that
-  ``preserves''
-  hypergraph isomorphisms of the original path labeling.\\
-  Secondly, line~\ref{ln:exit1} will execute iff the exit condition in
-  line~\ref{ln:3waycheck}, i.e. failure of three way intersection
-  preservation, becomes true in any iteration of the {\em \bf while}
-  loop.  Due to Lemma~\ref{lem:invar1} Invariant IV, the exit
-  condition does not occur if the input is a feasible ICPPL.\qed
-
-%   such that $\phi(x) = v$ where $v$ is the leaf considered in the
-%   first iterations of while. Clearly, $\phi$ is a renaming of
-%   vertices in hypergraph $\cF$ to those in hypergraph $\cF^\cl$. Thus
-%   the following facts can be observed in every iteration of the loop.
-
-%   \begin{enumerate}[\hspace{2mm}i. ] \vspace{\topshrink}
-%   \item all intersection cardinalities are preserved in this path
-%     labeling \vspace{\topshrink}
-%   \item element $x$ is exclusive in a hyperedge in $\cF$ since $v$ is
-%     exclusive in a hyperedge in $\cF^\cl$.
-%   \end{enumerate}
-
-%   Thus the exit condition is never rendered true after $x$ and $v$ are
-%   removed from their respective hyperedges. \qed
-
-% \noindent
-% This proof uses mathematical induction on the number
-%   of iterations $j$, $j \ge 0$, of the loop that executed
-%   without exiting. The base case, $j = 0$ is obviously true since the
-%   input is an ICPPL and the exit condition cannot hold true due to
-%   ICPPL property (iii).  Assume the algorithm executes till the end
-%   of $j-1$th iteration without exiting at line
-%  ~\ref{ln:3waycheck}. Consider the $j$th iteration. From Lemma
-%  ~\ref{lem:feasible} we know that $(\cF_j, \cl_j)$ and $(\cF_{j-1},
-%   \cl_{j-1})$ are feasible\remove[AS]{and from the proof in lemma
-%     lem:invar1 we know that $(\cF_{j-1}, \cl_{j-1})$ satisfies all the
-%     invariants defined in the lemma}.  Thus there exists a bijection
-%   $\phi: supp(\cF) \rightarrow V(T)$ such that the induced path
-%   % labeling on $\cF_{j-1}$ $\cl_{\phi[\cF_{j-1}]} = \cl_{j-1}$.
-%   labeling on $\cF_{j}$, $\cl_{\phi[\cF_{j}]}$ and on $\cF_{j-1}$,
-%   $\cl_{\phi[\cF_{j-1}]}$ are equal to $\cl_{j}$ and $\cl_{j-1}$
-%   respectively.  We need to prove that for any $R, R', R'' \in
-%   \cF_{j}$, $|R \cap R' \cap R''| = |\cl_j(R) \cap \cl_j(R') \cap
-%   \cl_j(R'')|$.
-%   The following are the possible cases that could arise. From argument
-%   above, $|\cl_j(R) \cap \cl_j(R') \cap \cl_j(R'')| =
-%   |\cl_{\phi[\cF_{j}]}(R) \cap \cl_{\phi[\cF_{j}]} (R') \cap
-%   \cl_{\phi[\cF_{j}]} (R'')|$
-
-%   \begin{enumerate}[a $|$]
-%   \item {\em None of the sets are new. $R, R', R'' \in \cF_{j-1}$.}
-%     We know $(\cF_{j-1}, \cl_{j-1})$ is feasible. Thus $|R \cap R'
-%     \cap R''| = |\cl_{j-1}(R) \cap \cl_{j-1}(R') \cap \cl_{j-1}(R'')|
-%     = |\cl_{j}(R) \cap \cl_{j}(R') \cap \cl_{j}(R'')|$.
-%   \item {\em Only one, say $R$, is a new set.}  Let $R = S_1 \cap S_2$
-%     ($S_1, S_2$ are defined in the proof of lemma
-%    ~\ref{lem:invar1}). Now we have $|R \cap R' \cap R''| = |S_1 \cap
-%     S_2 \cap R' \cap R''| = |\cl_{j-1}(S_1) \cap \cl_{j-1}(S_2) \cap
-%     \cl_{j-1}(R') \cap \cl_{j-1}(R'')| = |\cl_{j}(R) \cap \cl_{j}(R')
-%     \cap \cl_{j}(R'')|$. Thus proven. If $R$ is any of the other new
-%     sets, the same claim can be verified using lemma
-%    ~\ref{lem:setminuscard}.
-%     % \item []{\bf Case 3:}
-%   \item {\em At least two of $R, R', R''$ are new sets.}  The new sets
-%     are disjoint hence this case is vacuously true.
-%   \end{enumerate}
-%   \qed
-%   \tnote[E2]{remove the induction proof. just text saying x and v are
-%     exclusive in these sets therefore the intersection cardinalities
-%     don't change thus all invariants are still true} 
-\end{proof}
-
-As a result of Algorithm~\ref{perms} each leaf $v$ in $T$
-is such that there is exactly one set in $\cF$ with $v$ as a vertex in
-the path assigned to it.  In Algorithm~\ref{leafasgn} we identify
-elements in $supp(\cF)$ whose images are leaves in a hypergraph
-isomorphism if one exists.  Let $S \in \cF$ be such that $\cl(S)$ is a
-path with leaf and $v \in V(T)$ is the unique leaf incident on it.  We
-define a new path labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) =
-\set{v}$ where $x$ an arbitrary element from $S \setminus \bigcup_{\hS
-  \ne S} \hS$. In other words, $x$ is an element present in no other
-set in $\cF$ except $S$. This is intuitive since $v$ is present in no
-other path image under $\cl$ other than $\cl(S)$.  The element $x$ and
-leaf $v$ are then removed from the set $S$ and path $\cl(S)$
-respectively. After doing this for all leaves in $T$, all path images
-in the new path labeling $\cl_{new}$ except leaf labels (a path that
-has only a leaf is called the {\em leaf label} for the corresponding
-single element hyperedge or set) are paths from a new pruned tree $T_0
-= T \setminus \{v \mid v \text{ is a leaf in }
-T\}$. Algorithm~\ref{leafasgn} is now presented with details.
-
-\begin{algorithm}[h]
-  \caption{Leaf labeling from an ICPPL {\tt filter\_2($\cF, \cl, T$)}}
-  \label{leafasgn}
-  \begin{algorithmic}[\lndisplay]
-    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$
-    \COMMENT {Path images are such that no two path images share a
-      leaf.}
-    \STATE $j \assign 1$\\
-    \WHILE {there is a leaf $v$ in $T$ and a unique $S_1 \in
-      \cF_{j-1}$ such that $v \in \cl_{j-1}(S_1)$ }\label{uniqueleaf}
-    \STATE $\cF_j \assign \cF_{j-1} \setminus \{S_1\}$\\
-    \STATE for all $S \in \cF_{j-1}$ such that $S \ne S_1$ set
-    $\cl_j(S) \assign
-    \cl_{j-1}(S)$\\
-    \STATE $X \assign S_1 \setminus \bigcup_{S \in \cF_{j-1}, S \ne S_1}S$\\
-    \IF{$X$ is empty} \label{xempty} \STATE {\bf exit} \label{ln:exit2} \ENDIF
-    \STATE $x \assign $ arbitrary element from $X$\\
-    \STATE $\cF_j \assign \cF_j \cup \{\{x\}, S_1 \setminus \{x\}\} $\\
-    \STATE $\cl_j(\{x\}) \assign \{v\}$\\
-    \STATE $\cl_j(S_1 \setminus \{x\}) \assign \cl_{j-1}(S_1) \setminus \{v\}$\\
-    \STATE $j \assign j+1$\\
-    \ENDWHILE
-    \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
-    \RETURN $(\cF', \cl')$
-  \end{algorithmic}
-\end{algorithm}
-
-Suppose the input ICPPL $(\cF, \cl)$ is feasible, yet set $X$ in
-Algorithm~\ref{leafasgn} is empty in some iteration of the {\bf while}
-loop. This will abort our procedure of finding the hypergraph
-isomorphism. The following lemma shows that this will not happen.
-
-\begin{lemma}
-  \label{lem:xnotempty}
-  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{leafasgn} is
-  feasible, then for all iterations $j > 0$ of the {\em \bf while}
-  loop, the {\em \bf exit} statement in line~\ref{ln:exit2} does not
-  execute.
-\end{lemma}
-\begin{proof}
-  Assume $X$ is empty for some iteration $j > 0$. We know that $v$ is
-  an element of $\cl_{j-1}(S_1)$. Since it is uniquely present in
-  $\cl_{j-1}(S_1)$, it is clear that $v \in \cl_{j-1}(S_1) \setminus
-  \bigcup_{(S \in \cF_{j-1}) \wedge (S \ne S_1)}\cl_{j-1}(S)$.  Note
-  that for any $x \in S_1$ it is contained in at least two sets due to
-  our assumption about cardinality of $X$. Let $S_2 \in \cF_{j-1}$ be
-  another set that contains $x$. From the above argument, we know $v
-  \notin \cl_{j-1}(S_2)$. Therefore there cannot exist a hypergraph
-  isomorphism bijection that maps elements in $S_2$ to those in
-  $\cl_{j-1}(S_2)$. This contradicts our assumption that the input is
-  feasible. Thus $X$ cannot be empty if input is ICPPL and feasible.
-  \qed
-\end{proof}
-
-\begin{lemma}
-  \label{lem:invar3}
-  In Algorithm~\ref{leafasgn}, for all $j > 0$, at the end of the
-  $j$th iteration of the {\bf while} loop the four invariants given in
-  Lemma~\ref{lem:invar1} hold.
-\end{lemma}
-\begin{proof}
-  By Lemma~\ref{lem:xnotempty} we know that set $X$ will not be empty
-  in any iteration of the {\em \bf while} loop if input ICPPL $(\cF,
-  \cl)$ is feasible and $\cl_j$ is always computed for all $j >
-  0$. Also note that removing a leaf from any path keeps the new path
-  connected. Thus invariant I is obviously true. In every iteration $j
-  > 0$, we remove exactly one element $x$ from one set $S$ in $\cF$
-  and exactly one vertex $v$ which is a leaf from one path
-  $\cl_{j-1}(S)$ in $T$. This is because as seen in
-  Lemma~\ref{lem:xnotempty}, $x$ is exclusive to $S$ and $v$ is
-  exclusive to $\cl_{j-1}(S)$. Due to this fact, it is clear that the
-  intersection cardinality equations do not change, i.e., invariants
-  II, III, IV remain true. On the other hand, if the input ICPPL is
-  not feasible the invariants are vacuously true. \qed
-\end{proof}
-
-% \textcolor{cyan}{
-% \begin{lemma}
-%   \label{lem:notfeasibleexit}
-%   \tnote{IS THIS CORRECT?}  If input ICPPL $(\cF, \cl)$ is not
-%   feasible, then in one of the recursive calls to Algorithm --3--, the
-%   {\em \bf exit} statement in line x in Algorithm --1 or line y in
-%   Algorithm --2 will get executed.
-% \end{lemma}
-% }
-
-We have seen two filtering algorithms above, namely,
-Algorithm~\ref{perms} {\tt filter\_1} and Algorithm~\ref{leafasgn}
-{\tt filter\_2} which when executed serially repectively result in a
-new ICPPL on the same universe $U$ and tree $T$. We also proved that
-if the input is indeed feasible, these algorithms do indeed output the
-filtered ICPPL. Now we present the algorithmic characterization of a
-feasible tree path labeling by way of Algorithm~\ref{al:icppl-find-isomorph}.
-
-Algorithm~\ref{al:icppl-find-isomorph} computes a
-hypergraph isomorphism $\phi$ recursively using Algorithm~\ref{perms}
-and Algorithm~\ref{leafasgn} and pruning the leaves of the input
-tree. In brief, it is done as follows. Algorithm~\ref{leafasgn} gives
-us the leaf labels in $\cF_2$, i.e., the elements in $supp(\cF)$ that
-map to leaves in $T$, where $(\cF_2, \cl_2)$ is the output of
-Algorithm~\ref{leafasgn}. All leaves in $T$ are then pruned away. The
-leaf labels are removed from the path labeling $\cl_2$ and the
-corresponding elements are removed from the corresponding sets in
-$\cF_2$. This tree pruning algorithm is recursively called on the
-altered hypergraph $\cF'$, path label $\cl'$ and tree $T'$. The
-recursive call returns the bijection $\phi''$ for the rest of the
-elements in $supp(\cF)$ which along with the leaf labels $\phi'$ gives
-us the hypergraph isomorphism $\phi$.  The following lemma formalizes
-the characeterization of feasible path labeling.
-
-\begin{algorithm}[h]
-  \caption{{\tt get- hypergraph- isomorphism ($\cF, \cl, T$)}}
-  \label{al:icppl-find-isomorph}
-  \begin{algorithmic}[\lndisplay]
-
-    \IF{$T$ is empty}
-    \RETURN $\emptyset$\\
-    \ENDIF
-    \STATE $L \assign \{v \mid v \text{ is a leaf in }      T\}$\\
-    \STATE $(\cF_1, \cl_1) \assign$ {\tt filter\_1($\cF, \cl,
-      T$)}\\
-    \STATE $(\cF_2, \cl_2) \assign$ {\tt filter\_2($\cF_1,
-      \cl_1, T$)}\\
-
-    \STATE $(\cF', \cl') \assign (\cF_2, \cl_2)$\\
-    \STATE $\phi' \leftarrow \emptyset$
-
-    \FOR {every $v \in L$}
-    \STATE $\phi'(x) \assign v$ where $x \in \cl_2^{-1}(\{v\})$
-    \COMMENT {Copy the leaf labels to a one to one function $\phi':
-      supp(\cF) \rightarrow L$
-      }\\
-    \STATE Remove $\{x\}$ and $\{v\}$ from $\cF'$, $\cl'$  appropriately\\
-    \ENDFOR
-
-    \STATE $T' \assign T \setminus L$
-
-    \STATE $\phi'' \assign$ {\tt
-      get-hypergraph-isomorphism($\cF', \cl', T'$)}
-    \STATE $\phi \assign \phi'' \cup \phi'$ \\
-    \RETURN $\phi$
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma}
-  \label{lem:hyperiso}  %{lem:perm}
-  If $(\cF, \cl)$ is an ICPPL from a tree $T$ and
-  Algorithm~\ref{al:icppl-find-isomorph}, {\tt
-    get- hypergraph- isomorphism ($\cF, \cl, T$)} returns a non-empty
-  function, then there exists a hypergraph isomorphism $\phi :
-  supp(\cF) \rightarrow V(T)$ such that the $\phi$-induced tree path
-  labeling is equal to $\cl$ or $\cl_\phi = \cl$.
-\end{lemma}
-\begin{proof}
-  It is clear that in the end of every recursive call to
-  Algorithm~\ref{al:icppl-find-isomorph}, the function $\phi'$ is one
-  to one involving all the leaves in the tree passed as input to that
-  recursive call. Moreover, by Lemma~\ref{lem:noexit1} and
-  Lemma~\ref{lem:xnotempty} it is consistent with the tree path
-  labeling $\cl$ passed. The tree pruning is done by only removing
-  leaves in each call to the function and is done till the tree
-  becomes empty. Thus the returned function $\phi: supp(\cF)
-  \rightarrow V(T)$ is a union of mutually exclusive one to one
-  functions exhausting all vertices of the tree. In other words, it is
-  a bijection from $supp(\cF)$ to $V(T)$ inducing the given path
-  labeling $\cl$ and thus a hypergraph isomorphism. \qed
-\end{proof}
-
-\begin{theorem}
-  \label{th:charac}
-  A path labeling $(\cF, \cl)$ on tree $T$ is feasible iff it is an
-  ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
-  T)$ as input returns a non-empty function.
-\end{theorem}
-\begin{proof}
-  From Lemma~\ref{lem:hyperiso}, we know that if $(\cF, \cl)$ is an
-  ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
-  T)$ as input returns a non-empty function, $(\cF, \cl)$ is feasible.
-  Now consider the case where $(\cF, \cl)$ is feasible. i.e. there
-  exists a hypergraph isomorphism $\phi$ such that $\cl_\phi =
-  \cl$. Lemma~\ref{lem:noexit1} and Lemma~\ref{lem:xnotempty} show us
-  that filter 1 and filter 2 do not exit if input is feasible. Thus
-  Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
-  function.\qed
-\end{proof}
-
-\section{ Computing feasible TPL with special target trees \tnote{give
-    problem definition etc}}
-\label{sec:spltargettree}
-
-Section~\ref{sec:feasible} described properties that a TPL must have
-for it to be feasible. The next problem of interest is to test if a
-given hypergraph is a path hypergraph with respect to a given target
-tree\footnote{A larger problem would be to find if a given hypergraph
-  is a path graph on any tree. This problem is not addressed in this
-  thesis.}. In other words, the problem is to find out if a feasible
-tree path labeling exists from a given target tree for a given
-hypergraph. In this section we will see two special cases of this
-problem where the target tree is from a particular family of
-trees. The first one, where the tree is a path, is a shown to be
-equivalent to the well studied problem of consecutive-ones in
-Section~\ref{sec:icpplicpia}. The second one, where the tree is a
-$k$-subdivided tree\footnote{See Section~\ref{ch:prelims} for the
-  formal definition.}, has been solved using a polynomial time
-algorithm. The latter problem enforces some conditions on the
-hypergraph too which will be seen in Section~\ref{sec:ksubdivstar}.
-
-\subsection{Target tree is a Path}
-\label{sec:icpplicpia}
-
-Consider a special case of ICPPL with the following properties.
-\begin{enumerate}
-\item Given tree $T$ is a path. Hence, all path labels are interval
-  labels.
-\item Only pairwise intersection cardinality preservation is
-  sufficient. i.e. property (iii) in ICPPL is not enforced.
-\item The filter algorithms do not have {\em \bf exit} statements.
-\end{enumerate}
-This is called an Intersection Cardinality Preservation Interval
-Assignment (ICPIA) \cite{nsnrs09}. This structure and its algorithm is
-used in the next section for finding tree path labeling from a
-$k$-subdivided star due to this graph's close relationship with
-intervals.
-
-\subsection{Target tree is a $k$-subdivided Star}
-\label{sec:ksubdivstar}
-
-\begin{figure}[t] %[htbp] 
-  \centering
-  \begin{tabular}{lr}
-    (a) \includegraphics{cop-tree_thesis.1}&
-    (b) \includegraphics{cop-tree_thesis.2}
-  \end{tabular}
-  \caption{\figtabsize (a) $8$-subdivided star with 7 rays (b)
-    3-subdivided star with 3 rays}
-  \label{fig:kstar}
-\end{figure}
-
-In this section we consider the problem of assigning paths from a
-$k$-subdivided star $T$ to a given set system $\cF$.  We consider
-$\cF$ for which the overlap graph $\bO(\cF)$ is connected.  The
-overlap graph is well-known from the work of
-\cite{kklv10,nsnrs09,wlh02}.  We use the notation in
-\cite{kklv10}. Recall from Section~\ref{ch:prelims} that hyperedges
-$S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$
-and $S'$ have a non-empty intersection but neither of them is
-contained in the other. The overlap graph $\bO(\cF)$ is a graph in
-which the vertices correspond to the sets in $\cF$, and the vertices
-corresponding to the hyperedges $S$ and $S'$ are adjacent if and only
-if they overlap.  Note that the intersection graph of $\cF$,
-$\bI(\cF)$ is different from $\bO(\cF)$ and $\bO(\cF) \subseteq
-\bI(\cF)$.  A connected component of $\bO(\cF)$ is called an overlap
-component of $\cF$.  An interesting property of the overlap components
-is that any two distinct overlap components, say $\cO_1$ and $\cO_2$,
-are such that any two sets $S_1 \in \cO_1$ and $S_2 \in \cO_2$ are
-disjoint, or, w.l.o.g, all the sets in $\cO_1$ are contained within
-one set in $\cO_2$.  This containment relation naturally determines a
-decomposition of the overlap components into rooted containment trees.
-We consider the case when there is only one rooted containment tree,
-and we first present our algorithm when $\bO(\cF)$ is connected.  It
-is easy to see that once the path labeling to the overlap component in
-the root of the containment tree is achieved, the path labeling to the
-other overlap components in the rooted containment tree is essentially
-finding a path labeling when the target tree is a path: each target
-path is a path that is allocated to sets in the root overlap
-component.  Therefore, for the rest of this section, $\bO(\cF)$ is a
-connected graph. We also assume that all hyperedges are of cardinality
-at most $k+2$.
-
-Recall from Section~\ref{ch:prelims} that a $k$-subdivided star is a
-star with each edge subdivided $k$ times. Therefore, a $k$-subdivided
-star has a central vertex which we call the {\em root}, and each root
-to leaf path is called a {\em ray}. First, we observe that by removing
-the root $r$ from $T$, we get a collection of $p$ vertex disjoint
-paths of length $k+1$, $p$ being the number of leaves in $T$.  We
-denote the rays by $R_1, \ldots, R_p$, and the number of vertices in
-$R_i$, $i \in [p]$ is $k+2$.  Let $\seq{v_{i1},\ldots,v_{i(k+2)}=r}$
-denote the sequence of vertices in $R_i$, where $v_{i1}$ is the
-leaf. Note that $r$ is a common vertex to all $R_i$.
-
-In this section the given hypergraph, the $k$-subdivided star and the
-root of the star are denoted by $\cO$, $T$ and vertex $r$,
-respectively.
-% For each hyperedge $X \in \cO$, we will maintain a 2-tuple of non-negative
-% numbers $\seq{p_1(X), p_2(X)}$.  The numbers satisfy the property that
-% $p_1(X) + p_2(X) \leq |X|$, and at the end of path labeling, for each
-% $X$, $p_1(X) + p_2(X) = |X|$.  This signifies the algorithm tracking
-% the lengths of subpaths of the path assigned to $X$ from at most two
-% rays. We also maintain another parameter called the {\em residue} of
-% $X$ denoted by $s(X)=|X| - p_1(X)$. This signifies the residue path
-% length that must be assigned to $X$ which must be from another
-% ray. For instance, if $X$ is labeled a path from only one ray, then
-% $p_1(X) = |X|$, $p_2(X) = 0$ and $s(X) = 0$.
-
-% 
-% We iteratively consider each ray from which paths will be assigned to
-% hyperedges. At the beginning of each iteration hyperedges of $\cO$ are
-% classifed into the following disjoint sets.
-% \begin{enumerate}
-% \item [$\cL_1^i$] {\em Labeled without $r$.} Those that have been
-%   labeled with a path which does not contain $r$ in one of the
-%   previous iterations.\\  $\cL_1^i = \set{ X \mid p_1(X) = |X| \text{ and
-%     } p_2(X) = 0 \text{ and } s(X) = 0, X \in \cO}$
-% \item [$\cL_2^i$] {\em Labeled with $r$.} Those that have been labeled
-%   with two subpaths of $\cl(X)$ containing $r$ from two different rays
-%   in two previous iterations.\\ $\cL_2^i = \set{X \mid 0 < p_1\left(X\right),
-%     p_2\left(X\right) < |X| \text{ and } s(X) = 0, X \in \cO}$
-%   \item [$\cT_1^i$] {\em Type 1 / partially labeled.} Those that have
-%     been labeled with one path containing $r$ from a single ray in one
-%     of the previous iterations. Here, $p_1(X)$ denotes the length of
-%     the subpath of $\cl(X)$ that $X$ has been so far labeled
-%     with. Also, it is clear that such a path must start at $r$ and
-%     must be in a ray different from the one corresponding to $p_1(X)$.\\
-%     $\cT_1^i = \set{ X \mid 0 < p_1(X) < |X| \text{ and } p_2(X) = 0
-%       \text{ and } s(X) > 0, X \in \cO}$
-%   \item [$\cT_2^i$] {\em Type 2 / not labeled.} Those that have not been
-%     labeled with a path in any previous iteration.\\
-%     $\cT_2^i = \set{ X \mid p_1(X) = p_2(X) = 0 \text{ and } s(X) = |X|,
-%       X \in \cO}$
-% \end{enumerate}
-% \vspace{-2mm}
-% \begin{align*}
-%   \cO &= \cL_1^i \cup \cL_2^i \cup \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]\\
-%   \cO_i &= \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]
-% \end{align*}
-
-The set $\cO_i$ refers to the set of hyperedges $\cT_1^i \cup \cT_2^i$
-in the $i$th iteration.  Note that $\cO_1 = \cO$.  In the $i$th
-iteration, hyperedges from $\cO_i$ are assigned paths from $R_i$ using
-the following rules. Also the end of the iteration, $\cL_1^{i+1},
-\cL_2^{i+1}, \cT_1^{i+1}, \cT_2^{i+1}$ are set to $\cL_1^{i},
-\cL_2^{i}, \cT_1^{i}, \cT_2^{i}$ respectively, along with some
-case-specific changes mentioned in the rules below.
-
-\noindent
-Suppose the given hypergraph has a feasible path labeling to the given
-tree $T$. Let this labeling be $\cl$. Now we present an algorithm that
-lists the sequence of edges that maintain a set of properties
-necessary for them to have a feasible tree path labeling.
-\begin{enumerate}[I.]
-\item {\bf Step 1}
-  \begin{enumerate}
-  \item {\bf {There are type 1
-        edges, $\mathbf{|\cT_1^i| > 0}$}:} 
-% , and for each such $X$ assign the
-%     path in $R_i$ of length $s(X)$ starting at $v_{i,k+2}=r$.  $X$ is
-%     then removed from $\cO_i$, and $p_2(X)=s(X)$, and $s(X)=0$.If any
-%     intersection cardinality is violated because of this assignment,
-%     EXIT reporting the non-existence of a feasible assignment.
-  {If there is only one $\cT_1^i$ hyperedge, label it to the unique
-    path in $R_i$ of length $s(X)$ starting at $v_{i(k+2)}=r$. If
-    there are more than one $\cT_1^i$ hyperedges, for any pair of
-    hyperedges $S_1, S_2 \in \cT_1^i$, one of the following is
-    true. Let $\cl'(X) \subseteq R_j$ denote the subpath of length
-    $p_1(X)$ that has been assigned to $X$ from a previously
-    considered ray $R_j$.}
-  \begin{enumerate}[{\em Case }1. ]
-  \item $|\cl'(S_1) \cap \cl'(S_2)| = 1$ and $|S_1 \cap S_2| > 1$
-    i.e. they are in different rays and their residues must contain
-    their intersection. Hence assign both to $R_i$.
-  \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| >
-    |\cl'(S_1) \cap \cl'(S_2)|$ i.e. they are in the same rays and
-    their residues must contain their intersection. Hence assign both
-    to $R_i$.
-  \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| =
-    |\cl'(S_1) \cap \cl'(S_2)|$. i.e. the intersection cardinality has
-    been met. Hence assign only $S_1$ to $R_i$ and $S_2$ must be
-    assigned to some $i \ge p$.
-  \end{enumerate}
-  Remove the labeled edge or edges from $\cO_i$ and add it to
-  $\cO_{i+1}$. Also do not add it to $\cT_1^{i+1}$ and add to
-  $\cL_2^{i+1}$.
-\item {\bf {There are no type-1 edges nor} super-marginal type-2
-    hyperedges:} If all type-2 edges of length at most $k+1$, EXIT by
-  saying there is no ICPPL.  Otherwise, Find an $X$ of type-2 such
-  that $|X| \geq k+2$, assign it the unique path starting at
-  $v_{i(k+1)}=r$ of length $k+2$, and set $p_1(X)=k+2,
-  s(X)=|X|-(k+2)$, assign it the unique path starting at
-  $v_{i(k+1)}=r$ of length $k+2$, mark this as a type-1 hyperedge, and
-  add it to $\cO_{i+1}$ after removing it from $\cO_i$.
-  % \remove{Next, iteratively, consider each type-1 edge $X \in
-  %   \cO_i$, and if it intersects a previously assigned $Y$ such that
-  %   $\cl(Y) \subseteq R_i$, assign a path of length $s(X)$
-  %   containing $v_{i(k+2)}=r$, set $p_2(X)=s(X), s(X)=0$, and remove
-  %   $X$ from $\cO_i$.}\tnote{covered in step 2}
-
-\item {\bf There are super-marginal type-2 hyperedges {and no type-1
-      edges:}} Let $X$ be a super-marginal hyperedge. If $|X| \leq
-  k+2$, then $X$ is assigned the path of length $|X|$ starting at
-  $v_{i1}$, and $p_1(X) = |X|, p_2(X)=0$. $X$ is removed from $\cO_i$
-  and not added to $\cO_{i+1}$.  If $|X| > k+2$, then $p_1(X)=k+2$,
-  assign it the unique path starting at $v_{i(k+1)}=r$ of length $k+2$
-  and $s(X) = |X| - (k+2)$.  In this case, $X$ is classified as a
-  Type-1 edge, removed from $\cO_i$, and added to $\cO_{i+1}$.
-\end{enumerate}
-\item {\bf Step 2:} Iteratively, a hyperedge $X$ is selected from
-  $\cO_i$ that has an overlap with one of the hyperedges $Y$ such that
-  $\cl(Y) \subseteq R_i$, and a unique path is assigned to $X$.  The
-  path, say $U(X) \subseteq R_i$, that is assigned can be decided
-  unambiguously since the $X \overlap Y$, and all intersection
-  cardinalities can be preserved only at one of the ends of $\cl(Y)$.
-
-  \noindent
-  Let $\cl(X)$ denote the unique path assigned to $X$.  If $X$ is a
-  type-2 hyperedge and if the unique path of length $|X|$ does not
-  contain $r$, then $p_1(X) = |X|, p_2(X)=0$, and $X$ is removed from
-  $\cO_i$.  In the case when $\cl(X)$ has to contain $r$, then
-  $p_1(X)$ is the length of the path, $p_2(X)=0$, and $s(X) =
-  |X|-p_1(X)$. Further $X$ is classified as a type-1 hyperedge, added
-  to $\cO_{i+1}$ and removed from $\cO_i$.  In the case when $X$ is a
-  type-1 hyperedge, then we check if $U(X)$, which is of length $s(X)$
-  contains $r$. If it does, then we assign $\cl(X) \leftarrow \cl(X)
-  \cup U(X)$, remove $X$ from $\cO_i$ and do not add it to
-  $\cO_{i+1}$.  If not, then we {\bf Exit} reporting that an
-  assignment cannot be found.  The iteration ends when no hyperedge in
-  $\cO_i$ has an overlap with a hyperedge assigned to $R_i$.
-\end{enumerate}
-
-\noindent
-The order in which they are assigned is the sequence of hyperedges
-with the following properties. Note that if such a sequence cannot be
-computed, $\cO$ cannot have a feasible path labeling from $T$.
-
-\noindent
-\tnote{need to formalize the below properties} In the following
-lemmas we identify a set of necessary conditions for $\cF$ to have an
-ICPPL in the $k$-subdivided star $T$.  If during the execution of the
-above algorithm, one of these necessary conditions is not satisfied,
-the algorithm exits reporting the non-existence of an ICPPL.
-\begin{lemma}
-  Let all hyperedges in $\cO_i$ be type-1 edges.  Then there is a
-  maximal subset $\ccT_i \subseteq \cO_i$ with the following
-  properties:
-  \begin{enumerate}
-  \item $\ccT_i$ form an inclusion chain.
-  \item For all $X \in \ccT_i$, $s(X) \leq k+2$, and There is a an $X
-    \in \ccT_i$ such that $s(X)=k+2$.
-  \end{enumerate}
-\end{lemma}
-\begin{lemma}
-  If there are no \marginal type-2 edges in $\cO_i$, then there exists
-  at least $p-i$ type-2 hyperedges $X \in \cO_i$ such that $|X| \geq
-  k+2$.
-  % is a exist type-2 hyperedges of length at least $k+2$, and there
-  % must be at least $r-i$ such type-2 hyperedge, basically at least
-  % one per ray.  Each hyperedge contained in such an hyperedges of
-  % length at least $k+2$, is transitively related under the overlap
-  % relation to one that gets assigned a path that contains
-  % $r$. Basically, the proof will argue that otherwise, there will be
-  % a \marginal type-2 hyperedge, or that it is infeasible.  Also, it
-  % is here we assume that we assume that the size of the two
-  % universes is the same.
-\end{lemma}
-\begin{lemma}
-  At the end of Step-1 in the $i$-th iteration, if one hyperedge $X$
-  of type-2 is such that $\cl(X) \subseteq R_i$, then all other
-  hyperedges in $\cO_i$ are connected to $X$ in the overlap component.
-\end{lemma}
-\begin{lemma}
-  At the end of Step-2, If control has exit at any time, there is no
-  ICPPL. If control has not exit, then $R_i$ is saturated.  No
-  hyperedge of $\cO_{i+1}$ will get a path from $R_i$ in the future
-  iterations.  No type-2 hyperedge of $\cO_{i+1}$ will get a path from
-  $R_i$.  Basically $R_i$ is done.
-\end{lemma}
-\begin{lemma}
-  Finally, we need to prove that the assignment is an ICPPL. Secondly,
-  if there is a permutation then maps sets to paths, then it is indeed
-  an ICPPL, and our algorithm will basically find it.  It is a unique
-  assignment upto permutation of the leaves.
-\end{lemma}
-
-\begin{lemma}
-  \label{lem:sup-mar}
-  If $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
-  tree path labeling to tree $T$, then $\cl(X)$ will
-  contain a leaf in $T$.\\
-  On the otherhand, if $\cl(X)$ has a leaf in
-  $T$, $X$ is marginal but may not be super-marginal.
-\end{lemma}
-\begin{proof}
-  Suppose $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
-  path labeling from $T$.  Assume $\cl(X)$ does not have a leaf.  Let
-  $R_i$ be one of the rays (or the only ray) $\cl(X)$ is part of.
-  Since $X$ is in a connected overlap component, there exists $Y_1 \in
-  \cF$ and $X \nsubseteq Y_1$ such that $Y_1 \overlap X$ and $Y_1$ has
-  at least one vertex closer to the leaf in $R_i$ than any vertex in
-  $X$. Similarly with the same argument there exists $Y_2 \in \cF$
-  with same subset and overlap relation with $X$ except it has has at
-  least one vertex farther away from the leaf in $R_i$ than any vertex
-  in $X$. Clearly $Y_1 \cap X$ and $Y_2 \cap X$ cannot be part of same
-  inclusion chain which contradicts that assumption $X$ is
-  super-marginal. Thus the claim is proven.\qed
-\end{proof}
-
-\noindent
-Consider the overlap graph $\bO(\cF)$ of the given hypergraph
-$\cF$. Let $S_{sm} \in \cF$ be such that it is a super-marginal
-hyperedge.  Algorithm~\ref{al:ktree-label} uses $S_{sm}$ along with
-the overlap graph $\bO(\cF)$ to calculate the feasible tree path
-labeling to the $k$-subdivded tree $T$.
-
-\begin{algorithm}[h]
-  \caption{{\tt compute-ksubtree-path-labeling($X, \cF, T$)}}
-  \label{al:ktree-label}
-  \begin{algorithmic}[\lndisplay]
-    \IF{$X = S_{sm}$}
-    \STATE -- TBD --\\
-    \ELSE 
-    \STATE -- TBD --\\
-    \ENDIF
-
- \end{algorithmic}
-\end{algorithm}
-
-\section{ TPL with no restrictions}
-\label{sec:norestraint}
-
-{\tt abstract begin}
-  it is
-  known that if the given tree is a path a feasible assignment can be
-  found in polynomial time, and we observe that it can actually be
-  done in logspace. \tnote{[TRUE?]}
-{\tt abstract end}
-
-In Section \ref{prelims} we present the
-necessary preliminaries, in Section \ref{feasible} we present our
-characterization of feasible tree path assignments, and in Section
-\ref{decompos} we present the characterizing subproblems for finding a
-bijection between $U$ and $V(T)$ such that sets map to tree
-paths. Finally, in Section \ref{complexity} we conclude by showing
-that Tree Path Assignment is GI-Complete, and also observe that
-Consecutive Ones Testing is in Logspace.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC \subsection{Introduction}
-
-\annote{ We refer to this as the Tree Path Assignment
-problem for an input $(\cF,T)$ pair.}{Add this ``pair'' as a
-  definition in ch:prelims}
-
-\remove{It is an interesting fact that for a matrix with the COP, the
-intersection graph of the corresponding set system is an interval
-graph.  A similar connection to a subclass of chordal graphs, and this
-subclass contains interval graphs, exists for the generalization of
-COP.  In this case, the intersection graph of the corresponding set
-system must be a path graph. Chordal graphs are of great
-significance, extensively studied, and have several applications.  One
-of the well known and interesting properties of a chordal graphs is
-its connection with intersection graphs cite{mcg04}. For every chordal
-graph, there exists a tree and a family of subtrees of this tree such
-that the intersection graph of this family is isomorphic to the
-chordal graph cite{plr70,gav78,bp93}.  Certain format of these trees
-are called clique trees cite{apy92} of the graph which is a compact
-representation of the chordal graph. There has also been work done on
-the generalization of clique trees to clique hypergraphs cite{km02}.
-If the chordal graph can be represented as the intersection graph of
-paths in a tree, then the graph is called path graph cite{mcg04} or
-path chordal graph.  Therefore, it is clear that if there is a
-bijection from $U$ to $V(T)$ such that the sets map to paths, then the
-intersection graph of the set system is indeed a path chordal graph.
-This is, however, only a necessary condition and can be checked
-efficiently, as path graph recognition is polynomial
-solvable cite{gav78,aas93}.  Indeed, it is possible to construct a set
-system and tree, such that the intersection graph is a path chordal
-graph, but there is no bijection between $U$ and $V(T)$ such that the
-sets map to paths.  This connection indeed suggests that our problem
-is indeed as hard as path graph isomorphism.  Further path graph
-isomorphism is known be isomorphism-complete, see for example
-cite{kklv10}. }
-In the second part of this paper, we decompose our
-search for a bijection between $U$ and $V(T)$ into subproblems.  Each
-subproblem is on a set system in which for each set, there is another
-set in the set system with which the intersection is {\em strict}-
-there is a non-empty intersection, but neither is contained in the
-other.  This is in the spirit of results in \cite{wlh02,nsnrs09} where
-to test for COP in a given matrix, the COP problem is solved on an
-equivalent set of prime matrices.  \annote{Our decomposition localizes the
-challenge of path graph isomorphism to two problems.}{HOW?}
-
-Finally, we show that Tree Path Assignment is isomorphism-complete.
-We also point out Consecutive Ones Testing is in Logspace from two
-different results in the literature \cite{kklv10, mcc04}. To the best
-of our knowledge this observation has not been made earlier.
-
-% \noindent
-% It has been long known that interval graph recognition is in
-% logspace\cite{rei84}. Recently interval graph isomorphism was also
-% shown to be logspace decidable using a logspace canonization
-% algorithm by \cite{kklv10}.  This result is built on top of logspace
-% results of undirected graph connectivity \cite{rei08}, logspace
-% tractability using a certain logical formalism called FP+C and
-% modular decomposition of interval graphs\cite{lau10} etc.  Interval
-% graphs are closely connected to binary matrices with COP. The
-% maximal clique vertex incidence matrix (matrix with rows
-% representing maximal cliques and columns representing vertices of a
-% graph) has COP on columns iff the graph is an interval
-% graph\cite{fg65}. This follows from the interval graph
-% characterization by \cite{gh64}. Due to this close relation it is
-% natural to see if consecutive ones property can be
-% tested in logspace. \\
-% \noindent
-% We also explore some extensions of the interval assignment problem
-% in \cite{nsnrs09}, namely caterpillar path assignment problem.
-
-% We present a logspace algorithm here that uses the ICPIA
-% characterization of binary matrices with COP (set system associated
-% with such a matrix)\cite{nsnrs09}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC\subsection{Preliminaries} 
-\label{prelims} 
-In this paper, the collection
-$\F = \{S_i \mid S_i \subseteq U, S_i \ne \O, i \in I\}$ is a {\em set
-  system} of universe, $U = \{1, \ldots, n\}$. Moreoever, a set system
-is assumed to ``cover'' the universe,
-i.e. $ \bigcup_{i \in I}S_i = U$. \\
-
-\noindent
-The graph $T=(V,E)$ represents a given tree with $|V| = n$. Further,
-for simplicity, $V$ is defined as $\{1,\ldots,n\}$. All paths
-referred to in this paper are paths of $T$ unless explicitly specified. \\
-
-\noindent
-A {\em path assignment} $\A$ to $\F$ is defined as a set assignment
-where second universe is the vertex set $V$ of a given tree $T$ and
-every second subset in the ordered pairs is a path in this
-tree. Formally, the definition is as follows.
-\begin{align*}
-  \A = \{ (S_i,P_i) \mid S_i \in \F, P_i \subseteq V \text{
-    s.t. }T[P_i] \text{ is a path, } i \in I \}
-\end{align*}
-In other words, $P_i$ is the path on the tree $T$ assigned to $S_i$ in
-$\A$. As mentioned before for set systems, the paths cover the whole
-tree, i.e. $\bigcup_{i \in I}P_i = V$ \\
-
-\noindent
-Generalizing the definition of {\em feasibility} in \cite{nsnrs09} to
-a set assignment, a path assignment $\A$ is defined to be {\em
-  feasible} if there exists a bijection defined as follows.
-\begin{align}
-  \sigma: U \rightarrow V(T), \text{ such that }\sigma(S_i) = P_i
-  \text{ for all } i \in I, \sigma \text{ is a bijection}
-  \label{eq:stf}
-\end{align}
-% The assignment $\A$ is then called {\em Set Translation Feasible}.
-
-\noindent
-Let $X$ be a partially ordered set with $\preccurlyeq$ being the
-partial order on $X$.  $mub(X)$ represents an element in $X$ which is
-a maximal upper bound on $X$.  $X_m \in X$ is a maximal upper bound of
-$X$ if $\nexists X_q \in X$ such that $X_m
-\preccurlyeq X_q$. \\
-
-\noindent
-The set $I$ represents the index set $[m]$. If index $i$ is used
-without further qualification, it is meant to be $i \in I$. Any
-function, if not defined on a domain of sets, when applied on a set is
-understood as the function applied to each of its elements. i.e. for
-any function $f$ defined with domain $U$, the abuse of notation is as
-follows; $f(S)$ is used instead of $\hat f(S)$ where
-$\hat f(S) = \{y \mid y = f(x), x \in S\}$. \\
-
-\noindent
-When refering to a tree as $T$ it could be a reference to the tree
-itself, or the vertices of the tree. This will be clear from the
-context.\\
-
-\noindent
-Finally, an in-tree is a directed rooted tree in which all edges are
-directed toward to the root.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC\subsection{Characerization of Feasible Tree Path
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC  Assignment\tnote{Merge with the original charac section}} 
-\label{feasible} 
-Consider a path assignment $\cA =
-\{(S_i, P_i) \mid S_i \in \cF, P_i \text{ is a path from $T$}, i \in
-[m]\}$ to a set system $\cF = \{S_i \mid S_i \subseteq U, i \in
-[m]\}$, were $T$ is a given tree, $U$ is the set system's universe and
-$m$ is the number of sets in $\cF$. We call $\cA$ an {\em Intersection
-  Cardinality Preserving Path Assignment (ICPPA)} if it has the
-following properties.
-
-\begin{enumerate}
-\item [i.]  $|S_i| = |P_i|$ for all $i \in [m]$
-\item [ii.] $|S_i \cap S_j| = |P_i \cap P_j|$ for all $i,j \in [m]$
-  % \item [iii.] $|\bigcup_{i \in J} S_i| = |\bigcup_{i \in J} P_i|$
-  %   for all $J \subseteq I$\footnote{TBD: this is needed now to fix
-  %     some proofs}
-\item [iii.] $|S_i \cap S_j \cap S_k| = |P_i \cap P_j \cap P_k|$ for
-  all $i,j,k \in [m]$
-\end{enumerate}
-
-\begin{lemma}
-  \label{lem:setminuscard}
-  If $\cA$ is an ICPPA, and $(S_1, P_1),(S_2, P_2),(S_3, P_3) \in
-  \cA$, then $|S_1 \cap (S_2 \setminus S_3)| = |P_1 \cap (P_2
-  \setminus P_3)|$.
-\end{lemma}
-\begin{proof}
-  $|S_1 \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| =
-  |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to conditions (ii)
-  and (iii) of ICPPA, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
-  \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
-  = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. \qed
-\end{proof}
-
-\begin{lemma}
-  \label{lem:fourpaths} Consider four paths in a tree $Q_1, Q_2, Q_3,
-  Q_4$ such that they have nonempty pairwise intersection and $Q_1,
-  Q_2$ share a leaf. Then there exists distinct $i, j, k \in
-  \{1,2,3,4\}$ such that, $Q_1 \cap Q_2 \cap Q_3 \cap Q_4 = Q_i \cap
-  Q_j \cap Q_k$.
-\end{lemma}
-\begin{proof}
-  {\em Case 1:} w.l.o.g, consider $Q_3 \cap Q_4$ and let us call it
-  $Q$. This is clearly a path (intersection of two paths is a path).
-  % Since $Q_1, Q_2$ share a leaf, the following are paths $Q_1
-  % \setminus Q_2$, $Q_2 \setminus Q_1$, $Q_1 \cap Q_2$ and they are
-  % mutually disjoint.
-  Suppose $Q$ does not intersect with $Q_1 \setminus Q_2$, i.e. $Q
-  \cap (Q_1 \setminus Q_2) = \O$. Then $Q \cap Q_1 \cap Q_2 = Q \cap
-  Q_2$. Similarly, if $Q \cap (Q_2 \setminus Q_1) = \O$, $Q \cap Q_1
-  \cap Q_2 = Q \cap Q_1$. Thus it is clear that if the intersection of
-  any two paths does not intersect with any of the set differences of
-  the remaining two paths, the claim in the lemma is true.
-  % Note that $Q_1 \setminus Q_2$ and $Q_2
-  % \setminus Q_1$ are paths because $Q_1, Q_2$ share a leaf.\\
-  {\em Case 2:} Let us consider the compliment of the previous
-  case. i.e. the intersection of any two paths intersects with both
-  the set differences of the other two. First let us consider $Q \cap
-  (Q_1 \setminus Q_2) \ne \O$ and $Q \cap (Q_1 \setminus Q_2) \ne \O$,
-  where $Q = Q_3 \cap Q_4$. Since $Q_1$ and $Q_2$ share a leaf, there
-  is exactly one vertex at which they branch off from the path $Q_1
-  \cap Q_2$ into two paths $Q_1 \setminus Q_2$ and $Q_2 \setminus
-  Q_1$. Let this vertex be $v$. It is clear that if path $Q_3 \cap
-  Q_4$, must intersect with paths $Q_1 \setminus Q_2$ and $Q_2
-  \setminus Q_1$, it must contain $v$ since these are paths from a
-  tree. Moreover, $Q_3 \cap Q_4$ intersects with $Q_1 \cap Q_2$ at
-  exactly $v$ and only at $v$ which means that $Q_1 \cap Q_2$ does not
-  intersect with $Q_3 \setminus Q_4$ or $Q_4 \setminus Q_3$ which
-  contradicts initial condition of this case. Thus this
-  case cannot occur and case 1 is the only possible scenario. \\
-  Thus lemma is proven \qed
-\end{proof}
-
-\noindent
-In the remaining part of this section we show that a path assignment
-is feasible if and only if it is an ICPPA.  One direction of this
-claim is clear: that is a path assignment is feasible, then all
-intersection cardinalities are preserved, that is the path assignment
-is an ICPPA.  The reason is that a feasible path assignment has an
-associated bijection between $U$ and $V(T)$ such that the sets map to
-paths.  The rest of the section is devoted to constructively proving
-that it is sufficient, for the existence of a bijection, for a path
-assignment to be an ICPPA.  At a top-level, the constructive
-approaches refine the path assignment iteratively, such that at the
-end of each iteration we have a path assignment, and finally we have a
-family of bijections.  First we present and then prove the correctness
-of Algorithm \ref{perms}.  This algorithm refines the path assignment
-by considering pairs of paths that share a leaf.
-
-\begin{algorithm}[h]
-  \caption{Permutations from an ICPPA $\{(S_i,P_i) | i \in I\}$}
-  \label{perms}
-  \begin{algorithmic}
-    \STATE Let $\Pi_0=\{(S_i,P_i)| i \in I\}$\\
-    \STATE $j = 1$;
-    \label{shareleaf} \WHILE {There is $(P_1,Q_1), (P_2,Q_2) \in
-      \Pi_{j-1}$ with $Q_1$ and $Q_2$ having a common leaf} \STATE
-    $\Pi_j= \Pi_{j-1} \setminus \{(P_1,Q_1),(P_2,Q_2)\}$;
-    \label{setbreak}\STATE $\Pi_j = \Pi_j \cup \{(P_1 \cap P_2,Q_1
-    \cap Q_2), (P_1 \setminus P_2,Q_1 \setminus Q_2), (P_2 \setminus
-    P_1, Q_2 \setminus Q_1)\}$; \STATE $j = j+1$; \ENDWHILE \STATE
-    $\Pi = \Pi_j$; \STATE Return $\Pi$;
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma}
-  \label{lem:invar1}
-  In Algorithm \ref{perms}, at the end of $j$th iteration, $j \ge 0$,
-  of the while loop of Algorithm \ref{perms}, the following invariants
-  are maintained.
-  \begin{itemize}
-  \item {\em Invariant I:} $Q$ is a path in $T$ for each $(P,Q) \in
-    \Pi_j$
-  \item {\em Invariant II:} $|P|=|Q|$ for each $(P,Q) \in \Pi_j$
-  \item {\em Invariant III:} For any two $(P,Q), (P',Q') \in \Pi_j$,
-    $|P' \cap P''|=|Q' \cap Q''|$.
-  \item {\em Invariant IV:} For any three, $(P',Q'), (P'',Q''), (P, Q)
-    \in \Pi_j$, $|P' \cap P'' \cap P|=|Q' \cap Q'' \cap Q|$.
-  \end{itemize}
-\end{lemma}
-\begin{proof}
-  Proof is by induction on the number of iterations, $j$. In the rest
-  of the proof, the term ``new sets'' will refer to the new sets added
-  in $j$th iteration as defined in line \ref{setbreak} of Algorithm
-  \ref{perms}, i.e. the following three assignment pairs for some
-  $(P_1,Q_1), (P_2,Q_2) \in \Pi_{j-1}$ where $Q_1$ and $Q_2$ intersect
-  and share a leaf: $(P_1 \cap P_2, Q_1 \cap Q_2)$, or $(P_1 \setminus
-  P_2, Q_1 \setminus Q_2)$, or $(P_2 \setminus P_1, Q_2
-  \setminus Q_1)$.\\
-  \noindent
-  The base case, $\Pi_0 = \{(S_i,P_i) \mid i \in [m]\}$, is trivially
-  true since it is the input which is an ICPPA.  Assume the lemma is
-  true till the $j-1$ iteration. Consider $j$th
-  iteration:\\
-  \noindent
-  If $(P,Q)$, $(P',Q')$ and $(P'',Q'')$ are in $\Pi_{j}$ and
-  $\Pi_{j-1}$, all the invariants are
-  clearly true because they are from $j-1$ iteration.\\
-  If $(P,Q)$ is in $\Pi_{j}$ and not in $\Pi_{j-1}$, then it must be
-  one of the new sets added in $\Pi_j$. Since $(P_1,Q_1)$ and
-  $(P_2,Q_2)$ are from $\Pi_{j-1}$ and $Q_1,Q_2$ intersect and have a
-  common leaf, it can be verified that the
-  new sets are also paths. \\
-  By hypothesis for invariant III, invariant II also holds for $(P,Q)$
-  no matter which new set in $\Pi_j$ it
-  is.\\
-  To prove invariant III, if $(P,Q)$ and $(P',Q')$ are not in
-  $\Pi_{j-1}$, then they are both new sets and invariant III holds
-  trivially (new sets are disjoint). Next consider $(P,Q), (P',Q') \in
-  \Pi_j$ with only one of them, say $(P',Q')$, in $\Pi_{j-1}$. Then
-  $(P,Q)$ is one of the new sets added in line \ref{setbreak}. It is
-  easy to see that if $(P,Q)$ is $(P_1 \cap P_2, Q_1 \cap Q_2)$, then
-  due to invariant IV in hypothesis, invariant III becomes true in
-  this iteration. Similarly, using lemma \ref{lem:setminuscard}
-  invariant III is proven if $(P, Q)$ is $(P_1 \setminus P_2, Q_1
-  \setminus Q_2)$, or $(P_2 \setminus P_1, Q_2
-  \setminus Q_1)$.\\
-  To prove invariant IV, consider three assignments
-  $(P,Q),(P',Q'),(P'',Q'')$. If at least two of these pairs are in not
-  $\Pi_{j-1}$, then they are any two of the new sets. Note that these
-  new sets are disjoint and hence if $(P',Q'), (P'',Q'')$ are any of
-  these sets, $|P \cap P' \cap P''|=|Q \cap Q' \cap Q''|=0$ and
-  invariant IV is true. Now we consider the case if at most one of
-  $(P,Q),(P',Q'),(P'',Q'')$ is not in $\Pi_{j-1}$. If none of them are
-  not in $\Pi_{j-1}$ (i.e. all of them are in $\Pi_{j-1}$), invariant
-  IV is clearly true. Consider the case where exactly one of them is
-  not in $\Pi_{j-1}$. w.l.o.g let that be $(P,Q)$ and it could be any
-  one of the new sets. If $(P,Q)$ is $(P_1 \cap P_2, Q_1 \cap Q_2)$,
-  from lemma \ref{lem:fourpaths} and invariant III hypothesis,
-  invariant IV is proven. Similarly if $(P,Q)$ is any of the other new
-  sets, invariant IV is proven by also using lemma
-  \ref{lem:setminuscard}. \qed
-
-\end{proof}
-
-\noindent
-It can be observed that the output of algorithm \ref{perms} is such
-that every leaf is incident on at most a single path in the new set of
-assignments. This is due to the loop condition at line
-\ref{shareleaf}. Let $v_1$ be the leaf incident on path $P_i$. Assign
-to it any one element from $S_i \setminus \bigcup_{i \ne j}
-S_j$. Remove $(S_i, P_i)$ from assignments and add $(\{x_1\},
-\{v_1\}), (S_i \setminus \{x_1\}, P_i \setminus \{v_1\})$. Now all
-assignments except single leaf assignments are paths from the subtree
-$T_0 = T \setminus \{v \mid v \text{ is a leaf in } T\}$.
-
-% exclusively on the spine of the caterpillar. Consider the spine to
-% be an interval and run ICPIA algorithm on it to get the required
-% permutation. Algorithm \ref{leafasgn} makes this clear.
-
-\begin{algorithm}[h]
-  \caption{Leaf assignments from an ICPPA $\{(S_i,P_i) | i \in I\}$}
-  \label{leafasgn}
-  \begin{algorithmic}
-    \STATE Let $\Pi_0=\{(S_i,P_i)| i \in [m]\}$. Paths are such that
-    no
-    two paths $P_i, P_j, i \ne j$ share a leaf.\\
-    \STATE $j = 1$\\
-    \WHILE {there is a leaf $v$ and a unique $(S_{i_1}, P_{i_1})$ such
-      that $v \in P_{i_1}$}
-    \STATE $\Pi_j=   \Pi_{j-1} \setminus \{(S_{i_1}, P_{i_1})\}$\\
-    \STATE $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in I}S_i$
-    \IF{$X$ is empty} \STATE exit \ENDIF; \STATE Let $x = $ arbitrary
-    element from $X$ \vspace{2mm}
-
-    \STATE $\Pi_j = \Pi_j \cup \{(S_{i_1} \setminus \{x\}),(P_{i_1}
-    \setminus
-    \{v\}), (\{x\},\{v\})\}$\\
-    \STATE $j = j+1$\\
-    \ENDWHILE
-    \STATE $\Pi = \Pi_j$\\
-    \STATE Return $\Pi$\\
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma}
-  \label{lem:invar3}
-  In Algorithm \ref{leafasgn},for all $j \geq 0$, at the end of the
-  $j$th iteration the four invariants given in lemma \ref{lem:invar1}
-  are valid.
-  % Moreover, $X$ as defined in the algorithm is non-empty if this is
-  % an ICPPA.
-\end{lemma}
-\begin{proof}
-  First we see that $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in
-    I}S_i$ is non empty in every iteration for an ICPPA. Suppose $X$
-  is empty. We know that $v \in P_{i_1} \setminus \bigcup_{i \ne i_1,
-    i \in I}P_i$ since $v$ is in the unique path $P_{i_1}$. Since this
-  is an ICPPA $|S_{i_1}| = |P_{i_1}|$. For any $x \in S_{i_1}$ it is
-  contained in at least two sets due to our assumption. Let $S_{i_2}$
-  be a second set that contains $x$. We know $v \notin
-  P_{i_2}$. Therefore there cannot exist a permutation that maps
-  elements of $S_{i_2}$ to $P_{i_2}$. This contradicts our assumption
-  that this is an ICPPA. Thus $X$ cannot be empty.
-
-  We use mathematical induction on the number of iterations for this
-  proof. The term ``new sets'' will refer to the sets added in $\Pi_j$
-  in the $j$th iteration, i.e. $(P' \setminus \{x\},Q' \setminus
-  \{v\})$ and $(\{x\},\{v\})$ for some $(P',Q')$ in $\Pi_{j-1}$ such
-  that $v$ is a leaf and $Q'$ is the unique path
-  incident on it.\\
-  For $\Pi_0$ all invariants hold because it is output from algorithm
-  \ref{perms} which is an ICPPA. Hence base case is proved.  Assume
-  the lemma holds for $\Pi_{j-1}$. Consider $\Pi_j$ and any $(P,Q) \in
-  \Pi_j$. If $(P,Q) $ is in $ \Pi_j$ and $\Pi_{j-1}$ invariants I and
-  II are true because of induction assumption. If it is only in
-  $\Pi_j$, then it is $\{(P' \setminus \{x\}),(Q' \setminus \{v\})$ or
-  $(\{x\},\{v\})$ for some $(P',Q')$ in $\Pi_{j-1}$. By definition,
-  $x$ is an element in $P'$ (as defined in the algorithm) and $v$ is a
-  leaf in $Q'$. If $(P,Q)$ is $\{(P' \setminus \{x\}),(Q' \setminus
-  \{v\})$, $Q$ is a path since only a leaf is removed from path
-  $Q'$. We know $|P'| = |Q'|$, therefore $|P' \setminus \{x\}| = |Q'
-  \setminus \{v\}|$. Hence in this case invariants I and II are
-  obvious. It is easy to see these invariants hold if $(P,Q)$ is
-  $(\{x\},\{v\})$.
-
-  For invariant III consider $(P_1,Q_1),(P_2,Q_2)$ in $\Pi_j$. If both
-  of them are also in $\Pi_{j-1}$, claim is proved. If one of them is
-  not in $\Pi_{j-1}$ then it has to be $\{(P' \setminus \{x\}),(Q'
-  \setminus \{v\})$ or $(\{x\},\{v\})$ for some $(P',Q')$ in
-  $\Pi_{j-1}$. Since by definition, $Q'$ is the only path with $v$ and
-  $P'$ the only set with $x$ in the previous iteration, $|P_1 \cap (P'
-  \setminus \{x\})| = |P_1 \cap P'|$ and $|Q_1 \cap (Q' \setminus
-  \{v\})| = |Q_1 \cap Q'|$ and $|P_1 \cap \{x\}| = 0, Q_1 \cap \{v\} =
-  0$. Thus invariant III is also proven.
-
-  \noindent
-  To prove invariant IV, consider $(P_1,Q_1),(P_2,Q_2), (P_3,Q_3)$ in
-  $\Pi_j$. If exactly one of them, say $P_3 \notin \Pi_{j-1}$, it is
-  one of the new sets. By the same argument used to prove invariant
-  III, $|P_1 \cap P_2 \cap (P' \setminus \{x\})| = |P_1 \cap P_2 \cap
-  P'|$ and $|Q_1 \cap Q_2 \cap (Q' \setminus \{x\})| = |Q_1 \cap Q_2
-  \cap Q'|$. Since $P_1, P_2, P'$ are all in $\Pi_{j-1}$, by induction
-  hypothesis $|P_1 \cap P_2 \cap P'| = |Q_1 \cap Q_2 \cap Q'|$. Also
-  $|P_1 \cap P_2 \cap \{x\}| = 0, Q_1 \cap Q_2 \cap \{v\} = 0$.  If
-  two or more of them are not in $\Pi_{j-1}$, then it can be verified
-  that $|P_1 \cap P_2 \cap P_3| = |Q_1 \cap Q_2 \cap Q_3|$ since the
-  new sets in $\Pi_j$ are either disjoint or as follows: assuming
-  $P_1, P_2 \notin \Pi_{j-1}$ and new sets are derived from $(P', Q'),
-  (P'', Q'') \in \Pi_{j-1}$ with $x_1, x_2$ exclusively in $P_1, P_2$,
-  $(\{x_1\},\{v_1\}), (\{x_2\},\{v_2\}) \in \Pi_j $ thus $v_1, v_2$
-  are exclusively in $Q_1, Q_2$ resp. it follows that $|P_1 \cap P_2
-  \cap P_3| = |(P' \setminus \{x_1\}) \cap (P'' \setminus \{x_2\})
-  \cap P_3| = |P' \cap P'' \cap P_3| = |Q' \cap Q'' \cap Q_3| = |(Q'
-  \setminus \{v_1\} \cap Q'' \setminus \{v_2\} \cap Q_3| = |Q_1 \cap
-  Q_2 \cap Q_3|$. Thus invariant IV is also proven.  \qed
-\end{proof}
-
-\noindent
-Using algorithms \ref{perms} and \ref{leafasgn} we prove the following
-theorem.
-
-\begin{theorem}
-  \label{th:perm}
-  If $\cA$ is an ICPPA, then there exists a bijection $\sigma : U
-  \rightarrow V(T)$ such that $\sigma(S_i) = P_i$ for all $i \in I$
-\end{theorem}
-\begin{proof}
-  This is a contructive proof. First, the given ICPPA $\cA$ and tree
-  $T$ are given as input to Algorithm \ref{perms}. This yields a
-  ``filtered'' ICPPA as the output which is input to Algorithm
-  \ref{leafasgn}.  It can be observed that the output of Algorithm
-  \ref{leafasgn} is a set of interval assignments to sets and
-  one-to-one assignment of elements of $U$ to each leaf of $T$. To be
-  precise, it would be of the form $\cB_0 = \cA_0 \cup \cL_0$. The
-  leaf assignments are defined in $\cL_0 = \{ (x_i,v_i) \mid x_i \in
-  U, v_i \in T, x_i \ne x_j, v_i \ne v_j, i \ne j, i,j \in [k] \}$
-  where $k$ is the number of leaves in $T$. The path assignments are
-  defined in $\cA_0 \subseteq \{(S_i',P_i') \mid S_i' \subseteq U_0,
-  P_i' \text{ is a path from } T_0\}$ where $T_0$ is the tree obtained
-  by removing all the leaves in $T$ and $U_0 = U \setminus \{ x \mid x
-  \text{ is assigned to a leaf in }\cL_0 \}$. Now we have a subproblem
-  of finding the permutation for the path assignment $\cA_0$ which has
-  paths from tree $T_0$ and sets from universe $U_0$. Now we repeat
-  the procedure and the path assignment $\cA_0$ and tree $T_0$ is
-  given as input to Algorithm \ref{perms}. The output of this
-  algorithm is given to Algorithm \ref{leafasgn} to get a new union of
-  path and leaf assignments $\cB_1 = \cA_1 \cup \cL_1$ defined similar
-  to $\cB_0, \cL_0, \cA_0$. In general, the two algorithms are run on
-  path assignment $\cA_{i-1}$ with paths from tree $T_{i-1}$ to get a
-  new subproblem with path assignment $\cA_i$ and tree $T_{i}$. $T_i$
-  is the subtree of $T_{i-1}$ obtained by removing all its
-  leaves. More importantly, it gives leaf assignments $\cL_{i}$ to the
-  leaves in tree $T_{i-1}$. This is continued until we get a
-  subproblem with path assignment $\cA_{d-1}$ and tree $T_{d-1}$ for
-  some $d \le n$ which is just a path. From the last lemma we know
-  that $\cA_{d-1}$ is an ICPPA. Another observation is that an ICPPA
-  with all its tree paths being intervals (subpaths from a path) is
-  nothing but an ICPIA\cite{nsnrs09}.  Let $\cA_{d-1}$ be equal to
-  $\{(S_i'',P_i'') \mid S_i'' \subseteq U_{d-1}, P_i'' \text{ is a
-    path from } T_{d-1} \}$. It is true that the paths $P_i''$s may
-  not be precisely an interval in the sense of consecutive integers
-  because they are some nodes from a tree. However, it is easy to see
-  that the nodes of $T_{d-1}$ can be ordered from left to right and
-  ranked to get intervals $I_i$ for every path $P_i''$ as
-  follows. $I_i = \{[l,r] \mid l = \text{ the lowest rank of the nodes
-    in }P_i'', r = l+|P_i''|-1 \}$. Let asssignment $\cA_d$ be with
-  the renamed paths. $\cA_d = \{ (S_i'', I_i) \mid (S_i'', P_i'') \in
-  \cA_{d-1} \}$. What has been effectively done is renaming the nodes
-  in $T_{d-1}$ to get a tree $T_d$.  The ICPIA $\cA_d$ is now in the
-  format that the ICPIA algorithm requires which gives us the
-  permutation $\sigma' : U_{d-1} \rightarrow T_{d-1}$
-
-\noindent
-$\sigma'$ along with all the leaf assignments $\cL_i$ gives us the
-permutation for the original path assignment $\cA$.  More precisely,
-the permutation for tree path assignment $\cA$ is defined as
-follows. $\sigma: U \rightarrow T$ such that the following is
-maintained.
-\begin{align*}
-  \sigma(x) &= \sigma'(x),   \text{ if } x \in U_{d-1} \\
-  &= \cL_i(x), \text{ where $x$ is assigned to a leaf in a subproblem
-    $\cA_{i-1}, T_{i-1}$}
-\end{align*}
-
-\noindent
-To summarize, run algorithm \ref{perms} and \ref{leafasgn} on
-$T$. After the leaves have been assigned to specific elements from
-$U$, remove all leaves from $T$ to get new tree $T_0$. The leaf
-assignments are in $\cL_0$. Since only leaves were removed $T_0$ is
-indeed a tree. Repeat the algorithms on $T_0$ to get leaf assignments
-$\cL_{1}$. Remove the leaves in $T_0$ to get $T_1$ and so on until the
-pruned tree $T_d$ is a single path. Now run ICPIA algorithm on $T_d$
-to get permutation $\sigma'$. The relation $\cL_0 \cup \cL_1 \cup
-.. \cup \cL_{d} \cup \sigma'$ gives the bijection required in the
-original problem.\qed
-\end{proof}
-
-\subsection{Finding an assignment of tree paths to a set
-  system} \label{decompos} In the previous section we have shown that
-the problem of finding a Tree Path Asssignment to an input $(\cF,T)$
-is equivalent to finding an ICPPA to $\cF$ in tree $T$.  In this
-section we characterize those set systems that have an ICPPA in a
-given tree.  As a consequence of this characterization we identify two
-sub-problems that must be solved to obtain an ICPPA.  We do not solve
-the problem and in the next section show that finding an ICPPA in a
-given tree is GI-Complete.
-
-\noindent
-A set system can be concisely represented by a binary matrix where the
-row indices denote the universe of the set system and the column
-indices denote each of the sets. Let the binary matrix be $M$ with
-order $n \times m$, the set system be $\cF = \{S_i \mid i \in [m]\}$,
-universe of set system $U = \{x_1, \dots ,x_n\}$. If $M$ represents
-$\cF$, $|U| = n, |\cF| = m$. Thus $(i,j)$th element of $M$, $M_{ij} =
-1$ iff $x_i \in S_j$. If $\cF$ has a feasible tree path assignment
-(ICPPA) $\cA = \{(S_i,P_i) \mid i \in [m]\}$, then we say its
-corresponding matrix $M$ has an ICPPA. Conversly we say that a matrix
-$M$ has an ICPPA if there exists an ICPPA $\cA$ as defined
-above.\\
-\noindent
-We now define the strict intersection graph or overlap graph of
-$\cF$. This graph occurs at many places in the literature, see for
-example \cite{kklv10, wlh02, nsnrs09}.  The vertices of the graph
-correspond to the sets in $\cF$.  An edge is present between vertices
-of two sets iff the corresponding sets have a nonempty intersection
-and none is contained in the other. Formally, intersection graph is
-$G_f = (V_f, E_f)$ such that $V_f = \{v_i \mid S_i \in \cF\}$ and $E_f
-= \{(v_i, v_j) \mid S_i \cap S_j \ne \O \text{ and }S_i \nsubseteq
-S_j, S_j \nsubseteq S_i \}$.  We use this graph to decompose $M$ as
-described in \cite{wlh02,nsnrs09}.  A prime sub-matrix of $M$ is
-defined as the matrix formed by a set of columns of $M$ which
-correspond to a connected component of the graph $G_f$.  Let us denote
-the prime sub-matrices by $M_1,\ldots,M_p$ each corresponding to one
-of the $p$ components of $G_f$. Clearly, two distinct matrices have a
-distinct set of columns.  Let $col(M_i)$ be the set of columns in the
-sub-matrix $M_i$.  The support of a prime sub-matrix $M_i$ is defined
-as $supp(M_i) = \displaystyle \bigcup_{j \in col(M_i)}S_j$.  Note that
-for each $i$, $supp(M_i) \subseteq U$.  For a set of prime
-sub-matrices $X$ we define
-$supp(X) = \displaystyle \bigcup_{M \in X} supp(M)$. \\
-
-\noindent
-Consider the relation $\preccurlyeq$ on the prime sub-matrices $M_1,
-\ldots, M_p$ defined as follows:
-\begin{equation}
-  \nonumber \{(M_i,M_j) | \mbox{ A set } S \in
-  M_i \mbox{ is contained in a set } S' \in M_j\} \cup \{(M_i,M_i) | 1
-  \leq i \leq p\} 
-\end{equation}
-
-\noindent
-This relation is the same as that defined in \cite{nsnrs09}. The prime
-submatrices and the above relation can be defined for any set
-system. We will use this structure of prime submatrices to present our
-results on an an ICPPA for a set system $\cF$. Recall the following
-lemmas, and theorem that $\preccurlyeq$ is a partial order, from
-\cite{nsnrs09}.
-
-\begin{lemma} \label{lem:containment} Let $(M_i,M_j) \in
-  \preccurlyeq$.  Then there is a set $S' \in M_j$ such that for each
-  $S \in M_i$, $S \subseteq S'$.
-\end{lemma}
-\begin{lemma}
-  For each pair of prime sub-matrices, either $(M_i,M_j) \not\in
-  \preccurlyeq$ or $(M_j,M_i) \not\in \preccurlyeq$.
-  % If $(M_i,M_j) \in \preccurlyeq$ and $(M_j,M_i) \in \preccurlyeq$,
-  % then $i = j$ and $|M_i| = 1$.
-\end{lemma}
-\begin{lemma}
-  If $(M_i,M_j) \in \preccurlyeq $ and $(M_j,M_k) \in \preccurlyeq$,
-  then $(M_i,M_k) \in \preccurlyeq$.
-\end{lemma}
-\begin{lemma}
-  If $(M_i,M_j) \in \preccurlyeq$ and $(M_i,M_k) \in \preccurlyeq$,
-  then either $(M_j,M_k) \in \preccurlyeq$ or $(M_k,M_j) \in
-  \preccurlyeq$.
-\end{lemma}
-\begin{theorem} \label{thm:partitionold} $\preccurlyeq$ is a partial
-  order on the set of prime sub-matrices of $M$.  Further, it uniquely
-  partitions the prime sub-matrices of $M$ such that on each set in
-  the partition $\preccurlyeq$ induces a total order.
-\end{theorem}
-For the purposes of this paper, we refine the total order mentioned in
-Theorem \ref{thm:partitionold}. We do this by identifying an in-tree
-rooted at each maximal upper bound under $\preccurlyeq$.  Each of
-these in-trees will be on disjoint vertex sets, which in this case
-would be disjoint sets of prime-submatrices.  The in-trees are
-specified by selecting the appropriate edges from the Hasse diagram
-associated with $\preccurlyeq$.  Let $\cI$ be the following set:
-\begin{align*}
-  \cI = \{ (M_i,M_j) \in \preccurlyeq \mid \nexists M_k s.t. M_i
-  \preccurlyeq M_k, M_k \preccurlyeq M_j \} \cup \{ (M_i,M_i), i \in
-  [p] \}
-\end{align*}
-
-\begin{theorem} \label{thm:partition} Consider the directed graph $X$
-  whose vertices correspond to the prime sub-matrices, and the edges
-  are given by $\cI$.  Then, $X$ is a vertex disjoint collection of
-  in-trees and the root of each in-tree is a maximal upper bound in
-  $\preccurlyeq$.
-\end{theorem}
-\begin{proof}
-  To observe that $X$ is a collection of in-trees, we observe that for
-  vertices corresponding to maximal upper bounds, no out-going edge is
-  present in $I$.  Secondly, for each other element, exactly one
-  out-going edge is chosen, and for the minimal lower bound, there is
-  no in-coming edge.  Consequently, $X$ is acyclic, and since each
-  vertex has at most one edge leaving it, it follows that $X$ is a
-  collection of in-trees, and for each in-tree, the root is a maximal
-  upper bound in $\preccurlyeq$.  Hence the theorem.
-\end{proof}
-Let the partition of $X$ given by Theorem \ref{thm:partition} be
-$\{X_1,\ldots,X_r\}$.  Further, each in-tree itself can be layered
-based on the distance from the root.  The root is considered to be at
-level zero. For $j \geq 0$, Let $X_{i,j}$ denote the set of prime
-matrices in level $j$ of in-tree $X_i$.
-
-\begin{lemma}
-  \label{lem:subicppa}
-  Let $M$ be a matrix and let $X$ be the directed graph whose vertices
-  are in correspondence with the prime submatrices of $M$.  Further
-  let $\{X_1,\ldots,X_r\}$ be the partition of $X$ into in-trees as
-  defined above.  Then, matrix $M$ has an ICPPA in tree $T$ iff $T$
-  can be partitioned into vertex disjoint subtrees $\{T_1, T_2, \dots
-  T_r\}$ such that, for each $1 \leq i \leq r$, the set of prime
-  sub-matrices corresponding to vertices in $X_i$ has an ICPPA in
-  $T_i$.
-\end{lemma}
-\begin{proof}
-  Let us consider the reverse direction first.  Let us assume that $T$
-  can be partitioned into $T_1, \ldots, T_r$ such that for each $1
-  \leq i \leq r $, the set of prime sub-matrices corresponding to
-  vertices in $X_i$ has an ICPPA in $T_i$.  It is clear from the
-  properties of the partial order $\preccurlyeq$ that these ICPPAs
-  naturally yield an ICPPA of $M$ in $T$.  The main property used in
-  this inference is that for each $1 \leq i \neq j \leq r$, $supp(X_i)
-  \cap supp(X_j) = \phi$.
-
-\noindent
-To prove the forward direction, we show that if $M$ has an ICPPA, say
-$\cA$, in $T$, then there exists a partition of $T$ into vertex
-disjoint subtree $T_1, \ldots, T_r$ such that for each $1 \leq i \leq
-r$, the set of prime sub-matrices corresponding to vertices in $X_i$
-has an ICPPA in $T_i$.  For each $1 \leq i \leq r$, we define based on
-$\cA$ a subtree $T_i$ corresponding to $X_i$.  We then argue that the
-trees thus defined are vertex disjoint, and complete the proof.
-Consider $X_i$ and consider the prime sub-matrix in $X_{i,0}$.
-Consider the paths assigned under $\cA$ to the sets in the prime
-sub-matrix in $X_{i,0}$.  Since the component in $G_f$ corresponding
-to this matrix is a connected component, it follows that union of
-paths assigned to this prime-submatrix is a subtree of $T$.  We call
-this sub-tree $T_i$.  All other prime-submatrices in $X_i$ are
-assigned paths in $T_i$ since $\cA$ is an ICPPA, and the support of
-other prime sub-matrices in $X_i$ are contained in the support of the
-matrix in $X_{i,0}$.  Secondly, for each $1 \leq i \neq j \leq r$,
-$supp(X_i) \cap supp(X_j) = \phi$, and since $\cA$ is an ICPPA, it
-follows that $T_i$ and $T_j$ are vertex disjoint.  Finally, since $|U|
-= |V(T)|$, it follows that $T_1, \ldots, T_r$ is a partition of $T$
-into vertex disjoint sub-trees such that for each $1 \leq i \leq r$,
-the set of matrices corresponding to nodes in $X_i$ has an ICPPA in
-$T_i$.  Hence the lemma.
-\end{proof}
-The essence of the following lemma is that an ICPPA only needs to be
-assigned to the prime sub-matrix corresponding to the root of each
-in-tree, and all the other prime sub-matrices only need to have an
-Intersection Cardinality Preserving Interval Assignments (ICPIA).
-Recall, an ICPIA is an assignment of intervals to sets such that the
-cardinality of an assigned interval is same as the cardinality of the
-interval, and the cardinality of intersection of any two sets is same
-as the cardinality of the intersection of the corresponding intervals.
-It is shown in \cite{nsnrs09} that the existence of an ICPIA is a
-necessary and sufficient condition for a matrix to have COP.  We
-present the pseudo-code to test if $M$ has an ICPPA in $T$.
-\begin{lemma} \label{lem:rooticppa} Let $M$ be a matrix and let $X$ be
-  the directed graph whose vertices are in correspondence with the
-  prime submatrices of $M$.  Further let $\{X_1,\ldots,X_r\}$ be the
-  partition of $X$ into in-trees as defined earlier in this section.
-  Let $T$ be the given tree and let $\{T_1, \ldots, T_r\}$ be a given
-  partition of $T$ into vertex disjoint sub-trees.  Then, for each $1
-  \leq i \leq r$, the set of matrices corresponding to vertices of
-  $X_i$ has an ICPPA in $T_i$ if and only if the matrix in $X_{i,0}$
-  has an ICPPA in $T_i$ and all other matrices in $X_i$ have an {\bf
-    {\em ICPIA}}.
-\end{lemma}
-\begin{proof}
-  The proof is based on the following fact- $\preccurlyeq$ is a
-  partial order and $X$ is a digraph which is the disjoint union of
-  in-trees.  Each edge in the in-tree is a containment relationship
-  among the supports of the corresonding sub-matrices. Therefore, any
-  ICPPA to a prime sub-matrix that is not the root is contained in a
-  path assigned to the sets in the parent matrix.  Consequently, any
-  ICPPA to the prime sub-matrix that is not at the root is an ICPIA,
-  and any ICPIA can be used to construct an ICPPA to the matrices
-  corresponding to nodes in $X_i$ provided the matrix in the root has
-  an ICPPA in $T_i$.  Hence the lemma.
-\end{proof}
-Lemma \ref{lem:subicppa} and Lemma \ref{lem:rooticppa} point out two
-algorithmic challenges in finding an ICPPA for a given set system
-$\cF$ in a tree $T$.  Given $\cF$, finding $X$ and its partition
-$\{X_1,\ldots,X_r\}$ into in-trees can be done in polynomial time.  On
-the other hand, as per lemma \ref{lem:subicppa} we need to parition
-$T$ into vertex disjoint sub-trees $\{T_1, \ldots, T_r\}$ such that
-for each $i$, the set of matrices corresponding to nodes in $X_i$ have
-an ICPPA in $T_i$.  This seems to be a challenging step, and it must
-be remarked that this step is easy when $T$ itself is a path, as each
-individual $T_i$ would be sub-paths.  The second algorithmic challenge
-is identified by lemma \ref{lem:rooticppa} which is to assign an ICPPA
-from a given tree to the matrix associated with the root node of
-$X_i$.
-\begin{algorithm}[h]
-  \caption{Algorithm to find an ICPPA for a matrix $M$ on tree $T$:
-    $main\_ICPPA(M, T$)}
-  \label{al:icppa-main}
-  \begin{algorithmic}
-    \STATE Identify the prime sub-matrices. This is done by constructing the strict overlap graph and identify connected components.  Each connected component yields a prime sub-matrix.   \\
-    \STATE Construct the partial order $\preccurlyeq$ on the set of prime sub-matrices.  \\
-    \STATE Construct the partition $X_1,\ldots,X_r$ of the prime
-    sub-matrices induced by $\preccurlyeq$ \\
-    \STATE For each $1 \leq i \leq r$, Check if all matrices except
-    those in $X_{i,0}$ has an ICPIA.  If a matrix does not have ICPIA
-    exit with a negative answer.  To check for the existence of ICPIA,
-    use the result in \cite{nsnrs09}.  \STATE Find a partition of
-    $T_1, \ldots, T_r$ such that matrices in $X_{i,0}$ has an ICPPA in
-    $T_i$.  If not such partition exists, exit with negative answer.
-  \end{algorithmic}
-\end{algorithm}
-
-\subsection{Complexity of Tree Path Assignment-A
-  Discussion} 
-\label{complexity} 
-Recall that the input to the Tree
-Path Assignment question is an order pair $(\cF,T)$ where $\cF$ is a
-family of subsets of an universe $U$, and $T$ is a tree such that
-$|V(T)| = |U|$.  The question is to come up with a bijection from $U$
-to $V(T)$ such that the image of each set in $\cF$ is a path in $T$.
-We show that this problem is at least as hard as the problem of
-testing if two given path chordal graphs are isomorphic.
-\begin{theorem} \label{thm:isocomp} Tree Path Assignment is
-  isomorphism-complete.
-\end{theorem}
-\begin{proof}
-  It is well known (see for example \cite{kklv10}) that testing
-  isomorphism of path chordal graphs is isomorphism complete.  We show
-  a reduction of path chordal graph isomorphism to tree path
-  assignment.  Given $G_1$ and $G_2$ two path chordal graphs, let
-  $T_2$ be the clique tree of $G_2$ obtained from say \cite{gav78}.
-  The nodes of $T_2$ correspond to the maximal cliques of $G_2$ and
-  each vertex of $G_2$ corresponds to a path in $G_2$.  This is a
-  well-known characterization of path chordal graphs and $T_2$ can be
-  computed in polynomial time.  In $G_1$, let $S_v$ denote the maximal
-  cliques of $G_1$ that contain $v$.  This can be computed in
-  polynomial time as $G_1$ is path chordal, and all chordal graph only
-  have a linear number of maximal cliques.  The universe $U$
-  corresponds bijectively to the set of maximal cliques in $G_1$, and
-  $\cF = \{S_v | v \in V(G_1)\}$.  Now, we claim that $(\cF, T_2)$ has
-  a tree path assignment if and only if $G_1$ and $G_2$ are
-  isomorphic.  This is clear since for each vertex $v \in G_1$, there
-  is an associated $S_v$ which is the set of maximal cliques
-  containing $v$.  In $G_2$, each vertex corresponds to a path in
-  $T_2$, and the nodes on this path corresponds to the maximal cliques
-  in $G_2$.  Consequenlty, a tree path assignment will naturally yield
-  an isomorphism between $G_1$ and $G_2$, and vice versa.  Therefore,
-  Tree Path Assignment is isomorphism-complete.
-\end{proof}
-
-\subsubsection{Consecutive Ones Testing is in Logspace}
-While Tree Path Assignment is isomorphism-complete, it is polynomial
-time solvable when the given tree is a path.  Indeed, in this case we
-encounter a restatement of matrices with the COP.  The known
-approaches to testing for COP fall into two categories: those that
-provide a witness when the input matrix does not have COP, and those
-that do not provide a witness.  The first linear time algorithm for
-testing COP for a binary matrix was using a data structure called PQ
-trees, which represent all COP orderings of $M$, invented by
-\cite{bl76}. There is a PQ tree for a matrix iff the matrix has COP.
-Indeed, this is an algorithmic characterization of the consecutive
-ones property and the absence of the PQ-tree does not yield any
-witness to the reason for failure.  A closely related data structure
-is the generalized PQ tree in \cite{mcc04}.  In generalized PQ tree
-the P and Q nodes are called prime and linear nodes. Aside from that,
-it has a third type of node called degenerate nodes which is present
-only if the set system does not have COP \cite{mcc04}.  Using the idea
-of generalized PQ tree, \cite{mcc04} proves that checking for
-bipartiteness in the certain incomparability graph is sufficient to
-check for COP.  \cite{mcc04} invented a certificate to confirm when a
-binary matrix does not have COP.  \cite{mcc04} describes a graph
-called incompatibility graph of a set system $\cF$ which has vertices
-$(a,b), a \ne b$ for every $a, b \in U$, $U$ being the universe of the
-set system. There are edges $((a,b),(b,c))$ and $((b,a),(c,b))$ if
-there is a set $S \in \cF$ such that $a, c \in S$ and $b \notin S$. In
-other words the vertices of an edge in this graph represents two
-orderings that cannot occur in a consecutive ones ordering of $\cF$.
-\begin{theorem}[Theorem 6.1, \cite{mcc04}]
-  Let $\cF$ be an arbitrary set family on domain $V$. Then $\cF$ has
-  the consecutive ones property if and only if its incompatibility
-  graph is bipartite, and if it does not have the consecutive ones
-  property, the incompatibility graph has an odd cycle of length at
-  most $n+3$.
-\end{theorem}
-This theorem gives a certificate as to why a given matrix does not
-have COP.  Similarly, the approach of testing for an ICPIA in
-\cite{nsnrs09} also gives a different certificate- a prime sub-matrix
-that does not have an ICPIA.  Further, the above theorem can be used
-to check if a given matrix has COP in logspace by checking if its
-incompatibility graph is bipartite. \cite{rei84} showed that checking
-for bipartiteness can be done in logspace. Thus we conclude that
-consecutive ones testing can be done in logspace.
-
-\noindent
-More recently, \cite{kklv10} showed that interval graph isomorphism
-can be done in logspace. Their paper proves that a canon for interval
-graphs can be calculated in logspace using an interval hypergraph
-representation of the interval graph with each hyperedge being a set
-to which an interval shall be assigned by the canonization
-algorithm. An overlap graph (subgraph of intersection graph, edges
-define only strict intersections and no containment) of the hyperedges
-of the hypergraph is created and canons are computed for each overlap
-component. The overlap components define a tree like relation due to
-the fact that two overlap components are such that either all the
-hyperedges of one is disjoint from all in the other, or all of them
-are contained in one hyperedge in the other. This is similar to the
-containment tree defined in \cite{nsnrs09} and in this paper. Finally
-the canon for the whole graph is created using logspace tree
-canonization algorithm from \cite{sl92}. The interval labelling done
-in this process of canonization is exactly the same as the problem of
-assigning feasible intervals to a set system, and thus the problem of
-finding a COP ordering in a binary matrix \cite{nsnrs09}.
-
-\begin{theorem}[Theorem 4.7, \cite{kklv10}]
-  \label{th:canonlabel}
-  Given an interval hypergraph $\cH$, a canonical interval labeling
-  $l_H$ for $H$ can be computed in FL.
-\end{theorem}
-
-We present the following reduction to see that COP testing is indeed
-in logspace. Given a binary matrix $M$ of order $n \times m$, let $S_i
-= \{j \mid M[j,i]=1 \}$. Let $\cF = \{S_i \mid i \in [m] \}$ be this
-set system. Construct a hypergraph $\cH$ with its vertex set being
-$\{1, 2, \dots n\}$. The edge set of $\cH$ is isomorphic to
-$\cF$. Thus every edge in $\cH$ represents a set in the given set
-system $\cF$. Let this mapping be $\pi: E(\cH) \rightarrow \cF$. It is
-easy to see that if $M$ has COP, then $\cH$ is an interval
-hypergraph. From theorem \ref{th:canonlabel}, it is clear that the
-interval labeling $l_{\cH}: V(\cH) \rightarrow [n]$ can be calculated
-in logspace. Construct sets $I_i = \{ \l_{\cH}(x) \mid x \in E, E \in
-E(\cH), \pi(E) = S_i\}$, for all $i \in [m]$. Since $\cH$ is an
-interval hypergraph, $I_i$ is an interval for all $i \in [m]$, and is
-the interval assigned to $S_i$ if $M$ has COP.
-
-Now we have the following corollary.
-\begin{corollary}
-  \label{cor:coplog}
-  If a binary matrix $M$ has COP then the interval assignments to each
-  of its columns can be calculated in FL.
-\end{corollary}
-
-\noindent
-Finally, we conclude by asking about the complexity of Tree Path
-Assignment restricted to other subclasses of trees.  In particular, is
-Tree Path Assignment in caterpillars easier than Tree Path assignment
-in general trees.
-
-\chapter{Conclusion}
-\label{ch:conclusion}
-
-\begin{center}
-{\tt MOVED FROM INTRO SUBSECTION IN ch:myresearch CHAPTER begin}
-\end{center}
-
-% \annote{
-  It is an interesting fact that for a matrix with the COP, the
-  intersection graph of the corresponding set system is an interval
-  graph. % }{the following text could be in Further Study/Conclusion}
-A
-similar connection to a subclass of chordal graphs and a superclass of
-interval graphs exists for the generalization of COP.  In this case,
-the intersection graph of the corresponding set system must be a {\em
-  path graph}. Chordal graphs are of great significance, are
-extensively studied, and have several applications.  One of the well
-known and interesting properties of a chordal graphs is its connection
-with intersection graphs \cite{mcg04}. For every chordal graph, there
-exists a tree and a family of subtrees of this tree such that the
-intersection graph of this family is isomorphic to the chordal graph
-\cite{plr70,gav78,bp93}.  These trees when in a certain format, are
-called {\em clique trees} \cite{apy92} of the chordal graph. This is a
-compact representation of the chordal graph. There has also been work
-done on the generalization of clique trees to clique hypergraphs
-\cite{km02}.  If the chordal graph can be represented as the
-intersection graph of paths in a tree, then the graph is called path
-graph \cite{mcg04}.  Therefore, it is clear that if there is a
-bijection from $U$ to $V(T)$ such that for every set, the elements in
-it map to vertices of a unique path in $T$, then the intersection
-graph of the set system is indeed a path graph.  However, this is only
-a necessary condition and can be checked efficiently because path
-graph recognition is polynomial time
-solvable\cite{gav78,aas93}. Indeed, it is possible to construct a set
-system and tree, such that the intersection graph is a path graph, but
-there is no bijection between $U$ and $V(T)$ such that the sets map to
-paths. Path graph isomorphism is known be isomorphism-complete, see
-for example \cite{kklv10}. An interesting area of research would be to
-see what this result tells us about the complexity of the tree path
-labeling problem (not covered in this paper).
-
-\begin{center}
-{\tt MOVED FROM INTRO SUBSECTION IN ch:myresearch CHAPTER\\ end}
-\end{center}
-
-We give a characterization for feasible tree path labeling of path
-hypergraphs. The proof for this is constructive and computes a
-feasibility bijection mapping vertices of the hypergraph to vertices
-of the given target tree. This thesis also discovered an exponential algorithm
-that computes a feasible TPL to a given hypergraph if it is a path hypergraph.
-
-Our results have close connections to recognition of path graphs and
-to path graph isomorphism.  Graphs which can be represented as the
-intersection graph of paths in a tree are called {\em path
-  graphs}\cite{mcg04}. Thus, a hypergraph $\cF$ can be interpreted as
-paths in a tree, if and only if the intersection graph of $\cF$ is a
-{\em path graph}. Path graphs are a subclass of chordal graphs since
-chordal graphs are characterized as the intersection graphs of
-subtrees of a tree\cite{mcg04}.  Path graphs are well studied in the
-literature \cite{plr70}--\cite{mcg04}\tnote[TH]{Check if cite:bp93 is
-  indeed for path graphs}.  Path graph recognition can be done in
-polynomial time\cite{gav78,aas93}.  Clearly, this is a necessary
-condition in terms of the intersection graph of the input hypergraph
-$\cF$ in \FTPL. However, one can easily obtain a counterexample to
-show the insufficiency of this condition\tnote[TH]{Give a figure
-  showing insuff.}.  Path graph isomorphism is known be
-isomorphism-complete\cite{kklv10}. Therefore, \annote[TH]{it is
-  unlikely that we can solve the problem of finding feasible path
-  labeling $\cl$ for a given $\cF$ and }{WHY?} tree $T$.  It is
-definitely interesting to classify the kinds of trees and hypergraphs
-for which feasible path labelings can be\tnote[TH]{ELABORATE.}  found
-efficiently.  \annote[TH]{These results would form a natural
-  generalization of COP testing and interval graph isomorphism,
-  culminating in Graph Isomorphism }{REWORD} itself.  To this effect
-we consider TPL on {\kstar} and give a polynomial time solution for
-the same.
-
-While we address the computation of TPL
-for a given hypergraph% \footnote{The term {\em hypergraph} is used
-%   instead {\em set systems} are used interchangably in this
-%   section. Hypergraphs that have TPLs on some target tree are called
-%   {\em path hypergraphs}.}
-from a given target tree in this research, however optimization
-problems in TPL, akin to problems in Section~\ref{sec:optcop} for COP,
-is outside the scope of this thesis. Whether TPL on general trees is
-solvable in $P$ remains open. So do optimization opportunities in TPL
-(possible extensions to optimization for COP in the
-Section~\ref{sec:optcop}.\tnote[TH]{link to section}).
-
-%\newpage %DUMMY
-\tnote[TH]{ ||Complexity challenges:|| Canonization is an important
-  tool in graph isomorphism. Invention of a
-% deterministic method of canonization for any class of graphs naturally
-% results in an algorithm for isomorphism. All that is required is to
-% check if the canons of two graphs are the same. Thus complexity of
-% graph isomorphism can be studied by studying canonization
-% methods. While general graph ismorphism remains elusive in terms of
-% hardness, canonization has been studied for smaller classes of graphs
-% thus giving complexity/hardness results for them.
-
-% In 1992, \cite{sl92} made an important discovery that tree isomorphism
-% is in logspace. It was done by inventing a method of canonization of
-% trees using a logspace depth first traversal
-% algorithm. \tnote{confirm the statement of claim}.
-
-% \cite{kklv10} proved that interval graph canonization is also in
-% logspace thus drawing logspace conslusions about COT. 
-% % This result is
-% % built on the fact that 
-% Interval graphs are FP+C (fixed point with
-% counting) definable \tnote{elucidate} and \cite{l10} showed that
-% this implies that it captures PTIME. \tnote{elucidate} This result
-% along with that of undirected graph reachability being in logspace
-% \cite{rei08}, \cite{kklv10} proved their logspace result.
-}   %
-\tnote[TH]{ (1) the interesting survey in conclusion section of kklv10
-  (2) [[[COP helping some problems with hardness]]] having COP in the
-  input structure makes some problems less hard than in a general
-  input. -- min set cover problem is known to be NP hard. so is min
-  hitting
-  set. these two problems are equivalent cite-adp80\\
-  -- -- complexity results: log factor polynomial time approximable cite-f98, W[2]-complete with solution size as parameter cite-df99\\
-  -- a couple of generalizations:\\
-  -- -- min degree hypergraph (MDH) problem\\
-  -- -- red blue covering (RBSC) problem\\
-  ----- cite-d08phd Chapter 5.\\
-  ----- what are cite-d08phd results\\
-  ----- Rectangle stabbing \\
-  ----- see cite-d08phd Chapter 6?  }
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Appendices.
-
-% \appendix %%% NON NEIN NAHI ILLA NONE
-
-% \chapter{A}
-% \tnote{DO WE NEED THIS?}
-
-\newpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Bibliography.
-\begin{singlespace}
-%\begin{thebibliography}{10}
-\addcontentsline{toc}{chapter}{REFERENCES}
-
-{\footnotesize
-  \bibliographystyle{alpha} 
-  \bibliography{../lib/cop-variants__thesis}
-}
-%\end{thebibliography}
-\end{singlespace}
-%\newpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% List of papers
-
-% \listofpapers %%% NON NEIN NAHI ILLA NONE
-
-% \begin{enumerate}  
-% \item Authors....  \newblock
-%  Title...
-%   \newblock {\em Journal}, Volume,
-%   Page, (year).
-% \end{enumerate}
-
-\end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% DOCUMENTATION ON THE THESIS CLASS FILE ETC.
-
-% Before reading any further please note that you are strongly advised
-% against changing any of the formatting options used in the class
-% provided in this directory, unless you are absolutely sure that it
-% does not violate the IITM formatting guidelines.  \emph{Please do not
-%   change the margins or the spacing.}  If you do change the formatting
-% you are on your own (don't blame me if you need to reprint your entire
-% thesis).  In the case that you do change the formatting despite these
-% warnings, the least I ask is that you do not redistribute your style
-% files to your friends (or enemies).
-
-% It is also a good idea to take a quick look at the formatting
-% guidelines.  Your office or advisor should have a copy.  If they
-% don't, pester them, they really should have the formatting guidelines
-% readily available somewhere.
-
-% To generate PDF's with the links from the \verb+hyperref+ package use
-% the following command:
-% \begin{verbatim}
-% % dvipdfm -o thesis.pdf thesis.dvi
-% \end{verbatim}
-
-% \section{Package Options}
-
-% Use this thesis as a basic template to format your thesis.  The
-% \verb+iitmdiss+ class can be used by simply using something like this:
-% \begin{verbatim}
-% \documentclass[PhD]{iitmdiss}  
-% \end{verbatim}
-
-% To change the title page for different degrees just change the option
-% from \verb+PhD+ to one of \verb+MS+, \verb+MTech+ or \verb+BTech+.
-% The dual degree pages are not supported yet but should be quite easy
-% to add.  The title page formatting really depends on how large or
-% small your thesis title is.  Consequently it might require some hand
-% tuning.  Edit your version of \verb+iitmdiss.cls+ suitably to do this.
-% I recommend that this be done once your title is final.
-
-% To write a synopsis simply use the \verb+synopsis.tex+ file as a
-% simple template.  The synopsis option turns this on and can be used as
-% shown below.
-% \begin{verbatim}
-% \documentclass[PhD,synopsis]{iitmdiss}                                
-% \end{verbatim}
-
-% Once again the title page may require some small amount of fine
-% tuning.  This is again easily done by editing the class file.
-
-% This sample file uses the \verb+hyperref+ package that makes all
-% labels and references clickable in both the generated DVI and PDF
-% files.  These are very useful when reading the document online and do
-% not affect the output when the files are printed.
-
-% \section{Example Figures and tables}
-
-% Fig.~\ref{fig:iitm} shows a simple figure for illustration along with
-% a long caption.  The formatting of the caption text is automatically
-% single spaced and indented.  Table~\ref{tab:sample} shows a sample
-% table with the caption placed correctly.  The caption for this should
-% always be placed before the table as shown in the example.
-
-% \begin{figure}[htpb]
-%   \begin{center}
-%     \resizebox{50mm}{!} {\includegraphics *{iitm.eps}}
-%     \resizebox{50mm}{!} {\includegraphics *{iitm.eps}}
-%     \caption {Two IITM logos in a row.  This is also an
-%       illustration of a very long figure caption that wraps around two
-%       two lines.  Notice that the caption is single-spaced.}
-%   \label{fig:iitm}
-%   \end{center}
-% \end{figure}
-
-% \begin{table}[htbp]
-%   \caption{A sample table with a table caption placed
-%     appropriately. This caption is also very long and is
-%     single-spaced.  Also notice how the text is aligned.}
-%   \begin{center}
-%   \begin{tabular}[c]{|c|r|} \hline
-%     $x$ & $x^2$ \\ \hline
-%     1  &  1   \\
-%     2  &  4  \\
-%     3  &  9  \\
-%     4  &  16  \\
-%     5  &  25  \\
-%     6  &  36  \\
-%     7  &  49  \\
-%     8  &  64  \\ \hline
-%   \end{tabular}
-%   \label{tab:sample}
-%   \end{center}
-% \end{table}
-
-% \section{Bibliography with BIB\TeX}
-
-% I strongly recommend that you use BIB\TeX\ to automatically generate
-% your bibliography.  It makes managing your references much easier.  It
-% is an excellent way to organize your references and reuse them.  You
-% can use one set of entries for your references and cite them in your
-% thesis, papers and reports.  If you haven't used it anytime before
-% please invest some time learning how to use it.
-
-% I've included a simple example BIB\TeX\ file along in this directory
-% called \verb+refs.bib+.  The \verb+iitmdiss.cls+ class package which
-% is used in this thesis and for the synopsis uses the \verb+natbib+
-% package to format the references along with a customized bibliography
-% style provided as the \verb+iitm.bst+ file in the directory containing
-% \verb+thesis.tex+.  Documentation for the \verb+natbib+ package should
-% be available in your distribution of \LaTeX.  Basically, to cite the
-% author along with the author name and year use \verb+\cite{key}+ where
-% \verb+key+ is the citation key for your bibliography entry.  You can
-% also use \verb+\citet{key}+ to get the same effect.  To make the
-% citation without the author name in the main text but inside the
-% parenthesis use \verb+\citep{key}+.  The following paragraph shows how
-% citations can be used in text effectively.
-
-% More information on BIB\TeX\ is available in the book by
-% \cite{lamport:86}.  There are many
-% references~\citep{lamport:86,prabhu:xx} that explain how to use
-% BIB\TeX.  Read the \verb+natbib+ package documentation for more
-% details on how to cite things differently.
-
-% Here are other references for example.  \citet{viz:mayavi} presents a
-% Python based visualization system called MayaVi in a conference paper.
-% \citet{pan:pr:flat-fst} illustrates a journal article with multiple
-% authors.  Python~\citep{py:python} is a programming language and is
-% cited here to show how to cite something that is best identified with
-% a URL.
-
-% \section{Other useful \LaTeX\ packages}
-
-% The following packages might be useful when writing your thesis.
-
-% \begin{itemize}  
-% \item It is very useful to include line numbers in your document.
-%   That way, it is very easy for people to suggest corrections to your
-%   text.  I recommend the use of the \texttt{lineno} package for this
-%   purpose.  This is not a standard package but can be obtained on the
-%   internet.  The directory containing this file should contain a
-%   lineno directory that includes the package along with documentation
-%   for it.
-
-% \item The \texttt{listings} package should be available with your
-%   distribution of \LaTeX.  This package is very useful when one needs
-%   to list source code or pseudo-code.
-
-% \item For special figure captions the \texttt{ccaption} package may be
-%   useful.  This is specially useful if one has a figure that spans
-%   more than two pages and you need to use the same figure number.
-
-% \item The notation page can be entered manually or automatically
-%   generated using the \texttt{nomencl} package.
-
-% \end{itemize}
-
-% More details on how to use these specific packages are available along
-% with the documentation of the respective packages.
-
-% END OF DOCUMENTATION
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\theendnotes

### 2012-01-03 src/020chap2.tex
Cleaned up thesis.tex. Split thesis.tex into chapterwise subfiles for easy edit tex management. Content changes are mainly in chapter 1 and 3 although nothing significant. To being working on chapter 2 after this commit.

--- 2012-01-03 src/010chap1.tex
@@
-\chapter{Introduction}
-\label{ch:intro}
+\chapter{Consecutive-ones Property - a Survey of Important Results}
+\label{ch:copsurvey}
+\tnote{[i] TBD: have a few lines about organization of chapter\ \ \ \
+  \ [ii] Survey chapter: is the full fledged expansion of the survey in
+  introduction with details, observations, theorems etc.}
 
-Consecutive-ones property is a non-trivial property of binary matrices
-that has been studied widely in the literature for over past 50
-years. Detection of COP in a matrix is possible efficiently and there
-are several algorithms that achieve the same. This thesis documents
-the work done on an extension of COP extended from the equivalent
-interval assignment problem in \cite{nsnrs09}. These new results
-rigorously prove a natural extension (to trees) of their
-characterization as well as makes connections to graph isomorphism,
-namely path graph isomorphism.
+\section*{-- REAL CONTENT --}
 
-\section{Organization of the document}
-\label{sec:orgofdoc}
-This chapter (Chapter~\ref{ch:intro}) introduces the area of research
-and the problems addressed in this thesis.  Section~\ref{sec:problem}
-introduces the main problem of this thesis by way of an
-illustration. Section~\ref{sec:basicprelim} lays out a few general
-definitions that are helpful in understanding the rest of the chapter.
-Section~\ref{sec:background} gives a brief survey of COP
-and optimization problems related to it followed by motivation for the
-thesis in Section~\ref{sec:motive}.  Section~\ref{sec:results}
-presents a summary of our results on the extension of COP namely, the
-tree path labeling problem.
+%%%%%%                                              %%%%%%
+% [pressing]{\ \ \ \ \  is the second level of [pressing]%
+%%%%%%                                              %%%%%%
+\tnote[pressing]{ADD: As it will be described in detail later in this
+  document, isomorphism of certain % classes of graphs, namely chordal
+  % graphs, have a close relationship with consecutive-ones property
+  % and generalizations of it.  This is perhaps because of how closely
+  % COP of a matrix relates to properties of graphs derived from
+  % matrices as seen inz the following results.
+} \tnote[pressing]{\ \ \ \ \ ADD: peo exists iff chordal. lexicographic BFS
+  [tag:chordalGraph]} %
+\tnote[pressing]{\ \ \ \ \ ADD: A well known result in %perfect graph theory is that
+  % the maximal cliques of an interval graph $G$ can be linearly
+  % ordered such that for all $v \in V(G)$, cliques containing $v$ are
+  % consecutive in the ordering cite~{gh64}. This clearly means that a
+  % graph $G$ is an interval graph if and only if
+}%
+\tnote[pressing]{\ \ \ \ \ verify from paper the statement of claim.}%
+\tnote[pressing]{maximal clique vertex incidence matrix
+  of% $G$ has COP.  Also, maximal cliques of any
+  % chordal graph can be enumerated in polytime $O(m+n)$
+} \tnote[pressing]{\ \ \ \ \ citation?!!}%
+\tnote[pressing]{cite~{fg65} uses these results to give the first polynomial
+  time algorithm for COT.}%
+\tnote[pressing]{\ \ \ \ \ check. how do they use it?}%
+\tnote[pressing]{ A bipartite graph is
+  convex% (on $R$) if and only if its half adjacency matrix has COP
+  % on rows.  The results in cite~{bl76} on COT are based on the
+  % result that interval graphs are AT-free chordal graphs.
+}%
+\tnote[pressing]{\ \ \ \ \ the latter being Tucker's?}%
+\tnote[pressing]{\ \ \ \ \ (2) TBD survey
+  -- % see blue notes (in notebook) under Graph Isomorphism. namely
+  % citations in cite:aas93 (3) a brief on heirarchy of chordal
+  % graphs, path graphs, interval graphs, peo, clique tree etc. and
+  % results. (4) interval graphs are incomparability graphs. see
+  % Golumbic. [tag:classification] (5) finding min length hole in
+  % bipartite graph is polynomial time [Sec 3.3.2 in cite:d08phd] (6)
+  % theorem 2.2 in cite:d08phd - $G$ is union of v.d. caterpillars iff
+  % $M$ has COP in rows. $M$ is edge vertex incidence matrix $(*,2)$
+  % of $G$
+}%
 
-Chapter~\ref{ch:copsurvey} gives a more detailed survey briefed in
-Section~\ref{sec:background}.  Chapter~\ref{ch:myresearch} details all
-the results obtained to the problems of this thesis and finally the
-conclusion of the thesis is discussed in Chapter~\ref{ch:conclusion}.
+\section{Matrices with COP}
+\label{sec:surveycoptest}
+\tnote{Expand on ref:sec:copmatrices}
 
-% \tnote{testing 1 2 3...}%
-% \tnote[bogus1]{testing 1 2 3...}%
-% \tnote[important]{testing 1 2 3...}% 
-% \tnote[dire]{testing 1 2 3...}% 
-% \tnote[minor]{testing 1 2 3...}% 
-% \tnote[pressing]{testing 1 2 3...}% 
-% \tnote[UNDEF]{testing 1 2 3...}%
+Figure~\ref{fig:cop-matrix} shows examples of consecutive-ones
+property. \tnote[important]{move to general def section? if so, decide
+how to cross ref with repetition for completeness of chapter.}
 
-\section{Illustration of the problem}
-\label{sec:problem}
+\begin{figure}[t] %[htbp]
+  \centering
 
-A group of students, \Pa, \Pig, \Sn, \Wo, \Vi, \Li, \Ch, \Sa, \Fr,
-\Sc\ and \Lu\ enroll at the {\WSI} for a liberal arts programme.  As
-part of their semester thesis, they pick a body of work to study and
-form the namesake study groups, {\LLL}, {\GGG}, {\BBB} and
-{\TTT}\tnote[minor]{put bib entries for these works!}. A student will be
-in at least one study group and may be in more than one. For instance,
-as will be seen later, {\Fr} studies both {\LLL} and {\TTT} while \Wo\
-studies only \BBB.
-
-Let $U$ and $\cF$ represent the set of students and the set of study
-groups respectively and the integers $n$ and $m$ denote the total
-number students and study groups respectively. In relation to this
-example, these are defined in Table~\ref{tab:wsigroups}. Also given
-there is the study group allocation to students.
-
-\begin{table}[t]%[htbp]
-  \centering
-  { \footnotesize
-    \begin{tabular}{rcl}
-      $U $&$=$&$ \{\xPa,\; \xPi,\; \xSn,\; \xWo,\; \xVi,\; \xLi,\; \xCh,\;
-      \xSa,\; \xFr,\; \xSc,\; \xLu\}$\\
-      $\cF $&$=$&$ \{\xLLL, \xGGG, \xBBB, \xTTT\}$\\
-      $\xLLL $&$=$&$ \{\xCh,\;  \xSa,\;  \xFr,\;  \xSc,\;  \xLu \}$\\
-      $\xGGG $&$=$&$ \{\xPa,\;  \xPi,\;  \xVi,\;  \xCh \}$\\
-      $\xBBB $&$=$&$ \{\xSn,\;  \xPi,\;  \xWo \}$\\
-      $\xTTT $&$=$&$ \{\xVi,\;  \xLi,\;  \xCh,\;  \xFr \}$\\
-      $n $&$=$&$ |U| = 11$\\
-      $m $&$=$&$ |\cF| = 4$      
+  {\figtabsize
+    \begin{tabular}[h]{l|lcccl}
+      $M_1$: & $M_1'$: &&&& $M_2$:\\
+      &&&&&\\
+      \begin{tabular}[h]{llll}
+        $c_1$ & $c_2$ &$c_3$ &$c_4$\\
+        &&&\\
+        \un & 0   & \un & 0\\
+        0   & \un & 0   & \un \\
+        \un & 0   & 0   & \un
+      \end{tabular}
+      &
+      \begin{tabular}[h]{llll}
+        $c_3$ &$c_1$ &$c_4$& $c_2$\\
+        &&&\\
+        \un & \un & 0 & 0\\
+        0 & 0 & \un & \un \\
+        0 & \un & \un & 0
+      \end{tabular}
+      &&&&
+      \begin{tabular}[h]{llll}
+        $d_1$ & $d_2$ &$d_3$ &$d_4$\\
+        % &&&\\
+        &&&\\
+        \un & \un & 0 & 0\\
+        0 & \un & \un & 0 \\
+        0 & \un & 0 & \un 
+      \end{tabular}
     \end{tabular}
   }
-  \caption{\figtabsize Students and study groups in \WSI}
-  \label{tab:wsigroups}
-\end{table}
 
-The campus has a residential area {\residenceblock} that has $n$
-single occupancy apartments reserved for the study groups'
-accommodation.  All these apartments are located such that the streets
-connecting them do {\em not} form loops. Figure~\ref{fig:streetmap} shows
-the street map for {\residenceblock}. It may be noted that as a graph,
-it classifies as a tree.
+  \caption[\figtabsize Matrices with and without COP.]{\figtabsize
+    Matrices with and without COP. $M_1$ has COP because by permuting
+    its columns, $c_1$-$c_4$, one can obtain $M_1'$ where the {\un}s
+    in each row are consecutive. $M_2$, however, does not have COP
+    since no permutation of its columns, $d_1$-$d_4$, will arrange
+    {\un}s in each row consecutively \cite{d08phd}.}
 
-\begin{table}[t] %[htbp]
-  \centering
-  {\footnotesize    
-    \begin{tabular}{l||l}
-      \begin{tabular}{rcl}
-        $T $&$=$&$ \text{\em Street map tree of \residenceblock}$\\
-        $V(T) $&$=$&$ \{ 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11 \}$\\
-        $\cP $&$=$&$ \{R\xLLL, R\xGGG, R\xBBB, R\xTTT\}$\\
-        $R\xLLL $&$=$&$ \{9, 1, 5, 3, 11\}$\\
-        $R\xGGG $&$=$&$ \{7, 2, 6, 5\}$\\
-        $R\xBBB $&$=$&$ \{8, 2, 4 \}$\\
-        $R\xTTT $&$=$&$ \{10, 6, 5, 3\}$\\
-        $n $&$=$&$ |V| = 11$\\
-        $m $&$=$&$ |\cP| = 4$\\\\
-        $\cl $&$=$& {\em Study group to route mapping}\\        
-        $\cl(\mathbb{X}) $&$=$&  $R\mathbb{X}$ for all $\mathbb{X} \in
-        \cF$
-      \end{tabular} & 
-      \begin{tabular}{c}
-        {\em Apartment allocation }($\phi$)\\
-        \begin{tabular}{c|c}                      
-          1  & \xSa\\
-          2  & \xPi\\
-          3  & \xFr\\
-          4  & \xWo\\
-          5  & \xCh\\
-          6  & \xVi\\
-          7  & \xPa\\
-          8  & \xSn\\
-          9  & \xLu\\
-          10 & \xLi\\
-          11 & \xSc                     
-        \end{tabular}
-      \end{tabular}\\
-    \end{tabular}
-  }
-  \caption{\figtabsize A solution to \illustrationproblem} %study group accommodation problem
-  \label{tab:iltree}
-\end{table}
+  \label{fig:cop-matrix}
+\end{figure}
 
-A natural question would be to find how the students should be
-allocated apartments such that each study group has the least distance
-to travel for a discussion? More specifically, we are interested in
-the problem with additional conditions, namely, that all the students
-in a study group must be next to each other; in other words, for one
-student to reach another fellow study group member's apartment (for
-all study groups the student is part of), she must not have to pass
-the apartment of any student who is not in that study group. To
-further elucidate, the apartments of students of any study group must
-be arranged in an exclusive unfragmented path on the street
-map. Exclusivity here means that the path must not have apartments
-from other study groups (unless that apartment is also part of {\em
-  this} study group).
+\section{Optimization problems in COP}
+\label{sec:surveycopopt}
+\tnote{Expand on ref:sec:optcop}
 
-An intuitive approach to this problem would be to first find the paths
-that each study group decides to inhabit and then refine the
-allocation to individual students. A feasible allocation of exclusive
-routes to study groups is illustrated in Figure~\ref{fig:streetmap}.  The
-students' allocation of apartments that obeys this route allocation is
-shown in Figure~\ref{fig:streetmappathpeople}. Table~\ref{tab:iltree}
-shows the same solution set theoretically.  How this is
-algorithmically computed is the focus of this thesis.
+\section{********* COP in Relational Database Model}
+\label{sec:surveyrdbm}
+\tnote{Expand on sec:apprdbm} \tnote{(set systems theme)}
 
-%D<debug> figures inside tables
-% \begin{tabular}{cc}
-%  \begin{figure} %[t]%
-% %    \centering
-%     \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
-% %    \label{fig:streetmap}
-%     \caption{--placeholder--}
-%   \end{figure}
-%     &
-%   \begin{figure} %[t]%
-% %    \centering
-%     \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
-% %    \label{fig:streetmappath}
-%     \caption{--placeholder--}
-%   \end{figure}
-% \end{tabular}
+\section{********* COP in Graph Isomorphism}
+\label{sec:surveygraphiso}
+\tnote{Expand on sec:appgraphiso} \tnote{(canonization theme)}
 
-\begin{figure}[htbp] %[t]%
-  \centering
-  \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
-  \caption{\figtabsize {\residenceblock} street map.}
-  \label{fig:streetmap}
+\section{********* Certifying Algorithms}
+\label{sec:surveycertalgo}
+\tnote{Expand on sec:appcertalgo} \tnote{ (certification McC04 theme)}
 
-  \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
-  \caption[\figtabsize \residenceblock street map with study group
-    routes allocated.]{\figtabsize {\residenceblock} street map with study group
-    routes allocated. Routes are color coded as follows: red for
-    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
-    group, orange for \textcolor{yellow}{$\xBBB$} group, green for
-    \textcolor{green}{$\xTTT$} group}
-  \label{fig:streetmappath}
-%   \caption{\figtabsize (a) {\residenceblock} street map (b)
-%     {\residenceblock} street map with study group routes
-%     allocated. Routes are color coded as follows: red for
-%     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
-%     group, orange for \textcolor{yellow}{$\xBBB$} group, green for
-%     \textcolor{green}{$\xTTT$} group}
-%  \label{fig:streetmap}
+\section*{-- REFERENCE CONTENT --}
+%%%%%%%%%%%%% listing package for pretty formatting BEGIN
+% \lstloadlanguages{C}
+% \lstset{language=C, commentstyle=\scriptsize,
+%   numberstyle=\tiny, numbers=left,               % Line numbers
+%   stepnumber=2,numbersep=5pt,firstnumber=10, 
+%   showstringspaces=false,                    % No explicit spaces
+%   emph={printf},emphstyle=\underbar,      % Specific formatting
+%   emph=[2]{sum},emphstyle=[2]\color{blue},      % Sp format (multiple)
+%   emph=[3]{i},emphstyle=[3]\color{red}      % Sp format (multiple)
+%   }
 
-  \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
-  \caption[\figtabsize Solution to the student accommodation
-  problem.]{\figtabsize Individual allocation of apartments to
-    students in {\residenceblock} that meets the requirements stated
-    before.  The routes are color coded as follows: red for
-    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
-    group, orange for \textcolor{yellow}{$\xBBB$} group, green for
-    \textcolor{green}{$\xTTT$} group.\\ {\tiny {\em Peanuts images
-        {\copyright} Charles Schulz}}}% YellowOrange
-  \label{fig:streetmappathpeople}
-\end{figure} \tnote{[i] UPDATE IMAGE (b) 2-infinite-loop-BTWF.png. REMOVE
-  TITLES. ADD {$\xLLL$} {$\xGGG$} {$\xBBB$} {$\xTTT$} [ii]make
-  dashed/textured lines for routes. make it color agnostic.}
+%   \begin{singlespacing}
+%     % Text inside "lstlisting" env is verbatim.
+%     \begin{lstlisting}[keywordstyle=\textbf]      
+% int sum;
+% int i; /* for loop variable */ 
+% sum = 0; 
+% for (i=0, i<n; i++) { 
+%   sum += a[i]; 
+% } 
+% printf{"This is stupid."}
+%     \end{lstlisting}
+%   \end{singlespacing}
 
-As a special case, suppose all the apartments are on the same street
-or if they are all lined up on a single path, the street map becomes a
-tree that is just a path. Then the problem becomes what is called an
-{\em interval assignment problem}. The idea of interval assignment may
-not be obvious here; hence to see this, consider a different problem
-in {\WSI} where the classes for these study groups courses need to be
-scheduled during a day (or a week or any time period). Each study
-group has a bunch of courses associated with it some of which may be
-shared by two or more study groups. It is mandatory that a student who
-is a member of a study group takes all the courses associated with
-that group. There are slots during the day for classes to be held and
-the problem is to allocate class slots to courses such that all the
-classes of a study group are consecutive. It is debatable if this will
-not hamper the attention span and memory retention rate of the
-students but that is, regrettably, out of the scope of this
-thesis. The parallels between this class allocation problem and the
-accommodation problem can be seen as follows. The set $U$ here, are
-the courses offered (say Course 101 {\coneohone}, Course 102
-{\coneohtwo} and so on). In this variation of the problem, the
-collection $\cF$ is the set of study groups but the study groups are
-filled by course IDs (in place of students in the earlier
-example). For instance, Course 101 is mandatory for all study groups
-$\xLLL$, $\xGGG$, $\xBBB$, $\xTTT$ and Course 102 is mandatory for
-only the $\xLLL$ group) and so on. The sequence of class slots for the
-day (or week or any time period) is analogous to the street map in the
-accommodation problem. It is quite obvious now why this version of the
-problem (where the ``target graph'' is a path and not any
-tree\tnote[minor]{Allowing any tree in this example could be seen as a
-  scenario where there are parallel classes. A node falling in the
-  path between two other nodes would mean that the corresponding is
-  scheduled between the other two.}) is called an interval assignment
-problem.
+%%%%%%%%%%%%% listing package for pretty formatting END
 
-The interval assignment problem to a set system is equivalent to the
-consecutive-ones property (COP) problem in binary matrices\cite{wlh02,
-  nsnrs09}.  The COP problem is to rearrange rows (columns) of a
-binary matrix in such a way that every column (row) has its {\un}s
-occur consecutively. If this is possible the matrix is said to have
-the COP.  COP is a well researched combinatorial problem and has
-several positive results on tests for it and computing the COP
-permutation (i.e. the course schedule in the above illustration) which
-will be surveyed later in this document. Hence we are interested in
-extensions of COP, more specifically, the extension of interval
-assignment problem to tree path assignment problem (which is
-illustrated by the study group accommodation problem).
-
-\section{Basic preliminaries - general definitions and nomenclature}
-\label{sec:basicprelim}
-\tnote{(definitions theorems etc) needed if any - graph theory}
-
-\section{Consecutive-ones Property Testing - a Brief Survey}
-\label{sec:background}
-
-In this section, a brief survey of the consecutive-ones problem and
-its optimization problems is presented.
-
+{%\fontfamily{cmvtt}\selectfont
+\color{blue}
 \subsection{Matrices with COP}
@@
 
-% \begin{figure}[t] %[htbp]
-%   \centering
-
-%   {\figtabsize
-%     \begin{tabular}[h]{l|lcccl}
-%       $M_1$: & $M_1'$: &&&& $M_2$:\\
-%       &&&&&\\
-%       \begin{tabular}[h]{llll}
-%         $c_1$ & $c_2$ &$c_3$ &$c_4$\\
-%         &&&\\
-%         \un & 0   & \un & 0\\
-%         0   & \un & 0   & \un \\
-%         \un & 0   & 0   & \un
-%       \end{tabular}
-%       &
-%       \begin{tabular}[h]{llll}
-%         $c_3$ &$c_1$ &$c_4$& $c_2$\\
-%         &&&\\
-%         \un & \un & 0 & 0\\
-%         0 & 0 & \un & \un \\
-%         0 & \un & \un & 0
-%       \end{tabular}
-%       &&&&
-%       \begin{tabular}[h]{llll}
-%         $d_1$ & $d_2$ &$d_3$ &$d_4$\\
-%         % &&&\\
-%         &&&\\
-%         \un & \un & 0 & 0\\
-%         0 & \un & \un & 0 \\
-%         0 & \un & 0 & \un 
-%       \end{tabular}
-%     \end{tabular}
-%   }
-
-%   \caption[\figtabsize Matrices with and without COP.]{\figtabsize
-%     Matrices with and without COP. $M_1$ has COP because by permuting
-%     its columns, $c_1$-$c_4$, one can obtain $M_1'$ where the {\un}s
-%     in each row are consecutive. $M_2$, however, does not have COP
-%     since no permutation of its columns, $d_1$-$d_4$, will arrange
-%     {\un}s in each row consecutively \cite{d08phd}.}
-
-%   \label{fig:cop-matrix}
-% \end{figure}
-
 Recall that a matrix with COP is one whose rows (columns) can be
 rearranged so that the {\un}s in every column (row) are in consecutive
-rows (columns). % Figure~\ref{fig:cop-matrix} shows examples of this
-% property. 
-COP in binary matrices has several practical applications
+rows (columns). COP in binary matrices has several practical applications
 in diverse fields including scheduling \cite{hl06}, information
@@
 submatrix of at least $k$ rows is NP complete. \tnote{how is this
-  different from booth's 75 result??}  \tnote{where should this
-  go? cite|tz04 (approx submatrix with COP sparse matrices)}
+  different from booth's 75 result??}  \tnote{where should this go?
+  cite|tz04 (approx submatrix with COP sparse matrices)}
 
@@
 {\un}s. This is proved to be NP hard using a reduction from the
-Hamiltonian path problem.\tnote{Theorem 2.1 in jkckv}
-\tnote{(1) A connection of COP problem to the travelling salesman
-  problem is also introduced. what does this mean? -- COP can be used
-  as a tool to reorder $0.5T \le runs(M) \le T$. (2) The optimization
-  version of the $k$-run problem, i.e. minimization of number of
-  blocks of ones is proven to be NP complete by
-  cite:k77}\tnote{are these two the same?} \tnote{what is the
-  reduction?} \tnote{other problems similar to COP -- cite:ckl96
-  (ILP, circ ones, one drop) -- cite:th98 (generalization of COP -
-  minimax, biotonic column) Tucker}
+Hamiltonian path problem.\tnote{Theorem 2.1 in jkckv} \tnote{(1) A
+  connection of COP problem to the travelling salesman problem is also
+  introduced. what does this mean? -- COP can be used as a tool to
+  reorder $0.5T \le runs(M) \le T$. (2) The optimization version of
+  the $k$-run problem, i.e. minimization of number of blocks of ones
+  is proven to be NP complete by cite:k77}\tnote{are these two the
+  same?} \tnote{what is the reduction?} \tnote{other problems similar
+  to COP -- cite:ckl96 (ILP, circ ones, one drop) -- cite:th98
+  (generalization of COP - minimax, biotonic column) Tucker} }
 
-\section{********* Application of COP in Areas of Graph Theory and
-  Algorithms}
-\label{sec:appcopGTA}
-
-\subsection{********* COP in Relational Database Model}
-\label{sec:apprdbm}
-\tnote{(set systems theme)}
-
-\subsection{********* COP in Graph Isomorphism}
-\label{sec:appgraphiso}
-\tnote{(canonization theme)}
-
-\subsection{********* Certifying Algorithms}
-\label{sec:appcertalgo}
-\tnote{ (certification McC04 theme)}
-
-\section{Generalization of COP - the Motivation}
-\label{sec:motive}
-Section~\ref{sec:copmatrices} introduced a succinct characterization
-for consecutive-ones property which is solely based on the cardinality
-properties of the set representations of the matrix's columns
-\cite{nsnrs09}. This result is very relevant to this thesis because
-aside from it simplifying COT to a great degree, our generalization
-problem is motivated by their results.
-
-\cite{nsnrs09} characterizes interval assignments to the sets which
-can be obtained from a single permutation of the rows.  For an
-assignment to be feasible, the cardinality of the interval assigned to
-each set in the system must be same as the cardinality of the set, and
-the intersection cardinality of any two intervals must be same as the
-intersection cardinality of their corresponding sets.  While this is
-obviously a necessary condition, this result shows this is also
-sufficient.  \cite{nsnrs09} calls this an Intersection Cardinality
-Preserving Interval Assignment (ICPIA).  This paper generalizes the
-idea from \cite{wlh02} of decomposing a given binary matrix into prime
-matrices for COT and describes an algorithm to test if an ICPIA exists
-for a given set system.
-
-The equivalence of the problem of testing for the consecutive-ones
-property to the constraint statisfaction problem of interval
-assignment \cite{nsnrs09} or interval labeling \cite{kklv10} is as
-follows. Every column (row) of the binary matrix can be converted into
-a set of non-negative integers which are the indices of rows (columns)
-with {\un}s in that column (row). It is apparent that if the matrix
-has COP in columns (rows), then constructing such sets after applying
-the COP permutation to the rows (columns) of the matrix will result in
-sets with consecutive integers. In other words, after application of
-COP reordering, the sets are intervals. Indeed the problem now becomes
-finding interval assignments to a given set system such that there
-exists a permutation of the universe of set of row indices (column
-indices) which converts each set to its assigned interval.
-
-The problem of interest in this thesis, namely, tree path labeling
-problem, is a natural generalization of the interval assignment
-problem or the COT problem. The problem is defined as follows -- given
-a set system $\cF$ from a universe $U$ and a target tree $T$, does
-there exist a bijection from $U$ to the vertices of $T$ such that each
-set in the system maps to a path in $T$.  We refer to this as the
-{\CFTPL} problem or simply {\em tree path labeling} problem for an
-input set system and target tree pair -- $(\cF,T)$. The special case
-of the target tree being a path, is the interval assignment problem.
-We focus on generalizing the notion of an ICPIA \cite{nsnrs09} to
-characterize feasible path assignments.  We show that for a given set
-system $\cF$, a tree $T$, and an assignment of paths from $T$ to the
-sets, there is a feasible\endnote{The notion of {\em feasibility} is
-  formally defined in Section~\ref{ch:prelims}.}  bijection between
-$U$ and $V(T)$ if and only if the intersection cardinalities among any
-three sets (not necessarily distinct) is equal to that of the
-corresponding paths assigned to them and the input passes a filtering
-algorithm (described in this paper) successfully.  This algorithmic
-characterization gives a natural data structure that stores all the
-\remove{relevant} feasible bijections between $U$ and $V(T)$. This
-reduces the search space for the solution considerably from the
-universe of all possible bijections between $U$ and $V(T)$ to only
-those bijections that maintain the characterization.  Further, the
-filtering algorithm is also an efficient algorithm to test if a tree
-path labeling\endnote{The terms {\em tree path labeling} and {\em
-    tree path assignment} are, in informal language,
-  synonyms. Formally, the former refers to the bijection $\cl: \cF
-  \rightarrow \cP$. The latter refers to the set of ordered pairs
-  $\{(S, P) \mid S \in \cF, P \in \cP\}$. $\cP$ is a set of paths on
-  $T$.} is feasible.
-
-\section{Summary of New Results in this Thesis}
-\label{sec:results}
-
-We see in Section~\ref{sec:motive} that pairwise intersection
-cardinality preservation is necessary and sufficient for an interval
-assignment to be feasible for a given hypergraph\endnote{A {\em
-    hypergraph} is an alternate representation of a set system and
-  will be used in this thesis.}~\endnote{See Section~\ref{ch:prelims} for the
-  formal definition.} and thus is a characterization for COP
-\cite{nsnrs09}. \tnote{Refer to the survey section that will list the
-  theorems and/or lemma?} In our work we extend this characterization
-and find that trio-wise intersection cardinality preservation makes a
-tree path labeling\endnote{A {\em tree path labeling} $\cl$ is a
-  bijection of paths from the target tree $T$ to the hyperedges in
-  given hypergraph $\cF$.}~\endnotemark[4]  (TPL) feasible, which is a generalization of the COP
-problem. This problem is defined as follows.
-
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \FTPL}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
-        paths $\cP$ from $T$ and a
-        bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
-      \end{minipage}\\
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
-        such that $\phi$ when applied on any hyperedge in $\cF$ will
-        give
-        the path mapped to it by the given tree path labeling $\cl$.\\
-        { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every
-          hyperedge $S \in \cF$.}
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
-
-We give a necessary and sufficient condition by way of {\em
-  Intersection Cardinality Preservation Path Labeling} (ICPPL) and a
-filtering algorithm for {\FTPL} to output in affirmative. ICPPL
-captures the trio-wise cardinality property described
-earlier\endnote{See Section~\ref{sec:feasible} for the definition of
-  ICPPL.}. This characterization can be checked in polynomial time.  A
-relevant consequence of this constructive procedure is that it is
-sufficient to iteratively check if three-way intersection
-cardinalities are preserved.  In other words, in each iteration, it is
-sufficient to check if the intersection of any three hyperedges is of
-the same cardinality as the intersection of the corresponding paths.
-Thus this generalizes the well studied question of the feasible
-interval assignment problem which is the special case when the target
-tree $T$ is simply a path \cite{wlh02,nsnrs09}.
-
-Aside from checking if a given TPL is feasible, we also solve the
-problem of computing a feasible TPL for a given hypergraph and target
-tree, if one exists. This problem, {\CFTPL}, is defined as follows.
-
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \CFTPL}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$ and a tree $T$.\\
-      \end{minipage}\\
-
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a set of paths $\cP$ from $T$ and a bijection
-        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
-        {\bf true} on $(\cF, T, \cl)$.
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
-
-We present a polynomial time algorithm for {\CFTPL} when the target
-tree $T$ belongs to a special class of trees called {\em \kstars} and
-when the hyperedges in the hypergraph $\cF$ have at most $k+2$
-vertices. A couple of examples of {\kstars} are given in
-Figure~\ref{fig:ksubstar}.
-
-\begin{figure}[t]
-  \centering
-  \begin{tabular}[h]{ccccc}
-    \includegraphics[scale=0.3]{../img/star.pdf} &&&&
-    \includegraphics[scale=0.3]{../img/kstar.pdf}\\
-    (a) &&&& (b)
-  \end{tabular}
-  \caption{\figtabsize Examples of {\kstars}. (a) $k = 0$ (b) $k = 2$
-  }
-  \label{fig:ksubstar}
-\end{figure}
-
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \CFTPLKTREE}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$ such that every
-        hyperedge
-        $S \in \cF$ is of cardinality at most $k+2$ and a {\kstar} $T$.\\
-      \end{minipage}\\
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a set of paths $\cP$ from $T$ and a bijection
-        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
-        {\bf true} on $(\cF, T, \cl)$.
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
-
-\tnote{The following sentence seems out of place in this para} In
-spite of this being a restricted case, we believe that our results are
-of significant interest in understanding the nature of {\sc Graph
-  Isomorphism} which is polynomial time solvable in interval graphs
-while being hard on path graphs\cite{kklv10}. {\kstars} are a class
-of trees which are in many ways very close to intervals or paths. Each
-ray\endnote{The path from a leaf to the root, the vertex with highest
-  degree, is called a {\em ray} of the \kstar.}~\endnotemark[4] are independent except for the
-root\endnote{The vertex with maximum degree in a {\kstar} is called
-  {\em root}.}~\endnotemark[4] and hence can be
-considered as an independent interval till the root. Our algorithm
-builds on this fact and uses the interval assignment
-algorithm\cite{nsnrs09} up until ``reaching'' the root and then uses
-the trio-wise intersection cardinality (the extra condition in ICPPL
-that generalizes ICPIA) check to resolve the ambiguity about which ray
-the algorithm should ``grow'' the solution into in the next iteration.
-
-We also have an algorithm for solving {\CFTPL} with no restrictions on
-the target tree or set size which runs in exponential time.  This
-algorithm finds a path labeling from $T$ by decomposing the problem
-into subproblems of finding path labeling of subsets of $\cF$ from
-subtrees of $T$. Given the fact that binary matrices naturally
-represent a set system (see Section~\ref{sec:motive}) and that the
-{\em overlap}\tnote{Give informal definition of ``overlap''
-  in footnote/endnote and link to sec:prelims. ADD the definition in prelim.}
-relation between the sets involved is an obvious equivalence relation,
-$\cF$ quite naturally partitions into equivalence classes known as
-{\em overlap components}\tnote{Give informal definition of ``overlap
-  components'' in footnote and link to sec:prelims. ADD the definition
-  in prelim.}. In the context of COP, overlap components were used in
-\cite{wlh02} and \cite{kklv10}. Moreover, \cite{nsnrs09} discovered
-that these equivalence classes form a total order \tnote{Give informal
-  definition of ``total order'' in footnote?}. We extend this to TPL
-and find that when $\cF$ is a path hypergraph\endnote{If there exists
-  an FTPL for a hypergraph $\cF$, it is called a path hypergraph.},
-the classes can be partially ordered as an in-tree in polynomial
-time. Once $\cF$ is ``broken'' into overlap components, one must
-identify the subtree of $T$ that it needs to map to and this is the
-hard part which is currently open to be solved in polynomial time.
-
-\tnote{The connection of TPL to graph isomorphism will be made
-  later in the document}
-
-\theendnotes
+\theendnotes\tnote{remove if none.}

### 2012-01-03 src/030chap3.tex
Cleaned up thesis.tex. Split thesis.tex into chapterwise subfiles for easy edit tex management. Content changes are mainly in chapter 1 and 3 although nothing significant. To being working on chapter 2 after this commit.

--- 2011-12-31 src/thesis.tex
@@
-\documentclass[MS]             % MS thesis dissertation.
-              {iitmdiss_as}    % Hacked by hyperbolicme.
-%              {iitmdiss}      % If the powers to be are finicky.
-
-\usepackage{times}             % Font?
-\usepackage{t1enc}             % T1 encoding??  
-%\usepackage{boxedminipage}
-\usepackage{../lib/mythesislib}
-\usepackage{../lib/mylatexlib}
-% \usepackage{comment}           % DEBUG ONLY
-
-\def \mythesissubmissiondate {10 January 2012}
-\def \mythesissubmissionmonth {\MakeUppercase {January 2012}}
-
-\begin{document}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Title page
-\title{\mythesistitleNOTSYNPOSIS} 
-\author{\myname} 
-\date{
-  {\mythesissubmissionmonth}} 
-\department{\MakeUppercase{\mythesisdept}}
-%\department{\mythesisdept}
-
-%\nocite{*}
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Certificate
-\certificate
-
-\vspace*{0.5in}
-
-\noindent 
-This is to certify that the thesis titled {\bf \mythesistitleNOTSYNPOSIS},
-submitted by {\bf \myname}, to the {\bf Indian Institute of Technology
-Madras}, for the award of the degree of {\bf \mydegree}, is 
-a bona fide record of the research work done by her under our
-supervision.  The contents of this thesis, in full or in parts, have
-not been submitted to any other Institute or University for the award
-of any degree or diploma.
-
-\vspace*{1.5in}
-
-\begin{singlespacing}
-  \hspace*{-0.25in}
-  \parbox{3in}{
-    \noindent {\bf Dr.~\myadvisor} \\
-    \small
-    \noindent Research Guide\\
-    \noindent Associate Professor\\
-    \noindent Dept. of \mydept\\
-    \noindent IIT Madras -- 600 036
-  } \hfill
-  \parbox{1.25in}{               % {1.75in}
-    \small
-    \vspace*{0.75in}
-    \begin{tabular}[h]{r}
-    \noindent {Chennai}\\         
-    \noindent {\mythesissubmissiondate} 
-    \end{tabular}
-  }
-\end{singlespacing}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Acknowledgements
-\acknowledgements
-
-\tnote{WIP}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Abstract
-
-\begin{abstract}
-  \noindent {\em {\bf Keywords}: \hspace*{0.5em} \parbox[t]{4.4in}{consecutive
-    ones property, algorithmic graph theory, hypergraph isomorphism,
-    interval labeling}} \vspace*{24pt}
-
-  \def \tem {}
-  \noindent 
-  Consecutive-ones property is a non-trivial property of binary
-  matrices that has been studied widely in the literature for over
-  past 50 years. Detection of COP in a matrix is possible efficiently
-  and there are several algorithms that achieve the same. This thesis
-  documents the work done on an extension of COP extended from the
-  equivalent interval assignment problem in \cite{nsnrs09}. These new
-  results rigorously prove a natural extension (to trees) of their
-  characterization as well as makes connections to graph isomorphism,
-  namely path graph isomorphism.
-
-  \tnote{EXPAND. 
-  Abstract must be a brief about what results we have and how it fits
-  in the body of research.\\
-  -- Area: Broad to Specialized. i.e. Combinatorial
-    algorithms -> Matrix reorganization -> general data reorganization
-    (interval assignment) -> path assignment\\
-  -- Class of problems: say, data reorganization.\\
-  -- Nature of results: Is a generalization. We have a
-    Polynomial algorithm for a subset of the generalization.\\
-% maybe add - we explore a natural
-%     generalization of results on binary matrices with the |\tem
-%     consecutive-ones property|.  We consider the following constraint
-%     satisfaction problem. Given (i) a set system $\F \subseteq$
-%     $(2^{U} \setminus \emptyset)$ of a finite set $U$ of cardinality
-%     $n$, (ii) a tree $T$ of size $n$ and (iii) a bijection called
-%     |\tem tree path labeling|, $\cl$ mapping the sets in $\cF$ to
-%     paths in $T$, does there exist at least one bijection $\phi:U
-%     \rightarrow V(T)$ such that for each $S \in \cF$, $\{\phi(x) \mid
-%     x \in S\} = \cl(S)$?  A tree path labeling of a set system is
-%     called |\tem feasible| if there exists such a bijection $\phi$.
-%     We present an algorithmic characterization of feasible tree path
-%     labeling. COP is a special instance of tree path labeling problem
-%     when $T$ is a path.  We conclude with a polynomial time algorithm
-%     to find a feasible tree path labeling of a given set system when
-%     $T$ is a |\tem $k$-subdivided star|, set system has a single
-%     containment tree of overlap components and set size is limited to
-%     at most $k+2$.
-  }
-\end{abstract}
-
-%\pagebreak
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Table of contents etc.
-
-\begin{singlespace}
-\tableofcontents
-\thispagestyle{empty}
-
-\listoftables
-\addcontentsline{toc}{chapter}{LIST OF TABLES}
-\listoffigures
-\addcontentsline{toc}{chapter}{LIST OF FIGURES}
-\end{singlespace}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Abbreviations
-\abbreviations
-
-\noindent 
-\begin{tabbing}
-xxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
-\textbf{COP}   \> Consecutive-ones Property \\
-\textbf{COT}   \> Consecutive-ones Testing \\
-\textbf{ICPIA}   \> Intersection Cardinality Preservation Interval Assignment \\
-\textbf{ICPPL}   \> Intersection Cardinality Preserved Path Labeling \\
-% ADD MORE HERE
-\end{tabbing}
-
-\pagebreak
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Notation
-
-\chapter*{\centerline{NOTATION}}
-\addcontentsline{toc}{chapter}{NOTATION}
-
-\begin{singlespace}
-\begin{tabbing}
-xxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
-\textbf{$2^{U}$}  \> Powerset of set $U$ \\
-\end{tabbing}
-\end{singlespace}
-
-\pagebreak
-\clearpage
-
-% The main text will follow from this point so set the page numbering
-% to arabic from here on.
-\pagenumbering{arabic}
-
-\chapter{Introduction}
-\label{ch:intro}
-
-Consecutive-ones property is a non-trivial property of binary matrices
-that has been studied widely in the literature for over past 50
-years. Detection of COP in a matrix is possible efficiently and there
-are several algorithms that achieve the same. This thesis documents
-the work done on an extension of COP extended from the equivalent
-interval assignment problem in \cite{nsnrs09}. These new results
-rigorously prove a natural extension (to trees) of their
-characterization as well as makes connections to graph isomorphism,
-namely path graph isomorphism.
-
-\tnote{Have a para on Organization: Outline of document}
-Section~\ref{sec:background} gives a brief survey of COP and
-optimization problems related to it followed by motivation for the
-thesis in Section~\ref{sec:motive}.  Section~\ref{sec:results}
-presents a summary of our results on the extension of COP namely, the
-tree path labeling problem. \tnote[TH]{update for thesis}
-
-% \tnote{testing 1 2 3...}
-% \tnote[bogus1]{testing 1 2 3...}
-% \tnote[bogus2]{testing 1 2 3...}
-% \tnote[GTC]{testing 1 2 3...}
-% \tnote[bogus4]{testing 1 2 3...}
-% \tnote[TH]{testing 1 2 3...}
-
-\section{Illustration of the problem}
-\label{sec:problem}
-
-A group of students, \Pa, \Pig, \Sn, \Wo, \Vi, \Li, \Ch, \Sa, \Fr,
-\Sc\ and \Lu\ enroll at the {\WSI} for a liberal arts programme.  As
-part of their semester thesis, they pick a body of work to study and
-form the namesake study groups, {\LLL}, {\GGG}, {\BBB} and
-{\TTT}\tnote[TH]{put bib entries for these works!}. A student will be
-in at least one study group and may be in more than one. For instance,
-as will be seen later, {\Fr} studies both {\LLL} and {\TTT} while \Wo\
-studies only \BBB.
-
-Let $U$ and $\cF$ represent the set of students and the set of study
-groups respectively and the integers $n$ and $m$ denote the total
-number students and study groups respectively. In relation to this
-example, these are defined in Table~\ref{tab:wsigroups}. Also given
-there is the study group allocation to students.
-
-\begin{table}[t]%[htbp]
-  \centering
-  { \footnotesize
-  \begin{tabular}{rcl}
-    $U $&$=$&$ \{\xPa,\; \xPi,\; \xSn,\; \xWo,\; \xVi,\; \xLi,\; \xCh,\;
-    \xSa,\; \xFr,\; \xSc,\; \xLu\}$\\
-    $\cF $&$=$&$ \{\xLLL, \xGGG, \xBBB, \xTTT\}$\\
-    $\xLLL $&$=$&$ \{\xCh,\;  \xSa,\;  \xFr,\;  \xSc,\;  \xLu \}$\\
-    $\xGGG $&$=$&$ \{\xPa,\;  \xPi,\;  \xVi,\;  \xCh \}$\\
-    $\xBBB $&$=$&$ \{\xSn,\;  \xPi,\;  \xWo \}$\\
-    $\xTTT $&$=$&$ \{\xVi,\;  \xLi,\;  \xCh,\;  \xFr \}$\\
-    $n $&$=$&$ |U| = 11$\\
-    $m $&$=$&$ |\cF| = 4$      
-  \end{tabular}
-}
-  \caption{\figtabsize Students and study groups in \WSI}
-  \label{tab:wsigroups}
-\end{table}
-
-The campus has a residential area {\residenceblock} that has $n$
-single occupancy apartments reserved for the study groups'
-accommodation.  All these apartments are located such that the streets
-connecting them do {\em not} form
-loops. Fig~\ref{fig:streetmappathpeople} \tnote[TH]{fig:streetmap}
-shows the street map for {\residenceblock}. It may be noted that as a
-graph, it classifies as a tree.
-
-\begin{table}[htbp]
-  \centering
-  {\footnotesize    
-    \begin{tabular}{l||l}
-      \begin{tabular}{rcl}
-        $T $&$=$&$ \text{\em Street map tree of \residenceblock}$\\
-        $V(T) $&$=$&$ \{ 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11 \}$\\
-        $\cP $&$=$&$ \{R\xLLL, R\xGGG, R\xBBB, R\xTTT\}$\\
-        $R\xLLL $&$=$&$ \{9, 1, 5, 3, 11\}$\\
-        $R\xGGG $&$=$&$ \{7, 2, 6, 5\}$\\
-        $R\xBBB $&$=$&$ \{8, 2, 4 \}$\\
-        $R\xTTT $&$=$&$ \{10, 6, 5, 3\}$\\
-        $n $&$=$&$ |V| = 11$\\
-        $m $&$=$&$ |\cP| = 4$\\\\
-        $\cl $&$=$& {\em Study group to route mapping}\\        
-        $\cl(\mathbb{X}) $&$=$&  $R\mathbb{X}$ for all $\mathbb{X} \in
-        \cF$
-      \end{tabular} & 
-      \begin{tabular}{c}
-       {\em Apartment allocation }($\phi$)\\
-                    \begin{tabular}{c|c}                      
-                      1  & \xSa\\
-                      2  & \xPi\\
-                      3  & \xFr\\
-                      4  & \xWo\\
-                      5  & \xCh\\
-                      6  & \xVi\\
-                      7  & \xPa\\
-                      8  & \xSn\\
-                      9  & \xLu\\
-                      10 & \xLi\\
-                      11 & \xSc                     
-                    \end{tabular}
-      \end{tabular}\\
-    \end{tabular}
-  }
-  \caption{\figtabsize A solution to study group accommodation problem}
-  \label{tab:iltree}
-\end{table}
-
-A natural question would be to find how the students should be
-allocated apartments such that each study group has the {\em least
-  distance to travel} for a discussion? More specifically, we are
-interested in enforcing additional conditions, namely, that all the
-students in a study group must be next to each other; in other words,
-for one student to reach another fellow study group member's apartment
-(for all study groups the student is part of), she must not have to
-pass the apartment of any student who is not in that study group. To
-further elucidate, the apartments of students of any study group must
-be arranged in an exclusive unfragmented path on the street
-map. Exclusivity here means that the path must not have apartments
-from other study groups (unless that apartment is also part of {\em
-  this} study group).
-
-An intuitive approach to this problem would be to first find the paths
-that each study group decides to inhabit and then refine the
-allocation to individual students. A feasible allocation of exclusive
-routes to study groups is illustrated in
-Fig~\ref{fig:streetmappathpeople} \tnote[TH]{fig:streetmappath} and
-the students' allocation of apartments that obeys this route
-allocation is also shown. Table~\ref{tab:iltree} shows the same
-solution set theoretically.  How this is algorithmically computed is
-the focus of this thesis.
-
-\begin{figure}[htbp] %[t]%
- \centering
-  (a) \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
-  (b) \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
-
- \label{fig:streetmap}
- \caption{\figtabsize (a) {\residenceblock} street map (b)
-   {\residenceblock} street map with study group routes
-   allocated. Routes are color coded as follows: red for
-   \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
-   group, orange for \textcolor{yellow}{$\xBBB$} group, green for
-   \textcolor{green}{$\xTTT$} group}
-
- \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
-  \caption{\figtabsize Individual allocation of apartments to students
-    in {\residenceblock} that meets the requirements stated before.
-    The routes are color coded as follows: red for
-    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
-    group, orange for \textcolor{yellow}{$\xBBB$} group, green
-    for \textcolor{green}{$\xTTT$} group. {\tiny {\em Peanuts images
-        {\copyright} Charles Schulz}}}% YellowOrange
-  \label{fig:streetmappathpeople}
-\end{figure}
-\tnote{UPDATE IMAGE (b) 2-infinite-loop-BTWF.png. REMOVE TITLES. ADD
-  {$\xLLL$} {$\xGGG$} {$\xBBB$} {$\xTTT$}}
-
-\tnote[TH]{make dashed/textured lines for routes. make it color
-  agnostic.}
-
-As a special case, suppose all the apartments are on the same street
-or if they are all lined up on a single path, the street map becomes a
-tree that is just a path. Then the problem becomes what is called an
-{\em interval assignment problem}. The idea of interval assignment may
-not be obvious here; hence to see this, consider a different problem
-in {\WSI} where the classes for these study groups courses need to be
-scheduled during a day (or a week or any time period). Each study
-group has a bunch of courses associated with it some of which may be
-shared by two or more study groups. It is mandatory that a student who
-is a member of a study group takes all the courses associated with
-that group. There are slots during the day for classes to be held and
-the problem is to allocate class slots to courses such that all the
-classes of a study group are consecutive. It is debatable if this will
-not hamper the attention span and memory retention rate of the
-students but that is, regrettably, out of the scope of this
-thesis. The parallels between this class allocation problem and the
-accommodation problem can be seen as follows. The set $U$ here, are
-the courses offered (say Course 101 {\coneohone}, Course 102
-{\coneohtwo} and so on). In this variation of the problem, the
-collection $\cF$ is the set of study groups but the study groups are
-filled by course IDs (in place of students in the earlier
-example). For instance, Course 101 is mandatory for all study groups
-$\xLLL$, $\xGGG$, $\xBBB$, $\xTTT$ and Course 102 is mandatory for
-only the $\xLLL$ group) and so on. The sequence of class slots for the
-day (or week or any time period) is analogous to the street map in the
-accommodation problem. It is quite obvious now why this version of the
-problem (where the ``target graph'' is a path and not any
-tree\tnote[TH]{Allowing any tree in this example could be seen as a
-  scenario where there are parallel classes. A node falling in the
-  path between two other nodes would mean that the corresponding is
-  scheduled between the other two.}) is called an interval assignment
-problem.
-
-The interval assignment problem to a set system is equivalent to the
-consecutive-ones property (COP) problem in binary matrices\cite{wlh02,
-  nsnrs09}.  The COP problem is to rearrange rows (columns) of a
-binary matrix in such a way that every column (row) has its {\un}s
-occur consecutively. If this is possible the matrix is said to have
-the COP.  COP is a well researched combinatorial problem and has
-several positive results on tests for it and computing the COP
-permutation (i.e. the course schedule in the above illustration) which
-will be surveyed later in this document. Hence we are interested in
-extensions of COP, more specifically, the extension of interval
-assignment problem to tree path assignment problem (which is
-illustrated by the study group accommodation problem).
-
-\section{Basic preliminaries - general definitions and nomenclature}
-\tnote{(definitions theorems etc) needed if any
-    - graph theory}
-
-\section{Consecutive-ones Property Testing - a Brief Survey}
-\label{sec:background}
-
-In this section, a brief survey of the consecutive-ones problem and its
-optimization problems is presented.
-
-%%%%%%                              %%%%%%
-%THX is the second level of ``tnote''-ing%
-%%%%%%                              %%%%%%
-\tnote[TH]{ADD: As it will be described in detail later in this
-  document, isomorphism of certain % classes of graphs, namely chordal
-%   graphs, have a close relationship with consecutive-ones property and
-%   generalizations of it.  This is perhaps because of how closely COP
-%   of a matrix relates to properties of graphs derived from matrices as
-%   seen in the following results.
-}
-\tnote[THX]{ADD: peo exists iff chordal. lexicographic BFS
-  [tag:chordalGraph]} %
-\tnote[THX]{ADD: A well known result in %perfect graph theory is that
-%  the maximal cliques of  an interval graph $G$ can be linearly ordered
-%   such that for all $v \in V(G)$, cliques containing $v$ are
-%   consecutive in the ordering cite~{gh64}. This clearly means that a
-%   graph $G$ is an interval graph if and only if
-}% 
-\tnote[THX]{verify
-  from paper the statement of claim.}% 
-\tnote[TH]{maximal clique vertex
-  incidence matrix of%  $G$ has COP.  Also, maximal cliques of any
-%   chordal graph can be enumerated in polytime $O(m+n)$
-}
-\tnote[THX]{citation?!!}%  
-\tnote[TH]{cite~{fg65} uses these results to
-  give the first polynomial time algorithm for COT.}%
-\tnote[THX]{check. how do they use it?}%  
-\tnote[TH]{ A bipartite graph
-  is convex%  (on $R$) if and only if its half adjacency matrix has COP
-%   on rows.  The results in cite~{bl76} on COT are based on the result
-%   that interval graphs are AT-free chordal graphs.
-}%
-\tnote[THX]{the latter being Tucker's?}%  
-\tnote[THX]{(2) TBD survey
-  -- % see blue notes (in notebook) under Graph Isomorphism. namely
-%   citations in cite:aas93 (3) a brief on heirarchy of chordal graphs,
-%   path graphs, interval graphs, peo, clique tree etc. and results. (4)
-%   interval graphs are incomparability graphs. see
-%   Golumbic. [tag:classification] (5) finding min length hole in
-%   bipartite graph is polynomial time [Sec 3.3.2 in cite:d08phd] (6)
-%   theorem 2.2 in cite:d08phd - $G$ is union of v.d. caterpillars iff
-%   $M$ has COP in rows. $M$ is edge vertex incidence matrix $(*,2)$ of
-%   $G$
-}%
-
-\subsection{Matrices with COP}
-\label{sec:copmatrices}
-As seen earlier, the interval assignment problem (illustrated as the
-course scheduling problem in Section~\ref{sec:problem}), is a special
-case of the problem we address in this thesis, namely the tree path
-labeling problem (illustrated as the study group accommodation
-problem). The interval assignment problem and COP problem are equivalent
-problems. In this section we will see some of the results that
-exists in the literature today towards solving the COP problem and
-optimization problems surrounding it.
-
-\begin{figure}[t] %[htbp]
-  \centering
-
-  {\figtabsize
-    \begin{tabular}[h]{l|lcccl}
-      $M_1$: & $M_1'$: &&&& $M_2$:\\
-      &&&&&\\
-      \begin{tabular}[h]{llll}
-        $c_1$ & $c_2$ &$c_3$ &$c_4$\\
-        &&&\\
-        \un & 0   & \un & 0\\
-        0   & \un & 0   & \un \\
-        \un & 0   & 0   & \un
-      \end{tabular}
-      &
-      \begin{tabular}[h]{llll}
-        $c_3$ &$c_1$ &$c_4$& $c_2$\\
-        &&&\\
-        \un & \un & 0 & 0\\
-        0 & 0 & \un & \un \\
-        0 & \un & \un & 0
-      \end{tabular}
-      &&&&
-      \begin{tabular}[h]{llll}
-        $d_1$ & $d_2$ &$d_3$ &$d_4$\\
-        % &&&\\
-        &&&\\
-        \un & \un & 0 & 0\\
-        0 & \un & \un & 0 \\
-        0 & \un & 0 & \un 
-      \end{tabular}
-    \end{tabular}
-  }
-
-  \caption{\figtabsize Matrices with and without COP. $M_1$ has COP
-    because by permuting its columns, $c_1$-$c_4$, one can obtain
-    $M_1'$ where the {\un}s in each row are consecutive. $M_2$,
-    however, does not have COP since no permutation of its columns,
-    $d_1$-$d_4$, will arrange {\un}s in each row consecutively
-    \cite{d08phd}.}
-
-  \label{fig:cop-matrix}
-\end{figure}
-
-Recall that a matrix with COP is one whose rows (columns) can be
-rearranged so that the {\un}s in every column (row) are in consecutive
-rows (columns). Figure~\ref{fig:cop-matrix} shows examples of this
-property.  COP in binary matrices has several practical applications
-in diverse fields including scheduling \cite{hl06}, information
-retrieval \cite{k77} and computational biology \cite{abh98}.  Further,
-it is a tool in graph theory \cite{mcg04} for interval graph
-recognition, characterization of Hamiltonian graphs, planarity testing
-\cite{bl76} and in integer linear programming \cite{ht02,hl06}.
-
-The obvious first questions after being introduced to the consecutive
-ones property of binary matrices are if COP can be detected
-efficiently in a binary matrix and if so, can the COP permutation of
-the matrix also be computed efficiently?  Recognition of COP in a
-binary matrix is polynomial time solvable and the first such algorithm
-was given by \cite{fg65}.  A landmark result came a few years later
-when \cite{at72} discovered the families of forbidden submatrices that
-prevent a matrix from having COP and most, if not all, results that
-came later were based on this discovery which connected COP in binary
-matrices to convex bipartite graphs. In fact, the forbidden
-submatrices came as a corollary to the discovery that convex bipartite
-graphs are AT-free in \cite{at72}\tnote[TH]{check}. The first linear
-time algorithm for COP testing (COT) was invented by \cite{bl76} using
-a data structure called PQ trees.  Since then several COT algorithms
-have been invented -- some of which involved variations of PQ trees
-\cite{mm96,wlh01,mcc04}, some involved set theory and ICPIA
-\cite{wlh02,nsnrs09}, parallel COT algorithms\cite{as95,bs03,ly91} and
-certifying algorithms\cite{mcc04}.
-
-The construction of PQ trees in \cite{bl76} draws on the close
-relationship of matrices with COP to interval graphs. A PQ tree of a
-matrix is one that stores all row (column) permutations of the matrix
-that give the COP orders (there could be multiple orders of rows or
-columns) of the matrix. This is constructed using an elaborate linear
-time procedure and is also a test for planarity\tnote[TH]{check check
-  check. both interval graph and planarity in this paper?}.  PQR trees
-is a generalized data structure based on PQ trees \cite{mm96,mpt98}.
-\cite{tm05} describes an improved algorithm to build PQR
-trees. \tnote[TH]{improv in terms of what?}\cite{wlh02} describes the
-simpler algorithm for COT. Hsu also invented PC trees
-\cite{wlh01}\tnote[TH]{This result first appeared inproc ISAAC92}
-which is claimed to be much easier to implement. \cite{nsnrs09}
-describes a characterization of consecutive-ones property solely based
-on the cardinality properties of the set representations of the
-columns (rows); every column (row) is equivalent to a set that has the
-row (column) indices of the rows (columns) that have one entries in
-this column (row). This is interesting and relevant, especially to
-this thesis because it simplifies COT to a great degree. \tnote[TH]{it
-  reduces the solution search space. fill in the blanks.}
-
-\cite{mcc04} describes a different approach to COT. While all previous
-COT algorithms gave the COP order if the matrix has the property but
-exited stating negative if otherwise, this algorithm gives an evidence
-by way of a certificate of matrix even when it has no COP. This
-enables a user to verify the algorithm's result even when the answer
-is negative. This is significant from an implementation perspective
-because automated program verification is hard and manual verification
-is more viable. Hence having a certificate reinforces an
-implementation's credibility. Note that when the matrix {\em has} COP,
-the COP order is the certificate.  The internal machinery of this
-algorithm is related to the weighted betweenness problem
-addressed\tnote[TH]{in what way??} in \cite{co98}.  \tnote[TH]{expand
-  on the COP order graph creation and it having to be bipartite for M
-  to have COP. and thus an odd cycle being an evidence of no COP.}
-\tnote[TH]{ where should this go?: (1) cite|jlm97 (application of PQ trees
-  in graphics). (2) helly's theorem citation
-  19XXdgk-Hellystheorem-Danzer-Gruenbaum-Klee}
-
-\subsection{Optimization problems in COP}
-\label{sec:optcop}
-
-So far we have been concerned about matrices that have the consecutive
-ones property. However in real life applications, it is rare that data
-sets represented by binary matrices have COP, primarily due to the
-noisy nature of data available. At the same time, COP is not arbitrary
-and is a desirable property in practical data representation
-\cite{co98,jkckv04,k77}. In this context, there are several
-interesting problems when a matrix does not have COP but is ``close''
-to having COP or is allowed to be altered to have COP. These are the
-optimization problems related to a matrix which does not have
-COP. Some of the significant problems are surveyed in this section.
-
-\tnote[TH]{ -- sect 4.1 in cite:d08phd has many results
-  surveyed. hardness results, approx. results. results are usually for
-  a class of matrices $(a,b)$ where number {\un}s in columns and rows
-  are restriced to $a$ and $b$ . -- problem of flipping at most $k$
-  entries of $M$ to make it attain COP. this is NP complete
-  cite:b75-phd}\tnote[TH]{(1) scite:lb62 showed that interval graphs
-  are AT-free.  describe AT (2) show the close relationship b/w COP
-  and graphs sec 2.2, pg 31} \cite{at72} showed that a matrix that
-does not have COP have certain substructures that prevent it from
-having COP. Tucker classified these forbidden substructures into five
-classes of submatrices. This result is presented in the context of
-convex bipartite graphs which \cite{at72} proved to be
-AT-free\tnote[TH]{ check this up. give details. - doms'}. By
-definition, convex bipartite graph have half adjacency matrices that
-have COP on either rows or columns (graph is biconvex if it has COP on
-both)\cite{d08phd}. A half adjacency matrix is a binary matrix
-representing a bipartite graph as follows. The set of rows and the set
-of columns form the two partitions of the graph. Each row node is
-adjacent to those nodes that represent the columns that have {\un}s in
-the corresponding row. \cite{at72} proves that this bipartite graph
-has no asteroidal triple if and only if the matrix has COP and goes on
-to identify the forbidden substructures for these bipartite
-graphs. The matrices corresponding to these substructures are the
-forbidden submatrices.
-
-Once a matrix has been detected to not have COP (using any of the COT
-algorithms mentioned earlier), it is naturally of interest to find out
-the smallest forbidden substructure (in terms of number of rows and/or
-columns and/or number of entries that are {\un}s). \cite{d08phd}
-discusses a couple of algorithms which are efficient if the number of
-{\un}s in a row is small. This is of significance in the case of sparse
-matrices where this number is much lesser than the number of
-columns. $(*,\Delta)${\em -matrices} are matrices with no restriction
-on number of {\un}s in any column but has at most $\Delta$ {\un}s in
-any row. {\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are
-similar problems which deals with inducing COP on a matrix. In {\sc
-  Min COS-R (Min COS-C)} the question is to find the minimum number of
-rows (columns) that must be deleted to result in a matrix with COP.
-In the dual problem {\sc Max COS-R (Max COS-C)} the search is for the
-maximum number of rows (columns) that induces a submatrix with
-COP. Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
-a submatrix $M'$ with all columns\tnote[TH]{check if b75 deals with
-  COP col or COP row. also is it any submatrix with k less than r rows
-  or submatrix must have all columns?} but a maximum cardinality
-subset of rows such that $M'$ has COP is NP complete. \cite{hg02}
-corrects an error of the abridged proof of this reduction as given in
-\cite{gj79}.  \cite{d08phd} discusses all these problems in detail
-giving an extensive survey of the previously existing results which
-are almost exhaustively all approximation results and hardness
-results. Taking this further, \cite{d08phd} presents new results in
-the area of parameterized algorithms for this
-problem\tnote[TH]{elaborate - what are the results?}.
-
-Another problem is to find the minimum number of entries in the matrix
-that can be toggled to result in a matrix with COP.  \cite{v85}
-discusses approximation of {\sc COP Augmentation} which is the problem
-of changing of the minimum number of zero entries to {\un}s so that
-the resulting matrix has COP. As mentioned earlier, this problem is
-known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
-using a reduction to the longest path problem, \tnote[TH]{or is it a
-  survey of another result?  check.} that finding a Tucker's forbidden
-submatrix of at least $k$ rows is NP complete. \tnote[TH]{how is this
-  different from booth's 75 result??}  \tnote[TH]{where should this
-  go? cite|tz04 (approx submatrix with COP sparse matrices)}
-
-\cite{jkckv04} discusses the use of matrices with almost-COP (instead
-of one block of consecutive {\un}s, they have $x$ blocks, or {\em
-  runs}, of consecutive {\un}s and $x$ is not too large) in the
-storage of very large databases.  The problem is that of reordering of
-a binary matrix such that the resulting matrix has at most $k$ runs of
-{\un}s. This is proved to be NP hard using a reduction from the
-Hamiltonian path problem.\tnote[TH]{Theorem 2.1 in jkckv}
-\tnote[TH]{(1) A connection of COP problem to the travelling salesman
-  problem is also introduced. what does this mean? -- COP can be used
-  as a tool to reorder $0.5T \le runs(M) \le T$. (2) The optimization
-  version of the $k$-run problem, i.e. minimization of number of
-  blocks of ones is proven to be NP complete by
-  cite:k77}\tnote[TH]{are these two the same?} \tnote[TH]{what is the
-  reduction?} \tnote[TH]{other problems similar to COP -- cite:ckl96
-  (ILP, circ ones, one drop) -- cite:th98 (generalization of COP -
-  minimax, biotonic column) Tucker}
-
-\section{********* Application of COP in Areas of Graph Theory and
-  Algorithms}
-\label{sec:appcopGTA}
-
-\subsection{********* COP in Relational Database Model}
-\label{sec:apprdbm}
-\tnote{(set systems theme)}
-
-\subsection{********* COP in Graph Isomorphism}
-\label{sec:appgraphiso}
-\tnote{(canonization theme)}
-
-\subsection{********* Certifying Algorithms}
-\label{sec:appcertalgo}
-\tnote{ (certification McC04 theme)}
-
-\section{Generalization of COP - the Motivation}
-\label{sec:motive}
-Section~\ref{sec:copmatrices} introduced a succinct characterization
-for consecutive-ones property which is solely based on the cardinality
-properties of the set representations of the matrix's columns
-\cite{nsnrs09}. This result is very relevant to this thesis because
-aside from it simplifying COT to a great degree, our generalization
-problem is motivated by their results.
-
-\cite{nsnrs09} characterizes interval assignments to the sets which
-can be obtained from a single permutation of the rows.  For an
-assignment to be feasible, the cardinality of the interval assigned to
-each set in the system must be same as the cardinality of the set, and
-the intersection cardinality of any two intervals must be same as the
-intersection cardinality of their corresponding sets.  While this is
-obviously a necessary condition, this result shows this is also
-sufficient.  \cite{nsnrs09} calls this an Intersection Cardinality
-Preserving Interval Assignment (ICPIA).  This paper generalizes the
-idea from \cite{wlh02} of decomposing a given binary matrix into prime
-matrices for COT and describes an algorithm to test if an ICPIA exists
-for a given set system.
-
-The equivalence of the problem of testing for the consecutive-ones
-property to the constraint statisfaction problem of interval
-assignment \cite{nsnrs09} or interval labeling \cite{kklv10} is as
-follows. Every column (row) of the binary matrix can be converted into
-a set of non-negative integers which are the indices of rows (columns)
-with {\un}s in that column (row). It is apparent that if the matrix
-has COP in columns (rows), then constructing such sets after applying
-the COP permutation to the rows (columns) of the matrix will result in
-sets with consecutive integers. In other words, after application of
-COP reordering, the sets are intervals. Indeed the problem now becomes
-finding interval assignments to a given set system such that there
-exists a permutation of the universe of set of row indices (column
-indices) which converts each set to its assigned interval.
-
-The problem of interest in this thesis, namely, tree path labeling
-problem, is a natural generalization of the interval assignment
-problem or the COT problem. The problem is defined as follows -- given
-a set system $\cF$ from a universe $U$ and a target tree $T$, does
-there exist a bijection from $U$ to the vertices of $T$ such that each
-set in the system maps to a path in $T$.  We refer to this as the
-{\CFTPL} problem or simply {\em tree path labeling} problem for an
-input set system and target tree pair -- $(\cF,T)$. The special case
-of the target tree being a path, is the interval assignment problem.
-We focus on generalizing the notion of an ICPIA \cite{nsnrs09} to
-characterize feasible path assignments.  We show that for a given set
-system $\cF$, a tree $T$, and an assignment of paths from $T$ to the
-sets, there is a feasible\footnote{The notion of {\em feasibility} is
-  formally defined in Section~\ref{ch:prelims}.}  bijection between
-$U$ and $V(T)$ if and only if the intersection cardinalities among any
-three sets (not necessarily distinct) is equal to that of the
-corresponding paths assigned to them and the input passes a filtering
-algorithm (described in this paper) successfully.  This algorithmic
-characterization gives a natural data structure that stores all the
-\remove{relevant} feasible bijections between $U$ and $V(T)$. This
-reduces the search space for the solution considerably from the
-universe of all possible bijections between $U$ and $V(T)$ to only
-those bijections that maintain the characterization.  Further, the
-filtering algorithm is also an efficient algorithm to test if a tree
-path labeling\footnote{The terms {\em tree path labeling} and {\em
-    tree path assignment} are, in informal language,
-  synonyms. Formally, the former refers to the bijection $\cl: \cF
-  \rightarrow \cP$. The latter refers to the set of ordered pairs
-  $\{(S, P) \mid S \in \cF, P \in \cP\}$. $\cP$ is a set of paths on
-  $T$.} is feasible.
-
-\section{Summary of New Results in this Thesis}
-\label{sec:results}
-
-We see in Section~\ref{sec:motive} that pairwise intersection
-cardinality preservation is necessary and sufficient for an interval
-assignment to be feasible for a given hypergraph\footnote{A {\em
-    hypergraph} is an alternate representation of a set system and
-  will be used in this thesis. See Section~\ref{ch:prelims} for the
-  formal definition.} and thus is a characterization for COP
-\cite{nsnrs09}. \tnote{Refer to the survey section that will list the
-  theorems and/or lemma?} In our work we extend this characterization
-and find that trio-wise intersection cardinality preservation makes a
-tree path labeling \footnote{A {\em tree path labeling} $\cl$ is a
-  bijection of paths from the target tree $T$ to the hyperedges in
-  given hypergraph $\cF$.  See Section~\ref{ch:prelims} for the formal
-  definition.}  (TPL) feasible, which is a generalization of the COP
-problem. This problem is defined as follows.
-
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \FTPL}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
-        paths $\cP$ from $T$ and a
-        bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
-      \end{minipage}\\
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
-        such that $\phi$ when applied on any hyperedge in $\cF$ will
-        give
-        the path mapped to it by the given tree path labeling $\cl$.\\
-        { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every
-          hyperedge $S \in \cF$.}
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
-
-We give a necessary and sufficient condition by way of {\em
-  Intersection Cardinality Preservation Path Labeling} (ICPPL) and a
-filtering algorithm for {\FTPL} to output in affirmative. ICPPL
-captures the trio-wise cardinality property described
-earlier\footnote{See Section~\ref{sec:feasible} for the definition of
-  ICPPL.}. This characterization can be checked in polynomial time.  A
-relevant consequence of this constructive procedure is that it is
-sufficient to iteratively check if three-way intersection
-cardinalities are preserved.  In other words, in each iteration, it is
-sufficient to check if the intersection of any three hyperedges is of
-the same cardinality as the intersection of the corresponding paths.
-Thus this generalizes the well studied question of the feasible
-interval assignment problem which is the special case when the target
-tree $T$ is simply a path \cite{wlh02,nsnrs09}.
-
-Aside from checking if a given TPL is feasible, we also solve the
-problem of computing a feasible TPL for a given hypergraph and target
-tree, if one exists. This problem, {\CFTPL}, is defined as follows.
-
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \CFTPL}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$ and a tree $T$.\\
-      \end{minipage}\\
-
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a set of paths $\cP$ from $T$ and a bijection
-        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
-        {\bf true} on $(\cF, T, \cl)$.
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
-
-We present a polynomial time algorithm for {\CFTPL} when the target
-tree $T$ belongs to a special class of trees called {\em \kstar s} and
-when the hyperedges in the hypergraph $\cF$ have at most $k+2$
-vertices. A couple of examples of {\kstar s} are given in
-Figure~\ref{fig:ksubstar}.
-
-\begin{figure}[t]
-  \centering
-  \begin{tabular}[h]{ccccc}
-    \includegraphics[scale=0.3]{../img/star.pdf} &&&&
-    \includegraphics[scale=0.3]{../img/kstar.pdf}\\
-    (a) &&&& (b)
-  \end{tabular}
-  \caption{\figtabsize Examples of {\kstar s}. (a) $k = 0$ (b) $k = 2$
-  }
-  \label{fig:ksubstar}
-\end{figure}
-
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \CFTPLKTREE}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$ such that every
-        hyperedge
-        $S \in \cF$ is of cardinality at most $k+2$ and a {\kstar} $T$.\\
-      \end{minipage}\\
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a set of paths $\cP$ from $T$ and a bijection
-        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
-        {\bf true} on $(\cF, T, \cl)$.
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
-
-\tnote{The following sentence seems out of place in this para} In
-spite of this being a restricted case, we believe that our results are
-of significant interest in understanding the nature of {\sc Graph
-  Isomorphism} which is polynomial time solvable in interval graphs
-while being hard on path graphs\cite{kklv10}. {\kstar s} are a class
-of trees which are in many ways very close to intervals or paths. Each
-ray\footnote{The path from a leaf to the root, the vertex with highest
-  degree, is called a {\em ray} of the \kstar. See
-  Section~\ref{ch:prelims}.} are independent except for the
-root\footnote{The vertex with maximum degree in a {\kstar} is called
-  {\em root}. See Section~\ref{ch:prelims}.} and hence can be
-considered as an independent interval till the root. Our algorithm
-builds on this fact and uses the interval assignment
-algorithm\cite{nsnrs09} up until ``reaching'' the root and then uses
-the trio-wise intersection cardinality (the extra condition in ICPPL
-that generalizes ICPIA) check to resolve the ambiguity about which ray
-the algorithm should ``grow'' the solution into in the next iteration.
-
-We also have an algorithm for solving {\CFTPL} with no restrictions on
-the target tree or set size which runs in exponential time.  This
-algorithm finds a path labeling from $T$ by decomposing the problem
-into subproblems of finding path labeling of subsets of $\cF$ from
-subtrees of $T$. Given the fact that binary matrices naturally
-represent a set system (see Section~\ref{sec:motive}) and that the
-{\em overlap}\footnote{}\tnote{Give informal definition of ``overlap''
-  in footnote and link to sec:prelims. ADD the definition in prelim.}
-relation between the sets involved is an obvious equivalence relation,
-$\cF$ quite naturally partitions into equivalence classes known as
-{\em overlap components}\tnote{Give informal definition of ``overlap
-  components'' in footnote and link to sec:prelims. ADD the definition
-  in prelim.}. In the context of COP, overlap components were used in
-\cite{wlh02} and \cite{kklv10}. Moreover, \cite{nsnrs09} discovered
-that these equivalence classes form a total order \tnote{Give informal
-  definition of ``total order'' in footnote?}. We extend this to TPL
-and find that when $\cF$ is a path hypergraph\footnote{If there exists
-  an FTPL for a hypergraph $\cF$, it is called a path hypergraph.},
-the classes can be partially ordered as an in-tree in polynomial
-time. Once $\cF$ is ``broken'' into overlap components, one must
-identify the subtree of $T$ that it needs to map to and this is the
-hard part which is currently open to be solved in polynomial time.
-
-\tnote[TH]{The connection of TPL to graph isomorphism will be made
-  later in the document}
-
-\chapter{Consecutive-ones Property - a Survey of Important Results}
-\label{ch:copsurvey}
-\tnote{TBD: have a few lines about organization of chapter}
-
-\tnote{ Survey chapter: is the full fledged expansion of the survey in
-  introduction with details, observations, theorems etc.}
-
-\section{Matrices with COP}
-\label{sec:surveycoptest}
-\tnote{Expand on ref:sec:copmatrices}
-
-\section{Optimization problems in COP}
-\label{sec:surveycopopt}
-\tnote{Expand on ref:sec:optcop}
-
-\section{********* COP in Relational Database Model}
-\label{sec:surveyrdbm}
-\tnote{Expand on sec:apprdbm}
-\tnote{(set systems theme)}
-
-\section{********* COP in Graph Isomorphism}
-\label{sec:surveygraphiso}
-\tnote{Expand on sec:appgraphiso}
-\tnote{(canonization theme)}
-
-\section{********* Certifying Algorithms}
-\label{sec:surveycertalgo}
-\tnote{Expand on sec:appcertalgo}
-\tnote{ (certification McC04 theme)}
-
 \chapter{Tree Path Labeling of Path Hypergraphs - the New Results}
@@
 }
-%\end{tabular}
+% \end{tabular}
 
@@
 it. Hence we consider two special cases of \CFTPL -- namely, \CFTPLINT
-and \CFTPLKTREE on special target trees, intervals and \kstar, respectively.
-Section~\ref{sec:spltargettree}
-discusses these problems.
+and \CFTPLKTREE on special target trees, intervals and \kstar,
+respectively.  Section~\ref{sec:spltargettree} discusses these
+problems.
 
@@
 
-\CFTPLINT is nothing but the
-consecutive-ones property testing problem.
+\CFTPLINT is nothing but the consecutive-ones property testing
+problem.
 
@@
 
-\remove{We focus on the question of generalizing the notion of an ICPIA
-cite-nsnrs09 to characterize feasible path assignments.  We show
-that for a given set system $\cF$, a tree $T$, and an assignment of
-paths from $T$ to the sets, there is a feasible bijection between $U$
-and $V(T)$ if and only if all intersection cardinalities among any
-three sets (not necessarily distinct) is same as the intersection
-cardinality of the paths assigned to them and the input runs a
-filtering algorithm (described in this paper) successfully.  This
-characterization is proved constructively and it gives a natural data
-structure that stores all the relevant feasible bijections between $U$
-and $V(T)$.  Further, the filtering algorithm is also an efficient
-algorithm to test if a tree path labeling to the set system is
-feasible.  This generalizes the result in cite-nsnrs09.}
+\remove{We focus on the question of generalizing the notion of an
+  ICPIA cite-nsnrs09 to characterize feasible path assignments.  We
+  show that for a given set system $\cF$, a tree $T$, and an
+  assignment of paths from $T$ to the sets, there is a feasible
+  bijection between $U$ and $V(T)$ if and only if all intersection
+  cardinalities among any three sets (not necessarily distinct) is
+  same as the intersection cardinality of the paths assigned to them
+  and the input runs a filtering algorithm (described in this paper)
+  successfully.  This characterization is proved constructively and it
+  gives a natural data structure that stores all the relevant feasible
+  bijections between $U$ and $V(T)$.  Further, the filtering algorithm
+  is also an efficient algorithm to test if a tree path labeling to
+  the set system is feasible.  This generalizes the result in
+  cite-nsnrs09.}
 
@@
 a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
-\cite{kklv10}, written as $B=\phi(A)$.
-This is called {\em hypergraph isomorphism}.
+\cite{kklv10}, written as $B=\phi(A)$.  This is called {\em hypergraph
+  isomorphism}.
 
@@
 
-\add{Due to the equivalence of set system and hypergraph in the scope of
-this paper, we drop the subscript $_H$ in the notation and refer to
-both the structures by $\cF$.}
+\add{Due to the equivalence of set system and hypergraph in the scope
+  of this paper, we drop the subscript $_H$ in the notation and refer
+  to both the structures by $\cF$.}
 
@@
   path labeling} of the hypergraph $\cF$.  Note that there are two
-kinds of isomorphisms here.  One is the isomorphism of
-intersection graphs on $\cF$ and $\cP$, i.e. $\bI(\cF)$ and $\bI(\cP)$
-respectively. Second is the isomorphism
-between the hypergraphs $\cF$ and $\cP$.
-
-\add{To illustrate further, let $\cg: V(\cF) \rightarrow V(\cP)$ be the
-above mentioned isomorphism where $V(\cF)$ and $V(\cP)$ are the
+kinds of isomorphisms here.  One is the isomorphism of intersection
+graphs on $\cF$ and $\cP$, i.e. $\bI(\cF)$ and $\bI(\cP)$
+respectively. Second is the isomorphism between the hypergraphs $\cF$
+and $\cP$.
+
+\add{To illustrate further, let $\cg: V(\cF) \rightarrow V(\cP)$ be
+  the above mentioned isomorphism where $V(\cF)$ and $V(\cP)$ are the
   vertex sets that represent the hyperedges for each hypergraph
@@
 
-If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$
-is called a {\em path hypergraph} and $\cP$ is called {\em path
-  representation} of $\cF$. If this isomorphism is $\phi: supp(\cF)
-\rightarrow V(T)$, then it is clear that there is an {\em induced path
-labeling} $\cl_\phi: \cF \rightarrow \cP$ to the set system;
-$\cl_\phi(S) = \set{y \mid y = \phi(x), x \in S}$ for all $S \in \cF$. Recall that
+If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$ is called
+a {\em path hypergraph} and $\cP$ is called {\em path representation}
+of $\cF$. If this isomorphism is $\phi: supp(\cF) \rightarrow V(T)$,
+then it is clear that there is an {\em induced path labeling}
+$\cl_\phi: \cF \rightarrow \cP$ to the set system; $\cl_\phi(S) =
+\set{y \mid y = \phi(x), x \in S}$ for all $S \in \cF$. Recall that
 $supp(\cP) = V(T)$.
 
-\add{A graph} $G$ is a {\em path graph} if it is
-isomorphic to the intersection graph $\bI(\cP)$ of a path system
-$\cP$.  This isomorphism gives a bijection $\cl': V(G) \rightarrow
-\cP$. Moreover, for the purposes of this paper, we require that in a
-path labeling, $supp(\cP) = V(T)$.  If graph $G$ is also isomorphic
-to $\bI(\cF)$ for some hypergraph $\cF$, then clearly there is a
-bijection $\cl: \cF \rightarrow \cP$ such that $\cl(S) = \cl'(v_S)$
-where $v_S$ is the vertex corresponding to set $S$ in $\bI(\cF)$ for
-any $S \in \cF$. This bijection $\cl$ is called the {\em path
-  labeling} of the hypergraph $\cF$ and the path system $\cP$ may be
-alternatively denoted as $\cF^\cl$.
-
-A path labeling $(\cF, \cl)$ is defined to be {\em
-  feasible} if
-$\cF \cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
+\add{A graph} $G$ is a {\em path graph} if it is isomorphic to the
+intersection graph $\bI(\cP)$ of a path system $\cP$.  This
+isomorphism gives a bijection $\cl': V(G) \rightarrow \cP$. Moreover,
+for the purposes of this paper, we require that in a path labeling,
+$supp(\cP) = V(T)$.  If graph $G$ is also isomorphic to $\bI(\cF)$ for
+some hypergraph $\cF$, then clearly there is a bijection $\cl: \cF
+\rightarrow \cP$ such that $\cl(S) = \cl'(v_S)$ where $v_S$ is the
+vertex corresponding to set $S$ in $\bI(\cF)$ for any $S \in
+\cF$. This bijection $\cl$ is called the {\em path labeling} of the
+hypergraph $\cF$ and the path system $\cP$ may be alternatively
+denoted as $\cF^\cl$.
+
+A path labeling $(\cF, \cl)$ is defined to be {\em feasible} if $\cF
+\cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
 \rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
-\rightarrow \cF^\cl$ such that $\cl_\phi = \cl$. 
-\add{In this work}, we are given as input $\cF$ and a tree $T$, and the
-question is whether there is a path labeling $\cl$ to a set of paths
-in $T$.  We refer to such a solution path system by $\cF^\cl$.  A path
-labeling $(\cF, \cl)$ is defined to be {\em feasible} if
+\rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.  \add{In this work},
+we are given as input $\cF$ and a tree $T$, and the question is
+whether there is a path labeling $\cl$ to a set of paths in $T$.  We
+refer to such a solution path system by $\cF^\cl$.  A path labeling
+$(\cF, \cl)$ is defined to be {\em feasible} if
 % $\cF \cong \cF^\cl$ and this
@@
 
-{\bf Overlap Graphs and Marginal Hyperedges:} An {\em
-  overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is a graph such
-that its vertex set has a bijection to $\cF$ and there exists an edge
-between two of its vertices iff their corresponding hyperedges
-overlap. Two hyperedges $S$ and $S'$ are said to {\em overlap},
-denoted by $S \overlap S'$, if they have a non-empty intersection and
-neither is contained in the other; $S \overlap S' \text{ iff } S \cap
-S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq S$. Thus $\bO(\cF)$
-is a spanning subgraph of $\bI(\cF)$ and not necessarily
-connected. Each connected component of $\bO(\cF)$ is called an {\em
-  overlap component}.
+{\bf Overlap Graphs and Marginal Hyperedges:} An {\em overlap graph}\,
+$\bO(\cF)$ of a hypergraph $\cF$ is a graph such that its vertex set
+has a bijection to $\cF$ and there exists an edge between two of its
+vertices iff their corresponding hyperedges overlap. Two hyperedges
+$S$ and $S'$ are said to {\em overlap}, denoted by $S \overlap S'$, if
+they have a non-empty intersection and neither is contained in the
+other; $S \overlap S' \text{ iff } S \cap S' \ne \emptyset, S
+\nsubseteq S', S' \nsubseteq S$. Thus $\bO(\cF)$ is a spanning
+subgraph of $\bI(\cF)$ and not necessarily connected. Each connected
+component of $\bO(\cF)$ is called an {\em overlap component}.
 
@@
 
-{\bf $k$-subdivided star -- a special tree} A {\em star} graph
-is a complete bipartite graph $K_{1,p}$ which is clearly a tree and
-$p$ is the number of leaves. The vertex with maximum degree is called
-the {\em center} of the star and the edges are called {\em rays} of
-the star. A {\em $k$-subdivided star} is a star with all its rays
+{\bf $k$-subdivided star -- a special tree} A {\em star} graph is a
+complete bipartite graph $K_{1,p}$ which is clearly a tree and $p$ is
+the number of leaves. The vertex with maximum degree is called the
+{\em center} of the star and the edges are called {\em rays} of the
+star. A {\em $k$-subdivided star} is a star with all its rays
 subdivided exactly $k$ times. The definition of a {\em ray of a
@@
 \begin{proof}%[Proof of Lemma~\ref{lem:setminuscard}]
-  Let $P_i = \cl(S_i)$, for all $1 \le i \le  3$.
-  $|S_1 \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| =
-  |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to properties (ii)
-  and (iii) of ICPPL, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
-  \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
-  = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. \qed
+  Let $P_i = \cl(S_i)$, for all $1 \le i \le 3$.  $|S_1 \cap (S_2
+  \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| = |S_1 \cap S_2| -
+  |S_1 \cap S_2 \cap S_3|$. Due to properties (ii) and (iii) of ICPPL,
+  $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1 \cap P_2| - |P_1
+  \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3| = |P_1 \cap (P_2
+  \setminus P_3)|$. Thus lemma is proven. \qed
 \end{proof}
 
-In the remaining part of this section we show that $(\cF,
-\cl)$ is feasible if and only if it is an ICPPL and
+In the remaining part of this section we show that $(\cF, \cl)$ is
+feasible if and only if it is an ICPPL and
 Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
-function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two levels of
-filtering of $(\cF, \cl)$ to make it simpler while retaining the set
-of isomorphisms, if any, between $\cF$ and $\cF^\cl$.
-% One direction of this claim isclear: that if a path
-% labeling is feasible, then all intersection cardinalities are
-% preserved, i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
+function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two
+levels of filtering of $(\cF, \cl)$ to make it simpler while retaining
+the set of isomorphisms, if any, between $\cF$ and $\cF^\cl$.
+% One direction of this claim isclear: that if a path labeling is
+% feasible, then all intersection cardinalities are preserved,
+% i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
 % \annote{has no premature exit condition hence any input will go
@@
 % feasible path labeling has an associated bijection between
-% $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that
-% the sets map to paths, ``preserving'' the path labeling.  The rest
-% of the section is devoted to constructively proving that it is
+% $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that the
+% sets map to paths, ``preserving'' the path labeling.  The rest of
+% the section is devoted to constructively proving that it is
 % sufficient for a path labeling to be an ICPPL and pass the two
@@
       $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
-    \{S_1, S_2\})
-    \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1 \}$ \label{setbreak} 
-    \COMMENT {Remove $S_1$, $S_2$ and add the ``filtered'' sets}
-    \STATE {\bf for} every $S \in \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
+    \{S_1, S_2\}) \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2
+    \setminus S_1 \}$ \label{setbreak} \COMMENT {Remove $S_1$, $S_2$
+      and add the ``filtered'' sets} \STATE {\bf for} every $S \in
+    \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
     S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\
@@
     \STATE $\cl_j(S_1 \cap S_2) \assign \cl_{j-1}(S_1) \cap
-    \cl_{j-1}(S_2)$
-    \COMMENT {Carry forward the path labeling for all existing sets other than
-      $S_1$, $S_2$}
-    \STATE $\cl_j(S_1 \setminus S_2) \assign \cl_{j-1}(S_1) \setminus
-    \cl_{j-1}(S_2)$ 
-    \COMMENT {Define path labeling for new sets}
-    \STATE $\cl_j(S_2 \setminus S_1) \assign \cl_{j-1}(S_2) \setminus
-    \cl_{j-1}(S_1)$
-
-    \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of ICPPL}
+    \cl_{j-1}(S_2)$ \COMMENT {Carry forward the path labeling for all
+      existing sets other than $S_1$, $S_2$} \STATE $\cl_j(S_1
+    \setminus S_2) \assign \cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)$
+    \COMMENT {Define path labeling for new sets} \STATE $\cl_j(S_2
+    \setminus S_1) \assign \cl_{j-1}(S_2) \setminus \cl_{j-1}(S_1)$
+
+    \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of
+      ICPPL}
     \label{ln:3waycheck}
@@
 
-\end{proof}
-{\tt WALCOM end}
+\end{proof} {\tt WALCOM end}
 
@@
   in $j$th iteration in line~\ref{setbreak} of Algorithm~\ref{perms},
-  $S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1$ and its
-  images in $\cl_j$ where $\cl_{j-1}(S_1)$
+  $S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1$ and its images
+  in $\cl_j$ where $\cl_{j-1}(S_1)$
   and $\cl_{j-1}(S_2)$ intersect and share a leaf.\\
@@
   is the input ICPPL.  Assume the lemma is true till the $j-1$th
-  iteration. Let us consider the possible cases for each of the above invariants for
-  the $j$th iteration.
+  iteration. Let us consider the possible cases for each of the above
+  invariants for the $j$th iteration.
 
@@
       $R = S_1 \cap S_2$, $|R| = |S_1 \cap S_2| = |\cl_{j-1}(S_1) \cap
-      \cl_{j-1}(S_2)| = |\cl_j(S_1 \cap S_2)| = |\cl_j(R)|$.
-      If $R = S_1 \setminus S_2$, $|R| = |S_1 \setminus S_2| = |S_1| -
-      |S_1 \cap S_2| = |\cl_{j-1}(S_1)| - |\cl_{j-1}(S_1) \cap
+      \cl_{j-1}(S_2)| = |\cl_j(S_1 \cap S_2)| = |\cl_j(R)|$.  If $R =
+      S_1 \setminus S_2$, $|R| = |S_1 \setminus S_2| = |S_1| - |S_1
+      \cap S_2| = |\cl_{j-1}(S_1)| - |\cl_{j-1}(S_1) \cap
       \cl_{j-1}(S_2)| = |\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)| =
@@
 
-    \item {\em $R$ and $R'$ are new sets.} By definition, the new
-      sets and their path images in path label $\cl_j$ are disjoint so
-      $|R \cap R'| = |\cl_j(R) \cap \cl_j(R)| = 0$. Thus case proven.
+    \item {\em $R$ and $R'$ are new sets.} By definition, the new sets
+      and their path images in path label $\cl_j$ are disjoint so $|R
+      \cap R'| = |\cl_j(R) \cap \cl_j(R)| = 0$. Thus case proven.
     \end{enumerate}
@@
     % \begin{enumerate} [{Case 3.}1:]
-%     \item {\em $R$, $R'$ and $R''$ are not new sets.} Trivially
-%       true by induction hypothesis.
-%     \item {\em Only one, say $R$, is a new set.}
-%       If $R = S_1 \cap S_2$,  from Lemma~\ref{lem:fourpaths} and
-%       invariant III hypothesis,  this case is proven. Similarly if $R$
-%       is any of the other new  sets, the case is proven by also using
-%       Lemma ~\ref{lem:setminuscard}.
-%     \item {\em At least two of $R, R', R''$ are new sets.}
-%       The new sets are disjoint hence this case is vacuously true.
+    % \item {\em $R$, $R'$ and $R''$ are not new sets.} Trivially true
+    %   by induction hypothesis.
+    % \item {\em Only one, say $R$, is a new set.}  If $R = S_1 \cap
+    %   S_2$, from Lemma~\ref{lem:fourpaths} and invariant III
+    %   hypothesis, this case is proven. Similarly if $R$ is any of
+    %   the other new sets, the case is proven by also using Lemma
+    %   ~\ref{lem:setminuscard}.
+    % \item {\em At least two of $R, R', R''$ are new sets.}  The new
+    %   sets are disjoint hence this case is vacuously true.
     % \end{enumerate}
   \end{enumerate} \qed
-%\vspace{-6mm}
+  % \vspace{-6mm}
 
@@
   Since $(\cF,\cl)$ is feasible, by Lemma~\ref{lem:feasible}
-  $(\cF_j,\cl_j)$ for every iteration $j > 0$ is feasible.  % Therefore,
-%   every hypergraph isomorphism $\phi: supp(\cF) \rightarrow V(T)$ that
-%   induces $\cl$ on $\cF$ also induces $\cl_{j-1}$ and $\cl_{j}$ on
-%   $\cF_{j-1}$ and $\cF_{j}$ respectively, i.e., $\cl_{\phi[\cF_{j-1}]}
-%   = \cl_{j-1}$ and $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
-%   that for all $x \in supp(\cF)$, for all $v \in V(T)$ the following
-%   hold true.
+  $(\cF_j,\cl_j)$ for every iteration $j > 0$ is
+  feasible.  % Therefore,
+  % every hypergraph isomorphism $\phi: supp(\cF) \rightarrow V(T)$
+  % that induces $\cl$ on $\cF$ also induces $\cl_{j-1}$ and $\cl_{j}$
+  % on $\cF_{j-1}$ and $\cF_{j}$ respectively, i.e.,
+  % $\cl_{\phi[\cF_{j-1}]} = \cl_{j-1}$ and $\cl_{\phi[\cF_j]} =
+  % \cl_j$. Thus it can be seen that for all $x \in supp(\cF)$, for
+  % all $v \in V(T)$ the following hold true.
   Also, every hypergraph isomorphism $\phi: supp(\cF) \rightarrow
@@
   \in S$.
-% the following
-%   hold true.
+  % the following hold true.
   % \begin{enumerate}[i. ]
@@
   % first iterations of while. Clearly, $\phi$ is a renaming of
-%   vertices in hypergraph $\cF$ to those in hypergraph $\cF^\cl$. Thus
-%   the following facts can be observed in every iteration of the loop.
+  % vertices in hypergraph $\cF$ to those in hypergraph
+  % $\cF^\cl$. Thus the following facts can be observed in every
+  % iteration of the loop.
 
@@
   %   labeling \vspace{\topshrink}
-%   \item element $x$ is exclusive in a hyperedge in $\cF$ since $v$ is
-%     exclusive in a hyperedge in $\cF^\cl$.
+  % \item element $x$ is exclusive in a hyperedge in $\cF$ since $v$
+  %   is exclusive in a hyperedge in $\cF^\cl$.
   % \end{enumerate}
 
-%   Thus the exit condition is never rendered true after $x$ and $v$ are
-%   removed from their respective hyperedges. \qed
+  % Thus the exit condition is never rendered true after $x$ and $v$
+  % are removed from their respective hyperedges. \qed
 
   % \noindent
-% This proof uses mathematical induction on the number
-%   of iterations $j$, $j \ge 0$, of the loop that executed
-%   without exiting. The base case, $j = 0$ is obviously true since the
-%   input is an ICPPL and the exit condition cannot hold true due to
-%   ICPPL property (iii).  Assume the algorithm executes till the end
-%   of $j-1$th iteration without exiting at line
-%  ~\ref{ln:3waycheck}. Consider the $j$th iteration. From Lemma
-%  ~\ref{lem:feasible} we know that $(\cF_j, \cl_j)$ and $(\cF_{j-1},
-%   \cl_{j-1})$ are feasible\remove[AS]{and from the proof in lemma
-%     lem:invar1 we know that $(\cF_{j-1}, \cl_{j-1})$ satisfies all the
-%     invariants defined in the lemma}.  Thus there exists a bijection
-%   $\phi: supp(\cF) \rightarrow V(T)$ such that the induced path
+  % This proof uses mathematical induction on the number of iterations
+  % $j$, $j \ge 0$, of the loop that executed without exiting. The
+  % base case, $j = 0$ is obviously true since the input is an ICPPL
+  % and the exit condition cannot hold true due to ICPPL property
+  % (iii).  Assume the algorithm executes till the end of $j-1$th
+  % iteration without exiting at line ~\ref{ln:3waycheck}. Consider
+  % the $j$th iteration. From Lemma ~\ref{lem:feasible} we know that
+  % $(\cF_j, \cl_j)$ and $(\cF_{j-1}, \cl_{j-1})$ are
+  % feasible\remove[AS]{and from the proof in lemma lem:invar1 we know
+  %   that $(\cF_{j-1}, \cl_{j-1})$ satisfies all the invariants
+  %   defined in the lemma}.  Thus there exists a bijection $\phi:
+  % supp(\cF) \rightarrow V(T)$ such that the induced path
   %   % labeling on $\cF_{j-1}$ $\cl_{\phi[\cF_{j-1}]} = \cl_{j-1}$.
@@
   % \cF_{j}$, $|R \cap R' \cap R''| = |\cl_j(R) \cap \cl_j(R') \cap
-%   \cl_j(R'')|$.
-%   The following are the possible cases that could arise. From argument
-%   above, $|\cl_j(R) \cap \cl_j(R') \cap \cl_j(R'')| =
-%   |\cl_{\phi[\cF_{j}]}(R) \cap \cl_{\phi[\cF_{j}]} (R') \cap
-%   \cl_{\phi[\cF_{j}]} (R'')|$
+  % \cl_j(R'')|$.  The following are the possible cases that could
+  % arise. From argument above, $|\cl_j(R) \cap \cl_j(R') \cap
+  % \cl_j(R'')| = |\cl_{\phi[\cF_{j}]}(R) \cap \cl_{\phi[\cF_{j}]}
+  % (R') \cap \cl_{\phi[\cF_{j}]} (R'')|$
 
@@
   %   We know $(\cF_{j-1}, \cl_{j-1})$ is feasible. Thus $|R \cap R'
-%     \cap R''| = |\cl_{j-1}(R) \cap \cl_{j-1}(R') \cap \cl_{j-1}(R'')|
-%     = |\cl_{j}(R) \cap \cl_{j}(R') \cap \cl_{j}(R'')|$.
-%   \item {\em Only one, say $R$, is a new set.}  Let $R = S_1 \cap S_2$
-%     ($S_1, S_2$ are defined in the proof of lemma
-%    ~\ref{lem:invar1}). Now we have $|R \cap R' \cap R''| = |S_1 \cap
-%     S_2 \cap R' \cap R''| = |\cl_{j-1}(S_1) \cap \cl_{j-1}(S_2) \cap
-%     \cl_{j-1}(R') \cap \cl_{j-1}(R'')| = |\cl_{j}(R) \cap \cl_{j}(R')
-%     \cap \cl_{j}(R'')|$. Thus proven. If $R$ is any of the other new
-%     sets, the same claim can be verified using lemma
+  %   \cap R''| = |\cl_{j-1}(R) \cap \cl_{j-1}(R') \cap
+  %   \cl_{j-1}(R'')| = |\cl_{j}(R) \cap \cl_{j}(R') \cap
+  %   \cl_{j}(R'')|$.
+  % \item {\em Only one, say $R$, is a new set.}  Let $R = S_1 \cap
+  %   S_2$ ($S_1, S_2$ are defined in the proof of lemma
+  %   ~\ref{lem:invar1}). Now we have $|R \cap R' \cap R''| = |S_1
+  %   \cap S_2 \cap R' \cap R''| = |\cl_{j-1}(S_1) \cap \cl_{j-1}(S_2)
+  %   \cap \cl_{j-1}(R') \cap \cl_{j-1}(R'')| = |\cl_{j}(R) \cap
+  %   \cl_{j}(R') \cap \cl_{j}(R'')|$. Thus proven. If $R$ is any of
+  %   the other new sets, the same claim can be verified using lemma
   %   ~\ref{lem:setminuscard}.
   %     % \item []{\bf Case 3:}
-%   \item {\em At least two of $R, R', R''$ are new sets.}  The new sets
-%     are disjoint hence this case is vacuously true.
+  % \item {\em At least two of $R, R', R''$ are new sets.}  The new
+  %   sets are disjoint hence this case is vacuously true.
   % \end{enumerate}
-%   \qed
-%   \tnote[E2]{remove the induction proof. just text saying x and v are
-%     exclusive in these sets therefore the intersection cardinalities
-%     don't change thus all invariants are still true} 
+  % \qed \tnote[E2]{remove the induction proof. just text saying x and
+  %   v are exclusive in these sets therefore the intersection
+  %   cardinalities don't change thus all invariants are still true}
 \end{proof}
 
-As a result of Algorithm~\ref{perms} each leaf $v$ in $T$
-is such that there is exactly one set in $\cF$ with $v$ as a vertex in
-the path assigned to it.  In Algorithm~\ref{leafasgn} we identify
-elements in $supp(\cF)$ whose images are leaves in a hypergraph
-isomorphism if one exists.  Let $S \in \cF$ be such that $\cl(S)$ is a
-path with leaf and $v \in V(T)$ is the unique leaf incident on it.  We
-define a new path labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) =
-\set{v}$ where $x$ an arbitrary element from $S \setminus \bigcup_{\hS
-  \ne S} \hS$. In other words, $x$ is an element present in no other
-set in $\cF$ except $S$. This is intuitive since $v$ is present in no
-other path image under $\cl$ other than $\cl(S)$.  The element $x$ and
-leaf $v$ are then removed from the set $S$ and path $\cl(S)$
+As a result of Algorithm~\ref{perms} each leaf $v$ in $T$ is such that
+there is exactly one set in $\cF$ with $v$ as a vertex in the path
+assigned to it.  In Algorithm~\ref{leafasgn} we identify elements in
+$supp(\cF)$ whose images are leaves in a hypergraph isomorphism if one
+exists.  Let $S \in \cF$ be such that $\cl(S)$ is a path with leaf and
+$v \in V(T)$ is the unique leaf incident on it.  We define a new path
+labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) = \set{v}$ where
+$x$ an arbitrary element from $S \setminus \bigcup_{\hS \ne S}
+\hS$. In other words, $x$ is an element present in no other set in
+$\cF$ except $S$. This is intuitive since $v$ is present in no other
+path image under $\cl$ other than $\cl(S)$.  The element $x$ and leaf
+$v$ are then removed from the set $S$ and path $\cl(S)$
 respectively. After doing this for all leaves in $T$, all path images
@@
   \begin{algorithmic}[\lndisplay]
-    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$
-    \COMMENT {Path images are such that no two path images share a
-      leaf.}
+    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S
+    \in \cF_0$ \COMMENT {Path images are such that no two path images
+      share a leaf.}
     \STATE $j \assign 1$\\
@@
     \STATE $X \assign S_1 \setminus \bigcup_{S \in \cF_{j-1}, S \ne S_1}S$\\
-    \IF{$X$ is empty} \label{xempty} \STATE {\bf exit} \label{ln:exit2} \ENDIF
+    \IF{$X$ is empty} \label{xempty} \STATE {\bf
+      exit} \label{ln:exit2} \ENDIF
     \STATE $x \assign $ arbitrary element from $X$\\
@@
 %     \tnote{IS THIS CORRECT?}  If input ICPPL $(\cF, \cl)$ is not
-%   feasible, then in one of the recursive calls to Algorithm --3--, the
-%   {\em \bf exit} statement in line x in Algorithm --1 or line y in
-%   Algorithm --2 will get executed.
+%     feasible, then in one of the recursive calls to Algorithm --3--,
+%     the {\em \bf exit} statement in line x in Algorithm --1 or line
+%     y in Algorithm --2 will get executed.
 %   \end{lemma}
@@
 filtered ICPPL. Now we present the algorithmic characterization of a
-feasible tree path labeling by way of Algorithm~\ref{al:icppl-find-isomorph}.
-
-Algorithm~\ref{al:icppl-find-isomorph} computes a
-hypergraph isomorphism $\phi$ recursively using Algorithm~\ref{perms}
-and Algorithm~\ref{leafasgn} and pruning the leaves of the input
-tree. In brief, it is done as follows. Algorithm~\ref{leafasgn} gives
-us the leaf labels in $\cF_2$, i.e., the elements in $supp(\cF)$ that
-map to leaves in $T$, where $(\cF_2, \cl_2)$ is the output of
+feasible tree path labeling by way of
+Algorithm~\ref{al:icppl-find-isomorph}.
+
+Algorithm~\ref{al:icppl-find-isomorph} computes a hypergraph
+isomorphism $\phi$ recursively using Algorithm~\ref{perms} and
+Algorithm~\ref{leafasgn} and pruning the leaves of the input tree. In
+brief, it is done as follows. Algorithm~\ref{leafasgn} gives us the
+leaf labels in $\cF_2$, i.e., the elements in $supp(\cF)$ that map to
+leaves in $T$, where $(\cF_2, \cl_2)$ is the output of
 Algorithm~\ref{leafasgn}. All leaves in $T$ are then pruned away. The
@@
 
-    \FOR {every $v \in L$}
-    \STATE $\phi'(x) \assign v$ where $x \in \cl_2^{-1}(\{v\})$
-    \COMMENT {Copy the leaf labels to a one to one function $\phi':
-      supp(\cF) \rightarrow L$
+    \FOR {every $v \in L$} \STATE $\phi'(x) \assign v$ where $x \in
+    \cl_2^{-1}(\{v\})$ \COMMENT {Copy the leaf labels to a one to one
+      function $\phi': supp(\cF) \rightarrow L$
     }\\
@@
 
-    \STATE $\phi'' \assign$ {\tt
-      get-hypergraph-isomorphism($\cF', \cl', T'$)}
+    \STATE $\phi'' \assign$ {\tt get-hypergraph-isomorphism($\cF',
+      \cl', T'$)}
     \STATE $\phi \assign \phi'' \cup \phi'$ \\
@@
   If $(\cF, \cl)$ is an ICPPL from a tree $T$ and
-  Algorithm~\ref{al:icppl-find-isomorph}, {\tt
-    get- hypergraph- isomorphism ($\cF, \cl, T$)} returns a non-empty
-  function, then there exists a hypergraph isomorphism $\phi :
-  supp(\cF) \rightarrow V(T)$ such that the $\phi$-induced tree path
-  labeling is equal to $\cl$ or $\cl_\phi = \cl$.
+  Algorithm~\ref{al:icppl-find-isomorph}, {\tt get- hypergraph-
+    isomorphism ($\cF, \cl, T$)} returns a non-empty function, then
+  there exists a hypergraph isomorphism $\phi : supp(\cF) \rightarrow
+  V(T)$ such that the $\phi$-induced tree path labeling is equal to
+  $\cl$ or $\cl_\phi = \cl$.
 \end{lemma}
@@
 respectively.
-% For each hyperedge $X \in \cO$, we will maintain a 2-tuple of non-negative
-% numbers $\seq{p_1(X), p_2(X)}$.  The numbers satisfy the property that
-% $p_1(X) + p_2(X) \leq |X|$, and at the end of path labeling, for each
-% $X$, $p_1(X) + p_2(X) = |X|$.  This signifies the algorithm tracking
-% the lengths of subpaths of the path assigned to $X$ from at most two
-% rays. We also maintain another parameter called the {\em residue} of
-% $X$ denoted by $s(X)=|X| - p_1(X)$. This signifies the residue path
-% length that must be assigned to $X$ which must be from another
-% ray. For instance, if $X$ is labeled a path from only one ray, then
-% $p_1(X) = |X|$, $p_2(X) = 0$ and $s(X) = 0$.
+% For each hyperedge $X \in \cO$, we will maintain a 2-tuple of
+% non-negative numbers $\seq{p_1(X), p_2(X)}$.  The numbers satisfy
+% the property that $p_1(X) + p_2(X) \leq |X|$, and at the end of path
+% labeling, for each $X$, $p_1(X) + p_2(X) = |X|$.  This signifies the
+% algorithm tracking the lengths of subpaths of the path assigned to
+% $X$ from at most two rays. We also maintain another parameter called
+% the {\em residue} of $X$ denoted by $s(X)=|X| - p_1(X)$. This
+% signifies the residue path length that must be assigned to $X$ which
+% must be from another ray. For instance, if $X$ is labeled a path
+% from only one ray, then $p_1(X) = |X|$, $p_2(X) = 0$ and $s(X) = 0$.
 
 % 
-% We iteratively consider each ray from which paths will be assigned to
-% hyperedges. At the beginning of each iteration hyperedges of $\cO$ are
-% classifed into the following disjoint sets.
+% We iteratively consider each ray from which paths will be assigned
+% to hyperedges. At the beginning of each iteration hyperedges of
+% $\cO$ are classifed into the following disjoint sets.
 % \begin{enumerate}
@@
 %   labeled with a path which does not contain $r$ in one of the
-%   previous iterations.\\  $\cL_1^i = \set{ X \mid p_1(X) = |X| \text{ and
-%     } p_2(X) = 0 \text{ and } s(X) = 0, X \in \cO}$
-% \item [$\cL_2^i$] {\em Labeled with $r$.} Those that have been labeled
-%   with two subpaths of $\cl(X)$ containing $r$ from two different rays
-%   in two previous iterations.\\ $\cL_2^i = \set{X \mid 0 < p_1\left(X\right),
-%     p_2\left(X\right) < |X| \text{ and } s(X) = 0, X \in \cO}$
+%   previous iterations.\\ $\cL_1^i = \set{ X \mid p_1(X) = |X| \text{
+%       and } p_2(X) = 0 \text{ and } s(X) = 0, X \in \cO}$
+% \item [$\cL_2^i$] {\em Labeled with $r$.} Those that have been
+%   labeled with two subpaths of $\cl(X)$ containing $r$ from two
+%   different rays in two previous iterations.\\ $\cL_2^i = \set{X
+%     \mid 0 < p_1\left(X\right), p_2\left(X\right) < |X| \text{ and }
+%     s(X) = 0, X \in \cO}$
 % \item [$\cT_1^i$] {\em Type 1 / partially labeled.} Those that have
@@
 %     \text{ and } s(X) > 0, X \in \cO}$
-%   \item [$\cT_2^i$] {\em Type 2 / not labeled.} Those that have not been
+% \item [$\cT_2^i$] {\em Type 2 / not labeled.} Those that have not
+%   been
 %   labeled with a path in any previous iteration.\\
-%     $\cT_2^i = \set{ X \mid p_1(X) = p_2(X) = 0 \text{ and } s(X) = |X|,
-%       X \in \cO}$
+%   $\cT_2^i = \set{ X \mid p_1(X) = p_2(X) = 0 \text{ and } s(X) =
+%     |X|, X \in \cO}$
 % \end{enumerate}
@@
   \begin{enumerate}
-  \item {\bf {There are type 1
-        edges, $\mathbf{|\cT_1^i| > 0}$}:} 
-% , and for each such $X$ assign the
-%     path in $R_i$ of length $s(X)$ starting at $v_{i,k+2}=r$.  $X$ is
-%     then removed from $\cO_i$, and $p_2(X)=s(X)$, and $s(X)=0$.If any
-%     intersection cardinality is violated because of this assignment,
-%     EXIT reporting the non-existence of a feasible assignment.
+  \item {\bf {There are type 1 edges, $\mathbf{|\cT_1^i| > 0}$}:}
+    % , and for each such $X$ assign the path in $R_i$ of length
+    % $s(X)$ starting at $v_{i,k+2}=r$.  $X$ is then removed from
+    % $\cO_i$, and $p_2(X)=s(X)$, and $s(X)=0$.If any intersection
+    % cardinality is violated because of this assignment, EXIT
+    % reporting the non-existence of a feasible assignment.
     {If there is only one $\cT_1^i$ hyperedge, label it to the unique
@@
       |\cl'(S_1) \cap \cl'(S_2)|$ i.e. they are in the same rays and
-    their residues must contain their intersection. Hence assign both
-    to $R_i$.
+      their residues must contain their intersection. Hence assign
+      both to $R_i$.
     \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| =
-    |\cl'(S_1) \cap \cl'(S_2)|$. i.e. the intersection cardinality has
-    been met. Hence assign only $S_1$ to $R_i$ and $S_2$ must be
+      |\cl'(S_1) \cap \cl'(S_2)|$. i.e. the intersection cardinality
+      has been met. Hence assign only $S_1$ to $R_i$ and $S_2$ must be
       assigned to some $i \ge p$.
@@
   \item {\bf {There are no type-1 edges nor} super-marginal type-2
-    hyperedges:} If all type-2 edges of length at most $k+1$, EXIT by
-  saying there is no ICPPL.  Otherwise, Find an $X$ of type-2 such
-  that $|X| \geq k+2$, assign it the unique path starting at
+      hyperedges:} If all type-2 edges of length at most $k+1$, EXIT
+    by saying there is no ICPPL.  Otherwise, Find an $X$ of type-2
+    such that $|X| \geq k+2$, assign it the unique path starting at
     $v_{i(k+1)}=r$ of length $k+2$, and set $p_1(X)=k+2,
     s(X)=|X|-(k+2)$, assign it the unique path starting at
-  $v_{i(k+1)}=r$ of length $k+2$, mark this as a type-1 hyperedge, and
-  add it to $\cO_{i+1}$ after removing it from $\cO_i$.
+    $v_{i(k+1)}=r$ of length $k+2$, mark this as a type-1 hyperedge,
+    and add it to $\cO_{i+1}$ after removing it from $\cO_i$.
     % \remove{Next, iteratively, consider each type-1 edge $X \in
-  %   \cO_i$, and if it intersects a previously assigned $Y$ such that
+    %   \cO_i$, and if it intersects a previously assigned $Y$ such
+    %   that
     %   $\cl(Y) \subseteq R_i$, assign a path of length $s(X)$
-  %   containing $v_{i(k+2)}=r$, set $p_2(X)=s(X), s(X)=0$, and remove
+    %   containing $v_{i(k+2)}=r$, set $p_2(X)=s(X), s(X)=0$, and
+    %   remove
     %   $X$ from $\cO_i$.}\tnote{covered in step 2}
@@
     k+2$, then $X$ is assigned the path of length $|X|$ starting at
-  $v_{i1}$, and $p_1(X) = |X|, p_2(X)=0$. $X$ is removed from $\cO_i$
-  and not added to $\cO_{i+1}$.  If $|X| > k+2$, then $p_1(X)=k+2$,
-  assign it the unique path starting at $v_{i(k+1)}=r$ of length $k+2$
-  and $s(X) = |X| - (k+2)$.  In this case, $X$ is classified as a
-  Type-1 edge, removed from $\cO_i$, and added to $\cO_{i+1}$.
+    $v_{i1}$, and $p_1(X) = |X|, p_2(X)=0$. $X$ is removed from
+    $\cO_i$ and not added to $\cO_{i+1}$.  If $|X| > k+2$, then
+    $p_1(X)=k+2$, assign it the unique path starting at $v_{i(k+1)}=r$
+    of length $k+2$ and $s(X) = |X| - (k+2)$.  In this case, $X$ is
+    classified as a Type-1 edge, removed from $\cO_i$, and added to
+    $\cO_{i+1}$.
   \end{enumerate}
@@
 \noindent
-\tnote{need to formalize the below properties} In the following
-lemmas we identify a set of necessary conditions for $\cF$ to have an
-ICPPL in the $k$-subdivided star $T$.  If during the execution of the
-above algorithm, one of these necessary conditions is not satisfied,
-the algorithm exits reporting the non-existence of an ICPPL.
+\tnote{need to formalize the below properties} In the following lemmas
+we identify a set of necessary conditions for $\cF$ to have an ICPPL
+in the $k$-subdivided star $T$.  If during the execution of the above
+algorithm, one of these necessary conditions is not satisfied, the
+algorithm exits reporting the non-existence of an ICPPL.
 \begin{lemma}
@@
   \label{lem:sup-mar}
-  If $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
-  tree path labeling to tree $T$, then $\cl(X)$ will
+  If $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible tree
+  path labeling to tree $T$, then $\cl(X)$ will
   contain a leaf in $T$.\\
-  On the otherhand, if $\cl(X)$ has a leaf in
-  $T$, $X$ is marginal but may not be super-marginal.
+  On the otherhand, if $\cl(X)$ has a leaf in $T$, $X$ is marginal but
+  may not be super-marginal.
 \end{lemma}
@@
 
-{\tt abstract begin}
-  it is
-  known that if the given tree is a path a feasible assignment can be
-  found in polynomial time, and we observe that it can actually be
-  done in logspace. \tnote{[TRUE?]}
-{\tt abstract end}
-
-In Section \ref{prelims} we present the
-necessary preliminaries, in Section \ref{feasible} we present our
-characterization of feasible tree path assignments, and in Section
-\ref{decompos} we present the characterizing subproblems for finding a
-bijection between $U$ and $V(T)$ such that sets map to tree
-paths. Finally, in Section \ref{complexity} we conclude by showing
-that Tree Path Assignment is GI-Complete, and also observe that
-Consecutive Ones Testing is in Logspace.
+{\tt abstract begin} it is known that if the given tree is a path a
+feasible assignment can be found in polynomial time, and we observe
+that it can actually be done in logspace. \tnote{[TRUE?]}  {\tt
+  abstract end}
+
+In Section \ref{prelims} we present the necessary preliminaries, in
+Section \ref{feasible} we present our characterization of feasible
+tree path assignments, and in Section \ref{decompos} we present the
+characterizing subproblems for finding a bijection between $U$ and
+$V(T)$ such that sets map to tree paths. Finally, in Section
+\ref{complexity} we conclude by showing that Tree Path Assignment is
+GI-Complete, and also observe that Consecutive Ones Testing is in
+Logspace.
 
@@
 
-\annote{ We refer to this as the Tree Path Assignment
-problem for an input $(\cF,T)$ pair.}{Add this ``pair'' as a
-  definition in ch:prelims}
+\annote{ We refer to this as the Tree Path Assignment problem for an
+  input $(\cF,T)$ pair.}{Add this ``pair'' as a definition in
+  ch:prelims}
 
@@
   intersection graph of the corresponding set system is an interval
-graph.  A similar connection to a subclass of chordal graphs, and this
-subclass contains interval graphs, exists for the generalization of
-COP.  In this case, the intersection graph of the corresponding set
-system must be a path graph. Chordal graphs are of great
-significance, extensively studied, and have several applications.  One
-of the well known and interesting properties of a chordal graphs is
-its connection with intersection graphs cite{mcg04}. For every chordal
-graph, there exists a tree and a family of subtrees of this tree such
-that the intersection graph of this family is isomorphic to the
-chordal graph cite{plr70,gav78,bp93}.  Certain format of these trees
-are called clique trees cite{apy92} of the graph which is a compact
-representation of the chordal graph. There has also been work done on
-the generalization of clique trees to clique hypergraphs cite{km02}.
-If the chordal graph can be represented as the intersection graph of
-paths in a tree, then the graph is called path graph cite{mcg04} or
-path chordal graph.  Therefore, it is clear that if there is a
-bijection from $U$ to $V(T)$ such that the sets map to paths, then the
-intersection graph of the set system is indeed a path chordal graph.
-This is, however, only a necessary condition and can be checked
-efficiently, as path graph recognition is polynomial
-solvable cite{gav78,aas93}.  Indeed, it is possible to construct a set
-system and tree, such that the intersection graph is a path chordal
-graph, but there is no bijection between $U$ and $V(T)$ such that the
-sets map to paths.  This connection indeed suggests that our problem
-is indeed as hard as path graph isomorphism.  Further path graph
-isomorphism is known be isomorphism-complete, see for example
-cite{kklv10}. }
-In the second part of this paper, we decompose our
-search for a bijection between $U$ and $V(T)$ into subproblems.  Each
-subproblem is on a set system in which for each set, there is another
-set in the set system with which the intersection is {\em strict}-
-there is a non-empty intersection, but neither is contained in the
-other.  This is in the spirit of results in \cite{wlh02,nsnrs09} where
-to test for COP in a given matrix, the COP problem is solved on an
-equivalent set of prime matrices.  \annote{Our decomposition localizes the
-challenge of path graph isomorphism to two problems.}{HOW?}
+  graph.  A similar connection to a subclass of chordal graphs, and
+  this subclass contains interval graphs, exists for the
+  generalization of COP.  In this case, the intersection graph of the
+  corresponding set system must be a path graph. Chordal graphs are of
+  great significance, extensively studied, and have several
+  applications.  One of the well known and interesting properties of a
+  chordal graphs is its connection with intersection graphs
+  cite{mcg04}. For every chordal graph, there exists a tree and a
+  family of subtrees of this tree such that the intersection graph of
+  this family is isomorphic to the chordal graph
+  cite{plr70,gav78,bp93}.  Certain format of these trees are called
+  clique trees cite{apy92} of the graph which is a compact
+  representation of the chordal graph. There has also been work done
+  on the generalization of clique trees to clique hypergraphs
+  cite{km02}.  If the chordal graph can be represented as the
+  intersection graph of paths in a tree, then the graph is called path
+  graph cite{mcg04} or path chordal graph.  Therefore, it is clear
+  that if there is a bijection from $U$ to $V(T)$ such that the sets
+  map to paths, then the intersection graph of the set system is
+  indeed a path chordal graph.  This is, however, only a necessary
+  condition and can be checked efficiently, as path graph recognition
+  is polynomial solvable cite{gav78,aas93}.  Indeed, it is possible to
+  construct a set system and tree, such that the intersection graph is
+  a path chordal graph, but there is no bijection between $U$ and
+  $V(T)$ such that the sets map to paths.  This connection indeed
+  suggests that our problem is indeed as hard as path graph
+  isomorphism.  Further path graph isomorphism is known be
+  isomorphism-complete, see for example cite{kklv10}. } In the second
+part of this paper, we decompose our search for a bijection between
+$U$ and $V(T)$ into subproblems.  Each subproblem is on a set system
+in which for each set, there is another set in the set system with
+which the intersection is {\em strict}- there is a non-empty
+intersection, but neither is contained in the other.  This is in the
+spirit of results in \cite{wlh02,nsnrs09} where to test for COP in a
+given matrix, the COP problem is solved on an equivalent set of prime
+matrices.  \annote{Our decomposition localizes the challenge of path
+  graph isomorphism to two problems.}{HOW?}
 
@@
 
-\section{Complexity of Tree Path Assignment-A
-  Discussion} 
+\section{Complexity of Tree Path Assignment-A Discussion}
 \label{complexity}
-Recall that the input to the Tree
-Path Assignment question is an order pair $(\cF,T)$ where $\cF$ is a
-family of subsets of an universe $U$, and $T$ is a tree such that
-$|V(T)| = |U|$.  The question is to come up with a bijection from $U$
-to $V(T)$ such that the image of each set in $\cF$ is a path in $T$.
+Recall that the input to the Tree Path Assignment question is an order
+pair $(\cF,T)$ where $\cF$ is a family of subsets of an universe $U$,
+and $T$ is a tree such that $|V(T)| = |U|$.  The question is to come
+up with a bijection from $U$ to $V(T)$ such that the image of each set
+in $\cF$ is a path in $T$.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC  Assignment\tnote{Merge with the original charac section}} 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC Assignment\tnote{Merge
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%   with the original charac
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%   section}}
 % \label{feasible}
-% Consider a path assignment $\cA =
-% \{(S_i, P_i) \mid S_i \in \cF, P_i \text{ is a path from $T$}, i \in
-% [m]\}$ to a set system $\cF = \{S_i \mid S_i \subseteq U, i \in
-% [m]\}$, were $T$ is a given tree, $U$ is the set system's universe and
-% $m$ is the number of sets in $\cF$. We call $\cA$ an {\em Intersection
-%   Cardinality Preserving Path Assignment (ICPPA)} if it has the
-% following properties.
+% Consider a path assignment $\cA = \{(S_i, P_i) \mid S_i \in \cF, P_i
+% \text{ is a path from $T$}, i \in [m]\}$ to a set system $\cF =
+% \{S_i \mid S_i \subseteq U, i \in [m]\}$, were $T$ is a given tree,
+% $U$ is the set system's universe and $m$ is the number of sets in
+% $\cF$. We call $\cA$ an {\em Intersection Cardinality Preserving
+%   Path Assignment (ICPPA)} if it has the following properties.
 
@@
 %   % \item [iii.] $|\bigcup_{i \in J} S_i| = |\bigcup_{i \in J} P_i|$
-%   %   for all $J \subseteq I$\footnote{TBD: this is needed now to fix
+%   %   for all $J \subseteq I$\footnote{TBD: this is needed now to
+%   %     fix
 %   %     some proofs}
@@
 %   |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to conditions (ii)
-%   and (iii) of ICPPA, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
-%   \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
-%   = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. \qed
+%   and (iii) of ICPPA, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| =
+%   |P_1 \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2)
+%   \setminus P_3| = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is
+%   proven. \qed
 % \end{proof}
@@
 % \begin{lemma}
-%   \label{lem:fourpaths} Consider four paths in a tree $Q_1, Q_2, Q_3,
-%   Q_4$ such that they have nonempty pairwise intersection and $Q_1,
-%   Q_2$ share a leaf. Then there exists distinct $i, j, k \in
+%   \label{lem:fourpaths} Consider four paths in a tree $Q_1, Q_2,
+%   Q_3, Q_4$ such that they have nonempty pairwise intersection and
+%   $Q_1, Q_2$ share a leaf. Then there exists distinct $i, j, k \in
 %   \{1,2,3,4\}$ such that, $Q_1 \cap Q_2 \cap Q_3 \cap Q_4 = Q_i \cap
@@
 %   %   Since $Q_1, Q_2$ share a leaf, the following are paths $Q_1
-%   % \setminus Q_2$, $Q_2 \setminus Q_1$, $Q_1 \cap Q_2$ and they are
+%   %   \setminus Q_2$, $Q_2 \setminus Q_1$, $Q_1 \cap Q_2$ and they
+%   %   are
 %   %   mutually disjoint.
@@
 %   Q_2$. Similarly, if $Q \cap (Q_2 \setminus Q_1) = \O$, $Q \cap Q_1
-%   \cap Q_2 = Q \cap Q_1$. Thus it is clear that if the intersection of
-%   any two paths does not intersect with any of the set differences of
-%   the remaining two paths, the claim in the lemma is true.
+%   \cap Q_2 = Q \cap Q_1$. Thus it is clear that if the intersection
+%   of any two paths does not intersect with any of the set
+%   differences of the remaining two paths, the claim in the lemma is
+%   true.
 %   %   Note that $Q_1 \setminus Q_2$ and $Q_2
@@
 %   case. i.e. the intersection of any two paths intersects with both
-%   the set differences of the other two. First let us consider $Q \cap
-%   (Q_1 \setminus Q_2) \ne \O$ and $Q \cap (Q_1 \setminus Q_2) \ne \O$,
-%   where $Q = Q_3 \cap Q_4$. Since $Q_1$ and $Q_2$ share a leaf, there
-%   is exactly one vertex at which they branch off from the path $Q_1
-%   \cap Q_2$ into two paths $Q_1 \setminus Q_2$ and $Q_2 \setminus
-%   Q_1$. Let this vertex be $v$. It is clear that if path $Q_3 \cap
-%   Q_4$, must intersect with paths $Q_1 \setminus Q_2$ and $Q_2
-%   \setminus Q_1$, it must contain $v$ since these are paths from a
-%   tree. Moreover, $Q_3 \cap Q_4$ intersects with $Q_1 \cap Q_2$ at
-%   exactly $v$ and only at $v$ which means that $Q_1 \cap Q_2$ does not
-%   intersect with $Q_3 \setminus Q_4$ or $Q_4 \setminus Q_3$ which
-%   contradicts initial condition of this case. Thus this
+%   the set differences of the other two. First let us consider $Q
+%   \cap (Q_1 \setminus Q_2) \ne \O$ and $Q \cap (Q_1 \setminus Q_2)
+%   \ne \O$, where $Q = Q_3 \cap Q_4$. Since $Q_1$ and $Q_2$ share a
+%   leaf, there is exactly one vertex at which they branch off from
+%   the path $Q_1 \cap Q_2$ into two paths $Q_1 \setminus Q_2$ and
+%   $Q_2 \setminus Q_1$. Let this vertex be $v$. It is clear that if
+%   path $Q_3 \cap Q_4$, must intersect with paths $Q_1 \setminus Q_2$
+%   and $Q_2 \setminus Q_1$, it must contain $v$ since these are paths
+%   from a tree. Moreover, $Q_3 \cap Q_4$ intersects with $Q_1 \cap
+%   Q_2$ at exactly $v$ and only at $v$ which means that $Q_1 \cap
+%   Q_2$ does not intersect with $Q_3 \setminus Q_4$ or $Q_4 \setminus
+%   Q_3$ which contradicts initial condition of this case. Thus this
 %   case cannot occur and case 1 is the only possible scenario. \\
@@
 % claim is clear: that is a path assignment is feasible, then all
-% intersection cardinalities are preserved, that is the path assignment
-% is an ICPPA.  The reason is that a feasible path assignment has an
-% associated bijection between $U$ and $V(T)$ such that the sets map to
-% paths.  The rest of the section is devoted to constructively proving
-% that it is sufficient, for the existence of a bijection, for a path
-% assignment to be an ICPPA.  At a top-level, the constructive
-% approaches refine the path assignment iteratively, such that at the
-% end of each iteration we have a path assignment, and finally we have a
-% family of bijections.  First we present and then prove the correctness
-% of Algorithm \ref{perms}.  This algorithm refines the path assignment
-% by considering pairs of paths that share a leaf.
+% intersection cardinalities are preserved, that is the path
+% assignment is an ICPPA.  The reason is that a feasible path
+% assignment has an associated bijection between $U$ and $V(T)$ such
+% that the sets map to paths.  The rest of the section is devoted to
+% constructively proving that it is sufficient, for the existence of a
+% bijection, for a path assignment to be an ICPPA.  At a top-level,
+% the constructive approaches refine the path assignment iteratively,
+% such that at the end of each iteration we have a path assignment,
+% and finally we have a family of bijections.  First we present and
+% then prove the correctness of Algorithm \ref{perms}.  This algorithm
+% refines the path assignment by considering pairs of paths that share
+% a leaf.
 
@@
 %   \label{lem:invar1}
-%   In Algorithm \ref{perms}, at the end of $j$th iteration, $j \ge 0$,
-%   of the while loop of Algorithm \ref{perms}, the following invariants
-%   are maintained.
+%   In Algorithm \ref{perms}, at the end of $j$th iteration, $j \ge
+%   0$, of the while loop of Algorithm \ref{perms}, the following
+%   invariants are maintained.
 %   \begin{itemize}
@@
 %     $|P' \cap P''|=|Q' \cap Q''|$.
-%   \item {\em Invariant IV:} For any three, $(P',Q'), (P'',Q''), (P, Q)
-%     \in \Pi_j$, $|P' \cap P'' \cap P|=|Q' \cap Q'' \cap Q|$.
+%   \item {\em Invariant IV:} For any three, $(P',Q'), (P'',Q''), (P,
+%     Q) \in \Pi_j$, $|P' \cap P'' \cap P|=|Q' \cap Q'' \cap Q|$.
 %   \end{itemize}
@@
 % \begin{proof}
-%   Proof is by induction on the number of iterations, $j$. In the rest
-%   of the proof, the term ``new sets'' will refer to the new sets added
-%   in $j$th iteration as defined in line \ref{setbreak} of Algorithm
-%   \ref{perms}, i.e. the following three assignment pairs for some
-%   $(P_1,Q_1), (P_2,Q_2) \in \Pi_{j-1}$ where $Q_1$ and $Q_2$ intersect
-%   and share a leaf: $(P_1 \cap P_2, Q_1 \cap Q_2)$, or $(P_1 \setminus
-%   P_2, Q_1 \setminus Q_2)$, or $(P_2 \setminus P_1, Q_2
+%   Proof is by induction on the number of iterations, $j$. In the
+%   rest of the proof, the term ``new sets'' will refer to the new
+%   sets added in $j$th iteration as defined in line \ref{setbreak} of
+%   Algorithm \ref{perms}, i.e. the following three assignment pairs
+%   for some $(P_1,Q_1), (P_2,Q_2) \in \Pi_{j-1}$ where $Q_1$ and
+%   $Q_2$ intersect and share a leaf: $(P_1 \cap P_2, Q_1 \cap Q_2)$,
+%   or $(P_1 \setminus P_2, Q_1 \setminus Q_2)$, or $(P_2 \setminus
+%   P_1, Q_2
 %   \setminus Q_1)$.\\
 %   \noindent
-%   The base case, $\Pi_0 = \{(S_i,P_i) \mid i \in [m]\}$, is trivially
-%   true since it is the input which is an ICPPA.  Assume the lemma is
-%   true till the $j-1$ iteration. Consider $j$th
+%   The base case, $\Pi_0 = \{(S_i,P_i) \mid i \in [m]\}$, is
+%   trivially true since it is the input which is an ICPPA.  Assume
+%   the lemma is true till the $j-1$ iteration. Consider $j$th
 %   iteration:\\
@@
 %   one of the new sets added in $\Pi_j$. Since $(P_1,Q_1)$ and
-%   $(P_2,Q_2)$ are from $\Pi_{j-1}$ and $Q_1,Q_2$ intersect and have a
-%   common leaf, it can be verified that the
+%   $(P_2,Q_2)$ are from $\Pi_{j-1}$ and $Q_1,Q_2$ intersect and have
+%   a common leaf, it can be verified that the
 %   new sets are also paths. \\
-%   By hypothesis for invariant III, invariant II also holds for $(P,Q)$
-%   no matter which new set in $\Pi_j$ it
+%   By hypothesis for invariant III, invariant II also holds for
+%   $(P,Q)$ no matter which new set in $\Pi_j$ it
 %   is.\\
@@
 %   $\Pi_{j-1}$, then they are both new sets and invariant III holds
-%   trivially (new sets are disjoint). Next consider $(P,Q), (P',Q') \in
-%   \Pi_j$ with only one of them, say $(P',Q')$, in $\Pi_{j-1}$. Then
-%   $(P,Q)$ is one of the new sets added in line \ref{setbreak}. It is
-%   easy to see that if $(P,Q)$ is $(P_1 \cap P_2, Q_1 \cap Q_2)$, then
-%   due to invariant IV in hypothesis, invariant III becomes true in
-%   this iteration. Similarly, using lemma \ref{lem:setminuscard}
-%   invariant III is proven if $(P, Q)$ is $(P_1 \setminus P_2, Q_1
-%   \setminus Q_2)$, or $(P_2 \setminus P_1, Q_2
+%   trivially (new sets are disjoint). Next consider $(P,Q), (P',Q')
+%   \in \Pi_j$ with only one of them, say $(P',Q')$, in
+%   $\Pi_{j-1}$. Then $(P,Q)$ is one of the new sets added in line
+%   \ref{setbreak}. It is easy to see that if $(P,Q)$ is $(P_1 \cap
+%   P_2, Q_1 \cap Q_2)$, then due to invariant IV in hypothesis,
+%   invariant III becomes true in this iteration. Similarly, using
+%   lemma \ref{lem:setminuscard} invariant III is proven if $(P, Q)$
+%   is $(P_1 \setminus P_2, Q_1 \setminus Q_2)$, or $(P_2 \setminus
+%   P_1, Q_2
 %   \setminus Q_1)$.\\
 %   To prove invariant IV, consider three assignments
-%   $(P,Q),(P',Q'),(P'',Q'')$. If at least two of these pairs are in not
-%   $\Pi_{j-1}$, then they are any two of the new sets. Note that these
-%   new sets are disjoint and hence if $(P',Q'), (P'',Q'')$ are any of
-%   these sets, $|P \cap P' \cap P''|=|Q \cap Q' \cap Q''|=0$ and
-%   invariant IV is true. Now we consider the case if at most one of
-%   $(P,Q),(P',Q'),(P'',Q'')$ is not in $\Pi_{j-1}$. If none of them are
-%   not in $\Pi_{j-1}$ (i.e. all of them are in $\Pi_{j-1}$), invariant
-%   IV is clearly true. Consider the case where exactly one of them is
-%   not in $\Pi_{j-1}$. w.l.o.g let that be $(P,Q)$ and it could be any
-%   one of the new sets. If $(P,Q)$ is $(P_1 \cap P_2, Q_1 \cap Q_2)$,
-%   from lemma \ref{lem:fourpaths} and invariant III hypothesis,
-%   invariant IV is proven. Similarly if $(P,Q)$ is any of the other new
-%   sets, invariant IV is proven by also using lemma
+%   $(P,Q),(P',Q'),(P'',Q'')$. If at least two of these pairs are in
+%   not $\Pi_{j-1}$, then they are any two of the new sets. Note that
+%   these new sets are disjoint and hence if $(P',Q'), (P'',Q'')$ are
+%   any of these sets, $|P \cap P' \cap P''|=|Q \cap Q' \cap Q''|=0$
+%   and invariant IV is true. Now we consider the case if at most one
+%   of $(P,Q),(P',Q'),(P'',Q'')$ is not in $\Pi_{j-1}$. If none of
+%   them are not in $\Pi_{j-1}$ (i.e. all of them are in $\Pi_{j-1}$),
+%   invariant IV is clearly true. Consider the case where exactly one
+%   of them is not in $\Pi_{j-1}$. w.l.o.g let that be $(P,Q)$ and it
+%   could be any one of the new sets. If $(P,Q)$ is $(P_1 \cap P_2,
+%   Q_1 \cap Q_2)$, from lemma \ref{lem:fourpaths} and invariant III
+%   hypothesis, invariant IV is proven. Similarly if $(P,Q)$ is any of
+%   the other new sets, invariant IV is proven by also using lemma
 %   \ref{lem:setminuscard}. \qed
@@
 % It can be observed that the output of algorithm \ref{perms} is such
-% that every leaf is incident on at most a single path in the new set of
-% assignments. This is due to the loop condition at line
-% \ref{shareleaf}. Let $v_1$ be the leaf incident on path $P_i$. Assign
-% to it any one element from $S_i \setminus \bigcup_{i \ne j}
-% S_j$. Remove $(S_i, P_i)$ from assignments and add $(\{x_1\},
-% \{v_1\}), (S_i \setminus \{x_1\}, P_i \setminus \{v_1\})$. Now all
-% assignments except single leaf assignments are paths from the subtree
-% $T_0 = T \setminus \{v \mid v \text{ is a leaf in } T\}$.
+% that every leaf is incident on at most a single path in the new set
+% of assignments. This is due to the loop condition at line
+% \ref{shareleaf}. Let $v_1$ be the leaf incident on path
+% $P_i$. Assign to it any one element from $S_i \setminus \bigcup_{i
+%   \ne j} S_j$. Remove $(S_i, P_i)$ from assignments and add
+% $(\{x_1\}, \{v_1\}), (S_i \setminus \{x_1\}, P_i \setminus
+% \{v_1\})$. Now all assignments except single leaf assignments are
+% paths from the subtree $T_0 = T \setminus \{v \mid v \text{ is a
+%   leaf in } T\}$.
 
@@
 %     \STATE $j = 1$\\
-%     \WHILE {there is a leaf $v$ and a unique $(S_{i_1}, P_{i_1})$ such
-%       that $v \in P_{i_1}$}
+%     \WHILE {there is a leaf $v$ and a unique $(S_{i_1}, P_{i_1})$
+%       such that $v \in P_{i_1}$}
 %     \STATE $\Pi_j=   \Pi_{j-1} \setminus \{(S_{i_1}, P_{i_1})\}$\\
 %     \STATE $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in I}S_i$
-%     \IF{$X$ is empty} \STATE exit \ENDIF; \STATE Let $x = $ arbitrary
-%     element from $X$ \vspace{2mm}
+%     \IF{$X$ is empty} \STATE exit \ENDIF; \STATE Let $x = $
+%     arbitrary element from $X$ \vspace{2mm}
 
@@
 %   In Algorithm \ref{leafasgn},for all $j \geq 0$, at the end of the
-%   $j$th iteration the four invariants given in lemma \ref{lem:invar1}
-%   are valid.
-%   % Moreover, $X$ as defined in the algorithm is non-empty if this is
+%   $j$th iteration the four invariants given in lemma
+%   \ref{lem:invar1} are valid.
+%   %   Moreover, $X$ as defined in the algorithm is non-empty if this
+%   %   is
 %   %   an ICPPA.
@@
 %     I}S_i$ is non empty in every iteration for an ICPPA. Suppose $X$
-%   is empty. We know that $v \in P_{i_1} \setminus \bigcup_{i \ne i_1,
-%     i \in I}P_i$ since $v$ is in the unique path $P_{i_1}$. Since this
-%   is an ICPPA $|S_{i_1}| = |P_{i_1}|$. For any $x \in S_{i_1}$ it is
-%   contained in at least two sets due to our assumption. Let $S_{i_2}$
-%   be a second set that contains $x$. We know $v \notin
-%   P_{i_2}$. Therefore there cannot exist a permutation that maps
-%   elements of $S_{i_2}$ to $P_{i_2}$. This contradicts our assumption
-%   that this is an ICPPA. Thus $X$ cannot be empty.
+%   is empty. We know that $v \in P_{i_1} \setminus \bigcup_{i \ne
+%     i_1, i \in I}P_i$ since $v$ is in the unique path
+%   $P_{i_1}$. Since this is an ICPPA $|S_{i_1}| = |P_{i_1}|$. For any
+%   $x \in S_{i_1}$ it is contained in at least two sets due to our
+%   assumption. Let $S_{i_2}$ be a second set that contains $x$. We
+%   know $v \notin P_{i_2}$. Therefore there cannot exist a
+%   permutation that maps elements of $S_{i_2}$ to $P_{i_2}$. This
+%   contradicts our assumption that this is an ICPPA. Thus $X$ cannot
+%   be empty.
 
 %   We use mathematical induction on the number of iterations for this
-%   proof. The term ``new sets'' will refer to the sets added in $\Pi_j$
-%   in the $j$th iteration, i.e. $(P' \setminus \{x\},Q' \setminus
-%   \{v\})$ and $(\{x\},\{v\})$ for some $(P',Q')$ in $\Pi_{j-1}$ such
-%   that $v$ is a leaf and $Q'$ is the unique path
+%   proof. The term ``new sets'' will refer to the sets added in
+%   $\Pi_j$ in the $j$th iteration, i.e. $(P' \setminus \{x\},Q'
+%   \setminus \{v\})$ and $(\{x\},\{v\})$ for some $(P',Q')$ in
+%   $\Pi_{j-1}$ such that $v$ is a leaf and $Q'$ is the unique path
 %   incident on it.\\
-%   For $\Pi_0$ all invariants hold because it is output from algorithm
-%   \ref{perms} which is an ICPPA. Hence base case is proved.  Assume
-%   the lemma holds for $\Pi_{j-1}$. Consider $\Pi_j$ and any $(P,Q) \in
-%   \Pi_j$. If $(P,Q) $ is in $ \Pi_j$ and $\Pi_{j-1}$ invariants I and
-%   II are true because of induction assumption. If it is only in
-%   $\Pi_j$, then it is $\{(P' \setminus \{x\}),(Q' \setminus \{v\})$ or
-%   $(\{x\},\{v\})$ for some $(P',Q')$ in $\Pi_{j-1}$. By definition,
-%   $x$ is an element in $P'$ (as defined in the algorithm) and $v$ is a
-%   leaf in $Q'$. If $(P,Q)$ is $\{(P' \setminus \{x\}),(Q' \setminus
-%   \{v\})$, $Q$ is a path since only a leaf is removed from path
-%   $Q'$. We know $|P'| = |Q'|$, therefore $|P' \setminus \{x\}| = |Q'
-%   \setminus \{v\}|$. Hence in this case invariants I and II are
-%   obvious. It is easy to see these invariants hold if $(P,Q)$ is
-%   $(\{x\},\{v\})$.
-
-%   For invariant III consider $(P_1,Q_1),(P_2,Q_2)$ in $\Pi_j$. If both
-%   of them are also in $\Pi_{j-1}$, claim is proved. If one of them is
-%   not in $\Pi_{j-1}$ then it has to be $\{(P' \setminus \{x\}),(Q'
-%   \setminus \{v\})$ or $(\{x\},\{v\})$ for some $(P',Q')$ in
-%   $\Pi_{j-1}$. Since by definition, $Q'$ is the only path with $v$ and
-%   $P'$ the only set with $x$ in the previous iteration, $|P_1 \cap (P'
-%   \setminus \{x\})| = |P_1 \cap P'|$ and $|Q_1 \cap (Q' \setminus
-%   \{v\})| = |Q_1 \cap Q'|$ and $|P_1 \cap \{x\}| = 0, Q_1 \cap \{v\} =
-%   0$. Thus invariant III is also proven.
+%   For $\Pi_0$ all invariants hold because it is output from
+%   algorithm \ref{perms} which is an ICPPA. Hence base case is
+%   proved.  Assume the lemma holds for $\Pi_{j-1}$. Consider $\Pi_j$
+%   and any $(P,Q) \in \Pi_j$. If $(P,Q) $ is in $ \Pi_j$ and
+%   $\Pi_{j-1}$ invariants I and II are true because of induction
+%   assumption. If it is only in $\Pi_j$, then it is $\{(P' \setminus
+%   \{x\}),(Q' \setminus \{v\})$ or $(\{x\},\{v\})$ for some $(P',Q')$
+%   in $\Pi_{j-1}$. By definition, $x$ is an element in $P'$ (as
+%   defined in the algorithm) and $v$ is a leaf in $Q'$. If $(P,Q)$ is
+%   $\{(P' \setminus \{x\}),(Q' \setminus \{v\})$, $Q$ is a path since
+%   only a leaf is removed from path $Q'$. We know $|P'| = |Q'|$,
+%   therefore $|P' \setminus \{x\}| = |Q' \setminus \{v\}|$. Hence in
+%   this case invariants I and II are obvious. It is easy to see these
+%   invariants hold if $(P,Q)$ is $(\{x\},\{v\})$.
+
+%   For invariant III consider $(P_1,Q_1),(P_2,Q_2)$ in $\Pi_j$. If
+%   both of them are also in $\Pi_{j-1}$, claim is proved. If one of
+%   them is not in $\Pi_{j-1}$ then it has to be $\{(P' \setminus
+%   \{x\}),(Q' \setminus \{v\})$ or $(\{x\},\{v\})$ for some $(P',Q')$
+%   in $\Pi_{j-1}$. Since by definition, $Q'$ is the only path with
+%   $v$ and $P'$ the only set with $x$ in the previous iteration,
+%   $|P_1 \cap (P' \setminus \{x\})| = |P_1 \cap P'|$ and $|Q_1 \cap
+%   (Q' \setminus \{v\})| = |Q_1 \cap Q'|$ and $|P_1 \cap \{x\}| = 0,
+%   Q_1 \cap \{v\} = 0$. Thus invariant III is also proven.
 
 %   \noindent
-%   To prove invariant IV, consider $(P_1,Q_1),(P_2,Q_2), (P_3,Q_3)$ in
-%   $\Pi_j$. If exactly one of them, say $P_3 \notin \Pi_{j-1}$, it is
-%   one of the new sets. By the same argument used to prove invariant
-%   III, $|P_1 \cap P_2 \cap (P' \setminus \{x\})| = |P_1 \cap P_2 \cap
-%   P'|$ and $|Q_1 \cap Q_2 \cap (Q' \setminus \{x\})| = |Q_1 \cap Q_2
-%   \cap Q'|$. Since $P_1, P_2, P'$ are all in $\Pi_{j-1}$, by induction
-%   hypothesis $|P_1 \cap P_2 \cap P'| = |Q_1 \cap Q_2 \cap Q'|$. Also
-%   $|P_1 \cap P_2 \cap \{x\}| = 0, Q_1 \cap Q_2 \cap \{v\} = 0$.  If
-%   two or more of them are not in $\Pi_{j-1}$, then it can be verified
-%   that $|P_1 \cap P_2 \cap P_3| = |Q_1 \cap Q_2 \cap Q_3|$ since the
-%   new sets in $\Pi_j$ are either disjoint or as follows: assuming
-%   $P_1, P_2 \notin \Pi_{j-1}$ and new sets are derived from $(P', Q'),
-%   (P'', Q'') \in \Pi_{j-1}$ with $x_1, x_2$ exclusively in $P_1, P_2$,
+%   To prove invariant IV, consider $(P_1,Q_1),(P_2,Q_2), (P_3,Q_3)$
+%   in $\Pi_j$. If exactly one of them, say $P_3 \notin \Pi_{j-1}$, it
+%   is one of the new sets. By the same argument used to prove
+%   invariant III, $|P_1 \cap P_2 \cap (P' \setminus \{x\})| = |P_1
+%   \cap P_2 \cap P'|$ and $|Q_1 \cap Q_2 \cap (Q' \setminus \{x\})| =
+%   |Q_1 \cap Q_2 \cap Q'|$. Since $P_1, P_2, P'$ are all in
+%   $\Pi_{j-1}$, by induction hypothesis $|P_1 \cap P_2 \cap P'| =
+%   |Q_1 \cap Q_2 \cap Q'|$. Also $|P_1 \cap P_2 \cap \{x\}| = 0, Q_1
+%   \cap Q_2 \cap \{v\} = 0$.  If two or more of them are not in
+%   $\Pi_{j-1}$, then it can be verified that $|P_1 \cap P_2 \cap P_3|
+%   = |Q_1 \cap Q_2 \cap Q_3|$ since the new sets in $\Pi_j$ are
+%   either disjoint or as follows: assuming $P_1, P_2 \notin
+%   \Pi_{j-1}$ and new sets are derived from $(P', Q'), (P'', Q'') \in
+%   \Pi_{j-1}$ with $x_1, x_2$ exclusively in $P_1, P_2$,
 %   $(\{x_1\},\{v_1\}), (\{x_2\},\{v_2\}) \in \Pi_j $ thus $v_1, v_2$
@@
 
-% Using algorithms \ref{perms} and \ref{leafasgn} we prove the following
-% theorem.
+% Using algorithms \ref{perms} and \ref{leafasgn} we prove the
+% following theorem.
 
@@
 %   \ref{leafasgn} is a set of interval assignments to sets and
-%   one-to-one assignment of elements of $U$ to each leaf of $T$. To be
-%   precise, it would be of the form $\cB_0 = \cA_0 \cup \cL_0$. The
-%   leaf assignments are defined in $\cL_0 = \{ (x_i,v_i) \mid x_i \in
-%   U, v_i \in T, x_i \ne x_j, v_i \ne v_j, i \ne j, i,j \in [k] \}$
-%   where $k$ is the number of leaves in $T$. The path assignments are
-%   defined in $\cA_0 \subseteq \{(S_i',P_i') \mid S_i' \subseteq U_0,
-%   P_i' \text{ is a path from } T_0\}$ where $T_0$ is the tree obtained
-%   by removing all the leaves in $T$ and $U_0 = U \setminus \{ x \mid x
-%   \text{ is assigned to a leaf in }\cL_0 \}$. Now we have a subproblem
-%   of finding the permutation for the path assignment $\cA_0$ which has
-%   paths from tree $T_0$ and sets from universe $U_0$. Now we repeat
-%   the procedure and the path assignment $\cA_0$ and tree $T_0$ is
-%   given as input to Algorithm \ref{perms}. The output of this
-%   algorithm is given to Algorithm \ref{leafasgn} to get a new union of
-%   path and leaf assignments $\cB_1 = \cA_1 \cup \cL_1$ defined similar
-%   to $\cB_0, \cL_0, \cA_0$. In general, the two algorithms are run on
-%   path assignment $\cA_{i-1}$ with paths from tree $T_{i-1}$ to get a
-%   new subproblem with path assignment $\cA_i$ and tree $T_{i}$. $T_i$
-%   is the subtree of $T_{i-1}$ obtained by removing all its
-%   leaves. More importantly, it gives leaf assignments $\cL_{i}$ to the
-%   leaves in tree $T_{i-1}$. This is continued until we get a
-%   subproblem with path assignment $\cA_{d-1}$ and tree $T_{d-1}$ for
-%   some $d \le n$ which is just a path. From the last lemma we know
-%   that $\cA_{d-1}$ is an ICPPA. Another observation is that an ICPPA
-%   with all its tree paths being intervals (subpaths from a path) is
-%   nothing but an ICPIA\cite{nsnrs09}.  Let $\cA_{d-1}$ be equal to
+%   one-to-one assignment of elements of $U$ to each leaf of $T$. To
+%   be precise, it would be of the form $\cB_0 = \cA_0 \cup
+%   \cL_0$. The leaf assignments are defined in $\cL_0 = \{ (x_i,v_i)
+%   \mid x_i \in U, v_i \in T, x_i \ne x_j, v_i \ne v_j, i \ne j, i,j
+%   \in [k] \}$ where $k$ is the number of leaves in $T$. The path
+%   assignments are defined in $\cA_0 \subseteq \{(S_i',P_i') \mid
+%   S_i' \subseteq U_0, P_i' \text{ is a path from } T_0\}$ where
+%   $T_0$ is the tree obtained by removing all the leaves in $T$ and
+%   $U_0 = U \setminus \{ x \mid x \text{ is assigned to a leaf in
+%   }\cL_0 \}$. Now we have a subproblem of finding the permutation
+%   for the path assignment $\cA_0$ which has paths from tree $T_0$
+%   and sets from universe $U_0$. Now we repeat the procedure and the
+%   path assignment $\cA_0$ and tree $T_0$ is given as input to
+%   Algorithm \ref{perms}. The output of this algorithm is given to
+%   Algorithm \ref{leafasgn} to get a new union of path and leaf
+%   assignments $\cB_1 = \cA_1 \cup \cL_1$ defined similar to $\cB_0,
+%   \cL_0, \cA_0$. In general, the two algorithms are run on path
+%   assignment $\cA_{i-1}$ with paths from tree $T_{i-1}$ to get a new
+%   subproblem with path assignment $\cA_i$ and tree $T_{i}$. $T_i$ is
+%   the subtree of $T_{i-1}$ obtained by removing all its leaves. More
+%   importantly, it gives leaf assignments $\cL_{i}$ to the leaves in
+%   tree $T_{i-1}$. This is continued until we get a subproblem with
+%   path assignment $\cA_{d-1}$ and tree $T_{d-1}$ for some $d \le n$
+%   which is just a path. From the last lemma we know that $\cA_{d-1}$
+%   is an ICPPA. Another observation is that an ICPPA with all its
+%   tree paths being intervals (subpaths from a path) is nothing but
+%   an ICPIA\cite{nsnrs09}.  Let $\cA_{d-1}$ be equal to
 %   $\{(S_i'',P_i'') \mid S_i'' \subseteq U_{d-1}, P_i'' \text{ is a
@@
 %   not be precisely an interval in the sense of consecutive integers
-%   because they are some nodes from a tree. However, it is easy to see
-%   that the nodes of $T_{d-1}$ can be ordered from left to right and
-%   ranked to get intervals $I_i$ for every path $P_i''$ as
-%   follows. $I_i = \{[l,r] \mid l = \text{ the lowest rank of the nodes
-%     in }P_i'', r = l+|P_i''|-1 \}$. Let asssignment $\cA_d$ be with
-%   the renamed paths. $\cA_d = \{ (S_i'', I_i) \mid (S_i'', P_i'') \in
-%   \cA_{d-1} \}$. What has been effectively done is renaming the nodes
-%   in $T_{d-1}$ to get a tree $T_d$.  The ICPIA $\cA_d$ is now in the
-%   format that the ICPIA algorithm requires which gives us the
-%   permutation $\sigma' : U_{d-1} \rightarrow T_{d-1}$
+%   because they are some nodes from a tree. However, it is easy to
+%   see that the nodes of $T_{d-1}$ can be ordered from left to right
+%   and ranked to get intervals $I_i$ for every path $P_i''$ as
+%   follows. $I_i = \{[l,r] \mid l = \text{ the lowest rank of the
+%     nodes in }P_i'', r = l+|P_i''|-1 \}$. Let asssignment $\cA_d$ be
+%   with the renamed paths. $\cA_d = \{ (S_i'', I_i) \mid (S_i'',
+%   P_i'') \in \cA_{d-1} \}$. What has been effectively done is
+%   renaming the nodes in $T_{d-1}$ to get a tree $T_d$.  The ICPIA
+%   $\cA_d$ is now in the format that the ICPIA algorithm requires
+%   which gives us the permutation $\sigma' : U_{d-1} \rightarrow
+%   T_{d-1}$
 
@@
 %   $\sigma'$ along with all the leaf assignments $\cL_i$ gives us the
-% permutation for the original path assignment $\cA$.  More precisely,
-% the permutation for tree path assignment $\cA$ is defined as
-% follows. $\sigma: U \rightarrow T$ such that the following is
-% maintained.
+%   permutation for the original path assignment $\cA$.  More
+%   precisely, the permutation for tree path assignment $\cA$ is
+%   defined as follows. $\sigma: U \rightarrow T$ such that the
+%   following is maintained.
 %   \begin{align*}
 %     \sigma(x) &= \sigma'(x),   \text{ if } x \in U_{d-1} \\
-%   &= \cL_i(x), \text{ where $x$ is assigned to a leaf in a subproblem
-%     $\cA_{i-1}, T_{i-1}$}
+%     &= \cL_i(x), \text{ where $x$ is assigned to a leaf in a
+%       subproblem $\cA_{i-1}, T_{i-1}$}
 %   \end{align*}
@@
 %   $U$, remove all leaves from $T$ to get new tree $T_0$. The leaf
-% assignments are in $\cL_0$. Since only leaves were removed $T_0$ is
-% indeed a tree. Repeat the algorithms on $T_0$ to get leaf assignments
-% $\cL_{1}$. Remove the leaves in $T_0$ to get $T_1$ and so on until the
-% pruned tree $T_d$ is a single path. Now run ICPIA algorithm on $T_d$
-% to get permutation $\sigma'$. The relation $\cL_0 \cup \cL_1 \cup
-% .. \cup \cL_{d} \cup \sigma'$ gives the bijection required in the
-% original problem.\qed
+%   assignments are in $\cL_0$. Since only leaves were removed $T_0$
+%   is indeed a tree. Repeat the algorithms on $T_0$ to get leaf
+%   assignments $\cL_{1}$. Remove the leaves in $T_0$ to get $T_1$ and
+%   so on until the pruned tree $T_d$ is a single path. Now run ICPIA
+%   algorithm on $T_d$ to get permutation $\sigma'$. The relation
+%   $\cL_0 \cup \cL_1 \cup .. \cup \cL_{d} \cup \sigma'$ gives the
+%   bijection required in the original problem.\qed
 % \end{proof}
-
-\chapter{Conclusion}
-\label{ch:conclusion}
-
-\begin{center}
-{\tt MOVED FROM INTRO SUBSECTION IN ch:myresearch CHAPTER begin}
-\end{center}
-
-% \annote{
-  It is an interesting fact that for a matrix with the COP, the
-  intersection graph of the corresponding set system is an interval
-  graph. % }{the following text could be in Further Study/Conclusion}
-A
-similar connection to a subclass of chordal graphs and a superclass of
-interval graphs exists for the generalization of COP.  In this case,
-the intersection graph of the corresponding set system must be a {\em
-  path graph}. Chordal graphs are of great significance, are
-extensively studied, and have several applications.  One of the well
-known and interesting properties of a chordal graphs is its connection
-with intersection graphs \cite{mcg04}. For every chordal graph, there
-exists a tree and a family of subtrees of this tree such that the
-intersection graph of this family is isomorphic to the chordal graph
-\cite{plr70,gav78,bp93}.  These trees when in a certain format, are
-called {\em clique trees} \cite{apy92} of the chordal graph. This is a
-compact representation of the chordal graph. There has also been work
-done on the generalization of clique trees to clique hypergraphs
-\cite{km02}.  If the chordal graph can be represented as the
-intersection graph of paths in a tree, then the graph is called path
-graph \cite{mcg04}.  Therefore, it is clear that if there is a
-bijection from $U$ to $V(T)$ such that for every set, the elements in
-it map to vertices of a unique path in $T$, then the intersection
-graph of the set system is indeed a path graph.  However, this is only
-a necessary condition and can be checked efficiently because path
-graph recognition is polynomial time
-solvable\cite{gav78,aas93}. Indeed, it is possible to construct a set
-system and tree, such that the intersection graph is a path graph, but
-there is no bijection between $U$ and $V(T)$ such that the sets map to
-paths. Path graph isomorphism is known be isomorphism-complete, see
-for example \cite{kklv10}. An interesting area of research would be to
-see what this result tells us about the complexity of the tree path
-labeling problem (not covered in this paper).
-
-\begin{center}
-{\tt MOVED FROM INTRO SUBSECTION IN ch:myresearch CHAPTER\\ end}
-\end{center}
-
-We give a characterization for feasible tree path labeling of path
-hypergraphs. The proof for this is constructive and computes a
-feasibility bijection mapping vertices of the hypergraph to vertices
-of the given target tree. This thesis also discovered an exponential algorithm
-that computes a feasible TPL to a given hypergraph if it is a path hypergraph.
-
-Our results have close connections to recognition of path graphs and
-to path graph isomorphism.  Graphs which can be represented as the
-intersection graph of paths in a tree are called {\em path
-  graphs}\cite{mcg04}. Thus, a hypergraph $\cF$ can be interpreted as
-paths in a tree, if and only if the intersection graph of $\cF$ is a
-{\em path graph}. Path graphs are a subclass of chordal graphs since
-chordal graphs are characterized as the intersection graphs of
-subtrees of a tree\cite{mcg04}.  Path graphs are well studied in the
-literature \cite{plr70}--\cite{mcg04}\tnote[TH]{Check if cite:bp93 is
-  indeed for path graphs}.  Path graph recognition can be done in
-polynomial time\cite{gav78,aas93}.  Clearly, this is a necessary
-condition in terms of the intersection graph of the input hypergraph
-$\cF$ in \FTPL. However, one can easily obtain a counterexample to
-show the insufficiency of this condition\tnote[TH]{Give a figure
-  showing insuff.}.  Path graph isomorphism is known be
-isomorphism-complete\cite{kklv10}. Therefore, \annote[TH]{it is
-  unlikely that we can solve the problem of finding feasible path
-  labeling $\cl$ for a given $\cF$ and }{WHY?} tree $T$.  It is
-definitely interesting to classify the kinds of trees and hypergraphs
-for which feasible path labelings can be\tnote[TH]{ELABORATE.}  found
-efficiently.  \annote[TH]{These results would form a natural
-  generalization of COP testing and interval graph isomorphism,
-  culminating in Graph Isomorphism }{REWORD} itself.  To this effect
-we consider TPL on {\kstar} and give a polynomial time solution for
-the same.
-
-While we address the computation of TPL
-for a given hypergraph% \footnote{The term {\em hypergraph} is used
-%   instead {\em set systems} are used interchangably in this
-%   section. Hypergraphs that have TPLs on some target tree are called
-%   {\em path hypergraphs}.}
-from a given target tree in this research, however optimization
-problems in TPL, akin to problems in Section~\ref{sec:optcop} for COP,
-is outside the scope of this thesis. Whether TPL on general trees is
-solvable in $P$ remains open. So do optimization opportunities in TPL
-(possible extensions to optimization for COP in the
-Section~\ref{sec:optcop}.\tnote[TH]{link to section}).
-
-%\newpage %DUMMY
-\tnote[TH]{ ||Complexity challenges:|| Canonization is an important
-  tool in graph isomorphism. Invention of a
-% deterministic method of canonization for any class of graphs naturally
-% results in an algorithm for isomorphism. All that is required is to
-% check if the canons of two graphs are the same. Thus complexity of
-% graph isomorphism can be studied by studying canonization
-% methods. While general graph ismorphism remains elusive in terms of
-% hardness, canonization has been studied for smaller classes of graphs
-% thus giving complexity/hardness results for them.
-
-% In 1992, \cite{sl92} made an important discovery that tree isomorphism
-% is in logspace. It was done by inventing a method of canonization of
-% trees using a logspace depth first traversal
-% algorithm. \tnote{confirm the statement of claim}.
-
-% \cite{kklv10} proved that interval graph canonization is also in
-% logspace thus drawing logspace conslusions about COT. 
-% % This result is
-% % built on the fact that 
-% Interval graphs are FP+C (fixed point with
-% counting) definable \tnote{elucidate} and \cite{l10} showed that
-% this implies that it captures PTIME. \tnote{elucidate} This result
-% along with that of undirected graph reachability being in logspace
-% \cite{rei08}, \cite{kklv10} proved their logspace result.
-}   %
-\tnote[TH]{ (1) the interesting survey in conclusion section of kklv10
-  (2) [[[COP helping some problems with hardness]]] having COP in the
-  input structure makes some problems less hard than in a general
-  input. -- min set cover problem is known to be NP hard. so is min
-  hitting
-  set. these two problems are equivalent cite-adp80\\
-  -- -- complexity results: log factor polynomial time approximable cite-f98, W[2]-complete with solution size as parameter cite-df99\\
-  -- a couple of generalizations:\\
-  -- -- min degree hypergraph (MDH) problem\\
-  -- -- red blue covering (RBSC) problem\\
-  ----- cite-d08phd Chapter 5.\\
-  ----- what are cite-d08phd results\\
-  ----- Rectangle stabbing \\
-  ----- see cite-d08phd Chapter 6?  }
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Appendices.
-
-% \appendix %%% NON NEIN NAHI ILLA NONE
-
-% \chapter{A}
-% \tnote{DO WE NEED THIS?}
-
-\newpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Bibliography.
-\begin{singlespace}
-%\begin{thebibliography}{10}
-\addcontentsline{toc}{chapter}{REFERENCES}
-
-{\footnotesize
-  \bibliographystyle{alpha} 
-  \bibliography{../lib/cop-variants__thesis}
-}
-%\end{thebibliography}
-\end{singlespace}
-%\newpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% List of papers
-
-% \listofpapers %%% NON NEIN NAHI ILLA NONE
-
-% \begin{enumerate}  
-% \item Authors....  \newblock
-%  Title...
-%   \newblock {\em Journal}, Volume,
-%   Page, (year).
-% \end{enumerate}
-
-\end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% DOCUMENTATION ON THE THESIS CLASS FILE ETC.
-
-% Before reading any further please note that you are strongly advised
-% against changing any of the formatting options used in the class
-% provided in this directory, unless you are absolutely sure that it
-% does not violate the IITM formatting guidelines.  \emph{Please do not
-%   change the margins or the spacing.}  If you do change the formatting
-% you are on your own (don't blame me if you need to reprint your entire
-% thesis).  In the case that you do change the formatting despite these
-% warnings, the least I ask is that you do not redistribute your style
-% files to your friends (or enemies).
-
-% It is also a good idea to take a quick look at the formatting
-% guidelines.  Your office or advisor should have a copy.  If they
-% don't, pester them, they really should have the formatting guidelines
-% readily available somewhere.
-
-% To generate PDF's with the links from the \verb+hyperref+ package use
-% the following command:
-% \begin{verbatim}
-% % dvipdfm -o thesis.pdf thesis.dvi
-% \end{verbatim}
-
-% \section{Package Options}
-
-% Use this thesis as a basic template to format your thesis.  The
-% \verb+iitmdiss+ class can be used by simply using something like this:
-% \begin{verbatim}
-% \documentclass[PhD]{iitmdiss}  
-% \end{verbatim}
-
-% To change the title page for different degrees just change the option
-% from \verb+PhD+ to one of \verb+MS+, \verb+MTech+ or \verb+BTech+.
-% The dual degree pages are not supported yet but should be quite easy
-% to add.  The title page formatting really depends on how large or
-% small your thesis title is.  Consequently it might require some hand
-% tuning.  Edit your version of \verb+iitmdiss.cls+ suitably to do this.
-% I recommend that this be done once your title is final.
-
-% To write a synopsis simply use the \verb+synopsis.tex+ file as a
-% simple template.  The synopsis option turns this on and can be used as
-% shown below.
-% \begin{verbatim}
-% \documentclass[PhD,synopsis]{iitmdiss}                                
-% \end{verbatim}
-
-% Once again the title page may require some small amount of fine
-% tuning.  This is again easily done by editing the class file.
-
-% This sample file uses the \verb+hyperref+ package that makes all
-% labels and references clickable in both the generated DVI and PDF
-% files.  These are very useful when reading the document online and do
-% not affect the output when the files are printed.
-
-% \section{Example Figures and tables}
-
-% Fig.~\ref{fig:iitm} shows a simple figure for illustration along with
-% a long caption.  The formatting of the caption text is automatically
-% single spaced and indented.  Table~\ref{tab:sample} shows a sample
-% table with the caption placed correctly.  The caption for this should
-% always be placed before the table as shown in the example.
-
-% \begin{figure}[htpb]
-%   \begin{center}
-%     \resizebox{50mm}{!} {\includegraphics *{iitm.eps}}
-%     \resizebox{50mm}{!} {\includegraphics *{iitm.eps}}
-%     \caption {Two IITM logos in a row.  This is also an
-%       illustration of a very long figure caption that wraps around two
-%       two lines.  Notice that the caption is single-spaced.}
-%   \label{fig:iitm}
-%   \end{center}
-% \end{figure}
-
-% \begin{table}[htbp]
-%   \caption{A sample table with a table caption placed
-%     appropriately. This caption is also very long and is
-%     single-spaced.  Also notice how the text is aligned.}
-%   \begin{center}
-%   \begin{tabular}[c]{|c|r|} \hline
-%     $x$ & $x^2$ \\ \hline
-%     1  &  1   \\
-%     2  &  4  \\
-%     3  &  9  \\
-%     4  &  16  \\
-%     5  &  25  \\
-%     6  &  36  \\
-%     7  &  49  \\
-%     8  &  64  \\ \hline
-%   \end{tabular}
-%   \label{tab:sample}
-%   \end{center}
-% \end{table}
-
-% \section{Bibliography with BIB\TeX}
-
-% I strongly recommend that you use BIB\TeX\ to automatically generate
-% your bibliography.  It makes managing your references much easier.  It
-% is an excellent way to organize your references and reuse them.  You
-% can use one set of entries for your references and cite them in your
-% thesis, papers and reports.  If you haven't used it anytime before
-% please invest some time learning how to use it.
-
-% I've included a simple example BIB\TeX\ file along in this directory
-% called \verb+refs.bib+.  The \verb+iitmdiss.cls+ class package which
-% is used in this thesis and for the synopsis uses the \verb+natbib+
-% package to format the references along with a customized bibliography
-% style provided as the \verb+iitm.bst+ file in the directory containing
-% \verb+thesis.tex+.  Documentation for the \verb+natbib+ package should
-% be available in your distribution of \LaTeX.  Basically, to cite the
-% author along with the author name and year use \verb+\cite{key}+ where
-% \verb+key+ is the citation key for your bibliography entry.  You can
-% also use \verb+\citet{key}+ to get the same effect.  To make the
-% citation without the author name in the main text but inside the
-% parenthesis use \verb+\citep{key}+.  The following paragraph shows how
-% citations can be used in text effectively.
-
-% More information on BIB\TeX\ is available in the book by
-% \cite{lamport:86}.  There are many
-% references~\citep{lamport:86,prabhu:xx} that explain how to use
-% BIB\TeX.  Read the \verb+natbib+ package documentation for more
-% details on how to cite things differently.
-
-% Here are other references for example.  \citet{viz:mayavi} presents a
-% Python based visualization system called MayaVi in a conference paper.
-% \citet{pan:pr:flat-fst} illustrates a journal article with multiple
-% authors.  Python~\citep{py:python} is a programming language and is
-% cited here to show how to cite something that is best identified with
-% a URL.
-
-% \section{Other useful \LaTeX\ packages}
-
-% The following packages might be useful when writing your thesis.
-
-% \begin{itemize}  
-% \item It is very useful to include line numbers in your document.
-%   That way, it is very easy for people to suggest corrections to your
-%   text.  I recommend the use of the \texttt{lineno} package for this
-%   purpose.  This is not a standard package but can be obtained on the
-%   internet.  The directory containing this file should contain a
-%   lineno directory that includes the package along with documentation
-%   for it.
-
-% \item The \texttt{listings} package should be available with your
-%   distribution of \LaTeX.  This package is very useful when one needs
-%   to list source code or pseudo-code.
-
-% \item For special figure captions the \texttt{ccaption} package may be
-%   useful.  This is specially useful if one has a figure that spans
-%   more than two pages and you need to use the same figure number.
-
-% \item The notation page can be entered manually or automatically
-%   generated using the \texttt{nomencl} package.
-
-% \end{itemize}
-
-% More details on how to use these specific packages are available along
-% with the documentation of the respective packages.
-
-% END OF DOCUMENTATION
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-01-03 src/040chap4.tex
Cleaned up thesis.tex. Split thesis.tex into chapterwise subfiles for easy edit tex management. Content changes are mainly in chapter 1 and 3 although nothing significant. To being working on chapter 2 after this commit.

--- 2011-12-30 src/thesis.tex
@@
-\documentclass[MS]             % MS thesis dissertation.
-              {iitmdiss_as}    % Hacked by hyperbolicme.
-%              {iitmdiss}      % If the powers to be are finicky.
-
-\usepackage{times}             % Font?
-\usepackage{t1enc}             % T1 encoding??  
-%\usepackage{boxedminipage}
-\usepackage{../lib/mythesislib}
-\usepackage{../lib/mylatexlib}
-% \usepackage{comment}           % DEBUG ONLY
-
-\def \mythesissubmissiondate {10 January 2012}
-\def \mythesissubmissionmonth {\MakeUppercase {January 2012}}
-
-\begin{document}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Title page
-\title{\mythesistitleNOTSYNPOSIS} 
-\author{\myname} 
-\date{
-  {\mythesissubmissionmonth}} 
-\department{\MakeUppercase{\mythesisdept}}
-%\department{\mythesisdept}
-
-%\nocite{*}
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Certificate
-\certificate
-
-\vspace*{0.5in}
-
-\noindent 
-This is to certify that the thesis titled {\bf \mythesistitleNOTSYNPOSIS},
-submitted by {\bf \myname}, to the {\bf Indian Institute of Technology
-Madras}, for the award of the degree of {\bf \mydegree}, is 
-a bona fide record of the research work done by her under our
-supervision.  The contents of this thesis, in full or in parts, have
-not been submitted to any other Institute or University for the award
-of any degree or diploma.
-
-\vspace*{1.5in}
-
-\begin{singlespacing}
-  \hspace*{-0.25in}
-  \parbox{3in}{
-    \noindent {\bf Dr.~\myadvisor} \\
-    \small
-    \noindent Research Guide\\
-    \noindent Associate Professor\\
-    \noindent Dept. of \mydept\\
-    \noindent IIT Madras -- 600 036
-  } \hfill
-  \parbox{1.25in}{               % {1.75in}
-    \small
-    \vspace*{0.75in}
-    \begin{tabular}[h]{r}
-    \noindent {Chennai}\\         
-    \noindent {\mythesissubmissiondate} 
-    \end{tabular}
-  }
-\end{singlespacing}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Acknowledgements
-\acknowledgements
-
-\tnote{WIP}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Abstract
-
-\begin{abstract}
-  \noindent {\em {\bf Keywords}: \hspace*{0.5em} \parbox[t]{4.4in}{consecutive
-    ones property, algorithmic graph theory, hypergraph isomorphism,
-    interval labeling}} \vspace*{24pt}
-
-  \def \tem {}
-  \noindent 
-  Consecutive-ones property is a non-trivial property of binary
-  matrices that has been studied widely in the literature for over
-  past 50 years. Detection of COP in a matrix is possible efficiently
-  and there are several algorithms that achieve the same. This thesis
-  documents the work done on an extension of COP extended from the
-  equivalent interval assignment problem in \cite{nsnrs09}. These new
-  results rigorously prove a natural extension (to trees) of their
-  characterization as well as makes connections to graph isomorphism,
-  namely path graph isomorphism.
-
-  \tnote{EXPAND. 
-  Abstract must be a brief about what results we have and how it fits
-  in the body of research.\\
-  -- Area: Broad to Specialized. i.e. Combinatorial
-    algorithms -> Matrix reorganization -> general data reorganization
-    (interval assignment) -> path assignment\\
-  -- Class of problems: say, data reorganization.\\
-  -- Nature of results: Is a generalization. We have a
-    Polynomial algorithm for a subset of the generalization.\\
-% maybe add - we explore a natural
-%     generalization of results on binary matrices with the |\tem
-%     consecutive-ones property|.  We consider the following constraint
-%     satisfaction problem. Given (i) a set system $\F \subseteq$
-%     $(2^{U} \setminus \emptyset)$ of a finite set $U$ of cardinality
-%     $n$, (ii) a tree $T$ of size $n$ and (iii) a bijection called
-%     |\tem tree path labeling|, $\cl$ mapping the sets in $\cF$ to
-%     paths in $T$, does there exist at least one bijection $\phi:U
-%     \rightarrow V(T)$ such that for each $S \in \cF$, $\{\phi(x) \mid
-%     x \in S\} = \cl(S)$?  A tree path labeling of a set system is
-%     called |\tem feasible| if there exists such a bijection $\phi$.
-%     We present an algorithmic characterization of feasible tree path
-%     labeling. COP is a special instance of tree path labeling problem
-%     when $T$ is a path.  We conclude with a polynomial time algorithm
-%     to find a feasible tree path labeling of a given set system when
-%     $T$ is a |\tem $k$-subdivided star|, set system has a single
-%     containment tree of overlap components and set size is limited to
-%     at most $k+2$.
-  }
-\end{abstract}
-
-%\pagebreak
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Table of contents etc.
-
-\begin{singlespace}
-\tableofcontents
-\thispagestyle{empty}
-
-\listoftables
-\addcontentsline{toc}{chapter}{LIST OF TABLES}
-\listoffigures
-\addcontentsline{toc}{chapter}{LIST OF FIGURES}
-\end{singlespace}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Abbreviations
-\abbreviations
-
-\noindent 
-\begin{tabbing}
-xxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
-\textbf{COP}   \> Consecutive-ones Property \\
-\textbf{COT}   \> Consecutive-ones Testing \\
-\textbf{ICPIA}   \> Intersection Cardinality Preservation Interval Assignment \\
-\textbf{ICPPL}   \> Intersection Cardinality Preserved Path Labeling \\
-% ADD MORE HERE
-\end{tabbing}
-
-\pagebreak
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Notation
-
-\chapter*{\centerline{NOTATION}}
-\addcontentsline{toc}{chapter}{NOTATION}
-
-\begin{singlespace}
-\begin{tabbing}
-xxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
-\textbf{$2^{U}$}  \> Powerset of set $U$ \\
-\end{tabbing}
-\end{singlespace}
-
-\pagebreak
-\clearpage
-
-% The main text will follow from this point so set the page numbering
-% to arabic from here on.
-\pagenumbering{arabic}
-
-\chapter{Introduction}
-\label{ch:intro}
-
-Consecutive-ones property is a non-trivial property of binary matrices
-that has been studied widely in the literature for over past 50
-years. Detection of COP in a matrix is possible efficiently and there
-are several algorithms that achieve the same. This thesis documents
-the work done on an extension of COP extended from the equivalent
-interval assignment problem in \cite{nsnrs09}. These new results
-rigorously prove a natural extension (to trees) of their
-characterization as well as makes connections to graph isomorphism,
-namely path graph isomorphism.
-
-\tnote{Have a para on Organization: Outline of document}
-Section~\ref{sec:background} gives a brief survey of COP and
-optimization problems related to it followed by motivation for the
-thesis in Section~\ref{sec:motive}.  Section~\ref{sec:results}
-presents a summary of our results on the extension of COP namely, the
-tree path labeling problem. \tnote[TH]{update for thesis}
-
-% \tnote{testing 1 2 3...}
-% \tnote[bogus1]{testing 1 2 3...}
-% \tnote[bogus2]{testing 1 2 3...}
-% \tnote[GTC]{testing 1 2 3...}
-% \tnote[bogus4]{testing 1 2 3...}
-% \tnote[TH]{testing 1 2 3...}
-
-\section{Illustration of the problem}
-\label{sec:problem}
-
-A group of students, \Pa, \Pig, \Sn, \Wo, \Vi, \Li, \Ch, \Sa, \Fr,
-\Sc\ and \Lu\ enroll at the {\WSI} for a liberal arts programme.  As
-part of their semester thesis, they pick a body of work to study and
-form the namesake study groups, {\LLL}, {\GGG}, {\BBB} and
-{\TTT}\tnote[TH]{put bib entries for these works!}. A student will be
-in at least one study group and may be in more than one. For instance,
-as will be seen later, {\Fr} studies both {\LLL} and {\TTT} while \Wo\
-studies only \BBB.
-
-Let $U$ and $\cF$ represent the set of students and the set of study
-groups respectively and the integers $n$ and $m$ denote the total
-number students and study groups respectively. In relation to this
-example, these are defined in Table~\ref{tab:wsigroups}. Also given
-there is the study group allocation to students.
-
-\begin{table}[t]%[htbp]
-  \centering
-  { \footnotesize
-  \begin{tabular}{rcl}
-    $U $&$=$&$ \{\xPa,\; \xPi,\; \xSn,\; \xWo,\; \xVi,\; \xLi,\; \xCh,\;
-    \xSa,\; \xFr,\; \xSc,\; \xLu\}$\\
-    $\cF $&$=$&$ \{\xLLL, \xGGG, \xBBB, \xTTT\}$\\
-    $\xLLL $&$=$&$ \{\xCh,\;  \xSa,\;  \xFr,\;  \xSc,\;  \xLu \}$\\
-    $\xGGG $&$=$&$ \{\xPa,\;  \xPi,\;  \xVi,\;  \xCh \}$\\
-    $\xBBB $&$=$&$ \{\xSn,\;  \xPi,\;  \xWo \}$\\
-    $\xTTT $&$=$&$ \{\xVi,\;  \xLi,\;  \xCh,\;  \xFr \}$\\
-    $n $&$=$&$ |U| = 11$\\
-    $m $&$=$&$ |\cF| = 4$      
-  \end{tabular}
-}
-  \caption{\figtabsize Students and study groups in \WSI}
-  \label{tab:wsigroups}
-\end{table}
-
-The campus has a residential area {\residenceblock} that has $n$
-single occupancy apartments reserved for the study groups'
-accommodation.  All these apartments are located such that the streets
-connecting them do {\em not} form
-loops. Fig~\ref{fig:streetmappathpeople} \tnote[TH]{fig:streetmap}
-shows the street map for {\residenceblock}. It may be noted that as a
-graph, it classifies as a tree.
-
-\begin{table}[htbp]
-  \centering
-  {\footnotesize    
-    \begin{tabular}{l||l}
-      \begin{tabular}{rcl}
-        $T $&$=$&$ \text{\em Street map tree of \residenceblock}$\\
-        $V(T) $&$=$&$ \{ 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11 \}$\\
-        $\cP $&$=$&$ \{R\xLLL, R\xGGG, R\xBBB, R\xTTT\}$\\
-        $R\xLLL $&$=$&$ \{9, 1, 5, 3, 11\}$\\
-        $R\xGGG $&$=$&$ \{7, 2, 6, 5\}$\\
-        $R\xBBB $&$=$&$ \{8, 2, 4 \}$\\
-        $R\xTTT $&$=$&$ \{10, 6, 5, 3\}$\\
-        $n $&$=$&$ |V| = 11$\\
-        $m $&$=$&$ |\cP| = 4$\\\\
-        $\cl $&$=$& {\em Study group to route mapping}\\        
-        $\cl(\mathbb{X}) $&$=$&  $R\mathbb{X}$ for all $\mathbb{X} \in
-        \cF$
-      \end{tabular} & 
-      \begin{tabular}{c}
-       {\em Apartment allocation }($\phi$)\\
-                    \begin{tabular}{c|c}                      
-                      1  & \xSa\\
-                      2  & \xPi\\
-                      3  & \xFr\\
-                      4  & \xWo\\
-                      5  & \xCh\\
-                      6  & \xVi\\
-                      7  & \xPa\\
-                      8  & \xSn\\
-                      9  & \xLu\\
-                      10 & \xLi\\
-                      11 & \xSc                     
-                    \end{tabular}
-      \end{tabular}\\
-    \end{tabular}
-  }
-  \caption{\figtabsize A solution to study group accommodation problem}
-  \label{tab:iltree}
-\end{table}
-
-A natural question would be to find how the students should be
-allocated apartments such that each study group has the {\em least
-  distance to travel} for a discussion? More specifically, we are
-interested in enforcing additional conditions, namely, that all the
-students in a study group must be next to each other; in other words,
-for one student to reach another fellow study group member's apartment
-(for all study groups the student is part of), she must not have to
-pass the apartment of any student who is not in that study group. To
-further elucidate, the apartments of students of any study group must
-be arranged in an exclusive unfragmented path on the street
-map. Exclusivity here means that the path must not have apartments
-from other study groups (unless that apartment is also part of {\em
-  this} study group).
-
-An intuitive approach to this problem would be to first find the paths
-that each study group decides to inhabit and then refine the
-allocation to individual students. A feasible allocation of exclusive
-routes to study groups is illustrated in
-Fig~\ref{fig:streetmappathpeople} \tnote[TH]{fig:streetmappath} and
-the students' allocation of apartments that obeys this route
-allocation is also shown. Table~\ref{tab:iltree} shows the same
-solution set theoretically.  How this is algorithmically computed is
-the focus of this thesis.
-
-\begin{figure}[htbp] %[t]%
- \centering
-  (a) \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
-  (b) \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
-
- \label{fig:streetmap}
- \caption{\figtabsize (a) {\residenceblock} street map (b)
-   {\residenceblock} street map with study group routes
-   allocated. Routes are color coded as follows: red for
-   \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
-   group, orange for \textcolor{yellow}{$\xBBB$} group, green for
-   \textcolor{green}{$\xTTT$} group}
-
- \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
-  \caption{\figtabsize Individual allocation of apartments to students
-    in {\residenceblock} that meets the requirements stated before.
-    The routes are color coded as follows: red for
-    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
-    group, orange for \textcolor{yellow}{$\xBBB$} group, green
-    for \textcolor{green}{$\xTTT$} group. {\tiny {\em Peanuts images
-        {\copyright} Charles Schulz}}}% YellowOrange
-  \label{fig:streetmappathpeople}
-\end{figure}
-\tnote{UPDATE IMAGE (b) 2-infinite-loop-BTWF.png. REMOVE TITLES. ADD
-  {$\xLLL$} {$\xGGG$} {$\xBBB$} {$\xTTT$}}
-
-\tnote[TH]{make dashed/textured lines for routes. make it color
-  agnostic.}
-
-As a special case, suppose all the apartments are on the same street
-or if they are all lined up on a single path, the street map becomes a
-tree that is just a path. Then the problem becomes what is called an
-{\em interval assignment problem}. The idea of interval assignment may
-not be obvious here; hence to see this, consider a different problem
-in {\WSI} where the classes for these study groups courses need to be
-scheduled during a day (or a week or any time period). Each study
-group has a bunch of courses associated with it some of which may be
-shared by two or more study groups. It is mandatory that a student who
-is a member of a study group takes all the courses associated with
-that group. There are slots during the day for classes to be held and
-the problem is to allocate class slots to courses such that all the
-classes of a study group are consecutive. It is debatable if this will
-not hamper the attention span and memory retention rate of the
-students but that is, regrettably, out of the scope of this
-thesis. The parallels between this class allocation problem and the
-accommodation problem can be seen as follows. The set $U$ here, are
-the courses offered (say Course 101 {\coneohone}, Course 102
-{\coneohtwo} and so on). In this variation of the problem, the
-collection $\cF$ is the set of study groups but the study groups are
-filled by course IDs (in place of students in the earlier
-example). For instance, Course 101 is mandatory for all study groups
-$\xLLL$, $\xGGG$, $\xBBB$, $\xTTT$ and Course 102 is mandatory for
-only the $\xLLL$ group) and so on. The sequence of class slots for the
-day (or week or any time period) is analogous to the street map in the
-accommodation problem. It is quite obvious now why this version of the
-problem (where the ``target graph'' is a path and not any
-tree\tnote[TH]{Allowing any tree in this example could be seen as a
-  scenario where there are parallel classes. A node falling in the
-  path between two other nodes would mean that the corresponding is
-  scheduled between the other two.}) is called an interval assignment
-problem.
-
-The interval assignment problem to a set system is equivalent to the
-consecutive-ones property (COP) problem in binary matrices\cite{wlh02,
-  nsnrs09}.  The COP problem is to rearrange rows (columns) of a
-binary matrix in such a way that every column (row) has its {\un}s
-occur consecutively. If this is possible the matrix is said to have
-the COP.  COP is a well researched combinatorial problem and has
-several positive results on tests for it and computing the COP
-permutation (i.e. the course schedule in the above illustration) which
-will be surveyed later in this document. Hence we are interested in
-extensions of COP, more specifically, the extension of interval
-assignment problem to tree path assignment problem (which is
-illustrated by the study group accommodation problem).
-
-\section{Basic preliminaries - general definitions and nomenclature}
-\tnote{(definitions theorems etc) needed if any
-    - graph theory}
-
-\section{Consecutive-ones Property Testing - a Brief Survey}
-\label{sec:background}
-
-In this section, a brief survey of the consecutive-ones problem and its
-optimization problems is presented.
-
-%%%%%%                              %%%%%%
-%THX is the second level of ``tnote''-ing%
-%%%%%%                              %%%%%%
-\tnote[TH]{ADD: As it will be described in detail later in this
-  document, isomorphism of certain % classes of graphs, namely chordal
-%   graphs, have a close relationship with consecutive-ones property and
-%   generalizations of it.  This is perhaps because of how closely COP
-%   of a matrix relates to properties of graphs derived from matrices as
-%   seen in the following results.
-}
-\tnote[THX]{ADD: peo exists iff chordal. lexicographic BFS
-  [tag:chordalGraph]} %
-\tnote[THX]{ADD: A well known result in %perfect graph theory is that
-%  the maximal cliques of  an interval graph $G$ can be linearly ordered
-%   such that for all $v \in V(G)$, cliques containing $v$ are
-%   consecutive in the ordering cite~{gh64}. This clearly means that a
-%   graph $G$ is an interval graph if and only if
-}% 
-\tnote[THX]{verify
-  from paper the statement of claim.}% 
-\tnote[TH]{maximal clique vertex
-  incidence matrix of%  $G$ has COP.  Also, maximal cliques of any
-%   chordal graph can be enumerated in polytime $O(m+n)$
-}
-\tnote[THX]{citation?!!}%  
-\tnote[TH]{cite~{fg65} uses these results to
-  give the first polynomial time algorithm for COT.}%
-\tnote[THX]{check. how do they use it?}%  
-\tnote[TH]{ A bipartite graph
-  is convex%  (on $R$) if and only if its half adjacency matrix has COP
-%   on rows.  The results in cite~{bl76} on COT are based on the result
-%   that interval graphs are AT-free chordal graphs.
-}%
-\tnote[THX]{the latter being Tucker's?}%  
-\tnote[THX]{(2) TBD survey
-  -- % see blue notes (in notebook) under Graph Isomorphism. namely
-%   citations in cite:aas93 (3) a brief on heirarchy of chordal graphs,
-%   path graphs, interval graphs, peo, clique tree etc. and results. (4)
-%   interval graphs are incomparability graphs. see
-%   Golumbic. [tag:classification] (5) finding min length hole in
-%   bipartite graph is polynomial time [Sec 3.3.2 in cite:d08phd] (6)
-%   theorem 2.2 in cite:d08phd - $G$ is union of v.d. caterpillars iff
-%   $M$ has COP in rows. $M$ is edge vertex incidence matrix $(*,2)$ of
-%   $G$
-}%
-
-\subsection{Matrices with COP}
-\label{sec:copmatrices}
-As seen earlier, the interval assignment problem (illustrated as the
-course scheduling problem in Section~\ref{sec:problem}), is a special
-case of the problem we address in this thesis, namely the tree path
-labeling problem (illustrated as the study group accommodation
-problem). The interval assignment problem and COP problem are equivalent
-problems. In this section we will see some of the results that
-exists in the literature today towards solving the COP problem and
-optimization problems surrounding it.
-
-\begin{figure}[t] %[htbp]
-  \centering
-
-  {\figtabsize
-    \begin{tabular}[h]{l|lcccl}
-      $M_1$: & $M_1'$: &&&& $M_2$:\\
-      &&&&&\\
-      \begin{tabular}[h]{llll}
-        $c_1$ & $c_2$ &$c_3$ &$c_4$\\
-        &&&\\
-        \un & 0   & \un & 0\\
-        0   & \un & 0   & \un \\
-        \un & 0   & 0   & \un
-      \end{tabular}
-      &
-      \begin{tabular}[h]{llll}
-        $c_3$ &$c_1$ &$c_4$& $c_2$\\
-        &&&\\
-        \un & \un & 0 & 0\\
-        0 & 0 & \un & \un \\
-        0 & \un & \un & 0
-      \end{tabular}
-      &&&&
-      \begin{tabular}[h]{llll}
-        $d_1$ & $d_2$ &$d_3$ &$d_4$\\
-        % &&&\\
-        &&&\\
-        \un & \un & 0 & 0\\
-        0 & \un & \un & 0 \\
-        0 & \un & 0 & \un 
-      \end{tabular}
-    \end{tabular}
-  }
-
-  \caption{\figtabsize Matrices with and without COP. $M_1$ has COP
-    because by permuting its columns, $c_1$-$c_4$, one can obtain
-    $M_1'$ where the {\un}s in each row are consecutive. $M_2$,
-    however, does not have COP since no permutation of its columns,
-    $d_1$-$d_4$, will arrange {\un}s in each row consecutively
-    \cite{d08phd}.}
-
-  \label{fig:cop-matrix}
-\end{figure}
-
-Recall that a matrix with COP is one whose rows (columns) can be
-rearranged so that the {\un}s in every column (row) are in consecutive
-rows (columns). Figure~\ref{fig:cop-matrix} shows examples of this
-property.  COP in binary matrices has several practical applications
-in diverse fields including scheduling \cite{hl06}, information
-retrieval \cite{k77} and computational biology \cite{abh98}.  Further,
-it is a tool in graph theory \cite{mcg04} for interval graph
-recognition, characterization of Hamiltonian graphs, planarity testing
-\cite{bl76} and in integer linear programming \cite{ht02,hl06}.
-
-The obvious first questions after being introduced to the consecutive
-ones property of binary matrices are if COP can be detected
-efficiently in a binary matrix and if so, can the COP permutation of
-the matrix also be computed efficiently?  Recognition of COP in a
-binary matrix is polynomial time solvable and the first such algorithm
-was given by \cite{fg65}.  A landmark result came a few years later
-when \cite{at72} discovered the families of forbidden submatrices that
-prevent a matrix from having COP and most, if not all, results that
-came later were based on this discovery which connected COP in binary
-matrices to convex bipartite graphs. In fact, the forbidden
-submatrices came as a corollary to the discovery that convex bipartite
-graphs are AT-free in \cite{at72}\tnote[TH]{check}. The first linear
-time algorithm for COP testing (COT) was invented by \cite{bl76} using
-a data structure called PQ trees.  Since then several COT algorithms
-have been invented -- some of which involved variations of PQ trees
-\cite{mm96,wlh01,mcc04}, some involved set theory and ICPIA
-\cite{wlh02,nsnrs09}, parallel COT algorithms\cite{as95,bs03,ly91} and
-certifying algorithms\cite{mcc04}.
-
-The construction of PQ trees in \cite{bl76} draws on the close
-relationship of matrices with COP to interval graphs. A PQ tree of a
-matrix is one that stores all row (column) permutations of the matrix
-that give the COP orders (there could be multiple orders of rows or
-columns) of the matrix. This is constructed using an elaborate linear
-time procedure and is also a test for planarity\tnote[TH]{check check
-  check. both interval graph and planarity in this paper?}.  PQR trees
-is a generalized data structure based on PQ trees \cite{mm96,mpt98}.
-\cite{tm05} describes an improved algorithm to build PQR
-trees. \tnote[TH]{improv in terms of what?}\cite{wlh02} describes the
-simpler algorithm for COT. Hsu also invented PC trees
-\cite{wlh01}\tnote[TH]{This result first appeared inproc ISAAC92}
-which is claimed to be much easier to implement. \cite{nsnrs09}
-describes a characterization of consecutive-ones property solely based
-on the cardinality properties of the set representations of the
-columns (rows); every column (row) is equivalent to a set that has the
-row (column) indices of the rows (columns) that have one entries in
-this column (row). This is interesting and relevant, especially to
-this thesis because it simplifies COT to a great degree. \tnote[TH]{it
-  reduces the solution search space. fill in the blanks.}
-
-\cite{mcc04} describes a different approach to COT. While all previous
-COT algorithms gave the COP order if the matrix has the property but
-exited stating negative if otherwise, this algorithm gives an evidence
-by way of a certificate of matrix even when it has no COP. This
-enables a user to verify the algorithm's result even when the answer
-is negative. This is significant from an implementation perspective
-because automated program verification is hard and manual verification
-is more viable. Hence having a certificate reinforces an
-implementation's credibility. Note that when the matrix {\em has} COP,
-the COP order is the certificate.  The internal machinery of this
-algorithm is related to the weighted betweenness problem
-addressed\tnote[TH]{in what way??} in \cite{co98}.  \tnote[TH]{expand
-  on the COP order graph creation and it having to be bipartite for M
-  to have COP. and thus an odd cycle being an evidence of no COP.}
-\tnote[TH]{ where should this go?: (1) cite|jlm97 (application of PQ trees
-  in graphics). (2) helly's theorem citation
-  19XXdgk-Hellystheorem-Danzer-Gruenbaum-Klee}
-
-\subsection{Optimization problems in COP}
-\label{sec:optcop}
-
-So far we have been concerned about matrices that have the consecutive
-ones property. However in real life applications, it is rare that data
-sets represented by binary matrices have COP, primarily due to the
-noisy nature of data available. At the same time, COP is not arbitrary
-and is a desirable property in practical data representation
-\cite{co98,jkckv04,k77}. In this context, there are several
-interesting problems when a matrix does not have COP but is ``close''
-to having COP or is allowed to be altered to have COP. These are the
-optimization problems related to a matrix which does not have
-COP. Some of the significant problems are surveyed in this section.
-
-\tnote[TH]{ -- sect 4.1 in cite:d08phd has many results
-  surveyed. hardness results, approx. results. results are usually for
-  a class of matrices $(a,b)$ where number {\un}s in columns and rows
-  are restriced to $a$ and $b$ . -- problem of flipping at most $k$
-  entries of $M$ to make it attain COP. this is NP complete
-  cite:b75-phd}\tnote[TH]{(1) scite:lb62 showed that interval graphs
-  are AT-free.  describe AT (2) show the close relationship b/w COP
-  and graphs sec 2.2, pg 31} \cite{at72} showed that a matrix that
-does not have COP have certain substructures that prevent it from
-having COP. Tucker classified these forbidden substructures into five
-classes of submatrices. This result is presented in the context of
-convex bipartite graphs which \cite{at72} proved to be
-AT-free\tnote[TH]{ check this up. give details. - doms'}. By
-definition, convex bipartite graph have half adjacency matrices that
-have COP on either rows or columns (graph is biconvex if it has COP on
-both)\cite{d08phd}. A half adjacency matrix is a binary matrix
-representing a bipartite graph as follows. The set of rows and the set
-of columns form the two partitions of the graph. Each row node is
-adjacent to those nodes that represent the columns that have {\un}s in
-the corresponding row. \cite{at72} proves that this bipartite graph
-has no asteroidal triple if and only if the matrix has COP and goes on
-to identify the forbidden substructures for these bipartite
-graphs. The matrices corresponding to these substructures are the
-forbidden submatrices.
-
-Once a matrix has been detected to not have COP (using any of the COT
-algorithms mentioned earlier), it is naturally of interest to find out
-the smallest forbidden substructure (in terms of number of rows and/or
-columns and/or number of entries that are {\un}s). \cite{d08phd}
-discusses a couple of algorithms which are efficient if the number of
-{\un}s in a row is small. This is of significance in the case of sparse
-matrices where this number is much lesser than the number of
-columns. $(*,\Delta)${\em -matrices} are matrices with no restriction
-on number of {\un}s in any column but has at most $\Delta$ {\un}s in
-any row. {\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are
-similar problems which deals with inducing COP on a matrix. In {\sc
-  Min COS-R (Min COS-C)} the question is to find the minimum number of
-rows (columns) that must be deleted to result in a matrix with COP.
-In the dual problem {\sc Max COS-R (Max COS-C)} the search is for the
-maximum number of rows (columns) that induces a submatrix with
-COP. Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
-a submatrix $M'$ with all columns\tnote[TH]{check if b75 deals with
-  COP col or COP row. also is it any submatrix with k less than r rows
-  or submatrix must have all columns?} but a maximum cardinality
-subset of rows such that $M'$ has COP is NP complete. \cite{hg02}
-corrects an error of the abridged proof of this reduction as given in
-\cite{gj79}.  \cite{d08phd} discusses all these problems in detail
-giving an extensive survey of the previously existing results which
-are almost exhaustively all approximation results and hardness
-results. Taking this further, \cite{d08phd} presents new results in
-the area of parameterized algorithms for this
-problem\tnote[TH]{elaborate - what are the results?}.
-
-Another problem is to find the minimum number of entries in the matrix
-that can be toggled to result in a matrix with COP.  \cite{v85}
-discusses approximation of {\sc COP Augmentation} which is the problem
-of changing of the minimum number of zero entries to {\un}s so that
-the resulting matrix has COP. As mentioned earlier, this problem is
-known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
-using a reduction to the longest path problem, \tnote[TH]{or is it a
-  survey of another result?  check.} that finding a Tucker's forbidden
-submatrix of at least $k$ rows is NP complete. \tnote[TH]{how is this
-  different from booth's 75 result??}  \tnote[TH]{where should this
-  go? cite|tz04 (approx submatrix with COP sparse matrices)}
-
-\cite{jkckv04} discusses the use of matrices with almost-COP (instead
-of one block of consecutive {\un}s, they have $x$ blocks, or {\em
-  runs}, of consecutive {\un}s and $x$ is not too large) in the
-storage of very large databases.  The problem is that of reordering of
-a binary matrix such that the resulting matrix has at most $k$ runs of
-{\un}s. This is proved to be NP hard using a reduction from the
-Hamiltonian path problem.\tnote[TH]{Theorem 2.1 in jkckv}
-\tnote[TH]{(1) A connection of COP problem to the travelling salesman
-  problem is also introduced. what does this mean? -- COP can be used
-  as a tool to reorder $0.5T \le runs(M) \le T$. (2) The optimization
-  version of the $k$-run problem, i.e. minimization of number of
-  blocks of ones is proven to be NP complete by
-  cite:k77}\tnote[TH]{are these two the same?} \tnote[TH]{what is the
-  reduction?} \tnote[TH]{other problems similar to COP -- cite:ckl96
-  (ILP, circ ones, one drop) -- cite:th98 (generalization of COP -
-  minimax, biotonic column) Tucker}
-
-\section{********* Application of COP in Areas of Graph Theory and
-  Algorithms}
-\label{sec:appcopGTA}
-
-\subsection{********* COP in Relational Database Model}
-\label{sec:apprdbm}
-\tnote{(set systems theme)}
-
-\subsection{********* COP in Graph Isomorphism}
-\label{sec:appgraphiso}
-\tnote{(canonization theme)}
-
-\subsection{********* Certifying Algorithms}
-\label{sec:appcertalgo}
-\tnote{ (certification McC04 theme)}
-
-\section{Generalization of COP - the Motivation}
-\label{sec:motive}
-Section~\ref{sec:copmatrices} introduced a succinct characterization
-for consecutive-ones property which is solely based on the cardinality
-properties of the set representations of the matrix's columns
-\cite{nsnrs09}. This result is very relevant to this thesis because
-aside from it simplifying COT to a great degree, our generalization
-problem is motivated by their results.
-
-\cite{nsnrs09} characterizes interval assignments to the sets which
-can be obtained from a single permutation of the rows.  For an
-assignment to be feasible, the cardinality of the interval assigned to
-each set in the system must be same as the cardinality of the set, and
-the intersection cardinality of any two intervals must be same as the
-intersection cardinality of their corresponding sets.  While this is
-obviously a necessary condition, this result shows this is also
-sufficient.  \cite{nsnrs09} calls this an Intersection Cardinality
-Preserving Interval Assignment (ICPIA).  This paper generalizes the
-idea from \cite{wlh02} of decomposing a given binary matrix into prime
-matrices for COT and describes an algorithm to test if an ICPIA exists
-for a given set system.
-
-The equivalence of the problem of testing for the consecutive-ones
-property to the constraint statisfaction problem of interval
-assignment \cite{nsnrs09} or interval labeling \cite{kklv10} is as
-follows. Every column (row) of the binary matrix can be converted into
-a set of non-negative integers which are the indices of rows (columns)
-with {\un}s in that column (row). It is apparent that if the matrix
-has COP in columns (rows), then constructing such sets after applying
-the COP permutation to the rows (columns) of the matrix will result in
-sets with consecutive integers. In other words, after application of
-COP reordering, the sets are intervals. Indeed the problem now becomes
-finding interval assignments to a given set system such that there
-exists a permutation of the universe of set of row indices (column
-indices) which converts each set to its assigned interval.
-
-The problem of interest in this thesis, namely, tree path labeling
-problem, is a natural generalization of the interval assignment
-problem or the COT problem. The problem is defined as follows -- given
-a set system $\cF$ from a universe $U$ and a target tree $T$, does
-there exist a bijection from $U$ to the vertices of $T$ such that each
-set in the system maps to a path in $T$.  We refer to this as the
-{\CFTPL} problem or simply {\em tree path labeling} problem for an
-input set system and target tree pair -- $(\cF,T)$. The special case
-of the target tree being a path, is the interval assignment problem.
-We focus on generalizing the notion of an ICPIA \cite{nsnrs09} to
-characterize feasible path assignments.  We show that for a given set
-system $\cF$, a tree $T$, and an assignment of paths from $T$ to the
-sets, there is a feasible\footnote{The notion of {\em feasibility} is
-  formally defined in Section~\ref{ch:prelims}.}  bijection between
-$U$ and $V(T)$ if and only if the intersection cardinalities among any
-three sets (not necessarily distinct) is equal to that of the
-corresponding paths assigned to them and the input passes a filtering
-algorithm (described in this paper) successfully.  This algorithmic
-characterization gives a natural data structure that stores all the
-\remove{relevant} feasible bijections between $U$ and $V(T)$. This
-reduces the search space for the solution considerably from the
-universe of all possible bijections between $U$ and $V(T)$ to only
-those bijections that maintain the characterization.  Further, the
-filtering algorithm is also an efficient algorithm to test if a tree
-path labeling\footnote{The terms {\em tree path labeling} and {\em
-    tree path assignment} are, in informal language,
-  synonyms. Formally, the former refers to the bijection $\cl: \cF
-  \rightarrow \cP$. The latter refers to the set of ordered pairs
-  $\{(S, P) \mid S \in \cF, P \in \cP\}$. $\cP$ is a set of paths on
-  $T$.} is feasible.
-
-\section{Summary of New Results in this Thesis}
-\label{sec:results}
-
-We see in Section~\ref{sec:motive} that pairwise intersection
-cardinality preservation is necessary and sufficient for an interval
-assignment to be feasible for a given hypergraph\footnote{A {\em
-    hypergraph} is an alternate representation of a set system and
-  will be used in this thesis. See Section~\ref{ch:prelims} for the
-  formal definition.} and thus is a characterization for COP
-\cite{nsnrs09}. \tnote{Refer to the survey section that will list the
-  theorems and/or lemma?} In our work we extend this characterization
-and find that trio-wise intersection cardinality preservation makes a
-tree path labeling \footnote{A {\em tree path labeling} $\cl$ is a
-  bijection of paths from the target tree $T$ to the hyperedges in
-  given hypergraph $\cF$.  See Section~\ref{ch:prelims} for the formal
-  definition.}  (TPL) feasible, which is a generalization of the COP
-problem. This problem is defined as follows.
-
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \FTPL}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
-        paths $\cP$ from $T$ and a
-        bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
-      \end{minipage}\\
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
-        such that $\phi$ when applied on any hyperedge in $\cF$ will
-        give
-        the path mapped to it by the given tree path labeling $\cl$.\\
-        { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every
-          hyperedge $S \in \cF$.}
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
-
-We give a necessary and sufficient condition by way of {\em
-  Intersection Cardinality Preservation Path Labeling} (ICPPL) and a
-filtering algorithm for {\FTPL} to output in affirmative. ICPPL
-captures the trio-wise cardinality property described
-earlier\footnote{See Section~\ref{sec:feasible} for the definition of
-  ICPPL.}. This characterization can be checked in polynomial time.  A
-relevant consequence of this constructive procedure is that it is
-sufficient to iteratively check if three-way intersection
-cardinalities are preserved.  In other words, in each iteration, it is
-sufficient to check if the intersection of any three hyperedges is of
-the same cardinality as the intersection of the corresponding paths.
-Thus this generalizes the well studied question of the feasible
-interval assignment problem which is the special case when the target
-tree $T$ is simply a path \cite{wlh02,nsnrs09}.
-
-Aside from checking if a given TPL is feasible, we also solve the
-problem of computing a feasible TPL for a given hypergraph and target
-tree, if one exists. This problem, {\CFTPL}, is defined as follows.
-
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \CFTPL}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$ and a tree $T$.\\
-      \end{minipage}\\
-
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a set of paths $\cP$ from $T$ and a bijection
-        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
-        {\bf true} on $(\cF, T, \cl)$.
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
-
-We present a polynomial time algorithm for {\CFTPL} when the target
-tree $T$ belongs to a special class of trees called {\em \kstar s} and
-when the hyperedges in the hypergraph $\cF$ have at most $k+2$
-vertices. A couple of examples of {\kstar s} are given in
-Figure~\ref{fig:ksubstar}.
-
-\begin{figure}[t]
-  \centering
-  \begin{tabular}[h]{ccccc}
-    \includegraphics[scale=0.3]{../img/star.pdf} &&&&
-    \includegraphics[scale=0.3]{../img/kstar.pdf}\\
-    (a) &&&& (b)
-  \end{tabular}
-  \caption{\figtabsize Examples of {\kstar s}. (a) $k = 0$ (b) $k = 2$
-  }
-  \label{fig:ksubstar}
-\end{figure}
-
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \CFTPLKTREE}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$ such that every
-        hyperedge
-        $S \in \cF$ is of cardinality at most $k+2$ and a {\kstar} $T$.\\
-      \end{minipage}\\
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a set of paths $\cP$ from $T$ and a bijection
-        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
-        {\bf true} on $(\cF, T, \cl)$.
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
-
-\tnote{The following sentence seems out of place in this para} In
-spite of this being a restricted case, we believe that our results are
-of significant interest in understanding the nature of {\sc Graph
-  Isomorphism} which is polynomial time solvable in interval graphs
-while being hard on path graphs\cite{kklv10}. {\kstar s} are a class
-of trees which are in many ways very close to intervals or paths. Each
-ray\footnote{The path from a leaf to the root, the vertex with highest
-  degree, is called a {\em ray} of the \kstar. See
-  Section~\ref{ch:prelims}.} are independent except for the
-root\footnote{The vertex with maximum degree in a {\kstar} is called
-  {\em root}. See Section~\ref{ch:prelims}.} and hence can be
-considered as an independent interval till the root. Our algorithm
-builds on this fact and uses the interval assignment
-algorithm\cite{nsnrs09} up until ``reaching'' the root and then uses
-the trio-wise intersection cardinality (the extra condition in ICPPL
-that generalizes ICPIA) check to resolve the ambiguity about which ray
-the algorithm should ``grow'' the solution into in the next iteration.
-
-We also have an algorithm for solving {\CFTPL} with no restrictions on
-the target tree or set size which runs in exponential time.  This
-algorithm finds a path labeling from $T$ by decomposing the problem
-into subproblems of finding path labeling of subsets of $\cF$ from
-subtrees of $T$. Given the fact that binary matrices naturally
-represent a set system (see Section~\ref{sec:motive}) and that the
-{\em overlap}\footnote{}\tnote{Give informal definition of ``overlap''
-  in footnote and link to sec:prelims. ADD the definition in prelim.}
-relation between the sets involved is an obvious equivalence relation,
-$\cF$ quite naturally partitions into equivalence classes known as
-{\em overlap components}\tnote{Give informal definition of ``overlap
-  components'' in footnote and link to sec:prelims. ADD the definition
-  in prelim.}. In the context of COP, overlap components were used in
-\cite{wlh02} and \cite{kklv10}. Moreover, \cite{nsnrs09} discovered
-that these equivalence classes form a total order \tnote{Give informal
-  definition of ``total order'' in footnote?}. We extend this to TPL
-and find that when $\cF$ is a path hypergraph\footnote{If there exists
-  an FTPL for a hypergraph $\cF$, it is called a path hypergraph.},
-the classes can be partially ordered as an in-tree in polynomial
-time. Once $\cF$ is ``broken'' into overlap components, one must
-identify the subtree of $T$ that it needs to map to and this is the
-hard part which is currently open to be solved in polynomial time.
-
-\tnote[TH]{The connection of TPL to graph isomorphism will be made
-  later in the document}
-
-\chapter{Consecutive-ones Property - a Survey of Important Results}
-\label{ch:copsurvey}
-\tnote{TBD: have a few lines about organization of chapter}
-
-\tnote{ Survey chapter: is the full fledged expansion of the survey in
-  introduction with details, observations, theorems etc.}
-
-\section{Matrices with COP}
-\label{sec:surveycoptest}
-\tnote{Expand on ref:sec:copmatrices}
-
-\section{Optimization problems in COP}
-\label{sec:surveycopopt}
-\tnote{Expand on ref:sec:optcop}
-
-\section{********* COP in Relational Database Model}
-\label{sec:surveyrdbm}
-\tnote{Expand on sec:apprdbm}
-\tnote{(set systems theme)}
-
-\subsection{********* COP in Graph Isomorphism}
-\label{sec:surveygraphiso}
-\tnote{Expand on sec:appgraphiso}
-\tnote{(canonization theme)}
-
-\subsection{********* Certifying Algorithms}
-\label{sec:surveycertalgo}
-\tnote{Expand on sec:appcertalgo}
-\tnote{ (certification McC04 theme)}
-
-\chapter{Tree Path Labeling of Path Hypergraphs - the New Results}
-\label{ch:myresearch}
-
-\tnote{UPDATE TO FINAL VERSION. BELOW IS OLDER VERSION WITH INCOMPLETE
-  SEC 4.}  This chapter documents all the new results obtained by us in the
-area of tree path labeling of path hypergraphs which is the parent
-problem addressed in this thesis. In Section~\ref{sec:motive} we see
-that consecutive-ones property and its equivalent problem of interval
-labeling of a hypergraph is a special case of the general problem of tree
-path labeling of path hypergraphs.
-
-\tnote{UPDATE the following para to current skeleton.}The necessary
-preliminaries with definitions etc. are presented in
-Section~\ref{ch:prelims}. Section~\ref{sec:feasible} documents the
-characterization of a feasible path labeling for any path labeling
-from any target trees. Section~\ref{sec:spltargettree} describes the
-special case where the target tree is of a particular family of trees
-and specifically, Section~\ref{sec:ksubdivstar} describes a polynomial
-time algorithm to find the tree path labeling of a given set system
-from a given $k$-subdivided tree. Section~\ref{sec:norestraint}
-describes the general situation where the target tree has no
-restrictions and the algorithm to find a TPL, if any, in this case.
-
-\section{Introduction}
-
-The problem of consecutive-ones property testing can be easily seen as
-a simple constraint satisfaction problem involving a hypergraph or a
-system of sets from a universe. Every column (row) of the binary
-matrix can be converted into a set of non-negative integers which are
-the indices of the rows (columns) with {\un}s in that column
-(row). When observed in this context, if the matrix has the COP on
-columns (rows), a reordering of its rows (columns) will result in sets
-that have only consecutive integers. In other words, the sets after
-applying the COP row (column) permutation are intervals.  In this
-form, one can see that this is indeed the problem of finding interval
-assignments to the given set system \cite{nsnrs09} with a single
-permutation of the universe (set of row or column indices for COP of
-columns or rows, respectively) which permutes each set to its
-interval. The result in \cite{nsnrs09} characterizes interval
-assignments to the sets which can be obtained from a single
-permutation of the rows (columns). They show that for each set, the
-cardinality of the interval assigned to it must be same as the
-cardinality of the set, and the intersection cardinality of any two
-sets must be same as the intersection cardinality of the corresponding
-intervals. This is a necessary and sufficient
-condition. \tnote{(ICPIA) put reference to survey chapter}.  Finally,
-the idea of decomposing a given binary matrix into prime matrices to
-check for COP is adopted from cite|wlh02| to test if an ICPIA exists
-for a given set system.\tnote{REPEATITION from intro chapter. but idea
-  recall is necessary. so keep it but reword it to avoid monotone.}
-
-A natural generalization of the interval assignment problem is
-feasible tree path labeling problem of a set system. The problem is
-defined as follows -- given a set system $\cF$ from a universe $U$ and
-a tree $T$, does there exist a bijection from $U$ to the vertices of
-$T$ such that each set in the system maps to a path in $T$.  We refer
-to this as the {\em tree path labeling problem} for an input set
-system, target tree pair -- $(\cF,T)$. \tnote{USE THE FORMAL PROBLEM DEF BLOCK}
-%\begin{tabular}[t]{c}
-{\small
-\begin{minipage}[h]{5in}
- % \begin{singlespace}
- \vspace{2mm}
- {\large \FTPL}\\
- \begin{tabular}[t]{l|l}
- \hline\\
-    {\tt Input} & 
-    \begin{minipage}[t]{\probdefwidth}
-      A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
-      paths $\cP$ from $T$ and a
-      bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
-    \end{minipage}\\
-    {\tt Question} &
-    \begin{minipage}[t]{\probdefwidth}
-      Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
-      such that $\phi$ when applied on any hyperedge in $\cF$ will
-      give
-      the path mapped to it by the given tree path labeling $\cl$.\\
-      { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every hyperedge
-        $S \in \cF$.}
-    \end{minipage}\\
-  \end{tabular}
-  % \end{singlespace}
-\end{minipage}\\
-}
-%\end{tabular}
-
-As a special case if the tree
-$T$ is a path, the problem becomes the interval assignment problem. \tnote{USE THE FORMAL PROBLEM DEF BLOCK}
-We focus on the question of generalizing the notion of an ICPIA
-\cite{nsnrs09} to characterize feasible path assignments.  We show
-that for a given set system $\cF$, a tree $T$, and an assignment of
-paths from $T$ to the sets, there is a feasible bijection between $U$
-and $V(T)$ if and only if all intersection cardinalities among any
-three sets (not necessarily distinct) is same as the intersection
-cardinality of the paths assigned to them and the input runs a
-filtering algorithm (described in this paper) successfully.  This
-characterization is proved constructively and it gives a natural data
-structure that stores all the relevant feasible bijections between $U$
-and $V(T)$.  Further, the filtering algorithm is also an efficient
-algorithm to test if a tree path labeling to the set system is
-feasible.  This generalizes the result in \cite{nsnrs09}.
-
-In the later part of this paper, we focus on a new special case of the
-tree path labeling problem. Here the set system is such that every set
-is at most $k+2$ in size and for every pair of sets in it there exists
-a sequence of sets between them with consecutive sets in this sequence
-having a strict intersection -- i.e., non-empty intersection with
-neither being contained in the other. Moreover, the given tree is a
-{\em $k$-subdivided star}. We demonstrate a polynomial time algorithm
-to find a feasible path labeling in this case.
-
-\section{Preliminaries to new results}
-\label{ch:prelims}
-
-\tnote{TBD: have a few lines about organization of chapter}This
-section states definitions and basic facts necessary in the scope of
-this document.
-
-\tnote{Move some of this to general prelim in chapter 1}
-
-The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
-  system} of a universe $U$ with $|U| = n$.  The {\em support} of a
-set system $\F$ denoted by $supp(\cF)$ is the union of all the sets in
-$\F$; $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this
-paper, a set system is required to ``cover'' the universe; $
-supp(\cF) = U$.
-
-The graph $T$ represents a {\em target tree} with same
-number of vertices as elements in $U$; $|V(T)| = n$.  A {\em path
-  system}\, $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P
-\mid P \subseteq V, \text{ } T[P] \text{ is a path} \}$.
-
-% 
-% A set system $\cF$ can be alternatively represented by a {\em
-%   hypergraph} $\H_\cF$ whose vertex set is $supp(\cF)$ and
-% hyperedges are the sets in $\cF$. This is a known representation for
-% interval systems in literature \cite{bls99,kklv10}.  We extend this
-% definition here to path systems.
-
-A set system $\cF$ can be alternatively represented by a {\em
-  hypergraph}\, $\cF_H$ whose vertex set is $supp(\cF)$ and hyperedges
-are the sets in $\cF$. This is a known representation for interval
-systems in literature \cite{bls99,kklv10}.  We extend this definition
-here to path systems. Due to the equivalence of set system and
-hypergraph in the scope of this paper, we drop the subscript $_H$ in
-the notation and refer to both the structures by $\cF$.
-
-Two hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to
-each other, denoted by $\cF' \cong \cF''$, iff there exists a
-bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for all
-sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ iff $B$ is
-a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
-\cite{kklv10}. This is called {\em hypergraph
-  isomorphism}. \tnote{also extend $\phi$ to hyperedges -- see if
-  required }
-
-The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph $\cF$ is a
-graph such that its vertex set has a bijection to $\cF$ and there
-exists an edge between two vertices iff their corresponding hyperedges
-have a non-empty intersection \cite{mcg04}.
-
-If the intersection graphs of two hypergraphs are isomorphic,
-$\bI(\cF) \cong \bI(\cP)$ where $\cP$ is also a path system, then the
-bijection $\cl: \cF \rightarrow \cP$ due to this isomorphism is called
-a {\em path labeling} of the hypergraph $\cF$. To illustrate further,
-let $\cg: V(\cF) \rightarrow V(\cP)$ be the above mentioned
-isomorphism where $V(\cF)$ and $V(\cP)$ are the vertex sets that
-represent the hyperedges for each hypergraph respectively, $V(\cF) =
-\{ v_S \mid S \in \cF\}$ and $V(\cP) = \{ v_P \mid P \in \cP\}$. Then
-the path labeling $\cl$ is defined as follows: $\cl(S_1) = P_1$ iff
-$\cg (v_{S_1}) = v_{P_1}$. The path system $\cP$ may be alternatively
-denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In most scenarios in
-this paper, what is given are the pair $(\cF, \cl)$ and the target
-tree $T$; hence this notation will be used more often.
-
-If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$
-is called a {\em path hypergraph} and $\cP$ is called {\em path
-  representation} of $\cF$. If this isomorphism is $\phi: supp(\cF)
-\rightarrow V(T)$, then it is clear that there is an {\em induced path
-labeling} $\cl_\phi: \cF \rightarrow \cP$ to the set system;
-$\cl_\phi(S) = \set{y \mid y = \phi(x), x \in S}$ for all $S \in \cF$. Recall that
-$supp(\cP) = V(T)$.
-
-A path labeling $(\cF, \cl)$ is defined to be {\em
-  feasible} if
-$\cF \cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
-\rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
-\rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.
-
-An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is
-a graph such that its vertex set has a bijection to $\cF$ and there
-exists an edge between two of its vertices iff their corresponding
-hyperedges overlap. Two hyperedges $S$ and $S'$ are said to {\em
-  overlap}, denoted by $S \overlap S'$, if they have a non-empty
-intersection and neither is contained in the other; $S \overlap S'
-\text{ iff } S \cap S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq
-S$. Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$ and not
-necessarily connected. Each connected component of $\bO(\cF)$ is
-called an {\em overlap component}.
-
-A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
-\overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
-\cite{kklv10}. Additionally, if $S$ is such that it is contained in no
-other hyperedge in $\cF$, i.e., it is inclusion maximal then it is called
-{\em super-marginal}.
-
-A {\em star} graph is a complete bipartite graph $K_{1,p}$
-which is clearly a tree and $p$ is the number of leaves. The vertex
-with maximum degree is called the {\em center} of the star and the
-edges are called {\em rays} of the star. A {\em $k$-subdivided star}
-is a star with all its rays subdivided exactly $k$ times. The
-definition of a {\em ray of a $k$-subdivided star} is extended to the path
-from the center to a leaf. It is clear that all rays are of length $k+2$.
-
-\section{Characterization of Feasible Tree Path  Labeling} 
-\label{sec:feasible}
-
-In this section we give an algorithmic characterization of a
-feasibility of tree path labeling.  Consider a path labeling $(\cF,
-\cl)$ on the given tree $T$. We call $(\cF, \cl)$ an {\em Intersection
-  Cardinality Preserving Path Labeling (ICPPL)} if it has the
-following properties.
-
-\begin{enumerate}[{(\icpplpr\ }i) \ \ \ ]
-\item \label{pr:i} $|S| = |\cl(S)|$ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $S \in \cF$
-  % \vspace{\topshrink}
-\item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ \ \ \ \
-  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all distinct $S_1, S_2 \in \cF$
-  % \vspace{\topshrink}
-\item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
-  \cap \cl(S_3)|$ \ \ \ for all distinct $S_1, S_2, S_3 \in \cF$
-\end{enumerate}
-
-The following lemma is useful in subsequent arguments. 
-\begin{lemma}
-  \label{lem:setminuscard}
-  If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
-  (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
-  \cl(S_3))|$.
-\end{lemma}
-\begin{proof}%[Proof of Lemma~\ref{lem:setminuscard}]
-  Let $P_i = \cl(S_i)$, for all $1 \le i \le  3$.
-  $|S_1 \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| =
-  |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to properties (ii)
-  and (iii) of ICPPL, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
-  \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
-  = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. \qed
-\end{proof}
-
-In the remaining part of this section we show that $(\cF,
-\cl)$ is feasible if and only if it is an ICPPL and
-Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
-function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two levels of
-filtering of $(\cF, \cl)$ to make it simpler while retaining the set
-of isomorphisms, if any, between $\cF$ and $\cF^\cl$.
-% One direction of this claim isclear: that if a path
-% labeling is feasible, then all intersection cardinalities are
-% preserved, i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
-% \annote{has no premature exit condition hence any input will go
-%   through it}{Prove that the filtered sets has ICPPL iff input PL
-%   has ICPPL?}. Algorithm~\ref{leafasgn} has an exit condition at
-% line~\ref{xempty}. It can be easily verified that $X$ cannot be
-% empty if $\cl$ is a feasible path labeling. The reason is that a
-% feasible path labeling has an associated bijection between
-% $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that
-% the sets map to paths, ``preserving'' the path labeling.  The rest
-% of the section is devoted to constructively proving that it is
-% sufficient for a path labeling to be an ICPPL and pass the two
-% filtering algorithms.  To describe in brief, the constructive
-% approaches refine an ICPPL iteratively, such that at the end of each
-% iteration we have a ``filtered'' path labeling, and finally we have
-% a path labeling that defines a family of bijections from $supp(\cF)$
-% to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.
-First, we present Algorithm~\ref{perms} or {\tt filter\_1}, and prove
-its correctness.  This algorithm refines the path labeling by
-processing pairs of paths in $\cF^\cl$ that share a leaf until no two
-paths in the new path labeling share any leaf.
-
-\begin{algorithm}[h]
-  \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
-  \label{perms}
-  \begin{algorithmic}[\lndisplay]
-    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$\\
-    \STATE $j \assign 1$\\
-    \WHILE {there is $S_1, S_2 \in \cF_{j-1}$ such that
-      $\cl_{j-1}(S_1)$ and $\cl_{j-1}(S_2)$ have a common leaf in
-      $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
-    \{S_1, S_2\})
-    \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1 \}$ \label{setbreak} 
-    \COMMENT {Remove $S_1$, $S_2$ and add the ``filtered'' sets}
-    \STATE {\bf for} every $S \in \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
-    S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\
-
-    \STATE $\cl_j(S_1 \cap S_2) \assign \cl_{j-1}(S_1) \cap
-    \cl_{j-1}(S_2)$
-    \COMMENT {Carry forward the path labeling for all existing sets other than
-      $S_1$, $S_2$}
-    \STATE $\cl_j(S_1 \setminus S_2) \assign \cl_{j-1}(S_1) \setminus
-    \cl_{j-1}(S_2)$ 
-    \COMMENT {Define path labeling for new sets}
-    \STATE $\cl_j(S_2 \setminus S_1) \assign \cl_{j-1}(S_2) \setminus
-    \cl_{j-1}(S_1)$
-
-    \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of ICPPL}
-    \label{ln:3waycheck}
-    \STATE {\bf exit} \label{ln:exit1} \\
-    \ENDIF
-
-    \STATE $j \assign j+1$\\
-    \ENDWHILE
-    \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
-    \RETURN $(\cF', \cl')$
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma} 
- \label{lem:feasible} 
- In Algorithm~\ref{perms}, if input $(\cF, \cl)$ is a feasible path
- assignment then at the end of $j$th iteration of the {\bf while}
- loop, $j \ge 0$, $(\cF_j, \cl_j)$ is a feasible path assignment.
-\end{lemma}
-\begin{proof}%[Proof of Lemma~\ref{lem:feasible}]
-  We will prove this by mathematical induction on the number of
-  iterations. The base case $(\cF_0, \cl_0)$ is feasible since it is
-  the input itself which is given to be feasible. Assume the lemma is
-  true till $j-1$th iteration. i.e. every hypergraph isomorphism
-  $\phi: supp\left(\cF_{j-1}\right) \rightarrow V\left(T \right)$ that
-  defines $(\cF, \cl)$'s feasibility, is such that the induced path
-  labeling on $\cF_{j-1}$, $\cl_{\phi[{\cF_{j-1}}]}$ is equal to
-  $\cl_{j-1}$. We will prove that $\phi$ is also the bijection that
-  makes $(\cF_j, \cl_j)$ feasible. Note that $supp(\cF_{j-1}) =
-  supp(\cF_{j})$ since the new sets in $\cF_j$ are created from basic
-  set operations to the sets in $\cF_{j-1}$. For the same reason and
-  $\phi$ being a bijection, it is clear that when applying the $\phi$
-  induced path labeling on $\cF_j$, $ \cl_{\phi[{\cF_{j}}]}(S_1
-  \setminus S_2) = \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
-  \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Now observe that $ \cl_j(S_1
-  \setminus S_2) = \cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2) =
-  \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
-  \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Thus the induced path labeling
-  $\cl_{\phi[{\cF_{j}}]} = \cl_{j}$. Therefore lemma is proven.  \qed
-\end{proof}
-
-\begin{lemma}
-  \label{lem:invar1} In Algorithm~\ref{perms}, at the end of $j$th
-  iteration, $j \ge 0$, of the {\bf while} loop, the following
-  invariants are maintained.
-  \begin{enumerate}[I {\ }] %\vspace{\topshrink}
-  \item $\cl_j(R)$ is a path in $T$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ for all $R \in \cF_j$%\vspace{\topshrink}
-  \item $|R| = |\cl_j(R)|$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $R \in
-    \cF_j$%\vspace{\topshrink}
-  \item $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ for all $R, R' \in \cF_j$%\vspace{\topshrink}
-  \item $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
-    \cl_j(R'')|$, \ \ \ for all $R, R', R'' \in \cF_j$
-  \end{enumerate}
-\end{lemma}
-
-\begin{proof}
- Proof is by induction on the number of iterations, $j$. In this
-  proof, the term ``new sets'' will refer to the sets added to $\cF_j$
-  in $j$th iteration in line~\ref{setbreak} of Algorithm~\ref{perms},
-  $S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1$ and its
-  images in $\cl_j$ where $\cl_{j-1}(S_1)$
-  and $\cl_{j-1}(S_2)$ intersect and share a leaf.\\
-  The invariants are true in the base case $(\cF_0, \cl_0)$, since it
-  is the input ICPPL.  Assume the lemma is true till the $j-1$th
-  iteration. Let us consider the possible cases for each of the above invariants for
-  the $j$th iteration.
-
- \begin{enumerate}[\xbullet]
-  \item {\em Invariant} I/II
-    \begin{enumerate}[{I/II}a $|$] % \textbullet 
-    \item {\em $R$ is not a new set.} It is in $\cF_{j-1}$. Thus
-      trivially true by induction hypothesis.
-    \item {\em $R$ is a new set.} If $R$ is in $\cF_{j}$ and not in
-      $\cF_{j-1}$, then it must be one of the new sets added in
-      $\cF_j$. In this case, it is clear that for each new set, the
-      image under $\cl_j$ is a path since by definition the chosen
-      sets $S_1$, $S_2$ are from $\cF_{j-1}$ and due to the while loop
-      condition, $\cl_{j-1}(S_1)$, $\cl_{j-1}(S_2)$ have a
-      common leaf. Thus invariant I is proven.\\
-      Moreover, due to induction hypothesis of invariant III and the
-      definition of $l_j$ in terms of $l_{j-1}$, invariant II is
-      indeed true in the $j$th iteration for any of the new sets.  If
-      $R = S_1 \cap S_2$, $|R| = |S_1 \cap S_2| = |\cl_{j-1}(S_1) \cap
-      \cl_{j-1}(S_2)| = |\cl_j(S_1 \cap S_2)| = |\cl_j(R)|$.
-      If $R = S_1 \setminus S_2$, $|R| = |S_1 \setminus S_2| = |S_1| -
-      |S_1 \cap S_2| = |\cl_{j-1}(S_1)| - |\cl_{j-1}(S_1) \cap
-      \cl_{j-1}(S_2)| = |\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)| =
-      |\cl_j(S_1 \setminus S_2)|
-      = |\cl_j(R)|$. Similarly if $R = S_2 \setminus S_1$.\\
-    \end{enumerate}
-  \item {\em Invariant} III
-    \begin{enumerate}[{III}a $|$]
-    \item {\em $R$ and $R'$ are not new sets.} It is in
-      $\cF_{j-1}$. Thus trivially true by induction hypothesis.
-    \item {\em Only one, say $R$, is a new set.} Due to invariant IV
-      induction hypothesis, Lemma~\ref{lem:setminuscard} and
-      definition of $\cl_j$, it follows that invariant III is true no
-      matter which of the new sets $R$ is equal to. If $R = S_1 \cap
-      S_2$, $|R \cap R'| = |S_1 \cap S_2 \cap R'| = |\cl_{j-1}(S_1)
-      \cap \cl_{j-1}(S_2) \cap \cl_{j-1}(R')| = |\cl_j(S_1 \cap S_2)
-      \cap \cl_j(R')| = |\cl_j(R) \cap \cl_j(R')|$.  If $R = S_1
-      \setminus S_2$, $|R \cap R'| = |(S_1 \setminus S_2) \cap R'| =
-      |(\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)) \cap \cl_{j-1}(R')| =
-      |\cl_{j}(S_1 \cap S_2) \cap \cl_{j}(R')| = |\cl_{j}(R) \cap
-      \cl_{j}(R')|$. Similarly, if $R = S_2 \setminus
-      S_1$. Note $R'$ is not a new set.\\
-
-    \item {\em $R$ and $R'$ are new sets.} By definition, the new
-      sets and their path images in path label $\cl_j$ are disjoint so
-      $|R \cap R'| = |\cl_j(R) \cap \cl_j(R)| = 0$. Thus case proven.
-    \end{enumerate}
-  \item {\em Invariant} IV
-
-    Due to the condition in line~\ref{ln:3waycheck}, this invariant is
-    ensured at the end of every iteration.
-%     \begin{enumerate} [{Case 3.}1:]
-%     \item {\em $R$, $R'$ and $R''$ are not new sets.} Trivially
-%       true by induction hypothesis.
-%     \item {\em Only one, say $R$, is a new set.}
-%       If $R = S_1 \cap S_2$,  from Lemma~\ref{lem:fourpaths} and
-%       invariant III hypothesis,  this case is proven. Similarly if $R$
-%       is any of the other new  sets, the case is proven by also using
-%       Lemma ~\ref{lem:setminuscard}.
-%     \item {\em At least two of $R, R', R''$ are new sets.}
-%       The new sets are disjoint hence this case is vacuously true.
-%     \end{enumerate}
-  \end{enumerate} \qed
-%\vspace{-6mm}
-
-\end{proof}
-
-\begin{lemma}
-  \label{lem:noexit1}
-  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{perms} is
-  feasible, then the set of hypergraph isomorphism functions that
-  defines $(\cF, \cl)$'s feasibility is the same as the set that
-  defines $(\cF_j, \cl_j)$'s feasibility, if any.  Secondly, for any
-  iteration $j > 0$ of the {\em \bf while} loop, the {\em \bf exit}
-  statement in line~\ref{ln:exit1} will not execute.
-\end{lemma}
-\begin{proof}
-  Since $(\cF,\cl)$ is feasible, by Lemma~\ref{lem:feasible}
-  $(\cF_j,\cl_j)$ for every iteration $j > 0$ is feasible.  % Therefore,
-%   every hypergraph isomorphism $\phi: supp(\cF) \rightarrow V(T)$ that
-%   induces $\cl$ on $\cF$ also induces $\cl_{j-1}$ and $\cl_{j}$ on
-%   $\cF_{j-1}$ and $\cF_{j}$ respectively, i.e., $\cl_{\phi[\cF_{j-1}]}
-%   = \cl_{j-1}$ and $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
-%   that for all $x \in supp(\cF)$, for all $v \in V(T)$ the following
-%   hold true.
-  Also, every hypergraph isomorphism $\phi: supp(\cF) \rightarrow
-  V(T)$ that induces $\cl$ on $\cF$ also induces $\cl_{j}$ on
-  $\cF_{j}$, i.e., $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
-  that for all $x \in supp(\cF)$, for all $v \in V(T)$, if $(x,v) \in
-  \phi$ then $v \in \cl_{j}(S)$ for all $S \in \cF_{j}$ such that $x
-  \in S$.
-% the following
-%   hold true.
-%   \begin{enumerate}[i. ]
-%   \item If $(x,v) \in \phi$ then $v \in \cl_{j-1}(S)$ for all $S \in
-%     \cF_{j-1}$ such that $x \in S$.
-%   \item If $(x,v) \in \phi$ then $v \in \cl_{j}(S)$ for all $S \in
-%     \cF_{j}$ such that $x \in S$
-%   \end{enumerate}
-  In other words, filter 1 outputs a filtered path labeling that
-  ``preserves''
-  hypergraph isomorphisms of the original path labeling.\\
-  Secondly, line~\ref{ln:exit1} will execute iff the exit condition in
-  line~\ref{ln:3waycheck}, i.e. failure of three way intersection
-  preservation, becomes true in any iteration of the {\em \bf while}
-  loop.  Due to Lemma~\ref{lem:invar1} Invariant IV, the exit
-  condition does not occur if the input is a feasible ICPPL.\qed
-
-%   such that $\phi(x) = v$ where $v$ is the leaf considered in the
-%   first iterations of while. Clearly, $\phi$ is a renaming of
-%   vertices in hypergraph $\cF$ to those in hypergraph $\cF^\cl$. Thus
-%   the following facts can be observed in every iteration of the loop.
-
-%   \begin{enumerate}[\hspace{2mm}i. ] \vspace{\topshrink}
-%   \item all intersection cardinalities are preserved in this path
-%     labeling \vspace{\topshrink}
-%   \item element $x$ is exclusive in a hyperedge in $\cF$ since $v$ is
-%     exclusive in a hyperedge in $\cF^\cl$.
-%   \end{enumerate}
-
-%   Thus the exit condition is never rendered true after $x$ and $v$ are
-%   removed from their respective hyperedges. \qed
-
-% \noindent
-% This proof uses mathematical induction on the number
-%   of iterations $j$, $j \ge 0$, of the loop that executed
-%   without exiting. The base case, $j = 0$ is obviously true since the
-%   input is an ICPPL and the exit condition cannot hold true due to
-%   ICPPL property (iii).  Assume the algorithm executes till the end
-%   of $j-1$th iteration without exiting at line
-%  ~\ref{ln:3waycheck}. Consider the $j$th iteration. From Lemma
-%  ~\ref{lem:feasible} we know that $(\cF_j, \cl_j)$ and $(\cF_{j-1},
-%   \cl_{j-1})$ are feasible\remove[AS]{and from the proof in lemma
-%     lem:invar1 we know that $(\cF_{j-1}, \cl_{j-1})$ satisfies all the
-%     invariants defined in the lemma}.  Thus there exists a bijection
-%   $\phi: supp(\cF) \rightarrow V(T)$ such that the induced path
-%   % labeling on $\cF_{j-1}$ $\cl_{\phi[\cF_{j-1}]} = \cl_{j-1}$.
-%   labeling on $\cF_{j}$, $\cl_{\phi[\cF_{j}]}$ and on $\cF_{j-1}$,
-%   $\cl_{\phi[\cF_{j-1}]}$ are equal to $\cl_{j}$ and $\cl_{j-1}$
-%   respectively.  We need to prove that for any $R, R', R'' \in
-%   \cF_{j}$, $|R \cap R' \cap R''| = |\cl_j(R) \cap \cl_j(R') \cap
-%   \cl_j(R'')|$.
-%   The following are the possible cases that could arise. From argument
-%   above, $|\cl_j(R) \cap \cl_j(R') \cap \cl_j(R'')| =
-%   |\cl_{\phi[\cF_{j}]}(R) \cap \cl_{\phi[\cF_{j}]} (R') \cap
-%   \cl_{\phi[\cF_{j}]} (R'')|$
-
-%   \begin{enumerate}[a $|$]
-%   \item {\em None of the sets are new. $R, R', R'' \in \cF_{j-1}$.}
-%     We know $(\cF_{j-1}, \cl_{j-1})$ is feasible. Thus $|R \cap R'
-%     \cap R''| = |\cl_{j-1}(R) \cap \cl_{j-1}(R') \cap \cl_{j-1}(R'')|
-%     = |\cl_{j}(R) \cap \cl_{j}(R') \cap \cl_{j}(R'')|$.
-%   \item {\em Only one, say $R$, is a new set.}  Let $R = S_1 \cap S_2$
-%     ($S_1, S_2$ are defined in the proof of lemma
-%    ~\ref{lem:invar1}). Now we have $|R \cap R' \cap R''| = |S_1 \cap
-%     S_2 \cap R' \cap R''| = |\cl_{j-1}(S_1) \cap \cl_{j-1}(S_2) \cap
-%     \cl_{j-1}(R') \cap \cl_{j-1}(R'')| = |\cl_{j}(R) \cap \cl_{j}(R')
-%     \cap \cl_{j}(R'')|$. Thus proven. If $R$ is any of the other new
-%     sets, the same claim can be verified using lemma
-%    ~\ref{lem:setminuscard}.
-%     % \item []{\bf Case 3:}
-%   \item {\em At least two of $R, R', R''$ are new sets.}  The new sets
-%     are disjoint hence this case is vacuously true.
-%   \end{enumerate}
-%   \qed
-%   \tnote[E2]{remove the induction proof. just text saying x and v are
-%     exclusive in these sets therefore the intersection cardinalities
-%     don't change thus all invariants are still true} 
-\end{proof}
-
-As a result of Algorithm~\ref{perms} each leaf $v$ in $T$
-is such that there is exactly one set in $\cF$ with $v$ as a vertex in
-the path assigned to it.  In Algorithm~\ref{leafasgn} we identify
-elements in $supp(\cF)$ whose images are leaves in a hypergraph
-isomorphism if one exists.  Let $S \in \cF$ be such that $\cl(S)$ is a
-path with leaf and $v \in V(T)$ is the unique leaf incident on it.  We
-define a new path labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) =
-\set{v}$ where $x$ an arbitrary element from $S \setminus \bigcup_{\hS
-  \ne S} \hS$. In other words, $x$ is an element present in no other
-set in $\cF$ except $S$. This is intuitive since $v$ is present in no
-other path image under $\cl$ other than $\cl(S)$.  The element $x$ and
-leaf $v$ are then removed from the set $S$ and path $\cl(S)$
-respectively. After doing this for all leaves in $T$, all path images
-in the new path labeling $\cl_{new}$ except leaf labels (a path that
-has only a leaf is called the {\em leaf label} for the corresponding
-single element hyperedge or set) are paths from a new pruned tree $T_0
-= T \setminus \{v \mid v \text{ is a leaf in }
-T\}$. Algorithm~\ref{leafasgn} is now presented with details.
-
-\begin{algorithm}[h]
-  \caption{Leaf labeling from an ICPPL {\tt filter\_2($\cF, \cl, T$)}}
-  \label{leafasgn}
-  \begin{algorithmic}[\lndisplay]
-    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$
-    \COMMENT {Path images are such that no two path images share a
-      leaf.}
-    \STATE $j \assign 1$\\
-    \WHILE {there is a leaf $v$ in $T$ and a unique $S_1 \in
-      \cF_{j-1}$ such that $v \in \cl_{j-1}(S_1)$ }\label{uniqueleaf}
-    \STATE $\cF_j \assign \cF_{j-1} \setminus \{S_1\}$\\
-    \STATE for all $S \in \cF_{j-1}$ such that $S \ne S_1$ set
-    $\cl_j(S) \assign
-    \cl_{j-1}(S)$\\
-    \STATE $X \assign S_1 \setminus \bigcup_{S \in \cF_{j-1}, S \ne S_1}S$\\
-    \IF{$X$ is empty} \label{xempty} \STATE {\bf exit} \label{ln:exit2} \ENDIF
-    \STATE $x \assign $ arbitrary element from $X$\\
-    \STATE $\cF_j \assign \cF_j \cup \{\{x\}, S_1 \setminus \{x\}\} $\\
-    \STATE $\cl_j(\{x\}) \assign \{v\}$\\
-    \STATE $\cl_j(S_1 \setminus \{x\}) \assign \cl_{j-1}(S_1) \setminus \{v\}$\\
-    \STATE $j \assign j+1$\\
-    \ENDWHILE
-    \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
-    \RETURN $(\cF', \cl')$
-  \end{algorithmic}
-\end{algorithm}
-
-Suppose the input ICPPL $(\cF, \cl)$ is feasible, yet set $X$ in
-Algorithm~\ref{leafasgn} is empty in some iteration of the {\bf while}
-loop. This will abort our procedure of finding the hypergraph
-isomorphism. The following lemma shows that this will not happen.
-
-\begin{lemma}
-  \label{lem:xnotempty}
-  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{leafasgn} is
-  feasible, then for all iterations $j > 0$ of the {\em \bf while}
-  loop, the {\em \bf exit} statement in line~\ref{ln:exit2} does not
-  execute.
-\end{lemma}
-\begin{proof}
-  Assume $X$ is empty for some iteration $j > 0$. We know that $v$ is
-  an element of $\cl_{j-1}(S_1)$. Since it is uniquely present in
-  $\cl_{j-1}(S_1)$, it is clear that $v \in \cl_{j-1}(S_1) \setminus
-  \bigcup_{(S \in \cF_{j-1}) \wedge (S \ne S_1)}\cl_{j-1}(S)$.  Note
-  that for any $x \in S_1$ it is contained in at least two sets due to
-  our assumption about cardinality of $X$. Let $S_2 \in \cF_{j-1}$ be
-  another set that contains $x$. From the above argument, we know $v
-  \notin \cl_{j-1}(S_2)$. Therefore there cannot exist a hypergraph
-  isomorphism bijection that maps elements in $S_2$ to those in
-  $\cl_{j-1}(S_2)$. This contradicts our assumption that the input is
-  feasible. Thus $X$ cannot be empty if input is ICPPL and feasible.
-  \qed
-\end{proof}
-
-\begin{lemma}
-  \label{lem:invar3}
-  In Algorithm~\ref{leafasgn}, for all $j > 0$, at the end of the
-  $j$th iteration of the {\bf while} loop the four invariants given in
-  Lemma~\ref{lem:invar1} hold.
-\end{lemma}
-\begin{proof}
-  By Lemma~\ref{lem:xnotempty} we know that set $X$ will not be empty
-  in any iteration of the {\em \bf while} loop if input ICPPL $(\cF,
-  \cl)$ is feasible and $\cl_j$ is always computed for all $j >
-  0$. Also note that removing a leaf from any path keeps the new path
-  connected. Thus invariant I is obviously true. In every iteration $j
-  > 0$, we remove exactly one element $x$ from one set $S$ in $\cF$
-  and exactly one vertex $v$ which is a leaf from one path
-  $\cl_{j-1}(S)$ in $T$. This is because as seen in
-  Lemma~\ref{lem:xnotempty}, $x$ is exclusive to $S$ and $v$ is
-  exclusive to $\cl_{j-1}(S)$. Due to this fact, it is clear that the
-  intersection cardinality equations do not change, i.e., invariants
-  II, III, IV remain true. On the other hand, if the input ICPPL is
-  not feasible the invariants are vacuously true. \qed
-\end{proof}
-
-% \textcolor{cyan}{
-% \begin{lemma}
-%   \label{lem:notfeasibleexit}
-%   \tnote{IS THIS CORRECT?}  If input ICPPL $(\cF, \cl)$ is not
-%   feasible, then in one of the recursive calls to Algorithm --3--, the
-%   {\em \bf exit} statement in line x in Algorithm --1 or line y in
-%   Algorithm --2 will get executed.
-% \end{lemma}
-% }
-
-We have seen two filtering algorithms above, namely,
-Algorithm~\ref{perms} {\tt filter\_1} and Algorithm~\ref{leafasgn}
-{\tt filter\_2} which when executed serially repectively result in a
-new ICPPL on the same universe $U$ and tree $T$. We also proved that
-if the input is indeed feasible, these algorithms do indeed output the
-filtered ICPPL. Now we present the algorithmic characterization of a
-feasible tree path labeling by way of Algorithm~\ref{al:icppl-find-isomorph}.
-
-Algorithm~\ref{al:icppl-find-isomorph} computes a
-hypergraph isomorphism $\phi$ recursively using Algorithm~\ref{perms}
-and Algorithm~\ref{leafasgn} and pruning the leaves of the input
-tree. In brief, it is done as follows. Algorithm~\ref{leafasgn} gives
-us the leaf labels in $\cF_2$, i.e., the elements in $supp(\cF)$ that
-map to leaves in $T$, where $(\cF_2, \cl_2)$ is the output of
-Algorithm~\ref{leafasgn}. All leaves in $T$ are then pruned away. The
-leaf labels are removed from the path labeling $\cl_2$ and the
-corresponding elements are removed from the corresponding sets in
-$\cF_2$. This tree pruning algorithm is recursively called on the
-altered hypergraph $\cF'$, path label $\cl'$ and tree $T'$. The
-recursive call returns the bijection $\phi''$ for the rest of the
-elements in $supp(\cF)$ which along with the leaf labels $\phi'$ gives
-us the hypergraph isomorphism $\phi$.  The following lemma formalizes
-the characeterization of feasible path labeling.
-
-\begin{algorithm}[h]
-  \caption{{\tt get- hypergraph- isomorphism ($\cF, \cl, T$)}}
-  \label{al:icppl-find-isomorph}
-  \begin{algorithmic}[\lndisplay]
-
-    \IF{$T$ is empty}
-    \RETURN $\emptyset$\\
-    \ENDIF
-    \STATE $L \assign \{v \mid v \text{ is a leaf in }      T\}$\\
-    \STATE $(\cF_1, \cl_1) \assign$ {\tt filter\_1($\cF, \cl,
-      T$)}\\
-    \STATE $(\cF_2, \cl_2) \assign$ {\tt filter\_2($\cF_1,
-      \cl_1, T$)}\\
-
-    \STATE $(\cF', \cl') \assign (\cF_2, \cl_2)$\\
-    \STATE $\phi' \leftarrow \emptyset$
-
-    \FOR {every $v \in L$}
-    \STATE $\phi'(x) \assign v$ where $x \in \cl_2^{-1}(\{v\})$
-    \COMMENT {Copy the leaf labels to a one to one function $\phi':
-      supp(\cF) \rightarrow L$
-      }\\
-    \STATE Remove $\{x\}$ and $\{v\}$ from $\cF'$, $\cl'$  appropriately\\
-    \ENDFOR
-
-    \STATE $T' \assign T \setminus L$
-
-    \STATE $\phi'' \assign$ {\tt
-      get-hypergraph-isomorphism($\cF', \cl', T'$)}
-    \STATE $\phi \assign \phi'' \cup \phi'$ \\
-    \RETURN $\phi$
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma}
-  \label{lem:hyperiso}  %{lem:perm}
-  If $(\cF, \cl)$ is an ICPPL from a tree $T$ and
-  Algorithm~\ref{al:icppl-find-isomorph}, {\tt
-    get- hypergraph- isomorphism ($\cF, \cl, T$)} returns a non-empty
-  function, then there exists a hypergraph isomorphism $\phi :
-  supp(\cF) \rightarrow V(T)$ such that the $\phi$-induced tree path
-  labeling is equal to $\cl$ or $\cl_\phi = \cl$.
-\end{lemma}
-\begin{proof}
-  It is clear that in the end of every recursive call to
-  Algorithm~\ref{al:icppl-find-isomorph}, the function $\phi'$ is one
-  to one involving all the leaves in the tree passed as input to that
-  recursive call. Moreover, by Lemma~\ref{lem:noexit1} and
-  Lemma~\ref{lem:xnotempty} it is consistent with the tree path
-  labeling $\cl$ passed. The tree pruning is done by only removing
-  leaves in each call to the function and is done till the tree
-  becomes empty. Thus the returned function $\phi: supp(\cF)
-  \rightarrow V(T)$ is a union of mutually exclusive one to one
-  functions exhausting all vertices of the tree. In other words, it is
-  a bijection from $supp(\cF)$ to $V(T)$ inducing the given path
-  labeling $\cl$ and thus a hypergraph isomorphism. \qed
-\end{proof}
-
-\begin{theorem}
-  \label{th:charac}
-  A path labeling $(\cF, \cl)$ on tree $T$ is feasible iff it is an
-  ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
-  T)$ as input returns a non-empty function.
-\end{theorem}
-\begin{proof}
-  From Lemma~\ref{lem:hyperiso}, we know that if $(\cF, \cl)$ is an
-  ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
-  T)$ as input returns a non-empty function, $(\cF, \cl)$ is feasible.
-  Now consider the case where $(\cF, \cl)$ is feasible. i.e. there
-  exists a hypergraph isomorphism $\phi$ such that $\cl_\phi =
-  \cl$. Lemma~\ref{lem:noexit1} and Lemma~\ref{lem:xnotempty} show us
-  that filter 1 and filter 2 do not exit if input is feasible. Thus
-  Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
-  function.\qed
-\end{proof}
-
-\section{ Computing feasible TPL with special target trees \tnote{give
-    problem definition etc}}
-\label{sec:spltargettree}
-
-Section~\ref{sec:feasible} described properties that a TPL must have
-for it to be feasible. The next problem of interest is to test if a
-given hypergraph is a path hypergraph with respect to a given target
-tree\footnote{A larger problem would be to find if a given hypergraph
-  is a path graph on any tree. This problem is not addressed in this
-  thesis.}. In other words, the problem is to find out if a feasible
-tree path labeling exists from a given target tree for a given
-hypergraph. In this section we will see two special cases of this
-problem where the target tree is from a particular family of
-trees. The first one, where the tree is a path, is a shown to be
-equivalent to the well studied problem of consecutive-ones in
-Section~\ref{sec:icpplicpia}. The second one, where the tree is a
-$k$-subdivided tree\footnote{See Section~\ref{ch:prelims} for the
-  formal definition.}, has been solved using a polynomial time
-algorithm. The latter problem enforces some conditions on the
-hypergraph too which will be seen in Section~\ref{sec:ksubdivstar}.
-
-\subsection{Target tree is a Path}
-\label{sec:icpplicpia}
-
-Consider a special case of ICPPL with the following properties.
-\begin{enumerate}
-\item Given tree $T$ is a path. Hence, all path labels are interval
-  labels.
-\item Only pairwise intersection cardinality preservation is
-  sufficient. i.e. property (iii) in ICPPL is not enforced.
-\item The filter algorithms do not have {\em \bf exit} statements.
-\end{enumerate}
-This is called an Intersection Cardinality Preservation Interval
-Assignment (ICPIA) \cite{nsnrs09}. This structure and its algorithm is
-used in the next section for finding tree path labeling from a
-$k$-subdivided star due to this graph's close relationship with
-intervals.
-
-\subsection{Target tree is a $k$-subdivided Star}
-\label{sec:ksubdivstar}
-
-\begin{figure}[t] %[htbp] 
-  \centering
-  \begin{tabular}{lr}
-    (a) \includegraphics{cop-tree_thesis.1}&
-    (b) \includegraphics{cop-tree_thesis.2}
-  \end{tabular}
-  \caption{\figtabsize (a) $8$-subdivided star with 7 rays (b)
-    3-subdivided star with 3 rays}
-  \label{fig:kstar}
-\end{figure}
-
-In this section we consider the problem of assigning paths from a
-$k$-subdivided star $T$ to a given set system $\cF$.  We consider
-$\cF$ for which the overlap graph $\bO(\cF)$ is connected.  The
-overlap graph is well-known from the work of
-\cite{kklv10,nsnrs09,wlh02}.  We use the notation in
-\cite{kklv10}. Recall from Section~\ref{ch:prelims} that hyperedges
-$S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$
-and $S'$ have a non-empty intersection but neither of them is
-contained in the other. The overlap graph $\bO(\cF)$ is a graph in
-which the vertices correspond to the sets in $\cF$, and the vertices
-corresponding to the hyperedges $S$ and $S'$ are adjacent if and only
-if they overlap.  Note that the intersection graph of $\cF$,
-$\bI(\cF)$ is different from $\bO(\cF)$ and $\bO(\cF) \subseteq
-\bI(\cF)$.  A connected component of $\bO(\cF)$ is called an overlap
-component of $\cF$.  An interesting property of the overlap components
-is that any two distinct overlap components, say $\cO_1$ and $\cO_2$,
-are such that any two sets $S_1 \in \cO_1$ and $S_2 \in \cO_2$ are
-disjoint, or, w.l.o.g, all the sets in $\cO_1$ are contained within
-one set in $\cO_2$.  This containment relation naturally determines a
-decomposition of the overlap components into rooted containment trees.
-We consider the case when there is only one rooted containment tree,
-and we first present our algorithm when $\bO(\cF)$ is connected.  It
-is easy to see that once the path labeling to the overlap component in
-the root of the containment tree is achieved, the path labeling to the
-other overlap components in the rooted containment tree is essentially
-finding a path labeling when the target tree is a path: each target
-path is a path that is allocated to sets in the root overlap
-component.  Therefore, for the rest of this section, $\bO(\cF)$ is a
-connected graph. We also assume that all hyperedges are of cardinality
-at most $k+2$.
-
-Recall from Section~\ref{ch:prelims} that a $k$-subdivided star is a
-star with each edge subdivided $k$ times. Therefore, a $k$-subdivided
-star has a central vertex which we call the {\em root}, and each root
-to leaf path is called a {\em ray}. First, we observe that by removing
-the root $r$ from $T$, we get a collection of $p$ vertex disjoint
-paths of length $k+1$, $p$ being the number of leaves in $T$.  We
-denote the rays by $R_1, \ldots, R_p$, and the number of vertices in
-$R_i$, $i \in [p]$ is $k+2$.  Let $\seq{v_{i1},\ldots,v_{i(k+2)}=r}$
-denote the sequence of vertices in $R_i$, where $v_{i1}$ is the
-leaf. Note that $r$ is a common vertex to all $R_i$.
-
-In this section the given hypergraph, the $k$-subdivided star and the
-root of the star are denoted by $\cO$, $T$ and vertex $r$,
-respectively.
-% For each hyperedge $X \in \cO$, we will maintain a 2-tuple of non-negative
-% numbers $\seq{p_1(X), p_2(X)}$.  The numbers satisfy the property that
-% $p_1(X) + p_2(X) \leq |X|$, and at the end of path labeling, for each
-% $X$, $p_1(X) + p_2(X) = |X|$.  This signifies the algorithm tracking
-% the lengths of subpaths of the path assigned to $X$ from at most two
-% rays. We also maintain another parameter called the {\em residue} of
-% $X$ denoted by $s(X)=|X| - p_1(X)$. This signifies the residue path
-% length that must be assigned to $X$ which must be from another
-% ray. For instance, if $X$ is labeled a path from only one ray, then
-% $p_1(X) = |X|$, $p_2(X) = 0$ and $s(X) = 0$.
-
-% 
-% We iteratively consider each ray from which paths will be assigned to
-% hyperedges. At the beginning of each iteration hyperedges of $\cO$ are
-% classifed into the following disjoint sets.
-% \begin{enumerate}
-% \item [$\cL_1^i$] {\em Labeled without $r$.} Those that have been
-%   labeled with a path which does not contain $r$ in one of the
-%   previous iterations.\\  $\cL_1^i = \set{ X \mid p_1(X) = |X| \text{ and
-%     } p_2(X) = 0 \text{ and } s(X) = 0, X \in \cO}$
-% \item [$\cL_2^i$] {\em Labeled with $r$.} Those that have been labeled
-%   with two subpaths of $\cl(X)$ containing $r$ from two different rays
-%   in two previous iterations.\\ $\cL_2^i = \set{X \mid 0 < p_1\left(X\right),
-%     p_2\left(X\right) < |X| \text{ and } s(X) = 0, X \in \cO}$
-%   \item [$\cT_1^i$] {\em Type 1 / partially labeled.} Those that have
-%     been labeled with one path containing $r$ from a single ray in one
-%     of the previous iterations. Here, $p_1(X)$ denotes the length of
-%     the subpath of $\cl(X)$ that $X$ has been so far labeled
-%     with. Also, it is clear that such a path must start at $r$ and
-%     must be in a ray different from the one corresponding to $p_1(X)$.\\
-%     $\cT_1^i = \set{ X \mid 0 < p_1(X) < |X| \text{ and } p_2(X) = 0
-%       \text{ and } s(X) > 0, X \in \cO}$
-%   \item [$\cT_2^i$] {\em Type 2 / not labeled.} Those that have not been
-%     labeled with a path in any previous iteration.\\
-%     $\cT_2^i = \set{ X \mid p_1(X) = p_2(X) = 0 \text{ and } s(X) = |X|,
-%       X \in \cO}$
-% \end{enumerate}
-% \vspace{-2mm}
-% \begin{align*}
-%   \cO &= \cL_1^i \cup \cL_2^i \cup \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]\\
-%   \cO_i &= \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]
-% \end{align*}
-
-The set $\cO_i$ refers to the set of hyperedges $\cT_1^i \cup \cT_2^i$
-in the $i$th iteration.  Note that $\cO_1 = \cO$.  In the $i$th
-iteration, hyperedges from $\cO_i$ are assigned paths from $R_i$ using
-the following rules. Also the end of the iteration, $\cL_1^{i+1},
-\cL_2^{i+1}, \cT_1^{i+1}, \cT_2^{i+1}$ are set to $\cL_1^{i},
-\cL_2^{i}, \cT_1^{i}, \cT_2^{i}$ respectively, along with some
-case-specific changes mentioned in the rules below.
-
-\noindent
-Suppose the given hypergraph has a feasible path labeling to the given
-tree $T$. Let this labeling be $\cl$. Now we present an algorithm that
-lists the sequence of edges that maintain a set of properties
-necessary for them to have a feasible tree path labeling.
-\begin{enumerate}[I.]
-\item {\bf Step 1}
-  \begin{enumerate}
-  \item {\bf {There are type 1
-        edges, $\mathbf{|\cT_1^i| > 0}$}:} 
-% , and for each such $X$ assign the
-%     path in $R_i$ of length $s(X)$ starting at $v_{i,k+2}=r$.  $X$ is
-%     then removed from $\cO_i$, and $p_2(X)=s(X)$, and $s(X)=0$.If any
-%     intersection cardinality is violated because of this assignment,
-%     EXIT reporting the non-existence of a feasible assignment.
-  {If there is only one $\cT_1^i$ hyperedge, label it to the unique
-    path in $R_i$ of length $s(X)$ starting at $v_{i(k+2)}=r$. If
-    there are more than one $\cT_1^i$ hyperedges, for any pair of
-    hyperedges $S_1, S_2 \in \cT_1^i$, one of the following is
-    true. Let $\cl'(X) \subseteq R_j$ denote the subpath of length
-    $p_1(X)$ that has been assigned to $X$ from a previously
-    considered ray $R_j$.}
-  \begin{enumerate}[{\em Case }1. ]
-  \item $|\cl'(S_1) \cap \cl'(S_2)| = 1$ and $|S_1 \cap S_2| > 1$
-    i.e. they are in different rays and their residues must contain
-    their intersection. Hence assign both to $R_i$.
-  \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| >
-    |\cl'(S_1) \cap \cl'(S_2)|$ i.e. they are in the same rays and
-    their residues must contain their intersection. Hence assign both
-    to $R_i$.
-  \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| =
-    |\cl'(S_1) \cap \cl'(S_2)|$. i.e. the intersection cardinality has
-    been met. Hence assign only $S_1$ to $R_i$ and $S_2$ must be
-    assigned to some $i \ge p$.
-  \end{enumerate}
-  Remove the labeled edge or edges from $\cO_i$ and add it to
-  $\cO_{i+1}$. Also do not add it to $\cT_1^{i+1}$ and add to
-  $\cL_2^{i+1}$.
-\item {\bf {There are no type-1 edges nor} super-marginal type-2
-    hyperedges:} If all type-2 edges of length at most $k+1$, EXIT by
-  saying there is no ICPPL.  Otherwise, Find an $X$ of type-2 such
-  that $|X| \geq k+2$, assign it the unique path starting at
-  $v_{i(k+1)}=r$ of length $k+2$, and set $p_1(X)=k+2,
-  s(X)=|X|-(k+2)$, assign it the unique path starting at
-  $v_{i(k+1)}=r$ of length $k+2$, mark this as a type-1 hyperedge, and
-  add it to $\cO_{i+1}$ after removing it from $\cO_i$.
-  % \remove{Next, iteratively, consider each type-1 edge $X \in
-  %   \cO_i$, and if it intersects a previously assigned $Y$ such that
-  %   $\cl(Y) \subseteq R_i$, assign a path of length $s(X)$
-  %   containing $v_{i(k+2)}=r$, set $p_2(X)=s(X), s(X)=0$, and remove
-  %   $X$ from $\cO_i$.}\tnote{covered in step 2}
-
-\item {\bf There are super-marginal type-2 hyperedges {and no type-1
-      edges:}} Let $X$ be a super-marginal hyperedge. If $|X| \leq
-  k+2$, then $X$ is assigned the path of length $|X|$ starting at
-  $v_{i1}$, and $p_1(X) = |X|, p_2(X)=0$. $X$ is removed from $\cO_i$
-  and not added to $\cO_{i+1}$.  If $|X| > k+2$, then $p_1(X)=k+2$,
-  assign it the unique path starting at $v_{i(k+1)}=r$ of length $k+2$
-  and $s(X) = |X| - (k+2)$.  In this case, $X$ is classified as a
-  Type-1 edge, removed from $\cO_i$, and added to $\cO_{i+1}$.
-\end{enumerate}
-\item {\bf Step 2:} Iteratively, a hyperedge $X$ is selected from
-  $\cO_i$ that has an overlap with one of the hyperedges $Y$ such that
-  $\cl(Y) \subseteq R_i$, and a unique path is assigned to $X$.  The
-  path, say $U(X) \subseteq R_i$, that is assigned can be decided
-  unambiguously since the $X \overlap Y$, and all intersection
-  cardinalities can be preserved only at one of the ends of $\cl(Y)$.
-
-  \noindent
-  Let $\cl(X)$ denote the unique path assigned to $X$.  If $X$ is a
-  type-2 hyperedge and if the unique path of length $|X|$ does not
-  contain $r$, then $p_1(X) = |X|, p_2(X)=0$, and $X$ is removed from
-  $\cO_i$.  In the case when $\cl(X)$ has to contain $r$, then
-  $p_1(X)$ is the length of the path, $p_2(X)=0$, and $s(X) =
-  |X|-p_1(X)$. Further $X$ is classified as a type-1 hyperedge, added
-  to $\cO_{i+1}$ and removed from $\cO_i$.  In the case when $X$ is a
-  type-1 hyperedge, then we check if $U(X)$, which is of length $s(X)$
-  contains $r$. If it does, then we assign $\cl(X) \leftarrow \cl(X)
-  \cup U(X)$, remove $X$ from $\cO_i$ and do not add it to
-  $\cO_{i+1}$.  If not, then we {\bf Exit} reporting that an
-  assignment cannot be found.  The iteration ends when no hyperedge in
-  $\cO_i$ has an overlap with a hyperedge assigned to $R_i$.
-\end{enumerate}
-
-\noindent
-The order in which they are assigned is the sequence of hyperedges
-with the following properties. Note that if such a sequence cannot be
-computed, $\cO$ cannot have a feasible path labeling from $T$.
-
-\noindent
-\tnote{need to formalize the below properties} In the following
-lemmas we identify a set of necessary conditions for $\cF$ to have an
-ICPPL in the $k$-subdivided star $T$.  If during the execution of the
-above algorithm, one of these necessary conditions is not satisfied,
-the algorithm exits reporting the non-existence of an ICPPL.
-\begin{lemma}
-  Let all hyperedges in $\cO_i$ be type-1 edges.  Then there is a
-  maximal subset $\ccT_i \subseteq \cO_i$ with the following
-  properties:
-  \begin{enumerate}
-  \item $\ccT_i$ form an inclusion chain.
-  \item For all $X \in \ccT_i$, $s(X) \leq k+2$, and There is a an $X
-    \in \ccT_i$ such that $s(X)=k+2$.
-  \end{enumerate}
-\end{lemma}
-\begin{lemma}
-  If there are no \marginal type-2 edges in $\cO_i$, then there exists
-  at least $p-i$ type-2 hyperedges $X \in \cO_i$ such that $|X| \geq
-  k+2$.
-  % is a exist type-2 hyperedges of length at least $k+2$, and there
-  % must be at least $r-i$ such type-2 hyperedge, basically at least
-  % one per ray.  Each hyperedge contained in such an hyperedges of
-  % length at least $k+2$, is transitively related under the overlap
-  % relation to one that gets assigned a path that contains
-  % $r$. Basically, the proof will argue that otherwise, there will be
-  % a \marginal type-2 hyperedge, or that it is infeasible.  Also, it
-  % is here we assume that we assume that the size of the two
-  % universes is the same.
-\end{lemma}
-\begin{lemma}
-  At the end of Step-1 in the $i$-th iteration, if one hyperedge $X$
-  of type-2 is such that $\cl(X) \subseteq R_i$, then all other
-  hyperedges in $\cO_i$ are connected to $X$ in the overlap component.
-\end{lemma}
-\begin{lemma}
-  At the end of Step-2, If control has exit at any time, there is no
-  ICPPL. If control has not exit, then $R_i$ is saturated.  No
-  hyperedge of $\cO_{i+1}$ will get a path from $R_i$ in the future
-  iterations.  No type-2 hyperedge of $\cO_{i+1}$ will get a path from
-  $R_i$.  Basically $R_i$ is done.
-\end{lemma}
-\begin{lemma}
-  Finally, we need to prove that the assignment is an ICPPL. Secondly,
-  if there is a permutation then maps sets to paths, then it is indeed
-  an ICPPL, and our algorithm will basically find it.  It is a unique
-  assignment upto permutation of the leaves.
-\end{lemma}
-
-\begin{lemma}
-  \label{lem:sup-mar}
-  If $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
-  tree path labeling to tree $T$, then $\cl(X)$ will
-  contain a leaf in $T$.\\
-  On the otherhand, if $\cl(X)$ has a leaf in
-  $T$, $X$ is marginal but may not be super-marginal.
-\end{lemma}
-\begin{proof}
-  Suppose $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
-  path labeling from $T$.  Assume $\cl(X)$ does not have a leaf.  Let
-  $R_i$ be one of the rays (or the only ray) $\cl(X)$ is part of.
-  Since $X$ is in a connected overlap component, there exists $Y_1 \in
-  \cF$ and $X \nsubseteq Y_1$ such that $Y_1 \overlap X$ and $Y_1$ has
-  at least one vertex closer to the leaf in $R_i$ than any vertex in
-  $X$. Similarly with the same argument there exists $Y_2 \in \cF$
-  with same subset and overlap relation with $X$ except it has has at
-  least one vertex farther away from the leaf in $R_i$ than any vertex
-  in $X$. Clearly $Y_1 \cap X$ and $Y_2 \cap X$ cannot be part of same
-  inclusion chain which contradicts that assumption $X$ is
-  super-marginal. Thus the claim is proven.\qed
-\end{proof}
-
-\noindent
-Consider the overlap graph $\bO(\cF)$ of the given hypergraph
-$\cF$. Let $S_{sm} \in \cF$ be such that it is a super-marginal
-hyperedge.  Algorithm~\ref{al:ktree-label} uses $S_{sm}$ along with
-the overlap graph $\bO(\cF)$ to calculate the feasible tree path
-labeling to the $k$-subdivded tree $T$.
-
-\begin{algorithm}[h]
-  \caption{{\tt compute-ksubtree-path-labeling($X, \cF, T$)}}
-  \label{al:ktree-label}
-  \begin{algorithmic}[\lndisplay]
-    \IF{$X = S_{sm}$}
-    \STATE -- TBD --\\
-    \ELSE 
-    \STATE -- TBD --\\
-    \ENDIF
-
- \end{algorithmic}
-\end{algorithm}
-
-\section{ TPL with no restrictions}
-\label{sec:norestraint}
-
-{\tt abstract begin}
-  it is
-  known that if the given tree is a path a feasible assignment can be
-  found in polynomial time, and we observe that it can actually be
-  done in logspace. \tnote{[TRUE?]}
-{\tt abstract end}
-
-In Section \ref{prelims} we present the
-necessary preliminaries, in Section \ref{feasible} we present our
-characterization of feasible tree path assignments, and in Section
-\ref{decompos} we present the characterizing subproblems for finding a
-bijection between $U$ and $V(T)$ such that sets map to tree
-paths. Finally, in Section \ref{complexity} we conclude by showing
-that Tree Path Assignment is GI-Complete, and also observe that
-Consecutive Ones Testing is in Logspace.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC \subsection{Introduction}
-
-\annote{ We refer to this as the Tree Path Assignment
-problem for an input $(\cF,T)$ pair.}{Add this ``pair'' as a
-  definition in ch:prelims}
-
-\remove{It is an interesting fact that for a matrix with the COP, the
-intersection graph of the corresponding set system is an interval
-graph.  A similar connection to a subclass of chordal graphs, and this
-subclass contains interval graphs, exists for the generalization of
-COP.  In this case, the intersection graph of the corresponding set
-system must be a path graph. Chordal graphs are of great
-significance, extensively studied, and have several applications.  One
-of the well known and interesting properties of a chordal graphs is
-its connection with intersection graphs cite{mcg04}. For every chordal
-graph, there exists a tree and a family of subtrees of this tree such
-that the intersection graph of this family is isomorphic to the
-chordal graph cite{plr70,gav78,bp93}.  Certain format of these trees
-are called clique trees cite{apy92} of the graph which is a compact
-representation of the chordal graph. There has also been work done on
-the generalization of clique trees to clique hypergraphs cite{km02}.
-If the chordal graph can be represented as the intersection graph of
-paths in a tree, then the graph is called path graph cite{mcg04} or
-path chordal graph.  Therefore, it is clear that if there is a
-bijection from $U$ to $V(T)$ such that the sets map to paths, then the
-intersection graph of the set system is indeed a path chordal graph.
-This is, however, only a necessary condition and can be checked
-efficiently, as path graph recognition is polynomial
-solvable cite{gav78,aas93}.  Indeed, it is possible to construct a set
-system and tree, such that the intersection graph is a path chordal
-graph, but there is no bijection between $U$ and $V(T)$ such that the
-sets map to paths.  This connection indeed suggests that our problem
-is indeed as hard as path graph isomorphism.  Further path graph
-isomorphism is known be isomorphism-complete, see for example
-cite{kklv10}. }
-In the second part of this paper, we decompose our
-search for a bijection between $U$ and $V(T)$ into subproblems.  Each
-subproblem is on a set system in which for each set, there is another
-set in the set system with which the intersection is {\em strict}-
-there is a non-empty intersection, but neither is contained in the
-other.  This is in the spirit of results in \cite{wlh02,nsnrs09} where
-to test for COP in a given matrix, the COP problem is solved on an
-equivalent set of prime matrices.  \annote{Our decomposition localizes the
-challenge of path graph isomorphism to two problems.}{HOW?}
-
-Finally, we show that Tree Path Assignment is isomorphism-complete.
-We also point out Consecutive Ones Testing is in Logspace from two
-different results in the literature \cite{kklv10, mcc04}. To the best
-of our knowledge this observation has not been made earlier.
-
-% \noindent
-% It has been long known that interval graph recognition is in
-% logspace\cite{rei84}. Recently interval graph isomorphism was also
-% shown to be logspace decidable using a logspace canonization
-% algorithm by \cite{kklv10}.  This result is built on top of logspace
-% results of undirected graph connectivity \cite{rei08}, logspace
-% tractability using a certain logical formalism called FP+C and
-% modular decomposition of interval graphs\cite{lau10} etc.  Interval
-% graphs are closely connected to binary matrices with COP. The
-% maximal clique vertex incidence matrix (matrix with rows
-% representing maximal cliques and columns representing vertices of a
-% graph) has COP on columns iff the graph is an interval
-% graph\cite{fg65}. This follows from the interval graph
-% characterization by \cite{gh64}. Due to this close relation it is
-% natural to see if consecutive ones property can be
-% tested in logspace. \\
-% \noindent
-% We also explore some extensions of the interval assignment problem
-% in \cite{nsnrs09}, namely caterpillar path assignment problem.
-
-% We present a logspace algorithm here that uses the ICPIA
-% characterization of binary matrices with COP (set system associated
-% with such a matrix)\cite{nsnrs09}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC\subsection{Preliminaries} 
-\label{prelims} 
-In this paper, the collection
-$\F = \{S_i \mid S_i \subseteq U, S_i \ne \O, i \in I\}$ is a {\em set
-  system} of universe, $U = \{1, \ldots, n\}$. Moreoever, a set system
-is assumed to ``cover'' the universe,
-i.e. $ \bigcup_{i \in I}S_i = U$. \\
-
-\noindent
-The graph $T=(V,E)$ represents a given tree with $|V| = n$. Further,
-for simplicity, $V$ is defined as $\{1,\ldots,n\}$. All paths
-referred to in this paper are paths of $T$ unless explicitly specified. \\
-
-\noindent
-A {\em path assignment} $\A$ to $\F$ is defined as a set assignment
-where second universe is the vertex set $V$ of a given tree $T$ and
-every second subset in the ordered pairs is a path in this
-tree. Formally, the definition is as follows.
-\begin{align*}
-  \A = \{ (S_i,P_i) \mid S_i \in \F, P_i \subseteq V \text{
-    s.t. }T[P_i] \text{ is a path, } i \in I \}
-\end{align*}
-In other words, $P_i$ is the path on the tree $T$ assigned to $S_i$ in
-$\A$. As mentioned before for set systems, the paths cover the whole
-tree, i.e. $\bigcup_{i \in I}P_i = V$ \\
-
-\noindent
-Generalizing the definition of {\em feasibility} in \cite{nsnrs09} to
-a set assignment, a path assignment $\A$ is defined to be {\em
-  feasible} if there exists a bijection defined as follows.
-\begin{align}
-  \sigma: U \rightarrow V(T), \text{ such that }\sigma(S_i) = P_i
-  \text{ for all } i \in I, \sigma \text{ is a bijection}
-  \label{eq:stf}
-\end{align}
-% The assignment $\A$ is then called {\em Set Translation Feasible}.
-
-\noindent
-Let $X$ be a partially ordered set with $\preccurlyeq$ being the
-partial order on $X$.  $mub(X)$ represents an element in $X$ which is
-a maximal upper bound on $X$.  $X_m \in X$ is a maximal upper bound of
-$X$ if $\nexists X_q \in X$ such that $X_m
-\preccurlyeq X_q$. \\
-
-\noindent
-The set $I$ represents the index set $[m]$. If index $i$ is used
-without further qualification, it is meant to be $i \in I$. Any
-function, if not defined on a domain of sets, when applied on a set is
-understood as the function applied to each of its elements. i.e. for
-any function $f$ defined with domain $U$, the abuse of notation is as
-follows; $f(S)$ is used instead of $\hat f(S)$ where
-$\hat f(S) = \{y \mid y = f(x), x \in S\}$. \\
-
-\noindent
-When refering to a tree as $T$ it could be a reference to the tree
-itself, or the vertices of the tree. This will be clear from the
-context.\\
-
-\noindent
-Finally, an in-tree is a directed rooted tree in which all edges are
-directed toward to the root.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC\subsection{Characerization of Feasible Tree Path
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC  Assignment\tnote{Merge with the original charac section}} 
-\label{feasible} 
-Consider a path assignment $\cA =
-\{(S_i, P_i) \mid S_i \in \cF, P_i \text{ is a path from $T$}, i \in
-[m]\}$ to a set system $\cF = \{S_i \mid S_i \subseteq U, i \in
-[m]\}$, were $T$ is a given tree, $U$ is the set system's universe and
-$m$ is the number of sets in $\cF$. We call $\cA$ an {\em Intersection
-  Cardinality Preserving Path Assignment (ICPPA)} if it has the
-following properties.
-
-\begin{enumerate}
-\item [i.]  $|S_i| = |P_i|$ for all $i \in [m]$
-\item [ii.] $|S_i \cap S_j| = |P_i \cap P_j|$ for all $i,j \in [m]$
-  % \item [iii.] $|\bigcup_{i \in J} S_i| = |\bigcup_{i \in J} P_i|$
-  %   for all $J \subseteq I$\footnote{TBD: this is needed now to fix
-  %     some proofs}
-\item [iii.] $|S_i \cap S_j \cap S_k| = |P_i \cap P_j \cap P_k|$ for
-  all $i,j,k \in [m]$
-\end{enumerate}
-
-\begin{lemma}
-  \label{lem:setminuscard}
-  If $\cA$ is an ICPPA, and $(S_1, P_1),(S_2, P_2),(S_3, P_3) \in
-  \cA$, then $|S_1 \cap (S_2 \setminus S_3)| = |P_1 \cap (P_2
-  \setminus P_3)|$.
-\end{lemma}
-\begin{proof}
-  $|S_1 \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| =
-  |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to conditions (ii)
-  and (iii) of ICPPA, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
-  \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
-  = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. \qed
-\end{proof}
-
-\begin{lemma}
-  \label{lem:fourpaths} Consider four paths in a tree $Q_1, Q_2, Q_3,
-  Q_4$ such that they have nonempty pairwise intersection and $Q_1,
-  Q_2$ share a leaf. Then there exists distinct $i, j, k \in
-  \{1,2,3,4\}$ such that, $Q_1 \cap Q_2 \cap Q_3 \cap Q_4 = Q_i \cap
-  Q_j \cap Q_k$.
-\end{lemma}
-\begin{proof}
-  {\em Case 1:} w.l.o.g, consider $Q_3 \cap Q_4$ and let us call it
-  $Q$. This is clearly a path (intersection of two paths is a path).
-  % Since $Q_1, Q_2$ share a leaf, the following are paths $Q_1
-  % \setminus Q_2$, $Q_2 \setminus Q_1$, $Q_1 \cap Q_2$ and they are
-  % mutually disjoint.
-  Suppose $Q$ does not intersect with $Q_1 \setminus Q_2$, i.e. $Q
-  \cap (Q_1 \setminus Q_2) = \O$. Then $Q \cap Q_1 \cap Q_2 = Q \cap
-  Q_2$. Similarly, if $Q \cap (Q_2 \setminus Q_1) = \O$, $Q \cap Q_1
-  \cap Q_2 = Q \cap Q_1$. Thus it is clear that if the intersection of
-  any two paths does not intersect with any of the set differences of
-  the remaining two paths, the claim in the lemma is true.
-  % Note that $Q_1 \setminus Q_2$ and $Q_2
-  % \setminus Q_1$ are paths because $Q_1, Q_2$ share a leaf.\\
-  {\em Case 2:} Let us consider the compliment of the previous
-  case. i.e. the intersection of any two paths intersects with both
-  the set differences of the other two. First let us consider $Q \cap
-  (Q_1 \setminus Q_2) \ne \O$ and $Q \cap (Q_1 \setminus Q_2) \ne \O$,
-  where $Q = Q_3 \cap Q_4$. Since $Q_1$ and $Q_2$ share a leaf, there
-  is exactly one vertex at which they branch off from the path $Q_1
-  \cap Q_2$ into two paths $Q_1 \setminus Q_2$ and $Q_2 \setminus
-  Q_1$. Let this vertex be $v$. It is clear that if path $Q_3 \cap
-  Q_4$, must intersect with paths $Q_1 \setminus Q_2$ and $Q_2
-  \setminus Q_1$, it must contain $v$ since these are paths from a
-  tree. Moreover, $Q_3 \cap Q_4$ intersects with $Q_1 \cap Q_2$ at
-  exactly $v$ and only at $v$ which means that $Q_1 \cap Q_2$ does not
-  intersect with $Q_3 \setminus Q_4$ or $Q_4 \setminus Q_3$ which
-  contradicts initial condition of this case. Thus this
-  case cannot occur and case 1 is the only possible scenario. \\
-  Thus lemma is proven \qed
-\end{proof}
-
-\noindent
-In the remaining part of this section we show that a path assignment
-is feasible if and only if it is an ICPPA.  One direction of this
-claim is clear: that is a path assignment is feasible, then all
-intersection cardinalities are preserved, that is the path assignment
-is an ICPPA.  The reason is that a feasible path assignment has an
-associated bijection between $U$ and $V(T)$ such that the sets map to
-paths.  The rest of the section is devoted to constructively proving
-that it is sufficient, for the existence of a bijection, for a path
-assignment to be an ICPPA.  At a top-level, the constructive
-approaches refine the path assignment iteratively, such that at the
-end of each iteration we have a path assignment, and finally we have a
-family of bijections.  First we present and then prove the correctness
-of Algorithm \ref{perms}.  This algorithm refines the path assignment
-by considering pairs of paths that share a leaf.
-
-\begin{algorithm}[h]
-  \caption{Permutations from an ICPPA $\{(S_i,P_i) | i \in I\}$}
-  \label{perms}
-  \begin{algorithmic}
-    \STATE Let $\Pi_0=\{(S_i,P_i)| i \in I\}$\\
-    \STATE $j = 1$;
-    \label{shareleaf} \WHILE {There is $(P_1,Q_1), (P_2,Q_2) \in
-      \Pi_{j-1}$ with $Q_1$ and $Q_2$ having a common leaf} \STATE
-    $\Pi_j= \Pi_{j-1} \setminus \{(P_1,Q_1),(P_2,Q_2)\}$;
-    \label{setbreak}\STATE $\Pi_j = \Pi_j \cup \{(P_1 \cap P_2,Q_1
-    \cap Q_2), (P_1 \setminus P_2,Q_1 \setminus Q_2), (P_2 \setminus
-    P_1, Q_2 \setminus Q_1)\}$; \STATE $j = j+1$; \ENDWHILE \STATE
-    $\Pi = \Pi_j$; \STATE Return $\Pi$;
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma}
-  \label{lem:invar1}
-  In Algorithm \ref{perms}, at the end of $j$th iteration, $j \ge 0$,
-  of the while loop of Algorithm \ref{perms}, the following invariants
-  are maintained.
-  \begin{itemize}
-  \item {\em Invariant I:} $Q$ is a path in $T$ for each $(P,Q) \in
-    \Pi_j$
-  \item {\em Invariant II:} $|P|=|Q|$ for each $(P,Q) \in \Pi_j$
-  \item {\em Invariant III:} For any two $(P,Q), (P',Q') \in \Pi_j$,
-    $|P' \cap P''|=|Q' \cap Q''|$.
-  \item {\em Invariant IV:} For any three, $(P',Q'), (P'',Q''), (P, Q)
-    \in \Pi_j$, $|P' \cap P'' \cap P|=|Q' \cap Q'' \cap Q|$.
-  \end{itemize}
-\end{lemma}
-\begin{proof}
-  Proof is by induction on the number of iterations, $j$. In the rest
-  of the proof, the term ``new sets'' will refer to the new sets added
-  in $j$th iteration as defined in line \ref{setbreak} of Algorithm
-  \ref{perms}, i.e. the following three assignment pairs for some
-  $(P_1,Q_1), (P_2,Q_2) \in \Pi_{j-1}$ where $Q_1$ and $Q_2$ intersect
-  and share a leaf: $(P_1 \cap P_2, Q_1 \cap Q_2)$, or $(P_1 \setminus
-  P_2, Q_1 \setminus Q_2)$, or $(P_2 \setminus P_1, Q_2
-  \setminus Q_1)$.\\
-  \noindent
-  The base case, $\Pi_0 = \{(S_i,P_i) \mid i \in [m]\}$, is trivially
-  true since it is the input which is an ICPPA.  Assume the lemma is
-  true till the $j-1$ iteration. Consider $j$th
-  iteration:\\
-  \noindent
-  If $(P,Q)$, $(P',Q')$ and $(P'',Q'')$ are in $\Pi_{j}$ and
-  $\Pi_{j-1}$, all the invariants are
-  clearly true because they are from $j-1$ iteration.\\
-  If $(P,Q)$ is in $\Pi_{j}$ and not in $\Pi_{j-1}$, then it must be
-  one of the new sets added in $\Pi_j$. Since $(P_1,Q_1)$ and
-  $(P_2,Q_2)$ are from $\Pi_{j-1}$ and $Q_1,Q_2$ intersect and have a
-  common leaf, it can be verified that the
-  new sets are also paths. \\
-  By hypothesis for invariant III, invariant II also holds for $(P,Q)$
-  no matter which new set in $\Pi_j$ it
-  is.\\
-  To prove invariant III, if $(P,Q)$ and $(P',Q')$ are not in
-  $\Pi_{j-1}$, then they are both new sets and invariant III holds
-  trivially (new sets are disjoint). Next consider $(P,Q), (P',Q') \in
-  \Pi_j$ with only one of them, say $(P',Q')$, in $\Pi_{j-1}$. Then
-  $(P,Q)$ is one of the new sets added in line \ref{setbreak}. It is
-  easy to see that if $(P,Q)$ is $(P_1 \cap P_2, Q_1 \cap Q_2)$, then
-  due to invariant IV in hypothesis, invariant III becomes true in
-  this iteration. Similarly, using lemma \ref{lem:setminuscard}
-  invariant III is proven if $(P, Q)$ is $(P_1 \setminus P_2, Q_1
-  \setminus Q_2)$, or $(P_2 \setminus P_1, Q_2
-  \setminus Q_1)$.\\
-  To prove invariant IV, consider three assignments
-  $(P,Q),(P',Q'),(P'',Q'')$. If at least two of these pairs are in not
-  $\Pi_{j-1}$, then they are any two of the new sets. Note that these
-  new sets are disjoint and hence if $(P',Q'), (P'',Q'')$ are any of
-  these sets, $|P \cap P' \cap P''|=|Q \cap Q' \cap Q''|=0$ and
-  invariant IV is true. Now we consider the case if at most one of
-  $(P,Q),(P',Q'),(P'',Q'')$ is not in $\Pi_{j-1}$. If none of them are
-  not in $\Pi_{j-1}$ (i.e. all of them are in $\Pi_{j-1}$), invariant
-  IV is clearly true. Consider the case where exactly one of them is
-  not in $\Pi_{j-1}$. w.l.o.g let that be $(P,Q)$ and it could be any
-  one of the new sets. If $(P,Q)$ is $(P_1 \cap P_2, Q_1 \cap Q_2)$,
-  from lemma \ref{lem:fourpaths} and invariant III hypothesis,
-  invariant IV is proven. Similarly if $(P,Q)$ is any of the other new
-  sets, invariant IV is proven by also using lemma
-  \ref{lem:setminuscard}. \qed
-
-\end{proof}
-
-\noindent
-It can be observed that the output of algorithm \ref{perms} is such
-that every leaf is incident on at most a single path in the new set of
-assignments. This is due to the loop condition at line
-\ref{shareleaf}. Let $v_1$ be the leaf incident on path $P_i$. Assign
-to it any one element from $S_i \setminus \bigcup_{i \ne j}
-S_j$. Remove $(S_i, P_i)$ from assignments and add $(\{x_1\},
-\{v_1\}), (S_i \setminus \{x_1\}, P_i \setminus \{v_1\})$. Now all
-assignments except single leaf assignments are paths from the subtree
-$T_0 = T \setminus \{v \mid v \text{ is a leaf in } T\}$.
-
-% exclusively on the spine of the caterpillar. Consider the spine to
-% be an interval and run ICPIA algorithm on it to get the required
-% permutation. Algorithm \ref{leafasgn} makes this clear.
-
-\begin{algorithm}[h]
-  \caption{Leaf assignments from an ICPPA $\{(S_i,P_i) | i \in I\}$}
-  \label{leafasgn}
-  \begin{algorithmic}
-    \STATE Let $\Pi_0=\{(S_i,P_i)| i \in [m]\}$. Paths are such that
-    no
-    two paths $P_i, P_j, i \ne j$ share a leaf.\\
-    \STATE $j = 1$\\
-    \WHILE {there is a leaf $v$ and a unique $(S_{i_1}, P_{i_1})$ such
-      that $v \in P_{i_1}$}
-    \STATE $\Pi_j=   \Pi_{j-1} \setminus \{(S_{i_1}, P_{i_1})\}$\\
-    \STATE $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in I}S_i$
-    \IF{$X$ is empty} \STATE exit \ENDIF; \STATE Let $x = $ arbitrary
-    element from $X$ \vspace{2mm}
-
-    \STATE $\Pi_j = \Pi_j \cup \{(S_{i_1} \setminus \{x\}),(P_{i_1}
-    \setminus
-    \{v\}), (\{x\},\{v\})\}$\\
-    \STATE $j = j+1$\\
-    \ENDWHILE
-    \STATE $\Pi = \Pi_j$\\
-    \STATE Return $\Pi$\\
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma}
-  \label{lem:invar3}
-  In Algorithm \ref{leafasgn},for all $j \geq 0$, at the end of the
-  $j$th iteration the four invariants given in lemma \ref{lem:invar1}
-  are valid.
-  % Moreover, $X$ as defined in the algorithm is non-empty if this is
-  % an ICPPA.
-\end{lemma}
-\begin{proof}
-  First we see that $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in
-    I}S_i$ is non empty in every iteration for an ICPPA. Suppose $X$
-  is empty. We know that $v \in P_{i_1} \setminus \bigcup_{i \ne i_1,
-    i \in I}P_i$ since $v$ is in the unique path $P_{i_1}$. Since this
-  is an ICPPA $|S_{i_1}| = |P_{i_1}|$. For any $x \in S_{i_1}$ it is
-  contained in at least two sets due to our assumption. Let $S_{i_2}$
-  be a second set that contains $x$. We know $v \notin
-  P_{i_2}$. Therefore there cannot exist a permutation that maps
-  elements of $S_{i_2}$ to $P_{i_2}$. This contradicts our assumption
-  that this is an ICPPA. Thus $X$ cannot be empty.
-
-  We use mathematical induction on the number of iterations for this
-  proof. The term ``new sets'' will refer to the sets added in $\Pi_j$
-  in the $j$th iteration, i.e. $(P' \setminus \{x\},Q' \setminus
-  \{v\})$ and $(\{x\},\{v\})$ for some $(P',Q')$ in $\Pi_{j-1}$ such
-  that $v$ is a leaf and $Q'$ is the unique path
-  incident on it.\\
-  For $\Pi_0$ all invariants hold because it is output from algorithm
-  \ref{perms} which is an ICPPA. Hence base case is proved.  Assume
-  the lemma holds for $\Pi_{j-1}$. Consider $\Pi_j$ and any $(P,Q) \in
-  \Pi_j$. If $(P,Q) $ is in $ \Pi_j$ and $\Pi_{j-1}$ invariants I and
-  II are true because of induction assumption. If it is only in
-  $\Pi_j$, then it is $\{(P' \setminus \{x\}),(Q' \setminus \{v\})$ or
-  $(\{x\},\{v\})$ for some $(P',Q')$ in $\Pi_{j-1}$. By definition,
-  $x$ is an element in $P'$ (as defined in the algorithm) and $v$ is a
-  leaf in $Q'$. If $(P,Q)$ is $\{(P' \setminus \{x\}),(Q' \setminus
-  \{v\})$, $Q$ is a path since only a leaf is removed from path
-  $Q'$. We know $|P'| = |Q'|$, therefore $|P' \setminus \{x\}| = |Q'
-  \setminus \{v\}|$. Hence in this case invariants I and II are
-  obvious. It is easy to see these invariants hold if $(P,Q)$ is
-  $(\{x\},\{v\})$.
-
-  For invariant III consider $(P_1,Q_1),(P_2,Q_2)$ in $\Pi_j$. If both
-  of them are also in $\Pi_{j-1}$, claim is proved. If one of them is
-  not in $\Pi_{j-1}$ then it has to be $\{(P' \setminus \{x\}),(Q'
-  \setminus \{v\})$ or $(\{x\},\{v\})$ for some $(P',Q')$ in
-  $\Pi_{j-1}$. Since by definition, $Q'$ is the only path with $v$ and
-  $P'$ the only set with $x$ in the previous iteration, $|P_1 \cap (P'
-  \setminus \{x\})| = |P_1 \cap P'|$ and $|Q_1 \cap (Q' \setminus
-  \{v\})| = |Q_1 \cap Q'|$ and $|P_1 \cap \{x\}| = 0, Q_1 \cap \{v\} =
-  0$. Thus invariant III is also proven.
-
-  \noindent
-  To prove invariant IV, consider $(P_1,Q_1),(P_2,Q_2), (P_3,Q_3)$ in
-  $\Pi_j$. If exactly one of them, say $P_3 \notin \Pi_{j-1}$, it is
-  one of the new sets. By the same argument used to prove invariant
-  III, $|P_1 \cap P_2 \cap (P' \setminus \{x\})| = |P_1 \cap P_2 \cap
-  P'|$ and $|Q_1 \cap Q_2 \cap (Q' \setminus \{x\})| = |Q_1 \cap Q_2
-  \cap Q'|$. Since $P_1, P_2, P'$ are all in $\Pi_{j-1}$, by induction
-  hypothesis $|P_1 \cap P_2 \cap P'| = |Q_1 \cap Q_2 \cap Q'|$. Also
-  $|P_1 \cap P_2 \cap \{x\}| = 0, Q_1 \cap Q_2 \cap \{v\} = 0$.  If
-  two or more of them are not in $\Pi_{j-1}$, then it can be verified
-  that $|P_1 \cap P_2 \cap P_3| = |Q_1 \cap Q_2 \cap Q_3|$ since the
-  new sets in $\Pi_j$ are either disjoint or as follows: assuming
-  $P_1, P_2 \notin \Pi_{j-1}$ and new sets are derived from $(P', Q'),
-  (P'', Q'') \in \Pi_{j-1}$ with $x_1, x_2$ exclusively in $P_1, P_2$,
-  $(\{x_1\},\{v_1\}), (\{x_2\},\{v_2\}) \in \Pi_j $ thus $v_1, v_2$
-  are exclusively in $Q_1, Q_2$ resp. it follows that $|P_1 \cap P_2
-  \cap P_3| = |(P' \setminus \{x_1\}) \cap (P'' \setminus \{x_2\})
-  \cap P_3| = |P' \cap P'' \cap P_3| = |Q' \cap Q'' \cap Q_3| = |(Q'
-  \setminus \{v_1\} \cap Q'' \setminus \{v_2\} \cap Q_3| = |Q_1 \cap
-  Q_2 \cap Q_3|$. Thus invariant IV is also proven.  \qed
-\end{proof}
-
-\noindent
-Using algorithms \ref{perms} and \ref{leafasgn} we prove the following
-theorem.
-
-\begin{theorem}
-  \label{th:perm}
-  If $\cA$ is an ICPPA, then there exists a bijection $\sigma : U
-  \rightarrow V(T)$ such that $\sigma(S_i) = P_i$ for all $i \in I$
-\end{theorem}
-\begin{proof}
-  This is a contructive proof. First, the given ICPPA $\cA$ and tree
-  $T$ are given as input to Algorithm \ref{perms}. This yields a
-  ``filtered'' ICPPA as the output which is input to Algorithm
-  \ref{leafasgn}.  It can be observed that the output of Algorithm
-  \ref{leafasgn} is a set of interval assignments to sets and
-  one-to-one assignment of elements of $U$ to each leaf of $T$. To be
-  precise, it would be of the form $\cB_0 = \cA_0 \cup \cL_0$. The
-  leaf assignments are defined in $\cL_0 = \{ (x_i,v_i) \mid x_i \in
-  U, v_i \in T, x_i \ne x_j, v_i \ne v_j, i \ne j, i,j \in [k] \}$
-  where $k$ is the number of leaves in $T$. The path assignments are
-  defined in $\cA_0 \subseteq \{(S_i',P_i') \mid S_i' \subseteq U_0,
-  P_i' \text{ is a path from } T_0\}$ where $T_0$ is the tree obtained
-  by removing all the leaves in $T$ and $U_0 = U \setminus \{ x \mid x
-  \text{ is assigned to a leaf in }\cL_0 \}$. Now we have a subproblem
-  of finding the permutation for the path assignment $\cA_0$ which has
-  paths from tree $T_0$ and sets from universe $U_0$. Now we repeat
-  the procedure and the path assignment $\cA_0$ and tree $T_0$ is
-  given as input to Algorithm \ref{perms}. The output of this
-  algorithm is given to Algorithm \ref{leafasgn} to get a new union of
-  path and leaf assignments $\cB_1 = \cA_1 \cup \cL_1$ defined similar
-  to $\cB_0, \cL_0, \cA_0$. In general, the two algorithms are run on
-  path assignment $\cA_{i-1}$ with paths from tree $T_{i-1}$ to get a
-  new subproblem with path assignment $\cA_i$ and tree $T_{i}$. $T_i$
-  is the subtree of $T_{i-1}$ obtained by removing all its
-  leaves. More importantly, it gives leaf assignments $\cL_{i}$ to the
-  leaves in tree $T_{i-1}$. This is continued until we get a
-  subproblem with path assignment $\cA_{d-1}$ and tree $T_{d-1}$ for
-  some $d \le n$ which is just a path. From the last lemma we know
-  that $\cA_{d-1}$ is an ICPPA. Another observation is that an ICPPA
-  with all its tree paths being intervals (subpaths from a path) is
-  nothing but an ICPIA\cite{nsnrs09}.  Let $\cA_{d-1}$ be equal to
-  $\{(S_i'',P_i'') \mid S_i'' \subseteq U_{d-1}, P_i'' \text{ is a
-    path from } T_{d-1} \}$. It is true that the paths $P_i''$s may
-  not be precisely an interval in the sense of consecutive integers
-  because they are some nodes from a tree. However, it is easy to see
-  that the nodes of $T_{d-1}$ can be ordered from left to right and
-  ranked to get intervals $I_i$ for every path $P_i''$ as
-  follows. $I_i = \{[l,r] \mid l = \text{ the lowest rank of the nodes
-    in }P_i'', r = l+|P_i''|-1 \}$. Let asssignment $\cA_d$ be with
-  the renamed paths. $\cA_d = \{ (S_i'', I_i) \mid (S_i'', P_i'') \in
-  \cA_{d-1} \}$. What has been effectively done is renaming the nodes
-  in $T_{d-1}$ to get a tree $T_d$.  The ICPIA $\cA_d$ is now in the
-  format that the ICPIA algorithm requires which gives us the
-  permutation $\sigma' : U_{d-1} \rightarrow T_{d-1}$
-
-\noindent
-$\sigma'$ along with all the leaf assignments $\cL_i$ gives us the
-permutation for the original path assignment $\cA$.  More precisely,
-the permutation for tree path assignment $\cA$ is defined as
-follows. $\sigma: U \rightarrow T$ such that the following is
-maintained.
-\begin{align*}
-  \sigma(x) &= \sigma'(x),   \text{ if } x \in U_{d-1} \\
-  &= \cL_i(x), \text{ where $x$ is assigned to a leaf in a subproblem
-    $\cA_{i-1}, T_{i-1}$}
-\end{align*}
-
-\noindent
-To summarize, run algorithm \ref{perms} and \ref{leafasgn} on
-$T$. After the leaves have been assigned to specific elements from
-$U$, remove all leaves from $T$ to get new tree $T_0$. The leaf
-assignments are in $\cL_0$. Since only leaves were removed $T_0$ is
-indeed a tree. Repeat the algorithms on $T_0$ to get leaf assignments
-$\cL_{1}$. Remove the leaves in $T_0$ to get $T_1$ and so on until the
-pruned tree $T_d$ is a single path. Now run ICPIA algorithm on $T_d$
-to get permutation $\sigma'$. The relation $\cL_0 \cup \cL_1 \cup
-.. \cup \cL_{d} \cup \sigma'$ gives the bijection required in the
-original problem.\qed
-\end{proof}
-
-\subsection{Finding an assignment of tree paths to a set
-  system} \label{decompos} In the previous section we have shown that
-the problem of finding a Tree Path Asssignment to an input $(\cF,T)$
-is equivalent to finding an ICPPA to $\cF$ in tree $T$.  In this
-section we characterize those set systems that have an ICPPA in a
-given tree.  As a consequence of this characterization we identify two
-sub-problems that must be solved to obtain an ICPPA.  We do not solve
-the problem and in the next section show that finding an ICPPA in a
-given tree is GI-Complete.
-
-\noindent
-A set system can be concisely represented by a binary matrix where the
-row indices denote the universe of the set system and the column
-indices denote each of the sets. Let the binary matrix be $M$ with
-order $n \times m$, the set system be $\cF = \{S_i \mid i \in [m]\}$,
-universe of set system $U = \{x_1, \dots ,x_n\}$. If $M$ represents
-$\cF$, $|U| = n, |\cF| = m$. Thus $(i,j)$th element of $M$, $M_{ij} =
-1$ iff $x_i \in S_j$. If $\cF$ has a feasible tree path assignment
-(ICPPA) $\cA = \{(S_i,P_i) \mid i \in [m]\}$, then we say its
-corresponding matrix $M$ has an ICPPA. Conversly we say that a matrix
-$M$ has an ICPPA if there exists an ICPPA $\cA$ as defined
-above.\\
-\noindent
-We now define the strict intersection graph or overlap graph of
-$\cF$. This graph occurs at many places in the literature, see for
-example \cite{kklv10, wlh02, nsnrs09}.  The vertices of the graph
-correspond to the sets in $\cF$.  An edge is present between vertices
-of two sets iff the corresponding sets have a nonempty intersection
-and none is contained in the other. Formally, intersection graph is
-$G_f = (V_f, E_f)$ such that $V_f = \{v_i \mid S_i \in \cF\}$ and $E_f
-= \{(v_i, v_j) \mid S_i \cap S_j \ne \O \text{ and }S_i \nsubseteq
-S_j, S_j \nsubseteq S_i \}$.  We use this graph to decompose $M$ as
-described in \cite{wlh02,nsnrs09}.  A prime sub-matrix of $M$ is
-defined as the matrix formed by a set of columns of $M$ which
-correspond to a connected component of the graph $G_f$.  Let us denote
-the prime sub-matrices by $M_1,\ldots,M_p$ each corresponding to one
-of the $p$ components of $G_f$. Clearly, two distinct matrices have a
-distinct set of columns.  Let $col(M_i)$ be the set of columns in the
-sub-matrix $M_i$.  The support of a prime sub-matrix $M_i$ is defined
-as $supp(M_i) = \displaystyle \bigcup_{j \in col(M_i)}S_j$.  Note that
-for each $i$, $supp(M_i) \subseteq U$.  For a set of prime
-sub-matrices $X$ we define
-$supp(X) = \displaystyle \bigcup_{M \in X} supp(M)$. \\
-
-\noindent
-Consider the relation $\preccurlyeq$ on the prime sub-matrices $M_1,
-\ldots, M_p$ defined as follows:
-\begin{equation}
-  \nonumber \{(M_i,M_j) | \mbox{ A set } S \in
-  M_i \mbox{ is contained in a set } S' \in M_j\} \cup \{(M_i,M_i) | 1
-  \leq i \leq p\} 
-\end{equation}
-
-\noindent
-This relation is the same as that defined in \cite{nsnrs09}. The prime
-submatrices and the above relation can be defined for any set
-system. We will use this structure of prime submatrices to present our
-results on an an ICPPA for a set system $\cF$. Recall the following
-lemmas, and theorem that $\preccurlyeq$ is a partial order, from
-\cite{nsnrs09}.
-
-\begin{lemma} \label{lem:containment} Let $(M_i,M_j) \in
-  \preccurlyeq$.  Then there is a set $S' \in M_j$ such that for each
-  $S \in M_i$, $S \subseteq S'$.
-\end{lemma}
-\begin{lemma}
-  For each pair of prime sub-matrices, either $(M_i,M_j) \not\in
-  \preccurlyeq$ or $(M_j,M_i) \not\in \preccurlyeq$.
-  % If $(M_i,M_j) \in \preccurlyeq$ and $(M_j,M_i) \in \preccurlyeq$,
-  % then $i = j$ and $|M_i| = 1$.
-\end{lemma}
-\begin{lemma}
-  If $(M_i,M_j) \in \preccurlyeq $ and $(M_j,M_k) \in \preccurlyeq$,
-  then $(M_i,M_k) \in \preccurlyeq$.
-\end{lemma}
-\begin{lemma}
-  If $(M_i,M_j) \in \preccurlyeq$ and $(M_i,M_k) \in \preccurlyeq$,
-  then either $(M_j,M_k) \in \preccurlyeq$ or $(M_k,M_j) \in
-  \preccurlyeq$.
-\end{lemma}
-\begin{theorem} \label{thm:partitionold} $\preccurlyeq$ is a partial
-  order on the set of prime sub-matrices of $M$.  Further, it uniquely
-  partitions the prime sub-matrices of $M$ such that on each set in
-  the partition $\preccurlyeq$ induces a total order.
-\end{theorem}
-For the purposes of this paper, we refine the total order mentioned in
-Theorem \ref{thm:partitionold}. We do this by identifying an in-tree
-rooted at each maximal upper bound under $\preccurlyeq$.  Each of
-these in-trees will be on disjoint vertex sets, which in this case
-would be disjoint sets of prime-submatrices.  The in-trees are
-specified by selecting the appropriate edges from the Hasse diagram
-associated with $\preccurlyeq$.  Let $\cI$ be the following set:
-\begin{align*}
-  \cI = \{ (M_i,M_j) \in \preccurlyeq \mid \nexists M_k s.t. M_i
-  \preccurlyeq M_k, M_k \preccurlyeq M_j \} \cup \{ (M_i,M_i), i \in
-  [p] \}
-\end{align*}
-
-\begin{theorem} \label{thm:partition} Consider the directed graph $X$
-  whose vertices correspond to the prime sub-matrices, and the edges
-  are given by $\cI$.  Then, $X$ is a vertex disjoint collection of
-  in-trees and the root of each in-tree is a maximal upper bound in
-  $\preccurlyeq$.
-\end{theorem}
-\begin{proof}
-  To observe that $X$ is a collection of in-trees, we observe that for
-  vertices corresponding to maximal upper bounds, no out-going edge is
-  present in $I$.  Secondly, for each other element, exactly one
-  out-going edge is chosen, and for the minimal lower bound, there is
-  no in-coming edge.  Consequently, $X$ is acyclic, and since each
-  vertex has at most one edge leaving it, it follows that $X$ is a
-  collection of in-trees, and for each in-tree, the root is a maximal
-  upper bound in $\preccurlyeq$.  Hence the theorem.
-\end{proof}
-Let the partition of $X$ given by Theorem \ref{thm:partition} be
-$\{X_1,\ldots,X_r\}$.  Further, each in-tree itself can be layered
-based on the distance from the root.  The root is considered to be at
-level zero. For $j \geq 0$, Let $X_{i,j}$ denote the set of prime
-matrices in level $j$ of in-tree $X_i$.
-
-\begin{lemma}
-  \label{lem:subicppa}
-  Let $M$ be a matrix and let $X$ be the directed graph whose vertices
-  are in correspondence with the prime submatrices of $M$.  Further
-  let $\{X_1,\ldots,X_r\}$ be the partition of $X$ into in-trees as
-  defined above.  Then, matrix $M$ has an ICPPA in tree $T$ iff $T$
-  can be partitioned into vertex disjoint subtrees $\{T_1, T_2, \dots
-  T_r\}$ such that, for each $1 \leq i \leq r$, the set of prime
-  sub-matrices corresponding to vertices in $X_i$ has an ICPPA in
-  $T_i$.
-\end{lemma}
-\begin{proof}
-  Let us consider the reverse direction first.  Let us assume that $T$
-  can be partitioned into $T_1, \ldots, T_r$ such that for each $1
-  \leq i \leq r $, the set of prime sub-matrices corresponding to
-  vertices in $X_i$ has an ICPPA in $T_i$.  It is clear from the
-  properties of the partial order $\preccurlyeq$ that these ICPPAs
-  naturally yield an ICPPA of $M$ in $T$.  The main property used in
-  this inference is that for each $1 \leq i \neq j \leq r$, $supp(X_i)
-  \cap supp(X_j) = \phi$.
-
-\noindent
-To prove the forward direction, we show that if $M$ has an ICPPA, say
-$\cA$, in $T$, then there exists a partition of $T$ into vertex
-disjoint subtree $T_1, \ldots, T_r$ such that for each $1 \leq i \leq
-r$, the set of prime sub-matrices corresponding to vertices in $X_i$
-has an ICPPA in $T_i$.  For each $1 \leq i \leq r$, we define based on
-$\cA$ a subtree $T_i$ corresponding to $X_i$.  We then argue that the
-trees thus defined are vertex disjoint, and complete the proof.
-Consider $X_i$ and consider the prime sub-matrix in $X_{i,0}$.
-Consider the paths assigned under $\cA$ to the sets in the prime
-sub-matrix in $X_{i,0}$.  Since the component in $G_f$ corresponding
-to this matrix is a connected component, it follows that union of
-paths assigned to this prime-submatrix is a subtree of $T$.  We call
-this sub-tree $T_i$.  All other prime-submatrices in $X_i$ are
-assigned paths in $T_i$ since $\cA$ is an ICPPA, and the support of
-other prime sub-matrices in $X_i$ are contained in the support of the
-matrix in $X_{i,0}$.  Secondly, for each $1 \leq i \neq j \leq r$,
-$supp(X_i) \cap supp(X_j) = \phi$, and since $\cA$ is an ICPPA, it
-follows that $T_i$ and $T_j$ are vertex disjoint.  Finally, since $|U|
-= |V(T)|$, it follows that $T_1, \ldots, T_r$ is a partition of $T$
-into vertex disjoint sub-trees such that for each $1 \leq i \leq r$,
-the set of matrices corresponding to nodes in $X_i$ has an ICPPA in
-$T_i$.  Hence the lemma.
-\end{proof}
-The essence of the following lemma is that an ICPPA only needs to be
-assigned to the prime sub-matrix corresponding to the root of each
-in-tree, and all the other prime sub-matrices only need to have an
-Intersection Cardinality Preserving Interval Assignments (ICPIA).
-Recall, an ICPIA is an assignment of intervals to sets such that the
-cardinality of an assigned interval is same as the cardinality of the
-interval, and the cardinality of intersection of any two sets is same
-as the cardinality of the intersection of the corresponding intervals.
-It is shown in \cite{nsnrs09} that the existence of an ICPIA is a
-necessary and sufficient condition for a matrix to have COP.  We
-present the pseudo-code to test if $M$ has an ICPPA in $T$.
-\begin{lemma} \label{lem:rooticppa} Let $M$ be a matrix and let $X$ be
-  the directed graph whose vertices are in correspondence with the
-  prime submatrices of $M$.  Further let $\{X_1,\ldots,X_r\}$ be the
-  partition of $X$ into in-trees as defined earlier in this section.
-  Let $T$ be the given tree and let $\{T_1, \ldots, T_r\}$ be a given
-  partition of $T$ into vertex disjoint sub-trees.  Then, for each $1
-  \leq i \leq r$, the set of matrices corresponding to vertices of
-  $X_i$ has an ICPPA in $T_i$ if and only if the matrix in $X_{i,0}$
-  has an ICPPA in $T_i$ and all other matrices in $X_i$ have an {\bf
-    {\em ICPIA}}.
-\end{lemma}
-\begin{proof}
-  The proof is based on the following fact- $\preccurlyeq$ is a
-  partial order and $X$ is a digraph which is the disjoint union of
-  in-trees.  Each edge in the in-tree is a containment relationship
-  among the supports of the corresonding sub-matrices. Therefore, any
-  ICPPA to a prime sub-matrix that is not the root is contained in a
-  path assigned to the sets in the parent matrix.  Consequently, any
-  ICPPA to the prime sub-matrix that is not at the root is an ICPIA,
-  and any ICPIA can be used to construct an ICPPA to the matrices
-  corresponding to nodes in $X_i$ provided the matrix in the root has
-  an ICPPA in $T_i$.  Hence the lemma.
-\end{proof}
-Lemma \ref{lem:subicppa} and Lemma \ref{lem:rooticppa} point out two
-algorithmic challenges in finding an ICPPA for a given set system
-$\cF$ in a tree $T$.  Given $\cF$, finding $X$ and its partition
-$\{X_1,\ldots,X_r\}$ into in-trees can be done in polynomial time.  On
-the other hand, as per lemma \ref{lem:subicppa} we need to parition
-$T$ into vertex disjoint sub-trees $\{T_1, \ldots, T_r\}$ such that
-for each $i$, the set of matrices corresponding to nodes in $X_i$ have
-an ICPPA in $T_i$.  This seems to be a challenging step, and it must
-be remarked that this step is easy when $T$ itself is a path, as each
-individual $T_i$ would be sub-paths.  The second algorithmic challenge
-is identified by lemma \ref{lem:rooticppa} which is to assign an ICPPA
-from a given tree to the matrix associated with the root node of
-$X_i$.
-\begin{algorithm}[h]
-  \caption{Algorithm to find an ICPPA for a matrix $M$ on tree $T$:
-    $main\_ICPPA(M, T$)}
-  \label{al:icppa-main}
-  \begin{algorithmic}
-    \STATE Identify the prime sub-matrices. This is done by constructing the strict overlap graph and identify connected components.  Each connected component yields a prime sub-matrix.   \\
-    \STATE Construct the partial order $\preccurlyeq$ on the set of prime sub-matrices.  \\
-    \STATE Construct the partition $X_1,\ldots,X_r$ of the prime
-    sub-matrices induced by $\preccurlyeq$ \\
-    \STATE For each $1 \leq i \leq r$, Check if all matrices except
-    those in $X_{i,0}$ has an ICPIA.  If a matrix does not have ICPIA
-    exit with a negative answer.  To check for the existence of ICPIA,
-    use the result in \cite{nsnrs09}.  \STATE Find a partition of
-    $T_1, \ldots, T_r$ such that matrices in $X_{i,0}$ has an ICPPA in
-    $T_i$.  If not such partition exists, exit with negative answer.
-  \end{algorithmic}
-\end{algorithm}
-
-\subsection{Complexity of Tree Path Assignment-A
-  Discussion} 
-\label{complexity} 
-Recall that the input to the Tree
-Path Assignment question is an order pair $(\cF,T)$ where $\cF$ is a
-family of subsets of an universe $U$, and $T$ is a tree such that
-$|V(T)| = |U|$.  The question is to come up with a bijection from $U$
-to $V(T)$ such that the image of each set in $\cF$ is a path in $T$.
-We show that this problem is at least as hard as the problem of
-testing if two given path chordal graphs are isomorphic.
-\begin{theorem} \label{thm:isocomp} Tree Path Assignment is
-  isomorphism-complete.
-\end{theorem}
-\begin{proof}
-  It is well known (see for example \cite{kklv10}) that testing
-  isomorphism of path chordal graphs is isomorphism complete.  We show
-  a reduction of path chordal graph isomorphism to tree path
-  assignment.  Given $G_1$ and $G_2$ two path chordal graphs, let
-  $T_2$ be the clique tree of $G_2$ obtained from say \cite{gav78}.
-  The nodes of $T_2$ correspond to the maximal cliques of $G_2$ and
-  each vertex of $G_2$ corresponds to a path in $G_2$.  This is a
-  well-known characterization of path chordal graphs and $T_2$ can be
-  computed in polynomial time.  In $G_1$, let $S_v$ denote the maximal
-  cliques of $G_1$ that contain $v$.  This can be computed in
-  polynomial time as $G_1$ is path chordal, and all chordal graph only
-  have a linear number of maximal cliques.  The universe $U$
-  corresponds bijectively to the set of maximal cliques in $G_1$, and
-  $\cF = \{S_v | v \in V(G_1)\}$.  Now, we claim that $(\cF, T_2)$ has
-  a tree path assignment if and only if $G_1$ and $G_2$ are
-  isomorphic.  This is clear since for each vertex $v \in G_1$, there
-  is an associated $S_v$ which is the set of maximal cliques
-  containing $v$.  In $G_2$, each vertex corresponds to a path in
-  $T_2$, and the nodes on this path corresponds to the maximal cliques
-  in $G_2$.  Consequenlty, a tree path assignment will naturally yield
-  an isomorphism between $G_1$ and $G_2$, and vice versa.  Therefore,
-  Tree Path Assignment is isomorphism-complete.
-\end{proof}
-
-\subsubsection{Consecutive Ones Testing is in Logspace}
-While Tree Path Assignment is isomorphism-complete, it is polynomial
-time solvable when the given tree is a path.  Indeed, in this case we
-encounter a restatement of matrices with the COP.  The known
-approaches to testing for COP fall into two categories: those that
-provide a witness when the input matrix does not have COP, and those
-that do not provide a witness.  The first linear time algorithm for
-testing COP for a binary matrix was using a data structure called PQ
-trees, which represent all COP orderings of $M$, invented by
-\cite{bl76}. There is a PQ tree for a matrix iff the matrix has COP.
-Indeed, this is an algorithmic characterization of the consecutive
-ones property and the absence of the PQ-tree does not yield any
-witness to the reason for failure.  A closely related data structure
-is the generalized PQ tree in \cite{mcc04}.  In generalized PQ tree
-the P and Q nodes are called prime and linear nodes. Aside from that,
-it has a third type of node called degenerate nodes which is present
-only if the set system does not have COP \cite{mcc04}.  Using the idea
-of generalized PQ tree, \cite{mcc04} proves that checking for
-bipartiteness in the certain incomparability graph is sufficient to
-check for COP.  \cite{mcc04} invented a certificate to confirm when a
-binary matrix does not have COP.  \cite{mcc04} describes a graph
-called incompatibility graph of a set system $\cF$ which has vertices
-$(a,b), a \ne b$ for every $a, b \in U$, $U$ being the universe of the
-set system. There are edges $((a,b),(b,c))$ and $((b,a),(c,b))$ if
-there is a set $S \in \cF$ such that $a, c \in S$ and $b \notin S$. In
-other words the vertices of an edge in this graph represents two
-orderings that cannot occur in a consecutive ones ordering of $\cF$.
-\begin{theorem}[Theorem 6.1, \cite{mcc04}]
-  Let $\cF$ be an arbitrary set family on domain $V$. Then $\cF$ has
-  the consecutive ones property if and only if its incompatibility
-  graph is bipartite, and if it does not have the consecutive ones
-  property, the incompatibility graph has an odd cycle of length at
-  most $n+3$.
-\end{theorem}
-This theorem gives a certificate as to why a given matrix does not
-have COP.  Similarly, the approach of testing for an ICPIA in
-\cite{nsnrs09} also gives a different certificate- a prime sub-matrix
-that does not have an ICPIA.  Further, the above theorem can be used
-to check if a given matrix has COP in logspace by checking if its
-incompatibility graph is bipartite. \cite{rei84} showed that checking
-for bipartiteness can be done in logspace. Thus we conclude that
-consecutive ones testing can be done in logspace.
-
-\noindent
-More recently, \cite{kklv10} showed that interval graph isomorphism
-can be done in logspace. Their paper proves that a canon for interval
-graphs can be calculated in logspace using an interval hypergraph
-representation of the interval graph with each hyperedge being a set
-to which an interval shall be assigned by the canonization
-algorithm. An overlap graph (subgraph of intersection graph, edges
-define only strict intersections and no containment) of the hyperedges
-of the hypergraph is created and canons are computed for each overlap
-component. The overlap components define a tree like relation due to
-the fact that two overlap components are such that either all the
-hyperedges of one is disjoint from all in the other, or all of them
-are contained in one hyperedge in the other. This is similar to the
-containment tree defined in \cite{nsnrs09} and in this paper. Finally
-the canon for the whole graph is created using logspace tree
-canonization algorithm from \cite{sl92}. The interval labelling done
-in this process of canonization is exactly the same as the problem of
-assigning feasible intervals to a set system, and thus the problem of
-finding a COP ordering in a binary matrix \cite{nsnrs09}.
-
-\begin{theorem}[Theorem 4.7, \cite{kklv10}]
-  \label{th:canonlabel}
-  Given an interval hypergraph $\cH$, a canonical interval labeling
-  $l_H$ for $H$ can be computed in FL.
-\end{theorem}
-
-We present the following reduction to see that COP testing is indeed
-in logspace. Given a binary matrix $M$ of order $n \times m$, let $S_i
-= \{j \mid M[j,i]=1 \}$. Let $\cF = \{S_i \mid i \in [m] \}$ be this
-set system. Construct a hypergraph $\cH$ with its vertex set being
-$\{1, 2, \dots n\}$. The edge set of $\cH$ is isomorphic to
-$\cF$. Thus every edge in $\cH$ represents a set in the given set
-system $\cF$. Let this mapping be $\pi: E(\cH) \rightarrow \cF$. It is
-easy to see that if $M$ has COP, then $\cH$ is an interval
-hypergraph. From theorem \ref{th:canonlabel}, it is clear that the
-interval labeling $l_{\cH}: V(\cH) \rightarrow [n]$ can be calculated
-in logspace. Construct sets $I_i = \{ \l_{\cH}(x) \mid x \in E, E \in
-E(\cH), \pi(E) = S_i\}$, for all $i \in [m]$. Since $\cH$ is an
-interval hypergraph, $I_i$ is an interval for all $i \in [m]$, and is
-the interval assigned to $S_i$ if $M$ has COP.
-
-Now we have the following corollary.
-\begin{corollary}
-  \label{cor:coplog}
-  If a binary matrix $M$ has COP then the interval assignments to each
-  of its columns can be calculated in FL.
-\end{corollary}
-
-\noindent
-Finally, we conclude by asking about the complexity of Tree Path
-Assignment restricted to other subclasses of trees.  In particular, is
-Tree Path Assignment in caterpillars easier than Tree Path assignment
-in general trees.
-
 \chapter{Conclusion}
@@
 
-\begin{center}
-{\tt MOVED FROM INTRO SUBSECTION IN ch:myresearch CHAPTER begin}
+\begin{center} {\tt MOVED FROM INTRO SUBSECTION IN ch:myresearch
+    CHAPTER begin}
 \end{center}
@@
 graph. % }{the following text could be in Further Study/Conclusion}
-A
-similar connection to a subclass of chordal graphs and a superclass of
-interval graphs exists for the generalization of COP.  In this case,
-the intersection graph of the corresponding set system must be a {\em
-  path graph}. Chordal graphs are of great significance, are
+A similar connection to a subclass of chordal graphs and a superclass
+of interval graphs exists for the generalization of COP.  In this
+case, the intersection graph of the corresponding set system must be a
+{\em path graph}. Chordal graphs are of great significance, are
 extensively studied, and have several applications.  One of the well
@@
 
-\begin{center}
-{\tt MOVED FROM INTRO SUBSECTION IN ch:myresearch CHAPTER\\ end}
+\begin{center} {\tt MOVED FROM INTRO SUBSECTION IN ch:myresearch
+    CHAPTER\\ end}
 \end{center}
@@
 feasibility bijection mapping vertices of the hypergraph to vertices
-of the given target tree. This thesis also discovered an exponential algorithm
-that computes a feasible TPL to a given hypergraph if it is a path hypergraph.
+of the given target tree. This thesis also discovered an exponential
+algorithm that computes a feasible TPL to a given hypergraph if it is
+a path hypergraph.
 
@@
 subtrees of a tree\cite{mcg04}.  Path graphs are well studied in the
-literature \cite{plr70}--\cite{mcg04}\tnote[TH]{Check if cite:bp93 is
+literature \cite{plr70}--\cite{mcg04}\tnote{Check if cite:bp93 is
   indeed for path graphs}.  Path graph recognition can be done in
@@
 $\cF$ in \FTPL. However, one can easily obtain a counterexample to
-show the insufficiency of this condition\tnote[TH]{Give a figure
+show the insufficiency of this condition\tnote{Give a figure
   showing insuff.}.  Path graph isomorphism is known be
-isomorphism-complete\cite{kklv10}. Therefore, \annote[TH]{it is
+isomorphism-complete\cite{kklv10}. Therefore, \annote{it is
   unlikely that we can solve the problem of finding feasible path
@@
 definitely interesting to classify the kinds of trees and hypergraphs
-for which feasible path labelings can be\tnote[TH]{ELABORATE.}  found
-efficiently.  \annote[TH]{These results would form a natural
+for which feasible path labelings can be\tnote{ELABORATE.}  found
+efficiently.  \annote{These results would form a natural
   generalization of COP testing and interval graph isomorphism,
@@
 
-While we address the computation of TPL
-for a given hypergraph% \footnote{The term {\em hypergraph} is used
+While we address the computation of TPL for a given
+hypergraph% \footnote{The term {\em hypergraph} is used
 % instead {\em set systems} are used interchangably in this
@@
 (possible extensions to optimization for COP in the
-Section~\ref{sec:optcop}.\tnote[TH]{link to section}).
+Section~\ref{sec:optcop}.\tnote{link to section}).
 
-%\newpage %DUMMY
-\tnote[TH]{ ||Complexity challenges:|| Canonization is an important
+% \newpage %DUMMY
+\tnote{ ||Complexity challenges:|| Canonization is an important
   tool in graph isomorphism. Invention of a
-% deterministic method of canonization for any class of graphs naturally
-% results in an algorithm for isomorphism. All that is required is to
-% check if the canons of two graphs are the same. Thus complexity of
-% graph isomorphism can be studied by studying canonization
-% methods. While general graph ismorphism remains elusive in terms of
-% hardness, canonization has been studied for smaller classes of graphs
-% thus giving complexity/hardness results for them.
+  % deterministic method of canonization for any class of graphs
+  % naturally results in an algorithm for isomorphism. All that is
+  % required is to check if the canons of two graphs are the
+  % same. Thus complexity of graph isomorphism can be studied by
+  % studying canonization methods. While general graph ismorphism
+  % remains elusive in terms of hardness, canonization has been
+  % studied for smaller classes of graphs thus giving
+  % complexity/hardness results for them.
 
-% In 1992, \cite{sl92} made an important discovery that tree isomorphism
-% is in logspace. It was done by inventing a method of canonization of
-% trees using a logspace depth first traversal
+  % In 1992, \cite{sl92} made an important discovery that tree
+  % isomorphism is in logspace. It was done by inventing a method of
+  % canonization of trees using a logspace depth first traversal
   % algorithm. \tnote{confirm the statement of claim}.
@@
   % % built on the fact that
-% Interval graphs are FP+C (fixed point with
-% counting) definable \tnote{elucidate} and \cite{l10} showed that
-% this implies that it captures PTIME. \tnote{elucidate} This result
-% along with that of undirected graph reachability being in logspace
-% \cite{rei08}, \cite{kklv10} proved their logspace result.
-}   %
-\tnote[TH]{ (1) the interesting survey in conclusion section of kklv10
+  % Interval graphs are FP+C (fixed point with counting) definable
+  % \tnote{elucidate} and \cite{l10} showed that this implies that it
+  % captures PTIME. \tnote{elucidate} This result along with that of
+  % undirected graph reachability being in logspace \cite{rei08},
+  % \cite{kklv10} proved their logspace result.
+}%
+\tnote{ (1) the interesting survey in conclusion section of kklv10
   (2) [[[COP helping some problems with hardness]]] having COP in the
@@
   ----- see cite-d08phd Chapter 6?  }
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Appendices.
-
-% \appendix %%% NON NEIN NAHI ILLA NONE
-
-% \chapter{A}
-% \tnote{DO WE NEED THIS?}
-
-\newpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Bibliography.
-\begin{singlespace}
-%\begin{thebibliography}{10}
-\addcontentsline{toc}{chapter}{REFERENCES}
-
-{\footnotesize
-  \bibliographystyle{alpha} 
-  \bibliography{../lib/cop-variants__thesis}
-}
-%\end{thebibliography}
-\end{singlespace}
-%\newpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% List of papers
-
-% \listofpapers %%% NON NEIN NAHI ILLA NONE
-
-% \begin{enumerate}  
-% \item Authors....  \newblock
-%  Title...
-%   \newblock {\em Journal}, Volume,
-%   Page, (year).
-% \end{enumerate}
-
-\end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% DOCUMENTATION ON THE THESIS CLASS FILE ETC.
-
-% Before reading any further please note that you are strongly advised
-% against changing any of the formatting options used in the class
-% provided in this directory, unless you are absolutely sure that it
-% does not violate the IITM formatting guidelines.  \emph{Please do not
-%   change the margins or the spacing.}  If you do change the formatting
-% you are on your own (don't blame me if you need to reprint your entire
-% thesis).  In the case that you do change the formatting despite these
-% warnings, the least I ask is that you do not redistribute your style
-% files to your friends (or enemies).
-
-% It is also a good idea to take a quick look at the formatting
-% guidelines.  Your office or advisor should have a copy.  If they
-% don't, pester them, they really should have the formatting guidelines
-% readily available somewhere.
-
-% To generate PDF's with the links from the \verb+hyperref+ package use
-% the following command:
-% \begin{verbatim}
-% % dvipdfm -o thesis.pdf thesis.dvi
-% \end{verbatim}
-
-% \section{Package Options}
-
-% Use this thesis as a basic template to format your thesis.  The
-% \verb+iitmdiss+ class can be used by simply using something like this:
-% \begin{verbatim}
-% \documentclass[PhD]{iitmdiss}  
-% \end{verbatim}
-
-% To change the title page for different degrees just change the option
-% from \verb+PhD+ to one of \verb+MS+, \verb+MTech+ or \verb+BTech+.
-% The dual degree pages are not supported yet but should be quite easy
-% to add.  The title page formatting really depends on how large or
-% small your thesis title is.  Consequently it might require some hand
-% tuning.  Edit your version of \verb+iitmdiss.cls+ suitably to do this.
-% I recommend that this be done once your title is final.
-
-% To write a synopsis simply use the \verb+synopsis.tex+ file as a
-% simple template.  The synopsis option turns this on and can be used as
-% shown below.
-% \begin{verbatim}
-% \documentclass[PhD,synopsis]{iitmdiss}                                
-% \end{verbatim}
-
-% Once again the title page may require some small amount of fine
-% tuning.  This is again easily done by editing the class file.
-
-% This sample file uses the \verb+hyperref+ package that makes all
-% labels and references clickable in both the generated DVI and PDF
-% files.  These are very useful when reading the document online and do
-% not affect the output when the files are printed.
-
-% \section{Example Figures and tables}
-
-% Fig.~\ref{fig:iitm} shows a simple figure for illustration along with
-% a long caption.  The formatting of the caption text is automatically
-% single spaced and indented.  Table~\ref{tab:sample} shows a sample
-% table with the caption placed correctly.  The caption for this should
-% always be placed before the table as shown in the example.
-
-% \begin{figure}[htpb]
-%   \begin{center}
-%     \resizebox{50mm}{!} {\includegraphics *{iitm.eps}}
-%     \resizebox{50mm}{!} {\includegraphics *{iitm.eps}}
-%     \caption {Two IITM logos in a row.  This is also an
-%       illustration of a very long figure caption that wraps around two
-%       two lines.  Notice that the caption is single-spaced.}
-%   \label{fig:iitm}
-%   \end{center}
-% \end{figure}
-
-% \begin{table}[htbp]
-%   \caption{A sample table with a table caption placed
-%     appropriately. This caption is also very long and is
-%     single-spaced.  Also notice how the text is aligned.}
-%   \begin{center}
-%   \begin{tabular}[c]{|c|r|} \hline
-%     $x$ & $x^2$ \\ \hline
-%     1  &  1   \\
-%     2  &  4  \\
-%     3  &  9  \\
-%     4  &  16  \\
-%     5  &  25  \\
-%     6  &  36  \\
-%     7  &  49  \\
-%     8  &  64  \\ \hline
-%   \end{tabular}
-%   \label{tab:sample}
-%   \end{center}
-% \end{table}
-
-% \section{Bibliography with BIB\TeX}
-
-% I strongly recommend that you use BIB\TeX\ to automatically generate
-% your bibliography.  It makes managing your references much easier.  It
-% is an excellent way to organize your references and reuse them.  You
-% can use one set of entries for your references and cite them in your
-% thesis, papers and reports.  If you haven't used it anytime before
-% please invest some time learning how to use it.
-
-% I've included a simple example BIB\TeX\ file along in this directory
-% called \verb+refs.bib+.  The \verb+iitmdiss.cls+ class package which
-% is used in this thesis and for the synopsis uses the \verb+natbib+
-% package to format the references along with a customized bibliography
-% style provided as the \verb+iitm.bst+ file in the directory containing
-% \verb+thesis.tex+.  Documentation for the \verb+natbib+ package should
-% be available in your distribution of \LaTeX.  Basically, to cite the
-% author along with the author name and year use \verb+\cite{key}+ where
-% \verb+key+ is the citation key for your bibliography entry.  You can
-% also use \verb+\citet{key}+ to get the same effect.  To make the
-% citation without the author name in the main text but inside the
-% parenthesis use \verb+\citep{key}+.  The following paragraph shows how
-% citations can be used in text effectively.
-
-% More information on BIB\TeX\ is available in the book by
-% \cite{lamport:86}.  There are many
-% references~\citep{lamport:86,prabhu:xx} that explain how to use
-% BIB\TeX.  Read the \verb+natbib+ package documentation for more
-% details on how to cite things differently.
-
-% Here are other references for example.  \citet{viz:mayavi} presents a
-% Python based visualization system called MayaVi in a conference paper.
-% \citet{pan:pr:flat-fst} illustrates a journal article with multiple
-% authors.  Python~\citep{py:python} is a programming language and is
-% cited here to show how to cite something that is best identified with
-% a URL.
-
-% \section{Other useful \LaTeX\ packages}
-
-% The following packages might be useful when writing your thesis.
-
-% \begin{itemize}  
-% \item It is very useful to include line numbers in your document.
-%   That way, it is very easy for people to suggest corrections to your
-%   text.  I recommend the use of the \texttt{lineno} package for this
-%   purpose.  This is not a standard package but can be obtained on the
-%   internet.  The directory containing this file should contain a
-%   lineno directory that includes the package along with documentation
-%   for it.
-
-% \item The \texttt{listings} package should be available with your
-%   distribution of \LaTeX.  This package is very useful when one needs
-%   to list source code or pseudo-code.
-
-% \item For special figure captions the \texttt{ccaption} package may be
-%   useful.  This is specially useful if one has a figure that spans
-%   more than two pages and you need to use the same figure number.
-
-% \item The notation page can be entered manually or automatically
-%   generated using the \texttt{nomencl} package.
-
-% \end{itemize}
-
-% More details on how to use these specific packages are available along
-% with the documentation of the respective packages.
-
-% END OF DOCUMENTATION
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-01-03 src/thesis.tex
Cleaned up thesis.tex. Split thesis.tex into chapterwise subfiles for easy edit tex management. Content changes are mainly in chapter 1 and 3 although nothing significant. To being working on chapter 2 after this commit.

--- 2011-12-30 src/thesis.tex
@@
 \usepackage{t1enc}             % T1 encoding??  
-%\usepackage{boxedminipage}
+\usepackage[T1,OT1]{fontenc}
+\usepackage{listings}
+
 \usepackage{../lib/mythesislib}
@@
 % \usepackage{comment}         % DEBUG ONLY
+\includeonly{010chap1,020chap2}
+%\includeonly{0021ded}
 
@@
 \begin{document}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Title page
-\title{\mythesistitleNOTSYNPOSIS} 
-\author{\myname} 
-\date{
+%
+\title{\mythesistitleNOTSYNPOSIS} \author{\myname} \date{
   {\mythesissubmissionmonth}}
 \department{\MakeUppercase{\mythesisdept}}
-%\department{\mythesisdept}
-
-%\nocite{*}
 \maketitle
@@
 % Certificate
-\certificate
-
-\vspace*{0.5in}
-
-\noindent 
-This is to certify that the thesis titled {\bf \mythesistitleNOTSYNPOSIS},
-submitted by {\bf \myname}, to the {\bf Indian Institute of Technology
-Madras}, for the award of the degree of {\bf \mydegree}, is 
-a bona fide record of the research work done by her under our
-supervision.  The contents of this thesis, in full or in parts, have
-not been submitted to any other Institute or University for the award
-of any degree or diploma.
-
-\vspace*{1.5in}
-
-\begin{singlespacing}
-  \hspace*{-0.25in}
-  \parbox{3in}{
-    \noindent {\bf Dr.~\myadvisor} \\
-    \small
-    \noindent Research Guide\\
-    \noindent Associate Professor\\
-    \noindent Dept. of \mydept\\
-    \noindent IIT Madras -- 600 036
-  } \hfill
-  \parbox{1.25in}{               % {1.75in}
-    \small
-    \vspace*{0.75in}
-    \begin{tabular}[h]{r}
-    \noindent {Chennai}\\         
-    \noindent {\mythesissubmissiondate} 
-    \end{tabular}
-  }
-\end{singlespacing}
+%
+\include{001certificate}
 
@@
 % Acknowledgements
-\acknowledgements
+%
+\include{002ack}
 
-\tnote{WIP}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Dedication?
+%
+\include{0021ded}
 
@@
 % Abstract
-
-\begin{abstract}
-  \noindent {\em {\bf Keywords}: \hspace*{0.5em} \parbox[t]{4.4in}{consecutive
-    ones property, algorithmic graph theory, hypergraph isomorphism,
-    interval labeling}} \vspace*{24pt}
-
-  \def \tem {}
-  \noindent 
-  Consecutive-ones property is a non-trivial property of binary
-  matrices that has been studied widely in the literature for over
-  past 50 years. Detection of COP in a matrix is possible efficiently
-  and there are several algorithms that achieve the same. This thesis
-  documents the work done on an extension of COP extended from the
-  equivalent interval assignment problem in \cite{nsnrs09}. These new
-  results rigorously prove a natural extension (to trees) of their
-  characterization as well as makes connections to graph isomorphism,
-  namely path graph isomorphism.
-
-  \tnote{EXPAND. 
-  Abstract must be a brief about what results we have and how it fits
-  in the body of research.\\
-  -- Area: Broad to Specialized. i.e. Combinatorial
-    algorithms -> Matrix reorganization -> general data reorganization
-    (interval assignment) -> path assignment\\
-  -- Class of problems: say, data reorganization.\\
-  -- Nature of results: Is a generalization. We have a
-    Polynomial algorithm for a subset of the generalization.\\
-% maybe add - we explore a natural
-%     generalization of results on binary matrices with the |\tem
-%     consecutive-ones property|.  We consider the following constraint
-%     satisfaction problem. Given (i) a set system $\F \subseteq$
-%     $(2^{U} \setminus \emptyset)$ of a finite set $U$ of cardinality
-%     $n$, (ii) a tree $T$ of size $n$ and (iii) a bijection called
-%     |\tem tree path labeling|, $\cl$ mapping the sets in $\cF$ to
-%     paths in $T$, does there exist at least one bijection $\phi:U
-%     \rightarrow V(T)$ such that for each $S \in \cF$, $\{\phi(x) \mid
-%     x \in S\} = \cl(S)$?  A tree path labeling of a set system is
-%     called |\tem feasible| if there exists such a bijection $\phi$.
-%     We present an algorithmic characterization of feasible tree path
-%     labeling. COP is a special instance of tree path labeling problem
-%     when $T$ is a path.  We conclude with a polynomial time algorithm
-%     to find a feasible tree path labeling of a given set system when
-%     $T$ is a |\tem $k$-subdivided star|, set system has a single
-%     containment tree of overlap components and set size is limited to
-%     at most $k+2$.
-  }
-\end{abstract}
-
-%\pagebreak
+%
+\include{003abstract}
 
@@
 % Table of contents etc.
-
+%
 \begin{singlespace}
@@
   \thispagestyle{empty}
-
-\listoftables
-\addcontentsline{toc}{chapter}{LIST OF TABLES}
-\listoffigures
-\addcontentsline{toc}{chapter}{LIST OF FIGURES}
+  \listoftables \addcontentsline{toc}{chapter}{LIST OF TABLES}
+  \listoffigures \addcontentsline{toc}{chapter}{LIST OF FIGURES}
 \end{singlespace}
@@
 % Abbreviations
-\abbreviations
-
-\noindent 
-\begin{tabbing}
-xxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
-\textbf{COP}   \> Consecutive-ones Property \\
-\textbf{COT}   \> Consecutive-ones Testing \\
-\textbf{ICPIA}   \> Intersection Cardinality Preservation Interval Assignment \\
-\textbf{ICPPL}   \> Intersection Cardinality Preserved Path Labeling \\
-% ADD MORE HERE
-\end{tabbing}
-
-\pagebreak
+%
+\include{004abbrv}
 
@@
 % Notation
+%
+\include{005notation}
+%\pagebreak \clearpage
 
-\chapter*{\centerline{NOTATION}}
-\addcontentsline{toc}{chapter}{NOTATION}
+\pagenumbering{arabic} % The main text will follow from this point so
+                       %  set the page numbering to arabic from here on.
 
-\begin{singlespace}
-\begin{tabbing}
-xxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
-\textbf{$2^{U}$}  \> Powerset of set $U$ \\
-\end{tabbing}
-\end{singlespace}
-
-\pagebreak
-\clearpage
-
-% The main text will follow from this point so set the page numbering
-% to arabic from here on.
-\pagenumbering{arabic}
-
-\chapter{Introduction}
-\label{ch:intro}
-
-Consecutive-ones property is a non-trivial property of binary matrices
-that has been studied widely in the literature for over past 50
-years. Detection of COP in a matrix is possible efficiently and there
-are several algorithms that achieve the same. This thesis documents
-the work done on an extension of COP extended from the equivalent
-interval assignment problem in \cite{nsnrs09}. These new results
-rigorously prove a natural extension (to trees) of their
-characterization as well as makes connections to graph isomorphism,
-namely path graph isomorphism.
-
-\tnote{Have a para on Organization: Outline of document}
-Section~\ref{sec:background} gives a brief survey of COP and
-optimization problems related to it followed by motivation for the
-thesis in Section~\ref{sec:motive}.  Section~\ref{sec:results}
-presents a summary of our results on the extension of COP namely, the
-tree path labeling problem. \tnote[TH]{update for thesis}
-
-% \tnote{testing 1 2 3...}
-% \tnote[bogus1]{testing 1 2 3...}
-% \tnote[bogus2]{testing 1 2 3...}
-% \tnote[GTC]{testing 1 2 3...}
-% \tnote[bogus4]{testing 1 2 3...}
-% \tnote[TH]{testing 1 2 3...}
-
-\section{Illustration of the problem}
-\label{sec:problem}
-
-A group of students, \Pa, \Pig, \Sn, \Wo, \Vi, \Li, \Ch, \Sa, \Fr,
-\Sc\ and \Lu\ enroll at the {\WSI} for a liberal arts programme.  As
-part of their semester thesis, they pick a body of work to study and
-form the namesake study groups, {\LLL}, {\GGG}, {\BBB} and
-{\TTT}\tnote[TH]{put bib entries for these works!}. A student will be
-in at least one study group and may be in more than one. For instance,
-as will be seen later, {\Fr} studies both {\LLL} and {\TTT} while \Wo\
-studies only \BBB.
-
-Let $U$ and $\cF$ represent the set of students and the set of study
-groups respectively and the integers $n$ and $m$ denote the total
-number students and study groups respectively. In relation to this
-example, these are defined in Table~\ref{tab:wsigroups}. Also given
-there is the study group allocation to students.
-
-\begin{table}[t]%[htbp]
-  \centering
-  { \footnotesize
-  \begin{tabular}{rcl}
-    $U $&$=$&$ \{\xPa,\; \xPi,\; \xSn,\; \xWo,\; \xVi,\; \xLi,\; \xCh,\;
-    \xSa,\; \xFr,\; \xSc,\; \xLu\}$\\
-    $\cF $&$=$&$ \{\xLLL, \xGGG, \xBBB, \xTTT\}$\\
-    $\xLLL $&$=$&$ \{\xCh,\;  \xSa,\;  \xFr,\;  \xSc,\;  \xLu \}$\\
-    $\xGGG $&$=$&$ \{\xPa,\;  \xPi,\;  \xVi,\;  \xCh \}$\\
-    $\xBBB $&$=$&$ \{\xSn,\;  \xPi,\;  \xWo \}$\\
-    $\xTTT $&$=$&$ \{\xVi,\;  \xLi,\;  \xCh,\;  \xFr \}$\\
-    $n $&$=$&$ |U| = 11$\\
-    $m $&$=$&$ |\cF| = 4$      
-  \end{tabular}
-}
-  \caption{\figtabsize Students and study groups in \WSI}
-  \label{tab:wsigroups}
-\end{table}
-
-The campus has a residential area {\residenceblock} that has $n$
-single occupancy apartments reserved for the study groups'
-accommodation.  All these apartments are located such that the streets
-connecting them do {\em not} form
-loops. Fig~\ref{fig:streetmappathpeople} \tnote[TH]{fig:streetmap}
-shows the street map for {\residenceblock}. It may be noted that as a
-graph, it classifies as a tree.
-
-\begin{table}[htbp]
-  \centering
-  {\footnotesize    
-    \begin{tabular}{l||l}
-      \begin{tabular}{rcl}
-        $T $&$=$&$ \text{\em Street map tree of \residenceblock}$\\
-        $V(T) $&$=$&$ \{ 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11 \}$\\
-        $\cP $&$=$&$ \{R\xLLL, R\xGGG, R\xBBB, R\xTTT\}$\\
-        $R\xLLL $&$=$&$ \{9, 1, 5, 3, 11\}$\\
-        $R\xGGG $&$=$&$ \{7, 2, 6, 5\}$\\
-        $R\xBBB $&$=$&$ \{8, 2, 4 \}$\\
-        $R\xTTT $&$=$&$ \{10, 6, 5, 3\}$\\
-        $n $&$=$&$ |V| = 11$\\
-        $m $&$=$&$ |\cP| = 4$\\\\
-        $\cl $&$=$& {\em Study group to route mapping}\\        
-        $\cl(\mathbb{X}) $&$=$&  $R\mathbb{X}$ for all $\mathbb{X} \in
-        \cF$
-      \end{tabular} & 
-      \begin{tabular}{c}
-       {\em Apartment allocation }($\phi$)\\
-                    \begin{tabular}{c|c}                      
-                      1  & \xSa\\
-                      2  & \xPi\\
-                      3  & \xFr\\
-                      4  & \xWo\\
-                      5  & \xCh\\
-                      6  & \xVi\\
-                      7  & \xPa\\
-                      8  & \xSn\\
-                      9  & \xLu\\
-                      10 & \xLi\\
-                      11 & \xSc                     
-                    \end{tabular}
-      \end{tabular}\\
-    \end{tabular}
-  }
-  \caption{\figtabsize A solution to study group accommodation problem}
-  \label{tab:iltree}
-\end{table}
-
-A natural question would be to find how the students should be
-allocated apartments such that each study group has the {\em least
-  distance to travel} for a discussion? More specifically, we are
-interested in enforcing additional conditions, namely, that all the
-students in a study group must be next to each other; in other words,
-for one student to reach another fellow study group member's apartment
-(for all study groups the student is part of), she must not have to
-pass the apartment of any student who is not in that study group. To
-further elucidate, the apartments of students of any study group must
-be arranged in an exclusive unfragmented path on the street
-map. Exclusivity here means that the path must not have apartments
-from other study groups (unless that apartment is also part of {\em
-  this} study group).
-
-An intuitive approach to this problem would be to first find the paths
-that each study group decides to inhabit and then refine the
-allocation to individual students. A feasible allocation of exclusive
-routes to study groups is illustrated in
-Fig~\ref{fig:streetmappathpeople} \tnote[TH]{fig:streetmappath} and
-the students' allocation of apartments that obeys this route
-allocation is also shown. Table~\ref{tab:iltree} shows the same
-solution set theoretically.  How this is algorithmically computed is
-the focus of this thesis.
-
-\begin{figure}[htbp] %[t]%
- \centering
-  (a) \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
-  (b) \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
-
- \label{fig:streetmap}
- \caption{\figtabsize (a) {\residenceblock} street map (b)
-   {\residenceblock} street map with study group routes
-   allocated. Routes are color coded as follows: red for
-   \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
-   group, orange for \textcolor{yellow}{$\xBBB$} group, green for
-   \textcolor{green}{$\xTTT$} group}
-
- \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
-  \caption{\figtabsize Individual allocation of apartments to students
-    in {\residenceblock} that meets the requirements stated before.
-    The routes are color coded as follows: red for
-    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
-    group, orange for \textcolor{yellow}{$\xBBB$} group, green
-    for \textcolor{green}{$\xTTT$} group. {\tiny {\em Peanuts images
-        {\copyright} Charles Schulz}}}% YellowOrange
-  \label{fig:streetmappathpeople}
-\end{figure}
-\tnote{UPDATE IMAGE (b) 2-infinite-loop-BTWF.png. REMOVE TITLES. ADD
-  {$\xLLL$} {$\xGGG$} {$\xBBB$} {$\xTTT$}}
-
-\tnote[TH]{make dashed/textured lines for routes. make it color
-  agnostic.}
-
-As a special case, suppose all the apartments are on the same street
-or if they are all lined up on a single path, the street map becomes a
-tree that is just a path. Then the problem becomes what is called an
-{\em interval assignment problem}. The idea of interval assignment may
-not be obvious here; hence to see this, consider a different problem
-in {\WSI} where the classes for these study groups courses need to be
-scheduled during a day (or a week or any time period). Each study
-group has a bunch of courses associated with it some of which may be
-shared by two or more study groups. It is mandatory that a student who
-is a member of a study group takes all the courses associated with
-that group. There are slots during the day for classes to be held and
-the problem is to allocate class slots to courses such that all the
-classes of a study group are consecutive. It is debatable if this will
-not hamper the attention span and memory retention rate of the
-students but that is, regrettably, out of the scope of this
-thesis. The parallels between this class allocation problem and the
-accommodation problem can be seen as follows. The set $U$ here, are
-the courses offered (say Course 101 {\coneohone}, Course 102
-{\coneohtwo} and so on). In this variation of the problem, the
-collection $\cF$ is the set of study groups but the study groups are
-filled by course IDs (in place of students in the earlier
-example). For instance, Course 101 is mandatory for all study groups
-$\xLLL$, $\xGGG$, $\xBBB$, $\xTTT$ and Course 102 is mandatory for
-only the $\xLLL$ group) and so on. The sequence of class slots for the
-day (or week or any time period) is analogous to the street map in the
-accommodation problem. It is quite obvious now why this version of the
-problem (where the ``target graph'' is a path and not any
-tree\tnote[TH]{Allowing any tree in this example could be seen as a
-  scenario where there are parallel classes. A node falling in the
-  path between two other nodes would mean that the corresponding is
-  scheduled between the other two.}) is called an interval assignment
-problem.
-
-The interval assignment problem to a set system is equivalent to the
-consecutive-ones property (COP) problem in binary matrices\cite{wlh02,
-  nsnrs09}.  The COP problem is to rearrange rows (columns) of a
-binary matrix in such a way that every column (row) has its {\un}s
-occur consecutively. If this is possible the matrix is said to have
-the COP.  COP is a well researched combinatorial problem and has
-several positive results on tests for it and computing the COP
-permutation (i.e. the course schedule in the above illustration) which
-will be surveyed later in this document. Hence we are interested in
-extensions of COP, more specifically, the extension of interval
-assignment problem to tree path assignment problem (which is
-illustrated by the study group accommodation problem).
-
-\section{Basic preliminaries - general definitions and nomenclature}
-\tnote{(definitions theorems etc) needed if any
-    - graph theory}
-
-\section{Consecutive-ones Property Testing - a Brief Survey}
-\label{sec:background}
-
-In this section, a brief survey of the consecutive-ones problem and its
-optimization problems is presented.
-
-%%%%%%                              %%%%%%
-%THX is the second level of ``tnote''-ing%
-%%%%%%                              %%%%%%
-\tnote[TH]{ADD: As it will be described in detail later in this
-  document, isomorphism of certain % classes of graphs, namely chordal
-%   graphs, have a close relationship with consecutive-ones property and
-%   generalizations of it.  This is perhaps because of how closely COP
-%   of a matrix relates to properties of graphs derived from matrices as
-%   seen in the following results.
-}
-\tnote[THX]{ADD: peo exists iff chordal. lexicographic BFS
-  [tag:chordalGraph]} %
-\tnote[THX]{ADD: A well known result in %perfect graph theory is that
-%  the maximal cliques of  an interval graph $G$ can be linearly ordered
-%   such that for all $v \in V(G)$, cliques containing $v$ are
-%   consecutive in the ordering cite~{gh64}. This clearly means that a
-%   graph $G$ is an interval graph if and only if
-}% 
-\tnote[THX]{verify
-  from paper the statement of claim.}% 
-\tnote[TH]{maximal clique vertex
-  incidence matrix of%  $G$ has COP.  Also, maximal cliques of any
-%   chordal graph can be enumerated in polytime $O(m+n)$
-}
-\tnote[THX]{citation?!!}%  
-\tnote[TH]{cite~{fg65} uses these results to
-  give the first polynomial time algorithm for COT.}%
-\tnote[THX]{check. how do they use it?}%  
-\tnote[TH]{ A bipartite graph
-  is convex%  (on $R$) if and only if its half adjacency matrix has COP
-%   on rows.  The results in cite~{bl76} on COT are based on the result
-%   that interval graphs are AT-free chordal graphs.
-}%
-\tnote[THX]{the latter being Tucker's?}%  
-\tnote[THX]{(2) TBD survey
-  -- % see blue notes (in notebook) under Graph Isomorphism. namely
-%   citations in cite:aas93 (3) a brief on heirarchy of chordal graphs,
-%   path graphs, interval graphs, peo, clique tree etc. and results. (4)
-%   interval graphs are incomparability graphs. see
-%   Golumbic. [tag:classification] (5) finding min length hole in
-%   bipartite graph is polynomial time [Sec 3.3.2 in cite:d08phd] (6)
-%   theorem 2.2 in cite:d08phd - $G$ is union of v.d. caterpillars iff
-%   $M$ has COP in rows. $M$ is edge vertex incidence matrix $(*,2)$ of
-%   $G$
-}%
-
-\subsection{Matrices with COP}
-\label{sec:copmatrices}
-As seen earlier, the interval assignment problem (illustrated as the
-course scheduling problem in Section~\ref{sec:problem}), is a special
-case of the problem we address in this thesis, namely the tree path
-labeling problem (illustrated as the study group accommodation
-problem). The interval assignment problem and COP problem are equivalent
-problems. In this section we will see some of the results that
-exists in the literature today towards solving the COP problem and
-optimization problems surrounding it.
-
-\begin{figure}[t] %[htbp]
-  \centering
-
-  {\figtabsize
-    \begin{tabular}[h]{l|lcccl}
-      $M_1$: & $M_1'$: &&&& $M_2$:\\
-      &&&&&\\
-      \begin{tabular}[h]{llll}
-        $c_1$ & $c_2$ &$c_3$ &$c_4$\\
-        &&&\\
-        \un & 0   & \un & 0\\
-        0   & \un & 0   & \un \\
-        \un & 0   & 0   & \un
-      \end{tabular}
-      &
-      \begin{tabular}[h]{llll}
-        $c_3$ &$c_1$ &$c_4$& $c_2$\\
-        &&&\\
-        \un & \un & 0 & 0\\
-        0 & 0 & \un & \un \\
-        0 & \un & \un & 0
-      \end{tabular}
-      &&&&
-      \begin{tabular}[h]{llll}
-        $d_1$ & $d_2$ &$d_3$ &$d_4$\\
-        % &&&\\
-        &&&\\
-        \un & \un & 0 & 0\\
-        0 & \un & \un & 0 \\
-        0 & \un & 0 & \un 
-      \end{tabular}
-    \end{tabular}
-  }
-
-  \caption{\figtabsize Matrices with and without COP. $M_1$ has COP
-    because by permuting its columns, $c_1$-$c_4$, one can obtain
-    $M_1'$ where the {\un}s in each row are consecutive. $M_2$,
-    however, does not have COP since no permutation of its columns,
-    $d_1$-$d_4$, will arrange {\un}s in each row consecutively
-    \cite{d08phd}.}
-
-  \label{fig:cop-matrix}
-\end{figure}
-
-Recall that a matrix with COP is one whose rows (columns) can be
-rearranged so that the {\un}s in every column (row) are in consecutive
-rows (columns). Figure~\ref{fig:cop-matrix} shows examples of this
-property.  COP in binary matrices has several practical applications
-in diverse fields including scheduling \cite{hl06}, information
-retrieval \cite{k77} and computational biology \cite{abh98}.  Further,
-it is a tool in graph theory \cite{mcg04} for interval graph
-recognition, characterization of Hamiltonian graphs, planarity testing
-\cite{bl76} and in integer linear programming \cite{ht02,hl06}.
-
-The obvious first questions after being introduced to the consecutive
-ones property of binary matrices are if COP can be detected
-efficiently in a binary matrix and if so, can the COP permutation of
-the matrix also be computed efficiently?  Recognition of COP in a
-binary matrix is polynomial time solvable and the first such algorithm
-was given by \cite{fg65}.  A landmark result came a few years later
-when \cite{at72} discovered the families of forbidden submatrices that
-prevent a matrix from having COP and most, if not all, results that
-came later were based on this discovery which connected COP in binary
-matrices to convex bipartite graphs. In fact, the forbidden
-submatrices came as a corollary to the discovery that convex bipartite
-graphs are AT-free in \cite{at72}\tnote[TH]{check}. The first linear
-time algorithm for COP testing (COT) was invented by \cite{bl76} using
-a data structure called PQ trees.  Since then several COT algorithms
-have been invented -- some of which involved variations of PQ trees
-\cite{mm96,wlh01,mcc04}, some involved set theory and ICPIA
-\cite{wlh02,nsnrs09}, parallel COT algorithms\cite{as95,bs03,ly91} and
-certifying algorithms\cite{mcc04}.
-
-The construction of PQ trees in \cite{bl76} draws on the close
-relationship of matrices with COP to interval graphs. A PQ tree of a
-matrix is one that stores all row (column) permutations of the matrix
-that give the COP orders (there could be multiple orders of rows or
-columns) of the matrix. This is constructed using an elaborate linear
-time procedure and is also a test for planarity\tnote[TH]{check check
-  check. both interval graph and planarity in this paper?}.  PQR trees
-is a generalized data structure based on PQ trees \cite{mm96,mpt98}.
-\cite{tm05} describes an improved algorithm to build PQR
-trees. \tnote[TH]{improv in terms of what?}\cite{wlh02} describes the
-simpler algorithm for COT. Hsu also invented PC trees
-\cite{wlh01}\tnote[TH]{This result first appeared inproc ISAAC92}
-which is claimed to be much easier to implement. \cite{nsnrs09}
-describes a characterization of consecutive-ones property solely based
-on the cardinality properties of the set representations of the
-columns (rows); every column (row) is equivalent to a set that has the
-row (column) indices of the rows (columns) that have one entries in
-this column (row). This is interesting and relevant, especially to
-this thesis because it simplifies COT to a great degree. \tnote[TH]{it
-  reduces the solution search space. fill in the blanks.}
-
-\cite{mcc04} describes a different approach to COT. While all previous
-COT algorithms gave the COP order if the matrix has the property but
-exited stating negative if otherwise, this algorithm gives an evidence
-by way of a certificate of matrix even when it has no COP. This
-enables a user to verify the algorithm's result even when the answer
-is negative. This is significant from an implementation perspective
-because automated program verification is hard and manual verification
-is more viable. Hence having a certificate reinforces an
-implementation's credibility. Note that when the matrix {\em has} COP,
-the COP order is the certificate.  The internal machinery of this
-algorithm is related to the weighted betweenness problem
-addressed\tnote[TH]{in what way??} in \cite{co98}.  \tnote[TH]{expand
-  on the COP order graph creation and it having to be bipartite for M
-  to have COP. and thus an odd cycle being an evidence of no COP.}
-\tnote[TH]{ where should this go?: (1) cite|jlm97 (application of PQ trees
-  in graphics). (2) helly's theorem citation
-  19XXdgk-Hellystheorem-Danzer-Gruenbaum-Klee}
-
-\subsection{Optimization problems in COP}
-\label{sec:optcop}
-
-So far we have been concerned about matrices that have the consecutive
-ones property. However in real life applications, it is rare that data
-sets represented by binary matrices have COP, primarily due to the
-noisy nature of data available. At the same time, COP is not arbitrary
-and is a desirable property in practical data representation
-\cite{co98,jkckv04,k77}. In this context, there are several
-interesting problems when a matrix does not have COP but is ``close''
-to having COP or is allowed to be altered to have COP. These are the
-optimization problems related to a matrix which does not have
-COP. Some of the significant problems are surveyed in this section.
-
-\tnote[TH]{ -- sect 4.1 in cite:d08phd has many results
-  surveyed. hardness results, approx. results. results are usually for
-  a class of matrices $(a,b)$ where number {\un}s in columns and rows
-  are restriced to $a$ and $b$ . -- problem of flipping at most $k$
-  entries of $M$ to make it attain COP. this is NP complete
-  cite:b75-phd}\tnote[TH]{(1) scite:lb62 showed that interval graphs
-  are AT-free.  describe AT (2) show the close relationship b/w COP
-  and graphs sec 2.2, pg 31} \cite{at72} showed that a matrix that
-does not have COP have certain substructures that prevent it from
-having COP. Tucker classified these forbidden substructures into five
-classes of submatrices. This result is presented in the context of
-convex bipartite graphs which \cite{at72} proved to be
-AT-free\tnote[TH]{ check this up. give details. - doms'}. By
-definition, convex bipartite graph have half adjacency matrices that
-have COP on either rows or columns (graph is biconvex if it has COP on
-both)\cite{d08phd}. A half adjacency matrix is a binary matrix
-representing a bipartite graph as follows. The set of rows and the set
-of columns form the two partitions of the graph. Each row node is
-adjacent to those nodes that represent the columns that have {\un}s in
-the corresponding row. \cite{at72} proves that this bipartite graph
-has no asteroidal triple if and only if the matrix has COP and goes on
-to identify the forbidden substructures for these bipartite
-graphs. The matrices corresponding to these substructures are the
-forbidden submatrices.
-
-Once a matrix has been detected to not have COP (using any of the COT
-algorithms mentioned earlier), it is naturally of interest to find out
-the smallest forbidden substructure (in terms of number of rows and/or
-columns and/or number of entries that are {\un}s). \cite{d08phd}
-discusses a couple of algorithms which are efficient if the number of
-{\un}s in a row is small. This is of significance in the case of sparse
-matrices where this number is much lesser than the number of
-columns. $(*,\Delta)${\em -matrices} are matrices with no restriction
-on number of {\un}s in any column but has at most $\Delta$ {\un}s in
-any row. {\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are
-similar problems which deals with inducing COP on a matrix. In {\sc
-  Min COS-R (Min COS-C)} the question is to find the minimum number of
-rows (columns) that must be deleted to result in a matrix with COP.
-In the dual problem {\sc Max COS-R (Max COS-C)} the search is for the
-maximum number of rows (columns) that induces a submatrix with
-COP. Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
-a submatrix $M'$ with all columns\tnote[TH]{check if b75 deals with
-  COP col or COP row. also is it any submatrix with k less than r rows
-  or submatrix must have all columns?} but a maximum cardinality
-subset of rows such that $M'$ has COP is NP complete. \cite{hg02}
-corrects an error of the abridged proof of this reduction as given in
-\cite{gj79}.  \cite{d08phd} discusses all these problems in detail
-giving an extensive survey of the previously existing results which
-are almost exhaustively all approximation results and hardness
-results. Taking this further, \cite{d08phd} presents new results in
-the area of parameterized algorithms for this
-problem\tnote[TH]{elaborate - what are the results?}.
-
-Another problem is to find the minimum number of entries in the matrix
-that can be toggled to result in a matrix with COP.  \cite{v85}
-discusses approximation of {\sc COP Augmentation} which is the problem
-of changing of the minimum number of zero entries to {\un}s so that
-the resulting matrix has COP. As mentioned earlier, this problem is
-known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
-using a reduction to the longest path problem, \tnote[TH]{or is it a
-  survey of another result?  check.} that finding a Tucker's forbidden
-submatrix of at least $k$ rows is NP complete. \tnote[TH]{how is this
-  different from booth's 75 result??}  \tnote[TH]{where should this
-  go? cite|tz04 (approx submatrix with COP sparse matrices)}
-
-\cite{jkckv04} discusses the use of matrices with almost-COP (instead
-of one block of consecutive {\un}s, they have $x$ blocks, or {\em
-  runs}, of consecutive {\un}s and $x$ is not too large) in the
-storage of very large databases.  The problem is that of reordering of
-a binary matrix such that the resulting matrix has at most $k$ runs of
-{\un}s. This is proved to be NP hard using a reduction from the
-Hamiltonian path problem.\tnote[TH]{Theorem 2.1 in jkckv}
-\tnote[TH]{(1) A connection of COP problem to the travelling salesman
-  problem is also introduced. what does this mean? -- COP can be used
-  as a tool to reorder $0.5T \le runs(M) \le T$. (2) The optimization
-  version of the $k$-run problem, i.e. minimization of number of
-  blocks of ones is proven to be NP complete by
-  cite:k77}\tnote[TH]{are these two the same?} \tnote[TH]{what is the
-  reduction?} \tnote[TH]{other problems similar to COP -- cite:ckl96
-  (ILP, circ ones, one drop) -- cite:th98 (generalization of COP -
-  minimax, biotonic column) Tucker}
-
-\section{********* Application of COP in Areas of Graph Theory and
-  Algorithms}
-\label{sec:appcopGTA}
-
-\subsection{********* COP in Relational Database Model}
-\label{sec:apprdbm}
-\tnote{(set systems theme)}
-
-\subsection{********* COP in Graph Isomorphism}
-\label{sec:appgraphiso}
-\tnote{(canonization theme)}
-
-\subsection{********* Certifying Algorithms}
-\label{sec:appcertalgo}
-\tnote{ (certification McC04 theme)}
-
-\section{Generalization of COP - the Motivation}
-\label{sec:motive}
-Section~\ref{sec:copmatrices} introduced a succinct characterization
-for consecutive-ones property which is solely based on the cardinality
-properties of the set representations of the matrix's columns
-\cite{nsnrs09}. This result is very relevant to this thesis because
-aside from it simplifying COT to a great degree, our generalization
-problem is motivated by their results.
-
-\cite{nsnrs09} characterizes interval assignments to the sets which
-can be obtained from a single permutation of the rows.  For an
-assignment to be feasible, the cardinality of the interval assigned to
-each set in the system must be same as the cardinality of the set, and
-the intersection cardinality of any two intervals must be same as the
-intersection cardinality of their corresponding sets.  While this is
-obviously a necessary condition, this result shows this is also
-sufficient.  \cite{nsnrs09} calls this an Intersection Cardinality
-Preserving Interval Assignment (ICPIA).  This paper generalizes the
-idea from \cite{wlh02} of decomposing a given binary matrix into prime
-matrices for COT and describes an algorithm to test if an ICPIA exists
-for a given set system.
-
-The equivalence of the problem of testing for the consecutive-ones
-property to the constraint statisfaction problem of interval
-assignment \cite{nsnrs09} or interval labeling \cite{kklv10} is as
-follows. Every column (row) of the binary matrix can be converted into
-a set of non-negative integers which are the indices of rows (columns)
-with {\un}s in that column (row). It is apparent that if the matrix
-has COP in columns (rows), then constructing such sets after applying
-the COP permutation to the rows (columns) of the matrix will result in
-sets with consecutive integers. In other words, after application of
-COP reordering, the sets are intervals. Indeed the problem now becomes
-finding interval assignments to a given set system such that there
-exists a permutation of the universe of set of row indices (column
-indices) which converts each set to its assigned interval.
-
-The problem of interest in this thesis, namely, tree path labeling
-problem, is a natural generalization of the interval assignment
-problem or the COT problem. The problem is defined as follows -- given
-a set system $\cF$ from a universe $U$ and a target tree $T$, does
-there exist a bijection from $U$ to the vertices of $T$ such that each
-set in the system maps to a path in $T$.  We refer to this as the
-{\CFTPL} problem or simply {\em tree path labeling} problem for an
-input set system and target tree pair -- $(\cF,T)$. The special case
-of the target tree being a path, is the interval assignment problem.
-We focus on generalizing the notion of an ICPIA \cite{nsnrs09} to
-characterize feasible path assignments.  We show that for a given set
-system $\cF$, a tree $T$, and an assignment of paths from $T$ to the
-sets, there is a feasible\footnote{The notion of {\em feasibility} is
-  formally defined in Section~\ref{ch:prelims}.}  bijection between
-$U$ and $V(T)$ if and only if the intersection cardinalities among any
-three sets (not necessarily distinct) is equal to that of the
-corresponding paths assigned to them and the input passes a filtering
-algorithm (described in this paper) successfully.  This algorithmic
-characterization gives a natural data structure that stores all the
-\remove{relevant} feasible bijections between $U$ and $V(T)$. This
-reduces the search space for the solution considerably from the
-universe of all possible bijections between $U$ and $V(T)$ to only
-those bijections that maintain the characterization.  Further, the
-filtering algorithm is also an efficient algorithm to test if a tree
-path labeling\footnote{The terms {\em tree path labeling} and {\em
-    tree path assignment} are, in informal language,
-  synonyms. Formally, the former refers to the bijection $\cl: \cF
-  \rightarrow \cP$. The latter refers to the set of ordered pairs
-  $\{(S, P) \mid S \in \cF, P \in \cP\}$. $\cP$ is a set of paths on
-  $T$.} is feasible.
-
-\section{Summary of New Results in this Thesis}
-\label{sec:results}
-
-We see in Section~\ref{sec:motive} that pairwise intersection
-cardinality preservation is necessary and sufficient for an interval
-assignment to be feasible for a given hypergraph\footnote{A {\em
-    hypergraph} is an alternate representation of a set system and
-  will be used in this thesis. See Section~\ref{ch:prelims} for the
-  formal definition.} and thus is a characterization for COP
-\cite{nsnrs09}. \tnote{Refer to the survey section that will list the
-  theorems and/or lemma?} In our work we extend this characterization
-and find that trio-wise intersection cardinality preservation makes a
-tree path labeling \footnote{A {\em tree path labeling} $\cl$ is a
-  bijection of paths from the target tree $T$ to the hyperedges in
-  given hypergraph $\cF$.  See Section~\ref{ch:prelims} for the formal
-  definition.}  (TPL) feasible, which is a generalization of the COP
-problem. This problem is defined as follows.
-
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \FTPL}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
-        paths $\cP$ from $T$ and a
-        bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
-      \end{minipage}\\
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
-        such that $\phi$ when applied on any hyperedge in $\cF$ will
-        give
-        the path mapped to it by the given tree path labeling $\cl$.\\
-        { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every
-          hyperedge $S \in \cF$.}
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
-
-We give a necessary and sufficient condition by way of {\em
-  Intersection Cardinality Preservation Path Labeling} (ICPPL) and a
-filtering algorithm for {\FTPL} to output in affirmative. ICPPL
-captures the trio-wise cardinality property described
-earlier\footnote{See Section~\ref{sec:feasible} for the definition of
-  ICPPL.}. This characterization can be checked in polynomial time.  A
-relevant consequence of this constructive procedure is that it is
-sufficient to iteratively check if three-way intersection
-cardinalities are preserved.  In other words, in each iteration, it is
-sufficient to check if the intersection of any three hyperedges is of
-the same cardinality as the intersection of the corresponding paths.
-Thus this generalizes the well studied question of the feasible
-interval assignment problem which is the special case when the target
-tree $T$ is simply a path \cite{wlh02,nsnrs09}.
-
-Aside from checking if a given TPL is feasible, we also solve the
-problem of computing a feasible TPL for a given hypergraph and target
-tree, if one exists. This problem, {\CFTPL}, is defined as follows.
-
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \CFTPL}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$ and a tree $T$.\\
-      \end{minipage}\\
-
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a set of paths $\cP$ from $T$ and a bijection
-        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
-        {\bf true} on $(\cF, T, \cl)$.
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
-
-We present a polynomial time algorithm for {\CFTPL} when the target
-tree $T$ belongs to a special class of trees called {\em \kstar s} and
-when the hyperedges in the hypergraph $\cF$ have at most $k+2$
-vertices. A couple of examples of {\kstar s} are given in
-Figure~\ref{fig:ksubstar}.
-
-\begin{figure}[t]
-  \centering
-  \begin{tabular}[h]{ccccc}
-    \includegraphics[scale=0.3]{../img/star.pdf} &&&&
-    \includegraphics[scale=0.3]{../img/kstar.pdf}\\
-    (a) &&&& (b)
-  \end{tabular}
-  \caption{\figtabsize Examples of {\kstar s}. (a) $k = 0$ (b) $k = 2$
-  }
-  \label{fig:ksubstar}
-\end{figure}
-
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \CFTPLKTREE}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$ such that every
-        hyperedge
-        $S \in \cF$ is of cardinality at most $k+2$ and a {\kstar} $T$.\\
-      \end{minipage}\\
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a set of paths $\cP$ from $T$ and a bijection
-        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
-        {\bf true} on $(\cF, T, \cl)$.
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
-
-\tnote{The following sentence seems out of place in this para} In
-spite of this being a restricted case, we believe that our results are
-of significant interest in understanding the nature of {\sc Graph
-  Isomorphism} which is polynomial time solvable in interval graphs
-while being hard on path graphs\cite{kklv10}. {\kstar s} are a class
-of trees which are in many ways very close to intervals or paths. Each
-ray\footnote{The path from a leaf to the root, the vertex with highest
-  degree, is called a {\em ray} of the \kstar. See
-  Section~\ref{ch:prelims}.} are independent except for the
-root\footnote{The vertex with maximum degree in a {\kstar} is called
-  {\em root}. See Section~\ref{ch:prelims}.} and hence can be
-considered as an independent interval till the root. Our algorithm
-builds on this fact and uses the interval assignment
-algorithm\cite{nsnrs09} up until ``reaching'' the root and then uses
-the trio-wise intersection cardinality (the extra condition in ICPPL
-that generalizes ICPIA) check to resolve the ambiguity about which ray
-the algorithm should ``grow'' the solution into in the next iteration.
-
-We also have an algorithm for solving {\CFTPL} with no restrictions on
-the target tree or set size which runs in exponential time.  This
-algorithm finds a path labeling from $T$ by decomposing the problem
-into subproblems of finding path labeling of subsets of $\cF$ from
-subtrees of $T$. Given the fact that binary matrices naturally
-represent a set system (see Section~\ref{sec:motive}) and that the
-{\em overlap}\footnote{}\tnote{Give informal definition of ``overlap''
-  in footnote and link to sec:prelims. ADD the definition in prelim.}
-relation between the sets involved is an obvious equivalence relation,
-$\cF$ quite naturally partitions into equivalence classes known as
-{\em overlap components}\tnote{Give informal definition of ``overlap
-  components'' in footnote and link to sec:prelims. ADD the definition
-  in prelim.}. In the context of COP, overlap components were used in
-\cite{wlh02} and \cite{kklv10}. Moreover, \cite{nsnrs09} discovered
-that these equivalence classes form a total order \tnote{Give informal
-  definition of ``total order'' in footnote?}. We extend this to TPL
-and find that when $\cF$ is a path hypergraph\footnote{If there exists
-  an FTPL for a hypergraph $\cF$, it is called a path hypergraph.},
-the classes can be partially ordered as an in-tree in polynomial
-time. Once $\cF$ is ``broken'' into overlap components, one must
-identify the subtree of $T$ that it needs to map to and this is the
-hard part which is currently open to be solved in polynomial time.
-
-\tnote[TH]{The connection of TPL to graph isomorphism will be made
-  later in the document}
-
-\chapter{Consecutive-ones Property - a Survey of Important Results}
-\label{ch:copsurvey}
-\tnote{TBD: have a few lines about organization of chapter}
-
-\tnote{ Survey chapter: is the full fledged expansion of the survey in
-  introduction with details, observations, theorems etc.}
-
-\section{Matrices with COP}
-\label{sec:surveycoptest}
-\tnote{Expand on ref:sec:copmatrices}
-
-\section{Optimization problems in COP}
-\label{sec:surveycopopt}
-\tnote{Expand on ref:sec:optcop}
-
-\section{********* COP in Relational Database Model}
-\label{sec:surveyrdbm}
-\tnote{Expand on sec:apprdbm}
-\tnote{(set systems theme)}
-
-\subsection{********* COP in Graph Isomorphism}
-\label{sec:surveygraphiso}
-\tnote{Expand on sec:appgraphiso}
-\tnote{(canonization theme)}
-
-\subsection{********* Certifying Algorithms}
-\label{sec:surveycertalgo}
-\tnote{Expand on sec:appcertalgo}
-\tnote{ (certification McC04 theme)}
-
-\chapter{Tree Path Labeling of Path Hypergraphs - the New Results}
-\label{ch:myresearch}
-
-\tnote{UPDATE TO FINAL VERSION. BELOW IS OLDER VERSION WITH INCOMPLETE
-  SEC 4.}  This chapter documents all the new results obtained by us in the
-area of tree path labeling of path hypergraphs which is the parent
-problem addressed in this thesis. In Section~\ref{sec:motive} we see
-that consecutive-ones property and its equivalent problem of interval
-labeling of a hypergraph is a special case of the general problem of tree
-path labeling of path hypergraphs.
-
-\tnote{UPDATE the following para to current skeleton.}The necessary
-preliminaries with definitions etc. are presented in
-Section~\ref{ch:prelims}. Section~\ref{sec:feasible} documents the
-characterization of a feasible path labeling for any path labeling
-from any target trees. Section~\ref{sec:spltargettree} describes the
-special case where the target tree is of a particular family of trees
-and specifically, Section~\ref{sec:ksubdivstar} describes a polynomial
-time algorithm to find the tree path labeling of a given set system
-from a given $k$-subdivided tree. Section~\ref{sec:norestraint}
-describes the general situation where the target tree has no
-restrictions and the algorithm to find a TPL, if any, in this case.
-
-\section{Introduction}
-
-The problem of consecutive-ones property testing can be easily seen as
-a simple constraint satisfaction problem involving a hypergraph or a
-system of sets from a universe. Every column (row) of the binary
-matrix can be converted into a set of non-negative integers which are
-the indices of the rows (columns) with {\un}s in that column
-(row). When observed in this context, if the matrix has the COP on
-columns (rows), a reordering of its rows (columns) will result in sets
-that have only consecutive integers. In other words, the sets after
-applying the COP row (column) permutation are intervals.  In this
-form, one can see that this is indeed the problem of finding interval
-assignments to the given set system \cite{nsnrs09} with a single
-permutation of the universe (set of row or column indices for COP of
-columns or rows, respectively) which permutes each set to its
-interval. The result in \cite{nsnrs09} characterizes interval
-assignments to the sets which can be obtained from a single
-permutation of the rows (columns). They show that for each set, the
-cardinality of the interval assigned to it must be same as the
-cardinality of the set, and the intersection cardinality of any two
-sets must be same as the intersection cardinality of the corresponding
-intervals. This is a necessary and sufficient
-condition. \tnote{(ICPIA) put reference to survey chapter}.  Finally,
-the idea of decomposing a given binary matrix into prime matrices to
-check for COP is adopted from cite|wlh02| to test if an ICPIA exists
-for a given set system.\tnote{REPEATITION from intro chapter. but idea
-  recall is necessary. so keep it but reword it to avoid monotone.}
-
-A natural generalization of the interval assignment problem is
-feasible tree path labeling problem of a set system. The problem is
-defined as follows -- given a set system $\cF$ from a universe $U$ and
-a tree $T$, does there exist a bijection from $U$ to the vertices of
-$T$ such that each set in the system maps to a path in $T$.  We refer
-to this as the {\em tree path labeling problem} for an input set
-system, target tree pair -- $(\cF,T)$. \tnote{USE THE FORMAL PROBLEM DEF BLOCK}
-%\begin{tabular}[t]{c}
-{\small
-\begin{minipage}[h]{5in}
- % \begin{singlespace}
- \vspace{2mm}
- {\large \FTPL}\\
- \begin{tabular}[t]{l|l}
- \hline\\
-    {\tt Input} & 
-    \begin{minipage}[t]{\probdefwidth}
-      A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
-      paths $\cP$ from $T$ and a
-      bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
-    \end{minipage}\\
-    {\tt Question} &
-    \begin{minipage}[t]{\probdefwidth}
-      Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
-      such that $\phi$ when applied on any hyperedge in $\cF$ will
-      give
-      the path mapped to it by the given tree path labeling $\cl$.\\
-      { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every hyperedge
-        $S \in \cF$.}
-    \end{minipage}\\
-  \end{tabular}
-  % \end{singlespace}
-\end{minipage}\\
-}
-%\end{tabular}
-
-As a special case if the tree
-$T$ is a path, the problem becomes the interval assignment problem. \tnote{USE THE FORMAL PROBLEM DEF BLOCK}
-We focus on the question of generalizing the notion of an ICPIA
-\cite{nsnrs09} to characterize feasible path assignments.  We show
-that for a given set system $\cF$, a tree $T$, and an assignment of
-paths from $T$ to the sets, there is a feasible bijection between $U$
-and $V(T)$ if and only if all intersection cardinalities among any
-three sets (not necessarily distinct) is same as the intersection
-cardinality of the paths assigned to them and the input runs a
-filtering algorithm (described in this paper) successfully.  This
-characterization is proved constructively and it gives a natural data
-structure that stores all the relevant feasible bijections between $U$
-and $V(T)$.  Further, the filtering algorithm is also an efficient
-algorithm to test if a tree path labeling to the set system is
-feasible.  This generalizes the result in \cite{nsnrs09}.
-
-In the later part of this paper, we focus on a new special case of the
-tree path labeling problem. Here the set system is such that every set
-is at most $k+2$ in size and for every pair of sets in it there exists
-a sequence of sets between them with consecutive sets in this sequence
-having a strict intersection -- i.e., non-empty intersection with
-neither being contained in the other. Moreover, the given tree is a
-{\em $k$-subdivided star}. We demonstrate a polynomial time algorithm
-to find a feasible path labeling in this case.
-
-\section{Preliminaries to new results}
-\label{ch:prelims}
-
-\tnote{TBD: have a few lines about organization of chapter}This
-section states definitions and basic facts necessary in the scope of
-this document.
-
-\tnote{Move some of this to general prelim in chapter 1}
-
-The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
-  system} of a universe $U$ with $|U| = n$.  The {\em support} of a
-set system $\F$ denoted by $supp(\cF)$ is the union of all the sets in
-$\F$; $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this
-paper, a set system is required to ``cover'' the universe; $
-supp(\cF) = U$.
-
-The graph $T$ represents a {\em target tree} with same
-number of vertices as elements in $U$; $|V(T)| = n$.  A {\em path
-  system}\, $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P
-\mid P \subseteq V, \text{ } T[P] \text{ is a path} \}$.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Main body Chapters
 %
-% A set system $\cF$ can be alternatively represented by a {\em
-%   hypergraph} $\H_\cF$ whose vertex set is $supp(\cF)$ and
-% hyperedges are the sets in $\cF$. This is a known representation for
-% interval systems in literature \cite{bls99,kklv10}.  We extend this
-% definition here to path systems.
-
-A set system $\cF$ can be alternatively represented by a {\em
-  hypergraph}\, $\cF_H$ whose vertex set is $supp(\cF)$ and hyperedges
-are the sets in $\cF$. This is a known representation for interval
-systems in literature \cite{bls99,kklv10}.  We extend this definition
-here to path systems. Due to the equivalence of set system and
-hypergraph in the scope of this paper, we drop the subscript $_H$ in
-the notation and refer to both the structures by $\cF$.
-
-Two hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to
-each other, denoted by $\cF' \cong \cF''$, iff there exists a
-bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for all
-sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ iff $B$ is
-a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
-\cite{kklv10}. This is called {\em hypergraph
-  isomorphism}. \tnote{also extend $\phi$ to hyperedges -- see if
-  required }
-
-The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph $\cF$ is a
-graph such that its vertex set has a bijection to $\cF$ and there
-exists an edge between two vertices iff their corresponding hyperedges
-have a non-empty intersection \cite{mcg04}.
-
-If the intersection graphs of two hypergraphs are isomorphic,
-$\bI(\cF) \cong \bI(\cP)$ where $\cP$ is also a path system, then the
-bijection $\cl: \cF \rightarrow \cP$ due to this isomorphism is called
-a {\em path labeling} of the hypergraph $\cF$. To illustrate further,
-let $\cg: V(\cF) \rightarrow V(\cP)$ be the above mentioned
-isomorphism where $V(\cF)$ and $V(\cP)$ are the vertex sets that
-represent the hyperedges for each hypergraph respectively, $V(\cF) =
-\{ v_S \mid S \in \cF\}$ and $V(\cP) = \{ v_P \mid P \in \cP\}$. Then
-the path labeling $\cl$ is defined as follows: $\cl(S_1) = P_1$ iff
-$\cg (v_{S_1}) = v_{P_1}$. The path system $\cP$ may be alternatively
-denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In most scenarios in
-this paper, what is given are the pair $(\cF, \cl)$ and the target
-tree $T$; hence this notation will be used more often.
-
-If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$
-is called a {\em path hypergraph} and $\cP$ is called {\em path
-  representation} of $\cF$. If this isomorphism is $\phi: supp(\cF)
-\rightarrow V(T)$, then it is clear that there is an {\em induced path
-labeling} $\cl_\phi: \cF \rightarrow \cP$ to the set system;
-$\cl_\phi(S) = \set{y \mid y = \phi(x), x \in S}$ for all $S \in \cF$. Recall that
-$supp(\cP) = V(T)$.
-
-A path labeling $(\cF, \cl)$ is defined to be {\em
-  feasible} if
-$\cF \cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
-\rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
-\rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.
-
-An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is
-a graph such that its vertex set has a bijection to $\cF$ and there
-exists an edge between two of its vertices iff their corresponding
-hyperedges overlap. Two hyperedges $S$ and $S'$ are said to {\em
-  overlap}, denoted by $S \overlap S'$, if they have a non-empty
-intersection and neither is contained in the other; $S \overlap S'
-\text{ iff } S \cap S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq
-S$. Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$ and not
-necessarily connected. Each connected component of $\bO(\cF)$ is
-called an {\em overlap component}.
-
-A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
-\overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
-\cite{kklv10}. Additionally, if $S$ is such that it is contained in no
-other hyperedge in $\cF$, i.e., it is inclusion maximal then it is called
-{\em super-marginal}.
-
-A {\em star} graph is a complete bipartite graph $K_{1,p}$
-which is clearly a tree and $p$ is the number of leaves. The vertex
-with maximum degree is called the {\em center} of the star and the
-edges are called {\em rays} of the star. A {\em $k$-subdivided star}
-is a star with all its rays subdivided exactly $k$ times. The
-definition of a {\em ray of a $k$-subdivided star} is extended to the path
-from the center to a leaf. It is clear that all rays are of length $k+2$.
-
-\section{Characterization of Feasible Tree Path  Labeling} 
-\label{sec:feasible}
-
-In this section we give an algorithmic characterization of a
-feasibility of tree path labeling.  Consider a path labeling $(\cF,
-\cl)$ on the given tree $T$. We call $(\cF, \cl)$ an {\em Intersection
-  Cardinality Preserving Path Labeling (ICPPL)} if it has the
-following properties.
-
-\begin{enumerate}[{(\icpplpr\ }i) \ \ \ ]
-\item \label{pr:i} $|S| = |\cl(S)|$ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $S \in \cF$
-  % \vspace{\topshrink}
-\item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ \ \ \ \
-  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all distinct $S_1, S_2 \in \cF$
-  % \vspace{\topshrink}
-\item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
-  \cap \cl(S_3)|$ \ \ \ for all distinct $S_1, S_2, S_3 \in \cF$
-\end{enumerate}
-
-The following lemma is useful in subsequent arguments. 
-\begin{lemma}
-  \label{lem:setminuscard}
-  If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
-  (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
-  \cl(S_3))|$.
-\end{lemma}
-\begin{proof}%[Proof of Lemma~\ref{lem:setminuscard}]
-  Let $P_i = \cl(S_i)$, for all $1 \le i \le  3$.
-  $|S_1 \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| =
-  |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to properties (ii)
-  and (iii) of ICPPL, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
-  \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
-  = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. \qed
-\end{proof}
-
-In the remaining part of this section we show that $(\cF,
-\cl)$ is feasible if and only if it is an ICPPL and
-Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
-function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two levels of
-filtering of $(\cF, \cl)$ to make it simpler while retaining the set
-of isomorphisms, if any, between $\cF$ and $\cF^\cl$.
-% One direction of this claim isclear: that if a path
-% labeling is feasible, then all intersection cardinalities are
-% preserved, i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
-% \annote{has no premature exit condition hence any input will go
-%   through it}{Prove that the filtered sets has ICPPL iff input PL
-%   has ICPPL?}. Algorithm~\ref{leafasgn} has an exit condition at
-% line~\ref{xempty}. It can be easily verified that $X$ cannot be
-% empty if $\cl$ is a feasible path labeling. The reason is that a
-% feasible path labeling has an associated bijection between
-% $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that
-% the sets map to paths, ``preserving'' the path labeling.  The rest
-% of the section is devoted to constructively proving that it is
-% sufficient for a path labeling to be an ICPPL and pass the two
-% filtering algorithms.  To describe in brief, the constructive
-% approaches refine an ICPPL iteratively, such that at the end of each
-% iteration we have a ``filtered'' path labeling, and finally we have
-% a path labeling that defines a family of bijections from $supp(\cF)$
-% to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.
-First, we present Algorithm~\ref{perms} or {\tt filter\_1}, and prove
-its correctness.  This algorithm refines the path labeling by
-processing pairs of paths in $\cF^\cl$ that share a leaf until no two
-paths in the new path labeling share any leaf.
-
-\begin{algorithm}[h]
-  \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
-  \label{perms}
-  \begin{algorithmic}[\lndisplay]
-    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$\\
-    \STATE $j \assign 1$\\
-    \WHILE {there is $S_1, S_2 \in \cF_{j-1}$ such that
-      $\cl_{j-1}(S_1)$ and $\cl_{j-1}(S_2)$ have a common leaf in
-      $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
-    \{S_1, S_2\})
-    \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1 \}$ \label{setbreak} 
-    \COMMENT {Remove $S_1$, $S_2$ and add the ``filtered'' sets}
-    \STATE {\bf for} every $S \in \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
-    S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\
-
-    \STATE $\cl_j(S_1 \cap S_2) \assign \cl_{j-1}(S_1) \cap
-    \cl_{j-1}(S_2)$
-    \COMMENT {Carry forward the path labeling for all existing sets other than
-      $S_1$, $S_2$}
-    \STATE $\cl_j(S_1 \setminus S_2) \assign \cl_{j-1}(S_1) \setminus
-    \cl_{j-1}(S_2)$ 
-    \COMMENT {Define path labeling for new sets}
-    \STATE $\cl_j(S_2 \setminus S_1) \assign \cl_{j-1}(S_2) \setminus
-    \cl_{j-1}(S_1)$
-
-    \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of ICPPL}
-    \label{ln:3waycheck}
-    \STATE {\bf exit} \label{ln:exit1} \\
-    \ENDIF
-
-    \STATE $j \assign j+1$\\
-    \ENDWHILE
-    \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
-    \RETURN $(\cF', \cl')$
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma} 
- \label{lem:feasible} 
- In Algorithm~\ref{perms}, if input $(\cF, \cl)$ is a feasible path
- assignment then at the end of $j$th iteration of the {\bf while}
- loop, $j \ge 0$, $(\cF_j, \cl_j)$ is a feasible path assignment.
-\end{lemma}
-\begin{proof}%[Proof of Lemma~\ref{lem:feasible}]
-  We will prove this by mathematical induction on the number of
-  iterations. The base case $(\cF_0, \cl_0)$ is feasible since it is
-  the input itself which is given to be feasible. Assume the lemma is
-  true till $j-1$th iteration. i.e. every hypergraph isomorphism
-  $\phi: supp\left(\cF_{j-1}\right) \rightarrow V\left(T \right)$ that
-  defines $(\cF, \cl)$'s feasibility, is such that the induced path
-  labeling on $\cF_{j-1}$, $\cl_{\phi[{\cF_{j-1}}]}$ is equal to
-  $\cl_{j-1}$. We will prove that $\phi$ is also the bijection that
-  makes $(\cF_j, \cl_j)$ feasible. Note that $supp(\cF_{j-1}) =
-  supp(\cF_{j})$ since the new sets in $\cF_j$ are created from basic
-  set operations to the sets in $\cF_{j-1}$. For the same reason and
-  $\phi$ being a bijection, it is clear that when applying the $\phi$
-  induced path labeling on $\cF_j$, $ \cl_{\phi[{\cF_{j}}]}(S_1
-  \setminus S_2) = \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
-  \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Now observe that $ \cl_j(S_1
-  \setminus S_2) = \cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2) =
-  \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
-  \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Thus the induced path labeling
-  $\cl_{\phi[{\cF_{j}}]} = \cl_{j}$. Therefore lemma is proven.  \qed
-\end{proof}
-
-\begin{lemma}
-  \label{lem:invar1} In Algorithm~\ref{perms}, at the end of $j$th
-  iteration, $j \ge 0$, of the {\bf while} loop, the following
-  invariants are maintained.
-  \begin{enumerate}[I {\ }] %\vspace{\topshrink}
-  \item $\cl_j(R)$ is a path in $T$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ for all $R \in \cF_j$%\vspace{\topshrink}
-  \item $|R| = |\cl_j(R)|$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $R \in
-    \cF_j$%\vspace{\topshrink}
-  \item $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, \ \ \ \ \ \ \ \ \ \
-    \ \ \ \ \ \ \ \ \ \ for all $R, R' \in \cF_j$%\vspace{\topshrink}
-  \item $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
-    \cl_j(R'')|$, \ \ \ for all $R, R', R'' \in \cF_j$
-  \end{enumerate}
-\end{lemma}
-
-\begin{proof}
- Proof is by induction on the number of iterations, $j$. In this
-  proof, the term ``new sets'' will refer to the sets added to $\cF_j$
-  in $j$th iteration in line~\ref{setbreak} of Algorithm~\ref{perms},
-  $S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1$ and its
-  images in $\cl_j$ where $\cl_{j-1}(S_1)$
-  and $\cl_{j-1}(S_2)$ intersect and share a leaf.\\
-  The invariants are true in the base case $(\cF_0, \cl_0)$, since it
-  is the input ICPPL.  Assume the lemma is true till the $j-1$th
-  iteration. Let us consider the possible cases for each of the above invariants for
-  the $j$th iteration.
-
- \begin{enumerate}[\xbullet]
-  \item {\em Invariant} I/II
-    \begin{enumerate}[{I/II}a $|$] % \textbullet 
-    \item {\em $R$ is not a new set.} It is in $\cF_{j-1}$. Thus
-      trivially true by induction hypothesis.
-    \item {\em $R$ is a new set.} If $R$ is in $\cF_{j}$ and not in
-      $\cF_{j-1}$, then it must be one of the new sets added in
-      $\cF_j$. In this case, it is clear that for each new set, the
-      image under $\cl_j$ is a path since by definition the chosen
-      sets $S_1$, $S_2$ are from $\cF_{j-1}$ and due to the while loop
-      condition, $\cl_{j-1}(S_1)$, $\cl_{j-1}(S_2)$ have a
-      common leaf. Thus invariant I is proven.\\
-      Moreover, due to induction hypothesis of invariant III and the
-      definition of $l_j$ in terms of $l_{j-1}$, invariant II is
-      indeed true in the $j$th iteration for any of the new sets.  If
-      $R = S_1 \cap S_2$, $|R| = |S_1 \cap S_2| = |\cl_{j-1}(S_1) \cap
-      \cl_{j-1}(S_2)| = |\cl_j(S_1 \cap S_2)| = |\cl_j(R)|$.
-      If $R = S_1 \setminus S_2$, $|R| = |S_1 \setminus S_2| = |S_1| -
-      |S_1 \cap S_2| = |\cl_{j-1}(S_1)| - |\cl_{j-1}(S_1) \cap
-      \cl_{j-1}(S_2)| = |\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)| =
-      |\cl_j(S_1 \setminus S_2)|
-      = |\cl_j(R)|$. Similarly if $R = S_2 \setminus S_1$.\\
-    \end{enumerate}
-  \item {\em Invariant} III
-    \begin{enumerate}[{III}a $|$]
-    \item {\em $R$ and $R'$ are not new sets.} It is in
-      $\cF_{j-1}$. Thus trivially true by induction hypothesis.
-    \item {\em Only one, say $R$, is a new set.} Due to invariant IV
-      induction hypothesis, Lemma~\ref{lem:setminuscard} and
-      definition of $\cl_j$, it follows that invariant III is true no
-      matter which of the new sets $R$ is equal to. If $R = S_1 \cap
-      S_2$, $|R \cap R'| = |S_1 \cap S_2 \cap R'| = |\cl_{j-1}(S_1)
-      \cap \cl_{j-1}(S_2) \cap \cl_{j-1}(R')| = |\cl_j(S_1 \cap S_2)
-      \cap \cl_j(R')| = |\cl_j(R) \cap \cl_j(R')|$.  If $R = S_1
-      \setminus S_2$, $|R \cap R'| = |(S_1 \setminus S_2) \cap R'| =
-      |(\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)) \cap \cl_{j-1}(R')| =
-      |\cl_{j}(S_1 \cap S_2) \cap \cl_{j}(R')| = |\cl_{j}(R) \cap
-      \cl_{j}(R')|$. Similarly, if $R = S_2 \setminus
-      S_1$. Note $R'$ is not a new set.\\
-
-    \item {\em $R$ and $R'$ are new sets.} By definition, the new
-      sets and their path images in path label $\cl_j$ are disjoint so
-      $|R \cap R'| = |\cl_j(R) \cap \cl_j(R)| = 0$. Thus case proven.
-    \end{enumerate}
-  \item {\em Invariant} IV
-
-    Due to the condition in line~\ref{ln:3waycheck}, this invariant is
-    ensured at the end of every iteration.
-%     \begin{enumerate} [{Case 3.}1:]
-%     \item {\em $R$, $R'$ and $R''$ are not new sets.} Trivially
-%       true by induction hypothesis.
-%     \item {\em Only one, say $R$, is a new set.}
-%       If $R = S_1 \cap S_2$,  from Lemma~\ref{lem:fourpaths} and
-%       invariant III hypothesis,  this case is proven. Similarly if $R$
-%       is any of the other new  sets, the case is proven by also using
-%       Lemma ~\ref{lem:setminuscard}.
-%     \item {\em At least two of $R, R', R''$ are new sets.}
-%       The new sets are disjoint hence this case is vacuously true.
-%     \end{enumerate}
-  \end{enumerate} \qed
-%\vspace{-6mm}
-
-\end{proof}
-
-\begin{lemma}
-  \label{lem:noexit1}
-  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{perms} is
-  feasible, then the set of hypergraph isomorphism functions that
-  defines $(\cF, \cl)$'s feasibility is the same as the set that
-  defines $(\cF_j, \cl_j)$'s feasibility, if any.  Secondly, for any
-  iteration $j > 0$ of the {\em \bf while} loop, the {\em \bf exit}
-  statement in line~\ref{ln:exit1} will not execute.
-\end{lemma}
-\begin{proof}
-  Since $(\cF,\cl)$ is feasible, by Lemma~\ref{lem:feasible}
-  $(\cF_j,\cl_j)$ for every iteration $j > 0$ is feasible.  % Therefore,
-%   every hypergraph isomorphism $\phi: supp(\cF) \rightarrow V(T)$ that
-%   induces $\cl$ on $\cF$ also induces $\cl_{j-1}$ and $\cl_{j}$ on
-%   $\cF_{j-1}$ and $\cF_{j}$ respectively, i.e., $\cl_{\phi[\cF_{j-1}]}
-%   = \cl_{j-1}$ and $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
-%   that for all $x \in supp(\cF)$, for all $v \in V(T)$ the following
-%   hold true.
-  Also, every hypergraph isomorphism $\phi: supp(\cF) \rightarrow
-  V(T)$ that induces $\cl$ on $\cF$ also induces $\cl_{j}$ on
-  $\cF_{j}$, i.e., $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
-  that for all $x \in supp(\cF)$, for all $v \in V(T)$, if $(x,v) \in
-  \phi$ then $v \in \cl_{j}(S)$ for all $S \in \cF_{j}$ such that $x
-  \in S$.
-% the following
-%   hold true.
-%   \begin{enumerate}[i. ]
-%   \item If $(x,v) \in \phi$ then $v \in \cl_{j-1}(S)$ for all $S \in
-%     \cF_{j-1}$ such that $x \in S$.
-%   \item If $(x,v) \in \phi$ then $v \in \cl_{j}(S)$ for all $S \in
-%     \cF_{j}$ such that $x \in S$
-%   \end{enumerate}
-  In other words, filter 1 outputs a filtered path labeling that
-  ``preserves''
-  hypergraph isomorphisms of the original path labeling.\\
-  Secondly, line~\ref{ln:exit1} will execute iff the exit condition in
-  line~\ref{ln:3waycheck}, i.e. failure of three way intersection
-  preservation, becomes true in any iteration of the {\em \bf while}
-  loop.  Due to Lemma~\ref{lem:invar1} Invariant IV, the exit
-  condition does not occur if the input is a feasible ICPPL.\qed
-
-%   such that $\phi(x) = v$ where $v$ is the leaf considered in the
-%   first iterations of while. Clearly, $\phi$ is a renaming of
-%   vertices in hypergraph $\cF$ to those in hypergraph $\cF^\cl$. Thus
-%   the following facts can be observed in every iteration of the loop.
-
-%   \begin{enumerate}[\hspace{2mm}i. ] \vspace{\topshrink}
-%   \item all intersection cardinalities are preserved in this path
-%     labeling \vspace{\topshrink}
-%   \item element $x$ is exclusive in a hyperedge in $\cF$ since $v$ is
-%     exclusive in a hyperedge in $\cF^\cl$.
-%   \end{enumerate}
-
-%   Thus the exit condition is never rendered true after $x$ and $v$ are
-%   removed from their respective hyperedges. \qed
-
-% \noindent
-% This proof uses mathematical induction on the number
-%   of iterations $j$, $j \ge 0$, of the loop that executed
-%   without exiting. The base case, $j = 0$ is obviously true since the
-%   input is an ICPPL and the exit condition cannot hold true due to
-%   ICPPL property (iii).  Assume the algorithm executes till the end
-%   of $j-1$th iteration without exiting at line
-%  ~\ref{ln:3waycheck}. Consider the $j$th iteration. From Lemma
-%  ~\ref{lem:feasible} we know that $(\cF_j, \cl_j)$ and $(\cF_{j-1},
-%   \cl_{j-1})$ are feasible\remove[AS]{and from the proof in lemma
-%     lem:invar1 we know that $(\cF_{j-1}, \cl_{j-1})$ satisfies all the
-%     invariants defined in the lemma}.  Thus there exists a bijection
-%   $\phi: supp(\cF) \rightarrow V(T)$ such that the induced path
-%   % labeling on $\cF_{j-1}$ $\cl_{\phi[\cF_{j-1}]} = \cl_{j-1}$.
-%   labeling on $\cF_{j}$, $\cl_{\phi[\cF_{j}]}$ and on $\cF_{j-1}$,
-%   $\cl_{\phi[\cF_{j-1}]}$ are equal to $\cl_{j}$ and $\cl_{j-1}$
-%   respectively.  We need to prove that for any $R, R', R'' \in
-%   \cF_{j}$, $|R \cap R' \cap R''| = |\cl_j(R) \cap \cl_j(R') \cap
-%   \cl_j(R'')|$.
-%   The following are the possible cases that could arise. From argument
-%   above, $|\cl_j(R) \cap \cl_j(R') \cap \cl_j(R'')| =
-%   |\cl_{\phi[\cF_{j}]}(R) \cap \cl_{\phi[\cF_{j}]} (R') \cap
-%   \cl_{\phi[\cF_{j}]} (R'')|$
-
-%   \begin{enumerate}[a $|$]
-%   \item {\em None of the sets are new. $R, R', R'' \in \cF_{j-1}$.}
-%     We know $(\cF_{j-1}, \cl_{j-1})$ is feasible. Thus $|R \cap R'
-%     \cap R''| = |\cl_{j-1}(R) \cap \cl_{j-1}(R') \cap \cl_{j-1}(R'')|
-%     = |\cl_{j}(R) \cap \cl_{j}(R') \cap \cl_{j}(R'')|$.
-%   \item {\em Only one, say $R$, is a new set.}  Let $R = S_1 \cap S_2$
-%     ($S_1, S_2$ are defined in the proof of lemma
-%    ~\ref{lem:invar1}). Now we have $|R \cap R' \cap R''| = |S_1 \cap
-%     S_2 \cap R' \cap R''| = |\cl_{j-1}(S_1) \cap \cl_{j-1}(S_2) \cap
-%     \cl_{j-1}(R') \cap \cl_{j-1}(R'')| = |\cl_{j}(R) \cap \cl_{j}(R')
-%     \cap \cl_{j}(R'')|$. Thus proven. If $R$ is any of the other new
-%     sets, the same claim can be verified using lemma
-%    ~\ref{lem:setminuscard}.
-%     % \item []{\bf Case 3:}
-%   \item {\em At least two of $R, R', R''$ are new sets.}  The new sets
-%     are disjoint hence this case is vacuously true.
-%   \end{enumerate}
-%   \qed
-%   \tnote[E2]{remove the induction proof. just text saying x and v are
-%     exclusive in these sets therefore the intersection cardinalities
-%     don't change thus all invariants are still true} 
-\end{proof}
-
-As a result of Algorithm~\ref{perms} each leaf $v$ in $T$
-is such that there is exactly one set in $\cF$ with $v$ as a vertex in
-the path assigned to it.  In Algorithm~\ref{leafasgn} we identify
-elements in $supp(\cF)$ whose images are leaves in a hypergraph
-isomorphism if one exists.  Let $S \in \cF$ be such that $\cl(S)$ is a
-path with leaf and $v \in V(T)$ is the unique leaf incident on it.  We
-define a new path labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) =
-\set{v}$ where $x$ an arbitrary element from $S \setminus \bigcup_{\hS
-  \ne S} \hS$. In other words, $x$ is an element present in no other
-set in $\cF$ except $S$. This is intuitive since $v$ is present in no
-other path image under $\cl$ other than $\cl(S)$.  The element $x$ and
-leaf $v$ are then removed from the set $S$ and path $\cl(S)$
-respectively. After doing this for all leaves in $T$, all path images
-in the new path labeling $\cl_{new}$ except leaf labels (a path that
-has only a leaf is called the {\em leaf label} for the corresponding
-single element hyperedge or set) are paths from a new pruned tree $T_0
-= T \setminus \{v \mid v \text{ is a leaf in }
-T\}$. Algorithm~\ref{leafasgn} is now presented with details.
-
-\begin{algorithm}[h]
-  \caption{Leaf labeling from an ICPPL {\tt filter\_2($\cF, \cl, T$)}}
-  \label{leafasgn}
-  \begin{algorithmic}[\lndisplay]
-    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$
-    \COMMENT {Path images are such that no two path images share a
-      leaf.}
-    \STATE $j \assign 1$\\
-    \WHILE {there is a leaf $v$ in $T$ and a unique $S_1 \in
-      \cF_{j-1}$ such that $v \in \cl_{j-1}(S_1)$ }\label{uniqueleaf}
-    \STATE $\cF_j \assign \cF_{j-1} \setminus \{S_1\}$\\
-    \STATE for all $S \in \cF_{j-1}$ such that $S \ne S_1$ set
-    $\cl_j(S) \assign
-    \cl_{j-1}(S)$\\
-    \STATE $X \assign S_1 \setminus \bigcup_{S \in \cF_{j-1}, S \ne S_1}S$\\
-    \IF{$X$ is empty} \label{xempty} \STATE {\bf exit} \label{ln:exit2} \ENDIF
-    \STATE $x \assign $ arbitrary element from $X$\\
-    \STATE $\cF_j \assign \cF_j \cup \{\{x\}, S_1 \setminus \{x\}\} $\\
-    \STATE $\cl_j(\{x\}) \assign \{v\}$\\
-    \STATE $\cl_j(S_1 \setminus \{x\}) \assign \cl_{j-1}(S_1) \setminus \{v\}$\\
-    \STATE $j \assign j+1$\\
-    \ENDWHILE
-    \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
-    \RETURN $(\cF', \cl')$
-  \end{algorithmic}
-\end{algorithm}
-
-Suppose the input ICPPL $(\cF, \cl)$ is feasible, yet set $X$ in
-Algorithm~\ref{leafasgn} is empty in some iteration of the {\bf while}
-loop. This will abort our procedure of finding the hypergraph
-isomorphism. The following lemma shows that this will not happen.
-
-\begin{lemma}
-  \label{lem:xnotempty}
-  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{leafasgn} is
-  feasible, then for all iterations $j > 0$ of the {\em \bf while}
-  loop, the {\em \bf exit} statement in line~\ref{ln:exit2} does not
-  execute.
-\end{lemma}
-\begin{proof}
-  Assume $X$ is empty for some iteration $j > 0$. We know that $v$ is
-  an element of $\cl_{j-1}(S_1)$. Since it is uniquely present in
-  $\cl_{j-1}(S_1)$, it is clear that $v \in \cl_{j-1}(S_1) \setminus
-  \bigcup_{(S \in \cF_{j-1}) \wedge (S \ne S_1)}\cl_{j-1}(S)$.  Note
-  that for any $x \in S_1$ it is contained in at least two sets due to
-  our assumption about cardinality of $X$. Let $S_2 \in \cF_{j-1}$ be
-  another set that contains $x$. From the above argument, we know $v
-  \notin \cl_{j-1}(S_2)$. Therefore there cannot exist a hypergraph
-  isomorphism bijection that maps elements in $S_2$ to those in
-  $\cl_{j-1}(S_2)$. This contradicts our assumption that the input is
-  feasible. Thus $X$ cannot be empty if input is ICPPL and feasible.
-  \qed
-\end{proof}
-
-\begin{lemma}
-  \label{lem:invar3}
-  In Algorithm~\ref{leafasgn}, for all $j > 0$, at the end of the
-  $j$th iteration of the {\bf while} loop the four invariants given in
-  Lemma~\ref{lem:invar1} hold.
-\end{lemma}
-\begin{proof}
-  By Lemma~\ref{lem:xnotempty} we know that set $X$ will not be empty
-  in any iteration of the {\em \bf while} loop if input ICPPL $(\cF,
-  \cl)$ is feasible and $\cl_j$ is always computed for all $j >
-  0$. Also note that removing a leaf from any path keeps the new path
-  connected. Thus invariant I is obviously true. In every iteration $j
-  > 0$, we remove exactly one element $x$ from one set $S$ in $\cF$
-  and exactly one vertex $v$ which is a leaf from one path
-  $\cl_{j-1}(S)$ in $T$. This is because as seen in
-  Lemma~\ref{lem:xnotempty}, $x$ is exclusive to $S$ and $v$ is
-  exclusive to $\cl_{j-1}(S)$. Due to this fact, it is clear that the
-  intersection cardinality equations do not change, i.e., invariants
-  II, III, IV remain true. On the other hand, if the input ICPPL is
-  not feasible the invariants are vacuously true. \qed
-\end{proof}
-
-% \textcolor{cyan}{
-% \begin{lemma}
-%   \label{lem:notfeasibleexit}
-%   \tnote{IS THIS CORRECT?}  If input ICPPL $(\cF, \cl)$ is not
-%   feasible, then in one of the recursive calls to Algorithm --3--, the
-%   {\em \bf exit} statement in line x in Algorithm --1 or line y in
-%   Algorithm --2 will get executed.
-% \end{lemma}
-% }
-
-We have seen two filtering algorithms above, namely,
-Algorithm~\ref{perms} {\tt filter\_1} and Algorithm~\ref{leafasgn}
-{\tt filter\_2} which when executed serially repectively result in a
-new ICPPL on the same universe $U$ and tree $T$. We also proved that
-if the input is indeed feasible, these algorithms do indeed output the
-filtered ICPPL. Now we present the algorithmic characterization of a
-feasible tree path labeling by way of Algorithm~\ref{al:icppl-find-isomorph}.
-
-Algorithm~\ref{al:icppl-find-isomorph} computes a
-hypergraph isomorphism $\phi$ recursively using Algorithm~\ref{perms}
-and Algorithm~\ref{leafasgn} and pruning the leaves of the input
-tree. In brief, it is done as follows. Algorithm~\ref{leafasgn} gives
-us the leaf labels in $\cF_2$, i.e., the elements in $supp(\cF)$ that
-map to leaves in $T$, where $(\cF_2, \cl_2)$ is the output of
-Algorithm~\ref{leafasgn}. All leaves in $T$ are then pruned away. The
-leaf labels are removed from the path labeling $\cl_2$ and the
-corresponding elements are removed from the corresponding sets in
-$\cF_2$. This tree pruning algorithm is recursively called on the
-altered hypergraph $\cF'$, path label $\cl'$ and tree $T'$. The
-recursive call returns the bijection $\phi''$ for the rest of the
-elements in $supp(\cF)$ which along with the leaf labels $\phi'$ gives
-us the hypergraph isomorphism $\phi$.  The following lemma formalizes
-the characeterization of feasible path labeling.
-
-\begin{algorithm}[h]
-  \caption{{\tt get- hypergraph- isomorphism ($\cF, \cl, T$)}}
-  \label{al:icppl-find-isomorph}
-  \begin{algorithmic}[\lndisplay]
-
-    \IF{$T$ is empty}
-    \RETURN $\emptyset$\\
-    \ENDIF
-    \STATE $L \assign \{v \mid v \text{ is a leaf in }      T\}$\\
-    \STATE $(\cF_1, \cl_1) \assign$ {\tt filter\_1($\cF, \cl,
-      T$)}\\
-    \STATE $(\cF_2, \cl_2) \assign$ {\tt filter\_2($\cF_1,
-      \cl_1, T$)}\\
-
-    \STATE $(\cF', \cl') \assign (\cF_2, \cl_2)$\\
-    \STATE $\phi' \leftarrow \emptyset$
-
-    \FOR {every $v \in L$}
-    \STATE $\phi'(x) \assign v$ where $x \in \cl_2^{-1}(\{v\})$
-    \COMMENT {Copy the leaf labels to a one to one function $\phi':
-      supp(\cF) \rightarrow L$
-      }\\
-    \STATE Remove $\{x\}$ and $\{v\}$ from $\cF'$, $\cl'$  appropriately\\
-    \ENDFOR
-
-    \STATE $T' \assign T \setminus L$
-
-    \STATE $\phi'' \assign$ {\tt
-      get-hypergraph-isomorphism($\cF', \cl', T'$)}
-    \STATE $\phi \assign \phi'' \cup \phi'$ \\
-    \RETURN $\phi$
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma}
-  \label{lem:hyperiso}  %{lem:perm}
-  If $(\cF, \cl)$ is an ICPPL from a tree $T$ and
-  Algorithm~\ref{al:icppl-find-isomorph}, {\tt
-    get- hypergraph- isomorphism ($\cF, \cl, T$)} returns a non-empty
-  function, then there exists a hypergraph isomorphism $\phi :
-  supp(\cF) \rightarrow V(T)$ such that the $\phi$-induced tree path
-  labeling is equal to $\cl$ or $\cl_\phi = \cl$.
-\end{lemma}
-\begin{proof}
-  It is clear that in the end of every recursive call to
-  Algorithm~\ref{al:icppl-find-isomorph}, the function $\phi'$ is one
-  to one involving all the leaves in the tree passed as input to that
-  recursive call. Moreover, by Lemma~\ref{lem:noexit1} and
-  Lemma~\ref{lem:xnotempty} it is consistent with the tree path
-  labeling $\cl$ passed. The tree pruning is done by only removing
-  leaves in each call to the function and is done till the tree
-  becomes empty. Thus the returned function $\phi: supp(\cF)
-  \rightarrow V(T)$ is a union of mutually exclusive one to one
-  functions exhausting all vertices of the tree. In other words, it is
-  a bijection from $supp(\cF)$ to $V(T)$ inducing the given path
-  labeling $\cl$ and thus a hypergraph isomorphism. \qed
-\end{proof}
-
-\begin{theorem}
-  \label{th:charac}
-  A path labeling $(\cF, \cl)$ on tree $T$ is feasible iff it is an
-  ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
-  T)$ as input returns a non-empty function.
-\end{theorem}
-\begin{proof}
-  From Lemma~\ref{lem:hyperiso}, we know that if $(\cF, \cl)$ is an
-  ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
-  T)$ as input returns a non-empty function, $(\cF, \cl)$ is feasible.
-  Now consider the case where $(\cF, \cl)$ is feasible. i.e. there
-  exists a hypergraph isomorphism $\phi$ such that $\cl_\phi =
-  \cl$. Lemma~\ref{lem:noexit1} and Lemma~\ref{lem:xnotempty} show us
-  that filter 1 and filter 2 do not exit if input is feasible. Thus
-  Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
-  function.\qed
-\end{proof}
-
-\section{ Computing feasible TPL with special target trees \tnote{give
-    problem definition etc}}
-\label{sec:spltargettree}
-
-Section~\ref{sec:feasible} described properties that a TPL must have
-for it to be feasible. The next problem of interest is to test if a
-given hypergraph is a path hypergraph with respect to a given target
-tree\footnote{A larger problem would be to find if a given hypergraph
-  is a path graph on any tree. This problem is not addressed in this
-  thesis.}. In other words, the problem is to find out if a feasible
-tree path labeling exists from a given target tree for a given
-hypergraph. In this section we will see two special cases of this
-problem where the target tree is from a particular family of
-trees. The first one, where the tree is a path, is a shown to be
-equivalent to the well studied problem of consecutive-ones in
-Section~\ref{sec:icpplicpia}. The second one, where the tree is a
-$k$-subdivided tree\footnote{See Section~\ref{ch:prelims} for the
-  formal definition.}, has been solved using a polynomial time
-algorithm. The latter problem enforces some conditions on the
-hypergraph too which will be seen in Section~\ref{sec:ksubdivstar}.
-
-\subsection{Target tree is a Path}
-\label{sec:icpplicpia}
-
-Consider a special case of ICPPL with the following properties.
-\begin{enumerate}
-\item Given tree $T$ is a path. Hence, all path labels are interval
-  labels.
-\item Only pairwise intersection cardinality preservation is
-  sufficient. i.e. property (iii) in ICPPL is not enforced.
-\item The filter algorithms do not have {\em \bf exit} statements.
-\end{enumerate}
-This is called an Intersection Cardinality Preservation Interval
-Assignment (ICPIA) \cite{nsnrs09}. This structure and its algorithm is
-used in the next section for finding tree path labeling from a
-$k$-subdivided star due to this graph's close relationship with
-intervals.
-
-\subsection{Target tree is a $k$-subdivided Star}
-\label{sec:ksubdivstar}
-
-\begin{figure}[t] %[htbp] 
-  \centering
-  \begin{tabular}{lr}
-    (a) \includegraphics{cop-tree_thesis.1}&
-    (b) \includegraphics{cop-tree_thesis.2}
-  \end{tabular}
-  \caption{\figtabsize (a) $8$-subdivided star with 7 rays (b)
-    3-subdivided star with 3 rays}
-  \label{fig:kstar}
-\end{figure}
-
-In this section we consider the problem of assigning paths from a
-$k$-subdivided star $T$ to a given set system $\cF$.  We consider
-$\cF$ for which the overlap graph $\bO(\cF)$ is connected.  The
-overlap graph is well-known from the work of
-\cite{kklv10,nsnrs09,wlh02}.  We use the notation in
-\cite{kklv10}. Recall from Section~\ref{ch:prelims} that hyperedges
-$S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$
-and $S'$ have a non-empty intersection but neither of them is
-contained in the other. The overlap graph $\bO(\cF)$ is a graph in
-which the vertices correspond to the sets in $\cF$, and the vertices
-corresponding to the hyperedges $S$ and $S'$ are adjacent if and only
-if they overlap.  Note that the intersection graph of $\cF$,
-$\bI(\cF)$ is different from $\bO(\cF)$ and $\bO(\cF) \subseteq
-\bI(\cF)$.  A connected component of $\bO(\cF)$ is called an overlap
-component of $\cF$.  An interesting property of the overlap components
-is that any two distinct overlap components, say $\cO_1$ and $\cO_2$,
-are such that any two sets $S_1 \in \cO_1$ and $S_2 \in \cO_2$ are
-disjoint, or, w.l.o.g, all the sets in $\cO_1$ are contained within
-one set in $\cO_2$.  This containment relation naturally determines a
-decomposition of the overlap components into rooted containment trees.
-We consider the case when there is only one rooted containment tree,
-and we first present our algorithm when $\bO(\cF)$ is connected.  It
-is easy to see that once the path labeling to the overlap component in
-the root of the containment tree is achieved, the path labeling to the
-other overlap components in the rooted containment tree is essentially
-finding a path labeling when the target tree is a path: each target
-path is a path that is allocated to sets in the root overlap
-component.  Therefore, for the rest of this section, $\bO(\cF)$ is a
-connected graph. We also assume that all hyperedges are of cardinality
-at most $k+2$.
-
-Recall from Section~\ref{ch:prelims} that a $k$-subdivided star is a
-star with each edge subdivided $k$ times. Therefore, a $k$-subdivided
-star has a central vertex which we call the {\em root}, and each root
-to leaf path is called a {\em ray}. First, we observe that by removing
-the root $r$ from $T$, we get a collection of $p$ vertex disjoint
-paths of length $k+1$, $p$ being the number of leaves in $T$.  We
-denote the rays by $R_1, \ldots, R_p$, and the number of vertices in
-$R_i$, $i \in [p]$ is $k+2$.  Let $\seq{v_{i1},\ldots,v_{i(k+2)}=r}$
-denote the sequence of vertices in $R_i$, where $v_{i1}$ is the
-leaf. Note that $r$ is a common vertex to all $R_i$.
-
-In this section the given hypergraph, the $k$-subdivided star and the
-root of the star are denoted by $\cO$, $T$ and vertex $r$,
-respectively.
-% For each hyperedge $X \in \cO$, we will maintain a 2-tuple of non-negative
-% numbers $\seq{p_1(X), p_2(X)}$.  The numbers satisfy the property that
-% $p_1(X) + p_2(X) \leq |X|$, and at the end of path labeling, for each
-% $X$, $p_1(X) + p_2(X) = |X|$.  This signifies the algorithm tracking
-% the lengths of subpaths of the path assigned to $X$ from at most two
-% rays. We also maintain another parameter called the {\em residue} of
-% $X$ denoted by $s(X)=|X| - p_1(X)$. This signifies the residue path
-% length that must be assigned to $X$ which must be from another
-% ray. For instance, if $X$ is labeled a path from only one ray, then
-% $p_1(X) = |X|$, $p_2(X) = 0$ and $s(X) = 0$.
-
-% 
-% We iteratively consider each ray from which paths will be assigned to
-% hyperedges. At the beginning of each iteration hyperedges of $\cO$ are
-% classifed into the following disjoint sets.
-% \begin{enumerate}
-% \item [$\cL_1^i$] {\em Labeled without $r$.} Those that have been
-%   labeled with a path which does not contain $r$ in one of the
-%   previous iterations.\\  $\cL_1^i = \set{ X \mid p_1(X) = |X| \text{ and
-%     } p_2(X) = 0 \text{ and } s(X) = 0, X \in \cO}$
-% \item [$\cL_2^i$] {\em Labeled with $r$.} Those that have been labeled
-%   with two subpaths of $\cl(X)$ containing $r$ from two different rays
-%   in two previous iterations.\\ $\cL_2^i = \set{X \mid 0 < p_1\left(X\right),
-%     p_2\left(X\right) < |X| \text{ and } s(X) = 0, X \in \cO}$
-%   \item [$\cT_1^i$] {\em Type 1 / partially labeled.} Those that have
-%     been labeled with one path containing $r$ from a single ray in one
-%     of the previous iterations. Here, $p_1(X)$ denotes the length of
-%     the subpath of $\cl(X)$ that $X$ has been so far labeled
-%     with. Also, it is clear that such a path must start at $r$ and
-%     must be in a ray different from the one corresponding to $p_1(X)$.\\
-%     $\cT_1^i = \set{ X \mid 0 < p_1(X) < |X| \text{ and } p_2(X) = 0
-%       \text{ and } s(X) > 0, X \in \cO}$
-%   \item [$\cT_2^i$] {\em Type 2 / not labeled.} Those that have not been
-%     labeled with a path in any previous iteration.\\
-%     $\cT_2^i = \set{ X \mid p_1(X) = p_2(X) = 0 \text{ and } s(X) = |X|,
-%       X \in \cO}$
-% \end{enumerate}
-% \vspace{-2mm}
-% \begin{align*}
-%   \cO &= \cL_1^i \cup \cL_2^i \cup \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]\\
-%   \cO_i &= \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]
-% \end{align*}
-
-The set $\cO_i$ refers to the set of hyperedges $\cT_1^i \cup \cT_2^i$
-in the $i$th iteration.  Note that $\cO_1 = \cO$.  In the $i$th
-iteration, hyperedges from $\cO_i$ are assigned paths from $R_i$ using
-the following rules. Also the end of the iteration, $\cL_1^{i+1},
-\cL_2^{i+1}, \cT_1^{i+1}, \cT_2^{i+1}$ are set to $\cL_1^{i},
-\cL_2^{i}, \cT_1^{i}, \cT_2^{i}$ respectively, along with some
-case-specific changes mentioned in the rules below.
-
-\noindent
-Suppose the given hypergraph has a feasible path labeling to the given
-tree $T$. Let this labeling be $\cl$. Now we present an algorithm that
-lists the sequence of edges that maintain a set of properties
-necessary for them to have a feasible tree path labeling.
-\begin{enumerate}[I.]
-\item {\bf Step 1}
-  \begin{enumerate}
-  \item {\bf {There are type 1
-        edges, $\mathbf{|\cT_1^i| > 0}$}:} 
-% , and for each such $X$ assign the
-%     path in $R_i$ of length $s(X)$ starting at $v_{i,k+2}=r$.  $X$ is
-%     then removed from $\cO_i$, and $p_2(X)=s(X)$, and $s(X)=0$.If any
-%     intersection cardinality is violated because of this assignment,
-%     EXIT reporting the non-existence of a feasible assignment.
-  {If there is only one $\cT_1^i$ hyperedge, label it to the unique
-    path in $R_i$ of length $s(X)$ starting at $v_{i(k+2)}=r$. If
-    there are more than one $\cT_1^i$ hyperedges, for any pair of
-    hyperedges $S_1, S_2 \in \cT_1^i$, one of the following is
-    true. Let $\cl'(X) \subseteq R_j$ denote the subpath of length
-    $p_1(X)$ that has been assigned to $X$ from a previously
-    considered ray $R_j$.}
-  \begin{enumerate}[{\em Case }1. ]
-  \item $|\cl'(S_1) \cap \cl'(S_2)| = 1$ and $|S_1 \cap S_2| > 1$
-    i.e. they are in different rays and their residues must contain
-    their intersection. Hence assign both to $R_i$.
-  \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| >
-    |\cl'(S_1) \cap \cl'(S_2)|$ i.e. they are in the same rays and
-    their residues must contain their intersection. Hence assign both
-    to $R_i$.
-  \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| =
-    |\cl'(S_1) \cap \cl'(S_2)|$. i.e. the intersection cardinality has
-    been met. Hence assign only $S_1$ to $R_i$ and $S_2$ must be
-    assigned to some $i \ge p$.
-  \end{enumerate}
-  Remove the labeled edge or edges from $\cO_i$ and add it to
-  $\cO_{i+1}$. Also do not add it to $\cT_1^{i+1}$ and add to
-  $\cL_2^{i+1}$.
-\item {\bf {There are no type-1 edges nor} super-marginal type-2
-    hyperedges:} If all type-2 edges of length at most $k+1$, EXIT by
-  saying there is no ICPPL.  Otherwise, Find an $X$ of type-2 such
-  that $|X| \geq k+2$, assign it the unique path starting at
-  $v_{i(k+1)}=r$ of length $k+2$, and set $p_1(X)=k+2,
-  s(X)=|X|-(k+2)$, assign it the unique path starting at
-  $v_{i(k+1)}=r$ of length $k+2$, mark this as a type-1 hyperedge, and
-  add it to $\cO_{i+1}$ after removing it from $\cO_i$.
-  % \remove{Next, iteratively, consider each type-1 edge $X \in
-  %   \cO_i$, and if it intersects a previously assigned $Y$ such that
-  %   $\cl(Y) \subseteq R_i$, assign a path of length $s(X)$
-  %   containing $v_{i(k+2)}=r$, set $p_2(X)=s(X), s(X)=0$, and remove
-  %   $X$ from $\cO_i$.}\tnote{covered in step 2}
-
-\item {\bf There are super-marginal type-2 hyperedges {and no type-1
-      edges:}} Let $X$ be a super-marginal hyperedge. If $|X| \leq
-  k+2$, then $X$ is assigned the path of length $|X|$ starting at
-  $v_{i1}$, and $p_1(X) = |X|, p_2(X)=0$. $X$ is removed from $\cO_i$
-  and not added to $\cO_{i+1}$.  If $|X| > k+2$, then $p_1(X)=k+2$,
-  assign it the unique path starting at $v_{i(k+1)}=r$ of length $k+2$
-  and $s(X) = |X| - (k+2)$.  In this case, $X$ is classified as a
-  Type-1 edge, removed from $\cO_i$, and added to $\cO_{i+1}$.
-\end{enumerate}
-\item {\bf Step 2:} Iteratively, a hyperedge $X$ is selected from
-  $\cO_i$ that has an overlap with one of the hyperedges $Y$ such that
-  $\cl(Y) \subseteq R_i$, and a unique path is assigned to $X$.  The
-  path, say $U(X) \subseteq R_i$, that is assigned can be decided
-  unambiguously since the $X \overlap Y$, and all intersection
-  cardinalities can be preserved only at one of the ends of $\cl(Y)$.
-
-  \noindent
-  Let $\cl(X)$ denote the unique path assigned to $X$.  If $X$ is a
-  type-2 hyperedge and if the unique path of length $|X|$ does not
-  contain $r$, then $p_1(X) = |X|, p_2(X)=0$, and $X$ is removed from
-  $\cO_i$.  In the case when $\cl(X)$ has to contain $r$, then
-  $p_1(X)$ is the length of the path, $p_2(X)=0$, and $s(X) =
-  |X|-p_1(X)$. Further $X$ is classified as a type-1 hyperedge, added
-  to $\cO_{i+1}$ and removed from $\cO_i$.  In the case when $X$ is a
-  type-1 hyperedge, then we check if $U(X)$, which is of length $s(X)$
-  contains $r$. If it does, then we assign $\cl(X) \leftarrow \cl(X)
-  \cup U(X)$, remove $X$ from $\cO_i$ and do not add it to
-  $\cO_{i+1}$.  If not, then we {\bf Exit} reporting that an
-  assignment cannot be found.  The iteration ends when no hyperedge in
-  $\cO_i$ has an overlap with a hyperedge assigned to $R_i$.
-\end{enumerate}
-
-\noindent
-The order in which they are assigned is the sequence of hyperedges
-with the following properties. Note that if such a sequence cannot be
-computed, $\cO$ cannot have a feasible path labeling from $T$.
-
-\noindent
-\tnote{need to formalize the below properties} In the following
-lemmas we identify a set of necessary conditions for $\cF$ to have an
-ICPPL in the $k$-subdivided star $T$.  If during the execution of the
-above algorithm, one of these necessary conditions is not satisfied,
-the algorithm exits reporting the non-existence of an ICPPL.
-\begin{lemma}
-  Let all hyperedges in $\cO_i$ be type-1 edges.  Then there is a
-  maximal subset $\ccT_i \subseteq \cO_i$ with the following
-  properties:
-  \begin{enumerate}
-  \item $\ccT_i$ form an inclusion chain.
-  \item For all $X \in \ccT_i$, $s(X) \leq k+2$, and There is a an $X
-    \in \ccT_i$ such that $s(X)=k+2$.
-  \end{enumerate}
-\end{lemma}
-\begin{lemma}
-  If there are no \marginal type-2 edges in $\cO_i$, then there exists
-  at least $p-i$ type-2 hyperedges $X \in \cO_i$ such that $|X| \geq
-  k+2$.
-  % is a exist type-2 hyperedges of length at least $k+2$, and there
-  % must be at least $r-i$ such type-2 hyperedge, basically at least
-  % one per ray.  Each hyperedge contained in such an hyperedges of
-  % length at least $k+2$, is transitively related under the overlap
-  % relation to one that gets assigned a path that contains
-  % $r$. Basically, the proof will argue that otherwise, there will be
-  % a \marginal type-2 hyperedge, or that it is infeasible.  Also, it
-  % is here we assume that we assume that the size of the two
-  % universes is the same.
-\end{lemma}
-\begin{lemma}
-  At the end of Step-1 in the $i$-th iteration, if one hyperedge $X$
-  of type-2 is such that $\cl(X) \subseteq R_i$, then all other
-  hyperedges in $\cO_i$ are connected to $X$ in the overlap component.
-\end{lemma}
-\begin{lemma}
-  At the end of Step-2, If control has exit at any time, there is no
-  ICPPL. If control has not exit, then $R_i$ is saturated.  No
-  hyperedge of $\cO_{i+1}$ will get a path from $R_i$ in the future
-  iterations.  No type-2 hyperedge of $\cO_{i+1}$ will get a path from
-  $R_i$.  Basically $R_i$ is done.
-\end{lemma}
-\begin{lemma}
-  Finally, we need to prove that the assignment is an ICPPL. Secondly,
-  if there is a permutation then maps sets to paths, then it is indeed
-  an ICPPL, and our algorithm will basically find it.  It is a unique
-  assignment upto permutation of the leaves.
-\end{lemma}
-
-\begin{lemma}
-  \label{lem:sup-mar}
-  If $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
-  tree path labeling to tree $T$, then $\cl(X)$ will
-  contain a leaf in $T$.\\
-  On the otherhand, if $\cl(X)$ has a leaf in
-  $T$, $X$ is marginal but may not be super-marginal.
-\end{lemma}
-\begin{proof}
-  Suppose $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
-  path labeling from $T$.  Assume $\cl(X)$ does not have a leaf.  Let
-  $R_i$ be one of the rays (or the only ray) $\cl(X)$ is part of.
-  Since $X$ is in a connected overlap component, there exists $Y_1 \in
-  \cF$ and $X \nsubseteq Y_1$ such that $Y_1 \overlap X$ and $Y_1$ has
-  at least one vertex closer to the leaf in $R_i$ than any vertex in
-  $X$. Similarly with the same argument there exists $Y_2 \in \cF$
-  with same subset and overlap relation with $X$ except it has has at
-  least one vertex farther away from the leaf in $R_i$ than any vertex
-  in $X$. Clearly $Y_1 \cap X$ and $Y_2 \cap X$ cannot be part of same
-  inclusion chain which contradicts that assumption $X$ is
-  super-marginal. Thus the claim is proven.\qed
-\end{proof}
-
-\noindent
-Consider the overlap graph $\bO(\cF)$ of the given hypergraph
-$\cF$. Let $S_{sm} \in \cF$ be such that it is a super-marginal
-hyperedge.  Algorithm~\ref{al:ktree-label} uses $S_{sm}$ along with
-the overlap graph $\bO(\cF)$ to calculate the feasible tree path
-labeling to the $k$-subdivded tree $T$.
-
-\begin{algorithm}[h]
-  \caption{{\tt compute-ksubtree-path-labeling($X, \cF, T$)}}
-  \label{al:ktree-label}
-  \begin{algorithmic}[\lndisplay]
-    \IF{$X = S_{sm}$}
-    \STATE -- TBD --\\
-    \ELSE 
-    \STATE -- TBD --\\
-    \ENDIF
-
- \end{algorithmic}
-\end{algorithm}
-
-\section{ TPL with no restrictions}
-\label{sec:norestraint}
-
-{\tt abstract begin}
-  it is
-  known that if the given tree is a path a feasible assignment can be
-  found in polynomial time, and we observe that it can actually be
-  done in logspace. \tnote{[TRUE?]}
-{\tt abstract end}
-
-In Section \ref{prelims} we present the
-necessary preliminaries, in Section \ref{feasible} we present our
-characterization of feasible tree path assignments, and in Section
-\ref{decompos} we present the characterizing subproblems for finding a
-bijection between $U$ and $V(T)$ such that sets map to tree
-paths. Finally, in Section \ref{complexity} we conclude by showing
-that Tree Path Assignment is GI-Complete, and also observe that
-Consecutive Ones Testing is in Logspace.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC \subsection{Introduction}
-
-\annote{ We refer to this as the Tree Path Assignment
-problem for an input $(\cF,T)$ pair.}{Add this ``pair'' as a
-  definition in ch:prelims}
-
-\remove{It is an interesting fact that for a matrix with the COP, the
-intersection graph of the corresponding set system is an interval
-graph.  A similar connection to a subclass of chordal graphs, and this
-subclass contains interval graphs, exists for the generalization of
-COP.  In this case, the intersection graph of the corresponding set
-system must be a path graph. Chordal graphs are of great
-significance, extensively studied, and have several applications.  One
-of the well known and interesting properties of a chordal graphs is
-its connection with intersection graphs cite{mcg04}. For every chordal
-graph, there exists a tree and a family of subtrees of this tree such
-that the intersection graph of this family is isomorphic to the
-chordal graph cite{plr70,gav78,bp93}.  Certain format of these trees
-are called clique trees cite{apy92} of the graph which is a compact
-representation of the chordal graph. There has also been work done on
-the generalization of clique trees to clique hypergraphs cite{km02}.
-If the chordal graph can be represented as the intersection graph of
-paths in a tree, then the graph is called path graph cite{mcg04} or
-path chordal graph.  Therefore, it is clear that if there is a
-bijection from $U$ to $V(T)$ such that the sets map to paths, then the
-intersection graph of the set system is indeed a path chordal graph.
-This is, however, only a necessary condition and can be checked
-efficiently, as path graph recognition is polynomial
-solvable cite{gav78,aas93}.  Indeed, it is possible to construct a set
-system and tree, such that the intersection graph is a path chordal
-graph, but there is no bijection between $U$ and $V(T)$ such that the
-sets map to paths.  This connection indeed suggests that our problem
-is indeed as hard as path graph isomorphism.  Further path graph
-isomorphism is known be isomorphism-complete, see for example
-cite{kklv10}. }
-In the second part of this paper, we decompose our
-search for a bijection between $U$ and $V(T)$ into subproblems.  Each
-subproblem is on a set system in which for each set, there is another
-set in the set system with which the intersection is {\em strict}-
-there is a non-empty intersection, but neither is contained in the
-other.  This is in the spirit of results in \cite{wlh02,nsnrs09} where
-to test for COP in a given matrix, the COP problem is solved on an
-equivalent set of prime matrices.  \annote{Our decomposition localizes the
-challenge of path graph isomorphism to two problems.}{HOW?}
-
-Finally, we show that Tree Path Assignment is isomorphism-complete.
-We also point out Consecutive Ones Testing is in Logspace from two
-different results in the literature \cite{kklv10, mcc04}. To the best
-of our knowledge this observation has not been made earlier.
-
-% \noindent
-% It has been long known that interval graph recognition is in
-% logspace\cite{rei84}. Recently interval graph isomorphism was also
-% shown to be logspace decidable using a logspace canonization
-% algorithm by \cite{kklv10}.  This result is built on top of logspace
-% results of undirected graph connectivity \cite{rei08}, logspace
-% tractability using a certain logical formalism called FP+C and
-% modular decomposition of interval graphs\cite{lau10} etc.  Interval
-% graphs are closely connected to binary matrices with COP. The
-% maximal clique vertex incidence matrix (matrix with rows
-% representing maximal cliques and columns representing vertices of a
-% graph) has COP on columns iff the graph is an interval
-% graph\cite{fg65}. This follows from the interval graph
-% characterization by \cite{gh64}. Due to this close relation it is
-% natural to see if consecutive ones property can be
-% tested in logspace. \\
-% \noindent
-% We also explore some extensions of the interval assignment problem
-% in \cite{nsnrs09}, namely caterpillar path assignment problem.
-
-% We present a logspace algorithm here that uses the ICPIA
-% characterization of binary matrices with COP (set system associated
-% with such a matrix)\cite{nsnrs09}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC\subsection{Preliminaries} 
-\label{prelims} 
-In this paper, the collection
-$\F = \{S_i \mid S_i \subseteq U, S_i \ne \O, i \in I\}$ is a {\em set
-  system} of universe, $U = \{1, \ldots, n\}$. Moreoever, a set system
-is assumed to ``cover'' the universe,
-i.e. $ \bigcup_{i \in I}S_i = U$. \\
-
-\noindent
-The graph $T=(V,E)$ represents a given tree with $|V| = n$. Further,
-for simplicity, $V$ is defined as $\{1,\ldots,n\}$. All paths
-referred to in this paper are paths of $T$ unless explicitly specified. \\
-
-\noindent
-A {\em path assignment} $\A$ to $\F$ is defined as a set assignment
-where second universe is the vertex set $V$ of a given tree $T$ and
-every second subset in the ordered pairs is a path in this
-tree. Formally, the definition is as follows.
-\begin{align*}
-  \A = \{ (S_i,P_i) \mid S_i \in \F, P_i \subseteq V \text{
-    s.t. }T[P_i] \text{ is a path, } i \in I \}
-\end{align*}
-In other words, $P_i$ is the path on the tree $T$ assigned to $S_i$ in
-$\A$. As mentioned before for set systems, the paths cover the whole
-tree, i.e. $\bigcup_{i \in I}P_i = V$ \\
-
-\noindent
-Generalizing the definition of {\em feasibility} in \cite{nsnrs09} to
-a set assignment, a path assignment $\A$ is defined to be {\em
-  feasible} if there exists a bijection defined as follows.
-\begin{align}
-  \sigma: U \rightarrow V(T), \text{ such that }\sigma(S_i) = P_i
-  \text{ for all } i \in I, \sigma \text{ is a bijection}
-  \label{eq:stf}
-\end{align}
-% The assignment $\A$ is then called {\em Set Translation Feasible}.
-
-\noindent
-Let $X$ be a partially ordered set with $\preccurlyeq$ being the
-partial order on $X$.  $mub(X)$ represents an element in $X$ which is
-a maximal upper bound on $X$.  $X_m \in X$ is a maximal upper bound of
-$X$ if $\nexists X_q \in X$ such that $X_m
-\preccurlyeq X_q$. \\
-
-\noindent
-The set $I$ represents the index set $[m]$. If index $i$ is used
-without further qualification, it is meant to be $i \in I$. Any
-function, if not defined on a domain of sets, when applied on a set is
-understood as the function applied to each of its elements. i.e. for
-any function $f$ defined with domain $U$, the abuse of notation is as
-follows; $f(S)$ is used instead of $\hat f(S)$ where
-$\hat f(S) = \{y \mid y = f(x), x \in S\}$. \\
-
-\noindent
-When refering to a tree as $T$ it could be a reference to the tree
-itself, or the vertices of the tree. This will be clear from the
-context.\\
-
-\noindent
-Finally, an in-tree is a directed rooted tree in which all edges are
-directed toward to the root.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC\subsection{Characerization of Feasible Tree Path
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC  Assignment\tnote{Merge with the original charac section}} 
-\label{feasible} 
-Consider a path assignment $\cA =
-\{(S_i, P_i) \mid S_i \in \cF, P_i \text{ is a path from $T$}, i \in
-[m]\}$ to a set system $\cF = \{S_i \mid S_i \subseteq U, i \in
-[m]\}$, were $T$ is a given tree, $U$ is the set system's universe and
-$m$ is the number of sets in $\cF$. We call $\cA$ an {\em Intersection
-  Cardinality Preserving Path Assignment (ICPPA)} if it has the
-following properties.
-
-\begin{enumerate}
-\item [i.]  $|S_i| = |P_i|$ for all $i \in [m]$
-\item [ii.] $|S_i \cap S_j| = |P_i \cap P_j|$ for all $i,j \in [m]$
-  % \item [iii.] $|\bigcup_{i \in J} S_i| = |\bigcup_{i \in J} P_i|$
-  %   for all $J \subseteq I$\footnote{TBD: this is needed now to fix
-  %     some proofs}
-\item [iii.] $|S_i \cap S_j \cap S_k| = |P_i \cap P_j \cap P_k|$ for
-  all $i,j,k \in [m]$
-\end{enumerate}
-
-\begin{lemma}
-  \label{lem:setminuscard}
-  If $\cA$ is an ICPPA, and $(S_1, P_1),(S_2, P_2),(S_3, P_3) \in
-  \cA$, then $|S_1 \cap (S_2 \setminus S_3)| = |P_1 \cap (P_2
-  \setminus P_3)|$.
-\end{lemma}
-\begin{proof}
-  $|S_1 \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| =
-  |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to conditions (ii)
-  and (iii) of ICPPA, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
-  \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
-  = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. \qed
-\end{proof}
-
-\begin{lemma}
-  \label{lem:fourpaths} Consider four paths in a tree $Q_1, Q_2, Q_3,
-  Q_4$ such that they have nonempty pairwise intersection and $Q_1,
-  Q_2$ share a leaf. Then there exists distinct $i, j, k \in
-  \{1,2,3,4\}$ such that, $Q_1 \cap Q_2 \cap Q_3 \cap Q_4 = Q_i \cap
-  Q_j \cap Q_k$.
-\end{lemma}
-\begin{proof}
-  {\em Case 1:} w.l.o.g, consider $Q_3 \cap Q_4$ and let us call it
-  $Q$. This is clearly a path (intersection of two paths is a path).
-  % Since $Q_1, Q_2$ share a leaf, the following are paths $Q_1
-  % \setminus Q_2$, $Q_2 \setminus Q_1$, $Q_1 \cap Q_2$ and they are
-  % mutually disjoint.
-  Suppose $Q$ does not intersect with $Q_1 \setminus Q_2$, i.e. $Q
-  \cap (Q_1 \setminus Q_2) = \O$. Then $Q \cap Q_1 \cap Q_2 = Q \cap
-  Q_2$. Similarly, if $Q \cap (Q_2 \setminus Q_1) = \O$, $Q \cap Q_1
-  \cap Q_2 = Q \cap Q_1$. Thus it is clear that if the intersection of
-  any two paths does not intersect with any of the set differences of
-  the remaining two paths, the claim in the lemma is true.
-  % Note that $Q_1 \setminus Q_2$ and $Q_2
-  % \setminus Q_1$ are paths because $Q_1, Q_2$ share a leaf.\\
-  {\em Case 2:} Let us consider the compliment of the previous
-  case. i.e. the intersection of any two paths intersects with both
-  the set differences of the other two. First let us consider $Q \cap
-  (Q_1 \setminus Q_2) \ne \O$ and $Q \cap (Q_1 \setminus Q_2) \ne \O$,
-  where $Q = Q_3 \cap Q_4$. Since $Q_1$ and $Q_2$ share a leaf, there
-  is exactly one vertex at which they branch off from the path $Q_1
-  \cap Q_2$ into two paths $Q_1 \setminus Q_2$ and $Q_2 \setminus
-  Q_1$. Let this vertex be $v$. It is clear that if path $Q_3 \cap
-  Q_4$, must intersect with paths $Q_1 \setminus Q_2$ and $Q_2
-  \setminus Q_1$, it must contain $v$ since these are paths from a
-  tree. Moreover, $Q_3 \cap Q_4$ intersects with $Q_1 \cap Q_2$ at
-  exactly $v$ and only at $v$ which means that $Q_1 \cap Q_2$ does not
-  intersect with $Q_3 \setminus Q_4$ or $Q_4 \setminus Q_3$ which
-  contradicts initial condition of this case. Thus this
-  case cannot occur and case 1 is the only possible scenario. \\
-  Thus lemma is proven \qed
-\end{proof}
-
-\noindent
-In the remaining part of this section we show that a path assignment
-is feasible if and only if it is an ICPPA.  One direction of this
-claim is clear: that is a path assignment is feasible, then all
-intersection cardinalities are preserved, that is the path assignment
-is an ICPPA.  The reason is that a feasible path assignment has an
-associated bijection between $U$ and $V(T)$ such that the sets map to
-paths.  The rest of the section is devoted to constructively proving
-that it is sufficient, for the existence of a bijection, for a path
-assignment to be an ICPPA.  At a top-level, the constructive
-approaches refine the path assignment iteratively, such that at the
-end of each iteration we have a path assignment, and finally we have a
-family of bijections.  First we present and then prove the correctness
-of Algorithm \ref{perms}.  This algorithm refines the path assignment
-by considering pairs of paths that share a leaf.
-
-\begin{algorithm}[h]
-  \caption{Permutations from an ICPPA $\{(S_i,P_i) | i \in I\}$}
-  \label{perms}
-  \begin{algorithmic}
-    \STATE Let $\Pi_0=\{(S_i,P_i)| i \in I\}$\\
-    \STATE $j = 1$;
-    \label{shareleaf} \WHILE {There is $(P_1,Q_1), (P_2,Q_2) \in
-      \Pi_{j-1}$ with $Q_1$ and $Q_2$ having a common leaf} \STATE
-    $\Pi_j= \Pi_{j-1} \setminus \{(P_1,Q_1),(P_2,Q_2)\}$;
-    \label{setbreak}\STATE $\Pi_j = \Pi_j \cup \{(P_1 \cap P_2,Q_1
-    \cap Q_2), (P_1 \setminus P_2,Q_1 \setminus Q_2), (P_2 \setminus
-    P_1, Q_2 \setminus Q_1)\}$; \STATE $j = j+1$; \ENDWHILE \STATE
-    $\Pi = \Pi_j$; \STATE Return $\Pi$;
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma}
-  \label{lem:invar1}
-  In Algorithm \ref{perms}, at the end of $j$th iteration, $j \ge 0$,
-  of the while loop of Algorithm \ref{perms}, the following invariants
-  are maintained.
-  \begin{itemize}
-  \item {\em Invariant I:} $Q$ is a path in $T$ for each $(P,Q) \in
-    \Pi_j$
-  \item {\em Invariant II:} $|P|=|Q|$ for each $(P,Q) \in \Pi_j$
-  \item {\em Invariant III:} For any two $(P,Q), (P',Q') \in \Pi_j$,
-    $|P' \cap P''|=|Q' \cap Q''|$.
-  \item {\em Invariant IV:} For any three, $(P',Q'), (P'',Q''), (P, Q)
-    \in \Pi_j$, $|P' \cap P'' \cap P|=|Q' \cap Q'' \cap Q|$.
-  \end{itemize}
-\end{lemma}
-\begin{proof}
-  Proof is by induction on the number of iterations, $j$. In the rest
-  of the proof, the term ``new sets'' will refer to the new sets added
-  in $j$th iteration as defined in line \ref{setbreak} of Algorithm
-  \ref{perms}, i.e. the following three assignment pairs for some
-  $(P_1,Q_1), (P_2,Q_2) \in \Pi_{j-1}$ where $Q_1$ and $Q_2$ intersect
-  and share a leaf: $(P_1 \cap P_2, Q_1 \cap Q_2)$, or $(P_1 \setminus
-  P_2, Q_1 \setminus Q_2)$, or $(P_2 \setminus P_1, Q_2
-  \setminus Q_1)$.\\
-  \noindent
-  The base case, $\Pi_0 = \{(S_i,P_i) \mid i \in [m]\}$, is trivially
-  true since it is the input which is an ICPPA.  Assume the lemma is
-  true till the $j-1$ iteration. Consider $j$th
-  iteration:\\
-  \noindent
-  If $(P,Q)$, $(P',Q')$ and $(P'',Q'')$ are in $\Pi_{j}$ and
-  $\Pi_{j-1}$, all the invariants are
-  clearly true because they are from $j-1$ iteration.\\
-  If $(P,Q)$ is in $\Pi_{j}$ and not in $\Pi_{j-1}$, then it must be
-  one of the new sets added in $\Pi_j$. Since $(P_1,Q_1)$ and
-  $(P_2,Q_2)$ are from $\Pi_{j-1}$ and $Q_1,Q_2$ intersect and have a
-  common leaf, it can be verified that the
-  new sets are also paths. \\
-  By hypothesis for invariant III, invariant II also holds for $(P,Q)$
-  no matter which new set in $\Pi_j$ it
-  is.\\
-  To prove invariant III, if $(P,Q)$ and $(P',Q')$ are not in
-  $\Pi_{j-1}$, then they are both new sets and invariant III holds
-  trivially (new sets are disjoint). Next consider $(P,Q), (P',Q') \in
-  \Pi_j$ with only one of them, say $(P',Q')$, in $\Pi_{j-1}$. Then
-  $(P,Q)$ is one of the new sets added in line \ref{setbreak}. It is
-  easy to see that if $(P,Q)$ is $(P_1 \cap P_2, Q_1 \cap Q_2)$, then
-  due to invariant IV in hypothesis, invariant III becomes true in
-  this iteration. Similarly, using lemma \ref{lem:setminuscard}
-  invariant III is proven if $(P, Q)$ is $(P_1 \setminus P_2, Q_1
-  \setminus Q_2)$, or $(P_2 \setminus P_1, Q_2
-  \setminus Q_1)$.\\
-  To prove invariant IV, consider three assignments
-  $(P,Q),(P',Q'),(P'',Q'')$. If at least two of these pairs are in not
-  $\Pi_{j-1}$, then they are any two of the new sets. Note that these
-  new sets are disjoint and hence if $(P',Q'), (P'',Q'')$ are any of
-  these sets, $|P \cap P' \cap P''|=|Q \cap Q' \cap Q''|=0$ and
-  invariant IV is true. Now we consider the case if at most one of
-  $(P,Q),(P',Q'),(P'',Q'')$ is not in $\Pi_{j-1}$. If none of them are
-  not in $\Pi_{j-1}$ (i.e. all of them are in $\Pi_{j-1}$), invariant
-  IV is clearly true. Consider the case where exactly one of them is
-  not in $\Pi_{j-1}$. w.l.o.g let that be $(P,Q)$ and it could be any
-  one of the new sets. If $(P,Q)$ is $(P_1 \cap P_2, Q_1 \cap Q_2)$,
-  from lemma \ref{lem:fourpaths} and invariant III hypothesis,
-  invariant IV is proven. Similarly if $(P,Q)$ is any of the other new
-  sets, invariant IV is proven by also using lemma
-  \ref{lem:setminuscard}. \qed
-
-\end{proof}
-
-\noindent
-It can be observed that the output of algorithm \ref{perms} is such
-that every leaf is incident on at most a single path in the new set of
-assignments. This is due to the loop condition at line
-\ref{shareleaf}. Let $v_1$ be the leaf incident on path $P_i$. Assign
-to it any one element from $S_i \setminus \bigcup_{i \ne j}
-S_j$. Remove $(S_i, P_i)$ from assignments and add $(\{x_1\},
-\{v_1\}), (S_i \setminus \{x_1\}, P_i \setminus \{v_1\})$. Now all
-assignments except single leaf assignments are paths from the subtree
-$T_0 = T \setminus \{v \mid v \text{ is a leaf in } T\}$.
-
-% exclusively on the spine of the caterpillar. Consider the spine to
-% be an interval and run ICPIA algorithm on it to get the required
-% permutation. Algorithm \ref{leafasgn} makes this clear.
-
-\begin{algorithm}[h]
-  \caption{Leaf assignments from an ICPPA $\{(S_i,P_i) | i \in I\}$}
-  \label{leafasgn}
-  \begin{algorithmic}
-    \STATE Let $\Pi_0=\{(S_i,P_i)| i \in [m]\}$. Paths are such that
-    no
-    two paths $P_i, P_j, i \ne j$ share a leaf.\\
-    \STATE $j = 1$\\
-    \WHILE {there is a leaf $v$ and a unique $(S_{i_1}, P_{i_1})$ such
-      that $v \in P_{i_1}$}
-    \STATE $\Pi_j=   \Pi_{j-1} \setminus \{(S_{i_1}, P_{i_1})\}$\\
-    \STATE $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in I}S_i$
-    \IF{$X$ is empty} \STATE exit \ENDIF; \STATE Let $x = $ arbitrary
-    element from $X$ \vspace{2mm}
-
-    \STATE $\Pi_j = \Pi_j \cup \{(S_{i_1} \setminus \{x\}),(P_{i_1}
-    \setminus
-    \{v\}), (\{x\},\{v\})\}$\\
-    \STATE $j = j+1$\\
-    \ENDWHILE
-    \STATE $\Pi = \Pi_j$\\
-    \STATE Return $\Pi$\\
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma}
-  \label{lem:invar3}
-  In Algorithm \ref{leafasgn},for all $j \geq 0$, at the end of the
-  $j$th iteration the four invariants given in lemma \ref{lem:invar1}
-  are valid.
-  % Moreover, $X$ as defined in the algorithm is non-empty if this is
-  % an ICPPA.
-\end{lemma}
-\begin{proof}
-  First we see that $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in
-    I}S_i$ is non empty in every iteration for an ICPPA. Suppose $X$
-  is empty. We know that $v \in P_{i_1} \setminus \bigcup_{i \ne i_1,
-    i \in I}P_i$ since $v$ is in the unique path $P_{i_1}$. Since this
-  is an ICPPA $|S_{i_1}| = |P_{i_1}|$. For any $x \in S_{i_1}$ it is
-  contained in at least two sets due to our assumption. Let $S_{i_2}$
-  be a second set that contains $x$. We know $v \notin
-  P_{i_2}$. Therefore there cannot exist a permutation that maps
-  elements of $S_{i_2}$ to $P_{i_2}$. This contradicts our assumption
-  that this is an ICPPA. Thus $X$ cannot be empty.
-
-  We use mathematical induction on the number of iterations for this
-  proof. The term ``new sets'' will refer to the sets added in $\Pi_j$
-  in the $j$th iteration, i.e. $(P' \setminus \{x\},Q' \setminus
-  \{v\})$ and $(\{x\},\{v\})$ for some $(P',Q')$ in $\Pi_{j-1}$ such
-  that $v$ is a leaf and $Q'$ is the unique path
-  incident on it.\\
-  For $\Pi_0$ all invariants hold because it is output from algorithm
-  \ref{perms} which is an ICPPA. Hence base case is proved.  Assume
-  the lemma holds for $\Pi_{j-1}$. Consider $\Pi_j$ and any $(P,Q) \in
-  \Pi_j$. If $(P,Q) $ is in $ \Pi_j$ and $\Pi_{j-1}$ invariants I and
-  II are true because of induction assumption. If it is only in
-  $\Pi_j$, then it is $\{(P' \setminus \{x\}),(Q' \setminus \{v\})$ or
-  $(\{x\},\{v\})$ for some $(P',Q')$ in $\Pi_{j-1}$. By definition,
-  $x$ is an element in $P'$ (as defined in the algorithm) and $v$ is a
-  leaf in $Q'$. If $(P,Q)$ is $\{(P' \setminus \{x\}),(Q' \setminus
-  \{v\})$, $Q$ is a path since only a leaf is removed from path
-  $Q'$. We know $|P'| = |Q'|$, therefore $|P' \setminus \{x\}| = |Q'
-  \setminus \{v\}|$. Hence in this case invariants I and II are
-  obvious. It is easy to see these invariants hold if $(P,Q)$ is
-  $(\{x\},\{v\})$.
-
-  For invariant III consider $(P_1,Q_1),(P_2,Q_2)$ in $\Pi_j$. If both
-  of them are also in $\Pi_{j-1}$, claim is proved. If one of them is
-  not in $\Pi_{j-1}$ then it has to be $\{(P' \setminus \{x\}),(Q'
-  \setminus \{v\})$ or $(\{x\},\{v\})$ for some $(P',Q')$ in
-  $\Pi_{j-1}$. Since by definition, $Q'$ is the only path with $v$ and
-  $P'$ the only set with $x$ in the previous iteration, $|P_1 \cap (P'
-  \setminus \{x\})| = |P_1 \cap P'|$ and $|Q_1 \cap (Q' \setminus
-  \{v\})| = |Q_1 \cap Q'|$ and $|P_1 \cap \{x\}| = 0, Q_1 \cap \{v\} =
-  0$. Thus invariant III is also proven.
-
-  \noindent
-  To prove invariant IV, consider $(P_1,Q_1),(P_2,Q_2), (P_3,Q_3)$ in
-  $\Pi_j$. If exactly one of them, say $P_3 \notin \Pi_{j-1}$, it is
-  one of the new sets. By the same argument used to prove invariant
-  III, $|P_1 \cap P_2 \cap (P' \setminus \{x\})| = |P_1 \cap P_2 \cap
-  P'|$ and $|Q_1 \cap Q_2 \cap (Q' \setminus \{x\})| = |Q_1 \cap Q_2
-  \cap Q'|$. Since $P_1, P_2, P'$ are all in $\Pi_{j-1}$, by induction
-  hypothesis $|P_1 \cap P_2 \cap P'| = |Q_1 \cap Q_2 \cap Q'|$. Also
-  $|P_1 \cap P_2 \cap \{x\}| = 0, Q_1 \cap Q_2 \cap \{v\} = 0$.  If
-  two or more of them are not in $\Pi_{j-1}$, then it can be verified
-  that $|P_1 \cap P_2 \cap P_3| = |Q_1 \cap Q_2 \cap Q_3|$ since the
-  new sets in $\Pi_j$ are either disjoint or as follows: assuming
-  $P_1, P_2 \notin \Pi_{j-1}$ and new sets are derived from $(P', Q'),
-  (P'', Q'') \in \Pi_{j-1}$ with $x_1, x_2$ exclusively in $P_1, P_2$,
-  $(\{x_1\},\{v_1\}), (\{x_2\},\{v_2\}) \in \Pi_j $ thus $v_1, v_2$
-  are exclusively in $Q_1, Q_2$ resp. it follows that $|P_1 \cap P_2
-  \cap P_3| = |(P' \setminus \{x_1\}) \cap (P'' \setminus \{x_2\})
-  \cap P_3| = |P' \cap P'' \cap P_3| = |Q' \cap Q'' \cap Q_3| = |(Q'
-  \setminus \{v_1\} \cap Q'' \setminus \{v_2\} \cap Q_3| = |Q_1 \cap
-  Q_2 \cap Q_3|$. Thus invariant IV is also proven.  \qed
-\end{proof}
-
-\noindent
-Using algorithms \ref{perms} and \ref{leafasgn} we prove the following
-theorem.
-
-\begin{theorem}
-  \label{th:perm}
-  If $\cA$ is an ICPPA, then there exists a bijection $\sigma : U
-  \rightarrow V(T)$ such that $\sigma(S_i) = P_i$ for all $i \in I$
-\end{theorem}
-\begin{proof}
-  This is a contructive proof. First, the given ICPPA $\cA$ and tree
-  $T$ are given as input to Algorithm \ref{perms}. This yields a
-  ``filtered'' ICPPA as the output which is input to Algorithm
-  \ref{leafasgn}.  It can be observed that the output of Algorithm
-  \ref{leafasgn} is a set of interval assignments to sets and
-  one-to-one assignment of elements of $U$ to each leaf of $T$. To be
-  precise, it would be of the form $\cB_0 = \cA_0 \cup \cL_0$. The
-  leaf assignments are defined in $\cL_0 = \{ (x_i,v_i) \mid x_i \in
-  U, v_i \in T, x_i \ne x_j, v_i \ne v_j, i \ne j, i,j \in [k] \}$
-  where $k$ is the number of leaves in $T$. The path assignments are
-  defined in $\cA_0 \subseteq \{(S_i',P_i') \mid S_i' \subseteq U_0,
-  P_i' \text{ is a path from } T_0\}$ where $T_0$ is the tree obtained
-  by removing all the leaves in $T$ and $U_0 = U \setminus \{ x \mid x
-  \text{ is assigned to a leaf in }\cL_0 \}$. Now we have a subproblem
-  of finding the permutation for the path assignment $\cA_0$ which has
-  paths from tree $T_0$ and sets from universe $U_0$. Now we repeat
-  the procedure and the path assignment $\cA_0$ and tree $T_0$ is
-  given as input to Algorithm \ref{perms}. The output of this
-  algorithm is given to Algorithm \ref{leafasgn} to get a new union of
-  path and leaf assignments $\cB_1 = \cA_1 \cup \cL_1$ defined similar
-  to $\cB_0, \cL_0, \cA_0$. In general, the two algorithms are run on
-  path assignment $\cA_{i-1}$ with paths from tree $T_{i-1}$ to get a
-  new subproblem with path assignment $\cA_i$ and tree $T_{i}$. $T_i$
-  is the subtree of $T_{i-1}$ obtained by removing all its
-  leaves. More importantly, it gives leaf assignments $\cL_{i}$ to the
-  leaves in tree $T_{i-1}$. This is continued until we get a
-  subproblem with path assignment $\cA_{d-1}$ and tree $T_{d-1}$ for
-  some $d \le n$ which is just a path. From the last lemma we know
-  that $\cA_{d-1}$ is an ICPPA. Another observation is that an ICPPA
-  with all its tree paths being intervals (subpaths from a path) is
-  nothing but an ICPIA\cite{nsnrs09}.  Let $\cA_{d-1}$ be equal to
-  $\{(S_i'',P_i'') \mid S_i'' \subseteq U_{d-1}, P_i'' \text{ is a
-    path from } T_{d-1} \}$. It is true that the paths $P_i''$s may
-  not be precisely an interval in the sense of consecutive integers
-  because they are some nodes from a tree. However, it is easy to see
-  that the nodes of $T_{d-1}$ can be ordered from left to right and
-  ranked to get intervals $I_i$ for every path $P_i''$ as
-  follows. $I_i = \{[l,r] \mid l = \text{ the lowest rank of the nodes
-    in }P_i'', r = l+|P_i''|-1 \}$. Let asssignment $\cA_d$ be with
-  the renamed paths. $\cA_d = \{ (S_i'', I_i) \mid (S_i'', P_i'') \in
-  \cA_{d-1} \}$. What has been effectively done is renaming the nodes
-  in $T_{d-1}$ to get a tree $T_d$.  The ICPIA $\cA_d$ is now in the
-  format that the ICPIA algorithm requires which gives us the
-  permutation $\sigma' : U_{d-1} \rightarrow T_{d-1}$
-
-\noindent
-$\sigma'$ along with all the leaf assignments $\cL_i$ gives us the
-permutation for the original path assignment $\cA$.  More precisely,
-the permutation for tree path assignment $\cA$ is defined as
-follows. $\sigma: U \rightarrow T$ such that the following is
-maintained.
-\begin{align*}
-  \sigma(x) &= \sigma'(x),   \text{ if } x \in U_{d-1} \\
-  &= \cL_i(x), \text{ where $x$ is assigned to a leaf in a subproblem
-    $\cA_{i-1}, T_{i-1}$}
-\end{align*}
-
-\noindent
-To summarize, run algorithm \ref{perms} and \ref{leafasgn} on
-$T$. After the leaves have been assigned to specific elements from
-$U$, remove all leaves from $T$ to get new tree $T_0$. The leaf
-assignments are in $\cL_0$. Since only leaves were removed $T_0$ is
-indeed a tree. Repeat the algorithms on $T_0$ to get leaf assignments
-$\cL_{1}$. Remove the leaves in $T_0$ to get $T_1$ and so on until the
-pruned tree $T_d$ is a single path. Now run ICPIA algorithm on $T_d$
-to get permutation $\sigma'$. The relation $\cL_0 \cup \cL_1 \cup
-.. \cup \cL_{d} \cup \sigma'$ gives the bijection required in the
-original problem.\qed
-\end{proof}
-
-\subsection{Finding an assignment of tree paths to a set
-  system} \label{decompos} In the previous section we have shown that
-the problem of finding a Tree Path Asssignment to an input $(\cF,T)$
-is equivalent to finding an ICPPA to $\cF$ in tree $T$.  In this
-section we characterize those set systems that have an ICPPA in a
-given tree.  As a consequence of this characterization we identify two
-sub-problems that must be solved to obtain an ICPPA.  We do not solve
-the problem and in the next section show that finding an ICPPA in a
-given tree is GI-Complete.
-
-\noindent
-A set system can be concisely represented by a binary matrix where the
-row indices denote the universe of the set system and the column
-indices denote each of the sets. Let the binary matrix be $M$ with
-order $n \times m$, the set system be $\cF = \{S_i \mid i \in [m]\}$,
-universe of set system $U = \{x_1, \dots ,x_n\}$. If $M$ represents
-$\cF$, $|U| = n, |\cF| = m$. Thus $(i,j)$th element of $M$, $M_{ij} =
-1$ iff $x_i \in S_j$. If $\cF$ has a feasible tree path assignment
-(ICPPA) $\cA = \{(S_i,P_i) \mid i \in [m]\}$, then we say its
-corresponding matrix $M$ has an ICPPA. Conversly we say that a matrix
-$M$ has an ICPPA if there exists an ICPPA $\cA$ as defined
-above.\\
-\noindent
-We now define the strict intersection graph or overlap graph of
-$\cF$. This graph occurs at many places in the literature, see for
-example \cite{kklv10, wlh02, nsnrs09}.  The vertices of the graph
-correspond to the sets in $\cF$.  An edge is present between vertices
-of two sets iff the corresponding sets have a nonempty intersection
-and none is contained in the other. Formally, intersection graph is
-$G_f = (V_f, E_f)$ such that $V_f = \{v_i \mid S_i \in \cF\}$ and $E_f
-= \{(v_i, v_j) \mid S_i \cap S_j \ne \O \text{ and }S_i \nsubseteq
-S_j, S_j \nsubseteq S_i \}$.  We use this graph to decompose $M$ as
-described in \cite{wlh02,nsnrs09}.  A prime sub-matrix of $M$ is
-defined as the matrix formed by a set of columns of $M$ which
-correspond to a connected component of the graph $G_f$.  Let us denote
-the prime sub-matrices by $M_1,\ldots,M_p$ each corresponding to one
-of the $p$ components of $G_f$. Clearly, two distinct matrices have a
-distinct set of columns.  Let $col(M_i)$ be the set of columns in the
-sub-matrix $M_i$.  The support of a prime sub-matrix $M_i$ is defined
-as $supp(M_i) = \displaystyle \bigcup_{j \in col(M_i)}S_j$.  Note that
-for each $i$, $supp(M_i) \subseteq U$.  For a set of prime
-sub-matrices $X$ we define
-$supp(X) = \displaystyle \bigcup_{M \in X} supp(M)$. \\
-
-\noindent
-Consider the relation $\preccurlyeq$ on the prime sub-matrices $M_1,
-\ldots, M_p$ defined as follows:
-\begin{equation}
-  \nonumber \{(M_i,M_j) | \mbox{ A set } S \in
-  M_i \mbox{ is contained in a set } S' \in M_j\} \cup \{(M_i,M_i) | 1
-  \leq i \leq p\} 
-\end{equation}
-
-\noindent
-This relation is the same as that defined in \cite{nsnrs09}. The prime
-submatrices and the above relation can be defined for any set
-system. We will use this structure of prime submatrices to present our
-results on an an ICPPA for a set system $\cF$. Recall the following
-lemmas, and theorem that $\preccurlyeq$ is a partial order, from
-\cite{nsnrs09}.
-
-\begin{lemma} \label{lem:containment} Let $(M_i,M_j) \in
-  \preccurlyeq$.  Then there is a set $S' \in M_j$ such that for each
-  $S \in M_i$, $S \subseteq S'$.
-\end{lemma}
-\begin{lemma}
-  For each pair of prime sub-matrices, either $(M_i,M_j) \not\in
-  \preccurlyeq$ or $(M_j,M_i) \not\in \preccurlyeq$.
-  % If $(M_i,M_j) \in \preccurlyeq$ and $(M_j,M_i) \in \preccurlyeq$,
-  % then $i = j$ and $|M_i| = 1$.
-\end{lemma}
-\begin{lemma}
-  If $(M_i,M_j) \in \preccurlyeq $ and $(M_j,M_k) \in \preccurlyeq$,
-  then $(M_i,M_k) \in \preccurlyeq$.
-\end{lemma}
-\begin{lemma}
-  If $(M_i,M_j) \in \preccurlyeq$ and $(M_i,M_k) \in \preccurlyeq$,
-  then either $(M_j,M_k) \in \preccurlyeq$ or $(M_k,M_j) \in
-  \preccurlyeq$.
-\end{lemma}
-\begin{theorem} \label{thm:partitionold} $\preccurlyeq$ is a partial
-  order on the set of prime sub-matrices of $M$.  Further, it uniquely
-  partitions the prime sub-matrices of $M$ such that on each set in
-  the partition $\preccurlyeq$ induces a total order.
-\end{theorem}
-For the purposes of this paper, we refine the total order mentioned in
-Theorem \ref{thm:partitionold}. We do this by identifying an in-tree
-rooted at each maximal upper bound under $\preccurlyeq$.  Each of
-these in-trees will be on disjoint vertex sets, which in this case
-would be disjoint sets of prime-submatrices.  The in-trees are
-specified by selecting the appropriate edges from the Hasse diagram
-associated with $\preccurlyeq$.  Let $\cI$ be the following set:
-\begin{align*}
-  \cI = \{ (M_i,M_j) \in \preccurlyeq \mid \nexists M_k s.t. M_i
-  \preccurlyeq M_k, M_k \preccurlyeq M_j \} \cup \{ (M_i,M_i), i \in
-  [p] \}
-\end{align*}
-
-\begin{theorem} \label{thm:partition} Consider the directed graph $X$
-  whose vertices correspond to the prime sub-matrices, and the edges
-  are given by $\cI$.  Then, $X$ is a vertex disjoint collection of
-  in-trees and the root of each in-tree is a maximal upper bound in
-  $\preccurlyeq$.
-\end{theorem}
-\begin{proof}
-  To observe that $X$ is a collection of in-trees, we observe that for
-  vertices corresponding to maximal upper bounds, no out-going edge is
-  present in $I$.  Secondly, for each other element, exactly one
-  out-going edge is chosen, and for the minimal lower bound, there is
-  no in-coming edge.  Consequently, $X$ is acyclic, and since each
-  vertex has at most one edge leaving it, it follows that $X$ is a
-  collection of in-trees, and for each in-tree, the root is a maximal
-  upper bound in $\preccurlyeq$.  Hence the theorem.
-\end{proof}
-Let the partition of $X$ given by Theorem \ref{thm:partition} be
-$\{X_1,\ldots,X_r\}$.  Further, each in-tree itself can be layered
-based on the distance from the root.  The root is considered to be at
-level zero. For $j \geq 0$, Let $X_{i,j}$ denote the set of prime
-matrices in level $j$ of in-tree $X_i$.
-
-\begin{lemma}
-  \label{lem:subicppa}
-  Let $M$ be a matrix and let $X$ be the directed graph whose vertices
-  are in correspondence with the prime submatrices of $M$.  Further
-  let $\{X_1,\ldots,X_r\}$ be the partition of $X$ into in-trees as
-  defined above.  Then, matrix $M$ has an ICPPA in tree $T$ iff $T$
-  can be partitioned into vertex disjoint subtrees $\{T_1, T_2, \dots
-  T_r\}$ such that, for each $1 \leq i \leq r$, the set of prime
-  sub-matrices corresponding to vertices in $X_i$ has an ICPPA in
-  $T_i$.
-\end{lemma}
-\begin{proof}
-  Let us consider the reverse direction first.  Let us assume that $T$
-  can be partitioned into $T_1, \ldots, T_r$ such that for each $1
-  \leq i \leq r $, the set of prime sub-matrices corresponding to
-  vertices in $X_i$ has an ICPPA in $T_i$.  It is clear from the
-  properties of the partial order $\preccurlyeq$ that these ICPPAs
-  naturally yield an ICPPA of $M$ in $T$.  The main property used in
-  this inference is that for each $1 \leq i \neq j \leq r$, $supp(X_i)
-  \cap supp(X_j) = \phi$.
-
-\noindent
-To prove the forward direction, we show that if $M$ has an ICPPA, say
-$\cA$, in $T$, then there exists a partition of $T$ into vertex
-disjoint subtree $T_1, \ldots, T_r$ such that for each $1 \leq i \leq
-r$, the set of prime sub-matrices corresponding to vertices in $X_i$
-has an ICPPA in $T_i$.  For each $1 \leq i \leq r$, we define based on
-$\cA$ a subtree $T_i$ corresponding to $X_i$.  We then argue that the
-trees thus defined are vertex disjoint, and complete the proof.
-Consider $X_i$ and consider the prime sub-matrix in $X_{i,0}$.
-Consider the paths assigned under $\cA$ to the sets in the prime
-sub-matrix in $X_{i,0}$.  Since the component in $G_f$ corresponding
-to this matrix is a connected component, it follows that union of
-paths assigned to this prime-submatrix is a subtree of $T$.  We call
-this sub-tree $T_i$.  All other prime-submatrices in $X_i$ are
-assigned paths in $T_i$ since $\cA$ is an ICPPA, and the support of
-other prime sub-matrices in $X_i$ are contained in the support of the
-matrix in $X_{i,0}$.  Secondly, for each $1 \leq i \neq j \leq r$,
-$supp(X_i) \cap supp(X_j) = \phi$, and since $\cA$ is an ICPPA, it
-follows that $T_i$ and $T_j$ are vertex disjoint.  Finally, since $|U|
-= |V(T)|$, it follows that $T_1, \ldots, T_r$ is a partition of $T$
-into vertex disjoint sub-trees such that for each $1 \leq i \leq r$,
-the set of matrices corresponding to nodes in $X_i$ has an ICPPA in
-$T_i$.  Hence the lemma.
-\end{proof}
-The essence of the following lemma is that an ICPPA only needs to be
-assigned to the prime sub-matrix corresponding to the root of each
-in-tree, and all the other prime sub-matrices only need to have an
-Intersection Cardinality Preserving Interval Assignments (ICPIA).
-Recall, an ICPIA is an assignment of intervals to sets such that the
-cardinality of an assigned interval is same as the cardinality of the
-interval, and the cardinality of intersection of any two sets is same
-as the cardinality of the intersection of the corresponding intervals.
-It is shown in \cite{nsnrs09} that the existence of an ICPIA is a
-necessary and sufficient condition for a matrix to have COP.  We
-present the pseudo-code to test if $M$ has an ICPPA in $T$.
-\begin{lemma} \label{lem:rooticppa} Let $M$ be a matrix and let $X$ be
-  the directed graph whose vertices are in correspondence with the
-  prime submatrices of $M$.  Further let $\{X_1,\ldots,X_r\}$ be the
-  partition of $X$ into in-trees as defined earlier in this section.
-  Let $T$ be the given tree and let $\{T_1, \ldots, T_r\}$ be a given
-  partition of $T$ into vertex disjoint sub-trees.  Then, for each $1
-  \leq i \leq r$, the set of matrices corresponding to vertices of
-  $X_i$ has an ICPPA in $T_i$ if and only if the matrix in $X_{i,0}$
-  has an ICPPA in $T_i$ and all other matrices in $X_i$ have an {\bf
-    {\em ICPIA}}.
-\end{lemma}
-\begin{proof}
-  The proof is based on the following fact- $\preccurlyeq$ is a
-  partial order and $X$ is a digraph which is the disjoint union of
-  in-trees.  Each edge in the in-tree is a containment relationship
-  among the supports of the corresonding sub-matrices. Therefore, any
-  ICPPA to a prime sub-matrix that is not the root is contained in a
-  path assigned to the sets in the parent matrix.  Consequently, any
-  ICPPA to the prime sub-matrix that is not at the root is an ICPIA,
-  and any ICPIA can be used to construct an ICPPA to the matrices
-  corresponding to nodes in $X_i$ provided the matrix in the root has
-  an ICPPA in $T_i$.  Hence the lemma.
-\end{proof}
-Lemma \ref{lem:subicppa} and Lemma \ref{lem:rooticppa} point out two
-algorithmic challenges in finding an ICPPA for a given set system
-$\cF$ in a tree $T$.  Given $\cF$, finding $X$ and its partition
-$\{X_1,\ldots,X_r\}$ into in-trees can be done in polynomial time.  On
-the other hand, as per lemma \ref{lem:subicppa} we need to parition
-$T$ into vertex disjoint sub-trees $\{T_1, \ldots, T_r\}$ such that
-for each $i$, the set of matrices corresponding to nodes in $X_i$ have
-an ICPPA in $T_i$.  This seems to be a challenging step, and it must
-be remarked that this step is easy when $T$ itself is a path, as each
-individual $T_i$ would be sub-paths.  The second algorithmic challenge
-is identified by lemma \ref{lem:rooticppa} which is to assign an ICPPA
-from a given tree to the matrix associated with the root node of
-$X_i$.
-\begin{algorithm}[h]
-  \caption{Algorithm to find an ICPPA for a matrix $M$ on tree $T$:
-    $main\_ICPPA(M, T$)}
-  \label{al:icppa-main}
-  \begin{algorithmic}
-    \STATE Identify the prime sub-matrices. This is done by constructing the strict overlap graph and identify connected components.  Each connected component yields a prime sub-matrix.   \\
-    \STATE Construct the partial order $\preccurlyeq$ on the set of prime sub-matrices.  \\
-    \STATE Construct the partition $X_1,\ldots,X_r$ of the prime
-    sub-matrices induced by $\preccurlyeq$ \\
-    \STATE For each $1 \leq i \leq r$, Check if all matrices except
-    those in $X_{i,0}$ has an ICPIA.  If a matrix does not have ICPIA
-    exit with a negative answer.  To check for the existence of ICPIA,
-    use the result in \cite{nsnrs09}.  \STATE Find a partition of
-    $T_1, \ldots, T_r$ such that matrices in $X_{i,0}$ has an ICPPA in
-    $T_i$.  If not such partition exists, exit with negative answer.
-  \end{algorithmic}
-\end{algorithm}
-
-\subsection{Complexity of Tree Path Assignment-A
-  Discussion} 
-\label{complexity} 
-Recall that the input to the Tree
-Path Assignment question is an order pair $(\cF,T)$ where $\cF$ is a
-family of subsets of an universe $U$, and $T$ is a tree such that
-$|V(T)| = |U|$.  The question is to come up with a bijection from $U$
-to $V(T)$ such that the image of each set in $\cF$ is a path in $T$.
-We show that this problem is at least as hard as the problem of
-testing if two given path chordal graphs are isomorphic.
-\begin{theorem} \label{thm:isocomp} Tree Path Assignment is
-  isomorphism-complete.
-\end{theorem}
-\begin{proof}
-  It is well known (see for example \cite{kklv10}) that testing
-  isomorphism of path chordal graphs is isomorphism complete.  We show
-  a reduction of path chordal graph isomorphism to tree path
-  assignment.  Given $G_1$ and $G_2$ two path chordal graphs, let
-  $T_2$ be the clique tree of $G_2$ obtained from say \cite{gav78}.
-  The nodes of $T_2$ correspond to the maximal cliques of $G_2$ and
-  each vertex of $G_2$ corresponds to a path in $G_2$.  This is a
-  well-known characterization of path chordal graphs and $T_2$ can be
-  computed in polynomial time.  In $G_1$, let $S_v$ denote the maximal
-  cliques of $G_1$ that contain $v$.  This can be computed in
-  polynomial time as $G_1$ is path chordal, and all chordal graph only
-  have a linear number of maximal cliques.  The universe $U$
-  corresponds bijectively to the set of maximal cliques in $G_1$, and
-  $\cF = \{S_v | v \in V(G_1)\}$.  Now, we claim that $(\cF, T_2)$ has
-  a tree path assignment if and only if $G_1$ and $G_2$ are
-  isomorphic.  This is clear since for each vertex $v \in G_1$, there
-  is an associated $S_v$ which is the set of maximal cliques
-  containing $v$.  In $G_2$, each vertex corresponds to a path in
-  $T_2$, and the nodes on this path corresponds to the maximal cliques
-  in $G_2$.  Consequenlty, a tree path assignment will naturally yield
-  an isomorphism between $G_1$ and $G_2$, and vice versa.  Therefore,
-  Tree Path Assignment is isomorphism-complete.
-\end{proof}
-
-\subsubsection{Consecutive Ones Testing is in Logspace}
-While Tree Path Assignment is isomorphism-complete, it is polynomial
-time solvable when the given tree is a path.  Indeed, in this case we
-encounter a restatement of matrices with the COP.  The known
-approaches to testing for COP fall into two categories: those that
-provide a witness when the input matrix does not have COP, and those
-that do not provide a witness.  The first linear time algorithm for
-testing COP for a binary matrix was using a data structure called PQ
-trees, which represent all COP orderings of $M$, invented by
-\cite{bl76}. There is a PQ tree for a matrix iff the matrix has COP.
-Indeed, this is an algorithmic characterization of the consecutive
-ones property and the absence of the PQ-tree does not yield any
-witness to the reason for failure.  A closely related data structure
-is the generalized PQ tree in \cite{mcc04}.  In generalized PQ tree
-the P and Q nodes are called prime and linear nodes. Aside from that,
-it has a third type of node called degenerate nodes which is present
-only if the set system does not have COP \cite{mcc04}.  Using the idea
-of generalized PQ tree, \cite{mcc04} proves that checking for
-bipartiteness in the certain incomparability graph is sufficient to
-check for COP.  \cite{mcc04} invented a certificate to confirm when a
-binary matrix does not have COP.  \cite{mcc04} describes a graph
-called incompatibility graph of a set system $\cF$ which has vertices
-$(a,b), a \ne b$ for every $a, b \in U$, $U$ being the universe of the
-set system. There are edges $((a,b),(b,c))$ and $((b,a),(c,b))$ if
-there is a set $S \in \cF$ such that $a, c \in S$ and $b \notin S$. In
-other words the vertices of an edge in this graph represents two
-orderings that cannot occur in a consecutive ones ordering of $\cF$.
-\begin{theorem}[Theorem 6.1, \cite{mcc04}]
-  Let $\cF$ be an arbitrary set family on domain $V$. Then $\cF$ has
-  the consecutive ones property if and only if its incompatibility
-  graph is bipartite, and if it does not have the consecutive ones
-  property, the incompatibility graph has an odd cycle of length at
-  most $n+3$.
-\end{theorem}
-This theorem gives a certificate as to why a given matrix does not
-have COP.  Similarly, the approach of testing for an ICPIA in
-\cite{nsnrs09} also gives a different certificate- a prime sub-matrix
-that does not have an ICPIA.  Further, the above theorem can be used
-to check if a given matrix has COP in logspace by checking if its
-incompatibility graph is bipartite. \cite{rei84} showed that checking
-for bipartiteness can be done in logspace. Thus we conclude that
-consecutive ones testing can be done in logspace.
-
-\noindent
-More recently, \cite{kklv10} showed that interval graph isomorphism
-can be done in logspace. Their paper proves that a canon for interval
-graphs can be calculated in logspace using an interval hypergraph
-representation of the interval graph with each hyperedge being a set
-to which an interval shall be assigned by the canonization
-algorithm. An overlap graph (subgraph of intersection graph, edges
-define only strict intersections and no containment) of the hyperedges
-of the hypergraph is created and canons are computed for each overlap
-component. The overlap components define a tree like relation due to
-the fact that two overlap components are such that either all the
-hyperedges of one is disjoint from all in the other, or all of them
-are contained in one hyperedge in the other. This is similar to the
-containment tree defined in \cite{nsnrs09} and in this paper. Finally
-the canon for the whole graph is created using logspace tree
-canonization algorithm from \cite{sl92}. The interval labelling done
-in this process of canonization is exactly the same as the problem of
-assigning feasible intervals to a set system, and thus the problem of
-finding a COP ordering in a binary matrix \cite{nsnrs09}.
-
-\begin{theorem}[Theorem 4.7, \cite{kklv10}]
-  \label{th:canonlabel}
-  Given an interval hypergraph $\cH$, a canonical interval labeling
-  $l_H$ for $H$ can be computed in FL.
-\end{theorem}
-
-We present the following reduction to see that COP testing is indeed
-in logspace. Given a binary matrix $M$ of order $n \times m$, let $S_i
-= \{j \mid M[j,i]=1 \}$. Let $\cF = \{S_i \mid i \in [m] \}$ be this
-set system. Construct a hypergraph $\cH$ with its vertex set being
-$\{1, 2, \dots n\}$. The edge set of $\cH$ is isomorphic to
-$\cF$. Thus every edge in $\cH$ represents a set in the given set
-system $\cF$. Let this mapping be $\pi: E(\cH) \rightarrow \cF$. It is
-easy to see that if $M$ has COP, then $\cH$ is an interval
-hypergraph. From theorem \ref{th:canonlabel}, it is clear that the
-interval labeling $l_{\cH}: V(\cH) \rightarrow [n]$ can be calculated
-in logspace. Construct sets $I_i = \{ \l_{\cH}(x) \mid x \in E, E \in
-E(\cH), \pi(E) = S_i\}$, for all $i \in [m]$. Since $\cH$ is an
-interval hypergraph, $I_i$ is an interval for all $i \in [m]$, and is
-the interval assigned to $S_i$ if $M$ has COP.
-
-Now we have the following corollary.
-\begin{corollary}
-  \label{cor:coplog}
-  If a binary matrix $M$ has COP then the interval assignments to each
-  of its columns can be calculated in FL.
-\end{corollary}
-
-\noindent
-Finally, we conclude by asking about the complexity of Tree Path
-Assignment restricted to other subclasses of trees.  In particular, is
-Tree Path Assignment in caterpillars easier than Tree Path assignment
-in general trees.
-
-\chapter{Conclusion}
-\label{ch:conclusion}
-
-\begin{center}
-{\tt MOVED FROM INTRO SUBSECTION IN ch:myresearch CHAPTER begin}
-\end{center}
-
-% \annote{
-  It is an interesting fact that for a matrix with the COP, the
-  intersection graph of the corresponding set system is an interval
-  graph. % }{the following text could be in Further Study/Conclusion}
-A
-similar connection to a subclass of chordal graphs and a superclass of
-interval graphs exists for the generalization of COP.  In this case,
-the intersection graph of the corresponding set system must be a {\em
-  path graph}. Chordal graphs are of great significance, are
-extensively studied, and have several applications.  One of the well
-known and interesting properties of a chordal graphs is its connection
-with intersection graphs \cite{mcg04}. For every chordal graph, there
-exists a tree and a family of subtrees of this tree such that the
-intersection graph of this family is isomorphic to the chordal graph
-\cite{plr70,gav78,bp93}.  These trees when in a certain format, are
-called {\em clique trees} \cite{apy92} of the chordal graph. This is a
-compact representation of the chordal graph. There has also been work
-done on the generalization of clique trees to clique hypergraphs
-\cite{km02}.  If the chordal graph can be represented as the
-intersection graph of paths in a tree, then the graph is called path
-graph \cite{mcg04}.  Therefore, it is clear that if there is a
-bijection from $U$ to $V(T)$ such that for every set, the elements in
-it map to vertices of a unique path in $T$, then the intersection
-graph of the set system is indeed a path graph.  However, this is only
-a necessary condition and can be checked efficiently because path
-graph recognition is polynomial time
-solvable\cite{gav78,aas93}. Indeed, it is possible to construct a set
-system and tree, such that the intersection graph is a path graph, but
-there is no bijection between $U$ and $V(T)$ such that the sets map to
-paths. Path graph isomorphism is known be isomorphism-complete, see
-for example \cite{kklv10}. An interesting area of research would be to
-see what this result tells us about the complexity of the tree path
-labeling problem (not covered in this paper).
-
-\begin{center}
-{\tt MOVED FROM INTRO SUBSECTION IN ch:myresearch CHAPTER\\ end}
-\end{center}
-
-We give a characterization for feasible tree path labeling of path
-hypergraphs. The proof for this is constructive and computes a
-feasibility bijection mapping vertices of the hypergraph to vertices
-of the given target tree. This thesis also discovered an exponential algorithm
-that computes a feasible TPL to a given hypergraph if it is a path hypergraph.
-
-Our results have close connections to recognition of path graphs and
-to path graph isomorphism.  Graphs which can be represented as the
-intersection graph of paths in a tree are called {\em path
-  graphs}\cite{mcg04}. Thus, a hypergraph $\cF$ can be interpreted as
-paths in a tree, if and only if the intersection graph of $\cF$ is a
-{\em path graph}. Path graphs are a subclass of chordal graphs since
-chordal graphs are characterized as the intersection graphs of
-subtrees of a tree\cite{mcg04}.  Path graphs are well studied in the
-literature \cite{plr70}--\cite{mcg04}\tnote[TH]{Check if cite:bp93 is
-  indeed for path graphs}.  Path graph recognition can be done in
-polynomial time\cite{gav78,aas93}.  Clearly, this is a necessary
-condition in terms of the intersection graph of the input hypergraph
-$\cF$ in \FTPL. However, one can easily obtain a counterexample to
-show the insufficiency of this condition\tnote[TH]{Give a figure
-  showing insuff.}.  Path graph isomorphism is known be
-isomorphism-complete\cite{kklv10}. Therefore, \annote[TH]{it is
-  unlikely that we can solve the problem of finding feasible path
-  labeling $\cl$ for a given $\cF$ and }{WHY?} tree $T$.  It is
-definitely interesting to classify the kinds of trees and hypergraphs
-for which feasible path labelings can be\tnote[TH]{ELABORATE.}  found
-efficiently.  \annote[TH]{These results would form a natural
-  generalization of COP testing and interval graph isomorphism,
-  culminating in Graph Isomorphism }{REWORD} itself.  To this effect
-we consider TPL on {\kstar} and give a polynomial time solution for
-the same.
-
-While we address the computation of TPL
-for a given hypergraph% \footnote{The term {\em hypergraph} is used
-%   instead {\em set systems} are used interchangably in this
-%   section. Hypergraphs that have TPLs on some target tree are called
-%   {\em path hypergraphs}.}
-from a given target tree in this research, however optimization
-problems in TPL, akin to problems in Section~\ref{sec:optcop} for COP,
-is outside the scope of this thesis. Whether TPL on general trees is
-solvable in $P$ remains open. So do optimization opportunities in TPL
-(possible extensions to optimization for COP in the
-Section~\ref{sec:optcop}.\tnote[TH]{link to section}).
-
-%\newpage %DUMMY
-\tnote[TH]{ ||Complexity challenges:|| Canonization is an important
-  tool in graph isomorphism. Invention of a
-% deterministic method of canonization for any class of graphs naturally
-% results in an algorithm for isomorphism. All that is required is to
-% check if the canons of two graphs are the same. Thus complexity of
-% graph isomorphism can be studied by studying canonization
-% methods. While general graph ismorphism remains elusive in terms of
-% hardness, canonization has been studied for smaller classes of graphs
-% thus giving complexity/hardness results for them.
-
-% In 1992, \cite{sl92} made an important discovery that tree isomorphism
-% is in logspace. It was done by inventing a method of canonization of
-% trees using a logspace depth first traversal
-% algorithm. \tnote{confirm the statement of claim}.
-
-% \cite{kklv10} proved that interval graph canonization is also in
-% logspace thus drawing logspace conslusions about COT. 
-% % This result is
-% % built on the fact that 
-% Interval graphs are FP+C (fixed point with
-% counting) definable \tnote{elucidate} and \cite{l10} showed that
-% this implies that it captures PTIME. \tnote{elucidate} This result
-% along with that of undirected graph reachability being in logspace
-% \cite{rei08}, \cite{kklv10} proved their logspace result.
-}   %
-\tnote[TH]{ (1) the interesting survey in conclusion section of kklv10
-  (2) [[[COP helping some problems with hardness]]] having COP in the
-  input structure makes some problems less hard than in a general
-  input. -- min set cover problem is known to be NP hard. so is min
-  hitting
-  set. these two problems are equivalent cite-adp80\\
-  -- -- complexity results: log factor polynomial time approximable cite-f98, W[2]-complete with solution size as parameter cite-df99\\
-  -- a couple of generalizations:\\
-  -- -- min degree hypergraph (MDH) problem\\
-  -- -- red blue covering (RBSC) problem\\
-  ----- cite-d08phd Chapter 5.\\
-  ----- what are cite-d08phd results\\
-  ----- Rectangle stabbing \\
-  ----- see cite-d08phd Chapter 6?  }
+\include{010chap1}
+\include{020chap2}
+\include{030chap3}
+\include{040chap4}
 
@@
 % Appendices.
+%
+% \appendix %%% NON NEIN NAHI ILLA NONE
+% \chapter{A}
+% \newpage
 
-% \appendix %%% NON NEIN NAHI ILLA NONE
-
-% \chapter{A}
-% \tnote{DO WE NEED THIS?}
-
-\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Bibliography.
+%
 \begin{singlespace}
-%\begin{thebibliography}{10}
-\addcontentsline{toc}{chapter}{REFERENCES}
+  \addcontentsline{toc}{chapter}{REFERENCES \tnote[minor]{make names
+    in bib file uniform style w.r.t. firstname/initials}}
 
-{\footnotesize
-  \bibliographystyle{alpha} 
-  \bibliography{../lib/cop-variants__thesis}
+  {\footnotesize%
+    \bibliographystyle{alpha}%
+    \bibliography{../lib/cop-variants__thesis} %
   }
-%\end{thebibliography}
 \end{singlespace}
-%\newpage
 
@@
 % List of papers
-
+%
 % \listofpapers %%% NON NEIN NAHI ILLA NONE
-
 % \begin{enumerate}
-% \item Authors....  \newblock
-%  Title...
-%   \newblock {\em Journal}, Volume,
-%   Page, (year).
+% \item Authors....  \newblock Title...  \newblock {\em Journal},
+%   Volume, Page, (year).
 % \end{enumerate}

### 2012-01-04 src/020chap2.tex
Reordered chrono.
Added amsthm and changed theorem environs to comply.
Added atbeginend commands to space theorem environs.
Added a few \def strings.

Added placeholders for a couple of definitions.

Started editing. Intro. Added a couple of definitions.

Minor edits to comply with amsthm

Spacing hacks mostly.

Trivial.

@@
-\chapter{Consecutive-ones Property - a Survey of Important Results}
+\chapter{\CoP\,-- A Survey of Important Results}
 \label{ch:copsurvey}
-\tnote{[i] TBD: have a few lines about organization of chapter\ \ \ \
-  \ [ii] Survey chapter: is the full fledged expansion of the survey in
-  introduction with details, observations, theorems etc.}
 
-\section*{-- REAL CONTENT --}
+This chapter surveys several results that are significant to this
+thesis or to \COP in general. These predominantly pertain to
+characterizations of \COP, algorithmic tests to check for \COP (COT),
+optimization problems on binary matrices that do not have \COP and
+some applications of \COP.
+
+\tnote[important]{ADD: have a few lines about organization of chapter}
+
+\tnote{DEFINE COP SOMEWHERE}
+
+\section{\COP in Graph Theory}
+
+\COP is closely connected to several types of graphs by way of
+describing certain combinatorial graph properties. There are also
+certain graphs, like convex bipartite graphs, that are defined solely
+by some of its associated matrix having \COP.  In this section we will
+see the relevance of \cop to graphs.  To see this we introduce certain
+binary matrices that are used to define graphs in different
+ways. While adjacency matrix is perhaps the most commonly used such
+matrix, Definition~\ref{def:graphmatrices} defines this and a few
+more.
+
+\begin{definition}{\emph{Matrices that define
+      graphs.\cite[Def.~2.4]{d08phd}}} %
+  \label{def:graphmatrices} %
+  Let $G$ and $H$ be defined as follows. $G = (V,E_G)$ is a graph with
+  vertex set $V = \{v_i \mid i \in [n]\}$ and edge set $E_G \subseteq
+  \{(v_i,v_j) \mid i, j \in [n]\}$ such that $|E_G| = m$. $H = (A, B,
+  E_H)$ is a bipartite graph with partitions $A = \{a_i \mid i \in
+  [n_a]\}$ and $B = \{b_i \mid i \in [n_b]\}$.
+  \begin{enumerate}[{\ref{def:graphmatrices}}--i.]
+  \item \emph{Adjacency matrix} of $G$ is the symmetric $n \times n$
+    binary matrix $M$ with $m_{i,j} = \un$ \iff $(v_i,v_j) \in E_G$
+    for all $i,j \in [n]$.
+  \item \emph{Augmented adjacency matrix} of $G$ is obtained from its
+    adjacency matrix by setting all main diagonal elements to \un, \ie
+    $m_{i,i} = \un$ for all $i \in [n]$.
+  \item \emph{Maximal clique matrix} or \emph{vertex-clique incidence
+      matrix} of $G$ is the $n \times k$ binary matrix $M$ with
+    $m_{i,j} = \un$ \iff $v_i \in C_j$ for all $i \in [n], j \in [k]$
+    where $\{C_j \mid j \in [k]\}$ is the set of maximal cliques of
+    $G$.
+  \item \emph{Half adjacency matrix} of $H$ is the $n_a \times n_b$
+    binary matrix $M$ with $m_{i,j} = \un$ \iff $(a_i, b_j) \in E_H$
+  \end{enumerate}
+\end{definition}
+
+Now we will see in Definition~\ref{def:graphwithcop} certain graph
+classes that is related to \COP or \CROP.\\
+
+%\hrule
+\begin{definition}{\emph{Graphs that relate to
+      COP.\cite[Def.~2.5]{d08phd}}} %
+   \label{def:graphwithcop} %
+   Let $G$ be a graph and $H$ be a bipartite graph.
+
+   \begin{enumerate}[\ref{def:graphwithcop}--i.]
+   \item $G$ is \emph{convex-round} if its adjacency matrix has the
+     \CROP.
+   \item $G$ is \emph{concave-round} if its augmented adjacency matrix
+     has \CROP. \tnote{cite BHY00} \tnote[minor]{add CROP to glossary}
+   \item $G$ is an \emph{interval graph} if its vertices can be mapped
+     to intervals on the real line \stt two vertices are adjacent \iff
+     their corresponding intervals overlap \tnote{cite Ben59, Haj57}.
+     $G$ is an interval graph \iff its maximal clique matrix has COP
+     due to the result by \cite{gh64} which states that the maximal
+     cliques of interval graph $G$ can be linearly ordered \stt for
+     all $v \in V(G)$, cliques containing $v$ are consecutive in the
+     ordering.
+   \item $G$ is a \emph{unit interval graph} if it is an interval
+     graph \stt all intervals have the same length.
+   \item $G$ is a \emph{proper interval graph} if it is an interval
+     graph \stt no interval properly contains another.
+   \item $G$ is a \emph{circular-arc graph} if its vertices can be
+     mapped to a set of arcs on a circle \stt two vertices are
+     adjacent \iff their corresponding arcs overlap.
+   \item $H$ is \emph{convex bipartite} if its half adjacency matrix
+     has \COP on either rows or columns.
+   \item $H$ is \emph{biconvex bipartite} or \emph{doubly
+       convex}\cite{yc95} if its half adjacency matrix has \COP on
+     both rows and columns.
+   \item $H$ is \emph{circular convex} if its half adjacency matrix
+     has \CROP.
+   \end{enumerate}
+\end{definition}
+%\hrule
 

### 2012-01-04 src/030chap3.tex
Reordered chrono.
Added amsthm and changed theorem environs to comply.
Added atbeginend commands to space theorem environs.
Added a few \def strings.

Added placeholders for a couple of definitions.

Started editing. Intro. Added a couple of definitions.

Minor edits to comply with amsthm

Spacing hacks mostly.

Trivial.

@@
 
-\begin{enumerate}[{(\icpplpr\ }i) \ \ \ ]
-\item \label{pr:i} $|S| = |\cl(S)|$ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $S \in \cF$
+% \begin{enumerate}[{(\icpplpr\ }i) \ \ \ ]
+% \item \label{pr:i} $|S| = |\cl(S)|$ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
+%   \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $S \in \cF$
+%   % \vspace{\topshrink}
+% \item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ \ \ \ \
+%   \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all distinct $S_1, S_2 \in \cF$
+%   % \vspace{\topshrink}
+% \item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
+%   \cap \cl(S_3)|$ \ \ \ for all distinct $S_1, S_2, S_3 \in \cF$
+% \end{enumerate}
+\begin{enumerate}[{(\icpplpr\ }i)]
+\item \label{pr:i} $|S| = |\cl(S)|$ for all $S \in \cF$
   % \vspace{\topshrink}
-\item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ \ \ \ \
-  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all distinct $S_1, S_2 \in \cF$
+\item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ for all distinct $S_1, S_2 \in \cF$
   % \vspace{\topshrink}
 \item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
-  \cap \cl(S_3)|$ \ \ \ for all distinct $S_1, S_2, S_3 \in \cF$
+  \cap \cl(S_3)|$ for all distinct $S_1, S_2, S_3 \in \cF$
 \end{enumerate}
@@
   \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3| = |P_1 \cap (P_2
-  \setminus P_3)|$. Thus lemma is proven. \qed
+  \setminus P_3)|$. Thus lemma is proven. %\qed
 \end{proof}
@@
   \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Thus the induced path labeling
-  $\cl_{\phi[{\cF_{j}}]} = \cl_{j}$. Therefore lemma is proven.  \qed
+  $\cl_{\phi[{\cF_{j}}]} = \cl_{j}$. Therefore lemma is proven.  %\qed
 \end{proof}
@@
   \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
-  = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. \qed
+  = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. %\qed
 \end{proof}
@@
   case cannot occur and case 1 is the only possible scenario. \\
-  Thus lemma is proven \qed
+  Thus lemma is proven %\qed
 \end{proof}
@@
   sets, invariant IV is proven by also using lemma
-  \ref{lem:setminuscard}. \qed
+  \ref{lem:setminuscard}. %\qed
 
@@
     % \end{enumerate}
-  \end{enumerate} \qed
+  \end{enumerate} %\qed
   % \vspace{-6mm}
@@
   loop.  Due to Lemma~\ref{lem:invar1} Invariant IV, the exit
-  condition does not occur if the input is a feasible ICPPL.\qed
+  condition does not occur if the input is a feasible ICPPL.%\qed
 
@@
   % Thus the exit condition is never rendered true after $x$ and $v$
-  % are removed from their respective hyperedges. \qed
+  % are removed from their respective hyperedges. %\qed
 
@@
   % \end{enumerate}
-  % \qed \tnote[E2]{remove the induction proof. just text saying x and
+  % %\qed \tnote[E2]{remove the induction proof. just text saying x and
   %   v are exclusive in these sets therefore the intersection
@@
   \setminus \{v_1\} \cap Q'' \setminus \{v_2\} \cap Q_3| = |Q_1 \cap
-  Q_2 \cap Q_3|$. Thus invariant IV is also proven.  \qed
+  Q_2 \cap Q_3|$. Thus invariant IV is also proven.  %\qed
 \end{proof}
@@
 .. \cup \cL_{d} \cup \sigma'$ gives the bijection required in the
-original problem.\qed
+original problem.%\qed
 \end{proof}
@@
   feasible. Thus $X$ cannot be empty if input is ICPPL and feasible.
-  \qed
+  %\qed
 \end{proof}
@@
   II, III, IV remain true. On the other hand, if the input ICPPL is
-  not feasible the invariants are vacuously true. \qed
+  not feasible the invariants are vacuously true. %\qed
 \end{proof}
@@
   a bijection from $supp(\cF)$ to $V(T)$ inducing the given path
-  labeling $\cl$ and thus a hypergraph isomorphism. \qed
+  labeling $\cl$ and thus a hypergraph isomorphism. %\qed
 \end{proof}
@@
   Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
-  function.\qed
+  function.%\qed
 \end{proof}
@@
   inclusion chain which contradicts that assumption $X$ is
-  super-marginal. Thus the claim is proven.\qed
+  super-marginal. Thus the claim is proven.%\qed
 \end{proof}
@@
 %   \setminus P_3| = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is
-%   proven. \qed
+%   proven. %\qed
 % \end{proof}
@@
 %   case cannot occur and case 1 is the only possible scenario. \\
-%   Thus lemma is proven \qed
+%   Thus lemma is proven %\qed
 % \end{proof}
@@
 %   the other new sets, invariant IV is proven by also using lemma
-%   \ref{lem:setminuscard}. \qed
+%   \ref{lem:setminuscard}. %\qed
 
@@
 %   \setminus \{v_1\} \cap Q'' \setminus \{v_2\} \cap Q_3| = |Q_1 \cap
-%   Q_2 \cap Q_3|$. Thus invariant IV is also proven.  \qed
+%   Q_2 \cap Q_3|$. Thus invariant IV is also proven.  %\qed
 % \end{proof}
@@
 %   $\cL_0 \cup \cL_1 \cup .. \cup \cL_{d} \cup \sigma'$ gives the
-%   bijection required in the original problem.\qed
+%   bijection required in the original problem.%\qed
 % \end{proof}

### 2012-01-04 src/020chap2.tex
Added figure illustrating graph matrix definitions. Also - this is going SO slow. Fastfowrding on.

@@
 \end{definition}
+\qed
+
+\begin{figure}[t]
+  \centering
+  \begin{tabular}[h]{cccl}
+    \figtabsize
+    ${\bf G_1}$: 
+    \includegraphics[scale=0.2]{../img/ex1_concaveroundgr.pdf} &
+    ${\bf G_2}$: 
+    \includegraphics[scale=0.2]{../img/ex2_intervalgr.pdf} &
+    ${\bf H}$: 
+    \includegraphics[scale=0.2]{../img/ex2_intervalgr.pdf} &
+    \begin{tabular}[h]{l|llll}
+      % ${\bf B}$:
+      % &&&&\\
+      ${\bf C}$ &$w_1$ &$w_2$ &$w_3$ &$w_4$\\
+      \hline
+      $u_1$ &\un   &\un   &0     &0   \\
+      $u_2$ &\un   &\un   &\un   &\un \\
+      $u_3$ &0     &0     &\un   &\un \\
+      $u_4$ &0     &\un   &\un   &0   \\
+      $u_5$ &\un   &\un   &\un   &0   
+    \end{tabular}
+  \end{tabular}\\
+
+  \begin{tabular}{lll} 
+    \begin{tabular}[h]{l|lll lll}
+      % ${\bf A_1}$:  
+      %&&&&&&\\
+      ${\bf A_1}$ &$v_1$ &$v_2$ &$v_3$ &$v_4$ &$v_5$ &$v_6$\\
+      \hline
+      $v_1$     &0     &\un   &\un  &0     &\un   &\un \\
+      $v_2$     &\un   &0     &\un  &\un   &0     &\un \\
+      $v_3$     &\un   &\un   &0    &\un   &\un   &0   \\
+      $v_4$     &0     &\un   &\un  &0     &0     &0   \\
+      $v_5$     &\un   &0     &\un  &0     &0     &0   \\
+      $v_6$     &\un   &\un   &0    &0     &0     &0   
+    \end{tabular}
+    & 
+    \begin{tabular}[h]{l|llll}
+      % ${\bf B}$:  
+      %&&&&\\
+      ${\bf B}$&$c_1$ &$c_2$ &$c_3$ &$c_4$\\
+      \hline
+      $v_1$ &\un   &\un   &\un   &0   \\
+      $v_2$ &\un   &0     &0     &0   \\
+      $v_3$ &0     &\un   &0     &0   \\
+      $v_4$ &0     &0     &\un   &\un \\
+      $v_5$ &0     &\un   &\un   &0   \\
+      $v_6$ &0     &0     &0     &\un  
+    \end{tabular}
+    & 
+    \\    
+    && \\
+    \begin{tabular}[h]{l|llllll}
+      % ${\bf A_2}$:  
+      %&&&&&&\\
+      ${\bf A_2}$&$v_1$ &$v_2$ &$v_3$ &$v_4$ &$v_5$ &$v_6$\\
+      \hline
+      $v_1$ &\un   &\un   &\un  &0     &\un   &\un \\
+      $v_2$ &\un   &\un   &\un  &\un   &0     &\un \\
+      $v_3$ &\un   &\un   &\un  &\un   &\un   &0   \\
+      $v_4$ &0     &\un   &\un  &\un   &0     &0   \\
+      $v_5$ &\un   &0     &\un  &0     &\un   &0   \\
+      $v_6$ &\un   &\un   &0    &0     &0     &\un 
+    \end{tabular}    
+    &
+    \begin{tabular}[h]{l|llllll}
+      % ${\bf A_3}$:  
+      %&&&&&&\\
+      ${\bf A_2'}$ &$v_1$ &$v_6$ &$v_2$ &$v_4$ &$v_3$ &$v_5$\\
+      \hline
+      $v_1$ &\un   &\un   &\un  &0     &\un   &\un \\
+      $v_6$ &\un   &\un   &\un  &0     &0     &0   \\
+      $v_2$ &\un   &\un   &\un  &\un   &\un   &0   \\
+      $v_4$ &0     &0     &\un  &\un   &\un   &0   \\
+      $v_3$ &\un   &0     &\un  &\un   &\un   &\un \\
+      $v_5$ &\un   &0     &0    &0     &\un   &\un 
+    \end{tabular}
+   & \\
+
+% ex3_convexbipartitegr.png  
+  \end{tabular}
+
+  \caption[\figtabsize Matrices defined in
+  Def.~\ref{def:graphmatrices}]{\figtabsize $A_1$ is the
+    {\em adjacency matrix} and $A_2$ is the {\em augmented adjacency matrix} of
+    $G_1$. $A_2'$ is obtained from $A_2$ by permuting its rows
+    and columns to achieve {\em \CROP order}, \ie $A_2$ has
+    \CROP\,-- thus $G_1$
+    is {\em a concave-round graph} (Def.~\ref{def:graphwithcop}~\ref{def::concave-round})
+    and {\em a circular-arc graph} (Tab.~\ref{tab:graphmatrices})
+    $B$ is the maximal clique matrix of $G_2$ and has \COP\,-- thus $G_2$ is an
+    {\em interval graph}(Def.~\ref{def:graphwithcop}~\ref{def::concave-round}). $C$ is
+    the half adjacency matrix of bipartite graph $H$ and has \COP on
+    rows -- thus $H$ is {a convex bipartite graph}.} 
+  \label{fig:graphmatrices}
+\end{figure}
 
@@
 classes that is related to \COP or \CROP.\\
+
+{\tt tab~\ref{tab:graphmatrices} -}   \tnote{have an abridged version of table 2.1 in dom}
+
+\begin{table}[htbp]
+  \centering
+ \caption{\figtabsize **** Graph matrices}
+  \label{tab:graphmatrices}
+\end{table}
 
@@
      \CROP.
-   \item $G$ is \emph{concave-round} if its augmented adjacency matrix
+   \item \label{def::concave-round} $G$ is \emph{concave-round} if its augmented adjacency matrix
      has \CROP. \tnote{cite BHY00} \tnote[minor]{add CROP to glossary}
@@
 \end{definition}
+\qed
 %\hrule

### 2012-01-05 src/010chap1.tex
Reordered chrono.
Added amsthm and changed theorem environs to comply.
Added atbeginend commands to space theorem environs.
Added a few \def strings.

Added placeholders for a couple of definitions.

Started editing. Intro. Added a couple of definitions.

Minor edits to comply with amsthm

Spacing hacks mostly.

Trivial.
Chapter 2 edited. See diff for details. Too many to remember. or maybe too little.

@@
 
+\begin{definition}{\emph{\Cop}}
+  \tnote[important]{define cop here. refer to it in chap 2 or
+    duplicate it using slash def }
+  \begin{enumerate}
+  \item \Cop
+  \item Strong \cop
+  \item \Cop order or \cop permutation
+  \end{enumerate}
+\end{definition}
+
+\begin{figure}[t] %[htbp]
+  \centering
+  {\figtabsize
+    \begin{tabular}[h]{l|lcccl}
+      $M_1$: & $M_1'$: &&&& $M_2$:\\
+      &&&&&\\
+      \begin{tabular}[h]{llll}
+        $c_1$ & $c_2$ &$c_3$ &$c_4$\\
+        &&&\\
+        \un & 0   & \un & 0\\
+        0   & \un & 0   & \un \\
+        \un & 0   & 0   & \un
+      \end{tabular}
+      &
+      \begin{tabular}[h]{llll}
+        $c_3$ &$c_1$ &$c_4$& $c_2$\\
+        &&&\\
+        \un & \un & 0 & 0\\
+        0 & 0 & \un & \un \\
+        0 & \un & \un & 0
+      \end{tabular}
+      &&&&
+      \begin{tabular}[h]{llll}
+        $d_1$ & $d_2$ &$d_3$ &$d_4$\\
+        % &&&\\
+        &&&\\
+        \un & \un & 0 & 0\\
+        0 & \un & \un & 0 \\
+        0 & \un & 0 & \un 
+      \end{tabular}
+    \end{tabular}
+  }
+
+  \caption[\figtabsize Matrices with and without COP.]{\figtabsize
+    Matrices with and without COP. $M_1$ has COP because by permuting
+    its columns, $c_1$-$c_4$, one can obtain $M_1'$ where the {\un}s
+    in each row are consecutive. $M_2$, however, does not have COP
+    since no permutation of its columns, $d_1$-$d_4$, will arrange
+    {\un}s in each row consecutively \cite{d08phd}.}
+
+  \label{fig:cop-matrix}
+\end{figure}
+
+\begin{definition}{\emph{\Crop}}
+  \tnote[important]{define circ1 here. refer to it in chap 2 or
+    duplicate it using slash def }
+\end{definition}
+
+\begin{definition}{\emph{Binary matrix}}
+  Let $M$ be an $n \times m$ matrix. $m_{i,j}$ denotes its $(i,j)$th
+  element, i.e. element at $i$th row and $j$th column. $M$ is a binary
+  matrix if each of its element is 0 or \un.
+  In other words, for all $i \in [n]$ and $j \in [m]$, $m_{i,j} \in
+  \{0, \un\}$ 
+\end{definition}
+
 \section{Consecutive-ones Property Testing - a Brief Survey}
@@
 submatrices came as a corollary to the discovery that convex bipartite
-graphs are AT-free in \cite{at72}\tnote{check}. The first linear
+graphs are AT-free on at least one of the partitions in
+\cite{at72}\tnote{check the real meaning of AT FREE. }. The first linear
 time algorithm for COP testing (COT) was invented by \cite{bl76} using
@@
 
-\section{********* Application of COP in Areas of Graph Theory and
+\section{Application of COP in Areas of Graph Theory and
   Algorithms}
@@
 
-\subsection{********* COP in Relational Database Model}
+\temptext{Combine COP in Relational Database Model + COP in Graph
+  Isomorphism + Certifying Algorithms}
+
+%\subsection{COP in Relational Database Model}
 \label{sec:apprdbm}
@@
 
-\subsection{********* COP in Graph Isomorphism}
+%\subsection{}
 \label{sec:appgraphiso}
@@
 
-\subsection{********* Certifying Algorithms}
+%\subsection{}
 \label{sec:appcertalgo}

### 2012-01-05 src/020chap2.tex
Chapter 2 edited. See diff for details. Too many to remember. or maybe too little.

--- 2012-01-04 src/020chap2.tex
@@
+\definecolor{gray}{rgb}{0.8,0.8,0.8}
+\def\tblhcolor {gray}
+
 \chapter{\CoP\,-- A Survey of Important Results}
@@
 
-\begin{definition}{\emph{Matrices that define
-      graphs.\cite[Def.~2.4]{d08phd}}} %
-  \label{def:graphmatrices} %
+% \begin{definition}{\emph{Matrices that define
+%       graphs.\cite[Def.~2.4]{d08phd}}} %
+\begin{definition}
+  {\em Matrices that define graphs. \cite[Def.~2.4]{d08phd}.}  
+  \label{def:graphmatrices} 
   Let $G$ and $H$ be defined as follows. $G = (V,E_G)$ is a graph with
@@
   \item \emph{Half adjacency matrix} of $H$ is the $n_a \times n_b$
-    binary matrix $M$ with $m_{i,j} = \un$ \iff $(a_i, b_j) \in E_H$
+    binary matrix $M$ with $m_{i,j} = \un$ \iff $(a_i, b_j) \in E_H$.
   \end{enumerate}
 \end{definition}
+
+\figgraphmatrices
 
@@
 
-%\hrule
 \begin{definition}{\emph{Graphs that relate to
@@
     \CROP.
-   \item $G$ is \emph{concave-round} if its augmented adjacency matrix
-     has \CROP. \tnote{cite BHY00} \tnote[minor]{add CROP to glossary}
+  \item \label{def::concave-round} $G$ is \emph{concave-round} if its
+    augmented adjacency matrix has \CROP. \tnote{cite BHY00}
+    \tnote[minor]{add CROP to glossary}
   \item $G$ is an \emph{interval graph} if its vertices can be mapped
@@
     $G$ is an interval graph \iff its maximal clique matrix has COP
-     due to the result by \cite{gh64} which states that the maximal
-     cliques of interval graph $G$ can be linearly ordered \stt for
-     all $v \in V(G)$, cliques containing $v$ are consecutive in the
-     ordering.
+    \cite{fg65}\endnote{This follows \cite{gh64} which states that the
+      maximal cliques of interval graph $G$ can be linearly ordered
+      \stt for all $v \in V(G)$, cliques containing $v$ are
+      consecutive in the ordering \cite[Th. 8.1]{mcg04}.}
+    \begin{enumerate}[a.]
     \item $G$ is a \emph{unit interval graph} if it is an interval
-     graph \stt all intervals have the same length.
+      graph \stt all intervals have the same length.\endnotemark[2]
     \item $G$ is a \emph{proper interval graph} if it is an interval
-     graph \stt no interval properly contains another.
+      graph \stt no interval properly contains another.\endnote{The
+        set of unit interval graphs and the set of proper interval
+        graphs are the same} \tnote{cite {rob69,gar07} in endnote. pg
+        33 dom}
+    \end{enumerate}
  \item $G$ is a \emph{circular-arc graph} if its vertices can be
-     mapped to a set of arcs on a circle \stt two vertices are
-     adjacent \iff their corresponding arcs overlap.
-   \item $H$ is \emph{convex bipartite} if its half adjacency matrix
-     has \COP on either rows or columns.
+    mapped to a set of arcs on a circle \stt two vertices are adjacent
+    \iff their corresponding arcs overlap.\tnote[pressing]{how is
+      CO/ROP related?}
+  \item \label{def::convexbi} $H$ is \emph{convex bipartite on columns (rows)} if its half adjacency matrix
+    has \COP on rows (columns).
   \item $H$ is \emph{biconvex bipartite} or \emph{doubly
-       convex}\cite{yc95} if its half adjacency matrix has \COP on
-     both rows and columns.
-   \item $H$ is \emph{circular convex} if its half adjacency matrix
-     has \CROP.
+      convex}\cite{yc95} if its half adjacency matrix has \COP on both
+    rows and columns.
+  \item $H$ is \emph{circular convex} if its half adjacency matrix has
+    \CROP.
   \end{enumerate}
 \end{definition}
-%\hrule
 
-%%%%%%                                              %%%%%%
-% [pressing]{\ \ \ \ \  is the second level of [pressing]%
-%%%%%%                                              %%%%%%
-\tnote[pressing]{ADD: As it will be described in detail later in this
-  document, isomorphism of certain % classes of graphs, namely chordal
-  % graphs, have a close relationship with consecutive-ones property
-  % and generalizations of it.  This is perhaps because of how closely
-  % COP of a matrix relates to properties of graphs derived from
-  % matrices as seen inz the following results.
-} \tnote[pressing]{\ \ \ \ \ ADD: peo exists iff chordal. lexicographic BFS
-  [tag:chordalGraph]} %
-\tnote[pressing]{\ \ \ \ \ ADD: A well known result in %perfect graph theory is that
-  % the maximal cliques of an interval graph $G$ can be linearly
-  % ordered such that for all $v \in V(G)$, cliques containing $v$ are
-  % consecutive in the ordering cite~{gh64}. This clearly means that a
-  % graph $G$ is an interval graph if and only if
-}%
-\tnote[pressing]{\ \ \ \ \ verify from paper the statement of claim.}%
-\tnote[pressing]{maximal clique vertex incidence matrix
-  of% $G$ has COP.  Also, maximal cliques of any
-  % chordal graph can be enumerated in polytime $O(m+n)$
-} \tnote[pressing]{\ \ \ \ \ citation?!!}%
-\tnote[pressing]{cite~{fg65} uses these results to give the first polynomial
-  time algorithm for COT.}%
-\tnote[pressing]{\ \ \ \ \ check. how do they use it?}%
-\tnote[pressing]{ A bipartite graph is
-  convex% (on $R$) if and only if its half adjacency matrix has COP
-  % on rows.  The results in cite~{bl76} on COT are based on the
-  % result that interval graphs are AT-free chordal graphs.
-}%
-\tnote[pressing]{\ \ \ \ \ the latter being Tucker's?}%
-\tnote[pressing]{\ \ \ \ \ (2) TBD survey
-  -- % see blue notes (in notebook) under Graph Isomorphism. namely
-  % citations in cite:aas93 (3) a brief on heirarchy of chordal
-  % graphs, path graphs, interval graphs, peo, clique tree etc. and
-  % results. (4) interval graphs are incomparability graphs. see
-  % Golumbic. [tag:classification] (5) finding min length hole in
-  % bipartite graph is polynomial time [Sec 3.3.2 in cite:d08phd] (6)
-  % theorem 2.2 in cite:d08phd - $G$ is union of v.d. caterpillars iff
-  % $M$ has COP in rows. $M$ is edge vertex incidence matrix $(*,2)$
-  % of $G$
-}%
+Interval graphs and circular-arc graphs have a long history in
+research. The interest around them is due to their very desirable
+property that several problems that are NP-hard\tnote[minor]{all NPC
+  problems are NPH yes?} on general graphs are not so in these graph
+classes, \ie they are polynomial time solvable on them. In a similar
+fashion, a lot of problems that are hard on general matrices have
+efficient solutions on matrics with \COP or \CROP \cite[more citations
+pg.\,33]{d08phd}.
+
+Table~\ref{tab:graphmatrices} summarises the way these graphs are
+characterized by their matrices having \COP or \CROP.
+Our focus in this chapter (and thesis) is mainly \COP and 
+having seen how useful \COP is in identifying or characterizing many
+types of graphs, we will now see results that study recognition of \COP
+in matrices in the following section.
+
+\temptext{ tab~\ref{tab:graphmatrices} -  have an abridged version
+  of table 2.1 in dom. see notes for the possibility of a
+  ``circle diagram''.} 
+\begin{table}[htbp]
+  \centering
+ \caption{\figtabsize Graph matrices \temptext{PLACEHOLDER}}
+  \label{tab:graphmatrices}
+\end{table}
+
+\clearpage      % Print pending floats
 
@@
 \label{sec:surveycoptest}
-\tnote{Expand on ref:sec:copmatrices}
 
-Figure~\ref{fig:cop-matrix} shows examples of consecutive-ones
-property. \tnote[important]{move to general def section? if so, decide
-how to cross ref with repetition for completeness of chapter.}
+The most important questions with respect to a particular 
+property desired in a structure/object are perhaps the following.
+\begin{itemize}%[]
+\singlespacing
+\item Does the desired property exist in the given input?
+\item If the test is affirmative, what is a certificate of the affirmative?
+\item If the test is negative, what are the optimization possibilities
+  for the property in the input? In other words, how close to having the
+  property can the input be?
+\item If the test is negative, what is a certificate of the negative?
+\end{itemize}
 
-\begin{figure}[t] %[htbp]
+In this section and the rest of the chapter we see results that shaped the
+correponding areas respectively for \cop in binary matrices.
+
+\begin{enumerate}[a.]
+\singlespacing
+\item \label{q:testcop} Does a given binary matrix have \COP?
+\item \label{q:copperm} What is the \COP permutation for the given matrix with \COP?
+\item \label{q:copopt} What are the optimizations possible and practically useful on
+  the given matrix without \COP?
+\item \label{q:nocopcert} If algorithm for (\ref{q:testcop}) returns \textbf{false}, can a
+  certificate for this be computed?
+\end{enumerate}
+
+Without doubt, besides computing answers to these questions, we are
+interested in the efficiency of these computations in terms of
+computational complexity theory. Results towards
+questions~(\ref{q:testcop}) and (\ref{q:copperm}) are surveyed in this
+section. Those for question~(\ref{q:copopt}) are discussed in
+Section~\ref{sec:surveycopopt} and question~(\ref{q:nocopcert}) is discussed
+in Section~\ref{sec:surveycertalgo}.
+
+It may be noted that one way to design an algorithm to test for \COP
+is by deriving one from any interval graph recognition algorithm using the
+result {\sc hmpv00}\tnote{cite hmpv00} \tnote{in endnote put the theorem 2.7 - dom
+  pg 43} \cite{d08phd} which demonstrates how such a derivation can be
+done. However, this does not necessarily yield an efficient
+algorithm. We will see results that directly solve the problem on
+matrices since it is known that questions~(\ref{q:testcop}) and
+(\ref{q:copperm}) stated above for \COP are efficiently solvable.
+Table~\ref{tab:cophistory} gives a snapshot of these results.
+
+\begin{table}[htbp]
   \centering
+   \includegraphics[scale=0.5]{../img/cophistory.png} % TEMPORARY
 
-  {\figtabsize
-    \begin{tabular}[h]{l|lcccl}
-      $M_1$: & $M_1'$: &&&& $M_2$:\\
-      &&&&&\\
-      \begin{tabular}[h]{llll}
-        $c_1$ & $c_2$ &$c_3$ &$c_4$\\
-        &&&\\
-        \un & 0   & \un & 0\\
-        0   & \un & 0   & \un \\
-        \un & 0   & 0   & \un
-      \end{tabular}
-      &
-      \begin{tabular}[h]{llll}
-        $c_3$ &$c_1$ &$c_4$& $c_2$\\
-        &&&\\
-        \un & \un & 0 & 0\\
-        0 & 0 & \un & \un \\
-        0 & \un & \un & 0
-      \end{tabular}
-      &&&&
-      \begin{tabular}[h]{llll}
-        $d_1$ & $d_2$ &$d_3$ &$d_4$\\
-        % &&&\\
-        &&&\\
-        \un & \un & 0 & 0\\
-        0 & \un & \un & 0 \\
-        0 & \un & 0 & \un 
-      \end{tabular}
-    \end{tabular}
+  \caption[\figtabsize A brief history of \COP research]{\figtabsize A
+    brief history of \COP research \temptext{ OLD IMG - PLACEHOLDER - A LATEX
+    TABLE WILL REPLACE THIS + WILL BE UPDATED }}
+  \label{tab:cophistory}
+\end{table}
+
+\subsection{Tucker's forbidden submatrices for \COP}
+
+% \change{While it is remarkable that the first polynomial time
+%   algorithm for \COP testing came in }{The first polynomial time
+%   algorithm for \COP testing came in }\cite{fg65} \change{,}{.}
+% \change{it is safe to say that the most}{A few years later, a very}
+% significant result in understanding \COP came \remove{a few years
+%   later} from Tucker which gave a combinatorial (negative)
+% characterization of matrices with COP \cite{at72}.
+A polynomial time algorithm for \COP testing was seen for the first
+time in \cite{fg65} which uses overlapping properties of columns with
+\un s. A few years later, a deeply significant result based on very
+different ideas in understanding \COP came from Tucker which gave a
+combinatorial (negative) characterization of matrices with COP
+\cite{at72}. This result influenced most of the \COP results that
+followed in the literature \annote[pressing]{including linear time
+  algorithms}{did it? which ones?} for \COP recognition.
+
+\cite{at72} discovered certain forbidden structures for convex
+bipartite graphs\endnote{The terminology in \cite{at72} differs. It
+  uses the term {\em graphs with $V_1$-consecutive arragement} instead
+  of {\em convex bipartite graphs}.} and by definition of this graph
+class, this translates to a set of forbidden submatrices for matrices
+with \cop.  The following are the theorems from \cite{at72} that
+acheived this characterization.
+
+Theorem~\ref{th:tuckeratfree} proves that convex bipartite graphs
+cannot have {\em asteroidal triples}\endnote{If $G = (V,E)$ is a
+  graph, a set of three vertices from $V$ form an {\em asteroidal
+    triple} if between any two of them there exists a path in $G$ that
+  does not contain any vertex from the closed neighborhood of the
+  third vertex.} contained in the corresponding vertex
+partition\endnote{The partition corresponds to columns (rows) if its
+  half adjacency matrix has \COP columns (rows).}.
+Theorem~\ref{th:tuckerforbidden} shows what are the structures in a
+bipartite graph that force one of its vertex partitions to have
+asteriodal triples -- in other words, it identifies the subgraphs that
+prevent the graph from being convex bipartite.
+
+\begin{theorem}
+  (\cite[Th.~6]{at72}, \cite[Th.~2.3]{d08phd})\\
+  A bipartite graph $G = (V_1, V_2, E)$ is convex bipartite on
+  columns\endnote{Abridged to match terminology adopted in this
+    document. See previous note.} \iff $V_1$ contains no asteroidal
+  triple of $G$.
+  \label{th:tuckeratfree}
+\end{theorem}
+
+\begin{theorem}
+  (\cite[Th.~7]{at72}, \cite[Th.~2.4]{d08phd})\\
+  In a bipartite graph $G = (V_1, V_2, E)$ the vertex set $V_1$
+  contains no asteroidal triple \iff $G$ contains none of the graphs
+  $G_{I_k}$, $G_{II_k}$, $G_{III_k}$ (with $k \ge 1$), $G_{IV}$,
+  $G_{V}$ as shown in Figure~\ref{fig:forbiddensubgraphs} as subgraphs.
+  \label{th:tuckerforbidden}
+\end{theorem}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[scale=0.35]{../img/tuckersforbiddensubgraphs.png}  
+  \caption[\figtabsize Tucker's forbidden subgraphs]{\figtabsize Tucker's
+    forbidden subgraphs for convex bipartite graphs. \temptext{ PLACEHOLDER
+    IMG }} 
+  \label{fig:forbiddensubgraphs}
+\end{figure}
+
+Theorem~\ref{th:tuckeratfree} and Theorem~\ref{th:tuckerforbidden}
+result in the following Theorem~\ref{th:tuckercop} which characterizes
+matrices with \COP.
+
+\begin{theorem}
+  (\cite[Th.~9]{at72}, \cite[Th.~2.5]{d08phd})\\
+  A matrix $M$ has \COP \iff it contains none of the matrices 
+$M_{I_k}$, $M_{II_k}$, $M_{III_k}$ (with $k \ge 1$), $M_{IV}$,
+  $M_{V}$ as shown in Figure~\ref{fig:forbiddensubmatrices} as submatrices.
+  \label{th:tuckercop}
+\end{theorem}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[scale=0.35]{../img/tuckersforbiddensubmatrices.png}  
+  \caption[\figtabsize Tucker's forbidden submatrices]{\figtabsize Tucker's
+    forbidden submatrices for convex bipartite graphs. \temptext{ PLACEHOLDER
+    IMG }} 
+  \label{fig:forbiddensubmatrices}
+\end{figure}
+
+It can be verified that the matrices in
+Figure~\ref{fig:forbiddensubmatrices} are the half adjacency matrices
+of the graphs in Figure~\ref{fig:forbiddensubgraphs} respectively
+which is not surprising due to Definition~\ref{def:graphwithcop}~\ref{def::convexbi}.
+
+\temptext{
+  \begin{enumerate}[-I-]
+  \item PQ trees. Define. Concise flowchart of algorithm? [BL76]\\
+    The construction of PQ trees in \cite{bl76} draws on the close
+    relationship of matrices with COP to interval graphs. A PQ tree of
+    a matrix is one that stores all row (column) permutations of the
+    matrix that give the COP orders (there could be multiple orders of
+    rows or columns) of the matrix. This is constructed using an
+    elaborate linear time procedure and is also a test for
+    planarity\tnote{check check check. both interval graph and
+      planarity in this paper?}.
+  \item Variations - PQR, PC etc. [MM96, Hsu01, McC04]\\
+    PQR trees is a generalized data structure based on PQ trees
+    \cite{mm96,mpt98}.  \cite{tm05} describes an improved algorithm to
+    build PQR
+    trees. \tnote{improv in terms of what?}.\\
+    Generalized PQ trees McC04 - brief. Will be elaborated in
+    Section~\ref{sec:surveycertalgo}\\
+    Hsu also invented PC trees \cite{wlh01}\tnote{This result first
+      appeared inproc ISAAC92} which is claimed to be much easier to
+    implement.
+  \item Set theoretic characterizations [Hsu02, NS09]\\
+    \cite{wlh02} describes the simpler algorithm for
+    COT.\tnote[pressing]{in terms of what?}\\
+    \cite{nsnrs09} describes a characterization of consecutive-ones
+    property solely based on the cardinality properties of the set
+    representations of the columns (rows); every column (row) is
+    equivalent to a set that has the row (column) indices of the rows
+    (columns) that have one entries in this column (row). This is
+    interesting and relevant, especially to this thesis because it
+    simplifies COT to a great degree by reducing the solution search
+    space owing to the a simple set theoretic characterization.\\
+  \item Parallel algorithms [AS95, BS03, CY91]\\
+  \item
+  \end{enumerate}
+
 }
 
-  \caption[\figtabsize Matrices with and without COP.]{\figtabsize
-    Matrices with and without COP. $M_1$ has COP because by permuting
-    its columns, $c_1$-$c_4$, one can obtain $M_1'$ where the {\un}s
-    in each row are consecutive. $M_2$, however, does not have COP
-    since no permutation of its columns, $d_1$-$d_4$, will arrange
-    {\un}s in each row consecutively \cite{d08phd}.}
+\tnote[pressing]{ The results in cite~{bl76} on COT are based on the
+   result that interval graphs are AT-free chordal graphs.
+ -  Tucker's?}%
 
-  \label{fig:cop-matrix}
-\end{figure}
+\tnote[pressing]{cite~{fg65} uses these results to give the first polynomial
+  time algorithm for COT. WHICH RESULTS? \ \ \ \ \ check. how do they use it?}%
 
+\tnote[pressing]{ADD: peo exists iff chordal. lexicographic BFS
+  [tag:chordalGraph]} %
+
+\clearpage      % Print pending floats
 \section{Optimization problems in COP}
@@
 
-\section{********* COP in Relational Database Model}
-\label{sec:surveyrdbm}
-\tnote{Expand on sec:apprdbm} \tnote{(set systems theme)}
-
-\section{********* COP in Graph Isomorphism}
-\label{sec:surveygraphiso}
-\tnote{Expand on sec:appgraphiso} \tnote{(canonization theme)}
-
-\section{********* Certifying Algorithms}
-\label{sec:surveycertalgo}
-\tnote{Expand on sec:appcertalgo} \tnote{ (certification McC04 theme)}
-
-\section*{-- REFERENCE CONTENT --}
-%%%%%%%%%%%%% listing package for pretty formatting BEGIN
-% \lstloadlanguages{C}
-% \lstset{language=C, commentstyle=\scriptsize,
-%   numberstyle=\tiny, numbers=left,               % Line numbers
-%   stepnumber=2,numbersep=5pt,firstnumber=10, 
-%   showstringspaces=false,                    % No explicit spaces
-%   emph={printf},emphstyle=\underbar,      % Specific formatting
-%   emph=[2]{sum},emphstyle=[2]\color{blue},      % Sp format (multiple)
-%   emph=[3]{i},emphstyle=[3]\color{red}      % Sp format (multiple)
-%   }
-
-%   \begin{singlespacing}
-%     % Text inside "lstlisting" env is verbatim.
-%     \begin{lstlisting}[keywordstyle=\textbf]      
-% int sum;
-% int i; /* for loop variable */ 
-% sum = 0; 
-% for (i=0, i<n; i++) { 
-%   sum += a[i]; 
-% } 
-% printf{"This is stupid."}
-%     \end{lstlisting}
-%   \end{singlespacing}
-
-%%%%%%%%%%%%% listing package for pretty formatting END
-
-{%\fontfamily{cmvtt}\selectfont
-\color{blue}
-\subsection{Matrices with COP}
-\label{sec:copmatrices}
-As seen earlier, the interval assignment problem (illustrated as the
-course scheduling problem in Section~\ref{sec:problem}), is a special
-case of the problem we address in this thesis, namely the tree path
-labeling problem (illustrated as the \illustrationproblem). The
-interval assignment problem and COP problem are equivalent
-problems. In this section we will see some of the results that exists
-in the literature today towards solving the COP problem and
-optimization problems surrounding it.
-
-Recall that a matrix with COP is one whose rows (columns) can be
-rearranged so that the {\un}s in every column (row) are in consecutive
-rows (columns). COP in binary matrices has several practical applications
-in diverse fields including scheduling \cite{hl06}, information
-retrieval \cite{k77} and computational biology \cite{abh98}.  Further,
-it is a tool in graph theory \cite{mcg04} for interval graph
-recognition, characterization of Hamiltonian graphs, planarity testing
-\cite{bl76} and in integer linear programming \cite{ht02,hl06}.
-
-The obvious first questions after being introduced to the consecutive
-ones property of binary matrices are if COP can be detected
-efficiently in a binary matrix and if so, can the COP permutation of
-the matrix also be computed efficiently?  Recognition of COP in a
-binary matrix is polynomial time solvable and the first such algorithm
-was given by \cite{fg65}.  A landmark result came a few years later
-when \cite{at72} discovered the families of forbidden submatrices that
-prevent a matrix from having COP and most, if not all, results that
-came later were based on this discovery which connected COP in binary
-matrices to convex bipartite graphs. In fact, the forbidden
-submatrices came as a corollary to the discovery that convex bipartite
-graphs are AT-free in \cite{at72}\tnote{check}. The first linear
-time algorithm for COP testing (COT) was invented by \cite{bl76} using
-a data structure called PQ trees.  Since then several COT algorithms
-have been invented -- some of which involved variations of PQ trees
-\cite{mm96,wlh01,mcc04}, some involved set theory and ICPIA
-\cite{wlh02,nsnrs09}, parallel COT algorithms\cite{as95,bs03,ly91} and
-certifying algorithms\cite{mcc04}.
-
-The construction of PQ trees in \cite{bl76} draws on the close
-relationship of matrices with COP to interval graphs. A PQ tree of a
-matrix is one that stores all row (column) permutations of the matrix
-that give the COP orders (there could be multiple orders of rows or
-columns) of the matrix. This is constructed using an elaborate linear
-time procedure and is also a test for planarity\tnote{check check
-  check. both interval graph and planarity in this paper?}.  PQR trees
-is a generalized data structure based on PQ trees \cite{mm96,mpt98}.
-\cite{tm05} describes an improved algorithm to build PQR
-trees. \tnote{improv in terms of what?}\cite{wlh02} describes the
-simpler algorithm for COT. Hsu also invented PC trees
-\cite{wlh01}\tnote{This result first appeared inproc ISAAC92}
-which is claimed to be much easier to implement. \cite{nsnrs09}
-describes a characterization of consecutive-ones property solely based
-on the cardinality properties of the set representations of the
-columns (rows); every column (row) is equivalent to a set that has the
-row (column) indices of the rows (columns) that have one entries in
-this column (row). This is interesting and relevant, especially to
-this thesis because it simplifies COT to a great degree. \tnote{it
-  reduces the solution search space. fill in the blanks.}
-
-\cite{mcc04} describes a different approach to COT. While all previous
-COT algorithms gave the COP order if the matrix has the property but
-exited stating negative if otherwise, this algorithm gives an evidence
-by way of a certificate of matrix even when it has no COP. This
-enables a user to verify the algorithm's result even when the answer
-is negative. This is significant from an implementation perspective
-because automated program verification is hard and manual verification
-is more viable. Hence having a certificate reinforces an
-implementation's credibility. Note that when the matrix {\em has} COP,
-the COP order is the certificate.  The internal machinery of this
-algorithm is related to the weighted betweenness problem
-addressed\tnote{in what way??} in \cite{co98}.  \tnote{expand
-  on the COP order graph creation and it having to be bipartite for M
-  to have COP. and thus an odd cycle being an evidence of no COP.}
-\tnote{ where should this go?: (1) cite|jlm97 (application of PQ
-  trees in graphics). (2) helly's theorem citation
-  19XXdgk-Hellystheorem-Danzer-Gruenbaum-Klee}
-
-\subsection{Optimization problems in COP}
-\label{sec:optcop}
-
+%{\color{blue} \footnotesize
+\temptext{
 So far we have been concerned about matrices that have the consecutive
@@
   to COP -- cite:ckl96 (ILP, circ ones, one drop) -- cite:th98
-  (generalization of COP - minimax, biotonic column) Tucker} }
+  (generalization of COP - minimax, biotonic column) Tucker}
+
+}
+
+\section{Certifying Algorithms}
+\label{sec:surveycertalgo}
+%\tnote{Expand on sec:appcertalgo} \tnote{ (certification McC04
+%theme)}
+
+\temptext{ Certifying algorithm (Generalized PQ trees) [McC04] -- Move this to section on
+    certifying algo.
+
+\cite{mcc04} describes a different approach to COT. While all previous
+COT algorithms gave the COP order if the matrix has the property but
+exited stating negative if otherwise, this algorithm gives an evidence
+by way of a certificate of matrix even when it has no COP. This
+enables a user to verify the algorithm's result even when the answer
+is negative. This is significant from an implementation perspective
+because automated program verification is hard and manual verification
+is more viable. Hence having a certificate reinforces an
+implementation's credibility. Note that when the matrix {\em has} COP,
+the COP order is the certificate.  The internal machinery of this
+algorithm is related to the weighted betweenness problem
+addressed\tnote{in what way??} in \cite{co98}.  \tnote{expand
+  on the COP order graph creation and it having to be bipartite for M
+  to have COP. and thus an odd cycle being an evidence of no COP.}
+
+}
+
+\section{COP in Relational Database Model}
+\label{sec:surveyrdbm}
+\tnote{Expand on sec:apprdbm} \tnote{(set systems theme)}
+
+\section{COP in Graph Isomorphism}
+\label{sec:surveygraphiso}
+\tnote{Expand on sec:appgraphiso} \tnote{(canonization theme)}
+
+\temptext{
+The survey from kklv10 conclusion.
+}
+
+% \section{Other applications}
+% {\tt If any. \tnote{ where should this go?}: (1) cite|jlm97 (application of PQ
+%   trees in graphics). (2) helly's theorem citation
+%   19XXdgk-Hellystheorem-Danzer-Gruenbaum-Klee}}
 
 \theendnotes\tnote{remove if none.}
+
+%%%%%%%%%%%%% listing package for pretty formatting BEGIN
+% \lstloadlanguages{C}
+% \lstset{language=C, commentstyle=\scriptsize,
+%   numberstyle=\tiny, numbers=left,               % Line numbers
+%   stepnumber=2,numbersep=5pt,firstnumber=10, 
+%   showstringspaces=false,                    % No explicit spaces
+%   emph={printf},emphstyle=\underbar,      % Specific formatting
+%   emph=[2]{sum},emphstyle=[2]\color{blue},      % Sp format (multiple)
+%   emph=[3]{i},emphstyle=[3]\color{red}      % Sp format (multiple)
+%   }
+
+%   \begin{singlespacing}
+%     % Text inside "lstlisting" env is verbatim.
+%     \begin{lstlisting}[keywordstyle=\textbf]      
+% int sum;
+% int i; /* for loop variable */ 
+% sum = 0; 
+% for (i=0, i<n; i++) { 
+%   sum += a[i]; 
+% } 
+% printf{"This is stupid."}
+%     \end{lstlisting}
+%   \end{singlespacing}
+
+%%%%%%%%%%%%% listing package for pretty formatting END

### 2012-01-05 src/040chap4.tex
Chapter 2 edited. See diff for details. Too many to remember. or maybe too little.

@@
 \label{ch:conclusion}
+
+\tnote[pressing]{TBD survey
+  see blue notes (in notebook) under Graph Isomorphism. namely
+  citations in cite:aas93 (3) a brief on heirarchy of chordal
+  graphs, path graphs, interval graphs, peo, clique tree etc. and
+  results. (4) interval graphs are incomparability graphs. see
+  Golumbic. [tag:classification] (5) finding min length hole in
+  bipartite graph is polynomial time [Sec 3.3.2 in cite:d08phd] (6)
+  theorem 2.2 in cite:d08phd - $G$ is union of v.d. caterpillars iff
+  $M$ has COP in rows. $M$ is edge vertex incidence matrix $(*,2)$
+  of $G$
+}%
 

### 2012-01-05 src/thesis.tex
Reordered chrono.
Added amsthm and changed theorem environs to comply.
Added atbeginend commands to space theorem environs.
Added a few \def strings.

Added placeholders for a couple of definitions.

Started editing. Intro. Added a couple of definitions.

Minor edits to comply with amsthm

Spacing hacks mostly.

Trivial.
endnote flushing - TBD
Chapter 2 edited. See diff for details. Too many to remember. or maybe too little.

@@
 \usepackage{../lib/mylatexlib}
+\usepackage{thesisfiguresandtables} % Code for tables are in this
+                                % package if it is too long to inline
+                                % in .tex file.
+
 % \usepackage{comment}         % DEBUG ONLY
-\includeonly{010chap1,020chap2}
-%\includeonly{0021ded}
+%\includeonly{020chap2}         % DEBUG ONLY 
+%\includeonly{0021ded}         % DEBUG ONLY
 
@@
 %
+\include{figuresandtables} % Load floats. No rendering.
+
 \include{010chap1}
+% turning counter reset. perhaps universal endnote counter is good?
+% being schizo about it now. DECIDE! ENDNOTEMARKS DEPEND ON IT.
+\setcounter{endnote}{0} % need to empty the file too. how?! -
+                        % FIXED. turns out, just having notes is
+                        % enough. or maybe it was fluke.
 \include{020chap2}
@@
 \begin{singlespace}
-  \addcontentsline{toc}{chapter}{REFERENCES \tnote[minor]{make names
-    in bib file uniform style w.r.t. firstname/initials}}
+  \addcontentsline{toc}{chapter}{REFERENCES \tnote[minor]{(i)make names
+    in bib file uniform style w.r.t. firstname/initials (ii) have back
+  refs - i.e. pages that cite the item.}}
 

### 2012-01-06 src/0021ded.tex
Chapter 2 edited. See diff for details. Too many to remember. or maybe too little.
Added more stuff to chapter 2. see diff.

@@
   \vskip 5ex%
-  {\Large \noindent The rhetoric and mythos of science create the
+  {\large \noindent The rhetoric and mythos of science create the
     comforting image of linear progression toward truth.%
@@
   }%  
-  \vskip 5ex%
-  {\huge%
+  \vskip 15ex%
+  {\large%
     \noindent Art is the solution to chaos. \par %
+    {\normalsize \par \hfill \emph{citation needed.}}
   }%

### 2012-01-06 src/010chap1.tex
Added more stuff to chapter 2. see diff.

@@
 \label{sec:basicprelim}
-\tnote{(definitions theorems etc) needed if any - graph theory}
 
-\begin{definition}{\emph{\Cop}}
+\begin{definition}[\emph{Binary matrix}]
+  Let $M$ be an $n \times m$ matrix. $m_{i,j}$ denotes its $(i,j)$th
+  element, i.e. element at $i$th row and $j$th column. $M$ is a binary
+  matrix if each of its element is 0 or \un.
+  In other words, for all $i \in [n]$ and $j \in [m]$, $m_{i,j} \in
+  \{0, \un\}$ 
+\end{definition}
+
+\begin{definition}[\emph{set system}]
+\end{definition}
+
+\begin{definition}[\emph{\Cop}]
   \tnote[important]{define cop here. refer to it in chap 2 or
@@
   \item \Cop order or \cop permutation
+  \item \Cop for set systems\tnote{see mcc04}
   \end{enumerate}
@@
 
-\begin{figure}[t] %[htbp]
-  \centering
-  {\figtabsize
-    \begin{tabular}[h]{l|lcccl}
-      $M_1$: & $M_1'$: &&&& $M_2$:\\
-      &&&&&\\
-      \begin{tabular}[h]{llll}
-        $c_1$ & $c_2$ &$c_3$ &$c_4$\\
-        &&&\\
-        \un & 0   & \un & 0\\
-        0   & \un & 0   & \un \\
-        \un & 0   & 0   & \un
-      \end{tabular}
-      &
-      \begin{tabular}[h]{llll}
-        $c_3$ &$c_1$ &$c_4$& $c_2$\\
-        &&&\\
-        \un & \un & 0 & 0\\
-        0 & 0 & \un & \un \\
-        0 & \un & \un & 0
-      \end{tabular}
-      &&&&
-      \begin{tabular}[h]{llll}
-        $d_1$ & $d_2$ &$d_3$ &$d_4$\\
-        % &&&\\
-        &&&\\
-        \un & \un & 0 & 0\\
-        0 & \un & \un & 0 \\
-        0 & \un & 0 & \un 
-      \end{tabular}
-    \end{tabular}
-  }
+\figcopmatrix
+\tnote[important]{have a figure for set systems also}
 
-  \caption[\figtabsize Matrices with and without COP.]{\figtabsize
-    Matrices with and without COP. $M_1$ has COP because by permuting
-    its columns, $c_1$-$c_4$, one can obtain $M_1'$ where the {\un}s
-    in each row are consecutive. $M_2$, however, does not have COP
-    since no permutation of its columns, $d_1$-$d_4$, will arrange
-    {\un}s in each row consecutively \cite{d08phd}.}
-
-  \label{fig:cop-matrix}
-\end{figure}
-
-\begin{definition}{\emph{\Crop}}
+\begin{definition}[\emph{\Crop}]
   \tnote[important]{define circ1 here. refer to it in chap 2 or
@@
 
-\begin{definition}{\emph{Binary matrix}}
-  Let $M$ be an $n \times m$ matrix. $m_{i,j}$ denotes its $(i,j)$th
-  element, i.e. element at $i$th row and $j$th column. $M$ is a binary
-  matrix if each of its element is 0 or \un.
-  In other words, for all $i \in [n]$ and $j \in [m]$, $m_{i,j} \in
-  \{0, \un\}$ 
+\begin{definition}[\emph{Graph, Tree}]
+
+\end{definition}
+
+\begin{definition}[\emph{Maximal clique}]
+
 \end{definition}

### 2012-01-06 src/020chap2.tex
Added more stuff to chapter 2. see diff.

@@
 \tnote[important]{ADD: have a few lines about organization of chapter}
-
-\tnote{DEFINE COP SOMEWHERE}
 
@@
     $G$.
+    \label{def::maxcliquematrix}
   \item \emph{Half adjacency matrix} of $H$ is the $n_a \times n_b$
@@
       consecutive in the ordering \cite[Th. 8.1]{mcg04}.}
+
     \begin{enumerate}[a.]
@@
       CO/ROP related?}
-  \item \label{def::convexbi} $H$ is \emph{convex bipartite on columns (rows)} if its half adjacency matrix
-    has \COP on rows (columns).
+  \item $H$ is \emph{convex bipartite on columns (rows)} if its half
+    adjacency matrix has \COP on rows (columns).%
+    \label{def::convexbi}
   \item $H$ is \emph{biconvex bipartite} or \emph{doubly
@@
 
-\clearpage      % Print pending floats
+%\clearpage      % Print pending floats
 
@@
   the given matrix without \COP?
-\item \label{q:nocopcert} If algorithm for (\ref{q:testcop}) returns \textbf{false}, can a
-  certificate for this be computed?
+\item \label{q:nocopcert} If algorithm for (\ref{q:testcop}) returns
+  \textbf{false}, can a certificate for this be computed?
 \end{enumerate}
@@
 
+%\clearpage
 \subsection{Tucker's forbidden submatrices for \COP}
@@
 of the graphs in Figure~\ref{fig:forbiddensubgraphs} respectively
-which is not surprising due to Definition~\ref{def:graphwithcop}~\ref{def::convexbi}.
+which is not surprising due to
+Definition~\ref{def:graphwithcop}~\ref{def::convexbi}.
+
+%\clearpage
+\subsection{$PQ$ tree -- a linear COT algorithm}
+
+Booth and Lueker in their paper \cite{bl76} gave the first linear
+algorithm for \cop testing by way of an interval graph recognition
+algorithm. Their result has close relations to the characterization of
+interval graphs by \cite{gh64}. A graph $G$ is an interval graph \iff
+all its maximal cliques can be linearly ordered \stt for any vertex
+$v$ in $G$, all the cliques that $v$ is incident on are consecutive in
+this order. Clearly, this means that the maximal clique incidence
+matrix\footnote{
+  Definition~\ref{def:graphmatrices}~\ref{def::maxcliquematrix}} must
+have \COP on rows. This algorithm constructs a data structure called
+\PQtree. A \PQtree represents all the \COP orderings of the matrix it
+is associated with. \cite{bl76}'s algorithm uses the fact that if a
+matrix has \COP, a \PQtree for it can be constructed. It is
+interesting to note that aside from interval graph recognition and
+\COP testing, \PQtree is also useful in other applications like
+finding planar embeddings of planar graphs \cite{lec67,mcc04} and
+recognizing \CROP in a matrix.
+
+\begin{definition}[\emph{\PQtree \cite{bl76, mcc04}}]\\
+  A \PQtree of matrix $M$ with \COP, is a tree with the following properties.
+  \begin{enumerate}[i.]
+    \singlespacing
+  \item Each leaf uniquely represents a row (column) of $M$. The leaf
+    order of the tree gives a \COP order for column (row)\endnote{Note
+      that \COP order for column requires permutation of rows and vice
+      versa.} for $M$.
+  \item Every non-leaf node in the tree are labeled $P$ or $Q$.
+  \item \label{def::nodep} The children of $P$ nodes are
+    unordered. They can be permuted in any fashion to obtain a new
+    \COP order for $M$.
+  \item \label{def::nodeq} The children of $Q$ nodes are linearly
+    ordered. Their order can be reversed to get obtain a new \COP
+    order for $M$.
+  \end{enumerate}
+  \label{def:pqtree}
+\end{definition}
+
+Thus, effectively, there exists a bijection between set of matrices
+with \COP and the set of \PQtrees (accurately speaking, each matrix
+with \COP bijectively maps to an equivalence class of \PQtrees due to
+properties (\ref{def::nodep}) and (\ref{def::nodeq})). See
+Figure~\ref{fig:pqtree} for an example of \PQtree.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{tabular}[h]{lr}
+    \begin{tabular}[h]
+      {>{\columncolor{\tblhcolor}} l lllllllllll|}
+      \hline
+      $r_1$   &\un &0   &\un &0   &0   &0    &0   \\
+      $r_2$   &\un &0   &\un &0   &0   &0    &0   \\
+      $r_3$   &\un &0   &0   &\un &0   &0    &0   \\
+      $r_4$   &\un &0   &0   &\un &0   &0    &0   \\
+      $r_5$   &\un &0   &0   &0   &0   &0    &0   \\
+      $r_6$   &\un &\un &0   &0   &0   &0    &0   \\
+      $r_7$   &0   &\un &0   &0   &0   &\un  &0   \\
+      $r_8$   &0   &\un &0   &0   &\un &\un  &0   \\
+      $r_9$   &0   &\un &0   &0   &\un &\un  &0   \\
+      $r_{10}$ &0   &\un &0   &0   &\un &\un  &\un \\
+      $r_{11}$ &0   &\un &0   &0   &\un &0    &\un \\
+      \hline
+    \end{tabular}
+    &
+    \includegraphics[scale=0.3]{../img/pqtree.png}
+
+  \end{tabular}
+
+  \caption[\figtabsize An example for \PQtree]{\figtabsize An example for \PQtree.
+    Permuting the order of the left child of the root, we see that
+    \temptext{(d,a,b,c,e,f,g,h,i,j,k)} is a \COP order. Reversing the
+    order of the right child of the root, we see that
+    \temptext{(a,b,c,d,e,f,k,j,h,i,g) } is yet another \COP order. \temptext{PLACEHOLDER IMG.}
+    \cite[Fig.~1\endnote{\cite{mcc04}illustrates \COP on rows. Our
+      convention in this document is \COP on columns and thus example
+      matrix has been transposed. }]{mcc04}}
+  \label{fig:pqtree}
+\end{figure}
+
+The \cite{bl76} algorithm with input $n \times m$ matrix $M$ starts
+with a \PQtree for a vacuous $n \times 0$ matrix $M'$ (submatrix
+induced by 0 columns). This is known as a {\em universal} \PQtree
+which is one with its root as a $P$ node and only leaves as its
+children -- each leaf representative of a row of input (by definition
+of \COP for columns). This induced submatrix $M'$ vacuously has \COP.
+Each column is then added iteratively to $M'$ to check if the new $M'$
+has \COP.  By a complicated, but linear, procedure the algorithm does
+one of the following actions in each iteration: (a) declare that $M$
+has no \COP, or (b) modify the current \PQtree to represent the new
+$M'$ (which clearly, must have \COP, since if not, option (a) would
+have been executed).
 
 \temptext{
-  \begin{enumerate}[-I-]
-  \item PQ trees. Define. Concise flowchart of algorithm? [BL76]\\
-    The construction of PQ trees in \cite{bl76} draws on the close
-    relationship of matrices with COP to interval graphs. A PQ tree of
-    a matrix is one that stores all row (column) permutations of the
-    matrix that give the COP orders (there could be multiple orders of
-    rows or columns) of the matrix. This is constructed using an
-    elaborate linear time procedure and is also a test for
-    planarity\tnote{check check check. both interval graph and
-      planarity in this paper?}.
+  \begin{enumerate}[ -I-]
   \item  Variations - PQR, PC etc. [MM96, Hsu01, McC04]\\
@@
     space owing to the a simple set theoretic characterization.\\
-  \item Parallel algorithms [AS95, BS03, CY91]\\
-  \item
+  \item Parallel algorithms [AS95, BS03, CY91]
   \end{enumerate}
@@
 
-\clearpage      % Print pending floats
+%\clearpage      % Print pending floats
 \section{Optimization problems in COP}

### 2012-01-06 src/020chap2.tex
trivial commit before pushing

@@
 
+After the invention of \PQtrees, there has been several variants of
+the same with either simpler procedure of construction, like \PCtree
+\cite{wlh01}\endnote{First appeared in
+  \cite{wlh92_isaac}.\tnote{check}} or a generalization of it, like
+generalized \PQtree \cite{mcc04}. \PQRtree is another generalization
+of \PQtree by \cite{mm96,mpt98}. There algorithm was improved by
+\cite{tm05} \tnote{improved in terms of
+  what?}. % describes an improved algorithm to
+% build PQR trees.
+
+\temptext{Expand a little more on PC trees and PQR trees.}
+
+We will see more of \PQRtree in Section~\ref{sec:surveycertalgo}
+
 \temptext{
@@
   \item  Variations - PQR, PC etc. [MM96, Hsu01, McC04]\\
-    PQR trees is a generalized data structure based on PQ trees
-    \cite{mm96,mpt98}.  \cite{tm05} describes an improved algorithm to
-    build PQR
-    trees. \tnote{improv in terms of what?}.\\
     Generalized PQ trees McC04 - brief. Will be elaborated in

### 2012-01-07 src/020chap2.tex
chapter 2 - started set system side of things. see diff for more details.

@@
 
+\def\tabparboxw{10cm}
 \begin{table}[htbp]
   \centering
-   \includegraphics[scale=0.5]{../img/cophistory.png} % TEMPORARY
+  {
+    \fontencoding{OT1}\fontfamily     {cmss}\selectfont
+    % \sffamily
+    \footnotesize
+    % \includegraphics[scale=0.5]{../img/cophistory.png} % TEMPORARY
 
+   \begin{tabular}[h]{lll}
+     \hline
+     \hline
+     % 
+     1899 & \parbox[t]{\tabparboxw}{ First mention of \COP} & \cite{p59}  \\
+
+     \rowcolor[gray]{0.8} %
+     1951 & \parbox[t]{\tabparboxw}{ Heuristics for COT} & \cite{wsr51} \\
+
+     1965 & \parbox[t]{\tabparboxw}{ First polynomial time algorithm
+       for \COP testing} & \cite{fg65} \\
+
+     \rowcolor[gray]{0.8} %
+     1972 & \parbox[t]{\tabparboxw}{ Characterization for \COP -- forbidden
+       submatrices} &
+     \cite{at72} \\
+
+     1976 & \parbox[t]{\tabparboxw}{ First linear time algorithm for COT --
+       \PQtree} &
+     \cite{bl76} \\
+
+     \rowcolor[gray]{0.8} %
+     1992 & \parbox[t]{\tabparboxw}{ Linear
+       time algorithm COT without \PQtree} & \cite{wlh02}\endnote{First
+       published in \cite{wlh02isaac92}} \\
+
+     2001 & \parbox[t]{\tabparboxw}{\PCtree\ -- a simplification of
+       \PQtree} & \cite{wlh01,hm03} \\
+
+     \rowcolor[gray]{0.8} %     
+     1996 & \parbox[t]{\tabparboxw}{\PQRtree\ -- generalization of
+       \PQtree for any binary matrix regardless of its \COP status} &
+     \cite{mm96} \\
+
+     1998 & \parbox[t]{\tabparboxw}{ Almost linear time to construct
+       \PQRtree} & \cite{mpt98} \\
+
+     \rowcolor[gray]{0.8} %
+     2004 & \parbox[t]{\tabparboxw}{ A certifying algorithm for no
+       \COP.\\ Generalized \PQtree.} & \cite{mcc04} \\
+
+     2009 & \parbox[t]{\tabparboxw}{ Set theoretic, cardinality based
+       characterization of \COP\ -- ICPIA} & \cite{nsnrs09} \\
+
+     \rowcolor[gray]{0.8} %
+     2010 & \parbox[t]{\tabparboxw}{ Logspace \COP testing} &
+     \cite{kklv10}\\ 
+     \hline
+     \hline
+   \end{tabular}
+ }
  \caption[\figtabsize A brief history of \COP research]{\figtabsize A
-    brief history of \COP research \temptext{ OLD IMG - PLACEHOLDER - A LATEX
-    TABLE WILL REPLACE THIS + WILL BE UPDATED }}
+   brief history of \COP research }
  \label{tab:cophistory}
@@
 
-\begin{figure}[htbp]
-  \centering
-  \begin{tabular}[h]{lr}
-    \begin{tabular}[h]
-      {>{\columncolor{\tblhcolor}} l lllllllllll|}
-      \hline
-      $r_1$   &\un &0   &\un &0   &0   &0    &0   \\
-      $r_2$   &\un &0   &\un &0   &0   &0    &0   \\
-      $r_3$   &\un &0   &0   &\un &0   &0    &0   \\
-      $r_4$   &\un &0   &0   &\un &0   &0    &0   \\
-      $r_5$   &\un &0   &0   &0   &0   &0    &0   \\
-      $r_6$   &\un &\un &0   &0   &0   &0    &0   \\
-      $r_7$   &0   &\un &0   &0   &0   &\un  &0   \\
-      $r_8$   &0   &\un &0   &0   &\un &\un  &0   \\
-      $r_9$   &0   &\un &0   &0   &\un &\un  &0   \\
-      $r_{10}$ &0   &\un &0   &0   &\un &\un  &\un \\
-      $r_{11}$ &0   &\un &0   &0   &\un &0    &\un \\
-      \hline
-    \end{tabular}
-    &
-    \includegraphics[scale=0.3]{../img/pqtree.png}
-
-  \end{tabular}
-
-  \caption[\figtabsize An example for \PQtree]{\figtabsize An example for \PQtree.
-    Permuting the order of the left child of the root, we see that
-    \temptext{(d,a,b,c,e,f,g,h,i,j,k)} is a \COP order. Reversing the
-    order of the right child of the root, we see that
-    \temptext{(a,b,c,d,e,f,k,j,h,i,g) } is yet another \COP order. \temptext{PLACEHOLDER IMG.}
-    \cite[Fig.~1\endnote{\cite{mcc04}illustrates \COP on rows. Our
-      convention in this document is \COP on columns and thus example
-      matrix has been transposed. }]{mcc04}}
-  \label{fig:pqtree}
-\end{figure}
+\figpqtree
 
@@
 the same with either simpler procedure of construction, like \PCtree
-\cite{wlh01}\endnote{First appeared in
-  \cite{wlh92_isaac}.\tnote{check}} or a generalization of it, like
+\cite{wlh01} or a generalization of it, like
 generalized \PQtree \cite{mcc04}. \PQRtree is another generalization
@@
 
-\temptext{Expand a little more on PC trees and PQR trees.}
+\temptext{Expand a little more on PC trees and PQR trees.?}
 
-We will see more of \PQRtree in Section~\ref{sec:surveycertalgo}
+We will see more of generalized \PQtree in Section~\ref{sec:surveycertalgo}.
 
@@
   \begin{enumerate}[ -I-]
-  \item  Variations - PQR, PC etc. [MM96, Hsu01, McC04]\\
-    Generalized PQ trees McC04 - brief. Will be elaborated in
-    Section~\ref{sec:surveycertalgo}\\
-    Hsu also invented PC trees \cite{wlh01}\tnote{This result first
-      appeared inproc ISAAC92} which is claimed to be much easier to
-    implement.
-  \item Set theoretic characterizations [Hsu02, NS09]\\
-    \cite{wlh02} describes the simpler algorithm for
-    COT.\tnote[pressing]{in terms of what?}\\
-    \cite{nsnrs09} describes a characterization of consecutive-ones
-    property solely based on the cardinality properties of the set
-    representations of the columns (rows); every column (row) is
-    equivalent to a set that has the row (column) indices of the rows
-    (columns) that have one entries in this column (row). This is
-    interesting and relevant, especially to this thesis because it
-    simplifies COT to a great degree by reducing the solution search
-    space owing to the a simple set theoretic characterization.\\
    \item Parallel algorithms [AS95, BS03, CY91]
   \end{enumerate}
-
 }
@@
 
-\section{Certifying Algorithms}
+%\section{Certifying Algorithms}
+\section{COP for set systems}
 \label{sec:surveycertalgo}
@@
 
-\temptext{ Certifying algorithm (Generalized PQ trees) [McC04] -- Move this to section on
-    certifying algo.
+Section~\ref{xxx} mentions how a binary matrix naturally maps to a
+system of sets.  A set can be constructed for each column of matrix
+with its elements being those row indices at which the column has \un
+s. Thus the collection of sets corresponding each column of the matrix
+forms a set system with universe as the set of all row indices of the
+matrix.  This simple construction is formally described in
+Definition~\ref{def:matrixsetsystem}.
+
+\begin{definition}%
+  \label{def:matrixsetsystem}%
+  [\emph{Set system of a binary matrix}.] Let $M$ be a binary matrix
+  of order $n \times m$ % , $r_i, i \in [n]$ be the rows in $M$
+  and $\{c_i \mid i \in [m]\}$ be the columns in $M$. A set system
+  $\cF_{M} = \{S_i \mid S_i \subseteq [n], i \in [m]\}$ is defined
+  such that for every column $c_i$ of $M$, set $S_i = \{j \mid m_{ji}
+  = \un \}$. The collection $\cF_M$ is the set system of matrix $M$.
+\end{definition}
+
+Thus the idea of \cop naturally maps to set sytems as well\endnote{As
+  seen in Section~\ref{theCourseScheduleExample}} as
+Definition~\ref{def:copforsets} states. This is equivalent to \COP for
+matrices seen in Definition~\ref{def:copmatrix}.
+
+\begin{definition}%
+  \label{def:copforsets}%
+  [\emph{\Cop for set systems% \cite{mcc04}
+  }.] A set system $\cF$ with universe $U$ has the \cop if it is
+  possible to assign a linear order to $U$ where each set $S \in \cF$
+  is an interval in the linear order.\tnote[important]{NOT formal
+    enough.}
+\end{definition}
+
+\temptext{ Certifying algorithm (Generalized PQ trees) [McC04] -- Move
+  this to section on certifying algo.
+
+Set theoretic characterizations [Hsu02, NS09]\\
+    \cite{wlh02} describes the simpler algorithm for
+    COT.\tnote[pressing]{in terms of what?}\\
+    \cite{nsnrs09} describes a characterization of consecutive-ones
+    property solely based on the cardinality properties of the set
+    representations of the columns (rows); every column (row) is
+    equivalent to a set that has the row (column) indices of the rows
+    (columns) that have one entries in this column (row). This is
+    interesting and relevant, especially to this thesis because it
+    simplifies COT to a great degree by reducing the solution search
+    space owing to the a simple set theoretic characterization.\\
 

### 2012-01-11 src/010chap1.tex
chapter 2 - started set system side of things. see diff for more details.
chapter 2 in edit still. see diff for changes.

@@
 
-As a special case, suppose all the apartments are on the same street
-or if they are all lined up on a single path, the street map becomes a
-tree that is just a path. Then the problem becomes what is called an
-{\em interval assignment problem}. The idea of interval assignment may
-not be obvious here; hence to see this, consider a different problem
-in {\WSI} where the classes for these study groups courses need to be
+\subsection{Special case}
+\label{sec:courseschedule}
+As a special case of the study group accommodation problem, suppose
+all the apartments are on the same street or if they are all lined up
+on a single path, the street map becomes a tree that is just a
+path. Then the problem becomes what is called an {\em interval
+  assignment problem}. The idea of interval assignment may not be
+obvious here; hence to see this, consider a different problem in
+{\WSI} where the classes for these study groups courses need to be
 scheduled during a day (or a week or any time period). Each study
@@
 
-\begin{definition}[\emph{Binary matrix}]
+\begin{definition}[\emph{Binary matrix}.]
+  \label{def:binmatrix}
   Let $M$ be an $n \times m$ matrix. $m_{i,j}$ denotes its $(i,j)$th
@@
 
-\begin{definition}[\emph{set system}]
-\end{definition}
-
 \begin{definition}[\emph{\Cop}]
-  \tnote[important]{define cop here. refer to it in chap 2 or
-    duplicate it using slash def }
+    \label{def:copmatrix}
   \begin{enumerate}
@@
 \begin{definition}[\emph{\Crop}]
+  \label{def:cropmatrix}
   \tnote[important]{define circ1 here. refer to it in chap 2 or
@@
 
+\begin{definition}[\emph{set system}]
+    \label{def:setsystem}
+
+    also include overlapping sets.
+\end{definition}
+
 \begin{definition}[\emph{Graph, Tree}]
-
+  \label{def:graphtree}  
 \end{definition}
@@
 \begin{definition}[\emph{Maximal clique}]
+  \label{def:maxclique}    
+\end{definition}
 
-\end{definition}
+If $n$ is a positive integer, $[n]$ denotes the set $\{1,2,... n\}$.
 
@@
 graphs are AT-free on at least one of the partitions in
-\cite{at72}\tnote{check the real meaning of AT FREE. }. The first linear
-time algorithm for COP testing (COT) was invented by \cite{bl76} using
-a data structure called PQ trees.  Since then several COT algorithms
-have been invented -- some of which involved variations of PQ trees
-\cite{mm96,wlh01,mcc04}, some involved set theory and ICPIA
-\cite{wlh02,nsnrs09}, parallel COT algorithms\cite{as95,bs03,ly91} and
-certifying algorithms\cite{mcc04}.
+\cite{at72}. The first linear time algorithm for COP testing (COT) was
+invented by \cite{bl76} using a data structure called \PQtrees.  Since
+then several COT algorithms have been invented -- some of which
+involved variations of \PQtrees \cite{mm96,wlh01,mcc04}, some involved
+set theory and ICPIA \cite{wlh02,nsnrs09}, parallel COT
+algorithms\cite{as95,bs03,ly91} and certifying algorithms\cite{mcc04}.
 
-The construction of PQ trees in \cite{bl76} draws on the close
-relationship of matrices with COP to interval graphs. A PQ tree of a
+The construction of \PQtrees in \cite{bl76} draws on the close
+relationship of matrices with \COP to interval graphs. A PQ tree of a
 matrix is one that stores all row (column) permutations of the matrix
@@
 
-\tnote{ -- sect 4.1 in cite:d08phd has many results
-  surveyed. hardness results, approx. results. results are usually for
-  a class of matrices $(a,b)$ where number {\un}s in columns and rows
-  are restriced to $a$ and $b$ . -- problem of flipping at most $k$
-  entries of $M$ to make it attain COP. this is NP complete
-  cite:b75-phd}\tnote{(1) scite:lb62 showed that interval graphs
-  are AT-free.  describe AT (2) show the close relationship b/w COP
-  and graphs sec 2.2, pg 31} \cite{at72} showed that a matrix that
-does not have COP have certain substructures that prevent it from
-having COP. Tucker classified these forbidden substructures into five
-classes of submatrices. This result is presented in the context of
-convex bipartite graphs which \cite{at72} proved to be
-AT-free\tnote{ check this up. give details. - doms'}. By
+\cite{at72} showed that a matrix that does not have COP have certain
+substructures that prevent it from having COP. Tucker classified these
+forbidden substructures into five classes of submatrices. This result
+is presented in the context of convex bipartite graphs which
+\cite{at72} proved to be AT-free in one of the partitions. By
 definition, convex bipartite graph have half adjacency matrices that
@@
 the corresponding row. \cite{at72} proves that this bipartite graph
-has no asteroidal triple if and only if the matrix has COP and goes on
-to identify the forbidden substructures for these bipartite
-graphs. The matrices corresponding to these substructures are the
-forbidden submatrices.
+has no asteroidal triple in vertex partition corresponding to rows if
+and only if the matrix has COP on columns and goes on to identify the
+forbidden substructures for these bipartite graphs. The matrices
+corresponding to these substructures are the forbidden submatrices.
 
@@
 columns. $(*,\Delta)${\em -matrices} are matrices with no restriction
-on number of {\un}s in any column but has at most $\Delta$ {\un}s in
+on number of {\un}s in any column but have at most $\Delta$ {\un}s in
 any row. {\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are

### 2012-01-11 src/020chap2.tex
chapter 2 in edit still. see diff for changes.

@@
   Let $G$ be a graph and $H$ be a bipartite graph.
-
   \begin{enumerate}[\ref{def:graphwithcop}--i.]
@@
       consecutive in the ordering \cite[Th. 8.1]{mcg04}.}
-
     \begin{enumerate}[a.]
@@
 
-Interval graphs and circular-arc graphs have a long history in
-research. The interest around them is due to their very desirable
-property that several problems that are NP-hard\tnote[minor]{all NPC
-  problems are NPH yes?} on general graphs are not so in these graph
-classes, \ie they are polynomial time solvable on them. In a similar
-fashion, a lot of problems that are hard on general matrices have
-efficient solutions on matrics with \COP or \CROP \cite[more citations
-pg.\,33]{d08phd}.
+Interval graphs\endnote { \cite{mcc04} cites that the problem of
+  recognizing interval graphs has significance in molecular
+  biology. Interestingly, in the late 1950s, before the structure of
+  DNA was well-understood, Seymour Benzer was able to show that the
+  intersection graph of a large number of fragments of genetic
+  material was an interval graph \cite{sb59}. This was regarded as
+  compelling evidence that genetic information was somehow arranged
+  inside a structure that had a linear topology which we now know to
+  be true\tnote{linear arrangement of DNA/genes?}.} and circular-arc
+graphs have a long history in research.  The interest around them is
+due to their very desirable property that several problems that are
+NP-hard \tnote[minor]{all NPC problems are NPH yes?} on general
+graphs, like finding a maximum clique or minimum coloring or
+independent set, are polynomial time solvable in these graph classes
+\cite{clrs01}.
+In a similar fashion, a lot of problems that are hard on general
+matrices have efficient solutions on matrices with \COP or \CROP
+\cite[more citations pg.\,33]{d08phd}.
 
@@
 having seen how useful \COP is in identifying or characterizing many
-types of graphs, we will now see results that study recognition of \COP
+types of graphs\endnote{\COP has several other applications
+  \tnote{fill up.}}, we will now see results that study recognition of \COP
 in matrices in the following section.
@@
      % 
-     1899 & \parbox[t]{\tabparboxw}{ First mention of \COP} & \cite{p59}  \\
+     1899 & \parbox[t]{\tabparboxw}{ First mention of \COP
+       (archaeology)} & \parbox[t]{2cm}{\temptext{by petrie -- cite
+         kendall 69 pacific journal of mathematics 1969 }}  \\
 
@@
 
+The first polynomial time algorithm for \COP testing was by
+\cite{fg65} which uses overlapping properties of columns with \un
+s. Their result has close relations to the characterization of
+interval graphs by \cite{gh64}. A graph $G$ is an interval graph \iff
+all its maximal cliques can be linearly ordered \stt for any vertex
+$v$ in $G$, all the cliques that $v$ is incident on are consecutive in
+this order. Clearly, this means that the maximal clique incidence
+matrix\endnote{
+  Definition~\ref{def:graphmatrices}~\ref{def::maxcliquematrix}} must
+have \COP on rows.
+
+A few years later, a deeply significant result based on very
+different ideas in understanding \COP came from Tucker which gave a
+combinatorial (negative) characterization of matrices with COP
+\cite{at72}. This result influenced most of the \COP results that
+followed in the literature \annote[pressing]{including linear time
+  algorithms}{did it? which ones?} for \COP recognition.
+
 %\clearpage
@@
 % characterization of matrices with COP \cite{at72}.
-A polynomial time algorithm for \COP testing was seen for the first
-time in \cite{fg65} which uses overlapping properties of columns with
-\un s. A few years later, a deeply significant result based on very
-different ideas in understanding \COP came from Tucker which gave a
-combinatorial (negative) characterization of matrices with COP
-\cite{at72}. This result influenced most of the \COP results that
-followed in the literature \annote[pressing]{including linear time
-  algorithms}{did it? which ones?} for \COP recognition.
 
@@
 
-Theorem~\ref{th:tuckeratfree} proves that convex bipartite graphs
+Theorem~\ref{th:tuckeratfree} states that convex bipartite graphs
 cannot have {\em asteroidal triples}\endnote{If $G = (V,E)$ is a
@@
   half adjacency matrix has \COP columns (rows).}.
-Theorem~\ref{th:tuckerforbidden} shows what are the structures in a
+Theorem~\ref{th:tuckerforbidden} lists the structures in a
 bipartite graph that force one of its vertex partitions to have
@@
 %\clearpage
-\subsection{$PQ$ tree -- a linear COT algorithm}
+\subsection{Booth and Lueker's $PQ$ tree -- a linear COT algorithm}
 
 Booth and Lueker in their paper \cite{bl76} gave the first linear
-algorithm for \cop testing by way of an interval graph recognition
-algorithm. Their result has close relations to the characterization of
-interval graphs by \cite{gh64}. A graph $G$ is an interval graph \iff
-all its maximal cliques can be linearly ordered \stt for any vertex
-$v$ in $G$, all the cliques that $v$ is incident on are consecutive in
-this order. Clearly, this means that the maximal clique incidence
-matrix\footnote{
-  Definition~\ref{def:graphmatrices}~\ref{def::maxcliquematrix}} must
-have \COP on rows. This algorithm constructs a data structure called
-\PQtree. A \PQtree represents all the \COP orderings of the matrix it
-is associated with. \cite{bl76}'s algorithm uses the fact that if a
-matrix has \COP, a \PQtree for it can be constructed. It is
+algorithm\endnote{Time complexity is $O\left(m+n+f\right)$ where $m
+  \times n$ is the order of the input matrix and $f$ is the number of $\un$s in it.} for \cop testing while given a linear time interval graph
+recognition algorithm by a simplification of \tnote{cite - A. Lempel,
+  S. Even and I. Cederbaum, An Algorithm for Planarity Testing of
+  Graphs, Theory of Graphs, ed., P. Rosenstiehl, Gordon and Breach,
+  New York, (1967), 215-232.  }'s planarity test algorithm.
+% Hsu in wlh01 said BL76 gave a planarity test - "the induced PQ-tree
+% algorithm can considerably simplify Booth and Lueker’s modification
+% of Lempel, Even and Cederbaum’s planarity test. " 
+%
+% McConnell in hm03 said the first linear algo for interval graph
+% recog was by BL76 - "Benzer’s work and other applications of
+% interval graphs motivated a search for eﬃcient algorithms for
+% recognizing interval graphs, and for constructing a set of intervals
+% to represent the graph when it is one [8]. A linear-time algorithm
+% was given in 1976 by Booth and Lueker [2]." 
+\cite{bl76} introduces a data structure called \PQtree and their \COP
+testing algorithm is a constructive one that outputs a \PQtree if the
+input has \COP. A \PQtree represents all the \COP orderings of the
+matrix it is associated with. \cite{bl76}'s algorithm uses the fact
+that if a matrix has \COP, a \PQtree for it can be constructed. It is
 interesting to note that aside from interval graph recognition and
@@
 \begin{definition}[\emph{\PQtree \cite{bl76, mcc04}}]\\
-  A \PQtree of matrix $M$ with \COP, is a tree with the following properties.
+  A \PQtree of matrix $M$ with \COP on columns (rows), is a tree with the following properties.
   \begin{enumerate}[i.]
@@
       versa.} for $M$.
-  \item Every non-leaf node in the tree are labeled $P$ or $Q$.
+  \item Every non-leaf node in the tree is labeled $P$ or $Q$.
   \item \label{def::nodep} The children of $P$ nodes are
@@
   \item \label{def::nodeq} The children of $Q$ nodes are linearly
-    ordered. Their order can be reversed to get obtain a new \COP
+    ordered. Their order can be reversed to obtain a new \COP
     order for $M$.
@@
 
-After the invention of \PQtrees, there has been several variants of
-the same with either simpler procedure of construction, like \PCtree
-\cite{wlh01} or a generalization of it, like
-generalized \PQtree \cite{mcc04}. \PQRtree is another generalization
-of \PQtree by \cite{mm96,mpt98}. There algorithm was improved by
-\cite{tm05} \tnote{improved in terms of
-  what?}. % describes an improved algorithm to
+This algorithm is notoriously difficult to program. In the procedure
+to change the \PQtree at each iteration, nodes are considered from
+leaves to tree. At each node considered, it uses one of nine templates
+to determine how the tree must be altered in the vicinity of this
+node. Recognition of this template poses a difficult challenge in
+terms of implementing it. Each template is actually a representative
+of a larger class of similar templates, which must be dealt with
+explicitly by a program\cite{mcc04}.
+
+After the invention of \PQtrees, presumably due to the implementation
+challenge it posed, there has been several variants of the same in the
+literature, like \PCtree
+\cite{sh99,wlh01,hm03}, generalized \PQtree \cite{km89,mcc04}, \PQRtree
+\cite{mm96,mpt98} etc. Most of these are generalizations of \PQtree --
+for instance, \PCtree is generalized to matrices with \CROP, \PQRtree and
+generalized \PQtree are generalized to matrices and set systems with
+or without \COP. \cite{km89} invented a modified form of \PQtree a simpler
+incremental update of the tree only for recognizing interval graphs.
+\cite{kr88} constructed efficient parallel algorithms for manipulating
+PQ trees. \temptext{dom chapter 2 pg 40 Variations of PQTrees first
+  para - summarize.}
+
+In the next few following sections we will see some of these
+variations.
+%Their algorithm was improved by \cite{tm05}
+%\tnote{improved in terms of what?}. 
+% describes an improved algorithm to
 % build PQR trees.
 
-\temptext{Expand a little more on PC trees and PQR trees.?}
+\subsection{\PCtree -- a generalization of \PQtree}
 
-We will see more of generalized \PQtree in Section~\ref{sec:surveycertalgo}.
+\PCtree is a generalization of \PQtree. It is a data structure that is
+analogous to \PQtree but for matrices with \crop. \PCtree was
+introduced by \cite{sh99} for the purpose of planarity testing where
+this data structure represents partial embeddings of planar
+graphs. % noted from wlh01 abstract.
+In \cite{wlh01}, Hsu reintroduces \PCtree as a generalization of
+\PQtree and shows how it simplifies \cite{bl76}'s planarity test by
+making the \PQtree construction much less complicated. Later
+\cite{hm03}, discovers that \PCtree is a representation of all \crop
+orders of a matrix when it is unrooted. \PCtree presented in
+\cite{wlh01} is rooted; however the construction of \PCtree is the
+same in both results. The property of being unrooted is necessary in
+order to use \PCtree as a data structure for encoding circular
+ordering. Definition~\ref{def:pctree} defines \PCtree.
 
-\temptext{
-  \begin{enumerate}[ -I-]
-   \item Parallel algorithms [AS95, BS03, CY91]
+\begin{definition}
+  \label{def:pctree}
+  [\emph{\PCtree} \cite{wlh01,d08phd}.]  A \PCtree of matrix $M$ with
+  \CROP on columns (rows), is a tree with the following properties.
+  \begin{enumerate}[i.]
+    \singlespacing
+  \item Is unrooted -- thus it has (a) no parent child relationship
+    between nodes (b) there is no left to right (or vice versa)
+    ordering.
+  \item Each leaf uniquely represents a row (column) of $M$. The leaf
+    order of the tree gives a \CROP order for column (row) for
+    $M$. Moreover, any sequence obtained by considering the leaves in
+    clockwise or counter-clockwise order describes a \CROP order for
+    $M$.
+  \item Every non-leaf node in the tree is labeled $P$ or $C$.
+  \item The neighbors of $P$ nodes can be permuted in any fashion to
+    obtain a new \CROP order for $M$.
+  \item \label{def::pcmirror} The tree can be changed by applying the following
+    ``mirroring'' operation to obtain a new \CROP order for $M$. Root
+    the \PCtree at a neighbor of a $C$-node, $v$ and mirror the
+    subtree whose root is $v$ and finally unrooting the
+    tree. Mirroring a subtree is done by putting the children of every
+    node of the subtree in reverse order.
   \end{enumerate}
-}
+\end{definition}
 
-\tnote[pressing]{ The results in cite~{bl76} on COT are based on the
-   result that interval graphs are AT-free chordal graphs.
- -  Tucker's?}%
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[scale=0.4]{../img/pctree_hm03.png}
 
-\tnote[pressing]{cite~{fg65} uses these results to give the first polynomial
-  time algorithm for COT. WHICH RESULTS? \ \ \ \ \ check. how do they use it?}%
+  \centering
+  \includegraphics[scale=0.4]{../img/pctree_dom.png}
+  \caption{\figtabsize \PCtree \temptext{ PLACEHOLDER 
+IMGS \cite{hm03, d08phd}}}
+  \label{fig:pctree}
+\end{figure}
 
-\tnote[pressing]{ADD: peo exists iff chordal. lexicographic BFS
-  [tag:chordalGraph]} %
+As a data structure when \PCtree is compared with \PQtree, the
+differences are, (i) it is unrooted, (ii) it represents all \CROP order of a
+matrix (iii) it has $C$ nodes instead of $Q$ nodes which can be
+``mirrored'' (operation defined in Definition~\ref{def:pctree}~\ref{def::pcmirror}).
+The algorithms of construction of \PQtree in \cite{bl76} and that of
+\PCtree in \cite{wlh01, hm03} starkly differ since the latter is a
+much simplified procedure.
+
+% WLH01 abstract:
+% [..] The original implementation of the PQ-tree algorithms by Booth and
+% Lueker using nine templates in each bottom-up iteration is rather
+% complicated. Also the complexity analysis is rather intricate. We give
+% a very simple linear time PC-tree algorithm with the following
+% advantages: (1) it does not use any template; (2) at each iteration,
+% it does all necessary tree-modification operations in one batch and
+% does not involve the node-by-node bottom-up matching; (3) it can be
+% used naturally to test the circular ones property in matrices; (4) the
+% induced PQ-tree algorithm can considerably simplify Booth and Lueker’s
+% modification of Lempel, Even and Cederbaum’s planarity test.
+
+\tnote[FYI]{ The results in cite~{bl76} on COT are based on the
+   result that interval graphs are AT-free chordal graphs cite~{lb62}?
+ No they don't seem related. this is the stupid bl lb confusion.}%
 
 %\clearpage      % Print pending floats
-\section{Optimization problems in COP}
-\label{sec:surveycopopt}
-\tnote{Expand on ref:sec:optcop}
+%\section{Certifying Algorithms}
+\subsection{Generalized \PQtree\ --  COP for set systems}
+\label{sec:surveycertalgo}
 
-%{\color{blue} \footnotesize
-\temptext{
-So far we have been concerned about matrices that have the consecutive
-ones property. However in real life applications, it is rare that data
-sets represented by binary matrices have COP, primarily due to the
-noisy nature of data available. At the same time, COP is not arbitrary
-and is a desirable property in practical data representation
-\cite{co98,jkckv04,k77}. In this context, there are several
-interesting problems when a matrix does not have COP but is ``close''
-to having COP or is allowed to be altered to have COP. These are the
-optimization problems related to a matrix which does not have
-COP. Some of the significant problems are surveyed in this section.
-
-\tnote{ -- sect 4.1 in cite:d08phd has many results
-  surveyed. hardness results, approx. results. results are usually for
-  a class of matrices $(a,b)$ where number {\un}s in columns and rows
-  are restriced to $a$ and $b$ . -- problem of flipping at most $k$
-  entries of $M$ to make it attain COP. this is NP complete
-  cite:b75-phd}\tnote{(1) scite:lb62 showed that interval graphs
-  are AT-free.  describe AT (2) show the close relationship b/w COP
-  and graphs sec 2.2, pg 31} \cite{at72} showed that a matrix that
-does not have COP have certain substructures that prevent it from
-having COP. Tucker classified these forbidden substructures into five
-classes of submatrices. This result is presented in the context of
-convex bipartite graphs which \cite{at72} proved to be
-AT-free\tnote{ check this up. give details. - doms'}. By
-definition, convex bipartite graph have half adjacency matrices that
-have COP on either rows or columns (graph is biconvex if it has COP on
-both)\cite{d08phd}. A half adjacency matrix is a binary matrix
-representing a bipartite graph as follows. The set of rows and the set
-of columns form the two partitions of the graph. Each row node is
-adjacent to those nodes that represent the columns that have {\un}s in
-the corresponding row. \cite{at72} proves that this bipartite graph
-has no asteroidal triple if and only if the matrix has COP and goes on
-to identify the forbidden substructures for these bipartite
-graphs. The matrices corresponding to these substructures are the
-forbidden submatrices.
-
-Once a matrix has been detected to not have COP (using any of the COT
-algorithms mentioned earlier), it is naturally of interest to find out
-the smallest forbidden substructure (in terms of number of rows and/or
-columns and/or number of entries that are {\un}s). \cite{d08phd}
-discusses a couple of algorithms which are efficient if the number of
-{\un}s in a row is small. This is of significance in the case of
-sparse matrices where this number is much lesser than the number of
-columns. $(*,\Delta)${\em -matrices} are matrices with no restriction
-on number of {\un}s in any column but has at most $\Delta$ {\un}s in
-any row. {\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are
-similar problems which deals with inducing COP on a matrix. In {\sc
-  Min COS-R (Min COS-C)} the question is to find the minimum number of
-rows (columns) that must be deleted to result in a matrix with COP.
-In the dual problem {\sc Max COS-R (Max COS-C)} the search is for the
-maximum number of rows (columns) that induces a submatrix with
-COP. Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
-a submatrix $M'$ with all columns\tnote{check if b75 deals with
-  COP col or COP row. also is it any submatrix with k less than r rows
-  or submatrix must have all columns?} but a maximum cardinality
-subset of rows such that $M'$ has COP is NP complete. \cite{hg02}
-corrects an error of the abridged proof of this reduction as given in
-\cite{gj79}.  \cite{d08phd} discusses all these problems in detail
-giving an extensive survey of the previously existing results which
-are almost exhaustively all approximation results and hardness
-results. Taking this further, \cite{d08phd} presents new results in
-the area of parameterized algorithms for this
-problem\tnote{elaborate - what are the results?}.
-
-Another problem is to find the minimum number of entries in the matrix
-that can be toggled to result in a matrix with COP.  \cite{v85}
-discusses approximation of {\sc COP Augmentation} which is the problem
-of changing of the minimum number of zero entries to {\un}s so that
-the resulting matrix has COP. As mentioned earlier, this problem is
-known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
-using a reduction to the longest path problem, \tnote{or is it a
-  survey of another result?  check.} that finding a Tucker's forbidden
-submatrix of at least $k$ rows is NP complete. \tnote{how is this
-  different from booth's 75 result??}  \tnote{where should this go?
-  cite|tz04 (approx submatrix with COP sparse matrices)}
-
-\cite{jkckv04} discusses the use of matrices with almost-COP (instead
-of one block of consecutive {\un}s, they have $x$ blocks, or {\em
-  runs}, of consecutive {\un}s and $x$ is not too large) in the
-storage of very large databases.  The problem is that of reordering of
-a binary matrix such that the resulting matrix has at most $k$ runs of
-{\un}s. This is proved to be NP hard using a reduction from the
-Hamiltonian path problem.\tnote{Theorem 2.1 in jkckv} \tnote{(1) A
-  connection of COP problem to the travelling salesman problem is also
-  introduced. what does this mean? -- COP can be used as a tool to
-  reorder $0.5T \le runs(M) \le T$. (2) The optimization version of
-  the $k$-run problem, i.e. minimization of number of blocks of ones
-  is proven to be NP complete by cite:k77}\tnote{are these two the
-  same?} \tnote{what is the reduction?} \tnote{other problems similar
-  to COP -- cite:ckl96 (ILP, circ ones, one drop) -- cite:th98
-  (generalization of COP - minimax, biotonic column) Tucker}
-
-}
-
-%\section{Certifying Algorithms}
-\section{COP for set systems}
-\label{sec:surveycertalgo}
-%\tnote{Expand on sec:appcertalgo} \tnote{ (certification McC04
-%theme)}
-
-Section~\ref{xxx} mentions how a binary matrix naturally maps to a
-system of sets.  A set can be constructed for each column of matrix
-with its elements being those row indices at which the column has \un
-s. Thus the collection of sets corresponding each column of the matrix
-forms a set system with universe as the set of all row indices of the
-matrix.  This simple construction is formally described in
-Definition~\ref{def:matrixsetsystem}.
+Section~\ref{sec:motive} mentions how a binary matrix naturally maps
+to a system of sets.  A set can be constructed for each column of
+matrix with its elements being those row indices at which the column
+has \un s. Thus the collection of sets corresponding each column of
+the matrix forms a set system with universe as the set of all row
+indices of the matrix.  This simple construction is formally described
+in Definition~\ref{def:matrixsetsystem}.
 
@@
 Thus the idea of \cop naturally maps to set sytems as well\endnote{As
-  seen in Section~\ref{theCourseScheduleExample}} as
+  seen in Section~\ref{sec:courseschedule}} as
 Definition~\ref{def:copforsets} states. This is equivalent to \COP for
@@
 
-\temptext{ Certifying algorithm (Generalized PQ trees) [McC04] -- Move
-  this to section on certifying algo.
+\tnote{What does n89 say? download paper from cornell website.}
+Generalized \PQtree is a data structure defined in \cite{mcc04} for
+arbitrary set systems. If the set system has \COP, the generalized
+\PQtree is identical to its \PQtree.
+
+An important observation made by \cite{mm96} is that if $\cF$ is a set
+system with \COP then, aside from every set in $\cF$ being consecutive
+(after applying the \COP order) the following sets must also be
+consecutive for any $A, B \in \cF$ 
+\begin{enumerate}
+  \item The intersection $A \cap B$ 
+  \item The union $A \cup B$ if $A \cap B \ne \emptyset$
+  \item The relative complements $A \setminus B$ and $B \setminus A$
+    if $B \nsubseteq A$ and  $A \nsubseteq B$ respectively.
+  \item Additionally the sets $U$, singleton sets and the empty set $\emptyset$
+    (collectively called {\em trivial sets} in \cite{mm96}) are consecutive in all orderings
+    of the universe. $\emptyset$ is considered consecutive by convention.
+\end{enumerate}
+
+For any set system $\cF$, the smallest super set system containing
+$\cF$ and the trivial sets that is closed under the above operations
+is called the {\em weak closure} of $\cF$, denoted $\cW(\cF)$
+\cite[Def.~3.2]{mcc04} \cite[Def.~2]{mm96}.
+
+What is interesting is that the set of \COP permutations of $\cF$ is
+the same as that of $\cW(\cF)$. \cite[Th.~3]{mm96}
+
+The generalized \PQtree of $\cF$ is defined as the {\em decomposition
+  tree} of its weak closure $\cW(\cF)$. Decomposition tree is the
+inclusion tree (inclusion p.o. Hasse diagram) of the strong elements
+of $\cW(\cF)$ with labels {\em prime, linear} and {\em degenerate}.\cite{mcc04}
+
+\temptext{ Certifying algorithm (Generalized PQ trees) [McC04] --
 
@@
 
-\section{COP in Relational Database Model}
-\label{sec:surveyrdbm}
-\tnote{Expand on sec:apprdbm} \tnote{(set systems theme)}
+% \section{COP in Relational Database Model}
+% \label{sec:surveyrdbm}
+% \tnote{Expand on sec:apprdbm} \tnote{(set systems theme)}
+\section{Optimization problems in COP}
+\label{sec:surveycopopt}
+\tnote{Expand on ref:sec:optcop}
+
+\temptext{ So far we have been concerned about matrices that have the
+  consecutive ones property. However in real life applications, it is
+  rare that data sets represented by binary matrices have COP,
+  primarily due to the noisy nature of data available. At the same
+  time, COP is not arbitrary and is a desirable property in practical
+  data representation \cite{co98,jkckv04,k77}. In this context, there
+  are several interesting problems when a matrix does not have COP but
+  is ``close'' to having COP or is allowed to be altered to have
+  COP. These are the optimization problems related to a matrix which
+  does not have COP. Some of the significant problems are surveyed in
+  this section.
+
+\tnote{ -- sect 4.1 in cite:d08phd has many results surveyed. hardness
+  results, approx. results. results are usually for a class of
+  matrices $(a,b)$ where number {\un}s in columns and rows are
+  restriced to $a$ and $b$ . -- problem of flipping at most $k$
+  entries of $M$ to make it attain COP. this is NP complete
+  cite:b75-phd}%
+\tnote{(1) scite:lb62 showed that interval graphs are chordal and
+  AT-free.} %
+\temptext{couple of lines refering to tucker's submatrices. refer
+  earlier section.}
+
+Once a matrix has been detected to not have COP (using any of the COT
+algorithms mentioned earlier), it is naturally of interest
+
+\begin{enumerate}
+\item  to find out
+the smallest forbidden substructure (in terms of number of rows and/or
+columns and/or number of entries that are {\un}s). \cite{d08phd}
+discusses a couple of algorithms which are efficient if the number of
+{\un}s in a row is small. This is of significance in the case of
+sparse matrices where this number is much lesser than the number of
+columns.
+
+\item $(*,\Delta)${\em -matrices} are matrices with no restriction
+on number of {\un}s in any column but has at most $\Delta$ {\un}s in
+any row. 
+{\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are
+similar problems which deals with {\em inducing COP} on a matrix.
+
+\begin{enumerate}
+\item In the dual problem {\sc Max COS-R (Max COS-C)} the search is for the
+maximum number of rows (columns) that induces a submatrix with
+COP.
+\item In {\sc
+  Min COS-R (Min COS-C)} the question is to find the minimum number of
+rows (columns) that must be deleted to result in a matrix with COP.
+\end{enumerate}
+
+Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
+a submatrix $M'$ with all columns\tnote{check if b75 deals with
+  COP col or COP row. also is it any submatrix with k less than r rows
+  or submatrix must have all columns?} but a maximum cardinality
+subset of rows such that $M'$ has COP is NP complete. \cite{hg02}
+corrects an error of the abridged proof of this reduction as given in
+\cite{gj79}.  \cite{d08phd} discusses all these problems in detail
+giving an extensive survey of the previously existing results which
+are almost exhaustively all approximation results and hardness
+results. Taking this further, \cite{d08phd} presents new results in
+the area of parameterized algorithms for this
+problem\tnote{elaborate - what are the results?}.
+
+\item Another problem is to find the minimum number of entries in the matrix
+that can be toggled to result in a matrix with COP.  \cite{v85}
+discusses approximation of {\sc COP Augmentation} which is the problem
+of changing of the minimum number of zero entries to {\un}s so that
+the resulting matrix has COP. As mentioned earlier, this problem is
+known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
+using a reduction to the longest path problem, \tnote{or is it a
+  survey of another result?  check.} that finding a Tucker's forbidden
+submatrix of at least $k$ rows is NP complete. \tnote{how is this
+  different from booth's 75 result??}  \tnote{where should this go?
+  cite|tz04 (approx submatrix with COP sparse matrices)}
+
+\item \cite{jkckv04} discusses the use of matrices with almost-COP (instead
+of one block of consecutive {\un}s, they have $x$ blocks, or {\em
+  runs}, of consecutive {\un}s and $x$ is not too large) in the
+storage of very large databases.  The problem is that of reordering of
+a binary matrix such that the resulting matrix has at most $k$ runs of
+{\un}s. This is proved to be NP hard using a reduction from the
+Hamiltonian path problem.\tnote{Theorem 2.1 in jkckv} \tnote{(1) A
+  connection of COP problem to the travelling salesman problem is also
+  introduced. what does this mean? -- COP can be used as a tool to
+  reorder $0.5T \le runs(M) \le T$. (2) The optimization version of
+  the $k$-run problem, i.e. minimization of number of blocks of ones
+  is proven to be NP complete by cite:k77}\tnote{are these two the
+  same?} \tnote{what is the reduction?} \tnote{other problems similar
+  to COP -- cite:ckl96 (ILP, circ ones, one drop) -- cite:th98
+  (generalization of COP - minimax, biotonic column) Tucker}
+
+\end{enumerate}
+
+}
 
@@
 \tnote{Expand on sec:appgraphiso} \tnote{(canonization theme)}
+
+\tnote[pressing]{ADD: peo exists iff chordal. lexicographic BFS
+  [tag:chordalGraph]} %
 

### 2012-01-11 src/020chap2.tex
made modes in section for PQR trees

@@
 
-Thus, effectively, there exists a bijection between set of matrices
-with \COP and the set of \PQtrees (accurately speaking, each matrix
-with \COP bijectively maps to an equivalence class of \PQtrees due to
-properties (\ref{def::nodep}) and (\ref{def::nodeq})). See
-Figure~\ref{fig:pqtree} for an example of \PQtree.
+See Figure~\ref{fig:pqtree} for an example of \PQtree. It may be noted
+that there is no way an empty set of \COP orderings can be represented
+in this data structure. For this reason, \PQtree is undefined for
+matrices that do not have \COP.  Thus effectively, there exists a
+bijection between set of matrices with \COP and the set of \PQtrees
+(accurately speaking, each matrix with \COP bijectively maps to an
+equivalence class of \PQtrees resulting from properties (\ref{def::nodep}) and
+(\ref{def::nodeq})).
 
@@
 
-This algorithm is notoriously difficult to program. In the procedure
-to change the \PQtree at each iteration, nodes are considered from
-leaves to tree. At each node considered, it uses one of nine templates
-to determine how the tree must be altered in the vicinity of this
+Judging from notes in literature, this algorithm is apparently
+notoriously difficult to program. In the procedure to modify the
+\PQtree at each iteration, nodes are considered from leaves to
+tree. At each node considered, it uses one of nine templates to
+determine how the tree must be altered in the vicinity of this
 node. Recognition of this template poses a difficult challenge in
@@
 challenge it posed, there has been several variants of the same in the
-literature, like \PCtree
-\cite{sh99,wlh01,hm03}, generalized \PQtree \cite{km89,mcc04}, \PQRtree
-\cite{mm96,mpt98} etc. Most of these are generalizations of \PQtree --
-for instance, \PCtree is generalized to matrices with \CROP, \PQRtree and
-generalized \PQtree are generalized to matrices and set systems with
-or without \COP. \cite{km89} invented a modified form of \PQtree a simpler
-incremental update of the tree only for recognizing interval graphs.
-\cite{kr88} constructed efficient parallel algorithms for manipulating
-PQ trees. \temptext{dom chapter 2 pg 40 Variations of PQTrees first
-  para - summarize.}
+literature, like \PCtree \cite{sh99,wlh01,hm03}, generalized \PQtree
+\cite{km89,mcc04}, \PQRtree \cite{mm96,mpt98} etc. Most of these are
+generalizations of \PQtree -- for instance, \PCtree is generalized to
+matrices with \CROP, \PQRtree and generalized \PQtree are generalized
+to matrices and set systems with or without \COP. \cite{km89} invented
+a modified form of \PQtree a simpler incremental update of the tree
+only for recognizing interval graphs.  \cite{kr88} constructed
+efficient parallel algorithms for manipulating PQ trees. \temptext{dom
+  chapter 2 pg 40 Variations of PQTrees first para - summarize.}
 
@@
 % build PQR trees.
+
+%\section{Certifying Algorithms}
+\subsection{\PQRtree\ --  COP for set systems}
+\label{sec:surveycertalgo}
+
+Section~\ref{sec:motive} mentions how a binary matrix naturally maps
+to a system of sets.  A set can be constructed for each column of
+matrix with its elements being those row indices at which the column
+has \un s. Thus the collection of sets corresponding each column of
+the matrix forms a set system with universe as the set of all row
+indices of the matrix.  This simple construction is formally described
+in Definition~\ref{def:matrixsetsystem} along with the idea of \cop
+for set systems\endnote{As seen in Section~\ref{sec:courseschedule}}.
+
+\begin{definition}%
+  \label{def:matrixsetsystem}%
+  % [\emph{Set system of a binary matrix}.]
+  [\emph{\Cop for set systems}.]  Let $M$ be a binary matrix of order
+  $n \times m$ % , $r_i, i \in [n]$ be the rows in $M$
+  and $\{c_i \mid i \in [m]\}$ be the columns in $M$. A set system
+  $\cF_{M} = \{S_i \mid S_i \subseteq [n], i \in [m]\}$ is defined
+  such that for every column $c_i$ of $M$, set $S_i = \{j \mid m_{ji}
+  = \un \}$. The collection $\cF_M$ is the {\em set system of binary
+    matrix} $M$.\\
+  A set system $\cF$ from universe $U = \{1, 2, \ldots, n\}$ has the
+  \cop if it is possible to assign a linear order $\lambda$ to $U$
+  where each set $S \in \cF$ appears as a consecutive subsequence of
+  $\lambda$ (an interval) in the linear order.
+\end{definition}
+
+It is easy to see the equivalence of this definition to \COP for
+matrices in Definition~\ref{def:copmatrix}.
+
+Generalized \PQtree is a data structure defined in \cite{n89, mcc04} for
+arbitrary set systems. If the set system has \COP, the generalized
+\PQtree is identical to its \PQtree. We will use the notation from
+\cite{n89} and call generalized \PQtrees as \gPQtrees.
+
+A \gPQtree of a given set system $\cF$ from universe $U$ represents
+all the sets from $\power{U}$ that have trivial intersections with
+every set in $\cF$.
+
+Two sets $A$ and $B$ are said to have a trivial intersection if $A
+\cap B$ is one of the following -- $\emptyset$, $A$ or $B$. In other
+words, $A$ and $B$ are either disjoint or one is the subset of the
+other. {\em Trivial sets} are sets that have trivial intersections
+with any set in $\power{U}$. These are namely, $U$, singleton sets in
+$\power{U}$ and $\emptyset$\cite{n89, mm96}.
+
+An important observation made by \cite{mm96} is that if $\cF$ is a set
+system with \COP then, aside from every set in $\cF$ being consecutive
+(after applying the \COP order) the following sets must also be
+consecutive for any $A, B \in \cF$.
+\begin{enumerate}
+  \item The intersection $A \cap B$ 
+  \item The union $A \cup B$ if $A \cap B \ne \emptyset$
+  \item The relative complements $A \setminus B$ and $B \setminus A$
+    if $B \nsubseteq A$ and $A \nsubseteq B$ respectively.
+\end{enumerate}
+
+Additionally the {\em trivial sets} of universe $U$ denoted by
+$\cT(U)$, are namely -- $U$, singleton sets and the empty set
+$\emptyset$. Clearly, $\cT(U)$ are consecutive
+in all orderings of $U$. $\emptyset$ is considered consecutive by
+convention.
+
+Overlap components are computable in linear time \cite{mm95,wlh92}.
+
+For any set system $\cF$, the smallest super set system containing
+$\cF$ and the trivial sets that is closed under the above operations
+is called the {\em weak closure} of $\cF$, denoted $\cW(\cF)$
+\cite[Def.~3.2]{mcc04} \cite[Def.~2]{mm96}.
+
+What is interesting is that the set of \COP permutations of $\cF$ is
+the same as that of $\cW(\cF)$. \cite[Th.~3]{mm96}
+
+The generalized \PQtree of $\cF$ is defined as the {\em decomposition
+  tree} of its weak closure $\cW(\cF)$. Decomposition tree is the
+inclusion tree (inclusion p.o. Hasse diagram) of the strong elements
+of $\cW(\cF)$ with labels {\em prime, linear} and {\em degenerate}.\cite{mcc04}
+
+\temptext{ Certifying algorithm (Generalized PQ trees) [McC04] --
+
+Set theoretic characterizations [Hsu02, NS09]\\
+    \cite{wlh02} describes the simpler algorithm for
+    COT.\tnote[pressing]{in terms of what?}\\
+    \cite{nsnrs09} describes a characterization of consecutive-ones
+    property solely based on the cardinality properties of the set
+    representations of the columns (rows); every column (row) is
+    equivalent to a set that has the row (column) indices of the rows
+    (columns) that have one entries in this column (row). This is
+    interesting and relevant, especially to this thesis because it
+    simplifies COT to a great degree by reducing the solution search
+    space owing to the a simple set theoretic characterization.\\
+
+\cite{mcc04} describes a different approach to COT. While all previous
+COT algorithms gave the COP order if the matrix has the property but
+exited stating negative if otherwise, this algorithm gives an evidence
+by way of a certificate of matrix even when it has no COP. This
+enables a user to verify the algorithm's result even when the answer
+is negative. This is significant from an implementation perspective
+because automated program verification is hard and manual verification
+is more viable. Hence having a certificate reinforces an
+implementation's credibility. Note that when the matrix {\em has} COP,
+the COP order is the certificate.  The internal machinery of this
+algorithm is related to the weighted betweenness problem
+addressed\tnote{in what way??} in \cite{co98}.  \tnote{expand
+  on the COP order graph creation and it having to be bipartite for M
+  to have COP. and thus an odd cycle being an evidence of no COP.}
+
+}
 
@@
 %\clearpage      % Print pending floats
-%\section{Certifying Algorithms}
-\subsection{Generalized \PQtree\ --  COP for set systems}
-\label{sec:surveycertalgo}
-
-Section~\ref{sec:motive} mentions how a binary matrix naturally maps
-to a system of sets.  A set can be constructed for each column of
-matrix with its elements being those row indices at which the column
-has \un s. Thus the collection of sets corresponding each column of
-the matrix forms a set system with universe as the set of all row
-indices of the matrix.  This simple construction is formally described
-in Definition~\ref{def:matrixsetsystem}.
-
-\begin{definition}%
-  \label{def:matrixsetsystem}%
-  [\emph{Set system of a binary matrix}.] Let $M$ be a binary matrix
-  of order $n \times m$ % , $r_i, i \in [n]$ be the rows in $M$
-  and $\{c_i \mid i \in [m]\}$ be the columns in $M$. A set system
-  $\cF_{M} = \{S_i \mid S_i \subseteq [n], i \in [m]\}$ is defined
-  such that for every column $c_i$ of $M$, set $S_i = \{j \mid m_{ji}
-  = \un \}$. The collection $\cF_M$ is the set system of matrix $M$.
-\end{definition}
-
-Thus the idea of \cop naturally maps to set sytems as well\endnote{As
-  seen in Section~\ref{sec:courseschedule}} as
-Definition~\ref{def:copforsets} states. This is equivalent to \COP for
-matrices seen in Definition~\ref{def:copmatrix}.
-
-\begin{definition}%
-  \label{def:copforsets}%
-  [\emph{\Cop for set systems% \cite{mcc04}
-  }.] A set system $\cF$ with universe $U$ has the \cop if it is
-  possible to assign a linear order to $U$ where each set $S \in \cF$
-  is an interval in the linear order.\tnote[important]{NOT formal
-    enough.}
-\end{definition}
-
-\tnote{What does n89 say? download paper from cornell website.}
-Generalized \PQtree is a data structure defined in \cite{mcc04} for
-arbitrary set systems. If the set system has \COP, the generalized
-\PQtree is identical to its \PQtree.
-
-An important observation made by \cite{mm96} is that if $\cF$ is a set
-system with \COP then, aside from every set in $\cF$ being consecutive
-(after applying the \COP order) the following sets must also be
-consecutive for any $A, B \in \cF$ 
-\begin{enumerate}
-  \item The intersection $A \cap B$ 
-  \item The union $A \cup B$ if $A \cap B \ne \emptyset$
-  \item The relative complements $A \setminus B$ and $B \setminus A$
-    if $B \nsubseteq A$ and  $A \nsubseteq B$ respectively.
-  \item Additionally the sets $U$, singleton sets and the empty set $\emptyset$
-    (collectively called {\em trivial sets} in \cite{mm96}) are consecutive in all orderings
-    of the universe. $\emptyset$ is considered consecutive by convention.
-\end{enumerate}
-
-For any set system $\cF$, the smallest super set system containing
-$\cF$ and the trivial sets that is closed under the above operations
-is called the {\em weak closure} of $\cF$, denoted $\cW(\cF)$
-\cite[Def.~3.2]{mcc04} \cite[Def.~2]{mm96}.
-
-What is interesting is that the set of \COP permutations of $\cF$ is
-the same as that of $\cW(\cF)$. \cite[Th.~3]{mm96}
-
-The generalized \PQtree of $\cF$ is defined as the {\em decomposition
-  tree} of its weak closure $\cW(\cF)$. Decomposition tree is the
-inclusion tree (inclusion p.o. Hasse diagram) of the strong elements
-of $\cW(\cF)$ with labels {\em prime, linear} and {\em degenerate}.\cite{mcc04}
-
-\temptext{ Certifying algorithm (Generalized PQ trees) [McC04] --
-
-Set theoretic characterizations [Hsu02, NS09]\\
-    \cite{wlh02} describes the simpler algorithm for
-    COT.\tnote[pressing]{in terms of what?}\\
-    \cite{nsnrs09} describes a characterization of consecutive-ones
-    property solely based on the cardinality properties of the set
-    representations of the columns (rows); every column (row) is
-    equivalent to a set that has the row (column) indices of the rows
-    (columns) that have one entries in this column (row). This is
-    interesting and relevant, especially to this thesis because it
-    simplifies COT to a great degree by reducing the solution search
-    space owing to the a simple set theoretic characterization.\\
-
-\cite{mcc04} describes a different approach to COT. While all previous
-COT algorithms gave the COP order if the matrix has the property but
-exited stating negative if otherwise, this algorithm gives an evidence
-by way of a certificate of matrix even when it has no COP. This
-enables a user to verify the algorithm's result even when the answer
-is negative. This is significant from an implementation perspective
-because automated program verification is hard and manual verification
-is more viable. Hence having a certificate reinforces an
-implementation's credibility. Note that when the matrix {\em has} COP,
-the COP order is the certificate.  The internal machinery of this
-algorithm is related to the weighted betweenness problem
-addressed\tnote{in what way??} in \cite{co98}.  \tnote{expand
-  on the COP order graph creation and it having to be bipartite for M
-  to have COP. and thus an odd cycle being an evidence of no COP.}
-
-}
-
 % \section{COP in Relational Database Model}

### 2012-01-12 src/010chap1.tex
partative set theory TBD in chapter 2. beginning to work on chapter 3 now.

@@
   \{0, \un\}$ 
+\end{definition}
+
+\begin{definition}
+  A {\em permutation} $\lambda$ of a set $X = \{x_1, x_2, \ldots
+  x_n\}$ is a bijection $\lambda: X \rightarrow X$.  For simplicity,
+  sometimes, $\lambda$ is written as a sequence $x_{i_1}x_{i_2}\ldots
+  x_{i_n}$, where $i_j \in \{1, 2 \ldots n\}$ to mean that
+  $\lambda(x_j)= x_{i_j}$.\\
+  This document uses both notations as convenient in context.
 \end{definition}

### 2012-01-12 src/020chap2.tex
partative set theory TBD in chapter 2. beginning to work on chapter 3 now.

@@
   A set system $\cF$ from universe $U = \{1, 2, \ldots, n\}$ has the
-  \cop if it is possible to assign a linear order $\lambda$ to $U$
-  where each set $S \in \cF$ appears as a consecutive subsequence of
-  $\lambda$ (an interval) in the linear order.
+  {\em \cop} if it is possible to assign a linear order, \ie a
+  permutation $\sigma = u_1u_2\ldots u_n$ to $U$ \stt each set $S \in
+  \cF$ appears as a consecutive subsequence $u_{i}u_{i+1}\ldots
+  u_{i+k-1}$ (also called an interval) of $\lambda$ in the linear order, where
+  positive integers $k = |S|$ and $i \le n+1-k$.
 \end{definition}
@@
 
-Generalized \PQtree is a data structure defined in \cite{n89, mcc04} for
-arbitrary set systems. If the set system has \COP, the generalized
-\PQtree is identical to its \PQtree. We will use the notation from
-\cite{n89} and call generalized \PQtrees as \gPQtrees.
+Generalized \PQtree or \gPQtree is a data structure defined in \cite{n89} to
+represent all orthogonal sets of a set system $\cF$. {\em Orthogonal
+  sets} of $\cF$ are sets that do not overlap with any set in $\cF$. A
+data structure with the same name was later defined in \cite{mcc04} as
+part of a {\em substitution decomposition} for a set system
+$\cF$. \tnote{define substitutio decomp.} This is later used to define
+a certificate if $\cF$ has no \COP. \PQRtree is a data structure
+defined in \cite{mm96} as a data structure to represent any set system
+$\cF$ with additional information in their $R$-nodes if $\cF$ has no
+\COP.  All three of these data structures are proposed as
+generalizations of \cite{bl76}'s \PQtree and hence produce the
+$\PQtree$ if $\cF$ has \COP. As a whole, all these three data
+strutures are near-identical with their differences being notional.
+In this section, we will discuss the basic theory that they all hold
+and stick to the name of the data structure as \PQRtree.
 
-A \gPQtree of a given set system $\cF$ from universe $U$ represents
+A \PQRtree of a given set system $\cF$ from universe $U$ represents
 all the sets from $\power{U}$ that have trivial intersections with
-every set in $\cF$.
-
-Two sets $A$ and $B$ are said to have a trivial intersection if $A
-\cap B$ is one of the following -- $\emptyset$, $A$ or $B$. In other
-words, $A$ and $B$ are either disjoint or one is the subset of the
-other. {\em Trivial sets} are sets that have trivial intersections
-with any set in $\power{U}$. These are namely, $U$, singleton sets in
-$\power{U}$ and $\emptyset$\cite{n89, mm96}.
+every set in $\cF$.  Two sets $A$ and $B$ are said to have a {\em
+  trivial intersection} if $A \cap B$ is one of the following --
+$\emptyset$, $A$ or $B$. In other words, $A$ and $B$ are either (i)
+disjoint or (ii) one is the subset of the other. Such sets are also
+called {\em orthogonal sets} or said that $A$ is orthogonal to $B$. {\em Trivial subsets}
+of a universe $U$, denoted by $\cT(U)$, are sets that have trivial
+intersections with any set in $\power{U}$. Namely, $\cT(U)$ consists
+of $U$, singleton sets in $\power{U}$ and $\emptyset$\cite{n89, mm96}.
+In other words, the all sets presented in a \PQRtree share the
+property that they do not overlap with any set in $\cF$. We will now
+see how this helps in testing for \COP in $\cF$.
 
@@
 \end{enumerate}
+Also note that trivially, sets in $\cT(U)$ are consecutive in any
+permutation of $U$\endnote{$\emptyset$ is considered consecutive by
+  convention.}.
 
-Additionally the {\em trivial sets} of universe $U$ denoted by
-$\cT(U)$, are namely -- $U$, singleton sets and the empty set
-$\emptyset$. Clearly, $\cT(U)$ are consecutive
-in all orderings of $U$. $\emptyset$ is considered consecutive by
-convention.
-
-Overlap components are computable in linear time \cite{mm95,wlh92}.
-
-For any set system $\cF$, the smallest super set system containing
-$\cF$ and the trivial sets that is closed under the above operations
-is called the {\em weak closure} of $\cF$, denoted $\cW(\cF)$
-\cite[Def.~3.2]{mcc04} \cite[Def.~2]{mm96}.
+Any set system that is closed under the above operations and contains
+$\cT(U)$ is called {\complete} \cite{mm96} or a {\em weakly partative
+  family} \cite{mcc04}. Such a set system is easier to
+work with for reasons that will become apparent.\tnote{VAGUe!}
+For any set system $\cF$, the smallest super set system that is
+complete/weakly partative is called the {\em weak closure} of $\cF$,
+denoted $\cW(\cF)$ \cite[Def.~3.2]{mcc04}. \cite[Def.~2, smallest {\em
+  complete} supercollection]{mm96}.
 
 What is interesting is that the set of \COP permutations of $\cF$ is
-the same as that of $\cW(\cF)$. \cite[Th.~3]{mm96}
+the same as that of $\cW(\cF)$ \cite[Th.~3]{mm96}. The reason this is
+interesting is because it becomes apparent by this theory that it is
+easier to use $\cW(\cF)$ to create the $\PQRtree$ for $\cF$.
 
@@
 of $\cW(\cF)$ with labels {\em prime, linear} and {\em degenerate}.\cite{mcc04}
+
+Overlap components are computable in linear time \cite{mm95,wlh92}.
 

### 2012-01-12 src/030chap3.tex
Added more stuff to chapter 2. see diff.
partative set theory TBD in chapter 2. beginning to work on chapter 3 now.

@@
 permutation of the universe (set of row or column indices for COP of
-columns or rows, respectively) which permutes each set to its
+columns or rows, respectively) which permutes each set to an
 interval. The result in \cite{nsnrs09} characterizes interval
 assignments to the sets which can be obtained from a single
-permutation of the rows (columns). The cardinality of the interval
+permutation of the universe. The cardinality of the interval
 assigned to it must be same as the cardinality of the set, and the
@@
 assignment problem from any tree. We refer to this as the {\em tree
-  path labeling problem}. To elaborate, the problem is defined as
-follows -- given a hypergraph $\cF$ from universe (vertex set) $U$ and
-a target tree $T$, does there exist a bijection from $U$ to the
-vertices of $T$ such that when applied to elements of each hyperedge,
-it gives a path on $T$.  More formally, the problem definition is as
-defined by \CFTPL.
+  path labeling problem of path hypergraphs}. This is analogous to the
+interval labeling problem in literature \cite{kklv10} to interval
+hypergraphs. To elaborate, the problem is defined as follows -- given
+a hypergraph $\cF$ from universe (hypergraph vertex set) $U$ and a target tree
+$T$, does there exist a bijection $\phi$ from $U$ to the vertices of $T$ such
+that for each hyperedge when applied to its elements, $\phi$ gives a path on
+$T$.  More formally, the problem definition is as defined by \CFTPL.
 
@@
 problem, we were unable to discover an efficient algorithm for
-it. Hence we consider two special cases of \CFTPL -- namely, \CFTPLINT
-and \CFTPLKTREE on special target trees, intervals and \kstar,
+it. Hence we consider two special cases of \CFTPL\ -- namely, \CFTPLINT
+and \CFTPLKTREE on special target trees, intervals and \kstars,
 respectively.  Section~\ref{sec:spltargettree} discusses these
@@
 Section~\ref{sec:norestraint}. \tnote{put some more description
-  here. refer walcom.}
+  here. refer WG11.}
 
@@
 
-{\tt --- WALCOM}
+\temptext{ --- WG11 begin --- }
 
@@
 
+\temptext{ --- WG11 end --- }
+
 \section{Characterization of Feasible Tree Path Labeling}
@@
 
-% \begin{enumerate}[{(\icpplpr\ }i) \ \ \ ]
-% \item \label{pr:i} $|S| = |\cl(S)|$ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
-%   \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $S \in \cF$
-%   % \vspace{\topshrink}
-% \item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ \ \ \ \
-%   \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all distinct $S_1, S_2 \in \cF$
-%   % \vspace{\topshrink}
-% \item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
-%   \cap \cl(S_3)|$ \ \ \ for all distinct $S_1, S_2, S_3 \in \cF$
-% \end{enumerate}
 \begin{enumerate}[{(\icpplpr\ }i)]
@@
 
-{\tt WALCOM begin}
+\temptext{ --- WG11 begin --- }
 
@@
 
-\end{proof} {\tt WALCOM end}
+\end{proof} 
+\temptext{ --- WG11 end --- }
 
@@
 
-{\tt WALCOM begin}
+\temptext{ --- WG11 begin --- }
 \begin{lemma}
@@
 \end{proof}
-
-{\tt WALCOM end}
+\temptext{ --- WG11 end --- }
 
@@
 
-Consider a special case of ICPPL with the following properties.
-\begin{enumerate}
-\item Given tree $T$ is a path. Hence, all path labels are interval
-  labels.
-\item Only pairwise intersection cardinality preservation is
-  sufficient. i.e. property (iii) in ICPPL is not enforced.
-\item The filter algorithms do not have {\em \bf exit} statements.
-\end{enumerate}
-This is called an Intersection Cardinality Preservation Interval
-Assignment (ICPIA) \cite{nsnrs09}. This structure and its algorithm is
+Consider a special case of ICPPL with the following properties when the tree $T$
+is a path.  Hence, all path labels are can be viewed as intervals assigned to the
+sets in $\cF$.  It is shown, in \cite{nsnrs09}, that the filtering algorithms outlined
+above need only preserve pairwise intersection cardinalities, and higher level
+intersection cardinalities are preserved by the Helly Property of intervals.  Consequently,
+the filter algorithms do not need to ever evaluate the additional check to {\em \bf exit}.
+%\begin{enumerate}
+%\item Given tree $T$ is a path. Hence, all path labels are interval labels.
+%\item Only pairwise intersection cardinality
+%  preservation is sufficient. i.e. property (iii) in ICPPL is not enforced.
+%\item The filter algorithms do not have {\em \bf exit} statements.
+%\end{enumerate}
+%This is called an Intersection Cardinality Preservation Interval
+%Assignment (ICPIA) \cite{nsnrs09}. 
+This structure and its algorithm is
 used in the next section for finding tree path labeling from a
@@
 
+\temptext{repetition}
+
+In this section we consider the problem of assigning paths from a
+$k$-subdivided star $T$ to a given set system $\cF$ such that each set
+$X \in \cF$ is of cardinality at most $k+2$.  Secondly, we present our
+results only for the case when overlap graph $\bO(\cF)$ is connected.
+%The
+%overlap graph is well-known from the work of
+%\cite{kklv10,nsnrs09,wlh02}.  We use the notation in
+%\cite{kklv10}. Recall from Section~\ref{sec:prelims} that hyperedges
+%$S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$ and $S'$
+%have a non-empty intersection but neither of them is contained in the
+%other. The overlap graph $\bO(\cF)$ is a graph in which the vertices
+%correspond to the sets in $\cF$, and the vertices corresponding to the
+%hyperedges $S$ and $S'$ are adjacent if and only if they overlap.  Note
+%that the intersection graph of $\cF$, $\bI(\cF)$ is different from
+%$\bO(\cF)$ and $\bO(\cF) \subseteq \bI(\cF)$.  
+A connected component of $\bO(\cF)$ is called an overlap component of
+$\cF$.  An interesting property of the overlap components is that any
+two distinct overlap components, say $\cO_1$ and $\cO_2$, are such
+that any two sets $S_1 \in \cO_1$ and $S_2 \in \cO_2$ are disjoint,
+or, w.l.o.g, all the sets in $\cO_1$ are contained within one set in
+$\cO_2$.  This containment relation naturally determines a
+decomposition of the overlap components into rooted containment trees.
+We consider the case when there is only one rooted containment tree,
+and we first present our algorithm when $\bO(\cF)$ is connected.  It
+is easy to see that once the path labeling to the overlap component in
+the root of the containment tree is achieved, the path labeling to the
+other overlap components in the rooted containment tree is essentially
+finding a path labeling when the target tree is a path: each target
+path is a path that is allocated to sets in the root overlap
+component.  Therefore, for the rest of this section, $\bO(\cF)$ is a
+connected graph. Recall that we also consider the special case when
+all hyperedges are of cardinality at most $k+2$.  By definition, a
+$k$-subdivided star has a central vertex which we call the {\em root},
+and each root to leaf path is called a {\em ray}.  First, we observe
+that by removing the root $r$ from $T$, we get a collection of $p$
+vertex disjoint paths of length $k+1$, $p$ being the number of leaves
+in $T$.  We denote the rays by $R_1, \ldots, R_p$, and the number of
+vertices in $R_i$, $i \in [p]$ is $k+2$.  Let
+$\seq{v_{i1},\ldots,v_{i(k+2)}=r}$ denote the sequence of vertices in
+$R_i$, where $v_{i1}$ is the leaf. Note that $r$ is a common vertex to
+all $R_i$.
+
+\subsection{Description of the Algorithm}
+In this section the given hypergraph $\cF$, the $k$-subdivided star
+and the root of the star are denoted by $\cO$, $T$ and vertex $r$,
+respectively.  In particular, note that the vertices of $\cO$
+correspond to the sets in $\cF$, and the edges correspond to the
+overlap relation.
+
 \noindent
-Suppose the given hypergraph has a feasible path labeling to the given
-tree $T$. Let this labeling be $\cl$. Now we present an algorithm that
-lists the sequence of edges that maintain a set of properties
-necessary for them to have a feasible tree path labeling.
+For each hyperedge $X \in \cO$, we will maintain a 2-tuple of non-negative
+ numbers $\seq{p_1(X), p_2(X)}$.  The numbers satisfy the property that
+ $p_1(X) + p_2(X) \leq |X|$, and at the end of path labeling, for each
+ $X$, $p_1(X) + p_2(X) = |X|$.  This signifies the algorithm tracking
+ the lengths of subpaths of the path assigned to $X$ from at most two
+ rays. We also maintain another parameter called the {\em residue} of
+ $X$ denoted by $s(X)=|X| - p_1(X)$. This signifies the residue path
+ length that must be assigned to $X$ which must be from another
+ ray. For instance, if $X$ is labeled a path from only one ray, then
+ $p_1(X) = |X|$, $p_2(X) = 0$ and $s(X) = 0$.
+
+ \xnoindent The algorithm proceeds in iterations, and in the $i$-th
+ iteration, $i > 1$, a single hyperedge $X$ that overlaps with a
+ hyperedge that has been assigned a path is considered.  At the
+ beginning of each iteration hyperedges of $\cO$ are classifed into
+ the following disjoint sets.
+ \begin{enumerate}
+ \item [$\cL_1^i$] {\em Labeled without $r$.} Those that have been
+   labeled with a path which does not contain $r$ in one of the
+   previous iterations.\\  $\cL_1^i = \set{ X \mid p_1(X) = |X| \text{ and
+     } p_2(X) = 0 \text{ and } s(X) = 0, X \in \cO}$
+ \item [$\cL_2^i$] {\em Labeled with $r$.} Those that have been labeled
+   with two subpaths of $\cl(X)$ containing $r$ from two different rays
+   in two previous iterations.\\ $\cL_2^i = \set{X \mid 0 < p_1\left(X\right),
+     p_2\left(X\right) < |X|=p_1(X)+p_2(X) \text{ and } s(X) = 0, X \in \cO}$
+   \item [$\cT_1^i$] {\em Type 1 / partially labeled.} Those that have
+     been labeled with one path containing $r$ from a single ray in one
+     of the previous iterations. Here, $p_1(X)$ denotes the length of
+     the subpath of $\cl(X)$ that $X$ has been so far labeled
+     with. \\
+     $\cT_1^i = \set{ X \mid 0 < p_1(X) < |X| \text{ and } p_2(X) = 0
+       \text{ and } s(X) = |X|-p_1(X), X \in \cO}$
+   \item [$\cT_2^i$] {\em Type 2 / not labeled.} Those that have not been
+     labeled with a path in any previous iteration.\\
+     $\cT_2^i = \set{ X \mid p_1(X) = p_2(X) = 0 \text{ and } s(X) = |X|,
+       X \in \cO}$
+ \end{enumerate}
+% \vspace{-2mm}
+% \begin{align*}
+%   \cO &= \cL_1^i \cup \cL_2^i \cup \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]\\
+%   \cO_i &= \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]
+% \end{align*}
+
+\noindent
+The set $\cO_i$ refers to the set of hyperedges $\cT_1^i \cup \cT_2^i$
+in the $i$th iteration.  Note that $\cO_1 = \cO$.  In the $i$th
+iteration, hyperedges from $\cO_i$ are assigned paths from $T$ using
+the following rules. Also the end of the iteration, $\cL_1^{i+1},
+\cL_2^{i+1}, \cT_1^{i+1}, \cT_2^{i+1}$ are set to $\cL_1^{i},
+\cL_2^{i}, \cT_1^{i}, \cT_2^{i}$ respectively, along with some
+case-specific changes mentioned in the rules below.
+
+\noindent
 \begin{enumerate}[I.]
-\item {\bf Step 1}
+\item {\bf Iteration 1:} Let $S=\{X_1,\ldots,X_s\}$ denote the
+  super-marginal hyperedges from $\cO_1$.  If $|S|=s \neq p$, then
+  exit reporting failure.  Else, assign to each $X_j \in S$, the path
+  from $R_j$ such that the path contains the leaf in $R_j$.  This path
+  is refered to as $\cl(X_j)$.  Set $p_1(X_j)=|X|, p_2(X_j)=s(X_j)=0$.
+  Hyperedges in $S$ are not added to $\cO_2$ but are added to
+  $\cL_1^2$ and all other hyperedges are added to
+  $\cO_2$.
+\item {\bf Iteration $i$:} Let $X$ be a hyperedge from $\cO_i$ such
+  that there exists $Y \in \cL_1^i \cup \cL_2^i$ and $X \overlap
+  Y$. Further let $Z \in \cL_1^i \cup L_2^i$ such that $Z \overlap Y$.
+  If $X \in \cT_2^i$, and if there are multiple $Y$ candidates then
+  any $Y$ is selected.  On the other hand, if $X \in \cT_1^i$, then
+  $X$ has a partial path assignment, $\cl'(X)$ from a previous
+  iteration, say from ray $R_j$. Then, $Y$ is
+  chosen such that $X \cap Y$ has a non-empty intersection with a ray
+  different from $R_j$.  The key things that are done in assigning a
+  path to $X$ are as follows. The {\em end} of path $\cl(Y)$ where
+  $\cl(X)$ would overlap is found, and then based on this the
+  existence of a feasible assignment is decided.  It is important to
+  note that since $X \overlap Y$, $\cl(X) \overlap \cl(Y)$ in any
+  feasible assignment.  Therefore, the notion of the {\em end} at
+  which $\cl(X)$ and $\cl(Y)$ overlap is unambiguous, since for any
+  path, there are two end points.
   \begin{enumerate}
-  \item {\bf {There are type 1 edges, $\mathbf{|\cT_1^i| > 0}$}:}
-    % , and for each such $X$ assign the path in $R_i$ of length
-    % $s(X)$ starting at $v_{i,k+2}=r$.  $X$ is then removed from
-    % $\cO_i$, and $p_2(X)=s(X)$, and $s(X)=0$.If any intersection
-    % cardinality is violated because of this assignment, EXIT
-    % reporting the non-existence of a feasible assignment.
-    {If there is only one $\cT_1^i$ hyperedge, label it to the unique
-      path in $R_i$ of length $s(X)$ starting at $v_{i(k+2)}=r$. If
-      there are more than one $\cT_1^i$ hyperedges, for any pair of
-      hyperedges $S_1, S_2 \in \cT_1^i$, one of the following is
-      true. Let $\cl'(X) \subseteq R_j$ denote the subpath of length
-      $p_1(X)$ that has been assigned to $X$ from a previously
-      considered ray $R_j$.}
-    \begin{enumerate}[{\em Case }1. ]
-    \item $|\cl'(S_1) \cap \cl'(S_2)| = 1$ and $|S_1 \cap S_2| > 1$
-      i.e. they are in different rays and their residues must contain
-      their intersection. Hence assign both to $R_i$.
-    \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| >
-      |\cl'(S_1) \cap \cl'(S_2)|$ i.e. they are in the same rays and
-      their residues must contain their intersection. Hence assign
-      both to $R_i$.
-    \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| =
-      |\cl'(S_1) \cap \cl'(S_2)|$. i.e. the intersection cardinality
-      has been met. Hence assign only $S_1$ to $R_i$ and $S_2$ must be
-      assigned to some $i \ge p$.
+  \item \label{iendpoint} {\em End point of $\cl(Y)$ where $\cl(X)$ overlaps
+      depends on $X \cap Z$:} If $X \cap Z
+    \neq \emptyset$, then $\cl(X)$ has an overlap of $|X \cap Y|$ at that
+    end of $\cl(Y)$ at which $\cl(Y)$ and $\cl(Z)$ overlap.  If $X
+    \cap Z = \emptyset$, then $\cl(X)$ has an overlap of $|X \cap Y|$ at
+    that end of $\cl(Y)$ where $\cl(Y)$ and $\cl(Z)$ do not intersect.
+  \item {\em Any path of length $s(X)$ at the appropriate end contains
+      $r$:} If $X \in \cT_1^i$ then after finding the appropriate end
+    as in step~\ref{iendpoint} this the unique path of length $s(X)$
+    should end at $r$.  If not, we exit reporting failure.  Else,
+    $\cl(X)$ is computed as union of $\cl'(X)$ and this path. If any
+    three-way intersection cardinality is violated with this new
+    assignment, then exit, reporting failure.  Otherwise, $X$ is added
+    to $\cL_2^{i+1}$.  On the other hand, if $X \in \cT_2^i$, then
+    after step~\ref{iendpoint}, $\cl(X)$ or $\cl'(X)$ is unique up to
+    the root and including it. Clearly, the vertices $\cl(X)$ or
+    $\cl'(X)$ contains depends on $|X|$ and $|X \cap Y|$.  If any
+    three way intersection cardinality is violated due to this
+    assignment, exit, reporting failure.  Otherwise,
+    $p_1(X)$ is updated as the length of the assigned path, and $s(X)
+    = |X|-p_1(X)$.  If $s(X) > 0$, then $X$ is added to $\cT_1^{i+1}$.
+    If $s(X)=0$, then $X$ is added to $\cL_1^{i+1}$.
+  \item {\em The unique path of length $s(X)$ overlapping at the
+      appropriate end of $Y$ does not contain $r$:} In this case,
+    $\cl(X)$ is updated to include this path.  If any three way
+    intersection cardinality is violated, exit, reporting failure.
+    Otherwise, update $p_1(X)$ and $p_2(X)$ are appropriate, $X$ is
+    added to $\cL_1^{i+1}$ or $\cL_2^{i+1}$, as appropriate.
   \end{enumerate}
-    Remove the labeled edge or edges from $\cO_i$ and add it to
-    $\cO_{i+1}$. Also do not add it to $\cT_1^{i+1}$ and add to
-    $\cL_2^{i+1}$.
-  \item {\bf {There are no type-1 edges nor} super-marginal type-2
-      hyperedges:} If all type-2 edges of length at most $k+1$, EXIT
-    by saying there is no ICPPL.  Otherwise, Find an $X$ of type-2
-    such that $|X| \geq k+2$, assign it the unique path starting at
-    $v_{i(k+1)}=r$ of length $k+2$, and set $p_1(X)=k+2,
-    s(X)=|X|-(k+2)$, assign it the unique path starting at
-    $v_{i(k+1)}=r$ of length $k+2$, mark this as a type-1 hyperedge,
-    and add it to $\cO_{i+1}$ after removing it from $\cO_i$.
-    % \remove{Next, iteratively, consider each type-1 edge $X \in
-    %   \cO_i$, and if it intersects a previously assigned $Y$ such
-    %   that
-    %   $\cl(Y) \subseteq R_i$, assign a path of length $s(X)$
-    %   containing $v_{i(k+2)}=r$, set $p_2(X)=s(X), s(X)=0$, and
-    %   remove
-    %   $X$ from $\cO_i$.}\tnote{covered in step 2}
-
-  \item {\bf There are super-marginal type-2 hyperedges {and no type-1
-        edges:}} Let $X$ be a super-marginal hyperedge. If $|X| \leq
-    k+2$, then $X$ is assigned the path of length $|X|$ starting at
-    $v_{i1}$, and $p_1(X) = |X|, p_2(X)=0$. $X$ is removed from
-    $\cO_i$ and not added to $\cO_{i+1}$.  If $|X| > k+2$, then
-    $p_1(X)=k+2$, assign it the unique path starting at $v_{i(k+1)}=r$
-    of length $k+2$ and $s(X) = |X| - (k+2)$.  In this case, $X$ is
-    classified as a Type-1 edge, removed from $\cO_i$, and added to
-    $\cO_{i+1}$.
 \end{enumerate}
-\item {\bf Step 2:} Iteratively, a hyperedge $X$ is selected from
-  $\cO_i$ that has an overlap with one of the hyperedges $Y$ such that
-  $\cl(Y) \subseteq R_i$, and a unique path is assigned to $X$.  The
-  path, say $U(X) \subseteq R_i$, that is assigned can be decided
-  unambiguously since the $X \overlap Y$, and all intersection
-  cardinalities can be preserved only at one of the ends of $\cl(Y)$.
+
+\noindent {\bf Proof of Correctness and Analysis of Running Time:} It
+is clear that the algorithm runs in polynomial time, as at each step,
+at most three-way intersection cardinalities need to be checked.
+Further, finding super-marginal hyperedges can also be done in
+polynomial time, as it involves considering the overlap regions and
+checking if the inclusion partial order contains a single minimal
+element.  In particular, once the super-marginal edges are identified,
+each iteration involes finding the next hyperedge to consider, and
+testing for a path to that hyperedge.  To identify the next hyperedge
+to consider, we consider the breadth first layering of the hyperedges
+with the zeroeth layer consisting of the super-marginal hyperedges.
+Since $\cO$ is connected, it follows that all hyperedges of $\cO$ will
+be considered by the algorithm.  Once a hyperedge is considered, the
+path to be assigned to it can also be computed in constant time.  In
+particular, in the algorithm the path to be assigned to $X$ depends on
+$\cl(Y), \cl(Z)$, $s(X)$ and the presence or absence of $r$ in the
+candidate partial path $\cl'(X)$.  Therefore, once the super-marginal
+edges are identified, the running time of the algorithm is linear in
+the size of the input.  By the technique used for constructing prime
+matrices \cite{wlh02}, the super-marginal edges can be found in linear
+time in the input size.  Therefore, the algorithm can be implemented
+to run in linear time in the input size.
 
 \noindent
-  Let $\cl(X)$ denote the unique path assigned to $X$.  If $X$ is a
-  type-2 hyperedge and if the unique path of length $|X|$ does not
-  contain $r$, then $p_1(X) = |X|, p_2(X)=0$, and $X$ is removed from
-  $\cO_i$.  In the case when $\cl(X)$ has to contain $r$, then
-  $p_1(X)$ is the length of the path, $p_2(X)=0$, and $s(X) =
-  |X|-p_1(X)$. Further $X$ is classified as a type-1 hyperedge, added
-  to $\cO_{i+1}$ and removed from $\cO_i$.  In the case when $X$ is a
-  type-1 hyperedge, then we check if $U(X)$, which is of length $s(X)$
-  contains $r$. If it does, then we assign $\cl(X) \leftarrow \cl(X)
-  \cup U(X)$, remove $X$ from $\cO_i$ and do not add it to
-  $\cO_{i+1}$.  If not, then we {\bf Exit} reporting that an
-  assignment cannot be found.  The iteration ends when no hyperedge in
-  $\cO_i$ has an overlap with a hyperedge assigned to $R_i$.
+The proof of correctness uses the following main properties:
+\begin{enumerate}
+\item The $k$-subdivided star has a very symmetric structure.  This
+  symmetry is quantified based on the following observation -- either
+  there are no feasible path labelings of $\cO$ using paths from $T$,
+  or there are exactly $p!$ feasible path labelings.  In other words,
+  there is either no feasible assignment, or effectively a unique
+  assignment modulo symmetry.
+\item The $p$ super-marginal hyperedges, if they exist, will each be
+  assigned a path from distinct rays, and each such path contains the
+  leaf.
+\item For a candidate hyperedge $X$, the partial path assignment
+  $\cl'(X)$ is decided by its overlap with $\cl(Y)$ and cardinality of
+  intersection with $\cl(Z)$.
 \end{enumerate}
-
-\noindent
-The order in which they are assigned is the sequence of hyperedges
-with the following properties. Note that if such a sequence cannot be
-computed, $\cO$ cannot have a feasible path labeling from $T$.
-
-\noindent
-\tnote{need to formalize the below properties} In the following lemmas
-we identify a set of necessary conditions for $\cF$ to have an ICPPL
-in the $k$-subdivided star $T$.  If during the execution of the above
-algorithm, one of these necessary conditions is not satisfied, the
-algorithm exits reporting the non-existence of an ICPPL.
-\begin{lemma}
-  Let all hyperedges in $\cO_i$ be type-1 edges.  Then there is a
-  maximal subset $\ccT_i \subseteq \cO_i$ with the following
-  properties:
-  \begin{enumerate}
-  \item $\ccT_i$ form an inclusion chain.
-  \item For all $X \in \ccT_i$, $s(X) \leq k+2$, and There is a an $X
-    \in \ccT_i$ such that $s(X)=k+2$.
-  \end{enumerate}
-\end{lemma}
-\begin{lemma}
-  If there are no \marginal type-2 edges in $\cO_i$, then there exists
-  at least $p-i$ type-2 hyperedges $X \in \cO_i$ such that $|X| \geq
-  k+2$.
-  % is a exist type-2 hyperedges of length at least $k+2$, and there
-  % must be at least $r-i$ such type-2 hyperedge, basically at least
-  % one per ray.  Each hyperedge contained in such an hyperedges of
-  % length at least $k+2$, is transitively related under the overlap
-  % relation to one that gets assigned a path that contains
-  % $r$. Basically, the proof will argue that otherwise, there will be
-  % a \marginal type-2 hyperedge, or that it is infeasible.  Also, it
-  % is here we assume that we assume that the size of the two
-  % universes is the same.
-\end{lemma}
-\begin{lemma}
-  At the end of Step-1 in the $i$-th iteration, if one hyperedge $X$
-  of type-2 is such that $\cl(X) \subseteq R_i$, then all other
-  hyperedges in $\cO_i$ are connected to $X$ in the overlap component.
-\end{lemma}
-\begin{lemma}
-  At the end of Step-2, If control has exit at any time, there is no
-  ICPPL. If control has not exit, then $R_i$ is saturated.  No
-  hyperedge of $\cO_{i+1}$ will get a path from $R_i$ in the future
-  iterations.  No type-2 hyperedge of $\cO_{i+1}$ will get a path from
-  $R_i$.  Basically $R_i$ is done.
-\end{lemma}
-\begin{lemma}
-  Finally, we need to prove that the assignment is an ICPPL. Secondly,
-  if there is a permutation then maps sets to paths, then it is indeed
-  an ICPPL, and our algorithm will basically find it.  It is a unique
-  assignment upto permutation of the leaves.
-\end{lemma}
-
+These properties are formalized as follows:
 \begin{lemma}
   \label{lem:sup-mar}
-  If $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible tree
-  path labeling to tree $T$, then $\cl(X)$ will
-  contain a leaf in $T$.\\
-  On the otherhand, if $\cl(X)$ has a leaf in $T$, $X$ is marginal but
-  may not be super-marginal.
+  If $X \in \cF$ is super-marginal and $\cl$ is a feasible tree path
+  labeling to tree $T$, then $\cl(X)$ will contain a leaf in $T$.
 \end{lemma}
@@
   inclusion chain which contradicts that assumption $X$ is
-  super-marginal. Thus the claim is proven.%\qed
+  super-marginal. Thus the claim is proved.\qed
 \end{proof}
+\begin{lemma}
+  If $\cO$ does not have any super-marginal edges, then in any
+  feasible path labeling $\cl$ of $\cO$ with paths from $T$ is such
+  that, for any hyperedge $X$ for which $\cl(X)$ contains a leaf, $|X|
+  \geq k+3$.
+\end{lemma}
+\begin{proof}
+  The proof of this lemma is by contradiction.  Let $X$ be a
+  hyperedges such that $|X| \leq k+2$ and that $\cl(X)$ has a leaf.
+  This implies that the overlap regions with $X$, which are captured by
+  the overlap regions with $\cl(X)$, will form a single inclusion
+  chain. This shows that $X$ is a marginal hyperedge which
+  contradicts the assumption that $\cO$ does not have super-marginal
+  hyperedges. \qed
+\end{proof}
+This lemma is used to prove the next lemma for the case when for all
+$X \in \cO$, $|X| \leq k+2$.  The proof is left out as it just uses
+the previous lemma and the fact that the hyperedges in $X$ have at
+most $k+2$ elements.
+\begin{lemma}
+  If there is a feasible path labeling for $\cO$ in $T$, then there
+  are exactly $p$ super-marginal hyperedges.
+\end{lemma}
+These lemmas now are used to prove the following theorem.
+\begin{theorem}
+  Given $\cO$ and a $k$-subdivided star $T$, the above algorithm
+  decides correctly if there is a feasible path labeling $\cl$.
+\end{theorem}
+\begin{proof} {\em Outline.}
+%   If the algorithm outputs a path labeling $\cl$, then it is clear
+%   that it is an ICPPL. The reason is that the algorithm checks that
+%   three-way intersection cardinalities are preserved in each iteration
+%   which ensures \icpplpr~\ref{pr:iii}. Moreover, it is clear that
+%   $\cl(X)$ for any $X \in \cO$ is computed by maintaining
+%   \icpplpr~\ref{pr:i} and \icpplpr~\ref{pr:ii}. For such a labeling
+%   $\cl$, the proof that it is feasible is by induction on $k$. What
+%   needs to be shown is that Algorithm~\ref{al:icppl-find-isomorph}
+%   successfully runs on input $(\cO, \cl)$. In base case $k=0$, $T$ is
+%   a star. The claim is clear by observing that after Filter 1 and one
+%   iteration of Filter 2, all the leaves have found their pre-images
+%   from support $\cO$.  Therefore, in the induction step, it is clear
+%   that after Filter 1 and one iteration of Filter 2, the leaves are
+%   assigned pre-images.  Removing the leaves from $T$ and the
+%   pre-images from support of $\cO$, results in an ICPPL to a
+%   $(k-1)$-subdivided star.  Now we apply the induction hypothesis, and
+%   we get a isomorphism between the hypergraph $\cO$ and $\cO^\cl$.
+  If the algorithm outputs a path labeling $\cl$, then it is clear
+  that it is an ICPPL. The reason is that the algorithm checks that
+  three-way intersection cardinalities are preserved in each iteration
+  which ensures ICPPL \icpplpr~\ref{pr:iii}. Moreover, it is clear
+  that $\cl(X)$ for any $X \in \cO$ is computed by maintaining ICPPL
+  \icpplpr~\ref{pr:i} and ICPPL \icpplpr~\ref{pr:ii}. For such a
+  labeling $\cl$, the proof that it is feasible is by induction on
+  $k$. What needs to be shown is that
+  Algorithm~\ref{al:icppl-find-isomorph} successfully runs on input
+  $(\cO, \cl)$. In base case $k=0$, $T$ is a star. Also every set is
+  at most size 2 ($k+2$) size and thus overlaps are at most 1. If two
+  paths share a leaf in {\tt filter\_1} one must be of length 2 and
+  the other of length 1. Thus the exit condition is not met. Further,
+  it is also clear that the exit condition in {\tt filter\_2} is also
+  not met. Thus claim proven for base case.  Now assume the claim to
+  be true when target tree is a $(k-1)$-subdivided star. Consider the
+  case of a $k$-subdivided star.  We can show that after {\tt
+    filter\_1} and one iteration of a modified {\tt filter\_2} {\em
+    all} leaves are assigned pre-images.  Removing the leaves from $T$
+  and the pre-images from support of $\cO$, results in an ICPPL to a
+  $(k-1)$-subdivided star.  Now we apply the induction hypothesis, and
+  we get an isomorphism between the hypergraphs $\cO$ and $\cO^\cl$.
 
   \noindent
-Consider the overlap graph $\bO(\cF)$ of the given hypergraph
-$\cF$. Let $S_{sm} \in \cF$ be such that it is a super-marginal
-hyperedge.  Algorithm~\ref{al:ktree-label} uses $S_{sm}$ along with
-the overlap graph $\bO(\cF)$ to calculate the feasible tree path
-labeling to the $k$-subdivded tree $T$.
-
-\begin{algorithm}[h]
-  \caption{{\tt compute-ksubtree-path-labeling($X, \cF, T$)}}
-  \label{al:ktree-label}
-  \begin{algorithmic}[\lndisplay]
-    \IF{$X = S_{sm}$}
-    \STATE -- TBD --\\
-    \ELSE
-    \STATE -- TBD --\\
-    \ENDIF
-
-  \end{algorithmic}
-\end{algorithm}
+  In the reverse direction if there is a feasible path labeling $\cl$,
+  then we know that $\cl$ is unique up to isomorphism.  Therefore,
+  again by induction on $k$ it follows that the algorithm finds $\cl$.
+  \qed
+\end{proof}
+
+% \noindent
+% Consider the overlap graph $\bO(\cF)$ of the given hypergraph
+% $\cF$. Let $S_{sm} \in \cF$ be such that it is a super-marginal
+% hyperedge.  Algorithm~\ref{al:ktree-label} uses $S_{sm}$ along with
+% the overlap graph $\bO(\cF)$ to calculate the feasible tree path
+% labeling to the $k$-subdivded tree $T$.
+%
+% \begin{algorithm}[h]
+%   \caption{{\tt compute-ksubtree-path-labeling($X, \cF, T$)}}
+%   \label{al:ktree-label}
+%   \begin{algorithmic}[\lndisplay]
+%     \IF{$X = S_{sm}$}
+%     \STATE -- TBD --\\
+%     \ELSE
+%     \STATE -- TBD --\\
+%     \ENDIF
+%
+%   \end{algorithmic}
+% \end{algorithm}
+
+% \subsection{temp section from nsnsr09}
+
+% \begin{algorithm}
+%   \caption{Basic step in an algorithm to find an ICPIA for a prime
+%     matrix $M'$}
+%   \label{ds-algo}
+%   ICPIA(Set $S$, Integer $p > 0$)
+
+%   \noindent
+%   /* {\tt $S \cap S^i \not = \phi$ for some $i \in \{1,\ldots,p\}$,
+%     but $S \not\subseteq S^i$, $S^i \not\subseteq S$.  \\ Assigns to
+%     $S$ an interval $I$ such that $\{I^1,\ldots,I^p,I\}$ forms an
+%     ICPIA for $\{S^1,\ldots,S^p,S\}$.} */
+%   \begin{algorithmic}
+%     \STATE Let $|S \cap S^i| = z$. \\
+%     \STATE Let $I_l$ be the interval such that $|I_l \cap I^i| = z$, $|I_l| = |S|$ and the $z$ common elements are the smallest elements of $I^i$. \\
+%     \STATE Let $I_r$ be the interval such that $|I_r \cap I^i| = z$, $|I_r| = |S|$, and the $z$ common elements are the largest elements of $I^i$. \\
+%     \IF {$p == 1$}
+%     \STATE Assign $I_l$ to $S$ \\
+%     \STATE /* {\tt In this case, $I_r$ could also be assigned to $S$.  This will yield the {\bf other} ICPIA} */\\
+%     \ELSE \IF{$|I_l \cap I^q| = |S \cap S^q|$ for each $q \in
+%       \{1,\ldots,p\}$}
+%     \STATE Assign $I_l$ to $S$ and exit.\\
+%     \ENDIF \IF{$|I_r \cap I^q| = |S \cap S^q|$ for each $q \in
+%       \{1,\ldots,p\}$}
+%     \STATE Assign $I_r$ to $S$ and exit.\\
+%     \ENDIF \ENDIF \STATE Report no ICPIA and exit.
+%   \end{algorithmic}
+% \end{algorithm}
+
+% \begin{theorem}
+%   Algorithm~\ref{ds-algo} outputs an ICPIA to a prime matrix $M'$
+%   iff there is an ICPIA for $M'$.
+% \end{theorem}
+% \begin{proof}
+%   The only-if part of the theorem is straightforward.  We now show
+%   that if there is an ICPIA for $M'$, then Algorithm~\ref{ds-algo}
+%   will indeed discover it.  The key fact is that in $M'$ for each
+%   set $S$, there is another set $T \in M'$ such that $S \cap T \not
+%   = \phi$, and $S$ and $T$ are not contained in each other.  Due to
+%   this fact, there are exactly two ICPIAs for $M'$.  The two
+%   distinct ICPIAs differ based on the interval assigned to $S_1$,
+%   see Algorithm ~\ref{ds-algo}.  If $I_l$ is assigned to $S_1$, then
+%   we get one, and the other ICPIA is obtained by assigning $I_r$ to
+%   $S_1$.  For each subsequent set, say $S^j$, the interval to be
+%   assigned is forced.  It is forced due to the fact that the
+%   interval assigned to $S^j$ is based on the interval assigned to
+%   $S^i$, where $S^i \cap S^j \not = \phi$, and $S^i \not\subseteq
+%   S^j$, and $S^j \not\subseteq S^i$.  Given the fact that the
+%   algorithm is an exact implementation of these observations, it
+%   follows that Algorithm~\ref{ds-algo} finds an ICPIA if there is
+%   one.
+% \end{proof}
 

### 2012-01-14 src/0031ded.tex
Started cleaning up chapter 3. Deferring it though to work on Chapter 2. Need to accommodate N's new proofs for removal of algorithmic charac.
See diff for other changes left out in this comment.

\definecolor{dim}{rgb}{0.7,0.7,0.7}
\def \dimspace {0.125in}

{%
  \fontencoding{T1}%
  \fontfamily{bch} %{put} %{bch} % {pbk} % \fontfamily{cmdh}
  % \fontfamily{pag}
  \selectfont%

  \vglue -0.5in % this makes top margin 2in
%   \vskip -5ex%
%   \hspace{5in}\parbox{1in}{\raggedright Hello world}

  \hspace{-1in}%
  \parbox[c]{4in}{%
    \large \raggedright \noindent %
    \color{dim}{%
      The process of preparing programs for the digital computer is
      especially attractive, not only because it can be economically
      and scientifically rewarding, but also because it can be an
      aesthetic experience much like composing poetry or music.%
      {%
        \small \par \hfill \emph{The Art of Computer Programming},
        D.\,E.\,K.%
      }%
    }%
  }%

  \vspace{\dimspace}

  \hspace{1.5in}%
  \parbox[t]{4in}{%
    \Large \noindent \raggedleft%
    \color{dim}{%
      The rhetoric and mythos of science create the comforting image
      of linear progression toward truth.%
      {%
        \small \par \hfill%
        \textbf{mythos}; \emph{New Oxford American Dictionary, 2nd
          Ed.}%
      }%
    }%
  }
  \\
  \vspace{\dimspace}

  % \hspace{1in}
  {\centering
    \parbox{6in}{%
%      \Huge 
      \fontsize{27}{1.5}\selectfont
      \centering \noindent %
%      \textbf{
      \emph{L'art c'est la solution au chaos.} \par \noindent
      \noindent (Art is the solution to chaos.)
%      }
%       {%
%         \small \par \hfill%
%         \emph{citation needed.}%
%       }%
    }%
  }%

  \vspace{\dimspace}

  \hspace{-0.25in}%
  \parbox{2.5in}{%
    \Large 
    \raggedright \noindent%
    \color{dim}{%
      Kilroy wasn't Haar.  Free the group.  Nuke the kernel.  Power to
      the $n$.  $N = 1 \Rightarrow P = NP$ %
    }%
    {%
      \small \par \hfill%
      \emph{Concrete Mathematics}%
    }%
  }%
  \vspace{\dimspace}

  \hspace{1in}%
  \parbox{5in}{%
    \large \raggedleft \noindent%
    \color{dim}{%
      Anything that happens, happens.  Anything that, in happening,
      causes something else to happen, causes something else to
      happen.  Anything that, in happening, causes itself to happen
      again, happens again.  It doesn't necessarily do it in
      chronological order, though.  }%
    {%
      \small \par \hfill%
      \emph{The Salmon of Doubt}, D.\,N.\,A.%
    }%
  }%
  \vspace{\dimspace}

  \hspace{-0.50in}%
  \parbox{4in}{%
    \huge \noindent \raggedright%
    \color{dim}{%
      This is water, this is water.%
      {%
        \small \par \hfill%
        D.\,F.\,W.%
      }%
    }%
  }%

}

### 2012-01-14 src/020chap2.tex
Started cleaning up chapter 3. Deferring it though to work on Chapter 2. Need to accommodate N's new proofs for removal of algorithmic charac.
See diff for other changes left out in this comment.

@@
 
+    \tnote[important]{1899 by petrie -- cite
+         kendall 69 pacific journal of mathematics 1969 }
+
 \def\tabparboxw{10cm}
@@
 
-   \begin{tabular}[h]{lll}
+  \begin{tabular}{lll}
      \hline
@@
      1899 & \parbox[t]{\tabparboxw}{ First mention of \COP
-       (archaeology)} & \parbox[t]{2cm}{\temptext{by petrie -- cite
-         kendall 69 pacific journal of mathematics 1969 }}  \\
+       (archaeology)} & 
+       \\ % \parbox[t]{2cm}{}
 
@@
 
-A few years later, a deeply significant result based on very
-different ideas in understanding \COP came from Tucker which gave a
+A few years later, a deeply significant result based on very different
+ideas in understanding \COP came from Tucker which gave a
 combinatorial (negative) characterization of matrices with COP
 \cite{at72}. This result influenced most of the \COP results that
-followed in the literature \annote[pressing]{including linear time
-  algorithms}{did it? which ones?} for \COP recognition.
+followed in the literature including linear time
+algorithms\tnote[pressing]{linear? did it? which ones?} for \COP
+recognition.
 
@@
 
-\begin{definition}[\emph{\PQtree \cite{bl76, mcc04}}]\\
+\begin{definition}[\PQtree \cite{bl76, mcc04}]
   A \PQtree of matrix $M$ with \COP on columns (rows), is a tree with the following properties.
@@
 disjoint or (ii) one is the subset of the other. Such sets are also
-called {\em orthogonal sets} or said that $A$ is orthogonal to $B$. {\em Trivial subsets}
-of a universe $U$, denoted by $\cT(U)$, are sets that have trivial
-intersections with any set in $\power{U}$. Namely, $\cT(U)$ consists
-of $U$, singleton sets in $\power{U}$ and $\emptyset$\cite{n89, mm96}.
-In other words, the all sets presented in a \PQRtree share the
-property that they do not overlap with any set in $\cF$. We will now
-see how this helps in testing for \COP in $\cF$.
+called {\em orthogonal sets} or said that $A$ is orthogonal to
+$B$. {\em Trivial subsets} of a universe $U$, denoted by $\cT(U)$, are
+sets that have trivial intersections with any set in
+$\power{U}$. Namely, $\cT(U)$ consists of $U$, singleton sets in
+$\power{U}$ and $\emptyset$\cite{n89, mm96}.  In other words, the all
+sets presented in a \PQRtree share the property that they do not
+overlap with any set in $\cF$. We will now see how this helps in
+testing for \COP in $\cF$.
 
@@
 Any set system that is closed under the above operations and contains
-$\cT(U)$ is called {\complete} \cite{mm96} or a {\em weakly partative
+$\cT(U)$ is called {\em complete} \cite{mm96} or a {\em weakly partative
   family} \cite{mcc04}. Such a set system is easier to

### 2012-01-14 src/030chap3.tex
Started cleaning up chapter 3. Deferring it though to work on Chapter 2. Need to accommodate N's new proofs for removal of algorithmic charac.
See diff for other changes left out in this comment.

@@
 
-\tnote{UPDATE TO FINAL VERSION. BELOW IS OLDER VERSION WITH INCOMPLETE
-  SEC 4.}  This chapter documents all the new results obtained by us
+This chapter documents all the new results obtained by us
 in the area of tree path labeling of path hypergraphs which is the
@@
 
-\section{Introduction}
+\section{Summary of Proposed Problems}
 \label{sec:myresearchintro}
@@
 
-To characterize a feasible TPL, we consider the case were a tree path
+To characterize a feasible TPL, we consider the case where a tree path
 labeling is also given as input and we are required to test if the
@@
 problem, we were unable to discover an efficient algorithm for
-it. Hence we consider two special cases of \CFTPL\ -- namely, \CFTPLINT
-and \CFTPLKTREE on special target trees, intervals and \kstars,
+it. Hence we consider two special cases of the same -- namely, \CFTPLINT
+and \CFTPLKTREE on special target trees, namely, intervals and \kstars,
 respectively.  Section~\ref{sec:spltargettree} discusses these
@@
 polynomial time is presented in
-Section~\ref{sec:norestraint}. \tnote{put some more description
+Section~\ref{sec:norestraint}. \tnote[minor]{put some more description
   here. refer WG11.}
 
-\remove{We focus on the question of generalizing the notion of an
+\remove[minor]{We focus on the question of generalizing the notion of an
   ICPIA cite-nsnrs09 to characterize feasible path assignments.  We
@@
 
-{\bf Hypergraph Preliminaries:} The set $\F \subseteq (2^{U} \setminus
-\emptyset)$ is a {\em set system} of a universe $U$ with $|U| = n$.
-The {\em support} of a set system $\F$ denoted by $supp(\cF)$ is the
-union of all the sets in $\F$; $supp(\F) = \bigcup_{S \in \F}S$. For
-the purposes of this paper, a set system is required to ``cover'' the
-universe; $ supp(\cF) = U$.  A set system $\cF$ can also be visualized
-as a {\em hypergraph}\, vertex set is $supp(\cF)$ and hyperedges are
-the sets in $\cF$.  The {\em intersection graph}\, $\bI(\cF)$ of a
-hypergraph $\cF$ is a graph such that its vertex set has a bijection
-to $\cF$ and there exists an edge between two vertices iff their
+\subsection{Hypergraph Preliminaries}
+
+\begin{definition}[Set systems and hypergraphs] 
+  \noindent 
+  The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
+    system} of a universe $U$ with $|U| = n$.  The {\em support} of a
+  set system $\F$ denoted by $\supt{\cF}$\tnote[minor]{change all supp
+    references to supt command} is the union of all the sets in $\F$;
+  $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this paper, a
+  set system is required to ``cover'' the universe; $ supp(\cF) = U$.
+
+  \noindent
+  A set system $\cF$ can also be visualized as a {\em hypergraph}
+  whose vertex set is $supp(\cF)$ and hyperedges are the sets in
+  $\cF$. This is a known representation for
+  interval systems in literature \cite{bls99,kklv10}.  We extend this
+  definition here to path systems. Due to the equivalence of set
+  system and hypergraph in the scope of this paper, we drop the
+  subscript $_H$ in the notation and refer to both the structures by
+  $\cF$.
+
+  \noindent
+  The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph
+  $\cF$ is a graph such that its vertex set has a bijection to $\cF$
+  and there exists an edge between two vertices \iff their
   corresponding hyperedges have a non-empty intersection \cite{mcg04}.
 
+  \noindent 
   Two hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to
-each other, denoted by $\cF' \cong \cF''$, iff there exists a
-bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for all
-sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ iff $B$ is
-a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
-\cite{kklv10}, written as $B=\phi(A)$.  This is called {\em hypergraph
-  isomorphism}.
-
-\add{A set system $\cF$ can be alternatively} represented by a {\em
-  hypergraph}\, $\cF_H$ whose vertex set is $supp(\cF)$ and hyperedges
-are the sets in $\cF$. This is a known representation for interval
-systems in literature \cite{bls99,kklv10}.  We extend this definition
-here to path systems. Due to the equivalence of set system and
-hypergraph in the scope of this paper, we drop the subscript $_H$ in
-the notation and refer to both the structures by $\cF$.
-
-{\bf Path Hypergraph from a Tree:} The graph $T$ represents a {\em
-  target tree} with same number of vertices as elements in $U$;
-$|V(T)|=|U|=n$.  A {\em path system}\, $\cP$ is a set system of paths
-from $T$; $\cP \subseteq \{P \mid P \subseteq V, \text{ } T[P] \text{
-  is a path} \}$.  This generalizes the fact, from the literature
-\cite{bls99,kklv10}, that intervals can be viewed as sub-paths of a
-path.
-
-\add{Due to the equivalence of set system and hypergraph in the scope
-  of this paper, we drop the subscript $_H$ in the notation and refer
-  to both the structures by $\cF$.}
-
+  each other, denoted by $\cF' \cong \cF''$, \iff there exists a
+  bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for
+  all sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ \iff
+  $B$ is a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
+  \cite{kklv10}, written as $B=\phi(A)$.  This is called {\em
+    hypergraph isomorphism}.  
+\end{definition}
+
+\begin{definition}[Path Hypergraph from a Tree]
+  The graph $T$ represents a {\em target tree} with same number of
+  vertices as elements in $U$; $|V(T)|=|U|=n$.  A {\em path system}\,
+  $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P \mid P
+  \subseteq V, \text{ } T[P] \text{ is a path} \}$.  This generalizes
+  the fact, from the literature \cite{bls99,kklv10}, that intervals
+  can be viewed as sub-paths of a path.
+
+  \noindent
   If the intersection graphs of $\cF$ and $\cP$ (a path system) are
@@
   $\cl: \cF \rightarrow \cP$ due to this isomorphism is called a {\em
-  path labeling} of the hypergraph $\cF$.  Note that there are two
-kinds of isomorphisms here.  One is the isomorphism of intersection
-graphs on $\cF$ and $\cP$, i.e. $\bI(\cF)$ and $\bI(\cP)$
-respectively. Second is the isomorphism between the hypergraphs $\cF$
-and $\cP$.
-
+    path labeling} of the hypergraph $\cF$.\endnote{Note that there
+    are two kinds of isomorphisms here.  One is the isomorphism of
+    intersection graphs on $\cF$ and $\cP$, i.e. $\bI(\cF)$ and
+    $\bI(\cP)$ respectively. Second is the isomorphism between the
+    hypergraphs $\cF$ and $\cP$.}
+
+% The subset of the path labeling that
+%   maps elements of the universe to paths that contain only a leaf the
+%   target tree is  known as a {\em leaf labeling}.
+%   \ie if the target tree has the set of $k$ leaves $L = \{v_i, i \in
+%   [k]\}$, then 
+% $\cl': \rightarrow U' \rightarrow L$ is a leaf labeling where $U'
+% \subseteq U$, $|U'| = k$.
+
+%   $\cL_0 = \{ (x_i,v_i) \mid x_i \in
+%   U, v_i \in T, x_i \ne x_j, v_i \ne v_j, i \ne j, i,j \in [k] \}$
+%   where $k$ is the number of leaves in $T$
+
+  \noindent
   \add{To illustrate further, let $\cg: V(\cF) \rightarrow V(\cP)$ be
-  the above mentioned isomorphism where $V(\cF)$ and $V(\cP)$ are the
-  vertex sets that represent the hyperedges for each hypergraph
+    the above mentioned isomorphism where $V(\cF)$ and $V(\cP)$ are
+    the vertex sets that represent the hyperedges for each hypergraph
     respectively, $V(\cF) = \{ v_S \mid S \in \cF\}$ and $V(\cP) = \{
@@
     follows: $\cl(S_1) = P_1$ iff $\cg (v_{S_1}) = v_{P_1}$.  Just to
-  emphasize, for a path labeling $\cl$ of $\cF$ with $\cP$ as the path
-  system, $\cF^\cl$ is same as $\cP$.  The path system $\cP$ may be
-  alternatively denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In
-  most scenarios in this paper, what is given are the pair $(\cF,
-  \cl)$ and the target tree $T$; hence this notation will be used more
-  often.}
-
-If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$ is called
-a {\em path hypergraph} and $\cP$ is called {\em path representation}
-of $\cF$. If this isomorphism is $\phi: supp(\cF) \rightarrow V(T)$,
-then it is clear that there is an {\em induced path labeling}
-$\cl_\phi: \cF \rightarrow \cP$ to the set system; $\cl_\phi(S) =
-\set{y \mid y = \phi(x), x \in S}$ for all $S \in \cF$. Recall that
-$supp(\cP) = V(T)$.
-
+    emphasize, for a path labeling $\cl$ of $\cF$ with $\cP$ as the
+    path system, $\cF^\cl$ is same as $\cP$.  The path system $\cP$
+    may be alternatively denoted in terms of $\cF$ and $\cl$ as
+    $\cF^\cl$. In most scenarios in this paper, what is given are the
+    pair $(\cF, \cl)$ and the target tree $T$; hence this notation
+    will be used more often.}
+
+  \noindent
+  If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$ is
+  called a {\em path hypergraph} and $\cP$ is called {\em path
+    representation} of $\cF$. If this isomorphism is $\phi: supp(\cF)
+  \rightarrow V(T)$, then it is clear that there is an {\em induced
+    path labeling} $\cl_\phi: \cF \rightarrow \cP$ to the set system;
+  $\cl_\phi(S) = \set{y \mid y = \phi(x), x \in S}$ for all $S \in
+  \cF$. Recall that $supp(\cP) = V(T)$.
+
+  \noindent
   \add{A graph} $G$ is a {\em path graph} if it is isomorphic to the
   intersection graph $\bI(\cP)$ of a path system $\cP$.  This
-isomorphism gives a bijection $\cl': V(G) \rightarrow \cP$. Moreover,
-for the purposes of this paper, we require that in a path labeling,
-$supp(\cP) = V(T)$.  If graph $G$ is also isomorphic to $\bI(\cF)$ for
-some hypergraph $\cF$, then clearly there is a bijection $\cl: \cF
-\rightarrow \cP$ such that $\cl(S) = \cl'(v_S)$ where $v_S$ is the
-vertex corresponding to set $S$ in $\bI(\cF)$ for any $S \in
-\cF$. This bijection $\cl$ is called the {\em path labeling} of the
-hypergraph $\cF$ and the path system $\cP$ may be alternatively
-denoted as $\cF^\cl$.
-
+  isomorphism gives a bijection $\cl': V(G) \rightarrow
+  \cP$. Moreover, for the purposes of this paper, we require that in a
+  path labeling, $supp(\cP) = V(T)$.  If graph $G$ is also isomorphic
+  to $\bI(\cF)$ for some hypergraph $\cF$, then clearly there is a
+  bijection $\cl: \cF \rightarrow \cP$ such that $\cl(S) = \cl'(v_S)$
+  where $v_S$ is the vertex corresponding to set $S$ in $\bI(\cF)$ for
+  any $S \in \cF$. This bijection $\cl$ is called the {\em path
+    labeling} of the hypergraph $\cF$ and the path system $\cP$ may be
+  alternatively denoted as $\cF^\cl$.
+
+  \noindent
   A path labeling $(\cF, \cl)$ is defined to be {\em feasible} if $\cF
@@
   \rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
-\rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.  \add{In this work},
-we are given as input $\cF$ and a tree $T$, and the question is
-whether there is a path labeling $\cl$ to a set of paths in $T$.  We
-refer to such a solution path system by $\cF^\cl$.  A path labeling
-$(\cF, \cl)$ is defined to be {\em feasible} if
+  \rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.  \add{In this
+    work}, we are given as input $\cF$ and a tree $T$, and the
+  question is whether there is a path labeling $\cl$ to a set of paths
+  in $T$.  We refer to such a solution path system by $\cF^\cl$.  A
+  path labeling $(\cF, \cl)$ is defined to be {\em feasible} if
   % $\cF \cong \cF^\cl$ and this
   there is a hypergraph isomorphism $\phi: supp(\cF) \rightarrow
-supp(\cF^\cl)=V(T)$ induces a path labeling $\cl_\phi: \cF \rightarrow
-\cF^\cl$ such that $\cl_\phi = \cl$.
-
-{\bf Overlap Graphs and Marginal Hyperedges:} An {\em overlap graph}\,
-$\bO(\cF)$ of a hypergraph $\cF$ is a graph such that its vertex set
-has a bijection to $\cF$ and there exists an edge between two of its
-vertices iff their corresponding hyperedges overlap. Two hyperedges
-$S$ and $S'$ are said to {\em overlap}, denoted by $S \overlap S'$, if
-they have a non-empty intersection and neither is contained in the
-other; $S \overlap S' \text{ iff } S \cap S' \ne \emptyset, S
-\nsubseteq S', S' \nsubseteq S$. Thus $\bO(\cF)$ is a spanning
-subgraph of $\bI(\cF)$ and not necessarily connected. Each connected
-component of $\bO(\cF)$ is called an {\em overlap component}.
-
+  supp(\cF^\cl)=V(T)$ induces a path labeling $\cl_\phi: \cF
+  \rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.
+\end{definition}
+
+\begin{definition}[Overlap Graphs and Marginal Hyperedges]
+  An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is a graph
+  such that its vertex set has a bijection to $\cF$ and there exists
+  an edge between two of its vertices \iff their corresponding
+  hyperedges overlap. Two hyperedges $S$ and $S'$ are said to {\em
+    overlap}, denoted by $S \overlap S'$, if they have a non-empty
+  intersection and neither is contained in the other; $S \overlap S'
+  \text{ \iff } S \cap S' \ne \emptyset, S \nsubseteq S', S'
+  \nsubseteq S$. Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$
+  and not necessarily connected. Each connected component of
+  $\bO(\cF)$ is called an {\em overlap component}.
+
+  \noindent
   A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
   \overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
-\cite{kklv10}. Additionally, if $S$ is such that it is contained in no
-other marginal hyperedge in $\cF$, then it is called {\em
+  \cite{kklv10}. Additionally, if $S$ is such that it is contained in
+  no other marginal hyperedge in $\cF$, then it is called {\em
     super-marginal}.
-
-{\bf $k$-subdivided star -- a special tree} A {\em star} graph is a
-complete bipartite graph $K_{1,p}$ which is clearly a tree and $p$ is
-the number of leaves. The vertex with maximum degree is called the
-{\em center} of the star and the edges are called {\em rays} of the
-star. A {\em $k$-subdivided star} is a star with all its rays
+\end{definition}
+
+\begin{definition}[$k$-subdivided star] 
+  A {\em star} graph is a
+  complete bipartite graph $K_{1,p}$ which is clearly a tree and $p$
+  is the number of leaves. The vertex with maximum degree is called
+  the {\em center} of the star and the edges are called {\em rays} of
+  the star. A {\em $k$-subdivided star} is a star with all its rays
   subdivided exactly $k$ times. The definition of a {\em ray of a
@@
   leaf. It is clear that all rays are of length $k+2$.
+\end{definition}
 
@@
 
-A {\em path assignment} $\A$ to $\F$ is defined as a set assignment
-where second universe is the vertex set $V$ of a given tree $T$ and
-every second subset in the ordered pairs is a path in this
-tree. Formally, the definition is as follows.
-\begin{align*}
-  \A = \{ (S_i,P_i) \mid S_i \in \F, P_i \subseteq V \text{
-    s.t. }T[P_i] \text{ is a path, } i \in I \}
-\end{align*}
-In other words, $P_i$ is the path on the tree $T$ assigned to $S_i$ in
-$\A$. As mentioned before for set systems, the paths cover the whole
-tree, i.e. $\bigcup_{i \in I}P_i = V$ \\
-
-\noindent
-Generalizing the definition of {\em feasibility} in \cite{nsnrs09} to
-a set assignment, a path assignment $\A$ is defined to be {\em
-  feasible} if there exists a bijection defined as follows.
-\begin{align}
-  \sigma: U \rightarrow V(T), \text{ such that }\sigma(S_i) = P_i
-  \text{ for all } i \in I, \sigma \text{ is a bijection}
-  \label{eq:stf}
-\end{align}
+\remove[dire]{
+A path assignment }% $\A$ to $\F$ is defined as a set assignment
+% where second universe is the vertex set $V$ of a given tree $T$ and
+% every second subset in the ordered pairs is a path in this
+% tree. Formally, the definition is as follows. 
+% \begin{align*}
+%   \A = \{ (S_i,P_i) \mid S_i \in \F, P_i \subseteq V \text{
+%     s.t. }T[P_i] \text{ is a path, } i \in I \}
+% \end{align*}
+% In other words, $P_i$ is the path on the tree $T$ assigned to $S_i$ in
+% $\A$. As mentioned before for set systems, the paths cover the whole
+% tree, i.e. $\bigcup_{i \in I}P_i = V$ \\
+
+% Generalizing the definition of { feasibility} in cite{nsnrs09} to
+% a set assignment, a path assignment $\A$ is defined to be {
+%   feasible} if there exists a bijection defined as follows.
+% \begin{align}
+%   \sigma: U \rightarrow V(T), \text{ such that }\sigma(S_i) = P_i
+%   \text{ for all } i \in I, \sigma \text{ is a bijection}
+%   \label{eq:stf}
+% \end{align}
+
+\begin{definition}[Partially ordered sets]
 
@@
 $X$ if $\nexists X_q \in X$ such that $X_m
-\preccurlyeq X_q$. \\
+\preccurlyeq X_q$. 
+\end{definition}
 
@@
 
-In this section we give an algorithmic characterization of a
-feasibility of tree path labeling.  Consider a path labeling $(\cF,
+\tnote{\sc refer to Problem Defs here}
+In this section we give an algorithmic characterization of the
+feasibility of a tree path labeling.  Consider a path labeling $(\cF,
 \cl)$ on the given tree $T$. We call $(\cF, \cl)$ an {\em Intersection
@@
   % \vspace{\topshrink}
-\item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ for all distinct $S_1, S_2 \in \cF$
+\item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ for all
+  distinct $S_1, S_2 \in \cF$
   % \vspace{\topshrink}
@@
 \end{lemma}
-\begin{proof}%[Proof of Lemma~\ref{lem:setminuscard}]
+\begin{proof}\thesisspacing%[Proof of Lemma~\ref{lem:setminuscard}]
   Let $P_i = \cl(S_i)$, for all $1 \le i \le 3$.  $|S_1 \cap (S_2
@@
   \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3| = |P_1 \cap (P_2
-  \setminus P_3)|$. Thus lemma is proven. %\qed
+  \setminus P_3)|$. %\qed
 \end{proof}
@@
 feasible if and only if it is an ICPPL and
-Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
+Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty path
+hypergraph isomorphism
 function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two
-levels of filtering of $(\cF, \cl)$ to make it simpler while retaining
-the set of isomorphisms, if any, between $\cF$ and $\cF^\cl$.
+levels of filtering of $(\cF, \cl)$ to make it simpler in terms of set
+intersections while retaining the set of isomorphisms, if any, between
+$\cF$ and $\cF^\cl$.
 % One direction of this claim isclear: that if a path labeling is
 % feasible, then all intersection cardinalities are preserved,
-% i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
+% i.e. the path labeling is an ICPPL. Algorithm~\ref{al:filterone}
 % \annote{has no premature exit condition hence any input will go
-%   through it}{Prove that the filtered sets has ICPPL iff input PL
-%   has ICPPL?}. Algorithm~\ref{leafasgn} has an exit condition at
+%   through it}{Prove that the filtered sets has ICPPL \iff input PL
+%   has ICPPL?}. Algorithm~\ref{al:filtertwo} has an exit condition at
 % line~\ref{xempty}. It can be easily verified that $X$ cannot be
@@
 % to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.
-First, we present Algorithm~\ref{perms} or {\tt filter\_1}, and prove
-its correctness.  This algorithm refines the path labeling by
-processing pairs of paths in $\cF^\cl$ that share a leaf until no two
-paths in the new path labeling share any leaf.
+First, we present Algorithm~\ref{al:filterone} which describes {\tt
+  filter\_1}, and prove its correctness.  This algorithm refines the
+path labeling by processing pairs of paths in $\cF^\cl$ that share a
+leaf until no two paths in the new path labeling share any leaf.
 
@@
   \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
-  \label{perms}
+  \label{al:filterone} % {perms}
   \begin{algorithmic}[\lndisplay]
@@
   \label{lem:feasible}
-  In Algorithm~\ref{perms}, if input $(\cF, \cl)$ is a feasible path
+  In Algorithm~\ref{al:filterone}, if input $(\cF, \cl)$ is a feasible path
   assignment then at the end of $j$th iteration of the {\bf while}
@@
 \end{lemma}
-\begin{proof}%[Proof of Lemma~\ref{lem:feasible}]
+\begin{proof}\thesisspacing%[Proof of Lemma~\ref{lem:feasible}]
   We will prove this by mathematical induction on the number of
@@
   the input itself which is given to be feasible. Assume the lemma is
-  true till $j-1$th iteration. i.e. every hypergraph isomorphism
-  $\phi: supp\left(\cF_{j-1}\right) \rightarrow V\left(T \right)$ that
+  true till $j-1$th iteration. \ie every hypergraph isomorphism $\phi:
+  supp\left(\cF_{j-1}\right) \rightarrow V\left(T \right)$ that
   defines $(\cF, \cl)$'s feasibility, is such that the induced path
-  labeling on $\cF_{j-1}$, $\cl_{\phi[{\cF_{j-1}}]}$ is equal to
+  labeling on $\cF_{j-1}$, say denoted by $\cl_{\phi[{\cF_{j-1}}]}$, is equal to
   $\cl_{j-1}$. We will prove that $\phi$ is also the bijection that
@@
   supp(\cF_{j})$ since the new sets in $\cF_j$ are created from basic
-  set operations to the sets in $\cF_{j-1}$. For the same reason and
-  $\phi$ being a bijection, it is clear that when applying the $\phi$
-  induced path labeling on $\cF_j$, $ \cl_{\phi[{\cF_{j}}]}(S_1
-  \setminus S_2) = \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
+  set operations to the sets in $\cF_{j-1}$ adding or removing no
+  elements. For the same reason and $\phi$ being a bijection, it is
+  clear that when applying the $\phi$-induced path labeling on
+  $\cF_j$, $ \cl_{\phi[{\cF_{j}}]}(S_1 \setminus S_2) =
+  \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
   \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Now observe that $ \cl_j(S_1
@@
   \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Thus the induced path labeling
-  $\cl_{\phi[{\cF_{j}}]} = \cl_{j}$. Therefore lemma is proven.  %\qed
+  $\cl_{\phi[{\cF_{j}}]} = \cl_{j}$.  %\qed
 \end{proof}
 
-\temptext{ --- WG11 begin --- }
-
 \begin{lemma}
-  \label{lem:setminuscard}
-  If $\cA$ is an ICPPA, and $(S_1, P_1),(S_2, P_2),(S_3, P_3) \in
-  \cA$, then $|S_1 \cap (S_2 \setminus S_3)| = |P_1 \cap (P_2
-  \setminus P_3)|$.
-\end{lemma}
-\begin{proof}
-  $|S_1 \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| =
-  |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to conditions (ii)
-  and (iii) of ICPPA, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
-  \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
-  = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. %\qed
-\end{proof}
-
-\begin{lemma}
-  \label{lem:fourpaths} Consider four paths in a tree $Q_1, Q_2, Q_3,
-  Q_4$ such that they have nonempty pairwise intersection and $Q_1,
-  Q_2$ share a leaf. Then there exists distinct $i, j, k \in
-  \{1,2,3,4\}$ such that, $Q_1 \cap Q_2 \cap Q_3 \cap Q_4 = Q_i \cap
-  Q_j \cap Q_k$.
-\end{lemma}
-\begin{proof}
-  {\em Case 1:} w.l.o.g, consider $Q_3 \cap Q_4$ and let us call it
-  $Q$. This is clearly a path (intersection of two paths is a path).
-  % Since $Q_1, Q_2$ share a leaf, the following are paths $Q_1
-  % \setminus Q_2$, $Q_2 \setminus Q_1$, $Q_1 \cap Q_2$ and they are
-  % mutually disjoint.
-  Suppose $Q$ does not intersect with $Q_1 \setminus Q_2$, i.e. $Q
-  \cap (Q_1 \setminus Q_2) = \O$. Then $Q \cap Q_1 \cap Q_2 = Q \cap
-  Q_2$. Similarly, if $Q \cap (Q_2 \setminus Q_1) = \O$, $Q \cap Q_1
-  \cap Q_2 = Q \cap Q_1$. Thus it is clear that if the intersection of
-  any two paths does not intersect with any of the set differences of
-  the remaining two paths, the claim in the lemma is true.
-  % Note that $Q_1 \setminus Q_2$ and $Q_2
-  % \setminus Q_1$ are paths because $Q_1, Q_2$ share a leaf.\\
-  {\em Case 2:} Let us consider the compliment of the previous
-  case. i.e. the intersection of any two paths intersects with both
-  the set differences of the other two. First let us consider $Q \cap
-  (Q_1 \setminus Q_2) \ne \O$ and $Q \cap (Q_1 \setminus Q_2) \ne \O$,
-  where $Q = Q_3 \cap Q_4$. Since $Q_1$ and $Q_2$ share a leaf, there
-  is exactly one vertex at which they branch off from the path $Q_1
-  \cap Q_2$ into two paths $Q_1 \setminus Q_2$ and $Q_2 \setminus
-  Q_1$. Let this vertex be $v$. It is clear that if path $Q_3 \cap
-  Q_4$, must intersect with paths $Q_1 \setminus Q_2$ and $Q_2
-  \setminus Q_1$, it must contain $v$ since these are paths from a
-  tree. Moreover, $Q_3 \cap Q_4$ intersects with $Q_1 \cap Q_2$ at
-  exactly $v$ and only at $v$ which means that $Q_1 \cap Q_2$ does not
-  intersect with $Q_3 \setminus Q_4$ or $Q_4 \setminus Q_3$ which
-  contradicts initial condition of this case. Thus this
-  case cannot occur and case 1 is the only possible scenario. \\
-  Thus lemma is proven %\qed
-\end{proof}
-
-\begin{lemma}
-  \label{lem:invar1}
-  In Algorithm \ref{perms}, at the end of $j$th iteration, $j \ge 0$,
-  of the while loop of Algorithm \ref{perms}, the following invariants
-  are maintained.
-  \begin{itemize}
-  \item {\em Invariant I:} $Q$ is a path in $T$ for each $(P,Q) \in
-    \Pi_j$
-  \item {\em Invariant II:} $|P|=|Q|$ for each $(P,Q) \in \Pi_j$
-  \item {\em Invariant III:} For any two $(P,Q), (P',Q') \in \Pi_j$,
-    $|P' \cap P''|=|Q' \cap Q''|$.
-  \item {\em Invariant IV:} For any three, $(P',Q'), (P'',Q''), (P, Q)
-    \in \Pi_j$, $|P' \cap P'' \cap P|=|Q' \cap Q'' \cap Q|$.
-  \end{itemize}
-\end{lemma}
-\begin{proof}
-  Proof is by induction on the number of iterations, $j$. In the rest
-  of the proof, the term ``new sets'' will refer to the new sets added
-  in $j$th iteration as defined in line \ref{setbreak} of Algorithm
-  \ref{perms}, i.e. the following three assignment pairs for some
-  $(P_1,Q_1), (P_2,Q_2) \in \Pi_{j-1}$ where $Q_1$ and $Q_2$ intersect
-  and share a leaf: $(P_1 \cap P_2, Q_1 \cap Q_2)$, or $(P_1 \setminus
-  P_2, Q_1 \setminus Q_2)$, or $(P_2 \setminus P_1, Q_2
-  \setminus Q_1)$.\\
-  \noindent
-  The base case, $\Pi_0 = \{(S_i,P_i) \mid i \in [m]\}$, is trivially
-  true since it is the input which is an ICPPA.  Assume the lemma is
-  true till the $j-1$ iteration. Consider $j$th
-  iteration:\\
-  \noindent
-  If $(P,Q)$, $(P',Q')$ and $(P'',Q'')$ are in $\Pi_{j}$ and
-  $\Pi_{j-1}$, all the invariants are
-  clearly true because they are from $j-1$ iteration.\\
-  If $(P,Q)$ is in $\Pi_{j}$ and not in $\Pi_{j-1}$, then it must be
-  one of the new sets added in $\Pi_j$. Since $(P_1,Q_1)$ and
-  $(P_2,Q_2)$ are from $\Pi_{j-1}$ and $Q_1,Q_2$ intersect and have a
-  common leaf, it can be verified that the
-  new sets are also paths. \\
-  By hypothesis for invariant III, invariant II also holds for $(P,Q)$
-  no matter which new set in $\Pi_j$ it
-  is.\\
-  To prove invariant III, if $(P,Q)$ and $(P',Q')$ are not in
-  $\Pi_{j-1}$, then they are both new sets and invariant III holds
-  trivially (new sets are disjoint). Next consider $(P,Q), (P',Q') \in
-  \Pi_j$ with only one of them, say $(P',Q')$, in $\Pi_{j-1}$. Then
-  $(P,Q)$ is one of the new sets added in line \ref{setbreak}. It is
-  easy to see that if $(P,Q)$ is $(P_1 \cap P_2, Q_1 \cap Q_2)$, then
-  due to invariant IV in hypothesis, invariant III becomes true in
-  this iteration. Similarly, using lemma \ref{lem:setminuscard}
-  invariant III is proven if $(P, Q)$ is $(P_1 \setminus P_2, Q_1
-  \setminus Q_2)$, or $(P_2 \setminus P_1, Q_2
-  \setminus Q_1)$.\\
-  To prove invariant IV, consider three assignments
-  $(P,Q),(P',Q'),(P'',Q'')$. If at least two of these pairs are in not
-  $\Pi_{j-1}$, then they are any two of the new sets. Note that these
-  new sets are disjoint and hence if $(P',Q'), (P'',Q'')$ are any of
-  these sets, $|P \cap P' \cap P''|=|Q \cap Q' \cap Q''|=0$ and
-  invariant IV is true. Now we consider the case if at most one of
-  $(P,Q),(P',Q'),(P'',Q'')$ is not in $\Pi_{j-1}$. If none of them are
-  not in $\Pi_{j-1}$ (i.e. all of them are in $\Pi_{j-1}$), invariant
-  IV is clearly true. Consider the case where exactly one of them is
-  not in $\Pi_{j-1}$. w.l.o.g let that be $(P,Q)$ and it could be any
-  one of the new sets. If $(P,Q)$ is $(P_1 \cap P_2, Q_1 \cap Q_2)$,
-  from lemma \ref{lem:fourpaths} and invariant III hypothesis,
-  invariant IV is proven. Similarly if $(P,Q)$ is any of the other new
-  sets, invariant IV is proven by also using lemma
-  \ref{lem:setminuscard}. %\qed
-
-\end{proof} 
-\temptext{ --- WG11 end --- }
-
-\begin{lemma}
-  \label{lem:invar1} In Algorithm~\ref{perms}, at the end of $j$th
+  \label{lem:invar1} In Algorithm~\ref{al:filterone}, at the end of $j$th
   iteration, $j \ge 0$, of the {\bf while} loop, the following
@@
 
-\begin{proof}
+\begin{proof}\thesisspacing
   Proof is by induction on the number of iterations, $j$. In this
   proof, the term ``new sets'' will refer to the sets added to $\cF_j$
-  in $j$th iteration in line~\ref{setbreak} of Algorithm~\ref{perms},
+  in $j$th iteration in line~\ref{setbreak} of Algorithm~\ref{al:filterone},
   $S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1$ and its images
@@
   % \vspace{-6mm}
-
 \end{proof}
@@
   \label{lem:noexit1}
-  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{perms} is
+  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{al:filterone} is
   feasible, then the set of hypergraph isomorphism functions that
@@
 \end{lemma}
-\begin{proof}
+\begin{proof}\thesisspacing
   Since $(\cF,\cl)$ is feasible, by Lemma~\ref{lem:feasible}
@@
   hypergraph isomorphisms of the original path labeling.\\
-  Secondly, line~\ref{ln:exit1} will execute iff the exit condition in
+  Secondly, line~\ref{ln:exit1} will execute \iff the exit condition in
   line~\ref{ln:3waycheck}, i.e. failure of three way intersection
@@
 
-As a result of Algorithm~\ref{perms} each leaf $v$ in $T$ is such that
+As a result of Algorithm~\ref{al:filterone} each leaf $v$ in $T$ is such that
 there is exactly one set in $\cF$ with $v$ as a vertex in the path
-assigned to it.  In Algorithm~\ref{leafasgn} we identify elements in
+assigned to it.  In Algorithm~\ref{al:filtertwo} we identify elements in
 $supp(\cF)$ whose images are leaves in a hypergraph isomorphism if one
@@
 = T \setminus \{v \mid v \text{ is a leaf in }
-T\}$. Algorithm~\ref{leafasgn} is now presented with details.
+T\}$. Algorithm~\ref{al:filtertwo} is now presented with details.
+
+%\temptext{ --- WG11 begin --- } %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% \begin{lemma}
+%   \label{lem:setminuscard}
+%   If $\cA$ is an ICPPA, and $(S_1, P_1),(S_2, P_2),(S_3, P_3) \in
+%   \cA$, then $|S_1 \cap (S_2 \setminus S_3)| = |P_1 \cap (P_2
+%   \setminus P_3)|$.
+% \end{lemma}
+% \begin{proof}\thesisspacing
+%   $|S_1 \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| =
+%   |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to conditions (ii)
+%   and (iii) of ICPPA, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
+%   \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
+%   = |P_1 \cap (P_2 \setminus P_3)|$. Thus lemma is proven. %\qed
+% \end{proof}
+
+% \begin{lemma}
+%   \label{lem:invar1}
+%   In Algorithm \ref{al:filterone}, at the end of $j$th iteration, $j \ge 0$,
+%   of the while loop of Algorithm \ref{al:filterone}, the following invariants
+%   are maintained.
+%   \begin{itemize}
+%   \item {\em Invariant I:} $Q$ is a path in $T$ for each $(P,Q) \in
+%     \Pi_j$
+%   \item {\em Invariant II:} $|P|=|Q|$ for each $(P,Q) \in \Pi_j$
+%   \item {\em Invariant III:} For any two $(P,Q), (P',Q') \in \Pi_j$,
+%     $|P' \cap P''|=|Q' \cap Q''|$.
+%   \item {\em Invariant IV:} For any three, $(P',Q'), (P'',Q''), (P, Q)
+%     \in \Pi_j$, $|P' \cap P'' \cap P|=|Q' \cap Q'' \cap Q|$.
+%   \end{itemize}
+% \end{lemma}
+% \begin{proof}\thesisspacing
+%   Proof is by induction on the number of iterations, $j$. In the rest
+%   of the proof, the term ``new sets'' will refer to the new sets added
+%   in $j$th iteration as defined in line \ref{setbreak} of Algorithm
+%   \ref{al:filterone}, i.e. the following three assignment pairs for some
+%   $(P_1,Q_1), (P_2,Q_2) \in \Pi_{j-1}$ where $Q_1$ and $Q_2$ intersect
+%   and share a leaf: $(P_1 \cap P_2, Q_1 \cap Q_2)$, or $(P_1 \setminus
+%   P_2, Q_1 \setminus Q_2)$, or $(P_2 \setminus P_1, Q_2
+%   \setminus Q_1)$.\\
+%   \noindent
+%   The base case, $\Pi_0 = \{(S_i,P_i) \mid i \in [m]\}$, is trivially
+%   true since it is the input which is an ICPPA.  Assume the lemma is
+%   true till the $j-1$ iteration. Consider $j$th
+%   iteration:\\
+%   \noindent
+%   If $(P,Q)$, $(P',Q')$ and $(P'',Q'')$ are in $\Pi_{j}$ and
+%   $\Pi_{j-1}$, all the invariants are
+%   clearly true because they are from $j-1$ iteration.\\
+%   If $(P,Q)$ is in $\Pi_{j}$ and not in $\Pi_{j-1}$, then it must be
+%   one of the new sets added in $\Pi_j$. Since $(P_1,Q_1)$ and
+%   $(P_2,Q_2)$ are from $\Pi_{j-1}$ and $Q_1,Q_2$ intersect and have a
+%   common leaf, it can be verified that the
+%   new sets are also paths. \\
+%   By hypothesis for invariant III, invariant II also holds for $(P,Q)$
+%   no matter which new set in $\Pi_j$ it
+%   is.\\
+%   To prove invariant III, if $(P,Q)$ and $(P',Q')$ are not in
+%   $\Pi_{j-1}$, then they are both new sets and invariant III holds
+%   trivially (new sets are disjoint). Next consider $(P,Q), (P',Q') \in
+%   \Pi_j$ with only one of them, say $(P',Q')$, in $\Pi_{j-1}$. Then
+%   $(P,Q)$ is one of the new sets added in line \ref{setbreak}. It is
+%   easy to see that if $(P,Q)$ is $(P_1 \cap P_2, Q_1 \cap Q_2)$, then
+%   due to invariant IV in hypothesis, invariant III becomes true in
+%   this iteration. Similarly, using lemma \ref{lem:setminuscard}
+%   invariant III is proven if $(P, Q)$ is $(P_1 \setminus P_2, Q_1
+%   \setminus Q_2)$, or $(P_2 \setminus P_1, Q_2
+%   \setminus Q_1)$.\\
+%   To prove invariant IV, consider three assignments
+%   $(P,Q),(P',Q'),(P'',Q'')$. If at least two of these pairs are in not
+%   $\Pi_{j-1}$, then they are any two of the new sets. Note that these
+%   new sets are disjoint and hence if $(P',Q'), (P'',Q'')$ are any of
+%   these sets, $|P \cap P' \cap P''|=|Q \cap Q' \cap Q''|=0$ and
+%   invariant IV is true. Now we consider the case if at most one of
+%   $(P,Q),(P',Q'),(P'',Q'')$ is not in $\Pi_{j-1}$. If none of them are
+%   not in $\Pi_{j-1}$ (i.e. all of them are in $\Pi_{j-1}$), invariant
+%   IV is clearly true. Consider the case where exactly one of them is
+%   not in $\Pi_{j-1}$. w.l.o.g let that be $(P,Q)$ and it could be any
+%   one of the new sets. If $(P,Q)$ is $(P_1 \cap P_2, Q_1 \cap Q_2)$,
+%   from lemma \ref{lem:fourpaths} and invariant III hypothesis,
+%   invariant IV is proven. Similarly if $(P,Q)$ is any of the other new
+%   sets, invariant IV is proven by also using lemma
+%   \ref{lem:setminuscard}. %\qed
+
+% \end{proof} 
+%\temptext{ --- WG11 end --- }%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
   \caption{Leaf labeling from an ICPPL {\tt filter\_2($\cF, \cl, T$)}}
-  \label{leafasgn}
+  \label{al:filtertwo} % {leafasgn}
   \begin{algorithmic}[\lndisplay]
@@
 Suppose the input ICPPL $(\cF, \cl)$ is feasible, yet set $X$ in
-Algorithm~\ref{leafasgn} is empty in some iteration of the {\bf while}
+Algorithm~\ref{al:filtertwo} is empty in some iteration of the {\bf while}
 loop. This will abort our procedure of finding the hypergraph
@@
 
-\temptext{ --- WG11 begin --- }
-\begin{lemma}
-  \label{lem:invar3}
-  In Algorithm \ref{leafasgn},for all $j \geq 0$, at the end of the
-  $j$th iteration the four invariants given in lemma \ref{lem:invar1}
-  are valid.
-  % Moreover, $X$ as defined in the algorithm is non-empty if this is
-  % an ICPPA.
-\end{lemma}
-\begin{proof}
-  First we see that $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in
-    I}S_i$ is non empty in every iteration for an ICPPA. Suppose $X$
-  is empty. We know that $v \in P_{i_1} \setminus \bigcup_{i \ne i_1,
-    i \in I}P_i$ since $v$ is in the unique path $P_{i_1}$. Since this
-  is an ICPPA $|S_{i_1}| = |P_{i_1}|$. For any $x \in S_{i_1}$ it is
-  contained in at least two sets due to our assumption. Let $S_{i_2}$
-  be a second set that contains $x$. We know $v \notin
-  P_{i_2}$. Therefore there cannot exist a permutation that maps
-  elements of $S_{i_2}$ to $P_{i_2}$. This contradicts our assumption
-  that this is an ICPPA. Thus $X$ cannot be empty.
-
-  We use mathematical induction on the number of iterations for this
-  proof. The term ``new sets'' will refer to the sets added in $\Pi_j$
-  in the $j$th iteration, i.e. $(P' \setminus \{x\},Q' \setminus
-  \{v\})$ and $(\{x\},\{v\})$ for some $(P',Q')$ in $\Pi_{j-1}$ such
-  that $v$ is a leaf and $Q'$ is the unique path
-  incident on it.\\
-  For $\Pi_0$ all invariants hold because it is output from algorithm
-  \ref{perms} which is an ICPPA. Hence base case is proved.  Assume
-  the lemma holds for $\Pi_{j-1}$. Consider $\Pi_j$ and any $(P,Q) \in
-  \Pi_j$. If $(P,Q) $ is in $ \Pi_j$ and $\Pi_{j-1}$ invariants I and
-  II are true because of induction assumption. If it is only in
-  $\Pi_j$, then it is $\{(P' \setminus \{x\}),(Q' \setminus \{v\})$ or
-  $(\{x\},\{v\})$ for some $(P',Q')$ in $\Pi_{j-1}$. By definition,
-  $x$ is an element in $P'$ (as defined in the algorithm) and $v$ is a
-  leaf in $Q'$. If $(P,Q)$ is $\{(P' \setminus \{x\}),(Q' \setminus
-  \{v\})$, $Q$ is a path since only a leaf is removed from path
-  $Q'$. We know $|P'| = |Q'|$, therefore $|P' \setminus \{x\}| = |Q'
-  \setminus \{v\}|$. Hence in this case invariants I and II are
-  obvious. It is easy to see these invariants hold if $(P,Q)$ is
-  $(\{x\},\{v\})$.
-
-  For invariant III consider $(P_1,Q_1),(P_2,Q_2)$ in $\Pi_j$. If both
-  of them are also in $\Pi_{j-1}$, claim is proved. If one of them is
-  not in $\Pi_{j-1}$ then it has to be $\{(P' \setminus \{x\}),(Q'
-  \setminus \{v\})$ or $(\{x\},\{v\})$ for some $(P',Q')$ in
-  $\Pi_{j-1}$. Since by definition, $Q'$ is the only path with $v$ and
-  $P'$ the only set with $x$ in the previous iteration, $|P_1 \cap (P'
-  \setminus \{x\})| = |P_1 \cap P'|$ and $|Q_1 \cap (Q' \setminus
-  \{v\})| = |Q_1 \cap Q'|$ and $|P_1 \cap \{x\}| = 0, Q_1 \cap \{v\} =
-  0$. Thus invariant III is also proven.
-
-  \noindent
-  To prove invariant IV, consider $(P_1,Q_1),(P_2,Q_2), (P_3,Q_3)$ in
-  $\Pi_j$. If exactly one of them, say $P_3 \notin \Pi_{j-1}$, it is
-  one of the new sets. By the same argument used to prove invariant
-  III, $|P_1 \cap P_2 \cap (P' \setminus \{x\})| = |P_1 \cap P_2 \cap
-  P'|$ and $|Q_1 \cap Q_2 \cap (Q' \setminus \{x\})| = |Q_1 \cap Q_2
-  \cap Q'|$. Since $P_1, P_2, P'$ are all in $\Pi_{j-1}$, by induction
-  hypothesis $|P_1 \cap P_2 \cap P'| = |Q_1 \cap Q_2 \cap Q'|$. Also
-  $|P_1 \cap P_2 \cap \{x\}| = 0, Q_1 \cap Q_2 \cap \{v\} = 0$.  If
-  two or more of them are not in $\Pi_{j-1}$, then it can be verified
-  that $|P_1 \cap P_2 \cap P_3| = |Q_1 \cap Q_2 \cap Q_3|$ since the
-  new sets in $\Pi_j$ are either disjoint or as follows: assuming
-  $P_1, P_2 \notin \Pi_{j-1}$ and new sets are derived from $(P', Q'),
-  (P'', Q'') \in \Pi_{j-1}$ with $x_1, x_2$ exclusively in $P_1, P_2$,
-  $(\{x_1\},\{v_1\}), (\{x_2\},\{v_2\}) \in \Pi_j $ thus $v_1, v_2$
-  are exclusively in $Q_1, Q_2$ resp. it follows that $|P_1 \cap P_2
-  \cap P_3| = |(P' \setminus \{x_1\}) \cap (P'' \setminus \{x_2\})
-  \cap P_3| = |P' \cap P'' \cap P_3| = |Q' \cap Q'' \cap Q_3| = |(Q'
-  \setminus \{v_1\} \cap Q'' \setminus \{v_2\} \cap Q_3| = |Q_1 \cap
-  Q_2 \cap Q_3|$. Thus invariant IV is also proven.  %\qed
-\end{proof}
-
-Using algorithms \ref{perms} and \ref{leafasgn} we prove the following
-theorem.
-
-\begin{theorem}
-  \label{th:perm}
-  If $\cA$ is an ICPPA, then there exists a bijection $\sigma : U
-  \rightarrow V(T)$ such that $\sigma(S_i) = P_i$ for all $i \in I$
-\end{theorem}
-\begin{proof}
-  This is a contructive proof. First, the given ICPPA $\cA$ and tree
-  $T$ are given as input to Algorithm \ref{perms}. This yields a
-  ``filtered'' ICPPA as the output which is input to Algorithm
-  \ref{leafasgn}.  It can be observed that the output of Algorithm
-  \ref{leafasgn} is a set of interval assignments to sets and
-  one-to-one assignment of elements of $U$ to each leaf of $T$. To be
-  precise, it would be of the form $\cB_0 = \cA_0 \cup \cL_0$. The
-  leaf assignments are defined in $\cL_0 = \{ (x_i,v_i) \mid x_i \in
-  U, v_i \in T, x_i \ne x_j, v_i \ne v_j, i \ne j, i,j \in [k] \}$
-  where $k$ is the number of leaves in $T$. The path assignments are
-  defined in $\cA_0 \subseteq \{(S_i',P_i') \mid S_i' \subseteq U_0,
-  P_i' \text{ is a path from } T_0\}$ where $T_0$ is the tree obtained
-  by removing all the leaves in $T$ and $U_0 = U \setminus \{ x \mid x
-  \text{ is assigned to a leaf in }\cL_0 \}$. Now we have a subproblem
-  of finding the permutation for the path assignment $\cA_0$ which has
-  paths from tree $T_0$ and sets from universe $U_0$. Now we repeat
-  the procedure and the path assignment $\cA_0$ and tree $T_0$ is
-  given as input to Algorithm \ref{perms}. The output of this
-  algorithm is given to Algorithm \ref{leafasgn} to get a new union of
-  path and leaf assignments $\cB_1 = \cA_1 \cup \cL_1$ defined similar
-  to $\cB_0, \cL_0, \cA_0$. In general, the two algorithms are run on
-  path assignment $\cA_{i-1}$ with paths from tree $T_{i-1}$ to get a
-  new subproblem with path assignment $\cA_i$ and tree $T_{i}$. $T_i$
-  is the subtree of $T_{i-1}$ obtained by removing all its
-  leaves. More importantly, it gives leaf assignments $\cL_{i}$ to the
-  leaves in tree $T_{i-1}$. This is continued until we get a
-  subproblem with path assignment $\cA_{d-1}$ and tree $T_{d-1}$ for
-  some $d \le n$ which is just a path. From the last lemma we know
-  that $\cA_{d-1}$ is an ICPPA. Another observation is that an ICPPA
-  with all its tree paths being intervals (subpaths from a path) is
-  nothing but an ICPIA\cite{nsnrs09}.  Let $\cA_{d-1}$ be equal to
-  $\{(S_i'',P_i'') \mid S_i'' \subseteq U_{d-1}, P_i'' \text{ is a
-    path from } T_{d-1} \}$. It is true that the paths $P_i''$s may
-  not be precisely an interval in the sense of consecutive integers
-  because they are some nodes from a tree. However, it is easy to see
-  that the nodes of $T_{d-1}$ can be ordered from left to right and
-  ranked to get intervals $I_i$ for every path $P_i''$ as
-  follows. $I_i = \{[l,r] \mid l = \text{ the lowest rank of the nodes
-    in }P_i'', r = l+|P_i''|-1 \}$. Let asssignment $\cA_d$ be with
-  the renamed paths. $\cA_d = \{ (S_i'', I_i) \mid (S_i'', P_i'') \in
-  \cA_{d-1} \}$. What has been effectively done is renaming the nodes
-  in $T_{d-1}$ to get a tree $T_d$.  The ICPIA $\cA_d$ is now in the
-  format that the ICPIA algorithm requires which gives us the
-  permutation $\sigma' : U_{d-1} \rightarrow T_{d-1}$
-
-\noindent
-$\sigma'$ along with all the leaf assignments $\cL_i$ gives us the
-permutation for the original path assignment $\cA$.  More precisely,
-the permutation for tree path assignment $\cA$ is defined as
-follows. $\sigma: U \rightarrow T$ such that the following is
-maintained.
-\begin{align*}
-  \sigma(x) &= \sigma'(x),   \text{ if } x \in U_{d-1} \\
-  &= \cL_i(x), \text{ where $x$ is assigned to a leaf in a subproblem
-    $\cA_{i-1}, T_{i-1}$}
-\end{align*}
-
-\noindent
-To summarize, run algorithm \ref{perms} and \ref{leafasgn} on
-$T$. After the leaves have been assigned to specific elements from
-$U$, remove all leaves from $T$ to get new tree $T_0$. The leaf
-assignments are in $\cL_0$. Since only leaves were removed $T_0$ is
-indeed a tree. Repeat the algorithms on $T_0$ to get leaf assignments
-$\cL_{1}$. Remove the leaves in $T_0$ to get $T_1$ and so on until the
-pruned tree $T_d$ is a single path. Now run ICPIA algorithm on $T_d$
-to get permutation $\sigma'$. The relation $\cL_0 \cup \cL_1 \cup
-.. \cup \cL_{d} \cup \sigma'$ gives the bijection required in the
-original problem.%\qed
-\end{proof}
-\temptext{ --- WG11 end --- }
+%\temptext{ --- WG11 begin --- }%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \begin{lemma}
+%   \label{lem:invar3}
+%   In Algorithm \ref{al:filtertwo},for all $j \geq 0$, at the end of the
+%   $j$th iteration the four invariants given in lemma \ref{lem:invar1}
+%   are valid.
+%   % Moreover, $X$ as defined in the algorithm is non-empty if this is
+%   % an ICPPA.
+% \end{lemma}
+% \begin{proof}\thesisspacing
+%   First we see that $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in
+%     I}S_i$ is non empty in every iteration for an ICPPA. Suppose $X$
+%   is empty. We know that $v \in P_{i_1} \setminus \bigcup_{i \ne i_1,
+%     i \in I}P_i$ since $v$ is in the unique path $P_{i_1}$. Since this
+%   is an ICPPA $|S_{i_1}| = |P_{i_1}|$. For any $x \in S_{i_1}$ it is
+%   contained in at least two sets due to our assumption. Let $S_{i_2}$
+%   be a second set that contains $x$. We know $v \notin
+%   P_{i_2}$. Therefore there cannot exist a permutation that maps
+%   elements of $S_{i_2}$ to $P_{i_2}$. This contradicts our assumption
+%   that this is an ICPPA. Thus $X$ cannot be empty.
+
+%   We use mathematical induction on the number of iterations for this
+%   proof. The term ``new sets'' will refer to the sets added in $\Pi_j$
+%   in the $j$th iteration, i.e. $(P' \setminus \{x\},Q' \setminus
+%   \{v\})$ and $(\{x\},\{v\})$ for some $(P',Q')$ in $\Pi_{j-1}$ such
+%   that $v$ is a leaf and $Q'$ is the unique path
+%   incident on it.\\
+%   For $\Pi_0$ all invariants hold because it is output from algorithm
+%   \ref{al:filterone} which is an ICPPA. Hence base case is proved.  Assume
+%   the lemma holds for $\Pi_{j-1}$. Consider $\Pi_j$ and any $(P,Q) \in
+%   \Pi_j$. If $(P,Q) $ is in $ \Pi_j$ and $\Pi_{j-1}$ invariants I and
+%   II are true because of induction assumption. If it is only in
+%   $\Pi_j$, then it is $\{(P' \setminus \{x\}),(Q' \setminus \{v\})$ or
+%   $(\{x\},\{v\})$ for some $(P',Q')$ in $\Pi_{j-1}$. By definition,
+%   $x$ is an element in $P'$ (as defined in the algorithm) and $v$ is a
+%   leaf in $Q'$. If $(P,Q)$ is $\{(P' \setminus \{x\}),(Q' \setminus
+%   \{v\})$, $Q$ is a path since only a leaf is removed from path
+%   $Q'$. We know $|P'| = |Q'|$, therefore $|P' \setminus \{x\}| = |Q'
+%   \setminus \{v\}|$. Hence in this case invariants I and II are
+%   obvious. It is easy to see these invariants hold if $(P,Q)$ is
+%   $(\{x\},\{v\})$.
+
+%   For invariant III consider $(P_1,Q_1),(P_2,Q_2)$ in $\Pi_j$. If both
+%   of them are also in $\Pi_{j-1}$, claim is proved. If one of them is
+%   not in $\Pi_{j-1}$ then it has to be $\{(P' \setminus \{x\}),(Q'
+%   \setminus \{v\})$ or $(\{x\},\{v\})$ for some $(P',Q')$ in
+%   $\Pi_{j-1}$. Since by definition, $Q'$ is the only path with $v$ and
+%   $P'$ the only set with $x$ in the previous iteration, $|P_1 \cap (P'
+%   \setminus \{x\})| = |P_1 \cap P'|$ and $|Q_1 \cap (Q' \setminus
+%   \{v\})| = |Q_1 \cap Q'|$ and $|P_1 \cap \{x\}| = 0, Q_1 \cap \{v\} =
+%   0$. Thus invariant III is also proven.
+
+%   \noindent
+%   To prove invariant IV, consider $(P_1,Q_1),(P_2,Q_2), (P_3,Q_3)$ in
+%   $\Pi_j$. If exactly one of them, say $P_3 \notin \Pi_{j-1}$, it is
+%   one of the new sets. By the same argument used to prove invariant
+%   III, $|P_1 \cap P_2 \cap (P' \setminus \{x\})| = |P_1 \cap P_2 \cap
+%   P'|$ and $|Q_1 \cap Q_2 \cap (Q' \setminus \{x\})| = |Q_1 \cap Q_2
+%   \cap Q'|$. Since $P_1, P_2, P'$ are all in $\Pi_{j-1}$, by induction
+%   hypothesis $|P_1 \cap P_2 \cap P'| = |Q_1 \cap Q_2 \cap Q'|$. Also
+%   $|P_1 \cap P_2 \cap \{x\}| = 0, Q_1 \cap Q_2 \cap \{v\} = 0$.  If
+%   two or more of them are not in $\Pi_{j-1}$, then it can be verified
+%   that $|P_1 \cap P_2 \cap P_3| = |Q_1 \cap Q_2 \cap Q_3|$ since the
+%   new sets in $\Pi_j$ are either disjoint or as follows: assuming
+%   $P_1, P_2 \notin \Pi_{j-1}$ and new sets are derived from $(P', Q'),
+%   (P'', Q'') \in \Pi_{j-1}$ with $x_1, x_2$ exclusively in $P_1, P_2$,
+%   $(\{x_1\},\{v_1\}), (\{x_2\},\{v_2\}) \in \Pi_j $ thus $v_1, v_2$
+%   are exclusively in $Q_1, Q_2$ resp. it follows that $|P_1 \cap P_2
+%   \cap P_3| = |(P' \setminus \{x_1\}) \cap (P'' \setminus \{x_2\})
+%   \cap P_3| = |P' \cap P'' \cap P_3| = |Q' \cap Q'' \cap Q_3| = |(Q'
+%   \setminus \{v_1\} \cap Q'' \setminus \{v_2\} \cap Q_3| = |Q_1 \cap
+%   Q_2 \cap Q_3|$. Thus invariant IV is also proven.  %\qed
+% \end{proof}
+
+% Using algorithms \ref{al:filterone} and \ref{al:filtertwo} we prove the following
+% theorem.
+
+% \begin{theorem}
+%   \label{th:perm}
+%   If $\cA$ is an ICPPA, then there exists a bijection $\sigma : U
+%   \rightarrow V(T)$ such that $\sigma(S_i) = P_i$ for all $i \in I$
+% \end{theorem}
+% \begin{proof}\thesisspacing
+%   This is a contructive proof. First, the given ICPPA $\cA$ and tree
+%   $T$ are given as input to Algorithm \ref{al:filterone}. This yields a
+%   ``filtered'' ICPPA as the output which is input to Algorithm
+%   \ref{al:filtertwo}.  It can be observed that the output of Algorithm
+%   \ref{al:filtertwo} is a set of interval assignments to sets and
+%   one-to-one assignment of elements of $U$ to each leaf of $T$. To be
+%   precise, it would be of the form $\cB_0 = \cA_0 \cup \cL_0$. The
+%   leaf assignments are defined in $\cL_0 = \{ (x_i,v_i) \mid x_i \in
+%   U, v_i \in T, x_i \ne x_j, v_i \ne v_j, i \ne j, i,j \in [k] \}$
+%   where $k$ is the number of leaves in $T$. The path assignments are
+%   defined in $\cA_0 \subseteq \{(S_i',P_i') \mid S_i' \subseteq U_0,
+%   P_i' \text{ is a path from } T_0\}$ where $T_0$ is the tree obtained
+%   by removing all the leaves in $T$ and $U_0 = U \setminus \{ x \mid x
+%   \text{ is assigned to a leaf in }\cL_0 \}$. Now we have a subproblem
+%   of finding the permutation for the path assignment $\cA_0$ which has
+%   paths from tree $T_0$ and sets from universe $U_0$. Now we repeat
+%   the procedure and the path assignment $\cA_0$ and tree $T_0$ is
+%   given as input to Algorithm \ref{al:filterone}. The output of this
+%   algorithm is given to Algorithm \ref{al:filtertwo} to get a new union of
+%   path and leaf assignments $\cB_1 = \cA_1 \cup \cL_1$ defined similar
+%   to $\cB_0, \cL_0, \cA_0$. In general, the two algorithms are run on
+%   path assignment $\cA_{i-1}$ with paths from tree $T_{i-1}$ to get a
+%   new subproblem with path assignment $\cA_i$ and tree $T_{i}$. $T_i$
+%   is the subtree of $T_{i-1}$ obtained by removing all its
+%   leaves. More importantly, it gives leaf assignments $\cL_{i}$ to the
+%   leaves in tree $T_{i-1}$. This is continued until we get a
+%   subproblem with path assignment $\cA_{d-1}$ and tree $T_{d-1}$ for
+%   some $d \le n$ which is just a path. From the last lemma we know
+%   that $\cA_{d-1}$ is an ICPPA. Another observation is that an ICPPA
+%   with all its tree paths being intervals (subpaths from a path) is
+%   nothing but an ICPIA\cite{nsnrs09}.  Let $\cA_{d-1}$ be equal to
+%   $\{(S_i'',P_i'') \mid S_i'' \subseteq U_{d-1}, P_i'' \text{ is a
+%     path from } T_{d-1} \}$. It is true that the paths $P_i''$s may
+%   not be precisely an interval in the sense of consecutive integers
+%   because they are some nodes from a tree. However, it is easy to see
+%   that the nodes of $T_{d-1}$ can be ordered from left to right and
+%   ranked to get intervals $I_i$ for every path $P_i''$ as
+%   follows. $I_i = \{[l,r] \mid l = \text{ the lowest rank of the nodes
+%     in }P_i'', r = l+|P_i''|-1 \}$. Let asssignment $\cA_d$ be with
+%   the renamed paths. $\cA_d = \{ (S_i'', I_i) \mid (S_i'', P_i'') \in
+%   \cA_{d-1} \}$. What has been effectively done is renaming the nodes
+%   in $T_{d-1}$ to get a tree $T_d$.  The ICPIA $\cA_d$ is now in the
+%   format that the ICPIA algorithm requires which gives us the
+%   permutation $\sigma' : U_{d-1} \rightarrow T_{d-1}$
+
+% \noindent
+% $\sigma'$ along with all the leaf assignments $\cL_i$ gives us the
+% permutation for the original path assignment $\cA$.  More precisely,
+% the permutation for tree path assignment $\cA$ is defined as
+% follows. $\sigma: U \rightarrow T$ such that the following is
+% maintained.
+% \begin{align*}
+%   \sigma(x) &= \sigma'(x),   \text{ if } x \in U_{d-1} \\
+%   &= \cL_i(x), \text{ where $x$ is assigned to a leaf in a subproblem
+%     $\cA_{i-1}, T_{i-1}$}
+% \end{align*}
+
+% \noindent
+% To summarize, run algorithm \ref{al:filterone} and \ref{al:filtertwo} on
+% $T$. After the leaves have been assigned to specific elements from
+% $U$, remove all leaves from $T$ to get new tree $T_0$. The leaf
+% assignments are in $\cL_0$. Since only leaves were removed $T_0$ is
+% indeed a tree. Repeat the algorithms on $T_0$ to get leaf assignments
+% $\cL_{1}$. Remove the leaves in $T_0$ to get $T_1$ and so on until the
+% pruned tree $T_d$ is a single path. Now run ICPIA algorithm on $T_d$
+% to get permutation $\sigma'$. The relation $\cL_0 \cup \cL_1 \cup
+% .. \cup \cL_{d} \cup \sigma'$ gives the bijection required in the
+% original problem.%\qed
+% \end{proof}
+%\temptext{ --- WG11 end --- }%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
   \label{lem:xnotempty}
-  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{leafasgn} is
+  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{al:filtertwo} is
   feasible, then for all iterations $j > 0$ of the {\em \bf while}
@@
 \end{lemma}
-\begin{proof}
+\begin{proof}\thesisspacing
   Assume $X$ is empty for some iteration $j > 0$. We know that $v$ is
@@
   \label{lem:invar3}
-  In Algorithm~\ref{leafasgn}, for all $j > 0$, at the end of the
+  In Algorithm~\ref{al:filtertwo}, for all $j > 0$, at the end of the
   $j$th iteration of the {\bf while} loop the four invariants given in
@@
 \end{lemma}
-\begin{proof}
+\begin{proof}\thesisspacing
   By Lemma~\ref{lem:xnotempty} we know that set $X$ will not be empty
@@
 
-% \textcolor{cyan}{
-%   \begin{lemma}
-%     \label{lem:notfeasibleexit}
-%     \tnote{IS THIS CORRECT?}  If input ICPPL $(\cF, \cl)$ is not
-%     feasible, then in one of the recursive calls to Algorithm --3--,
-%     the {\em \bf exit} statement in line x in Algorithm --1 or line
-%     y in Algorithm --2 will get executed.
-%   \end{lemma}
-% }
-
 We have seen two filtering algorithms above, namely,
-Algorithm~\ref{perms} {\tt filter\_1} and Algorithm~\ref{leafasgn}
-{\tt filter\_2} which when executed serially repectively result in a
-new ICPPL on the same universe $U$ and tree $T$. We also proved that
-if the input is indeed feasible, these algorithms do indeed output the
-filtered ICPPL. Now we present the algorithmic characterization of a
-feasible tree path labeling by way of
-Algorithm~\ref{al:icppl-find-isomorph}.
+Algorithm~\ref{al:filterone} for {\tt filter\_1} and
+Algorithm~\ref{al:filtertwo} for {\tt filter\_2} which when executed
+in that order result in a new ICPPL on the same universe $U$ and
+target tree $T$. We also proved that if the input is indeed feasible,
+these algorithms executed in sequence give an ICPPL that preserves \change{the
+  same set}{a subset} of hypergraph isomorphisms as that of the input
+ICPPL. Now we present the algorithmic characterization of a feasible
+tree path labeling by way of Algorithm~\ref{al:icppl-find-isomorph}.
+
+\temptext{Prove that the same set of hypergraph isomorphisms are in
+  the filtered ICPPL? -- NOPE. we are only interested in testing
+  feasibility. not necessarily giving ALL bijections.}
+\begin{lemma}{\temptext{PROOF TBD.}}
+  Let $(\cF_2, \cl_2)$ be the output of {\tt filter\_2} and the ICPPL
+  input to {\tt filter\_1} preceeding its call is $(\cF, \cl)$ with
+  target tree $T$. If a bijection $\phi:\supt{\cF_2} \rightarrow V(T)$
+  is a hypergraph isomorphism of $(\cF_2, \cl_2)$, then it is a
+  hypergraph isomorphism of $(\cF, \cl)$.
+\end{lemma}
 
 Algorithm~\ref{al:icppl-find-isomorph} computes a hypergraph
-isomorphism $\phi$ recursively using Algorithm~\ref{perms} and
-Algorithm~\ref{leafasgn} and pruning the leaves of the input tree. In
-brief, it is done as follows. Algorithm~\ref{leafasgn} gives us the
-leaf labels in $\cF_2$, i.e., the elements in $supp(\cF)$ that map to
-leaves in $T$, where $(\cF_2, \cl_2)$ is the output of
-Algorithm~\ref{leafasgn}. All leaves in $T$ are then pruned away. The
-leaf labels are removed from the path labeling $\cl_2$ and the
-corresponding elements are removed from the corresponding sets in
-$\cF_2$. This tree pruning algorithm is recursively called on the
-altered hypergraph $\cF'$, path label $\cl'$ and tree $T'$. The
-recursive call returns the bijection $\phi''$ for the rest of the
-elements in $supp(\cF)$ which along with the leaf labels $\phi'$ gives
-us the hypergraph isomorphism $\phi$.  The following lemma formalizes
-the characeterization of feasible path labeling.
+isomorphism $\phi$ recursively using Algorithm~\ref{al:filterone} and
+Algorithm~\ref{al:filtertwo} and pruning the leaves of the input
+tree. In brief, it is done as follows. If the input is feasible, the
+output of {\tt filter\_1}, say $(\cF_1, \cl_1)$, gives a new ICPPL
+which preserves all the original hypergraph isomorphisms.  This ICPPL
+$(\cF_1, \cl_1)$ is then given to {\tt filter\_2} which gives the
+second filtered ICPPL, say $(\cF_2, \cl_2)$ such that a feasible leaf
+labeling is in singleton elements of $\cF_2$. These leaf labels are
+the elements in $supp(\cF)$ that map to leaves in $T$ in at least one
+hypergraph isomorphism of the input ICPPL $\cl$. It must be noted that
+at this point, the algorithm ``chooses'' one (or more - there could be
+more than one bijections that have the same leaf labels) hypergraph
+isomorphism $\phi'$, from the set of all isomorphisms resulting from the
+feasibility of the original input $(\cF, \cl)$. However, if there
+exists one (\ie input is feasible), it will be found -- thus
+performing the test required in \FTPL.
+
+Since their pre-images have been found, all leaves in $T$ are then
+pruned away. The leaf labels are removed from the path labeling
+$\cl_2$ and the corresponding elements are removed from the
+corresponding sets in $\cF_2$. It is now clear that we have a
+subproblem with a new hypergraph $\cF'$, new tree path labeling $\cl'$
+and target tree $T'$.  The tree pruning algorithm is recursively
+called on $\cF', \cl', T'$. The recursive call returns the bijection
+$\phi''$ for the rest of the elements in $supp(\cF)$ which along with
+the leaf labels $\phi'$ computed earlier gives us a hypergraph
+isomorphism $\phi$ for the input $\cF,\cl,T$.  The following lemma formalizes the
+characterization of feasible path labeling.
 
 \begin{algorithm}[h]
-  \caption{{\tt get- hypergraph- isomorphism ($\cF, \cl, T$)}}
+  \caption{{\tt get-hypergraph-isomorphism ($\cF, \cl, T$)}}
   \label{al:icppl-find-isomorph}
@@
   If $(\cF, \cl)$ is an ICPPL from a tree $T$ and
-  Algorithm~\ref{al:icppl-find-isomorph}, {\tt get- hypergraph-
-    isomorphism ($\cF, \cl, T$)} returns a non-empty function, then
-  there exists a hypergraph isomorphism $\phi : supp(\cF) \rightarrow
-  V(T)$ such that the $\phi$-induced tree path labeling is equal to
-  $\cl$ or $\cl_\phi = \cl$.
+  Algorithm~\ref{al:icppl-find-isomorph}, on input $\cF, \cl, T$
+  returns a non-empty function, then there exists a hypergraph
+  isomorphism $\phi : supp(\cF) \rightarrow V(T)$ such that the
+  $\phi$-induced tree path labeling is equal to $\cl$ or $\cl_\phi =
+  \cl$.
 \end{lemma}
-\begin{proof}
+\begin{proof}\thesisspacing
   It is clear that in the end of every recursive call to
-  Algorithm~\ref{al:icppl-find-isomorph}, the function $\phi'$ is one
-  to one involving all the leaves in the tree passed as input to that
-  recursive call. Moreover, by Lemma~\ref{lem:noexit1} and
-  Lemma~\ref{lem:xnotempty} it is consistent with the tree path
-  labeling $\cl$ passed. The tree pruning is done by only removing
-  leaves in each call to the function and is done till the tree
-  becomes empty. Thus the returned function $\phi: supp(\cF)
-  \rightarrow V(T)$ is a union of mutually exclusive one to one
-  functions exhausting all vertices of the tree. In other words, it is
-  a bijection from $supp(\cF)$ to $V(T)$ inducing the given path
-  labeling $\cl$ and thus a hypergraph isomorphism. %\qed
+  Algorithm~\ref{al:icppl-find-isomorph}, the function $\phi'$ is
+  one-to-one involving all the leaves in the input target tree $T$ (of
+  the current recursive call). Moreover, by Lemma~\ref{lem:noexit1}
+  and Lemma~\ref{lem:xnotempty} it is consistent with the input tree
+  path labeling $\cl$ (of the current recursive call). The tree
+  pruning is done by only removing leaves in each call to the function
+  and is done till the tree becomes empty. Thus the returned function
+  $\phi: supp(\cF) \rightarrow V(T)$ is a union of mutually exclusive
+  one-to-one functions exhausting all vertices of the tree. In other
+  words, it is a bijection from $supp(\cF)$ to $V(T)$ inducing the
+  given path labeling $\cl$ and thus a hypergraph isomorphism. %\qed
 \end{proof}
@@
   \label{th:charac}
-  A path labeling $(\cF, \cl)$ on tree $T$ is feasible iff it is an
+  A path labeling $(\cF, \cl)$ on tree $T$ is feasible \iff it is an
   ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
@@
 \end{theorem}
-\begin{proof}
+\begin{proof}\thesisspacing
   From Lemma~\ref{lem:hyperiso}, we know that if $(\cF, \cl)$ is an
-  ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
-  T)$ as input returns a non-empty function, $(\cF, \cl)$ is feasible.
-  Now consider the case where $(\cF, \cl)$ is feasible. i.e. there
-  exists a hypergraph isomorphism $\phi$ such that $\cl_\phi =
-  \cl$. Lemma~\ref{lem:noexit1} and Lemma~\ref{lem:xnotempty} show us
-  that filter 1 and filter 2 do not exit if input is feasible. Thus
+  ICPPL and if Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
+  T)$ as input returns a non-empty function, then $(\cF, \cl)$ is
+  feasible.  Now consider the case where $(\cF, \cl)$ is
+  feasible, \ie there exists a hypergraph isomorphism $\phi$ such
+  that $\cl_\phi = \cl$. Lemma~\ref{lem:noexit1} and
+  Lemma~\ref{lem:xnotempty} show us that {\tt filter\_1} and {\tt
+    filter\_2} do not exit if input is feasible. Thus
   Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
@@
 
-\section{ Computing feasible TPL with special target trees \tnote{give
-    problem definition etc}}
+\section{ Computing feasible TPL with special target trees}
 \label{sec:spltargettree}
 
+\tnote{\sc refer to Problem Defs here}
 Section~\ref{sec:feasible} described properties that a TPL must have
@@
 
-Consider a special case of ICPPL with the following properties when the tree $T$
-is a path.  Hence, all path labels are can be viewed as intervals assigned to the
-sets in $\cF$.  It is shown, in \cite{nsnrs09}, that the filtering algorithms outlined
-above need only preserve pairwise intersection cardinalities, and higher level
-intersection cardinalities are preserved by the Helly Property of intervals.  Consequently,
-the filter algorithms do not need to ever evaluate the additional check to {\em \bf exit}.
+Let us consider a special case of ICPPL with the following properties.
+The target tree $T$ is a path.  Hence, all path labels are can be viewed as
+intervals assigned to the sets in $\cF$.  It is shown, in
+\cite{nsnrs09}, that the filtering algorithms outlined above need only
+preserve pairwise intersection cardinalities, and higher level
+intersection cardinalities are preserved by the Helly Property of
+intervals.  Consequently, the filter algorithms do not need to ever
+evaluate the additional check to {\em \bf exit}.
 %\begin{enumerate}
@@
 %Assignment (ICPIA) \cite{nsnrs09}. 
-This structure and its algorithm is
-used in the next section for finding tree path labeling from a
-$k$-subdivided star due to this graph's close relationship with
-intervals.
+This structure and its algorithm is used in the next section for
+finding tree path labeling from a $k$-subdivided star due to this
+graph's close relationship with intervals.
+
+\tnote[dire]{ELABORATE}
 
@@
 \end{lemma}
-\begin{proof}
+\begin{proof}\thesisspacing
   Suppose $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
@@
 \end{lemma}
-\begin{proof}
+\begin{proof}\thesisspacing
   The proof of this lemma is by contradiction.  Let $X$ be a
@@
 \end{theorem}
-\begin{proof} {\em Outline.}
+\begin{proof}\thesisspacing {\em Outline.}
 %   If the algorithm outputs a path labeling $\cl$, then it is clear
@@
 %   Algorithm~\ref{ds-algo} outputs an ICPIA to a prime matrix $M'$
-%   iff there is an ICPIA for $M'$.
+%   \iff there is an ICPIA for $M'$.
 % \end{theorem}
@@
 $\cF$, $|U| = n, |\cF| = m$. Thus $(i,j)$th element of $M$, $M_{ij} =
-1$ iff $x_i \in S_j$. If $\cF$ has a feasible tree path assignment
+1$ \iff $x_i \in S_j$. If $\cF$ has a feasible tree path assignment
 (ICPPA) $\cA = \{(S_i,P_i) \mid i \in [m]\}$, then we say its
@@
 correspond to the sets in $\cF$.  An edge is present between vertices
-of two sets iff the corresponding sets have a nonempty intersection
+of two sets \iff the corresponding sets have a nonempty intersection
 and none is contained in the other. Formally, intersection graph is
@@
 \end{theorem}
-\begin{proof}
+\begin{proof}\thesisspacing
   To observe that $X$ is a collection of in-trees, we observe that for
@@
   let $\{X_1,\ldots,X_r\}$ be the partition of $X$ into in-trees as
-  defined above.  Then, matrix $M$ has an ICPPA in tree $T$ iff $T$
+  defined above.  Then, matrix $M$ has an ICPPA in tree $T$ \iff $T$
   can be partitioned into vertex disjoint subtrees $\{T_1, T_2, \dots
@@
 \end{lemma}
-\begin{proof}
+\begin{proof}\thesisspacing
   Let us consider the reverse direction first.  Let us assume that $T$
@@
 \end{lemma}
-\begin{proof}
+\begin{proof}\thesisspacing
   The proof is based on the following fact- $\preccurlyeq$ is a
@@
 trees, which represent all COP orderings of $M$, invented by
-\cite{bl76}. There is a PQ tree for a matrix iff the matrix has COP.
+\cite{bl76}. There is a PQ tree for a matrix \iff the matrix has COP.
 Indeed, this is an algorithmic characterization of the consecutive
@@
 % representing maximal cliques and columns representing vertices of a
-% graph) has COP on columns iff the graph is an interval
+% graph) has COP on columns \iff the graph is an interval
 % graph\cite{fg65}. This follows from the interval graph
@@
 % \end{lemma}
-% \begin{proof}
+% \begin{proof}\thesisspacing
 %   {\em Case 1:} w.l.o.g, consider $Q_3 \cap Q_4$ and let us call it
@@
 % and finally we have a family of bijections.  First we present and
-% then prove the correctness of Algorithm \ref{perms}.  This algorithm
+% then prove the correctness of Algorithm \ref{al:filterone}.  This algorithm
 % refines the path assignment by considering pairs of paths that share
@@
 %   \caption{Permutations from an ICPPA $\{(S_i,P_i) | i \in I\}$}
-%   \label{perms}
+%   \label{al:filterone}
 %   \begin{algorithmic}
@@
 %   \label{lem:invar1}
-%   In Algorithm \ref{perms}, at the end of $j$th iteration, $j \ge
-%   0$, of the while loop of Algorithm \ref{perms}, the following
+%   In Algorithm \ref{al:filterone}, at the end of $j$th iteration, $j \ge
+%   0$, of the while loop of Algorithm \ref{al:filterone}, the following
 %   invariants are maintained.
@@
 %   sets added in $j$th iteration as defined in line \ref{setbreak} of
-%   Algorithm \ref{perms}, i.e. the following three assignment pairs
+%   Algorithm \ref{al:filterone}, i.e. the following three assignment pairs
 %   for some $(P_1,Q_1), (P_2,Q_2) \in \Pi_{j-1}$ where $Q_1$ and
@@
 % \noindent
-% It can be observed that the output of algorithm \ref{perms} is such
+% It can be observed that the output of algorithm \ref{al:filterone} is such
 % that every leaf is incident on at most a single path in the new set
@@
 %   \caption{Leaf assignments from an ICPPA $\{(S_i,P_i) | i \in I\}$}
-%   \label{leafasgn}
+%   \label{al:filtertwo}
 %   \begin{algorithmic}
@@
 %   \label{lem:invar3}
-%   In Algorithm \ref{leafasgn},for all $j \geq 0$, at the end of the
+%   In Algorithm \ref{al:filtertwo},for all $j \geq 0$, at the end of the
 %   $j$th iteration the four invariants given in lemma
@@
 %   For $\Pi_0$ all invariants hold because it is output from
-%   algorithm \ref{perms} which is an ICPPA. Hence base case is
+%   algorithm \ref{al:filterone} which is an ICPPA. Hence base case is
 %   proved.  Assume the lemma holds for $\Pi_{j-1}$. Consider $\Pi_j$
@@
 
-% Using algorithms \ref{perms} and \ref{leafasgn} we prove the
+% Using algorithms \ref{al:filterone} and \ref{al:filtertwo} we prove the
 % following theorem.
@@
 %   This is a contructive proof. First, the given ICPPA $\cA$ and tree
-%   $T$ are given as input to Algorithm \ref{perms}. This yields a
+%   $T$ are given as input to Algorithm \ref{al:filterone}. This yields a
 %   ``filtered'' ICPPA as the output which is input to Algorithm
-%   \ref{leafasgn}.  It can be observed that the output of Algorithm
-%   \ref{leafasgn} is a set of interval assignments to sets and
+%   \ref{al:filtertwo}.  It can be observed that the output of Algorithm
+%   \ref{al:filtertwo} is a set of interval assignments to sets and
 %   one-to-one assignment of elements of $U$ to each leaf of $T$. To
@@
 %   path assignment $\cA_0$ and tree $T_0$ is given as input to
-%   Algorithm \ref{perms}. The output of this algorithm is given to
-%   Algorithm \ref{leafasgn} to get a new union of path and leaf
+%   Algorithm \ref{al:filterone}. The output of this algorithm is given to
+%   Algorithm \ref{al:filtertwo} to get a new union of path and leaf
 %   assignments $\cB_1 = \cA_1 \cup \cL_1$ defined similar to $\cB_0,
@@
 %   \noindent
-%   To summarize, run algorithm \ref{perms} and \ref{leafasgn} on
+%   To summarize, run algorithm \ref{al:filterone} and \ref{al:filtertwo} on
 %   $T$. After the leaves have been assigned to specific elements from

### 2012-01-14 src/050appa.tex
Started cleaning up chapter 3. Deferring it though to work on Chapter 2. Need to accommodate N's new proofs for removal of algorithmic charac.
See diff for other changes left out in this comment.

\chapter{More proofs \tnote{remove/modify}}
\begin{lemma}
  \label{lem:fourpaths} Consider four paths in a tree $Q_1, Q_2, Q_3,
  Q_4$ such that they have nonempty pairwise intersection and $Q_1,
  Q_2$ share a leaf. Then there exists distinct $i, j, k \in
  \{1,2,3,4\}$ such that, $Q_1 \cap Q_2 \cap Q_3 \cap Q_4 = Q_i \cap
  Q_j \cap Q_k$.
\end{lemma}
\begin{proof}\thesisspacing
  {\em Case 1:} w.l.o.g, consider $Q_3 \cap Q_4$ and let us call it
  $Q$. This is clearly a path (intersection of two paths is a path).
  % Since $Q_1, Q_2$ share a leaf, the following are paths $Q_1
  % \setminus Q_2$, $Q_2 \setminus Q_1$, $Q_1 \cap Q_2$ and they are
  % mutually disjoint.
  Suppose $Q$ does not intersect with $Q_1 \setminus Q_2$, i.e. $Q
  \cap (Q_1 \setminus Q_2) = \O$. Then $Q \cap Q_1 \cap Q_2 = Q \cap
  Q_2$. Similarly, if $Q \cap (Q_2 \setminus Q_1) = \O$, $Q \cap Q_1
  \cap Q_2 = Q \cap Q_1$. Thus it is clear that if the intersection of
  any two paths does not intersect with any of the set differences of
  the remaining two paths, the claim in the lemma is true.
  % Note that $Q_1 \setminus Q_2$ and $Q_2
  % \setminus Q_1$ are paths because $Q_1, Q_2$ share a leaf.\\
  {\em Case 2:} Let us consider the compliment of the previous
  case. i.e. the intersection of any two paths intersects with both
  the set differences of the other two. First let us consider $Q \cap
  (Q_1 \setminus Q_2) \ne \O$ and $Q \cap (Q_1 \setminus Q_2) \ne \O$,
  where $Q = Q_3 \cap Q_4$. Since $Q_1$ and $Q_2$ share a leaf, there
  is exactly one vertex at which they branch off from the path $Q_1
  \cap Q_2$ into two paths $Q_1 \setminus Q_2$ and $Q_2 \setminus
  Q_1$. Let this vertex be $v$. It is clear that if path $Q_3 \cap
  Q_4$, must intersect with paths $Q_1 \setminus Q_2$ and $Q_2
  \setminus Q_1$, it must contain $v$ since these are paths from a
  tree. Moreover, $Q_3 \cap Q_4$ intersects with $Q_1 \cap Q_2$ at
  exactly $v$ and only at $v$ which means that $Q_1 \cap Q_2$ does not
  intersect with $Q_3 \setminus Q_4$ or $Q_4 \setminus Q_3$ which
  contradicts initial condition of this case. Thus this
  case cannot occur and case 1 is the only possible scenario. \\
  Thus lemma is proven %\qed
\end{proof}

### 2012-01-16 src/020chap2.tex
Fixed a bug in chapter 2 (from last commit). Endnotes in chap 2 changed to footnotes. Tucker matrices coded. Tikz is going to be the way of the graphics in Infinite Jest.

@@
 
-% \begin{definition}{\emph{Matrices that define
-%       graphs.\cite[Def.~2.4]{d08phd}}} %
 \begin{definition}
@@
     $G$ is an interval graph \iff its maximal clique matrix has COP
-    \cite{fg65}\endnote{This follows \cite{gh64} which states that the
+    \cite{fg65}\footnote{This follows \cite{gh64} which states that the
       maximal cliques of interval graph $G$ can be linearly ordered
@@
     \item $G$ is a \emph{unit interval graph} if it is an interval
-      graph \stt all intervals have the same length.\endnotemark[2]
+      graph \stt all intervals have the same length.\footnotemark[2]
     \item $G$ is a \emph{proper interval graph} if it is an interval
-      graph \stt no interval properly contains another.\endnote{The
+      graph \stt no interval properly contains another.\footnote{The
         set of unit interval graphs and the set of proper interval
-        graphs are the same} \tnote{cite {rob69,gar07} in endnote. pg
+        graphs are the same} \tnote{cite {rob69,gar07} in footnote. pg
         33 dom}
@@
 
-Interval graphs\endnote { \cite{mcc04} cites that the problem of
+Interval graphs\footnote{\cite{mcc04} cites that the problem of
   recognizing interval graphs has significance in molecular
@@
   inside a structure that had a linear topology which we now know to
-  be true\tnote{linear arrangement of DNA/genes?}.} and circular-arc
-graphs have a long history in research.  The interest around them is
-due to their very desirable property that several problems that are
-NP-hard \tnote[minor]{all NPC problems are NPH yes?} on general
-graphs, like finding a maximum clique or minimum coloring or
-independent set, are polynomial time solvable in these graph classes
-\cite{clrs01}.
-In a similar fashion, a lot of problems that are hard on general
-matrices have efficient solutions on matrices with \COP or \CROP
-\cite[more citations pg.\,33]{d08phd}.
+  be true from the discovery of linear structure of DNA.}and
+circular-arc graphs have a long history in research.  The interest
+around them is due to their very desirable property that several
+problems that are NP-hard \tnote[minor]{all NPC problems are NPH yes?}
+on general graphs, like finding a maximum clique or minimum coloring
+or independent set, are polynomial time solvable in these graph
+classes \cite{clrs01}.  In a similar fashion, a lot of problems that
+are hard on general matrices have efficient solutions on matrices with
+\COP or \CROP \cite[more citations pg.\,33]{d08phd}.
 
@@
 having seen how useful \COP is in identifying or characterizing many
-types of graphs\endnote{\COP has several other applications
-  \tnote{fill up.}}, we will now see results that study recognition of \COP
+types of graphs\footnote{\COP has several other applications
+  \temptext{fill up.}}, we will now see results that study recognition of \COP
 in matrices in the following section.
@@
   \centering
+% \includegraphics[scale=0.1]{../img/circlediagram_domtab2.1__2.png} %
  \caption{\figtabsize Graph matrices \temptext{PLACEHOLDER}}
@@
 \end{table}
-
-%\clearpage      % Print pending floats
 
@@
 is by deriving one from any interval graph recognition algorithm using the
-result {\sc hmpv00}\tnote{cite hmpv00} \tnote{in endnote put the theorem 2.7 - dom
+result {\sc hmpv00}\tnote{cite hmpv00} \tnote{in footnote put the theorem 2.7 - dom
   pg 43} \cite{d08phd} which demonstrates how such a derivation can be
@@
   {
-    \fontencoding{OT1}\fontfamily     {cmss}\selectfont
-    % \sffamily
-    \footnotesize
-    % \includegraphics[scale=0.5]{../img/cophistory.png} % TEMPORARY
+    \fontencoding{OT1}\fontfamily{cmss}\selectfont%
+    \footnotesize%
 
@@
       \hline
-     % 
+
       1899 & \parbox[t]{\tabparboxw}{ First mention of \COP
-       (archaeology)} & 
-       \\ % \parbox[t]{2cm}{}
+        (archaeology)} &                  \\
 
@@
       1992 & \parbox[t]{\tabparboxw}{ Linear
-       time algorithm COT without \PQtree} & \cite{wlh02}\endnote{First
-       published in \cite{wlh02isaac92}} \\
+        time algorithm COT without \PQtree} & \cite{wlh02}%\footnote{First
+      % published in \cite{wlh02isaac92}}
+      \\
 
@@
 this order. Clearly, this means that the maximal clique incidence
-matrix\endnote{
+matrix\footnote{
   Definition~\ref{def:graphmatrices}~\ref{def::maxcliquematrix}} must
@@
 
-%\clearpage
 \subsection{Tucker's forbidden submatrices for \COP}
@@
 \cite{at72} discovered certain forbidden structures for convex
-bipartite graphs\endnote{The terminology in \cite{at72} differs. It
+bipartite graphs\footnote{The terminology in \cite{at72} differs. It
   uses the term {\em graphs with $V_1$-consecutive arragement} instead
@@
 Theorem~\ref{th:tuckeratfree} states that convex bipartite graphs
-cannot have {\em asteroidal triples}\endnote{If $G = (V,E)$ is a
+cannot have {\em asteroidal triples}\footnote{If $G = (V,E)$ is a
   graph, a set of three vertices from $V$ form an {\em asteroidal
@@
   third vertex.} contained in the corresponding vertex
-partition\endnote{The partition corresponds to columns (rows) if its
+partition\footnote{The partition corresponds to columns (rows) if its
   half adjacency matrix has \COP columns (rows).}.
@@
   A bipartite graph $G = (V_1, V_2, E)$ is convex bipartite on
-  columns\endnote{Abridged to match terminology adopted in this
+  columns\footnote{Abridged to match terminology adopted in this
     document. See previous note.} \iff $V_1$ contains no asteroidal
@@
   \centering
-  \includegraphics[scale=0.35]{../img/tuckersforbiddensubmatrices.png}  
+%  \includegraphics[scale=0.35]{../img/tuckersforbiddensubmatrices.png}  
+  \setlength\dashlinedash{1pt} %
+  \setlength\dashlinegap{0pt}  %  Make non zero if cline can be done
+                               %   with dashed lines.
+  \singlespacing
+  \begin{tabular}[h]{ll}
+%   \small 
+   \begin{tabular}[t]{|lll ll l|l} %                   MATRIX 1
+      \multicolumn{7}{l}{\normalsize \hspace{-0.5in} $M_{I_k}, k \ge 1$}              \\   
+      \multicolumn{6}{c}{\matrixbracetop{3.25}{$k+2$}}   \\
+      \cline{1-6} % \firsthline %      Begin matrix  %%%%%%
+      \un & \un & 0   & \multicolumn{2}{c}{\dots}  & 0    &
+        \hspace{-0.1in}\multirow{5}{*}{\matrixbraceright{2.5}{$k+2$}}   \\    
+      0   & \un & \un & 0 & ...& 0   \\
+      \multicolumn{6}{|c|}{\dots}                        \\
+      0   & \multicolumn{2}{c}{\dots}  & 0   & \un & \un \\
+      \cline{1-6} % \hdashline %%%%%%%%%%%%%%%%%%%%%%%%%%%%
+      \un & 0   & \multicolumn{2}{c}{\dots}  & 0   & \un \\
+      \cline{1-6} % \lasthline %      End matrix  %%%%%%%%%
+    \end{tabular}
+    &%
+%    \small
+    \begin{tabular}[t]{|lll ll ll|l} %                   MATRIX 2
+      \multicolumn{8}{l}{\normalsize \hspace{-0.5in}$M_{II_k}, k \ge 1$}                   \\
+      \multicolumn{7}{c}{\matrixbracetop{3.8}{$k+3$}}            \\ 
+      \cline{1-7}%  \firsthline %      Begin matrix  %%%%%%%%%%%%
+      \un & \un & 0   & \multicolumn{3}{c:}{\dots}       & 0    &
+        \hspace{-0.1in}\multirow{6}{*}{\matrixbraceright{3}{$k+3$}}         \\
+      0   & \un & \un & 0 & \multicolumn{2}{c:}{\dots}   & 0   \\ 
+      \multicolumn{6}{|c:}{\dots}                        &     \\
+      0   & \multicolumn{2}{c}{\dots}  & 0   & \un              &
+        \multicolumn{1}{l:}{\un} & 0                           \\
+      \cline{1-6}    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+      0   & \un & \multicolumn{4}{c}{\dots}              & \un \\
+      \un & \multicolumn{3}{c}{\dots}        & \un & 0   & \un \\
+      \cline{1-7}% \lasthline %      End matrix  %%%%%%%%%%%%%%%%
+    \end{tabular}
+    \\\\%
+%    \small
+    \begin{tabular}[t]{|lll ll l l|l} %                  MATRIX 3 
+      \multicolumn{8}{l}{\normalsize \hspace{-0.5in}$M_{III_k}, k \ge 1$}                  \\
+      \multicolumn{7}{c}{\matrixbracetop{3.8}{$k+3$}}            \\
+      \cline{1-7}%      \firsthline %      Begin matrix %%%%%%%%%
+      \un & \un & 0   & \multicolumn{3}{c:}{\dots}       & 0    &
+        \hspace{-0.1in}\multirow{5}{*}{\matrixbraceright{2.5}{$k+2$}}         \\
+      0   & \un & \un & 0 & \multicolumn{2}{c:}{\dots}   & 0   \\ 
+      \multicolumn{6}{|c:}{\dots}                        &     \\
+      0   & \multicolumn{2}{c}{\dots}  & 0   & \un              &
+        \multicolumn{1}{l:}{\un}                         & 0   \\
+      \cline{1-6}    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+      0   & \un & \multicolumn{2}{c}{\dots}  & \un & 0   & \un \\
+      \cline{1-7}%      \lasthline %      End matrix %%%%%%%%%%%%
+    \end{tabular}
+    &%
+%    \small
+    \begin{tabular}[t]{|llllll|}  %                     MATRIX 4
+      \multicolumn{6}{l}{\normalsize \hspace{-0.5in}$M_{IV}$} \\
+      \multicolumn{6}{l}{} \\
+      \firsthline
+      \un & \un & 0 & 0 & 0 & 0   \\
+      0 & 0 & \un & \un & 0 & 0   \\
+      0 & 0 & 0 & 0 & \un & \un   \\
+      \un & 0 & \un & 0 & \un & 0 \\
+      \lasthline
+    \end{tabular}
+    \\\\%
+%    \small
+    \begin{tabular}[t]{|lllll|}  %                     MATRIX 5
+      \multicolumn{5}{l}{\normalsize \hspace{-0.5in}$M_{V}$} \\
+      \multicolumn{5}{l}{} \\
+      \firsthline
+      \un & \un & 0   & 0   & 0   \\
+      0   & 0   & \un & \un & 0   \\
+      \un & \un & \un & \un & 0   \\
+      \un & 0   & \un & 0   & \un \\
+      \lasthline
+    \end{tabular}
+  \end{tabular}
+
   \caption[\figtabsize Tucker's forbidden submatrices]{\figtabsize Tucker's
-    forbidden submatrices for convex bipartite graphs. \temptext{ PLACEHOLDER
-    IMG }} 
+    forbidden submatrices for convex bipartite graphs. \cite{at72}} 
   \label{fig:forbiddensubmatrices}
@@
 
-%\clearpage
 \subsection{Booth and Lueker's $PQ$ tree -- a linear COT algorithm}
@@
 Booth and Lueker in their paper \cite{bl76} gave the first linear
-algorithm\endnote{Time complexity is $O\left(m+n+f\right)$ where $m
-  \times n$ is the order of the input matrix and $f$ is the number of $\un$s in it.} for \cop testing while given a linear time interval graph
-recognition algorithm by a simplification of \tnote{cite - A. Lempel,
-  S. Even and I. Cederbaum, An Algorithm for Planarity Testing of
-  Graphs, Theory of Graphs, ed., P. Rosenstiehl, Gordon and Breach,
-  New York, (1967), 215-232.  }'s planarity test algorithm.
+algorithm\footnote{Time complexity is $O\left(m+n+f\right)$ where $m
+  \times n$ is the order of the input matrix and $f$ is the number of
+  $\un$s in it.} for \cop testing while given a linear time interval
+graph recognition algorithm by a simplification of \tnote{cite -
+  A. Lempel, S. Even and I. Cederbaum, An Algorithm for Planarity
+  Testing of Graphs, Theory of Graphs, ed., P. Rosenstiehl, Gordon and
+  Breach, New York, (1967), 215-232.  }'s planarity test algorithm.
 % Hsu in wlh01 said BL76 gave a planarity test - "the induced PQ-tree
@@
 
+%\begin{minipage}[h]{5.5in}
 \begin{definition}[\PQtree \cite{bl76, mcc04}]
-  A \PQtree of matrix $M$ with \COP on columns (rows), is a tree with the following properties.
+  A \PQtree of matrix $M$ with \COP on columns (rows), is a tree with
+  the following properties.
   \begin{enumerate}[i.]
@@
   \item Each leaf uniquely represents a row (column) of $M$. The leaf
-    order of the tree gives a \COP order for column (row)\endnote{Note
+    order of the tree gives a \COP order for column (row)\footnote{Note
       that \COP order for column requires permutation of rows and vice
@@
 \end{definition}
+%\end{minipage}
 
@@
 
-%\section{Certifying Algorithms}
 \subsection{\PQRtree\ --  COP for set systems}
@@
 in Definition~\ref{def:matrixsetsystem} along with the idea of \cop
-for set systems\endnote{As seen in Section~\ref{sec:courseschedule}}.
+for set systems\footnote{As seen in Section~\ref{sec:courseschedule}}.
 
 \begin{definition}%
-  \label{def:matrixsetsystem}%
+  \label{def:matrixsetsystem}{\emph{\Cop for set systems}}%
   % [\emph{Set system of a binary matrix}.]
-  [\emph{\Cop for set systems}.]  Let $M$ be a binary matrix of order
+   Let $M$ be a binary matrix of order
   $n \times m$ % , $r_i, i \in [n]$ be the rows in $M$
@@
 generalizations of \cite{bl76}'s \PQtree and hence produce the
-$\PQtree$ if $\cF$ has \COP. As a whole, all these three data
+\PQtree if $\cF$ has \COP. As a whole, all these three data
 strutures are near-identical with their differences being notional.
@@
 Also note that trivially, sets in $\cT(U)$ are consecutive in any
-permutation of $U$\endnote{$\emptyset$ is considered consecutive by
+permutation of $U$\footnote{$\emptyset$ is considered consecutive by
   convention.}.
@@
 
-\tnote[FYI]{ The results in cite~{bl76} on COT are based on the
-   result that interval graphs are AT-free chordal graphs cite~{lb62}?
- No they don't seem related. this is the stupid bl lb confusion.}%
+% \tnote[FYI]{ The results in cite~{bl76} on COT are based on the
+%    result that interval graphs are AT-free chordal graphs cite~{lb62}?
+%  No they don't seem related. this is the stupid bl lb confusion.}%
 
-%\clearpage      % Print pending floats
-% \section{COP in Relational Database Model}
-% \label{sec:surveyrdbm}
-% \tnote{Expand on sec:apprdbm} \tnote{(set systems theme)}
 \section{Optimization problems in COP}
@@
 
-\theendnotes\tnote{remove if none.}
+%\theendnotes\tnote{remove if none.}
 

### 2012-01-16 src/030chap3.tex
Fixed a bug in chapter 2 (from last commit). Endnotes in chap 2 changed to footnotes. Tucker matrices coded. Tikz is going to be the way of the graphics in Infinite Jest.

@@
   $\cl: \cF \rightarrow \cP$ due to this isomorphism is called a {\em
-    path labeling} of the hypergraph $\cF$.\endnote{Note that there
+    path labeling} of the hypergraph $\cF$.\footnote{Note that there
     are two kinds of isomorphisms here.  One is the isomorphism of

### 2012-01-18 src/020chap2.tex
first draft of PQR trees done. next is cert algo and icpia.

@@
 
-\begin{definition}%
-  \label{def:matrixsetsystem}{\emph{\Cop for set systems}}%
+\begin{definition}[{\em \Cop for set systems}]%
+  \label{def:matrixsetsystem}%
   % [\emph{Set system of a binary matrix}.]
-   Let $M$ be a binary matrix of order
-  $n \times m$ % , $r_i, i \in [n]$ be the rows in $M$
-  and $\{c_i \mid i \in [m]\}$ be the columns in $M$. A set system
-  $\cF_{M} = \{S_i \mid S_i \subseteq [n], i \in [m]\}$ is defined
-  such that for every column $c_i$ of $M$, set $S_i = \{j \mid m_{ji}
-  = \un \}$. The collection $\cF_M$ is the {\em set system of binary
-    matrix} $M$.\\
-  A set system $\cF$ from universe $U = \{1, 2, \ldots, n\}$ has the
-  {\em \cop} if it is possible to assign a linear order, \ie a
-  permutation $\sigma = u_1u_2\ldots u_n$ to $U$ \stt each set $S \in
-  \cF$ appears as a consecutive subsequence $u_{i}u_{i+1}\ldots
-  u_{i+k-1}$ (also called an interval) of $\lambda$ in the linear order, where
-  positive integers $k = |S|$ and $i \le n+1-k$.
+  Let $M$ be a binary matrix of order $n \times m$ and $\{c_i \mid i
+  \in [m]\}$ be the columns in $M$.  A set system $\cF_{M} = \{S_i
+  \mid S_i \subseteq [n], i \in [m]\}$ is defined such that for every
+  column $c_i$ of $M$, set $S_i = \{j \mid m_{ji} = \un \}$. The
+  collection $\cF_M$ is the {\em set system of binary matrix} $M$. The
+  {\em binary matrix for set system} $\cF$ is conversely constructed
+  and denoted by $M^\cF$. Thus, $M^{\cF_M} = M$.
+
+  A set system $\cF$ from universe $U$, $|U| = n$ has the
+  {\em \cop} if there exists a linear order or 
+  permutation $\sigma = w_1w_2\ldots w_n$ that can be applied to $U$ \stt each set $S \in
+  \cF$ becomes a consecutive subsequence\footnote{Also termed an {\em interval}} $w_{i}w_{i+1}\ldots
+  w_{i+k-1}$ on $\sigma$ for some positive integer $i \le n+1-k$ where  $k = |S|$.
+
+  The set $\valid{U,\cF}$ respresents all \COP orders of $\cF$ in $U$.
+  \defnstop
 \end{definition}
@@
 
-Generalized \PQtree or \gPQtree is a data structure defined in \cite{n89} to
-represent all orthogonal sets of a set system $\cF$. {\em Orthogonal
-  sets} of $\cF$ are sets that do not overlap with any set in $\cF$. A
+Before proceeding to describe \PQRtree per se, we will see a few more
+terminologies that will make the subsequent discussion in this section
+simpler.
+
+\begin{definition}[\em Orthogonal sets \cite{n89, mm96,
+    mcc04}\footnote{\cite{mcc04} does not use the term ``mutually
+    orthogonal'' but refers to the same idea as ``sets that do not
+    overlap''. This terminology is also used in other literature like
+    \cite{nsnrs09,wlh02}.} ]
+  \label{def:orthosets}%
+  Let $\cF$ be a set system with universe $U$ and sets $A, B \in \cF$.
+  \begin{enumerate}
+  \item $A$ and $B$ are said to have a {\em trivial intersection} if
+    $A \cap B$ is one of the following.
+    \begin{enumerate}[i.]
+    \singlespacing
+    \item $\emptyset$
+    \item $A$
+    \item $B$\footnote{In other words, $A$ and $B$ are either disjoint
+        or one is the subset of the other.}
+    \end{enumerate}
+    % \vspace{-5mm}%
+
+  \item $A$ and $B$ are called {\em mutually orthogonal} or $A$ is {\em
+      orthogonal to} $B$ and vice versa, if they have a trivial
+    intersection.
+
+  \item {\em Trivial subsets} of a universe $U$, denoted by $\cT(U)$,
+    are sets that have trivial intersections with any set in
+    $\power{U}$. These sets are $U$, singleton sets in
+    $\power{U}$ and $\emptyset$ \cite{n89, mm96}.
+    Thus, $\cT(U) = \{U\} \  \bigcup \  \{ \{v\} \mid v \in U \} \  \bigcup \
+    \{\emptyset\}$.
+
+  \item {\em Orthogonal sets of a set system} $\cF$ with universe $U$
+    are subsets of $U$ that are orthogonal to all sets in $\cF$. The
+    set of all orthogonal sets to $\cF$ is denoted by
+    $\cF^{\bot}$\footnote{\cite[Def.~3.1]{mcc04} uses the term {\em non-overlapping
+    family} of $\cF$ and denotes it by $\cN(\cF)$. \label{mcc1}}.
+
+  \item $\cF$ is called {\em complete}\footnote{\cite{mcc04} calls
+      this a {\em weakly partative family}. \label{mcc2}} if the following hold
+    true for every pair of non-orthogonal\footnote{Or overlapping.} sets
+    $A, B$ in $\cF$.
+    \begin{enumerate}[i.]
+    \item $\cT(U) \subset \cF$
+    \item $A \cup B \in \cF$
+    \item $A \cap B \in \cF$
+    \item $A \setminus B \in \cF$
+    \item $B \setminus A \in \cF$
+    \end{enumerate}
+    In other words, $\cF$ contains all the trivial
+    subsets of $U$, $A \cup B$ and the partitions of $A \cup B$
+    defined by intersection and set difference.
+
+  \item $\overline{\cF}$ represents the smallest super set system of
+    $\cF$ that is complete\footnote{\cite[Def.~3.2]{mcc04} calls
+      this the {\em weak closure} of $\cF$ denoted by $\cW(\cF)$. \label{mcc3}}
+
+  \end{enumerate}
+  \defnstop
+\end{definition}
+
+Generalized \PQtree or \gPQtree is a data structure defined in
+\cite{n89} to represent all orthogonal sets of a set system $\cF$. A
 data structure with the same name was later defined in \cite{mcc04} as
-part of a {\em substitution decomposition} for a set system
-$\cF$. \tnote{define substitutio decomp.} This is later used to define
-a certificate if $\cF$ has no \COP. \PQRtree is a data structure
-defined in \cite{mm96} as a data structure to represent any set system
-$\cF$ with additional information in their $R$-nodes if $\cF$ has no
-\COP.  All three of these data structures are proposed as
-generalizations of \cite{bl76}'s \PQtree and hence produce the
-\PQtree if $\cF$ has \COP. As a whole, all these three data
-strutures are near-identical with their differences being notional.
-In this section, we will discuss the basic theory that they all hold
-and stick to the name of the data structure as \PQRtree.
+part of a {\em substitution decomposition} for a set system $\cF$
+\tnote{define substitutio decomp.} and subsequently \cite{mcc04} gives
+a new characterization of $\cF$ using a so-called incompatibility
+graph. \PQRtree is defined in \cite{mm96} as a data
+structure to represent any set system $\cF$ with additional
+information in their $R$-nodes if $\cF$ has no \COP.  All three of
+these data structures are proposed as generalizations of \cite{bl76}'s
+\PQtree and hence produce the \PQtree if $\cF$ has \COP. As a whole,
+all these three data strutures are largely identical with their
+differences being notional.  In this section, we will discuss the
+basic theory that they all hold. We will predominantly use the
+terminology from \cite{mm96} and refer to the data structure as
+\PQRtree.
 
-A \PQRtree of a given set system $\cF$ from universe $U$ represents
-all the sets from $\power{U}$ that have trivial intersections with
-every set in $\cF$.  Two sets $A$ and $B$ are said to have a {\em
-  trivial intersection} if $A \cap B$ is one of the following --
-$\emptyset$, $A$ or $B$. In other words, $A$ and $B$ are either (i)
-disjoint or (ii) one is the subset of the other. Such sets are also
-called {\em orthogonal sets} or said that $A$ is orthogonal to
-$B$. {\em Trivial subsets} of a universe $U$, denoted by $\cT(U)$, are
-sets that have trivial intersections with any set in
-$\power{U}$. Namely, $\cT(U)$ consists of $U$, singleton sets in
-$\power{U}$ and $\emptyset$\cite{n89, mm96}.  In other words, the all
-sets presented in a \PQRtree share the property that they do not
-overlap with any set in $\cF$. We will now see how this helps in
-testing for \COP in $\cF$.
+An important observation made by \cite{mm96} is presented now along
+with a few theorems that help in decomposing the \COP problem on $\cF$
+into subproblems.
 
-An important observation made by \cite{mm96} is that if $\cF$ is a set
-system with \COP then, aside from every set in $\cF$ being consecutive
-(after applying the \COP order) the following sets must also be
-consecutive for any $A, B \in \cF$.
+\begin{observation}[{\cite[Sec.~3]{mm96}}]
+  If $\cF$ is a set system with \COP then, after applying the \COP
+  order, not only must every set in $\cF$ be consecutive but the
+  following sets must also be consecutive for any $A, B \in \cF$.
 \begin{enumerate}
@@
     if $B \nsubseteq A$ and $A \nsubseteq B$ respectively.
+  \item Also note that trivially, sets in $\cT(U)$ are consecutive in
+    any permutation of $U$\footnote{$\emptyset$ is considered
+      consecutive by convention.}.
 \end{enumerate}
-Also note that trivially, sets in $\cT(U)$ are consecutive in any
-permutation of $U$\footnote{$\emptyset$ is considered consecutive by
-  convention.}.
+\end{observation}
 
-Any set system that is closed under the above operations and contains
-$\cT(U)$ is called {\em complete} \cite{mm96} or a {\em weakly partative
-  family} \cite{mcc04}. Such a set system is easier to
-work with for reasons that will become apparent.\tnote{VAGUe!}
-For any set system $\cF$, the smallest super set system that is
-complete/weakly partative is called the {\em weak closure} of $\cF$,
-denoted $\cW(\cF)$ \cite[Def.~3.2]{mcc04}. \cite[Def.~2, smallest {\em
-  complete} supercollection]{mm96}.
+\begin{theorem}[{\cite[Th.~3,6]{mm96}}]
+  \label{th:validcop}
+  For any set system $\cF$ we have the following. \par
+  \centering
+    $\valid{\cF} = \valid{\overline{\cF}}$ \\
+    $\cF^{\bot} = \overline{\cF^{\bot}} = (\overline{\cF})^{\bot}$
+\end{theorem}
 
-What is interesting is that the set of \COP permutations of $\cF$ is
-the same as that of $\cW(\cF)$ \cite[Th.~3]{mm96}. The reason this is
-interesting is because it becomes apparent by this theory that it is
-easier to use $\cW(\cF)$ to create the $\PQRtree$ for $\cF$.
+\begin{proposition}
+  \label{pr:validcop}
+  Any set that is consecutive on all the \COP permutations of $\cF$ is
+  present in $\overline{\cF}$.
+\end{proposition}
 
-The generalized \PQtree of $\cF$ is defined as the {\em decomposition
-  tree} of its weak closure $\cW(\cF)$. Decomposition tree is the
-inclusion tree (inclusion p.o. Hasse diagram) of the strong elements
-of $\cW(\cF)$ with labels {\em prime, linear} and {\em degenerate}.\cite{mcc04}
+Proposition~\ref{pr:validcop} is owing to the following theorem by
+which \cite{mm96} describes a way to decompose the problem of finding
+all \COP orders of $\cF$ into two subproblems using sets in
+$\overline{\cF} \cap \cF^{\bot}$.
 
-Overlap components are computable in linear time \cite{mm95,wlh92}.
+\begin{theorem}[{\cite[Th.~7]{mm96}}]
+  \label{th:mmdecomp}
+  For any set system $\cF$, and $\emptyset \ne H \in \overline{\cF}
+  \cap \cF^{\bot}$ we have the following.\par
+  \centering
+  $\valid{U,\cF} = \valid{U/H,\cF/H} * \valid{H,\cF \cap 2^H}$
+\end{theorem}
+
+\def \ff {\overline{\cF} \cap \cF^{\bot}}
+
+The idea behind Theorem~\ref{th:mmdecomp} is as follows. A permutation $\alpha$
+of $U$ is a composition of two permutations with respect to $H$ - i) a
+permutation $\gamma$ of $H$ and (ii) a permutation $\beta$ of $U/H$.
+
+For two mutually orthogonal sets $A, B$ such that $A \nsubseteq B$,
+$A/B$ is defined as the set obtained by removing all elements of $B$
+from $A$ and adding a representative element for $B$ in $A$. Being orthogonal,
+results in only the following three possibilities.
+\begin{enumerate}
+\item $A$ and $ B$ are disjoint: $A/B  = A$
+\item $A$ is a subset of $B$: $A/B$ is not defined
+\item $B$ is a subset of $A$: $A/B = A \setminus B \cup \{b\}$, where $b$
+is a new element not in $U$ added to represent $B$.
+\end{enumerate}
+
+We observe that this idea of decomposing the \COP problem into two
+subproblems in \cite{mm96} is very similar to the substitution
+decomposition of a set system given in \cite{mcc04}.\footnote{Using
+  the theory cited in footnotes~\ref{mcc1},\ref{mcc2},\ref{mcc3}.}
+
+The following corollary states how \PQRtree elegantly fits into this
+whole theory and help in computing all \COP orders of $\cF$.
+
+\begin{corollary}[{\cite[Cor.~8]{mm96}}]
+  Let $\cF$ be a set system with universe $U$ and $H$ is a non-empty
+  orthogonal set $H \in
+  \ff$. If there is a \PQRtree $T_1$ that encodes all permutations in
+  $\valid{U/H, \cF/H}$ and also a \PQRtree $T_2'$ that encodes all
+  permutations $\valid{H,\cF \cap \power{H}}$, then a \PQRtree $T$ for
+  $\cF$ can be obtained by replacing the leaf $h$ in $T_1$ by $T_2$.
+\end{corollary}
+
+%Finding elements in $\ff$ is done using \PQRtree.
+
+Thus we have a recursive algorithm that can compute the \PQRtree for
+$\cF$ provided we find an element from $\ff$ in each iteration. The
+non-empty sets in $\ff$ are called {\em node sets} since they form the
+nodes in the \PQRtree. They are calculated as follows.  One is by
+computing the overlap components of $\cF$. The overlap components is
+the partition that results from the {overlap} equivalence relation
+which is nothing but non-orthogonal equivalence relation\footnote{It
+  is easy to verify that this relation is indeed an equivalence
+  relation.}. Overlap components are linearly
+computable\cite{mm95,wlh92}. Once these elements are factored
+out, the rest of the node sets are obtained by identifying {\em twin}
+elements\footnote{\cite[Sec.~3]{kklv10} calls this {indistinguishable}
+  elements. The equivalence class is called a {\em slot}.}. Two
+elements $a, b \in U$ are twins if their membership in every set of
+$\cF$ is in tandem with each other, \ie $\{a,b\} \bot \cF$. This is
+clearly an equivalence relation and their equivalence classes is known
+to be computable in linear time\cite{wlh01,mm96-ref11} and even in
+logspace\cite{kklv10}.
+
+The recursion end condition is when one cannot find any more sets from
+$\ff$ that are non-trivial. This is when $\ff = \cT(U)$. This is the
+point where the parents of the leaves of the \PQRtree are created. The
+following theorem helps the algorithm decide whether a $P$, $Q$ or
+$R$ node must be created.
+
+\begin{theorem}[{\cite[Th.~9]{mm96},\cite[Th.~2.1]{mcc04}}]
+\label{th:pqrfinalmm}
+If $\cF$ is a set system with universe $U$ and $|U| \ge 3$ then one of
+the following statements hold.
+\begin{enumerate}
+\item \label{th::it1} $\overline{\cF} = \cT(U)$ 
+\item \label{th::it2}$\overline{\cF} = consec(\alpha)$\tnote{what is consec?} for
+  some permutation $\alpha$ on $U$ 
+\item \label{th::it3}$\overline{\cF} = \power{U}$ 
+\end{enumerate}
+\end{theorem}
+
+In case (\ref{th::it1}) all elements in $U$ are made children of a
+$P$ node. In case (\ref{th::it2}) all elements in $U$ are made
+children of a $Q$ node in the order given by $\alpha$. Finally case
+(\ref{th::it3}) is the one when no permutation of $U$ gives \COP. All
+elements of $U$ are in this case made children in an $R$ node.
+
+Theorem~\ref{th:pqrfinalmm} and the theory leading to it is very
+similar to \cite[Th.~2.1, 3.5. Also Th.~3.2, 3.3, 3.4]{mcc04} which categorizes
+the nodes in the above the three cases as {\em prime}, {\em linear}
+and {\em degenerate} respectively. Their generalized \PQtree is
+created in similar ways as \PQRtree above. This tree is in essense the
+Hasse diagram of what they call {\em strong elements} of
+$\overline{\cF}$. Strong elements of a set family are elements that do
+not overlap with any other elements in the family, \ie it is
+orthogonal to all other sets\cite[Def.~3.3]{mcc04}.
+
+\begin{theorem}[{\cite{mm96}, \cite[Th.~3.6]{mcc04}}]
+The set system $\cF$ has \COP \iff its \PQRtree has no $R$ nodes.  
+\end{theorem}
+
+% The generalized \PQtree of $\cF$ is defined as the {\em decomposition
+%   tree} of its weak closure $\cW(\cF)$. Decomposition tree is the
+% inclusion tree (inclusion p.o. Hasse diagram) of the strong elements
+% of $\cW(\cF)$ with labels {\em prime, linear} and {\em degenerate}.\cite{mcc04}
+
+% Overlap components are computable in linear time \cite{mm95,wlh92}.
+
+----- Cert algo
+
+------ ICPIA
 

### 2012-01-18 src/_region_.tex
first draft of PQR trees done. next is cert algo and icpia.

--- 2012-01-14 src/020chap2.tex
@@
-\definecolor{gray}{rgb}{0.8,0.8,0.8}
-\def\tblhcolor {gray}
+\message{ !name(020chap2.tex)}
+\message{ !name(020chap2.tex) !offset(707) }
+  \item The relative complements $A \setminus B$ and $B \setminus A$
+    if $B \nsubseteq A$ and $A \nsubseteq B$ respectively.
+  \item Also note that trivially, sets in $\cT(U)$ are consecutive in
+    any permutation of $U$\footnote{$\emptyset$ is considered
+      consecutive by convention.}.
+\end{enumerate}
+\end{observation}
 
-\chapter{\CoP\,-- A Survey of Important Results}
-\label{ch:copsurvey}
-
-This chapter surveys several results that are significant to this
-thesis or to \COP in general. These predominantly pertain to
-characterizations of \COP, algorithmic tests to check for \COP (COT),
-optimization problems on binary matrices that do not have \COP and
-some applications of \COP.
-
-\tnote[important]{ADD: have a few lines about organization of chapter}
-
-\section{\COP in Graph Theory}
-
-\COP is closely connected to several types of graphs by way of
-describing certain combinatorial graph properties. There are also
-certain graphs, like convex bipartite graphs, that are defined solely
-by some of its associated matrix having \COP.  In this section we will
-see the relevance of \cop to graphs.  To see this we introduce certain
-binary matrices that are used to define graphs in different
-ways. While adjacency matrix is perhaps the most commonly used such
-matrix, Definition~\ref{def:graphmatrices} defines this and a few
-more.
-
-% \begin{definition}{\emph{Matrices that define
-%       graphs.\cite[Def.~2.4]{d08phd}}} %
-\begin{definition}
-  {\em Matrices that define graphs. \cite[Def.~2.4]{d08phd}.}  
-  \label{def:graphmatrices} 
-  Let $G$ and $H$ be defined as follows. $G = (V,E_G)$ is a graph with
-  vertex set $V = \{v_i \mid i \in [n]\}$ and edge set $E_G \subseteq
-  \{(v_i,v_j) \mid i, j \in [n]\}$ such that $|E_G| = m$. $H = (A, B,
-  E_H)$ is a bipartite graph with partitions $A = \{a_i \mid i \in
-  [n_a]\}$ and $B = \{b_i \mid i \in [n_b]\}$.
-  \begin{enumerate}[{\ref{def:graphmatrices}}--i.]
-  \item \emph{Adjacency matrix} of $G$ is the symmetric $n \times n$
-    binary matrix $M$ with $m_{i,j} = \un$ \iff $(v_i,v_j) \in E_G$
-    for all $i,j \in [n]$.
-  \item \emph{Augmented adjacency matrix} of $G$ is obtained from its
-    adjacency matrix by setting all main diagonal elements to \un, \ie
-    $m_{i,i} = \un$ for all $i \in [n]$.
-  \item \emph{Maximal clique matrix} or \emph{vertex-clique incidence
-      matrix} of $G$ is the $n \times k$ binary matrix $M$ with
-    $m_{i,j} = \un$ \iff $v_i \in C_j$ for all $i \in [n], j \in [k]$
-    where $\{C_j \mid j \in [k]\}$ is the set of maximal cliques of
-    $G$.
-    \label{def::maxcliquematrix}
-  \item \emph{Half adjacency matrix} of $H$ is the $n_a \times n_b$
-    binary matrix $M$ with $m_{i,j} = \un$ \iff $(a_i, b_j) \in E_H$.
-  \end{enumerate}
-\end{definition}
-
-\figgraphmatrices
-
-Now we will see in Definition~\ref{def:graphwithcop} certain graph
-classes that is related to \COP or \CROP.\\
-
-\begin{definition}{\emph{Graphs that relate to
-      COP.\cite[Def.~2.5]{d08phd}}} %
-  \label{def:graphwithcop} %
-  Let $G$ be a graph and $H$ be a bipartite graph.
-  \begin{enumerate}[\ref{def:graphwithcop}--i.]
-  \item $G$ is \emph{convex-round} if its adjacency matrix has the
-    \CROP.
-  \item \label{def::concave-round} $G$ is \emph{concave-round} if its
-    augmented adjacency matrix has \CROP. \tnote{cite BHY00}
-    \tnote[minor]{add CROP to glossary}
-  \item $G$ is an \emph{interval graph} if its vertices can be mapped
-    to intervals on the real line \stt two vertices are adjacent \iff
-    their corresponding intervals overlap \tnote{cite Ben59, Haj57}.
-    $G$ is an interval graph \iff its maximal clique matrix has COP
-    \cite{fg65}\endnote{This follows \cite{gh64} which states that the
-      maximal cliques of interval graph $G$ can be linearly ordered
-      \stt for all $v \in V(G)$, cliques containing $v$ are
-      consecutive in the ordering \cite[Th. 8.1]{mcg04}.}
-    \begin{enumerate}[a.]
-    \item $G$ is a \emph{unit interval graph} if it is an interval
-      graph \stt all intervals have the same length.\endnotemark[2]
-    \item $G$ is a \emph{proper interval graph} if it is an interval
-      graph \stt no interval properly contains another.\endnote{The
-        set of unit interval graphs and the set of proper interval
-        graphs are the same} \tnote{cite {rob69,gar07} in endnote. pg
-        33 dom}
-    \end{enumerate}
- \item $G$ is a \emph{circular-arc graph} if its vertices can be
-    mapped to a set of arcs on a circle \stt two vertices are adjacent
-    \iff their corresponding arcs overlap.\tnote[pressing]{how is
-      CO/ROP related?}
-  \item $H$ is \emph{convex bipartite on columns (rows)} if its half
-    adjacency matrix has \COP on rows (columns).%
-    \label{def::convexbi}
-  \item $H$ is \emph{biconvex bipartite} or \emph{doubly
-      convex}\cite{yc95} if its half adjacency matrix has \COP on both
-    rows and columns.
-  \item $H$ is \emph{circular convex} if its half adjacency matrix has
-    \CROP.
-  \end{enumerate}
-\end{definition}
-
-Interval graphs\endnote { \cite{mcc04} cites that the problem of
-  recognizing interval graphs has significance in molecular
-  biology. Interestingly, in the late 1950s, before the structure of
-  DNA was well-understood, Seymour Benzer was able to show that the
-  intersection graph of a large number of fragments of genetic
-  material was an interval graph \cite{sb59}. This was regarded as
-  compelling evidence that genetic information was somehow arranged
-  inside a structure that had a linear topology which we now know to
-  be true\tnote{linear arrangement of DNA/genes?}.} and circular-arc
-graphs have a long history in research.  The interest around them is
-due to their very desirable property that several problems that are
-NP-hard \tnote[minor]{all NPC problems are NPH yes?} on general
-graphs, like finding a maximum clique or minimum coloring or
-independent set, are polynomial time solvable in these graph classes
-\cite{clrs01}.
-In a similar fashion, a lot of problems that are hard on general
-matrices have efficient solutions on matrices with \COP or \CROP
-\cite[more citations pg.\,33]{d08phd}.
-
-Table~\ref{tab:graphmatrices} summarises the way these graphs are
-characterized by their matrices having \COP or \CROP.
-Our focus in this chapter (and thesis) is mainly \COP and 
-having seen how useful \COP is in identifying or characterizing many
-types of graphs\endnote{\COP has several other applications
-  \tnote{fill up.}}, we will now see results that study recognition of \COP
-in matrices in the following section.
-
-\temptext{ tab~\ref{tab:graphmatrices} -  have an abridged version
-  of table 2.1 in dom. see notes for the possibility of a
-  ``circle diagram''.} 
-\begin{table}[htbp]
-  \centering
- \caption{\figtabsize Graph matrices \temptext{PLACEHOLDER}}
-  \label{tab:graphmatrices}
-\end{table}
-
-%\clearpage      % Print pending floats
-
-\section{Matrices with COP}
-\label{sec:surveycoptest}
-
-The most important questions with respect to a particular 
-property desired in a structure/object are perhaps the following.
-\begin{itemize}%[]
-\singlespacing
-\item Does the desired property exist in the given input?
-\item If the test is affirmative, what is a certificate of the affirmative?
-\item If the test is negative, what are the optimization possibilities
-  for the property in the input? In other words, how close to having the
-  property can the input be?
-\item If the test is negative, what is a certificate of the negative?
-\end{itemize}
-
-In this section and the rest of the chapter we see results that shaped the
-correponding areas respectively for \cop in binary matrices.
-
-\begin{enumerate}[a.]
-\singlespacing
-\item \label{q:testcop} Does a given binary matrix have \COP?
-\item \label{q:copperm} What is the \COP permutation for the given matrix with \COP?
-\item \label{q:copopt} What are the optimizations possible and practically useful on
-  the given matrix without \COP?
-\item \label{q:nocopcert} If algorithm for (\ref{q:testcop}) returns
-  \textbf{false}, can a certificate for this be computed?
-\end{enumerate}
-
-Without doubt, besides computing answers to these questions, we are
-interested in the efficiency of these computations in terms of
-computational complexity theory. Results towards
-questions~(\ref{q:testcop}) and (\ref{q:copperm}) are surveyed in this
-section. Those for question~(\ref{q:copopt}) are discussed in
-Section~\ref{sec:surveycopopt} and question~(\ref{q:nocopcert}) is discussed
-in Section~\ref{sec:surveycertalgo}.
-
-It may be noted that one way to design an algorithm to test for \COP
-is by deriving one from any interval graph recognition algorithm using the
-result {\sc hmpv00}\tnote{cite hmpv00} \tnote{in endnote put the theorem 2.7 - dom
-  pg 43} \cite{d08phd} which demonstrates how such a derivation can be
-done. However, this does not necessarily yield an efficient
-algorithm. We will see results that directly solve the problem on
-matrices since it is known that questions~(\ref{q:testcop}) and
-(\ref{q:copperm}) stated above for \COP are efficiently solvable.
-Table~\ref{tab:cophistory} gives a snapshot of these results.
-
-    \tnote[important]{1899 by petrie -- cite
-         kendall 69 pacific journal of mathematics 1969 }
-
-\def\tabparboxw{10cm}
-\begin{table}[htbp]
-  \centering
-  {
-    \fontencoding{OT1}\fontfamily     {cmss}\selectfont
-    % \sffamily
-    \footnotesize
-    % \includegraphics[scale=0.5]{../img/cophistory.png} % TEMPORARY
-
-  \begin{tabular}{lll}
-     \hline
-     \hline
-     % 
-     1899 & \parbox[t]{\tabparboxw}{ First mention of \COP
-       (archaeology)} & 
-       \\ % \parbox[t]{2cm}{}
-
-     \rowcolor[gray]{0.8} %
-     1951 & \parbox[t]{\tabparboxw}{ Heuristics for COT} & \cite{wsr51} \\
-
-     1965 & \parbox[t]{\tabparboxw}{ First polynomial time algorithm
-       for \COP testing} & \cite{fg65} \\
-
-     \rowcolor[gray]{0.8} %
-     1972 & \parbox[t]{\tabparboxw}{ Characterization for \COP -- forbidden
-       submatrices} &
-     \cite{at72} \\
-
-     1976 & \parbox[t]{\tabparboxw}{ First linear time algorithm for COT --
-       \PQtree} &
-     \cite{bl76} \\
-
-     \rowcolor[gray]{0.8} %
-     1992 & \parbox[t]{\tabparboxw}{ Linear
-       time algorithm COT without \PQtree} & \cite{wlh02}\endnote{First
-       published in \cite{wlh02isaac92}} \\
-
-     2001 & \parbox[t]{\tabparboxw}{\PCtree\ -- a simplification of
-       \PQtree} & \cite{wlh01,hm03} \\
-
-     \rowcolor[gray]{0.8} %     
-     1996 & \parbox[t]{\tabparboxw}{\PQRtree\ -- generalization of
-       \PQtree for any binary matrix regardless of its \COP status} &
-     \cite{mm96} \\
-
-     1998 & \parbox[t]{\tabparboxw}{ Almost linear time to construct
-       \PQRtree} & \cite{mpt98} \\
-
-     \rowcolor[gray]{0.8} %
-     2004 & \parbox[t]{\tabparboxw}{ A certifying algorithm for no
-       \COP.\\ Generalized \PQtree.} & \cite{mcc04} \\
-
-     2009 & \parbox[t]{\tabparboxw}{ Set theoretic, cardinality based
-       characterization of \COP\ -- ICPIA} & \cite{nsnrs09} \\
-
-     \rowcolor[gray]{0.8} %
-     2010 & \parbox[t]{\tabparboxw}{ Logspace \COP testing} &
-     \cite{kklv10}\\ 
-     \hline
-     \hline
-   \end{tabular}
- }
- \caption[\figtabsize A brief history of \COP research]{\figtabsize A
-   brief history of \COP research }
- \label{tab:cophistory}
-\end{table}
-
-The first polynomial time algorithm for \COP testing was by
-\cite{fg65} which uses overlapping properties of columns with \un
-s. Their result has close relations to the characterization of
-interval graphs by \cite{gh64}. A graph $G$ is an interval graph \iff
-all its maximal cliques can be linearly ordered \stt for any vertex
-$v$ in $G$, all the cliques that $v$ is incident on are consecutive in
-this order. Clearly, this means that the maximal clique incidence
-matrix\endnote{
-  Definition~\ref{def:graphmatrices}~\ref{def::maxcliquematrix}} must
-have \COP on rows.
-
-A few years later, a deeply significant result based on very different
-ideas in understanding \COP came from Tucker which gave a
-combinatorial (negative) characterization of matrices with COP
-\cite{at72}. This result influenced most of the \COP results that
-followed in the literature including linear time
-algorithms\tnote[pressing]{linear? did it? which ones?} for \COP
-recognition.
-
-%\clearpage
-\subsection{Tucker's forbidden submatrices for \COP}
-
-% \change{While it is remarkable that the first polynomial time
-%   algorithm for \COP testing came in }{The first polynomial time
-%   algorithm for \COP testing came in }\cite{fg65} \change{,}{.}
-% \change{it is safe to say that the most}{A few years later, a very}
-% significant result in understanding \COP came \remove{a few years
-%   later} from Tucker which gave a combinatorial (negative)
-% characterization of matrices with COP \cite{at72}.
-
-\cite{at72} discovered certain forbidden structures for convex
-bipartite graphs\endnote{The terminology in \cite{at72} differs. It
-  uses the term {\em graphs with $V_1$-consecutive arragement} instead
-  of {\em convex bipartite graphs}.} and by definition of this graph
-class, this translates to a set of forbidden submatrices for matrices
-with \cop.  The following are the theorems from \cite{at72} that
-acheived this characterization.
-
-Theorem~\ref{th:tuckeratfree} states that convex bipartite graphs
-cannot have {\em asteroidal triples}\endnote{If $G = (V,E)$ is a
-  graph, a set of three vertices from $V$ form an {\em asteroidal
-    triple} if between any two of them there exists a path in $G$ that
-  does not contain any vertex from the closed neighborhood of the
-  third vertex.} contained in the corresponding vertex
-partition\endnote{The partition corresponds to columns (rows) if its
-  half adjacency matrix has \COP columns (rows).}.
-Theorem~\ref{th:tuckerforbidden} lists the structures in a
-bipartite graph that force one of its vertex partitions to have
-asteriodal triples -- in other words, it identifies the subgraphs that
-prevent the graph from being convex bipartite.
-
-\begin{theorem}
-  (\cite[Th.~6]{at72}, \cite[Th.~2.3]{d08phd})\\
-  A bipartite graph $G = (V_1, V_2, E)$ is convex bipartite on
-  columns\endnote{Abridged to match terminology adopted in this
-    document. See previous note.} \iff $V_1$ contains no asteroidal
-  triple of $G$.
-  \label{th:tuckeratfree}
+\begin{theorem}[{\cite[Th~3]{mm96}}]
+  For any set system $\cF$, the set of all \COP permutations of $\cF$
+  is equal to the set of all \COP permutations of $\overline{\cF}$.
 \end{theorem}
 
-\begin{theorem}
-  (\cite[Th.~7]{at72}, \cite[Th.~2.4]{d08phd})\\
-  In a bipartite graph $G = (V_1, V_2, E)$ the vertex set $V_1$
-  contains no asteroidal triple \iff $G$ contains none of the graphs
-  $G_{I_k}$, $G_{II_k}$, $G_{III_k}$ (with $k \ge 1$), $G_{IV}$,
-  $G_{V}$ as shown in Figure~\ref{fig:forbiddensubgraphs} as subgraphs.
-  \label{th:tuckerforbidden}
-\end{theorem}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[scale=0.35]{../img/tuckersforbiddensubgraphs.png}  
-  \caption[\figtabsize Tucker's forbidden subgraphs]{\figtabsize Tucker's
-    forbidden subgraphs for convex bipartite graphs. \temptext{ PLACEHOLDER
-    IMG }} 
-  \label{fig:forbiddensubgraphs}
-\end{figure}
-
-Theorem~\ref{th:tuckeratfree} and Theorem~\ref{th:tuckerforbidden}
-result in the following Theorem~\ref{th:tuckercop} which characterizes
-matrices with \COP.
-
-\begin{theorem}
-  (\cite[Th.~9]{at72}, \cite[Th.~2.5]{d08phd})\\
-  A matrix $M$ has \COP \iff it contains none of the matrices 
-$M_{I_k}$, $M_{II_k}$, $M_{III_k}$ (with $k \ge 1$), $M_{IV}$,
-  $M_{V}$ as shown in Figure~\ref{fig:forbiddensubmatrices} as submatrices.
-  \label{th:tuckercop}
-\end{theorem}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[scale=0.35]{../img/tuckersforbiddensubmatrices.png}  
-  \caption[\figtabsize Tucker's forbidden submatrices]{\figtabsize Tucker's
-    forbidden submatrices for convex bipartite graphs. \temptext{ PLACEHOLDER
-    IMG }} 
-  \label{fig:forbiddensubmatrices}
-\end{figure}
-
-It can be verified that the matrices in
-Figure~\ref{fig:forbiddensubmatrices} are the half adjacency matrices
-of the graphs in Figure~\ref{fig:forbiddensubgraphs} respectively
-which is not surprising due to
-Definition~\ref{def:graphwithcop}~\ref{def::convexbi}.
-
-%\clearpage
-\subsection{Booth and Lueker's $PQ$ tree -- a linear COT algorithm}
-
-Booth and Lueker in their paper \cite{bl76} gave the first linear
-algorithm\endnote{Time complexity is $O\left(m+n+f\right)$ where $m
-  \times n$ is the order of the input matrix and $f$ is the number of $\un$s in it.} for \cop testing while given a linear time interval graph
-recognition algorithm by a simplification of \tnote{cite - A. Lempel,
-  S. Even and I. Cederbaum, An Algorithm for Planarity Testing of
-  Graphs, Theory of Graphs, ed., P. Rosenstiehl, Gordon and Breach,
-  New York, (1967), 215-232.  }'s planarity test algorithm.
-% Hsu in wlh01 said BL76 gave a planarity test - "the induced PQ-tree
-% algorithm can considerably simplify Booth and Lueker’s modification
-% of Lempel, Even and Cederbaum’s planarity test. " 
-%
-% McConnell in hm03 said the first linear algo for interval graph
-% recog was by BL76 - "Benzer’s work and other applications of
-% interval graphs motivated a search for eﬃcient algorithms for
-% recognizing interval graphs, and for constructing a set of intervals
-% to represent the graph when it is one [8]. A linear-time algorithm
-% was given in 1976 by Booth and Lueker [2]." 
-\cite{bl76} introduces a data structure called \PQtree and their \COP
-testing algorithm is a constructive one that outputs a \PQtree if the
-input has \COP. A \PQtree represents all the \COP orderings of the
-matrix it is associated with. \cite{bl76}'s algorithm uses the fact
-that if a matrix has \COP, a \PQtree for it can be constructed. It is
-interesting to note that aside from interval graph recognition and
-\COP testing, \PQtree is also useful in other applications like
-finding planar embeddings of planar graphs \cite{lec67,mcc04} and
-recognizing \CROP in a matrix.
-
-\begin{definition}[\PQtree \cite{bl76, mcc04}]
-  A \PQtree of matrix $M$ with \COP on columns (rows), is a tree with the following properties.
-  \begin{enumerate}[i.]
-    \singlespacing
-  \item Each leaf uniquely represents a row (column) of $M$. The leaf
-    order of the tree gives a \COP order for column (row)\endnote{Note
-      that \COP order for column requires permutation of rows and vice
-      versa.} for $M$.
-  \item Every non-leaf node in the tree is labeled $P$ or $Q$.
-  \item \label{def::nodep} The children of $P$ nodes are
-    unordered. They can be permuted in any fashion to obtain a new
-    \COP order for $M$.
-  \item \label{def::nodeq} The children of $Q$ nodes are linearly
-    ordered. Their order can be reversed to obtain a new \COP
-    order for $M$.
-  \end{enumerate}
-  \label{def:pqtree}
-\end{definition}
-
-See Figure~\ref{fig:pqtree} for an example of \PQtree. It may be noted
-that there is no way an empty set of \COP orderings can be represented
-in this data structure. For this reason, \PQtree is undefined for
-matrices that do not have \COP.  Thus effectively, there exists a
-bijection between set of matrices with \COP and the set of \PQtrees
-(accurately speaking, each matrix with \COP bijectively maps to an
-equivalence class of \PQtrees resulting from properties (\ref{def::nodep}) and
-(\ref{def::nodeq})).
-
-\figpqtree
-
-The \cite{bl76} algorithm with input $n \times m$ matrix $M$ starts
-with a \PQtree for a vacuous $n \times 0$ matrix $M'$ (submatrix
-induced by 0 columns). This is known as a {\em universal} \PQtree
-which is one with its root as a $P$ node and only leaves as its
-children -- each leaf representative of a row of input (by definition
-of \COP for columns). This induced submatrix $M'$ vacuously has \COP.
-Each column is then added iteratively to $M'$ to check if the new $M'$
-has \COP.  By a complicated, but linear, procedure the algorithm does
-one of the following actions in each iteration: (a) declare that $M$
-has no \COP, or (b) modify the current \PQtree to represent the new
-$M'$ (which clearly, must have \COP, since if not, option (a) would
-have been executed).
-
-Judging from notes in literature, this algorithm is apparently
-notoriously difficult to program. In the procedure to modify the
-\PQtree at each iteration, nodes are considered from leaves to
-tree. At each node considered, it uses one of nine templates to
-determine how the tree must be altered in the vicinity of this
-node. Recognition of this template poses a difficult challenge in
-terms of implementing it. Each template is actually a representative
-of a larger class of similar templates, which must be dealt with
-explicitly by a program\cite{mcc04}.
-
-After the invention of \PQtrees, presumably due to the implementation
-challenge it posed, there has been several variants of the same in the
-literature, like \PCtree \cite{sh99,wlh01,hm03}, generalized \PQtree
-\cite{km89,mcc04}, \PQRtree \cite{mm96,mpt98} etc. Most of these are
-generalizations of \PQtree -- for instance, \PCtree is generalized to
-matrices with \CROP, \PQRtree and generalized \PQtree are generalized
-to matrices and set systems with or without \COP. \cite{km89} invented
-a modified form of \PQtree a simpler incremental update of the tree
-only for recognizing interval graphs.  \cite{kr88} constructed
-efficient parallel algorithms for manipulating PQ trees. \temptext{dom
-  chapter 2 pg 40 Variations of PQTrees first para - summarize.}
-
-In the next few following sections we will see some of these
-variations.
-%Their algorithm was improved by \cite{tm05}
-%\tnote{improved in terms of what?}. 
-% describes an improved algorithm to
-% build PQR trees.
-
-%\section{Certifying Algorithms}
-\subsection{\PQRtree\ --  COP for set systems}
-\label{sec:surveycertalgo}
-
-Section~\ref{sec:motive} mentions how a binary matrix naturally maps
-to a system of sets.  A set can be constructed for each column of
-matrix with its elements being those row indices at which the column
-has \un s. Thus the collection of sets corresponding each column of
-the matrix forms a set system with universe as the set of all row
-indices of the matrix.  This simple construction is formally described
-in Definition~\ref{def:matrixsetsystem} along with the idea of \cop
-for set systems\endnote{As seen in Section~\ref{sec:courseschedule}}.
-
-\begin{definition}%
-  \label{def:matrixsetsystem}%
-  % [\emph{Set system of a binary matrix}.]
-  [\emph{\Cop for set systems}.]  Let $M$ be a binary matrix of order
-  $n \times m$ % , $r_i, i \in [n]$ be the rows in $M$
-  and $\{c_i \mid i \in [m]\}$ be the columns in $M$. A set system
-  $\cF_{M} = \{S_i \mid S_i \subseteq [n], i \in [m]\}$ is defined
-  such that for every column $c_i$ of $M$, set $S_i = \{j \mid m_{ji}
-  = \un \}$. The collection $\cF_M$ is the {\em set system of binary
-    matrix} $M$.\\
-  A set system $\cF$ from universe $U = \{1, 2, \ldots, n\}$ has the
-  {\em \cop} if it is possible to assign a linear order, \ie a
-  permutation $\sigma = u_1u_2\ldots u_n$ to $U$ \stt each set $S \in
-  \cF$ appears as a consecutive subsequence $u_{i}u_{i+1}\ldots
-  u_{i+k-1}$ (also called an interval) of $\lambda$ in the linear order, where
-  positive integers $k = |S|$ and $i \le n+1-k$.
-\end{definition}
-
-It is easy to see the equivalence of this definition to \COP for
-matrices in Definition~\ref{def:copmatrix}.
-
-Generalized \PQtree or \gPQtree is a data structure defined in \cite{n89} to
-represent all orthogonal sets of a set system $\cF$. {\em Orthogonal
-  sets} of $\cF$ are sets that do not overlap with any set in $\cF$. A
-data structure with the same name was later defined in \cite{mcc04} as
-part of a {\em substitution decomposition} for a set system
-$\cF$. \tnote{define substitutio decomp.} This is later used to define
-a certificate if $\cF$ has no \COP. \PQRtree is a data structure
-defined in \cite{mm96} as a data structure to represent any set system
-$\cF$ with additional information in their $R$-nodes if $\cF$ has no
-\COP.  All three of these data structures are proposed as
-generalizations of \cite{bl76}'s \PQtree and hence produce the
-$\PQtree$ if $\cF$ has \COP. As a whole, all these three data
-strutures are near-identical with their differences being notional.
-In this section, we will discuss the basic theory that they all hold
-and stick to the name of the data structure as \PQRtree.
-
-A \PQRtree of a given set system $\cF$ from universe $U$ represents
-all the sets from $\power{U}$ that have trivial intersections with
-every set in $\cF$.  Two sets $A$ and $B$ are said to have a {\em
-  trivial intersection} if $A \cap B$ is one of the following --
-$\emptyset$, $A$ or $B$. In other words, $A$ and $B$ are either (i)
-disjoint or (ii) one is the subset of the other. Such sets are also
-called {\em orthogonal sets} or said that $A$ is orthogonal to
-$B$. {\em Trivial subsets} of a universe $U$, denoted by $\cT(U)$, are
-sets that have trivial intersections with any set in
-$\power{U}$. Namely, $\cT(U)$ consists of $U$, singleton sets in
-$\power{U}$ and $\emptyset$\cite{n89, mm96}.  In other words, the all
-sets presented in a \PQRtree share the property that they do not
-overlap with any set in $\cF$. We will now see how this helps in
-testing for \COP in $\cF$.
-
-An important observation made by \cite{mm96} is that if $\cF$ is a set
-system with \COP then, aside from every set in $\cF$ being consecutive
-(after applying the \COP order) the following sets must also be
-consecutive for any $A, B \in \cF$.
-\begin{enumerate}
-  \item The intersection $A \cap B$ 
-  \item The union $A \cup B$ if $A \cap B \ne \emptyset$
-  \item The relative complements $A \setminus B$ and $B \setminus A$
-    if $B \nsubseteq A$ and $A \nsubseteq B$ respectively.
-\end{enumerate}
-Also note that trivially, sets in $\cT(U)$ are consecutive in any
-permutation of $U$\endnote{$\emptyset$ is considered consecutive by
-  convention.}.
-
-Any set system that is closed under the above operations and contains
-$\cT(U)$ is called {\em complete} \cite{mm96} or a {\em weakly partative
-  family} \cite{mcc04}. Such a set system is easier to
-work with for reasons that will become apparent.\tnote{VAGUe!}
-For any set system $\cF$, the smallest super set system that is
-complete/weakly partative is called the {\em weak closure} of $\cF$,
-denoted $\cW(\cF)$ \cite[Def.~3.2]{mcc04}. \cite[Def.~2, smallest {\em
-  complete} supercollection]{mm96}.
-
-What is interesting is that the set of \COP permutations of $\cF$ is
-the same as that of $\cW(\cF)$ \cite[Th.~3]{mm96}. The reason this is
-interesting is because it becomes apparent by this theory that it is
-easier to use $\cW(\cF)$ to create the $\PQRtree$ for $\cF$.
-
-The generalized \PQtree of $\cF$ is defined as the {\em decomposition
-  tree} of its weak closure $\cW(\cF)$. Decomposition tree is the
-inclusion tree (inclusion p.o. Hasse diagram) of the strong elements
-of $\cW(\cF)$ with labels {\em prime, linear} and {\em degenerate}.\cite{mcc04}
-
-Overlap components are computable in linear time \cite{mm95,wlh92}.
-
-\temptext{ Certifying algorithm (Generalized PQ trees) [McC04] --
-
-Set theoretic characterizations [Hsu02, NS09]\\
-    \cite{wlh02} describes the simpler algorithm for
-    COT.\tnote[pressing]{in terms of what?}\\
-    \cite{nsnrs09} describes a characterization of consecutive-ones
-    property solely based on the cardinality properties of the set
-    representations of the columns (rows); every column (row) is
-    equivalent to a set that has the row (column) indices of the rows
-    (columns) that have one entries in this column (row). This is
-    interesting and relevant, especially to this thesis because it
-    simplifies COT to a great degree by reducing the solution search
-    space owing to the a simple set theoretic characterization.\\
-
-\cite{mcc04} describes a different approach to COT. While all previous
-COT algorithms gave the COP order if the matrix has the property but
-exited stating negative if otherwise, this algorithm gives an evidence
-by way of a certificate of matrix even when it has no COP. This
-enables a user to verify the algorithm's result even when the answer
-is negative. This is significant from an implementation perspective
-because automated program verification is hard and manual verification
-is more viable. Hence having a certificate reinforces an
-implementation's credibility. Note that when the matrix {\em has} COP,
-the COP order is the certificate.  The internal machinery of this
-algorithm is related to the weighted betweenness problem
-addressed\tnote{in what way??} in \cite{co98}.  \tnote{expand
-  on the COP order graph creation and it having to be bipartite for M
-  to have COP. and thus an odd cycle being an evidence of no COP.}
-
-}
-
-\subsection{\PCtree -- a generalization of \PQtree}
-
-\PCtree is a generalization of \PQtree. It is a data structure that is
-analogous to \PQtree but for matrices with \crop. \PCtree was
-introduced by \cite{sh99} for the purpose of planarity testing where
-this data structure represents partial embeddings of planar
-graphs. % noted from wlh01 abstract.
-In \cite{wlh01}, Hsu reintroduces \PCtree as a generalization of
-\PQtree and shows how it simplifies \cite{bl76}'s planarity test by
-making the \PQtree construction much less complicated. Later
-\cite{hm03}, discovers that \PCtree is a representation of all \crop
-orders of a matrix when it is unrooted. \PCtree presented in
-\cite{wlh01} is rooted; however the construction of \PCtree is the
-same in both results. The property of being unrooted is necessary in
-order to use \PCtree as a data structure for encoding circular
-ordering. Definition~\ref{def:pctree} defines \PCtree.
-
-\begin{definition}
-  \label{def:pctree}
-  [\emph{\PCtree} \cite{wlh01,d08phd}.]  A \PCtree of matrix $M$ with
-  \CROP on columns (rows), is a tree with the following properties.
-  \begin{enumerate}[i.]
-    \singlespacing
-  \item Is unrooted -- thus it has (a) no parent child relationship
-    between nodes (b) there is no left to right (or vice versa)
-    ordering.
-  \item Each leaf uniquely represents a row (column) of $M$. The leaf
-    order of the tree gives a \CROP order for column (row) for
-    $M$. Moreover, any sequence obtained by considering the leaves in
-    clockwise or counter-clockwise order describes a \CROP order for
-    $M$.
-  \item Every non-leaf node in the tree is labeled $P$ or $C$.
-  \item The neighbors of $P$ nodes can be permuted in any fashion to
-    obtain a new \CROP order for $M$.
-  \item \label{def::pcmirror} The tree can be changed by applying the following
-    ``mirroring'' operation to obtain a new \CROP order for $M$. Root
-    the \PCtree at a neighbor of a $C$-node, $v$ and mirror the
-    subtree whose root is $v$ and finally unrooting the
-    tree. Mirroring a subtree is done by putting the children of every
-    node of the subtree in reverse order.
-  \end{enumerate}
-\end{definition}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[scale=0.4]{../img/pctree_hm03.png}
-
-  \centering
-  \includegraphics[scale=0.4]{../img/pctree_dom.png}
-  \caption{\figtabsize \PCtree \temptext{ PLACEHOLDER 
-IMGS \cite{hm03, d08phd}}}
-  \label{fig:pctree}
-\end{figure}
-
-As a data structure when \PCtree is compared with \PQtree, the
-differences are, (i) it is unrooted, (ii) it represents all \CROP order of a
-matrix (iii) it has $C$ nodes instead of $Q$ nodes which can be
-``mirrored'' (operation defined in Definition~\ref{def:pctree}~\ref{def::pcmirror}).
-The algorithms of construction of \PQtree in \cite{bl76} and that of
-\PCtree in \cite{wlh01, hm03} starkly differ since the latter is a
-much simplified procedure.
-
-% WLH01 abstract:
-% [..] The original implementation of the PQ-tree algorithms by Booth and
-% Lueker using nine templates in each bottom-up iteration is rather
-% complicated. Also the complexity analysis is rather intricate. We give
-% a very simple linear time PC-tree algorithm with the following
-% advantages: (1) it does not use any template; (2) at each iteration,
-% it does all necessary tree-modification operations in one batch and
-% does not involve the node-by-node bottom-up matching; (3) it can be
-% used naturally to test the circular ones property in matrices; (4) the
-% induced PQ-tree algorithm can considerably simplify Booth and Lueker’s
-% modification of Lempel, Even and Cederbaum’s planarity test.
-
-\tnote[FYI]{ The results in cite~{bl76} on COT are based on the
-   result that interval graphs are AT-free chordal graphs cite~{lb62}?
- No they don't seem related. this is the stupid bl lb confusion.}%
-
-%\clearpage      % Print pending floats
-% \section{COP in Relational Database Model}
-% \label{sec:surveyrdbm}
-% \tnote{Expand on sec:apprdbm} \tnote{(set systems theme)}
-\section{Optimization problems in COP}
-\label{sec:surveycopopt}
-\tnote{Expand on ref:sec:optcop}
-
-\temptext{ So far we have been concerned about matrices that have the
-  consecutive ones property. However in real life applications, it is
-  rare that data sets represented by binary matrices have COP,
-  primarily due to the noisy nature of data available. At the same
-  time, COP is not arbitrary and is a desirable property in practical
-  data representation \cite{co98,jkckv04,k77}. In this context, there
-  are several interesting problems when a matrix does not have COP but
-  is ``close'' to having COP or is allowed to be altered to have
-  COP. These are the optimization problems related to a matrix which
-  does not have COP. Some of the significant problems are surveyed in
-  this section.
-
-\tnote{ -- sect 4.1 in cite:d08phd has many results surveyed. hardness
-  results, approx. results. results are usually for a class of
-  matrices $(a,b)$ where number {\un}s in columns and rows are
-  restriced to $a$ and $b$ . -- problem of flipping at most $k$
-  entries of $M$ to make it attain COP. this is NP complete
-  cite:b75-phd}%
-\tnote{(1) scite:lb62 showed that interval graphs are chordal and
-  AT-free.} %
-\temptext{couple of lines refering to tucker's submatrices. refer
-  earlier section.}
-
-Once a matrix has been detected to not have COP (using any of the COT
-algorithms mentioned earlier), it is naturally of interest
-
-\begin{enumerate}
-\item  to find out
-the smallest forbidden substructure (in terms of number of rows and/or
-columns and/or number of entries that are {\un}s). \cite{d08phd}
-discusses a couple of algorithms which are efficient if the number of
-{\un}s in a row is small. This is of significance in the case of
-sparse matrices where this number is much lesser than the number of
-columns.
-
-\item $(*,\Delta)${\em -matrices} are matrices with no restriction
-on number of {\un}s in any column but has at most $\Delta$ {\un}s in
-any row. 
-{\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are
-similar problems which deals with {\em inducing COP} on a matrix.
-
-\begin{enumerate}
-\item In the dual problem {\sc Max COS-R (Max COS-C)} the search is for the
-maximum number of rows (columns) that induces a submatrix with
-COP.
-\item In {\sc
-  Min COS-R (Min COS-C)} the question is to find the minimum number of
-rows (columns) that must be deleted to result in a matrix with COP.
-\end{enumerate}
-
-Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
-a submatrix $M'$ with all columns\tnote{check if b75 deals with
-  COP col or COP row. also is it any submatrix with k less than r rows
-  or submatrix must have all columns?} but a maximum cardinality
-subset of rows such that $M'$ has COP is NP complete. \cite{hg02}
-corrects an error of the abridged proof of this reduction as given in
-\cite{gj79}.  \cite{d08phd} discusses all these problems in detail
-giving an extensive survey of the previously existing results which
-are almost exhaustively all approximation results and hardness
-results. Taking this further, \cite{d08phd} presents new results in
-the area of parameterized algorithms for this
-problem\tnote{elaborate - what are the results?}.
-
-\item Another problem is to find the minimum number of entries in the matrix
-that can be toggled to result in a matrix with COP.  \cite{v85}
-discusses approximation of {\sc COP Augmentation} which is the problem
-of changing of the minimum number of zero entries to {\un}s so that
-the resulting matrix has COP. As mentioned earlier, this problem is
-known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
-using a reduction to the longest path problem, \tnote{or is it a
-  survey of another result?  check.} that finding a Tucker's forbidden
-submatrix of at least $k$ rows is NP complete. \tnote{how is this
-  different from booth's 75 result??}  \tnote{where should this go?
-  cite|tz04 (approx submatrix with COP sparse matrices)}
-
-\item \cite{jkckv04} discusses the use of matrices with almost-COP (instead
-of one block of consecutive {\un}s, they have $x$ blocks, or {\em
-  runs}, of consecutive {\un}s and $x$ is not too large) in the
-storage of very large databases.  The problem is that of reordering of
-a binary matrix such that the resulting matrix has at most $k$ runs of
-{\un}s. This is proved to be NP hard using a reduction from the
-Hamiltonian path problem.\tnote{Theorem 2.1 in jkckv} \tnote{(1) A
-  connection of COP problem to the travelling salesman problem is also
-  introduced. what does this mean? -- COP can be used as a tool to
-  reorder $0.5T \le runs(M) \le T$. (2) The optimization version of
-  the $k$-run problem, i.e. minimization of number of blocks of ones
-  is proven to be NP complete by cite:k77}\tnote{are these two the
-  same?} \tnote{what is the reduction?} \tnote{other problems similar
-  to COP -- cite:ckl96 (ILP, circ ones, one drop) -- cite:th98
-  (generalization of COP - minimax, biotonic column) Tucker}
-
-\end{enumerate}
-
-}
-
-\section{COP in Graph Isomorphism}
-\label{sec:surveygraphiso}
-\tnote{Expand on sec:appgraphiso} \tnote{(canonization theme)}
-
-\tnote[pressing]{ADD: peo exists iff chordal. lexicographic BFS
-  [tag:chordalGraph]} %
-
-\temptext{
-The survey from kklv10 conclusion.
-}
-
-% \section{Other applications}
-% {\tt If any. \tnote{ where should this go?}: (1) cite|jlm97 (application of PQ
-%   trees in graphics). (2) helly's theorem citation
-%   19XXdgk-Hellystheorem-Danzer-Gruenbaum-Klee}}
-
-\theendnotes\tnote{remove if none.}
-
-%%%%%%%%%%%%% listing package for pretty formatting BEGIN
-% \lstloadlanguages{C}
-% \lstset{language=C, commentstyle=\scriptsize,
-%   numberstyle=\tiny, numbers=left,               % Line numbers
-%   stepnumber=2,numbersep=5pt,firstnumber=10, 
-%   showstringspaces=false,                    % No explicit spaces
-%   emph={printf},emphstyle=\underbar,      % Specific formatting
-%   emph=[2]{sum},emphstyle=[2]\color{blue},      % Sp format (multiple)
-%   emph=[3]{i},emphstyle=[3]\color{red}      % Sp format (multiple)
-%   }
-
-%   \begin{singlespacing}
-%     % Text inside "lstlisting" env is verbatim.
-%     \begin{lstlisting}[keywordstyle=\textbf]      
-% int sum;
-% int i; /* for loop variable */ 
-% sum = 0; 
-% for (i=0, i<n; i++) { 
-%   sum += a[i]; 
-% } 
-% printf{"This is stupid."}
-%     \end{lstlisting}
-%   \end{singlespacing}
-
-%%%%%%%%%%%%% listing package for pretty formatting END
+\message{ !name(020chap2.tex) !offset(-17) }

### 2012-01-18 src/thesis.tex
Added more stuff to chapter 2. see diff.
chapter 2 in edit still. see diff for changes.
partative set theory TBD in chapter 2. beginning to work on chapter 3 now.
Started cleaning up chapter 3. Deferring it though to work on Chapter 2. Need to accommodate N's new proofs for removal of algorithmic charac.
See diff for other changes left out in this comment.
Fixed a bug in chapter 2 (from last commit). Endnotes in chap 2 changed to footnotes. Tucker matrices coded. Tikz is going to be the way of the graphics in Infinite Jest.
first draft of PQR trees done. next is cert algo and icpia.

@@
 
+\usepackage{../lib/mylatexlib}
 \usepackage{../lib/mythesislib}
-\usepackage{../lib/mylatexlib}
 \usepackage{thesisfiguresandtables} % Code for tables are in this
@@
                                 % in .tex file.
+%\usepackage[inline] 
+%\usepackage[margins] 
+\usepackage[finalnew]
+         {../lib/trackchanges} % For tracking changes.
+                               % :Package options:
+                               % [finalold]
+                               %   Ignore all of the edits. 
+                               %   The document will look as if
+                               %   the edits had not been added. 
+                               % [finalnew]
+                               %   Accept all of the edits. 
+                               %   Notes will not be shown in the
+                               %   final output. 
+                               % [footnotes]
+                               %   Added text will be shown
+                               %   inline. Removed text and notes
+                               %   will be shown as footnotes.  
+                               %   This is the default option.
+                               % [margins]
+                               %   Added text will be shown
+                               %   inline. Removed text and notes
+                               %   will be shown in the
+                               %   margin. Margin notes will be
+                               %   aligned with the edits when
+                               %   possible. 
+                               % [inline]
+                               %   All changes and notes will be
+                               %   shown inline.
+
+\addeditor{fyi}                % use "fyi" - for general notes with no
+%                                 action points.
+\addeditor{important}          % needs attention level 2
+\addeditor{dire}               % needs attention level infinity NOW!
+\addeditor{minor}              % least priority TBDs that
+                               %  need to go into thesis  
+\addeditor{pressing}           % needs attention level 1
 
 % \usepackage{comment}         % DEBUG ONLY
-%\includeonly{020chap2}         % DEBUG ONLY 
-%\includeonly{0021ded}         % DEBUG ONLY
+\includeonly{020chap2}         % DEBUG ONLY 
+%\includeonly{030chap3,050appa}         % DEBUG ONLY 
+%\includeonly{0031ded}         % DEBUG ONLY
 
-\def \mythesissubmissiondate {10 January 2012}
+\def \mythesissubmissiondate {13 January 2012}
 \def \mythesissubmissionmonth {\MakeUppercase {January 2012}}
@@
 %
-\title{\mythesistitleNOTSYNPOSIS} \author{\myname} \date{
-  {\mythesissubmissionmonth}}
+\title{\mythesistitleNOTSYNPOSIS \tnote[fyi]{Synopsis title is ``Generalization
+  of Consecutive Ones Property of Binary Matrices''}}% 
+\author{\myname}% 
+\date{{\mythesissubmissionmonth}}%
 \department{\MakeUppercase{\mythesisdept}}
 \maketitle
+
+\nocite{*}
 
@@
 %
-\include{0021ded}
+\include{0031ded}%{0021ded}
 
@@
 % being schizo about it now. DECIDE! ENDNOTEMARKS DEPEND ON IT.
-\setcounter{endnote}{0} % need to empty the file too. how?! -
+%\setcounter{endnote}{0} % need to empty the file too. how?! -
                         % FIXED. turns out, just having notes is
@@
 %
-% \appendix %%% NON NEIN NAHI ILLA NONE
-% \chapter{A}
-% \newpage
+\appendix %%% NON NEIN NAHI ILLA NONE. But maybe.
+\include{050appa}
+
+%\newpage
 

### 2012-01-24 src/0031ded.tex
Did some over all clean up. Added fancy chapter headings. Begun work on opt survey.

@@
 \definecolor{dim}{rgb}{0.7,0.7,0.7}
-\def \dimspace {0.125in}
+\def \dimspace {0in} %{0.125in}
+\def \dimspacemore {0.5in} %{0.125in}
+\def \authspace {-0.25in}
+\def \brightspace {0.375in}
 
 {%
+  \thispagestyle{empty}
   \fontencoding{T1}%
@@
 
-  \vglue -0.5in % this makes top margin 2in
+  \vglue -1in % this changes top margin 
 %   \vskip -5ex%
@@
 
-  \hspace{-1in}%
-  \parbox[c]{4in}{%
+  \hspace{-0.75in}%
+  \parbox[t]{6.5in}{%
+    \Large \noindent \raggedright%
+    \color{dim}{%
+      % One teacher called these moments 'mathematical
+      % experiences'. What I didn't know then was that
+      [...] a mathematical experience was aesthetic in nature, an
+      epiphany in Joyce's original sense. These moments appeared in
+      proof-completions, or maybe algorithms. Or like a gorgeously
+      simple solution you suddenly see after filling half a notebook
+      with gnarly attempted solutions. It was really an experience of
+      what I think Yeats called ``the click of a well made box''.%
+      % The word I always think of is as 'click'.
+      {%
+        \vspace{\authspace}
+        \small \par \hfill%
+        D.\,F.\,W.%
+      }%
+    }%
+  }%
+
+  \vspace{\dimspace}
+
+  \hspace{1in}%
+  \parbox{4in}{ }{%
+    \LARGE 
+    \raggedleft \noindent%
+    \color{dim}{%
+      Math Graffiti: Kilroy wasn't Haar.  Free the group.  Nuke the kernel.  Power to
+      the $n$.  $N = 1 \Rightarrow P = NP$ %
+      {%
+%        \vspace{\authspace}%
+        \small \par \hfill%
+        \emph{Concrete Mathematics (margin notes)}%
+      }%
+    }%
+  }%
+
+  \vspace{\dimspace}
+
+  \hspace{-1.25in}%
+%  \ding{125}
+  \parbox[t]{5in}{%
     \large \raggedright \noindent %
@@
       {%
+%        \vspace{\authspace}%
         \small \par \hfill \emph{The Art of Computer Programming},
         D.\,E.\,K.%
+      }%
+    }%
+  }%
+
+  \vspace{\brightspace}
+
+  {%
+      \centering%
+      \hspace{-1.25in}
+      \parbox{8in}{%
+      \fontsize{36}{1.5}\selectfont%
+      \centering \noindent%
+      \emph{L'art c'est la solution au chaos.}%
+      \par \noindent%
+      \normalsize (Art is the solution to chaos.)%
+    }%
+  }%
+
+  \vspace{\brightspace}
+
+ \hspace{-0.25in}%
+  \parbox[t]{6.5in}{%
+    \LARGE  \noindent \raggedleft%
+    \color{dim}{%
+      \textbf{myth\textperiodcentered os} {|{\huge $'$mi $\theta$
+          $\overline{o}$s}|} \\
+      a set of beliefs or assumptions about something
+      \emph{ :\\ 
+      the rhetoric and mythos of science create the comforting image
+      of linear progression toward truth.}%
+      {%
+%        \vspace{\authspace}%
+        \small \par \hfill%
+        \emph{New Oxford American Dictionary, 2nd
+          Ed.}%
+     }%
+    }%
+  }
+
+  \vspace{\dimspace}
+
+  \hspace{-1in}%
+  \parbox{6in}{%
+    \singlespacing
+    \large \raggedright \noindent%
+    \color{dim}{%
+      Anything that happens, happens.  Anything that, in happening,
+      causes something else to happen, causes something else to
+      happen.  Anything that, in happening, causes itself to happen
+      again, happens again.  It doesn't necessarily do it in
+      chronological order, though.%
+      {%
+%       \vspace{\authspace}%
+        \small \par \hfill%
+        \emph{The Salmon of Doubt}, D.\,N.\,A.%
       }%
@@
   \hspace{1.5in}%
-  \parbox[t]{4in}{%
-    \Large \noindent \raggedleft%
-    \color{dim}{%
-      The rhetoric and mythos of science create the comforting image
-      of linear progression toward truth.%
-      {%
-        \small \par \hfill%
-        \textbf{mythos}; \emph{New Oxford American Dictionary, 2nd
-          Ed.}%
-      }%
-    }%
-  }
-  \\
-  \vspace{\dimspace}
-
-  % \hspace{1in}
-  {\centering
-    \parbox{6in}{%
-%      \Huge 
-      \fontsize{27}{1.5}\selectfont
-      \centering \noindent %
-%      \textbf{
-      \emph{L'art c'est la solution au chaos.} \par \noindent
-      \noindent (Art is the solution to chaos.)
-%      }
-%       {%
-%         \small \par \hfill%
-%         \emph{citation needed.}%
-%       }%
-    }%
-  }%
-
-  \vspace{\dimspace}
-
-  \hspace{-0.25in}%
-  \parbox{2.5in}{%
+  \parbox{4in}{%
+    %\huge 
     \Large 
-    \raggedright \noindent%
-    \color{dim}{%
-      Kilroy wasn't Haar.  Free the group.  Nuke the kernel.  Power to
-      the $n$.  $N = 1 \Rightarrow P = NP$ %
-    }%
-    {%
-      \small \par \hfill%
-      \emph{Concrete Mathematics}%
-    }%
-  }%
-  \vspace{\dimspace}
-
-  \hspace{1in}%
-  \parbox{5in}{%
-    \large \raggedleft \noindent%
-    \color{dim}{%
-      Anything that happens, happens.  Anything that, in happening,
-      causes something else to happen, causes something else to
-      happen.  Anything that, in happening, causes itself to happen
-      again, happens again.  It doesn't necessarily do it in
-      chronological order, though.  }%
-    {%
-      \small \par \hfill%
-      \emph{The Salmon of Doubt}, D.\,N.\,A.%
-    }%
-  }%
-  \vspace{\dimspace}
-
-  \hspace{-0.50in}%
-  \parbox{4in}{%
-    \huge \noindent \raggedright%
+    \noindent \raggedleft%
     \color{dim}{%
@@
       {%
+%        \vspace{\authspace}%
         \small \par \hfill%
@@
   }%
+  % xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx
+  % xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx
+  % xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx
+  % xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx
+  % xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx
+  % xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx
+  % xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx
+  % xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx
+
+  \vskip -02.5in % this changes bottom margin
 

### 2012-01-24 src/010chap1.tex
Did some over all clean up. Added fancy chapter headings. Begun work on opt survey.

@@
 \label{sec:orgofdoc}
-This chapter (Chapter~\ref{ch:intro}) introduces the area of research
-and the problems addressed in this thesis.  Section~\ref{sec:problem}
-introduces the main problem of this thesis by way of an
+Chapter~\ref{ch:intro} introduces the area of research and the
+problems addressed in this thesis.  Chapter~\ref{ch:copsurvey} gives a
+more detailed survey briefed in Section~\ref{sec:background}.
+Chapter~\ref{ch:myresearch} details all the results obtained to the
+problems of this thesis and finally the conclusion of the thesis is
+discussed in Chapter~\ref{ch:conclusion}.
+
+In this chapter, Section~\ref{sec:problem} introduces the main problem
+of this thesis by way of an
 illustration. Section~\ref{sec:basicprelim} lays out a few general
 definitions that are helpful in understanding the rest of the chapter.
-Section~\ref{sec:background} gives a brief survey of COP
-and optimization problems related to it followed by motivation for the
+Section~\ref{sec:background} gives a brief survey of COP and
+optimization problems related to it followed by motivation for the
 thesis in Section~\ref{sec:motive}.  Section~\ref{sec:results}
@@
 tree path labeling problem.
-
-Chapter~\ref{ch:copsurvey} gives a more detailed survey briefed in
-Section~\ref{sec:background}.  Chapter~\ref{ch:myresearch} details all
-the results obtained to the problems of this thesis and finally the
-conclusion of the thesis is discussed in Chapter~\ref{ch:conclusion}.
 
@@
 part of their semester thesis, they pick a body of work to study and
-form the namesake study groups, {\LLL}, {\GGG}, {\BBB} and
-{\TTT}\tnote[minor]{put bib entries for these works!}. A student will be
+form the namesake study groups, {\LLL} \cite{wallace99}, {\GGG} \cite{wallace96}, {\BBB}
+\cite{wallace00} and
+{\TTT} \cite{wallace10}. A student will be
 in at least one study group and may be in more than one. For instance,
@@
 
-\begin{table}[t] %[htbp]
+\begin{table}[b] %[htbp]
   \centering
@@
 
-%D<debug> figures inside tables
-% \begin{tabular}{cc}
-%  \begin{figure} %[t]%
-% %    \centering
-%     \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
-% %    \label{fig:streetmap}
-%     \caption{--placeholder--}
-%   \end{figure}
-%     &
-%   \begin{figure} %[t]%
-% %    \centering
-%     \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
-% %    \label{fig:streetmappath}
-%     \caption{--placeholder--}
-%   \end{figure}
-% \end{tabular}
+% <debug> figures inside tables -- doesn't work. see git history to
+% know what was tried.
 
@@
   \centering
-  \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
+  \vspace{-10mm}
+  \includegraphics[scale=0.45]{../img/1_infinite_loop__cropped.pdf}
   \caption{\figtabsize {\residenceblock} street map.}
@@
 
-  \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
+  \includegraphics[scale=0.45]{../img/2_infinite_loop_BTWF__cropped.pdf}
   \caption[\figtabsize \residenceblock street map with study group
     routes allocated.]{\figtabsize {\residenceblock} street map with study group
-    routes allocated. Routes are color coded as follows: red for
-    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
-    group, orange for \textcolor{yellow}{$\xBBB$} group, green for
-    \textcolor{green}{$\xTTT$} group}
-  \label{fig:streetmappath}
-%   \caption{\figtabsize (a) {\residenceblock} street map (b)
-%     {\residenceblock} street map with study group routes
-%     allocated. Routes are color coded as follows: red for
+    routes allocated. % Routes are color coded as follows: red for
 %     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
 %     group, orange for \textcolor{yellow}{$\xBBB$} group, green for
-%     \textcolor{green}{$\xTTT$} group}
-%  \label{fig:streetmap}
+%     \textcolor{green}{$\xTTT$} group
+  }
+  \label{fig:streetmappath}
 
-  \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
+  \includegraphics[scale=0.5]{../img/3_infinite_loop.pdf}
   \caption[\figtabsize Solution to the student accommodation
@@
     students in {\residenceblock} that meets the requirements stated
-    before.  The routes are color coded as follows: red for
-    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
-    group, orange for \textcolor{yellow}{$\xBBB$} group, green for
-    \textcolor{green}{$\xTTT$} group.\\ {\tiny {\em Peanuts images
+    before.  % The routes are color coded as follows: red for
+%     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
+%     group, orange for \textcolor{yellow}{$\xBBB$} group, green for
+%     \textcolor{green}{$\xTTT$} group.
+    \\ {\tiny {\em Peanuts images
         {\copyright} Charles Schulz}}}% YellowOrange
   \label{fig:streetmappathpeople}
-\end{figure} \tnote{[i] UPDATE IMAGE (b) 2-infinite-loop-BTWF.png. REMOVE
-  TITLES. ADD {$\xLLL$} {$\xGGG$} {$\xBBB$} {$\xTTT$} [ii]make
-  dashed/textured lines for routes. make it color agnostic.}
+\end{figure}
 
@@
 system $\cF$, a tree $T$, and an assignment of paths from $T$ to the
-sets, there is a feasible\endnote{The notion of {\em feasibility} is
+sets, there is a feasible\footnote{The notion of {\em feasibility} is
   formally defined in Section~\ref{ch:prelims}.}  bijection between
@@
 filtering algorithm is also an efficient algorithm to test if a tree
-path labeling\endnote{The terms {\em tree path labeling} and {\em
+path labeling\footnote{The terms {\em tree path labeling} and {\em
     tree path assignment} are, in informal language,
@@
 cardinality preservation is necessary and sufficient for an interval
-assignment to be feasible for a given hypergraph\endnote{A {\em
+assignment to be feasible for a given hypergraph\footnote{A {\em
     hypergraph} is an alternate representation of a set system and
-  will be used in this thesis.}~\endnote{See Section~\ref{ch:prelims} for the
+  will be used in this thesis.}~\footnote{See Section~\ref{ch:prelims} for the
   formal definition.} and thus is a characterization for COP
@@
 and find that trio-wise intersection cardinality preservation makes a
-tree path labeling\endnote{A {\em tree path labeling} $\cl$ is a
+tree path labeling\footnote{A {\em tree path labeling} $\cl$ is a
   bijection of paths from the target tree $T$ to the hyperedges in
-  given hypergraph $\cF$.}~\endnotemark[4]  (TPL) feasible, which is a generalization of the COP
+  given hypergraph $\cF$.}~\footnotemark[4]  (TPL) feasible, which is a generalization of the COP
 problem. This problem is defined as follows.
@@
 captures the trio-wise cardinality property described
-earlier\endnote{See Section~\ref{sec:feasible} for the definition of
+earlier\footnote{See Section~\ref{sec:feasible} for the definition of
   ICPPL.}. This characterization can be checked in polynomial time.  A
@@
 of trees which are in many ways very close to intervals or paths. Each
-ray\endnote{The path from a leaf to the root, the vertex with highest
-  degree, is called a {\em ray} of the \kstar.}~\endnotemark[4] are independent except for the
-root\endnote{The vertex with maximum degree in a {\kstar} is called
-  {\em root}.}~\endnotemark[4] and hence can be
+ray\footnote{The path from a leaf to the root, the vertex with highest
+  degree, is called a {\em ray} of the \kstar.}~\footnotemark[4] are independent except for the
+root\footnote{The vertex with maximum degree in a {\kstar} is called
+  {\em root}.}~\footnotemark[4] and hence can be
 considered as an independent interval till the root. Our algorithm
@@
   definition of ``total order'' in footnote?}. We extend this to TPL
-and find that when $\cF$ is a path hypergraph\endnote{If there exists
+and find that when $\cF$ is a path hypergraph\footnote{If there exists
   an FTPL for a hypergraph $\cF$, it is called a path hypergraph.},
@@
 
-\theendnotes
+%\theendnotes

### 2012-01-24 src/020chap2.tex
Did some over all clean up. Added fancy chapter headings. Begun work on opt survey.

@@
 
-\chapter{\CoP\,-- A Survey of Important Results}
+%\chapter{\CoP\,-- A Survey of Important Results}
+\chapter{Consecutive-ones Property -- A Survey of Important Results}
 \label{ch:copsurvey}
@@
 having seen how useful \COP is in identifying or characterizing many
-types of graphs\footnote{\COP has several other applications
-  \temptext{fill up.}}, we will now see results that study recognition of \COP
+types of graphs, we will now see results that study recognition of \COP
 in matrices in the following section.
 
-\temptext{ tab~\ref{tab:graphmatrices} -  have an abridged version
-  of table 2.1 in dom. see notes for the possibility of a
-  ``circle diagram''.} 
-\begin{table}[htbp]
-  \centering
-% \includegraphics[scale=0.1]{../img/circlediagram_domtab2.1__2.png} %
- \caption{\figtabsize Graph matrices \temptext{PLACEHOLDER}}
-  \label{tab:graphmatrices}
-\end{table}
+\tabgraphmatrices
 
@@
 It may be noted that one way to design an algorithm to test for \COP
-is by deriving one from any interval graph recognition algorithm using the
-result {\sc hmpv00}\tnote{cite hmpv00} \tnote{in footnote put the theorem 2.7 - dom
-  pg 43} \cite{d08phd} which demonstrates how such a derivation can be
-done. However, this does not necessarily yield an efficient
-algorithm. We will see results that directly solve the problem on
-matrices since it is known that questions~(\ref{q:testcop}) and
-(\ref{q:copperm}) stated above for \COP are efficiently solvable.
+is by deriving one from any interval graph recognition algorithm using
+the result {\sc hmpv00}\tnote{cite hmpv00} \tnote{in footnote put the
+  theorem 2.7 - dom pg 43} \cite{d08phd} which demonstrates how such a
+derivation can be done. However, this does not necessarily yield an
+efficient algorithm. We will see results that directly solve the
+problem on matrices since it is known that questions~(\ref{q:testcop})
+and (\ref{q:copperm}) stated above for \COP are efficiently solvable.
 Table~\ref{tab:cophistory} gives a snapshot of these results.
 
-\tnote[important]{1899 by petrie -- cite
-         kendall 69 pacific journal of mathematics 1969 }
-
-\def\tabparboxw{10cm}
-\begin{table}[htbp]
-  \centering
-  {
-    \fontencoding{OT1}\fontfamily{cmss}\selectfont%
-    \footnotesize%
-
-    \begin{tabular}{lll}
-      \hline
-      \hline
-
-      1899 & \parbox[t]{\tabparboxw}{ First mention of \COP
-        (archaeology)} &                  \\
-
-      \rowcolor[gray]{0.8} %
-      1951 & \parbox[t]{\tabparboxw}{ Heuristics for COT} & \cite{wsr51} \\
-
-      1965 & \parbox[t]{\tabparboxw}{ First polynomial time algorithm
-        for \COP testing} & \cite{fg65} \\
-
-      \rowcolor[gray]{0.8} %
-      1972 & \parbox[t]{\tabparboxw}{ Characterization for \COP -- forbidden
-        submatrices} &
-      \cite{at72} \\
-
-      1976 & \parbox[t]{\tabparboxw}{ First linear time algorithm for COT --
-        \PQtree} &
-      \cite{bl76} \\
-
-      \rowcolor[gray]{0.8} %
-      1992 & \parbox[t]{\tabparboxw}{ Linear
-        time algorithm COT without \PQtree} & \cite{wlh02}%\footnote{First
-      % published in \cite{wlh02isaac92}}
-      \\
-
-      2001 & \parbox[t]{\tabparboxw}{\PCtree\ -- a simplification of
-        \PQtree} & \cite{wlh01,hm03} \\
-
-      \rowcolor[gray]{0.8} %     
-      1996 & \parbox[t]{\tabparboxw}{\PQRtree\ -- generalization of
-        \PQtree for any binary matrix regardless of its \COP status} &
-      \cite{mm96} \\
-
-      1998 & \parbox[t]{\tabparboxw}{ Almost linear time to construct
-        \PQRtree} & \cite{mpt98} \\
-
-      \rowcolor[gray]{0.8} %
-      2004 & \parbox[t]{\tabparboxw}{ A certifying algorithm for no
-        \COP.\\ Generalized \PQtree.} & \cite{mcc04} \\
-
-      2009 & \parbox[t]{\tabparboxw}{ Set theoretic, cardinality based
-        characterization of \COP\ -- ICPIA} & \cite{nsnrs09} \\
-
-      \rowcolor[gray]{0.8} %
-      2010 & \parbox[t]{\tabparboxw}{ Logspace \COP testing} &
-      \cite{kklv10}\\ 
-      \hline
-      \hline
-    \end{tabular}
-  }
-\caption[\figtabsize A brief history of \COP research]{\figtabsize A
-  brief history of \COP research }
-\label{tab:cophistory}
-\end{table}
+\tabcophistory
 
@@
 \subsection{Tucker's forbidden submatrices for \COP}
-
+\label{sec:tucker}
 % \change{While it is remarkable that the first polynomial time
@@
 
-\begin{theorem}
-  (\cite[Th.~6]{at72}, \cite[Th.~2.3]{d08phd})\\
+\begin{theoremsansproof}
+  [{\cite[Th.~6]{at72}, \cite[Th.~2.3]{d08phd}}]%\\
   A bipartite graph $G = (V_1, V_2, E)$ is convex bipartite on
@@
   \label{th:tuckeratfree}
-\end{theorem}
-
-\begin{theorem}
-  (\cite[Th.~7]{at72}, \cite[Th.~2.4]{d08phd})\\
+\end{theoremsansproof}
+
+\begin{theoremsansproof}
+  [{\cite[Th.~7]{at72}, \cite[Th.~2.4]{d08phd}}]%\\
   In a bipartite graph $G = (V_1, V_2, E)$ the vertex set $V_1$
@@
   \label{th:tuckerforbidden}
-\end{theorem}
+\end{theoremsansproof}
 
@@
 
-\begin{theorem}
-  (\cite[Th.~9]{at72}, \cite[Th.~2.5]{d08phd})\\
+\begin{theoremsansproof}
+  [{\cite[Th.~9]{at72}, \cite[Th.~2.5]{d08phd}}]%\\
   A matrix $M$ has \COP \iff it contains none of the matrices 
@@
   \label{th:tuckercop}
-\end{theorem}
-
-\begin{figure}[htbp]
-  \centering
-%  \includegraphics[scale=0.35]{../img/tuckersforbiddensubmatrices.png}  
-  \setlength\dashlinedash{1pt} %
-  \setlength\dashlinegap{0pt}  %  Make non zero if cline can be done
-                               %   with dashed lines.
-  \singlespacing
-  \begin{tabular}[h]{ll}
-%   \small 
-   \begin{tabular}[t]{|lll ll l|l} %                   MATRIX 1
-      \multicolumn{7}{l}{\normalsize \hspace{-0.5in} $M_{I_k}, k \ge 1$}              \\   
-      \multicolumn{6}{c}{\matrixbracetop{3.25}{$k+2$}}   \\
-      \cline{1-6} % \firsthline %      Begin matrix  %%%%%%
-      \un & \un & 0   & \multicolumn{2}{c}{\dots}  & 0    &
-        \hspace{-0.1in}\multirow{5}{*}{\matrixbraceright{2.5}{$k+2$}}   \\    
-      0   & \un & \un & 0 & ...& 0   \\
-      \multicolumn{6}{|c|}{\dots}                        \\
-      0   & \multicolumn{2}{c}{\dots}  & 0   & \un & \un \\
-      \cline{1-6} % \hdashline %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-      \un & 0   & \multicolumn{2}{c}{\dots}  & 0   & \un \\
-      \cline{1-6} % \lasthline %      End matrix  %%%%%%%%%
-    \end{tabular}
-    &%
-%    \small
-    \begin{tabular}[t]{|lll ll ll|l} %                   MATRIX 2
-      \multicolumn{8}{l}{\normalsize \hspace{-0.5in}$M_{II_k}, k \ge 1$}                   \\
-      \multicolumn{7}{c}{\matrixbracetop{3.8}{$k+3$}}            \\ 
-      \cline{1-7}%  \firsthline %      Begin matrix  %%%%%%%%%%%%
-      \un & \un & 0   & \multicolumn{3}{c:}{\dots}       & 0    &
-        \hspace{-0.1in}\multirow{6}{*}{\matrixbraceright{3}{$k+3$}}         \\
-      0   & \un & \un & 0 & \multicolumn{2}{c:}{\dots}   & 0   \\ 
-      \multicolumn{6}{|c:}{\dots}                        &     \\
-      0   & \multicolumn{2}{c}{\dots}  & 0   & \un              &
-        \multicolumn{1}{l:}{\un} & 0                           \\
-      \cline{1-6}    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-      0   & \un & \multicolumn{4}{c}{\dots}              & \un \\
-      \un & \multicolumn{3}{c}{\dots}        & \un & 0   & \un \\
-      \cline{1-7}% \lasthline %      End matrix  %%%%%%%%%%%%%%%%
-    \end{tabular}
-    \\\\%
-%    \small
-    \begin{tabular}[t]{|lll ll l l|l} %                  MATRIX 3 
-      \multicolumn{8}{l}{\normalsize \hspace{-0.5in}$M_{III_k}, k \ge 1$}                  \\
-      \multicolumn{7}{c}{\matrixbracetop{3.8}{$k+3$}}            \\
-      \cline{1-7}%      \firsthline %      Begin matrix %%%%%%%%%
-      \un & \un & 0   & \multicolumn{3}{c:}{\dots}       & 0    &
-        \hspace{-0.1in}\multirow{5}{*}{\matrixbraceright{2.5}{$k+2$}}         \\
-      0   & \un & \un & 0 & \multicolumn{2}{c:}{\dots}   & 0   \\ 
-      \multicolumn{6}{|c:}{\dots}                        &     \\
-      0   & \multicolumn{2}{c}{\dots}  & 0   & \un              &
-        \multicolumn{1}{l:}{\un}                         & 0   \\
-      \cline{1-6}    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-      0   & \un & \multicolumn{2}{c}{\dots}  & \un & 0   & \un \\
-      \cline{1-7}%      \lasthline %      End matrix %%%%%%%%%%%%
-    \end{tabular}
-    &%
-%    \small
-    \begin{tabular}[t]{|llllll|}  %                     MATRIX 4
-      \multicolumn{6}{l}{\normalsize \hspace{-0.5in}$M_{IV}$} \\
-      \multicolumn{6}{l}{} \\
-      \firsthline
-      \un & \un & 0 & 0 & 0 & 0   \\
-      0 & 0 & \un & \un & 0 & 0   \\
-      0 & 0 & 0 & 0 & \un & \un   \\
-      \un & 0 & \un & 0 & \un & 0 \\
-      \lasthline
-    \end{tabular}
-    \\\\%
-%    \small
-    \begin{tabular}[t]{|lllll|}  %                     MATRIX 5
-      \multicolumn{5}{l}{\normalsize \hspace{-0.5in}$M_{V}$} \\
-      \multicolumn{5}{l}{} \\
-      \firsthline
-      \un & \un & 0   & 0   & 0   \\
-      0   & 0   & \un & \un & 0   \\
-      \un & \un & \un & \un & 0   \\
-      \un & 0   & \un & 0   & \un \\
-      \lasthline
-    \end{tabular}
-  \end{tabular}
-
-  \caption[\figtabsize Tucker's forbidden submatrices]{\figtabsize Tucker's
-    forbidden submatrices for convex bipartite graphs. \cite{at72}} 
-  \label{fig:forbiddensubmatrices}
-\end{figure}
+\end{theoremsansproof}
+
+\figforbiddensubmatrices
 
@@
   {\em binary matrix for set system} $\cF$ is conversely constructed
-  and denoted by $M^\cF$. Thus, $M^{\cF_M} = M$.
-
+  and denoted by $M^\cF$. Thus, $M^{\cF_M} = M$.%
+  \par\noindent%
   A set system $\cF$ from universe $U$, $|U| = n$ has the
@@
   w_{i+k-1}$ on $\sigma$ for some positive integer $i \le n+1-k$ where  $k = |S|$.
-
+  \par\noindent%
   The set $\valid{U,\cF}$ respresents all \COP orders of $\cF$ in $U$.
-  \defnstop
+  \dstop
 \end{definition}
@@
       this the {\em weak closure} of $\cF$ denoted by $\cW(\cF)$. \label{mcc3}}
-
-  \end{enumerate}
-  \defnstop
+  \end{enumerate}
+  \dstop
 \end{definition}
@@
 
-\begin{theorem}[{\cite[Th.~3,6]{mm96}}]
+\begin{theoremsansproof}[{\cite[Th.~3,6]{mm96}}]
   \label{th:validcop}
@@
     $\cF^{\bot} = \overline{\cF^{\bot}} = (\overline{\cF})^{\bot}$
-\end{theorem}
+\end{theoremsansproof}
 
@@
 
-\begin{theorem}[{\cite[Th.~7]{mm96}}]
+\begin{theoremsansproof}[{\cite[Th.~7]{mm96}}]
   \label{th:mmdecomp}
@@
   $\valid{U,\cF} = \valid{U/H,\cF/H} * \valid{H,\cF \cap 2^H}$
-\end{theorem}
+\end{theoremsansproof}
 
@@
 subproblems in \cite{mm96} is very similar to the substitution
-decomposition of a set system given in \cite{mcc04}.\footnote{Using
+decomposition of a set system given in \cite[Sec.~4]{mcc04}.\footnote{Using
   the theory cited in footnotes~\ref{mcc1},\ref{mcc2},\ref{mcc3}.}
@@
 
-\begin{theorem}[{\cite[Th.~9]{mm96},\cite[Th.~2.1]{mcc04}}]
+\begin{theoremsansproof}[{\cite[Th.~9]{mm96},\cite[Th.~2.1]{mcc04}}]
 \label{th:pqrfinalmm}
@@
 \end{enumerate}
-\end{theorem}
+\end{theoremsansproof}
 
@@
 
-\begin{theorem}[{\cite{mm96}, \cite[Th.~3.6]{mcc04}}]
+\begin{theoremsansproof}[{\cite{mm96}, \cite[Th.~3.6]{mcc04}}]
 The set system $\cF$ has \COP \iff its \PQRtree has no $R$ nodes.  
-\end{theorem}
-
-% The generalized \PQtree of $\cF$ is defined as the {\em decomposition
-%   tree} of its weak closure $\cW(\cF)$. Decomposition tree is the
-% inclusion tree (inclusion p.o. Hasse diagram) of the strong elements
-% of $\cW(\cF)$ with labels {\em prime, linear} and {\em degenerate}.\cite{mcc04}
-
-% Overlap components are computable in linear time \cite{mm95,wlh92}.
-
------ Cert algo
-
------- ICPIA
-
-\temptext{ Certifying algorithm (Generalized PQ trees) [McC04] --
-
-Set theoretic characterizations [Hsu02, NS09]\\
-    \cite{wlh02} describes the simpler algorithm for
-    COT.\tnote[pressing]{in terms of what?}\\
-    \cite{nsnrs09} describes a characterization of consecutive-ones
-    property solely based on the cardinality properties of the set
-    representations of the columns (rows); every column (row) is
-    equivalent to a set that has the row (column) indices of the rows
-    (columns) that have one entries in this column (row). This is
-    interesting and relevant, especially to this thesis because it
-    simplifies COT to a great degree by reducing the solution search
-    space owing to the a simple set theoretic characterization.\\
-
-\cite{mcc04} describes a different approach to COT. While all previous
-COT algorithms gave the COP order if the matrix has the property but
-exited stating negative if otherwise, this algorithm gives an evidence
-by way of a certificate of matrix even when it has no COP. This
-enables a user to verify the algorithm's result even when the answer
-is negative. This is significant from an implementation perspective
-because automated program verification is hard and manual verification
-is more viable. Hence having a certificate reinforces an
-implementation's credibility. Note that when the matrix {\em has} COP,
-the COP order is the certificate.  The internal machinery of this
-algorithm is related to the weighted betweenness problem
-addressed\tnote{in what way??} in \cite{co98}.  \tnote{expand
-  on the COP order graph creation and it having to be bipartite for M
-  to have COP. and thus an odd cycle being an evidence of no COP.}
-
+\end{theoremsansproof}
+
+Thus \PQRtree gives a data structure that encodes possible linear
+orderings of a set that demonstrates the \COP property in it or
+narrows it down to parts of the universe in $R$ nodes that prevent the
+set system from having \COP.
+
+As mentioned before, we observe the three data strutures of \PQRtree,
+\gPQtree and generalized \PQtree to be equivalentand the theory of all
+these data structures is summarised in Table~\ref{tab:pqrcomparison}.
+
+We will now see one more generalization of \PQtree before seeing other
+approaches to solving \COP testing including ICPIA which is a set
+theoretic characterization of \COP.
+
+\begin{table}[t]
+  \def\colwidth{4cm}
+  \centering
+  \fontencoding{OT1}\fontfamily{cmss}\selectfont%
+  \footnotesize%
+  \setlength\extrarowheight{0.1in}
+
+  \begin{tabular}{l >{\columncolor{\tblhcolor}}l l}
+
+    \rowcolor[gray]{0.8} %                 Header row colored
+    \normalsize \gPQtree \cite{n89} & 
+    \normalsize \PQRtree \cite{mm96} & 
+    \normalsize generalized \PQtree \cite{mcc04}
+    \\
+
+    Trivial intersections &
+    Trivial sets $\cT(U)$ & 
+    \\\hline
+
+    Trivially intersecting sets &
+    Mutually orthogonal sets, $A \bot B$ &
+    Non-overlapping sets
+    \\\hline
+
+      &
+    Complete collection &
+    Weakly partative family
+    \\\hline
+
+    &
+    $\overline{\cF}$&
+    Weak closure, $\cW(\cF)$
+    \\\hline
+
+    &
+    $\cF^{\bot}$&
+    Non-overlapping family, $\cN(\cF)$
+    \\\hline
+
+    &
+    \parbox[t]{\colwidth}
+    { 
+      Sets in $\cF$ orthogonal to all other sets in $\cF$ 
+    }&
+    Strong elements
+    \\\hline
+
+    &
+%     \parbox[t]{\colwidth}
+%       {}
+    \temptext{PQR tree def. }
+    &
+    \parbox[t]{\colwidth}
+    {
+      Decomposition tree of $\cW(\cF)$, $T(\cW(\cF))$
+      \temptext{PQR tree def. }\tnote[pressing]{in terms of theory}
     }
+    \\\hline
+
+    &
+    Node sets, $(\cF \cap \cF^\bot) \setminus \{\emptyset\}$
+    \\\hline
+
+    &
+    \parbox[t]{\colwidth}
+    {
+      $P$-node $\Leftrightarrow \cF = \cT(\cF)$\\
+      $Q$-node $\Leftrightarrow \cF = consec(\alpha)$\\
+      $R$-node $\Leftrightarrow \cF = \power{U}$\\
+    }&
+    \temptext{give def of P Q R nodes}
+    \\\lasthline
+
+  \end{tabular}
+  \caption[\figtabsize Comparison of theory of \PQRtree, \gPQtree, generalized
+  \PQtree]{\figtabsize Comparison of theory of \cite{mm96} \PQRtree, \cite{n89} \gPQtree,
+    \cite{mcc04} generalized \PQtree}
+  \label{tab:pqrcomparison}
+\end{table}
 
@@
 
-\PCtree is a generalization of \PQtree. It is a data structure that is
+\PCtree is another generalization of \PQtree. It is a data structure that is
 analogous to \PQtree but for matrices with \crop. \PCtree was
@@
 
+\subsection{ICPIA - a set cardinality based \COP test}
+
+\cite{nsnrs09} describes a characterization of consecutive-ones
+property solely based on the cardinality properties of the sets in the
+set system and does not use any variants of \PQtrees.
+
+\begin{definition}[Intersection Cardinality Preserving Interval
+  Assignment (ICPIA)]
+  \label{def:icpia}
+  Let $\cF = \{A_i \mid A_i \subseteq U, i \in [m]\}$ be a set system
+  from universe $U$ and the set of ordered pairs $\Pi = \{(A_i, B_i)
+  \mid B_i  \text{ is an interval from } [n], i \in [m]\}$ be an {\em interval
+    assigment} of $\cF$, then it is called an {\em ICPIA} if it has
+  the following properties.
+  \begin{enumerate}[i.]
+  \item $|A_i| = |B_i|$ for all $i \in [m]$
+  \item $|A_i \cap A_j| = |B_i \cap B_j|$ for all $i, j \in [m]$
+  \end{enumerate}
+\end{definition}
+
+\begin{theoremsansproof}[{\cite[Th.~1]{nsnrs09}}]
+  \label{th:icpia1}
+  If an interval assignment $\Pi$ is feasible, then it is an ICPIA.
+\end{theoremsansproof}
+
+The necessity of ICPIA for a set system to have \COP given in
+Theorem~\ref{th:inpia1} is fairly obvious. It turns out that it is
+sufficient too.  This is demostrated by two algorithms which work as
+follows. The first one iterates over the assignment
+$\Pi$ %set system $\cF$
+and changes it in each iteration till the altered
+assignment %set system
+has no more overlapping intervals\footnote{Intervals are the second
+  element in each ordered pair of $\Pi$. Def.~\ref{def:icpia}} %sets
+\cite[Alg.~1]{nsnrs09}. This is acheived by replacing two overlapping
+assignment pairs $(A_i, B_i)$ and $(A_j, B_j)$ with the partitions of
+$(A_i \cup A_j,B_i \cup B_j)$ induced by their overlaps -- $(A_i
+\setminus A_j, B_i \setminus B_j)$, $(A_i \cap A_j, B_i \cap B_j)$ and
+$(A_j \setminus A_i, B_j \setminus B_i)$.
+
+To see it in terms of the theory explained in
+Section~\ref{sec:surveycertalgo}, this results in the weak closure of
+$\cF$. Let $\cF$, $\cI$ be the set system and interval
+system\footnote{A set system where all the sets are intervals.},
+respectively, involved in the assignment $\Pi$ and $\cF'$, $\cI'$ be
+the same for the output of the algorithm $\Pi'$.  It can be observed
+that $\cF' \cup \cF$ is $\overline{\cF}$ or the weak closure of
+$\cF$. The analog is true for $\cI \cup \cI'$.  Moreover, this output
+$\cF'$, $\cI'$ contains the strong elements of $\overline{\cF}$,
+$\overline{\cI}$ respectively. It turns out that $\cI'$ is also an
+interval system and that $\Pi'$ is an ICPIA \cite[Lem.~2]{nsnrs09}.
+
+The next algorithm \cite[Alg.~2]{nsnrs09} further refines $\Pi'$ to
+$\Pi''$ which represents the family of permutations that yield the
+\COP orders represented by $\Pi$ of $\cF$. The subset relation Hasse
+diagram of $I'$ is created with each node representing the
+corresponding assignment pair of the interval from $\Pi'$. Since all intervals in
+$\Pi'$ are non-overlapping, this is a tree. Notice that this means any
+two intervals are now either disjoint or one is contained in the
+other. Hence this tree is called a {\em containment tree}.  The
+algorithm traverses this tree in post order fashion where the post
+order function is to replace an assignment pair $(X,Y)$  in $\Pi'$ by their child
+assignment pairs in the containment tree, say $(X_1, Y_1),(X_2, Y_2)
+\ldots (X_k,Y_k)$. The result $\Pi''$ gives the
+\COP order for $\Pi$. These two algorithms prove that ICPIA is a
+sufficient condition for the feasibility of the interval assignment.
+
+\begin{theoremsansproof}[{\cite[Th.~2]{nsnrs09}}]
+  \label{th:icpia2}
+  Let  $\cF = \{A_i \mid A_i \subseteq U, i \in [m]\}$ be a set system
+  from universe $U$, $|U| = n$
+  and $\Pi = \{(A_i,B_i) \mid B_i \text{ is an interval from } [n], i
+  \in [m]\}$ be an ICPIA. Then there exists a permutation $\sigma: [n]
+  \rightarrow [n]$ \stt $\sigma(A_i) = B_i$.
+\end{theoremsansproof}
+
+What we saw so far in this section is checking for the feasibility of
+a given interval assignment. ICPIA can also be used to find an
+interval assignment when only the set system $\cF$ is given. Part of
+the approach is similar to the overlap component idea in \cite{fg65,
+  wlh02}. Once the overlap components are factored out, they can be
+independently assigned a subinterval from to which the sets in the
+component need to be mapped. This is simple because the overlap
+components are disjoint from each other. Now the subproblem is finding
+\COP order for each overlap component. In each component, the set
+whose intersections with the rest of the sets in that component forms
+a single inclusion chain is chosen first and assigned the
+leftmost\footnote{By convention. Rightmost can also be chosen
+  analogously to get a different \COP order.} interval. The next
+candidate set is chosen from this set's overlapping sets. The interval
+assigned is calculated by simple intersection cardinality
+means\footnote{Assume that the next set overlaps on one of the
+  ``ends'' of the current set and calculate the interval by shifting
+  left or right of the current interval and adjusting its size to
+  match the next set's cardinality.}. Once the next set is assigned an
+interval, the current assigned sets are all checked for ICPIA. This is
+a backtracking algorithm -- if at any point ICPIA fails, another
+overlapping set is chosen. If all overlapping set attempts fail ICPIA,
+a different ``first set'' is chosen. This is a remarkably simple
+algorithm due to its obvious simplicity in implementation.
+
+\subsection{Other \COP testing algorithms}
+There are more \COP test algorithms in the literature. We make a
+mention of one of them here.  A simple linear time algorithm was
+presented by \cite{hmpv00}. They use something called {\em Lex-BFS
+  ordering} of the vertices of a graph to decide in linear time
+whether the graph is an interval graph. They also show how any
+interval graph recognition algorithm can be used to recognize matrices
+with \COP. This is shown in Theorem~\ref{th:intgraphcop}.
+
+\begin{theoremsansproof}[{\cite[Th.~2]{hmpv00}}]
+  \label{th:intgraphcop}
+  For a binary matrix $M$ the following statements are equivalent.
+  \begin{enumerate}
+  \item The row adjacency graph $G_r(M)$ is an interval graph and $M$
+    is its maximal clique matrix.
+  \item The columns of $M$ are maximal and $M$ has the \COP for rows.
+  \end{enumerate}
+\end{theoremsansproof}
+
 \section{Optimization problems in COP}
 \label{sec:surveycopopt}
-\tnote{Expand on ref:sec:optcop}
+
+\COP is a very beneficial property since it simplifies the structure
+of the input\footnote{Binary matrix or set system.} leading to
+efficient algorithms to otherwise hard problems.  While
+Section~\ref{sec:surveycoptest} discusses matrices with \COP and how
+to compute \COP orders, this section describes the problems that are
+of interest when a matrix does not have this property.  The central
+questions in this area are (i) how close is the matrix to having \COP,
+mainly in terms of Tucker's forbidden submatrices, and (ii) how
+optimally can one alter the matrix to attain \COP.  With regard to the
+latter question, recent literature indicate that there has been a lot
+of interest in matrix modification problems to make a matrix have \COP
+\cite{hg02, tz07}. However, to tackle this problem it is more than
+insightful to solve question (i).
+
+To study the smallest hamming distance \temptext{(?)} between the
+input matrix and one with \COP, identifying the Tucker's forbidden
+substructures\footnote{See Section~\ref{sec:tucker} for detailed
+  discussion on Tucker's forbidden substructures for matrices with
+  \COP.} is of prime importance.
+
+\subsection{Finding forbidden submatrices}
+It is known from Theorem~\ref{th:tuckercop}
+that when a matrix does not have \COP, it must have one of Tucker's
+forbidden submatrices. However, none of the \COP test algorithms discussed in
+Section~\ref{sec:surveycoptest} outputs the forbidden submatrix when
+the test is negative. This certainly gives a hint to how hard it is to find
+these substructures in the input. \temptext{Fill neg results on
+  finding forbidden submartrices -- boo75, some highlighted stuff from v85}
+
+\temptext{TEXT INTRODUCING SECTION}
+
+\subsection{Incompatibility graph - a certificate for no \COP}
+
+When a \COP test algorithm reports negative, it is hard to verify
+this. While \cite{at72} gives forbidden substructures for \COP, there
+are no efficient algorithms to find these strutures in a matrix. In
+fact, \cite{v82} shows that even finding a forbidden submatrix with at
+least $k$ rows in an input matrix is NP-complete. This was proven by
+polynomially reducing the longest path problem for graphs to this
+problem.
+
+\cite{mcc04} describes another characterization for matrices without
+\COP which is efficiently computable. This is done by way of an {\em
+  incompatibility graph} of a set system\footnote{Since \COP in
+  matrices and set systems are equivalent problems as seen in
+  Section~\ref{sec:surveycertalgo} we will use the term set system
+  here.}. The construction of this graph is based on the following
+observation.
+
+\begin{observation} 
+  Consider elements $a, b, c \in U$ and a set system $\cF$ with
+  universe $U$. If there is at least one set $S \in \cF$ such that $a,
+  c \in \cF$ but $b \notin S$, then it is impossible to have a \COP
+  order that will place $b$ in between $a$ and $c$. \tnote{also cite
+    the inbetweenness problem}
+\end{observation}
+
+For the sake of argument, suppose there exists such a \COP
+order. Clearly, it will not succeed in mapping $S$ to an interval
+since $b \notin S$ which is a contradiction to the assumption that
+this is a \COP order.  The ordered pair $(x,y)$ denotes that in the
+\COP order $y$ comes after $x$. Based on above observations, this
+means $(a,b)$ and $(b,c)$ are not compatible in any \COP order. So are
+$(c,b)$ and $(b,a)$.  Thus a binary relation {\em is incompatible
+  with} on $U \times U \setminus \{(a,a) \mid a \in U\}$ can be
+defined based on the membership of elements of $U$ in the sets in
+$\cF$.  \cite{mcc04} creates an incompatiblity graph with vertices
+denoting ordered pairs and edges denoting the incompatibility
+relation.
+
+\begin{definition}[{\em Incompatibility graph \cite[Def.~6.1]{mcc04}}]
+  Let $\cF$ be a set system with universe $U$. Let $A_{\cF} = \{(a,b)
+  \mid a, b \in U, a \ne b\}$. The {\em incompatibility graph}
+  $G_I^\cF$ of $\cF$ is an undirected graph defined as follows.
+%  \singlespacing
+  \begin{enumerate}
+  \item   The vertex set of $G_I^\cF$ is $ A_{\cF}$
+  \item  The edge set of $G_I^\cF$ are pairs of the following forms
+    \begin{itemize}
+    \item $\{(a,b),(b,a)\}$ for all $a, b \in U$
+    \item $\{(a,b),(b,c)\}$ if there exists $S \in \cF$ \stt $a, c
+      \in S$ and $b \notin S$, for all $a, b, c \in U$
+    \end{itemize}
+  \end{enumerate}
+  \dstop
+\end{definition}
+
+Thus an edge between two vertices $(a,b), (b,c)$ in $G_I^{\cF}$ means
+that there exists no linear order on $U$ that can place $b$ after $a$
+and place $c$ after $b$.  For any \COP order, there must not be any
+incompatible pairs. Thus it must consist of an independent set $I$ of
+the incompatibility graph. Moreover the independent set must have
+exactly half the vertex set of $G_I^\cF$ since the \COP order must
+involve all elements in $U$. Secondly, the reverse of a \COP order is
+also a \COP order. Thus $A_\cF \setminus I$ must also be an
+independent set. In other words, $G_I^\cF$ must be bipartite with
+partitions being of equal size if $\cF$ has \COP. An odd cycle is a
+forbidden structure in a bipartite graph, hence the same is forbidden
+in the incompatibility graph of a set system with \COP. In other
+words, if the set system has no \COP, its incompatibility graph must
+have an odd cycle.  There is also a requirement on the length of this
+odd cycle which is formally stated in the following
+theorem. \tnote{all refrences to ``following''/abov theorem must be
+  chagned to slashrefs. }
+
+\begin{theoremsansproof}[{\cite[Th.~6.1]{mcc04}, \cite[Corrected by
+    Th.~2.6, Proof p. 44--47]{d08phd}}] 
+  Let $\cF$ be a set family with universe $U$. Then $\cF$ has \COP
+  \iff its incompatibility graph is bipartite and if it does not have
+  the \COP, the incompatibility graph has an odd cycle of length at
+  most $n+3$.
+\end{theoremsansproof}
+
+Each edge of the cycle is labeled with a set from $\cF$ that documents
+the incompatibility.
+
+\temptext{how does it give clues
+  about the $R$ node? See dom.}
+
+%     \cite{mcc04} describes a different approach to COT. While all
+%     previous COT algorithms gave the COP order if the matrix has the
+%     property but exited stating negative if otherwise, this algorithm
+%     gives an evidence by way of a certificate of matrix even when it
+%     has no COP. This enables a user to verify the algorithm's result
+%     even when the answer is negative. This is significant from an
+%     implementation perspective because automated program verification
+%     is hard and manual verification is more viable. Hence having a
+%     certificate reinforces an implementation's credibility. Note that
+%     when the matrix {\em has} COP, the COP order is the certificate.
+%     The internal machinery of this algorithm is related to the
+%     weighted betweenness problem addressed\tnote{in what way??} in
+%     \cite{co98}.  \tnote{expand on the COP order graph creation and it
+%       having to be bipartite for M to have COP. and thus an odd cycle
+%       being an evidence of no COP.}
+
+\temptext{\textbf{REST OF THE SUBSECTIONS}}
 

### 2012-01-25 src/020chap2.tex
* Regular edits. See diff for changes.
* Formatting/layout related changes. stuff i'll forget later.
-------
basically:
- added two side printing
- added headers
- added tag/label markers for draft
- made all chapters/chapter*s start on odd page

technically:
- Added contents to Toc
- increased vglue in ded.tex. adjusted one hspace too. to make them fit in.
- began adding short titles for Toc. (chapter 2)
- added twoside option in LoadClass
- adjusted margins in \RequirePackage{geometry} - reduced left margin. otherwise doc bloats up with a lot of margin whitespace on twoside.
- added headings

@@
 %\chapter{\CoP\,-- A Survey of Important Results}
-\chapter{Consecutive-ones Property -- A Survey of Important Results}
+\chapter[Consecutive-ones Property -- A Survey]{Consecutive-ones Property -- A Survey of Important Results}
 \label{ch:copsurvey}
@@
   \par\noindent%
-  A set system $\cF$ from universe $U$, $|U| = n$ has the
-  {\em \cop} if there exists a linear order or 
-  permutation $\sigma = w_1w_2\ldots w_n$ that can be applied to $U$ \stt each set $S \in
-  \cF$ becomes a consecutive subsequence\footnote{Also termed an {\em interval}} $w_{i}w_{i+1}\ldots
-  w_{i+k-1}$ on $\sigma$ for some positive integer $i \le n+1-k$ where  $k = |S|$.
+  A set system $\cF$ from universe $U$, $|U| = n$ has the {\em \cop}
+  if there exists a linear order or permutation $\sigma = w_1w_2\ldots
+  w_n$ that can be applied to $U$ \stt each set $S \in \cF$ becomes a
+  consecutive subsequence\footnote{Also termed an {\em interval}}
+  $w_{i}w_{i+1}\ldots w_{i+k-1}$ on $\sigma$ for some positive integer
+  $i \le n+1-k$ where $k = |S|$.
   \par\noindent%
@@
     overlap''. This terminology is also used in other literature like
-    \cite{nsnrs09,wlh02}.} ]
+    \cite{nsnrs09,wlh02}.} etc.]
   \label{def:orthosets}%
@@
     $\cF$ that is complete\footnote{\cite[Def.~3.2]{mcc04} calls
-      this the {\em weak closure} of $\cF$ denoted by $\cW(\cF)$. \label{mcc3}}
+      this the {\em weak closure} of $\cF$ denoted by
+      $\cW(\cF)$. \label{mcc3}}
+
+  \item \label{def::slashop} The binary operator $/$ is defined as
+    follows. Let $A, B$ be two sets from set sysetm $\cF$.
+    \begin{align*}
+      A/B &= A \setminus B \cup \{b\}\\
+      \cF/B &= \{ S/B \mid S \in \cF, S \nsubseteq B \}
+    \end{align*}
+    $b$ is a new element added as a representative for $B$.
  \end{enumerate}
@@
 
+\temptext{narrative is abrupt from here. add some explanation for the
+  theorem and proposition below?}
+
 \begin{theoremsansproof}[{\cite[Th.~3,6]{mm96}}]
@@
 all \COP orders of $\cF$ into two subproblems using sets in
-$\overline{\cF} \cap \cF^{\bot}$.
+$\overline{\cF} \cap \cF^{\bot}$. The operator $/$ used is defined in
+Definition~\ref{def:orthosets}~(\ref{def::slashop}).
 
@@
 \def \ff {\overline{\cF} \cap \cF^{\bot}}
-
 The idea behind Theorem~\ref{th:mmdecomp} is as follows. A permutation $\alpha$
-of $U$ is a composition of two permutations with respect to $H$ - i) a
-permutation $\gamma$ of $H$ and (ii) a permutation $\beta$ of $U/H$.
+of $U$ is a composition of two permutations with respect to $H$ - (i) a
+permutation $\gamma$ of $H$ and (ii) a permutation $\beta$ of $U
+\setminus H$.
 
@@
 from $A$ and adding a representative element for $B$ in $A$. Being orthogonal,
-results in only the following three possibilities.
-\begin{enumerate}
+$A/B$ results in only the following three possibilities.
+\begin{enumerate}[i. ]
 \item $A$ and $ B$ are disjoint: $A/B  = A$
@@
 
-The following corollary states how \PQRtree elegantly fits into this
-whole theory and help in computing all \COP orders of $\cF$.
+The following corollary, Corollary~\ref{cor:pqrrecursive} states how
+\PQRtree elegantly fits into this whole theory and help in computing
+all \COP orders of $\cF$.
 
 \begin{corollary}[{\cite[Cor.~8]{mm96}}]
+  \label{cor:pqrrecursive}
   Let $\cF$ be a set system with universe $U$ and $H$ is a non-empty
-  orthogonal set $H \in
-  \ff$. If there is a \PQRtree $T_1$ that encodes all permutations in
-  $\valid{U/H, \cF/H}$ and also a \PQRtree $T_2'$ that encodes all
-  permutations $\valid{H,\cF \cap \power{H}}$, then a \PQRtree $T$ for
-  $\cF$ can be obtained by replacing the leaf $h$ in $T_1$ by $T_2$.
+  orthogonal set $H \in \ff$. If there is a \PQRtree $T_1$ that
+  encodes all permutations in $\valid{U/H, \cF/H}$ and also a \PQRtree
+  $T_2$ that encodes all permutations $\valid{H,\cF \cap \power{H}}$,
+  then a \PQRtree $T$ for $\cF$ can be obtained by replacing the leaf
+  $h$ in $T_1$ by $T_2$, where $h$ is the representative element for $H$.
 \end{corollary}
@@
 non-empty sets in $\ff$ are called {\em node sets} since they form the
-nodes in the \PQRtree. They are calculated as follows.  One is by
-computing the overlap components of $\cF$. The overlap components is
-the partition that results from the {overlap} equivalence relation
-which is nothing but non-orthogonal equivalence relation\footnote{It
-  is easy to verify that this relation is indeed an equivalence
-  relation.}. Overlap components are linearly
-computable\cite{mm95,wlh92}. Once these elements are factored
-out, the rest of the node sets are obtained by identifying {\em twin}
+nodes in the \PQRtree, as alluded to by
+Corollary~\ref{cor:pqrrecursive}. They are calculated as follows.
+First method is by computing the overlap components of $\cF$. The
+overlap components is the partition that results from the {\em
+  overlap} equivalence relation which is nothing but {\em
+  non-orthogonal} equivalence relation\footnote{It is easy to verify
+  that this relation is indeed an equivalence relation.}. Overlap
+components are linearly computable\cite{mm95,wlh92}. Once these
+elements are factored out, the rest of the node sets are obtained by
+the second method which is identifying {\em twin}
 elements\footnote{\cite[Sec.~3]{kklv10} calls this {indistinguishable}
@@
 $\ff$ that are non-trivial. This is when $\ff = \cT(U)$. This is the
-point where the parents of the leaves of the \PQRtree are created. The
+point where the parents of the leaves of the final \PQRtree are created. The
 following theorem helps the algorithm decide whether a $P$, $Q$ or
@@
 
-\begin{theoremsansproof}[{\cite[Th.~9]{mm96},\cite[Th.~2.1]{mcc04}}]
+\begin{theoremsansproof}[{\cite[Th.~9]{mm96}}] %,\cite[Th.~2.1]{mcc04}
 \label{th:pqrfinalmm}
@@
 
-In case (\ref{th::it1}) all elements in $U$ are made children of a
-$P$ node. In case (\ref{th::it2}) all elements in $U$ are made
-children of a $Q$ node in the order given by $\alpha$. Finally case
-(\ref{th::it3}) is the one when no permutation of $U$ gives \COP. All
-elements of $U$ are in this case made children in an $R$ node.
+In case (\ref{th::it1}) of Theorem~\ref{th:pqrfinalmm} all elements in
+$U$ are made children of a $P$ node. In case (\ref{th::it2}) all
+elements in $U$ are made children of a $Q$ node in the order given by
+$\alpha$. Finally, case (\ref{th::it3}) is the situation where no permutation
+of $U$ gives \COP. All elements of $U$ are in this case made children
+in an $R$ node.
 
 Theorem~\ref{th:pqrfinalmm} and the theory leading to it is very
-similar to \cite[Th.~2.1, 3.5. Also Th.~3.2, 3.3, 3.4]{mcc04} which categorizes
-the nodes in the above the three cases as {\em prime}, {\em linear}
-and {\em degenerate} respectively. Their generalized \PQtree is
-created in similar ways as \PQRtree above. This tree is in essense the
-Hasse diagram of what they call {\em strong elements} of
-$\overline{\cF}$. Strong elements of a set family are elements that do
-not overlap with any other elements in the family, \ie it is
-orthogonal to all other sets\cite[Def.~3.3]{mcc04}.
+similar to \cite[esp. Th.~2.1, 3.5. also Th.~3.2, 3.3, 3.4]{mcc04}
+which categorizes the nodes in the above the three cases as {\em
+  prime}, {\em linear} and {\em degenerate}
+respectively. \cite{mcc04}'s generalized \PQtree is created in similar
+ways as \PQRtree above. This tree is, in essence, the Hasse diagram of
+what they call {\em strong elements} of weak closure,
+$\cW(\cF)$\footnote{\ie $\overline{\cF}$}. Strong elements of a set
+family are elements that do not overlap with any other elements in the
+family, \ie it is orthogonal to all other sets\cite[Def.~3.3]{mcc04}.
 
@@
 As mentioned before, we observe the three data strutures of \PQRtree,
-\gPQtree and generalized \PQtree to be equivalentand the theory of all
-these data structures is summarised in Table~\ref{tab:pqrcomparison}.
-
-We will now see one more generalization of \PQtree before seeing other
-approaches to solving \COP testing including ICPIA which is a set
-theoretic characterization of \COP.
+\gPQtree and generalized \PQtree to be equivalent. The terminologies
+of the theory of all these data structures is summarised in
+Table~\ref{tab:pqrcomparison}.
+
+We will now see, in brief, one more generalization of \PQtree before
+seeing another approaches to solving \COP testing called ICPIA which
+is a set theoretic characterization of \COP.
 
@@
 
+\temptext{narrative is abrupt}
+
 \cite{nsnrs09} describes a characterization of consecutive-ones
@@
   \mid B_i  \text{ is an interval from } [n], i \in [m]\}$ be an {\em interval
-    assigment} of $\cF$, then it is called an {\em ICPIA} if it has
+    assignment} of $\cF$, then $\Pi$ is called an {\em ICPIA} if it has
   the following properties.
@@
 
-The necessity of ICPIA for a set system to have \COP given in
-Theorem~\ref{th:inpia1} is fairly obvious. It turns out that it is
-sufficient too.  This is demostrated by two algorithms which work as
-follows. The first one iterates over the assignment
-$\Pi$ %set system $\cF$
+The necessity condition of Theorem~\ref{th:icpia1} is fairly
+obvious. It turns out that it is sufficient too.  This is demostrated
+by two algorithms which work as follows. The first one iterates over
+the assignment $\Pi$ %set system $\cF$
 and changes it in each iteration till the altered
@@
   wlh02}. Once the overlap components are factored out, they can be
-independently assigned a subinterval from to which the sets in the
+independently assigned a subinterval to which the sets in the
 component need to be mapped. This is simple because the overlap
@@
 \COP order for each overlap component. In each component, the set
-whose intersections with the rest of the sets in that component forms
-a single inclusion chain is chosen first and assigned the
+whose overlaps with the rest of the sets in that component forms
+a single inclusion chain\footnote{A set of sets form a {\em single
+  inclusion chain} when the Hasse diagram of their subset relation
+forms a single chain. In this particular reference, the set of sets is
+the set of overlaps.} is chosen first and assigned the
 leftmost\footnote{By convention. Rightmost can also be chosen
@@
 
+\temptext{ abrupt: Add a section closing para. does this method (above)
+  lead to efficient algo? }
+
 \section{Optimization problems in COP}
@@
 
-\temptext{TEXT INTRODUCING SECTION}
+In this section we will describe \cite{d08phd}'s approximation
+algorithm for finding forbidden submatrix. We will use $\Tuckerset$ to
+refer to the set of Tucker's forbidden submatrices shown in
+Figure~\ref{fig:forbiddensubmatrices}.
+
+A corollary to one of Tucker's theorems presented in
+Theorem~\ref{th:tuckeratfree} is given in
+Corollary~\ref{cor:tuckertheorem}. This is central to \cite{d08phd}'s
+algorithm.
+
+\begin{corollary}[{\cite[Cor.~3.1]{d08phd}}]
+  \label{cor:tuckertheorem}
+  A matrix $M$ has \COP \iff its representing bipartite graph
+  $G_M$\footnote{Graph whose half adjacency matrix
+    (Def.~\ref{def:xxxx}) is $M$.} does not contain an asteroidal
+  triple whose three vertices correspond to columns of M.
+\end{corollary}
+
+\temptext{
+\cite[Ch.~3]{d08phd} describes two main algorithms - one is an
+approximation algorithm and the other one is an exact algorithm.
+
+\begin{theoremsansproof}[{\cite[Th.~3.1]{d08phd}}] 
+\label{th:forbiddenalg}
+Let $M$ be a $(*,\Delta)$-matrix of size $m
+\times n$. A forbidden submatrix from $\Tuckerset$ in $M$ that has a minimum
+number of rows can be found in $O(\Delta^3m^2n\dot(m+n^2))$
+time. Within the same time, one can also find a forbidden submatrix
+from $\Tuckerset$ in $M$ that has a minimum number of columns, a minimum number
+of rows and columns, or a minimum number of entries.
+
+\end{theoremsansproof}
+
+}
 
@@
   Section~\ref{sec:surveycertalgo} we will use the term set system
-  here.}. The construction of this graph is based on the following
+  in this section, in line with \cite{mcc04}'s convention.}. The construction of this graph is based on the following
 observation.
@@
   Consider elements $a, b, c \in U$ and a set system $\cF$ with
-  universe $U$. If there is at least one set $S \in \cF$ such that $a,
-  c \in \cF$ but $b \notin S$, then it is impossible to have a \COP
-  order that will place $b$ in between $a$ and $c$. \tnote{also cite
-    the inbetweenness problem}
+  universe $U$. If there is at least one set $S \in \cF$ \stt $a, c
+  \in S$ but $b \notin S$, then it is impossible to have a \COP order
+  that will place $b$ in between $a$ and $c$. \tnote{also cite the
+    inbetweenness problem}
 \end{observation}

### 2012-01-27 src/010chap1.tex
See diff please can't recall.

@@
 part of their semester thesis, they pick a body of work to study and
-form the namesake study groups, {\LLL} \cite{wallace99}, {\GGG} \cite{wallace96}, {\BBB}
-\cite{wallace00} and
-{\TTT} \cite{wallace10}. A student will be
-in at least one study group and may be in more than one. For instance,
-as will be seen later, {\Fr} studies both {\LLL} and {\TTT} while \Wo\
-studies only \BBB.
+form the namesake study groups, {\LLL} \cite{wallace99}, {\GGG}
+\cite{wallace96}, {\BBB} \cite{wallace00} and {\TTT}
+\cite{wallace10}. A student will be in at least one study group and
+may be in more than one. For instance, as will be seen later, {\Fr}
+studies both {\LLL} and {\TTT} while \Wo\ studies only \BBB.
 
@@
   }
-  \caption{\figtabsize A solution to \illustrationproblem} %study group accommodation problem
+  \caption{\figtabsize A solution to \illustrationproblem}
   \label{tab:iltree}
@@
 
-\section{Basic preliminaries - general definitions and nomenclature}
+\section[Basic preliminaries]{Basic preliminaries} 
 \label{sec:basicprelim}
@@
 
-\section{Consecutive-ones Property Testing - a Brief Survey}
+\section[Brief Survey]{Consecutive-ones Property - a Brief Survey}
 \label{sec:background}
@@
 
-\section{Application of COP in Areas of Graph Theory and
-  Algorithms}
-\label{sec:appcopGTA}
+% \section[Some applications of \COP]{Application of COP in Areas of Graph Theory and
+%   Algorithms}
+% \label{sec:appcopGTA}
 
-\temptext{Combine COP in Relational Database Model + COP in Graph
-  Isomorphism + Certifying Algorithms}
+% \temptext{Combine COP in Relational Database Model + COP in Graph
+%   Isomorphism + Certifying Algorithms}
 
-%\subsection{COP in Relational Database Model}
-\label{sec:apprdbm}
-\tnote{(set systems theme)}
+% %\subsection{COP in Relational Database Model}
+% \label{sec:apprdbm}
+% \tnote{(set systems theme)}
 
-%\subsection{}
-\label{sec:appgraphiso}
-\tnote{(canonization theme)}
+% %\subsection{}
+% \label{sec:appgraphiso}
+% \tnote{(canonization theme)}
 
-%\subsection{}
-\label{sec:appcertalgo}
-\tnote{ (certification McC04 theme)}
+% %\subsection{}
+% \label{sec:appcertalgo}
+% \tnote{ (certification McC04 theme)}
 
-\section{Generalization of COP - the Motivation}
+\section[Generalization of COP]{Generalization of COP - the Motivation}
 \label{sec:motive}
@@
 
-\section{Summary of New Results in this Thesis}
+\section[Summary of new results]{Summary of New Results in this Thesis}
 \label{sec:results}

### 2012-01-27 src/020chap2.tex
See diff please can't recall.

@@
 %\chapter{\CoP\,-- A Survey of Important Results}
-\chapter[Consecutive-ones Property -- A Survey]{Consecutive-ones Property -- A Survey of Important Results}
+\chapter[COP -- A Survey]{Consecutive-ones Property -- A Survey of
+  Important Results} 
 \label{ch:copsurvey}
@@
 
-\subsection{Booth and Lueker's $PQ$ tree -- a linear COT algorithm}
+\subsection{Booth and Lueker's $PQ$-tree -- a linear COT algorithm} %[\PQtree]
 
@@
 \label{sec:surveygraphiso}
-\tnote{Expand on sec:appgraphiso} \tnote{(canonization theme)}
+\tnote{(canonization theme)}
 

### 2012-01-27 src/030chap3.tex
Did some over all clean up. Added fancy chapter headings. Begun work on opt survey.
See diff please can't recall.

@@
-\chapter{Tree Path Labeling of Path Hypergraphs - the New Results}
+\chapter[Tree Path Labeling]{Tree Path Labeling of Path
+  Hypergraphs -- New Results}
 \label{ch:myresearch}
@@
 
-\section{Summary of Proposed Problems}
+\section[Summary of problems]{Summary of Proposed Problems}
 \label{sec:myresearchintro}
@@
 
-\section{Preliminaries to new results}
+\section[Preliminaries]{Preliminaries to new results}
 \label{ch:prelims}
@@
 
-\section{Characterization of Feasible Tree Path Labeling}
+\section[Characterization of FTPL]{Characterization of Feasible Tree Path Labeling}
 \label{sec:feasible}
@@
 
-\section{ Computing feasible TPL with special target trees}
+\section[Special target trees]{ Computing feasible TPL with special target trees}
 \label{sec:spltargettree}
@@
 
-\section{Complexity of Tree Path Assignment-A Discussion}
+\section[Complexity]{Complexity of Tree Path Assignment-A Discussion}
 \label{complexity}

### 2012-01-28 src/010chap1.tex
cleaned up tnotes

@@
 problem (where the ``target graph'' is a path and not any
-tree\tnote[minor]{Allowing any tree in this example could be seen as a
-  scenario where there are parallel classes. A node falling in the
-  path between two other nodes would mean that the corresponding is
-  scheduled between the other two.}) is called an interval assignment
+tree) is called an interval assignment
 problem.
@@
   \item \Cop order or \cop permutation
-  \item \Cop for set systems\tnote{see mcc04}
   \end{enumerate}
@@
   \label{def:cropmatrix}
-  \tnote[important]{define circ1 here. refer to it in chap 2 or
-    duplicate it using slash def }
 \end{definition}
@@
 columns) of the matrix. This is constructed using an elaborate linear
-time procedure and is also a test for planarity\tnote{check check
-  check. both interval graph and planarity in this paper?}.  PQR trees
-is a generalized data structure based on PQ trees \cite{mm96,mpt98}.
+time procedure and is also a test for planarity.  PQR trees is a
+generalized data structure based on PQ trees \cite{mm96,mpt98}.
 \cite{tm05} describes an improved algorithm to build PQR
-trees. \tnote{improv in terms of what?}\cite{wlh02} describes the
-simpler algorithm for COT. Hsu also invented PC trees
-\cite{wlh01}\tnote{This result first appeared inproc ISAAC92}
-which is claimed to be much easier to implement. \cite{nsnrs09}
-describes a characterization of consecutive-ones property solely based
-on the cardinality properties of the set representations of the
-columns (rows); every column (row) is equivalent to a set that has the
-row (column) indices of the rows (columns) that have one entries in
-this column (row). This is interesting and relevant, especially to
-this thesis because it simplifies COT to a great degree. \tnote{it
-  reduces the solution search space. fill in the blanks.}
+trees. \cite{wlh02} describes the simpler algorithm for COT. Hsu also
+invented PC trees \cite{wlh01} which is claimed to be much easier to
+implement. \cite{nsnrs09} describes a characterization of
+consecutive-ones property solely based on the cardinality properties
+of the set representations of the columns (rows); every column (row)
+is equivalent to a set that has the row (column) indices of the rows
+(columns) that have one entries in this column (row). This is
+interesting and relevant, especially to this thesis because it
+simplifies COT to a great degree.
 
@@
 algorithm is related to the weighted betweenness problem
-addressed\tnote{in what way??} in \cite{co98}.  \tnote{expand
-  on the COP order graph creation and it having to be bipartite for M
-  to have COP. and thus an odd cycle being an evidence of no COP.}
-\tnote{ where should this go?: (1) cite|jlm97 (application of PQ
-  trees in graphics). (2) helly's theorem citation
-  19XXdgk-Hellystheorem-Danzer-Gruenbaum-Klee}
+addressed in \cite{co98}.
 
@@
 COP. Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
-a submatrix $M'$ with all columns\tnote{check if b75 deals with
+a submatrix $M'$ with all columns\temptext{ check if b75 deals with
   COP col or COP row. also is it any submatrix with k less than r rows
-  or submatrix must have all columns?} but a maximum cardinality
+  or submatrix must have all columns? } but a maximum cardinality
 subset of rows such that $M'$ has COP is NP complete. \cite{hg02}
@@
 the area of parameterized algorithms for this
-problem\tnote{elaborate - what are the results?}.
+problem.
 
@@
 known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
-using a reduction to the longest path problem, \tnote{or is it a
+using a reduction to the longest path problem, \temptext{ or is it a
   survey of another result?  check.} that finding a Tucker's forbidden
-submatrix of at least $k$ rows is NP complete. \tnote{how is this
-  different from booth's 75 result??}  \tnote{where should this
-  go? cite|tz04 (approx submatrix with COP sparse matrices)}
+submatrix of at least $k$ rows is NP complete. \temptext{ how is this
+  different from booth's 75 result??}
 
@@
 {\un}s. This is proved to be NP hard using a reduction from the
-Hamiltonian path problem.\tnote{Theorem 2.1 in jkckv}
-\tnote{(1) A connection of COP problem to the travelling salesman
-  problem is also introduced. what does this mean? -- COP can be used
-  as a tool to reorder $0.5T \le runs(M) \le T$. (2) The optimization
-  version of the $k$-run problem, i.e. minimization of number of
-  blocks of ones is proven to be NP complete by
-  cite:k77}\tnote{are these two the same?} \tnote{what is the
-  reduction?} \tnote{other problems similar to COP -- cite:ckl96
-  (ILP, circ ones, one drop) -- cite:th98 (generalization of COP -
-  minimax, biotonic column) Tucker}
+Hamiltonian path problem.
 
@@
     hypergraph} is an alternate representation of a set system and
-  will be used in this thesis.}~\footnote{See Section~\ref{ch:prelims} for the
-  formal definition.} and thus is a characterization for COP
-\cite{nsnrs09}. \tnote{Refer to the survey section that will list the
-  theorems and/or lemma?} In our work we extend this characterization
-and find that trio-wise intersection cardinality preservation makes a
-tree path labeling\footnote{A {\em tree path labeling} $\cl$ is a
-  bijection of paths from the target tree $T$ to the hyperedges in
-  given hypergraph $\cF$.}~\footnotemark[4]  (TPL) feasible, which is a generalization of the COP
-problem. This problem is defined as follows.
+  will be used in this thesis.}~\footnote{See Section~\ref{ch:prelims}
+  for the formal definition.} and thus is a characterization for COP
+\cite{nsnrs09}. In our work we extend this characterization and find
+that trio-wise intersection cardinality preservation makes a tree path
+labeling\footnote{A {\em tree path labeling} $\cl$ is a bijection of
+  paths from the target tree $T$ to the hyperedges in given hypergraph
+  $\cF$.}~\footnotemark[4] (TPL) feasible, which is a generalization
+of the COP problem. This problem is defined as follows.
 
@@
 
-\tnote{The following sentence seems out of place in this para} In
+\temptext{The following sentence seems out of place in this para} In
 spite of this being a restricted case, we believe that our results are
@@
 represent a set system (see Section~\ref{sec:motive}) and that the
-{\em overlap}\tnote{Give informal definition of ``overlap''
-  in footnote/endnote and link to sec:prelims. ADD the definition in prelim.}
-relation between the sets involved is an obvious equivalence relation,
-$\cF$ quite naturally partitions into equivalence classes known as
-{\em overlap components}\tnote{Give informal definition of ``overlap
-  components'' in footnote and link to sec:prelims. ADD the definition
-  in prelim.}. In the context of COP, overlap components were used in
-\cite{wlh02} and \cite{kklv10}. Moreover, \cite{nsnrs09} discovered
-that these equivalence classes form a total order \tnote{Give informal
-  definition of ``total order'' in footnote?}. We extend this to TPL
-and find that when $\cF$ is a path hypergraph\footnote{If there exists
-  an FTPL for a hypergraph $\cF$, it is called a path hypergraph.},
-the classes can be partially ordered as an in-tree in polynomial
+{\em overlap} relation between the sets involved is an obvious
+equivalence relation, $\cF$ quite naturally partitions into
+equivalence classes known as {\em overlap components}. In the context
+of COP, overlap components were used in \cite{wlh02} and
+\cite{kklv10}. Moreover, \cite{nsnrs09} discovered that these
+equivalence classes form a total order. We extend this to TPL and find
+that when $\cF$ is a path hypergraph\footnote{If there exists an FTPL
+  for a hypergraph $\cF$, it is called a path hypergraph.}, the
+classes can be partially ordered as an in-tree in polynomial
 time. Once $\cF$ is ``broken'' into overlap components, one must
@@
 hard part which is currently open to be solved in polynomial time.
-
-\tnote{The connection of TPL to graph isomorphism will be made
-  later in the document}
-
-%\theendnotes

### 2012-01-28 src/020chap2.tex
cleaned up tnotes

@@
 
-\tnote[important]{ADD: have a few lines about organization of chapter}
+\temptext{ADD: have a few lines about organization of chapter}
 
@@
   \item \label{def::concave-round} $G$ is \emph{concave-round} if its
-    augmented adjacency matrix has \CROP. \tnote{cite BHY00}
-    \tnote[minor]{add CROP to glossary}
+    augmented adjacency matrix has \CROP. \tnote[pressing]{cite BHY00}
   \item $G$ is an \emph{interval graph} if its vertices can be mapped
     to intervals on the real line \stt two vertices are adjacent \iff
-    their corresponding intervals overlap \tnote{cite Ben59, Haj57}.
-    $G$ is an interval graph \iff its maximal clique matrix has COP
-    \cite{fg65}\footnote{This follows \cite{gh64} which states that the
-      maximal cliques of interval graph $G$ can be linearly ordered
-      \stt for all $v \in V(G)$, cliques containing $v$ are
+    their corresponding intervals overlap \tnote[pressing]{cite Ben59,
+      Haj57}.  $G$ is an interval graph \iff its maximal clique matrix
+    has COP \cite{fg65}\footnote{This follows \cite{gh64} which states
+      that the maximal cliques of interval graph $G$ can be linearly
+      ordered \stt for all $v \in V(G)$, cliques containing $v$ are
       consecutive in the ordering \cite[Th. 8.1]{mcg04}.}
@@
         set of unit interval graphs and the set of proper interval
-        graphs are the same} \tnote{cite {rob69,gar07} in footnote. pg
-        33 dom}
+        graphs are the same} \tnote[pressing]{cite {rob69,gar07} in
+        footnote. pg 33 dom}
     \end{enumerate}
@@
 around them is due to their very desirable property that several
-problems that are NP-hard \tnote[minor]{all NPC problems are NPH yes?}
-on general graphs, like finding a maximum clique or minimum coloring
-or independent set, are polynomial time solvable in these graph
-classes \cite{clrs01}.  In a similar fashion, a lot of problems that
-are hard on general matrices have efficient solutions on matrices with
-\COP or \CROP \cite[more citations pg.\,33]{d08phd}.
+problems that are NP-complete on general graphs, like finding a
+maximum clique or minimum coloring or independent set, are polynomial
+time solvable in these graph classes \cite{clrs01}.  In a similar
+fashion, a lot of problems that are hard on general matrices have
+efficient solutions on matrices with \COP or \CROP \cite[more
+citations pg.\,33]{d08phd}.
 
@@
 is by deriving one from any interval graph recognition algorithm using
-the result {\sc hmpv00}\tnote{cite hmpv00} \tnote{in footnote put the
-  theorem 2.7 - dom pg 43} \cite{d08phd} which demonstrates how such a
-derivation can be done. However, this does not necessarily yield an
-efficient algorithm. We will see results that directly solve the
-problem on matrices since it is known that questions~(\ref{q:testcop})
-and (\ref{q:copperm}) stated above for \COP are efficiently solvable.
-Table~\ref{tab:cophistory} gives a snapshot of these results.
+the result \cite{hmpv00} \cite[Th~2.7]{d08phd} which demonstrates how
+such a derivation can be done. However, this does not necessarily
+yield an efficient algorithm. We will see results that directly solve
+the problem on matrices since it is known that
+questions~(\ref{q:testcop}) and (\ref{q:copperm}) stated above for
+\COP are efficiently solvable.  Table~\ref{tab:cophistory} gives a
+snapshot of these results.
 
@@
 \label{sec:tucker}
-% \change{While it is remarkable that the first polynomial time
-%   algorithm for \COP testing came in }{The first polynomial time
-%   algorithm for \COP testing came in }\cite{fg65} \change{,}{.}
-% \change{it is safe to say that the most}{A few years later, a very}
-% significant result in understanding \COP came \remove{a few years
-%   later} from Tucker which gave a combinatorial (negative)
-% characterization of matrices with COP \cite{at72}.
 
@@
   $\un$s in it.} for \cop testing while given a linear time interval
-graph recognition algorithm by a simplification of \tnote{cite -
+graph recognition algorithm by a simplification of \tnote[important]{cite -
   A. Lempel, S. Even and I. Cederbaum, An Algorithm for Planarity
@@
 part of a {\em substitution decomposition} for a set system $\cF$
-\tnote{define substitutio decomp.} and subsequently \cite{mcc04} gives
+\tnote[pressing]{define substitutio decomp.} and subsequently \cite{mcc04} gives
 a new characterization of $\cF$ using a so-called incompatibility
@@
 \item \label{th::it1} $\overline{\cF} = \cT(U)$ 
-\item \label{th::it2}$\overline{\cF} = consec(\alpha)$\tnote{what is consec?} for
+\item \label{th::it2}$\overline{\cF} = consec(\alpha)$\tnote[pressing]{what is consec?} for
   some permutation $\alpha$ on $U$ 
@@
       Decomposition tree of $\cW(\cF)$, $T(\cW(\cF))$
-      \temptext{PQR tree def. }\tnote[pressing]{in terms of theory}
+      \temptext{PQR tree def. }
     }
@@
   \in S$ but $b \notin S$, then it is impossible to have a \COP order
-  that will place $b$ in between $a$ and $c$. \tnote{also cite the
+  that will place $b$ in between $a$ and $c$. \tnote[important]{also cite the
     inbetweenness problem}
@@
 odd cycle which is formally stated in the following
-theorem. \tnote{all refrences to ``following''/abov theorem must be
+theorem. \tnote[pressing]{all refrences to ``following''/abov theorem must be
   chagned to slashrefs. }
@@
 
-\temptext{ So far we have been concerned about matrices that have the
-  consecutive ones property. However in real life applications, it is
-  rare that data sets represented by binary matrices have COP,
-  primarily due to the noisy nature of data available. At the same
-  time, COP is not arbitrary and is a desirable property in practical
-  data representation \cite{co98,jkckv04,k77}. In this context, there
-  are several interesting problems when a matrix does not have COP but
-  is ``close'' to having COP or is allowed to be altered to have
-  COP. These are the optimization problems related to a matrix which
-  does not have COP. Some of the significant problems are surveyed in
-  this section.
-
-\tnote{ -- sect 4.1 in cite:d08phd has many results surveyed. hardness
+\temptext{ -- sect 4.1 in cite:d08phd has many results surveyed. hardness
   results, approx. results. results are usually for a class of
@@
   entries of $M$ to make it attain COP. this is NP complete
-  cite:b75-phd}%
-\tnote{(1) scite:lb62 showed that interval graphs are chordal and
-  AT-free.} %
-\temptext{couple of lines refering to tucker's submatrices. refer
-  earlier section.}
-
-Once a matrix has been detected to not have COP (using any of the COT
-algorithms mentioned earlier), it is naturally of interest
-
-\begin{enumerate}
-\item  to find out
-the smallest forbidden substructure (in terms of number of rows and/or
-columns and/or number of entries that are {\un}s). \cite{d08phd}
-discusses a couple of algorithms which are efficient if the number of
-{\un}s in a row is small. This is of significance in the case of
-sparse matrices where this number is much lesser than the number of
-columns.
-
-\item $(*,\Delta)${\em -matrices} are matrices with no restriction
-on number of {\un}s in any column but has at most $\Delta$ {\un}s in
-any row. 
-{\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are
-similar problems which deals with {\em inducing COP} on a matrix.
-
-\begin{enumerate}
-\item In the dual problem {\sc Max COS-R (Max COS-C)} the search is for the
-maximum number of rows (columns) that induces a submatrix with
-COP.
-\item In {\sc
-  Min COS-R (Min COS-C)} the question is to find the minimum number of
-rows (columns) that must be deleted to result in a matrix with COP.
-\end{enumerate}
-
-Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
-a submatrix $M'$ with all columns\tnote{check if b75 deals with
-  COP col or COP row. also is it any submatrix with k less than r rows
-  or submatrix must have all columns?} but a maximum cardinality
-subset of rows such that $M'$ has COP is NP complete. \cite{hg02}
-corrects an error of the abridged proof of this reduction as given in
-\cite{gj79}.  \cite{d08phd} discusses all these problems in detail
-giving an extensive survey of the previously existing results which
-are almost exhaustively all approximation results and hardness
-results. Taking this further, \cite{d08phd} presents new results in
-the area of parameterized algorithms for this
-problem\tnote{elaborate - what are the results?}.
-
-\item Another problem is to find the minimum number of entries in the matrix
+  cite:b75-phd%
+\\(1) scite:lb62 showed that interval graphs are chordal and
+  AT-free. %
+\\couple of lines refering to tucker's submatrices. refer
+  earlier section.
+
+Another problem is to find the minimum number of entries in the matrix
 that can be toggled to result in a matrix with COP.  \cite{v85}
@@
 known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
-using a reduction to the longest path problem, \tnote{or is it a
+using a reduction to the longest path problem, \tnote[pressing]{or is it a
   survey of another result?  check.} that finding a Tucker's forbidden
-submatrix of at least $k$ rows is NP complete. \tnote{how is this
-  different from booth's 75 result??}  \tnote{where should this go?
+submatrix of at least $k$ rows is NP complete. \tnote[pressing]{how is this
+  different from booth's 75 result??}  \tnote[pressing]{where should this go?
   cite|tz04 (approx submatrix with COP sparse matrices)}
 
-\item \cite{jkckv04} discusses the use of matrices with almost-COP (instead
+\cite{jkckv04} discusses the use of matrices with almost-COP (instead
 of one block of consecutive {\un}s, they have $x$ blocks, or {\em
@@
 {\un}s. This is proved to be NP hard using a reduction from the
-Hamiltonian path problem.\tnote{Theorem 2.1 in jkckv} \tnote{(1) A
-  connection of COP problem to the travelling salesman problem is also
-  introduced. what does this mean? -- COP can be used as a tool to
-  reorder $0.5T \le runs(M) \le T$. (2) The optimization version of
-  the $k$-run problem, i.e. minimization of number of blocks of ones
-  is proven to be NP complete by cite:k77}\tnote{are these two the
-  same?} \tnote{what is the reduction?} \tnote{other problems similar
-  to COP -- cite:ckl96 (ILP, circ ones, one drop) -- cite:th98
-  (generalization of COP - minimax, biotonic column) Tucker}
-
-\end{enumerate}
-
+Hamiltonian path problem.\tnote[pressing]{Theorem 2.1 in jkckv}
+\tnote[pressing]{(1) A connection of COP problem to the travelling
+  salesman problem is also introduced. what does this mean? -- COP can
+  be used as a tool to reorder $0.5T \le runs(M) \le T$. (2) The
+  optimization version of the $k$-run problem, i.e. minimization of
+  number of blocks of ones is proven to be NP complete by
+  cite:k77}\tnote{are these two the same?} \tnote{what is the
+  reduction?} \tnote{other problems similar to COP -- cite:ckl96 (ILP,
+  circ ones, one drop) -- cite:th98 (generalization of COP - minimax,
+  biotonic column) Tucker}
 }
@@
 }
-
-% \section{Other applications}
-% {\tt If any. \tnote{ where should this go?}: (1) cite|jlm97 (application of PQ
-%   trees in graphics). (2) helly's theorem citation
-%   19XXdgk-Hellystheorem-Danzer-Gruenbaum-Klee}}
-
-%\theendnotes\tnote{remove if none.}
 

### 2012-01-28 src/thesis.tex
Did some over all clean up. Added fancy chapter headings. Begun work on opt survey.
* Regular edits. See diff for changes.
* Formatting/layout related changes. stuff i'll forget later.
-------
basically:
- added two side printing
- added headers
- added tag/label markers for draft
- made all chapters/chapter*s start on odd page

technically:
- Added contents to Toc
- increased vglue in ded.tex. adjusted one hspace too. to make them fit in.
- began adding short titles for Toc. (chapter 2)
- added twoside option in LoadClass
- adjusted margins in \RequirePackage{geometry} - reduced left margin.

@@
                                 % in .tex file.
-%\usepackage[inline] 
+
+%% Drafting libraries %%
+%
+\usepackage[color,notcite,notref]{showkeys}   % To show keys used - useful in
+                               % drafts. 
+                               % Options: color, notref, notcite,
+                               % draft(default), final
+
+\newcommand{\xclearpage}{\cleardoublepage} % cleardoublepage for
+                                % twoside. keep empty for one side.
+\usepackage[inline] 
 %\usepackage[margins] 
-\usepackage[finalnew]
+%\usepackage[finalnew]
          {../lib/trackchanges} % For tracking changes.
@@
 % \usepackage{comment}         % DEBUG ONLY
-\includeonly{020chap2}         % DEBUG ONLY 
+%\includeonly{010chap1}         % DEBUG ONLY 
+%\includeonly{020chap2}         % DEBUG ONLY 
 %\includeonly{030chap3,050appa}         % DEBUG ONLY 
@@
 
-\def \mythesissubmissiondate {13 January 2012}
+\def \mythesissubmissiondate {31 January 2012}
 \def \mythesissubmissionmonth {\MakeUppercase {January 2012}}
+
+%\renewcommand\sectionmark[1]{\markright{\thesection. #1}}
+\renewcommand\chaptermark[1]{\markboth{\chaptername\ \thesection. #1}{}}
+\renewcommand\sectionmark[1]{\markright{\thesection. #1}}
 
@@
 %
+
+\xclearpage% \cleardoublepage
+\setcounter{page}{1}
 \include{001certificate}
@@
 %
+\xclearpage
 \include{002ack}
@@
 %
-\include{0031ded}%{0021ded}
+\xclearpage
+\include{0031ded}
 
@@
 %
+\xclearpage
 \include{003abstract}
@@
 \begin{singlespace}
+  \xclearpage
   \tableofcontents
   \thispagestyle{empty}
-  \listoftables \addcontentsline{toc}{chapter}{LIST OF TABLES}
-  \listoffigures \addcontentsline{toc}{chapter}{LIST OF FIGURES}
+  \xclearpage
+  \listoftables %\addcontentsline{toc}{chapter}{LIST OF TABLES}
+  \xclearpage
+  \listoffigures %\addcontentsline{toc}{chapter}{LIST OF FIGURES}
 \end{singlespace}
@@
 %
+\xclearpage
 \include{004abbrv}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Notation
-%
-\include{005notation}
-%\pagebreak \clearpage
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% % Notation
+% %
+% \xclearpage
+% \include{005notation}
 
+\xclearpage
+\pagestyle{fancy}%headings}   % For running headers.
+\fancyhf{}%            % Clears headers
+\fancyhead[LE,RO]{\scriptsize \slshape \rightmark} %{\slshape \rightmark}
+\fancyhead[LO,RE]{\scriptsize \slshape \leftmark} %{\slshape \leftmark}
 \pagenumbering{arabic} % The main text will follow from this point so
@@
 %
-\include{figuresandtables} % Load floats. No rendering.
-
 \include{010chap1}
@@
                         % enough. or maybe it was fluke.
+\xclearpage
 \include{020chap2}
+\xclearpage
 \include{030chap3}
+\xclearpage
 \include{040chap4}
@@
 \begin{singlespace}
-  \addcontentsline{toc}{chapter}{REFERENCES \tnote[minor]{(i)make names
+%  \addcontentsline{toc}{chapter}{REFERENCES }
+
+  \tnote[minor]{(i)make names
     in bib file uniform style w.r.t. firstname/initials (ii) have back
-  refs - i.e. pages that cite the item.}}
-
+  refs - i.e. pages that cite the item.}
   {\footnotesize%

### 2012-01-29 src/010chap1.tex
- a whole bunch of cleanups. see diff. - removed problem defs from chap 1. they were duplicates from chap 3. - CODED FIGURES IN TIKZ. forbidden subgraphs, graphmatrix example graphs. - BEGAN CHAPTER 3 EDITS.
Formatting/layout changes
-------

- Added FrenchLayout to make lists denser
- Added hanging indent for Definitions
- Compressed theorem env's above below spaces
- Creaded env for problem definition minipages.
- moved all clearpage/page counter and other prechapter settings to the chapter tex files.
- Fixed the page number bug in draft printout. removed \fancyhf{} to avoid resetting page n

@@
 the problem is to allocate class slots to courses such that all the
-classes of a study group are consecutive. It is debatable if this will
-not hamper the attention span and memory retention rate of the
-students but that is, regrettably, out of the scope of this
-thesis. The parallels between this class allocation problem and the
+classes of a study group are consecutive. % \footnote{It is debatable if this will
+% not hamper the attention span and memory retention rate of the
+% students but that is undoubtedly out of the scope of this
+% thesis.}
+The parallels between this class allocation problem and the
 accommodation problem can be seen as follows. The set $U$ here, are
@@
 
+If $n$ is a positive integer, $[n]$ denotes the set $\{1,2,... n\}$.
+
 \begin{definition}[\emph{Binary matrix}.]
@@
   element, i.e. element at $i$th row and $j$th column. $M$ is a binary
-  matrix if each of its element is 0 or \un.
-  In other words, for all $i \in [n]$ and $j \in [m]$, $m_{i,j} \in
-  \{0, \un\}$ 
+  matrix if each of its element is 0 or \un.  In other words, for all
+  $i \in [n]$ and $j \in [m]$, $m_{i,j} \in \{0, \un\}$
 \end{definition}
@@
   x_n\}$ is a bijection $\lambda: X \rightarrow X$.  For simplicity,
-  sometimes, $\lambda$ is written as a sequence $x_{i_1}x_{i_2}\ldots
-  x_{i_n}$, where $i_j \in \{1, 2 \ldots n\}$ to mean that
-  $\lambda(x_j)= x_{i_j}$.\\
+  $\lambda$ may be written as a sequence $x_{i_1}x_{i_2}\ldots x_{i_n}$,
+  where $i_j \in [n]$ to mean that $\lambda(x_j)= x_{i_j}$.
   This document uses both notations as convenient in context.
@@
   \label{def:graphtree}
+
 \end{definition}
+When refering to a tree as $T$ it could be a reference to the tree
+itself, or the vertices of the tree. This will be clear from the
+context.
 
@@
 \end{definition}
-
-If $n$ is a positive integer, $[n]$ denotes the set $\{1,2,... n\}$.
 
@@
   $\cF$.}~\footnotemark[4] (TPL) feasible, which is a generalization
-of the COP problem. This problem is defined as follows.
+of the COP problem. This problem is defined by \FTPL (See
+Section~\ref{sec:myresearchintro} for problem definition). %as follows.
 
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \FTPL}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
-        paths $\cP$ from $T$ and a
-        bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
-      \end{minipage}\\
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
-        such that $\phi$ when applied on any hyperedge in $\cF$ will
-        give
-        the path mapped to it by the given tree path labeling $\cl$.\\
-        { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every
-          hyperedge $S \in \cF$.}
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
+% {\small
+%   \begin{minipage}[h]{5in}
+%     % \begin{singlespace}
+%     \vspace{2mm}
+%     {\large \FTPL}\\
+%     \begin{tabular}[t]{l|l}
+%       \hline\\
+%       {\tt Input} & 
+%       \begin{minipage}[t]{\probdefwidth}
+%         A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
+%         paths $\cP$ from $T$ and a
+%         bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
+%       \end{minipage}\\
+%       {\tt Question} &
+%       \begin{minipage}[t]{\probdefwidth}
+%         Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
+%         such that $\phi$ when applied on any hyperedge in $\cF$ will
+%         give
+%         the path mapped to it by the given tree path labeling $\cl$.\\
+%         { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every
+%           hyperedge $S \in \cF$.}
+%       \end{minipage}\\
+%     \end{tabular}
+%     % \end{singlespace}
+%   \end{minipage}\\
+% }
 
@@
 problem of computing a feasible TPL for a given hypergraph and target
-tree, if one exists. This problem, {\CFTPL}, is defined as follows.
+tree, if one exists. This problem is the {\CFTPL} problem (See
+Section~\ref{sec:myresearchintro} for problem definition). %, is defined as follows.
 
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \CFTPL}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$ and a tree $T$.\\
-      \end{minipage}\\
+% {\small
+%   \begin{minipage}[h]{5in}
+%     % \begin{singlespace}
+%     \vspace{2mm}
+%     {\large \CFTPL}\\
+%     \begin{tabular}[t]{l|l}
+%       \hline\\
+%       {\tt Input} & 
+%       \begin{minipage}[t]{\probdefwidth}
+%         A hypergraph $\cF$ with vertex set $U$ and a tree $T$.\\
+%       \end{minipage}\\
 
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a set of paths $\cP$ from $T$ and a bijection
-        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
-        {\bf true} on $(\cF, T, \cl)$.
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
+%       {\tt Question} &
+%       \begin{minipage}[t]{\probdefwidth}
+%         Does there exist a set of paths $\cP$ from $T$ and a bijection
+%         $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
+%         {\bf true} on $(\cF, T, \cl)$.
+%       \end{minipage}\\
+%     \end{tabular}
+%     % \end{singlespace}
+%   \end{minipage}\\
+% }
 
@@
 when the hyperedges in the hypergraph $\cF$ have at most $k+2$
-vertices. A couple of examples of {\kstars} are given in
+vertices (\CFTPLKTREE -- See Section~\ref{sec:myresearchintro} for
+problem definition). A couple of examples of {\kstars} are given in
 Figure~\ref{fig:ksubstar}.
@@
 
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \CFTPLKTREE}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$ such that every
-        hyperedge
-        $S \in \cF$ is of cardinality at most $k+2$ and a {\kstar} $T$.\\
-      \end{minipage}\\
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
-        Does there exist a set of paths $\cP$ from $T$ and a bijection
-        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
-        {\bf true} on $(\cF, T, \cl)$.
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
+% {\small
+%   \begin{minipage}[h]{5in}
+%     % \begin{singlespace}
+%     \vspace{2mm}
+%     {\large \CFTPLKTREE}\\
+%     \begin{tabular}[t]{l|l}
+%       \hline\\
+%       {\tt Input} & 
+%       \begin{minipage}[t]{\probdefwidth}
+%         A hypergraph $\cF$ with vertex set $U$ such that every
+%         hyperedge
+%         $S \in \cF$ is of cardinality at most $k+2$ and a {\kstar} $T$.\\
+%       \end{minipage}\\
+%       {\tt Question} &
+%       \begin{minipage}[t]{\probdefwidth}
+%         Does there exist a set of paths $\cP$ from $T$ and a bijection
+%         $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
+%         {\bf true} on $(\cF, T, \cl)$.
+%       \end{minipage}\\
+%     \end{tabular}
+%     % \end{singlespace}
+%   \end{minipage}\\
+% }
 
-\temptext{The following sentence seems out of place in this para} In
-spite of this being a restricted case, we believe that our results are
-of significant interest in understanding the nature of {\sc Graph
+In spite of this being a restricted case, we believe that our results
+are of significant interest in understanding the nature of {\sc Graph
   Isomorphism} which is polynomial time solvable in interval graphs
-while being hard on path graphs\cite{kklv10}. {\kstars} are a class
-of trees which are in many ways very close to intervals or paths. Each
+while being hard on path graphs\cite{kklv10}. {\kstars} are a class of
+trees which are in many ways very close to intervals or paths. Each
 ray\footnote{The path from a leaf to the root, the vertex with highest
-  degree, is called a {\em ray} of the \kstar.}~\footnotemark[4] are independent except for the
-root\footnote{The vertex with maximum degree in a {\kstar} is called
-  {\em root}.}~\footnotemark[4] and hence can be
-considered as an independent interval till the root. Our algorithm
-builds on this fact and uses the interval assignment
+  degree, is called a {\em ray} of the \kstar.}~\footnotemark[4] are
+independent except for the root\footnote{The vertex with maximum
+  degree in a {\kstar} is called {\em root}.}~\footnotemark[4] and
+hence can be considered as an independent interval till the root. Our
+algorithm builds on this fact and uses the interval assignment
 algorithm\cite{nsnrs09} up until ``reaching'' the root and then uses

### 2012-01-29 src/020chap2.tex
- a whole bunch of cleanups. see diff. - removed problem defs from chap 1. they were duplicates from chap 3. - CODED FIGURES IN TIKZ. forbidden subgraphs, graphmatrix example graphs. - BEGAN CHAPTER 3 EDITS.
Formatting/layout changes
-------

- Added FrenchLayout to make lists denser
- Added hanging indent for Definitions
- Compressed theorem env's above below spaces
- Creaded env for problem definition minipages.
- moved all clearpage/page counter and other prechapter settings to the chapter tex files.
- Fixed the page number bug in draft printout. removed \fancyhf{} to avoid resetting page n

@@
-\definecolor{gray}{rgb}{0.8,0.8,0.8}
-\def\tblhcolor {gray}
+\xclearpage
 
@@
 
-\temptext{ADD: have a few lines about organization of chapter}
+\tnote[dire]{ADD: have a few lines about organization of chapter}
 
@@
 certain graphs, like convex bipartite graphs, that are defined solely
-by some of its associated matrix having \COP.  In this section we will
+by one of its associated matrix having \COP.  In this section we will
 see the relevance of \cop to graphs.  To see this we introduce certain
@@
 
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[scale=0.35]{../img/tuckersforbiddensubgraphs.png}  
-  \caption[\figtabsize Tucker's forbidden subgraphs]{\figtabsize Tucker's
-    forbidden subgraphs for convex bipartite graphs. \temptext{ PLACEHOLDER
-    IMG }} 
-  \label{fig:forbiddensubgraphs}
-\end{figure}
+\figforbiddensubgraphs
 
@@
 \cite{km89,mcc04}, \PQRtree \cite{mm96,mpt98} etc. Most of these are
-generalizations of \PQtree -- for instance, \PCtree is generalized to
+generalizations of \PQtree\ -- for instance, \PCtree is generalized to
 matrices with \CROP, \PQRtree and generalized \PQtree are generalized
@@
 only for recognizing interval graphs.  \cite{kr88} constructed
-efficient parallel algorithms for manipulating PQ trees. \temptext{dom
+efficient parallel algorithms for manipulating \PQtrees. \temptext{dom
   chapter 2 pg 40 Variations of PQTrees first para - summarize.}
 
-In the next few following sections we will see some of these
+In the next few sections we will see some of these
 variations.
@@
 
-\begin{definition}
+\begin{definition}[\PCtree \cite{wlh01,d08phd}]
   \label{def:pctree}
-  [\emph{\PCtree} \cite{wlh01,d08phd}.]  A \PCtree of matrix $M$ with
+   A \PCtree of matrix $M$ with
   \CROP on columns (rows), is a tree with the following properties.
@@
 \subsection{ICPIA - a set cardinality based \COP test}
-
+\label{sec:icpiasurvey}
 \temptext{narrative is abrupt}

### 2012-01-29 src/030chap3.tex
- a whole bunch of cleanups. see diff. - removed problem defs from chap 1. they were duplicates from chap 3. - CODED FIGURES IN TIKZ. forbidden subgraphs, graphmatrix example graphs. - BEGAN CHAPTER 3 EDITS.
Formatting/layout changes
-------

- Added FrenchLayout to make lists denser
- Added hanging indent for Definitions
- Compressed theorem env's above below spaces
- Creaded env for problem definition minipages.
- moved all clearpage/page counter and other prechapter settings to the chapter tex files.
- Fixed the page number bug in draft printout. removed \fancyhf{} to avoid resetting page n

@@
+\xclearpage
+
 \chapter[Tree Path Labeling]{Tree Path Labeling of Path
@@
 
-This chapter documents all the new results obtained by us
-in the area of tree path labeling of path hypergraphs which is the
-parent problem addressed in this thesis.
+This chapter documents all the new results obtained by us in the area
+of tree path labeling of path hypergraphs which is the problem
+addressed in this thesis. The organization of the chapter is as
+follows.
 
@@
 presented in Section~\ref{ch:prelims}. Section~\ref{sec:feasible}
-documents the characterization of a feasible path labeling for any
-path labeling from any target trees. Section~\ref{sec:spltargettree}
-describes two special cases where the target tree is of a particular
-family of trees. The first one is shown to be equivalent to COP
-testing in Section~\ref{sec:icpplicpia}. Section~\ref{sec:ksubdivstar}
-discusses the second special case and presents a polynomial time
-algorithm to find the tree path labeling of a given set system from a
-given \kstar. Section~\ref{sec:norestraint} discusses the plain
-vanilla version where the target tree has no restrictions and the
-algorithm to find a feasible TPL, if any, in this case.
+documents the characterization of a feasible path labeling from any
+target trees. Section~\ref{sec:spltargettree} describes two special
+cases where the target tree is of a particular family of trees. The
+first one is shown to be equivalent to COP testing in
+Section~\ref{sec:icpplicpia}. Section~\ref{sec:ksubdivstar} discusses
+the second special case and presents a polynomial time algorithm to
+find the tree path labeling of a given set system from a given
+\kstar. Section~\ref{sec:norestraint} discusses the plain vanilla
+version where the target tree has no restrictions and the algorithm to
+find a feasible TPL, if any, in this case.
 
@@
 special case of the general problem of tree path labeling of path
-hypergraphs.  The problem of consecutive-ones property testing can be
-easily seen as a simple constraint satisfaction problem involving a
-hypergraph or a system of sets from a universe. Every column (row) of
-the binary matrix can be converted into a set of non-negative integers
-which are the indices of the rows (columns) with {\un}s in that column
+hypergraphs. Moreover, Section~\ref{sec:icpiasurvey} describes how the
+problem of consecutive-ones property testing can be easily seen as a
+simple constraint satisfaction problem involving a hypergraph or a
+system of sets from a universe. Every column (row) of the binary
+matrix can be converted into a set of non-negative integers which are
+the indices of the rows (columns) with {\un}s in that column
 (row). When observed in this context, if the matrix has the COP on
@@
 form, one can see that this is indeed the problem of finding interval
-assignments to the given set system \cite{nsnrs09} with a single
-permutation of the universe (set of row or column indices for COP of
-columns or rows, respectively) which permutes each set to an
-interval. The result in \cite{nsnrs09} characterizes interval
-assignments to the sets which can be obtained from a single
-permutation of the universe. The cardinality of the interval
-assigned to it must be same as the cardinality of the set, and the
-intersection cardinality of any two sets must be same as the
+assignments to the given set system with a single permutation of the
+universe (set of row or column indices for COP of columns or rows,
+respectively) which permutes each set to an interval. The result in
+\cite{nsnrs09} (see Theorems~\ref{th:icpia2} and \ref{th:icpia1})
+characterizes interval assignments to the sets which can be obtained
+from a single permutation of the universe. The cardinality of the
+interval assigned to it must be same as the cardinality of the set,
+and the intersection cardinality of any two sets must be same as the
 intersection cardinality of the corresponding intervals. This is a
-necessary and sufficient condition. \tnote{(ICPIA) put reference to
-  survey chapter}
-
-Naturally, intervals are paths from a tree with maximum degree
-two. Thus the interval assignment problem can be generalized into path
+necessary and sufficient condition.
+
+Intervals are paths from a tree with maximum degree two. Thus the
+interval assignment problem can be naturally generalized into path
 assignment problem from any tree. We refer to this as the {\em tree
@@
 interval labeling problem in literature \cite{kklv10} to interval
-hypergraphs. To elaborate, the problem is defined as follows -- given
-a hypergraph $\cF$ from universe (hypergraph vertex set) $U$ and a target tree
-$T$, does there exist a bijection $\phi$ from $U$ to the vertices of $T$ such
-that for each hyperedge when applied to its elements, $\phi$ gives a path on
-$T$.  More formally, the problem definition is as defined by \CFTPL.
-
-{\small
-  \begin{minipage}[h]{5in}
-    \vspace{2mm}
-    {\large \CFTPL}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$ and a tree $T$.\\
-      \end{minipage}\\
-
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
+hypergraphs. In brief, the problem is defined as follows -- given a
+hypergraph $\cF$ from universe $U$ (\ie hypergraph vertex set) and a
+target tree $T$, does there exist a bijection $\phi$ from $U$ to the
+vertices of $T$ such that for each hyperedge when $\phi$ is applied to
+its elements it gives a path on $T$.  More formally, the problem
+definition is given by \CFTPL.
+
+\begin{problemdef}{\CFTPL}{A hypergraph $\cF$ with vertex set $U$ and
+    a tree $T$.}
   Does there exist a set of paths $\cP$ from $T$ and a bijection
-        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
-        {\bf true} on $(\cF, T, \cl)$.
-      \end{minipage}\\
-    \end{tabular}
-  \end{minipage}\\
-}
+  $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns {\bf
+    true} on $(\cF, T, \cl)$.
+\end{problemdef}
 
@@
 
-%\begin{tabular}[t]{c}
-{\small
-\begin{minipage}[h]{5in}
- % \begin{singlespace}
- \vspace{2mm}
- {\large \FTPL}\\
- \begin{tabular}[t]{l|l}
- \hline\\
-    {\tt Input} & 
-    \begin{minipage}[t]{\probdefwidth}
-      A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
-      paths $\cP$ from $T$ and a
-      bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
-    \end{minipage}\\
-    {\tt Question} &
-    \begin{minipage}[t]{\probdefwidth}
-      Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
-      such that $\phi$ when applied on any hyperedge in $\cF$ will
-      give
+\begin{problemdef}{\FTPL}{A hypergraph $\cF$ with vertex set $U$, a
+    tree $T$, a set of paths $\cP$ from $T$ and a bijection
+    $\cl$~$:$~$\cF \rightarrow \cP$.}
+
+  Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$ such
+  that $\phi$ when applied on any hyperedge in $\cF$ will give
   the path mapped to it by the given tree path labeling $\cl$.\\
-      { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every hyperedge
-        $S \in \cF$.}
-    \end{minipage}\\
-  \end{tabular}
-  % \end{singlespace}
-\end{minipage}\\
-}
-% \end{tabular}
+  { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every hyperedge $S
+    \in \cF$.}
+\end{problemdef}
 
@@
 problem, we were unable to discover an efficient algorithm for
-it. Hence we consider two special cases of the same -- namely, \CFTPLINT
-and \CFTPLKTREE on special target trees, namely, intervals and \kstars,
-respectively.  Section~\ref{sec:spltargettree} discusses these
-problems.
-
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \CFTPLINT}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$ and a tree $T$ with
-        maximum degree 2.\\
-      \end{minipage}\\
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
+it. However, we consider two special cases of the same on special
+target trees -- namely, \CFTPLINT and \CFTPLKTREE on intervals and
+\kstars, respectively.  Section~\ref{sec:spltargettree} discusses
+these problems and their polynomial time solutions.  \CFTPLINT is a
+demonstrated to be the known problem of \COP testing. The polynomial
+time solution to \CFTPLKTREE is a new result by us.
+
+\begin{problemdef}{\CFTPLINT}{ A hypergraph $\cF$ with vertex set $U$
+    and a tree $T$ with maximum degree 2.}
+
   Does there exist a set of paths $\cP$ from $T$ and a bijection
-        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
-        {\bf true} on $(\cF, T, \cl)$.
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
-
-{\small
-  \begin{minipage}[h]{5in}
-    % \begin{singlespace}
-    \vspace{2mm}
-    {\large \CFTPLKTREE}\\
-    \begin{tabular}[t]{l|l}
-      \hline\\
-      {\tt Input} & 
-      \begin{minipage}[t]{\probdefwidth}
-        A hypergraph $\cF$ with vertex set $U$ such that every
-        hyperedge
-        $S \in \cF$ is of cardinality at most $k+2$ and a {\kstar} $T$.\\
-      \end{minipage}\\
-      {\tt Question} &
-      \begin{minipage}[t]{\probdefwidth}
+  $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns {\bf
+    true} on $(\cF, T, \cl)$.  
+\end{problemdef}
+
+\begin{problemdef}{\CFTPLKTREE}{A hypergraph $\cF$ with vertex set $U$
+    such that every hyperedge $S \in \cF$ is of cardinality at most
+    $k+2$ and a {\kstar} $T$.}
+
   Does there exist a set of paths $\cP$ from $T$ and a bijection
-        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
-        {\bf true} on $(\cF, T, \cl)$.
-      \end{minipage}\\
-    \end{tabular}
-    % \end{singlespace}
-  \end{minipage}\\
-}
-
-\CFTPLINT is nothing but the consecutive-ones property testing
-problem.
+  $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns {\bf
+    true} on $(\cF, T, \cl)$.
+\end{problemdef}
+
+% \CFTPLINT is nothing but the consecutive-ones property testing
+% problem.
 
@@
 polynomial time is presented in
-Section~\ref{sec:norestraint}. \tnote[minor]{put some more description
-  here. refer WG11.}
-
-\remove[minor]{We focus on the question of generalizing the notion of an
+Section~\ref{sec:norestraint}.
+
+\section[Preliminaries]{Preliminaries to new results}
+\label{ch:prelims}
+
+This section describes ideas and terminologies necessary in this
+chapter. Some of these may have been seen earlier in this document,
+but is recalled for the sake of completeness of the chapter.
+
+\subsection{Set systems and hypergraphs}
+%\begin{definition}[Set systems and hypergraphs]
+
+  The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
+    system} of a universe $U$ with $|U| = n$.  The {\em support} of a
+  set system $\F$ denoted by $\supt{\cF}$ is the union of all the sets
+  in $\F$; $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this
+  paper, a set system is required to ``cover'' the universe; $
+  supp(\cF) = U$.
+  \par
+%  \noindent
+  A set system $\cF$ can also be visualized as a {\em hypergraph}
+  whose vertex set is $supp(\cF)$ and hyperedges are the sets in
+  $\cF$. This is a known representation for
+  interval systems in literature \cite{bls99,kklv10}.  We extend this
+  definition here to path systems. Due to the equivalence of set
+  system and hypergraph in the scope of this paper, we drop the
+  subscript $_H$ in the notation and refer to both the structures by
+  $\cF$.
+  \par
+%  \noindent
+  The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph
+  $\cF$ is a graph such that its vertex set has a bijection to $\cF$
+  and there exists an edge between two vertices \iff their
+  corresponding hyperedges have a non-empty intersection \cite{mcg04}.
+  \par
+%  \noindent 
+  Two hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to
+  each other, denoted by $\cF' \cong \cF''$, \iff there exists a
+  bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for
+  all sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ \iff
+  $B$ is a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
+  \cite{kklv10}, written as $B=\phi(A)$.  This is called {\em
+    hypergraph isomorphism}.  
+%\end{definition}
+
+\subsection{Path Hypergraph from a Tree}
+%\begin{definition}[Path Hypergraph from a Tree]
+  The graph $T$ represents a {\em target tree} with same number of
+  vertices as elements in $U$; $|V(T)|=|U|=n$.  A {\em path system}\,
+  $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P \mid P
+  \subseteq V, \text{ } T[P] \text{ is a path} \}$.  This generalizes
+  the fact, from the literature \cite{bls99,kklv10}, that intervals
+  can be viewed as sub-paths of a path.
+
+%  \noindent
+  If the intersection graphs of $\cF$ and $\cP$ (a path system) are
+  isomorphic, $\bI(\cF) \cong \bI(\cP)$, then the associated bijection
+  $\cl: \cF \rightarrow \cP$ due to this isomorphism is called a {\em
+    path labeling} of the hypergraph $\cF$.\footnote{Note that there
+    are two kinds of isomorphisms here.  One is the isomorphism of
+    intersection graphs on $\cF$ and $\cP$, i.e. $\bI(\cF)$ and
+    $\bI(\cP)$ respectively. Second is the isomorphism between the
+    hypergraphs $\cF$ and $\cP$.}
+
+% The subset of the path labeling that
+%   maps elements of the universe to paths that contain only a leaf the
+%   target tree is  known as a {\em leaf labeling}.
+%   \ie if the target tree has the set of $k$ leaves $L = \{v_i, i \in
+%   [k]\}$, then 
+% $\cl': \rightarrow U' \rightarrow L$ is a leaf labeling where $U'
+% \subseteq U$, $|U'| = k$.
+
+%   $\cL_0 = \{ (x_i,v_i) \mid x_i \in
+%   U, v_i \in T, x_i \ne x_j, v_i \ne v_j, i \ne j, i,j \in [k] \}$
+%   where $k$ is the number of leaves in $T$
+
+%  \noindent
+  \add{To illustrate further, let $\cg: V(\cF) \rightarrow V(\cP)$ be
+    the above mentioned isomorphism where $V(\cF)$ and $V(\cP)$ are
+    the vertex sets that represent the hyperedges for each hypergraph
+    respectively, $V(\cF) = \{ v_S \mid S \in \cF\}$ and $V(\cP) = \{
+    v_P \mid P \in \cP\}$. Then the path labeling $\cl$ is defined as
+    follows: $\cl(S_1) = P_1$ iff $\cg (v_{S_1}) = v_{P_1}$.  Just to
+    emphasize, for a path labeling $\cl$ of $\cF$ with $\cP$ as the
+    path system, $\cF^\cl$ is same as $\cP$.  The path system $\cP$
+    may be alternatively denoted in terms of $\cF$ and $\cl$ as
+    $\cF^\cl$. In most scenarios in this paper, what is given are the
+    pair $(\cF, \cl)$ and the target tree $T$; hence this notation
+    will be used more often.}
+
+%  \noindent
+  If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$ is
+  called a {\em path hypergraph} and $\cP$ is called {\em path
+    representation} of $\cF$. If this isomorphism is $\phi: supp(\cF)
+  \rightarrow V(T)$, then it is clear that there is an {\em induced
+    path labeling} $\cl_\phi: \cF \rightarrow \cP$ to the set system;
+  $\cl_\phi(S) = \set{y \mid y = \phi(x), x \in S}$ for all $S \in
+  \cF$. Recall that $supp(\cP) = V(T)$.
+
+%  \noindent
+  \add{A graph} $G$ is a {\em path graph} if it is isomorphic to the
+  intersection graph $\bI(\cP)$ of a path system $\cP$.  This
+  isomorphism gives a bijection $\cl': V(G) \rightarrow
+  \cP$. Moreover, for the purposes of this paper, we require that in a
+  path labeling, $supp(\cP) = V(T)$.  If graph $G$ is also isomorphic
+  to $\bI(\cF)$ for some hypergraph $\cF$, then clearly there is a
+  bijection $\cl: \cF \rightarrow \cP$ such that $\cl(S) = \cl'(v_S)$
+  where $v_S$ is the vertex corresponding to set $S$ in $\bI(\cF)$ for
+  any $S \in \cF$. This bijection $\cl$ is called the {\em path
+    labeling} of the hypergraph $\cF$ and the path system $\cP$ may be
+  alternatively denoted as $\cF^\cl$.
+
+%  \noindent
+  A path labeling $(\cF, \cl)$ is defined to be {\em feasible} if $\cF
+  \cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
+  \rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
+  \rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.  \add{In this
+    work}, we are given as input $\cF$ and a tree $T$, and the
+  question is whether there is a path labeling $\cl$ to a set of paths
+  in $T$.  We refer to such a solution path system by $\cF^\cl$.  A
+  path labeling $(\cF, \cl)$ is defined to be {\em feasible} if
+  % $\cF \cong \cF^\cl$ and this
+  there is a hypergraph isomorphism $\phi: supp(\cF) \rightarrow
+  supp(\cF^\cl)=V(T)$ induces a path labeling $\cl_\phi: \cF
+  \rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.
+%\end{definition}
+
+\subsection{Overlap Graphs and Marginal Hyperedges}
+%\begin{definition}[Overlap Graphs and Marginal Hyperedges]
+  An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is a graph
+  such that its vertex set has a bijection to $\cF$ and there exists
+  an edge between two of its vertices \iff their corresponding
+  hyperedges overlap. Two hyperedges $S$ and $S'$ are said to {\em
+    overlap}, denoted by $S \overlap S'$, if they have a non-empty
+  intersection and neither is contained in the other; $S \overlap S'
+  \text{ \iff } S \cap S' \ne \emptyset, S \nsubseteq S', S'
+  \nsubseteq S$. Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$
+  and not necessarily connected. Each connected component of
+  $\bO(\cF)$ is called an {\em overlap component}.
+
+%  \noindent
+  A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
+  \overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
+  \cite{kklv10}. Additionally, if $S$ is such that it is contained in
+  no other marginal hyperedge in $\cF$, then it is called {\em
+    super-marginal}.
+%\end{definition}
+
+\subsection{Miscellaneous}
+%\begin{definition}[$k$-subdivided star] 
+  A {\em star} graph is a
+  complete bipartite graph $K_{1,p}$ which is clearly a tree and $p$
+  is the number of leaves. The vertex with maximum degree is called
+  the {\em center} of the star and the edges are called {\em rays} of
+  the star. A {\em $k$-subdivided star} is a star with all its rays
+  subdivided exactly $k$ times. The definition of a {\em ray of a
+    $k$-subdivided star} is extended to the path from the center to a
+  leaf. It is clear that all rays are of length $k+2$.
+%\end{definition}
+
+% A path assignment $\A$ to $\F$ is defined as a set assignment
+% where second universe is the vertex set $V$ of a given tree $T$ and
+% every second subset in the ordered pairs is a path in this
+% tree. Formally, the definition is as follows. 
+% \begin{align*}
+%   \A = \{ (S_i,P_i) \mid S_i \in \F, P_i \subseteq V \text{
+%     s.t. }T[P_i] \text{ is a path, } i \in I \}
+% \end{align*}
+% In other words, $P_i$ is the path on the tree $T$ assigned to $S_i$ in
+% $\A$. As mentioned before for set systems, the paths cover the whole
+% tree, i.e. $\bigcup_{i \in I}P_i = V$ \\
+
+% Generalizing the definition of { feasibility} in cite{nsnrs09} to
+% a set assignment, a path assignment $\A$ is defined to be {
+%   feasible} if there exists a bijection defined as follows.
+% \begin{align}
+%   \sigma: U \rightarrow V(T), \text{ such that }\sigma(S_i) = P_i
+%   \text{ for all } i \in I, \sigma \text{ is a bijection}
+%   \label{eq:stf}
+% \end{align}
+
+%\begin{definition}[Partially ordered sets]
+
+  Let $X$ be a partially ordered set with $\preccurlyeq$ being the
+  partial order on $X$.  $X_m \in X$ is a maximal upper bound of $X$
+  if $\nexists X_q \in X$ such that $X_m \preccurlyeq X_q$.  A maximal
+  upper bound on $X$ is denoted by $mub(X)$.
+  %represents an element in $X$ which is a maximal upper bound on $X$.
+%\end{definition}
+
+An {\em in-tree} is a directed rooted tree in which all edges are
+directed toward to the root.
+
+The set $I$ represents the index set $[m]$. If index $i$ is used
+without further qualification, it is intended that $i \in I$.
+
+\section[Characterization of FTPL]{Characterization of Feasible Tree
+  Path Labeling}
+\label{sec:feasible}
+
+\tnote{\sc refer to Problem Defs here}
+In this section we give an algorithmic characterization of the
+feasibility of a tree path labeling.  
+We focus on the question of generalizing the notion of an
   ICPIA cite-nsnrs09 to characterize feasible path assignments.  We
@@
   the set system is feasible.  This generalizes the result in
-  cite-nsnrs09.}
-
-\section[Preliminaries]{Preliminaries to new results}
-\label{ch:prelims}
-
-This section states definitions and basic facts necessary in the scope
-of this document.
-
-\subsection{Hypergraph Preliminaries}
-
-\begin{definition}[Set systems and hypergraphs] 
-  \noindent 
-  The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
-    system} of a universe $U$ with $|U| = n$.  The {\em support} of a
-  set system $\F$ denoted by $\supt{\cF}$\tnote[minor]{change all supp
-    references to supt command} is the union of all the sets in $\F$;
-  $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this paper, a
-  set system is required to ``cover'' the universe; $ supp(\cF) = U$.
-
-  \noindent
-  A set system $\cF$ can also be visualized as a {\em hypergraph}
-  whose vertex set is $supp(\cF)$ and hyperedges are the sets in
-  $\cF$. This is a known representation for
-  interval systems in literature \cite{bls99,kklv10}.  We extend this
-  definition here to path systems. Due to the equivalence of set
-  system and hypergraph in the scope of this paper, we drop the
-  subscript $_H$ in the notation and refer to both the structures by
-  $\cF$.
-
-  \noindent
-  The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph
-  $\cF$ is a graph such that its vertex set has a bijection to $\cF$
-  and there exists an edge between two vertices \iff their
-  corresponding hyperedges have a non-empty intersection \cite{mcg04}.
-
-  \noindent 
-  Two hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to
-  each other, denoted by $\cF' \cong \cF''$, \iff there exists a
-  bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for
-  all sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ \iff
-  $B$ is a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
-  \cite{kklv10}, written as $B=\phi(A)$.  This is called {\em
-    hypergraph isomorphism}.  
-\end{definition}
-
-\begin{definition}[Path Hypergraph from a Tree]
-  The graph $T$ represents a {\em target tree} with same number of
-  vertices as elements in $U$; $|V(T)|=|U|=n$.  A {\em path system}\,
-  $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P \mid P
-  \subseteq V, \text{ } T[P] \text{ is a path} \}$.  This generalizes
-  the fact, from the literature \cite{bls99,kklv10}, that intervals
-  can be viewed as sub-paths of a path.
-
-  \noindent
-  If the intersection graphs of $\cF$ and $\cP$ (a path system) are
-  isomorphic, $\bI(\cF) \cong \bI(\cP)$, then the associated bijection
-  $\cl: \cF \rightarrow \cP$ due to this isomorphism is called a {\em
-    path labeling} of the hypergraph $\cF$.\footnote{Note that there
-    are two kinds of isomorphisms here.  One is the isomorphism of
-    intersection graphs on $\cF$ and $\cP$, i.e. $\bI(\cF)$ and
-    $\bI(\cP)$ respectively. Second is the isomorphism between the
-    hypergraphs $\cF$ and $\cP$.}
-
-% The subset of the path labeling that
-%   maps elements of the universe to paths that contain only a leaf the
-%   target tree is  known as a {\em leaf labeling}.
-%   \ie if the target tree has the set of $k$ leaves $L = \{v_i, i \in
-%   [k]\}$, then 
-% $\cl': \rightarrow U' \rightarrow L$ is a leaf labeling where $U'
-% \subseteq U$, $|U'| = k$.
-
-%   $\cL_0 = \{ (x_i,v_i) \mid x_i \in
-%   U, v_i \in T, x_i \ne x_j, v_i \ne v_j, i \ne j, i,j \in [k] \}$
-%   where $k$ is the number of leaves in $T$
-
-  \noindent
-  \add{To illustrate further, let $\cg: V(\cF) \rightarrow V(\cP)$ be
-    the above mentioned isomorphism where $V(\cF)$ and $V(\cP)$ are
-    the vertex sets that represent the hyperedges for each hypergraph
-    respectively, $V(\cF) = \{ v_S \mid S \in \cF\}$ and $V(\cP) = \{
-    v_P \mid P \in \cP\}$. Then the path labeling $\cl$ is defined as
-    follows: $\cl(S_1) = P_1$ iff $\cg (v_{S_1}) = v_{P_1}$.  Just to
-    emphasize, for a path labeling $\cl$ of $\cF$ with $\cP$ as the
-    path system, $\cF^\cl$ is same as $\cP$.  The path system $\cP$
-    may be alternatively denoted in terms of $\cF$ and $\cl$ as
-    $\cF^\cl$. In most scenarios in this paper, what is given are the
-    pair $(\cF, \cl)$ and the target tree $T$; hence this notation
-    will be used more often.}
-
-  \noindent
-  If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$ is
-  called a {\em path hypergraph} and $\cP$ is called {\em path
-    representation} of $\cF$. If this isomorphism is $\phi: supp(\cF)
-  \rightarrow V(T)$, then it is clear that there is an {\em induced
-    path labeling} $\cl_\phi: \cF \rightarrow \cP$ to the set system;
-  $\cl_\phi(S) = \set{y \mid y = \phi(x), x \in S}$ for all $S \in
-  \cF$. Recall that $supp(\cP) = V(T)$.
-
-  \noindent
-  \add{A graph} $G$ is a {\em path graph} if it is isomorphic to the
-  intersection graph $\bI(\cP)$ of a path system $\cP$.  This
-  isomorphism gives a bijection $\cl': V(G) \rightarrow
-  \cP$. Moreover, for the purposes of this paper, we require that in a
-  path labeling, $supp(\cP) = V(T)$.  If graph $G$ is also isomorphic
-  to $\bI(\cF)$ for some hypergraph $\cF$, then clearly there is a
-  bijection $\cl: \cF \rightarrow \cP$ such that $\cl(S) = \cl'(v_S)$
-  where $v_S$ is the vertex corresponding to set $S$ in $\bI(\cF)$ for
-  any $S \in \cF$. This bijection $\cl$ is called the {\em path
-    labeling} of the hypergraph $\cF$ and the path system $\cP$ may be
-  alternatively denoted as $\cF^\cl$.
-
-  \noindent
-  A path labeling $(\cF, \cl)$ is defined to be {\em feasible} if $\cF
-  \cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
-  \rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
-  \rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.  \add{In this
-    work}, we are given as input $\cF$ and a tree $T$, and the
-  question is whether there is a path labeling $\cl$ to a set of paths
-  in $T$.  We refer to such a solution path system by $\cF^\cl$.  A
-  path labeling $(\cF, \cl)$ is defined to be {\em feasible} if
-  % $\cF \cong \cF^\cl$ and this
-  there is a hypergraph isomorphism $\phi: supp(\cF) \rightarrow
-  supp(\cF^\cl)=V(T)$ induces a path labeling $\cl_\phi: \cF
-  \rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.
-\end{definition}
-
-\begin{definition}[Overlap Graphs and Marginal Hyperedges]
-  An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is a graph
-  such that its vertex set has a bijection to $\cF$ and there exists
-  an edge between two of its vertices \iff their corresponding
-  hyperedges overlap. Two hyperedges $S$ and $S'$ are said to {\em
-    overlap}, denoted by $S \overlap S'$, if they have a non-empty
-  intersection and neither is contained in the other; $S \overlap S'
-  \text{ \iff } S \cap S' \ne \emptyset, S \nsubseteq S', S'
-  \nsubseteq S$. Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$
-  and not necessarily connected. Each connected component of
-  $\bO(\cF)$ is called an {\em overlap component}.
-
-  \noindent
-  A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
-  \overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
-  \cite{kklv10}. Additionally, if $S$ is such that it is contained in
-  no other marginal hyperedge in $\cF$, then it is called {\em
-    super-marginal}.
-\end{definition}
-
-\begin{definition}[$k$-subdivided star] 
-  A {\em star} graph is a
-  complete bipartite graph $K_{1,p}$ which is clearly a tree and $p$
-  is the number of leaves. The vertex with maximum degree is called
-  the {\em center} of the star and the edges are called {\em rays} of
-  the star. A {\em $k$-subdivided star} is a star with all its rays
-  subdivided exactly $k$ times. The definition of a {\em ray of a
-    $k$-subdivided star} is extended to the path from the center to a
-  leaf. It is clear that all rays are of length $k+2$.
-\end{definition}
-
-\temptext{ --- WG11 begin --- }
-
-\remove[dire]{
-A path assignment }% $\A$ to $\F$ is defined as a set assignment
-% where second universe is the vertex set $V$ of a given tree $T$ and
-% every second subset in the ordered pairs is a path in this
-% tree. Formally, the definition is as follows. 
-% \begin{align*}
-%   \A = \{ (S_i,P_i) \mid S_i \in \F, P_i \subseteq V \text{
-%     s.t. }T[P_i] \text{ is a path, } i \in I \}
-% \end{align*}
-% In other words, $P_i$ is the path on the tree $T$ assigned to $S_i$ in
-% $\A$. As mentioned before for set systems, the paths cover the whole
-% tree, i.e. $\bigcup_{i \in I}P_i = V$ \\
-
-% Generalizing the definition of { feasibility} in cite{nsnrs09} to
-% a set assignment, a path assignment $\A$ is defined to be {
-%   feasible} if there exists a bijection defined as follows.
-% \begin{align}
-%   \sigma: U \rightarrow V(T), \text{ such that }\sigma(S_i) = P_i
-%   \text{ for all } i \in I, \sigma \text{ is a bijection}
-%   \label{eq:stf}
-% \end{align}
-
-\begin{definition}[Partially ordered sets]
-
-Let $X$ be a partially ordered set with $\preccurlyeq$ being the
-partial order on $X$.  $mub(X)$ represents an element in $X$ which is
-a maximal upper bound on $X$.  $X_m \in X$ is a maximal upper bound of
-$X$ if $\nexists X_q \in X$ such that $X_m
-\preccurlyeq X_q$. 
-\end{definition}
-
-\noindent
-The set $I$ represents the index set $[m]$. If index $i$ is used
-without further qualification, it is meant to be $i \in I$. Any
-function, if not defined on a domain of sets, when applied on a set is
-understood as the function applied to each of its elements. i.e. for
-any function $f$ defined with domain $U$, the abuse of notation is as
-follows; $f(S)$ is used instead of $\hat f(S)$ where
-$\hat f(S) = \{y \mid y = f(x), x \in S\}$. \\
-
-\noindent
-When refering to a tree as $T$ it could be a reference to the tree
-itself, or the vertices of the tree. This will be clear from the
-context.\\
-
-\noindent
-Finally, an in-tree is a directed rooted tree in which all edges are
-directed toward to the root.
-
-\temptext{ --- WG11 end --- }
-
-\section[Characterization of FTPL]{Characterization of Feasible Tree Path Labeling}
-\label{sec:feasible}
-
-\tnote{\sc refer to Problem Defs here}
-In this section we give an algorithmic characterization of the
-feasibility of a tree path labeling.  Consider a path labeling $(\cF,
+  cite-nsnrs09.
+
+Consider a path labeling $(\cF,
 \cl)$ on the given tree $T$. We call $(\cF, \cl)$ an {\em Intersection

### 2012-01-31 src/010chap1.tex
- Figure 3.1 for hypergraphs added. Preliminaries section complete. - Characterization of FTPL section moreorless done except section for worked out example -- changed h.isomorphism induced tpl notation to \cp_{\phi} since \cl is confusing because it is the input in most places - Filled up some definitions in Chap 1 - Added asteroidal triples def in chapter 2

@@
 
+\tnote[dire]{
 \section[Basic preliminaries]{Basic preliminaries} 
@@
 
-\begin{definition}[\emph{Binary matrix}.]
+\begin{definition}[Binary matrix]
   \label{def:binmatrix}
   Let $M$ be an $n \times m$ matrix. $m_{i,j}$ denotes its $(i,j)$th
-  element, i.e. element at $i$th row and $j$th column. $M$ is a binary
-  matrix if each of its element is 0 or \un.  In other words, for all
+  element, i.e. element at $i$th row and $j$th column. $M$ is a \textbf{binary
+  matrix} if each of its element is 0 or \un.  In other words, for all
   $i \in [n]$ and $j \in [m]$, $m_{i,j} \in \{0, \un\}$
@@
 
-\begin{definition}
-  A {\em permutation} $\lambda$ of a set $X = \{x_1, x_2, \ldots
+\begin{definition}[Permutation]
+  A \textbf{permutation} $\lambda$ of a set $X = \{x_1, x_2, \ldots
   x_n\}$ is a bijection $\lambda: X \rightarrow X$.  For simplicity,
@@
 
-\begin{definition}[\emph{\Cop}]
+\begin{definition}[\Cop]
   \label{def:copmatrix}
-  \begin{enumerate}
-  \item \Cop
-  \item Strong \cop
-  \item \Cop order or \cop permutation
-  \end{enumerate}
+  Let $M$ be a binary matrix.
+  \begin{itemize}
+  \item \textbf{\Cop}
+  \item \textbf{Strong \cop}
+  \item \textbf{\Cop order} or \cop permutation
+  \end{itemize}
 \end{definition}
@@
 \figcopmatrix
-\tnote[important]{have a figure for set systems also}
 
-\begin{definition}[\emph{\Crop}]
+\begin{definition}[\Crop]
   \label{def:cropmatrix}
@@
 
-\begin{definition}[\emph{set system}]
+\begin{definition}[Set system]
   \label{def:setsystem}
 
-    also include overlapping sets.
+  Let $U$ be a universe with $|U| = n$.  
+  \begin{itemize}
+  \item The set $\F \subseteq (2^{U} \setminus \emptyset)$ is called a
+    \textbf{set system} with universe $U$.
+  \item Two sets $S, S' \in \cF$ are said to \textbf{overlap}, denoted
+    by $S \overlap S'$, if they have a non-empty intersection and
+    neither is contained in the other.
+    \[S \overlap S' \text{ \iff\ } S \cap S' \ne \emptyset, S
+    \nsubseteq S', S' \nsubseteq S\]
+  \end{itemize}
 \end{definition}
 
-\begin{definition}[\emph{Graph, Tree}]
+\begin{definition}[Partial order]
+  \label{def:poset}    
+  Let $A$ be a set and $R$ be a binary relation $R \subseteq A \times
+  A$.
+
+  \textbf{poset}
+
+  Let $X$ be a partially ordered set with $\preccurlyeq$ being the
+  partial order on $X$.
+
+  \begin{itemize}
+  \item \textbf{hasse diagram}
+  \item \textbf{maximal upper bound}  $X_m \in X$ is a \textbf{maximal upper bound} of $X$
+  if $\nexists X_q \in X$ such that $X_m \preccurlyeq X_q$.  A maximal
+  upper bound on $X$ is denoted by $mub(X)$.
+
+  \item A set of sets form a \textbf{single
+      inclusion chain} when the Hasse diagram of their subset relation
+    forms a single chain.
+  \end{itemize}
+\end{definition}
+
+\begin{definition}[Graph, Tree]
   \label{def:graphtree}  
+  \tnote[dire]{take from doms}
+\end{definition}
 
-\end{definition}
 When refering to a tree as $T$ it could be a reference to the tree
@@
 
-\begin{definition}[\emph{Maximal clique}]
+\begin{definition}[Intersection graph]
+  \label{def:intersectiongraph}
+  Let $\cF$ be a set system. Then its \textbf{intersection graph}
+  $\bI(\cF)$ is a graph \stt its vertex set has a bijection to $\cF$
+  and there exists an edge between two vertices \iff their
+  corresponding sets in $\cF$ have a non-empty intersection.
+\end{definition}
+
+\begin{definition}[Interval graph, Path graph]
+  \label{def:pathgraph}
+
+  \begin{itemize}
+  \item A graph $G$ is an \textbf{interval graph} if there a set of intervals
+    $\cI$ \stt $G$ is isomorphic to the intersection graph of
+    $\cI$. \[G \cong \bI(\cI)\]
+  \item A graph $G$ is a \textbf{path graph} if there exists a tree $T$ and a
+    set of paths from $T$, $\cP$ \stt $G$ is isomorphic to the
+    intersection graph of $\cP$. \[G \cong \bI(\cP)\]
+  \end{itemize}
+\end{definition}
+Moreover, a \textbf{chordal graph} is characterized as an intersection
+graph of subtrees of a tree.
+
+\begin{definition}[Maximal clique]
   \label{def:maxclique}    
 \end{definition}
+}
 

### 2012-01-31 src/020chap2.tex
- Figure 3.1 for hypergraphs added. Preliminaries section complete. - Characterization of FTPL section moreorless done except section for worked out example -- changed h.isomorphism induced tpl notation to \cp_{\phi} since \cl is confusing because it is the input in most places - Filled up some definitions in Chap 1 - Added asteroidal triples def in chapter 2

@@
 Theorem~\ref{th:tuckeratfree} states that convex bipartite graphs
-cannot have {\em asteroidal triples}\footnote{If $G = (V,E)$ is a
-  graph, a set of three vertices from $V$ form an {\em asteroidal
-    triple} if between any two of them there exists a path in $G$ that
-  does not contain any vertex from the closed neighborhood of the
-  third vertex.} contained in the corresponding vertex
-partition\footnote{The partition corresponds to columns (rows) if its
-  half adjacency matrix has \COP columns (rows).}.
-Theorem~\ref{th:tuckerforbidden} lists the structures in a
-bipartite graph that force one of its vertex partitions to have
-asteriodal triples -- in other words, it identifies the subgraphs that
-prevent the graph from being convex bipartite.
+cannot have {\em asteroidal triples} (see
+Definition~\ref{def:asteroidal}) contained in the corresponding vertex
+partition. This partition of the bipartite graph corresponds to
+columns (rows) if its half adjacency matrix has \COP columns (rows).
+Theorem~\ref{th:tuckerforbidden} lists the structures in a bipartite
+graph that force one of its vertex partitions to have asteriodal
+triples -- in other words, it identifies the subgraphs that prevent
+the graph from being convex bipartite.
+
+\begin{definition}[Asteroidal triple]
+  \label{def:asteroidal}
+  % If $G = (V,E)$ is a graph, a set of three vertices from $V$ form
+  % an {\em asteroidal triple} if between any two of them there exists
+  % a path in $G$ that does not contain any vertex from the closed
+  % neighborhood of the third vertex.
+  An {\em asteroidal triple} in a graph $G$ is a tuple of vertices
+  $\{v_1,v_2,v_3\} \subset V$ such that between every pair of vertices
+  from this tuple there exists a path that avoids the closed
+  neighbourbood of the third vertex.
+\end{definition}
 
@@
 \COP order for each overlap component. In each component, the set
-whose overlaps with the rest of the sets in that component forms
-a single inclusion chain\footnote{A set of sets form a {\em single
-  inclusion chain} when the Hasse diagram of their subset relation
-forms a single chain. In this particular reference, the set of sets is
-the set of overlaps.} is chosen first and assigned the
-leftmost\footnote{By convention. Rightmost can also be chosen
-  analogously to get a different \COP order.} interval. The next
-candidate set is chosen from this set's overlapping sets. The interval
-assigned is calculated by simple intersection cardinality
-means\footnote{Assume that the next set overlaps on one of the
-  ``ends'' of the current set and calculate the interval by shifting
-  left or right of the current interval and adjusting its size to
-  match the next set's cardinality.}. Once the next set is assigned an
-interval, the current assigned sets are all checked for ICPIA. This is
-a backtracking algorithm -- if at any point ICPIA fails, another
-overlapping set is chosen. If all overlapping set attempts fail ICPIA,
-a different ``first set'' is chosen. This is a remarkably simple
-algorithm due to its obvious simplicity in implementation.
+whose overlaps with the rest of the sets in that component forms a
+single inclusion chain (see Definition~\ref{def:poset}. In this
+particular reference, the set of sets is the set of overlaps.)  is
+chosen first and assigned the leftmost\footnote{By
+  convention. Rightmost can also be chosen analogously to get a
+  different \COP order.} interval. The next candidate set is chosen
+from this set's overlapping sets. The interval assigned is calculated
+by simple intersection cardinality means\footnote{Assume that the next
+  set overlaps on one of the ``ends'' of the current set and calculate
+  the interval by shifting left or right of the current interval and
+  adjusting its size to match the next set's cardinality.}. Once the
+next set is assigned an interval, the current assigned sets are all
+checked for ICPIA. This is a backtracking algorithm -- if at any point
+ICPIA fails, another overlapping set is chosen. If all overlapping set
+attempts fail ICPIA, a different ``first set'' is chosen. This is a
+remarkably simple algorithm due to its obvious simplicity in
+implementation.
 

### 2012-01-31 src/030chap3.tex
- Figure 3.1 for hypergraphs added. Preliminaries section complete. - Characterization of FTPL section moreorless done except section for worked out example -- changed h.isomorphism induced tpl notation to \cp_{\phi} since \cl is confusing because it is the input in most places - Filled up some definitions in Chap 1 - Added asteroidal triples def in chapter 2

@@
 
-\subsection{Set systems and hypergraphs}
-%\begin{definition}[Set systems and hypergraphs]
-
-  The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
-    system} of a universe $U$ with $|U| = n$.  The {\em support} of a
-  set system $\F$ denoted by $\supt{\cF}$ is the union of all the sets
-  in $\F$; $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this
-  paper, a set system is required to ``cover'' the universe; $
-  supp(\cF) = U$.
+A note on attribution -- the machinery for labeling of hypergraphs
+has been adopted from \cite{kklv10} and adapted to the needs of our
+work. Some classic graph definitions are from \cite{mcg04}.
+
+\subsection{Set systems and Hypergraphs}
+\label{sec:setsyshypergraph}
+
+Consider a universe $U$ with $|U| = n$.  The set $\F \subseteq (2^{U}
+\setminus \emptyset)$ is a {\em set system} with universe $U$.  The
+{\em support of set system} $\F$ denoted by $\supt{\cF}$, is the union
+of all the sets in $\F$.
+\[supp(\F) = \bigcup_{S \in \F}S\]%
+By convention, the set system $\cF$ ``covers'' its universe:
+$supp(\cF) = U$. If the sets in $\cF$ are intervals or paths of a
+tree, then it is called an interval system or a path system,
+respectively.
 \par
-%  \noindent
 A set system $\cF$ can also be visualized as a {\em hypergraph}
-  whose vertex set is $supp(\cF)$ and hyperedges are the sets in
-  $\cF$. This is a known representation for
-  interval systems in literature \cite{bls99,kklv10}.  We extend this
-  definition here to path systems. Due to the equivalence of set
-  system and hypergraph in the scope of this paper, we drop the
-  subscript $_H$ in the notation and refer to both the structures by
-  $\cF$.
+$\cF_H$ whose vertex set is $U$ and hyperedges are the sets in
+$\cF$. This is a known representation for interval systems in
+literature \cite{bls99,kklv10}.  We extend this definition here to
+path systems. Due to the equivalence of set system and hypergraph as
+far as this thesis is concerned, we drop the subscript $_H$ in the
+notation and refer to both the structures by $\cF$. Thus hypergraphs
+and set systems are synonyms. So are hyperedges and sets.
 \par
-%  \noindent
-  The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph
-  $\cF$ is a graph such that its vertex set has a bijection to $\cF$
-  and there exists an edge between two vertices \iff their
-  corresponding hyperedges have a non-empty intersection \cite{mcg04}.
-  \par
-%  \noindent 
-  Two hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to
-  each other, denoted by $\cF' \cong \cF''$, \iff there exists a
-  bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for
-  all sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ \iff
-  $B$ is a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
-  \cite{kklv10}, written as $B=\phi(A)$.  This is called {\em
-    hypergraph isomorphism}.  
-%\end{definition}
-
-\subsection{Path Hypergraph from a Tree}
-%\begin{definition}[Path Hypergraph from a Tree]
+The notion of {\em hypergraph isomorphism} is central to the work of
+this thesis. This is defined by Definition~\ref{def:hisomorph}.
+
+\begin{definition}[Hypergraph isomorphism]
+  \label{def:hisomorph}
+  Let $\cF'$, $\cF''$ be two hypergraphs.  They are said to be {\em
+    isomorphic} to each other, denoted by $\cF' \cong \cF''$, \iff
+  there exists a bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$
+  such that for all sets $A \subseteq supp(\cF')$, $A$ is a hyperedge
+  in $\cF'$ \iff $B$ is a hyperedge in $\cF''$ where $B = \{\phi(x)
+  \mid x \in A\}$ (this is also written as $B=\phi(A)$).
+\end{definition}
+
+Figure~\ref{fig:hypergraphs} shows an example of hypergraphs and
+hypergraph isomorphism.
+
+Recall intersection graph and path graph from
+Definition~\ref{def:intersectiongraph} and \ref{def:pathgraph}. The
+{\em intersection graph} $\bI(\cF)$ of a hypergraph or set system
+$\cF$ is a graph such that its vertex set has a bijection to $\cF$ and
+there exists an edge between two vertices \iff their corresponding
+hyperedges have a non-empty intersection. A graph $G$ is a path graph
+if it is isomorphic to the intersection graph of a path system.
+
+\begin{figure}[htb]
+  \centering
+
+  \begin{tabular}[h]{l}
+
+    \begin{tabular}[h]{ll}
+      \begin{tabular}[t]{l}
+        $\cF_1:$ \\
+        \hypergraphi%
+      \end{tabular} 
+      &     
+      \begin{tabular}[t]{l}
+        $\cF_2:$\\
+        \hypergraphii %        
+      \end{tabular}
+    \end{tabular}
+    \\
+    \begin{tabular}[h]{ll}
+      \begin{tabular}[t]{l}
+        \\ \\
+        $\cP$:\\
+        \pathhypergraphi %       
+      \end{tabular}    
+      &     
+      \begin{tabular}[t]{l}
+        $\cF_3:$\\
+        \hypergraphiii %        
+      \end{tabular}
+    \end{tabular}
+    \\  
+    \begin{tabular}[h]{l}
+      $\cI$:\\
+      \intervalhypergraphi %
+    \end{tabular}
+  \end{tabular}
+
+  \caption[\figtabsize Hypergraphs and set systems]{\figtabsize
+    Hypergraphs and set systems. $\cF_1$ and $\cF_2$ are isomorphic
+    hypergraphs. The subscript of each element is such that if $\phi$
+    is the hypergraph isomorphism, $\phi(x_i)=y_i$. $\cI$ and $\cP$
+    are an interval system and a path system respectively which are both
+    isomorphic to $\cF_1$ (and by transitivity, isomorphic to $\cF_2$). Here too the
+    subscript of elements indicate the isomorphism bijection. $\cF_3$
+    is a hypergraph that is not isomorphic to any of the other
+    hypergraphs -- moreover, it is one that cannot have a feasible
+    path labeling to any tree since it does not have the Helly's
+    property anymore due to the addition of hyperedge $\{x_1,
+    x_4\}$. }
+  \label{fig:hypergraphs}
+\end{figure}
+
+\subsection{Path Labeling and Path Hypergraphs}
+\label{sec:pathhypergraph}
 The graph $T$ represents a {\em target tree} with same number of
-  vertices as elements in $U$; $|V(T)|=|U|=n$.  A {\em path system}\,
-  $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P \mid P
-  \subseteq V, \text{ } T[P] \text{ is a path} \}$.  This generalizes
-  the fact, from the literature \cite{bls99,kklv10}, that intervals
-  can be viewed as sub-paths of a path.
-
-%  \noindent
-  If the intersection graphs of $\cF$ and $\cP$ (a path system) are
-  isomorphic, $\bI(\cF) \cong \bI(\cP)$, then the associated bijection
-  $\cl: \cF \rightarrow \cP$ due to this isomorphism is called a {\em
-    path labeling} of the hypergraph $\cF$.\footnote{Note that there
-    are two kinds of isomorphisms here.  One is the isomorphism of
-    intersection graphs on $\cF$ and $\cP$, i.e. $\bI(\cF)$ and
-    $\bI(\cP)$ respectively. Second is the isomorphism between the
-    hypergraphs $\cF$ and $\cP$.}
-
+vertices as elements in $U$; $|V(T)|=|U|=n$.  A {\em path system}
+$\cP$ is a set system of paths from $T$. 
+\[\cP \subseteq \{P \mid P \subseteq V, \text{ } T[P] \text{ is a
+  path} \}\]
+
+A special mapping of paths from target tree $T$ to hyperedges in $\cF$
+is called a {\em path labeling}. Definition~\ref{def:pathlabel} gives
+details. Note that the term ``tree path labeling'' is synonymous
+with ``path labeling'' in this document since we only consider trees and
+not general graphs as source of paths.
+
+\begin{definition}[Path labeling of a hypergraph]
+  \label{def:pathlabel}
+  Consider a hypergraph $\cF$ % with universe $U$
+  and a path system $\cP$ from tree $T$ such that $\bI(\cF) \cong
+  \bI(\cP)$.  Then the associated bijection $\cl: \cF
+  \rightarrow \cP$ due to this isomorphism is called a {\em path
+    labeling} of $\cF$ from $T$. The path system $\cP$ is
+  alternatively denoted as $\cF^\cl$. The path labeling may also be
+  denoted by ordered pairs $(\cF,\cl)$ or $(\cF,\cl,T)$.
+\end{definition}
+% If the intersection graphs of $\cF$ and $\cP$ are isomorphic,
+% $\bI(\cF) \cong \bI(\cP)$, then the associated bijection $\cl: \cF
+% \rightarrow \cP$ due to this isomorphism is called a {\em path
+%   labeling} of the hypergraph $\cF$.
+To elaborate on the phrase ``associated bijection'' in
+Definition~\ref{def:pathlabel}, let $\cg: V(\bI(\cF)) \rightarrow
+V(\bI(\cP))$ be the given intersection graph isomorphism, where
+$V(\bI(\cF)) = \{ v_S \mid S \in \cF\}$ and $V(\bI(\cP)) = \{v_P \mid
+P \in \cP\}$. Then the associated bijection $\cl: \cF \rightarrow \cP$
+is defined as follows. For all $S' \in \cF$ and $P' \in \cP$, $\cl(S')
+= P'$ \iff $\cg(v_{S'}) = v_{P'}$.  Thus, it can be seen that a path
+labeling preserves the non-empty intersection property between every
+pair of hyperedges in $\cF$.
+
+At this point, it is worth comparing Definition~\ref{def:pathlabel}
+with Definition~\ref{def:hisomorph} seen in the previous section.  It
+may be noted that there are two kinds of isomorphisms here. One is the
+isomorphism of intersection graphs of $\cF$ and $\cP$, i.e. $\bI(\cF)$
+and $\bI(\cP)$ and second is the isomorphism between the hypergraphs
+$\cF$ and $\cP$ themselves. It is easy to verify that given a path
+labeling of hypergraph $\cF$, there need not be a hypergraph
+isomorphism involved. In fact, checking if an induced hypergraph
+isomorphism results from a given path labeling is the problem of
+\CFTPL.  However, a hypergraph isomorphism always induces a path
+labeling as shown by the following Definition~\ref{def:inducedpl}.
+
+\begin{definition}[Hypergraph isomorphism induced path labeling]
+  \label{def:inducedpl}
+  Let $\cF$ be a hypergraph and $\cP$ be a path system \stt $\cF \cong
+  \cP$. Let  $\phi: \supt{\cF} \rightarrow \supt{\cP}$ be this hypergraph
+  isomorphism function. % described in
+  % Definition~\ref{def:pathhypergraph}.
+  Then there is an {\em induced path labeling} % denoted by $\cl_\phi:
+%   \cF \rightarrow \cP$ 
+  denoted by $\cp_\phi: \cF \rightarrow \cP$
+  of the hypergraph \stt % $\cl_\phi(S) = \phi(S)$
+  $\cp_\phi(S) = \phi(S)$
+  (see Definition~\ref{def:hisomorph}) for all $S \in \cF$.  
+\end{definition}
+
+Combining the ideas of path labeling, hypergraph isomorphism and its
+induced path labeling we now define {\em feasiblity} of a path
+labeling in Definition~\ref{def:feasibletpl}.
+
+\begin{definition}[Feasible path labeling]
+  \label{def:feasibletpl}
+  Let $\cF$ be a hypergraph and $(\cF, \cl)$ be a path
+  labeling. $(\cF, \cl)$ is called a {\em feasible path labeling} if
+  the following conditions hold.
+  \begin{itemize}
+  \item $\cF \cong \cF^\cl$
+  \item if the above hypergraph isomorphism is $\phi: supp(\cF)
+    \rightarrow supp(\cF^\cl)$, then $\cp_\phi = \cl$
+  \end{itemize}
+
+  This isomorphism $\phi$ is called the \textbf{feasibility hypergraph
+  isomorphism} or feasibility isomorphism.
+\end{definition}
+
+One of the problems solved in this thesis \FTPL characterizes
+feasibile tree path labeling.
+
+Now we will see the definition of a {\em path hypergraph} which is a
+simple idea extended from interval hypergraphs in \cite{kklv10}.
+
+\begin{definition}[Path hypergraph]
+  \label{def:pathhypergraph}
+  Let $\cF$ be a set system. $\cF$ is called a {\em path hypergraph}
+  if there exists a target tree $T$ and a path system $\cP$ from $T$
+  \stt $\cF \cong \cP$. If $\cF$ is thus a path hypergraph, $\cP$ is
+  called {\em path representation} of $\cF$.
+\end{definition}
+
+In other words, if there exists a feasible tree path labeling for a
+given hypergraph, it is a path hypergraph.
+
+%% XXXXXX
 % The subset of the path labeling that
@@
 % \subseteq U$, $|U'| = k$.
-
 %   $\cL_0 = \{ (x_i,v_i) \mid x_i \in
@@
 
-%  \noindent
-  \add{To illustrate further, let $\cg: V(\cF) \rightarrow V(\cP)$ be
-    the above mentioned isomorphism where $V(\cF)$ and $V(\cP)$ are
-    the vertex sets that represent the hyperedges for each hypergraph
-    respectively, $V(\cF) = \{ v_S \mid S \in \cF\}$ and $V(\cP) = \{
-    v_P \mid P \in \cP\}$. Then the path labeling $\cl$ is defined as
-    follows: $\cl(S_1) = P_1$ iff $\cg (v_{S_1}) = v_{P_1}$.  Just to
-    emphasize, for a path labeling $\cl$ of $\cF$ with $\cP$ as the
-    path system, $\cF^\cl$ is same as $\cP$.  The path system $\cP$
-    may be alternatively denoted in terms of $\cF$ and $\cl$ as
-    $\cF^\cl$. In most scenarios in this paper, what is given are the
-    pair $(\cF, \cl)$ and the target tree $T$; hence this notation
-    will be used more often.}
-
-%  \noindent
-  If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$ is
-  called a {\em path hypergraph} and $\cP$ is called {\em path
-    representation} of $\cF$. If this isomorphism is $\phi: supp(\cF)
-  \rightarrow V(T)$, then it is clear that there is an {\em induced
-    path labeling} $\cl_\phi: \cF \rightarrow \cP$ to the set system;
-  $\cl_\phi(S) = \set{y \mid y = \phi(x), x \in S}$ for all $S \in
-  \cF$. Recall that $supp(\cP) = V(T)$.
-
-%  \noindent
-  \add{A graph} $G$ is a {\em path graph} if it is isomorphic to the
-  intersection graph $\bI(\cP)$ of a path system $\cP$.  This
-  isomorphism gives a bijection $\cl': V(G) \rightarrow
-  \cP$. Moreover, for the purposes of this paper, we require that in a
-  path labeling, $supp(\cP) = V(T)$.  If graph $G$ is also isomorphic
-  to $\bI(\cF)$ for some hypergraph $\cF$, then clearly there is a
-  bijection $\cl: \cF \rightarrow \cP$ such that $\cl(S) = \cl'(v_S)$
-  where $v_S$ is the vertex corresponding to set $S$ in $\bI(\cF)$ for
-  any $S \in \cF$. This bijection $\cl$ is called the {\em path
-    labeling} of the hypergraph $\cF$ and the path system $\cP$ may be
-  alternatively denoted as $\cF^\cl$.
-
-%  \noindent
-  A path labeling $(\cF, \cl)$ is defined to be {\em feasible} if $\cF
-  \cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
-  \rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
-  \rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.  \add{In this
-    work}, we are given as input $\cF$ and a tree $T$, and the
-  question is whether there is a path labeling $\cl$ to a set of paths
-  in $T$.  We refer to such a solution path system by $\cF^\cl$.  A
-  path labeling $(\cF, \cl)$ is defined to be {\em feasible} if
-  % $\cF \cong \cF^\cl$ and this
-  there is a hypergraph isomorphism $\phi: supp(\cF) \rightarrow
-  supp(\cF^\cl)=V(T)$ induces a path labeling $\cl_\phi: \cF
-  \rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.
-%\end{definition}
-
 \subsection{Overlap Graphs and Marginal Hyperedges}
-%\begin{definition}[Overlap Graphs and Marginal Hyperedges]
-  An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is a graph
-  such that its vertex set has a bijection to $\cF$ and there exists
-  an edge between two of its vertices \iff their corresponding
-  hyperedges overlap. Two hyperedges $S$ and $S'$ are said to {\em
-    overlap}, denoted by $S \overlap S'$, if they have a non-empty
-  intersection and neither is contained in the other; $S \overlap S'
-  \text{ \iff } S \cap S' \ne \emptyset, S \nsubseteq S', S'
-  \nsubseteq S$. Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$
-  and not necessarily connected. Each connected component of
-  $\bO(\cF)$ is called an {\em overlap component}.
-
-%  \noindent
+\label{sec:overlapmarginal}
+Recall the notion of {\em overlap} from
+Definition~\ref{def:setsystem}.  Two hyperedges $S$ and $S'$ are said
+to overlap, denoted by $S \overlap S'$, if they have a non-empty
+intersection and neither is contained in the other.
+
+An {\em overlap graph}  $\bO(\cF)$ of a hypergraph $\cF$ is a graph
+such that its vertex set has a bijection to $\cF$ and there exists an
+edge between two of its vertices \iff their corresponding hyperedges
+overlap.  Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$ and not
+necessarily connected. Each connected component of $\bO(\cF)$ is
+called an {\em overlap component}.
+
 A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
 \overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
-  \cite{kklv10}. Additionally, if $S$ is such that it is contained in
-  no other marginal hyperedge in $\cF$, then it is called {\em
-    super-marginal}.
-%\end{definition}
+\cite{kklv10} (see Definition\ref{def:poset}). Additionally, if $S$ is
+such that it is contained in no other marginal hyperedge in $\cF$,
+then it is called {\em super-marginal}.
 
 \subsection{Miscellaneous}
-%\begin{definition}[$k$-subdivided star] 
-  A {\em star} graph is a
-  complete bipartite graph $K_{1,p}$ which is clearly a tree and $p$
-  is the number of leaves. The vertex with maximum degree is called
-  the {\em center} of the star and the edges are called {\em rays} of
-  the star. A {\em $k$-subdivided star} is a star with all its rays
-  subdivided exactly $k$ times. The definition of a {\em ray of a
-    $k$-subdivided star} is extended to the path from the center to a
-  leaf. It is clear that all rays are of length $k+2$.
-%\end{definition}
+\label{sec:miscprelim}
+A {\em star} graph is a complete bipartite graph $K_{1,\,p}$ (which is
+clearly a tree with $p$ leaves). The vertex with maximum degree in a
+star is called its {\em center} and the edges are called {\em
+  rays}. This thesis is interested in a star-like graph which is
+defined in Definition~\ref{def:ksubstar}
+
+\begin{definition}[$k$-subdivided star]
+  \label{def:ksubstar}
+  A {\em $k$-subdivided star} is a star with all its rays subdivided
+  exactly $k$ times. A {\em ray} of a $k$-subdivided star is the path
+  from the center vertex to a leaf. It is clear that all rays of a
+  $k$-subdivided star are of length $k+2$.
+\end{definition}
 
@@
 
-%\begin{definition}[Partially ordered sets]
-
-  Let $X$ be a partially ordered set with $\preccurlyeq$ being the
-  partial order on $X$.  $X_m \in X$ is a maximal upper bound of $X$
-  if $\nexists X_q \in X$ such that $X_m \preccurlyeq X_q$.  A maximal
-  upper bound on $X$ is denoted by $mub(X)$.
-  %represents an element in $X$ which is a maximal upper bound on $X$.
-%\end{definition}
-
+\begin{definition}[In-tree]
 An {\em in-tree} is a directed rooted tree in which all edges are
 directed toward to the root.  
-
-The set $I$ represents the index set $[m]$. If index $i$ is used
-without further qualification, it is intended that $i \in I$.
+\end{definition}
+In-trees are used in Section~\ref{sec:decompos} which describes our
+solution to \CFTPL problem.
+
+Finally, note that as a convention in this thesis, the set $I$
+represents the index set $[m]$ or $[n]$. If index $i$ is used without further
+qualification, it is intended that $i \in I$.
 
@@
 
-\tnote{\sc refer to Problem Defs here}
-In this section we give an algorithmic characterization of the
-feasibility of a tree path labeling.  
-We focus on the question of generalizing the notion of an
-  ICPIA cite-nsnrs09 to characterize feasible path assignments.  We
-  show that for a given set system $\cF$, a tree $T$, and an
-  assignment of paths from $T$ to the sets, there is a feasible
-  bijection between $U$ and $V(T)$ if and only if all intersection
-  cardinalities among any three sets (not necessarily distinct) is
-  same as the intersection cardinality of the paths assigned to them
-  and the input runs a filtering algorithm (described in this paper)
-  successfully.  This characterization is proved constructively and it
-  gives a natural data structure that stores all the relevant feasible
-  bijections between $U$ and $V(T)$.  Further, the filtering algorithm
-  is also an efficient algorithm to test if a tree path labeling to
-  the set system is feasible.  This generalizes the result in
-  cite-nsnrs09.
-
-Consider a path labeling $(\cF,
-\cl)$ on the given tree $T$. We call $(\cF, \cl)$ an {\em Intersection
-  Cardinality Preserving Path Labeling (ICPPL)} if it has the
-following properties.
-
-\begin{enumerate}[{(\icpplpr\ }i)]
+In this section we discuss the \FTPL problem introduced in
+Section~\ref{sec:myresearchintro} and give a set cardinality based
+characterization for it.
+
+For this characterization, we generalize the notion of the ICPIA
+\cite{nsnrs09}.  We show that for a given hypergraph $\cF$ with
+universe $U$, a target tree $T$, and path labeling from $T$ to $\cF$, there
+is a feasible bijection between $U$ and $V(T)$ \iff all intersection
+cardinalities among any three sets (not necessarily distinct) is same
+as the intersection cardinality of the paths assigned to them.  This
+characterization is proved constructively and the algorithm outputs a
+feasible bijection between $U$ and $V(T)$.
+
+We now introduce a special kind of path labeling called {\em
+  Intersection Cardinality Preserving Path Labeling (ICPPL)} in
+Definition~\ref{def:icppl}.
+
+\begin{definition}
+  \label{def:icppl}
+  Consider a path labeling $(\cF, \cl)$ on the given tree $T$. We call
+  $(\cF, \cl)$ an {\em Intersection Cardinality Preserving Path
+    Labeling (ICPPL)} if it has the following properties.
+  \par
+  \parbox{\linewidth}{%
+    \hangindent \defindent%
+    \begin{enumerate}[i. ] %{(\icpplpr\ }
+      \singlespacing
     \item \label{pr:i} $|S| = |\cl(S)|$ for all $S \in \cF$
-  % \vspace{\topshrink}
-\item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ for all
-  distinct $S_1, S_2 \in \cF$
-  % \vspace{\topshrink}
-\item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
-  \cap \cl(S_3)|$ for all distinct $S_1, S_2, S_3 \in \cF$
+    \item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ for
+    all distinct $S_1, S_2 \in \cF$
+    \item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap
+    \cl(S_2) \cap \cl(S_3)|$ for all distinct $S_1, S_2, S_3 \in \cF$
     \end{enumerate}
-
-The following lemma is useful in subsequent arguments.
-\begin{lemma}
-  \label{lem:setminuscard}
-  If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
-  (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
-  \cl(S_3))|$.
-\end{lemma}
-\begin{proof}\thesisspacing%[Proof of Lemma~\ref{lem:setminuscard}]
-  Let $P_i = \cl(S_i)$, for all $1 \le i \le 3$.  $|S_1 \cap (S_2
-  \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| = |S_1 \cap S_2| -
-  |S_1 \cap S_2 \cap S_3|$. Due to properties (ii) and (iii) of ICPPL,
-  $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1 \cap P_2| - |P_1
-  \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3| = |P_1 \cap (P_2
-  \setminus P_3)|$. %\qed
-\end{proof}
+  }
+\end{definition}
 
 In the remaining part of this section we show that $(\cF, \cl)$ is
-feasible if and only if it is an ICPPL and
-Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty path
-hypergraph isomorphism
-function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two
-levels of filtering of $(\cF, \cl)$ to make it simpler in terms of set
-intersections while retaining the set of isomorphisms, if any, between
-$\cF$ and $\cF^\cl$.
+feasible if and only if it is an ICPPL%  and
+% Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty path
+% hypergraph isomorphism
+% function
+.  Algorithm~\ref{al:icppl-find-isomorph} takes $(\cF, \cl)$ as input
+and computes the feasibility bijection if it is feasible or else
+reports failure. This algorithm recursively does two levels of
+filtering of $(\cF, \cl)$ to make it simpler in terms of set
+intersections while retaining a non-empty subset of isomorphisms, if
+any, between $\cF$ and $\cF^\cl$. These two filters are {\tt
+  filter\_1} in Algorithm~\ref{al:filterone} and {\tt filter\_2} in
+Algorithm~\ref{al:filtertwo}.
 % One direction of this claim isclear: that if a path labeling is
@@
 % to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.
+
 First, we present Algorithm~\ref{al:filterone} which describes {\tt
@@
 
-    \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of
-      ICPPL}
-    \label{ln:3waycheck}
-    \STATE {\bf exit} \label{ln:exit1} \\
-    \ENDIF
+%     \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of
+%       ICPPL}
+%     \label{ln:3waycheck}
+%     \STATE {\bf exit} \label{ln:exit1} \\
+%     \ENDIF
 
@@
 
+The following lemma is useful in subsequent arguments.
+\begin{lemma}
+  \label{lem:setminuscard}
+  If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
+  (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
+  \cl(S_3))|$.
+\end{lemma}
+\begin{proof}\thesisspacing%[Proof of Lemma~\ref{lem:setminuscard}]
+  Let $P_i = \cl(S_i)$, for all $1 \le i \le 3$.  $|S_1 \cap (S_2
+  \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| = |S_1 \cap S_2| -
+  |S_1 \cap S_2 \cap S_3|$. Due to properties (ii) and (iii) of ICPPL,
+  $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1 \cap P_2| - |P_1
+  \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3| = |P_1 \cap (P_2
+  \setminus P_3)|$. %\qed
+\end{proof}
+
+Now we will prove the effectiveness of Algorithm~\ref{al:filterone}.
+Lemma~\ref{lem:feasible} proves that {\tt filter\_1} does not destroy
+the feasibility of the input, if it is indeed feasible. Following
+this, Lemma~\ref{lem:invar1} and Lemma~\ref{lem:noexit1} show that if
+the input is an ICPPL the result of {\tt filter\_1} is also an ICPPL
+and they both have the same set of feasibility hypergraph
+isomorphisms.
+
 \begin{lemma}
@@
   In Algorithm~\ref{al:filterone}, if input $(\cF, \cl)$ is a feasible path
-  assignment then at the end of $j$th iteration of the {\bf while}
-  loop, $j \ge 0$, $(\cF_j, \cl_j)$ is a feasible path assignment.
+  labeling then at the end of $j$th iteration of the {\bf while}
+  loop, $j \ge 0$, $(\cF_j, \cl_j)$ is a feasible path labeling.
 \end{lemma}
@@
   the input itself which is given to be feasible. Assume the lemma is
-  true till $j-1$th iteration. \ie every hypergraph isomorphism $\phi:
+  true till $j-1$th iteration \ie every hypergraph isomorphism $\phi:
   supp\left(\cF_{j-1}\right) \rightarrow V\left(T \right)$ that
   defines $(\cF, \cl)$'s feasibility, is such that the induced path
-  labeling on $\cF_{j-1}$, say denoted by $\cl_{\phi[{\cF_{j-1}}]}$, is equal to
-  $\cl_{j-1}$. We will prove that $\phi$ is also the bijection that
-  makes $(\cF_j, \cl_j)$ feasible. Note that $supp(\cF_{j-1}) =
-  supp(\cF_{j})$ since the new sets in $\cF_j$ are created from basic
-  set operations to the sets in $\cF_{j-1}$ adding or removing no
-  elements. For the same reason and $\phi$ being a bijection, it is
-  clear that when applying the $\phi$-induced path labeling on
-  $\cF_j$, $ \cl_{\phi[{\cF_{j}}]}(S_1 \setminus S_2) =
-  \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
-  \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Now observe that $ \cl_j(S_1
+  labeling on $\cF_{j-1}$, say denoted by $\cp_{\phi[{\cF_{j-1}}]}$,
+  is equal to $\cl_{j-1}$.
+% Note that any hypergraph isomorphism of
+% $\cF$ (say, one that defines the feasibility of $(\cF,\cl)$) can be
+% ``projected'' to $\cF_{j-1}$ since their supports are equal. If $\phi$
+% is a hypergraph isomorphism of to $(\cF, \cl)$, then 
+  We will prove that $\phi$ is also the bijection that makes $(\cF_j,
+  \cl_j)$ feasible. Note that $supp(\cF_{j-1}) = supp(\cF_{j})$ since
+  the new sets in $\cF_j$ are created from basic set operations to the
+  sets in $\cF_{j-1}$ adding or removing no elements. For the same
+  reason and $\phi$ being a bijection, it is clear that when applying
+  the $\phi$-induced path labeling on $\cF_j$, $
+  \cp_{\phi[{\cF_{j}}]}(S_1 \setminus S_2) =
+  \cp_{\phi[{\cF_{j-1}}]}(S_1) \setminus
+  \cp_{\phi[{\cF_{j-1}}]}(S_2)$. Now observe that $ \cl_j(S_1
   \setminus S_2) = \cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2) =
-  \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
-  \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Thus the induced path labeling
-  $\cl_{\phi[{\cF_{j}}]} = \cl_{j}$.  %\qed
+  \cp_{\phi[{\cF_{j-1}}]}(S_1) \setminus
+  \cp_{\phi[{\cF_{j-1}}]}(S_2)$. Thus the induced path labeling
+  $\cp_{\phi[{\cF_{j}}]} = \cl_{j}$.  %\qed
 \end{proof}
@@
 \begin{lemma}
-  \label{lem:invar1} In Algorithm~\ref{al:filterone}, at the end of $j$th
-  iteration, $j \ge 0$, of the {\bf while} loop, the following
-  invariants are maintained.
+  \label{lem:invar1} Given that the input to Algorithm~\ref{al:filterone} is
+  an ICPPL, at the end of $j$th iteration, $j \ge 0$, of the {\bf
+    while} loop, the following invariants are maintained.
   \begin{enumerate}[I {\ }] %\vspace{\topshrink}
+    \singlespacing
   \item $\cl_j(R)$ is a path in $T$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
@@
 \end{lemma}
-
 \begin{proof}\thesisspacing
@@
   in $\cl_j$ where $\cl_{j-1}(S_1)$
-  and $\cl_{j-1}(S_2)$ intersect and share a leaf.\\
+  and $\cl_{j-1}(S_2)$ intersect and share a leaf.
   The invariants are true in the base case $(\cF_0, \cl_0)$, since it
@@
 
-    Due to the condition in line~\ref{ln:3waycheck}, this invariant is
-    ensured at the end of every iteration.
+    \tnote[dire]{Due to the condition in line~\ref{ln:3waycheck}, this invariant is
+    ensured at the end of every iteration.}
     % \begin{enumerate} [{Case 3.}1:]
@@
   defines $(\cF, \cl)$'s feasibility is the same as the set that
-  defines $(\cF_j, \cl_j)$'s feasibility, if any.  Secondly, for any
-  iteration $j > 0$ of the {\em \bf while} loop, the {\em \bf exit}
-  statement in line~\ref{ln:exit1} will not execute.
+  defines $(\cF_j, \cl_j)$'s feasibility, if any.%   Secondly, for any
+%   iteration $j > 0$ of the {\em \bf while} loop, the {\em \bf exit}
+%   statement in line~\ref{ln:exit1} will not execute.
 \end{lemma}
@@
   % \end{enumerate}
-  In other words, filter 1 outputs a filtered path labeling that
+  In other words, {\tt filter\_1} outputs a filtered path labeling that
   ``preserves''
-  hypergraph isomorphisms of the original path labeling.\\
-  Secondly, line~\ref{ln:exit1} will execute \iff the exit condition in
-  line~\ref{ln:3waycheck}, i.e. failure of three way intersection
-  preservation, becomes true in any iteration of the {\em \bf while}
-  loop.  Due to Lemma~\ref{lem:invar1} Invariant IV, the exit
-  condition does not occur if the input is a feasible ICPPL.%\qed
-
+  hypergraph isomorphisms of the original path labeling.
+%   Secondly, line~\ref{ln:exit1} will execute \iff the exit condition in
+%   line~\ref{ln:3waycheck}, i.e. failure of three way intersection
+%   preservation, becomes true in any iteration of the {\em \bf while}
+%   loop.  Due to Lemma~\ref{lem:invar1} Invariant IV, the exit
+%   condition does not occur if the input is a feasible ICPPL.%\qed
   % such that $\phi(x) = v$ where $v$ is the leaf considered in the
@@
   % iteration of the loop.
-
   % \begin{enumerate}[\hspace{2mm}i. ] \vspace{\topshrink}
@@
   % \end{enumerate}
-
   % Thus the exit condition is never rendered true after $x$ and $v$
@@
 
-As a result of Algorithm~\ref{al:filterone} each leaf $v$ in $T$ is such that
-there is exactly one set in $\cF$ with $v$ as a vertex in the path
-assigned to it.  In Algorithm~\ref{al:filtertwo} we identify elements in
-$supp(\cF)$ whose images are leaves in a hypergraph isomorphism if one
-exists.  Let $S \in \cF$ be such that $\cl(S)$ is a path with leaf and
-$v \in V(T)$ is the unique leaf incident on it.  We define a new path
-labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) = \set{v}$ where
-$x$ an arbitrary element from $S \setminus \bigcup_{\hS \ne S}
-\hS$. In other words, $x$ is an element present in no other set in
-$\cF$ except $S$. This is intuitive since $v$ is present in no other
-path image under $\cl$ other than $\cl(S)$.  The element $x$ and leaf
-$v$ are then removed from the set $S$ and path $\cl(S)$
+As a result of {\tt filter\_1} each leaf $v$ in $T$ is such that there
+is exactly one set in $\cF$ with $v$ as a vertex in the path assigned
+to it.  {\tt filter\_2}, in Algorithm~\ref{al:filtertwo}, identifies
+elements in $supp(\cF)$ whose images are leaves in a hypergraph
+isomorphism if one exists.  Let $S \in \cF$ be such that $\cl(S)$ is a
+path with leaf and $v \in V(T)$ is the unique leaf incident on it.  We
+define a new path labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) =
+\set{v}$ where $x$ an arbitrary element from $S \setminus \bigcup_{\hS
+  \ne S} \hS$. In other words, $x$ is an element present in no other
+set in $\cF$ except $S$. This is intuitive since $v$ is present in no
+other path image under $\cl$ other than $\cl(S)$.  The element $x$ and
+leaf $v$ are then removed from the set $S$ and path $\cl(S)$
 respectively. After doing this for all leaves in $T$, all path images
@@
     \STATE $X \assign S_1 \setminus \bigcup_{S \in \cF_{j-1}, S \ne S_1}S$\\
-    \IF{$X$ is empty} \label{xempty} \STATE {\bf
-      exit} \label{ln:exit2} \ENDIF
-    \STATE $x \assign $ arbitrary element from $X$\\
+%     \IF{$X$ is empty} \label{xempty} \STATE {\bf
+%       exit} \label{ln:exit2} \ENDIF
+    \STATE \label{xassign} $x \assign $ arbitrary element from $X$\\
     \STATE $\cF_j \assign \cF_j \cup \{\{x\}, S_1 \setminus \{x\}\} $\\
@@
 
-Suppose the input ICPPL $(\cF, \cl)$ is feasible, yet set $X$ in
-Algorithm~\ref{al:filtertwo} is empty in some iteration of the {\bf while}
-loop. This will abort our procedure of finding the hypergraph
-isomorphism. The following lemma shows that this will not happen.
+Suppose the input $(\cF, \cl)$ is feasible, yet set $X$ in
+Algorithm~\ref{al:filtertwo} is empty in some iteration of the {\bf
+  while} loop. Then clearly, the algorithm will have a problem in line~\ref{xassign}.
+% This will abort our procedure of finding the hypergraph
+% isomorphism. 
+The following Lemma~\ref{lem:xnotempty} shows that this cannot
+happen. Following that Lemma~\ref{lem:invar3} shows that given an
+ICPPL input, the filtered output path labeling after {\tt filter\_2}
+is also an ICPPL.
 
@@
   feasible, then for all iterations $j > 0$ of the {\em \bf while}
-  loop, the {\em \bf exit} statement in line~\ref{ln:exit2} does not
-  execute.
+  loop, the set $X$ will be non-empty.
+% the {\em \bf exit} statement in line~\ref{ln:exit2} does not
+%   execute.
 \end{lemma}
@@
 target tree $T$. We also proved that if the input is indeed feasible,
-these algorithms executed in sequence give an ICPPL that preserves \change{the
-  same set}{a subset} of hypergraph isomorphisms as that of the input
-ICPPL. Now we present the algorithmic characterization of a feasible
-tree path labeling by way of Algorithm~\ref{al:icppl-find-isomorph}.
-
-\temptext{Prove that the same set of hypergraph isomorphisms are in
-  the filtered ICPPL? -- NOPE. we are only interested in testing
-  feasibility. not necessarily giving ALL bijections.}
-\begin{lemma}{\temptext{PROOF TBD.}}
-  Let $(\cF_2, \cl_2)$ be the output of {\tt filter\_2} and the ICPPL
-  input to {\tt filter\_1} preceeding its call is $(\cF, \cl)$ with
-  target tree $T$. If a bijection $\phi:\supt{\cF_2} \rightarrow V(T)$
-  is a hypergraph isomorphism of $(\cF_2, \cl_2)$, then it is a
-  hypergraph isomorphism of $(\cF, \cl)$.
-\end{lemma}
+these algorithms executed in sequence give an ICPPL that preserves a
+non-empty subset of set of feasibility hypergraph isomorphisms of the
+original input. % Now we present the algorithmic characterization of a feasible
+% tree path labeling by way of Algorithm~\ref{al:icppl-find-isomorph}.
+Now we present Algorithm~\ref{al:icppl-find-isomorph} which uses these
+two filters to compute one such isomorphism.
+
+%% IS THIS LEMMA NECESSARY? 31 JAN 2012 %%
+% \begin{lemma}{\temptext{PROOF TBD.}}
+%   Let $(\cF_2, \cl_2)$ be the output of {\tt filter\_2} and the ICPPL
+%   input to {\tt filter\_1} preceeding its call is $(\cF, \cl)$ with
+%   target tree $T$. If a bijection $\phi:\supt{\cF_2} \rightarrow V(T)$
+%   is a hypergraph isomorphism of $(\cF_2, \cl_2)$, then it is a
+%   hypergraph isomorphism of $(\cF, \cl)$.
+% \end{lemma}
 
@@
 the leaf labels $\phi'$ computed earlier gives us a hypergraph
-isomorphism $\phi$ for the input $\cF,\cl,T$.  The following lemma formalizes the
-characterization of feasible path labeling.
+isomorphism $\phi$ for the input $\cF,\cl,T$.  
+Lemma~\ref{lem:hyperiso} proves the correctness of this computation.
 
@@
   \label{lem:hyperiso} %{lem:perm}
+%   If $(\cF, \cl)$ is an ICPPL from a tree $T$ and
+%   Algorithm~\ref{al:icppl-find-isomorph}, on input $\cF, \cl, T$
+%   returns a non-empty function, then there exists a hypergraph
+%   isomorphism $\phi : supp(\cF) \rightarrow V(T)$ such that the
+%   $\phi$-induced tree path labeling is equal to $\cl$ or $\cl_\phi =
+%   \cl$.
   If $(\cF, \cl)$ is an ICPPL from a tree $T$ and
-  Algorithm~\ref{al:icppl-find-isomorph}, on input $\cF, \cl, T$
-  returns a non-empty function, then there exists a hypergraph
+  Algorithm~\ref{al:icppl-find-isomorph}, on input $(\cF, \cl, T)$
+  returns a non-empty function $\phi$, then $\phi$ is a hypergraph
   isomorphism $\phi : supp(\cF) \rightarrow V(T)$ such that the
-  $\phi$-induced tree path labeling is equal to $\cl$ or $\cl_\phi =
+  $\phi$-induced tree path labeling is equal to $\cl$; $\cp_\phi =
   \cl$.
@@
 
+Finally we have Theorem~\ref{th:charac} which puts together all that
+we saw so far in this section to prove that ICPPL is indeed a
+characterization of a feasible path labeling.
 \begin{theorem}
@@
 \end{proof}
+
+\section{Solution to \illustrationproblem}
+
+\tnote[dire]{ inthis section we run through the algorithm in the
+  problem introduced in chapter 1.}
 
@@
   \begin{tabular}{lr}
-    (a) \includegraphics{cop-tree_thesis.1}&
-    (b) \includegraphics{cop-tree_thesis.2}
+%     (a) \includegraphics{cop-tree_thesis.1}&
+%     (b) \includegraphics{cop-tree_thesis.2}
+    (a) \ksubstari & (b) \ksubstarii
   \end{tabular}
@@
 \subsection{Description of the Algorithm}
+\label{sec:ksubstaralgo}
 In this section the given hypergraph $\cF$, the $k$-subdivided star
@@
   inclusion chain which contradicts that assumption $X$ is
-  super-marginal. Thus the claim is proved.\qed
+  super-marginal. Thus the claim is proved.
 \end{proof}
@@
   contradicts the assumption that $\cO$ does not have super-marginal
-  hyperedges. \qed
+  hyperedges. 
 \end{proof}
@@
   again by induction on $k$ it follows that the algorithm finds $\cl$.
-  \qed
+
 \end{proof}
@@
 Section \ref{feasible} we present our characterization of feasible
-tree path assignments, and in Section \ref{decompos} we present the
+tree path assignments, and in Section \ref{sec:decompos} we present the
 characterizing subproblems for finding a bijection between $U$ and
 $V(T)$ such that sets map to tree paths. Finally, in Section
-\ref{complexity} we conclude by showing that Tree Path Assignment is
+\ref{sec:tplcomplexity} we conclude by showing that Tree Path Assignment is
 GI-Complete, and also observe that Consecutive Ones Testing is in
@@
 \subsection{Finding an assignment of tree paths to a set
-  system} \label{decompos} In the previous section we have shown that
+  system}% 
+\label{sec:decompos}% 
+In the previous section we have shown that
 the problem of finding a Tree Path Asssignment to an input $(\cF,T)$
@@
 \section[Complexity]{Complexity of Tree Path Assignment-A Discussion}
-\label{complexity}
+\label{sec:tplcomplexity}
 Recall that the input to the Tree Path Assignment question is an order
@@
 \subsection{Consecutive Ones Testing is in Logspace}
+\label{sec:cotlogspace}
 While Tree Path Assignment is isomorphism-complete, it is polynomial

### 2012-02-01 src/010chap1.tex
- completed chapter 1. am. not. working. on. this. anymore. -- binary matrix example has been transposed to show cop on rows. - added section with illustration problem worked out in chapter 3. this shit is done.

@@
 
-\tnote[dire]{
 \section[Basic preliminaries]{Basic preliminaries} 
@@
 
+Here we see some basic definitions and conventions necessary for the contents
+of this thesis.
+
+\subsection{Matrices}
 If $n$ is a positive integer, $[n]$ denotes the set $\{1,2,... n\}$.
@@
   Let $M$ be an $n \times m$ matrix. $m_{i,j}$ denotes its $(i,j)$th
-  element, i.e. element at $i$th row and $j$th column. $M$ is a \textbf{binary
-  matrix} if each of its element is 0 or \un.  In other words, for all
+  element, i.e. element at $i$th row and $j$th column. $M$ is a
+  \textbf{binary matrix} if each of its element is 0 or \un; for all
   $i \in [n]$ and $j \in [m]$, $m_{i,j} \in \{0, \un\}$
@@
   A \textbf{permutation} $\lambda$ of a set $X = \{x_1, x_2, \ldots
-  x_n\}$ is a bijection $\lambda: X \rightarrow X$.  For simplicity,
-  $\lambda$ may be written as a sequence $x_{i_1}x_{i_2}\ldots x_{i_n}$,
-  where $i_j \in [n]$ to mean that $\lambda(x_j)= x_{i_j}$.
-  This document uses both notations as convenient in context.
+  x_n\}$ is a bijection $\lambda: X \rightarrow X$.  $\lambda$ may be
+  written as a sequence $x_{i_1}x_{i_2}\ldots x_{i_n}$, where $i_j \in
+  [n]$ to mean that $\lambda(x_j)= x_{i_j}$.  This document uses both
+  notations as convenient in context.
 \end{definition}
 
-\begin{definition}[\Cop]
+The idea of \cop of binary matrices was mentioned a few times in
+earlier sections. Now we will see the formal definition of \COP in
+Definition~\ref{{def:copmatrix}}.
+
+\begin{definition}[\Cop (on columns)]
   \label{def:copmatrix}
   Let $M$ be a binary matrix.
-  \begin{itemize}
-  \item \textbf{\Cop}
-  \item \textbf{Strong \cop}
-  \item \textbf{\Cop order} or \cop permutation
-  \end{itemize}
+  \begin{enumerate}
+  \hangindent \defindent
+  \item A \textbf{block of \un s (block of 0s)} in a column of $M$ is a maximal set of
+    consecutive \un-entries (0-entries) in this column.
+  \item $M$ has the \textbf{strong \cop
+      (strong \COP)} if in every column the \un s appear consecutively,
+    \ie if every column contains at most one block of \un s.
+  \item $M$ has the \textbf{\cop} if its rows can be
+    permuted in such a way that the resulting matrix has the strong \COP.
+  \item If an ordering for the rows of $M$ gives the strong \COP, it
+    is called a \textbf{\COP order} or \textbf{\COP permutation}.
+  \end{enumerate}
 \end{definition}
+
+When the roles of rows and columns are exchanged in
+Definition~\ref{def:copmatrix}, it defines \COP on rows. Both
+are equivalent properties (for the purposes of this thesis) and both
+conventions are seen the literature.
+Figure~\ref{fig:cop-matrix} gives an example of \COP on columns.
 
@@
 
-\begin{definition}[\Crop]
+A less restrictive property of binary matrices that is a
+generalization of the \COP is \crop (\CROP). The matrix can be
+visualized to be wrapped around a horizontal cylinder and demands that
+after some row permutations, if required, in every column the \un s
+appear consecutively on the cylinder -- note that this implies the 0s
+also appear consecutively.
+
+\begin{definition}[\Crop (on columns)]
   \label{def:cropmatrix}
+  Let $M$ be a binary matrix.
+  \begin{enumerate}
+  \hangindent \defindent
+  \item $M$ has the \textbf{strong \crop (strong \CROP)} if in every
+    column the \un s appear consecutively or the 0s appear
+    consecutively or both.
+  \item $M$ has the \textbf{\crop (\CROP)} if its rows can be permuted
+    in such a way that the resulting matrix has the strong \CROP.
+  \item If an ordering for the rows of $M$ gives the strong \CROP,
+    then it is called the \textbf{\CROP ordering} or \textbf{\CROP
+      permutation}.
+  \end{enumerate}
 \end{definition}
+
+\subsection{Sets}
 
@@
   \label{def:setsystem}
-
   Let $U$ be a universe with $|U| = n$.  
-  \begin{itemize}
+  \begin{enumerate}
+  \hangindent \defindent
   \item The set $\F \subseteq (2^{U} \setminus \emptyset)$ is called a
@@
     \nsubseteq S', S' \nsubseteq S\]
-  \end{itemize}
+  \end{enumerate}
 \end{definition}
 
-\begin{definition}[Partial order]
+\begin{definition}
   \label{def:poset}    
-  Let $A$ be a set and $R$ be a binary relation $R \subseteq A \times
-  A$.
+%   Let $A$ be a set and $R$ be a binary relation $R \subseteq A \times
+%   A$.
 
-  \textbf{poset}
-
+%   \textbf{poset}
   Let $X$ be a partially ordered set with $\preccurlyeq$ being the
@@
 
-  \begin{itemize}
-  \item \textbf{hasse diagram}
-  \item \textbf{maximal upper bound}  $X_m \in X$ is a \textbf{maximal upper bound} of $X$
-  if $\nexists X_q \in X$ such that $X_m \preccurlyeq X_q$.  A maximal
-  upper bound on $X$ is denoted by $mub(X)$.
+  \begin{enumerate}
+  \hangindent \defindent
+%  \item \textbf{hasse diagram}
+  \item An element $X_m \in X$ is a
+    \textbf{maximal upper bound} of $X$ if $\nexists X_q \in X$ such
+    that $X_m \preccurlyeq X_q$.  A maximal upper bound on $X$ is
+    denoted by $mub(X)$.
 
@@
     forms a single chain.
-  \end{itemize}
+  \end{enumerate}
 \end{definition}
 
-\begin{definition}[Graph, Tree]
+\subsection{Graphs}
+
+\begin{definition}[Graph, Tree, Path]
   \label{def:graphtree}
-  \tnote[dire]{take from doms}
+  \begin{enumerate}
+    \hangindent \defindent
+  \item An \textbf{(undirected) graph} is $G = (V,E)$ is \stt $V$ is a
+    finite set and $E$ is a set of ordered pairs, $E \subseteq V
+    \times V$. An element in $V$ is called a \textbf{vertex
+      (\emph{pl.} vertices)} and an element from $E$ is called an
+    \textbf{edge}. $V$ and $E$ may be written as $V(G)$ and $E(G)$
+    respectively. If $u,v \in V$ and $(u,v) \in E$ then $v$ is said to
+    be \textbf{adjacent} to $u$ and vice versa. A \textbf{subgraph} of
+    $G$ is a graph $G' = (V',E')$ \stt $V' \subseteq V$ and $E'
+    \subseteq E$.
+
+  \item A \textbf{path} is a graph $P = (V,E)$ with vertex set $V =
+    \{v_1,\ldots,v_n\}$ and edge set $E= \{ (v_1,v_2),(v_2,v_3),
+    \ldots,(v_{n-2},v_{n-1}),(v_{n-1},v_{n})\}$. The vertices $v_1,
+    v_n$ are called the \textbf{endpoints} of $P$. The graph $P$ is
+    called a \textbf{cycle} if it has an additional edge $(v_n,v_1)$.
+
+  \item A \textbf{tree} is a graph $T$ that has no subgraph that is a
+    cycle.
+  \end{enumerate}
 \end{definition}
 
-When refering to a tree as $T$ it could be a reference to the tree
-itself, or the vertices of the tree. This will be clear from the
-context.
+% When refering to a tree as $T$ it could be a reference to the tree
+% itself, or the vertices of the tree. This will be clear from the
+% context.
 
@@
   \label{def:pathgraph}
-
-  \begin{itemize}
-  \item A graph $G$ is an \textbf{interval graph} if there a set of intervals
+  Let $G$ be a graph.
+  \begin{enumerate}
+  \hangindent \defindent
+  \item $G$ is an \textbf{interval graph} if there a set of intervals
     $\cI$ \stt $G$ is isomorphic to the intersection graph of
     $\cI$. \[G \cong \bI(\cI)\]
-  \item A graph $G$ is a \textbf{path graph} if there exists a tree $T$ and a
+  \item $G$ is a \textbf{path graph} if there exists a tree $T$ and a
     set of paths from $T$, $\cP$ \stt $G$ is isomorphic to the
     intersection graph of $\cP$. \[G \cong \bI(\cP)\]
-  \end{itemize}
+  \item $G$ is a \textbf{chordal graph} if it has no induced cycle of
+    length greater than 3. Moreover, a \textbf{chordal graph} is
+    characterized as an intersection graph of subtrees of a tree.
+  \end{enumerate}
 \end{definition}
-Moreover, a \textbf{chordal graph} is characterized as an intersection
-graph of subtrees of a tree.
-
-\begin{definition}[Maximal clique]
-  \label{def:maxclique}    
-\end{definition}
-}
 

### 2012-02-01 src/030chap3.tex
- completed chapter 1. am. not. working. on. this. anymore. -- binary matrix example has been transposed to show cop on rows. - added section with illustration problem worked out in chapter 3. this shit is done.

@@
 
-\tnote[dire]{ inthis section we run through the algorithm in the
-  problem introduced in chapter 1.}
-
+In this section we run through the filtering and tree pruning
+algorithm seen in Section~\ref{sec:feasible} on an example. The
+example problem is the \illustrationproblem posed in
+Section~\ref{sec:problem}.  % Chapter~\ref{ch:intro}
+The solution is depicted in Figure~\ref{fig:solpartI},
+Figure~\ref{fig:solpartIII}, Figure~\ref{fig:solpartIVa},
+Figure~\ref{fig:solpartIVb}, Figure~\ref{fig:solpartVI},
+Figure~\ref{fig:solpartVII}, Figure~\ref{fig:solpartVIII} and the
+figures along with their captions are self-explanatory.
+
+\begin{figure}[htb]
+  \centering
+  \begin{tabular}[h]{c|cc}
+    $T:$ \infiniteloopI & $\cF:$ \studygroupsI
+  \end{tabular}
+
+  \caption[\figtabsize Problem solution part 1]{\figtabsize The
+    problem is shown again here using the input described in
+    Section~\ref{sec:problem}. The target tree is $T$ with vertex set
+    $\{1,2,\ldots,11\}$ on the left hand side along with the path
+    system (hypergraph) -- $\{7,2,6,5\}, \{8,2,4\}, \{10,6,5,3\},
+    \{9,1,5,3,11\}$. The hypergraph is $\cF$ with hyperedges $\{\xPa,
+    \xPi, \xVi, \xCh\}, \{\xSn, \xWo, \xPi\}, \{\xCh, \xFr, \xVi,
+    \xLi\}, \{\xLu, \xSa, \xSc, \xCh, \xFr\}$ shown in the right hand
+    side. In subsequent illustrations, the gray boxes which contain
+    the elements of universe $U$, will be assigned to nodes and
+    removed from the right hand side figure and placed next to the
+    vertex of $T$ in the left hand side figure.  The path labeling is
+    implicitly given by the color-coded Venn diagrams (the order of
+    listing of the hyperedges in this caption is also done respective
+    to the input path labeling).}
+  \label{fig:solpartI}  
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \begin{tabular}[h]{c|cc}
+    \infiniteloopIII 
+    & 
+    \studygroupsIII 
+    &
+    \studygroupsIIItable
+  \end{tabular}
+  \caption[\figtabsize Problem solution part 2]{\figtabsize It can be observed in
+    Figure~\ref{fig:solpartI}, that none of the leaves are shared by
+    more than one path. Hence, \filteri does not alter the
+    hypergraphs. \filterii finds the leaf assignments as shown --
+    elements \xPa, \xSn, \xLu, \xSc, \xLi, \xWo are assigned to
+    vertices 7, 8, 9, 11, 10, 4 respectively. Computation of these
+    leaf labels are deterministic except for $\xLu, \xSc$ where the
+    algorithm arbritarily choses them from $X = \{\xLu, \xSa, \xSc\}$
+    and for $\xSn, \xWo$ arbitrarily choses from $X = \{\xSn,
+    \xWo\}$. Here we show only one such choice and clearly one could
+    have chosen another leaf label (\eg $\xSa$ on 9 or $\xWo$ on 8) to
+    get a different feasible hypergraph isomorphism.  After
+    this, the leaves of $T$ are pruned and the hypergraph updated
+    accordingly by \algoiii subroutine -- the pruned leaves are shown
+    with dotted edges and the corresponding Venn diagrams show the
+    updated hypergraphs. $\phi_0$ is this leaf assignment and the
+    algorithm proceeds to solve the subproblem with the new
+    hypergraphs and path labeling which is: $\{2,6,5\}, \{2\},
+    \{6,5,3\}, \{1,5,3\}$ labeled with $\{\xPi, \xVi, \xCh\},
+    \{\xPi\}, \{\xCh, \xFr, \xVi\}, \{\xSa, \xCh, \xFr\}$
+    respectively.}
+  \label{fig:solpartIII}  
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \begin{tabular}[h]{c|cc}
+    \infiniteloopIVa
+    & 
+    \studygroupsIVa
+  \end{tabular}
+  \caption[\figtabsize Problem solution part 3]{\figtabsize In the subproblem shown in
+    Figure~\ref{fig:solpartIII}, it can be noted that there are two
+    leaves which are shared by more than one paths -- 2 and 3. Hence
+    \filteri filters this out one by one. Here we show the handling of
+    leaf 2 which was shared by paths $\{2,6,5\}$ ($=P_1$, say) and
+    $\{2\}$($=P_2$, say). \filteri removes $P_1$ and adds
+    $\{6,5\}$. Note that only one set is added and removed since $P_2
+    \subset P_1$. Accordingly, \filteri alters the hypergraph on the
+  right hand side by removing and adding $\{\xPi, \xVi, \xCh\}$ and $\{\xVi, \xCh\}$ respectively.}
+  \label{fig:solpartIVa}  
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \begin{tabular}[h]{c|cc}
+    \infiniteloopIVb
+    & 
+    \studygroupsIVb
+  \end{tabular}
+  \caption[\figtabsize Problem solution part 4]{\figtabsize After handling leaf 2, \filteri proceeds to
+    handle leaf 3 which is shared by paths $\{1,5,3\}$ and
+    $\{6,5,3\}$. \filteri removes these two paths and adds their
+    intersection and set differences -- $\{5,3\}$, $\{1\}$ and
+    $\{6\}$. Correspondingly, hyperedges $\{\xSa, \xCh, \xFr\}$ and $\{\xVi,
+    \xCh, \xFr\}$ are removed and new hyperedges $\{\xCh, \xFr\}$, $\{\xSa\}$
+    and $\{\xVi\}$ respectively are added. There are no more leaves
+    shared and the algorithm proceeds to \filterii.}
+  \label{fig:solpartIVb}  
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \begin{tabular}[h]{c|cc}
+    \infiniteloopVI
+    & 
+    \studygroupsVI
+    &
+    \studygroupsVItable
+  \end{tabular}
+  \caption[\figtabsize Problem solution part 5]{\figtabsize \filterii computes the leaf assignment $\phi_1$ -- $\xPi,
+    \xSa, \xFr$ on 2,1,3 respectively. These assignments are
+    deterministic since $X$ consisted of only single elements for each
+  of these leaves. After this the rest of the labeling is fairly
+  obvious but we present it in Figure~\ref{fig:solpartVII} for the
+  sake of completeness.}
+  \label{fig:solpartVI}  
+\end{figure}
+\clearpage
+
+\begin{figure}[htb]
+  \centering
+  \begin{tabular}[h]{c|cc}
+
+    \infiniteloopVII 
+    &
+    &
+    \studygroupsVIItable
+  \end{tabular}
+  \caption[\figtabsize Problem solution part 6]{\figtabsize The last part of the isomorphism is computed as
+  the leaf label $\phi_2$ which maps $\xCh, \xVi$ to 5, 6
+  respectively. Figure~\ref{fig:solpartVIII} shows the final
+  feasibility isomorphism which is the union of all the leaf labels.}
+  \label{fig:solpartVII}  
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \begin{tabular}[h]{c|cc}
+    \infiniteloopVIII
+    &
+    &
+    \studygroupsVIIItable
+  \end{tabular}
+  \caption[\figtabsize Problem solution part 7]{\figtabsize The solution to the problem presented in
+    Figure~\ref{fig:solpartI}. The feasibility hypergraph isomorphism
+    $\phi$ is computed as the union of the leaf labels
+    $\phi_0,\phi_1,\phi_2$. One can easily verify that this
+    isomorphism does indeed induce the path labeling given in the
+    original input.}
+  \label{fig:solpartVIII}
+\end{figure}
+
+%\clearpage
 \section[Special target trees]{ Computing feasible TPL with special target trees}

### 2012-02-01 src/thesis.tex
- a whole bunch of cleanups. see diff. - removed problem defs from chap 1. they were duplicates from chap 3. - CODED FIGURES IN TIKZ. forbidden subgraphs, graphmatrix example graphs. - BEGAN CHAPTER 3 EDITS.
Formatting/layout changes
-------

- Added FrenchLayout to make lists denser
- Added hanging indent for Definitions
- Compressed theorem env's above below spaces
- Creaded env for problem definition minipages.
- moved all clearpage/page counter and other prechapter settings to the chapter tex files.
- Fixed the page number bug in draft printout. removed \fancyhf{} to avoid resetting page n

@@
 %
-\usepackage[color,notcite,notref]{showkeys}   % To show keys used - useful in
+\usepackage[color,notcite,notref,final]{showkeys}   % To show keys used - useful in
                                % drafts. 
@@
                                 % twoside. keep empty for one side.
-\usepackage[inline] 
+%\usepackage[inline] 
 %\usepackage[margins] 
-%\usepackage[finalnew]
+\usepackage[finalnew]
          {../lib/trackchanges} % For tracking changes.
@@
 % \usepackage{comment}         % DEBUG ONLY
-%\includeonly{010chap1}         % DEBUG ONLY 
+\includeonly{010chap1}         % DEBUG ONLY 
 %\includeonly{020chap2}         % DEBUG ONLY 
+%\includeonly{010chap1, 020chap2}         % DEBUG ONLY 
 %\includeonly{030chap3,050appa}         % DEBUG ONLY 
 %\includeonly{0031ded}         % DEBUG ONLY
+%\includeonly{020chap2}         % DEBUG ONLY 
+%\includeonly{_tempscrap}         % DEBUG ONLY 
+%\figforbiddensubgraphs
 
@@
 %
-
-\xclearpage% \cleardoublepage
-\setcounter{page}{1}
 \include{001certificate}
@@
 %
-\xclearpage
 \include{002ack}
@@
 %
-\xclearpage
 \include{0031ded}
@@
 %
-\xclearpage
 \include{003abstract}
@@
   \xclearpage
-  \listoftables %\addcontentsline{toc}{chapter}{LIST OF TABLES}
+  \listoftables 
   \xclearpage
-  \listoffigures %\addcontentsline{toc}{chapter}{LIST OF FIGURES}
+  \listoffigures
 \end{singlespace}
@@
 %
-\xclearpage
 \include{004abbrv}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Main body Chapters
+%
 \xclearpage
 \pagestyle{fancy}%headings}   % For running headers.
-\fancyhf{}%            % Clears headers
+%\fancyhf{}%            % Clears headers -- ALSO REMOVES PAGE
+                        % NUMBERS. IF USED, PUT PG# IN FOOTER.
 \fancyhead[LE,RO]{\scriptsize \slshape \rightmark} %{\slshape \rightmark}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Main body Chapters
-%
 \include{010chap1}
-% turning counter reset. perhaps universal endnote counter is good?
-% being schizo about it now. DECIDE! ENDNOTEMARKS DEPEND ON IT.
-%\setcounter{endnote}{0} % need to empty the file too. how?! -
-                        % FIXED. turns out, just having notes is
-                        % enough. or maybe it was fluke.
-\xclearpage
+
 \include{020chap2}
-\xclearpage
+
 \include{030chap3}
-\xclearpage
+
 \include{040chap4}
@@
 
-%\newpage
+\include{_tempscrap}
 
@@
 %
-\begin{singlespace}
-%  \addcontentsline{toc}{chapter}{REFERENCES }
-
-  \tnote[minor]{(i)make names
-    in bib file uniform style w.r.t. firstname/initials (ii) have back
-  refs - i.e. pages that cite the item.}
-  {\footnotesize%
-    \bibliographystyle{alpha}%
-    \bibliography{../lib/cop-variants__thesis} %
-  }
-\end{singlespace}
+\include{biblio}
 

### 2012-02-02 src/0031ded.tex
* Regular edits. See diff for changes.
* Formatting/layout related changes. stuff i'll forget later.
-------
basically:
- added two side printing
- added headers
- added tag/label markers for draft
- made all chapters/chapter*s start on odd page

technically:
- Added contents to Toc
- increased vglue in ded.tex. adjusted one hspace too. to make them fit in.
- began adding short titles for Toc. (chapter 2)
- added twoside option in LoadClass
- adjusted margins in \RequirePackage{geometry} - reduced left margin. otherwise doc bloats up with a lot of margin whitespace on twoside.
- added headings

--- 2012-01-14 src/0031ded.tex
@@
-\definecolor{dim}{rgb}{0.7,0.7,0.7}
-\def \dimspace {0.125in}
+\xclearpage
 
 {% 
-  \fontencoding{T1}%
-  \fontfamily{bch} %{put} %{bch} % {pbk} % \fontfamily{cmdh}
-  % \fontfamily{pag}
-  \selectfont%
-
-  \vglue -0.5in % this makes top margin 2in
-%   \vskip -5ex%
-%   \hspace{5in}\parbox{1in}{\raggedright Hello world}
-
-  \hspace{-1in}%
-  \parbox[c]{4in}{%
-    \large \raggedright \noindent %
-    \color{dim}{%
-      The process of preparing programs for the digital computer is
-      especially attractive, not only because it can be economically
-      and scientifically rewarding, but also because it can be an
-      aesthetic experience much like composing poetry or music.%
+  \thispagestyle{empty}%
+  \parbox[c]{\linewidth}%
   {%
-        \small \par \hfill \emph{The Art of Computer Programming},
-        D.\,E.\,K.%
+    \vspace{0.5\linewidth}% 
+    \centering% 
+    % To you, Z., for your patience.%
+    \large%
+    {\em For Z.}%
   }%
 }%
-  }%
-
-  \vspace{\dimspace}
-
-  \hspace{1.5in}%
-  \parbox[t]{4in}{%
-    \Large \noindent \raggedleft%
-    \color{dim}{%
-      The rhetoric and mythos of science create the comforting image
-      of linear progression toward truth.%
-      {%
-        \small \par \hfill%
-        \textbf{mythos}; \emph{New Oxford American Dictionary, 2nd
-          Ed.}%
-      }%
-    }%
-  }
-  \\
-  \vspace{\dimspace}
-
-  % \hspace{1in}
-  {\centering
-    \parbox{6in}{%
-%      \Huge 
-      \fontsize{27}{1.5}\selectfont
-      \centering \noindent %
-%      \textbf{
-      \emph{L'art c'est la solution au chaos.} \par \noindent
-      \noindent (Art is the solution to chaos.)
-%      }
-%       {%
-%         \small \par \hfill%
-%         \emph{citation needed.}%
-%       }%
-    }%
-  }%
-
-  \vspace{\dimspace}
-
-  \hspace{-0.25in}%
-  \parbox{2.5in}{%
-    \Large 
-    \raggedright \noindent%
-    \color{dim}{%
-      Kilroy wasn't Haar.  Free the group.  Nuke the kernel.  Power to
-      the $n$.  $N = 1 \Rightarrow P = NP$ %
-    }%
-    {%
-      \small \par \hfill%
-      \emph{Concrete Mathematics}%
-    }%
-  }%
-  \vspace{\dimspace}
-
-  \hspace{1in}%
-  \parbox{5in}{%
-    \large \raggedleft \noindent%
-    \color{dim}{%
-      Anything that happens, happens.  Anything that, in happening,
-      causes something else to happen, causes something else to
-      happen.  Anything that, in happening, causes itself to happen
-      again, happens again.  It doesn't necessarily do it in
-      chronological order, though.  }%
-    {%
-      \small \par \hfill%
-      \emph{The Salmon of Doubt}, D.\,N.\,A.%
-    }%
-  }%
-  \vspace{\dimspace}
-
-  \hspace{-0.50in}%
-  \parbox{4in}{%
-    \huge \noindent \raggedright%
-    \color{dim}{%
-      This is water, this is water.%
-      {%
-        \small \par \hfill%
-        D.\,F.\,W.%
-      }%
-    }%
-  }%
-
-}

### 2012-02-02 src/010chap1.tex
-- sent to n for review
- moved ded to quot. added ded.
- pqtree, pctree figures completed.
- fixed a few missing bibs
- random clean up including temptexts. see diff.

@@
 COP. Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
-a submatrix $M'$ with all columns\temptext{ check if b75 deals with
-  COP col or COP row. also is it any submatrix with k less than r rows
-  or submatrix must have all columns? } but a maximum cardinality
+a submatrix $M'$ with all columns but a maximum cardinality
 subset of rows such that $M'$ has COP is NP complete. \cite{hg02}
@@
 known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
-using a reduction to the longest path problem, \temptext{ or is it a
-  survey of another result?  check.} that finding a Tucker's forbidden
-submatrix of at least $k$ rows is NP complete. \temptext{ how is this
-  different from booth's 75 result??}
+using a reduction to the longest path problem, that finding a Tucker's
+forbidden submatrix of at least $k$ rows is NP complete.
 

### 2012-02-02 src/020chap2.tex
-- sent to n for review
- moved ded to quot. added ded.
- pqtree, pctree figures completed.
- fixed a few missing bibs
- random clean up including temptexts. see diff.

@@
 thesis or to \COP in general. These predominantly pertain to
-characterizations of \COP, algorithmic tests to check for \COP (COT),
+characterizations of \COP, algorithmic tests to check for \COP,
 optimization problems on binary matrices that do not have \COP and
@@
 
-\tnote[dire]{ADD: have a few lines about organization of chapter}
+The chapter is organized as follows. Section~\ref{sec:copgraphtheory}
+discusses the close connections of matrices with \COP in graph
+theory. Section~\ref{sec:surveycoptest} discusses the important
+results in the area of \cop testing (COT) -- we describe the history
+of this problem's several solutions including Tucker's forbidden
+submatrices, \PQtree, \PQRtree and ICPIA. \tnote[dire]{mention the section
+  that compares pqr and gen pq tree} 
+Section~\ref{sec:surveycopopt} briefly touches upon the scenario of
+algorithmically verifying that a matrix indeed does not have \COP. Its
+significance lies in the fact that almost all \COT algorithms, upon
+detection of no \COP, simply reports the same without any certificate
+for verification during algorithm implementation. The two approaches
+presented in this section gives a certificate.
 
 \section{\COP in Graph Theory}
+\label{sec:copgraphtheory}
 
@@
       Haj57}.  $G$ is an interval graph \iff its maximal clique matrix
-    has COP \cite{fg65}\footnote{This follows \cite{gh64} which states
+    has COP \cite{fg65} (This follows \cite{gh64} which states
       that the maximal cliques of interval graph $G$ can be linearly
       ordered \stt for all $v \in V(G)$, cliques containing $v$ are
-      consecutive in the ordering \cite[Th. 8.1]{mcg04}.}
+      consecutive in the ordering \cite[Th. 8.1]{mcg04}.)
     \begin{enumerate}[a.]
     \item $G$ is a \emph{unit interval graph} if it is an interval
-      graph \stt all intervals have the same length.\footnotemark[2]
+      graph \stt all intervals have the same length.
     \item $G$ is a \emph{proper interval graph} if it is an interval
-      graph \stt no interval properly contains another.\footnote{The
-        set of unit interval graphs and the set of proper interval
-        graphs are the same} \tnote[pressing]{cite {rob69,gar07} in
-        footnote. pg 33 dom}
-    \end{enumerate}
+      graph \stt no interval properly contains another.
+    \end{enumerate}
+    The set of unit interval graphs and the set of proper interval
+    graphs coincide \cite{rob69, gar07}.
  \item $G$ is a \emph{circular-arc graph} if its vertices can be
@@
 this order. Clearly, this means that the maximal clique incidence
-matrix\footnote{
-  Definition~\ref{def:graphmatrices}~\ref{def::maxcliquematrix}} must
+matrix (see
+Definition~\ref{def:graphmatrices}~\ref{def::maxcliquematrix}) must
 have \COP on rows.
@@
 Booth and Lueker in their paper \cite{bl76} gave the first linear
-algorithm\footnote{Time complexity is $O\left(m+n+f\right)$ where $m
-  \times n$ is the order of the input matrix and $f$ is the number of
-  $\un$s in it.} for \cop testing while given a linear time interval
-graph recognition algorithm by a simplification of \tnote[important]{cite -
-  A. Lempel, S. Even and I. Cederbaum, An Algorithm for Planarity
-  Testing of Graphs, Theory of Graphs, ed., P. Rosenstiehl, Gordon and
-  Breach, New York, (1967), 215-232.  }'s planarity test algorithm.
+algorithm for \cop testing while given a linear time interval graph
+recognition algorithm by a simplification of \cite{lec67}'s planarity
+test algorithm. This \COT algorithm has time complexity
+$O\left(m+n+f\right)$ where $m \times n$ is the order of the input
+matrix and $f$ is the number of $\un$s in it.
 % Hsu in wlh01 said BL76 gave a planarity test - "the induced PQ-tree
@@
   \item Each leaf uniquely represents a row (column) of $M$. The leaf
-    order of the tree gives a \COP order for column (row)\footnote{Note
-      that \COP order for column requires permutation of rows and vice
-      versa.} for $M$.
+    order of the tree gives a \COP order for column (row) (\COP order
+    for column requires permutation of rows and vice versa.) for $M$.
   \item Every non-leaf node in the tree is labeled $P$ or $Q$.
@@
 to matrices and set systems with or without \COP. \cite{km89} invented
-a modified form of \PQtree a simpler incremental update of the tree
-only for recognizing interval graphs.  \cite{kr88} constructed
-efficient parallel algorithms for manipulating \PQtrees. \temptext{dom
-  chapter 2 pg 40 Variations of PQTrees first para - summarize.}
-
-In the next few sections we will see some of these
-variations.
-%Their algorithm was improved by \cite{tm05}
-%\tnote{improved in terms of what?}. 
-% describes an improved algorithm to
-% build PQR trees.
+a modified form of \PQtree, called $MPQ$-trees, a simpler incremental
+update of the tree only for recognizing interval graphs.  \cite{kr88}
+constructed efficient parallel algorithms for manipulating \PQtrees.
+
+In the next few sections we will see some of these variations.
 
@@
 in Definition~\ref{def:matrixsetsystem} along with the idea of \cop
-for set systems\footnote{As seen in Section~\ref{sec:courseschedule}}.
+for set systems (As seen in Section~\ref{sec:courseschedule}).
 
@@
   w_n$ that can be applied to $U$ \stt each set $S \in \cF$ becomes a
-  consecutive subsequence\footnote{Also termed an {\em interval}}
+  consecutive subsequence (slso termed an {\em interval})
   $w_{i}w_{i+1}\ldots w_{i+k-1}$ on $\sigma$ for some positive integer
@@
   \item $A$ and $B$ are said to have a {\em trivial intersection} if
-    $A \cap B$ is one of the following.
-    \begin{enumerate}[i.]
-    \singlespacing
-    \item $\emptyset$
-    \item $A$
-    \item $B$\footnote{In other words, $A$ and $B$ are either disjoint
-        or one is the subset of the other.}
-    \end{enumerate}
-    % \vspace{-5mm}%
+    $A \cap B$ is $\emptyset$ or $A$ or $B$.  In other words, $A$ and
+    $B$ are either disjoint or one is the subset of the other.
 
@@
 
-\temptext{narrative is abrupt from here. add some explanation for the
+\tnote[dire]{narrative is abrupt from here. add some explanation for the
   theorem and proposition below?}
@@
   overlap} equivalence relation which is nothing but {\em
-  non-orthogonal} equivalence relation\footnote{It is easy to verify
-  that this relation is indeed an equivalence relation.}. Overlap
+  non-orthogonal} equivalence relation (It is easy to verify
+  that this relation is indeed an equivalence relation.). Overlap
 components are linearly computable\cite{mm95,wlh92}. Once these
@@
 what they call {\em strong elements} of weak closure,
-$\cW(\cF)$\footnote{\ie $\overline{\cF}$}. Strong elements of a set
+$\cW(\cF)$(\ie $\overline{\cF}$). Strong elements of a set
 family are elements that do not overlap with any other elements in the
@@
 %       {}
-    \temptext{PQR tree def. }
+    \tnote[dire]{PQR tree def. }
     &
@@
       Decomposition tree of $\cW(\cF)$, $T(\cW(\cF))$
-      \temptext{PQR tree def. }
+      \tnote[dire]{PQR tree def. }
     }
@@
     }&
-    \temptext{give def of P Q R nodes}
+    \tnote[dire]{give def of P Q R nodes}
     \\\lasthline
@@
   \centering
-  \includegraphics[scale=0.4]{../img/pctree_hm03.png}
-
-  \centering
-  \includegraphics[scale=0.4]{../img/pctree_dom.png}
-  \caption{\figtabsize \PCtree \temptext{ PLACEHOLDER 
-IMGS \cite{hm03, d08phd}}}
+  %\includegraphics[scale=0.4]{../img/pctree_hm03.png}
+
+  \begin{tabular}[h]{lcc}
+    \pctreegraphmatrix & 
+    \pctreegraphi &
+    \pctreegraphii
+  \end{tabular}
+  \caption[\figtabsize \PCtree]{\figtabsize \PCtree with the
+    associated matrix. The dark circular nodes are $P$-nodes and the
+    double circular nodes are $C$-nodes.  The two \PCtrees are both
+    for the same matrix. The second tree is obtained by performing the
+    ``mirroring'' operation on the subtree (highlighted in the figure)
+    rooted at the bottom right $C$-node.  \cite[Fig.~2.11]{d08phd}}
   \label{fig:pctree}
@@
 \label{sec:icpiasurvey}
-\temptext{narrative is abrupt}
+\tnote[dire]{narrative is abrupt}
 
@@
 assignment %set system
-has no more overlapping intervals\footnote{Intervals are the second
-  element in each ordered pair of $\Pi$. Def.~\ref{def:icpia}} %sets
+has no more overlapping intervals (intervals are the second
+  element in each ordered pair of $\Pi$. Def.~\ref{def:icpia}) %sets
 \cite[Alg.~1]{nsnrs09}. This is acheived by replacing two overlapping
@@
 $\cF$. Let $\cF$, $\cI$ be the set system and interval
-system\footnote{A set system where all the sets are intervals.},
+system (A set system where all the sets are intervals.),
 respectively, involved in the assignment $\Pi$ and $\cF'$, $\cI'$ be
@@
 particular reference, the set of sets is the set of overlaps.)  is
-chosen first and assigned the leftmost\footnote{By
+chosen first and assigned the leftmost\footnote{This is by
   convention. Rightmost can also be chosen analogously to get a
@@
 from this set's overlapping sets. The interval assigned is calculated
-by simple intersection cardinality means\footnote{Assume that the next
-  set overlaps on one of the ``ends'' of the current set and calculate
-  the interval by shifting left or right of the current interval and
-  adjusting its size to match the next set's cardinality.}. Once the
-next set is assigned an interval, the current assigned sets are all
-checked for ICPIA. This is a backtracking algorithm -- if at any point
-ICPIA fails, another overlapping set is chosen. If all overlapping set
+by simple intersection cardinality means. Assume that the next set
+overlaps on one of the ``ends'' of the current set and calculate the
+interval by shifting left or right of the current interval and
+adjusting its size to match the next set's cardinality. Once the next
+set is assigned an interval, the current assigned sets are all checked
+for ICPIA. This is a backtracking algorithm -- if at any point ICPIA
+fails, another overlapping set is chosen. If all overlapping set
 attempts fail ICPIA, a different ``first set'' is chosen. This is a
@@
 
-\temptext{ abrupt: Add a section closing para. does this method (above)
+\tnote[dire]{ abrupt: Add a section closing para. does this method (above)
   lead to efficient algo? }
 
-\section{Optimization problems in COP}
+\section{Matrices without \COP}
 \label{sec:surveycopopt}
@@
 \COP is a very beneficial property since it simplifies the structure
-of the input\footnote{Binary matrix or set system.} leading to
+of the input (binary matrix or set system) leading to
 efficient algorithms to otherwise hard problems.  While
 Section~\ref{sec:surveycoptest} discusses matrices with \COP and how
-to compute \COP orders, this section describes the problems that are
-of interest when a matrix does not have this property.  The central
-questions in this area are (i) how close is the matrix to having \COP,
-mainly in terms of Tucker's forbidden submatrices, and (ii) how
-optimally can one alter the matrix to attain \COP.  With regard to the
-latter question, recent literature indicate that there has been a lot
-of interest in matrix modification problems to make a matrix have \COP
-\cite{hg02, tz07}. However, to tackle this problem it is more than
-insightful to solve question (i).
-
-To study the smallest hamming distance \temptext{(?)} between the
-input matrix and one with \COP, identifying the Tucker's forbidden
-substructures\footnote{See Section~\ref{sec:tucker} for detailed
-  discussion on Tucker's forbidden substructures for matrices with
-  \COP.} is of prime importance.
+to compute \COP orders, this section describes a couple of problems
+when the input does not have \COP and identify their forbidden
+substructures that prevents them from having \COP.
+%describes the problems  that are
+% of interest when a matrix does not have this property
 
@@
 the test is negative. This certainly gives a hint to how hard it is to find
-these substructures in the input. \temptext{Fill neg results on
+these substructures in the input. \tnote[dire]{Fill neg results on
   finding forbidden submartrices -- boo75, some highlighted stuff from v85}
@@
   \label{cor:tuckertheorem}
-  A matrix $M$ has \COP \iff its representing bipartite graph
-  $G_M$\footnote{Graph whose half adjacency matrix
-    (Def.~\ref{def:xxxx}) is $M$.} does not contain an asteroidal
-  triple whose three vertices correspond to columns of M.
+  A matrix $M$ has \COP \iff its representing bipartite graph $G_M$
+  (the graph whose half adjacency matrix is $M$) does not contain an
+  asteroidal triple whose three vertices correspond to columns of M.
 \end{corollary}
@@
   matrices and set systems are equivalent problems as seen in
-  Section~\ref{sec:surveycertalgo} we will use the term set system
-  in this section, in line with \cite{mcc04}'s convention.}. The construction of this graph is based on the following
-observation.
+  Section~\ref{sec:surveycertalgo} we will use the term set system in
+  this section, in line with \cite{mcc04}'s convention.}. The
+construction of this graph is based on the following observation.
 
@@
 
-\temptext{how does it give clues
+\tnote[dire]{how does it give clues
   about the $R$ node? See dom.}
@@
 
-\temptext{\textbf{REST OF THE SUBSECTIONS}}
-
-\temptext{ -- sect 4.1 in cite:d08phd has many results surveyed. hardness
+\tnote[dire]{ -- sect 4.1 in cite:d08phd has many results surveyed. hardness
   results, approx. results. results are usually for a class of
@@
 
-\section{COP in Graph Isomorphism}
-\label{sec:surveygraphiso}
-\tnote{(canonization theme)}
-
-\tnote[pressing]{ADD: peo exists iff chordal. lexicographic BFS
-  [tag:chordalGraph]} %
-
-\temptext{
-The survey from kklv10 conclusion.
-}
-
 %%%%%%%%%%%%% listing package for pretty formatting BEGIN

### 2012-02-02 src/040chap4.tex
- a whole bunch of cleanups. see diff. - removed problem defs from chap 1. they were duplicates from chap 3. - CODED FIGURES IN TIKZ. forbidden subgraphs, graphmatrix example graphs. - BEGAN CHAPTER 3 EDITS.
Formatting/layout changes
-------

- Added FrenchLayout to make lists denser
- Added hanging indent for Definitions
- Compressed theorem env's above below spaces
- Creaded env for problem definition minipages.
- moved all clearpage/page counter and other prechapter settings to the chapter tex files.
- Fixed the page number bug in draft printout. removed \fancyhf{} to avoid resetting page n

@@
+\xclearpage
+
 \chapter{Conclusion}
@@
 }%
+
+There are optimization problems for matrices that do not have \COP
+which we only mention here but do not discuss further in this
+thesis. The central questions in this area are (i) how close is the
+matrix to having \COP, mainly in terms of Tucker's forbidden
+submatrices, and (ii) how optimally can one alter the matrix to attain
+\COP.  With regard to the latter question, recent literature indicate
+that there has been a lot of interest in matrix modification problems
+to make a matrix have \COP \cite{hg02, tz07}. However, to tackle this
+problem it is more than insightful to solve question (i).
+
+To study the smallest hamming distance \temptext{(?)} between the
+input matrix and one with \COP, identifying the Tucker's forbidden
+substructures (See Section~\ref{sec:tucker} for detailed
+  discussion on Tucker's forbidden substructures for matrices with
+  \COP) is of prime importance.
 
@@
   ----- see cite-d08phd Chapter 6?  }
+
+\temptext{
+COP in Graph Isomorphism  -- The survey from kklv10 conclusion.
+}

### 2012-02-02 src/0032quote.tex
-- sent to n for review
    - moved ded to quot. added ded.
    - pqtree, pctree figures completed.
    - fixed a few missing bibs
    - random clean up including temptexts. see diff.

(missd a file)

\xclearpage

\definecolor{dim}{rgb}{0.7,0.7,0.7}
\def \dimspace {0in} %{0.125in}
\def \dimspacemore {0.5in} %{0.125in}
\def \authspace {-0.25in}
\def \brightspace {0.375in}

{%
  \thispagestyle{empty}
  \fontencoding{T1}%
  \fontfamily{bch} %{put} %{bch} % {pbk} % \fontfamily{cmdh}
  % \fontfamily{pag}
  \selectfont%

  \vglue -1.5in % this changes top margin 
%   \vskip -5ex%
%   \hspace{5in}\parbox{1in}{\raggedright Hello world}

  \hspace{-0.75in}%
  \parbox[t]{6.5in}{%
    \Large \noindent \raggedright%
    \color{dim}{%
      % One teacher called these moments 'mathematical
      % experiences'. What I didn't know then was that
      [...] a mathematical experience was aesthetic in nature, an
      epiphany in Joyce's original sense. These moments appeared in
      proof-completions, or maybe algorithms. Or like a gorgeously
      simple solution you suddenly see after filling half a notebook
      with gnarly attempted solutions. It was really an experience of
      what I think Yeats called ``the click of a well made box''.%
      % The word I always think of is as 'click'.
      {%
        \vspace{\authspace}
        \small \par \hfill%
        D.\,F.\,W.%
      }%
    }%
  }%

  \vspace{\dimspace}

  \hspace{1in}%
  \parbox{4in}{ }{%
    \LARGE 
    \raggedleft \noindent%
    \color{dim}{%
      Math Graffiti: Kilroy wasn't Haar.  Free the group.  Nuke the kernel.  Power to
      the $n$.  $N = 1 \Rightarrow P = NP$ %
      {%
%        \vspace{\authspace}%
        \small \par \hfill%
        \emph{Concrete Mathematics (margin notes)}%
      }%
    }%
  }%

  \vspace{\dimspace}

  \hspace{-1in}%
%  \ding{125}
  \parbox[t]{5in}{%
    \large \raggedright \noindent %
    \color{dim}{%
      The process of preparing programs for the digital computer is
      especially attractive, not only because it can be economically
      and scientifically rewarding, but also because it can be an
      aesthetic experience much like composing poetry or music.%
      {%
%        \vspace{\authspace}%
        \small \par \hfill \emph{The Art of Computer Programming},
        D.\,E.\,K.%
      }%
    }%
  }%

  \vspace{\brightspace}

  {%
      \centering%
      \hspace{-1.25in}
      \parbox{8in}{%
      \fontsize{36}{1.5}\selectfont%
      \centering \noindent%
      \emph{L'art c'est la solution au chaos.}%
      \par \noindent%
      \normalsize (Art is the solution to chaos.)%
    }%
  }%

  \vspace{\brightspace}

 \hspace{-0.25in}%
  \parbox[t]{6.5in}{%
    \LARGE  \noindent \raggedleft%
    \color{dim}{%
      \textbf{myth\textperiodcentered os} {|{\huge $'$mi $\theta$
          $\overline{o}$s}|} \\
      a set of beliefs or assumptions about something
      \emph{ :\\ 
      the rhetoric and mythos of science create the comforting image
      of linear progression toward truth.}%
      {%
%        \vspace{\authspace}%
        \small \par \hfill%
        \emph{New Oxford American Dictionary, 2nd
          Ed.}%
     }%
    }%
  }

  \vspace{\dimspace}

  \hspace{-1in}%
  \parbox{6in}{%
    \singlespacing
    \large \raggedright \noindent%
    \color{dim}{%
      Anything that happens, happens.  Anything that, in happening,
      causes something else to happen, causes something else to
      happen.  Anything that, in happening, causes itself to happen
      again, happens again.  It doesn't necessarily do it in
      chronological order, though.%
      {%
%       \vspace{\authspace}%
        \small \par \hfill%
        \emph{The Salmon of Doubt}, D.\,N.\,A.%
      }%
    }%
  }%

  \vspace{\dimspace}

  \hspace{1.5in}%
  \parbox{4in}{%
    %\huge 
    \Large 
    \noindent \raggedleft%
    \color{dim}{%
      Everything in its right place.%
      {%
        \small \par \hfill%
       \emph{Kid A, Radiohead}, T.\,Y.%
      }%
    }%
  }%

%   \hspace{1.5in}%
%   \parbox{4in}{%
%     %\huge 
%     \Large 
%     \noindent \raggedleft%
%     \color{dim}{%
%       This is water, this is water.%
%       {%
% %        \vspace{\authspace}%
%         \small \par \hfill%
%         D.\,F.\,W.%
%       }%
%     }%
%   }%
  % xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx
  % xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx
  % xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx
  % xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx
  % xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx
  % xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx
  % xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx
  % xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx

  \vskip -02.5in % this changes bottom margin (hopefully)

}

### 2012-02-04 src/010chap1.tex
first draft of chapter 4 done. and hopefully the last draft.

@@
     $\cI$. \[G \cong \bI(\cI)\]
-  \item $G$ is a \textbf{path graph} if there exists a tree $T$ and a
-    set of paths from $T$, $\cP$ \stt $G$ is isomorphic to the
-    intersection graph of $\cP$. \[G \cong \bI(\cP)\]
+  \item $G$ is a \textbf{path graph} if there exists a (undirected
+    unrooted) tree $T$ and a set of paths from $T$, $\cP$ \stt $G$ is
+    isomorphic to the intersection graph of $\cP$. \[G \cong
+    \bI(\cP)\] Path graphs are also known as \uvgraphs (Undirected
+    tree where paths intersect in a Vertex) \cite{mw86}.
   \item $G$ is a \textbf{chordal graph} if it has no induced cycle of
@@
 \end{definition}
+
+The set of interval graphs is a subset of set of path graphs which is
+a subset of set of chordal graphs. In fact, there is another class
+between chordal graphs and path graphs called RDV-graphs (Rooted
+Directed tree where paths intersect in a Vertex)  \cite{mw86} which is
+defined exactly as a path graph except that the trees are rooted and
+directed.
 

### 2012-02-04 src/020chap2.tex
first draft of chapter 4 done. and hopefully the last draft.

@@
 \subsection{Finding forbidden submatrices}
+\label{sec:forbiddensubalgo}
+
+\tnote[dire]{
+To study the smallest hamming distance \temptext{(?)} between the
+input matrix and one with \COP, identifying the Tucker's forbidden
+substructures (See Section~\ref{sec:tucker} for detailed
+  discussion on Tucker's forbidden substructures for matrices with
+  \COP) is of prime importance and we saw an algorithm for it in
+  Section~\ref{sec:forbiddensubalgo}. 
+}
+
 It is known from Theorem~\ref{th:tuckercop}

### 2012-02-04 src/030chap3.tex
first draft of chapter 4 done. and hopefully the last draft.

@@
     path labeling to any tree since it does not have the Helly's
-    property anymore due to the addition of hyperedge $\{x_1,
+    property \cite[Pr.~4.7]{mcg04} anymore due to the addition of hyperedge $\{x_1,
     x_4\}$. }
@@
 
-\remove{It is an interesting fact that for a matrix with the COP, the
-  intersection graph of the corresponding set system is an interval
-  graph.  A similar connection to a subclass of chordal graphs, and
-  this subclass contains interval graphs, exists for the
-  generalization of COP.  In this case, the intersection graph of the
-  corresponding set system must be a path graph. Chordal graphs are of
-  great significance, extensively studied, and have several
-  applications.  One of the well known and interesting properties of a
-  chordal graphs is its connection with intersection graphs
-  cite{mcg04}. For every chordal graph, there exists a tree and a
-  family of subtrees of this tree such that the intersection graph of
-  this family is isomorphic to the chordal graph
-  cite{plr70,gav78,bp93}.  Certain format of these trees are called
-  clique trees cite{apy92} of the graph which is a compact
-  representation of the chordal graph. There has also been work done
-  on the generalization of clique trees to clique hypergraphs
-  cite{km02}.  If the chordal graph can be represented as the
-  intersection graph of paths in a tree, then the graph is called path
-  graph cite{mcg04} or path chordal graph.  Therefore, it is clear
-  that if there is a bijection from $U$ to $V(T)$ such that the sets
-  map to paths, then the intersection graph of the set system is
-  indeed a path chordal graph.  This is, however, only a necessary
-  condition and can be checked efficiently, as path graph recognition
-  is polynomial solvable cite{gav78,aas93}.  Indeed, it is possible to
-  construct a set system and tree, such that the intersection graph is
-  a path chordal graph, but there is no bijection between $U$ and
-  $V(T)$ such that the sets map to paths.  This connection indeed
-  suggests that our problem is indeed as hard as path graph
-  isomorphism.  Further path graph isomorphism is known be
-  isomorphism-complete, see for example cite{kklv10}. } In the second
-part of this paper, we decompose our search for a bijection between
-$U$ and $V(T)$ into subproblems.  Each subproblem is on a set system
-in which for each set, there is another set in the set system with
-which the intersection is {\em strict}- there is a non-empty
-intersection, but neither is contained in the other.  This is in the
-spirit of results in \cite{wlh02,nsnrs09} where to test for COP in a
-given matrix, the COP problem is solved on an equivalent set of prime
-matrices.  \annote{Our decomposition localizes the challenge of path
-  graph isomorphism to two problems.}{HOW?}
-
-Finally, we show that Tree Path Assignment is isomorphism-complete.
+In the second part of this paper, we decompose our search for a
+bijection between $U$ and $V(T)$ into subproblems.  Each subproblem is
+on a set system in which for each set, there is another set in the set
+system with which the intersection is {\em strict}- there is a
+non-empty intersection, but neither is contained in the other.  This
+is in the spirit of results in \cite{wlh02,nsnrs09} where to test for
+COP in a given matrix, the COP problem is solved on an equivalent set
+of prime matrices.
+
+% Finally, we show that Tree Path Assignment is isomorphism-complete.
 We also point out Consecutive Ones Testing is in Logspace from two

### 2012-02-04 src/040chap4.tex
first draft of chapter 4 done. and hopefully the last draft.

--- 2012-01-03 src/040chap4.tex
@@
-\chapter{Conclusion}
+\xclearpage
+
+\def\p{{\sc P}\xspace}
+\def\gicomplete{{\sc GI}-complete\xspace}
+\def\logspace{{\sc logspace\xspace}\xspace}
+\def\ptime{{\sc ptime}\xspace}
+\def\fpplusc{{\sc FP+C}\xspace}
+
+\chapter[Conclusion]{Conclusion and Further Research}
 \label{ch:conclusion}
 
-\begin{center} {\tt MOVED FROM INTRO SUBSECTION IN ch:myresearch
-    CHAPTER begin}
-\end{center}
+In this thesis, we surveyed \cop testing algorithms and as new results
+we gave a characterization for a generalization of \COP, namely,
+feasible tree path labeling of path hypergraphs. The proof for this is
+constructive and computes a feasibility bijection mapping vertices of
+the hypergraph to vertices of the given target tree. We also described
+an exponential algorithm that computes a feasible TPL to a given
+hypergraph if it is a path hypergraph.  In this chapter we will see a
+few ideas where tree path labeling, in our experience, shows promising
+scope for futher research.
 
-% \annote{
-It is an interesting fact that for a matrix with the COP, the
-intersection graph of the corresponding set system is an interval
-graph. % }{the following text could be in Further Study/Conclusion}
-A similar connection to a subclass of chordal graphs and a superclass
-of interval graphs exists for the generalization of COP.  In this
-case, the intersection graph of the corresponding set system must be a
-{\em path graph}. Chordal graphs are of great significance, are
-extensively studied, and have several applications.  One of the well
-known and interesting properties of a chordal graphs is its connection
-with intersection graphs \cite{mcg04}. For every chordal graph, there
-exists a tree and a family of subtrees of this tree such that the
-intersection graph of this family is isomorphic to the chordal graph
-\cite{plr70,gav78,bp93}.  These trees when in a certain format, are
-called {\em clique trees} \cite{apy92} of the chordal graph. This is a
-compact representation of the chordal graph. There has also been work
-done on the generalization of clique trees to clique hypergraphs
-\cite{km02}.  If the chordal graph can be represented as the
-intersection graph of paths in a tree, then the graph is called path
-graph \cite{mcg04}.  Therefore, it is clear that if there is a
-bijection from $U$ to $V(T)$ such that for every set, the elements in
-it map to vertices of a unique path in $T$, then the intersection
-graph of the set system is indeed a path graph.  However, this is only
-a necessary condition and can be checked efficiently because path
-graph recognition is polynomial time
-solvable\cite{gav78,aas93}. Indeed, it is possible to construct a set
-system and tree, such that the intersection graph is a path graph, but
-there is no bijection between $U$ and $V(T)$ such that the sets map to
-paths. Path graph isomorphism is known be isomorphism-complete, see
-for example \cite{kklv10}. An interesting area of research would be to
-see what this result tells us about the complexity of the tree path
-labeling problem (not covered in this paper).
+\parindent 0pt
 
-\begin{center} {\tt MOVED FROM INTRO SUBSECTION IN ch:myresearch
-    CHAPTER\\ end}
-\end{center}
+\textbf{Graph isomorphism and logspace canonization.}  Canonization is
+an important tool in graph isomorphism. Invention of a deterministic
+method of canonization for any class of graphs naturally results in an
+algorithm for isomorphism. All that is required is to check if the
+canons of two graphs are the same. Thus one way to study the
+complexity of graph isomorphism is by studying canonization
+methods. While general graph ismorphism remains elusive in terms of
+hardness, canonization has been studied for smaller classes of graphs
+thus giving complexity or hardness results for them -- some of the
+canonization results are \cite{sl92,dlntw09,adkk09,kklv10}.
 
-We give a characterization for feasible tree path labeling of path
-hypergraphs. The proof for this is constructive and computes a
-feasibility bijection mapping vertices of the hypergraph to vertices
-of the given target tree. This thesis also discovered an exponential
-algorithm that computes a feasible TPL to a given hypergraph if it is
-a path hypergraph.
+Ten years ago, \cite{sl92} made an important discovery that tree
+isomorphism is in \logspace. It was done by inventing a method of
+canonization of trees using a logspace depth first traversal
+algorithm.  Less than a couple of years ago, \cite{kklv10} proved that
+interval graph canonization is also in \logspace thus drawing logspace
+conclusions about \COT. Interval graphs are \fpplusc (fixed point
+logic with counting) definable and \cite{l10} showed that this implies
+that it captures \ptime. This result along with that of undirected
+graph reachability being in logspace \cite{rei08} helped \cite{kklv10}
+conclude their logspace result.
 
-Our results have close connections to recognition of path graphs and
-to path graph isomorphism.  Graphs which can be represented as the
-intersection graph of paths in a tree are called {\em path
-  graphs}\cite{mcg04}. Thus, a hypergraph $\cF$ can be interpreted as
-paths in a tree, if and only if the intersection graph of $\cF$ is a
-{\em path graph}. Path graphs are a subclass of chordal graphs since
-chordal graphs are characterized as the intersection graphs of
-subtrees of a tree\cite{mcg04}.  Path graphs are well studied in the
-literature \cite{plr70}--\cite{mcg04}\tnote{Check if cite:bp93 is
-  indeed for path graphs}.  Path graph recognition can be done in
-polynomial time\cite{gav78,aas93}.  Clearly, this is a necessary
-condition in terms of the intersection graph of the input hypergraph
-$\cF$ in \FTPL. However, one can easily obtain a counterexample to
-show the insufficiency of this condition\tnote{Give a figure
-  showing insuff.}.  Path graph isomorphism is known be
-isomorphism-complete\cite{kklv10}. Therefore, \annote{it is
-  unlikely that we can solve the problem of finding feasible path
-  labeling $\cl$ for a given $\cF$ and }{WHY?} tree $T$.  It is
-definitely interesting to classify the kinds of trees and hypergraphs
-for which feasible path labelings can be\tnote{ELABORATE.}  found
-efficiently.  \annote{These results would form a natural
-  generalization of COP testing and interval graph isomorphism,
-  culminating in Graph Isomorphism }{REWORD} itself.  To this effect
-we consider TPL on {\kstar} and give a polynomial time solution for
-the same.
+How the work in this thesis connects to graph isomorphism and
+canonization is as follows. For a matrix with the \COP, the
+intersection graph of the input hypergraph is an interval graph.  A
+similar connection exists for \TPL to path graphs which is a subclass
+of chordal graphs and a superclass of interval graphs. Graphs which
+can be represented as the intersection graph of paths in a tree are
+path graphs (as seen in Section~\ref{sec:basicprelim}).  % In TPL, the
+% intersection graph of the input hypergraph must be a path graph.  % Our
+% results have close connections to recognition of path graphs and to
+% path graph isomorphism.  
+Thus, a hypergraph $\cF$ can be labeled with paths in a tree (path
+labeling as defined in Section~\ref{ch:prelims}), \iff the
+intersection graph $\bI(\cF)$ is a path graph. Path graphs are a
+subclass of chordal graphs since chordal graphs are characterized as
+the intersection graphs of subtrees of a tree \cite[Ch.~4,
+Sec.~5]{mcg04}. Chordal graphs are of great significance, are
+extensively studied, and have several applications. For every chordal
+graph, there exists a tree and a family of subtrees of this tree such
+that the intersection graph of this family is isomorphic to the
+chordal graph \cite{jrw72phd,gav74b,bun74}.  These trees when in a
+certain format, are called {\em clique trees} \cite{ppy92} of the
+chordal graph. This is a compact representation of the chordal
+graph. There has also been work done on the generalization of clique
+trees to clique hypergraphs \cite{km02}.  Path graphs are well studied
+in the literature (characterization for \uvgraphs was perhaps first
+mentioned by \cite{plr70}).
+% Clearly, this is a necessary condition in terms of the intersection
+% graph of the input hypergraph $\cF$ in \FTPL. However, one can easily
+% obtain a counterexample to show the insufficiency of this
+% condition\tnote{Give a figure showing insuff.}.
+Indeed, path graph recognition is a necessary condition for \CFTPL but
+not sufficient.  We know that path graph recognition can be done in
+polynomial time \cite[first algorithm.
+$O\left(pn^3\right)$]{gav78},\cite[$O\left(p(m+n)\right)$]{aas93}
+\cite{sb94}, \cite[linear algorithm]{db95}, \cite[PR-trees]{cha11}
+\footnote{$n$ is number of vertices, $m$ is number of edges, $p$ is
+  number of maximal cliques in the input graph.}.  Path graph
+isomorphism is also known be \gicomplete \cite{bpt96}.  It would be
+interesting to pursue research to answer the following questions in
+relation to TPL and graph isomorphism.
+\begin{enumerate}
+\item What does path graph isomorphism hardness tell us about the
+  complexity of \CFTPL?
+\item Can TPL help in computing a canon for path graphs, especially a
+  logspace canon?
+\end{enumerate}
 
-While we address the computation of TPL for a given
-hypergraph% \footnote{The term {\em hypergraph} is used
-% instead {\em set systems} are used interchangably in this
-% section. Hypergraphs that have TPLs on some target tree are called
-% {\em path hypergraphs}.}
-from a given target tree in this research, however optimization
-problems in TPL, akin to problems in Section~\ref{sec:optcop} for COP,
-is outside the scope of this thesis. Whether TPL on general trees is
-solvable in $P$ remains open. So do optimization opportunities in TPL
-(possible extensions to optimization for COP in the
-Section~\ref{sec:optcop}.\tnote{link to section}).
+%\noindent
+\textbf{Complexity of TPL in other tree classes.}  We know that on
+intervals, \CFTPL is easy (Section~\ref{sec:icpplicpia}).  We showed
+that \CFTPL on {\kstar} is also polynomial time solvable in
+Section~\ref{sec:ksubdivstar}. However, due to its close connections
+with path graph isomorphism which is \gicomplete \cite{kklv10}, we
+conjecture that \CFTPL on general trees cannot do better than
+\gicomplete. Our algorithm in Section~\ref{sec:decompos} is unable to
+solve it in \p. It would therefore, be certainly of interest to
+classify the kinds of trees and hypergraphs for which feasible path
+labelings can be found efficiently, or better still, to solve \CFTPL
+for general trees itself in \p-time or prove that it is not in \p.
 
-% \newpage %DUMMY
-\tnote{ ||Complexity challenges:|| Canonization is an important
-  tool in graph isomorphism. Invention of a
-  % deterministic method of canonization for any class of graphs
-  % naturally results in an algorithm for isomorphism. All that is
-  % required is to check if the canons of two graphs are the
-  % same. Thus complexity of graph isomorphism can be studied by
-  % studying canonization methods. While general graph ismorphism
-  % remains elusive in terms of hardness, canonization has been
-  % studied for smaller classes of graphs thus giving
-  % complexity/hardness results for them.
+%\noindent
+\textbf{Optimization problems in TPL.}  There are optimization
+problems for matrices that do not have \COP which we only mention in
+Section~\ref{sec:optcop} but did not discuss further since it is not
+in the scope of the main work of this thesis. We mention them here
+again to encourage further analogous research in optimization problems
+of \TPL.  The central questions in this area are (i) how close is the
+matrix to having \COP, mainly in terms of Tucker's forbidden
+submatrices, and (ii) how optimally can one alter the matrix to attain
+\COP.  With regard to the latter question, recent literature indicate
+that there has been a lot of interest in matrix modification problems
+to make a matrix have \COP \cite{hg02, tz07}. While we address the
+computation of TPL for a given hypergraph from a given target tree in
+this research, however optimization opportunities in TPL are yet to be
+explored and their complexity remains open. We suggest a few
+optimization ideas here and there could be many more such questions
+that could be asked by a motivated researcher. 
+\begin{enumerate}
+\item If \CFTPL is not possible with the input target tree $T$, 
+  can $T$ be changed to get a feasible TPL?
+\item What is the characterization of a hypergraph that will fail the
+  \CFTPL problem for any target tree? \ie what is the characterization
+  of a hypergraph that is not a path hypergraph? Can a forbidden
+  substructure be found?
+\item If a given hypergraph is not a path hypergraph, can at most $k$
+  vertices be deleted from the hypergraph to make it a path
+  hypergraph? A dual problem would be to select at least $k$ vertices
+  to induce a path hypergraph? A different flavour of the problem
+  would be deleting (or selecting) hyperedges in the input to get a
+  path hypergraph (analogous to {\sc Min COS-R, Min COS-C, Max
+    COS-R, Max COS-C} problems in \COP).
+\item Can at most $k$ vertex memberships in hyperedges be
+  reversed to get a path hypergraph? (analogous to {\sc COP
+    Augmentation} problem).
+\end{enumerate}
+It is unlikely that these problems will have efficient exact
+algorithms since their \COP counterparts are all NP hard
+problems. Thus the scope of research would be to find approximation
+algorithms or fixed parameter algorithms or hardness results.
 
-  % In 1992, \cite{sl92} made an important discovery that tree
-  % isomorphism is in logspace. It was done by inventing a method of
-  % canonization of trees using a logspace depth first traversal
-  % algorithm. \tnote{confirm the statement of claim}.
+% %\noindent
 
-  % \cite{kklv10} proved that interval graph canonization is also in
-  % logspace thus drawing logspace conslusions about COT.
-  % % This result is
-  % % built on the fact that
-  % Interval graphs are FP+C (fixed point with counting) definable
-  % \tnote{elucidate} and \cite{l10} showed that this implies that it
-  % captures PTIME. \tnote{elucidate} This result along with that of
-  % undirected graph reachability being in logspace \cite{rei08},
-  % \cite{kklv10} proved their logspace result.
-}%
-\tnote{ (1) the interesting survey in conclusion section of kklv10
-  (2) [[[COP helping some problems with hardness]]] having COP in the
-  input structure makes some problems less hard than in a general
-  input. -- min set cover problem is known to be NP hard. so is min
-  hitting
-  set. these two problems are equivalent cite-adp80\\
-  -- -- complexity results: log factor polynomial time approximable cite-f98, W[2]-complete with solution size as parameter cite-df99\\
-  -- a couple of generalizations:\\
-  -- -- min degree hypergraph (MDH) problem\\
-  -- -- red blue covering (RBSC) problem\\
-  ----- cite-d08phd Chapter 5.\\
-  ----- what are cite-d08phd results\\
-  ----- Rectangle stabbing \\
-  ----- see cite-d08phd Chapter 6?  }
+% \temptext{ 
+%   (2) [[[COP helping some problems with hardness]]] having COP in the
+%   input structure makes some problems less hard than in a general
+%   input. -- min set cover problem is known to be NP hard. so is min
+%   hitting
+%   set. these two problems are equivalent cite-adp80\\
+%   -- -- complexity results: log factor polynomial time approximable
+%   cite-f98, W[2]-complete with solution size as parameter cite-df99\\ 
+%   -- a couple of generalizations:\\
+%   -- -- min degree hypergraph (MDH) problem\\
+%   -- -- red blue covering (RBSC) problem\\
+%   ----- cite-d08phd Chapter 5.\\
+%   ----- what are cite-d08phd results\\
+%   ----- Rectangle stabbing \\
+%   ----- see cite-d08phd Chapter 6?  }
+
+%\noindent
+\textbf{x-dimensional generalization of \COP.} As seen in the chapters
+of this thesis \TPL is a graph theoretic generalization of
+\COP. Intervals are target trees with maximum degree 2 and general
+trees have no bound on maximum degree. Another way to look at
+intervals is using geometry. Intervals can be imagined in one
+dimension as line segments along a single line in the Euclidean space
+and the elements of hypergraph universe are coordinate points on this
+line segment. The generalization of this visualization is obvious --
+how does the interval labeling property generalize when each hyperedge
+must be mapped to coordinate points bounded by a rectilinear 2-D shape
+on a single plane, say rectangles (analogous to coordinate points on
+line segments in interval labeling problem). One hardness result is
+already known -- the intersection graphs of axis-parallel boxes in
+$\bR^d$ is GI-complete \cite{ueh08}.

### 2012-02-20 src/010chap1.tex
- incorporated new changes to Characterization algorithm. no more exit    statements in algo 1 and 2. however it is not identitcal to n's    version. kept some of the old lemmas because it seems to elucidate    (to me) the thesis. - partial editing done of general tree section (prime matrix theory)    -- TBD - not sure about status of special tree = path section -- TBD
- changed notation for induced path labeling
- changed filter 1 and 2 names.
- poset definition

@@
 
+\def\posetr{\le}
+
+\begin{definition}[Poset]
+  \label{def:posetdef}    
+  Let $\posetr$ be a binary relation on a finite set $X$. $\posetr$ is a \textbf{
+    partial order} and $\left< X,\posetr \right>$  is a \textbf{ poset}
+  if the following hold true for any $a, b, c \in X$.
+ \parindent  \defindent 
+%  \hangindent \defindent
+  \begin{enumerate}
+  \item $a \posetr a$ (reflexivity)
+  \item If $a \posetr b$ and $b \posetr a$ then $a = b$ (antisymmetry)
+  \item If $a \posetr b$ and $b \posetr c$ then $a \posetr c$
+    (transitivity)
+  \end{enumerate}
+
+  An element $b$ \textbf{covers} $a$ provided that there exists no
+  third element $x \in X$ \stt $a \posetr x \posetr b$.  A directed
+  (or undirected, but with direction implied by horizontal positioning
+  of nodes) graph that represents $\posetr$ in terms of its covering
+  relation is called the \textbf{Hasse diagram} of the poset.
+\end{definition}
+
 \begin{definition}
   \label{def:poset}    
-%   Let $A$ be a set and $R$ be a binary relation $R \subseteq A \times
-%   A$.
-
-%   \textbf{poset}
   Let $X$ be a partially ordered set with $\preccurlyeq$ being the
@@
 \cite{kklv10}. Moreover, \cite{nsnrs09} discovered that these
-equivalence classes form a total order. We extend this to TPL and find
+equivalence classes form a total order in the case of \COP. We extend this to TPL and find
 that when $\cF$ is a path hypergraph\footnote{If there exists an FTPL

### 2012-02-20 src/030chap3.tex
- incorporated new changes to Characterization algorithm. no more exit    statements in algo 1 and 2. however it is not identitcal to n's    version. kept some of the old lemmas because it seems to elucidate    (to me) the thesis. - partial editing done of general tree section (prime matrix theory)    -- TBD - not sure about status of special tree = path section -- TBD
- changed notation for induced path labeling
- changed filter 1 and 2 names.
- poset definition

@@
 presented in Section~\ref{ch:prelims}. Section~\ref{sec:feasible}
-documents the characterization of a feasible path labeling from any
-target trees. Section~\ref{sec:spltargettree} describes two special
-cases where the target tree is of a particular family of trees. The
-first one is shown to be equivalent to COP testing in
+documents a characterization for feasible path labelings. In
+Section~\ref{sec:wallacesol} we demonstrate the algorithm described in
+Section~\ref{sec:feasible} by working out the solution to an example.
+Section~\ref{sec:spltargettree} describes two special cases where the
+target tree is of a particular family of trees. The first one is shown
+to be equivalent to COP testing in
 Section~\ref{sec:icpplicpia}. Section~\ref{sec:ksubdivstar} discusses
@@
 version where the target tree has no restrictions and the algorithm to
-find a feasible TPL, if any, in this case.
+find a feasible TPL, if any, in this case.  Finally, in Section
+\ref{sec:tplcomplexity} we discuss the complexity of the algorithms
+seen thus far and we conclude by showing that \cop testing, which is a
+special case of \TPL, is in \logspace.
 
@@
 \end{definition}
-In-trees are used in Section~\ref{sec:decompos} which describes our
-solution to \CFTPL problem.
+In-trees are used in Section~\ref{sec:norestraint} which describes our
+(albeit non-polynomial time) solution to \CFTPL problem.
 
@@
 In the remaining part of this section we show that $(\cF, \cl)$ is
-feasible if and only if it is an ICPPL%  and
-% Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty path
-% hypergraph isomorphism
-% function
-.  Algorithm~\ref{al:icppl-find-isomorph} takes $(\cF, \cl)$ as input
-and computes the feasibility bijection if it is feasible or else
-reports failure. This algorithm recursively does two levels of
-filtering of $(\cF, \cl)$ to make it simpler in terms of set
-intersections while retaining a non-empty subset of isomorphisms, if
-any, between $\cF$ and $\cF^\cl$. These two filters are {\tt
-  filter\_1} in Algorithm~\ref{al:filterone} and {\tt filter\_2} in
+feasible if and only if it is an ICPPL.
+Algorithm~\ref{al:icppl-find-isomorph} takes $(\cF, \cl)$ as input and
+computes the feasibility bijection if it is feasible or else reports
+failure. This algorithm recursively does two levels of filtering of
+$(\cF, \cl)$ to make it simpler in terms of set intersections while
+retaining a non-empty subset of isomorphisms, if any, between $\cF$
+and $\cF^\cl$. These two filters are \filteri in
+Algorithm~\ref{al:filterone} and \filterii in
 Algorithm~\ref{al:filtertwo}.
@@
 
-First, we present Algorithm~\ref{al:filterone} which describes {\tt
-  filter\_1}, and prove its correctness.  This algorithm refines the
+First, we present Algorithm~\ref{al:filterone} which describes \filteri, and prove its correctness.  This algorithm refines the
 path labeling by processing pairs of paths in $\cF^\cl$ that share a
@@
 \begin{algorithm}[h]
-  \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
+  \caption{Refine ICPPL {\tt \filteri($\cF, \cl, T$)}}
   \label{al:filterone} % {perms}
@@
 
-The following lemma is useful in subsequent arguments.
+In order to prove the correctness of Algorithm~\ref{al:filterone}, we
+present a few lemma which are useful in subsequent arguments.
+
+The first two lemma below are from results in \cite{nsnrs09}.  They
+have been reworded to use the terminology adopted in this
+thesis. Lemma~\ref{lem:intersection-cardinality} shows that the
+preservation of pairwise intersection cardinality in an interval
+assignment is sufficient to preserve three way intersection
+cardinality. Lemma~\ref{lem:icpia} shows that if a path labeling (i)
+has target tree as a path \ie it is an interval labeling, and (ii) it
+preserves pairwise intersection cardinalities \ie it is an ICPIA,
+then it is a feasible path labeling.
+
+\begin{lemma}[{\cite[Lem.~1]{nsnrs09}}]
+  \label{lem:intersection-cardinality} 
+  Let $P$ be a path, $S_1, S_2, S_3$ be 3 sets, and $T_1, T_2, T_3$ be
+  paths from $P$, such that $|S_i \cap S_j |=|T_i \cap T_j|, 1 \leq
+  i,j \leq 3$.  Then, $|S_1 \cap S_2 \cap S_3| = |T_1 \cap T_2 \cap
+  T_3|$.
+\end{lemma}
+\begin{lemma}[{\cite[Th.~2]{nsnrs09}}]
+  \label{lem:icpia} 
+  A path labeling $(\cF, \cl)$ with target tree $T$ as a path is
+  feasible iff it is an ICPIA.
+\end{lemma}
+
+Next, we see Lemma~\ref{lem:setminuscard} which is a direct
+consequence of Definition~\ref{def:icppl} and is analogous to
+\cite[Cor.~1]{nsnrs09}.
+
 \begin{lemma}
   \label{lem:setminuscard}
-  If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
+  If $\cl$ is an \ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
   (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
@@
 
+For interval labeling, \cite{nsnrs09} shows that if it is an ICPIA \ie
+pairwise intersection cardinalities are preserved, then $d$-way
+intersection cardinalities are preseved for $2 < d < n$. Thus, in a
+way, pairwise intersections ``capture'' feasibility of an interval
+labeling. In the same vein, the following lemma indicates that
+three-way intersection cardinality ``captures'' $d$-way intersection
+cardinalities for $3 < d < n$ in a tree path labeling.
+
+\begin{lemma}
+  \label{lem:4-way}
+  Let $(\cF, \cl)$ be an \ICPPL and $\cl(S_i)=P_i$, $S_i \in \cF$, $1
+  \leq i \leq 4$.  Then, $|\cap_{i=1}^4 S_i| = |\cap_{i=1}^4 P_i|$.
+\end{lemma}
+\begin{proof}
+  Consider the sets $S_2 \cap S_1$, $S_3 \cap S_1$, and $S_4 \cap
+  S_1$, and their images $P_2 \cap P_1$, $P_3 \cap P_1$, and $P_4 \cap
+  P_1$ respectively.  It is clear that this path labeling preserves
+  pairwise intersection cardinalities.  In this case, it is clear that
+  the 3 sets are assigned paths in the path $P_1$.  Now by applying
+  Lemma~\ref{lem:intersection-cardinality}, it follows that
+  $|\cap_{i=1}^4 S_i| =| \cap_{i=1}^4 P_i|$.
+\end{proof}
+\begin{corollary} 
+  \label{corr:4-way} 
+  Let $(\cF, \cl)$ be an \ICPPL and $\cl(S_i)=P_i$, $S_i \in \cF$, $1
+  \leq i \leq 4$. If $P_i$, $1
+  \leq i \leq 4$ are such that $P_1 \cap P_2$, $P_1
+  \setminus P_2$, $P_2 \setminus P_1$ are also paths, then:
+  \begin{align*}
+   |(S_1 \setminus S_2) \cap S_3 \cap S_4| &= |(P_1 \setminus P_2) \cap P_3 \cap P_4|\\
+   |(S_2 \setminus S_1) \cap S_3 \cap S_4| &= |(P_2 \setminus P_1) \cap P_3 \cap P_4| 
+  \end{align*}
+\end{corollary}
+\begin{proof}
+  It is clear that $|(P_1 \setminus P_2) \cap P_3 \cap P_4| =
+  |\displaystyle \cap_{i=1}^4 P_i| - | P_2 \cap P_3 \cap P_4|$.  Since
+  we have an ICPPL, from Lemma~\ref{lem:4-way}, it follows that the
+  intersection cardinalities are preserved.
+\end{proof}
+\begin{lemma}
+  \label{lem:priv-card}
+  Let $(\cF, \cl)$ be an \ICPPL and $\cl(S) = P$ for some $S \in \cF$.  If sets
+  $S_{priv}$ and $P_{priv}$ contain only those elements of $S$ and $P$
+  that occur in no set other than $S$ and $P$, respectively, then
+  $|S_{priv}|=|P_{priv}|$.
+\end{lemma}
+\begin{proof}
+  Consider the set system $ \cF' = \{S_i \cap S \mid S_i \in \cF, i
+  \in [n]\}$ and path labeling $\cl'(S_i \cap S) = P_i \cap P$.  Since
+  $(\cF, \cl)$ is an \ICPPL, it can be easily verified that $(\cF',
+  \cl')$ is an ICPIA on path (interval) $P$.  Let $S_{two}$ and
+  $P_{two}$ be the elements of $S$ and $P$ which are in one {\em
+    other} set and path, respectively \ie $S_{two} = S \cap (\cup_{S_i
+    \ne S} S_i)$ and $P_{two}$ is defined analogously. It may be noted
+  that $S_{two} = \supt{\cF'}$ and $P_{two} = \supt{\cF'^{\cl'}}$.
+  From Lemma~\ref{lem:icpia} it follows that there is a bijection
+  $\phi$ from $S_{two}$ to $P_{two}$ such that for each $i$, the image
+  of $S_i \cap S$ under $\phi$ is $P_i \cap P$ (definition of
+  feasibility).  Moreover, this means $|S_{two}| = |P_{two}|$. By
+  definition, $S_{priv} = S \setminus S_{two}$ and $P_{priv} = P \setminus P_{two}$.
+  Thus $|S_{priv}| = |S \setminus S_{two}| = |P \setminus
+  P_{two}|=|P_{priv}|$.  
+\end{proof}
+
 Now we will prove the effectiveness of Algorithm~\ref{al:filterone}.
-Lemma~\ref{lem:feasible} proves that {\tt filter\_1} does not destroy
+Lemma~\ref{lem:feasible} proves that \filteri does not destroy
 the feasibility of the input, if it is indeed feasible. Following
 this, Lemma~\ref{lem:invar1} and Lemma~\ref{lem:noexit1} show that if
-the input is an ICPPL the result of {\tt filter\_1} is also an ICPPL
+the input is an ICPPL the result of \filteri is also an ICPPL
 and they both have the same set of feasibility hypergraph
@@
 \end{lemma}
-\begin{proof}\thesisspacing%[Proof of Lemma~\ref{lem:feasible}]
+\begin{proof}\thesisspacing
   We will prove this by mathematical induction on the number of
@@
 
-    \tnote[dire]{Due to the condition in line~\ref{ln:3waycheck}, this invariant is
-    ensured at the end of every iteration.}
+    This invariant is ensured at the end of every iteration due to
+    Lemma~\ref{lem:4-way} and Corollary~\ref{corr:4-way}.
    % \begin{enumerate} [{Case 3.}1:]
@@
     % \end{enumerate}
-  \end{enumerate} %\qed
-  % \vspace{-6mm}
+  \end{enumerate} 
 \end{proof}
@@
   % \end{enumerate}
-  In other words, {\tt filter\_1} outputs a filtered path labeling that
+  In other words, \filteri outputs a filtered path labeling that
   ``preserves''
@@
 
-As a result of {\tt filter\_1} each leaf $v$ in $T$ is such that there
+As a result of \filteri each leaf $v$ in $T$ is such that there
 is exactly one set in $\cF$ with $v$ as a vertex in the path assigned
-to it.  {\tt filter\_2}, in Algorithm~\ref{al:filtertwo}, identifies
+to it.  \filterii, in Algorithm~\ref{al:filtertwo}, identifies
 elements in $supp(\cF)$ whose images are leaves in a hypergraph
@@
 %   \label{lem:setminuscard}
-%   If $\cA$ is an ICPPA, and $(S_1, P_1),(S_2, P_2),(S_3, P_3) \in
+%   If $\cA$ is an ICPPL, and $(S_1, P_1),(S_2, P_2),(S_3, P_3) \in
 %   \cA$, then $|S_1 \cap (S_2 \setminus S_3)| = |P_1 \cap (P_2
@@
 %   |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to conditions (ii)
-%   and (iii) of ICPPA, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
+%   and (iii) of ICPPL, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1
 %   \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3|
@@
 %   The base case, $\Pi_0 = \{(S_i,P_i) \mid i \in [m]\}$, is trivially
-%   true since it is the input which is an ICPPA.  Assume the lemma is
+%   true since it is the input which is an ICPPL.  Assume the lemma is
 %   true till the $j-1$ iteration. Consider $j$th
@@
 \begin{algorithm}[h]
-  \caption{Leaf labeling from an ICPPL {\tt filter\_2($\cF, \cl, T$)}}
+  \caption{Leaf labeling from an ICPPL {\tt \filterii($\cF, \cl, T$)}}
   \label{al:filtertwo} % {leafasgn}
@@
 happen. Following that Lemma~\ref{lem:invar3} shows that given an
-ICPPL input, the filtered output path labeling after {\tt filter\_2}
+ICPPL input, the filtered output path labeling after \filterii
 is also an ICPPL.
@@
 %   % Moreover, $X$ as defined in the algorithm is non-empty if this is
-%   % an ICPPA.
+%   % an ICPPL.
 % \end{lemma}
@@
 %   First we see that $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in
-%     I}S_i$ is non empty in every iteration for an ICPPA. Suppose $X$
+%     I}S_i$ is non empty in every iteration for an ICPPL. Suppose $X$
 %   is empty. We know that $v \in P_{i_1} \setminus \bigcup_{i \ne i_1,
 %     i \in I}P_i$ since $v$ is in the unique path $P_{i_1}$. Since this
-%   is an ICPPA $|S_{i_1}| = |P_{i_1}|$. For any $x \in S_{i_1}$ it is
+%   is an ICPPL $|S_{i_1}| = |P_{i_1}|$. For any $x \in S_{i_1}$ it is
 %   contained in at least two sets due to our assumption. Let $S_{i_2}$
@@
 %   elements of $S_{i_2}$ to $P_{i_2}$. This contradicts our assumption
-%   that this is an ICPPA. Thus $X$ cannot be empty.
+%   that this is an ICPPL. Thus $X$ cannot be empty.
 
@@
 %   For $\Pi_0$ all invariants hold because it is output from algorithm
-%   \ref{al:filterone} which is an ICPPA. Hence base case is proved.  Assume
+%   \ref{al:filterone} which is an ICPPL. Hence base case is proved.  Assume
 %   the lemma holds for $\Pi_{j-1}$. Consider $\Pi_j$ and any $(P,Q) \in
@@
 %   \label{th:perm}
-%   If $\cA$ is an ICPPA, then there exists a bijection $\sigma : U
+%   If $\cA$ is an ICPPL, then there exists a bijection $\sigma : U
 %   \rightarrow V(T)$ such that $\sigma(S_i) = P_i$ for all $i \in I$
@@
 % \begin{proof}\thesisspacing
-%   This is a contructive proof. First, the given ICPPA $\cA$ and tree
+%   This is a contructive proof. First, the given ICPPL $\cA$ and tree
 %   $T$ are given as input to Algorithm \ref{al:filterone}. This yields a
-%   ``filtered'' ICPPA as the output which is input to Algorithm
+%   ``filtered'' ICPPL as the output which is input to Algorithm
 %   \ref{al:filtertwo}.  It can be observed that the output of Algorithm
@@
 %   some $d \le n$ which is just a path. From the last lemma we know
-%   that $\cA_{d-1}$ is an ICPPA. Another observation is that an ICPPA
+%   that $\cA_{d-1}$ is an ICPPL. Another observation is that an ICPPL
 %   with all its tree paths being intervals (subpaths from a path) is
@@
 \begin{proof}\thesisspacing
-  Assume $X$ is empty for some iteration $j > 0$. We know that $v$ is
-  an element of $\cl_{j-1}(S_1)$. Since it is uniquely present in
-  $\cl_{j-1}(S_1)$, it is clear that $v \in \cl_{j-1}(S_1) \setminus
-  \bigcup_{(S \in \cF_{j-1}) \wedge (S \ne S_1)}\cl_{j-1}(S)$.  Note
-  that for any $x \in S_1$ it is contained in at least two sets due to
-  our assumption about cardinality of $X$. Let $S_2 \in \cF_{j-1}$ be
-  another set that contains $x$. From the above argument, we know $v
-  \notin \cl_{j-1}(S_2)$. Therefore there cannot exist a hypergraph
-  isomorphism bijection that maps elements in $S_2$ to those in
-  $\cl_{j-1}(S_2)$. This contradicts our assumption that the input is
-  feasible. Thus $X$ cannot be empty if input is ICPPL and feasible.
+This directly follows from Lemma~\ref{lem:priv-card} since $X$ is
+nothing but $S_{1_{priv}}$.
+%   Assume $X$ is empty for some iteration $j > 0$. We know that $v$ is
+%   an element of $\cl_{j-1}(S_1)$. Since it is uniquely present in
+%   $\cl_{j-1}(S_1)$, it is clear that $v \in \cl_{j-1}(S_1) \setminus
+%   \bigcup_{(S \in \cF_{j-1}) \wedge (S \ne S_1)}\cl_{j-1}(S)$.  Note
+%   that for any $x \in S_1$ it is contained in at least two sets due to
+%   our assumption about cardinality of $X$. Let $S_2 \in \cF_{j-1}$ be
+%   another set that contains $x$. From the above argument, we know $v
+%   \notin \cl_{j-1}(S_2)$. Therefore there cannot exist a hypergraph
+%   isomorphism bijection that maps elements in $S_2$ to those in
+%   $\cl_{j-1}(S_2)$. This contradicts our assumption that the input is
+%   feasible. Thus $X$ cannot be empty if input is ICPPL and feasible.
   %\qed
@@
 \begin{proof}\thesisspacing
-  By Lemma~\ref{lem:xnotempty} we know that set $X$ will not be empty
-  in any iteration of the {\em \bf while} loop if input ICPPL $(\cF,
-  \cl)$ is feasible and $\cl_j$ is always computed for all $j >
-  0$. Also note that removing a leaf from any path keeps the new path
-  connected. Thus invariant I is obviously true. In every iteration $j
-  > 0$, we remove exactly one element $x$ from one set $S$ in $\cF$
-  and exactly one vertex $v$ which is a leaf from one path
-  $\cl_{j-1}(S)$ in $T$. This is because as seen in
-  Lemma~\ref{lem:xnotempty}, $x$ is exclusive to $S$ and $v$ is
-  exclusive to $\cl_{j-1}(S)$. Due to this fact, it is clear that the
-  intersection cardinality equations do not change, i.e., invariants
-  II, III, IV remain true. On the other hand, if the input ICPPL is
-  not feasible the invariants are vacuously true. %\qed
+  Following Lemma~\ref{lem:xnotempty}, we know that set $X$ will not
+  be empty in any iteration of the {\em \bf while} loop if input ICPPL
+  $(\cF, \cl)$ is feasible and thus $\cl_j$ is always computed for all
+  $j > 0$. Also note that removing a leaf from any path keeps the new
+  path connected. Thus invariant I is obviously true. In every
+  iteration $j > 0$, we remove exactly one element $x$ from one set
+  $S$ in $\cF$ and exactly one vertex $v$ which is a leaf from one
+  path $\cl_{j-1}(S)$ in $T$. This is because $x$ is exclusive to $S$
+  and $v$ is exclusive to $\cl_{j-1}(S)$. Due to this fact, it is
+  clear that the intersection cardinality equations do not change,
+  i.e., invariants II, III, IV remain true. % On the other hand, if the
+%   input ICPPL is not feasible the invariants are vacuously true.
+  % \qed
 \end{proof}
@@
 We have seen two filtering algorithms above, namely,
-Algorithm~\ref{al:filterone} for {\tt filter\_1} and
-Algorithm~\ref{al:filtertwo} for {\tt filter\_2} which when executed
+Algorithm~\ref{al:filterone} for \filteri and
+Algorithm~\ref{al:filtertwo} for \filterii which when executed
 in that order result in a new ICPPL on the same universe $U$ and
@@
 % \begin{lemma}{\temptext{PROOF TBD.}}
-%   Let $(\cF_2, \cl_2)$ be the output of {\tt filter\_2} and the ICPPL
-%   input to {\tt filter\_1} preceeding its call is $(\cF, \cl)$ with
+%   Let $(\cF_2, \cl_2)$ be the output of \filterii and the ICPPL
+%   input to \filteri preceeding its call is $(\cF, \cl)$ with
 %   target tree $T$. If a bijection $\phi:\supt{\cF_2} \rightarrow V(T)$
@@
 tree. In brief, it is done as follows. If the input is feasible, the
-output of {\tt filter\_1}, say $(\cF_1, \cl_1)$, gives a new ICPPL
+output of \filteri, say $(\cF_1, \cl_1)$, gives a new ICPPL
 which preserves all the original hypergraph isomorphisms.  This ICPPL
-$(\cF_1, \cl_1)$ is then given to {\tt filter\_2} which gives the
+$(\cF_1, \cl_1)$ is then given to \filterii which gives the
 second filtered ICPPL, say $(\cF_2, \cl_2)$ such that a feasible leaf
@@
     \STATE $L \assign \{v \mid v \text{ is a leaf in }      T\}$\\
-    \STATE $(\cF_1, \cl_1) \assign$ {\tt filter\_1($\cF, \cl,
+    \STATE $(\cF_1, \cl_1) \assign$ {\tt \filteri($\cF, \cl,
       T$)}\\
-    \STATE $(\cF_2, \cl_2) \assign$ {\tt filter\_2($\cF_1,
+    \STATE $(\cF_2, \cl_2) \assign$ {\tt \filterii($\cF_1,
       \cl_1, T$)}\\
@@
   T)$ as input returns a non-empty function, then $(\cF, \cl)$ is
-  feasible.  Now consider the case where $(\cF, \cl)$ is
-  feasible, \ie there exists a hypergraph isomorphism $\phi$ such
-  that $\cl_\phi = \cl$. Lemma~\ref{lem:noexit1} and
-  Lemma~\ref{lem:xnotempty} show us that {\tt filter\_1} and {\tt
-    filter\_2} do not exit if input is feasible. Thus
+  feasible.  Now consider the case where $(\cF, \cl)$ is feasible, \ie
+  there exists a hypergraph isomorphism $\phi$ such that $\cl_\phi =
+  \cl$. Lemma~\ref{lem:noexit1} and Lemma~\ref{lem:xnotempty} show us
+  that \filteri and \filterii do not exit if input is feasible. Thus
   Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
@@
 \section{Solution to \illustrationproblem}
+\label{sec:wallacesol}
 
@@
 
-\tnote{\sc refer to Problem Defs here}
 Section~\ref{sec:feasible} described properties that a TPL must have
@@
 tree\footnote{A larger problem would be to find if a given hypergraph
-  is a path graph on any tree. This problem is not addressed in this
-  thesis.}. In other words, the problem is to find out if a feasible
-tree path labeling exists from a given target tree for a given
-hypergraph. In this section we will see two special cases of this
-problem where the target tree is from a particular family of
-trees. The first one, where the tree is a path, is a shown to be
-equivalent to the well studied problem of consecutive-ones in
+  is a path hypergraph with respect to any tree. This problem is not
+  addressed in this thesis.}.  In other words, the problem is to find
+out if a feasible tree path labeling exists from a given target tree
+for a given hypergraph. In this section we will see two special cases
+of this problem where the target tree is from a particular family of
+trees. The first one, where the tree is a path as described by the
+\CFTPLINT problem (see Section~\ref{sec:myresearchintro}) is a shown
+to be equivalent to the well studied problem of consecutive-ones in
 Section~\ref{sec:icpplicpia}. The second one, where the tree is a
-$k$-subdivided tree\footnote{See Section~\ref{ch:prelims} for the
-  formal definition.}, has been solved using a polynomial time
-algorithm. The latter problem enforces some conditions on the
+$k$-subdivided tree as described by \CFTPLKTREE (see
+Section~\ref{sec:myresearchintro}), has been solved using a polynomial
+time algorithm. The latter problem enforces some conditions on the
 hypergraph too which will be seen in Section~\ref{sec:ksubdivstar}.
@@
 
-Let us consider a special case of ICPPL with the following properties.
-The target tree $T$ is a path.  Hence, all path labels are can be viewed as
-intervals assigned to the sets in $\cF$.  It is shown, in
-\cite{nsnrs09}, that the filtering algorithms outlined above need only
-preserve pairwise intersection cardinalities, and higher level
-intersection cardinalities are preserved by the Helly Property of
-intervals.  Consequently, the filter algorithms do not need to ever
-evaluate the additional check to {\em \bf exit}.
+In Section~\ref{sec:motive} it is described how \COP is a
+special case of TPL as described by the \CFTPLINT problem. In this
+section we will dwell on this in more detail.
+
+\annote[dire]{The special case of TPL where the target tree $T$ is a
+path. Hence, all path labels are can be viewed as intervals assigned
+to the sets in $\cF$. }{needs work}
+
+It is shown, in \cite{nsnrs09}, that the
+filtering algorithms outlined above need only preserve pairwise
+intersection cardinalities, and higher level intersection
+cardinalities are preserved by the Helly Property of intervals.
+% Consequently, the filter algorithms do not need to ever evaluate the
+% additional check to {\em \bf exit}.
 %\begin{enumerate}
@@
   \begin{tabular}{lr}
-%     (a) \includegraphics{cop-tree_thesis.1}&
-%     (b) \includegraphics{cop-tree_thesis.2}
    (a) \ksubstari & (b) \ksubstarii
@@
 In this section we consider the problem of assigning paths from a
-$k$-subdivided star $T$ to a given set system $\cF$.  We consider
+$k$-subdivided star $T$ to a given set system $\cF$ as described by
+the \CFTPLKTREE problem.  We consider
 $\cF$ for which the overlap graph $\bO(\cF)$ is connected.  The
@@
 
-\temptext{repetition}
+\tnote[dire]{REPEAT!!
 
@@
 all $R_i$.
-
-\subsection{Description of the Algorithm}
-\label{sec:ksubstaralgo}
+}
+
+\noindent {\bf Description of the Algorithm.}
 In this section the given hypergraph $\cF$, the $k$-subdivided star
@@
   at most size 2 ($k+2$) size and thus overlaps are at most 1. If two
-  paths share a leaf in {\tt filter\_1} one must be of length 2 and
-  the other of length 1. Thus the exit condition is not met. Further,
-  it is also clear that the exit condition in {\tt filter\_2} is also
-  not met. Thus claim proven for base case.  Now assume the claim to
-  be true when target tree is a $(k-1)$-subdivided star. Consider the
-  case of a $k$-subdivided star.  We can show that after {\tt
-    filter\_1} and one iteration of a modified {\tt filter\_2} {\em
-    all} leaves are assigned pre-images.  Removing the leaves from $T$
-  and the pre-images from support of $\cO$, results in an ICPPL to a
-  $(k-1)$-subdivided star.  Now we apply the induction hypothesis, and
-  we get an isomorphism between the hypergraphs $\cO$ and $\cO^\cl$.
+  paths share a leaf in \filteri one must be of length 2 and the other
+  of length 1. Thus the exit condition is not met. Further, it is also
+  clear that the exit condition in \filterii is also not met. Thus
+  claim proven for base case.  Now assume the claim to be true when
+  target tree is a $(k-1)$-subdivided star. Consider the case of a
+  $k$-subdivided star.  We can show that after \filteri and one
+  iteration of a modified \filterii {\em all} leaves are assigned
+  pre-images.  Removing the leaves from $T$ and the pre-images from
+  support of $\cO$, results in an ICPPL to a $(k-1)$-subdivided star.
+  Now we apply the induction hypothesis, and we get an isomorphism
+  between the hypergraphs $\cO$ and $\cO^\cl$.
 
@@
   again by induction on $k$ it follows that the algorithm finds $\cl$.
-
 \end{proof}
@@
 
-\section{ TPL with no restrictions}
+\section{ TPL on arbitrary trees}
 \label{sec:norestraint}
 
-{\tt abstract begin} it is known that if the given tree is a path a
-feasible assignment can be found in polynomial time, and we observe
-that it can actually be done in logspace. \tnote{[TRUE?]}  {\tt
-  abstract end}
-
-In Section \ref{prelims} we present the necessary preliminaries, in
-Section \ref{feasible} we present our characterization of feasible
-tree path assignments, and in Section \ref{sec:decompos} we present the
-characterizing subproblems for finding a bijection between $U$ and
-$V(T)$ such that sets map to tree paths. Finally, in Section
-\ref{sec:tplcomplexity} we conclude by showing that Tree Path Assignment is
-GI-Complete, and also observe that Consecutive Ones Testing is in
-Logspace.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC \subsection{Introduction}
-
-\annote{ We refer to this as the Tree Path Assignment problem for an
-  input $(\cF,T)$ pair.}{Add this ``pair'' as a definition in
-  ch:prelims}
-
-In the second part of this paper, we decompose our search for a
-bijection between $U$ and $V(T)$ into subproblems.  Each subproblem is
-on a set system in which for each set, there is another set in the set
-system with which the intersection is {\em strict}- there is a
-non-empty intersection, but neither is contained in the other.  This
-is in the spirit of results in \cite{wlh02,nsnrs09} where to test for
-COP in a given matrix, the COP problem is solved on an equivalent set
-of prime matrices.
+In this section we describe an algorithm for the \CFTPL problem where
+the target tree can be any arbitrary tree. In essence, we provide a
+characterization for path hypergraphs (see
+Section~\ref{sec:pathhypergraph} for definition).  We first study some
+properties of the overlap components of the input hypergraph and
+introduce some theory in that area. This is in the same vein as the
+theory used in \cite{wlh02,nsnrs09} but we extend it to \TPL.  We use
+this theory to decompose \CFTPL into subproblems.  Each subproblem is
+on a sub-hypergraph of the input, in which for each hyperedge there is
+another hyperedge in the sub-hypergraph that overlaps with it. These
+sub-hypergraphs are called overlap components (see
+Section~\ref{sec:overlapmarginal} for definition). As a consequence of
+this characterization, aside from the overlap component subproblems,
+we identify two other subproblems that must be solved to obtain an
+ICPPL.  The inefficiency of our algorithm in terms of polynomial time
+solvability comes from these two subproblems. We leave these subproblems
+open to be solved in \p and only mention the obvious brute force
+method of solution.
+
+% with which the intersection is {\em
+%   strict}- there is a non-empty intersection, but neither is contained
+% in the other.  % In
+% \cite{wlh02,nsnrs09} where to test for COP in a given matrix, the COP
+% problem is solved on an equivalent set of prime matrices.
+
+% \annote{ We refer to this as the Tree Path Assignment problem for an
+%   input $(\cF,T)$ pair.}{Add this ``pair'' as a definition in
+%   ch:prelims 0 - or should we?}
+
+% In the second part of this paper, we decompose our search for a
+% bijection between $U$ and $V(T)$ into subproblems.
 
 % Finally, we show that Tree Path Assignment is isomorphism-complete.
-We also point out Consecutive Ones Testing is in Logspace from two
-different results in the literature \cite{kklv10, mcc04}. To the best
-of our knowledge this observation has not been made earlier.
-
-\subsection{Finding an assignment of tree paths to a set
-  system}% 
-\label{sec:decompos}% 
-In the previous section we have shown that
-the problem of finding a Tree Path Asssignment to an input $(\cF,T)$
-is equivalent to finding an ICPPA to $\cF$ in tree $T$.  In this
-section we characterize those set systems that have an ICPPA in a
-given tree.  As a consequence of this characterization we identify two
-sub-problems that must be solved to obtain an ICPPA.  We do not solve
-the problem and in the next section show that finding an ICPPA in a
-given tree is GI-Complete.
-
+
+%\noindent {\bf Finding an assignment of tree paths to a set system.} 
+As described by Definition~\ref{def:matrixsetsystem} a set system or
+hypergraph can be concisely represented by a binary matrix
+\stt the row indices of the matrix denote the vertex set of the
+hypergraph and each column bijectively correspond to a hyperedge. And
+thus the ideas of feasible \TPL and \ICPPL are also defined for a
+binary matrix. If a \TPL $(\cF,\cl,T)$ is feasible, \ie it is an ICPPL
+(see Section~\ref{xxxxx} \tnote[dire]{SECTION NUMBER}) and if $M$ is
+the representative matrix for hypergraph $\cF$, then we say that $M$
+has an \ICPPL and vice versa.
+
+Consider the overlap graph and overlap components of $\bO(\cF)$
+defined in Section~\ref{sec:overlapmarginal}.  We use this to
+decompose $M$ as described in \cite{wlh02,nsnrs09}.  A {\em prime
+  submatrix} of $M$ is defined as the matrix formed by a set of
+columns of $M$ which correspond to an overlap component.  Let us
+denote the prime submatrices of $M$ by $M_1,\ldots,M_p$, each
+corresponding to one of the $p$ overlap components of $\cF$. Clearly,
+two distinct prime submatrices have mutually exclusive sets of columns
+of $M$.  Let $col(M_i)$ be the set of columns in the submatrix $M_i$.
+The support of a prime submatrix $M_i$ and support of a
+set of prime submatrices $X$ are defined as follows. Note that for each $i$,
+$supp(M_i) \subseteq U$ where $U$ is the vertex set of hypergraph $\cF$.
+\begin{align*}
+supp(M_i) &= \displaystyle \bigcup_{j \in col(M_i)}S_j\\  
+supp(X) &= \displaystyle  \bigcup_{M \in X} supp(M)
+\end{align*}
+
+% A
+% set system can be concisely represented by a binary matrix where the
+% row indices denote the universe of the set system and the column
+% indices denote each of the sets. Let the binary matrix be $M$ with
+% order $n \times m$, the set system be $\cF = \{S_i \mid i \in [m]\}$,
+% universe of set system $U = \{x_1, \dots ,x_n\}$. If $M$ represents
+% $\cF$, $|U| = n, |\cF| = m$. Thus $(i,j)$th element of $M$, $M_{ij} =
+% 1$ \iff $x_i \in S_j$.
+
+% If $\cF$ has a feasible tree path assignment
+% (ICPPL) $\cA = \{(S_i,P_i) \mid i \in [m]\}$, then we say its
+% corresponding matrix $M$ has an ICPPL. Conversly we say that a matrix
+% $M$ has an ICPPL if there exists an ICPPL $\cA$ as defined
+% above.\\
+%\noindent
+
+% We now define the strict intersection graph or overlap graph of
+% $\cF$. This graph occurs at many places in the literature, see for
+% example \cite{kklv10, wlh02, nsnrs09}.  The vertices of the graph
+% correspond to the sets in $\cF$.  An edge is present between vertices
+% of two sets \iff the corresponding sets have a nonempty intersection
+% and none is contained in the other. Formally, intersection graph is
+% $G_f = (V_f, E_f)$ such that $V_f = \{v_i \mid S_i \in \cF\}$ and $E_f
+% = \{(v_i, v_j) \mid S_i \cap S_j \ne \O \text{ and }S_i \nsubseteq
+% S_j, S_j \nsubseteq S_i \}$.
+
+\def \primeless {\preccurlyeq}
 \noindent
-A set system can be concisely represented by a binary matrix where the
-row indices denote the universe of the set system and the column
-indices denote each of the sets. Let the binary matrix be $M$ with
-order $n \times m$, the set system be $\cF = \{S_i \mid i \in [m]\}$,
-universe of set system $U = \{x_1, \dots ,x_n\}$. If $M$ represents
-$\cF$, $|U| = n, |\cF| = m$. Thus $(i,j)$th element of $M$, $M_{ij} =
-1$ \iff $x_i \in S_j$. If $\cF$ has a feasible tree path assignment
-(ICPPA) $\cA = \{(S_i,P_i) \mid i \in [m]\}$, then we say its
-corresponding matrix $M$ has an ICPPA. Conversly we say that a matrix
-$M$ has an ICPPA if there exists an ICPPA $\cA$ as defined
-above.\\
-\noindent
-We now define the strict intersection graph or overlap graph of
-$\cF$. This graph occurs at many places in the literature, see for
-example \cite{kklv10, wlh02, nsnrs09}.  The vertices of the graph
-correspond to the sets in $\cF$.  An edge is present between vertices
-of two sets \iff the corresponding sets have a nonempty intersection
-and none is contained in the other. Formally, intersection graph is
-$G_f = (V_f, E_f)$ such that $V_f = \{v_i \mid S_i \in \cF\}$ and $E_f
-= \{(v_i, v_j) \mid S_i \cap S_j \ne \O \text{ and }S_i \nsubseteq
-S_j, S_j \nsubseteq S_i \}$.  We use this graph to decompose $M$ as
-described in \cite{wlh02,nsnrs09}.  A prime sub-matrix of $M$ is
-defined as the matrix formed by a set of columns of $M$ which
-correspond to a connected component of the graph $G_f$.  Let us denote
-the prime sub-matrices by $M_1,\ldots,M_p$ each corresponding to one
-of the $p$ components of $G_f$. Clearly, two distinct matrices have a
-distinct set of columns.  Let $col(M_i)$ be the set of columns in the
-sub-matrix $M_i$.  The support of a prime sub-matrix $M_i$ is defined
-as $supp(M_i) = \displaystyle \bigcup_{j \in col(M_i)}S_j$.  Note that
-for each $i$, $supp(M_i) \subseteq U$.  For a set of prime
-sub-matrices $X$ we define
-$supp(X) = \displaystyle \bigcup_{M \in X} supp(M)$. \\
-
-\noindent
-Consider the relation $\preccurlyeq$ on the prime sub-matrices $M_1,
-\ldots, M_p$ defined as follows:
-\begin{equation}
-  \nonumber \{(M_i,M_j) | \mbox{ A set } S \in
-  M_i \mbox{ is contained in a set } S' \in M_j\} \cup \{(M_i,M_i) | 1
-  \leq i \leq p\} 
-\end{equation}
-
-\noindent
+Consider the binary relation $\primeless$ defined on the set of
+prime submatrices $\bP = \{M_i \mid i \in [p]\}$ as follows.
+\begin{align*}
+  \primeless \ = \ &\{(M_i,M_j) \mid \exists S \in
+  M_i \text{ and } \exists S' \in M_j \text{ \stt\ } S \subseteq S'\} \\
+  \cup \  &\{(M_i,M_i) \mid i \in [p]\}
+\end{align*}
+
+% $  \primeless = \{(M_i,M_j) \mid \text{ a set } S \in
+%   M_i \text{ is contained in a set } S' \in M_j\} \cup \{(M_i,M_i) | 1
+%   \leq i \leq p\} $
+
 This relation is the same as that defined in \cite{nsnrs09}. The prime
-submatrices and the above relation can be defined for any set
-system. We will use this structure of prime submatrices to present our
-results on an an ICPPA for a set system $\cF$. Recall the following
-lemmas, and theorem that $\preccurlyeq$ is a partial order, from
-\cite{nsnrs09}.
-
-\begin{lemma} \label{lem:containment} Let $(M_i,M_j) \in
-  \preccurlyeq$.  Then there is a set $S' \in M_j$ such that for each
-  $S \in M_i$, $S \subseteq S'$.
+submatrices and the above relation can be defined for any hypergraph
+since an overlap graph is defined for any hypergraph. We will use this structure of prime submatrices to present our
+results on an ICPPL for a hypergraph $\cF$. 
+Now we present a few lemma and a theorem that show that $\primeless$ is
+a partial order. An alternate notation for $(M_i,M_j) \in \primeless$
+is $M_i \primeless M_j$.
+
+% Recall the following
+% lemmas, and theorem that $\preccurlyeq$ is a partial order, from
+% \cite{nsnrs09}.
+
+\begin{lemma}[{\cite[Lem.~3]{nsnrs09}}] %
+  \label{lem:containment} If % $(M_i,M_j) \in
+%   \preccurlyeq$
+  $M_i \primeless M_j$, then there is a set $S' \in M_j$ such that for
+  each $S \in M_i$, $S \subseteq S'$.
 \end{lemma}
-\begin{lemma}
-  For each pair of prime sub-matrices, either $(M_i,M_j) \not\in
-  \preccurlyeq$ or $(M_j,M_i) \not\in \preccurlyeq$.
-  % If $(M_i,M_j) \in \preccurlyeq$ and $(M_j,M_i) \in \preccurlyeq$,
-  % then $i = j$ and $|M_i| = 1$.
+\begin{lemma}[{\cite[Lem.~4]{nsnrs09}}] %
+  For each pair of prime submatrices $M_i, M_j$, either $M_i \not\primeless M_j$
+  or $M_j \not\primeless M_i$.
+% $(M_i,M_j) \not\in
+%   \preccurlyeq$ or $(M_j,M_i) \not\in \preccurlyeq$.
 \end{lemma}
-\begin{lemma}
-  If $(M_i,M_j) \in \preccurlyeq $ and $(M_j,M_k) \in \preccurlyeq$,
-  then $(M_i,M_k) \in \preccurlyeq$.
+\begin{lemma}[{\cite[Lem.~5]{nsnrs09}}] %
+  If $M_i \primeless M_j$ and $M_i \primeless M_k$,
+  then $M_i \primeless M_k$.
+%   If $(M_i,M_j) \in \preccurlyeq $ and $(M_j,M_k) \in \preccurlyeq$,
+%   then $(M_i,M_k) \in \preccurlyeq$.
 \end{lemma}
-\begin{lemma}
-  If $(M_i,M_j) \in \preccurlyeq$ and $(M_i,M_k) \in \preccurlyeq$,
-  then either $(M_j,M_k) \in \preccurlyeq$ or $(M_k,M_j) \in
-  \preccurlyeq$.
+\begin{lemma}[{\cite[Lem.~6]{nsnrs09}}] %
+  If $M_i \primeless M_j$ and $M_i \primeless M_k$,
+  then either $M_j \primeless M_k$ or $M_k \primeless M_j$.
+%   If $(M_i,M_j) \in \preccurlyeq$ and $(M_i,M_k) \in \preccurlyeq$,
+%   then either $(M_j,M_k) \in \preccurlyeq$ or $(M_k,M_j) \in
+%   \preccurlyeq$.
 \end{lemma}
-\begin{theorem} \label{thm:partitionold} $\preccurlyeq$ is a partial
-  order on the set of prime sub-matrices of $M$.  Further, it uniquely
-  partitions the prime sub-matrices of $M$ such that on each set in
-  the partition $\preccurlyeq$ induces a total order.
+
+\begin{theorem} [{\cite[Th.~4]{nsnrs09}}] %
+  \label{thm:partitionold} %
+  Let $M$ be a binary matrix and $\bP$ be the set of prime submatrices
+  of $M$. Then the following hold true.
+  \parindent \defindent
+  \begin{enumerate}[i. ]
+  \item $\primeless$ is a partial order on $\bP$.
+  \item $\primeless$ uniquely partitions $\bP$ into $\bP_1, \bP_2
+    \ldots ,\bP_r$ for some $r \ge 1$.
+  \item \label{thm:i:totalorder} If $M$ has \COP, then for each $i \in
+    [r]$, $\primeless$ induces a total order in $\bP_i$.
+  \end{enumerate}
+% $\preccurlyeq$ is a partial
+%   order on the set of prime submatrices of $M$.  Further, it uniquely
+%   partitions the prime submatrices of $M$ such that on each set in
+%   the partition $\preccurlyeq$ induces a total order.
 \end{theorem}
-For the purposes of this paper, we refine the total order mentioned in
-Theorem \ref{thm:partitionold}. We do this by identifying an in-tree
-rooted at each maximal upper bound under $\preccurlyeq$.  Each of
-these in-trees will be on disjoint vertex sets, which in this case
-would be disjoint sets of prime-submatrices.  The in-trees are
-specified by selecting the appropriate edges from the Hasse diagram
-associated with $\preccurlyeq$.  Let $\cI$ be the following set:
+
+% \begin{theorem} % [{\cite[Th.~4]{nsnrs09}}] %
+%   \label{thm:partitionnew}%
+%   \parindent \defindent
+%   \begin{enumerate}[i. ]
+%   \item $\primeless$ is a partial order on $\bP$.
+%   \item $\primeless$ uniquely partitions $\bP$ into $\bP_1, \bP_2
+%     \ldots ,\bP_k$ for some $k \ge 1$.
+%   \item If $M$ has \COP, then for each $i \in [k]$, $\primeless$
+%     induces a total order in $\bP_i$.
+%   \end{enumerate}
+% \end{theorem}
+
+%   Let $M$ be a binary matrix for a hypergraph
+%   with \TPL and $\bP$ be the set of prime submatrices of $M$. Then
+%   Theorem~\ref{thm:partitionold} holds true with (iii) changed as
+%   follows -- for each $i \in [k]$, $\primeless$ induces an in-tree
+%   in $\bP_i$.
+Since we are concerned with the generalization of \COP, we generalize
+the condition~(\ref{thm:i:totalorder}) in
+Theorem~\ref{thm:partitionold} from total order to in-trees as
+follows.  The in-tree is derived from the Hasse diagram of
+$\primeless$. Each of these in-trees will be on disjoint vertex sets,
+which in this case would be disjoint sets of prime-submatrices
+$\bP_i$, $i \in [r]$ for some $r > 0$.  For each $\bP_i$, the in-tree
+is rooted at a maximal upper bound under $\preccurlyeq$.  The in-trees
+are specified by selecting the appropriate edges from the Hasse
+diagram associated with $\preccurlyeq$. This is defined in
+Theorem~\ref{thm:partition} before which we see the covering relation
+for $\primeless$ defined below.
+% %D replace \cI with \primeless_c
+% Let $\primeless_c$ be the following set:
 \begin{align*}
-  \cI = \{ (M_i,M_j) \in \preccurlyeq \mid \nexists M_k s.t. M_i
-  \preccurlyeq M_k, M_k \preccurlyeq M_j \} \cup \{ (M_i,M_i), i \in
-  [p] \}
+  \primeless_c \ \  = \ \ &\{ (M_i,M_j) \mid M_i \primeless M_j \text{ \stt\ } \nexists
+  M_k \text{ s.t. } M_i \primeless M_k \text{, } M_k \primeless M_j \} \\
+  \cup \ \ &\{ (M_i,M_i), i \in [p] \}
 \end{align*}
 
-\begin{theorem} \label{thm:partition} Consider the directed graph $X$
-  whose vertices correspond to the prime sub-matrices, and the edges
-  are given by $\cI$.  Then, $X$ is a vertex disjoint collection of
-  in-trees and the root of each in-tree is a maximal upper bound in
-  $\preccurlyeq$.
+It is easy to see that $\primeless_c$ is the covering relation of the partial
+order $\primeless$.
+
+\begin{theorem}%
+  \label{thm:partition} %
+  Let $M$ be a binary matrix and $\bP$ be the set of prime matrices of
+  $M$. The partial order $\primeless$ and its covering relation
+  $\primeless_c$ are as defined earlier.
+  Then the directed graph $X = (\bP, \primeless_c)$ whose vertex
+  set is the set of prime submatrices and the edges are given
+  by the covering relation $\primeless_c$, is a vertex
+  disjoint collection of in-trees and the root of each in-tree is a
+  maximal upper bound in $\primeless$.
 \end{theorem}
@@
   To observe that $X$ is a collection of in-trees, we observe that for
-  vertices corresponding to maximal upper bounds, no out-going edge is
-  present in $I$.  Secondly, for each other element, exactly one
-  out-going edge is chosen, and for the minimal lower bound, there is
-  no in-coming edge.  Consequently, $X$ is acyclic, and since each
+  vertices corresponding to maximal upper bounds, no outgoing edge is
+  present in $\primeless_c$.  Secondly, for each other element, exactly one
+  outgoing edge is chosen, and for the minimal lower bound, there is
+  no incoming edge.  Consequently, $X$ is acyclic, and since each
   vertex has at most one edge leaving it, it follows that $X$ is a
   collection of in-trees, and for each in-tree, the root is a maximal
-  upper bound in $\preccurlyeq$.  Hence the theorem.
+  upper bound in $\primeless$.  
 \end{proof}
@@
   let $\{X_1,\ldots,X_r\}$ be the partition of $X$ into in-trees as
-  defined above.  Then, matrix $M$ has an ICPPA in tree $T$ \iff $T$
+  defined above.  Then, matrix $M$ has an ICPPL in tree $T$ \iff $T$
   can be partitioned into vertex disjoint subtrees $\{T_1, T_2, \dots
   T_r\}$ such that, for each $1 \leq i \leq r$, the set of prime
-  sub-matrices corresponding to vertices in $X_i$ has an ICPPA in
+  submatrices corresponding to vertices in $X_i$ has an ICPPL in
   $T_i$.
@@
   can be partitioned into $T_1, \ldots, T_r$ such that for each $1
-  \leq i \leq r $, the set of prime sub-matrices corresponding to
-  vertices in $X_i$ has an ICPPA in $T_i$.  It is clear from the
-  properties of the partial order $\preccurlyeq$ that these ICPPAs
-  naturally yield an ICPPA of $M$ in $T$.  The main property used in
+  \leq i \leq r $, the set of prime submatrices corresponding to
+  vertices in $X_i$ has an ICPPL in $T_i$.  It is clear from the
+  properties of the partial order $\preccurlyeq$ that these ICPPLs
+  naturally yield an ICPPL of $M$ in $T$.  The main property used in
   this inference is that for each $1 \leq i \neq j \leq r$, $supp(X_i)
@@
 \noindent
-To prove the forward direction, we show that if $M$ has an ICPPA, say
+To prove the forward direction, we show that if $M$ has an ICPPL, say
 $\cA$, in $T$, then there exists a partition of $T$ into vertex
 disjoint subtree $T_1, \ldots, T_r$ such that for each $1 \leq i \leq
-r$, the set of prime sub-matrices corresponding to vertices in $X_i$
-has an ICPPA in $T_i$.  For each $1 \leq i \leq r$, we define based on
+r$, the set of prime submatrices corresponding to vertices in $X_i$
+has an ICPPL in $T_i$.  For each $1 \leq i \leq r$, we define based on
 $\cA$ a subtree $T_i$ corresponding to $X_i$.  We then argue that the
 trees thus defined are vertex disjoint, and complete the proof.
-Consider $X_i$ and consider the prime sub-matrix in $X_{i,0}$.
+Consider $X_i$ and consider the prime submatrix in $X_{i,0}$.
 Consider the paths assigned under $\cA$ to the sets in the prime
-sub-matrix in $X_{i,0}$.  Since the component in $G_f$ corresponding
+submatrix in $X_{i,0}$.  Since the component in $G_f$ corresponding
 to this matrix is a connected component, it follows that union of
@@
 this sub-tree $T_i$.  All other prime-submatrices in $X_i$ are
-assigned paths in $T_i$ since $\cA$ is an ICPPA, and the support of
-other prime sub-matrices in $X_i$ are contained in the support of the
+assigned paths in $T_i$ since $\cA$ is an ICPPL, and the support of
+other prime submatrices in $X_i$ are contained in the support of the
 matrix in $X_{i,0}$.  Secondly, for each $1 \leq i \neq j \leq r$,
-$supp(X_i) \cap supp(X_j) = \phi$, and since $\cA$ is an ICPPA, it
+$supp(X_i) \cap supp(X_j) = \phi$, and since $\cA$ is an ICPPL, it
 follows that $T_i$ and $T_j$ are vertex disjoint.  Finally, since $|U|
@@
 into vertex disjoint sub-trees such that for each $1 \leq i \leq r$,
-the set of matrices corresponding to nodes in $X_i$ has an ICPPA in
+the set of matrices corresponding to nodes in $X_i$ has an ICPPL in
 $T_i$.  Hence the lemma.
 \end{proof}
-The essence of the following lemma is that an ICPPA only needs to be
-assigned to the prime sub-matrix corresponding to the root of each
-in-tree, and all the other prime sub-matrices only need to have an
+The essence of the following lemma is that an ICPPL only needs to be
+assigned to the prime submatrix corresponding to the root of each
+in-tree, and all the other prime submatrices only need to have an
 Intersection Cardinality Preserving Interval Assignments (ICPIA).
@@
 necessary and sufficient condition for a matrix to have COP.  We
-present the pseudo-code to test if $M$ has an ICPPA in $T$.
+present the pseudo-code to test if $M$ has an ICPPL in $T$.
 \begin{lemma} \label{lem:rooticppa} Let $M$ be a matrix and let $X$ be
@@
   \leq i \leq r$, the set of matrices corresponding to vertices of
-  $X_i$ has an ICPPA in $T_i$ if and only if the matrix in $X_{i,0}$
-  has an ICPPA in $T_i$ and all other matrices in $X_i$ have an {\bf
+  $X_i$ has an ICPPL in $T_i$ if and only if the matrix in $X_{i,0}$
+  has an ICPPL in $T_i$ and all other matrices in $X_i$ have an {\bf
     {\em ICPIA}}.
@@
   in-trees.  Each edge in the in-tree is a containment relationship
-  among the supports of the corresonding sub-matrices. Therefore, any
-  ICPPA to a prime sub-matrix that is not the root is contained in a
+  among the supports of the corresonding submatrices. Therefore, any
+  ICPPL to a prime submatrix that is not the root is contained in a
   path assigned to the sets in the parent matrix.  Consequently, any
-  ICPPA to the prime sub-matrix that is not at the root is an ICPIA,
-  and any ICPIA can be used to construct an ICPPA to the matrices
+  ICPPL to the prime submatrix that is not at the root is an ICPIA,
+  and any ICPIA can be used to construct an ICPPL to the matrices
   corresponding to nodes in $X_i$ provided the matrix in the root has
-  an ICPPA in $T_i$.  Hence the lemma.
+  an ICPPL in $T_i$.  Hence the lemma.
 \end{proof}
 Lemma \ref{lem:subicppa} and Lemma \ref{lem:rooticppa} point out two
-algorithmic challenges in finding an ICPPA for a given set system
+algorithmic challenges in finding an ICPPL for a given set system
 $\cF$ in a tree $T$.  Given $\cF$, finding $X$ and its partition
@@
 for each $i$, the set of matrices corresponding to nodes in $X_i$ have
-an ICPPA in $T_i$.  This seems to be a challenging step, and it must
+an ICPPL in $T_i$.  This seems to be a challenging step, and it must
 be remarked that this step is easy when $T$ itself is a path, as each
 individual $T_i$ would be sub-paths.  The second algorithmic challenge
-is identified by lemma \ref{lem:rooticppa} which is to assign an ICPPA
+is identified by lemma \ref{lem:rooticppa} which is to assign an ICPPL
 from a given tree to the matrix associated with the root node of
@@
 \begin{algorithm}[h]
-  \caption{Algorithm to find an ICPPA for a matrix $M$ on tree $T$:
-    $main\_ICPPA(M, T$)}
+  \caption{Algorithm to find an ICPPL for a matrix $M$ on tree $T$:
+    $main\_ICPPL(M, T$)}
   \label{al:icppa-main}
   \begin{algorithmic}
-    \STATE Identify the prime sub-matrices. This is done by constructing the strict overlap graph and identify connected components.  Each connected component yields a prime sub-matrix.   \\
-    \STATE Construct the partial order $\preccurlyeq$ on the set of prime sub-matrices.  \\
+    \STATE Identify the prime submatrices. This is done by
+    constructing the strict overlap graph and identify connected
+    components.  Each connected component yields a prime submatrix.
+    \\ 
+    \STATE Construct the partial order $\preccurlyeq$ on the set of
+    prime submatrices.  \\ 
     \STATE Construct the partition $X_1,\ldots,X_r$ of the prime
-    sub-matrices induced by $\preccurlyeq$ \\
+    submatrices induced by $\preccurlyeq$ \\
     \STATE For each $1 \leq i \leq r$, Check if all matrices except
@@
     use the result in \cite{nsnrs09}.  \STATE Find a partition of
-    $T_1, \ldots, T_r$ such that matrices in $X_{i,0}$ has an ICPPA in
+    $T_1, \ldots, T_r$ such that matrices in $X_{i,0}$ has an ICPPL in
     $T_i$.  If not such partition exists, exit with negative answer.
@@
 in $\cF$ is a path in $T$.
+
+\tnote[dire]{it is known that if the given tree is a path a
+feasible assignment can be found in polynomial time, and we observe
+that it can actually be done in logspace. }
+
+We also point out Consecutive Ones Testing is in Logspace from two
+different results in the literature \cite{kklv10, mcc04}. To the best
+of our knowledge this observation has not been made earlier.
 
@@
 have COP.  Similarly, the approach of testing for an ICPIA in
-\cite{nsnrs09} also gives a different certificate- a prime sub-matrix
+\cite{nsnrs09} also gives a different certificate- a prime submatrix
 that does not have an ICPIA.  Further, the above theorem can be used
@@
 % $\cF$. We call $\cA$ an {\em Intersection Cardinality Preserving
-%   Path Assignment (ICPPA)} if it has the following properties.
+%   Path Assignment (ICPPL)} if it has the following properties.
 
@@
 %   |S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to conditions (ii)
-%   and (iii) of ICPPA, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| =
+%   and (iii) of ICPPL, $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| =
 %   |P_1 \cap P_2| - |P_1 \cap P_2 \cap P_3| = |(P_1 \cap P_2)
@@
 % In the remaining part of this section we show that a path assignment
-% is feasible if and only if it is an ICPPA.  One direction of this
+% is feasible if and only if it is an ICPPL.  One direction of this
 % claim is clear: that is a path assignment is feasible, then all
 % intersection cardinalities are preserved, that is the path
-% assignment is an ICPPA.  The reason is that a feasible path
+% assignment is an ICPPL.  The reason is that a feasible path
 % assignment has an associated bijection between $U$ and $V(T)$ such
@@
 % constructively proving that it is sufficient, for the existence of a
-% bijection, for a path assignment to be an ICPPA.  At a top-level,
+% bijection, for a path assignment to be an ICPPL.  At a top-level,
 % the constructive approaches refine the path assignment iteratively,
@@
 % \begin{algorithm}[h]
-%   \caption{Permutations from an ICPPA $\{(S_i,P_i) | i \in I\}$}
+%   \caption{Permutations from an ICPPL $\{(S_i,P_i) | i \in I\}$}
 %   \label{al:filterone}
@@
 %   The base case, $\Pi_0 = \{(S_i,P_i) \mid i \in [m]\}$, is
-%   trivially true since it is the input which is an ICPPA.  Assume
+%   trivially true since it is the input which is an ICPPL.  Assume
 %   the lemma is true till the $j-1$ iteration. Consider $j$th
@@
 % \begin{algorithm}[h]
-%   \caption{Leaf assignments from an ICPPA $\{(S_i,P_i) | i \in I\}$}
+%   \caption{Leaf assignments from an ICPPL $\{(S_i,P_i) | i \in I\}$}
 %   \label{al:filtertwo}
@@
 %   %   is
-%   %   an ICPPA.
+%   %   an ICPPL.
 % \end{lemma}
@@
 %   First we see that $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in
-%     I}S_i$ is non empty in every iteration for an ICPPA. Suppose $X$
+%     I}S_i$ is non empty in every iteration for an ICPPL. Suppose $X$
 %   is empty. We know that $v \in P_{i_1} \setminus \bigcup_{i \ne
 %     i_1, i \in I}P_i$ since $v$ is in the unique path
-%   $P_{i_1}$. Since this is an ICPPA $|S_{i_1}| = |P_{i_1}|$. For any
+%   $P_{i_1}$. Since this is an ICPPL $|S_{i_1}| = |P_{i_1}|$. For any
 %   $x \in S_{i_1}$ it is contained in at least two sets due to our
@@
 %   permutation that maps elements of $S_{i_2}$ to $P_{i_2}$. This
-%   contradicts our assumption that this is an ICPPA. Thus $X$ cannot
+%   contradicts our assumption that this is an ICPPL. Thus $X$ cannot
 %   be empty.
@@
 %   For $\Pi_0$ all invariants hold because it is output from
-%   algorithm \ref{al:filterone} which is an ICPPA. Hence base case is
+%   algorithm \ref{al:filterone} which is an ICPPL. Hence base case is
 %   proved.  Assume the lemma holds for $\Pi_{j-1}$. Consider $\Pi_j$
@@
 % \begin{proof}
-%   This is a contructive proof. First, the given ICPPA $\cA$ and tree
+%   This is a contructive proof. First, the given ICPPL $\cA$ and tree
 %   $T$ are given as input to Algorithm \ref{al:filterone}. This yields a
-%   ``filtered'' ICPPA as the output which is input to Algorithm
+%   ``filtered'' ICPPL as the output which is input to Algorithm
 %   \ref{al:filtertwo}.  It can be observed that the output of Algorithm
@@
 %   which is just a path. From the last lemma we know that $\cA_{d-1}$
-%   is an ICPPA. Another observation is that an ICPPA with all its
+%   is an ICPPL. Another observation is that an ICPPL with all its
 %   tree paths being intervals (subpaths from a path) is nothing but

### 2012-02-24 src/003abstract.tex
- a whole bunch of cleanups. see diff. - removed problem defs from chap 1. they were duplicates from chap 3. - CODED FIGURES IN TIKZ. forbidden subgraphs, graphmatrix example graphs. - BEGAN CHAPTER 3 EDITS.
Formatting/layout changes
-------

- Added FrenchLayout to make lists denser
- Added hanging indent for Definitions
- Compressed theorem env's above below spaces
- Creaded env for problem definition minipages.
- moved all clearpage/page counter and other prechapter settings to the chapter tex files.
- Fixed the page number bug in draft printout. removed \fancyhf{} to avoid resetting page n

@@
+\xclearpage
+
 \begin{abstract}
@@
   \noindent
-  Consecutive-ones property is a non-trivial property of binary
-  matrices that has been studied widely in the literature for over
-  past 50 years. Detection of COP in a matrix is possible efficiently
-  and there are several algorithms that achieve the same. This thesis
-  documents the work done on an extension of COP extended from the
-  equivalent interval assignment problem in \cite{nsnrs09}. These new
-  results rigorously prove a natural extension (to trees) of their
-  characterization as well as makes connections to graph isomorphism,
-  namely path graph isomorphism.
+  Consecutive-ones property is a non-trivial combinatorial property of
+  binary matrices that has been studied widely in the literature for
+  over past 50 years. Detection of COP in a matrix is possible
+  efficiently and there are several algorithms that achieve the
+  same. This thesis documents the work done on an extension of COP.
+  We extend \COP from the equivalent interval assignment problem to what we
+  call {\em Tree Path Labeling (TPL)} problem. Inspired by the idea
+  that \COP is akin to assigning intervals to sets in a set system, we
+  ask the question what if the assignments to sets are made not of
+  intervals but of paths from any tree?  While we show that TPL is in
+  NP, we obtain a more efficient algorithm for a special case tree
+  class called {\em $k$-subdivided stars} as well as for
+  characterizing a feasible TPL. For extension of \COP to
+  arbitrary trees, we analyse the structure of a hypergraph (set
+  system / columns of matrix) using the idea of prime submatrices and
+  yield an algorithm for a feasible path labeling if one indeed
+  exists.  These new results rigorously prove the natural extension
+  (to trees) of the ICPIA characterization as well as makes
+  connections to graph isomorphism, namely path graph isomorphism.
 
-  \tnote{EXPAND.  Abstract must be a brief about what results we have
-    and how it fits
-    in the body of research.\\
-    -- Area: Broad to Specialized. i.e. Combinatorial algorithms ->
-    Matrix reorganization -> general data reorganization
-    (interval assignment) -> path assignment\\
-    -- Class of problems: say, data reorganization.\\
-    -- Nature of results: Is a generalization. We have a
-    Polynomial algorithm for a subset of the generalization.\\
-    % maybe add - we explore a natural generalization of results on
-    % binary matrices with the |\tem consecutive-ones property|.  We
-    % consider the following constraint satisfaction problem. Given
-    % (i) a set system $\F \subseteq$ $(2^{U} \setminus \emptyset)$ of
-    % a finite set $U$ of cardinality $n$, (ii) a tree $T$ of size $n$
-    % and (iii) a bijection called |\tem tree path labeling|, $\cl$
-    % mapping the sets in $\cF$ to paths in $T$, does there exist at
-    % least one bijection $\phi:U \rightarrow V(T)$ such that for each
-    % $S \in \cF$, $\{\phi(x) \mid x \in S\} = \cl(S)$?  A tree path
-    % labeling of a set system is called |\tem feasible| if there
-    % exists such a bijection $\phi$.  We present an algorithmic
-    % characterization of feasible tree path labeling. COP is a
-    % special instance of tree path labeling problem when $T$ is a
-    % path.  We conclude with a polynomial time algorithm to find a
-    % feasible tree path labeling of a given set system when $T$ is a
-    % |\tem $k$-subdivided star|, set system has a single containment
-    % tree of overlap components and set size is limited to at most
-    % $k+2$.
-  }
+  \noindent
+  Moreover, this thesis discusses the significance of \COP and the
+  evolution of existing data structures and algorithms used in
+  detection of \COP over the years -- from \PQtrees to Intersection
+  Cardinality Preserving Interval Assignment (ICPIA) data
+  structure. We draw conclusions and connections of which some of them
+  (for instance, \PQRtree and generalized \PQtree) have not be
+  observed before, to the best of our knowledge. This includes the
+  fact that \COP can be tested in logspace.
+
 \end{abstract}

### 2012-02-24 src/020chap2.tex
All content done except acknowledgement. TBD: - spell check - unusual whitespaces, wandering floats

--- 2012-02-02 src/020chap2.tex
@@
 of this problem's several solutions including Tucker's forbidden
-submatrices, \PQtree, \PQRtree and ICPIA. \tnote[dire]{mention the section
-  that compares pqr and gen pq tree} 
+submatrices, \PQtree, \PQRtree and ICPIA.
+
+It is worth noting that we make a very interesting observation about
+\PQRtree, generalised \PQtree and \gPQtree\ -- that the theory that
+all these independent data structures are built upon are very similar
+and equivalent. This is exhaustively described in
+Section~\ref{sec:surveycertalgo} and curiously, this comparison has
+not been encountered by us in the literature.
+
 Section~\ref{sec:surveycopopt} briefly touches upon the scenario of
@@
   \item \label{def::concave-round} $G$ is \emph{concave-round} if its
-    augmented adjacency matrix has \CROP. \tnote[pressing]{cite BHY00}
+    augmented adjacency matrix has \CROP. \cite{bhy00}
   \item $G$ is an \emph{interval graph} if its vertices can be mapped
     to intervals on the real line \stt two vertices are adjacent \iff
-    their corresponding intervals overlap \tnote[pressing]{cite Ben59,
-      Haj57}.  $G$ is an interval graph \iff its maximal clique matrix
+    their corresponding intervals overlap \cite{sb59}.  $G$ is an interval graph \iff its maximal clique matrix
     has COP \cite{fg65} (This follows \cite{gh64} which states
@@
     mapped to a set of arcs on a circle \stt two vertices are adjacent
-    \iff their corresponding arcs overlap.\tnote[pressing]{how is
-      CO/ROP related?}
+    \iff their corresponding arcs overlap.
   \item $H$ is \emph{convex bipartite on columns (rows)} if its half
@@
 the result \cite{hmpv00} \cite[Th~2.7]{d08phd} which demonstrates how
-such a derivation can be done. However, this does not necessarily
-yield an efficient algorithm. We will see results that directly solve
-the problem on matrices since it is known that
-questions~(\ref{q:testcop}) and (\ref{q:copperm}) stated above for
-\COP are efficiently solvable.  Table~\ref{tab:cophistory} gives a
-snapshot of these results.
+such a derivation can be done.  They use something called {\em Lex-BFS
+  ordering} of the vertices of a graph to decide in linear time
+whether the graph is an interval graph. They also show how any
+interval graph recognition algorithm can be used to recognize matrices
+with \COP. This is shown in Theorem~\ref{th:intgraphcop}.
+
+\begin{theoremsansproof}[{\cite[Th.~2]{hmpv00}}]
+  \label{th:intgraphcop}
+  For a binary matrix $M$ the following statements are equivalent.
+  \begin{enumerate}
+  \item The row adjacency graph $G_r(M)$ is an interval graph and $M$
+    is its maximal clique matrix.
+  \item The columns of $M$ are maximal and $M$ has the \COP for rows.
+  \end{enumerate}
+\end{theoremsansproof}
+
+However, this does not necessarily yield an efficient algorithm. We
+will see results that directly solve the problem on matrices since it
+is known that questions~(\ref{q:testcop}) and (\ref{q:copperm}) stated
+above for \COP are efficiently solvable.  Table~\ref{tab:cophistory}
+gives a snapshot of these results.
 
@@
 \cite{at72}. This result influenced most of the \COP results that
-followed in the literature including linear time
-algorithms\tnote[pressing]{linear? did it? which ones?} for \COP
+followed in the literature including linear time algorithms for \COP
 recognition.
@@
 
-\subsection{\PQRtree\ --  COP for set systems}
+\subsection{\PQRtree\ --  COT for set systems}
 \label{sec:surveycertalgo}
@@
   w_n$ that can be applied to $U$ \stt each set $S \in \cF$ becomes a
-  consecutive subsequence (slso termed an {\em interval})
+  consecutive subsequence (also termed an {\em interval})
   $w_{i}w_{i+1}\ldots w_{i+k-1}$ on $\sigma$ for some positive integer
@@
 
-  \item {\em Orthogonal sets of a set system} $\cF$ with universe $U$
-    are subsets of $U$ that are orthogonal to all sets in $\cF$. The
-    set of all orthogonal sets to $\cF$ is denoted by
-    $\cF^{\bot}$\footnote{\cite[Def.~3.1]{mcc04} uses the term {\em non-overlapping
-    family} of $\cF$ and denotes it by $\cN(\cF)$. \label{mcc1}}.
+  \item {\em Orthogonal sets of a set
+      system}\footnote{\cite[Def.~3.1]{mcc04} uses the term {\em
+        non-overlapping family} of $\cF$ and denotes it by
+      $\cN(\cF)$. \label{mcc1}} $\cF$ with universe $U$ are subsets of
+    $U$ that are orthogonal to all sets in $\cF$. The set of all
+    orthogonal sets to $\cF$ is denoted by $\cF^{\bot}$.
 
   \item $\cF$ is called {\em complete}\footnote{\cite{mcc04} calls
-      this a {\em weakly partative family}. \label{mcc2}} if the following hold
-    true for every pair of non-orthogonal\footnote{Or overlapping.} sets
-    $A, B$ in $\cF$.
-    \begin{enumerate}[i.]
-    \item $\cT(U) \subset \cF$
-    \item $A \cup B \in \cF$
-    \item $A \cap B \in \cF$
-    \item $A \setminus B \in \cF$
-    \item $B \setminus A \in \cF$
-    \end{enumerate}
+      this a {\em weakly partative family}. \label{mcc2}} if the
+    following hold true for every pair of non-orthogonal (overlapping)
+    sets $A, B$ in $\cF$.
+
+    \begin{tabular}[h]{lllll}
+%    \begin{enumerate}[i.]
+    (i) $\cT(U) \subset \cF$ &
+    (ii) $A \cup B \in \cF$ &
+    (iii) $A \cap B \in \cF$ &
+    (iv) $A \setminus B \in \cF$ &
+    (v) $B \setminus A \in \cF$
+%   \end{enumerate}      
+    \end{tabular}
+
     In other words, $\cF$ contains all the trivial
@@
 \end{definition}
-
-Generalized \PQtree or \gPQtree is a data structure defined in
-\cite{n89} to represent all orthogonal sets of a set system $\cF$. A
-data structure with the same name was later defined in \cite{mcc04} as
-part of a {\em substitution decomposition} for a set system $\cF$
-\tnote[pressing]{define substitutio decomp.} and subsequently \cite{mcc04} gives
-a new characterization of $\cF$ using a so-called incompatibility
-graph. \PQRtree is defined in \cite{mm96} as a data
-structure to represent any set system $\cF$ with additional
-information in their $R$-nodes if $\cF$ has no \COP.  All three of
-these data structures are proposed as generalizations of \cite{bl76}'s
-\PQtree and hence produce the \PQtree if $\cF$ has \COP. As a whole,
-all these three data strutures are largely identical with their
-differences being notional.  In this section, we will discuss the
-basic theory that they all hold. We will predominantly use the
-terminology from \cite{mm96} and refer to the data structure as
-\PQRtree.
 
@@
 
-\tnote[dire]{narrative is abrupt from here. add some explanation for the
-  theorem and proposition below?}
+The following theorem gives a very interesting property of orthogonal
+sets and $\overline{\cF}$
 
@@
 the second method which is identifying {\em twin}
-elements\footnote{\cite[Sec.~3]{kklv10} calls this {indistinguishable}
+elements\footnote{\cite[Sec.~3]{kklv10} calls this {\em indistinguishable}
   elements. The equivalence class is called a {\em slot}.}. Two
@@
 clearly an equivalence relation and their equivalence classes is known
-to be computable in linear time\cite{wlh01,mm96-ref11} and even in
+to be computable in linear time\cite{wlh01} and even in
 logspace\cite{kklv10}.
@@
 \item \label{th::it1} $\overline{\cF} = \cT(U)$ 
-\item \label{th::it2}$\overline{\cF} = consec(\alpha)$\tnote[pressing]{what is consec?} for
-  some permutation $\alpha$ on $U$ 
+\item \label{th::it2}$\overline{\cF} = consec(\alpha)$ for
+  some permutation $\alpha$ on $U$ ($consect(\alpha)$ is a linear order)
 \item \label{th::it3}$\overline{\cF} = \power{U}$ 
@@
 
+\subsubsection{Comparision with generalized \PQtree}
+
+Stepping back a little, the reader may have noticed the cross
+references between \PQRtree and other data structures in footnotes
+earlier.  Generalized \PQtree or \gPQtree is a data structure defined
+in \cite{n89} to represent all orthogonal sets of a set system
+$\cF$. A data structure with the same name was later defined in
+\cite{mcc04} as part of a {\em substitution decomposition} for a set
+system $\cF$ and subsequently \cite{mcc04} gives a new
+characterization of $\cF$ using a so-called incompatibility graph
+(this is discussed in more detail in
+Section~\ref{sec:incompatibilitygr}). As seen above, \PQRtree is
+defined in \cite{mm96} as a data structure to represent any set system
+$\cF$ with additional information in their $R$-nodes if $\cF$ has no
+\COP. All three of these data structures are proposed as
+generalizations of \cite{bl76}'s \PQtree and hence produce the \PQtree
+if $\cF$ has \COP. As a whole, all these three data strutures are
+largely identical with their differences being notional. We discussed
+the basic theory that they all hold and predominantly used the
+terminology from \cite{mm96}. To the best of our knowledge this
+comparison study has not been made earlier in the literature.  We
+observe the three data strutures of \PQRtree, \gPQtree and generalized
+\PQtree to be equivalent. The terminologies of the theory of all these
+data structures is summarised in Table~\ref{tab:pqrcomparison}.
+
+Going back to the theory we were discussing earlier,
 Theorem~\ref{th:pqrfinalmm} and the theory leading to it is very
@@
 ways as \PQRtree above. This tree is, in essence, the Hasse diagram of
-what they call {\em strong elements} of weak closure,
-$\cW(\cF)$(\ie $\overline{\cF}$). Strong elements of a set
-family are elements that do not overlap with any other elements in the
-family, \ie it is orthogonal to all other sets\cite[Def.~3.3]{mcc04}.
+what they call {\em strong elements} of weak closure, $\cW(\cF)$(\ie
+$\overline{\cF}$). Strong elements of a set family are elements that
+do not overlap with any other elements in the family, \ie it is
+orthogonal to all other sets\cite[Def.~3.3]{mcc04}.
 
@@
 set system from having \COP.
-
-As mentioned before, we observe the three data strutures of \PQRtree,
-\gPQtree and generalized \PQtree to be equivalent. The terminologies
-of the theory of all these data structures is summarised in
-Table~\ref{tab:pqrcomparison}.
 
@@
 %       {}
-    \tnote[dire]{PQR tree def. }
+    \PQRtree
     &
@@
       Decomposition tree of $\cW(\cF)$, $T(\cW(\cF))$
-      \tnote[dire]{PQR tree def. }
     }
@@
       $R$-node $\Leftrightarrow \cF = \power{U}$\\
-    }&
-    \tnote[dire]{give def of P Q R nodes}
+    }
+    &
+    \parbox[t]{\colwidth}
+    {
+      {\em Prime}: For every $I$, $\emptyset \ne I \subset [k],
+      \cup_{i \in I}S_i \notin \cF$\\
+      {\em Linear}: There exists an ordering of $[k]$
+      s.t. $\emptyset \ne I \subset [k]$ implies $\cup_{i \in I}S_i
+      \in \cF$ only if members of $I$ are consecutive in the ordering\\
+      {\em Degenerate}: For every $I$, $\emptyset \ne I \subset [k],
+      \cup_{i \in I}S_i \in \cF$\\
+    }
+
     \\\lasthline
@@
 \label{sec:icpiasurvey}
-\tnote[dire]{narrative is abrupt}
-
-\cite{nsnrs09} describes a characterization of consecutive-ones
-property solely based on the cardinality properties of the sets in the
-set system and does not use any variants of \PQtrees.
+In this section we see a characterization that does not use
+\PQtree. \cite{nsnrs09} describes a characterization of \cop solely
+based on the cardinality properties of the sets in the set system and
+does not use any variants of \PQtrees.
 
@@
 
-\subsection{Other \COP testing algorithms}
-There are more \COP test algorithms in the literature. We make a
-mention of one of them here.  A simple linear time algorithm was
-presented by \cite{hmpv00}. They use something called {\em Lex-BFS
-  ordering} of the vertices of a graph to decide in linear time
-whether the graph is an interval graph. They also show how any
-interval graph recognition algorithm can be used to recognize matrices
-with \COP. This is shown in Theorem~\ref{th:intgraphcop}.
-
-\begin{theoremsansproof}[{\cite[Th.~2]{hmpv00}}]
-  \label{th:intgraphcop}
-  For a binary matrix $M$ the following statements are equivalent.
-  \begin{enumerate}
-  \item The row adjacency graph $G_r(M)$ is an interval graph and $M$
-    is its maximal clique matrix.
-  \item The columns of $M$ are maximal and $M$ has the \COP for rows.
-  \end{enumerate}
-\end{theoremsansproof}
-
-\tnote[dire]{ abrupt: Add a section closing para. does this method (above)
-  lead to efficient algo? }
-
 \section{Matrices without \COP}
@@
 \subsection{Finding forbidden submatrices}
-It is known from Theorem~\ref{th:tuckercop}
-that when a matrix does not have \COP, it must have one of Tucker's
-forbidden submatrices. However, none of the \COP test algorithms discussed in
-Section~\ref{sec:surveycoptest} outputs the forbidden submatrix when
-the test is negative. This certainly gives a hint to how hard it is to find
-these substructures in the input. \tnote[dire]{Fill neg results on
-  finding forbidden submartrices -- boo75, some highlighted stuff from v85}
-
-In this section we will describe \cite{d08phd}'s approximation
+\label{sec:forbiddensubalgo}
+
+It is known from Theorem~\ref{th:tuckercop} that when a matrix does
+not have \COP, it must have one of Tucker's forbidden submatrices (See
+Section~\ref{sec:tucker} for detailed discussion on Tucker's forbidden
+substructures for matrices with \COP). However, none of the \COP test
+algorithms discussed in Section~\ref{sec:surveycoptest} outputs the
+forbidden submatrix when the test is negative. This perhaps gives a
+hint about how hard it is to find these substructures in the input
+matrix. Using a reduction to the longest path problem, \cite{v85}
+proves that finding a Tucker's forbidden submatrix of at least $k$
+rows is NP complete. An example of an optimization problem for a
+matrix without \COP is finding if deleting at most $k$ rows from input
+will restore \COP. This is known to be NP complete
+\cite{b75-phd}. There are several such optimization problems that try
+to find ways to get \COP on the input matrix with certain alterations
+and finding smallest forbidden submatrix is at the heart of these problems.
+
+In this section we will summarize \cite{d08phd}'s approximation
 algorithm for finding forbidden submatrix. We will use $\Tuckerset$ to
@@
 
-\temptext{
-\cite[Ch.~3]{d08phd} describes two main algorithms - one is an
-approximation algorithm and the other one is an exact algorithm.
+Using Corollary~\ref{cor:tuckertheorem}, a forbidden submatrix can be
+found as follows. For every vertex triple $x, y, z$ in $G_M$ that
+correspond to columns of $M$, determine the sum of the lengths of
+three shortest paths connecting each pair of these vertices (there are
+3 possible pairs) avoiding the neighborhood of third vertex. If all
+three paths exist, then $x, y, z$ indeed form an asteroidal
+triple. Among all such triples, select the triple $u, v, w$ where the
+sum is minimum and return the rows and columns of $M$ that correspond
+to the vertices in the three shortest paths described earlier for this triple. This
+return value forms a submatrix and it obviously must contain a
+submatrix from $\Tuckerset$ because it represents an asteroidal
+triple. It may be noted that this procedure does not always return a
+submatrix of smallest size because the shortest paths computed may not
+be vertex disjoint. The following claim shows that this certainly
+gives a reasonable approximation of the solution. The claim has been
+abridged to summarize the main idea behind this algorithm.
+Note that a $(x,y)$-matrix is one with at most $x$ number of {\un}s in every
+column and at most $y$ number of {\un}s in every row. If $x$ or $y$ is given as
+$*$, it denotes that number of {\un}s is unbounded.
+
+\begin{proposition}[{\cite[Pr.~3.1, {\em abridged}]{d08phd}}] 
+  \label{prop:tuckerdomapprox}
+  Let $M$ be a $(*,\Delta)$-matrix of size $m \times n$ that contains
+  a forbidden submatrix $M'$ of size $m' \times n'$ from $\Tuckerset$. Then
+  this algorithm computes a submatrix that belongs to $\Tuckerset$
+  with at most $m'+5$ rows and $n'+3$ columns in $O(\Delta m n^2 +
+  n^3)$ time.
+\end{proposition}
+
+In fact, the approximation mentioned in
+Proposition~\ref{prop:tuckerdomapprox} is the worst case when
+the actual solution $M'$ is of type $M_{IV}$. The approximation is
+slightly better for the other types of submatrices from $\Tuckerset$,
+the best case being an exact solution when $M'$ is $M_{I_k}$ or
+$M_{II_k}$. Thus this algorithm works accurately when the solution is of
+type $M_{I_k}$ or $M_{II_k}$.
+
+Next, notice that Tucker matrix types $M_{IV}$ and $M_{V}$ are
+constant size -- $4 \times 6$ and $4 \times 5$ respectively. Thus they
+can be searched for in polynomial time even with brute force method as
+Corollary~\ref{cor:mivmvpolytime} states.
+
+\begin{corollary}[{\cite[Cor.~3.3, {\em abridged}]{d08phd}}] 
+  \label{cor:mivmvpolytime}
+  Let $M$ be a binary $(*, \Delta)$-matrix of size $m \times n$ and
+  $M'$ be a binary matrix from $\Tuckerset$. A submatrix in $M$ that
+  is isomorphic to $M'$ can be found in $O(\Delta^3m^2n^3)$ time or
+  $O(\Delta^4m^2n)$ time when $M'$ is of type $M_{IV}$ or $M_V$
+  respectively.
+\end{corollary}
+
+Their algorithm for finding $M_{III_k}$ type submatrix is slightly
+involved and we only present here the analysis result in
+Proposition~\ref{cor:miiitime}. The algorithm is derived from the
+structural similarities between the first three types of Tucker
+matrices and the fact that $M_{I_k}$ is a chordless cycle of length
+$2k+4$.
+
+\begin{proposition}[{\cite[Pr.~3.6]{d08phd}}] 
+  \label{cor:miiitime}
+  Let $M$ be a $(*,\Delta)$-matrix $M$ of size $m \times n$. Then a
+  minimum size submatrix of type $M_{III_k}$ in $M$ can be found in
+  $O(\Delta^3m^3n+\Delta^2m^2n^2)$ time.
+\end{proposition}
+
+All the above algorithmic results are combined to obtain the final
+single algorithm for finding forbidden
+submatrix. Theorem~\ref{th:forbiddenalg} gives this result.
 
@@
   \label{th:forbiddenalg}
-Let $M$ be a $(*,\Delta)$-matrix of size $m
-\times n$. A forbidden submatrix from $\Tuckerset$ in $M$ that has a minimum
-number of rows can be found in $O(\Delta^3m^2n\dot(m+n^2))$
-time. Within the same time, one can also find a forbidden submatrix
-from $\Tuckerset$ in $M$ that has a minimum number of columns, a minimum number
-of rows and columns, or a minimum number of entries.
-
+  Let $M$ be a $(*,\Delta)$-matrix of size $m \times n$. A forbidden
+  submatrix from $\Tuckerset$ in $M$ that has a minimum number of rows
+  can be found in $O(\Delta^3m^2n\dot(m+n^2))$ time. Within the same
+  time, one can also find a forbidden submatrix from $\Tuckerset$ in
+  $M$ that has a minimum number of columns, a minimum number of rows
+  and columns, or a minimum number of entries.
 \end{theoremsansproof}
 
-}
+The procedure used to prove Theorem~\ref{th:forbiddenalg} is as
+follows. 
+\begin{enumerate}
+\item Run the algorithm mentioned in
+  Propostition~\ref{prop:tuckerdomapprox} to obtain a submatrix $A$
+  which is an approximation of the smallest forbidden submatrix. (If
+  the actual solution is $M_{I_k}$ or $M_{II_k}$ this will be an exact
+  solution.)
+\item Run the algorithm from Propostition~\ref{cor:miiitime} to find
+  an induced $M_{III_k}$ and let this be $B$.
+\item Run the algorithm from Corollary~\ref{cor:mivmvpolytime} twice
+  -- once each for finding $M_{IV}$ and finding $M_{V}$. Let the
+  results be $C$ and $D$ respectively.
+\item Return the minimum size matrix among $A, B, C, D$ (minimum size
+  being one of the following: number of rows, number of columns, sum
+  of number of rows and columns, number of {\un} entries).
+\end{enumerate}
+
+In the next section we see yet another forbidden substructure for
+\COP which is easier to compute.
 
 \subsection{Incompatibility graph - a certificate for no \COP}
+\label{sec:incompatibilitygr}
 
@@
 \COP which is efficiently computable. This is done by way of an {\em
-  incompatibility graph} of a set system\footnote{Since \COP in
+  incompatibility graph} of a set system (since \COP in
   matrices and set systems are equivalent problems as seen in
   Section~\ref{sec:surveycertalgo} we will use the term set system in
-  this section, in line with \cite{mcc04}'s convention.}. The
+  this section, in line with \cite{mcc04}'s convention). The
 construction of this graph is based on the following observation.
@@
 \begin{definition}[{\em Incompatibility graph \cite[Def.~6.1]{mcc04}}]
+  \label{def:incompatibility}
   Let $\cF$ be a set system with universe $U$. Let $A_{\cF} = \{(a,b)
@@
 have an odd cycle.  There is also a requirement on the length of this
-odd cycle which is formally stated in the following
-theorem. \tnote[pressing]{all refrences to ``following''/abov theorem must be
-  chagned to slashrefs. }
+odd cycle which is formally stated in Theorem~\ref{th:mccbipartite}.
 
@@
     Th.~2.6, Proof p. 44--47]{d08phd}}] 
+  \label{th:mccbipartite}
   Let $\cF$ be a set family with universe $U$. Then $\cF$ has \COP
@@
 
-Each edge of the cycle is labeled with a set from $\cF$ that documents
-the incompatibility.
-
-\tnote[dire]{how does it give clues
-  about the $R$ node? See dom.}
+Searching for an odd cycle to recognize a bipartite graph is a known
+and polynomially solved problem. \cite{mcc04}'s algorithm also labels
+each edge of the cycle with a set from $\cF$ that documents the
+incompatibility.
+
+% \tnote[dire]{how does it give clues
+%   about the $R$ node? See dom.}
 
@@
 
-\tnote[dire]{ -- sect 4.1 in cite:d08phd has many results surveyed. hardness
-  results, approx. results. results are usually for a class of
-  matrices $(a,b)$ where number {\un}s in columns and rows are
-  restriced to $a$ and $b$ . -- problem of flipping at most $k$
-  entries of $M$ to make it attain COP. this is NP complete
-  cite:b75-phd%
-\\(1) scite:lb62 showed that interval graphs are chordal and
-  AT-free. %
-\\couple of lines refering to tucker's submatrices. refer
-  earlier section.
-
-Another problem is to find the minimum number of entries in the matrix
-that can be toggled to result in a matrix with COP.  \cite{v85}
-discusses approximation of {\sc COP Augmentation} which is the problem
-of changing of the minimum number of zero entries to {\un}s so that
-the resulting matrix has COP. As mentioned earlier, this problem is
-known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
-using a reduction to the longest path problem, \tnote[pressing]{or is it a
-  survey of another result?  check.} that finding a Tucker's forbidden
-submatrix of at least $k$ rows is NP complete. \tnote[pressing]{how is this
-  different from booth's 75 result??}  \tnote[pressing]{where should this go?
-  cite|tz04 (approx submatrix with COP sparse matrices)}
-
-\cite{jkckv04} discusses the use of matrices with almost-COP (instead
-of one block of consecutive {\un}s, they have $x$ blocks, or {\em
-  runs}, of consecutive {\un}s and $x$ is not too large) in the
-storage of very large databases.  The problem is that of reordering of
-a binary matrix such that the resulting matrix has at most $k$ runs of
-{\un}s. This is proved to be NP hard using a reduction from the
-Hamiltonian path problem.\tnote[pressing]{Theorem 2.1 in jkckv}
-\tnote[pressing]{(1) A connection of COP problem to the travelling
-  salesman problem is also introduced. what does this mean? -- COP can
-  be used as a tool to reorder $0.5T \le runs(M) \le T$. (2) The
-  optimization version of the $k$-run problem, i.e. minimization of
-  number of blocks of ones is proven to be NP complete by
-  cite:k77}\tnote{are these two the same?} \tnote{what is the
-  reduction?} \tnote{other problems similar to COP -- cite:ckl96 (ILP,
-  circ ones, one drop) -- cite:th98 (generalization of COP - minimax,
-  biotonic column) Tucker}
-}
+% \tnote[dire]{ -- sect 4.1 in cite:d08phd has many results surveyed. hardness
+%   results, approx. results. results are usually for a class of
+%   matrices $(a,b)$ where number {\un}s in columns and rows are
+%   restriced to $a$ and $b$ . -- problem of flipping at most $k$
+%   entries of $M$ to make it attain COP. this is NP complete
+%   cite:b75-phd%
+% \\(1) scite:lb62 showed that interval graphs are chordal and
+%   AT-free. %
+% \\couple of lines refering to tucker's submatrices. refer
+%   earlier section.
+
+% Another problem is to find the minimum number of entries in the matrix
+% that can be toggled to result in a matrix with COP.  \cite{v85}
+% discusses approximation of {\sc COP Augmentation} which is the problem
+% of changing of the minimum number of zero entries to {\un}s so that
+% the resulting matrix has COP. As mentioned earlier, this problem is
+% known to be NP complete due to \cite{b75-phd}.  \tnote[pressing]{where should this go?
+%   cite|tz04 (approx submatrix with COP sparse matrices)}
+% }
 

### 2012-02-24 src/030chap3.tex
All content done except acknowledgement. TBD: - spell check - unusual whitespaces, wandering floats

@@
 Combining the ideas of path labeling, hypergraph isomorphism and its
-induced path labeling we now define {\em feasiblity} of a path
+induced path labeling we now define {\em feasibility} of a path
 labeling in Definition~\ref{def:feasibletpl}.
@@
 One of the problems solved in this thesis \FTPL characterizes
-feasibile tree path labeling.
+feasible tree path labeling.
 
@@
 pairwise intersection cardinalities are preserved, then $d$-way
-intersection cardinalities are preseved for $2 < d < n$. Thus, in a
+intersection cardinalities are preserved for $2 < d < n$. Thus, in a
 way, pairwise intersections ``capture'' feasibility of an interval
@@
 
+In Figure~\ref{fig:solpartI}, the original problem from
+Section~\ref{sec:problem} is restated more formally. The target tree
+is $T$ with vertex set $\{1,2,\ldots,11\}$ on the left hand side along
+with the path system (hypergraph) -- $\{7,2,6,5\}, \{8,2,4\},
+\{10,6,5,3\}, \{9,1,5,3,11\}$. The hypergraph is $\cF$ with hyperedges
+$\{\xPa, \xPi, \xVi, \xCh\}, \{\xSn, \xWo, \xPi\}, \{\xCh, \xFr, \xVi,
+\xLi\}, \{\xLu, \xSa, \xSc, \xCh, \xFr\}$ shown in the right hand
+side. In subsequent illustrations, the gray boxes which contain the
+elements of universe $U$, will be assigned to nodes and removed from
+the right hand side figure and placed next to the vertex of $T$ in the
+left hand side figure.  The path labeling is implicitly given by the
+color-coded Venn diagrams (the order of listing of the hyperedges in
+given above is also done respective to the input path labeling).
+
 \begin{figure}[htb]
@@
     problem is shown again here using the input described in
-    Section~\ref{sec:problem}. The target tree is $T$ with vertex set
-    $\{1,2,\ldots,11\}$ on the left hand side along with the path
-    system (hypergraph) -- $\{7,2,6,5\}, \{8,2,4\}, \{10,6,5,3\},
-    \{9,1,5,3,11\}$. The hypergraph is $\cF$ with hyperedges $\{\xPa,
-    \xPi, \xVi, \xCh\}, \{\xSn, \xWo, \xPi\}, \{\xCh, \xFr, \xVi,
-    \xLi\}, \{\xLu, \xSa, \xSc, \xCh, \xFr\}$ shown in the right hand
-    side. In subsequent illustrations, the gray boxes which contain
-    the elements of universe $U$, will be assigned to nodes and
-    removed from the right hand side figure and placed next to the
-    vertex of $T$ in the left hand side figure.  The path labeling is
-    implicitly given by the color-coded Venn diagrams (the order of
-    listing of the hyperedges in this caption is also done respective
-    to the input path labeling).}
+    Section~\ref{sec:problem}.}
   \label{fig:solpartI}  
 \end{figure}
+
+Now see
+Figure~\ref{fig:solpartIII} to see how the algorithm proceeds.
+Before that it can be observed in Figure~\ref{fig:solpartI}, that none of the
+leaves are shared by more than one path.  Hence,
+\filteri does not alter the hypergraphs. \filterii finds the leaf
+assignments as shown -- elements \xPa, \xSn, \xLu, \xSc, \xLi, \xWo
+are assigned to vertices 7, 8, 9, 11, 10, 4 respectively. Computation
+of these leaf labels are deterministic except for $\xLu, \xSc$ where
+the algorithm arbitrarily chooses them from $X = \{\xLu, \xSa, \xSc\}$
+and for $\xSn, \xWo$ arbitrarily chooses from $X = \{\xSn,
+\xWo\}$. Here we show only one such choice and clearly one could have
+chosen another leaf label (\eg $\xSa$ on 9 or $\xWo$ on 8) to get a
+different feasible hypergraph isomorphism.  After this, the leaves of
+$T$ are pruned and the hypergraph updated accordingly by \algoiii
+subroutine -- the pruned leaves are shown with dotted edges and the
+corresponding Venn diagrams show the updated hypergraphs. $\phi_0$ is
+this leaf assignment and the algorithm proceeds to solve the
+subproblem with the new hypergraphs and path labeling which is:
+$\{2,6,5\}, \{2\}, \{6,5,3\}, \{1,5,3\}$ labeled with $\{\xPi, \xVi,
+\xCh\}, \{\xPi\}, \{\xCh, \xFr, \xVi\}, \{\xSa, \xCh, \xFr\}$
+respectively.
 
@@
   \end{tabular}
-  \caption[\figtabsize Problem solution part 2]{\figtabsize It can be observed in
-    Figure~\ref{fig:solpartI}, that none of the leaves are shared by
-    more than one path. Hence, \filteri does not alter the
-    hypergraphs. \filterii finds the leaf assignments as shown --
-    elements \xPa, \xSn, \xLu, \xSc, \xLi, \xWo are assigned to
-    vertices 7, 8, 9, 11, 10, 4 respectively. Computation of these
-    leaf labels are deterministic except for $\xLu, \xSc$ where the
-    algorithm arbritarily choses them from $X = \{\xLu, \xSa, \xSc\}$
-    and for $\xSn, \xWo$ arbitrarily choses from $X = \{\xSn,
-    \xWo\}$. Here we show only one such choice and clearly one could
-    have chosen another leaf label (\eg $\xSa$ on 9 or $\xWo$ on 8) to
-    get a different feasible hypergraph isomorphism.  After
-    this, the leaves of $T$ are pruned and the hypergraph updated
-    accordingly by \algoiii subroutine -- the pruned leaves are shown
-    with dotted edges and the corresponding Venn diagrams show the
-    updated hypergraphs. $\phi_0$ is this leaf assignment and the
-    algorithm proceeds to solve the subproblem with the new
-    hypergraphs and path labeling which is: $\{2,6,5\}, \{2\},
-    \{6,5,3\}, \{1,5,3\}$ labeled with $\{\xPi, \xVi, \xCh\},
-    \{\xPi\}, \{\xCh, \xFr, \xVi\}, \{\xSa, \xCh, \xFr\}$
-    respectively.}
+  \caption[\figtabsize Problem solution part 2]{\figtabsize \filterii
+    finds leaf assignments as shown.}
   \label{fig:solpartIII}  
 \end{figure}
+
+In the subproblem shown in Figure~\ref{fig:solpartIII}, it can be
+noted that there are two leaves which are shared by more than one
+paths -- 2 and 3. Hence \filteri filters this out one by one. 
+In Figure~\ref{fig:solpartIVa}, we
+show the handling of leaf 2 which was shared by paths $\{2,6,5\}$
+($=P_1$, say) and $\{2\}$($=P_2$, say). \filteri removes $P_1$ and
+adds $\{6,5\}$. Note that only one set is added and removed since $P_2
+\subset P_1$. Accordingly, \filteri alters the hypergraph on the right
+hand side by removing and adding $\{\xPi, \xVi, \xCh\}$ and $\{\xVi,
+\xCh\}$ respectively.
 
@@
   \end{tabular}
-  \caption[\figtabsize Problem solution part 3]{\figtabsize In the subproblem shown in
-    Figure~\ref{fig:solpartIII}, it can be noted that there are two
-    leaves which are shared by more than one paths -- 2 and 3. Hence
-    \filteri filters this out one by one. Here we show the handling of
-    leaf 2 which was shared by paths $\{2,6,5\}$ ($=P_1$, say) and
-    $\{2\}$($=P_2$, say). \filteri removes $P_1$ and adds
-    $\{6,5\}$. Note that only one set is added and removed since $P_2
-    \subset P_1$. Accordingly, \filteri alters the hypergraph on the
-  right hand side by removing and adding $\{\xPi, \xVi, \xCh\}$ and $\{\xVi, \xCh\}$ respectively.}
+  \caption[\figtabsize Problem solution part 3]{\figtabsize \filteri
+    handling shared leaf 2.}
   \label{fig:solpartIVa}  
 \end{figure}
+
+After handling leaf 2, \filteri proceeds to handle leaf 3 which is
+shared by paths $\{1,5,3\}$ and $\{6,5,3\}$. This is illustrated in
+Figure~\ref{fig:solpartIVb}. \filteri removes these two paths and adds
+their intersection and set differences -- $\{5,3\}$, $\{1\}$ and
+$\{6\}$. Correspondingly, hyperedges $\{\xSa, \xCh, \xFr\}$ and
+$\{\xVi, \xCh, \xFr\}$ are removed and new hyperedges $\{\xCh,
+\xFr\}$, $\{\xSa\}$ and $\{\xVi\}$ respectively are added. There are
+no more leaves shared and the algorithm proceeds to \filterii.
 
@@
   \end{tabular}
-  \caption[\figtabsize Problem solution part 4]{\figtabsize After handling leaf 2, \filteri proceeds to
-    handle leaf 3 which is shared by paths $\{1,5,3\}$ and
-    $\{6,5,3\}$. \filteri removes these two paths and adds their
-    intersection and set differences -- $\{5,3\}$, $\{1\}$ and
-    $\{6\}$. Correspondingly, hyperedges $\{\xSa, \xCh, \xFr\}$ and $\{\xVi,
-    \xCh, \xFr\}$ are removed and new hyperedges $\{\xCh, \xFr\}$, $\{\xSa\}$
-    and $\{\xVi\}$ respectively are added. There are no more leaves
-    shared and the algorithm proceeds to \filterii.}
+  \caption[\figtabsize Problem solution part 4]{\figtabsize \filteri
+    handles shared leaf 3.}
   \label{fig:solpartIVb}  
 \end{figure}
+
+Next, Figure~\ref{fig:solpartVI} shows how \filterii computes the leaf
+assignment $\phi_1$ -- $\xPi, \xSa, \xFr$ on 2,1,3 respectively. These
+assignments are deterministic since $X$ consisted of only single
+elements for each of these leaves. After this the rest of the labeling
+is fairly obvious but we present it in Figure~\ref{fig:solpartVII} for
+the sake of completeness.
 
@@
   \end{tabular}
-  \caption[\figtabsize Problem solution part 5]{\figtabsize \filterii computes the leaf assignment $\phi_1$ -- $\xPi,
-    \xSa, \xFr$ on 2,1,3 respectively. These assignments are
-    deterministic since $X$ consisted of only single elements for each
-  of these leaves. After this the rest of the labeling is fairly
-  obvious but we present it in Figure~\ref{fig:solpartVII} for the
-  sake of completeness.}
+  \caption[\figtabsize Problem solution part 5]{\figtabsize \filterii
+    computes the leaf assignment $\phi_1$.}
   \label{fig:solpartVI}  
 \end{figure}
-\clearpage
+%\clearpage
+
+In Figure~\ref{fig:solpartVII}, the last part of the isomorphism is
+computed as the leaf label $\phi_2$ which maps $\xCh, \xVi$ to 5, 6
+respectively. Figure~\ref{fig:solpartVIII} shows the final feasibility
+isomorphism which is the union of all the leaf labels. The feasibility
+hypergraph isomorphism $\phi$ is computed as the union of the leaf
+labels $\phi_0,\phi_1,\phi_2$. One can easily verify that this
+isomorphism does indeed induce the path labeling given in the original
+input.
 
@@
   \end{tabular}
-  \caption[\figtabsize Problem solution part 6]{\figtabsize The last part of the isomorphism is computed as
-  the leaf label $\phi_2$ which maps $\xCh, \xVi$ to 5, 6
-  respectively. Figure~\ref{fig:solpartVIII} shows the final
-  feasibility isomorphism which is the union of all the leaf labels.}
+  \caption[\figtabsize Problem solution part 6]{\figtabsize Last part
+    of solution computed. }
   \label{fig:solpartVII}  
@@
   \caption[\figtabsize Problem solution part 7]{\figtabsize The solution to the problem presented in
-    Figure~\ref{fig:solpartI}. The feasibility hypergraph isomorphism
-    $\phi$ is computed as the union of the leaf labels
-    $\phi_0,\phi_1,\phi_2$. One can easily verify that this
-    isomorphism does indeed induce the path labeling given in the
-    original input.}
+    Figure~\ref{fig:solpartI}. }
   \label{fig:solpartVIII}
@@
 
-In Section~\ref{sec:motive} it is described how \COP is a
-special case of TPL as described by the \CFTPLINT problem. In this
-section we will dwell on this in more detail.
-
-\annote[dire]{The special case of TPL where the target tree $T$ is a
-path. Hence, all path labels are can be viewed as intervals assigned
-to the sets in $\cF$. }{needs work}
-
-It is shown, in \cite{nsnrs09}, that the
-filtering algorithms outlined above need only preserve pairwise
-intersection cardinalities, and higher level intersection
-cardinalities are preserved by the Helly Property of intervals.
+In Section~\ref{sec:motive} it is described how \COP is a special case
+of TPL as described by the \CFTPLINT problem.  This special case of
+TPL is where the target tree $T$ is a path. All path labels can now be
+viewed as intervals assigned to the hyperedges in input hypergraph
+$\cF$.  It is shown, in \cite{nsnrs09}, that the filtering algorithms
+outlined above need only preserve pairwise intersection cardinalities,
+and higher level intersection cardinalities are preserved by the Helly
+Property of intervals. Thus ICPIA as described in
+Section~\ref{sec:icpiasurvey} is a sufficient and necessary condition
+for \COP and this is solvable in polynomial time.
 % Consequently, the filter algorithms do not need to ever evaluate the
@@
 %Assignment (ICPIA) \cite{nsnrs09}.
-This structure and its algorithm is used in the next section for
-finding tree path labeling from a $k$-subdivided star due to this
-graph's close relationship with intervals.
-
-\tnote[dire]{ELABORATE}
+
+Moreover, this special case structure and its algorithm is used in the
+next section for finding tree path labeling from a $k$-subdivided star
+due to this tree's close relationship with intervals.
 
@@
 
-\tnote[dire]{REPEAT!!
-
-In this section we consider the problem of assigning paths from a
-$k$-subdivided star $T$ to a given set system $\cF$ such that each set
-$X \in \cF$ is of cardinality at most $k+2$.  Secondly, we present our
-results only for the case when overlap graph $\bO(\cF)$ is connected.
+% \tnote[dire]{REPEAT!!
+
+% In this section we consider the problem of assigning paths from a
+% $k$-subdivided star $T$ to a given set system $\cF$ such that each set
+% $X \in \cF$ is of cardinality at most $k+2$.  Secondly, we present our
+% results only for the case when overlap graph $\bO(\cF)$ is connected.
+
 %The
@@
 %$\bO(\cF)$ and $\bO(\cF) \subseteq \bI(\cF)$.
-A connected component of $\bO(\cF)$ is called an overlap component of
-$\cF$.  An interesting property of the overlap components is that any
-two distinct overlap components, say $\cO_1$ and $\cO_2$, are such
-that any two sets $S_1 \in \cO_1$ and $S_2 \in \cO_2$ are disjoint,
-or, w.l.o.g, all the sets in $\cO_1$ are contained within one set in
-$\cO_2$.  This containment relation naturally determines a
-decomposition of the overlap components into rooted containment trees.
-We consider the case when there is only one rooted containment tree,
-and we first present our algorithm when $\bO(\cF)$ is connected.  It
-is easy to see that once the path labeling to the overlap component in
-the root of the containment tree is achieved, the path labeling to the
-other overlap components in the rooted containment tree is essentially
-finding a path labeling when the target tree is a path: each target
-path is a path that is allocated to sets in the root overlap
-component.  Therefore, for the rest of this section, $\bO(\cF)$ is a
-connected graph. Recall that we also consider the special case when
-all hyperedges are of cardinality at most $k+2$.  By definition, a
-$k$-subdivided star has a central vertex which we call the {\em root},
-and each root to leaf path is called a {\em ray}.  First, we observe
-that by removing the root $r$ from $T$, we get a collection of $p$
-vertex disjoint paths of length $k+1$, $p$ being the number of leaves
-in $T$.  We denote the rays by $R_1, \ldots, R_p$, and the number of
-vertices in $R_i$, $i \in [p]$ is $k+2$.  Let
-$\seq{v_{i1},\ldots,v_{i(k+2)}=r}$ denote the sequence of vertices in
-$R_i$, where $v_{i1}$ is the leaf. Note that $r$ is a common vertex to
-all $R_i$.
-}
+
+% A connected component of $\bO(\cF)$ is called an overlap component of
+% $\cF$.  An interesting property of the overlap components is that any
+% two distinct overlap components, say $\cO_1$ and $\cO_2$, are such
+% that any two sets $S_1 \in \cO_1$ and $S_2 \in \cO_2$ are disjoint,
+% or, w.l.o.g, all the sets in $\cO_1$ are contained within one set in
+% $\cO_2$.  This containment relation naturally determines a
+% decomposition of the overlap components into rooted containment trees.
+% We consider the case when there is only one rooted containment tree,
+% and we first present our algorithm when $\bO(\cF)$ is connected.  It
+% is easy to see that once the path labeling to the overlap component in
+% the root of the containment tree is achieved, the path labeling to the
+% other overlap components in the rooted containment tree is essentially
+% finding a path labeling when the target tree is a path: each target
+% path is a path that is allocated to sets in the root overlap
+% component.  Therefore, for the rest of this section, $\bO(\cF)$ is a
+% connected graph. Recall that we also consider the special case when
+% all hyperedges are of cardinality at most $k+2$.  By definition, a
+% $k$-subdivided star has a central vertex which we call the {\em root},
+% and each root to leaf path is called a {\em ray}.  First, we observe
+% that by removing the root $r$ from $T$, we get a collection of $p$
+% vertex disjoint paths of length $k+1$, $p$ being the number of leaves
+% in $T$.  We denote the rays by $R_1, \ldots, R_p$, and the number of
+% vertices in $R_i$, $i \in [p]$ is $k+2$.  Let
+% $\seq{v_{i1},\ldots,v_{i(k+2)}=r}$ denote the sequence of vertices in
+% $R_i$, where $v_{i1}$ is the leaf. Note that $r$ is a common vertex to
+% all $R_i$.
+%}
 
@@
  hyperedge that has been assigned a path is considered.  At the
- beginning of each iteration hyperedges of $\cO$ are classifed into
+ beginning of each iteration hyperedges of $\cO$ are classified into
  the following disjoint sets.
@@
   from $R_j$ such that the path contains the leaf in $R_j$.  This path
-  is refered to as $\cl(X_j)$.  Set $p_1(X_j)=|X|, p_2(X_j)=s(X_j)=0$.
+  is referred to as $\cl(X_j)$.  Set $p_1(X_j)=|X|, p_2(X_j)=s(X_j)=0$.
   Hyperedges in $S$ are not added to $\cO_2$ but are added to
@@
 element.  In particular, once the super-marginal edges are identified,
-each iteration involes finding the next hyperedge to consider, and
+each iteration involves finding the next hyperedge to consider, and
 testing for a path to that hyperedge.  To identify the next hyperedge
 to consider, we consider the breadth first layering of the hyperedges
-with the zeroeth layer consisting of the super-marginal hyperedges.
+with the zeroth layer consisting of the super-marginal hyperedges.
 Since $\cO$ is connected, it follows that all hyperedges of $\cO$ will
@@
 As described by Definition~\ref{def:matrixsetsystem} a set system or
-hypergraph can be concisely represented by a binary matrix
-\stt the row indices of the matrix denote the vertex set of the
-hypergraph and each column bijectively correspond to a hyperedge. And
-thus the ideas of feasible \TPL and \ICPPL are also defined for a
-binary matrix. If a \TPL $(\cF,\cl,T)$ is feasible, \ie it is an ICPPL
-(see Section~\ref{xxxxx} \tnote[dire]{SECTION NUMBER}) and if $M$ is
-the representative matrix for hypergraph $\cF$, then we say that $M$
-has an \ICPPL and vice versa.
+hypergraph can be concisely represented by a binary matrix \stt the
+row indices of the matrix denote the vertex set of the hypergraph and
+each column bijectively correspond to a hyperedge. And thus the ideas
+of feasible \TPL and \ICPPL are also defined for a binary matrix. If a
+\TPL $(\cF,\cl,T)$ is feasible, \ie it is an ICPPL (see
+Section~\ref{sec:feasible}) and if $M$ is the representative matrix
+for hypergraph $\cF$, then we say that $M$ has an \ICPPL and vice
+versa.
 
@@
 of $M$.  Let $col(M_i)$ be the set of columns in the submatrix $M_i$.
-The support of a prime submatrix $M_i$ and support of a
-set of prime submatrices $X$ are defined as follows. Note that for each $i$,
-$supp(M_i) \subseteq U$ where $U$ is the vertex set of hypergraph $\cF$.
+The support of a prime submatrix $M_i$ and support of a set of prime
+submatrices $X$ are defined as follows. Note that for each $i$,
+$supp(M_i) \subseteq U$ where $U$ is the vertex set of hypergraph
+$\cF$.
 \begin{align*}
@@
 
-\def \primeless {\preccurlyeq}
 \noindent
@@
 submatrices and the above relation can be defined for any hypergraph
-since an overlap graph is defined for any hypergraph. We will use this structure of prime submatrices to present our
-results on an ICPPL for a hypergraph $\cF$. 
-Now we present a few lemma and a theorem that show that $\primeless$ is
-a partial order. An alternate notation for $(M_i,M_j) \in \primeless$
-is $M_i \primeless M_j$.
+since an overlap graph is defined for any hypergraph. We will use this
+structure of prime submatrices to present our results on an ICPPL for
+a hypergraph $\cF$.  Now we present a few lemma and a theorem that
+show that $\primeless$ is a partial order.  These results are from
+\cite{nsnrs09} but are independent of the \COP property of $M$ and is
+valid for any binary matrix $M$ except Theorem~\ref{thm:partitionold}
+which we subsequently extend in Theorem~\ref{thm:partition} to
+accommodate \ICPPL.  Note that an alternate notation for $(M_i,M_j)
+\in \primeless$ is $M_i \primeless M_j$.
 
@@
   \item $\primeless$ is a partial order on $\bP$.
-  \item $\primeless$ uniquely partitions $\bP$ into $\bP_1, \bP_2
-    \ldots ,\bP_r$ for some $r \ge 1$.
+  \item \label{thm:i:chainpartition} $\primeless$ uniquely partitions $\bP$ into $\bP_1, \bP_2
+    \ldots ,\bP_r$ for some $r \ge 1$ as follows. 
   \item \label{thm:i:totalorder} If $M$ has \COP, then for each $i \in
@@
   \end{enumerate}
-% $\preccurlyeq$ is a partial
-%   order on the set of prime submatrices of $M$.  Further, it uniquely
-%   partitions the prime submatrices of $M$ such that on each set in
-%   the partition $\preccurlyeq$ induces a total order.
 \end{theorem}
 
-% \begin{theorem} % [{\cite[Th.~4]{nsnrs09}}] %
-%   \label{thm:partitionnew}%
-%   \parindent \defindent
-%   \begin{enumerate}[i. ]
-%   \item $\primeless$ is a partial order on $\bP$.
-%   \item $\primeless$ uniquely partitions $\bP$ into $\bP_1, \bP_2
-%     \ldots ,\bP_k$ for some $k \ge 1$.
-%   \item If $M$ has \COP, then for each $i \in [k]$, $\primeless$
-%     induces a total order in $\bP_i$.
-%   \end{enumerate}
-% \end{theorem}
-
-%   Let $M$ be a binary matrix for a hypergraph
-%   with \TPL and $\bP$ be the set of prime submatrices of $M$. Then
-%   Theorem~\ref{thm:partitionold} holds true with (iii) changed as
-%   follows -- for each $i \in [k]$, $\primeless$ induces an in-tree
-%   in $\bP_i$.
-Since we are concerned with the generalization of \COP, we generalize
-the condition~(\ref{thm:i:totalorder}) in
-Theorem~\ref{thm:partitionold} from total order to in-trees as
-follows.  The in-tree is derived from the Hasse diagram of
-$\primeless$. Each of these in-trees will be on disjoint vertex sets,
-which in this case would be disjoint sets of prime-submatrices
+\cite{nsnrs09} does not explicitly mention how the partitioning of
+$\bP$ is done in Theorem~\ref{thm:partitionold} and merely mention
+that the total order in (\ref{thm:i:totalorder}) is obvious. We call
+this partitioning in Theorem~\ref{thm:partitionold} point
+(\ref{thm:i:chainpartition}) as {\em chain partitioning}. We explain
+what chain partitioning is in the following text. For that, first we
+construct the Hasse diagram of $\primeless$ which we observe is a
+collection of in-trees in the following Theorem~\ref{thm:partition}.
+Each of these in-trees have disjoint vertex sets,
+which in this case would be disjoint sets of prime submatrices
 $\bP_i$, $i \in [r]$ for some $r > 0$.  For each $\bP_i$, the in-tree
@@
 for $\primeless$ defined below.
-% %D replace \cI with \primeless_c
-% Let $\primeless_c$ be the following set:
 \begin{align*}
@@
 order $\primeless$.
+
+% Since we are concerned with the generalization of \COP, we generalize
+% the condition~(\ref{thm:i:totalorder}) in
+% Theorem~\ref{thm:partitionold} from total order to in-trees as
+% follows.
 
@@
   vertices corresponding to maximal upper bounds, no outgoing edge is
-  present in $\primeless_c$.  Secondly, for each other element, exactly one
-  outgoing edge is chosen, and for the minimal lower bound, there is
-  no incoming edge.  Consequently, $X$ is acyclic, and since each
-  vertex has at most one edge leaving it, it follows that $X$ is a
-  collection of in-trees, and for each in-tree, the root is a maximal
-  upper bound in $\primeless$.  
+  present in $\primeless_c$.  Secondly, for each other element,
+  exactly one outgoing edge is chosen, and for the minimal lower
+  bound, there is no incoming edge.  Consequently, $X$ is acyclic, and
+  since each vertex has at most one edge leaving it, it follows that
+  $X$ is a collection of vertex disjoint in-trees, and for each
+  in-tree, the root is a maximal upper bound in $\primeless$.
 \end{proof}
-Let the partition of $X$ given by Theorem \ref{thm:partition} be
-$\{X_1,\ldots,X_r\}$.  Further, each in-tree itself can be layered
-based on the distance from the root.  The root is considered to be at
-level zero. For $j \geq 0$, Let $X_{i,j}$ denote the set of prime
-matrices in level $j$ of in-tree $X_i$.
+
+We call the partitioning of set of prime matrices $\bP$ given by
+Theorem~\ref{thm:partition} as {\em containment partitioning}. The term
+signifies the containment property described in
+Lemma~\ref{lem:containment} and the fact that each containment parts
+are disjoint from one another.
+
+Let the containment partition of set of prime matrices $\bP$ be: 
+\[\{\bP_i \mid i \in [r], \bP_i \text{ is the vertex set of an
+  in-tree of } X\}\]
+
+Further, the vertices of each in-tree can be
+layered based on their distance from the root, the root being the
+maximal upper bound.  The root is considered
+to be at level zero. For $j \geq 0$, let $\bP_{i,j}$ denote the set of
+prime matrices in level $j$ of in-tree $\bP_i$. The set $\bP_{i,0}$
+has only one prime submatrix which is the maximal upper bound in
+$\bP_i$. Hence $\bP_{i,0}$ is denoted by $mub(\bP_i)$.
+
+Now we describe how the chain partitioning alluded to in
+Theorem~\ref{thm:partitionold} is obtained. Each in-tree described
+above is decomposed by the following rules into chains. Traverse the
+tree level by level from bottom to top. At each level, if there is a
+node with degree greater than 2, the chain that contains this node and
+all its descendants is one chain part. Remove this chain from the tree
+and recurse till reaching the root. Now it is clear why each of these
+partitions are totally ordered -- it is so because they are chains of
+the Hasse diagram. Now we continue with the description of
+$\primeless$ for TPLs.
+
+The next Lemma~\ref{lem:subicppa} and Lemma~\ref{lem:rooticppa} show why it
+is necessary to assign a subtree of the target tree only to the $mub$
+of each in-tree and all other matrices reduce to interval assignment
+problems.
+
+% CHAIN PARTITIONS
+% each leaf to a degree >= two vertex...
+% without include the degree >= two vertex
+% that is one chain  
+% do tis for all leaf
+% and remove the chain..
+% iterate
 
@@
   are in correspondence with the prime submatrices of $M$.  Further
-  let $\{X_1,\ldots,X_r\}$ be the partition of $X$ into in-trees as
-  defined above.  Then, matrix $M$ has an ICPPL in tree $T$ \iff $T$
-  can be partitioned into vertex disjoint subtrees $\{T_1, T_2, \dots
-  T_r\}$ such that, for each $1 \leq i \leq r$, the set of prime
-  submatrices corresponding to vertices in $X_i$ has an ICPPL in
-  $T_i$.
+  let $\{\bP_i \mid i \in [r]\}$ be the partition of $X$ into in-trees
+  as defined above.  Then, matrix $M$ has an ICPPL on target tree $T$
+  \iff $T$ can be partitioned into vertex disjoint subtrees $\{T_i
+  \mid i \in [r]\}$ such that, for each $i \in [r]$, the set of prime
+  submatrices $\bP_i$ has an ICPPL on subtree $T_i$.
 \end{lemma}
@@
   Let us consider the reverse direction first.  Let us assume that $T$
-  can be partitioned into $T_1, \ldots, T_r$ such that for each $1
-  \leq i \leq r $, the set of prime submatrices corresponding to
-  vertices in $X_i$ has an ICPPL in $T_i$.  It is clear from the
-  properties of the partial order $\preccurlyeq$ that these ICPPLs
-  naturally yield an ICPPL of $M$ in $T$.  The main property used in
-  this inference is that for each $1 \leq i \neq j \leq r$, $supp(X_i)
-  \cap supp(X_j) = \phi$.
-
-\noindent
-To prove the forward direction, we show that if $M$ has an ICPPL, say
-$\cA$, in $T$, then there exists a partition of $T$ into vertex
-disjoint subtree $T_1, \ldots, T_r$ such that for each $1 \leq i \leq
-r$, the set of prime submatrices corresponding to vertices in $X_i$
-has an ICPPL in $T_i$.  For each $1 \leq i \leq r$, we define based on
-$\cA$ a subtree $T_i$ corresponding to $X_i$.  We then argue that the
-trees thus defined are vertex disjoint, and complete the proof.
-Consider $X_i$ and consider the prime submatrix in $X_{i,0}$.
-Consider the paths assigned under $\cA$ to the sets in the prime
-submatrix in $X_{i,0}$.  Since the component in $G_f$ corresponding
-to this matrix is a connected component, it follows that union of
-paths assigned to this prime-submatrix is a subtree of $T$.  We call
-this sub-tree $T_i$.  All other prime-submatrices in $X_i$ are
-assigned paths in $T_i$ since $\cA$ is an ICPPL, and the support of
-other prime submatrices in $X_i$ are contained in the support of the
-matrix in $X_{i,0}$.  Secondly, for each $1 \leq i \neq j \leq r$,
-$supp(X_i) \cap supp(X_j) = \phi$, and since $\cA$ is an ICPPL, it
-follows that $T_i$ and $T_j$ are vertex disjoint.  Finally, since $|U|
-= |V(T)|$, it follows that $T_1, \ldots, T_r$ is a partition of $T$
-into vertex disjoint sub-trees such that for each $1 \leq i \leq r$,
-the set of matrices corresponding to nodes in $X_i$ has an ICPPL in
-$T_i$.  Hence the lemma.
+  can be partitioned into $T_1, \ldots, T_r$ such that for each $i \in
+  [r]$, the set of prime submatrices $\bP_i$ has an ICPPL on $T_i$.
+  It is clear from the properties of the partial order $\preccurlyeq$
+  that this naturally yield an ICPPL of $M$ on $T$.  The main property
+  used in this inference is that for each $1 \leq i \neq j \leq r$,
+  $supp(X_i) \cap supp(X_j) = \emptyset$.  To prove the forward
+  direction, we show that if $M$ has an ICPPL, say $\cl$, on $T$ then
+  there exists a partition of $T$ into vertex disjoint subtree $T_1,
+  \ldots, T_r$ such that for each $i \in [r]$, the set of prime
+  submatrices $\bP_i$ has an ICPPL on $T_i$.  For each $i \in [r]$, we
+  define based on $\cl$ a subtree $T_i$ corresponding to $X_i$.  We
+  then argue that the subtrees thus defined are vertex disjoint and
+  prove the claim.  In partition $\bP_i$, consider prime submatrix
+  $mub(\bP_i)$ and the paths labeled under $\cl$ to the hyperedges in
+  the prime submatrix $mub(\bP_i)$.  Since the component in $\bO(M)$
+  corresponding to this matrix (this notation is analogous to overlap
+  graph $\bO(\cF)$ for a hypergraph $\cF$) is a connected component,
+  it follows that the paths labeled to hyperedges in this prime
+  submatrix form a (connected) subtree of $T$.  We call this subtree
+  $T_i$.  All other prime submatrices in $\bP_i$ are assigned merely
+  paths in $T_i$. This follows from Lemma~\ref{lem:containment} along
+  with the facts that $\cl$ is an ICPPL and $M_x \primeless
+  mub(\bP_i)$ for any prime submatrix $M_x \in \bP_i$.  Secondly, for
+  each $1 \leq i \neq j \leq r$, $\supt{\bP_i} \cap \supt{\bP_j} =
+  \emptyset$ and $\cl$ is an ICPPL. Thus it follows that $T_i$ and
+  $T_j$ are vertex disjoint.  Finally, since $|U| = |V(T)|$, it
+  follows that $T_1, \ldots, T_r$ is indeed a partition of $T$. Thus
+  the claim is proven.
 \end{proof}
+
 The essence of the following lemma is that an ICPPL only needs to be
@@
 necessary and sufficient condition for a matrix to have COP.  We
-present the pseudo-code to test if $M$ has an ICPPL in $T$.
-\begin{lemma} \label{lem:rooticppa} Let $M$ be a matrix and let $X$ be
-  the directed graph whose vertices are in correspondence with the
-  prime submatrices of $M$.  Further let $\{X_1,\ldots,X_r\}$ be the
-  partition of $X$ into in-trees as defined earlier in this section.
-  Let $T$ be the given tree and let $\{T_1, \ldots, T_r\}$ be a given
-  partition of $T$ into vertex disjoint sub-trees.  Then, for each $1
-  \leq i \leq r$, the set of matrices corresponding to vertices of
-  $X_i$ has an ICPPL in $T_i$ if and only if the matrix in $X_{i,0}$
-  has an ICPPL in $T_i$ and all other matrices in $X_i$ have an {\bf
-    {\em ICPIA}}.
+present the pseudocode to test if $M$ has an ICPPL in $T$.
+
+\begin{lemma} 
+  \label{lem:rooticppa}%
+  Let $M$ be a matrix, $X$ be the directed graph whose vertices are in
+  correspondence with the prime submatrices of $M$ and
+  $\{\bP_1,\ldots,\bP_r\}$ be the partition of $X$ into in-trees by
+  $\primeless$ defined earlier.  Let $T$ be the target tree and let
+  $\{T_1, \ldots, T_r\}$ be a given partition of $T$ into vertex
+  disjoint subtrees.  Then, for each $1 \leq i \leq r$, the set of
+  prime submatrices $\bP_i$ has an ICPPL in $T_i$ if and only if the
+  prime submatrix $mub(\bP_i)$ has an ICPPL on $T_i$ and all other
+  matrices in $\bP_i$ have an ICPIA.
 \end{lemma}
 \begin{proof}\thesisspacing
-  The proof is based on the following fact- $\preccurlyeq$ is a
-  partial order and $X$ is a digraph which is the disjoint union of
+  The proof is based on the fact that $\preccurlyeq$ is a partial
+  order and $X$ is a directed graph which is a disjoint union of
   in-trees.  Each edge in the in-tree is a containment relationship
-  among the supports of the corresonding submatrices. Therefore, any
-  ICPPL to a prime submatrix that is not the root is contained in a
-  path assigned to the sets in the parent matrix.  Consequently, any
-  ICPPL to the prime submatrix that is not at the root is an ICPIA,
-  and any ICPIA can be used to construct an ICPPL to the matrices
-  corresponding to nodes in $X_i$ provided the matrix in the root has
-  an ICPPL in $T_i$.  Hence the lemma.
+  among the supports of the corresponding submatrices. Therefore, any
+  ICPPL to a prime submatrix that is not the $mub$ is contained in a
+  path assigned to the sets in the parent matrix.  This follows from
+  Lemma~\ref{lem:containment} and since $M_x \primeless mub(\bP_i)$
+  for any prime submatrix $M_x \in \bP_i$.  Consequently, any ICPPL to
+  the prime submatrix that is not $mub(\bP_i)$ is an ICPIA, and any
+  ICPIA can be used to construct an ICPPL to the matrices
+  corresponding to nodes in $\bP_i$ provided $mub(\bP_i)$ has
+  an ICPPL in $T_i$. 
 \end{proof}
 Lemma \ref{lem:subicppa} and Lemma \ref{lem:rooticppa} point out two
-algorithmic challenges in finding an ICPPL for a given set system
-$\cF$ in a tree $T$.  Given $\cF$, finding $X$ and its partition
-$\{X_1,\ldots,X_r\}$ into in-trees can be done in polynomial time.  On
-the other hand, as per lemma \ref{lem:subicppa} we need to parition
-$T$ into vertex disjoint sub-trees $\{T_1, \ldots, T_r\}$ such that
-for each $i$, the set of matrices corresponding to nodes in $X_i$ have
-an ICPPL in $T_i$.  This seems to be a challenging step, and it must
-be remarked that this step is easy when $T$ itself is a path, as each
-individual $T_i$ would be sub-paths.  The second algorithmic challenge
-is identified by lemma \ref{lem:rooticppa} which is to assign an ICPPL
-from a given tree to the matrix associated with the root node of
-$X_i$.
+algorithmic challenges in finding an ICPPL for a given hypergraph
+$\cF$ on a target tree $T$.  
+\begin{enumerate}
+\item Finding $X$ and its partition into in-trees
+  $\{\bP_1,\ldots,\bP_r\}$, can be done in polynomial time.  On the
+  other hand, as per Lemma~\ref{lem:subicppa} we need to partition $T$
+  into vertex disjoint subtrees $\{T_1, \ldots, T_r\}$ such that for
+  each $i \in [r]$, the prime submatrices in $\bP_i$ have an ICPPL in
+  $T_i$ (to be accurate, the prime submatrix $mub(\bP_i)$ has an ICPPL
+  in $T_i$ and the rest of the prime submatrices have an ICPIA as
+  claimed in Lemma~\ref{lem:rooticppa}).  This seems to be a
+  challenging step, and it must be remarked that this step is easy
+  when $T$ itself is a path, as each individual $T_i$ would be
+  sub-paths.
+\item The second algorithmic challenge is identified by
+  Lemma~\ref{lem:rooticppa} which is to compute an ICPPL from a given
+  tree to the matrix associated with $mub(\bP_i)$.
+\end{enumerate}
+
+\begin{figure}[h]
+  \centering
+
+  \begin{tabular}[h]{l}
+     \tplmatrix %        
+%      \\     
+%      \tpltreepaths %       
+  \end{tabular}
+
+  \caption[\figtabsize Partial order on prime submatrices -
+  containment partition]{\figtabsize
+    Partial order on prime submatrices. The table on top illustrates
+    the structural properties of prime matrices and partial order
+    $\primeless$. A hypergraph $\cF$ with FTPL has hyperedges $S_1, \ldots,
+    S_{12}$ which are represented by the binary matrix $M$. Empty
+    entries denote zeros.
+%     The graph
+%     below is the target tree with each set $S_i, 1 \le i \le 12$ shown
+%     as tree paths and intervals. Note that the sets that do not belong
+%     to the $mub$s $M_1$ and $M_4$ are shown as intervals.
+  }
+  \label{fig:tplpotheory-matrix}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+
+  \begin{tabular}[h]{l}
+%      \tplmatrix %        
+%      \\     
+     \tpltreepaths %       
+  \end{tabular}
+
+  \caption[\figtabsize Partial order on prime submatrices - target
+  tree]{\figtabsize This graph is the target tree for the matrix in
+    Fig.~\ref{fig:tplpotheory-matrix} with each column $S_i, 1 \le i
+    \le 12$ shown as tree paths and intervals.}
+  \label{fig:tplpotheory-graph}
+\end{figure}
+
+Figure~\ref{fig:tplpotheory-matrix} and
+Figure~\ref{fig:tplpotheory-graph} illustrate the containment
+partitions of a binary matrix and the claim of
+Theorem~\ref{lem:rooticppa}. In Figure~\ref{fig:tplpotheory-matrix},
+the feasibility bijection is explicitly shown by labeling the rows
+(which represents elements in hypergraph's universe) with the vertices
+of the target tree $T$, $v_1, \ldots, v_{12}$.  $M$ decomposed into
+prime submatrices $M_1, \ldots, M_6$ which is collectively represented
+by $\cP$. It can be verified that they correspond to the overlap
+components in $\bO(\cF)$. The containment partitions of $\cP$ are
+$\bP_1, \bP_2$ as shown. Note that the empty cells of the matrix
+represent 0s -- they are left empty to make the containment partitions
+more obvious.  Figure~\ref{fig:tplpotheory-graph} shows the target
+tree of this TPL. The rectangles depict the sets allocated to the
+columns in the $mub$ of $M$'s prime matrices, namely $M_1$ and
+$M_4$. The rest of the columns are shown as intervals illustrating
+Theorem~\ref{lem:rooticppa}.
+
 \begin{algorithm}[h]
@@
   \label{al:icppa-main}
-  \begin{algorithmic}
+  \begin{algorithmic}[\lndisplay]
     \STATE Identify the prime submatrices. This is done by
     constructing the strict overlap graph and identify connected
-    components.  Each connected component yields a prime submatrix.
-    \\ 
+    components.  Each connected component yields a prime submatrix.%
     \STATE Construct the partial order $\preccurlyeq$ on the set of
-    prime submatrices.  \\ 
-    \STATE Construct the partition $X_1,\ldots,X_r$ of the prime
-    submatrices induced by $\preccurlyeq$ \\
-    \STATE For each $1 \leq i \leq r$, Check if all matrices except
-    those in $X_{i,0}$ has an ICPIA.  If a matrix does not have ICPIA
-    exit with a negative answer.  To check for the existence of ICPIA,
-    use the result in \cite{nsnrs09}.  \STATE Find a partition of
-    $T_1, \ldots, T_r$ such that matrices in $X_{i,0}$ has an ICPPL in
-    $T_i$.  If not such partition exists, exit with negative answer.
+    prime submatrices.%
+    \STATE Construct the partition $\bP_1, \ldots, \bP_r$ of the prime
+    submatrices induced by $\preccurlyeq$%
+    \STATE For each $1 \leq i \leq r$, check if all matrices except
+    $mub(\bP_i)$ has an ICPIA.  If a matrix does not have ICPIA exit
+    with a negative answer.  To check for the existence of ICPIA, use
+    the result in \cite{nsnrs09}.%
+    \STATE \label{ln:findtreepartition} Find a partition of $T_1, \ldots, T_r$ such that matrices
+    in $mub(\bP_i)$ has an ICPPL in $T_i$.  If not such partition
+    exists, exit with negative answer.%
+
+    \FOR {each $1 \leq i \leq r$}%
+    \STATE \label{ln:findrooticppl} Find the ICPPL for $mub(\bP_i)$ on $T_i$%
+    \FOR {each prime submatrix $M_x$ other than $mub(\bP_i)$ in $\bP_i$}%
+    \STATE Find ICPIA for $M_x$ %     
+    \ENDFOR
+    \ENDFOR
   \end{algorithmic}
@@
 
-\section[Complexity]{Complexity of Tree Path Assignment-A Discussion}
+Algorithm~\ref{al:icppa-main} consolidates all that we described above
+and gives the procedure to solve \CFTPL. It may be noted that all operations other than in
+line~\ref{ln:findtreepartition} and line~\ref{ln:findrooticppl} are
+known to be polynomial time solvable.
+
+\clearpage % to flush all floats
+
+\section[Complexity]{Complexity of Tree Path Assignment -- A Discussion}
 \label{sec:tplcomplexity}
-Recall that the input to the Tree Path Assignment question is an order
-pair $(\cF,T)$ where $\cF$ is a family of subsets of an universe $U$,
-and $T$ is a tree such that $|V(T)| = |U|$.  The question is to come
-up with a bijection from $U$ to $V(T)$ such that the image of each set
-in $\cF$ is a path in $T$.
-
-\tnote[dire]{it is known that if the given tree is a path a
-feasible assignment can be found in polynomial time, and we observe
-that it can actually be done in logspace. }
-
-We also point out Consecutive Ones Testing is in Logspace from two
-different results in the literature \cite{kklv10, mcc04}. To the best
-of our knowledge this observation has not been made earlier.
+
+In this section, we briefly make some observations about complexity of
+TPL and \COP. First we present our conclusions on time complexity of
+TPL and then on space complexity of \COP. The latter is a conclusion
+based on existing literature. However to the best of our knowledge
+this particular observation about \COP's space complexity has not been
+made earlier.
+
+\subsection{Time complexity of TPL}
+\label{sec:tpltime}
+Complexity of \CFTPL for arbitrary trees is in \np. If a solution to
+\CFTPL is given, it can be verified using \ICPPL and computing bounded
+(in this case, up to 3-way) intersection cardinalities is polynomial
+time solvable. Hence the claim. Beyond this, however, we do not have
+conclusive complexity results for \CFTPL.
+
+Nonetheless, the problems \CFTPLINT and \CFTPLKTREE are polynomially
+solvable as we saw in Section~\ref{sec:spltargettree}.
+
+Further, we observe that \COP testing can in fact be done in
+logspace. We use two different results in the literature \cite{kklv10,
+  mcc04} to conclude that \COP is in logspace. We see this in the next
+section.
 
@@
 \label{sec:cotlogspace}
-While Tree Path Assignment is isomorphism-complete, it is polynomial
-time solvable when the given tree is a path.  Indeed, in this case we
-encounter a restatement of matrices with the COP.  The known
-approaches to testing for COP fall into two categories: those that
-provide a witness when the input matrix does not have COP, and those
-that do not provide a witness.  The first linear time algorithm for
-testing COP for a binary matrix was using a data structure called PQ
-trees, which represent all COP orderings of $M$, invented by
-\cite{bl76}. There is a PQ tree for a matrix \iff the matrix has COP.
-Indeed, this is an algorithmic characterization of the consecutive
-ones property and the absence of the PQ-tree does not yield any
-witness to the reason for failure.  A closely related data structure
-is the generalized PQ tree in \cite{mcc04}.  In generalized PQ tree
-the P and Q nodes are called prime and linear nodes. Aside from that,
+TPL is polynomial time solvable when the given tree is a path. This is
+the case that is equivalent to \COP testing of binary matrices.
+% Indeed, in this case we encounter a restatement of matrices with the
+% COP. 
+The known approaches to testing for COP fall into two categories:
+those that provide a witness when the input matrix does not have COP,
+and those that do not provide a witness.  We mention this aspect of
+\COT in Section~\ref{sec:surveycopopt}.
+Section~\ref{sec:surveycertalgo} described generalized \PQtree where
+$P$ and $Q$ nodes are called prime and linear nodes. Aside from that,
 it has a third type of node called degenerate nodes which is present
-only if the set system does not have COP \cite{mcc04}.  Using the idea
-of generalized PQ tree, \cite{mcc04} proves that checking for
+only if the set system does not have COP.
+Section~\ref{sec:surveycertalgo} further described that using the idea
+of generalized \PQtree, \cite{mcc04} proves that checking for
 bipartiteness in the certain incomparability graph is sufficient to
 check for COP.  \cite{mcc04} invented a certificate to confirm when a
-binary matrix does not have COP.  \cite{mcc04} describes a graph
-called incompatibility graph of a set system $\cF$ which has vertices
-$(a,b), a \ne b$ for every $a, b \in U$, $U$ being the universe of the
-set system. There are edges $((a,b),(b,c))$ and $((b,a),(c,b))$ if
-there is a set $S \in \cF$ such that $a, c \in S$ and $b \notin S$. In
-other words the vertices of an edge in this graph represents two
-orderings that cannot occur in a consecutive ones ordering of $\cF$.
-\begin{theorem}[Theorem 6.1, \cite{mcc04}]
-  Let $\cF$ be an arbitrary set family on domain $V$. Then $\cF$ has
-  the consecutive ones property if and only if its incompatibility
-  graph is bipartite, and if it does not have the consecutive ones
-  property, the incompatibility graph has an odd cycle of length at
+binary matrix does not have COP.  Recall \cite{mcc04}'s
+incompatibility graph of a set system $\cF$ from
+Definition~\ref{def:incompatibility} -- it has vertices $(a,b), a \ne
+b$ for every $a, b \in U$, $U$ being the universe of the set
+system. There are edges $((a,b),(b,c))$ and $((b,a),(c,b))$ if there
+is a set $S \in \cF$ such that $a, c \in S$ and $b \notin S$. In other
+words the vertices of an edge in this graph represents two orderings
+that cannot occur in a consecutive ones ordering of $\cF$.  In the
+same section, we also discuss Theorem~\ref{th:mccbipartite} which
+describes the linear certificate for no \COP on the incompatibility
+graph.  This is restated below in Theorem~\ref{th:mccbipartiterecall}.
+
+\begin{theoremsansproof}[{\cite[Th.~6.1]{mcc04}, \cite[Corrected by
+    Th.~2.6, Proof p. 44--47]{d08phd}}, Theorem~\ref{th:mccbipartite}] 
+  \label{th:mccbipartiterecall}
+  Let $\cF$ be a set family with universe $U$. Then $\cF$ has \COP
+  \iff its incompatibility graph is bipartite and if it does not have
+  the \COP, the incompatibility graph has an odd cycle of length at
   most $n+3$.
-\end{theorem}
-This theorem gives a certificate as to why a given matrix does not
-have COP.  Similarly, the approach of testing for an ICPIA in
-\cite{nsnrs09} also gives a different certificate- a prime submatrix
-that does not have an ICPIA.  Further, the above theorem can be used
-to check if a given matrix has COP in logspace by checking if its
-incompatibility graph is bipartite. \cite{rei84} showed that checking
-for bipartiteness can be done in logspace. Thus we conclude that
-consecutive ones testing can be done in logspace.
-
-\noindent
-More recently, \cite{kklv10} showed that interval graph isomorphism
-can be done in logspace. Their paper proves that a canon for interval
-graphs can be calculated in logspace using an interval hypergraph
-representation of the interval graph with each hyperedge being a set
-to which an interval shall be assigned by the canonization
+\end{theoremsansproof}
+
+Thus checking if a set system/hypergraph/binary matrix's
+incompatibility graph is bipartite is sufficient to conclude the
+existence of \COP. The following result makes this possible in
+logspace. \cite{rei84} showed that checking for bipartiteness can be
+done in logspace. Thus we conclude that consecutive ones testing can
+be done in logspace.
+
+We also observed logspace complexity of \COP using a more recent and
+independent result -- \cite{kklv10} showed that interval graph
+isomorphism can be done in logspace. Their paper proves that a canon
+for interval graphs can be calculated in logspace using an interval
+hypergraph representation of the interval graph with each hyperedge
+being a set to which an interval shall be assigned by the canonization
 algorithm. An overlap graph (subgraph of intersection graph, edges
@@
 the canon for the whole graph is created using logspace tree
-canonization algorithm from \cite{sl92}. The interval labelling done
-in this process of canonization is exactly the same as the problem of
+canonization algorithm from \cite{sl92}. The interval labeling done in
+this process of canonization is exactly the same as the problem of
 assigning feasible intervals to a set system, and thus the problem of
@@
 
-\begin{theorem}[Theorem 4.7, \cite{kklv10}]
+\begin{theorem}[{\cite[Th.~4.7]{kklv10}}]
   \label{th:canonlabel}
   Given an interval hypergraph $\cH$, a canonical interval labeling
-  $l_H$ for $H$ can be computed in FL.
+  $\cl$ for $\cH$ can be computed in {\sc FL}.
 \end{theorem}
 
-We present the following reduction to see that COP testing is indeed
-in logspace. Given a binary matrix $M$ of order $n \times m$, let $S_i
-= \{j \mid M[j,i]=1 \}$. Let $\cF = \{S_i \mid i \in [m] \}$ be this
-set system. Construct a hypergraph $\cH$ with its vertex set being
-$\{1, 2, \dots n\}$. The edge set of $\cH$ is isomorphic to
-$\cF$. Thus every edge in $\cH$ represents a set in the given set
-system $\cF$. Let this mapping be $\pi: E(\cH) \rightarrow \cF$. It is
-easy to see that if $M$ has COP, then $\cH$ is an interval
-hypergraph. From theorem \ref{th:canonlabel}, it is clear that the
-interval labeling $l_{\cH}: V(\cH) \rightarrow [n]$ can be calculated
-in logspace. Construct sets $I_i = \{ \l_{\cH}(x) \mid x \in E, E \in
-E(\cH), \pi(E) = S_i\}$, for all $i \in [m]$. Since $\cH$ is an
-interval hypergraph, $I_i$ is an interval for all $i \in [m]$, and is
-the interval assigned to $S_i$ if $M$ has COP.
-
-Now we have the following corollary.
+We know the equivalence of hypergraphs and set systems (see
+Definition~\ref{def:matrixsetsystem}). Thus interval labeling of
+hypergraphs obviously translates to finding \COP order of a binary
+matrix which leads to the following corollary.
+
+% We present the following reduction to see that COP testing is indeed
+% in logspace. Given a binary matrix $M$ of order $n \times m$, let $S_i
+% = \{j \mid M[j,i]=1 \}$. Let $\cF = \{S_i \mid i \in [m] \}$ be this
+% set system. Construct a hypergraph $\cH$ with its vertex set being
+% $\{1, 2, \dots n\}$. The edge set of $\cH$ is isomorphic to
+% $\cF$. Thus every edge in $\cH$ represents a set in the given set
+% system $\cF$. Let this mapping be $\pi: E(\cH) \rightarrow \cF$. It is
+% easy to see that if $M$ has COP, then $\cH$ is an interval
+% hypergraph. From theorem \ref{th:canonlabel}, it is clear that the
+% interval labeling $l_{\cH}: V(\cH) \rightarrow [n]$ can be calculated
+% in logspace. Construct sets $I_i = \{ \l_{\cH}(x) \mid x \in E, E \in
+% E(\cH), \pi(E) = S_i\}$, for all $i \in [m]$. Since $\cH$ is an
+% interval hypergraph, $I_i$ is an interval for all $i \in [m]$, and is
+% the interval assigned to $S_i$ if $M$ has COP.
 \begin{corollary}
@@
   If a binary matrix $M$ has COP then the interval assignments to each
-  of its columns can be calculated in FL.
+  of its columns can be calculated in {\sc FL}.
 \end{corollary}
 
-\noindent
-Finally, we conclude by asking about the complexity of Tree Path
-Assignment restricted to other subclasses of trees.  In particular, is
-Tree Path Assignment in caterpillars easier than Tree Path assignment
-in general trees.
+Thus we again conclude that \COP testing is indeed in logspace.
 

### 2012-02-24 src/010chap1.tex
spell check completed final version
shit just got really real.

@@
 tree path labeling problem.
-
-% \tnote{testing 1 2 3...}%
-% \tnote[bogus1]{testing 1 2 3...}%
-% \tnote[important]{testing 1 2 3...}% 
-% \tnote[dire]{testing 1 2 3...}% 
-% \tnote[minor]{testing 1 2 3...}% 
-% \tnote[pressing]{testing 1 2 3...}% 
-% \tnote[UNDEF]{testing 1 2 3...}%
 
@@
 earlier sections. Now we will see the formal definition of \COP in
-Definition~\ref{{def:copmatrix}}.
+Definition~\ref{def:copmatrix}.
 
@@
 The equivalence of the problem of testing for the consecutive-ones
-property to the constraint statisfaction problem of interval
+property to the constraint satisfaction problem of interval
 assignment \cite{nsnrs09} or interval labeling \cite{kklv10} is as

### 2012-02-24 src/020chap2.tex
spell check completed final version
shit just got really real.

@@
 In this section and the rest of the chapter we see results that shaped the
-correponding areas respectively for \cop in binary matrices.
+corresponding areas respectively for \cop in binary matrices.
 
@@
 bipartite graphs\footnote{The terminology in \cite{at72} differs. It
-  uses the term {\em graphs with $V_1$-consecutive arragement} instead
+  uses the term {\em graphs with $V_1$-consecutive arrangement} instead
   of {\em convex bipartite graphs}.} and by definition of this graph
@@
 with \cop.  The following are the theorems from \cite{at72} that
-acheived this characterization.
+achieved this characterization.
 
@@
 Theorem~\ref{th:tuckerforbidden} lists the structures in a bipartite
-graph that force one of its vertex partitions to have asteriodal
+graph that force one of its vertex partitions to have asteroidal
 triples -- in other words, it identifies the subgraphs that prevent
@@
   from this tuple there exists a path that avoids the closed
-  neighbourbood of the third vertex.
+  neighbourhood of the third vertex.
 \end{definition}
@@
   \par\noindent%
-  The set $\valid{U,\cF}$ respresents all \COP orders of $\cF$ in $U$.
+  The set $\valid{U,\cF}$ represents all \COP orders of $\cF$ in $U$.
   \dstop
@@
   \item $\cF$ is called {\em complete}\footnote{\cite{mcc04} calls
-      this a {\em weakly partative family}. \label{mcc2}} if the
+      this a {\em weakly partitive family}. \label{mcc2}} if the
     following hold true for every pair of non-orthogonal (overlapping)
@@
   \item \label{def::slashop} The binary operator $/$ is defined as
-    follows. Let $A, B$ be two sets from set sysetm $\cF$.
+    follows. Let $A, B$ be two sets from set system $\cF$.
     \begin{align*}
@@
 \item \label{th::it2}$\overline{\cF} = consec(\alpha)$ for
-  some permutation $\alpha$ on $U$ ($consect(\alpha)$ is a linear order)
+  some permutation $\alpha$ on $U$ ($consec(\alpha)$ is a linear order)
 \item \label{th::it3}$\overline{\cF} = \power{U}$ 
@@
 
-\subsubsection{Comparision with generalized \PQtree}
+\subsubsection{Comparison with generalized \PQtree}
 
@@
 generalizations of \cite{bl76}'s \PQtree and hence produce the \PQtree
-if $\cF$ has \COP. As a whole, all these three data strutures are
+if $\cF$ has \COP. As a whole, all these three data structures are
 largely identical with their differences being notional. We discussed
@@
 comparison study has not been made earlier in the literature.  We
-observe the three data strutures of \PQRtree, \gPQtree and generalized
+observe the three data structures of \PQRtree, \gPQtree and generalized
 \PQtree to be equivalent. The terminologies of the theory of all these
@@
     Complete collection &
-    Weakly partative family
+    Weakly partitive family
     \\\hline
@@
 The necessity condition of Theorem~\ref{th:icpia1} is fairly
-obvious. It turns out that it is sufficient too.  This is demostrated
+obvious. It turns out that it is sufficient too.  This is demonstrated
 by two algorithms which work as follows. The first one iterates over
@@
   element in each ordered pair of $\Pi$. Def.~\ref{def:icpia}) %sets
-\cite[Alg.~1]{nsnrs09}. This is acheived by replacing two overlapping
+\cite[Alg.~1]{nsnrs09}. This is achieved by replacing two overlapping
 assignment pairs $(A_i, B_i)$ and $(A_j, B_j)$ with the partitions of
@@
 \item Run the algorithm mentioned in
-  Propostition~\ref{prop:tuckerdomapprox} to obtain a submatrix $A$
+  Proposition~\ref{prop:tuckerdomapprox} to obtain a submatrix $A$
   which is an approximation of the smallest forbidden submatrix. (If
@@
   solution.)
-\item Run the algorithm from Propostition~\ref{cor:miiitime} to find
+\item Run the algorithm from Proposition~\ref{cor:miiitime} to find
   an induced $M_{III_k}$ and let this be $B$.
@@
 this. While \cite{at72} gives forbidden substructures for \COP, there
-are no efficient algorithms to find these strutures in a matrix. In
+are no efficient algorithms to find these structures in a matrix. In
 fact, \cite{v82} shows that even finding a forbidden submatrix with at
@@
 \COP which is efficiently computable. This is done by way of an {\em
-  incompatibility graph} of a set system (since \COP in
-  matrices and set systems are equivalent problems as seen in
+  incompatibility graph} of a set system (since \COP in matrices and
+set systems are equivalent problems as seen in
 Section~\ref{sec:surveycertalgo} we will use the term set system in
 this section, in line with \cite{mcc04}'s convention). The
-construction of this graph is based on the following observation.
+construction of this graph is based on the following observation. A
+similar observation was made in \cite{co98} in relation to the
+betweenness problem in computational biology.
 
@@
   \in S$ but $b \notin S$, then it is impossible to have a \COP order
-  that will place $b$ in between $a$ and $c$. \tnote[important]{also cite the
-    inbetweenness problem}
+  that will place $b$ in between $a$ and $c$. 
 \end{observation}
@@
 defined based on the membership of elements of $U$ in the sets in
-$\cF$.  \cite{mcc04} creates an incompatiblity graph with vertices
+$\cF$.  \cite{mcc04} creates an incompatibility graph with vertices
 denoting ordered pairs and edges denoting the incompatibility

### 2012-02-24 src/030chap3.tex
spell check completed final version
shit just got really real.

@@
   A path labeling $(\cF, \cl)$ with target tree $T$ as a path is
-  feasible iff it is an ICPIA.
+  feasible \iff it is an ICPIA.
 \end{lemma}
@@
 
-Since their pre-images have been found, all leaves in $T$ are then
+Since their preimages have been found, all leaves in $T$ are then
 pruned away. The leaf labels are removed from the path labeling

### 2012-02-24 src/040chap4.tex
- incorporated new changes to Characterization algorithm. no more exit    statements in algo 1 and 2. however it is not identitcal to n's    version. kept some of the old lemmas because it seems to elucidate    (to me) the thesis. - partial editing done of general tree section (prime matrix theory)    -- TBD - not sure about status of special tree = path section -- TBD
- changed notation for induced path labeling
- changed filter 1 and 2 names.
- poset definition
spell check completed final version
shit just got really real.

@@
 \xclearpage
-
-\def\p{{\sc P}\xspace}
-\def\gicomplete{{\sc GI}-complete\xspace}
-\def\logspace{{\sc logspace\xspace}\xspace}
-\def\ptime{{\sc ptime}\xspace}
-\def\fpplusc{{\sc FP+C}\xspace}
 
@@
 few ideas where tree path labeling, in our experience, shows promising
-scope for futher research.
+scope for further research.
 
@@
 complexity of graph isomorphism is by studying canonization
-methods. While general graph ismorphism remains elusive in terms of
+methods. While general graph isomorphism remains elusive in terms of
 hardness, canonization has been studied for smaller classes of graphs
@@
 conjecture that \CFTPL on general trees cannot do better than
-\gicomplete. Our algorithm in Section~\ref{sec:decompos} is unable to
+\gicomplete. Our algorithm in Section~\ref{sec:norestraint} is unable to
 solve it in \p. It would therefore, be certainly of interest to

### 2012-02-27 src/030chap3.tex
minor changes in TPL arbit trees -- more text around the figure. see diff.

@@
 \begin{enumerate}
-\item Finding $X$ and its partition into in-trees
+\item \label{it:challenge1} Finding $X$ and its partition into in-trees
   $\{\bP_1,\ldots,\bP_r\}$, can be done in polynomial time.  On the
@@
   sub-paths.
-\item The second algorithmic challenge is identified by
-  Lemma~\ref{lem:rooticppa} which is to compute an ICPPL from a given
-  tree to the matrix associated with $mub(\bP_i)$.
+\item \label{it:challenge2} The second algorithmic challenge is identified by
+  Lemma~\ref{lem:rooticppa} which is to compute an ICPPL from the
+  chosen subtree $T_i$
+  to the matrix associated with $mub(\bP_i)$.
 \end{enumerate}
+
+Figure~\ref{fig:tplpotheory-matrix} and
+Figure~\ref{fig:tplpotheory-graph} illustrate the containment
+partitions of a binary matrix and the claim of
+Lemma~\ref{lem:rooticppa}. In Figure~\ref{fig:tplpotheory-matrix}, the
+feasibility bijection is explicitly shown by labeling the rows (which
+represents elements in hypergraph's universe) with the vertices of the
+target tree $T$, $v_1, \ldots, v_{12}$.  $M$ decomposed into prime
+submatrices $M_1, \ldots, M_6$ which is collectively represented by
+$\bP$. It can be verified that they correspond to the overlap
+components in $\bO(\cF)$. The containment partitions of $\bP$ are
+$\bP_1, \bP_2$ as shown. Note that the empty cells of the matrix
+represent 0s -- they are left empty to make the containment partitions
+more obvious.  Figure~\ref{fig:tplpotheory-graph} shows the target
+tree of this TPL. The rectangles depict the tree paths allocated to
+the columns in the $mub$ of $M$'s prime matrices, namely $M_1$ and
+$M_4$. The rest of the columns are allocated intervals illustrating
+Lemma~\ref{lem:rooticppa}.
+
+Note that in this example the matrix has been explicitly labeled with
+the vertices of the tree which indicates the solution but arriving at
+this solution poses the above mentioned challenges.
+Challenge~\ref{it:challenge1} in this example can be seen as
+follows. Figure~\ref{fig:tplpotheory-graph} shows that $\bP_1$ has
+been allocated the top subtree and $\bP_2$ has been allocated the
+bottom subtree. Computationally this is not an easy decision. There is
+no obvious indication (as far as we know) that the tree must be
+decomposed at edge $(v_3, v_{11})$. If this was given, then one can
+make the decision based on the difference between the subtrees, namely
+sizes of the containment parts and the subtrees and conclude that
+$\bP_1$ must be allocated the subtree $\{v_1, \ldots, v_7\}$ and
+$\bP_2$ to the rest. What if the subtrees were of the same sizes, say
+vertex $v_{14}$ in $T$, the fourteenth row in $M$ and column $S_9$ in
+$M$ were not present? One has to try to solve
+challenge~\ref{it:challenge2} given above to conclude that the bottom
+(top) subtree indeed does not give an ICPPL to $\bP_1$ ($\bP_2$) and
+thus $\bP_1$ ($\bP_2$) but be allocated the top (bottom) subtree.
+Solving challenge~\ref{it:challenge2} is not easy (as far as we know)
+either. By brute force, one can try all possible path labeling and
+check for ICPPL. If all of them fail, one must backtrack to
+challenge~\ref{it:challenge1} and try another subtree for the
+containment part $\bP_i$.
+
+Algorithm~\ref{al:icppa-main} consolidates all that we described in
+this section and gives the procedure to solve \CFTPL. It may be noted
+that all operations other than in line~\ref{ln:findtreepartition} and
+line~\ref{ln:findrooticppl} are known to be polynomial time solvable.
 
@@
   containment partition]{\figtabsize
-    Partial order on prime submatrices. The table on top illustrates
+    Partial order on prime submatrices. The table illustrates
     the structural properties of prime matrices and partial order
@@
 
-Figure~\ref{fig:tplpotheory-matrix} and
-Figure~\ref{fig:tplpotheory-graph} illustrate the containment
-partitions of a binary matrix and the claim of
-Theorem~\ref{lem:rooticppa}. In Figure~\ref{fig:tplpotheory-matrix},
-the feasibility bijection is explicitly shown by labeling the rows
-(which represents elements in hypergraph's universe) with the vertices
-of the target tree $T$, $v_1, \ldots, v_{12}$.  $M$ decomposed into
-prime submatrices $M_1, \ldots, M_6$ which is collectively represented
-by $\cP$. It can be verified that they correspond to the overlap
-components in $\bO(\cF)$. The containment partitions of $\cP$ are
-$\bP_1, \bP_2$ as shown. Note that the empty cells of the matrix
-represent 0s -- they are left empty to make the containment partitions
-more obvious.  Figure~\ref{fig:tplpotheory-graph} shows the target
-tree of this TPL. The rectangles depict the sets allocated to the
-columns in the $mub$ of $M$'s prime matrices, namely $M_1$ and
-$M_4$. The rest of the columns are shown as intervals illustrating
-Theorem~\ref{lem:rooticppa}.
-
-\begin{algorithm}[h]
+\begin{algorithm}[htb]
   \caption{Algorithm to find an ICPPL for a matrix $M$ on tree $T$:
@@
 \end{algorithm}
-
-Algorithm~\ref{al:icppa-main} consolidates all that we described above
-and gives the procedure to solve \CFTPL. It may be noted that all operations other than in
-line~\ref{ln:findtreepartition} and line~\ref{ln:findrooticppl} are
-known to be polynomial time solvable.
 

### 2012-02-27 src/010chap1.tex
proof read TILL Sec 3.2.2

@@
   third element $x \in X$ \stt $a \posetr x \posetr b$.  A directed
-  (or undirected, but with direction implied by horizontal positioning
+  (or undirected, but with direction implied by vertical positioning
   of nodes) graph that represents $\posetr$ in terms of its covering
@@
   \label{def:poset} 
-  Let $X$ be a partially ordered set with $\preccurlyeq$ being the
+  Let $X$ be a partially ordered set with $\posetr$ being the
   partial order on $X$.
@@
 %  \item \textbf{hasse diagram}
-  \item An element $X_m \in X$ is a
-    \textbf{maximal upper bound} of $X$ if $\nexists X_q \in X$ such
-    that $X_m \preccurlyeq X_q$.  A maximal upper bound on $X$ is
+  \item An element $x_m \in X$ is a
+    \textbf{maximal upper bound} of $X$ if $\nexists a \in X$ such
+    that $x_m \posetr a$.  A maximal upper bound on $X$ is
     denoted by $mub(X)$.
@@
   \hangindent \defindent
-  \item $G$ is an \textbf{interval graph} if there a set of intervals
-    $\cI$ \stt $G$ is isomorphic to the intersection graph of
+\item $G$ is an \textbf{interval graph} if there exists a set of
+  intervals $\cI$ \stt $G$ is isomorphic to the intersection graph of
   $\cI$. \[G \cong \bI(\cI)\]
   \item $G$ is a \textbf{path graph} if there exists a (undirected
-    unrooted) tree $T$ and a set of paths from $T$, $\cP$ \stt $G$ is
+    unrooted) tree $T$ and a set of paths $\cP$ from $T$ \stt $G$ is
     isomorphic to the intersection graph of $\cP$. \[G \cong
@@
   \item $G$ is a \textbf{chordal graph} if it has no induced cycle of
-    length greater than 3. Moreover, a \textbf{chordal graph} is
-    characterized as an intersection graph of subtrees of a tree.
+    length greater than 3. Moreover, a chordal graph is characterized
+    as an intersection graph of subtrees of a tree.
   \end{enumerate}
@@
 
-The set of interval graphs is a subset of set of path graphs which is
-a subset of set of chordal graphs. In fact, there is another class
-between chordal graphs and path graphs called RDV-graphs (Rooted
+The set of interval graphs is a subset of the set of path graphs which
+is a subset of the set of chordal graphs. In fact, there is another
+class between chordal graphs and path graphs called RDV-graphs (Rooted
 Directed tree where paths intersect in a Vertex) \cite{mw86} which is
 defined exactly as a path graph except that the trees are rooted and
-directed.
+directed. There are other variants of path graphs studied in the
+literature with unrooted directed trees or where intersection of paths
+are in edges rather than vertices (DV, UE, DE, RDE) \cite{gm03}.
+% Recognizing clique graphs of directed edge path 
+% graphs
+% UV DuallyChordal Swarcter and Bornstein [10] 
+% DV DuallyDV Prisner and Swarcter [8] 
+% RDV DuallyRDV Prisner and Swarcter [8] 
+% UE 
+% DE DuallyDE This paper 
+% RDE = RDV DuallyRDV Prisner and Swarcter [8]
 
@@
 submatrices came as a corollary to the discovery that convex bipartite
-graphs are AT-free on at least one of the partitions in
+graphs are AT-free on at least one of the vertex partitions in
 \cite{at72}. The first linear time algorithm for COP testing (COT) was
@@
 The construction of \PQtrees in \cite{bl76} draws on the close
-relationship of matrices with \COP to interval graphs. A PQ tree of a
+relationship of matrices with \COP to interval graphs. A \PQtree of a
 matrix is one that stores all row (column) permutations of the matrix
@@
 columns) of the matrix. This is constructed using an elaborate linear
-time procedure and is also a test for planarity.  PQR trees is a
-generalized data structure based on PQ trees \cite{mm96,mpt98}.
-\cite{tm05} describes an improved algorithm to build PQR
-trees. \cite{wlh02} describes the simpler algorithm for COT. Hsu also
-invented PC trees \cite{wlh01} which is claimed to be much easier to
-implement. \cite{nsnrs09} describes a characterization of
+time procedure and is also a test for planarity.  \PQRtree is a
+generalized data structure based on \PQtrees \cite{mm96,mpt98}.
+\cite{tm05} describes an improved algorithm to build
+\PQRtrees. \cite{wlh02} describes the simpler algorithm for COT. Hsu
+also invented \PCtrees \cite{wlh01} which is claimed to be much easier
+to implement. \cite{nsnrs09} describes a characterization of
 consecutive-ones property solely based on the cardinality properties
@@
 exited stating negative if otherwise, this algorithm gives an evidence
-by way of a certificate of matrix even when it has no COP. This
-enables a user to verify the algorithm's result even when the answer
-is negative. This is significant from an implementation perspective
+by way of a certificate even when it has no COP. This enables a user
+to verify the algorithm's result even when the answer is
+negative. This is significant from an implementation perspective
 because automated program verification is hard and manual verification
@@
 the COP order is the certificate.  The internal machinery of this
-algorithm is related to the weighted betweenness problem
-addressed in \cite{co98}.
+algorithm is related to the weighted betweenness problem addressed in
+\cite{co98}.
 
@@
 of changing of the minimum number of zero entries to {\un}s so that
-the resulting matrix has COP. As mentioned earlier, this problem is
-known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
-using a reduction to the longest path problem, that finding a Tucker's
-forbidden submatrix of at least $k$ rows is NP complete.
+the resulting matrix has COP. This problem is known to be NP complete
+due to \cite{b75-phd}. \cite{v82} proves, using a reduction to the
+longest path problem, that finding a Tucker's forbidden submatrix of
+at least $k$ rows is NP complete.
 
@@
 characterization gives a natural data structure that stores all the
-\remove{relevant} feasible bijections between $U$ and $V(T)$. This
-reduces the search space for the solution considerably from the
-universe of all possible bijections between $U$ and $V(T)$ to only
-those bijections that maintain the characterization.  Further, the
-filtering algorithm is also an efficient algorithm to test if a tree
-path labeling\footnote{The terms {\em tree path labeling} and {\em
-    tree path assignment} are, in informal language,
-  synonyms. Formally, the former refers to the bijection $\cl: \cF
-  \rightarrow \cP$. The latter refers to the set of ordered pairs
-  $\{(S, P) \mid S \in \cF, P \in \cP\}$. $\cP$ is a set of paths on
-  $T$.} is feasible.
+feasible bijections between $U$ and $V(T)$. This reduces the search
+space for the solution considerably from the universe of all possible
+bijections between $U$ and $V(T)$ to only those bijections that
+maintain the characterization.  Further, the filtering algorithm is
+also an efficient algorithm to test if a tree path
+labeling\footnote{The terms {\em tree path labeling} and {\em tree
+    path assignment} are, in informal language, synonyms. Formally,
+  the former refers to the bijection $\cl: \cF \rightarrow \cP$. The
+  latter refers to the set of ordered pairs $\{(S, P) \mid S \in \cF,
+  P \in \cP\}$. $\cP$ is a set of paths on $T$.} is feasible.
 
@@
     hypergraph} is an alternate representation of a set system and
-  will be used in this thesis.}~\footnote{See Section~\ref{ch:prelims}
-  for the formal definition.} and thus is a characterization for COP
-\cite{nsnrs09}. In our work we extend this characterization and find
-that trio-wise intersection cardinality preservation makes a tree path
-labeling\footnote{A {\em tree path labeling} $\cl$ is a bijection of
-  paths from the target tree $T$ to the hyperedges in given hypergraph
-  $\cF$.}~\footnotemark[4] (TPL) feasible, which is a generalization
-of the COP problem. This problem is defined by \FTPL (See
-Section~\ref{sec:myresearchintro} for problem definition). %as follows.
+  will be used in this thesis, in line with \cite{kklv10} (see
+  Section~\ref{ch:prelims} for the formal definition).} and thus is a
+characterization for COP \cite{nsnrs09}. In our work we extend this
+characterization and find that trio-wise intersection cardinality
+preservation makes a tree path labeling\footnote{A {\em tree path
+    labeling} $\cl$ is a bijection of paths from the target tree $T$
+  to the hyperedges in given hypergraph $\cF$ (see
+  Section~\ref{ch:prelims} for the formal
+  definition).}% ~\footnotemark[4]
+(TPL) feasible, which is a generalization of the COP problem. This
+problem is defined by \FTPL (see Section~\ref{sec:myresearchintro} for
+problem definition). %as follows.
 
@@
 captures the trio-wise cardinality property described
-earlier\footnote{See Section~\ref{sec:feasible} for the definition of
-  ICPPL.}. This characterization can be checked in polynomial time.  A
+earlier (see Section~\ref{sec:feasible} for the definition of
+  ICPPL.). This characterization can be checked in polynomial time.  A
 relevant consequence of this constructive procedure is that it is
@@
 problem of computing a feasible TPL for a given hypergraph and target
-tree, if one exists. This problem is the {\CFTPL} problem (See
+tree, if one exists. This problem is the {\CFTPL} problem (see
 Section~\ref{sec:myresearchintro} for problem definition). %, is defined as follows.
@@
 when the hyperedges in the hypergraph $\cF$ have at most $k+2$
-vertices (\CFTPLKTREE -- See Section~\ref{sec:myresearchintro} for
+vertices (\CFTPLKTREE -- see Section~\ref{sec:myresearchintro} for
 problem definition). A couple of examples of {\kstars} are given in
@@
 ray\footnote{The path from a leaf to the root, the vertex with highest
-  degree, is called a {\em ray} of the \kstar.}~\footnotemark[4] are
+  degree, is called a {\em ray} of the \kstar (see
+  Section~\ref{ch:prelims} for the formal definition).} are
 independent except for the root\footnote{The vertex with maximum
-  degree in a {\kstar} is called {\em root}.}~\footnotemark[4] and
-hence can be considered as an independent interval till the root. Our
-algorithm builds on this fact and uses the interval assignment
+  degree in a {\kstar} is called {\em root} (see
+  Section~\ref{ch:prelims} for the formal definition).} and hence can
+be considered as an independent interval till the root. Our algorithm
+builds on this fact and uses the interval assignment
 algorithm\cite{nsnrs09} up until ``reaching'' the root and then uses
@@
 We also have an algorithm for solving {\CFTPL} with no restrictions on
-the target tree or set size which runs in exponential time.  This
-algorithm finds a path labeling from $T$ by decomposing the problem
-into subproblems of finding path labeling of subsets of $\cF$ from
-subtrees of $T$. Given the fact that binary matrices naturally
+the target tree or set cardinality which runs in exponential time.
+This algorithm finds a path labeling from $T$ by decomposing the
+problem into subproblems of finding path labeling of subsets of $\cF$
+from subtrees of $T$. Given the fact that binary matrices naturally
 represent a set system (see Section~\ref{sec:motive}) and that the
-{\em overlap} relation between the sets involved is an obvious
-equivalence relation, $\cF$ quite naturally partitions into
-equivalence classes known as {\em overlap components}. In the context
-of COP, overlap components were used in \cite{wlh02} and
-\cite{kklv10}. Moreover, \cite{nsnrs09} discovered that these
-equivalence classes form a total order in the case of \COP. We extend this to TPL and find
-that when $\cF$ is a path hypergraph\footnote{If there exists an FTPL
-  for a hypergraph $\cF$, it is called a path hypergraph.}, the
-classes can be partially ordered as an in-tree in polynomial
-time. Once $\cF$ is ``broken'' into overlap components, one must
-identify the subtree of $T$ that it needs to map to and this is the
-hard part which is currently open to be solved in polynomial time.
+{\em overlap} relation (see Section~\ref{ch:prelims} for the formal
+definition) between the sets involved is an obvious equivalence
+relation, $\cF$ quite naturally partitions into equivalence classes
+known as {\em overlap components} (see Section~\ref{ch:prelims} for
+the formal definition). In the context of COP, overlap components were
+used in \cite{wlh02} and \cite{kklv10}. Moreover, \cite{nsnrs09}
+discovered that these equivalence classes have a certain total order
+in the case of \COP (see Section~\ref{sec:norestraint} for the formal
+definition).  We extend this to TPL and find that when $\cF$ is a path
+hypergraph\footnote{If there exists an FTPL for a hypergraph $\cF$, it
+  is called a {\em path hypergraph}.}, the classes can be partially ordered
+as an in-tree in polynomial time. Once $\cF$ is ``broken'' into
+overlap components, one must identify the subtree of $T$ that it needs
+to map to and this is the hard part which is currently open to be
+solved in polynomial time.

### 2012-02-27 src/020chap2.tex
proof read TILL Sec 3.2.2

@@
 thesis or to \COP in general. These predominantly pertain to
-characterizations of \COP, algorithmic tests to check for \COP,
-optimization problems on binary matrices that do not have \COP and
-some applications of \COP.
+characterizations of \COP, algorithmic tests to check for \COP
+and certificates for binary matrices that do not have \COP.
 
@@
 certain graphs, like convex bipartite graphs, that are defined solely
-by one of its associated matrix having \COP.  In this section we will
+by one of its associated matrices having \COP.  In this section we will
 see the relevance of \cop to graphs.  To see this we introduce certain
@@
 \begin{definition}
-  {\em Matrices that define graphs. \cite[Def.~2.4]{d08phd}.}  
+  [Matrices that define graphs {\cite[Def.~2.4]{d08phd}}]  
   \label{def:graphmatrices} 
@@
   \begin{enumerate}[{\ref{def:graphmatrices}}--i.]
-  \item \emph{Adjacency matrix} of $G$ is the symmetric $n \times n$
+  \item \textbf{Adjacency matrix} of $G$ is the symmetric $n \times n$
     binary matrix $M$ with $m_{i,j} = \un$ \iff $(v_i,v_j) \in E_G$
     for all $i,j \in [n]$.
-  \item \emph{Augmented adjacency matrix} of $G$ is obtained from its
+  \item \textbf{Augmented adjacency matrix} of $G$ is obtained from its
     adjacency matrix by setting all main diagonal elements to \un, \ie
     $m_{i,i} = \un$ for all $i \in [n]$.
-  \item \emph{Maximal clique matrix} or \emph{vertex-clique incidence
+  \item \textbf{Maximal clique matrix} or \textbf{vertex-clique incidence
       matrix} of $G$ is the $n \times k$ binary matrix $M$ with
@@
     \label{def::maxcliquematrix}
-  \item \emph{Half adjacency matrix} of $H$ is the $n_a \times n_b$
+  \item \textbf{Half adjacency matrix} of $H$ is the $n_a \times n_b$
     binary matrix $M$ with $m_{i,j} = \un$ \iff $(a_i, b_j) \in E_H$.
@@
 
-\begin{definition}{\emph{Graphs that relate to
-      COP.\cite[Def.~2.5]{d08phd}}} %
+\begin{definition}[Graphs that relate to
+  COP {\cite[Def.~2.5]{d08phd}}] %
   \label{def:graphwithcop} %
@@
   \begin{enumerate}[\ref{def:graphwithcop}--i.]
-  \item $G$ is \emph{convex-round} if its adjacency matrix has the
+  \item $G$ is \textbf{convex-round} if its adjacency matrix has the
     \CROP.
-  \item \label{def::concave-round} $G$ is \emph{concave-round} if its
+  \item \label{def::concave-round} $G$ is \textbf{concave-round} if its
     augmented adjacency matrix has \CROP. \cite{bhy00}
-  \item $G$ is an \emph{interval graph} if its vertices can be mapped
+  \item $G$ is an \textbf{interval graph} if its vertices can be mapped
     to intervals on the real line \stt two vertices are adjacent \iff
@@
     \begin{enumerate}[a.]
-    \item $G$ is a \emph{unit interval graph} if it is an interval
+    \item $G$ is a \textbf{unit interval graph} if it is an interval
       graph \stt all intervals have the same length.
-    \item $G$ is a \emph{proper interval graph} if it is an interval
+    \item $G$ is a \textbf{proper interval graph} if it is an interval
       graph \stt no interval properly contains another.
@@
     graphs coincide \cite{rob69, gar07}.
- \item $G$ is a \emph{circular-arc graph} if its vertices can be
+ \item $G$ is a \textbf{circular-arc graph} if its vertices can be
     mapped to a set of arcs on a circle \stt two vertices are adjacent
     \iff their corresponding arcs overlap.
-  \item $H$ is \emph{convex bipartite on columns (rows)} if its half
+  \item $H$ is \textbf{convex bipartite on columns (rows)} if its half
     adjacency matrix has \COP on rows (columns).%
     \label{def::convexbi}
-  \item $H$ is \emph{biconvex bipartite} or \emph{doubly
+  \item $H$ is \textbf{biconvex bipartite} or \textbf{doubly
       convex}\cite{yc95} if its half adjacency matrix has \COP on both
     rows and columns.
-  \item $H$ is \emph{circular convex} if its half adjacency matrix has
+  \item $H$ is \textbf{circular convex} if its half adjacency matrix has
     \CROP.
@@
 questions~(\ref{q:testcop}) and (\ref{q:copperm}) are surveyed in this
-section. Those for question~(\ref{q:copopt}) are discussed in
-Section~\ref{sec:surveycopopt} and question~(\ref{q:nocopcert}) is discussed
-in Section~\ref{sec:surveycertalgo}.
-
-It may be noted that one way to design an algorithm to test for \COP
-is by deriving one from any interval graph recognition algorithm using
-the result \cite{hmpv00} \cite[Th~2.7]{d08phd} which demonstrates how
-such a derivation can be done.  They use something called {\em Lex-BFS
-  ordering} of the vertices of a graph to decide in linear time
-whether the graph is an interval graph. They also show how any
-interval graph recognition algorithm can be used to recognize matrices
-with \COP. This is shown in Theorem~\ref{th:intgraphcop}.
+section. Those for question~(\ref{q:nocopcert}) is discussed in
+Section~\ref{sec:surveycopopt}. Question~(\ref{q:copopt}) is discussed
+in brief in Section~\ref{sec:optcop} -- it is not elaborated in this
+chapter since the focus of this thesis is characterization and
+recognition rather than optimization.
+
+One way to design an algorithm to test for \COP is by deriving one
+from any interval graph recognition algorithm using the result
+\cite{hmpv00} \cite[Th~2.7]{d08phd}, restated below in
+Theorem~\ref{th:intgraphcop}, which demonstrates how such a derivation
+can be done.  They use something called {\em Lex-BFS ordering} of the
+vertices of a graph to decide in linear time whether the graph is an
+interval graph. They also show how any interval graph recognition
+algorithm can be used to recognize matrices with \COP.
 
@@
 However, this does not necessarily yield an efficient algorithm. We
-will see results that directly solve the problem on matrices since it
-is known that questions~(\ref{q:testcop}) and (\ref{q:copperm}) stated
-above for \COP are efficiently solvable.  Table~\ref{tab:cophistory}
-gives a snapshot of these results.
+will see results that directly solve the problem of \COT on matrices
+since it is known that questions~(\ref{q:testcop}) and
+(\ref{q:copperm}) stated above for \COP are efficiently solvable.
+Table~\ref{tab:cophistory} gives a snapshot of these results.
 
@@
   % neighborhood of the third vertex.
-  An {\em asteroidal triple} in a graph $G$ is a tuple of vertices
+  An \textbf{asteroidal triple} in a graph $G$ is a tuple of vertices
   $\{v_1,v_2,v_3\} \subset V$ such that between every pair of vertices
@@
 %\begin{minipage}[h]{5.5in}
-\begin{definition}[\PQtree \cite{bl76, mcc04}]
+\begin{definition} [\PQtree \cite{bl76, mcc04}]
   A \PQtree of matrix $M$ with \COP on columns (rows), is a tree with
@@
 \PQtree at each iteration, nodes are considered from leaves to
-tree. At each node considered, it uses one of nine templates to
+the root. At each node considered, it uses one of nine templates to
 determine how the tree must be altered in the vicinity of this
@@
 
-\begin{definition}[{\em \Cop for set systems}]%
+\begin{definition}[\Cop for set systems]%
   \label{def:matrixsetsystem}%
@@
   column $c_i$ of $M$, set $S_i = \{j \mid m_{ji} = \un \}$. The
-  collection $\cF_M$ is the {\em set system of binary matrix} $M$. The
-  {\em binary matrix for set system} $\cF$ is conversely constructed
+  collection $\cF_M$ is the \textbf{set system of binary matrix} $M$. The
+  \textbf{binary matrix for set system} $\cF$ is conversely constructed
   and denoted by $M^\cF$. Thus, $M^{\cF_M} = M$.%
   \par\noindent%
-  A set system $\cF$ from universe $U$, $|U| = n$ has the {\em \cop}
+  A set system $\cF$ from universe $U$, $|U| = n$ has the \textbf{\cop}
   if there exists a linear order or permutation $\sigma = w_1w_2\ldots
   w_n$ that can be applied to $U$ \stt each set $S \in \cF$ becomes a
-  consecutive subsequence (also termed an {\em interval})
+  consecutive subsequence (also termed an \textbf{interval})
   $w_{i}w_{i+1}\ldots w_{i+k-1}$ on $\sigma$ for some positive integer
@@
 
-\begin{definition}[\em Orthogonal sets \cite{n89, mm96,
+\begin{definition}[Orthogonal sets \cite{n89, mm96,
     mcc04}\footnote{\cite{mcc04} does not use the term ``mutually
@@
   \begin{enumerate}
-  \item $A$ and $B$ are said to have a {\em trivial intersection} if
+  \item $A$ and $B$ are said to have a \textbf{ trivial intersection} if
     $A \cap B$ is $\emptyset$ or $A$ or $B$.  In other words, $A$ and
@@
 
-  \item $A$ and $B$ are called {\em mutually orthogonal} or $A$ is {\em
+  \item $A$ and $B$ are called \textbf{ mutually orthogonal} or $A$ is \textbf{
       orthogonal to} $B$ and vice versa, if they have a trivial
@@
 
-  \item {\em Trivial subsets} of a universe $U$, denoted by $\cT(U)$,
+  \item \textbf{ Trivial subsets} of a universe $U$, denoted by $\cT(U)$,
     are sets that have trivial intersections with any set in
@@
   \item {\em Orthogonal sets of a set
-      system}\footnote{\cite[Def.~3.1]{mcc04} uses the term {\em
+      system}\footnote{\cite[Def.~3.1]{mcc04} uses the term \textbf{
         non-overlapping family} of $\cF$ and denotes it by
@@
 
-  \item $\cF$ is called {\em complete}\footnote{\cite{mcc04} calls
-      this a {\em weakly partitive family}. \label{mcc2}} if the
+  \item $\cF$ is called \textbf{ complete}\footnote{\cite{mcc04} calls
+      this a \textbf{ weakly partitive family}. \label{mcc2}} if the
     following hold true for every pair of non-orthogonal (overlapping)
@@
     $\cF$ that is complete\footnote{\cite[Def.~3.2]{mcc04} calls
-      this the {\em weak closure} of $\cF$ denoted by
+      this the \textbf{ weak closure} of $\cF$ denoted by
       $\cW(\cF)$. \label{mcc3}}
@@
   \item Also note that trivially, sets in $\cT(U)$ are consecutive in
-    any permutation of $U$\footnote{$\emptyset$ is considered
-      consecutive by convention.}.
+    any permutation of $U$ ($\emptyset$ is considered
+      consecutive by convention.).
 \end{enumerate}
@@
 The following theorem gives a very interesting property of orthogonal
-sets and $\overline{\cF}$
+sets and $\overline{\cF}$.
 
@@
 
-The following corollary, Corollary~\ref{cor:pqrrecursive} states how
+The following Corollary~\ref{cor:pqrrecursive} states how
 \PQRtree elegantly fits into this whole theory and help in computing
-all \COP orders of $\cF$.
+all \COP orders of $\cF$, if any.
 
@@
   then a \PQRtree $T$ for $\cF$ can be obtained by replacing the leaf
-  $h$ in $T_1$ by $T_2$, where $h$ is the representative element for $H$.
+  $h$ in $T_1$ by $T_2$, where $h$ is the representative element for
+  $H$ in $T_i$.
 \end{corollary}
@@
 \item \label{th::it2}$\overline{\cF} = consec(\alpha)$ for
-  some permutation $\alpha$ on $U$ ($consec(\alpha)$ is a linear order)
+  some permutation $\alpha$ on $U$ ($consec(\alpha)$ denotes a linear order)
 \item \label{th::it3}$\overline{\cF} = \power{U}$ 
@@
 similar to \cite[esp. Th.~2.1, 3.5. also Th.~3.2, 3.3, 3.4]{mcc04}
-which categorizes the nodes in the above the three cases as {\em
-  prime}, {\em linear} and {\em degenerate}
-respectively. \cite{mcc04}'s generalized \PQtree is created in similar
-ways as \PQRtree above. This tree is, in essence, the Hasse diagram of
-what they call {\em strong elements} of weak closure, $\cW(\cF)$(\ie
-$\overline{\cF}$). Strong elements of a set family are elements that
-do not overlap with any other elements in the family, \ie it is
-orthogonal to all other sets\cite[Def.~3.3]{mcc04}.
+which categorizes the nodes in the three cases of
+Theorem~\ref{th:pqrfinalmm} as {\em prime}, {\em linear} and {\em
+  degenerate} respectively. \cite{mcc04}'s generalized \PQtree is
+created in similar ways as \PQRtree above. This tree is, in essence,
+the Hasse diagram of what they call {\em strong elements} of weak
+closure, $\cW(\cF)$(\ie $\overline{\cF}$). Strong elements of a set
+family are elements that do not overlap with any other elements in the
+family, \ie it is orthogonal to all other sets\cite[Def.~3.3]{mcc04}.
 
@@
 We will now see, in brief, one more generalization of \PQtree before
-seeing another approaches to solving \COP testing called ICPIA which
+seeing another approach to solving \COP testing called ICPIA which
 is a set theoretic characterization of \COP.
@@
     {
-      {\em Prime}: For every $I$, $\emptyset \ne I \subset [k],
+      \textbf{Prime}: For every $I$, $\emptyset \ne I \subset [k],
       \cup_{i \in I}S_i \notin \cF$\\
-      {\em Linear}: There exists an ordering of $[k]$
+      \textbf{Linear}: There exists an ordering of $[k]$
       s.t. $\emptyset \ne I \subset [k]$ implies $\cup_{i \in I}S_i
       \in \cF$ only if members of $I$ are consecutive in the ordering\\
-      {\em Degenerate}: For every $I$, $\emptyset \ne I \subset [k],
+      \textbf{Degenerate}: For every $I$, $\emptyset \ne I \subset [k],
       \cup_{i \in I}S_i \in \cF$\\
@@
 To see it in terms of the theory explained in
-Section~\ref{sec:surveycertalgo}, this results in the weak closure of
+Section~\ref{sec:surveycertalgo}, this results in the set of strong
+elements of the weak closure of
 $\cF$. Let $\cF$, $\cI$ be the set system and interval
-system (A set system where all the sets are intervals.),
+system (a set system where all the sets are intervals),
 respectively, involved in the assignment $\Pi$ and $\cF'$, $\cI'$ be
@@
 that $\cF' \cup \cF$ is $\overline{\cF}$ or the weak closure of
-$\cF$. The analog is true for $\cI \cup \cI'$.  Moreover, this output
+$\cF$ without the set unions. The analog is true for $\cI \cup \cI'$.  Moreover, this output
 $\cF'$, $\cI'$ contains the strong elements of $\overline{\cF}$,
@@
 \COP orders represented by $\Pi$ of $\cF$. The subset relation Hasse
-diagram of $I'$ is created with each node representing the
+diagram of $\cI'$ is created with each node representing the
 corresponding assignment pair of the interval from $\Pi'$. Since all intervals in
@@
 What we saw so far in this section is checking for the feasibility of
-a given interval assignment. ICPIA can also be used to find an
+a given interval assignment. ICPIA can also be used to compute an
 interval assignment when only the set system $\cF$ is given. Part of
@@
 whose overlaps with the rest of the sets in that component forms a
-single inclusion chain (see Definition~\ref{def:poset}. In this
-particular reference, the set of sets is the set of overlaps.)  is
-chosen first and assigned the leftmost\footnote{This is by
-  convention. Rightmost can also be chosen analogously to get a
-  different \COP order.} interval. The next candidate set is chosen
-from this set's overlapping sets. The interval assigned is calculated
-by simple intersection cardinality means. Assume that the next set
-overlaps on one of the ``ends'' of the current set and calculate the
-interval by shifting left or right of the current interval and
-adjusting its size to match the next set's cardinality. Once the next
-set is assigned an interval, the current assigned sets are all checked
-for ICPIA. This is a backtracking algorithm -- if at any point ICPIA
-fails, another overlapping set is chosen. If all overlapping set
-attempts fail ICPIA, a different ``first set'' is chosen. This is a
-remarkably simple algorithm due to its obvious simplicity in
-implementation.
+single inclusion chain (see Definition~\ref{def:poset}; in this
+particular reference, the set of sets is the set of overlaps) is
+chosen first and assigned the leftmost interval. Chosing leftmost
+interval is only a convention -- rightmost can also be chosen
+analogously to get a different \COP order. The next candidate set is
+chosen from this set's overlapping sets. The interval assigned is
+calculated by simple intersection cardinality means. Assume that the
+next set overlaps on one of the ``ends'' of the current set and
+calculate the interval by shifting left or right of the current
+interval and adjusting its size to match the next set's
+cardinality. Once the next set is assigned an interval, the current
+assigned sets are all checked for ICPIA. This is a backtracking
+algorithm -- if at any point ICPIA fails, another overlapping set is
+chosen. If all overlapping set attempts fail ICPIA, a different
+``first set'' is chosen. This is a remarkably simple algorithm due to
+its obvious simplicity in implementation.
+
+The work of this thesis described in Chapter~\ref{ch:myresearch} is
+largely an extension of the ICPIA characterization. In the new work,
+we adopt the notation used in \cite{kklv10} (also described in that
+chapter).
 
@@
 
+% When a \COP test algorithm reports negative, it is not easy to verify
+% this. We already saw that \cite{at72} gives forbidden substructures
+% for \COP. \cite{v82} shows that the maximization problem of finding a forbidden submatrix with at
+% least $k$ rows in an input matrix is NP-complete. This was proven by
+% polynomially reducing the longest path problem for graphs to this
+% problem. For most cases, the minimization problem is more helpful.
+% Section~\ref{sec:forbiddensubalgo} gave a polynomial
+% algorithm for finding minimum size forbidden submatrix.
+
 It is known from Theorem~\ref{th:tuckercop} that when a matrix does
@@
 algorithms discussed in Section~\ref{sec:surveycoptest} outputs the
-forbidden submatrix when the test is negative. This perhaps gives a
-hint about how hard it is to find these substructures in the input
-matrix. Using a reduction to the longest path problem, \cite{v85}
-proves that finding a Tucker's forbidden submatrix of at least $k$
-rows is NP complete. An example of an optimization problem for a
+forbidden submatrix when the test is negative.  Using a reduction to
+the longest path problem, \cite{v82} proves that finding a Tucker's
+forbidden submatrix of at least $k$ rows is NP complete. This is a
+maximization problem.  An example of an optimization problem for a
 matrix without \COP is finding if deleting at most $k$ rows from input
@@
 to find ways to get \COP on the input matrix with certain alterations
-and finding smallest forbidden submatrix is at the heart of these problems.
-
-In this section we will summarize \cite{d08phd}'s approximation
-algorithm for finding forbidden submatrix. We will use $\Tuckerset$ to
-refer to the set of Tucker's forbidden submatrices shown in
+and finding smallest forbidden submatrix is at the heart of these
+problems. This minimization problem of finding smallest forbidden
+submatrix is polynomial time solvable.
+
+In this section we will summarize \cite{d08phd}'s algorithm for
+finding forbidden submatrix. We will use $\Tuckerset$ to refer to the
+set of Tucker's forbidden submatrices shown in
 Figure~\ref{fig:forbiddensubmatrices}.
@@
 sum is minimum and return the rows and columns of $M$ that correspond
-to the vertices in the three shortest paths described earlier for this triple. This
+to the vertices in the three shortest paths calculated earlier for this triple. This
 return value forms a submatrix and it obviously must contain a
@@
 In fact, the approximation mentioned in
-Proposition~\ref{prop:tuckerdomapprox} is the worst case when
-the actual solution $M'$ is of type $M_{IV}$. The approximation is
-slightly better for the other types of submatrices from $\Tuckerset$,
-the best case being an exact solution when $M'$ is $M_{I_k}$ or
-$M_{II_k}$. Thus this algorithm works accurately when the solution is of
-type $M_{I_k}$ or $M_{II_k}$.
+Proposition~\ref{prop:tuckerdomapprox} is the worst case which happens
+to be when the actual solution $M'$ is of type $M_{IV}$. The
+approximation is slightly better for the other types of submatrices
+from $\Tuckerset$, the best case being an exact solution when $M'$ is
+$M_{I_k}$ or $M_{II_k}$. Thus this algorithm works accurately when the
+solution is of type $M_{I_k}$ or $M_{II_k}$.
 
@@
   $M$ that has a minimum number of columns, a minimum number of rows
-  and columns, or a minimum number of entries.
+  and columns, or a minimum number of \un entries.
 \end{theoremsansproof}
@@
 
-In the next section we see yet another forbidden substructure for
-\COP which is easier to compute.
+Note that this algorithm has a factor $\Delta$ in the time complexity
+analysis, $\Delta$ being the maximum number of {\un}s in a row. For
+sparse matrices this is much lesser than the total number of columns
+$m$. When the matrix is not sparse, the worst case is when $\Delta$ is
+almost equal to $m$ (if they are equal quantities, that row can be
+disregarded from input matrix since a row with all {\un}s is trivially
+consecutive). Thus this is a polynomial time algorithm although a more
+efficient term in the running efficiency is desirable.
+
+In the next section we see yet another forbidden substructure for \COP
+which is can be computed in linear time.
 
@@
 
-When a \COP test algorithm reports negative, it is hard to verify
-this. While \cite{at72} gives forbidden substructures for \COP, there
-are no efficient algorithms to find these structures in a matrix. In
-fact, \cite{v82} shows that even finding a forbidden submatrix with at
-least $k$ rows in an input matrix is NP-complete. This was proven by
-polynomially reducing the longest path problem for graphs to this
-problem.
-
 \cite{mcc04} describes another characterization for matrices without
-\COP which is efficiently computable. This is done by way of an {\em
+\COP which is linearly computable. This is done by way of an {\em
   incompatibility graph} of a set system (since \COP in matrices and
@@
 since $b \notin S$ which is a contradiction to the assumption that
-this is a \COP order.  The ordered pair $(x,y)$ denotes that in the
-\COP order $y$ comes after $x$. Based on above observations, this
-means $(a,b)$ and $(b,c)$ are not compatible in any \COP order. So are
-$(c,b)$ and $(b,a)$.  Thus a binary relation {\em is incompatible
-  with} on $U \times U \setminus \{(a,a) \mid a \in U\}$ can be
-defined based on the membership of elements of $U$ in the sets in
-$\cF$.  \cite{mcc04} creates an incompatibility graph with vertices
-denoting ordered pairs and edges denoting the incompatibility
+this is a \COP order.  Let the ordered pair $(x,y)$ denote the
+condition that in the \COP order $y$ comes after $x$. Based on above
+observations, this means $(a,b)$ and $(b,c)$ are not compatible in any
+\COP order. So are $(c,b)$ and $(b,a)$.  Thus a binary relation {\em
+  is incompatible with} on $U \times U \setminus \{(a,a) \mid a \in
+U\}$ can be defined based on the membership of elements of $U$ in the
+sets in $\cF$.  \cite{mcc04} creates an incompatibility graph with
+vertices denoting ordered pairs and edges denoting the incompatibility
 relation.
 
-\begin{definition}[{\em Incompatibility graph \cite[Def.~6.1]{mcc04}}]
+\begin{definition}[{Incompatibility graph \cite[Def.~6.1]{mcc04}}]
   \label{def:incompatibility}
@@
 exactly half the vertex set of $G_I^\cF$ since the \COP order must
-involve all elements in $U$. Secondly, the reverse of a \COP order is
+involve all elements in $U$ and each vertex denotes a pair of elements. Secondly, the reverse of a \COP order is
 also a \COP order. Thus $A_\cF \setminus I$ must also be an

### 2012-02-27 src/030chap3.tex
proof read TILL Sec 3.2.2

@@
 find a feasible TPL, if any, in this case.  Finally, in Section
-\ref{sec:tplcomplexity} we discuss the complexity of the algorithms
-seen thus far and we conclude by showing that \cop testing, which is a
+\ref{sec:tplcomplexity} we discuss the complexity of \TPL
+and we conclude by showing that \cop testing, which is a
 special case of \TPL, is in \logspace.
@@
 respectively) which permutes each set to an interval. The result in
-\cite{nsnrs09} (see Theorems~\ref{th:icpia2} and \ref{th:icpia1})
+\cite{nsnrs09} (see Theorems~\ref{th:icpia1} and \ref{th:icpia2})
 characterizes interval assignments to the sets which can be obtained
@@
 \kstars, respectively.  Section~\ref{sec:spltargettree} discusses
-these problems and their polynomial time solutions.  \CFTPLINT is a
-demonstrated to be the known problem of \COP testing. The polynomial
-time solution to \CFTPLKTREE is a new result by us.
+these problems and their polynomial time solutions.  \CFTPLINT is the
+known problem of \COP testing. The polynomial time solution to
+\CFTPLKTREE is a new result by us.
 
@@
 
-An algorithm for \CFTPL on general trees which is less efficient than
-polynomial time is presented in
-Section~\ref{sec:norestraint}.
+An algorithm for \CFTPL on arbitrary trees is presented in
+Section~\ref{sec:norestraint} which would run in polynomial time if
+two subproblems we identify are polynomial time solvable.
 

### 2012-02-28 src/002ack.tex
finished proof reading. added ack.

\xclearpage
\acknowledgements

Like any body of work, big or small, this thesis and the study behind
it could not have been done by me without the support of many people
and things professionally and personally. I believe everyone and
everything I encountered in the last two years and seven months has directly or
indirectly helped me conclude this work. Following are the most
obvious and direct help that I explicitly acknowledge.

I have deep gratitude for my research advisor
Dr.\;N.\;S.\;Narayanaswamy. I have been very fortunate to be
associated with a teacher with such remarkable patience and tenacity
with his students that brings out the possible best in them. He let us
work with the right kind of liberty that is needed to discover and
practice one's own method of research and discipline.  I would like to
thank him much for all his brilliant ideas, many discussions (at times
repeated and with no judgement) and perpetual words of
encouragement. I doubt if I would have learned my research without his
persevering guidance.  I would like to thank all fellow students in my
research group for frequent technical discussions and faithful
reviews; also my friends outside AIDB lab, for all caffeine-induced
discussions and de-stressing arguments.

Institute of Mathematical Sciences, for all the mind-bending lectures
that set a foundation to my research. I would especially like to thank
Prof.\;R.\;Ramanujam for motivating and encouraging my interest in
interdisciplinary research.

My husband, Zabil, for his amazing support and encouragement to take a
break from my career and pursue my higher education. Moreover, for
putting up patiently with an approximate solution to our two-body
problem in these years.

My parents, Srinivasan and Seena, for always being there and helping
me in ways only parents can help. My sister, Renu, for her precocious optimism. Devi and Rinchen, for being a
hotline for procrastination-induced panic attacks and reminding me
that I can indeed complete my M.\;S.\;! LMN -- Lakshmi, Mahathi and
Nari, for being the closest thing to a family on campus always and
especially whenever I defaulted to being a hermit (usually a famished
one).  Jai, J.\;K.\;, Shivani and Sourab for having an open door
policy for me any day any time that I was in Chennai and incentivizing
my visits with great fun and awesome food.

And last but certainly not the least, IITM and its wonderful campus
for inspiring me to unlearn my old methods of learning and helping me
discover and enjoy research.

### 2012-02-28 src/030chap3.tex
finished proof reading. added ack.

@@
   \bI(\cP)$.  Then the associated bijection $\cl: \cF
-  \rightarrow \cP$ due to this isomorphism is called a {\em path
+  \rightarrow \cP$ due to this isomorphism is called a \textbf{path
     labeling} of $\cF$ from $T$. The path system $\cP$ is
@@
   % Definition~\ref{def:pathhypergraph}.
-  Then there is an {\em induced path labeling} % denoted by $\cl_\phi:
+  Then there is an \textbf{induced path labeling} % denoted by $\cl_\phi:
 %   \cF \rightarrow \cP$ 
@@
   Let $\cF$ be a hypergraph and $(\cF, \cl)$ be a path
-  labeling. $(\cF, \cl)$ is called a {\em feasible path labeling} if
+  labeling. $(\cF, \cl)$ is called a \textbf{ feasible path labeling} if
   the following conditions hold.
@@
 
-Now we will see the definition of a {\em path hypergraph} which is a
+Now we will see the definition of a \emph{ path hypergraph} which is a
 simple idea extended from interval hypergraphs in \cite{kklv10}.
@@
   \label{def:pathhypergraph}
-  Let $\cF$ be a set system. $\cF$ is called a {\em path hypergraph}
+  Let $\cF$ be a set system. $\cF$ is called a \textbf{path hypergraph}
   if there exists a target tree $T$ and a path system $\cP$ from $T$
   \stt $\cF \cong \cP$. If $\cF$ is thus a path hypergraph, $\cP$ is
-  called {\em path representation} of $\cF$.
+  called \textbf{path representation} of $\cF$.
 \end{definition}
@@
   \label{def:ksubstar}
-  A {\em $k$-subdivided star} is a star with all its rays subdivided
-  exactly $k$ times. A {\em ray} of a $k$-subdivided star is the path
+  A \textbf{$k$-subdivided star} is a star with all its rays subdivided
+  exactly $k$ times. A \textbf{ray} of a $k$-subdivided star is the path
   from the center vertex to a leaf. It is clear that all rays of a
@@
 \begin{definition}[In-tree]
-An {\em in-tree} is a directed rooted tree in which all edges are
+An \textbf{in-tree} is a directed rooted tree in which all edges are
 directed toward to the root.  
@@
 In-trees are used in Section~\ref{sec:norestraint} which describes our
-(albeit non-polynomial time) solution to \CFTPL problem.
+solution to \CFTPL problem.
 
@@
   Consider a path labeling $(\cF, \cl)$ on the given tree $T$. We call
-  $(\cF, \cl)$ an {\em Intersection Cardinality Preserving Path
+  $(\cF, \cl)$ an \textbf{Intersection Cardinality Preserving Path
     Labeling (ICPPL)} if it has the following properties.
@@
   feasible.  Now consider the case where $(\cF, \cl)$ is feasible, \ie
-  there exists a hypergraph isomorphism $\phi$ such that $\cl_\phi =
-  \cl$. Lemma~\ref{lem:noexit1} and Lemma~\ref{lem:xnotempty} show us
-  that \filteri and \filterii do not exit if input is feasible. Thus
-  Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
-  function.%\qed
+  there exists a hypergraph isomorphism $\phi$ such that $\cp_\phi =
+  \cl$. 
+  Algorithm~\ref{al:icppl-find-isomorph} then clearly returns a non-empty
+  function because the target tree is non-empty.%\qed
 \end{proof}
@@
 the algorithm arbitrarily chooses them from $X = \{\xLu, \xSa, \xSc\}$
-and for $\xSn, \xWo$ arbitrarily chooses from $X = \{\xSn,
+and for $\xSn, \xWo$ arbitrarily choosen from $X = \{\xSn,
 \xWo\}$. Here we show only one such choice and clearly one could have
@@
 given hypergraph is a path hypergraph with respect to a given target
-tree\footnote{A larger problem would be to find if a given hypergraph
-  is a path hypergraph with respect to any tree. This problem is not
-  addressed in this thesis.}.  In other words, the problem is to find
-out if a feasible tree path labeling exists from a given target tree
-for a given hypergraph. In this section we will see two special cases
-of this problem where the target tree is from a particular family of
-trees. The first one, where the tree is a path as described by the
-\CFTPLINT problem (see Section~\ref{sec:myresearchintro}) is a shown
-to be equivalent to the well studied problem of consecutive-ones in
-Section~\ref{sec:icpplicpia}. The second one, where the tree is a
-$k$-subdivided tree as described by \CFTPLKTREE (see
-Section~\ref{sec:myresearchintro}), has been solved using a polynomial
-time algorithm. The latter problem enforces some conditions on the
-hypergraph too which will be seen in Section~\ref{sec:ksubdivstar}.
+tree\footnote{A larger problem would be to test if a given hypergraph
+  is a path hypergraph where target tree is not given as input. This
+  problem is not addressed in this thesis and is discussed as further
+  research in Chapter~\ref{ch:conclusion}.}.  In other words, the
+problem is to find out if a feasible tree path labeling exists from a
+given target tree for a given hypergraph. In this section we will see
+two special cases of this problem where the target tree is from a
+particular family of trees. The first one, where the tree is a path as
+described by the \CFTPLINT problem (see
+Section~\ref{sec:myresearchintro} for definition) in
+Section~\ref{sec:icpplicpia} is known to be equivalent to the well
+studied problem of consecutive-ones. The second one, where the tree is
+a $k$-subdivided tree as described by \CFTPLKTREE (see
+Section~\ref{sec:myresearchintro} for definition), has been solved
+using a polynomial time algorithm. The latter problem enforces some
+conditions on the hypergraph too which will be seen in
+Section~\ref{sec:ksubdivstar}.
 
@@
 
-In this section the given hypergraph, the $k$-subdivided star and the
-root of the star are denoted by $\cO$, $T$ and vertex $r$,
-respectively.
 % For each hyperedge $X \in \cO$, we will maintain a 2-tuple of
@@
 % \end{align*}
-
-The set $\cO_i$ refers to the set of hyperedges $\cT_1^i \cup \cT_2^i$
-in the $i$th iteration.  Note that $\cO_1 = \cO$.  In the $i$th
-iteration, hyperedges from $\cO_i$ are assigned paths from $R_i$ using
-the following rules. Also the end of the iteration, $\cL_1^{i+1},
-\cL_2^{i+1}, \cT_1^{i+1}, \cT_2^{i+1}$ are set to $\cL_1^{i},
-\cL_2^{i}, \cT_1^{i}, \cT_2^{i}$ respectively, along with some
-case-specific changes mentioned in the rules below.
 

### 2012-03-02 src/declarationtext.tex
additional files created for submission

%\certificate
\xclearpage% \cleardoublepage
\setcounter{page}{1}

%\def\certheadingstr{Thesis Certificate}%{\sc \centering Thesis Certificate}
\section*{Declaration by Research Scholar}

\vspace*{0.5in}

\noindent 
This is to declare that the thesis titled {\bf
  \mythesistitleNOTSYNPOSIS} submitted by me to the {\bf Indian
  Institute of Technology Madras}, for the award of the degree of {\bf
  \mydegree}, is a bona fide record of the research work done by me
under the supervision of {\bf Dr.~\myadvisor}, Associate Professor at
Department of \mydept, IIT Madras.  The contents of this thesis, in
full or in parts, have not been submitted and will not be submitted to
any other Institute or University for the award of any degree or
diploma.

\vspace*{1.5in}

\begin{singlespacing}
  \hspace*{-0.25in}
  \parbox{3in}{
    \noindent {Anju Srinivasan} \\
    \small
    \noindent Research Scholar} \hfill
  \parbox{1.25in}{               % {1.75in}
    \small
    \vspace*{0.75in}
    \begin{tabular}[h]{r}
    \noindent {Chennai}\\         
    \noindent {\mythesissubmissiondate} 
    \end{tabular}
  }
\end{singlespacing}

\vspace*{1.5in}

\begin{singlespacing}
  \hspace*{-0.25in}
  \parbox{3in}{
    \noindent{\em Forwarded:}\\
    \noindent {\bf Dr.~\myadvisor} \\
    \small
    \noindent Research Guide} 
\end{singlespacing}

### 2013-03-21 src/response_to_reviewers.tex
Reviewer 1 (CRS) changes made.

\documentclass{article}
\usepackage{../lib/mythesislib}
\usepackage{../lib/mylatexlib}
\begin{document}

\title{Response to M.S. Thesis Reviewers\\ 
 {\em Thesis Title ``Generalization of the Consecutive-ones Property''}}
\author{Anju Srinivasan}
\date{25 March 2013}

\maketitle

\section{Reviewers}

\begin{enumerate}
\item  {\bf Dr. C. R. Subramaniam}\\ Institute of Mathematical Sciences, Chennai
\item  {\bf Dr. Subhas Chandra Nandy}\\ Indian Statistical Institute, Kolkata
\end{enumerate}

\section{Response to Reviewer 1}

\subsection{Observations}
The following are a few observations mentioned in the review document
with my responses/changes.

\begin{enumerate}
\item Observation: (in Page 1, second last paragraph) {\bf The result on characterization of TPL in
    the thesis is a variant of the 1978 JCT(B) result due to Fournier
  who showed that two hypergraphs $H_1 = (V_1, X)$ and $H_2 = V_2, Y$
such that $|X| = |Y| = m$ are isomorphic iff there is a bijection
$\phi$ on $I = \{1, ... , m\}$ s.t. for every $L \subset I,
|\bigcap_{i \in L}X_i| = |\bigcap_{j=\phi(i),i \in L}Y_j|$} \\
Response: This is an important observation and was not cited in the thesis. Also \cite{f80}
generalizes \cite{br72} and \cite{fg65} by characterizing the
isomorphism of two hypergraphs by means of equicardinality of certain
edge intersections and the exclusion of certain pairs of
subhypergraphs. TPL characterization is a special case of this with
one of the hypergraphs having hyperedges that are paths from a tree
and the characterization only uses edge intersections of at most 3
hyperedges. This has been added in the Conclusions chapter Page 72
under {\em Graph isomorphism and logspace canonization}.

\item Observation: (in Page 2, first paragraph) {\bf A $k$-subdivided tree is
  a tree with a single central vertex with a number of paths each of
  $k$ edges emanating from it.}\\
Response: $k$-subdivided as defined in this thesis (Section
3.2.4, Page 43) is such that each
of the aforementioned paths are of length ``$k+1$'' edges since each such
path has $k$ nodes apart from the central node and the leaf node, i.e.
$k+2$ nodes. No change in the thesis.
\end{enumerate}

\subsection{Comments}
Responses to the comments in review document Page 2 items 0 to 7.

\begin{enumerate}
\item[0] Comment: {\bf In thesis Page 7, Definition 1.3.7: After $x_m
  \le a$, you need to add $x_m \ne a$.}\\ 
Response: True, earlier definition did not accomodate the fact $x_m
\le x_m$. Suggested change made in Page 7 to the definiton of Poset.

\item Comment: {\bf In thesis Page 7, Definition
    1.3.8: $E$ is a set of ``unordered'' pairs. You need to add the
    word ``unordered''}\\
Response: Suggested change made in to definition of Undirected Graph Page 7 Definition 1.3.8 (1).

\item Comment: {\bf In thesis Page 8, A tree is a {\em connected} and
    acyclic graph}\\
Response: The word ``connected'' has been added to the definition of Tree in
Page 7 Definition 1.3.8 (3).

\item Comment: {\bf In thesis Page 8, Definition 1.3.10 you need to
    add that one is looking at intervals of integers. Also define
    isomorphism between graphs for the sake of completeness}.\\
Response: Definition of Interval graph has been changed to say ``a set
of intervals (of integers)'' in Page 8, Definition 1.3.10. Graph
isomorphism definition has been added in Page 7, Definition 1.3.8(4)
as follows: A \textbf{graph isomorphism} between two graphs $G_1$ and
    $G_2$ is a bijection $\phi: V(G_1) \rightarrow V(G_2)$ such that
    $u, v \in V(G_1)$ are adjacent in $G_1$ if and only if $\phi(u),
    \phi(v)$ are adjacent in $G_2$. If such a bijection exists, it is
    said that $G_1$ is \textbf{isomorphic} to $G_2$, denoted by $G_1
    \cong G_2$.

\item Comment: {\bf In thesis Page 8, Definition 1.3.10 definition of
path graph: You need to emphasise that $\cP$ forms a set system}\\
Response: This change has been made - ``a set system of paths $\cP$''
in Page 8 Definition 1.3.10(2).

\item Comment: {\bf In thesis Page 12, third para of Section 1.5: you
    need to write it more precisely and carefully conveying the
    idea. As it is it is not clear what you are saying.}\\
  Response: The problem {\sc Compute feasible tree path labeling} and
  its result have been described formally and precisely in Chapter 3
  (New results). The aforementioned paragraph describes the same
  informally. These informal sentences have been footnoted with the
  corresponding definitions and theorems from Chapter 3. This change
  is in Page 12.

\item Comment: {\bf In thesis Page 28, Definition 7 (of binary
    operator /): it is not defined precisely. What do you mean by
    ``$b$ is a new element added..''?}.\\
Response: This was in thesis Page 26. $A/B$ is created by removing the elements of second operand set $B$ from
the first operand set $A$ and introducing a new element {\em not} in the
universe to $A$ to represent set $B$. This is a set operation 
mentioned in \cite{mm96} (they use ``$B$'' to denote the new element to represent
set $B$ which is also confusing) which is eventually used to decompose
permutation of elements in $A$ to permutation of elements in $A/B$ and
permutation of elements in $B$.
Perhaps it would be clearer if
$b$ is replaced by $x_B$. This change has been made in Page 26
Definition 2.2.4 (7) and other mentions of this operation ``/'' Page
27 item (iii) and $h$ has been replaced with $x_H$ in Corollary
2.2.8. There was also a typo in Page 27 paragraph under Corollary
2.2.8 item (ii) where ``$U \setminus H$'' was written instead of $U / H$.
This has also been corrected.

\item Comment: {\bf In thesis Page 39 Definition 2.3.1: the definition
  of $A_\cF$ is not clear. Where is the dependence on $\cF$?}\\
Response: This was in thesis Page 36. It is true, there is no
dependence in $\cF$. It is only dependent on the universe $U$ of
$\cF$. All occurences of $A_\cF$ has been changed to $A_\U$ in Page 36.
\end{enumerate}

\section{Response to Reviewer 2}

\bibliographystyle{alpha}%
\bibliography{../lib/cop-variants__thesis} %

\end{document}

### 2013-03-21 src/010chap1.tex
Reviewer 1 (CRS) changes made.

@@
     \textbf{maximal upper bound} of $X$ if $\nexists a \in X$ such
-    that $x_m \posetr a$.  A maximal upper bound on $X$ is
+    that $x_m \ne a$ and $x_m \posetr a$.  A maximal upper bound on $X$ is
     denoted by $mub(X)$.
@@
 
-\begin{definition}[Graph, Tree, Path]
+\begin{definition}[Graph, Tree, Path, Graph Isomorphism]
   \label{def:graphtree}
@@
     \hangindent \defindent
-  \item An \textbf{(undirected) graph} is $G = (V,E)$ is \stt $V$ is a
-    finite set and $E$ is a set of ordered pairs, $E \subseteq V
+  \item An \textbf{(undirected)} \textbf{graph} is $G = (V,E)$ is \stt $V$ is a
+    finite set and $E$ is a set of unordered pairs, $E \subseteq V
     \times V$. An element in $V$ is called a \textbf{vertex
@@
 
-  \item A \textbf{tree} is a graph $T$ that has no subgraph that is a
+  \item A \textbf{tree} is a connected graph $T$ that has no subgraph that is a
     cycle.
+
+  \item A \textbf{graph isomorphism} between two graphs $G_1$ and
+    $G_2$ is a bijection $\phi: V(G_1) \rightarrow V(G_2)$ such that
+    $u, v \in V(G_1)$ are adjacent in $G_1$ if and only if $\phi(u),
+    \phi(v)$ are adjacent in $G_2$. If such a bijection exists, it is
+    said that $G_1$ is \textbf{isomorphic} to $G_2$, denoted by $G_1
+    \cong G_2$.
   \end{enumerate}
@@
 \item $G$ is an \textbf{interval graph} if there exists a set of
-  intervals $\cI$ \stt $G$ is isomorphic to the intersection graph of
+  intervals (of integers) $\cI$ \stt $G$ is isomorphic to the intersection graph of
   $\cI$. \[G \cong \bI(\cI)\]
   \item $G$ is a \textbf{path graph} if there exists a (undirected
-    unrooted) tree $T$ and a set of paths $\cP$ from $T$ \stt $G$ is
+    unrooted) tree $T$ and a set system of paths $\cP$ from $T$ \stt $G$ is
     isomorphic to the intersection graph of $\cP$. \[G \cong
@@
 set in the system maps to a path in $T$.  We refer to this as the
-{\CFTPL} problem or simply {\em tree path labeling} problem for an
-input set system and target tree pair -- $(\cF,T)$. The special case
-of the target tree being a path, is the interval assignment problem.
-We focus on generalizing the notion of an ICPIA \cite{nsnrs09} to
-characterize feasible path assignments.  We show that for a given set
-system $\cF$, a tree $T$, and an assignment of paths from $T$ to the
-sets, there is a feasible\footnote{The notion of {\em feasibility} is
-  formally defined in Section~\ref{ch:prelims}.}  bijection between
-$U$ and $V(T)$ if and only if the intersection cardinalities among any
-three sets (not necessarily distinct) is equal to that of the
-corresponding paths assigned to them and the input passes a filtering
-algorithm (described in this paper) successfully.  This algorithmic
-characterization gives a natural data structure that stores all the
-feasible bijections between $U$ and $V(T)$. This reduces the search
-space for the solution considerably from the universe of all possible
-bijections between $U$ and $V(T)$ to only those bijections that
-maintain the characterization.  Further, the filtering algorithm is
-also an efficient algorithm to test if a tree path
-labeling\footnote{The terms {\em tree path labeling} and {\em tree
-    path assignment} are, in informal language, synonyms. Formally,
-  the former refers to the bijection $\cl: \cF \rightarrow \cP$. The
-  latter refers to the set of ordered pairs $\{(S, P) \mid S \in \cF,
-  P \in \cP\}$. $\cP$ is a set of paths on $T$.} is feasible.
+{\CFTPL} problem \footnote{Problem is formally defined in
+  Section~\ref{sec:myresearchintro}.} or simply {\em tree path
+  labeling} problem for an input set system and target tree pair --
+$(\cF,T)$. The special case of the target tree being a path, is the
+interval assignment problem.  We focus on generalizing the notion of
+an ICPIA \cite{nsnrs09} to characterize feasible path assignments.  We
+show that for a given set system $\cF$, a tree $T$, and an assignment
+of paths from $T$ to the sets, there is a feasible\footnote{The notion
+  of {\em feasibility} is formally defined in
+  Section~\ref{ch:prelims}.}  bijection between $U$ and $V(T)$ if and
+only if the intersection cardinalities among any three sets (not
+necessarily distinct) is equal to that of the corresponding paths
+assigned to them and the input passes a filtering algorithm (described
+in this thesis) successfully\footnote{Formally defined by
+  Theorem~\ref{th:charac}}.  This algorithmic characterization gives a
+natural data structure that stores all the feasible bijections between
+$U$ and $V(T)$. This reduces the search space for the solution
+considerably from the universe of all possible bijections between $U$
+and $V(T)$ to only those bijections that maintain the
+characterization.  Further, the filtering algorithm is also an
+efficient algorithm to test if a tree path labeling\footnote{The terms
+  {\em tree path labeling} and {\em tree path assignment} are, in
+  informal language, synonyms. Formally, the former refers to the
+  bijection $\cl: \cF \rightarrow \cP$. The latter refers to the set
+  of ordered pairs $\{(S, P) \mid S \in \cF, P \in \cP\}$. $\cP$ is a
+  set of paths on $T$.} is feasible.
 

### 2013-03-21 src/020chap2.tex
Reviewer 1 (CRS) changes made.

@@
     \begin{align*}
-      A/B &= A \setminus B \cup \{b\}\\
+      A/B &= A \setminus B \cup \{x_B\}\\
       \cF/B &= \{ S/B \mid S \in \cF, S \nsubseteq B \}
     \end{align*}
-    $b$ is a new element added as a representative for $B$.
+    $x_B \notin U$ is a new element introduced as shown above as a
+    representative for $B$ after removing elements of $B$.
  \end{enumerate}
@@
 of $U$ is a composition of two permutations with respect to $H$ - (i) a
-permutation $\gamma$ of $H$ and (ii) a permutation $\beta$ of $U
-\setminus H$.
+permutation $\gamma$ of $H$ and (ii) a permutation $\beta$ of $U / H$.
 
@@
 \item $A$ is a subset of $B$: $A/B$ is not defined
-\item $B$ is a subset of $A$: $A/B = A \setminus B \cup \{b\}$, where $b$
+\item $B$ is a subset of $A$: $A/B = A \setminus B \cup \{x_B\}$, where $x_B$
 is a new element not in $U$ added to represent $B$.
@@
   then a \PQRtree $T$ for $\cF$ can be obtained by replacing the leaf
-  $h$ in $T_1$ by $T_2$, where $h$ is the representative element for
+  $x_h$ in $T_1$ by $T_2$, where $x_H$ is the representative element for
   $H$ in $T_i$.
@@
 largely an extension of the ICPIA characterization. In the new work,
-we adopt the notation used in \cite{kklv10} (also described in that
-chapter).
+we adopt the notation used in \cite{kklv10} namely, that of hypergraph
+isomorphism (also described in that chapter).
 
@@
   \label{def:incompatibility}
-  Let $\cF$ be a set system with universe $U$. Let $A_{\cF} = \{(a,b)
+  Let $\cF$ be a set system with universe $U$. Let $A_U = \{(a,b)
   \mid a, b \in U, a \ne b\}$. The {\em incompatibility graph}
@@
   \begin{enumerate}
-  \item   The vertex set of $G_I^\cF$ is $ A_{\cF}$
+  \item   The vertex set of $G_I^\cF$ is $A_U$
   \item  The edge set of $G_I^\cF$ are pairs of the following forms
@@
 involve all elements in $U$ and each vertex denotes a pair of elements. Secondly, the reverse of a \COP order is
-also a \COP order. Thus $A_\cF \setminus I$ must also be an
+also a \COP order. Thus $A_U \setminus I$ must also be an
 independent set. In other words, $G_I^\cF$ must be bipartite with

### 2013-03-21 src/040chap4.tex
Reviewer 1 (CRS) changes made.

@@
 graph. There has also been work done on the generalization of clique
-trees to clique hypergraphs \cite{km02}.  Path graphs are well studied
+trees to clique hypergraphs \cite{km02}.  Connections between
+hypergraph isomorphism and properties of set systems is also explored
+in \cite{br72} by extending Whitney's theorem of graph isomorphism to
+hypergraphs.  Moreover, \cite{f80} generalizes \cite{br72} and
+\cite{fg65} by characterizing the isomorphism of two hypergraphs by
+means of equicardinality of certain edge intersections and the
+exclusion of certain pairs of subhypergraphs.  Thus ICPPL is also a
+variant of the characterization in \cite{f80} with one of the
+hypergraphs having hyperedges that are paths from a tree (path
+hypergraph) and ICPPL uses only equicardinality of edge intersections
+of at most three hyperedges.
+
+Path graphs are well studied
 in the literature (characterization for \uvgraphs was perhaps first

### 2013-03-22 src/response_to_reviewers.tex
Most of reviewer 2 comments done.

@@
 \documentclass{article}
+\usepackage[normalem]{ulem}
 \usepackage{../lib/mythesislib}
 \usepackage{../lib/mylatexlib}
+%
 \begin{document}
 
-\title{Response to M.S. Thesis Reviewers\\ 
- {\em Thesis Title ``Generalization of the Consecutive-ones Property''}}
-\author{Anju Srinivasan}
+\title{Response to M.S. Thesis Review Comments\\ 
+ {Thesis Title: {\em Generalization of the Consecutive-ones Property}}}
+\author{Anju Srinivasan\\IIT Madras, Chennai - 600036.}
 \date{25 March 2013}
@@
 
-\subsection{Observations}
-The following are a few observations mentioned in the review document
-with my responses/changes.
-
-\begin{enumerate}
-\item Observation: (in Page 1, second last paragraph) {\bf The result on characterization of TPL in
-    the thesis is a variant of the 1978 JCT(B) result due to Fournier
-  who showed that two hypergraphs $H_1 = (V_1, X)$ and $H_2 = V_2, Y$
-such that $|X| = |Y| = m$ are isomorphic iff there is a bijection
-$\phi$ on $I = \{1, ... , m\}$ s.t. for every $L \subset I,
-|\bigcap_{i \in L}X_i| = |\bigcap_{j=\phi(i),i \in L}Y_j|$} \\
-Response: This is an important observation and was not cited in the thesis. Also \cite{f80}
-generalizes \cite{br72} and \cite{fg65} by characterizing the
-isomorphism of two hypergraphs by means of equicardinality of certain
-edge intersections and the exclusion of certain pairs of
-subhypergraphs. TPL characterization is a special case of this with
-one of the hypergraphs having hyperedges that are paths from a tree
-and the characterization only uses edge intersections of at most 3
-hyperedges. This has been added in the Conclusions chapter Page 72
-under {\em Graph isomorphism and logspace canonization}.
-
-\item Observation: (in Page 2, first paragraph) {\bf A $k$-subdivided tree is
-  a tree with a single central vertex with a number of paths each of
-  $k$ edges emanating from it.}\\
-Response: $k$-subdivided as defined in this thesis (Section
-3.2.4, Page 43) is such that each
-of the aforementioned paths are of length ``$k+1$'' edges since each such
-path has $k$ nodes apart from the central node and the leaf node, i.e.
-$k+2$ nodes. No change in the thesis.
-\end{enumerate}
-
-\subsection{Comments}
+\subsection{Comments on corrections}
 Responses to the comments in review document Page 2 items 0 to 7.
@@
 \begin{enumerate}
-\item[0] Comment: {\bf In thesis Page 7, Definition 1.3.7: After $x_m
+\item[0.] Comment: {\bf In thesis Page 7, Definition 1.3.7: After $x_m
     \le a$, you need to add $x_m \ne a$.}\\
@@
 
-\item Comment: {\bf In thesis Page 7, Definition
-    1.3.8: $E$ is a set of ``unordered'' pairs. You need to add the
+\item Comment: {\bf In thesis Page 7, Definition 1.3.8: $E$ is a set
+    of ``unordered'' pairs. You need to add the
     word ``unordered''}\\
-Response: Suggested change made in to definition of Undirected Graph Page 7 Definition 1.3.8 (1).
+  Response: Suggested change made in to definition of Undirected Graph
+  Page 7 Definition 1.3.8 (1).
 
@@
     acyclic graph}\\
-Response: The word ``connected'' has been added to the definition of Tree in
-Page 7 Definition 1.3.8 (3).
+  Response: The word ``connected'' has been added to the definition of
+  Tree in Page 7 Definition 1.3.8 (3).
 
@@
     isomorphism between graphs for the sake of completeness}.\\
-Response: Definition of Interval graph has been changed to say ``a set
-of intervals (of integers)'' in Page 8, Definition 1.3.10. Graph
+  Response: Definition of Interval graph has been changed to say ``a
+  set of intervals (of integers)'' in Page 8, Definition 1.3.10. Graph
   isomorphism definition has been added in Page 7, Definition 1.3.8(4)
-as follows: A \textbf{graph isomorphism} between two graphs $G_1$ and
-    $G_2$ is a bijection $\phi: V(G_1) \rightarrow V(G_2)$ such that
-    $u, v \in V(G_1)$ are adjacent in $G_1$ if and only if $\phi(u),
+  as follows:
+
+  {\em A \textbf{graph isomorphism} between two graphs $G_1$ and $G_2$
+    is a bijection $\phi: V(G_1) \rightarrow V(G_2)$ such that $u, v
+    \in V(G_1)$ are adjacent in $G_1$ if and only if $\phi(u),
     \phi(v)$ are adjacent in $G_2$. If such a bijection exists, it is
     said that $G_1$ is \textbf{isomorphic} to $G_2$, denoted by $G_1
-    \cong G_2$.
+    \cong G_2$.}
 
@@
     path graph: You need to emphasise that $\cP$ forms a set system}\\
-Response: This change has been made - ``a set system of paths $\cP$''
-in Page 8 Definition 1.3.10(2).
+  Response: This change has been made - ``a set system of paths
+  $\cP$'' in Page 8 Definition 1.3.10(2).
 
@@
     ``$b$ is a new element added..''?}.\\
-Response: This was in thesis Page 26. $A/B$ is created by removing the elements of second operand set $B$ from
-the first operand set $A$ and introducing a new element {\em not} in the
-universe to $A$ to represent set $B$. This is a set operation 
-mentioned in \cite{mm96} (they use ``$B$'' to denote the new element to represent
-set $B$ which is also confusing) which is eventually used to decompose
-permutation of elements in $A$ to permutation of elements in $A/B$ and
-permutation of elements in $B$.
-Perhaps it would be clearer if
-$b$ is replaced by $x_B$. This change has been made in Page 26
-Definition 2.2.4 (7) and other mentions of this operation ``/'' Page
-27 item (iii) and $h$ has been replaced with $x_H$ in Corollary
-2.2.8. There was also a typo in Page 27 paragraph under Corollary
-2.2.8 item (ii) where ``$U \setminus H$'' was written instead of $U / H$.
-This has also been corrected.
+  Response: This was in thesis Page 26. $A/B$ is created by removing
+  the elements of second operand set $B$ from the first operand set
+  $A$ and introducing a new element {\em not} in the universe to $A$
+  to represent set $B$. This is a set operation mentioned in
+  \cite{mm96} (they use ``$B$'' to denote the new element to represent
+  set $B$ which is also confusing) which is eventually used to
+  decompose permutation of elements in $A$ to permutation of elements
+  in $A/B$ and permutation of elements in $B$.  Perhaps it would be
+  clearer if $b$ is replaced by $x_B$. This change has been made in
+  Page 26 Definition 2.2.4 (7) and other mentions of this operation
+  ``/'' Page 27 item (iii) and $h$ has been replaced with $x_H$ in
+  Corollary 2.2.8. There was also a typo in Page 27 paragraph under
+  Corollary 2.2.8 item (ii) where ``$U \setminus H$'' was written
+  instead of $U / H$.  This has also been corrected.
 
@@
   dependence in $\cF$. It is only dependent on the universe $U$ of
-$\cF$. All occurences of $A_\cF$ has been changed to $A_\U$ in Page 36.
+  $\cF$. All occurences of $A_\cF$ has been changed to $A_U$ in Page
+  36.
+\end{enumerate}
+
+\subsection{Other comments}
+The following are a few observations mentioned in the review document
+with my responses/changes.
+
+\begin{enumerate}
+\item Observation: (in Page 1, second last paragraph) {\bf The result
+    on characterization of TPL in the thesis is a variant of the 1978
+    JCT(B) result due to Fournier who showed that two hypergraphs $H_1
+    = (V_1, X)$ and $H_2 = V_2, Y$ such that $|X| = |Y| = m$ are
+    isomorphic iff there is a bijection $\phi$ on $I = \{1, ... , m\}$
+    s.t. for every $L \subset I,
+    |\bigcap_{i \in L}X_i| = |\bigcap_{j=\phi(i),i \in L}Y_j|$} \\
+  Response: This is an important observation and was not cited in the
+  thesis. Also \cite{f80} generalizes \cite{br72} and \cite{fg65} by
+  characterizing the isomorphism of two hypergraphs by means of
+  equicardinality of certain edge intersections and the exclusion of
+  certain pairs of subhypergraphs. TPL characterization is a special
+  case of this with one of the hypergraphs having hyperedges that are
+  paths from a tree and the characterization only uses edge
+  intersections of at most 3 hyperedges. This has been added in the
+  Conclusions chapter Page 72 under {\em Graph isomorphism and
+    logspace canonization}.
+
+\item Observation: (in Page 2, first paragraph) {\bf A $k$-subdivided
+    tree is a tree with a single central vertex with a number of paths
+    each of $k$ edges emanating from it.}\\
+  Response: $k$-subdivided as defined in this thesis (Section 3.2.4,
+  Page 43) is such that each of the aforementioned paths are of length
+  ``$k+1$'' edges since each such path has $k$ nodes apart from the
+  central node and the leaf node, i.e.  $k+2$ nodes. No change in the
+  thesis.
 \end{enumerate}
@@
 \section{Response to Reviewer 2}
+
+\section{Other comments}
+\begin{enumerate}
+\item Observation: {\bf The definition of $k$-subdivided star is not
+    given in the thesis. What I understood is that it has a root node $v$
+    of degree $k$, the degree of all other nodes is at most 2, and
+    each path from root to leaf is $k+2$.}\\
+Response: The definition of $k$-subdevided star was given in Section
+3.2.4, Definition 3.2.6. Perhaps this has not been sufficiently
+referred to, hence references to this definition has been made in
+Chapter 1 Page 13 second last paragraph where it is first referred and
+in Chapter 3 Page 39 second paragraph where {\sc Compute
+  $k$-subdivided Star Path Labeling} problem is defined.
+
+The reviewer's understanding is 
+correct except that the root node can have any degree, thus a
+$k$-subdivided star is a collection of graphs with degree of all
+non-root nodes as at most 2, each path from root to leaf is $k+2$ and 
+the root having any degree. See thesis Figure 1.5 in page 13 for examples.
+
+\item Comment: {\bf The proof of Lemma 3.3.4 is unclear.}\\
+Response: The proof was correct but terse. It has been elaborated to as
+follows:
+
+  {\em 
+  Lemma:   Let $(\cF, \cl)$ be an \ICPPL and $\cl(S_i)=P_i$, $S_i \in \cF$, $1
+  \leq i \leq 4$.  Then, $|\cap_{i=1}^4 S_i| = |\cap_{i=1}^4 P_i|$.
+
+  Proof: Consider the set of set intersections with $S_1$ $\cF' = \{S_2 \cap S_1, S_3
+  \cap S_1, S_4 \cap S_1\}$ and let $\cl'$ be a tree path labeling of
+  $\cF'$ such that $\cl'(S_i \cap S_1) = P_i \cap P_1, 2 \le i \le
+  4$. Clearly, $P_2 \cap P_1$, $P_3 \cap P_1$, and $P_4 \cap P_1$ are
+  subpaths of path $P_1$, thus equivalent to intervals.  Due to the
+  three way intersection cardinality preservation property of the
+  ICPPL $(\cF, \cl)$, this new tree path labeling $(\cF', \cl')$
+  preserves pairwise intersection cardinalities.  Now by applying
+  Lemma 3.3.1 %~\ref{lem:intersection-cardinality} 
+  to the sets in $\cF'$ and
+  their corresponding path images due to $\cl'$, it follows that
+  $|\cap_{i=1}^4 S_i| =| \cap_{i=1}^4 P_i|$.}
+
+\item Comment: {\bf While studying Feasible Tree Path Labeling Problem
+  in Section 3.3, two filters are mentioned and their pseudocodes
+  given. If possible, state the problem in words. ... Explicitly state
+the roles of these two filters in the algorithm with proper
+figures.}\\
+Response:
+
+\item Comment: {\bf It is shown that TREE PATH LABELING PROBLEM on
+    arbitrary trees can be solved polynomial time if the two
+    subproblems, identified in section 3.6 in the thesis are
+    polynomial time solvable. The formal definition of these two
+    subproblems are not given. However, in page 71 item numbers 1 and
+    2 the subproblems are described with several notations, where the
+    motivation is unclear.}\\
+Response:
+
+\item Comment: {\bf ... it is shown that consecutive 1s property
+    testing for a \sout {graph} [binary matrix] is solvable in LOGSPACE. {\em This is a very
+      important contribution.}}\\
+Response: Since this is a conclusion from earlier results, it has not
+been mentioned as a ``new'' result. It is mentioned in the abstract as a
+derived result.
+\end{enumerate}
 

### 2013-03-22 src/010chap1.tex
Most of reviewer 2 comments done.

@@
 We present a polynomial time algorithm for {\CFTPL} when the target
-tree $T$ belongs to a special class of trees called {\em \kstars} and
+tree $T$ belongs to a special class of trees called {\em \kstars} (see
+Definition~\ref{def:ksubstar} for the formal definition) and
 when the hyperedges in the hypergraph $\cF$ have at most $k+2$

### 2013-03-22 src/030chap3.tex
Most of reviewer 2 comments done.

@@
 target trees -- namely, \CFTPLINT and \CFTPLKTREE on intervals and
-\kstars, respectively.  Section~\ref{sec:spltargettree} discusses
+\kstars (see Definition~\ref{def:ksubstar}), respectively.  Section~\ref{sec:spltargettree} discusses
 these problems and their polynomial time solutions.  \CFTPLINT is the
@@
 \begin{proof}
-  Consider the sets $S_2 \cap S_1$, $S_3 \cap S_1$, and $S_4 \cap
-  S_1$, and their images $P_2 \cap P_1$, $P_3 \cap P_1$, and $P_4 \cap
-  P_1$ respectively.  It is clear that this path labeling preserves
-  pairwise intersection cardinalities.  In this case, it is clear that
-  the 3 sets are assigned paths in the path $P_1$.  Now by applying
-  Lemma~\ref{lem:intersection-cardinality}, it follows that
+  Consider the set of set intersections with $S_1$ $\cF' = \{S_2 \cap S_1, S_3
+  \cap S_1, S_4 \cap S_1\}$ and let $\cl'$ be a tree path labeling of
+  $\cF'$ such that $\cl'(S_i \cap S_1) = P_i \cap P_1, 2 \le i \le
+  4$. Clearly, $P_2 \cap P_1$, $P_3 \cap P_1$, and $P_4 \cap P_1$ are
+  subpaths of path $P_1$, thus equivalent to intervals.  Due to the
+  three way intersection cardinality preservation property of the
+  ICPPL $(\cF, \cl)$, this new tree path labeling $(\cF', \cl')$
+  preserves pairwise intersection cardinalities.  Now by applying
+  Lemma~\ref{lem:intersection-cardinality} to the sets in $\cF'$ and
+  their corresponding path images due to $\cl'$, it follows that
   $|\cap_{i=1}^4 S_i| =| \cap_{i=1}^4 P_i|$.
@@
 In this section we consider the problem of assigning paths from a
-$k$-subdivided star $T$ to a given set system $\cF$ as described by
-the \CFTPLKTREE problem.  We consider
+$k$-subdivided star (see Definition~\ref{def:ksubstar}) $T$ to a given
+set system $\cF$ as described by the \CFTPLKTREE problem.  We consider
 $\cF$ for which the overlap graph $\bO(\cF)$ is connected.  The

### 2013-03-23 src/response_to_reviewers.tex
All changes made as per review. Viva presentation added.

@@
-\documentclass{article}
+\documentclass[12pt,a4paper]{article}
+\usepackage{fullpage}
 \usepackage[normalem]{ulem}
@@
 \usepackage{../lib/mylatexlib}
+\usepackage{setspace}
 %
+
 \begin{document}
@@
 
+\onehalfspacing
+
 \section{Reviewers}
@@
 \end{enumerate}
+The reviewers careful comments are acknowledged and
+appreciated. Responses to the same are given in following sections.
 
@@
     JCT(B) result due to Fournier who showed that two hypergraphs $H_1
-    = (V_1, X)$ and $H_2 = V_2, Y$ such that $|X| = |Y| = m$ are
+    = (V_1, X)$ and $H_2 = (V_2, Y)$ such that $|X| = |Y| = m$ are
     isomorphic iff there is a bijection $\phi$ on $I = \{1, ... , m\}$
@@
 \item Observation: (in Page 2, first paragraph) {\bf A $k$-subdivided
-    tree is a tree with a single central vertex with a number of paths
+    \sout{tree} [star] is a tree with a single central vertex with a number of paths
     each of $k$ edges emanating from it.}\\
-  Response: $k$-subdivided as defined in this thesis (Section 3.2.4,
+  Response: $k$-subdivided star as defined in this thesis (Section 3.2.4,
   Page 43) is such that each of the aforementioned paths are of length
@@
 
-\section{Other comments}
 \begin{enumerate}
@@
 figures.}\\
-Response:
+Response: The problems are described in words but they were part of a
+long narrative. It has been summarised now in Page 44 before going
+into details of the algorithms as follows:
+
+{\em \begin{enumerate}
+\item [Filter 1.] \filteri Refine $\cF$ such that the resulting
+  labeling will not have paths that share a leaf thus each leaf being
+  unique to a path. This is done by breaking the paths into subpaths and their
+  corresponding preimage sets as described in Algorithm 1.
+\item [Filter 2.] \filterii Find the element in universe $U$ that maps to each
+  leaf in $T$ as described in Algorithm 2. 
+\end{enumerate}
+Remove the leaves from $T$ and their corresponding preimages from $U$
+and call the filters again. This is repeated until the resulting
+truncated tree is a path. The remaining mapping can be found using
+ICPIA.
+}
+
+Hopefully this makes the role of the filter algorithms clear.
+I felt that describing details of the pseudocode in words in the narrative would
+be tedious and redundant and the pseudocode has comments that describe what is being
+done whenever it is not obvious. Regarding figures, Section 3.4 in
+page 52 has an example worked out in detail with figures. This was
+added with the intention of illustrating the filter algorithms.
 
@@
     motivation is unclear.}\\
-Response:
+Response: The prime submatrices partial order theory and notations are
+required for a formal definition.
+The two subproblems are defined now as \FOCPS and \FMFT and motivation is
+given in Page 61 before setting up the theory and formal defintion is
+given in page 66 after setting up the theory required to define it.
+
+{\em
+
+\begin{problemdef}{\FOCPS}{A hypergraph $\cF$ with its in-tree
+    partitions $\{\bP_1,\ldots,\bP_r\}$ and
+    a tree $T$.}
+  Compute a partition of $T$ into subtrees $\{T_1, \ldots, T_r\}$ such
+  that there exists an \ICPPL for $mub(\bP_i)$ from subtree
+  $T_i$ for all $i \in [r]$ and there exists an ICPIA for all other
+  prime submatrices in $\bP_i$ as claimed in Lemma 3.6.8.
+\end{problemdef}
+
+\begin{problemdef}{\FMFT}{An in-tree partition $\bP_i$ of a hypergraph
+    $\cF$ and a subtree $T_i$ of a tree $T$.}
+  Compute a feasible TPL for $mub(\bP_i)$ from subtree $T_i$.
+\end{problemdef}
+}
 
 \item Comment: {\bf ... it is shown that consecutive 1s property
-    testing for a \sout {graph} [binary matrix] is solvable in LOGSPACE. {\em This is a very
+    testing for a \sout {graph} [hypergraph] is solvable in LOGSPACE. {\em This is a very
       important contribution.}}\\
 Response: Since this is a conclusion from earlier results, it has not
-been mentioned as a ``new'' result. It is mentioned in the abstract as a
+been mentioned as a completely ``new'' result. It is mentioned in the abstract as a
 derived result.

### 2013-03-23 src/030chap3.tex
All changes made as per review. Viva presentation added.

@@
 retaining a non-empty subset of isomorphisms, if any, between $\cF$
-and $\cF^\cl$. These two filters are \filteri in
-Algorithm~\ref{al:filterone} and \filterii in
-Algorithm~\ref{al:filtertwo}.
+and $\cF^\cl$. These two filters are \filteri and \filterii which in
+summary are as follows:
+\begin{enumerate}
+\item \filteri Refine $\cF$ such that the resulting
+  labeling will not have paths that share a leaf thus each leaf being
+  unique to a path. This is done by breaking the paths into subpaths and their
+  corresponding preimage sets as described in Algorithm~\ref{al:filterone}.
+\item \filterii Find the element in universe $U$ that maps to each
+  leaf in $T$ as described in Algorithm~\ref{al:filtertwo}. 
+\end{enumerate}
+Remove the leaves from $T$ and their corresponding preimages from $U$
+and call the filters again. This is repeated until the resulting
+truncated tree is a path. The remaining mapping can be found using
+ICPIA.
+
 % One direction of this claim isclear: that if a path labeling is
@@
 
+The two subproblems \FOCPS and \FMFT are informally defined as
+follows. The precise definition is given after setting up the theory
+and notations required in the following paragraphs.
+\begin{enumerate}
+\item \FOCPS - {\em Dividing the tree $T$ to subtrees for overlap
+    component partitions: } Using the overlap partial ordering of
+  hyperedges in $\cF$, the hyperedges can be divided into partitions
+  of sub-hypergraphs (called partitions of prime submatrices) that are
+  disjoint from each other. Being disjoint they must be assigned to
+  disjoint subtrees of $T$ such that \FMFT can be solved successfully
+  and the other sub-hypergraphs in the partition will have an ICPIA
+  in accordance with the overlap property. Finding the right partition
+  of $T$ into subtrees is the first subproblem.
+
+\item \FMFT - {\em Finding ICPPL for $mub$ of each sub-hypergraph
+    partition: } In each partition of sub-hypergraph, due to the
+  overlap partial ordering, there will be an $mub$ sub-hypergraph that
+  contains all the other sub-hypergraphs in that partition. If $T_i$
+  is the subgraph assigned to this partition by solving \FOCPS, $T_i$
+  is essentially the tree for the $mub$'s TPL. Computing an \ICPPL for
+  this $mub$ sub-hypergraph from $T_i$ is the second subproblem.
+\end{enumerate}
+
 % with which the intersection is {\em
@@
 %\noindent {\bf Finding an assignment of tree paths to a set system.}
+
 As described by Definition~\ref{def:matrixsetsystem} a set system or
@@
 algorithmic challenges in finding an ICPPL for a given hypergraph
-$\cF$ on a target tree $T$.  
+$\cF$ on a target tree $T$, namely subproblems \FOCPS and \FMFT respectively.
+
 \begin{enumerate}
-\item \label{it:challenge1} Finding $X$ and its partition into in-trees
-  $\{\bP_1,\ldots,\bP_r\}$, can be done in polynomial time.  On the
-  other hand, as per Lemma~\ref{lem:subicppa} we need to partition $T$
-  into vertex disjoint subtrees $\{T_1, \ldots, T_r\}$ such that for
-  each $i \in [r]$, the prime submatrices in $\bP_i$ have an ICPPL in
-  $T_i$ (to be accurate, the prime submatrix $mub(\bP_i)$ has an ICPPL
-  in $T_i$ and the rest of the prime submatrices have an ICPIA as
-  claimed in Lemma~\ref{lem:rooticppa}).  This seems to be a
-  challenging step, and it must be remarked that this step is easy
-  when $T$ itself is a path, as each individual $T_i$ would be
-  sub-paths.
-\item \label{it:challenge2} The second algorithmic challenge is identified by
-  Lemma~\ref{lem:rooticppa} which is to compute an ICPPL from the
-  chosen subtree $T_i$
-  to the matrix associated with $mub(\bP_i)$.
+
+\item \label{it:challenge1} \FOCPS - Finding $X$ and its partition into
+  in-trees $\{\bP_1,\ldots,\bP_r\}$, can be done in polynomial time.
+  On the other hand, as per Lemma~\ref{lem:subicppa} we need to
+  partition $T$ into vertex disjoint subtrees $\{T_1, \ldots, T_r\}$
+  such that for each $i \in [r]$, the prime submatrices in $\bP_i$
+  have an ICPPL in $T_i$ (to be accurate, the prime submatrix
+  $mub(\bP_i)$ has an ICPPL in $T_i$ and the rest of the prime
+  submatrices have an ICPIA as claimed in Lemma~\ref{lem:rooticppa}).
+  This seems to be a challenging step, and it must be remarked that
+  this step is easy when $T$ itself is a path, as each individual
+  $T_i$ would be sub-paths.
+\item \label{it:challenge2} \FMFT - The second algorithmic challenge is
+  identified by Lemma~\ref{lem:rooticppa} which is to compute an ICPPL
+  from the chosen subtree $T_i$ to the matrix associated with
+  $mub(\bP_i)$.
 \end{enumerate}
+
+The formal definitions of the problems are given below.
+
+\begin{problemdef}{\FOCPS}{A hypergraph $\cF$ with its in-tree
+    partitions $\{\bP_1,\ldots,\bP_r\}$ and
+    a tree $T$.}
+  Compute a partition of $T$ into subtrees $\{T_1, \ldots, T_r\}$ such
+  that there exists an \ICPPL for $mub(\bP_i)$ from subtree
+  $T_i$ for all $i \in [r]$ and there exists an ICPIA for all other
+  prime submatrices in $\bP_i$ as claimed in Lemma~\ref{lem:rooticppa}.
+\end{problemdef}
+
+\begin{problemdef}{\FMFT}{An in-tree partition $\bP_i$ of a hypergraph
+    $\cF$ and a subtree $T_i$ of a tree $T$.}
+  Compute a feasible TPL for $mub(\bP_i)$ from subtree $T_i$.
+\end{problemdef}
 

### 2013-04-12 src/030chap3.tex
corrected an error in lemma lem:priv-card

@@
   \in [n]\}$ and path labeling $\cl'(S_i \cap S) = P_i \cap P$.  Since
-  $(\cF, \cl)$ is an \ICPPL, it can be easily verified that $(\cF',
+  $(\cF, \cl)$ is an ICPPL, it can be easily verified that $(\cF',
   \cl')$ is an ICPIA on path (interval) $P$.  Let $S_{two}$ and
   $P_{two}$ be the elements of $S$ and $P$ which are in one {\em
-    other} set and path, respectively \ie $S_{two} = S \cap (\cup_{S_i
+    other} set and path, respectively i.e. $S_{two} = S \cap (\cup_{S_i
     \ne S} S_i)$ and $P_{two}$ is defined analogously. It may be noted
-  that $S_{two} = \supt{\cF'}$ and $P_{two} = \supt{\cF'^{\cl'}}$.
+  that $S = \supt({\cF'})$ and $P = \supt({\cF'}^{\cl'})$.
   From Lemma~\ref{lem:icpia} it follows that there is a bijection
-  $\phi$ from $S_{two}$ to $P_{two}$ such that for each $i$, the image
+  $\phi$ from $S$ to $P$ such that for each $i$, the image
   of $S_i \cap S$ under $\phi$ is $P_i \cap P$ (definition of
-  feasibility).  Moreover, this means $|S_{two}| = |P_{two}|$. By
+  feasibility).  Moreover, this means that in $\phi$,
+the elements of $S_{two}$ are mapped to the elements of $P_{two}$.  The proof is of this claim is by contradiction. Let us assume that there is an $x \in S_{two}$ is mapped to
+a $v \in P_{priv}$ under $\phi$.
+Since $x \in S_{two}$, let $S_i$ be another set that contains $x$, other than $S$.
+Then it means that $\phi(S_i \cap S)$ which is same as $\cl'(S_i \cap S) =P_i \cap P$ contains $v$.
+This implies that $v \in P_i \cap P$, contradiction to the fact that $v \in P_{priv}$.
+Therefore, under the bijection $\phi$, the image of the set $S_{two}$ is the set $P_{two}$.
+Conseqently, it follows that $|S_{two}| = |P_{two}|$.  
+By
   definition, $S_{priv} = S \setminus S_{two}$ and $P_{priv} = P \setminus P_{two}$.
